### 2018-07-10 proposal_draft_2018.Rmd
first commit

---
title: "Kelp morphometric properties and the link to hydrodynamic modelling"
author:
- affiliation: University of the Western Cape
  name: Ross Coppin, AJ Smit, Christo Rautenbach
date: "2018"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    toc: no
bibliography: mycollection2.bib   
---

\newpage
## Background

Marine ecosystems are maintained by a variety of complex interactions between abiotic and
biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology [@Doney2012]. These abiotic and biotic variables, the interactions between them, and the various ecological processes, ultimately determine the community composition and ecological functioning of all ecosystems. Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and over-fishing [@Blamey2015]. Temperature and wave exposure have been recognised as important variables with regards to climate-driven changes within the ocean. In order to persist and survive within variable and changing environments, organisms must either migrate, adapt, or die.

Seaweeds are sessile organisms which are unable to migrate to new areas when local environmental conditions become unsuitable and therefore are forced to adapt to new conditions in order to avoid expiration. The main form of mortality for seaweeds is through mechanical dislodgment by wave action. Seaweeds, particularly brown seaweeds, are able to undergo rapid morphological adaptation to the hydrodynamic environment. This allows seaweeds to reduce mortality through mechanical dislodgement by inducing morphological changes which reduces overall drag. Seaweeds that are unable to avoid mechanical dislodgement either raft out to sea or wash up onto beaches, otherwise known as beach-cast. However, not all the beach-cast may have originated from a nearby kelp population and may have originated from other sites or regions of the coast through a combination of rafting and ocean currents. Therefore, because kelp morphology is specific to its local environment the morphological features may be able to indicate, within a certain amount of probability, what site or region it most likely originated from. In other words, beach-cast kelp may be used as proxy for investigating the flow of  coastal currents. Using kelp as a proxy for determining its original location will be calibrated by means of a hydrodynamic model which will be designed from already existing SWAN and Delf3D models. This combined approach will allow investigation into flow regimes around the west and south-west coasts of South Africa and the role they play in sub-tidal and beach ecology. For instance, micro-plastics are recognised as a threat to marine life, however very little is known about how micro-plastics may be transported along the coast.The coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation.

The increased use of plastic in society over the past half century has resulted in large amounts of plastic litter in both the marine and terrestrial environment. The problems associated with large plastic debris have received attention for many decades, whereas those connected to marine micro-plastics comparatively received very little attention. However, today it has become a prioritized area among political organizations, agencies and NGOs around the world. The micro-plastic debris present in the ocean are derived from marine and terrestrial sources, however there is little understanding of how microplastics may be distributed with ocean currents. Therefore, the coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation. This in turn may inform management decisions with regards to eliminating and managing marine pollution in South Africa.

## Rationale

Seaweeds, browns in particular, are capable of adapting morphological characteristics to persist in changing and variable ocean environments. Changes in morphology have been shown to increase photosynthetic ability, enhance nutrient uptake, and reduce drag with regards to changes in temperature and wave exposure. Wave exposure has been shown to be an important driver of seaweed morphology, as the main mechanism of seaweed mortality is through the mechanism of dislodgment. Changing morphology, such as decreasing frond area or strengthening through stipe circumference increase/reduces drag and increases the probability of survival. However, locally adapted seaweed may still be dislodged in pulse disturbance events such as storms, and may raft far distances and wash up on beaches. An example would be *Ecklonia maxima* which is a conspicuous brown seaweed along the coastline and beaches around South Africa. Therefore, the beach-cast may not always originate from adjacent kelp populations but rather from other regions which the individual is adapted for. Therefore, the kelp morphology may act as a proxy for investigating coastal currents and changes thereof. The advances in ocean hydrodynamic modelling has made great progress and has been applied in a variety of ways. Therefore, this study will use advances in hydrodynamic modelling in combination with kelp morphological characteristics to investigate coastal currents. Once the model has been established it may be applied in other ways, such as investigating the transport of micorplastics along the South African coastline. The harms of microplastics to the marine environment has gained much traction in recent years, but research in South Africa is lacking. The coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation. This in turn may inform management decisions with regards to eliminating and managing marine pollution in South Africa.

## Research problem

Pollution is a huge environmental problem that affects both terrestrial and marine ecosystems. Pollution from land enters the sea where it could harm or kill marine organisms or is transported by ocean currents to other coastal areas. In these areas the pollution could re-enter the ocean or be blown by wind into terrestrail ecoystems where it could once again be ingested by oragnsims causing harm or even death. Pollution is therefore a major threat to the environment and marine organisms. In recent years the affect of microplastics on the ocean has gained much traction with scientists and politictions alike. Microplastics cosiste of tiny particles of plastic and other pollution and are thereofore are difficult to detect.

Currently there is no hydrodynamic model that is able to determine dispersion, source and accumulation of microplastics along the South African coastline. Given the significant deterimental affect microplastics play in the ocean, it is important that such a mechanism be developed that will aid in better management of marine pollution in South Africa. Furthermore, this project allows for a multidicisplinery approach to be taken by combining ecology and coastal oceanography.

## Preliminary literature review

Abiotic and biotic factors interact in complex ways which indirectly determine behavioral and ecophysiological responses in organisms. For example, when storms or strong currents form in sub-tidal habitats, sea urchins form aggregations in order to reduce overall drag to avoid being swept away by currents. Organisms that are motile my migrate into more environmentally suitable areas when conditions become unfavorable or food sources become depleted. In changing environments migration may also allow organisms to extend their distributions. For example, ocean warming off the coast of Western Australia has allowed tropical fish species to extend their distribution into areas that were previously characterised as temperate reefs. Sessile organisms are unable to migrate into more environmentally suitable areas and are forced to either adapt or suffer expiration. Sessile organisms may respond to changing environmental conditions through changes in physiology. For example, plants may produce heat shock proteins that help buffer the affect of temperature increases. Sessile organisms may also adapt their morphology,in order to persist in changing and harsh environments, such as seaweeds.

Temperature and wave exposure have been shown to be important drivers of seaweed distribution, physiological functioning, ability to recover, population dynamics and morphology. Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10- 20 m s-1 with accelerations of 400 m s-2 [@Friedland1995] are the biggest threat to kelp survival. Kelps are able to rapidly adapt their morphological characteristics to reduce drag and avoid dislodgment [@Blanchette1997]. For example a study by @Koehletal2008 showed that transplanted *Nereocystis luetkeana* plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Another study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and was a combination of *in situ* sampling and transplant of juvenile plants. The results showed that morphology differed between wave-sheltered and wave-exposed sites (thin thallus at sheltered sites and a narrow, thick thallus with a thick stipe at exposed sites), and was consistent with previous studies. Juveniles transplanted into wave exposed sites under went rapid morphological adaption, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaption.

Kelp morphology may be distinct to a particular region with a specific hydrodynamic environment and has the ability to raft far distances using coastal currents, and may accumulate as beach-cast in areas far from its original location. Therefore, kelp morphology may be used as a proxy for determining its original location as well as aid in characterising coastal currents. However, this approach must be combined with advances in hydrodynamic modelling for a quantitative outcome.

Advances in numerical modelling has gained much traction in recent years and has been applied in a variety of ways with regards to ecological studies. For example, a study by @wang2009 used the Delft3D-Flow model to assess the hydraulic suitability of a stream as a spawing ground for the Chinese Sturgeon (*Acipenser sinensis*) in the Yangtze River. The authors calculated the horizontal mean vorticity which was used to assess the hydraulic environment of spawning ground. The flow field state was determined through model simulation and field-measured data used to validate the model. The results added to existing scientific database for spawning ground hydraulic environmental protection. Different numerical models can often be integrated to model across ecosystem levels. For example a study by @leon2003 used integrated physical (Delft3D hydrological model) and bio-chemical (Agricultural Non-point Source model) processes models to investigate the possible impact on the Lake Malawi water quality due to management actions performed at the watershed level.

Since wave energy is an important driver in marine ecosystems, particularly kelp, the advances in hydrodynamic modelling offer a new opportunity for multifactoral and quantitative approach to research in marine ecosystems. The Delf3D and SWAN models have been used successfully in previous studies regarding brine plume discharge, impacts of storms, affects of climate change on the hydrological environment etc. The models have not been designed for shallow environments (<6m) and therefore may not be suitable to model coastal hydrological environments. However these models may be adjusted to suit coastal waters if they are combined with a new numerical model which can be calibrated to suit these needs.

In recent years there has been growing attention on microplastics and the deterimental role they play in the enviornment, particulary the ocean. Microplastics are tiny plastic granuales used as scrubbers in cosmetics and air-blasting, and small plastic fragments that originate from larger pieces of plastic known as macroplastics. The potential harms of small pieces of plastic was highlighted in the 1970's and renewed interest has lead to research showing that microplastics in the ocean are widespread. The small size of microplastics makes them bio-available to biota in the food-web which may cause problems with an organisms physiological functioning. Furthermore, the relatively large surface area and composition of microplastics provides an environmeht that is able of adhering to organic pollutants. In other words microplastics also act as a vector for transport and assimilation of organic polluants. Therefore, microplastics may also aid in the assimilation of organic pollutans into all levels of the food web.

Therefore, this study not only enables research into the ecological affects of the hydrological environment on an important habitat-forming organism, it also offers the opportunity to improve on current hydrological numerical models to suit coastal environments. This in turn will allow investigation into the flow and accumulation of microplastics which are regarded as a major threat to marine life. Furthermore, the calibrated model could be applicable to other ecological studies such as dispersal of benthic flora and fauna, climate change studies, forecasting etc.

## Aims of research

The aim of the project is to investigate coastal flow regimes along the west coast and south-west coast of South Africa and the role this may play in transport of kelp beach-cast and microplastics. This aim will be met through the following objectives:

A) Determine if the hydrodynamic environment is the main driver of kelp morphology and if this is specific to a location
B) Simulate kelp rafting by means of a hydrodynamic modelling and calibrate this model with *in situ* beach-cast morphometric data.
C) Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.

## Research design and methodology

A combined quantitative ecology and oceanography approach will be used for this study. Morphological characteristics and their relation to the hydrodynamic environment must first be established and the findings will form the basis for the proceeding research. Once the relationship has been established it will be used to calibrate future hydrodynamic models developed during the course of the study. Once the hydrodynamic model has been established and calibrated it will be used to investigate the flow of microplastics along the South African coastline.

# Morphology

Morphometrics collection Between October 2014 and April 2015, morphological measurements of *Laminaria pallida* and *Ecklonia maxima* were collected at 19 sites along the Western Cape coast of South Africa (Refer to morphology bar graphs). These varying morphometrics allowed measurements such as weight, length and thickness to be compared between sites. Because the macroalgae differ in construction, species-specific morphometrics were included. These sites span across the majority of the Western Cape coast, in varying thermal and wave energy regimes. Sites were chosen to represent an array of morphological differences seen within each *Ecklonia maxima* and *Laminaria*. Sites were also chosen to reflect locations that displayed variable wave and temperature regimes, to allow us to robustly test our hypothesis of environmental drivers influencing kelp morphology. St. Helena Bay and Betty’s Bay constituted the north western and south eastern boundary sites respectively. These sites are roughly 300km apart, and are confidently lie within separate marine provinces, as outlined above. Cape Peninsula provides an interesting topographical boundary that shelters the coast in False Bay. Sites were therefore chosen to represent an array of environments, from offshore reefs (Batsata Rock), to sheltered intertidal zones (Miller’s Point). West of Cape Point, a number of sites were chosen to highlight the presence of upwelling (Oudekraal, Kommetjie), as well as kelps growing in protected bays (Hout Bay).

# Hydrodynamic model

The hydrodynamic model for coastal flow will be developed using already existing models that have been developed. These models will consist of the SWAN and Delf3D models and the apporach is to use aspects of each model, that are related to coastal flow, from which further development will take place. Once the model has been developed it will be used to predict kelp morphology at random sites which will then be sampled. The *in situ* kelp morphology data collected at the predicted site will be used to calibrate the model. This process will be repeated until the model can predict coastal flow and kelp morphology within a certian degree of probability.

# Microplastics

The calibrated model developed in previous chapters will be used to determine flow, source and accumulatiom points along the South African coastline. Certian characteristcs of microplastics such as bouyancy, flow rate, fluid dynamics ect. Which characteristics are important for modeling dispersion of microplastics will be achieved by investigating the literature and  experimental simulations. Once the model has been established it will once again be calibrated through _in situ_ sampling.

## Time-frame

## References

### 2018-07-10 proposal_mini_2018.Rmd
first commit

--- 2018-07-10 proposal_draft_2018.Rmd
@@
 ---
-title: "Kelp morphometric properties and the link to hydrodynamic modelling"
+title: "Kelp morphometric properties and its link to the hydrodynamic modelling"
 author:
-- affiliation: University of the Western Cape
-  name: Ross Coppin, AJ Smit, Christo Rautenbach
-date: "2018"
+  name: Ross Coppin, A.J Smit and Christo Rautenbach
+date: "May 2018"
 output:
-  pdf_document:
-    toc: yes
-    toc_depth: 4
-  word_document:
-    fig_caption: yes
-    fig_height: 7
-    fig_width: 7
-    highlight: zenburn
-    toc: no
-bibliography: mycollection2.bib   
+  word_document: default
+  html_document:
+    df_print: paged
+  pdf_document: default
 ---
 
-\newpage
-## Background
+# Background
 
 Marine ecosystems are maintained by a variety of complex interactions between abiotic and
-biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology [@Doney2012].
+biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology.
 These abiotic and biotic variables, the interactions between them, and the various ecological processes, ultimately determine the community composition and ecological functioning of all ecosystems.
-Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and over-fishing [@Blamey2015].
+Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and over-fishing.
 Temperature and wave exposure have been recognised as important variables with regards to climate-driven changes within the ocean.
@@
 Seaweeds, particularly brown seaweeds, are able to undergo rapid morphological adaptation to the hydrodynamic environment.
-This allows seaweeds to reduce mortality through mechanical dislodgement by inducing morphological changes which reduces overall drag.
-Seaweeds that are unable to avoid mechanical dislodgement either raft out to sea or wash up onto beaches, otherwise known as beach-cast.
-However, not all the beach-cast may have originated from a nearby kelp population and may have originated from other sites or regions of the coast through a combination of rafting and ocean currents.
+This allows seaweeds to reduce mortality through mechanical dislodgement by inducing morphology which reduces overall drag.
+Seaweeds that are unable to avoid mechanical dislodgement either raft out to sea or wash up onto beaches.
+However, not all the beach-cast kelp may have originated from a nearby kelp population and may have originated from other sites or regions of the coast through rafting and ocean currents.
 Therefore, because kelp morphology is specific to its local environment the morphological features may be able to indicate, within a certain amount of probability, what site or region it most likely originated from.
@@
 This combined approach will allow investigation into flow regimes around the west and south-west coasts of South Africa and the role they play in sub-tidal and beach ecology.
-For instance, micro-plastics are recognised as a threat to marine life, however very little is known about how micro-plastics may be transported along the coast.The coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation.
+For instance, micro-plastics are recognised as a threat to marine life, however very little is known about how micro-plastics may be transported along the coast.
 
@@
 Therefore, the coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation.
-This in turn may inform management decisions with regards to eliminating and managing marine pollution in South Africa.
+This in turn may inform management descions with regards to elimating and managing marine pollution in South Africa.
 
-## Rationale
+# Aims and objectives
 
-Seaweeds, browns in particular, are capable of adapting morphological characteristics to persist in changing and variable ocean environments.
-Changes in morphology have been shown to increase photosynthetic ability, enhance nutrient uptake, and reduce drag with regards to changes in temperature and wave exposure.
-Wave exposure has been shown to be an important driver of seaweed morphology, as the main mechanism of seaweed mortality is through the mechanism of dislodgment.
-Changing morphology, such as decreasing frond area or strengthening through stipe circumference increase/reduces drag and increases the probability of survival.
-However, locally adapted seaweed may still be dislodged in pulse disturbance events such as storms, and may raft far distances and wash up on beaches.
-An example would be *Ecklonia maxima* which is a conspicuous brown seaweed along the coastline and beaches around South Africa.
-Therefore, the beach-cast may not always originate from adjacent kelp populations but rather from other regions which the individual is adapted for.
-Therefore, the kelp morphology may act as a proxy for investigating coastal currents and changes thereof.
-The advances in ocean hydrodynamic modelling has made great progress and has been applied in a variety of ways.
-Therefore, this study will use advances in hydrodynamic modelling in combination with kelp morphological characteristics to investigate coastal currents.
-Once the model has been established it may be applied in other ways, such as investigating the transport of micorplastics along the South African coastline.
-The harms of microplastics to the marine environment has gained much traction in recent years, but research in South Africa is lacking.
-The coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation.
-This in turn may inform management decisions with regards to eliminating and managing marine pollution in South Africa.
-
-## Research problem
-
-Pollution is a huge environmental problem that affects both terrestrial and marine ecosystems.
-Pollution from land enters the sea where it could harm or kill marine organisms or is transported by ocean currents to other coastal areas.
-In these areas the pollution could re-enter the ocean or be blown by wind into terrestrail ecoystems where it could once again be ingested by oragnsims causing harm or even death.
-Pollution is therefore a major threat to the environment and marine organisms.
-In recent years the affect of microplastics on the ocean has gained much traction with scientists and politictions alike.
-Microplastics cosiste of tiny particles of plastic and other pollution and are thereofore are difficult to detect.
-
-Currently there is no hydrodynamic model that is able to determine dispersion, source and accumulation of microplastics along the South African coastline.
-Given the significant deterimental affect microplastics play in the ocean, it is important that such a mechanism be developed that will aid in better management of marine pollution in South Africa.
-Furthermore, this project allows for a multidicisplinery approach to be taken by combining ecology and coastal oceanography.
-
-## Preliminary literature review
-
-Abiotic and biotic factors interact in complex ways which indirectly determine behavioral and ecophysiological responses in organisms.
-For example, when storms or strong currents form in sub-tidal habitats, sea urchins form aggregations in order to reduce overall drag to avoid being swept away by currents.
-Organisms that are motile my migrate into more environmentally suitable areas when conditions become unfavorable or food sources become depleted.
-In changing environments migration may also allow organisms to extend their distributions.
-For example, ocean warming off the coast of Western Australia has allowed tropical fish species to extend their distribution into areas that were previously characterised as temperate reefs.
-Sessile organisms are unable to migrate into more environmentally suitable areas and are forced to either adapt or suffer expiration.
-Sessile organisms may respond to changing environmental conditions through changes in physiology.
-For example, plants may produce heat shock proteins that help buffer the affect of temperature increases.
-Sessile organisms may also adapt their morphology,in order to persist in changing and harsh environments, such as seaweeds.
-
-Temperature and wave exposure have been shown to be important drivers of seaweed distribution, physiological functioning, ability to recover, population dynamics and morphology.
-Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10- 20 m s-1 with accelerations of 400 m s-2 [@Friedland1995] are the biggest threat to kelp survival.
-Kelps are able to rapidly adapt their morphological characteristics to reduce drag and avoid dislodgment [@Blanchette1997].
-For example a study by @Koehletal2008 showed that transplanted *Nereocystis luetkeana* plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days.
-Another study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and was a combination of *in situ* sampling and transplant of juvenile plants.
-The results showed that morphology differed between wave-sheltered and wave-exposed sites (thin thallus at sheltered sites and a narrow, thick thallus with a thick stipe at exposed sites), and was consistent with previous studies.
-Juveniles transplanted into wave exposed sites under went rapid morphological adaption, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaption.
-
-Kelp morphology may be distinct to a particular region with a specific hydrodynamic environment and has the ability to raft far distances using coastal currents, and may accumulate as beach-cast in areas far from its original location.
-Therefore, kelp morphology may be used as a proxy for determining its original location as well as aid in characterising coastal currents.
-However, this approach must be combined with advances in hydrodynamic modelling for a quantitative outcome.
-
-Advances in numerical modelling has gained much traction in recent years and has been applied in a variety of ways with regards to ecological studies.
-For example, a study by @wang2009 used the Delft3D-Flow model to assess the hydraulic suitability of a stream as a spawing ground for the Chinese Sturgeon (*Acipenser sinensis*) in the Yangtze River.
-The authors calculated the horizontal mean vorticity which was used to assess the hydraulic environment of spawning ground.
-The flow field state was determined through model simulation and field-measured data used to validate the model.
-The results added to existing scientific database for spawning ground hydraulic environmental protection.
-Different numerical models can often be integrated to model across ecosystem levels.
-For example a study by @leon2003 used integrated physical (Delft3D hydrological model) and bio-chemical (Agricultural Non-point Source model) processes models to investigate the possible impact on the Lake Malawi water quality due to management actions performed at the watershed level.
-
-Since wave energy is an important driver in marine ecosystems, particularly kelp, the advances in hydrodynamic modelling offer a new opportunity for multifactoral and quantitative approach to research in marine ecosystems.
-The Delf3D and SWAN models have been used successfully in previous studies regarding brine plume discharge, impacts of storms, affects of climate change on the hydrological environment etc.
-The models have not been designed for shallow environments (<6m) and therefore may not be suitable to model coastal hydrological environments.
-However these models may be adjusted to suit coastal waters if they are combined with a new numerical model which can be calibrated to suit these needs.
-
-In recent years there has been growing attention on microplastics and the deterimental role they play in the enviornment, particulary the ocean.
-Microplastics are tiny plastic granuales used as scrubbers in cosmetics and air-blasting, and small plastic fragments that originate from larger pieces of plastic known as macroplastics.
-The potential harms of small pieces of plastic was highlighted in the 1970's and renewed interest has lead to research showing that microplastics in the ocean are widespread.
-The small size of microplastics makes them bio-available to biota in the food-web which may cause problems with an organisms physiological functioning.
-Furthermore, the relatively large surface area and composition of microplastics provides an environmeht that is able of adhering to organic pollutants.
-In other words microplastics also act as a vector for transport and assimilation of organic polluants.
-Therefore, microplastics may also aid in the assimilation of organic pollutans into all levels of the food web.
-
-Therefore, this study not only enables research into the ecological affects of the hydrological environment on an important habitat-forming organism, it also offers the opportunity to improve on current hydrological numerical models to suit coastal environments.
-This in turn will allow investigation into the flow and accumulation of microplastics which are regarded as a major threat to marine life.
-Furthermore, the calibrated model could be applicable to other ecological studies such as dispersal of benthic flora and fauna, climate change studies, forecasting etc.
-
-## Aims of research
-
-The aim of the project is to investigate coastal flow regimes along the west coast and south-west coast of South Africa and the role this may play in transport of kelp beach-cast and microplastics.
+The aim of the project is to investigate coastal flow regimes along the west coast and south-west coast of South Africa and the role this may play in transport of microplastics.
 This aim will be met through the following objectives:
 
-A) Determine if the hydrodynamic environment is the main driver of kelp morphology and if this is specific to a location
-B) Simulate kelp rafting by means of a hydrodynamic modelling and calibrate this model with *in situ* beach-cast morphometric data.
-C) Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.
+1) Determine if the hydrodynamic environment is the main driver of kelp morphology and if this is specific to an area
+2) Simulate kelp rafting by means of a hydrodynamic modelling and calibrate this model with *in situ* morphometric data.
+3) Use calibrated hydrodynamic model to investigate dispersal of microplastics along the South African coastline.
 
-## Research design and methodology
+# Methods
+## Kelp morphology
 
-A combined quantitative ecology and oceanography approach will be used for this study.
-Morphological characteristics and their relation to the hydrodynamic environment must first be established and the findings will form the basis for the proceeding research.
-Once the relationship has been established it will be used to calibrate future hydrodynamic models developed during the course of the study.
-Once the hydrodynamic model has been established and calibrated it will be used to investigate the flow of microplastics along the South African coastline.
-
-# Morphology
-
-Morphometrics collection Between October 2014 and April 2015, morphological measurements of *Laminaria pallida* and *Ecklonia maxima* were collected at 19 sites along the Western Cape coast of South Africa (Refer to morphology bar graphs).
+Between October 2014 and April 2015, morphological measurements of *Laminaria pallida* and *Ecklonia maxima* were collected at 19 sites along the Western Cape coast of South Africa.
 These varying morphometrics allowed measurements such as weight, length and thickness to be compared between sites.
@@
 These sites span across the majority of the Western Cape coast, in varying thermal and wave energy regimes.
-Sites were chosen to represent an array of morphological differences seen within each *Ecklonia maxima* and *Laminaria*.
-Sites were also chosen to reflect locations that displayed variable wave and temperature regimes, to allow us to robustly test our hypothesis of environmental drivers influencing kelp morphology.
-St.
-Helena Bay and Betty’s Bay constituted the north western and south eastern boundary sites respectively.
-These sites are roughly 300km apart, and are confidently lie within separate marine provinces, as outlined above.
-Cape Peninsula provides an interesting topographical boundary that shelters the coast in False Bay.
-Sites were therefore chosen to represent an array of environments, from offshore reefs (Batsata Rock), to sheltered intertidal zones (Miller’s Point).
-West of Cape Point, a number of sites were chosen to highlight the presence of upwelling (Oudekraal, Kommetjie), as well as kelps growing in protected bays (Hout Bay).
+Sampling of shallower kelp individuals is currenlty taking place which will allow a deeper understanding of the hydrodynamic environment in relation to kelp morphology.
 
-# Hydrodynamic model
+## Hydrodynamic model
 
-The hydrodynamic model for coastal flow will be developed using already existing models that have been developed.
-These models will consist of the SWAN and Delf3D models and the apporach is to use aspects of each model, that are related to coastal flow, from which further development will take place.
-Once the model has been developed it will be used to predict kelp morphology at random sites which will then be sampled.
-The *in situ* kelp morphology data collected at the predicted site will be used to calibrate the model.
-This process will be repeated until the model can predict coastal flow and kelp morphology within a certian degree of probability.
+A hydrodynamic model will be developed for this study using already existing models (SWAN and Delf3d), and simulations for dispersal of kelp rafts will be performed.
+Morphometric measurements will be taken *in situ* after the simulation has been run in order to verify, and where neccessary, calibrate the model.
 
-# Microplastics
+## Dispersal of microplastisics
 
-The calibrated model developed in previous chapters will be used to determine flow, source and accumulatiom points along the South African coastline.
-Certian characteristcs of microplastics such as bouyancy, flow rate, fluid dynamics ect.
-Which characteristics are important for modeling dispersion of microplastics will be achieved by investigating the literature and  experimental simulations.
-Once the model has been established it will once again be calibrated through _in situ_ sampling.
+The calibrated model will be used to simulate the dispersal of microplastics around the coast as well as to determine possible areas of accumulation.
 
-## Time-frame
+## Progress to date
 
-## References
+The kelp morphometrics have already been sampled and a paper using this data is currently being written up.
+Morphometrics are currenlty being sampled for shallower kelp individuals (<1m), which will add to the current data set.
+A full proposal is also in development.

### 2018-08-01 chapter_1.Rmd
added updated rmd file

--- 2018-07-10 proposal_draft_2018.Rmd
@@
 ---
-title: "Kelp morphometric properties and the link to hydrodynamic modelling"
-author:
-- affiliation: University of the Western Cape
-  name: Ross Coppin, AJ Smit, Christo Rautenbach
-date: "2018"
+title: "Kelp morphometric properties and its link to the hydrodynamic modelling"
+author: "Ross Coppin"
+date: "26 February 2017"
 output:
+  word_document:
+    toc: yes
   pdf_document:
+    number_sections: yes
     toc: yes
-    toc_depth: 4
-  word_document:
-    fig_caption: yes
-    fig_height: 7
-    fig_width: 7
-    highlight: zenburn
-    toc: no
 bibliography: mycollection2.bib
@@
 
-\newpage
-## Background
+## Summary
 
-Marine ecosystems are maintained by a variety of complex interactions between abiotic and
-biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology [@Doney2012].
+Marine ecosystems are maintained by a variety of complex interactions between abiotic and biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology [@Doney2012].
 These abiotic and biotic variables, the interactions between them, and the various ecological processes, ultimately determine the community composition and ecological functioning of all ecosystems.
@@
 Therefore, the coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation.
-This in turn may inform management decisions with regards to eliminating and managing marine pollution in South Africa.
+This in turn may inform management descions with regards to elimating and managing marine pollution in South Africa.
 
-## Rationale
+## Background
 
@@
 
-## Research problem
-
 Pollution is a huge environmental problem that affects both terrestrial and marine ecosystems.
@@
 Microplastics cosiste of tiny particles of plastic and other pollution and are thereofore are difficult to detect.
-
 Currently there is no hydrodynamic model that is able to determine dispersion, source and accumulation of microplastics along the South African coastline.
@@
 Furthermore, this project allows for a multidicisplinery approach to be taken by combining ecology and coastal oceanography.
-
-## Preliminary literature review
 
@@
 
+#Kelp morphology drivers
+
+#Light
+
+Light is an important factor for kelp survival, however if light is limited or excessive this may negatively impact kelp survival or growth.
+Much of the past research into the role light plays into the functioning of kelp [@Bruhn1996);@Belseth2012].
+For instance, solar ultraviolate radiation has been shown to aﬀect sub-canopy Ecklonia radiata sporophytes when the canopy of mature *Ecklonia radiata* was removed [@Wood 1987].
+The sub-canopy sporophytes experienced tissue damage,photopigment destruction,reducedgrowthand decreased survivorship, thus inhibiting their settlement and survival [@Wood 1987].
+Laboratory experiments revealed that the UV component of radiation, rather than intense radiation itself, was responsible for the eﬀects mentioned above.
+High light stress has negative eﬀects, such as photoinhibition and photo-damage on *Ecklonia cava* sporophytes [@Altamirano2004].
+@Altamirano2004 found that *Ecklonia cava* is more vulnerable to light stress conditions, and less likely to recover under unfavourable conditions [@Altamirano2004].
+@Bolton1985 showed that under sub-saturating irradiances and supra- optimal temperatures *Ecklonia maxima* to showed a decrease in reproductive rates and an increase in cell production.
+A interesting ﬁnding of this study was that despite the decrease in reproductive rates, the ﬁnal egge production per female was greater under these conditons.
+The authors interpreted this an ecological adaption that may increase survival rates under times of stress or non - ideal conditions [@Bolton1985].
+Light is also important for survival and functioning of benthic algal species; a study by @Toohey2004 investiagted the possible link of light and thallus scour of the kelp *Ecklonia radiata canopy* on the associated foliose algal assemblage.
+The authors concluded that a reduction in light, rather than thallus scour has an eﬀect on foliose algal assemblage.
+Reduced light was associated with reduced diveristy and an marked increase abundance in certian foliose species, which were able to acclimiate more light under low irradiances compared to other species [@Toohey2004].
+Kelp may also compete intra-specﬁcally or inter-speciﬁcally for light.
+For instance, @Clark2004 investiagted eﬀect of canopy shading of *Macrocystis pyrifera* and understorey canopy shading of *Pterygophora californica* individually and in combination of benthic macro-algae assemblage.
+Under all treatments there was a sigiﬁance change in macro-algal assemblage, with certian species able to recover in abundance more readily than others.
+The authors devided these species into two categories; “shade-tolerant” and “light adapted”, and conlcuded that certian foliose species are better adapted for diﬀerent light conidtions in accordance with canopy cover.
+In South Africa, the eﬀect of *Ecklonia maxima* harvesting on sub-story kelp, kelp population structure and kelp recruitement was investigated by @Rothman2006.
+The results showed no signiﬁcant diﬀerence in the three factors mentioned earlier and @Rothman2006 concluded that factors other than light play a role in structuring kelp communities.
+
+#Nutrients
+
+Nutrients Theimportanceofnutrientsinthefunctioningofkelpsiswellunderstood [@Dayton1985; @Gaylord2012].
+Dissolved nitrogen, and in particular nitrate, are important; however research has also placed emphasis on phosphate and other trace compounds for functioning of kelps [@Dayton1985].
+Additionally, some kelps have the ability to store inorganic nitrogen in order to compensate for periods of low nutrient availability, which has been observed for Laminaria and Macrocystis [@Dayton1985; @Gaylord2012].
+Nutrient stratiﬁcation is also an important factor, particularly for canopy type kelps.
+The concentration of nutrients at the surface is important to the functioning and maintenance of the canopy.
+For instance kelp canopies in California often deteriorate in the summer months when surface nitrate levels are low [@Jackson1977].
+Water motion is important in the assimilation of nutrients from the water column, and kelps have been shown to adapt blade morphology in order to create more turbulence around the boundary layer of the frond to enhance nutrient assimilation (Wheeler 1980).
+Temperature has also been closely linked with nutrient concentrations.
+Nutrients are often in higher concentrations in the water column during low temperature events.
+This is often an indication of an “up-welling” event, which brings cold and nutrient rich waters from the bottom to the surface of the water column.
+However temperature can play a direct role in the uptake of nutrients through eﬀects on algal metabolism; however this may vary from species to species [@Raven1988].
+However most studies investigate multiple characteristics in one study (i.e.
+temperature and light and water motion), while studies investigating the direct role of temperature in nutrient assimilation is lacking in marine algae is lacking.
+
+#Depth
+
+Depth does not aﬀect kelp ecosystems directly, however a change in depth often causes ﬂuctuations or changes in other environmental variables such as water motion, light and temperature.
+Water motion also decreases with depth, and some kelps better suited to deeper environments (*L.
+pallida*) replace those in the shallows (*E.
+maxima*) (Dayton 1985; Gerard 1982).
+The increase in depth can lead to a decrease in sunlight penetration, with some species better adapted for low-light conditions than others, such as (*L.
+pallida*).
+Temperature may also change along a depth gradient due to a reduction in sunlight penetration.
+(Dayton 1985; Gerard 1982).
+Therefore depth does not directly play a role in kelp functioning but may alter more inﬂuential factors such as light and water motion.
+
+#Sedimentation
+
+Sedimentation and scour can be highly destructive to kelp forests with most detrimental impacts on the development stages such as spores, gametophytes and young plants (Dayton 1985), and has been linked to seasonal trends relating to wind direction [@Moore1972].
+@Devinny1978 found that sedimentation greatly reduced spore settlement and which in turn greatly reduces the probability of survival for Macrocystis Pyrifera gametophytes.
+Some species seem more resistant to the eﬀects of sedimentation such as Laminaria farlowii and Eisena aborea compared to others such as Macrocystis Pyrifera (Dayton 1985).
+The impacts of sedimentation is not limited to the sub-tidal but has also been shown to eﬀect inter-tidal kelp as well.
+For example a study by [@Schiel2006] investigated the eﬀects of sedimentation on the post-settlement stages of habitat-forming seaweeds.
+The results showed high mortality of germlings at sites that had high sedimentation, and was further supported through laboratory tests.
+Grazing was found to be negligible and mortality of gametophytes was attributed to sedimentation, although its eﬀects did vary with molluscan grazing [@Schiel2006].
+The presence of kelps themselves can also eﬀect the degree of sedimentation.
+For instance, @Wernberg2005 investigated the ability of Ecklonia radiata (canopy forming species) to modify the physical environment and the related eﬀects on associated foliose algae.
+An increase in kelp densities was marked by a signiﬁcant reduction of light and sedimentation, and found that sedimentation did not have a signiﬁcant eﬀect on associated foliose algae [@Wernberg2005].
+Therefore it is often diﬃcult to separate the direct eﬀects of each of the environmental variables as they are often related to one another.
+However, sea surface temperatures is one of the main predictors of marine biological systems, as it has been shown to relate to physiological and ecological processes at all levels [@Andersen2013; @Bolton1985; @Bolton2012; @Dayton1985].
+Kelp forests are no exception, with temperature being the main limiting factor [@Dayton 1985].
+
+# Temperature
+
+Temperature Biogeographical shifts of kelp species can already be seen in areas where ocean warming is occurring.
+In Western Australia an increase in temperature has caused a biogeographical shift in kelp and urchins, which in turn has caused many kelp forests to shift to the less productive algal turfs [@Kelly2005];a review of the literature shows a consensus that biogeographical distributions of marine species are largely temperature related [@Dayton1985; @Jeﬀree1994; @Schils2006; @Sorte2010; @Bolton2012].
+Other examples include Norway and eastern North America where ocean warming has facilitated kelp die-oﬀ and an increase in herbivore grazing rates [@Dayton1985].
+The increase in grazing rates and the colonization of other algal species into these areas make it almost impossible for kelps to recover [@Kelly2005].
+The majority of kelp species are artic and temperate organisms, and the warming of ocean temperatures is expected to cause a poleward biogeographical shift of species [@Bolton2012].
+There is evidence to suggest that South African kelp forests are expanding due to ocean cooling [@Bolton2012], possibly driven by an intensiﬁcation and increase in coastal upwelling [@Blamey2012, @Blamey2015].
+In South Africa there has been a biogeographical shift eastward along the coast due to a change in inshore temperature regime, making South Africa no exception to changing ocean temperatures [@Bolton2012].
+Macroalgae, such as kelps, can react to an increase in surface temperatures in one of three ways: they can migrate, adapt and die [@Biskup2014].
+A study by @Biskup2014 investigated the functional response of two kelp species (*Laminaria ochroleuca* and *Saccorhiza polyschides*) to rising sea temperatures.
+The functional responses of Saccorhiza polyschides was measured for both the subtidal and intertidal habitats, to see what aﬀect non- optimal conditions (intertidal zone) had on the kelps [@Rinde2005].
+The study found that Laminaria ochroleuca exhibited a poor ability to acclimatise and was dependent on the kelp’s life history traits [@Biskup2014].
+Therefore annual kelp species are more likely to survive under non-ideal condition, and the intertidal Saccorhiza polyschides, compared to the subtidal, showed a higher physiological ﬂexibility to changing conditions [@Biskup2014].
+This may be because the intertidal zone undergoes far more change than the subtidal and therefore kelps in the intertidal are forced to adapt to harsher conditions where ﬂuctuations in temperature, sunlight, turbidity and water motion are common.
+The eﬀects on temperature have also been investigated by @Wernberg2010.
+The study looked at resilience of kelp beds along a latitudinal temperature gradient.
+Kelp abundance is likely to decline with the predicted warming of ocean waters @Wernberg2010 and although kelps have the ability to acclimatize and adjust their metabolic performance, which in turn allows them to change their physiological performance to mitigate the seasonal ﬂuctuations in temperature, this acclimatization is done at a cost @Wernberg2010.
+It is therefore important to understand the ecological implications of this cost in order to better understand the future impacts of rising sea temperatures @Wernberg2010.
+Results from the study showed that kelp beds in the northern regions had a lower resilience to perturbations, such as physical and biological disturbance, disease, competition from invasive species and decreasing water quality.
+This was as a result of low abundance of recruits, reduced physiological responsiveness and reliance on mature adult kelp for net recruitment, growth of recruits and recovery of canopy loss @Wernberg2010.
+Even though @Wernberg2010 acknowledged that it was diﬃcult to exclude the inﬂuences of other environmental parameters (sunlight, nutrients and wave exposure), temperature was most likely the main environmental driver of these eﬀects.
+While kelps may be able to maintain populations in the warmer upper limits of their distributions, the cost of acclimation will make them more susceptible to other stressors such as disturbance from anthropogenic sources, disease, competition and decreasing water quality.
+Changes in temperature can also impact the distribution of kelps.
+
 ## Aims of research
@@
 C) Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.
-
-## Research design and methodology
-
-A combined quantitative ecology and oceanography approach will be used for this study.
-Morphological characteristics and their relation to the hydrodynamic environment must first be established and the findings will form the basis for the proceeding research.
-Once the relationship has been established it will be used to calibrate future hydrodynamic models developed during the course of the study.
-Once the hydrodynamic model has been established and calibrated it will be used to investigate the flow of microplastics along the South African coastline.
-
-# Morphology
-
-Morphometrics collection Between October 2014 and April 2015, morphological measurements of *Laminaria pallida* and *Ecklonia maxima* were collected at 19 sites along the Western Cape coast of South Africa (Refer to morphology bar graphs).
-These varying morphometrics allowed measurements such as weight, length and thickness to be compared between sites.
-Because the macroalgae differ in construction, species-specific morphometrics were included.
-These sites span across the majority of the Western Cape coast, in varying thermal and wave energy regimes.
-Sites were chosen to represent an array of morphological differences seen within each *Ecklonia maxima* and *Laminaria*.
-Sites were also chosen to reflect locations that displayed variable wave and temperature regimes, to allow us to robustly test our hypothesis of environmental drivers influencing kelp morphology.
-St.
-Helena Bay and Betty’s Bay constituted the north western and south eastern boundary sites respectively.
-These sites are roughly 300km apart, and are confidently lie within separate marine provinces, as outlined above.
-Cape Peninsula provides an interesting topographical boundary that shelters the coast in False Bay.
-Sites were therefore chosen to represent an array of environments, from offshore reefs (Batsata Rock), to sheltered intertidal zones (Miller’s Point).
-West of Cape Point, a number of sites were chosen to highlight the presence of upwelling (Oudekraal, Kommetjie), as well as kelps growing in protected bays (Hout Bay).
-
-# Hydrodynamic model
-
-The hydrodynamic model for coastal flow will be developed using already existing models that have been developed.
-These models will consist of the SWAN and Delf3D models and the apporach is to use aspects of each model, that are related to coastal flow, from which further development will take place.
-Once the model has been developed it will be used to predict kelp morphology at random sites which will then be sampled.
-The *in situ* kelp morphology data collected at the predicted site will be used to calibrate the model.
-This process will be repeated until the model can predict coastal flow and kelp morphology within a certian degree of probability.
-
-# Microplastics
-
-The calibrated model developed in previous chapters will be used to determine flow, source and accumulatiom points along the South African coastline.
-Certian characteristcs of microplastics such as bouyancy, flow rate, fluid dynamics ect.
-Which characteristics are important for modeling dispersion of microplastics will be achieved by investigating the literature and  experimental simulations.
-Once the model has been established it will once again be calibrated through _in situ_ sampling.
-
-## Time-frame
-
-## References

### 2018-08-01 chapter_1.Rmd
removed some sections. I am keeping the background literature with regards to biotic drivers and overall kelp ecology for now. I think it may be important for the reader to understand the complexity of temperate reefs.

@@
   pdf_document:
+    highlight: zenburn
     number_sections: yes
     toc: yes
-bibliography: mycollection2.bib
 ---
@@
 For instance, solar ultraviolate radiation has been shown to aﬀect sub-canopy Ecklonia radiata sporophytes when the canopy of mature *Ecklonia radiata* was removed [@Wood 1987].
-The sub-canopy sporophytes experienced tissue damage,photopigment destruction,reducedgrowthand decreased survivorship, thus inhibiting their settlement and survival [@Wood 1987].
+The sub-canopy sporophytes experienced tissue damage,photopigment destruction,reduced growthand decreased survivorship, thus inhibiting their settlement and survival [@Wood 1987].
 Laboratory experiments revealed that the UV component of radiation, rather than intense radiation itself, was responsible for the eﬀects mentioned above.
@@
 The authors interpreted this an ecological adaption that may increase survival rates under times of stress or non - ideal conditions [@Bolton1985].
-Light is also important for survival and functioning of benthic algal species; a study by @Toohey2004 investiagted the possible link of light and thallus scour of the kelp *Ecklonia radiata canopy* on the associated foliose algal assemblage.
-The authors concluded that a reduction in light, rather than thallus scour has an eﬀect on foliose algal assemblage.
-Reduced light was associated with reduced diveristy and an marked increase abundance in certian foliose species, which were able to acclimiate more light under low irradiances compared to other species [@Toohey2004].
-Kelp may also compete intra-specﬁcally or inter-speciﬁcally for light.
-For instance, @Clark2004 investiagted eﬀect of canopy shading of *Macrocystis pyrifera* and understorey canopy shading of *Pterygophora californica* individually and in combination of benthic macro-algae assemblage.
-Under all treatments there was a sigiﬁance change in macro-algal assemblage, with certian species able to recover in abundance more readily than others.
-The authors devided these species into two categories; “shade-tolerant” and “light adapted”, and conlcuded that certian foliose species are better adapted for diﬀerent light conidtions in accordance with canopy cover.
-In South Africa, the eﬀect of *Ecklonia maxima* harvesting on sub-story kelp, kelp population structure and kelp recruitement was investigated by @Rothman2006.
-The results showed no signiﬁcant diﬀerence in the three factors mentioned earlier and @Rothman2006 concluded that factors other than light play a role in structuring kelp communities.
 
@@
 
-Nutrients Theimportanceofnutrientsinthefunctioningofkelpsiswellunderstood [@Dayton1985; @Gaylord2012].
+The importance of nutrients in the functioning of kelps is well understood [@Dayton1985; @Gaylord2012].
 Dissolved nitrogen, and in particular nitrate, are important; however research has also placed emphasis on phosphate and other trace compounds for functioning of kelps [@Dayton1985].
@@
 Therefore depth does not directly play a role in kelp functioning but may alter more inﬂuential factors such as light and water motion.
-
-#Sedimentation
-
-Sedimentation and scour can be highly destructive to kelp forests with most detrimental impacts on the development stages such as spores, gametophytes and young plants (Dayton 1985), and has been linked to seasonal trends relating to wind direction [@Moore1972].
-@Devinny1978 found that sedimentation greatly reduced spore settlement and which in turn greatly reduces the probability of survival for Macrocystis Pyrifera gametophytes.
-Some species seem more resistant to the eﬀects of sedimentation such as Laminaria farlowii and Eisena aborea compared to others such as Macrocystis Pyrifera (Dayton 1985).
-The impacts of sedimentation is not limited to the sub-tidal but has also been shown to eﬀect inter-tidal kelp as well.
-For example a study by [@Schiel2006] investigated the eﬀects of sedimentation on the post-settlement stages of habitat-forming seaweeds.
-The results showed high mortality of germlings at sites that had high sedimentation, and was further supported through laboratory tests.
-Grazing was found to be negligible and mortality of gametophytes was attributed to sedimentation, although its eﬀects did vary with molluscan grazing [@Schiel2006].
-The presence of kelps themselves can also eﬀect the degree of sedimentation.
-For instance, @Wernberg2005 investigated the ability of Ecklonia radiata (canopy forming species) to modify the physical environment and the related eﬀects on associated foliose algae.
-An increase in kelp densities was marked by a signiﬁcant reduction of light and sedimentation, and found that sedimentation did not have a signiﬁcant eﬀect on associated foliose algae [@Wernberg2005].
-Therefore it is often diﬃcult to separate the direct eﬀects of each of the environmental variables as they are often related to one another.
-However, sea surface temperatures is one of the main predictors of marine biological systems, as it has been shown to relate to physiological and ecological processes at all levels [@Andersen2013; @Bolton1985; @Bolton2012; @Dayton1985].
-Kelp forests are no exception, with temperature being the main limiting factor [@Dayton 1985].
 
@@
 
+# Wave exposure
+
 ## Aims of research

### 2018-08-01 chapter_1.tex
removed some sections. I am keeping the background literature with regards to biotic drivers and overall kelp ecology for now. I think it may be important for the reader to understand the complexity of temperate reefs.

--- 2018-07-10 proposal_draft_2018.Rmd
@@
----
-title: "Kelp morphometric properties and the link to hydrodynamic modelling"
-author:
-- affiliation: University of the Western Cape
-  name: Ross Coppin, AJ Smit, Christo Rautenbach
-date: "2018"
-output:
-  pdf_document:
-    toc: yes
-    toc_depth: 4
-  word_document:
-    fig_caption: yes
-    fig_height: 7
-    fig_width: 7
-    highlight: zenburn
-    toc: no
-bibliography: mycollection2.bib   
----
+\documentclass[]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Kelp morphometric properties and its link to the hydrodynamic modelling},
+            pdfauthor={Ross Coppin},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-\newpage
-## Background
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
 
-Marine ecosystems are maintained by a variety of complex interactions between abiotic and
-biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology [@Doney2012].
-These abiotic and biotic variables, the interactions between them, and the various ecological processes, ultimately determine the community composition and ecological functioning of all ecosystems.
-Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and over-fishing [@Blamey2015].
-Temperature and wave exposure have been recognised as important variables with regards to climate-driven changes within the ocean.
-In order to persist and survive within variable and changing environments, organisms must either migrate, adapt, or die.
+%%% Change title format to be more compact
+\usepackage{titling}
 
-Seaweeds are sessile organisms which are unable to migrate to new areas when local environmental conditions become unsuitable and therefore are forced to adapt to new conditions in order to avoid expiration.
-The main form of mortality for seaweeds is through mechanical dislodgment by wave action.
-Seaweeds, particularly brown seaweeds, are able to undergo rapid morphological adaptation to the hydrodynamic environment.
-This allows seaweeds to reduce mortality through mechanical dislodgement by inducing morphological changes which reduces overall drag.
-Seaweeds that are unable to avoid mechanical dislodgement either raft out to sea or wash up onto beaches, otherwise known as beach-cast.
-However, not all the beach-cast may have originated from a nearby kelp population and may have originated from other sites or regions of the coast through a combination of rafting and ocean currents.
-Therefore, because kelp morphology is specific to its local environment the morphological features may be able to indicate, within a certain amount of probability, what site or region it most likely originated from.
-In other words, beach-cast kelp may be used as proxy for investigating the flow of  coastal currents.
-Using kelp as a proxy for determining its original location will be calibrated by means of a hydrodynamic model which will be designed from already existing SWAN and Delf3D models.
-This combined approach will allow investigation into flow regimes around the west and south-west coasts of South Africa and the role they play in sub-tidal and beach ecology.
-For instance, micro-plastics are recognised as a threat to marine life, however very little is known about how micro-plastics may be transported along the coast.The coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation.
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
 
-The increased use of plastic in society over the past half century has resulted in large amounts of plastic litter in both the marine and terrestrial environment.
-The problems associated with large plastic debris have received attention for many decades, whereas those connected to marine micro-plastics comparatively received very little attention.
-However, today it has become a prioritized area among political organizations, agencies and NGOs around the world.
-The micro-plastic debris present in the ocean are derived from marine and terrestrial sources, however there is little understanding of how microplastics may be distributed with ocean currents.
-Therefore, the coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation.
-This in turn may inform management decisions with regards to eliminating and managing marine pollution in South Africa.
+\setlength{\droptitle}{-2em}
 
-## Rationale
+  \title{Kelp morphometric properties and its link to the hydrodynamic modelling}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{Ross Coppin}
+    \preauthor{\centering\large\emph}
+  \postauthor{\par}
+      \predate{\centering\large\emph}
+  \postdate{\par}
+    \date{26 February 2017}
 
-Seaweeds, browns in particular, are capable of adapting morphological characteristics to persist in changing and variable ocean environments.
-Changes in morphology have been shown to increase photosynthetic ability, enhance nutrient uptake, and reduce drag with regards to changes in temperature and wave exposure.
-Wave exposure has been shown to be an important driver of seaweed morphology, as the main mechanism of seaweed mortality is through the mechanism of dislodgment.
-Changing morphology, such as decreasing frond area or strengthening through stipe circumference increase/reduces drag and increases the probability of survival.
-However, locally adapted seaweed may still be dislodged in pulse disturbance events such as storms, and may raft far distances and wash up on beaches.
-An example would be *Ecklonia maxima* which is a conspicuous brown seaweed along the coastline and beaches around South Africa.
-Therefore, the beach-cast may not always originate from adjacent kelp populations but rather from other regions which the individual is adapted for.
-Therefore, the kelp morphology may act as a proxy for investigating coastal currents and changes thereof.
-The advances in ocean hydrodynamic modelling has made great progress and has been applied in a variety of ways.
-Therefore, this study will use advances in hydrodynamic modelling in combination with kelp morphological characteristics to investigate coastal currents.
-Once the model has been established it may be applied in other ways, such as investigating the transport of micorplastics along the South African coastline.
-The harms of microplastics to the marine environment has gained much traction in recent years, but research in South Africa is lacking.
-The coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation.
-This in turn may inform management decisions with regards to eliminating and managing marine pollution in South Africa.
+\begin{document}
+\maketitle
 
-## Research problem
+{
+\setcounter{tocdepth}{2}
+\tableofcontents
+}
+\subsection{Summary}\label{summary}
 
-Pollution is a huge environmental problem that affects both terrestrial and marine ecosystems.
-Pollution from land enters the sea where it could harm or kill marine organisms or is transported by ocean currents to other coastal areas.
-In these areas the pollution could re-enter the ocean or be blown by wind into terrestrail ecoystems where it could once again be ingested by oragnsims causing harm or even death.
-Pollution is therefore a major threat to the environment and marine organisms.
-In recent years the affect of microplastics on the ocean has gained much traction with scientists and politictions alike.
-Microplastics cosiste of tiny particles of plastic and other pollution and are thereofore are difficult to detect.
+Marine ecosystems are maintained by a variety of complex interactions
+between abiotic and biotic variables such as temperature, wave exposure,
+pH, competition, and processes such as top-down and bottom-up control,
+predator-prey relationships and phenology {[}@Doney2012{]}. These
+abiotic and biotic variables, the interactions between them, and the
+various ecological processes, ultimately determine the community
+composition and ecological functioning of all ecosystems. Climate
+directly and indirectly affects the way in which abiotic and biotic
+variables interact, but is often compounded by other impacts such as
+habitat destruction, pollution, and over-fishing {[}@Blamey2015{]}.
+Temperature and wave exposure have been recognised as important
+variables with regards to climate-driven changes within the ocean. In
+order to persist and survive within variable and changing environments,
+organisms must either migrate, adapt, or die.
 
-Currently there is no hydrodynamic model that is able to determine dispersion, source and accumulation of microplastics along the South African coastline.
-Given the significant deterimental affect microplastics play in the ocean, it is important that such a mechanism be developed that will aid in better management of marine pollution in South Africa.
-Furthermore, this project allows for a multidicisplinery approach to be taken by combining ecology and coastal oceanography.
+Seaweeds are sessile organisms which are unable to migrate to new areas
+when local environmental conditions become unsuitable and therefore are
+forced to adapt to new conditions in order to avoid expiration. The main
+form of mortality for seaweeds is through mechanical dislodgment by wave
+action. Seaweeds, particularly brown seaweeds, are able to undergo rapid
+morphological adaptation to the hydrodynamic environment. This allows
+seaweeds to reduce mortality through mechanical dislodgement by inducing
+morphology which reduces overall drag. Seaweeds that are unable to avoid
+mechanical dislodgement either raft out to sea or wash up onto beaches.
+However, not all the beach-cast kelp may have originated from a nearby
+kelp population and may have originated from other sites or regions of
+the coast through rafting and ocean currents. Therefore, because kelp
+morphology is specific to its local environment the morphological
+features may be able to indicate, within a certain amount of
+probability, what site or region it most likely originated from. In
+other words, beach-cast kelp may be used as proxy for investigating the
+flow of coastal currents. Using kelp as a proxy for determining its
+original location will be calibrated by means of a hydrodynamic model
+which will be designed from already existing SWAN and Delf3D models.
+This combined approach will allow investigation into flow regimes around
+the west and south-west coasts of South Africa and the role they play in
+sub-tidal and beach ecology. For instance, micro-plastics are recognised
+as a threat to marine life, however very little is known about how
+micro-plastics may be transported along the coast.
 
-## Preliminary literature review
+The increased use of plastic in society over the past half century has
+resulted in large amounts of plastic litter in both the marine and
+terrestrial environment. The problems associated with large plastic
+debris have received attention for many decades, whereas those connected
+to marine micro-plastics comparatively received very little attention.
+However, today it has become a prioritized area among political
+organizations, agencies and NGOs around the world. The micro-plastic
+debris present in the ocean are derived from marine and terrestrial
+sources, however there is little understanding of how microplastics may
+be distributed with ocean currents. Therefore, the coupling of kelp
+morphology and the ability to simulate hydrodynamic processes can
+greatly improve our understanding of transport pathways and likely
+locations of accumulation. This in turn may inform management descions
+with regards to elimating and managing marine pollution in South Africa.
 
-Abiotic and biotic factors interact in complex ways which indirectly determine behavioral and ecophysiological responses in organisms.
-For example, when storms or strong currents form in sub-tidal habitats, sea urchins form aggregations in order to reduce overall drag to avoid being swept away by currents.
-Organisms that are motile my migrate into more environmentally suitable areas when conditions become unfavorable or food sources become depleted.
-In changing environments migration may also allow organisms to extend their distributions.
-For example, ocean warming off the coast of Western Australia has allowed tropical fish species to extend their distribution into areas that were previously characterised as temperate reefs.
-Sessile organisms are unable to migrate into more environmentally suitable areas and are forced to either adapt or suffer expiration.
-Sessile organisms may respond to changing environmental conditions through changes in physiology.
-For example, plants may produce heat shock proteins that help buffer the affect of temperature increases.
-Sessile organisms may also adapt their morphology,in order to persist in changing and harsh environments, such as seaweeds.
+\subsection{Background}\label{background}
 
-Temperature and wave exposure have been shown to be important drivers of seaweed distribution, physiological functioning, ability to recover, population dynamics and morphology.
-Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10- 20 m s-1 with accelerations of 400 m s-2 [@Friedland1995] are the biggest threat to kelp survival.
-Kelps are able to rapidly adapt their morphological characteristics to reduce drag and avoid dislodgment [@Blanchette1997].
-For example a study by @Koehletal2008 showed that transplanted *Nereocystis luetkeana* plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days.
-Another study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and was a combination of *in situ* sampling and transplant of juvenile plants.
-The results showed that morphology differed between wave-sheltered and wave-exposed sites (thin thallus at sheltered sites and a narrow, thick thallus with a thick stipe at exposed sites), and was consistent with previous studies.
-Juveniles transplanted into wave exposed sites under went rapid morphological adaption, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaption.
+Seaweeds, browns in particular, are capable of adapting morphological
+characteristics to persist in changing and variable ocean environments.
+Changes in morphology have been shown to increase photosynthetic
+ability, enhance nutrient uptake, and reduce drag with regards to
+changes in temperature and wave exposure. Wave exposure has been shown
+to be an important driver of seaweed morphology, as the main mechanism
+of seaweed mortality is through the mechanism of dislodgment. Changing
+morphology, such as decreasing frond area or strengthening through stipe
+circumference increase/reduces drag and increases the probability of
+survival. However, locally adapted seaweed may still be dislodged in
+pulse disturbance events such as storms, and may raft far distances and
+wash up on beaches. An example would be \emph{Ecklonia maxima} which is
+a conspicuous brown seaweed along the coastline and beaches around South
+Africa. Therefore, the beach-cast may not always originate from adjacent
+kelp populations but rather from other regions which the individual is
+adapted for. Therefore, the kelp morphology may act as a proxy for
+investigating coastal currents and changes thereof. The advances in
+ocean hydrodynamic modelling has made great progress and has been
+applied in a variety of ways. Therefore, this study will use advances in
+hydrodynamic modelling in combination with kelp morphological
+characteristics to investigate coastal currents. Once the model has been
+established it may be applied in other ways, such as investigating the
+transport of micorplastics along the South African coastline. The harms
+of microplastics to the marine environment has gained much traction in
+recent years, but research in South Africa is lacking. The coupling of
+kelp morphology and the ability to simulate hydrodynamic processes can
+greatly improve our understanding of transport pathways and likely
+locations of accumulation. This in turn may inform management decisions
+with regards to eliminating and managing marine pollution in South
+Africa.
 
-Kelp morphology may be distinct to a particular region with a specific hydrodynamic environment and has the ability to raft far distances using coastal currents, and may accumulate as beach-cast in areas far from its original location.
-Therefore, kelp morphology may be used as a proxy for determining its original location as well as aid in characterising coastal currents.
-However, this approach must be combined with advances in hydrodynamic modelling for a quantitative outcome.
+Pollution is a huge environmental problem that affects both terrestrial
+and marine ecosystems. Pollution from land enters the sea where it could
+harm or kill marine organisms or is transported by ocean currents to
+other coastal areas. In these areas the pollution could re-enter the
+ocean or be blown by wind into terrestrail ecoystems where it could once
+again be ingested by oragnsims causing harm or even death. Pollution is
+therefore a major threat to the environment and marine organisms. In
+recent years the affect of microplastics on the ocean has gained much
+traction with scientists and politictions alike. Microplastics cosiste
+of tiny particles of plastic and other pollution and are thereofore are
+difficult to detect. Currently there is no hydrodynamic model that is
+able to determine dispersion, source and accumulation of microplastics
+along the South African coastline. Given the significant deterimental
+affect microplastics play in the ocean, it is important that such a
+mechanism be developed that will aid in better management of marine
+pollution in South Africa. Furthermore, this project allows for a
+multidicisplinery approach to be taken by combining ecology and coastal
+oceanography.
 
-Advances in numerical modelling has gained much traction in recent years and has been applied in a variety of ways with regards to ecological studies.
-For example, a study by @wang2009 used the Delft3D-Flow model to assess the hydraulic suitability of a stream as a spawing ground for the Chinese Sturgeon (*Acipenser sinensis*) in the Yangtze River.
-The authors calculated the horizontal mean vorticity which was used to assess the hydraulic environment of spawning ground.
-The flow field state was determined through model simulation and field-measured data used to validate the model.
-The results added to existing scientific database for spawning ground hydraulic environmental protection.
-Different numerical models can often be integrated to model across ecosystem levels.
-For example a study by @leon2003 used integrated physical (Delft3D hydrological model) and bio-chemical (Agricultural Non-point Source model) processes models to investigate the possible impact on the Lake Malawi water quality due to management actions performed at the watershed level.
+Abiotic and biotic factors interact in complex ways which indirectly
+determine behavioral and ecophysiological responses in organisms. For
+example, when storms or strong currents form in sub-tidal habitats, sea
+urchins form aggregations in order to reduce overall drag to avoid being
+swept away by currents. Organisms that are motile my migrate into more
+environmentally suitable areas when conditions become unfavorable or
+food sources become depleted. In changing environments migration may
+also allow organisms to extend their distributions. For example, ocean
+warming off the coast of Western Australia has allowed tropical fish
+species to extend their distribution into areas that were previously
+characterised as temperate reefs. Sessile organisms are unable to
+migrate into more environmentally suitable areas and are forced to
+either adapt or suffer expiration. Sessile organisms may respond to
+changing environmental conditions through changes in physiology. For
+example, plants may produce heat shock proteins that help buffer the
+affect of temperature increases. Sessile organisms may also adapt their
+morphology,in order to persist in changing and harsh environments, such
+as seaweeds.
 
-Since wave energy is an important driver in marine ecosystems, particularly kelp, the advances in hydrodynamic modelling offer a new opportunity for multifactoral and quantitative approach to research in marine ecosystems.
-The Delf3D and SWAN models have been used successfully in previous studies regarding brine plume discharge, impacts of storms, affects of climate change on the hydrological environment etc.
-The models have not been designed for shallow environments (<6m) and therefore may not be suitable to model coastal hydrological environments.
-However these models may be adjusted to suit coastal waters if they are combined with a new numerical model which can be calibrated to suit these needs.
+Temperature and wave exposure have been shown to be important drivers of
+seaweed distribution, physiological functioning, ability to recover,
+population dynamics and morphology. Mechanical forces generated by the
+hydrodynamic environment, in the form of sudden strong ocean currents or
+storms, between 10- 20 m s-1 with accelerations of 400 m s-2
+{[}@Friedland1995{]} are the biggest threat to kelp survival. Kelps are
+able to rapidly adapt their morphological characteristics to reduce drag
+and avoid dislodgment {[}@Blanchette1997{]}. For example a study by
+@Koehletal2008 showed that transplanted \emph{Nereocystis luetkeana}
+plants from a wave sheltered site to a wave exposed site changed their
+morphology to flat blades and narrow laterals that are less prone to
+drag forces in 4-5 days. Another study by @FowlerWalker2006 tested for
+differences in morphology of \emph{Ecklonia radiata} between
+wave-sheltered and wave-exposed sites and was a combination of \emph{in
+situ} sampling and transplant of juvenile plants. The results showed
+that morphology differed between wave-sheltered and wave-exposed sites
+(thin thallus at sheltered sites and a narrow, thick thallus with a
+thick stipe at exposed sites), and was consistent with previous studies.
+Juveniles transplanted into wave exposed sites under went rapid
+morphological adaption, whilst the opposite was true for wave-sheltered
+sites which showed slower morphological adaption.
 
-In recent years there has been growing attention on microplastics and the deterimental role they play in the enviornment, particulary the ocean.
-Microplastics are tiny plastic granuales used as scrubbers in cosmetics and air-blasting, and small plastic fragments that originate from larger pieces of plastic known as macroplastics.
-The potential harms of small pieces of plastic was highlighted in the 1970's and renewed interest has lead to research showing that microplastics in the ocean are widespread.
-The small size of microplastics makes them bio-available to biota in the food-web which may cause problems with an organisms physiological functioning.
-Furthermore, the relatively large surface area and composition of microplastics provides an environmeht that is able of adhering to organic pollutants.
-In other words microplastics also act as a vector for transport and assimilation of organic polluants.
-Therefore, microplastics may also aid in the assimilation of organic pollutans into all levels of the food web.
+Kelp morphology may be distinct to a particular region with a specific
+hydrodynamic environment and has the ability to raft far distances using
+coastal currents, and may accumulate as beach-cast in areas far from its
+original location. Therefore, kelp morphology may be used as a proxy for
+determining its original location as well as aid in characterising
+coastal currents. However, this approach must be combined with advances
+in hydrodynamic modelling for a quantitative outcome.
 
-Therefore, this study not only enables research into the ecological affects of the hydrological environment on an important habitat-forming organism, it also offers the opportunity to improve on current hydrological numerical models to suit coastal environments.
-This in turn will allow investigation into the flow and accumulation of microplastics which are regarded as a major threat to marine life.
-Furthermore, the calibrated model could be applicable to other ecological studies such as dispersal of benthic flora and fauna, climate change studies, forecasting etc.
+Advances in numerical modelling has gained much traction in recent years
+and has been applied in a variety of ways with regards to ecological
+studies. For example, a study by @wang2009 used the Delft3D-Flow model
+to assess the hydraulic suitability of a stream as a spawing ground for
+the Chinese Sturgeon (\emph{Acipenser sinensis}) in the Yangtze River.
+The authors calculated the horizontal mean vorticity which was used to
+assess the hydraulic environment of spawning ground. The flow field
+state was determined through model simulation and field-measured data
+used to validate the model. The results added to existing scientific
+database for spawning ground hydraulic environmental protection.
+Different numerical models can often be integrated to model across
+ecosystem levels. For example a study by @leon2003 used integrated
+physical (Delft3D hydrological model) and bio-chemical (Agricultural
+Non-point Source model) processes models to investigate the possible
+impact on the Lake Malawi water quality due to management actions
+performed at the watershed level.
 
-## Aims of research
+Since wave energy is an important driver in marine ecosystems,
+particularly kelp, the advances in hydrodynamic modelling offer a new
+opportunity for multifactoral and quantitative approach to research in
+marine ecosystems. The Delf3D and SWAN models have been used
+successfully in previous studies regarding brine plume discharge,
+impacts of storms, affects of climate change on the hydrological
+environment etc. The models have not been designed for shallow
+environments (\textless{}6m) and therefore may not be suitable to model
+coastal hydrological environments. However these models may be adjusted
+to suit coastal waters if they are combined with a new numerical model
+which can be calibrated to suit these needs.
 
-The aim of the project is to investigate coastal flow regimes along the west coast and south-west coast of South Africa and the role this may play in transport of kelp beach-cast and microplastics.
-This aim will be met through the following objectives:
+In recent years there has been growing attention on microplastics and
+the deterimental role they play in the enviornment, particulary the
+ocean. Microplastics are tiny plastic granuales used as scrubbers in
+cosmetics and air-blasting, and small plastic fragments that originate
+from larger pieces of plastic known as macroplastics. The potential
+harms of small pieces of plastic was highlighted in the 1970's and
+renewed interest has lead to research showing that microplastics in the
+ocean are widespread. The small size of microplastics makes them
+bio-available to biota in the food-web which may cause problems with an
+organisms physiological functioning. Furthermore, the relatively large
+surface area and composition of microplastics provides an environmeht
+that is able of adhering to organic pollutants. In other words
+microplastics also act as a vector for transport and assimilation of
+organic polluants. Therefore, microplastics may also aid in the
+assimilation of organic pollutans into all levels of the food web.
 
-A) Determine if the hydrodynamic environment is the main driver of kelp morphology and if this is specific to a location
-B) Simulate kelp rafting by means of a hydrodynamic modelling and calibrate this model with *in situ* beach-cast morphometric data.
-C) Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.
+Therefore, this study not only enables research into the ecological
+affects of the hydrological environment on an important habitat-forming
+organism, it also offers the opportunity to improve on current
+hydrological numerical models to suit coastal environments. This in turn
+will allow investigation into the flow and accumulation of microplastics
+which are regarded as a major threat to marine life. Furthermore, the
+calibrated model could be applicable to other ecological studies such as
+dispersal of benthic flora and fauna, climate change studies,
+forecasting etc.
 
-## Research design and methodology
+\section{Kelp morphology drivers}\label{kelp-morphology-drivers}
 
-A combined quantitative ecology and oceanography approach will be used for this study.
-Morphological characteristics and their relation to the hydrodynamic environment must first be established and the findings will form the basis for the proceeding research.
-Once the relationship has been established it will be used to calibrate future hydrodynamic models developed during the course of the study.
-Once the hydrodynamic model has been established and calibrated it will be used to investigate the flow of microplastics along the South African coastline.
+\section{Light}\label{light}
 
-# Morphology
+Light is an important factor for kelp survival, however if light is
+limited or excessive this may negatively impact kelp survival or growth.
+Much of the past research into the role light plays into the functioning
+of kelp {[}@Bruhn1996);@Belseth2012{]}. For instance, solar ultraviolate
+radiation has been shown to aﬀect sub-canopy Ecklonia radiata
+sporophytes when the canopy of mature \emph{Ecklonia radiata} was
+removed {[}@Wood 1987{]}. The sub-canopy sporophytes experienced tissue
+damage,photopigment destruction,reducedgrowthand decreased survivorship,
+thus inhibiting their settlement and survival {[}@Wood 1987{]}.
+Laboratory experiments revealed that the UV component of radiation,
+rather than intense radiation itself, was responsible for the eﬀects
+mentioned above. High light stress has negative eﬀects, such as
+photoinhibition and photo-damage on \emph{Ecklonia cava} sporophytes
+{[}@Altamirano2004{]}. @Altamirano2004 found that \emph{Ecklonia cava}
+is more vulnerable to light stress conditions, and less likely to
+recover under unfavourable conditions {[}@Altamirano2004{]}. @Bolton1985
+showed that under sub-saturating irradiances and supra- optimal
+temperatures \emph{Ecklonia maxima} to showed a decrease in reproductive
+rates and an increase in cell production. A interesting ﬁnding of this
+study was that despite the decrease in reproductive rates, the ﬁnal egge
+production per female was greater under these conditons. The authors
+interpreted this an ecological adaption that may increase survival rates
+under times of stress or non - ideal conditions {[}@Bolton1985{]}. Light
+is also important for survival and functioning of benthic algal species;
+a study by @Toohey2004 investiagted the possible link of light and
+thallus scour of the kelp \emph{Ecklonia radiata canopy} on the
+associated foliose algal assemblage. The authors concluded that a
+reduction in light, rather than thallus scour has an eﬀect on foliose
+algal assemblage. Reduced light was associated with reduced diveristy
+and an marked increase abundance in certian foliose species, which were
+able to acclimiate more light under low irradiances compared to other
+species {[}@Toohey2004{]}. Kelp may also compete intra-specﬁcally or
+inter-speciﬁcally for light. For instance, @Clark2004 investiagted eﬀect
+of canopy shading of \emph{Macrocystis pyrifera} and understorey canopy
+shading of \emph{Pterygophora californica} individually and in
+combination of benthic macro-algae assemblage. Under all treatments
+there was a sigiﬁance change in macro-algal assemblage, with certian
+species able to recover in abundance more readily than others. The
+authors devided these species into two categories; ``shade-tolerant''
+and ``light adapted'', and conlcuded that certian foliose species are
+better adapted for diﬀerent light conidtions in accordance with canopy
+cover. In South Africa, the eﬀect of \emph{Ecklonia maxima} harvesting
+on sub-story kelp, kelp population structure and kelp recruitement was
+investigated by @Rothman2006. The results showed no signiﬁcant diﬀerence
+in the three factors mentioned earlier and @Rothman2006 concluded that
+factors other than light play a role in structuring kelp communities.
 
-Morphometrics collection Between October 2014 and April 2015, morphological measurements of *Laminaria pallida* and *Ecklonia maxima* were collected at 19 sites along the Western Cape coast of South Africa (Refer to morphology bar graphs).
-These varying morphometrics allowed measurements such as weight, length and thickness to be compared between sites.
-Because the macroalgae differ in construction, species-specific morphometrics were included.
-These sites span across the majority of the Western Cape coast, in varying thermal and wave energy regimes.
-Sites were chosen to represent an array of morphological differences seen within each *Ecklonia maxima* and *Laminaria*.
-Sites were also chosen to reflect locations that displayed variable wave and temperature regimes, to allow us to robustly test our hypothesis of environmental drivers influencing kelp morphology.
-St.
-Helena Bay and Betty’s Bay constituted the north western and south eastern boundary sites respectively.
-These sites are roughly 300km apart, and are confidently lie within separate marine provinces, as outlined above.
-Cape Peninsula provides an interesting topographical boundary that shelters the coast in False Bay.
-Sites were therefore chosen to represent an array of environments, from offshore reefs (Batsata Rock), to sheltered intertidal zones (Miller’s Point).
-West of Cape Point, a number of sites were chosen to highlight the presence of upwelling (Oudekraal, Kommetjie), as well as kelps growing in protected bays (Hout Bay).
+\section{Nutrients}\label{nutrients}
 
-# Hydrodynamic model
+Nutrients
+Theimportanceofnutrientsinthefunctioningofkelpsiswellunderstood
+{[}@Dayton1985; @Gaylord2012{]}. Dissolved nitrogen, and in particular
+nitrate, are important; however research has also placed emphasis on
+phosphate and other trace compounds for functioning of kelps
+{[}@Dayton1985{]}. Additionally, some kelps have the ability to store
+inorganic nitrogen in order to compensate for periods of low nutrient
+availability, which has been observed for Laminaria and Macrocystis
+{[}@Dayton1985; @Gaylord2012{]}. Nutrient stratiﬁcation is also an
+important factor, particularly for canopy type kelps. The concentration
+of nutrients at the surface is important to the functioning and
+maintenance of the canopy. For instance kelp canopies in California
+often deteriorate in the summer months when surface nitrate levels are
+low {[}@Jackson1977{]}. Water motion is important in the assimilation of
+nutrients from the water column, and kelps have been shown to adapt
+blade morphology in order to create more turbulence around the boundary
+layer of the frond to enhance nutrient assimilation (Wheeler 1980).
+Temperature has also been closely linked with nutrient concentrations.
+Nutrients are often in higher concentrations in the water column during
+low temperature events. This is often an indication of an ``up-welling''
+event, which brings cold and nutrient rich waters from the bottom to the
+surface of the water column. However temperature can play a direct role
+in the uptake of nutrients through eﬀects on algal metabolism; however
+this may vary from species to species {[}@Raven1988{]}. However most
+studies investigate multiple characteristics in one study
+(i.e.~temperature and light and water motion), while studies
+investigating the direct role of temperature in nutrient assimilation is
+lacking in marine algae is lacking.
 
-The hydrodynamic model for coastal flow will be developed using already existing models that have been developed.
-These models will consist of the SWAN and Delf3D models and the apporach is to use aspects of each model, that are related to coastal flow, from which further development will take place.
-Once the model has been developed it will be used to predict kelp morphology at random sites which will then be sampled.
-The *in situ* kelp morphology data collected at the predicted site will be used to calibrate the model.
-This process will be repeated until the model can predict coastal flow and kelp morphology within a certian degree of probability.
+\section{Depth}\label{depth}
 
-# Microplastics
+Depth does not aﬀect kelp ecosystems directly, however a change in depth
+often causes ﬂuctuations or changes in other environmental variables
+such as water motion, light and temperature. Water motion also decreases
+with depth, and some kelps better suited to deeper environments
+(\emph{L. pallida}) replace those in the shallows (\emph{E. maxima})
+(Dayton 1985; Gerard 1982). The increase in depth can lead to a decrease
+in sunlight penetration, with some species better adapted for low-light
+conditions than others, such as (\emph{L. pallida}). Temperature may
+also change along a depth gradient due to a reduction in sunlight
+penetration. (Dayton 1985; Gerard 1982). Therefore depth does not
+directly play a role in kelp functioning but may alter more inﬂuential
+factors such as light and water motion.
 
-The calibrated model developed in previous chapters will be used to determine flow, source and accumulatiom points along the South African coastline.
-Certian characteristcs of microplastics such as bouyancy, flow rate, fluid dynamics ect.
-Which characteristics are important for modeling dispersion of microplastics will be achieved by investigating the literature and  experimental simulations.
-Once the model has been established it will once again be calibrated through _in situ_ sampling.
+\section{Sedimentation}\label{sedimentation}
 
-## Time-frame
+Sedimentation and scour can be highly destructive to kelp forests with
+most detrimental impacts on the development stages such as spores,
+gametophytes and young plants (Dayton 1985), and has been linked to
+seasonal trends relating to wind direction {[}@Moore1972{]}.
+@Devinny1978 found that sedimentation greatly reduced spore settlement
+and which in turn greatly reduces the probability of survival for
+Macrocystis Pyrifera gametophytes. Some species seem more resistant to
+the eﬀects of sedimentation such as Laminaria farlowii and Eisena aborea
+compared to others such as Macrocystis Pyrifera (Dayton 1985). The
+impacts of sedimentation is not limited to the sub-tidal but has also
+been shown to eﬀect inter-tidal kelp as well. For example a study by
+{[}@Schiel2006{]} investigated the eﬀects of sedimentation on the
+post-settlement stages of habitat-forming seaweeds. The results showed
+high mortality of germlings at sites that had high sedimentation, and
+was further supported through laboratory tests. Grazing was found to be
+negligible and mortality of gametophytes was attributed to
+sedimentation, although its eﬀects did vary with molluscan grazing
+{[}@Schiel2006{]}. The presence of kelps themselves can also eﬀect the
+degree of sedimentation. For instance, @Wernberg2005 investigated the
+ability of Ecklonia radiata (canopy forming species) to modify the
+physical environment and the related eﬀects on associated foliose algae.
+An increase in kelp densities was marked by a signiﬁcant reduction of
+light and sedimentation, and found that sedimentation did not have a
+signiﬁcant eﬀect on associated foliose algae {[}@Wernberg2005{]}.
+Therefore it is often diﬃcult to separate the direct eﬀects of each of
+the environmental variables as they are often related to one another.
+However, sea surface temperatures is one of the main predictors of
+marine biological systems, as it has been shown to relate to
+physiological and ecological processes at all levels {[}@Andersen2013;
+@Bolton1985; @Bolton2012; @Dayton1985{]}. Kelp forests are no exception,
+with temperature being the main limiting factor {[}@Dayton 1985{]}.
 
-## References
+\section{Temperature}\label{temperature}
+
+Temperature Biogeographical shifts of kelp species can already be seen
+in areas where ocean warming is occurring. In Western Australia an
+increase in temperature has caused a biogeographical shift in kelp and
+urchins, which in turn has caused many kelp forests to shift to the less
+productive algal turfs {[}@Kelly2005{]};a review of the literature shows
+a consensus that biogeographical distributions of marine species are
+largely temperature related {[}@Dayton1985; @Jeﬀree1994; @Schils2006;
+@Sorte2010; @Bolton2012{]}. Other examples include Norway and eastern
+North America where ocean warming has facilitated kelp die-oﬀ and an
+increase in herbivore grazing rates {[}@Dayton1985{]}. The increase in
+grazing rates and the colonization of other algal species into these
+areas make it almost impossible for kelps to recover {[}@Kelly2005{]}.
+The majority of kelp species are artic and temperate organisms, and the
+warming of ocean temperatures is expected to cause a poleward
+biogeographical shift of species {[}@Bolton2012{]}. There is evidence to
+suggest that South African kelp forests are expanding due to ocean
+cooling {[}@Bolton2012{]}, possibly driven by an intensiﬁcation and
+increase in coastal upwelling {[}@Blamey2012, @Blamey2015{]}. In South
+Africa there has been a biogeographical shift eastward along the coast
+due to a change in inshore temperature regime, making South Africa no
+exception to changing ocean temperatures {[}@Bolton2012{]}. Macroalgae,
+such as kelps, can react to an increase in surface temperatures in one
+of three ways: they can migrate, adapt and die {[}@Biskup2014{]}. A
+study by @Biskup2014 investigated the functional response of two kelp
+species (\emph{Laminaria ochroleuca} and \emph{Saccorhiza polyschides})
+to rising sea temperatures. The functional responses of Saccorhiza
+polyschides was measured for both the subtidal and intertidal habitats,
+to see what aﬀect non- optimal conditions (intertidal zone) had on the
+kelps {[}@Rinde2005{]}. The study found that Laminaria ochroleuca
+exhibited a poor ability to acclimatise and was dependent on the kelp's
+life history traits {[}@Biskup2014{]}. Therefore annual kelp species are
+more likely to survive under non-ideal condition, and the intertidal
+Saccorhiza polyschides, compared to the subtidal, showed a higher
+physiological ﬂexibility to changing conditions {[}@Biskup2014{]}. This
+may be because the intertidal zone undergoes far more change than the
+subtidal and therefore kelps in the intertidal are forced to adapt to
+harsher conditions where ﬂuctuations in temperature, sunlight, turbidity
+and water motion are common. The eﬀects on temperature have also been
+investigated by @Wernberg2010. The study looked at resilience of kelp
+beds along a latitudinal temperature gradient. Kelp abundance is likely
+to decline with the predicted warming of ocean waters @Wernberg2010 and
+although kelps have the ability to acclimatize and adjust their
+metabolic performance, which in turn allows them to change their
+physiological performance to mitigate the seasonal ﬂuctuations in
+temperature, this acclimatization is done at a cost @Wernberg2010. It is
+therefore important to understand the ecological implications of this
+cost in order to better understand the future impacts of rising sea
+temperatures @Wernberg2010. Results from the study showed that kelp beds
+in the northern regions had a lower resilience to perturbations, such as
+physical and biological disturbance, disease, competition from invasive
+species and decreasing water quality. This was as a result of low
+abundance of recruits, reduced physiological responsiveness and reliance
+on mature adult kelp for net recruitment, growth of recruits and
+recovery of canopy loss @Wernberg2010. Even though @Wernberg2010
+acknowledged that it was diﬃcult to exclude the inﬂuences of other
+environmental parameters (sunlight, nutrients and wave exposure),
+temperature was most likely the main environmental driver of these
+eﬀects. While kelps may be able to maintain populations in the warmer
+upper limits of their distributions, the cost of acclimation will make
+them more susceptible to other stressors such as disturbance from
+anthropogenic sources, disease, competition and decreasing water
+quality. Changes in temperature can also impact the distribution of
+kelps.
+
+\subsection{Aims of research}\label{aims-of-research}
+
+The aim of the project is to investigate coastal flow regimes along the
+west coast and south-west coast of South Africa and the role this may
+play in transport of kelp beach-cast and microplastics. This aim will be
+met through the following objectives:
+
+\begin{enumerate}
+\def\labelenumi{\Alph{enumi})}
+\tightlist
+\item
+  Determine if the hydrodynamic environment is the main driver of kelp
+  morphology and if this is specific to a location
+\item
+  Simulate kelp rafting by means of a hydrodynamic modelling and
+  calibrate this model with \emph{in situ} beach-cast morphometric data.
+\item
+  Use the calibrated hydrodynamic model to investigate dispersal of
+  microplastics along the west coast and south-west coast of South
+  Africa.
+\end{enumerate}
+
+\end{document}

### 2018-08-28 chapter_1.Rmd
updated bib. file, added hyrodynamic modelling section

@@
 Changes in morphology have been shown to increase photosynthetic ability, enhance nutrient uptake, and reduce drag with regards to changes in temperature and wave exposure.
-Wave exposure has been shown to be an important driver of seaweed morphology, as the main mechanism of seaweed mortality is through the mechanism of dislodgment.
-Changing morphology, such as decreasing frond area or strengthening through stipe circumference increase/reduces drag and increases the probability of survival.
+Wave exposure has been shown to be an important driver of seaweed morphology, as the main mechanism of seaweed mortality is through the dislodgment.
+Changing morphology reduces drag and increases the probability of survival.
 However, locally adapted seaweed may still be dislodged in pulse disturbance events such as storms, and may raft far distances and wash up on beaches.
@@
 
+Other than temperature, wave exposure is also recognised as an important driver of the marine environment.
+Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition and productivity of benthic and rocky shore communties.
+For example, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake through increased turbulence in the boundry layer [@lobban1994].
+@lewis found that degrees of wave exposure vertical structure rocky shore communities...**work in progress**...
+
+# Hydrodynamic modelling
+
+Traditional ecological measures of wave exposure usually incorportates integrative measures of hyrdodynamic conditions at a particular site.
+More specifically, it is the integration of mechanical processes and the influence that is has on the ecology of nearshore communities.
+
+Wave exposure may be modelled through various methods which range from simple cartographic to more advanced numerical wave models.
+Cartographical models can be qualitative or quantitative and were designed for the need of wave exposure measures to explain ecological distributions.
+A simple set of calculations on coastline and wind data, and relatively small input data sets are required.
+These are regarded as "fetch-based models", which measure the length of open water associated with a site along a straight line.
+The output of such an approach is a simplified estimate of the potential wave energy for a specific set of sites.
+Advances in cartographical methods using fetch-based models has allowed for wave exposure measurments for larger areas, and has been suggested as a method for predicting macroalgal community structure [@burrows2008].
+The study...**insert detailed description here**...
+
+@lindegarth2005 critized this approach, arguing that the choice of wave exposure method  can influence ecological inference.
+The authors also highlighted the need for objective, reproducible and quantitative studies comparing exposure indices [@lindegarth2005].
+A study by @sundbald...**include description of study here**...Other authors, such as @hill2010, have argued that these simple measures can be improved upon by including bathymetery data which allows the incorporation of diffraction into the calculation.
+Diffraction is topographically induced variations in wave direction.
+A model incorporating this complexity was developed by @isaeus2004, and is known as the "simplified wave model" (SWM).
+The model uses measurements of wind strength, fetch and empirically derived algorithims to mimic diffraction.
+
+Numerical models are able to incorporate various physical process, as well as their complexity of interactions.
+
 ## Aims of research

### 2018-08-28 proposal_draft_2018.Rmd
updated bib. file, added hyrodynamic modelling section

@@
 output:
-  pdf_document:
-    toc: yes
-    toc_depth: 4
   word_document:
@@
     toc: no
-bibliography: mycollection2.bib   
+  pdf_document:
+    toc: yes
+    toc_depth: 4
+bibliography: mycollection.bib   
 ---
@@
 These abiotic and biotic variables, the interactions between them, and the various ecological processes, ultimately determine the community composition and ecological functioning of all ecosystems.
-Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and over-fishing [@Blamey2015].
+Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and over-fishing [@Blamey2015a].
 Temperature and wave exposure have been recognised as important variables with regards to climate-driven changes within the ocean.

### 2018-08-28 chapter_1.Rmd
edits

@@
 
-#Kelp morphology drivers
+#Abiotic drivers
 
@@
 
-Temperature Biogeographical shifts of kelp species can already be seen in areas where ocean warming is occurring.
-In Western Australia an increase in temperature has caused a biogeographical shift in kelp and urchins, which in turn has caused many kelp forests to shift to the less productive algal turfs [@Kelly2005];a review of the literature shows a consensus that biogeographical distributions of marine species are largely temperature related [@Dayton1985; @Jeﬀree1994; @Schils2006; @Sorte2010; @Bolton2012].
-Other examples include Norway and eastern North America where ocean warming has facilitated kelp die-oﬀ and an increase in herbivore grazing rates [@Dayton1985].
-The increase in grazing rates and the colonization of other algal species into these areas make it almost impossible for kelps to recover [@Kelly2005].
-The majority of kelp species are artic and temperate organisms, and the warming of ocean temperatures is expected to cause a poleward biogeographical shift of species [@Bolton2012].
+Temperature is a driver of kelp species distributions and ecophysiological processes, as well as a lesser role in morphological adaption...**example here**...The majority of kelp species are artic and temperate organisms, and the warming of ocean temperatures is expected to cause a poleward biogeographical shift of species [@Bolton2012].
 There is evidence to suggest that South African kelp forests are expanding due to ocean cooling [@Bolton2012], possibly driven by an intensiﬁcation and increase in coastal upwelling [@Blamey2012, @Blamey2015].
@@
 The study looked at resilience of kelp beds along a latitudinal temperature gradient.
-Kelp abundance is likely to decline with the predicted warming of ocean waters @Wernberg2010 and although kelps have the ability to acclimatize and adjust their metabolic performance, which in turn allows them to change their physiological performance to mitigate the seasonal ﬂuctuations in temperature, this acclimatization is done at a cost @Wernberg2010.
-It is therefore important to understand the ecological implications of this cost in order to better understand the future impacts of rising sea temperatures @Wernberg2010.
-Results from the study showed that kelp beds in the northern regions had a lower resilience to perturbations, such as physical and biological disturbance, disease, competition from invasive species and decreasing water quality.
-This was as a result of low abundance of recruits, reduced physiological responsiveness and reliance on mature adult kelp for net recruitment, growth of recruits and recovery of canopy loss @Wernberg2010.
-Even though @Wernberg2010 acknowledged that it was diﬃcult to exclude the inﬂuences of other environmental parameters (sunlight, nutrients and wave exposure), temperature was most likely the main environmental driver of these eﬀects.
-While kelps may be able to maintain populations in the warmer upper limits of their distributions, the cost of acclimation will make them more susceptible to other stressors such as disturbance from anthropogenic sources, disease, competition and decreasing water quality.
-Changes in temperature can also impact the distribution of kelps.
+Kelp abundance is likely to decline with the predicted warming of ocean waters @Wernberg2010 and although kelps have the ability to acclimatize and adjust their metabolic performance, which in turn allows them to change their physiological performance to mitigate the seasonal ﬂuctuations in temperature, this acclimatization is done at a cost @Wernberg2010...**link to paragrapgh on kelp morphology**...
 
@@
 Advances in cartographical methods using fetch-based models has allowed for wave exposure measurments for larger areas, and has been suggested as a method for predicting macroalgal community structure [@burrows2008].
-The study...**insert detailed description here**...
+An example of such a model is the "BioEx model" which was developed by @baardseth1970 to estimate wave exposure over large regions.
+BioEx requires frequency, strength and direction of winds, weighted by degree of exposure within various directions.
+BioEx is calculated as the sum of the index developed at different spatial scales (local, fjord and open).
+This method has been used in mapping of marine coastal biodiveristy [@Rinde2004].
 
@@
 
-Numerical models are able to incorporate various physical process, as well as their complexity of interactions.
+Advances in numerical modelling have been founded on physical wave theory on how a wave "behaves".
+This approach is based on a theoretical persepctive rather than the need to answer ecological questions.
+Besides diffraction, these models incorporate more complexity by including wind forcing, wave-to-wave interactions and loss of energy due to friction and wave breaking.
+Numerical models have a variety of applications and are often incorporated within hydrodynamic general circulation models and are used operationally for forecasting the sea state [@hasselmann1988; @booij1999; @smith2001].
+The downside of advanced numerical models is that are computationally intensive which creates limitations for large scale simulations.
+Therefore, their application along long stretches of variable coastline, inshore environments and ocean-wide simulations limited due to the poor spatial coverage.
+However, this models can be designed for local or site specific coverage, provided the correct data is available.
 

### 2018-08-28 notes/paper_summaries.Rmd
moved some files around

---
title: "Paper summaries"
output: html_notebook
---

# @Cavanaugh2011

In southern California, growth is fastest in winter and spring when nutrient levels are high and competition for light and space is low (due to low algal biomass), and slowest during summer when nutrients are low and competition for light and space is high due to well-developed algal canopies (Zimmer- man & Kremer 1986, Reed et al. 2008). The relatively low capacity of giant kelp to store nutrients (~30 d, Zimmerman & Kremer 1986) causes populations to respond rapidly to fluctuations in nutrient supply. Much like growth, the recruitment of giant kelp in southern California and elsewhere responds strongly to fluctuations in nutrients and light as determined by biotic and abiotic processes (Dayton et al. 1984, Reed & Foster 1984, Reed et al. 2008)

Giant-kelp mortality occurs in the form of senes- cence, grazing, and wave-driven disturbance (Graham et al. 2007). Reed et al. (2008) found that both frond losses and plant mortality were correlated to wave heights in kelp forests near Santa Barbara, California, although correlations were stronger at the plant level. High mortality can also result from prolonged con - ditions of low nutrients such as those associated with El Niño events (Dayton & Tegner 1990, Dayton et al. 1999, Edwards 2004). The biomass dynamics of giant kelp reflect the interplay of these physical and bio - logical forcings that control patterns of recruitment, growth, and mortality.

The relative importance of resource availability (light and nutrients) versus physical disturbance (waves) in controlling the biomass dynamics of giant kelp remains an open question. For example, Dayton et al. (1999) found that large-scale, low-frequency changes in nutrient availability had the largest effects on kelp populations in San Diego; however, recent analyses of kelp forests in central and southern California during 2001 to 2009 (a period lacking any major nutrient-poor El Niño conditions) showed that wave disturbance ex - plained more variability in kelp biomass and produc- tion than either nutrient availability or grazing (Reed et al. 2008, D. C. Reed et al. unpubl.). It is clear that the apparent influence that each of these physical forcings has on kelp populations is dependent on the spatial and temporal scales of observation (Edwards 2004).

While a particular kelp forest may be nutrient-limited at a given time, another forest in a more nutrient-rich region may have its dynamics controlled by wave dis- turbance. The vast majority of long-term kelp studies have been made at the local scale (<500 m2) and so it has been difficult to test how their conclusions apply to larger areas (>1 km2)

### 2018-08-28 proposal_draft_2018.Rmd
moved some files around

@@
 For example a study by @leon2003 used integrated physical (Delft3D hydrological model) and bio-chemical (Agricultural Non-point Source model) processes models to investigate the possible impact on the Lake Malawi water quality due to management actions performed at the watershed level.
+Traditional ecological measures of wave exposure usually incorportates integrative measures of hyrdodynamic conditions at a particular site.
+More specifically, it is the integration of mechanical processes and the influence that is has on the ecology of nearshore communities.
+
+Wave exposure may be modelled through various methods which range from simple cartographic to more advanced numerical wave models.
+Cartographical models can be qualitative or quantitative and were designed for the need of wave exposure measures to explain ecological distributions.
+A simple set of calculations on coastline and wind data, and relatively small input data sets are required.
+These are regarded as "fetch-based models", which measure the length of open water associated with a site along a straight line.
+The output of such an approach is a simplified estimate of the potential wave energy for a specific set of sites.
+Advances in cartographical methods using fetch-based models has allowed for wave exposure measurments for larger areas, and has been suggested as a method for predicting macroalgal community structure [@burrows2008].
+An example of such a model is the "BioEx model" which was developed by @baardseth1970 to estimate wave exposure over large regions.
+BioEx requires frequency, strength and direction of winds, weighted by degree of exposure within various directions.
+BioEx is calculated as the sum of the index developed at different spatial scales (local, fjord and open).
+This method has been used in mapping of marine coastal biodiveristy [@Rinde2004].
+
+@lindegarth2005 critized this approach, arguing that the choice of wave exposure method  can influence ecological inference.
+The authors also highlighted the need for objective, reproducible and quantitative studies comparing exposure indices [@lindegarth2005].
+A study by @sundbald...**include description of study here**...Other authors, such as @hill2010, have argued that these simple measures can be improved upon by including bathymetery data which allows the incorporation of diffraction into the calculation.
+Diffraction is topographically induced variations in wave direction.
+A model incorporating this complexity was developed by @isaeus2004, and is known as the "simplified wave model" (SWM).
+The model uses measurements of wind strength, fetch and empirically derived algorithims to mimic diffraction.
+
+Advances in numerical modelling have been founded on physical wave theory on how a wave "behaves".
+This approach is based on a theoretical persepctive rather than the need to answer ecological questions.
+Besides diffraction, these models incorporate more complexity by including wind forcing, wave-to-wave interactions and loss of energy due to friction and wave breaking.
+Numerical models have a variety of applications and are often incorporated within hydrodynamic general circulation models and are used operationally for forecasting the sea state [@hasselmann1988; @booij1999; @smith2001].
+The downside of advanced numerical models is that are computationally intensive which creates limitations for large scale simulations.
+Therefore, their application along long stretches of variable coastline, inshore environments and ocean-wide simulations limited due to the poor spatial coverage.
+However, this models can be designed for local or site specific coverage, provided the correct data is available.
 
@@
 
-In recent years there has been growing attention on microplastics and the deterimental role they play in the enviornment, particulary the ocean.
+In recent years there has been growing attention on plastic pollution, particulary the ocean.
+Plastic pollution can range from macroplastics to microplastics.
 Microplastics are tiny plastic granuales used as scrubbers in cosmetics and air-blasting, and small plastic fragments that originate from larger pieces of plastic known as macroplastics.
-The potential harms of small pieces of plastic was highlighted in the 1970's and renewed interest has lead to research showing that microplastics in the ocean are widespread.
-The small size of microplastics makes them bio-available to biota in the food-web which may cause problems with an organisms physiological functioning.
-Furthermore, the relatively large surface area and composition of microplastics provides an environmeht that is able of adhering to organic pollutants.
-In other words microplastics also act as a vector for transport and assimilation of organic polluants.
-Therefore, microplastics may also aid in the assimilation of organic pollutans into all levels of the food web.
-
+The potential harms of plastic was highlighted in the 1970's and renewed interest has lead to research showing that plastics in the ocean are widespread.
+Plastic pollution is  bio-available to biota in the food-web which may cause problems with an organisms physiological functioning.
 Therefore, this study not only enables research into the ecological affects of the hydrological environment on an important habitat-forming organism, it also offers the opportunity to improve on current hydrological numerical models to suit coastal environments.

### 2018-08-30 chapter_2.Rmd
house keeping

---
title: 'Chapter 2:The shape of kelp: environmental influences on kelp morphology'
author: "Ross Coppin"
date: "30 August 2018"
output:
  pdf_document: 
    highlight: zenburn
  word_document: default
---
## Introduction

Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera [@Bolton2010], nevertheless form the basis of one of the most productive ecosystems globally [@Mann1973]. Kelps generally have a dependence on cool, temperate and arctic seawater temperatures [@Santelices2007; @Bolton2010], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002]. Outside of these latitudes, they are also found in cool, deep water towards the tropics [@Graham2007]: this deeper tropical water, due to its water clarity, also allows for a minimum photon flux rate of 50 ??? and a nitrate concentration of ??? [@Luning1990; @Dayton1999; @Zimmerman1985], the two other environmental variables that support kelp productivity. Although kelps have a low taxonomic diversity, their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude other turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Dunton1987; @Duggins1989; @Steneck2002], each depending on a wide suite of ecological services provided by the kelp habitat [@Gaines1987; @Bologna1993; @Levin1994; @Anderson1997].

Say something about kelp habitats that have a physical influence on the hydrodynamics within the beds.

The biogeographic distribution of kelp is limited by seawater temperature [@Bolton2010], where increasing temperature gradients reduce kelp distribution. Due to this limiting factor, the two main species of kelps in southern African waters, _Ecklonia maxima_ and _Laminaria pallida_, are distributed along a section of the south coast from De Hoop, extending west around the Cape Peninsula, and thriving north into Namibia [@Molloy1990; @Stegenga1997]. This distribution follows a temperate gradient, where sea temperatures increase as one moves south from Namibia, around Cape Point and towards De Hoop. Although the two species occur together for the majority of the coast, their basic morphologies and resource needs vary to a degree. The larger species, _Ecklonia maxima_, is distributed from Lüderitz to Cape Agulhas [@Bolton1985; @Probyn1985; @Bolton1987; @Bolton2012]. Characterised by a large distal swollen bulb filled with gas, and smooth fronds, this species grows to approximately 10 meters [@Bolton1987]. There was, however, a 17-meter specimen collected in 2015 off Cape Point (Smit, unpubl. data).This species of kelp not only dominate the biomass of the South African nearshore, but plays an important ecological role [@Bustamante1996]. The estimated productivity of _Ecklonia maxima_ within South Africa varies between 350 and 1500g Cm-2yr-1 [@Mann1982]. Across the majority of the coastline, _Laminaria pallida_ remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 meters [@Field1980; @Bolton1987; @Molloy1996]. This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 meters [@Field1980; @Molloy1990; @Molloy1996; @Stegenga1997]. Towards the north along the west coast, from around Hondeklipbaai, _Laminaria pallida_ replaces _Ecklonia maxima_ as the dominant kelp species [@Velimirov1977; @Stegenga1997] and it also occupies increasingly shallower subtidal regions. The northern populations also exhibit an increase in stipe hollowness, compared to the solid stipe morphs in the species’ southern distributions [@Molloy1996]. This variation in morphology was thought to represent two distinct species, with the northern populations formerly described as _Laminaria schinzii_ Foslie [@Molloy1996]. Genetic work has subsequently shown that the two morphs are in fact the same species [@Rothman2017]. In southern African waters, the primary production of _Laminaria pallida_ is between 120 and 1900g C m2yr1, similar to that of _Ecklonia maxima_ [@Mann1982]. Primary production is not the only pathway.

The aim of this study is, therefore, to understand how environmental drivers such as temperature and wave energy can influence morphoplasticity in two species of kelps around South Africa. This will be achieved by initially understanding the variation in abiotic parameters, and morphometrics of _Ecklonia maxima_ and _Laminaria pallida_, around the Western Cape coast. Thereafter we will look at how the abiotic parameters both correlate and influence each other in the nearshore environment of our study region. Finally, we will investigate which abiotic parameters best describe the morphological variation of the two kelp species, using statistical tools. It is predicted that higher wave energy environments will exhibit kelps with reduced undulations, possessing strap-like blades, opposed to low flow regimes that will show greater undulations in frond morphology. An increase in temperature is predicted to influence specific morphological traits that rely on nutrient uptake, as temperature and nutrients share an inverse relationship. An increase in temperature is also predicted to reduce the dominance of the colder water kelp, _Laminaria pallida_, at depth, with a total absence of _Laminaria pallida_ as one approaches the eastern boundary of sample sites.

### Study area

Due to the Cape Peninsula’s temperate latitude, winter months bring an increased frequency of frontal depressions that originate from the Southern Ocean [@Reason2006]. These low pressures are joined by large swells with increased wave energy. The nearshore environment, with the accompanied biota, therefore experiences high wave energy events, with increased frequency in winter. The large peninsula acts as an obstruction for large south westerly swells, providing decreased wave energy along the west side of False Bay [@Shipley1964]. Conversely, the west coast of Cape Point is battered by these large swells at full force. Multiple sites, therefore, exist where kelps grow in diverse temperature and wave energy climates, in close proximity.

\newpage
## Methods
### Morphometrics collection

Between October 2014 and April 2015, morphological measurements of _Laminaria pallida_ and _Ecklonia maxima_ were collected at 18 sites along the Western Cape coast of South Africa (Fig. 1). Eleven samples were collected per morphology, for each species (Table 1, 2). These varying morphometrics allowed measurements such as weight, length and thickness to be compared between sites. Because the macroalgae differ in morphological features, species-specific morphometrics were included. These sites span across the majority of the south-west coast, in varying thermal and wave energy regimes.

```{r, echo= FALSE, message=FALSE, warning=FALSE}
library(knitr)
morph <- read.csv("tables/Morphology table.csv")

colnames(morph) <- c("Morphometric", "Unit measurement")

morph_eck <- morph[(1:10), ]
morph_lam <- morph[(11:19), ]

```
```{r, echo= FALSE, message=FALSE, warning=FALSE}

kable(morph_eck, row.names = FALSE, caption = "A list of morphology measurements that were collected to investigate how environmental drivers may influence the morphology of _Ecklonia maxima _. Units for each morphometric are included.")

```
```{r, echo= FALSE, message=FALSE, warning=FALSE}

kable(morph_lam, row.names = FALSE, caption = "A list of morphology measurements that were collected to investigate how environmental drivers may influence the morphology of _Laminaria pallida_. Units for each morphometric are included.")

```

### Sites

Sites were chosen to represent an array of morphological differences seen within _Ecklonia maxima_ and _Laminaria pallida_. Sites were also chosen to reflect locations that displayed variable wave and temperature regimes (Fig. 2-5), to allow us to robustly test our hypothesis of environmental drivers influencing kelp morphology. St. Helena Bay and Betty’s Bay constituted the north western and south eastern boundary sites respectively. These sites are roughly 300km apart, and lie within separate marine provinces, as outlined above. The Cape Peninsula provides an interesting topographical boundary that shelters the coast in False Bay. Sites were therefore chosen to represent an array of environments, from offshore reefs (Batsata Rock), to sheltered intertidal zones (Miller’s Point). West of Cape Point, a number of sites were chosen to highlight the presence of upwelling (Oudekraal, Kommetjie), as well as kelps growing in protected bays (Hout Bay).

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="A map indicating the sites where morphometric measurements were collected for _Ecklonia maxima_ and _Laminaria pallida_.", fig.height=8, fig.width=12}

library(tidyverse)
library(vegan)
library(RColorBrewer)
library(gridExtra)
library(lubridate)
library(GGally)
library(ggpubr)
library(ggrepel)

# Load map data
load("maps/south_africa_coast.Rdata")
load("maps/sa_provinces.Rdata")
load("maps/site_list.Rdata")

# remove De Hoop
kelp_sites <- read_csv("metadata/sites_complete.csv")
kelp_sites_map <- subset(kelp_sites, !site == "De Hoop")

# Subset into 4 regions
west__up_site <- subset(kelp_sites, site == "St Helena Bay" | site == "Sea Point" 
                    | site == "Oudekraal" | site == "Hout Bay")

west__down_site <- subset(kelp_sites, site == "Soetwater" | site == "Olifantsbos" | site == "Scarborough" | site == "Kommetjie")

east_up_site <- subset(kelp_sites, site == "Roman Rock" | site == "A-Frame" | site == "Miller's Point" | site == "Baboon Rock" | site == "Kalk Bay")

east_down_site <- subset(kelp_sites, site == "Batsata Rock" | site == "Bordjies Deep" | site == "Buffelsbaai South" 
                         | site == "Buffelsbaai" | site == "Betty's Bay")

# plot map
ggplot(data = kelp_sites_map, aes(x = lon, y = lat)) +
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = kelp_sites_map)  +
  coord_fixed(xlim = c(17.65, 19.5), ylim = c(-32.75, -34.6)) +
  geom_label_repel(data = west__up_site, aes(x = lon, y = lat, label = site), 
                   size = 3, box.padding = 0.5, nudge_x = -0.5, nudge_y = 0.2, segment.alpha = 0.4) +
  geom_label_repel(data = west__down_site, aes(x = lon, y = lat, label = site), 
                   size = 3, box.padding = 0.5, nudge_x = -0.5, nudge_y = -0.2, segment.alpha = 0.4) +
  geom_label_repel(data = east_up_site, aes(x = lon, y = lat, label = site), 
                   size = 3, box.padding = 0.5, nudge_x = 0.5, nudge_y = 0.3, segment.alpha = 0.4) +
  geom_label_repel(data = east_down_site, aes(x = lon, y = lat, label = site), 
                   size = 3, box.padding = 0.5, nudge_x = 0.35, nudge_y = -0.15, segment.alpha = 0.4) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

```

### Abiotic parameters

In order to compare abiotic parameters for sites around the coast, large historical databases for both temperature and wave energy were accessed. Shallow water temperatures were sourced from the South African Coastal Temperature Network (SACTN) website (https://github.com/ajsmit/SACTN). Seven different organisations within South Africa contribute to the SACTN, where _in situ_ temperature measurements are made around the South African coast using either hand-held thermometers or digital temperature recorders positioned underwater. The mean duration of the 135 daily time series is 19.7 years, and these _in situ_ data are preferred over satellite SST, which have shown to exhibit large biases [@Smit2013]. Linear interpolated SST were calculated for sites where _in situ_ recorders were absent. Wave energy data formed part of the South African Coastal Vulnerability Assessment, presented to the Department of Environmental Affairs [@Theron2014]. These data are first forecasted using NOAA Wave Watch III (WWIII), with National Centers for Environmental Predictions (NCEP) product as the numerical input [@benassai2006implementation]. Hindcast data from WWIII span from 1994-2013 at a 3-hour resolution. The data are then used to model short –crested waves generated by the wind into the coastal environment, using Simulating Waves in the Nearshore [SWAN; @booij1997swan]. SWAN allows one to extract wave parameters from specific gridded locations in the nearshore. For False Bay, a resolution of 200 meters was modelled, at both 7 meter and 15 meter contours. A 200 meter resolution was used as False Bay was nested within a larger grid area for the research from where the data were sourced.  For Table Bay and east of Cape Hangklip the resolution 500 meters at 7 meters and 15 meters. For this study the 7 meter contours were used.

###Statistical analyses

To compare how kelp morphology varies around the coast, boxplots were constructed to summarise descriptive statistics for all the morphometrics of both species.These boxplots highlight five different descriptive statistics (Minimum, 25^th^ percentile, median, 75^th^ percentile and maximum), as well as the interquartile range. These allow us to visually identify variations and differences in morphology, and to provide evidence that kelp morphologies vary around the coast.Pairwise correlations were plotted to compare the abiotic parameters and understand if wave and temperature parameters correlate with one another around the coast. Therefore fluctuations such as minimum, maximum, range and standard deviations were included as temperature parameters, and standard deviations were included as wave parameters. Median calculations were made for wind and wave direction, as issues arise when calculating mean and standard deviation for compass metrics. Redundancy Analyses (RDA) were performed to understand how kelp morphology is driven by environmental drivers. An RDA performs multiple linear regressions between explanatory and response variables. This allows the user to calculate the amount of variation in response variables explained for by explanatory variables. Therefore response variables are influenced by explanatory variables.  Response variables were represented by morphology measurements, with wave and temperature variables selected as explanatory variables. Temperature and wave parameters were modelled separately, to fully understand and tease apart which abiotic variables most strongly explain kelp morphology variation. This was therefore performed for each species, equating to four RDA’s in total.

\newpage
## Results

```{r, echo= FALSE, message=FALSE, warning=FALSE}

# Complete lon/lats
kelp_sites <- read_csv("metadata/sites_complete.csv")

# Ecklonia
eck_data <- read_csv("data/Ecklonia_morphometrics.csv") %>%
  mutate(site = replace(site, site == "Buffels Bay", "Buffelsbaai")) %>%
  mutate(site = replace(site, site == "St. Helena Bay", "St Helena Bay")) %>%
  # Not using Rocky Bank in this study
  filter(site != "Rocky Bank")

# Laminaria
lam_data <- read_csv("data/Laminaria_morphometrics.csv") %>%
  mutate(site = replace(site, site == "Buffels Bay", "Buffelsbaai")) %>%
  # NB: Even though I am changing the name, there will still be no lon/lat values for this site
  # We need to decide how we want to handle this. Give it the same lon/lat as Buffelsbaai?
  mutate(site = replace(site, site == "Buffels Bay South", "Buffelsbaai South")) %>%
  # NB: This may also not be an appropriate change
  mutate(site = replace(site, site == "Bordjiestif North", "Bordjies Deep")) %>%
  # Not using Rocky Bank in this study
  filter(site != "Rocky Bank")

# NB: (RWS) I have removed Roman Rock as this has less than a year of temperature data

# The SACTN monthly data
load("data/SACTNmonthly_v4.1.Rdata")

# The missing site temperatures
load("data/interp_sites.Rdata")

# The patched Roman Rock temperatures
load("data/rr_patch.Rdata")

## Subset temperatures for collection sites
temps <- SACTNmonthly_v4.1 %>% 
  filter(site %in% eck_data$site | site %in% lam_data$site) %>% 
  select(site, date, temp) %>% 
  rbind(., interp_sites) %>% 
  filter(site!= "Roman Rock") %>% 
  rbind(., rr_patch)
```

```{r, echo= FALSE, message=FALSE, warning=FALSE, results='hide'}

## Up next we need to calculate the relevant temperature statistics for each site. These are: mean, range, and sd, for Annual, February, and August climatologies.

# Monthly clims
temp_feb_aug <- temps %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  group_by(site, date) %>% 
  summarise(mean_temp = mean(temp, na.rm = T),
            min_temp = min(temp, na.rm = T),
            max_temp = max(temp, na.rm = T),
            range_temp = range(temp, na.rm = T)[2]-range(temp, na.rm = T)[1],
            sd_temp = sd(temp, na.rm = T)) %>%
  filter(date %in% c("Feb", "Aug")) %>% 
  ungroup()

# Annual clims
temp_ann <- temps %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  group_by(site, date) %>% 
  summarise(mean_temp = mean(temp, na.rm = T),
            min_temp = min(temp, na.rm = T),
            max_temp = max(temp, na.rm = T),
            range_temp = range(temp, na.rm = T)[2]-range(temp, na.rm = T)[1],
            sd_temp = sd(temp, na.rm = T)) %>% 
  ungroup() %>% 
  select(-date) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  mutate(date = "Ann") %>% 
  select(site, date, everything())

# Combine
temp_clim <- rbind(temp_feb_aug, temp_ann) %>% 
  group_by(site) %>% 
  gather(key = statistic, value = value, 
         mean_temp:sd_temp) %>% 
  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
  select(-date, -statistic) %>% 
  spread(key = clim_stat, value = value)

# Here we need to convert the thermal metrics to Euclidian distances.

env3 <- temp_clim %>% 
  dplyr::ungroup(site) %>% 
  dplyr::select(-site) %>% 
  vegdist(method = 'euclidian')
env3.mat <- as.matrix(env3)
env3.diag <- diag(env3.mat[-1, -nrow(env3.mat)])
env3.diag <- append(0, env3.diag, after = 1)

# Wave data

# Load the loading functions
source("func/load.wave.R")

# Then load the data
fb_model <- load.model("data/wave_data/FB")
tb_model <- load.model("data/wave_data/TB")
he_model <- load.model("data/wave_data/HE")
sh_model <- load.model("data/wave_data/SH")

# Combine
wave_model <- rbind(fb_model, tb_model, he_model, sh_model)
rm(fb_model, tb_model)

# The site list key
wave_sites <- read_csv("metadata/sites_complete.csv") %>% 
  select(site, wave_7, wave_15)

# Calculate the necessary stats
# Monthly clims
wave_feb_aug <- wave_model %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  filter(site!= "FB25") %>% # FB_25 has issues with 'hs' and 'tp'
  group_by(site, date) %>% 
  summarise_all(funs(mean = mean, sd = sd), na.rm = T) %>%
  filter(date %in% c("Feb", "Aug")) %>% 
  ungroup()

# Annual clims
wave_ann <- wave_model %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  filter(site != "FB25") %>% # FB_25 has issues with 'hs' and 'tp'
  group_by(site, date) %>% 
  summarise_all(funs(mean = mean, sd = sd), na.rm = T) %>%
  ungroup() %>% 
  select(-date) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean), na.rm = T) %>% 
  mutate(date = "Ann") %>% 
  select(site, date, everything())

# Subset directions out
dirs <- subset(wave_model, select = c(1, 2, 5, 6))

# Seasons
dirs_feb_aug <- dirs %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  group_by(site, date) %>% 
  summarise(dir_min = min(dir, na.rm = T),
            dir_max = max(dir, na.rm = T),
            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
            dir_median = median(dir, na.rm = T), 
            dirw_min = min(dirw, na.rm = T),
            dirw_max = max(dirw, na.rm = T),
            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
            dirw_median = median(dirw, na.rm = T)) %>%
  filter(date %in% c("Feb", "Aug")) %>% 
  ungroup()

# Annual
dirs_ann <- dirs %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  group_by(site, date) %>% 
  summarise(dir_min = min(dir, na.rm = T),
            dir_max = max(dir, na.rm = T),
            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
            dir_median = median(dir, na.rm = T), 
            dirw_min = min(dirw, na.rm = T),
            dirw_max = max(dirw, na.rm = T),
            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
            dirw_median = median(dirw, na.rm = T)) %>% 
  ungroup() %>% 
  select(-date) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  mutate(date = "Ann") %>% 
  select(site, date, everything())

# subset useless columns
dirs_feb_aug <- subset(dirs_feb_aug, select = c(1, 2, 6, 10))
dirs_ann <- subset(dirs_ann, select = c(1, 2, 6, 10))

# combine all direction metrics
dirs_clim <- rbind(dirs_feb_aug, dirs_ann) %>% 
  group_by(site) %>% 
  gather(key = statistic, value = value, 
         3:4) %>% 
  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
  select(-date, -statistic) %>% 
  spread(key = clim_stat, value = value) %>% 
  left_join(wave_sites, by = c("site" = "wave_7")) %>% 
  ungroup() %>% 
  select(-site, wave_15) %>%
  rename(site = site.y) %>% 
  select(site, everything()) %>% 
  na.omit()

# Combine
# NB: This only uses the wave data from the 7 metre contour
wave_clim <- rbind(wave_feb_aug, wave_ann) %>% 
  group_by(site) %>% 
  gather(key = statistic, value = value, 
         hs_mean:spw_sd) %>% 
  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
  select(-date, -statistic) %>% 
  spread(key = clim_stat, value = value) %>% 
  left_join(wave_sites, by = c("site" = "wave_7")) %>% 
  ungroup() %>% 
  select(-site, wave_15) %>%
  rename(site = site.y) %>% 
  select(site, everything()) %>% 
  na.omit()

# combine all wave metrics
wave_clim <- cbind(wave_clim, dirs_clim)

# remove all useless direction metrics from original wave_clim
wave_clim <- subset(wave_clim, select = -c(2,3,4,5,12,13,14,15,22,23,24,25,32,33,40))

## First create mean vlaues from each site for each species
# Ecklonia
eck_mean <- eck_data %>% 
  select(site, frondMass:totalLength) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean(., na.rm = T)))

# Laminaria
lam_mean <- lam_data %>% 
  select(site, laminaWeight:totalLength) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean(., na.rm = T)))

# Correct errors
# No weight measurements for Laminaria at Sea Point for some reason...
lam_mean$laminaWeight[lam_mean$site == "Sea Point"] <- NA

## Merge the biotic and abiotic variables
# Ecklonia
eck_all <- eck_mean %>% 
  left_join(temp_clim, by = "site") %>%
  left_join(wave_clim, by = "site") %>% # With wave data
  na.omit()
eck_bio <- eck_all %>% 
  select(frondMass:totalLength) %>% 
  decostand(method = "standardize")
eck_abio <- eck_all %>% 
  # select(Ann_mean_temp:Feb_sd_temp) %>% # Without wave data
  select(Ann_mean_temp:Feb_tp_sd) %>% # With wave data
  decostand(method = "standardize")

# Laminaria
lam_all <- lam_mean %>% 
  left_join(temp_clim, by = "site") %>%
  left_join(wave_clim, by = "site") %>% # With wave data
  na.omit()
lam_bio <- lam_all %>% 
  select(laminaWeight:totalLength) %>% 
  decostand(method = "standardize")
lam_abio <- lam_all %>% 
  # select(Ann_mean_temp:Feb_sd_temp) %>% # Without wave data
  select(Ann_mean_temp:Feb_tp_sd) %>% # With wave data
  decostand(method = "standardize")

```

```{r, echo= FALSE, message=FALSE, warning=FALSE}
# Merge site list with temp_clim
site_temp <- merge(temp_clim, kelp_sites, by = "site")

# Remove unnec columns
site_temp <- subset(site_temp, select = -c(19:21))

# gather data into less columns
site_temp <- gather(data = site_temp, "stat", "value", 2:16)

# Split columns
site_temp <- site_temp %>% 
  separate(stat, c("season", "statistic", "abioPara"))

# change season key
season <- c("Ann" = "Annual", "Feb" = "February", "Aug" = "August")
statistic <- c("mean" = "Mean", "sd" = "SD", "range" = "Range", "max" = "Max", "min" = "Min")

```

```{r, echo= FALSE, message=FALSE, warning=FALSE}

maxTemp <- subset(site_temp, statistic == "max")
minTemp <- subset(site_temp, statistic == "min")
meanTemp <- subset(site_temp, statistic == "mean")
sdTemp <- subset(site_temp, statistic == "sd")
rangeTemp <- subset(site_temp, statistic == "range")

max <- ggplot(data = maxTemp, aes(x = lon, y = lat)) +
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = maxTemp, aes(col = value), size = 2) +
  scale_color_gradient("Max temp (°)", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(season), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

min <- ggplot(data = minTemp, aes(x = lon, y = lat)) +
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = minTemp, aes(col = value), size = 2) +
  scale_color_gradient("Min temp (°)", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(season), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

mean <- ggplot(data = meanTemp, aes(x = lon, y = lat)) +
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = meanTemp, aes(col = value), size = 2) +
  scale_color_gradient("Mean temp (°)", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(season), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

sd <- ggplot(data = sdTemp, aes(x = lon, y = lat)) +
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = sdTemp, aes(col = value), size = 2) +
  scale_color_gradient("SD temp (°)", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(season), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

range <- ggplot(data = rangeTemp, aes(x = lon, y = lat)) +
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = rangeTemp, aes(col = value), size = 2) +
  scale_color_gradient("Range temp (°)", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(season), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

```
### Temperature parameters

The temperature parameters around the Western Cape coast vary among sites and seasons. During February and August, the west side of Cape point experiences decreased temperatures (Fig. 2).  This region also experiences lower mean and maximum temperatures when compared to False Bay, a region known as east of Cape Point. The range of temperatures within False Bay are larger in winter (August) compared to summer (February) (Fig. 3). The sites north of Kommetjie on the west side display larger temperatures ranges than sites south of Kommetjie on the west side.

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Morphometric site locations represented by various temperature parameters. Temperature parameters include minimum, maximum and mean temperatures (° Celsius). These site locations are coloured by the temperature statistic relative to the legends provided. Each temperature parameter is also divided into August (winter) and February (summer) as well as the annual mean.", fig.height=8, fig.width=12, fig.pos="H"}

grid.arrange(min, max, mean, ncol = 1)
ggarrange(min, max, mean, ncol = 1, nrow = 1, labels = "AUTO", legend = "bottom", common.legend = TRUE, align = "none", heights = c(1.5,1))
```

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Morphometric site locations represented by various temperature parameters. Temperature parameters include temperature range as well as standard deviation (° Celsius). These site locations are coloured by the temperature statistic relative to the legends provided. Each temperature parameter is also divided into August (winter) and February (summer) as well as the annual mean.", fig.height=8, fig.width=12, fig.pos="H"}

grid.arrange(range, sd, ncol = 1)

```

```{r, echo= FALSE, message=FALSE, warning=FALSE}

# Wave parameters

# merge sites with wave_clim
site_wave <- merge(wave_clim, kelp_sites, by = "site")

# remove unecc columns
site_wave <- subset(site_wave, select = -c(28:30))
# gather data into less columns

site_wave <- gather(data = site_wave, "stat", "value", 2:25)
# split columns
site_wave <- site_wave %>% 
  separate(stat, c("season", "abioPara", "statistic"))

# change season key
season <- c("Ann" = "Annual", "Feb" = "February", "Aug" = "August")
statistic <- c("mean" = "Mean", "sd" = "SD", "range" = "Range", "max" = "Max", "min" = "Min", "median" = "Median")
abioPar <- c("dir" = "Direction", "dirw" = "Wind direction", "hs" = "Sig. wave height", "tp" = "Wave period", "spw" = "Spw")

# fix incorrect

```

```{r, echo= FALSE, message=FALSE, warning=FALSE}

## MEthod 2 (shows individual legends, still squashed)

medianDir <- subset(site_wave, statistic == "median" & abioPara == "dir")
medianDirw <- subset(site_wave, statistic == "median" & abioPara == "dirw")
meanHs <- subset(site_wave, statistic == "mean" & abioPara == "hs")
sdHs <- subset(site_wave, statistic == "sd" & abioPara == "hs")
meanSpw <- subset(site_wave, statistic == "mean" & abioPara == "spw")
sdSpw <- subset(site_wave, statistic == "sd" & abioPara == "spw")
meanTp <- subset(site_wave, statistic == "mean" & abioPara == "tp")
sdTp <- subset(site_wave, statistic == "sd" & abioPara == "tp")

# meanDir
medianDir1 <- ggplot(data = medianDir, aes(x = lon, y = lat)) +
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = medianDir, aes(col = value), size = 2) +
  scale_color_gradient("medianDir", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(c(season)), scales = "free", ncol = 3) + 
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

# meanDirw 
medianDirw1 <- ggplot(data = medianDirw , aes(x = lon, y = lat)) + 
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = medianDirw , aes(col = value), size = 2) +
  scale_color_gradient("medianDirw ", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(c(season)), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

# meanHs

meanHs1 <- ggplot(data = meanHs, aes(x = lon, y = lat)) + 
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = meanHs, aes(col = value), size = 2) +
  scale_color_gradient("meanHs", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(c(season)), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

# sdHs
sdHs1 <- ggplot(data = sdHs, aes(x = lon, y = lat)) + 
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = sdHs, aes(col = value), size = 2) +
  scale_color_gradient("sdHs", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(c(season)), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

# meanSpw
meanSpw1 <- ggplot(data = meanSpw, aes(x = lon, y = lat)) + 
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = meanSpw, aes(col = value), size = 2) +
  scale_color_gradient("meanSpw", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(c(season)), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

# sdSpw
sdSpw1 <- ggplot(data = sdSpw, aes(x = lon, y = lat)) + 
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = sdSpw, aes(col = value), size = 2) +
  scale_color_gradient("sdSpw", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(c(season)), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

# meanTp
meanTp1 <- ggplot(data = meanTp, aes(x = lon, y = lat)) + 
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = meanTp, aes(col = value), size = 2) +
  scale_color_gradient("meanTp", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(c(season)), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

# sdTp
sdTp1 <- ggplot(data = sdTp, aes(x = lon, y = lat)) + 
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = sdTp, aes(col = value), size = 2) +
  scale_color_gradient("sdTp", low = "blue", high = "red") +
  coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
  facet_wrap(~ season, labeller = as_labeller(c(season)), scales = "free", ncol = 3) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())
```

### Wave and wind parameters

Not only do the temperature parameters display seasonal fluctuations, but we see similar observations for wave parameters. The west side of Cape Point exhibit increased measurements of median direction, SD and mean significant wave height compared to within False Bay (Fig. 4). Sites within False Bay however display increased standard deviations of wave period. Median wind direction is northerly in winter (Fig. 5), and turns southerly in summer.

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Wave parameters relative to each morphometric collection site around the Western Cape coast. These wave parameters include mean and standard deviation of wave direction (° True north), significant wave height (Meters) as well as wave period (Seconds). Sites are colour-coded by the parameter statistic provided by the legend. Each wave parameter is also divided into August (winter), February (summer) and annual means, to visualise seasonal differences.", fig.height=8, fig.width=12, fig.pos="H"}

grid.arrange(medianDir1, meanHs1, sdHs1, meanTp1, sdTp1)

```

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Wind parameters relative to each morphometric collection site around the Western Cape coast. These wind parameters include mean and standard deviation of wind direction (° True north) as well as wind speed (Meters/ second). Sites are colour-coded by the parameter statistic provided by the legend. Each wave parameter is also divided into August (winter), February (summer) and annual means, to visualise seasonal differences.", fig.height=8, fig.width=12, fig.pos="H"}

grid.arrange(medianDirw1, meanSpw1, sdSpw1)

```

### Morphologies

#### _Laminaria pallida_
Lamina length for laminaria pallida showed no geographical pattern moving from west to east. Kommetjie, Olifantsbos and Batsata Rock showed great variability in lamina length, and both Buffels Bay and Betty’s Bay were visually different to Miller’s Point and Roman Rock, when comparing the summary data represented by the boxplots (Fig.6). Lamina thickness showed great variation across sites, with Baboon Rock, Miller’s Point, A-Frame and Roman rock displaying large lamina thickness, significantly different to the rest of the sites. Lamina weight was observed to vary for Baboon Rock and Betty’s Bay. Neighbouring sites A-Frame and Roman Rock also showed visual difference when comparing boxplot summary statistics. An increase in the number of digits was observed as one moved from Cape Point north along the western side of False Bay. This ceased at Batstata Rock, which exhibited significantly less digits compared to the previous site, Bordjiestif North. Stipe diameter showed some geographical grouping, with west of Cape Point sites exhibiting larger stipe diameters compared to False Bay sites.

Stipe diameter decreased as one rounded the point into False Bay, where a sudden, significant difference was seen between Bordjiestif North and Batsata Rock. Greater variation of stipe length was observed for sites outside of False Bay (Kommetjie, Olifantsbos and Betty’s Bay). Baboon Rock, Miller’s Point, A-Frame and Roman Rock were again grouped together exhibiting the lowest stipe lengths, that were all significantly different to sites found west of Cape Point. Stipe mass displays similar patterns to stipe length, with larger stipe lengths west of Cape Point compared to within False Bay. The thallus mass of _Laminaria pallida_ was observed to be greater for sites around Cape Point, with an observed difference between Batsata Rock and Bordjiestif North, similar to the number of digit patterns. Larger total lengths were observed around Cape Point and Betty’s Bay, with the smallest total lengths at sites that exhibited the greatest lamina thicknesses (Baboon Rock, Miller’s Point, A-Frame and Roman Rock).

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Boxplots representing the different _Laminaria pallida_ morphometrics measured around the Western Cape coastline, with the X-axis depicting the specific morphology measured, with units provided. Boxplots represent the minimum, 25^th^ percentile, median and 75^th^ percentile of the morphometrics measured. Interquartile range can be deduced as the different between the 75^th^ and 25^th^ percentiles, and dots represent outliers in the data. Sites are ordered sequentially on the Y-axis by location along the coast. The top site is Sea Point and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.height=8, fig.width=12, fig.pos="H"}

# Laminaria morphologies

# load data
lam <- read.csv("data/Laminaria_morphometrics.csv", sep = ",")

# remove unnecessary columns
lam <- subset(lam, select = -c(1,2,4))

# alter dataframe for more friendly plotting
lam_gather <- gather(lam, "Morphology", "Measurement", 2:10, na.rm = TRUE)

# Replace morphology names
lam_gather <- lam_gather %>%
  mutate(Morphology = replace(Morphology, 
                              Morphology == "digits", "Number of digits")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "laminaLength", "Lamina length (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "laminaThickness", "Lamina thickness (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "laminaWeight", "Lamina weight (g)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "stipeDiameter", "Stipe diameter (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "stipeLength", "Stipe Length (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "stipeMass", "Stipe mass (g)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "thallusMass", "Thallus mass (g)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "totalLength", "Total length (mm)"))

# Mean morphology per site

lam_gather_mean <- lam_gather %>%
  group_by(site, Morphology) %>%
  summarise(mean = mean(Measurement, na.rm = TRUE)) %>%
  ungroup()

# Remove Rocky Bank
lam_gather_mean <- subset(lam_gather_mean, !site == "Rocky Bank")

# Order site names by location west to east
positionsLam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point", "Baboon Rock", 
                  "Batsata Rock", "Bordjiestif North", "Buffels Bay South", 
                  "Buffels Bay", "Olifantsbos", "Kommetjie", "Sea Point")

# Plot morphologies
ggplot(data = lam_gather, aes(x = site, y = Measurement)) +
  geom_boxplot() +
  facet_wrap(~Morphology, scales = "free_x", strip.position = "bottom") +
  coord_flip() + 
  scale_x_discrete(limits = positionsLam) + 
  theme(strip.background = element_blank(), strip.placement = "outside", text = element_text(size = 4.5)) +
  theme_bw() +
  xlab("Site") +
  ylab("Morphology measurement")

```

#### _Ecklonia maxima_

For frond length, frond mass, stipe circumferences, stipe length and total length of _Ecklonia maxima_ morphometrics, we see a gradual increase in value as one moves south from St. Helena Bay to Kommetjie and Soetwater (Fig. 7). Hout Bay, Kommetjie and Soetwater show similarities in their stipes lengths to Buffels Bay, Batsata Rock and Betty’s Bay, and are significantly different to west of Cape Point counterpart sites such as Oudekraal and Scarborough. A difference for the majority of the morphologies are seen between Soetwater and Scarborough, with significant differences for epiphyte length, frond length, frond mass, stipe length, stipe mass and total length.  A separation along the same stretch of coast (west of Cape Point) for stipe mass shows that Hout Bay, Kommetjie and Soetwater are again significantly different to neighboring west side sites such as Oudekraal and Scarborough.

```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Boxplots representing the different _Ecklonia maxima_ morphometrics measured around the Western Cape coastline, with the X-axis depicting the specific morphology measured, with units provided. Boxplots represent the minimum, 25^th^ percentile, median and 75^th^ percentile of the morphometrics measured. Interquartile range can be deduced as the different between the 75^th^ and 25^th^ percentiles, and dots represent outliers in the data. Sites are ordered sequentially on the Y-axis by location along the coast. The top site is St. Helena Bay and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.height=8, fig.width=12, fig.pos="H"}

# Ecklonia morphologies

# load data
eck <- read.csv("data/Ecklonia_morphometrics.csv", sep = ",")

# remove unnecessary columns
eck <- subset(eck, select = -c(1,2,3,5,7,8))

# alter dataframe for more friendly plotting
eck_gather <- gather(eck, "Morphology", "Measurement", 2:12, na.rm = TRUE)

# Replace morphology names
eck_gather <- eck_gather %>%
  mutate(Morphology = replace(Morphology, 
                              Morphology == "depth", "Depth (m)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "epiLength", "Epiphyte length (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "frondLength", "Frond length (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "frondMass", "Frond mass (g)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "primaryLength", "Primary length (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "primaryWidth", "Primary width (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "stipeCirc", "Stipe circumference (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "stipeLength", "Stipe length (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "stipeMass", "Stipe mass (g)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "totalLength", "Total length (mm)")) %>% 
  mutate(Morphology = replace(Morphology, 
                              Morphology == "tufts", "Number of tufts"))

# Mean morphology per site

eck_gather_mean <- eck_gather %>%
  group_by(site, Morphology) %>%
  summarise(mean = mean(Measurement, na.rm = TRUE)) %>%
  ungroup()

# Order site names by location west to east
positionsEck <- c("Betty's Bay", "Kalk Bay", "A-Frame", 
                  "Miller's Point", "Baboon Rock", "Batsata Rock", 
                  "Buffels Bay", "Scarborough", 
                  "Soetwater","Kommetjie","Hout Bay", 
                  "Oudekraal", "St. Helena Bay")

# Remove Depth and Rocky Bank site
eck_gather <- subset(eck_gather, !Morphology == "Depth (m)")
eck_gather <- subset(eck_gather, !site == "Rocky Bank" | !site == "De Hoop")

# Plot morphologies
ggplot(data = eck_gather, aes(x = site, y = Measurement)) +
  geom_boxplot() +
  facet_wrap(~Morphology, scales = "free_x", strip.position = "bottom", ncol = 5) +
  coord_flip() + 
  scale_x_discrete(limits = positionsEck) +
  theme(strip.background = element_blank(), strip.placement = "outside", text = element_text(size = 4.5)) +
  theme_bw() +
  xlab("Site") +
  ylab("Morphology measurement")

```

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide"}

# merge all abiotic data
combine <- merge(temp_clim, wave_clim, by = "site")

# Gather data
abio_cor <- gather(combine, "parameter", "measurement", 2:40)

# Order by season/ parameter
abio_cor <- plyr::arrange(abio_cor, parameter)

# Subset by season
ann_abio_cor <- abio_cor[c(1:221), ]
aug_abio_cor <- abio_cor[c(222:442), ]
feb_abio_cor <- abio_cor[c(443:663), ]

# Change describing terms in rows
# Annual
ann_abio_cor <-  ann_abio_cor %>%
  mutate(parameter = replace(parameter, parameter == "Ann_dir_median", "Median dir")) %>%
  mutate(parameter = replace(parameter, parameter == "Ann_dirw_median", "Median wind dir")) %>%
  mutate(parameter = replace(parameter, parameter == "Ann_hs_mean", "Mean Hs")) %>%
  mutate(parameter = replace(parameter, parameter == "Ann_hs_sd", "SD Hs")) %>%
  mutate(parameter = replace(parameter, parameter == "Ann_tp_mean", "Mean Tp")) %>%
  mutate(parameter = replace(parameter, parameter == "Ann_tp_sd", "SD Tp")) %>%
  mutate(parameter = replace(parameter, parameter == "Ann_spw_mean", "Mean sPw")) %>%
  mutate(parameter = replace(parameter, parameter == "Ann_spw_sd", "SD sPw")) %>%
  mutate(parameter = replace(parameter, parameter == "Ann_mean_temp", "Mean temp")) %>%
  mutate(parameter = replace(parameter, parameter == "Ann_sd_temp", "SD temp")) %>% 
  mutate(parameter = replace(parameter, parameter == "Ann_range_temp", "Range temp")) %>% 
  mutate(parameter = replace(parameter, parameter == "Ann_max_temp", "Max temp")) %>% 
  mutate(parameter = replace(parameter, parameter == "Ann_min_temp", "Min temp"))

# August
aug_abio_cor <- aug_abio_cor %>%
  mutate(parameter = replace(parameter, parameter == "Aug_dir_median", "Median dir")) %>%
  mutate(parameter = replace(parameter, parameter == "Aug_dirw_median", "Median wind dir")) %>%
  mutate(parameter = replace(parameter, parameter == "Aug_hs_mean", "Mean Hs")) %>%
  mutate(parameter = replace(parameter, parameter == "Aug_hs_sd", "SD Hs")) %>%
  mutate(parameter = replace(parameter, parameter == "Aug_tp_mean", "Mean Tp")) %>%
  mutate(parameter = replace(parameter, parameter == "Aug_tp_sd", "SD Tp")) %>%
  mutate(parameter = replace(parameter, parameter == "Aug_spw_mean", "Mean sPw")) %>%
  mutate(parameter = replace(parameter, parameter == "Aug_spw_sd", "SD sPw")) %>%
  mutate(parameter = replace(parameter, parameter == "Aug_mean_temp", "Mean temp")) %>%
  mutate(parameter = replace(parameter, parameter == "Aug_sd_temp", "SD temp")) %>% 
  mutate(parameter = replace(parameter, parameter == "Aug_range_temp", "Range temp")) %>% 
  mutate(parameter = replace(parameter, parameter == "Aug_max_temp", "Max temp")) %>% 
  mutate(parameter = replace(parameter, parameter == "Aug_min_temp", "Min temp"))

# February
feb_abio_cor <- feb_abio_cor %>%
  mutate(parameter = replace(parameter, parameter == "Feb_dir_median", "Median dir")) %>%
  mutate(parameter = replace(parameter, parameter == "Feb_dirw_median", "Median wind dir")) %>%
  mutate(parameter = replace(parameter, parameter == "Feb_hs_mean", "Mean Hs")) %>%
  mutate(parameter = replace(parameter, parameter == "Feb_hs_sd", "SD Hs")) %>%
  mutate(parameter = replace(parameter, parameter == "Feb_tp_mean", "Mean Tp")) %>%
  mutate(parameter = replace(parameter, parameter == "Feb_tp_sd", "SD Tp")) %>%
  mutate(parameter = replace(parameter, parameter == "Feb_spw_mean", "Mean sPw")) %>%
  mutate(parameter = replace(parameter, parameter == "Feb_spw_sd", "SD sPw")) %>%
  mutate(parameter = replace(parameter, parameter == "Feb_mean_temp", "Mean temp")) %>%
  mutate(parameter = replace(parameter, parameter == "Feb_sd_temp", "SD temp")) %>% 
  mutate(parameter = replace(parameter, parameter == "Feb_range_temp", "Range temp")) %>% 
  mutate(parameter = replace(parameter, parameter == "Feb_max_temp", "Max temp")) %>% 
  mutate(parameter = replace(parameter, parameter == "Feb_min_temp", "Min temp"))

# Spread sideways for each season
ann_abio_cor <- spread(ann_abio_cor, "parameter", "measurement")
aug_abio_cor <- spread(aug_abio_cor, "parameter", "measurement")
feb_abio_cor <- spread(feb_abio_cor, "parameter", "measurement")

```

### Abiotic correlations

There are various strong correlations within the temperature parameters, and within the wave parameters on an annual time scale (Fig. 8). SD Significant wave height (Hs) and mean Hs showed a strong positive correlate(0.92), as well as  SD wave period (Tp) and mean Hs with a strong negative correlation(-0.9) For SD wind speed and mean wind speed a strong positive correlation existed (0.967). Minimum and mean temperatures correlate strongly (0.918), as well as SD wind speed and median wind direction (0.999). A strong correlation exists between SD temperature and temperature range (0.942). Median wind direction and mean wind speed are also correlated well (0.954). We however see no strong correlations between wave and temperature parameters.

These correlations vary through the three timescales (Annual, August and February). Although there are no strong correlations between temperature and wave parameters, we see interesting differences between the seasonal timescales The mean temperature and mean wind speed are weakly correlated in August (0.253)(Fig. 9), but are observed to be more strongly negatively correlated in February (-0.788)(Fig. 10).  The same is seen for minimum temperature and median wind direction in August (0.129) compared to February (-0.715).

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="A correlation graph depicting the relationships that various temperature and wave parameters share, at an annual timescale. The left Y-axis as well as bottom X-axis represent the measurements for the parameters, while the top X-axis and right Y-axis depict the abiotic parameter name. The top right section of the graph represents correlation coefficients between abiotic parameters. The bottom left section provides individual linear regressions between each abiotic parameter, with a fitted line. Each point of each linear graph represents a collection site. The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.height=8, fig.width=12, fig.pos="H"}
# # ggpairs to plot correlation coefficients
# # Annual
ggpairs(data = ann_abio_cor, columns = 2:14, 
        upper = list(continuous = wrap("cor", size = 3)), 
        lower = list(continuous = "smooth"), 
        diag = list(continuous = "densityDiag")) + 
  theme(text = element_text(size = 9)) + 
  ggtitle("Annual")

```

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="A correlation graph depicting the relationships that various temperature and wave parameters share, during August. The left Y-axis as well as bottom X-axis represent the measurements for the parameters, while the top X-axis and right Y-axis depict the abiotic parameter name. The top right section of the graph represents correlation coefficients between abiotic parameters. The bottom left section provides individual linear regressions between each abiotic parameter, with a fitted line. Each point of each linear graph represents a collection site. The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.height=8, fig.width=12, fig.pos="H"}

# # August
ggpairs(data = aug_abio_cor, columns = 2:14, 
        upper = list(continuous = wrap("cor", size = 3)), 
        lower = list(continuous = "smooth"), 
        diag = list(continuous = "densityDiag")) + 
  theme(text = element_text(size = 9)) + 
  ggtitle("August")

```

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="A correlation graph depicting the relationships that various temperature and wave parameters share, during February. The left Y-axis as well as bottom X-axis represent the measurements for the parameters, while the top X-axis and right Y-axis depict the abiotic parameter name. The top right section of the graph represents correlation coefficients between abiotic parameters. The bottom left section provides individual linear regressions between each abiotic parameter, with a fitted line. Each point of each linear graph represents a collection site. The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.height=8, fig.width=12, fig.pos="H"}

# # February
ggpairs(data = feb_abio_cor, columns = 2:14, 
        upper = list(continuous = wrap("cor", size = 3)), 
        lower = list(continuous = "smooth"),
       diag = list(continuous = "densityDiag")) + 
  theme(text = element_text(size = 9)) + 
  ggtitle("February")

```

### Redundancy analyses

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide"}

#### USing RDA's to see how wave parameters and temperature parameters
# influence the morphology pf kelp species

###

# rename eck_all to eck-site0

eck_site0 <- eck_all

# Subset site names
eck_site <- subset(eck_site0, select = c(1))

# Subset wave for ecklonia sites (Annual only)
eck_wave <- subset(eck_site0, select = c(27:32, 45, 46))

# standardise wave measurements
eck_wave <- eck_wave %>%
  decostand(method = "standardize")

# subset ecklonia morpho's
eck_wave_bio <- subset(eck_site0, select = c(2:11))

# standardise measurements
eck_wave_bio <- eck_wave_bio %>%
  decostand(method = "standardize")

# Force site names as column 0
eck_wave <- cbind(eck_site, eck_wave)

eck_wave <- eck_wave %>%
  remove_rownames %>%
  column_to_rownames(var = "site")

eck_wave_bio <- cbind(eck_site, eck_wave_bio)

eck_wave_bio <- eck_wave_bio%>%
  remove_rownames %>%
  column_to_rownames(var = "site")

# run RDA
eck_wave_RDA <- rda(eck_wave_bio ~ ., data = eck_wave)

```

#### Waves as a driver of _Ecklonia maxima_ morphometrics

The morphometrics of _Ecklonia maxima_ were explained more by wave parameters (75%), than temperature parameters (66%), when separate RDA’s were constructed (Fig. 11).  The first two axes for wave parameters driving _Ecklonia maxima_ morphology explained 34% of the variation. Stipe circumference was positively influenced by both mean and SD annual significant wave height as well as annual mean wave direction. There was also a negative influence by annual SD of wave period and wave direction (which correlate with one another) on stipe circumference. The primary length of _Ecklonia maxima_ was influenced positively by annual mean wind direction, and negatively explained by annual SD wind direction.

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="An RDA, represented by the first two RDA axes, depict the influence that annual wave parameters have on the morphology of _Ecklonia maxima_ sporophytes. Environmental (explanatory) variables, in this case annual wave parameters, are represented by blue vectors extending from the origin. Response variables, in this case _Ecklonia maxima_ morphologies, are representing by red points ordinated across the plane. Sites are represented in black. The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes", fig.height=8, fig.width=12, fig.pos="H"}
# Plot scaling = 2
plot(eck_wave_RDA, scaling  = 2)

# Summary
summary(eck_wave_RDA)
```

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap=""}
## Temp

# Subset temps for ecklonia sites (for seasons)
eck_temp_season <- subset(eck_site0, select = c(17:26))

# standardise wave measurements
eck_temp_season <- eck_temp_season %>%
  decostand(method = "standardize")

# Force site names as column 0
eck_temp_season <- cbind(eck_site, eck_temp_season)

eck_temp_season <- eck_temp_season %>%
  remove_rownames %>%
  column_to_rownames(var = "site")

# run RDA
eck_temp_season_RDA <- rda(eck_wave_bio ~ ., data = eck_temp_season)

```

#### Temperature as a driver of _Ecklonia maxima_ morphometrics

Although temperature parameters for _Ecklonia maxima_ showed some influence of the morphometric data, the explanation of morphology was not strong, with an R^2^ value of 0.662 and a negative adjusted R^2^ value (-0.138)(Fig. 12). Adjusted R^2^ values are modified R^2^ values to include the number predictors in a model. A negative adjusted R squared values equate to low explanatory power.  This is supported by the lack of relationships between environmental (temperature) and response (morphology) variables (Fig. 12).

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="An RDA, represented by the first two RDA axes, depict the influence that seasonal temperature parameters have on the morphology of _Ecklonia maxima_ sporophytes. Environmental (explanatory) variables, in this case seasonal temperature parameters, are represented by blue vectors extending from the origin. Response variables, in this case _Ecklonia maxima_ morphologies, are representing by red points ordinated across the plane. Sites are represented in black. The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.height=8, fig.width=12, fig.pos="H"}
# Plot scaling = 2
plot(eck_temp_season_RDA, scaling = 2)

# Summary
summary(eck_temp_season_RDA)

```

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide"}
###### Laminaria

# rename lam_all to lam_site

lam_site0 <- lam_all

# Subset site names
lam_site <- subset(lam_site0, select = c(1))

# Subset wave for Lam sites (Annual only)
lam_wave <- subset(lam_site0, select = c(26:31, 44, 45))

# standardise wave measurements
lam_wave <- lam_wave %>%
  decostand(method = "standardize")

# subset Lam morpho's
lam_wave_bio <- subset(lam_site0, select = c(2:10))

# standardise measurements
lam_wave_bio <- lam_wave_bio %>%
  decostand(method = "standardize")

# Force site names as column 0
lam_wave <- cbind(lam_site, lam_wave)

lam_wave <- lam_wave %>%
  remove_rownames %>%
  column_to_rownames(var = "site")

lam_wave_bio <- cbind(lam_site, lam_wave_bio)

lam_wave_bio <- lam_wave_bio%>%
  remove_rownames %>%
  column_to_rownames(var = "site")

# run RDA
lam_wave_RDA <- rda(lam_wave_bio ~ ., data = lam_wave)

```
#### Waves as a driver of _Laminaria pallida_ morphometrics

For _Laminaria pallida_, the percentage of explained data by wave parameters was 89%, compared to temperature parameters explaining 82% (Fig. 13). By using adjusted R^2^ values, the first two axes for wave parameters driving _Laminaria pallida_ morphology explained 57% of the variation of morphology in this species. An increase in annual SD of wave period and wave direction was a strong influence on lamina thickness of _Laminaria pallida_. Annual mean wave period (which negatively correlates with both annual SD of wave period and wave direction) saw a strong positive influence on lamina length, and a negative influence on lamina thickness. Stipe mass of _Laminaria pallida_ was influenced by annual mean wind direction, as well as annual mean significant wave height, wave direction and SD of significant wave height. The Total length of _Laminaria pallida_ specimens were similarly explained by annual mean significant wave height, wave direction and SD of significant wave height.

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="An RDA, represented by the first two RDA axes, depict the influence that annual wave parameters have on the morphology of _Laminaria pallida_ sporophytes. Environmental (explanatory) variables, in this case annual wave parameters, are represented by blue vectors extending from the origin. Response variables, in this case _Laminaria pallida_ morphologies, are representing by red points ordinated across the plane. Sites are represented in black. The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.height=8, fig.width=12, fig.pos="H"}
# Plot scaling = 2
plot(lam_wave_RDA, scaling  = 2)

# Summary
summary(lam_wave_RDA)

```

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide"}
## Temp

# Subset temps for Lam sites (seasons)
lam_temp_season <- subset(lam_site0, select = c(16:25))

# standardise wave measurements
lam_temp_season <- lam_temp_season %>%
  decostand(method = "standardize")

# Force site names as column 0
lam_temp_season <- cbind(lam_site, lam_temp_season)

lam_temp_season <- lam_temp_season %>%
  remove_rownames %>%
  column_to_rownames(var = "site")

# run RDA
lam_temp_season_RDA <- rda(lam_wave_bio ~ ., data = lam_temp_season)
```

#### Temperature as a driver of _Laminaria pallida_ morphometrics

Provided by adjusted R^2^ values, the first two axes of temperature effect on _Laminaria pallida_ morphologies amounted to only 11%, but exhibited a stronger influence than on _Ecklonia maxima_ (Fig. 14). The mean and minimum temperatures during February were observed to negatively influence _Laminaria pallida_ morphologies such as stipe mass, stipe length and stipe diameter, as well as thallus mass and total length. Maximum august temperatures however were observed to positively explain stipe diameter of _Laminaria pallida_.

```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide", fig.cap="An RDA, represented by the first two RDA axes, depict the influence that seasonal temperature parameters have on the morphology of _Laminaria pallida_ sporophytes. Environmental (explanatory) variables, in this case seasonal temperature parameters, are represented by blue vectors extending from the origin. Response variables, in this case _Laminaria pallida_ morphologies, are representing by red points ordinated across the plane. Sites are represented in black. The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.height=8, fig.width=12, fig.pos="H"}

# Plot scaling = 2
plot(lam_temp_season_RDA, scaling = 2)

# Summary
summary(lam_temp_season_RDA)

```

\newpage
## Discussion

Kelps, in particular brown seaweeds, are robust and resilient organisms that have been shown to adapt their morphology to suit local environmental conditions. Wave exposure has been shown to be the main driver of morphological variation in brown seaweeds as a strategy to reduce drag and ultimately dislodgement. Although wave exposure is regarded as an important abiotic variable that drives kelp morphology it does not act independently. For instance, a study by @Wernberg2005 investigated the effect of wave exposure on the morphology of _Ecklonia radiata_ of six locations along 1100 km of the southwest Australian coastline. The authors concluded that wave exposure was the main driver of kelp morphological adaption in the study, however the authors also noted the variation in morphology between sites was not consistent and concluded that wave parameters may play different roles in determining kelp morphology. The results from this current study confirm the idea that kelp morphology is driven by specific wave parameters, particulary significant wave height, wave period and wave direction.

Seasonal variations in significant wave height (Hs), wave period and wave direction were observed from the data (Fig. 4, 5). The direction of swell swings to the south west in winter, generated by strong low pressures that originate from the southern ocean [@Reason2006]. These swells were found to correlate with increased wave period (Tp), wave direction and Hs. False Bay is therefore shielded by these south westerly swells by the Cape Peninsula [@Shipley1964]. In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an explanation for increased variability of Hs and Tp. The wind speed and wind direction support the presence of upwelling in summer, where southerly winds blow parallel to the coast and trigger upwelling [@Field1980]. This is supported by the decreased in temperature in summer (Fig. 2, 3), specifically along the west side of Cape Point. We however see a differentiation of median wind directions in summer, with the western coast of False Bay experiencing south easterly winds, compared to the west side of Cape Point experiencing south westerly winds. It is hypothesised that the topography and elevation along the Cape Peninsula channels, shields winds along the strip of land. This is however absent in winter, where strong northerly winds are experiences from St. Helena Bay to Betty’s Bay [@Field1980]. St. Helena bay experiences decreased median wave directions, and is protected by a headland, similar to the sites found along the western side of False Bay, that are shielded by the Cape Peninsula. There is however increased variability in the wave direction from both False Bay sites and St. Helena Bay, which is encouraged by refraction of waves.

Morphological adaption due to water motion may manifest itself in a number of ways to high wave energy environments. For instance, reduction of blade thickness, blade elongation, increase of stipe length, and stipe circumference increase and force of attachment. Although this study did not measure force of attachment, other morphological respnses to wave parameters was evident. An increase in mean annual wave period saw an increase of lamina length of _Laminaria pallida_, but a decrease in lamina thickness (Fig. 13). By decreasing the thickness of the lamina, and increasing the length of the lamina (directly increasing surface area), a larger more flexible kelp can survive in environments with greater wave period. Increased thickness of cortical tissues within the lamina aid photosynthetic ability [@Aminina2012], however increased wave exposure deceases the boundary diffusion layer [@Raven1981; @Hurd2011]. Therefore _Laminaria pallida_ in increased wave period sites may be able to reduce the need for thick lamina for photosynthesis, as the increased wave period provide longer wave events, which decreases the diffusion boundary layer and allow easier nutrient uptake. Conversely we see that sites with reduced wave height and wave period, such as Baboon Rock, Miller’s Point, A-Frame and Roman Rock displayed the greatest lamina thickness, with better photosynthetic ability, for _Laminaria pallida_. Pace [@Pace1972] however studied the effect of wave exposure on _Macrocystis pyrifera_ morphologies, and found that lamina thickness had a strong positive relationship with wave exposure. This may be species specific, with _Macrocystis pyrifera_ attaining larger sizes, but with incomparable stipe and lamina morphologies to _Ecklonia maxima_ and _Laminaria pallida_.

Lamina thickness showed positive correlation with annual SD wave period (Fig. 13). Although this variable may just negatively correlate with annual mean wave period, another reason may exist. The greatest variation of wave period and wave direction occurs in August along the western side of False Bay. Kelps at these sites may have developed thicker lamina, with a possible strategy of maximising photosynthesis at the expense of maybe being dislodged through rare increased wave energy events. The ‘spreading out’ of the lamina may also suggest a ‘go with the flow’ tactic, to reduce breaking by being less stiff [@Friedland1995]. @Denny1997 explored the stress forces that wave period played on _Nereocystis luetkeana_ by exposing various kelp plant sizes to a range of wave period that would be found in coastal environments. Results showed that for shorter kelps (smaller stipe lengths), the maximum stress on a plant occurs at 10s wave periods, but only at 5.5s for larger kelps that are similar in length to the water depth.  Similar results were observed for kelps with larger blades, with maximal stress forces occurring at approximately 8s, compared to a stress ratio of only 0.4 at 16s wave periods.

Mean annual significant wave height (Hs) was observed to influence _Ecklonia maxima_ stipe circumference, with increased stipe circumference at larger Hs (Fig. 11). An increase in Hs saw greater wave actions as waves are larger, as well as greater wave periods that are positively correlated with Hs. An increase in stipe circumference provides _Ecklonia maxima_ with a more rigid structure to withstand the onslaught of waves.  These greater stipe circumferences of _Ecklonia maxima_ were observed at west side sites such as Soetwater and Kommetjie. While these sites experience increased Hs, they also are subject to seasonal upwelling events [@Andrews1980], that bring with it cold clean water. _Ecklonia maxima_ are thought to be influenced greatly by light attenuation [@Rothman2017], and thus have developed increased (stronger) stipe circumferences to be able to cope in large Hs environments, to access cleaner, nutrient-filled water. The same patterns are observed in _Laminaria pallida_ morphologies, that display increased stipe diameters at Soetwater and Kommetjie as well (Fig. 13).  This increase in stipe diameter correlates with stipe mass of _Laminaria pallida_, providing evidence that this species has developed thicker, strong stipes to persist in this region. Denny et al. (1997) found that the stress ratio decreased as stipe lengths increases, in a fixed depth. Therefore adult _Ecklonia maxima_ sporophytes would experience a lower stress ratio compared to adult _Laminaria pallida_ sporophytes in the same environment, where _Laminaria pallida_ is often a subsurface species to _Ecklonia maxima_.

Stipe circumference for _Ecklonia maxima_ were explained by annual wave direction (Fig. 11).  For _Ecklonia maxima_, the sporophytes may be producing stipes that have a greater circumference to cope with increased wave energy. The west side of the peninsula is more exposed to significant wave heights during winter due to the direction of swell, while other sites such as Betty's Bay and Basata rock are more protected due to their location along the coast and the presence of headlands which refract wave parameters.

Temperature parameters were not found to be a significant driver of kelp morphology in this study. Although temperature plays an important role in distribution of kelp and physiological functioning of adults and gametophytes, there is little evidence that temperature is a driver of morphological variation in kelps. Although turbidity is not a direct wave or temperature parameter, turbidity is influence by various abiotic variables. Increased swells around sandy beaches increase sedimentation and turbidity [@Yang2005], however wind direction and speed can encourage upwelling, with brings to the surface cold, clean, nutrient-rich water [@Boyd1998]. Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events [@Boyd1998]. We however see a lack of upwelling events for sites such as Betty’s Bay and Batsata Rock. These two sites display _Ecklonia maxima_ sporophytes with the largest total lengths (Fig. 7), as well as Betty’s Bay displaying the largest total length of _Laminaria pallida_ sporophytes(Fig. 6). The effect of depth on stipe length is regarded as insignificant in this study as all kelp were sampled at simialr depths. Stipe elongation has been shown to be a morphological adaption in wave exposed environment, as a longer kelp has time to extend back and forth in swell reducing the tension on the holdfast. Stipe lengths of _Ecklonia maxima_ from Kommetjie, Hout Bay and Soetwater are similar to Batsata rock and Betty’s Bay, but differs to west side neighbours such as Oudekraal and Scarborough (Fig. 7). Oudekraal and Scarborough experience more variability in wave direction, however they are observed to experience smaller annual mean significant wave heights, when compared to Kommetjie, Hout Bay, Soetwater, as well as Batsata Rock and Betty’s Bay (Fig. 4).@Friedland1995 found that longer kelps were able to reduce drag and tension as the sinusoidal displacement of most waves would not ‘stretch’ a kelp out long enough before falling. The ability to grow longer stipes would provide the kelp with a structure that could surpass most wave heights without the risk of dislodgement.

### 2018-08-30 func/load.wave.R

@@ -1,0 +1,45 @@
+############################################################################
+### "func/load.model.R"
+## This script calculates R2 values between all in situ and modelled sites
+# 1. Load all libraries and functions used in this script
+# 2. Create csv definition for model data
+# 3. Create a load function
+############################################################################
+
+
+# 1. Load all libraries and functions used in this script  ----------------
+
+# library(doMC); doMC::registerDoMC(4)
+library(tidyverse)
+library(lubridate)
+
+
+# 2. Create csv definition for model data ---------------------------------
+
+read.model <-function(file_dir){
+
+  # Determine site names
+  site_type <- sapply(strsplit(as.character(file_dir), "/"), "[[", 3)
+  site_num <- sapply(strsplit(as.character(file_dir), "/"), "[[", 4)
+  site_num <- sapply(strsplit(site_num, ".txt"), "[[", 1)
+  site_name <- paste(site_type, site_num, sep = "")
+
+  # Load and process site data
+  df <- read.csv(file_dir, col.names = c("date", "hs", "tp", "dir", "dirw", "spw"), sep = "", header = F) %>%
+    mutate(date = as.POSIXct(as.character(date), "%Y%m%d%H%M", tz = "Africa/Johannesburg")) %>%
+    mutate(site = site_name) %>%
+    select(site, everything())
+
+  # Finish
+  return(df)
+}
+
+
+# 3. Create a load function -----------------------------------------------
+
+load.model <- function(directory){
+  files_list <- dir(directory, full.names = TRUE)
+  dat <- plyr::ldply(files_list, read.model, .progress = "text")
+  return(dat)
+}
+

### 2018-08-30 LaTeX/kelp_morph.tex
more additions...

\documentclass[utf8]{frontiersSCNS} % for Science, Engineering and Humanities
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

%%% BEGIN added by AJS
\usepackage{xltxtra} % also loads 'fontspec'
\defaultfontfeatures{Ligatures=TeX}
\setmainfont{Minion Pro}
\setsansfont[Scale=MatchLowercase]{Myriad Pro}
\setmonofont[Scale=MatchLowercase]{Ubuntu Mono}
% \setmathfont{Minion Math}

\frenchspacing
\usepackage{textgreek}
\usepackage{ragged2e} % for '\RaggedRight' macro (allows hyphenation)
\usepackage{booktabs} % nice tables without vertical lines
\setlength\heavyrulewidth{0.1em}
\setlength\lightrulewidth{0.0625em}
\usepackage{siunitx}
    \sisetup{%
        detect-mode,
        group-digits            = false,
        input-symbols           = ( ) [ ] - + < > * §,
        table-align-text-post   = false,
        round-mode              = places,
        round-precision         = 3
        }

        % SECTION, SUBSECETC.TITLES
        \usepackage[compact]{titlesec}
        \titleformat{\chapter}
          {\normalfont\LARGE\sffamily\bfseries}
          {\thechapter}
          {1em}
          {}
        \titleformat{\section}
          {\normalfont\LARGE\sffamily\bfseries}
          {\thesection}
          {1em}
          {}
        \titleformat{\subsection}
          {\normalfont\Large\sffamily\bfseries}
          {\thesubsection}
          {1em}
          {}
        \titleformat{\subsubsection}
          {\normalfont\large\sffamily\bfseries\slshape}
          {\thesubsubsection}
          {1em}
          {}
        % \titlespacing*{<command>}{<left>}{<before-sep>}{<after-sep>}
        \titlespacing*{\chapter}
          {0pt}
          {1.2ex plus 1ex minus .2ex}
          {0.5ex plus .1ex minus .1ex}
        \titlespacing*{\section}
          {0pt}
          {1.2ex plus 1ex minus .2ex}
          {0.5ex plus .1ex minus .1ex}
        \titlespacing*{\subsection}
          {0pt}
          {1.2ex plus 1ex minus .2ex}
          {0.5ex plus .1ex minus .1ex}
        \titlespacing*{\subsubsection}
          {0pt}
          {1.2ex plus 1ex minus .2ex}
          {0.5ex plus .1ex minus .1ex}
 %%% END added by AJS

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Smit {et~al.}} %use et al only if is more than 1 author
\def\Authors{Albertus J. Smit\,$^{1,2,*}$, Ross C. Coppin\,$^{1}$, Robert W. Schlegel\,$^{1}$ and Christo Rautenbach\,$^{3}$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa \\
$^{2}$South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa \\
$^{3}$South African Weather Service, Marine Section, Cape Town, South Africa }
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Albertus J. Smit}

\def\corrEmail{ajsmit@uwc.ac.za}

\begin{document}
\onecolumn
\firstpage{1}

\title[The shape of kelp]{The shape of kelp: environmental influences on kelp morphology}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}

%%% Leave the Abstract empty if your article does not require one, please see the Summary Table for full details.
\section{}
Two species of kelp, \emph{Ecklonia maxima} and \emph{Laminaria pallida}, dominate the nearshore environment around the South African Western Cape coast. These structure producing kelps provide habitat for various organisms, and are distributed across a region where the thermal and wave energy regimes vary significantly. Across this distribution, the morphology of each species displays a disparity in shape and size. The morphology of macroalgae is known to be influenced by wave exposure, however waves are made up of various constituents, such as wave height, wave direction and period. Not much is understood on specific wave parameters and their influence on macroalgal morphology. We therefore used existing thermal and wave energy data for 18 sites across the Western Cape coast, and investigated how individual abiotic parameters influence the morphology of each species. While temperature influences kelp distributions most strongly, localised variation in the morphology of kelps are driven by wave parameters. The biggest influences were found to be wave direction, significant wave height and wave period, which explain much of the total length, lamina thickness and stipe circumference of each species. Turbidity correlates with wave energy, where sedimentation of the water column through excessive wave movements reduces light penetration. The expansion of \emph{Laminaria pallida} lamina area was found in high turbidity environments. These sites also displayed decreased lamina thickness (which positively correlates with photosynthetic ability). \emph{Laminaria pallida} sporophytes in highly turbid environments, where the boundary diffuse layer is reduced and nutrient uptake is easier, are thought to rely on greater light-capturing surface area instead of photosynthetic ability. We can therefore see how specific wave parameters are able to influence the morphology of kelps, and confirms the idea of morphoplasticity in South African kelp species.

\tiny
 \keyFont{ \section{Keywords:} kelp, \emph{Ecklonia maxima}, \emph{Laminaria pallida}, morphometrics, wave energy, seawater temperature}
\end{abstract}

\section{Introduction}
Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of \num{112} species in \num{33} genera \citep{bolton2010}, nevertheless form the basis of one of the most productive ecosystems globally \citep{mann1973}. Kelps generally have a dependence on cool, temperate and arctic seawater temperatures \citep{santelices2007,bolton2010}, and dominate the nearshore biomass in these regions within the rocky shallow coasts in both hemispheres \citep{steneck2002}. Outside of these latitudes, they are also found in cool, deep water towards the tropics \citep{graham2007}: this deeper tropical water, due to its water clarity, also allows for a minimum photon flux rate of \SI{50}{\micro\mol\per\meter\squared\per\second} and a nitrate concentration of \SI{>2}{\micro\mole\per\liter} \citep{luning1990,dayton1999,zimmerman1985}, the two other environmental variables that support kelp productivity.

[Say something about the thermal range in these parts of the world---from Lüning and from a plot of geographical distribution of kelps; Graham et al., 2007 set a 23°C maximum monthly climatological temperature to constrain the upper thermal limit...]

[Say something about their size range across the species...]

Although kelps have a low taxonomic diversity, their size and complex morphology provide a heterogeneous habitat structure \citep{steneck2002} that accommodate a multitude other turf and sub-canopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates \citep{mann1973,dunton1987,duggins1989,steneck2002}, each depending on a wide suite of ecological services provided by the kelp habitat \citep{gaines1987,bologna1993,levin1994,anderson1997}.

[Say something about kelp habitats that have a physical influence on the hydrodynamics within the beds...]

The biogeographic distribution of kelp is limited by seawater temperature (Bolton 2010), where increasing temperature gradients reduce kelp distribution. Due to this limiting factor, the two main species of kelps in southern African waters, \emph{Ecklonia maxima} and \emph{Laminaria pallida}, are distributed along a section of the south coast from De Hoop, extending west around the Cape Peninsula, and thriving north into Namibia (Molloy 1990, Stegenga et al. 1997). This distribution follows a temperate temperature gradient, where sea temperatures increase as one moves south from Namibia, around Cape Point and towards De Hoop. Although the two species occur together for the majority of the coast, their basic morphologies and resource needs vary to a degree. The larger species, \emph{Ecklonia maxima}, is distributed from Lüderitz to Cape Agulhas (Bolton and Levitt 1985, Probyn and McQuaid 1985, Bolton and Anderson 1987, Bolton et al. 2012). Characterised by a large distal swollen bulb filled with gas, and smooth fronds, this species grows to approximately 10 meters (Bolton and Anderson 1987). There was, however, a 17-meter specimen collected in 2015 off Cape Point (Smit, unpubl. data). This species of kelp not only dominate the biomass of the South African nearshore, but plays an important ecological role (Bustamante and Branch 1996). The estimated productivity of \emph{Ecklonia maxima} within South Africa varies between 350 and 1500g C m\textsuperscript{-2yr-1} (Mann 1982).

Across the majority of the coastline, \emph{Laminaria pallida} remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 meters (Field et al. 1980, Bolton and Anderson 1987, Molloy and Bolton 1996). This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 meters (Field et al. 1980, Molloy 1990, Molloy and Bolton 1996, Stegenga et al. 1997). Towards the north along the west coast, from around Hondeklipbaai, \emph{Laminaria pallida} replaces \emph{Ecklonia maxima} as the dominant kelp species (Velimirov et al. 1977, Stegenga et al. 1997) and it also occupies increasingly shallower subtidal regions. The northern populations also exhibit an increase in stipe hollowness, compared to the solid stipe morphs in the species' southern distributions (Molloy and Bolton 1996). This variation in morphology was thought to represent two distinct species, with the northern populations formerly described as \emph{Laminaria schinzii} Foslie (Molloy and Bolton 1996). Genetic work has subsequently shown that the two morphs are in fact the same species (Rothman et al. 2017). In southern African waters, the primary production of \emph{Laminaria pallida} is between 120 and 1900g C m\textsuperscript{-2yr-1}, similar to that of \emph{Ecklonia maxima} (Mann 1982). Primary production is not the only pathway by which these two species influence the environment. Fresh and detrital matter from these kelps not only provide support for organisms in the immediate sandy and rocky shores of South Africa, but connect to distant ecosystems (Krumhansl and Scheibling 2012).

\section{Methods}

\section{Results}

\subsection{Conclusions}

\section*{Conflict of Interest Statement}
The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}

\section*{Funding}
The research was partly funded by the South African National Research Foundation (http://www.nrf.ac.za) programme ``Thermal characteristics of the
South African nearshore: implications for biodiversity'' (CPRR14072378735).

Aside from providing funding, the funders had no role in the study design, data collection and analysis, decision to publish, or preparation of the manuscript.

\section*{Acknowledgements}
Robert Schlegel is thanked for his assistance in the preparation of some of Figure 1.

\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
\bibliography{kelp_morph}

%%% Make sure to upload the bib file along with the tex file and PDF
%%% Please see the test.bib file for some examples of references

\clearpage
\nolinenumbers

\end{document}

### 2018-08-30 chapter_1.Rmd
more additions...

@@
 ---
-title: "Kelp morphometric properties and its link to the hydrodynamic modelling"
+title: "Kelp morphometric properties and the link to hydrodynamic modelling"
 author: "Ross Coppin"
@@
 
-In recent years there has been growing attention on microplastics and the deterimental role they play in the enviornment, particulary the ocean.
-Microplastics are tiny plastic granuales used as scrubbers in cosmetics and air-blasting, and small plastic fragments that originate from larger pieces of plastic known as macroplastics.
-The potential harms of small pieces of plastic was highlighted in the 1970's and renewed interest has lead to research showing that microplastics in the ocean are widespread.
-The small size of microplastics makes them bio-available to biota in the food-web which may cause problems with an organisms physiological functioning.
+In recent years there has been growing attention on plastic pollution, particulary in the ocean.
+Plastic pollution can be in the form of macro- and microplastics.
+Microplastics are tiny plastic granuales used as scrubbers in cosmetics and air-blasting, and small plastic fragments that originate from larger pieces of plastic known as macroplastics, while macroplastics...**insert definition here**...
+The potential harms of  of plastic pollution in the marine environment was highlighted in the 1970's and renewed interest has lead to research showing that plastic pollution in the ocean are widespread.
+Plastics may become bio-available to biota in the food-web which may cause problems with an organisms physiological functioning.
 Furthermore, the relatively large surface area and composition of microplastics provides an environmeht that is able of adhering to organic pollutants.
 In other words microplastics also act as a vector for transport and assimilation of organic polluants.
-Therefore, microplastics may also aid in the assimilation of organic pollutans into all levels of the food web.
 
@@
 @Bolton1985 showed that under sub-saturating irradiances and supra- optimal temperatures *Ecklonia maxima* to showed a decrease in reproductive rates and an increase in cell production.
-A interesting ﬁnding of this study was that despite the decrease in reproductive rates, the ﬁnal egge production per female was greater under these conditons.
+An additional ﬁnding of this study was that despite the decrease in reproductive rates, the ﬁnal egge production per female was greater under these conditons.
 The authors interpreted this an ecological adaption that may increase survival rates under times of stress or non - ideal conditions [@Bolton1985].
+
+#Depth
+
+Depth does not aﬀect kelp ecosystems directly, however a change in depth often causes ﬂuctuations or changes in other environmental variables such as water motion, light and temperature.
+Water motion also decreases with depth, and some kelps better suited to deeper environments (*L.
+pallida*) replace those in the shallows (*E.
+maxima*) (Dayton 1985; Gerard 1982).
+The increase in depth can lead to a decrease in sunlight penetration, with some species better adapted for low-light conditions than others, such as (*L.
+pallida*).
+Temperature may also change along a depth gradient due to a reduction in sunlight penetration.
+(Dayton 1985; Gerard 1982).
+Therefore depth does not directly play a role in kelp functioning but may alter more inﬂuential factors such as light and water motion.
 
@@
 This is often an indication of an “up-welling” event, which brings cold and nutrient rich waters from the bottom to the surface of the water column.
-However temperature can play a direct role in the uptake of nutrients through eﬀects on algal metabolism; however this may vary from species to species [@Raven1988].
-However most studies investigate multiple characteristics in one study (i.e.
-temperature and light and water motion), while studies investigating the direct role of temperature in nutrient assimilation is lacking in marine algae is lacking.
-
-#Depth
-
-Depth does not aﬀect kelp ecosystems directly, however a change in depth often causes ﬂuctuations or changes in other environmental variables such as water motion, light and temperature.
-Water motion also decreases with depth, and some kelps better suited to deeper environments (*L.
-pallida*) replace those in the shallows (*E.
-maxima*) (Dayton 1985; Gerard 1982).
-The increase in depth can lead to a decrease in sunlight penetration, with some species better adapted for low-light conditions than others, such as (*L.
-pallida*).
-Temperature may also change along a depth gradient due to a reduction in sunlight penetration.
-(Dayton 1985; Gerard 1982).
-Therefore depth does not directly play a role in kelp functioning but may alter more inﬂuential factors such as light and water motion.
+Temperature can play a direct role in the uptake of nutrients through eﬀects on algal metabolism; however this may vary from species to species [@Raven1988].
 
@@
 The study looked at resilience of kelp beds along a latitudinal temperature gradient.
-Kelp abundance is likely to decline with the predicted warming of ocean waters @Wernberg2010 and although kelps have the ability to acclimatize and adjust their metabolic performance, which in turn allows them to change their physiological performance to mitigate the seasonal ﬂuctuations in temperature, this acclimatization is done at a cost @Wernberg2010...**link to paragrapgh on kelp morphology**...
+Kelp abundance is likely to decline with the predicted warming of ocean waters @Wernberg2010 and although kelps have the ability to acclimatize and adjust their metabolic performance, which in turn allows them to change their physiological performance to mitigate the seasonal ﬂuctuations in temperature, this acclimatization is done at a cost @Wernberg2010...**link to paragraph on kelp morphology**...
 

### 2018-09-06 chapter_2.Rmd
updated rmd file added

@@
 ---
-title: 'Chapter 2:The shape of kelp: environmental influences on kelp morphology'
-author: "Ross Coppin"
-date: "30 August 2018"
+title: "The shape of kelp: environmental influences on kelp morphology"
 output:
+  word_document:
+    fig_caption: yes
+    fig_height: 7
+    fig_width: 7
+    highlight: zenburn
+    toc: no
+  html_notebook:
+    fig_caption: yes
+    toc: yes
   pdf_document:
-    highlight: zenburn
-  word_document: default
+    toc: yes
+  bookdown::html_document2:
+    fig_caption: yes
+    fig_retina: 2
+    fig_width: 7
+    highlight: espresso
+    theme: sandstone
+    toc_float: no
+fontsize: 12pt
+geometry: margin=1in
+language: Australian
+mainfont: PT Serif
+monofont: PT Mono
+csl: LaTeX/ecography.csl
+papersize: A4
+sansfont: PT Sans
+tables: yes
+bibliography: LaTex/mycollection.bib
 ---
+
+\newpage
 ## Introduction
@@
 Kelps generally have a dependence on cool, temperate and arctic seawater temperatures [@Santelices2007; @Bolton2010], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
-Outside of these latitudes, they are also found in cool, deep water towards the tropics [@Graham2007]: this deeper tropical water, due to its water clarity, also allows for a minimum photon flux rate of 50 ???
-and a nitrate concentration of ???
-[@Luning1990; @Dayton1999; @Zimmerman1985], the two other environmental variables that support kelp productivity.
+Outside of these latitudes, they are also found in cool, deep water towards the tropics [@Graham2007]: this deeper tropical water, due to its water clarity, also allows for suitable photon flux and a nitrate concentration [@Luning1990; @Dayton1999; @Zimmerman1984], the two other environmental variables that support kelp productivity.
 Although kelps have a low taxonomic diversity, their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude other turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Dunton1987; @Duggins1989; @Steneck2002], each depending on a wide suite of ecological services provided by the kelp habitat [@Gaines1987; @Bologna1993; @Levin1994; @Anderson1997].
@@
 Although the two species occur together for the majority of the coast, their basic morphologies and resource needs vary to a degree.
-The larger species, _Ecklonia maxima_, is distributed from Lüderitz to Cape Agulhas [@Bolton1985; @Probyn1985; @Bolton1987; @Bolton2012].
+The larger species, _Ecklonia maxima_, is distributed from Lüderitz to Cape Agulhas [@Bolton1985a; @Probyn1985; @Bolton1987; @Bolton2012].
 Characterised by a large distal swollen bulb filled with gas, and smooth fronds, this species grows to approximately 10 meters [@Bolton1987].
@@
 The estimated productivity of _Ecklonia maxima_ within South Africa varies between 350 and 1500g Cm-2yr-1 [@Mann1982].
-Across the majority of the coastline, _Laminaria pallida_ remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 meters [@Field1980; @Bolton1987; @Molloy1996].
-This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 meters [@Field1980; @Molloy1990; @Molloy1996; @Stegenga1997].
+Across the majority of the coastline, _Laminaria pallida_ remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 meters [@Field1980b; @Bolton1987; @Molloy1996].
+This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 meters [@Field1980b; @Molloy1990; @Molloy1996; @Stegenga1997].
 Towards the north along the west coast, from around Hondeklipbaai, _Laminaria pallida_ replaces _Ecklonia maxima_ as the dominant kelp species [@Velimirov1977; @Stegenga1997] and it also occupies increasingly shallower subtidal regions.
@@
 Wave energy data formed part of the South African Coastal Vulnerability Assessment, presented to the Department of Environmental Affairs [@Theron2014].
-These data are first forecasted using NOAA Wave Watch III (WWIII), with National Centers for Environmental Predictions (NCEP) product as the numerical input [@benassai2006implementation].
+These data are first forecasted using NOAA Wave Watch III (WWIII), with National Centers for Environmental Predictions (NCEP) product as the numerical input [@benassai2006].
 Hindcast data from WWIII span from 1994-2013 at a 3-hour resolution.
-The data are then used to model short –crested waves generated by the wind into the coastal environment, using Simulating Waves in the Nearshore [SWAN; @booij1997swan].
+The data are then used to model short –crested waves generated by the wind into the coastal environment, using Simulating Waves in the Nearshore [SWAN; @booij1996].
 SWAN allows one to extract wave parameters from specific gridded locations in the nearshore.
@@
   # NB: This may also not be an appropriate change
-  mutate(site = replace(site, site == "Bordjiestif North", "Bordjies Deep")) %>%
+  mutate(site = replace(site, site == "Bordjies reef North", "Bordjies Deep")) %>%
   # Not using Rocky Bank in this study
@@
   mutate(date = lubridate::month(date, label = TRUE)) %>% 
-  filter(site!= "FB25") %>% # FB_25 has issues with 'hs' and 'tp'
+  filter(site != "FB25") %>% # FB_25 has issues with 'hs' and 'tp'
   group_by(site, date) %>% 
@@
   coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
-  facet_wrap(~ season, labeller = as_labeller(season), scales = "free", ncol = 3) +
+  facet_wrap(~ season, labeller = as_labeller(season), ncol = 3) +
   theme(axis.title = element_blank(),
@@
   coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
-  facet_wrap(~ season, labeller = as_labeller(c(season)), scales = "free", ncol = 3) + 
+  facet_wrap(~ season, labeller = as_labeller(c(season)), ncol = 3) + 
   theme(axis.title = element_blank(),
@@
   coord_fixed(xlim = c(17.85, 19), ylim = c(-32.75, -34.4)) +
-  facet_wrap(~ season, labeller = as_labeller(c(season)), scales = "free", ncol = 3) +
+  facet_wrap(~ season, labeller = as_labeller(c(season)), ncol = 3) +
   theme(axis.title = element_blank(),
@@
 Not only do the temperature parameters display seasonal fluctuations, but we see similar observations for wave parameters.
-The west side of Cape Point exhibit increased measurements of median direction, SD and mean significant wave height compared to within False Bay (Fig.
+The west side of Cape Point exhibit increased median direction, SD and mean significant wave height compared to within False Bay (Fig.
 4).
@@
 Lamina length for laminaria pallida showed no geographical pattern moving from west to east.
-Kommetjie, Olifantsbos and Batsata Rock showed great variability in lamina length, and both Buffels Bay and Betty’s Bay were visually different to Miller’s Point and Roman Rock, when comparing the summary data represented by the boxplots (Fig.6).
+Kommetjie, Olifantsbos and Batsata Rock showed great variability in lamina length, and both Buffels Bay and Betty’s Bay were visually different to Miller’s Point and Roman Rock, when comparing summary data (Fig.6).
 Lamina thickness showed great variation across sites, with Baboon Rock, Miller’s Point, A-Frame and Roman rock displaying large lamina thickness, significantly different to the rest of the sites.
@@
 An increase in the number of digits was observed as one moved from Cape Point north along the western side of False Bay.
-This ceased at Batstata Rock, which exhibited significantly less digits compared to the previous site, Bordjiestif North.
+This ceased at Batstata Rock, which exhibited significantly less digits compared to the previous site, Bordjies reef North.
 Stipe diameter showed some geographical grouping, with west of Cape Point sites exhibiting larger stipe diameters compared to False Bay sites.
 
-Stipe diameter decreased as one rounded the point into False Bay, where a sudden, significant difference was seen between Bordjiestif North and Batsata Rock.
+Stipe diameter decreased as one rounded the point into False Bay, where a sudden, significant difference was seen between Bordjies reef North and Batsata Rock.
 Greater variation of stipe length was observed for sites outside of False Bay (Kommetjie, Olifantsbos and Betty’s Bay).
@@
 Stipe mass displays similar patterns to stipe length, with larger stipe lengths west of Cape Point compared to within False Bay.
-The thallus mass of _Laminaria pallida_ was observed to be greater for sites around Cape Point, with an observed difference between Batsata Rock and Bordjiestif North, similar to the number of digit patterns.
+The thallus mass of _Laminaria pallida_ was observed to be greater for sites around Cape Point, with an observed difference between Batsata Rock and Bordjies reef North, similar to the number of digit patterns.
 Larger total lengths were observed around Cape Point and Betty’s Bay, with the smallest total lengths at sites that exhibited the greatest lamina thicknesses (Baboon Rock, Miller’s Point, A-Frame and Roman Rock).
@@
 positionsLam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point", "Baboon Rock", 
-                  "Batsata Rock", "Bordjiestif North", "Buffels Bay South", 
+                  "Batsata Rock", "Bordjies reef North", "Buffels Bay South", 
                   "Buffels Bay", "Olifantsbos", "Kommetjie", "Sea Point")
@@
 For instance, a study by @Wernberg2005 investigated the effect of wave exposure on the morphology of _Ecklonia radiata_ of six locations along 1100 km of the southwest Australian coastline.
-The authors concluded that wave exposure was the main driver of kelp morphological adaption in the study, however the authors also noted the variation in morphology between sites was not consistent and concluded that wave parameters may play different roles in determining kelp morphology.
-The results from this current study confirm the idea that kelp morphology is driven by specific wave parameters, particulary significant wave height, wave period and wave direction.
+The authors concluded that wave exposure was the main driver of kelp morphological adaptation in the study, however the authors also noted the variation in morphology between sites was not consistent and concluded that wave parameters may play different roles in determining kelp morphology.
+The results from this current study confirm the idea that kelp morphology is driven by specific wave parameters, particularly significant wave height, wave period and wave direction.
 
@@
 In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an explanation for increased variability of Hs and Tp.
-The wind speed and wind direction support the presence of upwelling in summer, where southerly winds blow parallel to the coast and trigger upwelling [@Field1980].
+The wind speed and wind direction support the presence of upwelling in summer, where southerly winds blow parallel to the coast and trigger upwelling [@Field1980a].
 This is supported by the decreased in temperature in summer (Fig.
@@
 This is however absent in winter, where strong northerly winds are experiences from St.
-Helena Bay to Betty’s Bay [@Field1980].
+Helena Bay to Betty’s Bay [@Field1980a].
 St.
@@
 
-Morphological adaption due to water motion may manifest itself in a number of ways to high wave energy environments.
+Morphological adaptation due to water motion may manifest itself in a number of ways to high wave energy environments.
 For instance, reduction of blade thickness, blade elongation, increase of stipe length, and stipe circumference increase and force of attachment.
-Although this study did not measure force of attachment, other morphological respnses to wave parameters was evident.
+Although this study did not measure force of attachment, other morphological responses to wave parameters was evident.
 An increase in mean annual wave period saw an increase of lamina length of _Laminaria pallida_, but a decrease in lamina thickness (Fig.
@@
 6).
-The effect of depth on stipe length is regarded as insignificant in this study as all kelp were sampled at simialr depths.
-Stipe elongation has been shown to be a morphological adaption in wave exposed environment, as a longer kelp has time to extend back and forth in swell reducing the tension on the holdfast.
+The effect of depth on stipe length is regarded as insignificant in this study as all kelp were sampled at similar depths.
+Stipe elongation has been shown to be a morphological adaptation in wave exposed environment, as a longer kelp has time to extend back and forth in swell reducing the tension on the holdfast.
 Stipe lengths of _Ecklonia maxima_ from Kommetjie, Hout Bay and Soetwater are similar to Batsata rock and Betty’s Bay, but differs to west side neighbours such as Oudekraal and Scarborough (Fig.
@@
 The ability to grow longer stipes would provide the kelp with a structure that could surpass most wave heights without the risk of dislodgement.
+
+# References

### 2018-09-10 chapter_2.Rmd
minor edits and additions

@@
     fig_width: 7
-    highlight: zenburn
     toc: no

### 2018-09-10 chapter_1.Rmd
additional writing

@@
 
-#Abiotic drivers
+#Kelp environmental drivers
+
+The important environmental drivers of kelp idividuals and communities include light, substrata, salinity, sedimentation, nitrients, temperature and water motion.
+Although studies have investigated the effects of important environmental drivers, the roles these factors play is often difficult to evaluate as such factors may never be fully independent of each other, i.e.
+environmental factors are to some extent temperature dependent.
+Multifactorial studies have attempt to explain combined affects, however these studies are often limited to investigating combination of two or three environmental drivers as inclusion of too many factors can lead to results that are difficult to interept.
+Environmetal factors are highly variable on temporal and spatial scales, and their effects may also dependent on life-stage adding a further layer of complexity to investigations.
 
@@
 pallida*) replace those in the shallows (*E.
-maxima*) (Dayton 1985; Gerard 1982).
+maxima*) [@Dayton1985; @Gerard1982].
 The increase in depth can lead to a decrease in sunlight penetration, with some species better adapted for low-light conditions than others, such as (*L.
 pallida*).
-Temperature may also change along a depth gradient due to a reduction in sunlight penetration.
-(Dayton 1985; Gerard 1982).
+Temperature may also change along a depth gradient due to a reduction in sunlight penetration [@Dayton1985; @Gerard1982].
 Therefore depth does not directly play a role in kelp functioning but may alter more inﬂuential factors such as light and water motion.
@@
 Other than temperature, wave exposure is also recognised as an important driver of the marine environment.
-Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition and productivity of benthic and rocky shore communties.
+Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity of macroalgae communties.
+The width, vertical zonation and diversity of algal communities often change predictably along gradients of exposure.
+This is also true for variation in morphological characters of individuals may also vary with wave exposure, such as frond width and degree of 'branching' [@moss1948].
+@moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed).
+The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid in relation to wave exposure between study sites.
+
 For example, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake through increased turbulence in the boundry layer [@lobban1994].

### 2018-09-10 chapter_1.tex
additional writing

@@
   \usepackage[utf8]{inputenc}
+  \usepackage{eurosym}
 \else % if luatex or xelatex
@@
   \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
 \fi
@@
 \hypersetup{unicode=true,
-            pdftitle={Kelp morphometric properties and its link to the hydrodynamic modelling},
+            pdftitle={Kelp morphometric properties and the link to hydrodynamic modelling},
             pdfauthor={Ross Coppin},
@@
 
-  \title{Kelp morphometric properties and its link to the hydrodynamic modelling}
+  \title{Kelp morphometric properties and the link to hydrodynamic modelling}
     \pretitle{\vspace{\droptitle}\centering\huge}
@@
 to be an important driver of seaweed morphology, as the main mechanism
-of seaweed mortality is through the mechanism of dislodgment. Changing
-morphology, such as decreasing frond area or strengthening through stipe
-circumference increase/reduces drag and increases the probability of
-survival. However, locally adapted seaweed may still be dislodged in
-pulse disturbance events such as storms, and may raft far distances and
-wash up on beaches. An example would be \emph{Ecklonia maxima} which is
-a conspicuous brown seaweed along the coastline and beaches around South
-Africa. Therefore, the beach-cast may not always originate from adjacent
-kelp populations but rather from other regions which the individual is
-adapted for. Therefore, the kelp morphology may act as a proxy for
-investigating coastal currents and changes thereof. The advances in
-ocean hydrodynamic modelling has made great progress and has been
-applied in a variety of ways. Therefore, this study will use advances in
-hydrodynamic modelling in combination with kelp morphological
-characteristics to investigate coastal currents. Once the model has been
-established it may be applied in other ways, such as investigating the
-transport of micorplastics along the South African coastline. The harms
-of microplastics to the marine environment has gained much traction in
-recent years, but research in South Africa is lacking. The coupling of
-kelp morphology and the ability to simulate hydrodynamic processes can
-greatly improve our understanding of transport pathways and likely
-locations of accumulation. This in turn may inform management decisions
-with regards to eliminating and managing marine pollution in South
-Africa.
+of seaweed mortality is through the dislodgment. Changing morphology
+reduces drag and increases the probability of survival. However, locally
+adapted seaweed may still be dislodged in pulse disturbance events such
+as storms, and may raft far distances and wash up on beaches. An example
+would be \emph{Ecklonia maxima} which is a conspicuous brown seaweed
+along the coastline and beaches around South Africa. Therefore, the
+beach-cast may not always originate from adjacent kelp populations but
+rather from other regions which the individual is adapted for.
+Therefore, the kelp morphology may act as a proxy for investigating
+coastal currents and changes thereof. The advances in ocean hydrodynamic
+modelling has made great progress and has been applied in a variety of
+ways. Therefore, this study will use advances in hydrodynamic modelling
+in combination with kelp morphological characteristics to investigate
+coastal currents. Once the model has been established it may be applied
+in other ways, such as investigating the transport of micorplastics
+along the South African coastline. The harms of microplastics to the
+marine environment has gained much traction in recent years, but
+research in South Africa is lacking. The coupling of kelp morphology and
+the ability to simulate hydrodynamic processes can greatly improve our
+understanding of transport pathways and likely locations of
+accumulation. This in turn may inform management decisions with regards
+to eliminating and managing marine pollution in South Africa.
 
@@
 
-In recent years there has been growing attention on microplastics and
-the deterimental role they play in the enviornment, particulary the
-ocean. Microplastics are tiny plastic granuales used as scrubbers in
-cosmetics and air-blasting, and small plastic fragments that originate
-from larger pieces of plastic known as macroplastics. The potential
-harms of small pieces of plastic was highlighted in the 1970's and
-renewed interest has lead to research showing that microplastics in the
-ocean are widespread. The small size of microplastics makes them
-bio-available to biota in the food-web which may cause problems with an
-organisms physiological functioning. Furthermore, the relatively large
-surface area and composition of microplastics provides an environmeht
-that is able of adhering to organic pollutants. In other words
-microplastics also act as a vector for transport and assimilation of
-organic polluants. Therefore, microplastics may also aid in the
-assimilation of organic pollutans into all levels of the food web.
+In recent years there has been growing attention on plastic pollution,
+particulary in the ocean. Plastic pollution can be in the form of macro-
+and microplastics. Microplastics are tiny plastic granuales used as
+scrubbers in cosmetics and air-blasting, and small plastic fragments
+that originate from larger pieces of plastic known as macroplastics,
+while macroplastics\ldots{}\textbf{insert definition here}\ldots{} The
+potential harms of of plastic pollution in the marine environment was
+highlighted in the 1970's and renewed interest has lead to research
+showing that plastic pollution in the ocean are widespread. Plastics may
+become bio-available to biota in the food-web which may cause problems
+with an organisms physiological functioning. Furthermore, the relatively
+large surface area and composition of microplastics provides an
+environmeht that is able of adhering to organic pollutants. In other
+words microplastics also act as a vector for transport and assimilation
+of organic polluants.
 
@@
 
-\section{Kelp morphology drivers}\label{kelp-morphology-drivers}
+\section{Kelp environmental drivers}\label{kelp-environmental-drivers}
+
+The important environmental drivers of kelp idividuals and communities
+include light, substrata, salinity, sedimentation, nitrients,
+temperature and water motion. Although studies have investigated the
+effects of important environmental drivers, the roles these factors play
+is often difficult to evaluate as such factors may never be fully
+independent of each other, i.e.~environmental factors are to some extent
+temperature dependent. Multifactorial studies have attempt to explain
+combined affects, however these studies are often limited to
+investigating combination of two or three environmental drivers as
+inclusion of too many factors can lead to results that are difficult to
+interept. Environmetal factors are highly variable on temporal and
+spatial scales, and their effects may also dependent on life-stage
+adding a further layer of complexity to investigations.
 
@@
 of kelp {[}@Bruhn1996);@Belseth2012{]}. For instance, solar ultraviolate
-radiation has been shown to aﬀect sub-canopy Ecklonia radiata
+radiation has been shown to aï¬\euro{}ect sub-canopy Ecklonia radiata
 sporophytes when the canopy of mature \emph{Ecklonia radiata} was
 removed {[}@Wood 1987{]}. The sub-canopy sporophytes experienced tissue
-damage,photopigment destruction,reducedgrowthand decreased survivorship,
-thus inhibiting their settlement and survival {[}@Wood 1987{]}.
-Laboratory experiments revealed that the UV component of radiation,
-rather than intense radiation itself, was responsible for the eﬀects
-mentioned above. High light stress has negative eﬀects, such as
-photoinhibition and photo-damage on \emph{Ecklonia cava} sporophytes
-{[}@Altamirano2004{]}. @Altamirano2004 found that \emph{Ecklonia cava}
-is more vulnerable to light stress conditions, and less likely to
-recover under unfavourable conditions {[}@Altamirano2004{]}. @Bolton1985
-showed that under sub-saturating irradiances and supra- optimal
-temperatures \emph{Ecklonia maxima} to showed a decrease in reproductive
-rates and an increase in cell production. A interesting ﬁnding of this
-study was that despite the decrease in reproductive rates, the ﬁnal egge
-production per female was greater under these conditons. The authors
-interpreted this an ecological adaption that may increase survival rates
-under times of stress or non - ideal conditions {[}@Bolton1985{]}. Light
-is also important for survival and functioning of benthic algal species;
-a study by @Toohey2004 investiagted the possible link of light and
-thallus scour of the kelp \emph{Ecklonia radiata canopy} on the
-associated foliose algal assemblage. The authors concluded that a
-reduction in light, rather than thallus scour has an eﬀect on foliose
-algal assemblage. Reduced light was associated with reduced diveristy
-and an marked increase abundance in certian foliose species, which were
-able to acclimiate more light under low irradiances compared to other
-species {[}@Toohey2004{]}. Kelp may also compete intra-specﬁcally or
-inter-speciﬁcally for light. For instance, @Clark2004 investiagted eﬀect
-of canopy shading of \emph{Macrocystis pyrifera} and understorey canopy
-shading of \emph{Pterygophora californica} individually and in
-combination of benthic macro-algae assemblage. Under all treatments
-there was a sigiﬁance change in macro-algal assemblage, with certian
-species able to recover in abundance more readily than others. The
-authors devided these species into two categories; ``shade-tolerant''
-and ``light adapted'', and conlcuded that certian foliose species are
-better adapted for diﬀerent light conidtions in accordance with canopy
-cover. In South Africa, the eﬀect of \emph{Ecklonia maxima} harvesting
-on sub-story kelp, kelp population structure and kelp recruitement was
-investigated by @Rothman2006. The results showed no signiﬁcant diﬀerence
-in the three factors mentioned earlier and @Rothman2006 concluded that
-factors other than light play a role in structuring kelp communities.
+damage,photopigment destruction,reduced growthand decreased
+survivorship, thus inhibiting their settlement and survival {[}@Wood
+1987{]}. Laboratory experiments revealed that the UV component of
+radiation, rather than intense radiation itself, was responsible for the
+eï¬\euro{}ects mentioned above. High light stress has negative
+eï¬\euro{}ects, such as photoinhibition and photo-damage on
+\emph{Ecklonia cava} sporophytes {[}@Altamirano2004{]}. @Altamirano2004
+found that \emph{Ecklonia cava} is more vulnerable to light stress
+conditions, and less likely to recover under unfavourable conditions
+{[}@Altamirano2004{]}. @Bolton1985 showed that under sub-saturating
+irradiances and supra- optimal temperatures \emph{Ecklonia maxima} to
+showed a decrease in reproductive rates and an increase in cell
+production. An additional ï¬nding of this study was that despite the
+decrease in reproductive rates, the ï¬nal egge production per female
+was greater under these conditons. The authors interpreted this an
+ecological adaption that may increase survival rates under times of
+stress or non - ideal conditions {[}@Bolton1985{]}.
+
+\section{Depth}\label{depth}
+
+Depth does not aï¬\euro{}ect kelp ecosystems directly, however a change
+in depth often causes ï¬‚uctuations or changes in other environmental
+variables such as water motion, light and temperature. Water motion also
+decreases with depth, and some kelps better suited to deeper
+environments (\emph{L. pallida}) replace those in the shallows (\emph{E.
+maxima}) (Dayton 1985; Gerard 1982). The increase in depth can lead to a
+decrease in sunlight penetration, with some species better adapted for
+low-light conditions than others, such as (\emph{L. pallida}).
+Temperature may also change along a depth gradient due to a reduction in
+sunlight penetration. (Dayton 1985; Gerard 1982). Therefore depth does
+not directly play a role in kelp functioning but may alter more
+inï¬‚uential factors such as light and water motion.
 
@@
 
-Nutrients
-Theimportanceofnutrientsinthefunctioningofkelpsiswellunderstood
-{[}@Dayton1985; @Gaylord2012{]}. Dissolved nitrogen, and in particular
-nitrate, are important; however research has also placed emphasis on
-phosphate and other trace compounds for functioning of kelps
+The importance of nutrients in the functioning of kelps is well
+understood {[}@Dayton1985; @Gaylord2012{]}. Dissolved nitrogen, and in
+particular nitrate, are important; however research has also placed
+emphasis on phosphate and other trace compounds for functioning of kelps
 {[}@Dayton1985{]}. Additionally, some kelps have the ability to store
@@
 availability, which has been observed for Laminaria and Macrocystis
-{[}@Dayton1985; @Gaylord2012{]}. Nutrient stratiﬁcation is also an
+{[}@Dayton1985; @Gaylord2012{]}. Nutrient stratiï¬cation is also an
 important factor, particularly for canopy type kelps. The concentration
@@
 Nutrients are often in higher concentrations in the water column during
-low temperature events. This is often an indication of an ``up-welling''
-event, which brings cold and nutrient rich waters from the bottom to the
-surface of the water column. However temperature can play a direct role
-in the uptake of nutrients through eﬀects on algal metabolism; however
-this may vary from species to species {[}@Raven1988{]}. However most
-studies investigate multiple characteristics in one study
-(i.e.~temperature and light and water motion), while studies
-investigating the direct role of temperature in nutrient assimilation is
-lacking in marine algae is lacking.
-
-\section{Depth}\label{depth}
-
-Depth does not aﬀect kelp ecosystems directly, however a change in depth
-often causes ﬂuctuations or changes in other environmental variables
-such as water motion, light and temperature. Water motion also decreases
-with depth, and some kelps better suited to deeper environments
-(\emph{L. pallida}) replace those in the shallows (\emph{E. maxima})
-(Dayton 1985; Gerard 1982). The increase in depth can lead to a decrease
-in sunlight penetration, with some species better adapted for low-light
-conditions than others, such as (\emph{L. pallida}). Temperature may
-also change along a depth gradient due to a reduction in sunlight
-penetration. (Dayton 1985; Gerard 1982). Therefore depth does not
-directly play a role in kelp functioning but may alter more inﬂuential
-factors such as light and water motion.
-
-\section{Sedimentation}\label{sedimentation}
-
-Sedimentation and scour can be highly destructive to kelp forests with
-most detrimental impacts on the development stages such as spores,
-gametophytes and young plants (Dayton 1985), and has been linked to
-seasonal trends relating to wind direction {[}@Moore1972{]}.
-@Devinny1978 found that sedimentation greatly reduced spore settlement
-and which in turn greatly reduces the probability of survival for
-Macrocystis Pyrifera gametophytes. Some species seem more resistant to
-the eﬀects of sedimentation such as Laminaria farlowii and Eisena aborea
-compared to others such as Macrocystis Pyrifera (Dayton 1985). The
-impacts of sedimentation is not limited to the sub-tidal but has also
-been shown to eﬀect inter-tidal kelp as well. For example a study by
-{[}@Schiel2006{]} investigated the eﬀects of sedimentation on the
-post-settlement stages of habitat-forming seaweeds. The results showed
-high mortality of germlings at sites that had high sedimentation, and
-was further supported through laboratory tests. Grazing was found to be
-negligible and mortality of gametophytes was attributed to
-sedimentation, although its eﬀects did vary with molluscan grazing
-{[}@Schiel2006{]}. The presence of kelps themselves can also eﬀect the
-degree of sedimentation. For instance, @Wernberg2005 investigated the
-ability of Ecklonia radiata (canopy forming species) to modify the
-physical environment and the related eﬀects on associated foliose algae.
-An increase in kelp densities was marked by a signiﬁcant reduction of
-light and sedimentation, and found that sedimentation did not have a
-signiﬁcant eﬀect on associated foliose algae {[}@Wernberg2005{]}.
-Therefore it is often diﬃcult to separate the direct eﬀects of each of
-the environmental variables as they are often related to one another.
-However, sea surface temperatures is one of the main predictors of
-marine biological systems, as it has been shown to relate to
-physiological and ecological processes at all levels {[}@Andersen2013;
-@Bolton1985; @Bolton2012; @Dayton1985{]}. Kelp forests are no exception,
-with temperature being the main limiting factor {[}@Dayton 1985{]}.
+low temperature events. This is often an indication of an
+â\euro{}œup-wellingâ\euro{} event, which brings cold and nutrient rich
+waters from the bottom to the surface of the water column. Temperature
+can play a direct role in the uptake of nutrients through eï¬\euro{}ects
+on algal metabolism; however this may vary from species to species
+{[}@Raven1988{]}.
 
@@
 
-Temperature Biogeographical shifts of kelp species can already be seen
-in areas where ocean warming is occurring. In Western Australia an
-increase in temperature has caused a biogeographical shift in kelp and
-urchins, which in turn has caused many kelp forests to shift to the less
-productive algal turfs {[}@Kelly2005{]};a review of the literature shows
-a consensus that biogeographical distributions of marine species are
-largely temperature related {[}@Dayton1985; @Jeﬀree1994; @Schils2006;
-@Sorte2010; @Bolton2012{]}. Other examples include Norway and eastern
-North America where ocean warming has facilitated kelp die-oﬀ and an
-increase in herbivore grazing rates {[}@Dayton1985{]}. The increase in
-grazing rates and the colonization of other algal species into these
-areas make it almost impossible for kelps to recover {[}@Kelly2005{]}.
-The majority of kelp species are artic and temperate organisms, and the
-warming of ocean temperatures is expected to cause a poleward
-biogeographical shift of species {[}@Bolton2012{]}. There is evidence to
-suggest that South African kelp forests are expanding due to ocean
-cooling {[}@Bolton2012{]}, possibly driven by an intensiﬁcation and
-increase in coastal upwelling {[}@Blamey2012, @Blamey2015{]}. In South
-Africa there has been a biogeographical shift eastward along the coast
-due to a change in inshore temperature regime, making South Africa no
-exception to changing ocean temperatures {[}@Bolton2012{]}. Macroalgae,
-such as kelps, can react to an increase in surface temperatures in one
-of three ways: they can migrate, adapt and die {[}@Biskup2014{]}. A
-study by @Biskup2014 investigated the functional response of two kelp
-species (\emph{Laminaria ochroleuca} and \emph{Saccorhiza polyschides})
-to rising sea temperatures. The functional responses of Saccorhiza
+Temperature is a driver of kelp species distributions and
+ecophysiological processes, as well as a lesser role in morphological
+adaption\ldots{}\textbf{example here}\ldots{}The majority of kelp
+species are artic and temperate organisms, and the warming of ocean
+temperatures is expected to cause a poleward biogeographical shift of
+species {[}@Bolton2012{]}. There is evidence to suggest that South
+African kelp forests are expanding due to ocean cooling
+{[}@Bolton2012{]}, possibly driven by an intensiï¬cation and increase
+in coastal upwelling {[}@Blamey2012, @Blamey2015{]}. In South Africa
+there has been a biogeographical shift eastward along the coast due to a
+change in inshore temperature regime, making South Africa no exception
+to changing ocean temperatures {[}@Bolton2012{]}. Macroalgae, such as
+kelps, can react to an increase in surface temperatures in one of three
+ways: they can migrate, adapt and die {[}@Biskup2014{]}. A study by
+@Biskup2014 investigated the functional response of two kelp species
+(\emph{Laminaria ochroleuca} and \emph{Saccorhiza polyschides}) to
+rising sea temperatures. The functional responses of Saccorhiza
 polyschides was measured for both the subtidal and intertidal habitats,
-to see what aﬀect non- optimal conditions (intertidal zone) had on the
-kelps {[}@Rinde2005{]}. The study found that Laminaria ochroleuca
-exhibited a poor ability to acclimatise and was dependent on the kelp's
-life history traits {[}@Biskup2014{]}. Therefore annual kelp species are
-more likely to survive under non-ideal condition, and the intertidal
-Saccorhiza polyschides, compared to the subtidal, showed a higher
-physiological ﬂexibility to changing conditions {[}@Biskup2014{]}. This
-may be because the intertidal zone undergoes far more change than the
-subtidal and therefore kelps in the intertidal are forced to adapt to
-harsher conditions where ﬂuctuations in temperature, sunlight, turbidity
-and water motion are common. The eﬀects on temperature have also been
-investigated by @Wernberg2010. The study looked at resilience of kelp
-beds along a latitudinal temperature gradient. Kelp abundance is likely
-to decline with the predicted warming of ocean waters @Wernberg2010 and
-although kelps have the ability to acclimatize and adjust their
-metabolic performance, which in turn allows them to change their
-physiological performance to mitigate the seasonal ﬂuctuations in
-temperature, this acclimatization is done at a cost @Wernberg2010. It is
-therefore important to understand the ecological implications of this
-cost in order to better understand the future impacts of rising sea
-temperatures @Wernberg2010. Results from the study showed that kelp beds
-in the northern regions had a lower resilience to perturbations, such as
-physical and biological disturbance, disease, competition from invasive
-species and decreasing water quality. This was as a result of low
-abundance of recruits, reduced physiological responsiveness and reliance
-on mature adult kelp for net recruitment, growth of recruits and
-recovery of canopy loss @Wernberg2010. Even though @Wernberg2010
-acknowledged that it was diﬃcult to exclude the inﬂuences of other
-environmental parameters (sunlight, nutrients and wave exposure),
-temperature was most likely the main environmental driver of these
-eﬀects. While kelps may be able to maintain populations in the warmer
-upper limits of their distributions, the cost of acclimation will make
-them more susceptible to other stressors such as disturbance from
-anthropogenic sources, disease, competition and decreasing water
-quality. Changes in temperature can also impact the distribution of
-kelps.
+to see what aï¬\euro{}ect non- optimal conditions (intertidal zone) had
+on the kelps {[}@Rinde2005{]}. The study found that Laminaria ochroleuca
+exhibited a poor ability to acclimatise and was dependent on the
+kelpâ\euro{}™s life history traits {[}@Biskup2014{]}. Therefore annual
+kelp species are more likely to survive under non-ideal condition, and
+the intertidal Saccorhiza polyschides, compared to the subtidal, showed
+a higher physiological ï¬‚exibility to changing conditions
+{[}@Biskup2014{]}. This may be because the intertidal zone undergoes far
+more change than the subtidal and therefore kelps in the intertidal are
+forced to adapt to harsher conditions where ï¬‚uctuations in
+temperature, sunlight, turbidity and water motion are common. The
+eï¬\euro{}ects on temperature have also been investigated by
+@Wernberg2010. The study looked at resilience of kelp beds along a
+latitudinal temperature gradient. Kelp abundance is likely to decline
+with the predicted warming of ocean waters @Wernberg2010 and although
+kelps have the ability to acclimatize and adjust their metabolic
+performance, which in turn allows them to change their physiological
+performance to mitigate the seasonal ï¬‚uctuations in temperature, this
+acclimatization is done at a cost @Wernberg2010\ldots{}\textbf{link to
+paragraph on kelp morphology}\ldots{}
+
+\section{Wave exposure}\label{wave-exposure}
+
+Other than temperature, wave exposure is also recognised as an important
+driver of the marine environment. Like temperature, wave exposure can
+act on regional scales and has been shown to play a role in determing
+distribution, abundance, diversity, composition and productivity of
+benthic and rocky shore communties. For example, increasing degrees of
+exposure may positively influence the amount of area available to trap
+light on macroalgal fronds, as well as increasing nutrient uptake
+through increased turbulence in the boundry layer {[}@lobban1994{]}.
+@lewis found that degrees of wave exposure vertical structure rocky
+shore communities\ldots{}\textbf{work in progress}\ldots{}
+
+\section{Hydrodynamic modelling}\label{hydrodynamic-modelling}
+
+Traditional ecological measures of wave exposure usually incorportates
+integrative measures of hyrdodynamic conditions at a particular site.
+More specifically, it is the integration of mechanical processes and the
+influence that is has on the ecology of nearshore communities.
+
+Wave exposure may be modelled through various methods which range from
+simple cartographic to more advanced numerical wave models.
+Cartographical models can be qualitative or quantitative and were
+designed for the need of wave exposure measures to explain ecological
+distributions. A simple set of calculations on coastline and wind data,
+and relatively small input data sets are required. These are regarded as
+``fetch-based models'', which measure the length of open water
+associated with a site along a straight line. The output of such an
+approach is a simplified estimate of the potential wave energy for a
+specific set of sites. Advances in cartographical methods using
+fetch-based models has allowed for wave exposure measurments for larger
+areas, and has been suggested as a method for predicting macroalgal
+community structure {[}@burrows2008{]}. An example of such a model is
+the ``BioEx model'' which was developed by @baardseth1970 to estimate
+wave exposure over large regions. BioEx requires frequency, strength and
+direction of winds, weighted by degree of exposure within various
+directions. BioEx is calculated as the sum of the index developed at
+different spatial scales (local, fjord and open). This method has been
+used in mapping of marine coastal biodiveristy {[}@Rinde2004{]}.
+
+@lindegarth2005 critized this approach, arguing that the choice of wave
+exposure method can influence ecological inference. The authors also
+highlighted the need for objective, reproducible and quantitative
+studies comparing exposure indices {[}@lindegarth2005{]}. A study by
+@sundbald\ldots{}\textbf{include description of study here}\ldots{}Other
+authors, such as @hill2010, have argued that these simple measures can
+be improved upon by including bathymetery data which allows the
+incorporation of diffraction into the calculation. Diffraction is
+topographically induced variations in wave direction. A model
+incorporating this complexity was developed by @isaeus2004, and is known
+as the ``simplified wave model'' (SWM). The model uses measurements of
+wind strength, fetch and empirically derived algorithims to mimic
+diffraction.
+
+Advances in numerical modelling have been founded on physical wave
+theory on how a wave ``behaves''. This approach is based on a
+theoretical persepctive rather than the need to answer ecological
+questions. Besides diffraction, these models incorporate more complexity
+by including wind forcing, wave-to-wave interactions and loss of energy
+due to friction and wave breaking. Numerical models have a variety of
+applications and are often incorporated within hydrodynamic general
+circulation models and are used operationally for forecasting the sea
+state {[}@hasselmann1988; @booij1999; @smith2001{]}. The downside of
+advanced numerical models is that are computationally intensive which
+creates limitations for large scale simulations. Therefore, their
+application along long stretches of variable coastline, inshore
+environments and ocean-wide simulations limited due to the poor spatial
+coverage. However, this models can be designed for local or site
+specific coverage, provided the correct data is available.
 

### 2018-09-20 chapter_1.Rmd
Minor edits to chapter_1 script and time-frame added to proposal

@@
 
-The important environmental drivers of kelp idividuals and communities include light, substrata, salinity, sedimentation, nitrients, temperature and water motion.
+The important environmental drivers of kelp individuals and communities include light, substrata, salinity, sedimentation, nitrients, temperature and wave exposure.
 Although studies have investigated the effects of important environmental drivers, the roles these factors play is often difficult to evaluate as such factors may never be fully independent of each other, i.e.
-environmental factors are to some extent temperature dependent.
-Multifactorial studies have attempt to explain combined affects, however these studies are often limited to investigating combination of two or three environmental drivers as inclusion of too many factors can lead to results that are difficult to interept.
-Environmetal factors are highly variable on temporal and spatial scales, and their effects may also dependent on life-stage adding a further layer of complexity to investigations.
+environmental factors are to some extent dependent on one another.
+Multifactorial studies have attempted to explain combined affects, however these studies are often limited to investigating combination of two or three environmental drivers as inclusion of too many factors can lead to results that are difficult to interept.
+Environmetal factors are highly variable on temporal and spatial scales and their effects may also be dependent on the life-stage of the organism, adding a further layer of complexity to investigations.
 
@@
 
+#Substrata
+#Salinity
 #Depth
@@
 
+#Sedimentation
 #Nutrients
@@
 Other than temperature, wave exposure is also recognised as an important driver of the marine environment.
-Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity of macroalgae communties.
+Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of macroalgae communties.
 The width, vertical zonation and diversity of algal communities often change predictably along gradients of exposure.
@@
 
-For example, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake through increased turbulence in the boundry layer [@lobban1994].
+Increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake through increased turbulence in the boundry layer [@lobban1994].
 @lewis found that degrees of wave exposure vertical structure rocky shore communities...**work in progress**...

### 2018-09-20 proposal_draft_2018.Rmd
Minor edits to chapter_1 script and time-frame added to proposal

@@
 The downside of advanced numerical models is that are computationally intensive which creates limitations for large scale simulations.
-Therefore, their application along long stretches of variable coastline, inshore environments and ocean-wide simulations limited due to the poor spatial coverage.
-However, this models can be designed for local or site specific coverage, provided the correct data is available.
+Therefore, their application along long stretches of variable coastline, inshore environments and ocean-wide simulations are limited due to the poor spatial coverage.
+However, these models can be designed for local or site specific coverage, provided the correct data are available.
 
@@
 A combined quantitative ecology and oceanography approach will be used for this study.
-Morphological characteristics and their relation to the hydrodynamic environment must first be established and the findings will form the basis for the proceeding research.
+Morphological characteristics and their relation to the hydrodynamic environment must first be established and the findings will form the basis for the proceeding chapters.
 Once the relationship has been established it will be used to calibrate future hydrodynamic models developed during the course of the study.
@@
 
+Months | Description
+------------------------------ | -------------------------------------------
+April 2018 - September 2018 | Develop and write proposal for submission to department
+September 2018 - December 2018 | Acceptable level of progress on literature review. The first data chapter should also be complete and submission of first paper
+January 2019 - May 2019 | Become competent in relevant software (Python). Data analyses for the 2nd data chapter should ideally be near completion. Substantial additions to literature review. 
+May 2019 - August 2019| Completion of 2nd data chapter and relevant paper should also be near completion. 
+August 2019 - December 2019 | Complete analyses and write-up for final data chapter. Significant progress on the relevant paper. Literature review should be near completion. 
+January 2020 - April 2020 | This time will be used to make final edits and complete thesis write-up
+
 ## References

### 2018-09-20 chapter_1.Rmd
added to wave exposure section

@@
 
-Other than temperature, wave exposure is also recognised as an important driver of the marine environment.
-Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of macroalgae communties.
-The width, vertical zonation and diversity of algal communities often change predictably along gradients of exposure.
-This is also true for variation in morphological characters of individuals may also vary with wave exposure, such as frond width and degree of 'branching' [@moss1948].
+Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception.
+Wave exposure has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of macroalgae communties.
+For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure.
+Wave exposure may also drive macroalgae communties indirectly through the alteration in affect of another environmental driver.
+For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundry layer around the frond [@lobban1994].
+The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgment, which ultimately leads to expiration.
+Wave exposure is a complex abiotic variable which varies spatially and temporarly in the marine enviroment.
+Furthermore, the degree to which a macroalgae community is exposed which depend on local site characteristics, such as bathymetry and local wind patterns.
+Despite this fact, macrolagae have been able to persist in often harsh and variable wave environments.
+Macroalage are sessile organisms and incapabable of migrating when local conditions become unsuitable.
+Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological variation.
 @moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed).
-The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid in relation to wave exposure between study sites.
-
-Increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake through increased turbulence in the boundry layer [@lobban1994].
-@lewis found that degrees of wave exposure vertical structure rocky shore communities...**work in progress**...
+The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid concentrations.
 

### 2018-09-20 chapter_1.Rmd
edited and added to the wave exposure section.

@@
 output:
-  word_document:
-    toc: yes
   pdf_document:
@@
     number_sections: yes
+    toc: yes
+  html_document:
+    df_print: paged
+    toc: yes
+  word_document:
     toc: yes
@@
 Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological variation.
+This morphological variation allows macroalgae to reduce the amount of drag experienced during times of high wave energy or pulse distrubance events, such as storms.
+This may be achieved through different strategies and are species-specific, which can be directly attributed the high diversity in morphological characters of algal communities.
+For example some algae have fronds and others are articulated corraline, and therefore these species would need to adpat their morphology differently in order to persist.
 @moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed).

### 2018-09-20 chapter_1.tex
edited and added to the wave exposure section.

@@
   \usepackage[utf8]{inputenc}
-  \usepackage{eurosym}
 \else % if luatex or xelatex
@@
   \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
 \fi
@@
 
-The important environmental drivers of kelp idividuals and communities
+The important environmental drivers of kelp individuals and communities
 include light, substrata, salinity, sedimentation, nitrients,
-temperature and water motion. Although studies have investigated the
+temperature and wave exposure. Although studies have investigated the
 effects of important environmental drivers, the roles these factors play
@@
 independent of each other, i.e.~environmental factors are to some extent
-temperature dependent. Multifactorial studies have attempt to explain
-combined affects, however these studies are often limited to
+dependent on one another. Multifactorial studies have attempted to
+explain combined affects, however these studies are often limited to
 investigating combination of two or three environmental drivers as
@@
 interept. Environmetal factors are highly variable on temporal and
-spatial scales, and their effects may also dependent on life-stage
-adding a further layer of complexity to investigations.
+spatial scales and their effects may also be dependent on the life-stage
+of the organism, adding a further layer of complexity to investigations.
 
@@
 of kelp {[}@Bruhn1996);@Belseth2012{]}. For instance, solar ultraviolate
-radiation has been shown to aï¬\euro{}ect sub-canopy Ecklonia radiata
+radiation has been shown to aﬀect sub-canopy Ecklonia radiata
 sporophytes when the canopy of mature \emph{Ecklonia radiata} was
@@
 radiation, rather than intense radiation itself, was responsible for the
-eï¬\euro{}ects mentioned above. High light stress has negative
-eï¬\euro{}ects, such as photoinhibition and photo-damage on
-\emph{Ecklonia cava} sporophytes {[}@Altamirano2004{]}. @Altamirano2004
-found that \emph{Ecklonia cava} is more vulnerable to light stress
-conditions, and less likely to recover under unfavourable conditions
-{[}@Altamirano2004{]}. @Bolton1985 showed that under sub-saturating
-irradiances and supra- optimal temperatures \emph{Ecklonia maxima} to
-showed a decrease in reproductive rates and an increase in cell
-production. An additional ï¬nding of this study was that despite the
-decrease in reproductive rates, the ï¬nal egge production per female
-was greater under these conditons. The authors interpreted this an
-ecological adaption that may increase survival rates under times of
-stress or non - ideal conditions {[}@Bolton1985{]}.
+eﬀects mentioned above. High light stress has negative eﬀects, such as
+photoinhibition and photo-damage on \emph{Ecklonia cava} sporophytes
+{[}@Altamirano2004{]}. @Altamirano2004 found that \emph{Ecklonia cava}
+is more vulnerable to light stress conditions, and less likely to
+recover under unfavourable conditions {[}@Altamirano2004{]}. @Bolton1985
+showed that under sub-saturating irradiances and supra- optimal
+temperatures \emph{Ecklonia maxima} to showed a decrease in reproductive
+rates and an increase in cell production. An additional ﬁnding of this
+study was that despite the decrease in reproductive rates, the ﬁnal egge
+production per female was greater under these conditons. The authors
+interpreted this an ecological adaption that may increase survival rates
+under times of stress or non - ideal conditions {[}@Bolton1985{]}.
+
+\section{Substrata}\label{substrata}
+
+\section{Salinity}\label{salinity}
 
@@
 
-Depth does not aï¬\euro{}ect kelp ecosystems directly, however a change
-in depth often causes ï¬‚uctuations or changes in other environmental
-variables such as water motion, light and temperature. Water motion also
-decreases with depth, and some kelps better suited to deeper
-environments (\emph{L. pallida}) replace those in the shallows (\emph{E.
-maxima}) (Dayton 1985; Gerard 1982). The increase in depth can lead to a
+Depth does not aﬀect kelp ecosystems directly, however a change in depth
+often causes ﬂuctuations or changes in other environmental variables
+such as water motion, light and temperature. Water motion also decreases
+with depth, and some kelps better suited to deeper environments
+(\emph{L. pallida}) replace those in the shallows (\emph{E. maxima})
+{[}@Dayton1985; @Gerard1982{]}. The increase in depth can lead to a
 decrease in sunlight penetration, with some species better adapted for
@@
 Temperature may also change along a depth gradient due to a reduction in
-sunlight penetration. (Dayton 1985; Gerard 1982). Therefore depth does
-not directly play a role in kelp functioning but may alter more
-inï¬‚uential factors such as light and water motion.
+sunlight penetration {[}@Dayton1985; @Gerard1982{]}. Therefore depth
+does not directly play a role in kelp functioning but may alter more
+inﬂuential factors such as light and water motion.
+
+\section{Sedimentation}\label{sedimentation}
 
@@
 availability, which has been observed for Laminaria and Macrocystis
-{[}@Dayton1985; @Gaylord2012{]}. Nutrient stratiï¬cation is also an
+{[}@Dayton1985; @Gaylord2012{]}. Nutrient stratiﬁcation is also an
 important factor, particularly for canopy type kelps. The concentration
@@
 Nutrients are often in higher concentrations in the water column during
-low temperature events. This is often an indication of an
-â\euro{}œup-wellingâ\euro{} event, which brings cold and nutrient rich
-waters from the bottom to the surface of the water column. Temperature
-can play a direct role in the uptake of nutrients through eï¬\euro{}ects
-on algal metabolism; however this may vary from species to species
-{[}@Raven1988{]}.
+low temperature events. This is often an indication of an ``up-welling''
+event, which brings cold and nutrient rich waters from the bottom to the
+surface of the water column. Temperature can play a direct role in the
+uptake of nutrients through eﬀects on algal metabolism; however this may
+vary from species to species {[}@Raven1988{]}.
 
@@
 African kelp forests are expanding due to ocean cooling
-{[}@Bolton2012{]}, possibly driven by an intensiï¬cation and increase
-in coastal upwelling {[}@Blamey2012, @Blamey2015{]}. In South Africa
-there has been a biogeographical shift eastward along the coast due to a
+{[}@Bolton2012{]}, possibly driven by an intensiﬁcation and increase in
+coastal upwelling {[}@Blamey2012, @Blamey2015{]}. In South Africa there
+has been a biogeographical shift eastward along the coast due to a
 change in inshore temperature regime, making South Africa no exception
@@
 polyschides was measured for both the subtidal and intertidal habitats,
-to see what aï¬\euro{}ect non- optimal conditions (intertidal zone) had
-on the kelps {[}@Rinde2005{]}. The study found that Laminaria ochroleuca
-exhibited a poor ability to acclimatise and was dependent on the
-kelpâ\euro{}™s life history traits {[}@Biskup2014{]}. Therefore annual
-kelp species are more likely to survive under non-ideal condition, and
-the intertidal Saccorhiza polyschides, compared to the subtidal, showed
-a higher physiological ï¬‚exibility to changing conditions
-{[}@Biskup2014{]}. This may be because the intertidal zone undergoes far
-more change than the subtidal and therefore kelps in the intertidal are
-forced to adapt to harsher conditions where ï¬‚uctuations in
-temperature, sunlight, turbidity and water motion are common. The
-eï¬\euro{}ects on temperature have also been investigated by
-@Wernberg2010. The study looked at resilience of kelp beds along a
-latitudinal temperature gradient. Kelp abundance is likely to decline
-with the predicted warming of ocean waters @Wernberg2010 and although
-kelps have the ability to acclimatize and adjust their metabolic
-performance, which in turn allows them to change their physiological
-performance to mitigate the seasonal ï¬‚uctuations in temperature, this
-acclimatization is done at a cost @Wernberg2010\ldots{}\textbf{link to
-paragraph on kelp morphology}\ldots{}
+to see what aﬀect non- optimal conditions (intertidal zone) had on the
+kelps {[}@Rinde2005{]}. The study found that Laminaria ochroleuca
+exhibited a poor ability to acclimatise and was dependent on the kelp's
+life history traits {[}@Biskup2014{]}. Therefore annual kelp species are
+more likely to survive under non-ideal condition, and the intertidal
+Saccorhiza polyschides, compared to the subtidal, showed a higher
+physiological ﬂexibility to changing conditions {[}@Biskup2014{]}. This
+may be because the intertidal zone undergoes far more change than the
+subtidal and therefore kelps in the intertidal are forced to adapt to
+harsher conditions where ﬂuctuations in temperature, sunlight, turbidity
+and water motion are common. The eﬀects on temperature have also been
+investigated by @Wernberg2010. The study looked at resilience of kelp
+beds along a latitudinal temperature gradient. Kelp abundance is likely
+to decline with the predicted warming of ocean waters @Wernberg2010 and
+although kelps have the ability to acclimatize and adjust their
+metabolic performance, which in turn allows them to change their
+physiological performance to mitigate the seasonal ﬂuctuations in
+temperature, this acclimatization is done at a cost
+@Wernberg2010\ldots{}\textbf{link to paragraph on kelp
+morphology}\ldots{}
 
@@
 Other than temperature, wave exposure is also recognised as an important
-driver of the marine environment. Like temperature, wave exposure can
-act on regional scales and has been shown to play a role in determing
-distribution, abundance, diversity, composition and productivity of
-benthic and rocky shore communties. For example, increasing degrees of
-exposure may positively influence the amount of area available to trap
-light on macroalgal fronds, as well as increasing nutrient uptake
-through increased turbulence in the boundry layer {[}@lobban1994{]}.
-@lewis found that degrees of wave exposure vertical structure rocky
-shore communities\ldots{}\textbf{work in progress}\ldots{}
+driver of the marine environment, and macroalgae are not exception. Wave
+exposure has been shown to play a role in determing distribution,
+abundance, diversity, composition, growth {[}@Cousens1982{]} and
+productivity {[}@pedersen2012a{]} of macroalgae communties. For example,
+the width, vertical zonation and diversity of algal communities often
+change predictably along gradients of wave exposure. Wave exposure may
+also drive macroalgae communties indirectly through the alteration in
+affect of another environmental driver. For instance, increasing degrees
+of exposure may positively influence the amount of area available to
+trap light on macroalgal fronds, as well as increasing nutrient uptake
+due to increased turbulence in the boundry layer around the frond
+{[}@lobban1994{]}. The most important direct effect of wave exposure on
+macroalgal communities is through mechanical dislodgment, which
+ultimately leads to expiration. Wave exposure is a complex abiotic
+variable which varies spatially and temporarly in the marine enviroment.
+Furthermore, the degree to which a macroalgae community is exposed which
+depend on local site characteristics, such as bathymetry and local wind
+patterns. Despite this fact, macrolagae have been able to persist in
+often harsh and variable wave environments. Macroalage are sessile
+organisms and incapabable of migrating when local conditions become
+unsuitable. Therefore, macroalgae must adapt to the local wave climate
+in order to persist and survive, and achieve this through morphological
+variation.
+
+\begin{quote}
+keeping this for now @moss1948 investigated the anatomy, chemical
+composistion of \emph{Fucus spiralis} at three sites that varied in wave
+exposure (sheltered, medium exposure and exposed). The authors found
+that individuals in exposed sites showed less branching of thalli as
+well as variation in organic nitrogen, mannitol, laminarian and alginic
+acid concentrations.
+\end{quote}
 

### 2018-09-21 chapter_1.Rmd
more additions and edits to the wave exposure section

@@
 Wave exposure is a complex abiotic variable which varies spatially and temporarly in the marine enviroment.
-Furthermore, the degree to which a macroalgae community is exposed which depend on local site characteristics, such as bathymetry and local wind patterns.
+Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns.
 Despite this fact, macrolagae have been able to persist in often harsh and variable wave environments.
 Macroalage are sessile organisms and incapabable of migrating when local conditions become unsuitable.
-Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological variation.
-This morphological variation allows macroalgae to reduce the amount of drag experienced during times of high wave energy or pulse distrubance events, such as storms.
+Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation.
+The morphology of macroalgae are not fixed gentic traits.
+A study by @??
+showed that *species name* was able to rapidly adapt to the local wave environment when transplanted from a low wave exposure site to a high wave exposure site.
+Advances in genetic techniques and taxonomy have revealed that species delination based on morphology has been inaccurate, and organisms that were once considered two seperate species are actually one species.**insert example here**.
+The morphological adaption that macroalgae display are therefore driven by local or site conditions, and therefore individuals must be morphologically flexible to persist in stochastic environments.
 This may be achieved through different strategies and are species-specific, which can be directly attributed the high diversity in morphological characters of algal communities.
-For example some algae have fronds and others are articulated corraline, and therefore these species would need to adpat their morphology differently in order to persist.
+For example some algae have fronds and others are articulated corraline, and therefore these species would need to adapt their morphology differently in order to persist.
 @moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed).
 The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid concentrations.
+**insert more examples here**.
 
@@
 
+Wave exposure may be modelled through various methods which range from simple cartographic to more advanced numerical wave models.
 Traditional ecological measures of wave exposure usually incorportates integrative measures of hyrdodynamic conditions at a particular site.
 More specifically, it is the integration of mechanical processes and the influence that is has on the ecology of nearshore communities.
-
-Wave exposure may be modelled through various methods which range from simple cartographic to more advanced numerical wave models.
 Cartographical models can be qualitative or quantitative and were designed for the need of wave exposure measures to explain ecological distributions.

### 2018-09-21 proposal_draft_2018.Rmd
more additions and edits to the wave exposure section
some minor edits and additions to the hydrodynamic model section

@@
 This process will be repeated until the model can predict coastal flow and kelp morphology within a certian degree of probability.
+Kelp morphology is driven by regional and local environmental factors, the spatial scales used will be dependent on the spatial patterns of the various abiotic variables being investigated.
+Fully nested hierarchical design at three spatial scales will be used: regions, locations (within regions) and sites (within locations).
 
@@
 August 2019 - December 2019 | Complete analyses and write-up for final data chapter. Significant progress on the relevant paper. Literature review should be near completion. 
-January 2020 - April 2020 | This time will be used to make final edits and complete thesis write-up
+January 2020 - June 2020 | Complete of final data chapter and final chapter.
+June 2020 - December 2020 | This time will be used to make final edits and complete thesis write-up
 

### 2018-09-24 proposal_draft_2018.Rmd
added to the preliminary literature review

@@
 \newpage
-## Background
+# Background
 
@@
 
-## Rationale
+# Rationale
 
 Seaweeds, browns in particular, are capable of adapting morphological characteristics to persist in changing and variable ocean environments.
-Changes in morphology have been shown to increase photosynthetic ability, enhance nutrient uptake, and reduce drag with regards to changes in temperature and wave exposure.
+Changes in morphology have been shown to increase photosynthetic ability, enhance nutrient uptake, and reduce drag with regards to changes in wave exposure.
 Wave exposure has been shown to be an important driver of seaweed morphology, as the main mechanism of seaweed mortality is through the mechanism of dislodgment.
@@
 
-## Research problem
+# Research problem
 
@@
 
-## Preliminary literature review
+# Preliminary literature review
 
 Abiotic and biotic factors interact in complex ways which indirectly determine behavioral and ecophysiological responses in organisms.
-For example, when storms or strong currents form in sub-tidal habitats, sea urchins form aggregations in order to reduce overall drag to avoid being swept away by currents.
-Organisms that are motile my migrate into more environmentally suitable areas when conditions become unfavorable or food sources become depleted.
-In changing environments migration may also allow organisms to extend their distributions.
-For example, ocean warming off the coast of Western Australia has allowed tropical fish species to extend their distribution into areas that were previously characterised as temperate reefs.
+Organisms that are mobile may migrate into more environmentally suitable areas when conditions become unfavorable or food sources become depleted.
 Sessile organisms are unable to migrate into more environmentally suitable areas and are forced to either adapt or suffer expiration.
@@
 
-Temperature and wave exposure have been shown to be important drivers of seaweed distribution, physiological functioning, ability to recover, population dynamics and morphology.
+Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception.
+Wave exposure has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of macroalgae communties.
+For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure.
+Wave exposure may also drive macroalgae communties indirectly through the alteration in affect of another environmental driver.
+For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundry layer around the frond [@lobban1994].
+The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgment, which ultimately leads to expiration.
 Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10- 20 m s-1 with accelerations of 400 m s-2 [@Friedland1995] are the biggest threat to kelp survival.
@@
 
-Kelp morphology may be distinct to a particular region with a specific hydrodynamic environment and has the ability to raft far distances using coastal currents, and may accumulate as beach-cast in areas far from its original location.
-Therefore, kelp morphology may be used as a proxy for determining its original location as well as aid in characterising coastal currents.
-However, this approach must be combined with advances in hydrodynamic modelling for a quantitative outcome.
+Wave exposure is a complex abiotic variable which varies spatially and temporarly in the marine enviroment.
+Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns.
+Despite this fact, macrolagae have been able to persist in often harsh and variable wave environments.
+Macroalage are sessile organisms and incapabable of migrating when local conditions become unsuitable.
+Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation.
+The morphology of macroalgae are not fixed gentic traits.
+A study by @??
+showed that *species name* was able to rapidly adapt to the local wave environment when transplanted from a low wave exposure site to a high wave exposure site.
+Advances in genetic techniques and taxonomy have revealed that species delination based on morphology has been inaccurate, and organisms that were once considered two seperate species are actually one species.**insert example here**.
+The morphological adaption that macroalgae display are therefore driven by local or site conditions, and therefore individuals must be morphologically flexible to persist in stochastic environments.
+This may be achieved through different strategies and are species-specific, which can be directly attributed the high diversity in morphological characters of algal communities.
+For example some algae have fronds and others are articulated corraline, and therefore these species would need to adapt their morphology differently in order to persist.
+@moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed).
+The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid concentrations.
+**insert more examples here**.
 
@@
 However these models may be adjusted to suit coastal waters if they are combined with a new numerical model which can be calibrated to suit these needs.
+Numerical modelling for this study will be undertaken using the Delft3D suite of numerical models developed by Deltares.
+The suite of models allows for the simulation of flows, sediment transport, waves, water quality, morphological developments and ecology.
+
+The Delft3D-WAVE module uses the SWAN (Simulating Waves Nearshore) numerical model developed at the Delft University of Technology to simulate the generation and propagation of wind-generated waves in coastal environments.
+The SWAN model is based on the discrete spectral action balance equation and is fully spectral (in all directions and frequencies).
+The latter implies that short-crested random wave fields propagating simultaneously from widely different directions can be accommodated (e.g.
+a wind sea with super-imposed swell).
+Delft3D-FLOW is a multi-dimensional (2D or 3D) hydrodynamic (and transport) simulation program which calculates non-steady
+flow and transport phenomena that result from tidal and meteorological forcing on a rectilinear or a curvilinear, boundary fitted grid.
+In 3D simulations, the vertical grid is defined following the sigma coordinate approach.
+The model solves the shallow water equations, consisting of the momentum and the continuity equations
 
@@
 
-## Research design and methodology
+# Research design and methodology
 
@@
 
-# Morphology
+## Morphology
 
@@
 
-# Hydrodynamic model
+## Hydrodynamic model
 
@@
 These models will consist of the SWAN and Delf3D models and the apporach is to use aspects of each model, that are related to coastal flow, from which further development will take place.
+
 Once the model has been developed it will be used to predict kelp morphology at random sites which will then be sampled.
@@
 
-# Microplastics
+## Microplastics
 

### 2018-09-24 proposal_draft_2018.Rmd
more references added

@@
 Wave exposure may also drive macroalgae communties indirectly through the alteration in affect of another environmental driver.
-For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundry layer around the frond [@lobban1994].
+For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundry layer around the frond [@lobban1995].
 The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgment, which ultimately leads to expiration.
@@
 showed that *species name* was able to rapidly adapt to the local wave environment when transplanted from a low wave exposure site to a high wave exposure site.
-Advances in genetic techniques and taxonomy have revealed that species delination based on morphology has been inaccurate, and organisms that were once considered two seperate species are actually one species.**insert example here**.
+Advances in genetic techniques and taxonomy have revealed that species delination based on morphology has been inaccurate, and organisms that were once considered two seperate species are actually one species.
 The morphological adaption that macroalgae display are therefore driven by local or site conditions, and therefore individuals must be morphologically flexible to persist in stochastic environments.
@@
 The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid concentrations.
-**insert more examples here**.
 
@@
 The output of such an approach is a simplified estimate of the potential wave energy for a specific set of sites.
-Advances in cartographical methods using fetch-based models has allowed for wave exposure measurments for larger areas, and has been suggested as a method for predicting macroalgal community structure [@burrows2008].
+Advances in cartographical methods using fetch-based models has allowed for wave exposure measurments for larger areas, and has been suggested as a method for predicting macroalgal community structure [@Burrows2008].
 An example of such a model is the "BioEx model" which was developed by @baardseth1970 to estimate wave exposure over large regions.
@@
 The authors also highlighted the need for objective, reproducible and quantitative studies comparing exposure indices [@lindegarth2005].
-A study by @sundbald...**include description of study here**...Other authors, such as @hill2010, have argued that these simple measures can be improved upon by including bathymetery data which allows the incorporation of diffraction into the calculation.
+A study by @sundbald.
+Other authors, such as @hill2010, have argued that these simple measures can be improved upon by including bathymetery data which allows the incorporation of diffraction into the calculation.
 Diffraction is topographically induced variations in wave direction.

### 2018-09-24 proposal_draft_2018.Rmd
references updated
more house cleaning

@@
 output:
+  pdf_document:
+    toc: yes
+    toc_depth: 4
   word_document:
@@
     highlight: zenburn
-    toc: no
-  pdf_document:
     toc: yes
-    toc_depth: 4
 bibliography: mycollection.bib   
@@
 The authors also highlighted the need for objective, reproducible and quantitative studies comparing exposure indices [@lindegarth2005].
-A study by @sundbald.
 Other authors, such as @hill2010, have argued that these simple measures can be improved upon by including bathymetery data which allows the incorporation of diffraction into the calculation.
@@
 
+\newpage
 ## References

### 2018-09-24 chapter_2.Rmd
more additions to proposal and chapter 2

@@
 
-Say something about kelp habitats that have a physical influence on the hydrodynamics within the beds.
+Kelps are sessile species and have to adapt to local environmental conditions in order to persist and survive.
+The resilence of kelps is evident through their distribution, which spans a range of environmental conditions that vary spatially and temporarly.
+Wave exposure and temperature are regarded as important environmental drivers of kelps, which play a role in the growth, reproduction, and survival of kelp species.
+Wave exposure is recognised as the main cause of kelp mortality, which may manifest itself through dislodgment of individuals or breakage of important morphological structures.
+These mechanisms usually occur during times of high wave energy or pulse disturbance events, such as storms.
+Kelps are able to reduce the probablity of dislodgment or breakage through morphological adaptation, which may involve several strategies.
+Kelps can either reduce drag, increase attachment strength or increase flexibility in order to survive high wave energy environments.
+**insert examples??*
 

### 2018-09-24 proposal_draft_2018.Rmd
more additions to proposal and chapter 2

@@
     toc: yes
+  pdf_document:
+    toc: yes
+    toc_depth: 4
 bibliography: mycollection.bib   
@@
 
-Abiotic and biotic factors interact in complex ways which indirectly determine behavioral and ecophysiological responses in organisms.
-Organisms that are mobile may migrate into more environmentally suitable areas when conditions become unfavorable or food sources become depleted.
-Sessile organisms are unable to migrate into more environmentally suitable areas and are forced to either adapt or suffer expiration.
-Sessile organisms may respond to changing environmental conditions through changes in physiology.
-For example, plants may produce heat shock proteins that help buffer the affect of temperature increases.
-Sessile organisms may also adapt their morphology,in order to persist in changing and harsh environments, such as seaweeds.
-
-Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception.
+Wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception.
 Wave exposure has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of macroalgae communties.
@@
 
-Wave exposure is a complex abiotic variable which varies spatially and temporarly in the marine enviroment.
-Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns.
-Despite this fact, macrolagae have been able to persist in often harsh and variable wave environments.
+Wave exposure is a complex abiotic variable which varies spatially and temporarly in the marine enviroment [@Sundblad2014; @Bekkby2008a].
+Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns [@Sundblad2014; @Bekkby2008a].
+Despite this fact, macrolagae have been able to persist in often harsh and variable wave environments [@Hatcher1987; @Friedland1995; @Charrier2012; @@Blanchette1997].
 Macroalage are sessile organisms and incapabable of migrating when local conditions become unsuitable.
-Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation.
+Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation [@Charrier2012].
 The morphology of macroalgae are not fixed gentic traits.
@@
 Advances in genetic techniques and taxonomy have revealed that species delination based on morphology has been inaccurate, and organisms that were once considered two seperate species are actually one species.
-The morphological adaption that macroalgae display are therefore driven by local or site conditions, and therefore individuals must be morphologically flexible to persist in stochastic environments.
+The morphological adaption that macroalgae display are therefore mostly driven by local or site conditions, and therefore individuals must be morphologically flexible to persist in stochastic environments.
 This may be achieved through different strategies and are species-specific, which can be directly attributed the high diversity in morphological characters of algal communities.

### 2018-09-26 chapter_1.Rmd
more added to wave exposure section
chnages and additions to chaper 1 and 2, also updated bibliography

@@
 Advances in genetic techniques and taxonomy have revealed that species delination based on morphology has been inaccurate, and organisms that were once considered two seperate species are actually one species.**insert example here**.
-The morphological adaption that macroalgae display are therefore driven by local or site conditions, and therefore individuals must be morphologically flexible to persist in stochastic environments.
+The morphological adaption that macroalgae display are therefore driven by site conditions, therefore individuals must be morphologically flexible to persist in stochastic environments.
 This may be achieved through different strategies and are species-specific, which can be directly attributed the high diversity in morphological characters of algal communities.
 For example some algae have fronds and others are articulated corraline, and therefore these species would need to adapt their morphology differently in order to persist.
+In general, flat strap-like blades are common in areas that are exposed to high wave energy, while at proected sites blade morphology is wide and undulated.
+
 @moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed).
 The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid concentrations.
-**insert more examples here**.
 
@@
 This method has been used in mapping of marine coastal biodiveristy [@Rinde2004].
-
 @lindegarth2005 critized this approach, arguing that the choice of wave exposure method  can influence ecological inference.
@@
 Advances in numerical modelling have been founded on physical wave theory on how a wave "behaves".
-This approach is based on a theoretical persepctive rather than the need to answer ecological questions.
+This approach is based on a theoretical perspective rather than the need to answer ecological questions.
 Besides diffraction, these models incorporate more complexity by including wind forcing, wave-to-wave interactions and loss of energy due to friction and wave breaking.

### 2018-09-26 chapter_1.tex
chnages and additions to chaper 1 and 2, also updated bibliography

@@
 variable which varies spatially and temporarly in the marine enviroment.
-Furthermore, the degree to which a macroalgae community is exposed which
-depend on local site characteristics, such as bathymetry and local wind
-patterns. Despite this fact, macrolagae have been able to persist in
-often harsh and variable wave environments. Macroalage are sessile
+Furthermore, the degree to which a macroalgae community is exposed, is
+dependent on local site characteristics, such as bathymetry and local
+wind patterns. Despite this fact, macrolagae have been able to persist
+in often harsh and variable wave environments. Macroalage are sessile
 organisms and incapabable of migrating when local conditions become
@@
 in order to persist and survive, and achieve this through morphological
-variation.
+adaptation. The morphology of macroalgae are not fixed gentic traits. A
+study by @?? showed that \emph{species name} was able to rapidly adapt
+to the local wave environment when transplanted from a low wave exposure
+site to a high wave exposure site. Advances in genetic techniques and
+taxonomy have revealed that species delination based on morphology has
+been inaccurate, and organisms that were once considered two seperate
+species are actually one species.\textbf{insert example here}. The
+morphological adaption that macroalgae display are therefore driven by
+site conditions, therefore individuals must be morphologically flexible
+to persist in stochastic environments. This may be achieved through
+different strategies and are species-specific, which can be directly
+attributed the high diversity in morphological characters of algal
+communities. For example some algae have fronds and others are
+articulated corraline, and therefore these species would need to adapt
+their morphology differently in order to persist. In general, flat
+strap-like blades are common in areas that are exposed to high wave
+energy, while at proected sites blade morphology is wide and undulated.
 
-\begin{quote}
-keeping this for now @moss1948 investigated the anatomy, chemical
-composistion of \emph{Fucus spiralis} at three sites that varied in wave
-exposure (sheltered, medium exposure and exposed). The authors found
-that individuals in exposed sites showed less branching of thalli as
-well as variation in organic nitrogen, mannitol, laminarian and alginic
-acid concentrations.
-\end{quote}
+@moss1948 investigated the anatomy, chemical composistion of \emph{Fucus
+spiralis} at three sites that varied in wave exposure (sheltered, medium
+exposure and exposed). The authors found that individuals in exposed
+sites showed less branching of thalli as well as variation in organic
+nitrogen, mannitol, laminarian and alginic acid concentrations.
 
@@
 
-Traditional ecological measures of wave exposure usually incorportates
-integrative measures of hyrdodynamic conditions at a particular site.
-More specifically, it is the integration of mechanical processes and the
+Wave exposure may be modelled through various methods which range from
+simple cartographic to more advanced numerical wave models. Traditional
+ecological measures of wave exposure usually incorportates integrative
+measures of hyrdodynamic conditions at a particular site. More
+specifically, it is the integration of mechanical processes and the
 influence that is has on the ecology of nearshore communities.
-
-Wave exposure may be modelled through various methods which range from
-simple cartographic to more advanced numerical wave models.
 Cartographical models can be qualitative or quantitative and were

### 2018-09-26 chapter_2.Rmd
chnages and additions to chaper 1 and 2, also updated bibliography

@@
 The resilence of kelps is evident through their distribution, which spans a range of environmental conditions that vary spatially and temporarly.
-Wave exposure and temperature are regarded as important environmental drivers of kelps, which play a role in the growth, reproduction, and survival of kelp species.
+Wave exposure and temperature are regarded as important environmental drivers with regards to macroalgae, and play a role in the growth, reproduction, and survival of kelp species.
 Wave exposure is recognised as the main cause of kelp mortality, which may manifest itself through dislodgment of individuals or breakage of important morphological structures.
@@
 Kelps can either reduce drag, increase attachment strength or increase flexibility in order to survive high wave energy environments.
-**insert examples??*
 

### 2018-10-02 chapter_1.Rmd
minor edits
added section on wave theory to chapter 1. More to come.

@@
 Although studies have investigated the effects of important environmental drivers, the roles these factors play is often difficult to evaluate as such factors may never be fully independent of each other, i.e.
-environmental factors are to some extent dependent on one another.
+environmental factors are to some extent are dependent on one another.
 Multifactorial studies have attempted to explain combined affects, however these studies are often limited to investigating combination of two or three environmental drivers as inclusion of too many factors can lead to results that are difficult to interept.
@@
 
+# Ocean and coastal waves
+
+Regardless of the location around the world, waves are a feature of any coastline and the marine environment and are important manifestations of energy in the ocean.
+Waves are not the movement of water particles but is rather the movement and propagation of energy through the ocean.
+The source of this energy can be formed locally, such as wind-driven waves, and or from distant locations in the ocean, such as storms.
+This energy is then transferred from deeper water into shallower water where it plays a role in driving beach morphology.
+Sand and rock are eroded with each passing wave, thereby playing a role in shaping coastlines.
+Energy that is left-over after these processes is transferred into heat energy, and heats up the sand and rocks on which each wave hits.
+This energy can also be harnessed in simple ways such as a surfer catching a wave, or more complex ways such as capturing energy from the ocean environment for electricity production.
+However, waves are stochastic in nature and therefore the energy that propagates through the ocean is not always consistent.
+Therefore, it is often difficult to quantify and predict wave energy in the marine environment.
+Despite this fact, waves are gaining more recognition for the role they play in shaping coastlines, beaches and hence the communities and organisms that depend on these systems.
+
+Waves manifest themselves in different ways, which is also dependent on the energy creating force, and can be classified into different categories.
+For instance, "chop" are produced by local winds, while "tsunamis"" are rare waves that are formed during a earthquakes or landslides and can be produced from thousands of kilometers away.
+Waves in the ocean are known as "swell" and are produced from far distant storms, while the most consistent form of waves that interacts with the coastlines, usually twice a day, are called "tides".
+Tides are produced from differences in gravitational forces between the ocean, earth's crust, sun and the moon.
+These classifications can be broken down further and will be investigated later in this chapter.
+The focus of this chapter will be understanding wave theory and how waves form and propogate into shallow-coastal waters.
+This will be essential to later chapters that aim to model wave energy and callibrate the produced model with biotic parameters.
+
+## Generating and restoring forces
+
+Waves are formed due to the constant interaction bewteen two forces, which are know as the "generating source" and the "restoration force".
+The generating force is the force which pushes water from one layer up into the other, for example pushing water up across the boundry of air.
+This can also occur in the ocean layers.
+The ocean layers through differences in desnsity, which in turn is partly driven by abiotic parameters such as temperature and salinity, which creates a boundry for wave energy to move along.
+Examples of such boundries are along the ocean surface, pycnoclines, and the sea-floor.
+**optional addition of pycnocline details**.
+In other words, waves occur along the boundries formed in the ocean by various abiotic processes, and therefore density boundries are essential to the propagation of energy through the ocean.
+
+The generating force and restoring force occur along these boundries, the generating force pushes water up across ocean boundries while the restoring force pulls it back to where the boundry was orginally, trying to restore the balance in energy.
+This "tug of war" between the two forces creates an oscillating motion between the boundry layers which acts as a point of disturbance, sending out energy in all directions.
+The disturbance energy will continue to manifest itself provided the tug and pull between the two forces is occuring.
+Once the generating force stops, and ultimately the energy from the point of disturbance, the waves dissipate and water restores to its orginal state.
+A simlple example would be blowing on water in a cup.
+The blowing of air onto the surface of the water in the cup creates a point of disturbance that pushes the air boundry into the water boundry.
+The restoring force is the surface tension of the water that is maintained through hydrogen bonds between water molecules.
+Once blowing on the water in the cup has stopped, the ripples or "waves" in the cup begin to diminish in size until the surface tension returns to its orginal state.
+The hydrogen bonds between the water molecules are stronger than the force of gravity, and therefore the force of the surface tension returns the water to its orginal state.
+These waves are know as "capillary waves" and is essentially residual energy after the generating source has stopped.
+The restoring force can also be in the form of gravity and are know as "gravity waves".
+Using the same example, if one blows too hard, the water boundry is pushed up into the air boundry, breaking the hydrogen bonds between water molecules which allows gravity to return the water to its orginal state.
+All waves in the ocean are either capillary waves or gravity waves, and their classification will be dependent on the restoring force involved.
+
+In nature, there are three kinds of generating forces.
+These are wind, displacement of large volumes of water and uneven forces of gravitational attraction bewtween the Earth, Moon and Sun.
+Different generating forces are associated with different wave heights, periods and the type of wave produced **insert appropriate figure**.
+Wind-generated waves comprise of capillary waves, chop, swell and seiche.
+Some seiche can form from landslides and earthquakes but most of these waves are know as a tsunami.
+Swell has create the waves with the large heights (up to 100m), while tides can create the tallest.
+
+# Wave physics and scales
+
+Waves have a number of charateristics which are dipected in figure ??, and is an idealised representation of what a wave is.
+The *amplitude* is the vertical distance from its midline or *equilibrium surface* or *still water level* to its highest point known as the *crest*.
+The *trough* is the same distance of the amplitude, however the measurement is taken from the equilibrium surface to the lowest point.
+
 # Hydrodynamic modelling

### 2018-10-02 chapter_1.Rmd
moved bibliograpghy file to LaTeX folder. Added to wave theory section. Updated pdf and word outputs.

@@
     toc: yes
+bibliography: LaTeX/mycollection.bib
 ---
 
-## Summary
+\newpage
+
+# Summary
 
@@
 
-## Background
+\newpage
+
+# Background
 
@@
 
-#Light
+##Light
 
 Light is an important factor for kelp survival, however if light is limited or excessive this may negatively impact kelp survival or growth.
-Much of the past research into the role light plays into the functioning of kelp [@Bruhn1996);@Belseth2012].
-For instance, solar ultraviolate radiation has been shown to aﬀect sub-canopy Ecklonia radiata sporophytes when the canopy of mature *Ecklonia radiata* was removed [@Wood 1987].
-The sub-canopy sporophytes experienced tissue damage,photopigment destruction,reduced growthand decreased survivorship, thus inhibiting their settlement and survival [@Wood 1987].
-Laboratory experiments revealed that the UV component of radiation, rather than intense radiation itself, was responsible for the eﬀects mentioned above.
-High light stress has negative eﬀects, such as photoinhibition and photo-damage on *Ecklonia cava* sporophytes [@Altamirano2004].
+Much of the past research into the role light plays into the functioning of kelp [@Bruhn1996;@Belseth2012].
+For instance, solar ultraviolate radiation has been shown to affect sub-canopy Ecklonia radiata sporophytes when the canopy of mature *Ecklonia radiata* was removed [@Wood1987].
+The sub-canopy sporophytes experienced tissue damage,photopigment destruction,reduced growthand decreased survivorship, thus inhibiting their settlement and survival [@Wood1987].
+Laboratory experiments revealed that the UV component of radiation, rather than intense radiation itself, was responsible for the effects mentioned above.
+High light stress has negative effects, such as photoinhibition and photo-damage on *Ecklonia cava* sporophytes [@Altamirano2004].
 @Altamirano2004 found that *Ecklonia cava* is more vulnerable to light stress conditions, and less likely to recover under unfavourable conditions [@Altamirano2004].
-@Bolton1985 showed that under sub-saturating irradiances and supra- optimal temperatures *Ecklonia maxima* to showed a decrease in reproductive rates and an increase in cell production.
-An additional ﬁnding of this study was that despite the decrease in reproductive rates, the ﬁnal egge production per female was greater under these conditons.
-The authors interpreted this an ecological adaption that may increase survival rates under times of stress or non - ideal conditions [@Bolton1985].
+@Bolton1985a showed that under sub-saturating irradiances and supra- optimal temperatures *Ecklonia maxima* to showed a decrease in reproductive rates and an increase in cell production.
+An additional finding of this study was that despite the decrease in reproductive rates, the final egge production per female was greater under these conditons.
+The authors interpreted this an ecological adaption that may increase survival rates under times of stress or non - ideal conditions [@Bolton1985a].
 
-#Substrata
-#Salinity
-#Depth
+##Substrata
+##Salinity
+##Depth
 
-Depth does not aﬀect kelp ecosystems directly, however a change in depth often causes ﬂuctuations or changes in other environmental variables such as water motion, light and temperature.
+Depth does not affect kelp ecosystems directly, however a change in depth often causes fluctuations or changes in other environmental variables such as water motion, light and temperature.
 Water motion also decreases with depth, and some kelps better suited to deeper environments (*L.
@@
 Temperature may also change along a depth gradient due to a reduction in sunlight penetration [@Dayton1985; @Gerard1982].
-Therefore depth does not directly play a role in kelp functioning but may alter more inﬂuential factors such as light and water motion.
+Therefore depth does not directly play a role in kelp functioning but may alter more influential factors such as light and water motion.
 
-#Sedimentation
-#Nutrients
+##Sedimentation
+##Nutrients
 
@@
 Additionally, some kelps have the ability to store inorganic nitrogen in order to compensate for periods of low nutrient availability, which has been observed for Laminaria and Macrocystis [@Dayton1985; @Gaylord2012].
-Nutrient stratiﬁcation is also an important factor, particularly for canopy type kelps.
+Nutrient stratification is also an important factor, particularly for canopy type kelps.
 The concentration of nutrients at the surface is important to the functioning and maintenance of the canopy.
@@
 This is often an indication of an “up-welling” event, which brings cold and nutrient rich waters from the bottom to the surface of the water column.
-Temperature can play a direct role in the uptake of nutrients through eﬀects on algal metabolism; however this may vary from species to species [@Raven1988].
+Temperature can play a direct role in the uptake of nutrients through effects on algal metabolism; however this may vary from species to species [@Raven1988].
 
-# Temperature
+##Temperature
 
 Temperature is a driver of kelp species distributions and ecophysiological processes, as well as a lesser role in morphological adaption...**example here**...The majority of kelp species are artic and temperate organisms, and the warming of ocean temperatures is expected to cause a poleward biogeographical shift of species [@Bolton2012].
-There is evidence to suggest that South African kelp forests are expanding due to ocean cooling [@Bolton2012], possibly driven by an intensiﬁcation and increase in coastal upwelling [@Blamey2012, @Blamey2015].
+There is evidence to suggest that South African kelp forests are expanding due to ocean cooling [@Bolton2012], possibly driven by an intensification and increase in coastal upwelling [@Blamey2012, @Blamey2015a].
 In South Africa there has been a biogeographical shift eastward along the coast due to a change in inshore temperature regime, making South Africa no exception to changing ocean temperatures [@Bolton2012].
@@
 A study by @Biskup2014 investigated the functional response of two kelp species (*Laminaria ochroleuca* and *Saccorhiza polyschides*) to rising sea temperatures.
-The functional responses of Saccorhiza polyschides was measured for both the subtidal and intertidal habitats, to see what aﬀect non- optimal conditions (intertidal zone) had on the kelps [@Rinde2005].
+The functional responses of Saccorhiza polyschides was measured for both the subtidal and intertidal habitats, to see what affect non- optimal conditions (intertidal zone) had on the kelps [@Rinde2005].
 The study found that Laminaria ochroleuca exhibited a poor ability to acclimatise and was dependent on the kelp’s life history traits [@Biskup2014].
-Therefore annual kelp species are more likely to survive under non-ideal condition, and the intertidal Saccorhiza polyschides, compared to the subtidal, showed a higher physiological ﬂexibility to changing conditions [@Biskup2014].
-This may be because the intertidal zone undergoes far more change than the subtidal and therefore kelps in the intertidal are forced to adapt to harsher conditions where ﬂuctuations in temperature, sunlight, turbidity and water motion are common.
-The eﬀects on temperature have also been investigated by @Wernberg2010.
+Therefore annual kelp species are more likely to survive under non-ideal condition, and the intertidal Saccorhiza polyschides, compared to the subtidal, showed a higher physiological flexibility to changing conditions [@Biskup2014].
+This may be because the intertidal zone undergoes far more change than the subtidal and therefore kelps in the intertidal are forced to adapt to harsher conditions where fluctuations in temperature, sunlight, turbidity and water motion are common.
+The effects on temperature have also been investigated by @Wernberg2010.
 The study looked at resilience of kelp beds along a latitudinal temperature gradient.
-Kelp abundance is likely to decline with the predicted warming of ocean waters @Wernberg2010 and although kelps have the ability to acclimatize and adjust their metabolic performance, which in turn allows them to change their physiological performance to mitigate the seasonal ﬂuctuations in temperature, this acclimatization is done at a cost @Wernberg2010...**link to paragraph on kelp morphology**...
+Kelp abundance is likely to decline with the predicted warming of ocean waters @Wernberg2010 and although kelps have the ability to acclimatize and adjust their metabolic performance, which in turn allows them to change their physiological performance to mitigate the seasonal fluctuations in temperature, this acclimatization is done at a cost @Wernberg2010...**link to paragraph on kelp morphology**...
 
-# Wave exposure
+##Wave exposure
 
@@
 
-# Ocean and coastal waves
+#Ocean and coastal waves
 
@@
 
-# Wave physics and scales
+## Wave physics and scales
 
@@
 The *amplitude* is the vertical distance from its midline or *equilibrium surface* or *still water level* to its highest point known as the *crest*.
+The equalibrium surface is the level the ocean would be if there were no waves, for a wave to form a disturbance must occur below or above this line.
 The *trough* is the same distance of the amplitude, however the measurement is taken from the equilibrium surface to the lowest point.
+Wave height is the vertical distance from crest to trough, and is equal to twice the the amplitude.
+The *wavelength* is the horizontal distance from a crest/trough to the next crest/trough respectively.
+It is important to note that the energy progpgating from a disturbance will not reach the ocean floor in deep-water environments.
+The depth below a wave where the water, and anything in the water, feels no motion or disturbance is known as the *wave base*.
+The wave base is calculated by descending vertical from the equalibrium surface by a value equal to halve the wavelength.
 
-# Hydrodynamic modelling
+As mentioned previously, the water particles in the ocean move in a circular orbit and hence return to their orginal position.
+This is because waves in the ocean represent moving energy, not moving water.
+
+##Hydrodynamic modelling
 
@@
 The authors also highlighted the need for objective, reproducible and quantitative studies comparing exposure indices [@lindegarth2005].
-A study by @sundbald...**include description of study here**...Other authors, such as @hill2010, have argued that these simple measures can be improved upon by including bathymetery data which allows the incorporation of diffraction into the calculation.
+Other authors, such as @hill2010, have argued that these simple measures can be improved upon by including bathymetery data which allows the incorporation of diffraction into the calculation.
 Diffraction is topographically induced variations in wave direction.
@@
 
-## Aims of research
+#Aims of research
 
@@
 C) Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.
+
+#References

### 2018-10-03 chapter_1.Rmd
minor changes to chapter 1

@@
 output:
-  pdf_document:
-    highlight: zenburn
-    number_sections: yes
+  word_document:
     toc: yes
@@
     toc: yes
-  word_document:
+  pdf_document:
+    highlight: zenburn
+    number_sections: yes
     toc: yes
@@
 The depth below a wave where the water, and anything in the water, feels no motion or disturbance is known as the *wave base*.
-The wave base is calculated by descending vertical from the equalibrium surface by a value equal to halve the wavelength.
-
-As mentioned previously, the water particles in the ocean move in a circular orbit and hence return to their orginal position.
+The wave base is calculated by descending vertical from the equalibrium surface by a value equal to halve the wavelength.The water particles in the ocean move in a circular orbit and hence return to their orginal position.
 This is because waves in the ocean represent moving energy, not moving water.
 
-##Hydrodynamic modelling
+##Types of waves
+
+There are a variety of waves that form in the ocean and all differ in terms of period or wavelength (??image??).
+The longest wave that can form in the ocean are known as *trans-tidal waves*, and are generated by fluctuations in magnitism between the Earth's crust and atmosphere.
+The magnetic push and pull between the Moon and the Sun creates waves with a slightly shorter wavelength, known as tides.
+Their period and wavelength can also range from a few hours to more than a day, and from a few hundred to a thousand kilometers respectively.
+*Storm surge* tend to have a slightly shorter wavelength and period compared to tides.
+When low atmospheric pressure systems and high wind speeds in a storm it elevates the ocean surface, generating storm surge, which may cause flooding in coastal areas as it approaches the coastline.
+*Tsunami's* are on the lower end of the scale, and are generated by earhquakes or submarine 'land' slides.
+Their random nature makes them difficult to predict and increase amplitude as the approach the coast.
+This can make them considerbly large along coastal areas and often cause immense damage and loss of life.
+
+# Hydrodynamic modelling
 

### 2018-10-03 chapter_2.Rmd
updated outputs and house cleaning for chapter 2

@@
 Kelps are sessile species and have to adapt to local environmental conditions in order to persist and survive.
-The resilence of kelps is evident through their distribution, which spans a range of environmental conditions that vary spatially and temporarly.
+The resilience of kelps is evident through their distribution, which spans a range of environmental conditions that vary spatially and temporally.
 Wave exposure and temperature are regarded as important environmental drivers with regards to macroalgae, and play a role in the growth, reproduction, and survival of kelp species.
@@
 These mechanisms usually occur during times of high wave energy or pulse disturbance events, such as storms.
-Kelps are able to reduce the probablity of dislodgment or breakage through morphological adaptation, which may involve several strategies.
+Kelps are able to reduce the probability of dislodgment or breakage through morphological adaptation, which may involve several strategies.
 Kelps can either reduce drag, increase attachment strength or increase flexibility in order to survive high wave energy environments.

### 2018-10-03 proposal_draft_2018.Rmd
chapter outline added and references updated. Word and pdf outputs updated. Note that the word out is not great...

@@
     toc_depth: 4
-bibliography: mycollection.bib   
+bibliography: LaTeX/mycollection.bib   
 ---
@@
 BioEx is calculated as the sum of the index developed at different spatial scales (local, fjord and open).
-This method has been used in mapping of marine coastal biodiveristy [@Rinde2004].
 
@@
 A) Determine if the hydrodynamic environment is the main driver of kelp morphology and if this is specific to a location
-B) Simulate kelp rafting by means of a hydrodynamic modelling and calibrate this model with *in situ* beach-cast morphometric data.
-C) Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.
+B) Determine quantitative relationships with kelp morphology and the hydrodynamic environment to create a biological index.
+C) Use *in situ* morphological data to calibrate this to a specific area, or ideally site, within a reasonable amount of probability
+D) Simulate kelp rafting by means of a hydrodynamic modelling and calibrate this model with *in situ* beach-cast morphometric data.
+E) Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.
 
@@
 These models will consist of the SWAN and Delf3D models and the apporach is to use aspects of each model, that are related to coastal flow, from which further development will take place.
-
 Once the model has been developed it will be used to predict kelp morphology at random sites which will then be sampled.
@@
 Once the model has been established it will once again be calibrated through _in situ_ sampling.
+
+## Outline of Chapters
+
+Chapter and sections| Title
+-----------------------|------------------------------------------------------------------------
+Chapter 1 | Summary
+Chapter 1.1| Background
+Chapter 1.2| Kelp environmental drivers
+Chapter 1.3| Ocean and coastal waves
+Chapter 1.4| Hydrodynamic modelling
+Chapter 1.5| Aims and objectives
+References | Reference list for chapter 
+Chapter 2 | Abiotic drivers of morphological adaption of *Ecklonia maxima* along the South African coastline
+Chapter 2.1 | Introduction 
+Chapter 2.1.2| Waves and kelp morphology 
+Chapter 2.1.3| The South African context 
+Chapter 2.2 | Aims and objectives
+Chapter 2.3 | Methods
+Chapter 2.4 | Results
+Chapter 2.5 | Discussion
+References | Reference list for chapter
+Chapter 3 | Using kelp morphology as an index for calibrating coastal-flow model
+Chapter 3.1 | Introduction
+Chapter 3.1.2 | The SWAN model design and application
+Chapter 3.1.3 | SWAN as a platform for further development
+Chapter 3.2 | Aims and objectives
+Chapter 3.3 | Methods
+Chapter 3.4 | Results
+Chapter 3.5 | Discussion
+References | Reference list for chapter
+Chapter 4 | Application of modified coastal-flow model
+Chapter 4.1 | Introduction
+Chapter 4.1.2 | Plastic plloution in the marine environment
+Chapter 4.1.3 | Application of hydrodynamic modelling in dispersal of plastics
+Chapter 4.1.4 | Plastic pollution in SOuth Africa
+Chapter 4.2 | Aims and objectives
+Chapter 4.3 | Methods
+Chapter 4.4 | Results
+Chapter 4.5 | Discussion
+References | Reference list for chapter
+Chapter 5 | Synthesis
+Chapter 5.1 | Synthesis of Chapters
+Chapter 5.2 | Challenges and limitations
+Chapter 5.3 | Gaps and future research
+Chapter 5.4 | Concluding remarks
 

### 2018-10-09 chapter_2.Rmd
added toc, updated outputs

@@
     fig_width: 7
-    toc: no
+    toc: yes
   html_notebook:

### 2018-10-16 chapter_2.Rmd
moved map to intro as suggested. paragraph added

@@
 
-The aim of this study is, therefore, to understand how environmental drivers such as temperature and wave energy can influence morphoplasticity in two species of kelps around South Africa.
-This will be achieved by initially understanding the variation in abiotic parameters, and morphometrics of _Ecklonia maxima_ and _Laminaria pallida_, around the Western Cape coast.
-Thereafter we will look at how the abiotic parameters both correlate and influence each other in the nearshore environment of our study region.
-Finally, we will investigate which abiotic parameters best describe the morphological variation of the two kelp species, using statistical tools.
-It is predicted that higher wave energy environments will exhibit kelps with reduced undulations, possessing strap-like blades, opposed to low flow regimes that will show greater undulations in frond morphology.
-An increase in temperature is predicted to influence specific morphological traits that rely on nutrient uptake, as temperature and nutrients share an inverse relationship.
-An increase in temperature is also predicted to reduce the dominance of the colder water kelp, _Laminaria pallida_, at depth, with a total absence of _Laminaria pallida_ as one approaches the eastern boundary of sample sites.
-
-### Study area
-
-Due to the Cape Peninsula’s temperate latitude, winter months bring an increased frequency of frontal depressions that originate from the Southern Ocean [@Reason2006].
-These low pressures are joined by large swells with increased wave energy.
-The nearshore environment, with the accompanied biota, therefore experiences high wave energy events, with increased frequency in winter.
-The large peninsula acts as an obstruction for large south westerly swells, providing decreased wave energy along the west side of False Bay [@Shipley1964].
-Conversely, the west coast of Cape Point is battered by these large swells at full force.
-Multiple sites, therefore, exist where kelps grow in diverse temperature and wave energy climates, in close proximity.
-
-\newpage
-## Methods
-### Morphometrics collection
-
-Between October 2014 and April 2015, morphological measurements of _Laminaria pallida_ and _Ecklonia maxima_ were collected at 18 sites along the Western Cape coast of South Africa (Fig.
-1).
-Eleven samples were collected per morphology, for each species (Table 1, 2).
-These varying morphometrics allowed measurements such as weight, length and thickness to be compared between sites.
-Because the macroalgae differ in morphological features, species-specific morphometrics were included.
-These sites span across the majority of the south-west coast, in varying thermal and wave energy regimes.
-
-```{r, echo= FALSE, message=FALSE, warning=FALSE}
-library(knitr)
-morph <- read.csv("tables/Morphology table.csv")
-
-colnames(morph) <- c("Morphometric", "Unit measurement")
-
-morph_eck <- morph[(1:10), ]
-morph_lam <- morph[(11:19), ]
-
-```
-```{r, echo= FALSE, message=FALSE, warning=FALSE}
-
-kable(morph_eck, row.names = FALSE, caption = "A list of morphology measurements that were collected to investigate how environmental drivers may influence the morphology of _Ecklonia maxima _.
-Units for each morphometric are included.")
-
-```
-```{r, echo= FALSE, message=FALSE, warning=FALSE}
-
-kable(morph_lam, row.names = FALSE, caption = "A list of morphology measurements that were collected to investigate how environmental drivers may influence the morphology of _Laminaria pallida_.
-Units for each morphometric are included.")
-
-```
-
-### Sites
-
-Sites were chosen to represent an array of morphological differences seen within _Ecklonia maxima_ and _Laminaria pallida_.
-Sites were also chosen to reflect locations that displayed variable wave and temperature regimes (Fig.
-2-5), to allow us to robustly test our hypothesis of environmental drivers influencing kelp morphology.
-St.
-Helena Bay and Betty’s Bay constituted the north western and south eastern boundary sites respectively.
-These sites are roughly 300km apart, and lie within separate marine provinces, as outlined above.
-The Cape Peninsula provides an interesting topographical boundary that shelters the coast in False Bay.
-Sites were therefore chosen to represent an array of environments, from offshore reefs (Batsata Rock), to sheltered intertidal zones (Miller’s Point).
-West of Cape Point, a number of sites were chosen to highlight the presence of upwelling (Oudekraal, Kommetjie), as well as kelps growing in protected bays (Hout Bay).
-
 ```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="A map indicating the sites where morphometric measurements were collected for _Ecklonia maxima_ and _Laminaria pallida_.", fig.height=8, fig.width=12}
@@
 ```
+
+The aim of this study is, therefore, to understand how environmental drivers such as temperature and wave energy can influence morphoplasticity in two species of kelps around South Africa.
+This will be achieved by initially understanding the variation in abiotic parameters, and morphometrics of _Ecklonia maxima_ and _Laminaria pallida_, around the Western Cape coast.
+Thereafter we will look at how the abiotic parameters both correlate and influence each other in the nearshore environment of our study region.
+Finally, we will investigate which abiotic parameters best describe the morphological variation of the two kelp species, using statistical tools.
+It is predicted that higher wave energy environments will exhibit kelps with reduced undulations, possessing strap-like blades, opposed to low flow regimes that will show greater undulations in frond morphology.
+An increase in temperature is predicted to influence specific morphological traits that rely on nutrient uptake, as temperature and nutrients share an inverse relationship.
+An increase in temperature is also predicted to reduce the dominance of the colder water kelp, _Laminaria pallida_, at depth, with a total absence of _Laminaria pallida_ as one approaches the eastern boundary of sample sites.
+
+### Study area
+
+Due to the Cape Peninsula’s temperate latitude, winter months bring an increased frequency of frontal depressions that originate from the Southern Ocean [@Reason2006].
+These low pressures are joined by large swells with increased wave energy.
+The nearshore environment, with the accompanied biota, therefore experiences high wave energy events, with increased frequency in winter [@Veitch2018].
+The large peninsula acts as an obstruction for large south westerly swells, providing decreased wave energy along the west side of False Bay [@Shipley1964].
+Conversely, the west coast of Cape Point is battered by these large swells at full force.
+Multiple sites, therefore, exist where kelps grow in diverse temperature and wave energy climates, in close proximity.
+
+In Figure 1 the total and seasonal wave roses of the directional wave buoy just offshore from Cape Town is given form the years 2000 till end 2017 (the period for which directional wave spectra was available).
+It is due to the near consistent South-westerly swell waves and the complex orography around the peninsula, that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.
+For details on the seasonality of the wave climate around the Cape peninsula please refer to [@Veitch2018].
+
+**insert wave rider images here**
+
+\newpage
+## Methods
+### Morphometrics collection
+
+Between October 2014 and April 2015, morphological measurements of _Laminaria pallida_ and _Ecklonia maxima_ were collected at 18 sites along the Western Cape coast of South Africa (Fig.
+1).
+Eleven samples were collected per morphology, for each species (Table 1, 2).
+These varying morphometrics allowed measurements such as weight, length and thickness to be compared between sites.
+Because the macroalgae differ in morphological features, species-specific morphometrics were included.
+These sites span across the majority of the south-west coast, in varying thermal and wave energy regimes.
+
+```{r, echo= FALSE, message=FALSE, warning=FALSE}
+library(knitr)
+morph <- read.csv("tables/Morphology table.csv")
+
+colnames(morph) <- c("Morphometric", "Unit measurement")
+
+morph_eck <- morph[(1:10), ]
+morph_lam <- morph[(11:19), ]
+
+```
+```{r, echo= FALSE, message=FALSE, warning=FALSE}
+
+kable(morph_eck, row.names = FALSE, caption = "A list of morphology measurements that were collected to investigate how environmental drivers may influence the morphology of _Ecklonia maxima _.
+Units for each morphometric are included.")
+
+```
+```{r, echo= FALSE, message=FALSE, warning=FALSE}
+
+kable(morph_lam, row.names = FALSE, caption = "A list of morphology measurements that were collected to investigate how environmental drivers may influence the morphology of _Laminaria pallida_.
+Units for each morphometric are included.")
+
+```
+
+### Sites
+
+Sites were chosen to represent an array of morphological differences seen within _Ecklonia maxima_ and _Laminaria pallida_.
+Sites were also chosen to reflect locations that displayed variable wave and temperature regimes (Fig.
+2-5), to allow us to robustly test our hypothesis of environmental drivers influencing kelp morphology.
+St.
+Helena Bay and Betty’s Bay constituted the north western and south eastern boundary sites respectively.
+These sites are roughly 300km apart, and lie within separate marine provinces, as outlined above.
+The Cape Peninsula provides an interesting topographical boundary that shelters the coast in False Bay.
+Sites were therefore chosen to represent an array of environments, from offshore reefs (Batsata Rock), to sheltered intertidal zones (Miller’s Point).
+West of Cape Point, a number of sites were chosen to highlight the presence of upwelling (Oudekraal, Kommetjie), as well as kelps growing in protected bays (Hout Bay).
 

### 2018-10-16 proposal_draft_2018.Rmd
house cleaning

@@
     toc: yes
-  pdf_document:
-    toc: yes
-    toc_depth: 4
 bibliography: LaTeX/mycollection.bib   

### 2018-10-18 chapter_2.Rmd
more of Christo's additions added. some minor glitches in transferring additions have been correct. images to follow

@@
 
+In Figure 2 the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
+These values have been computed using model output from approximately 20 year of simulated nearshore wave conditions.
+Hindcast global wave parameters were refracted into the nearshore via a spectral phase-averaged numerical wave modelling code SWAN (Simulating, WAves in the Nearshore - 3rd generation).
+The model is fully spectral in all frequencies and directions (0o -360o) (can solve swell and local wind generated waves simultaneously, propagating in different direction) and was used to model the propagation of short crested swell waves, waves generated by wind, non-linear wave-wave interaction and dissipation and depth induced breaking and refraction.
+Wave energy dissipation due to bottom friction and whitecapping is also accounted for (Deltares, 2014).
+
+Time series output from 1997 till 2013 of this model was then produced at the 7m and 15m depth contour lines.
+Each nearshore output point thus contained a 17-year timeseries of significant wave heights, peak periods and mean wave directions (with both swell and wind waves).
+This data set is available on a 500m resolution around most of the South African coastline and on a 200m resolution with False Bay as provided by the South African Department of Environmental Affairs (DEA) and Council for Scientific and Industrial Research (CSIR)(@heron2014).
+The total 17-year average energy, for each output location, was then computed using linear wave theory, taking into account deep, intermediate and shallow water wave approximations (based on wave length) for the wave energy (Holthuijsen, 2007; Kasiulis, Punys and Kofoed, 2015; Foteinis *et al.*, 2017) which is a function of wave height and period.
+
+The directional sheltering effect of the Cape Peninsula, against the dominant swell direction, (given in Figure 1) is clear observed in the wave exposures presented in Figure 2.
+The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits.
+It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world (Norderhaug *et al.*, 2012; Liberti, Carillo and Sannino, 2013; Sundblad *et al.*, 2014).
+Kelp and other ecosystems adapt to their typical environmental exposure and thus what is classified as extreme abiotic coastal forcings (temperature, winds, waves and currents) will vary around the world depending on the geographical setting.
+
+From Figure 2 the western periphery of the Cape peninsula is almost continuously producing high coastal wave exposures while the Eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures.
+Here the marked seasonality, with higher energy waves during winter, may be clearly observed once more.
+
+*Figure 2: Coastal wave energy exposure around the Cape Peninsula for the total 17-year nearshore wave record, (b) only winter months and (c) only summer months. *
+
+In Figure 3 the propagation of a typical offshore wave spectrum is given as produced from a single time-step in SWAN.
+Figure 3 is presented to clarify the averaged wave exposure maps presented in Figure 2.
+Tracing the wave height contours into False Bay its clear why this Bay’s western periphery is predominantly sheltered.
+It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions.
+This results in positive and negative wave exposure anomalies all around the peninsula.
+
+*Figure 3: SWAN numerical output for an example typical offshore wave conditions refracting around the Cape Peninsula.  *
+
 \newpage

### 2018-10-18 chapter_2.Rmd
christo's additional figures added. all figures have been converted to png format. All images can be located in the "figures" folder. PDF output updated.

@@
 output:
+  pdf_document:
+    toc: yes
+  html_notebook:
+    fig_caption: yes
+    toc: yes
   word_document:
@@
     fig_width: 7
-    toc: yes
-  html_notebook:
-    fig_caption: yes
-    toc: yes
-  pdf_document:
     toc: yes
@@
 
-**insert wave rider images here**
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Wave roses of the Cape Point wave rider buoy situated at 18.29oE, 34.2oS in 70m water depth.
+(a) total wave record, (b) winter waves and (c) summer waves."}
+
+library(png)
+library(grid)
+library(gridExtra)
+
+img_twr <-  rasterGrob(as.raster(readPNG("figures/total_wave_rose_2.png")), interpolate = FALSE)
+img_wwr <-  rasterGrob(as.raster(readPNG("figures/winter_wave_rose_2.png")), interpolate = FALSE)
+img_swr <-  rasterGrob(as.raster(readPNG("figures/summer_wave_rose_2.png")), interpolate = FALSE)
+
+grid.arrange(img_twr, img_wwr,img_swr, ncol = 3)
+
+```
 
@@
 The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits.
-It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world (Norderhaug *et al.*, 2012; Liberti, Carillo and Sannino, 2013; Sundblad *et al.*, 2014).
+It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world [Norderhaug2012; @Liberti2013; @Sundblad2014].
 Kelp and other ecosystems adapt to their typical environmental exposure and thus what is classified as extreme abiotic coastal forcings (temperature, winds, waves and currents) will vary around the world depending on the geographical setting.
@@
 
-*Figure 2: Coastal wave energy exposure around the Cape Peninsula for the total 17-year nearshore wave record, (b) only winter months and (c) only summer months. *
+```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Coastal wave energy exposure around the Cape Peninsula for the (a) total 17-year nearshore wave record, (b) only winter months and (c) only summer months."}
+
+img_we_total <-  rasterGrob(as.raster(readPNG("figures/coastal_we_total.png")), interpolate = FALSE)
+img_we_winter <-  rasterGrob(as.raster(readPNG("figures/coastal_we_winter.png")), interpolate = FALSE)
+img_we_summer <-  rasterGrob(as.raster(readPNG("figures/coastal_we_summer.png")), interpolate = FALSE)
+
+grid.arrange(img_we_total, img_we_winter,img_we_summer, ncol = 2,nrow = 2)
+
+```
 
@@
 
-*Figure 3: SWAN numerical output for an example typical offshore wave conditions refracting around the Cape Peninsula.  *
+```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="SWAN numerical output for an example typical offshore wave conditions refracting around the Cape Peninsula."}
+
+knitr::include_graphics('figures/regional_waves.png')
+
+```
 

### 2018-10-18 chapter_2.Rmd
sorted out issue of multiple graphs plotting for temperature

@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Wave roses of the Cape Point wave rider buoy situated at 18.29oE, 34.2oS in 70m water depth.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Wave roses of the Cape Point wave rider buoy situated at 18.29oE, message=FALSE, warning=FALSE, 34.2oS in 70m water depth.
 (a) total wave record, (b) winter waves and (c) summer waves."}
@@
 
-grid.arrange(img_twr, img_wwr,img_swr, ncol = 3)
+grid.arrange(img_twr, img_wwr,img_swr, ncol = 2, nrow = 2)
 
@@
 
-kable(morph_eck, row.names = FALSE, caption = "A list of morphology measurements that were collected to investigate how environmental drivers may influence the morphology of _Ecklonia maxima _.
+t1 <- kable(morph_eck, row.names = FALSE, caption = "A list of morphology measurements that were collected to investigate how environmental drivers may influence the morphology of _Ecklonia maxima _.
 Units for each morphometric are included.")
@@
 
-kable(morph_lam, row.names = FALSE, caption = "A list of morphology measurements that were collected to investigate how environmental drivers may influence the morphology of _Laminaria pallida_.
+t2 <- kable(morph_lam, row.names = FALSE, caption = "A list of morphology measurements that were collected to investigate how environmental drivers may influence the morphology of _Laminaria pallida_.
 Units for each morphometric are included.")
@@
 ```
-
 ```{r, echo= FALSE, message=FALSE, warning=FALSE, results='hide'}
@@
 ```
-
 ```{r, echo= FALSE, message=FALSE, warning=FALSE}
@@
 ```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Morphometric site locations represented by various temperature parameters.
-Temperature parameters include minimum, maximum and mean temperatures (° Celsius).
+Temperature parameters include minimum, maximum, mean, range and standard deviation (° Celsius).
 These site locations are coloured by the temperature statistic relative to the legends provided.
@@
 
-grid.arrange(min, max, mean, ncol = 1)
-ggarrange(min, max, mean, ncol = 1, nrow = 1, labels = "AUTO", legend = "bottom", common.legend = TRUE, align = "none", heights = c(1.5,1))
-```
-
-```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Morphometric site locations represented by various temperature parameters.
-Temperature parameters include temperature range as well as standard deviation (° Celsius).
-These site locations are coloured by the temperature statistic relative to the legends provided.
-Each temperature parameter is also divided into August (winter) and February (summer) as well as the annual mean.", fig.height=8, fig.width=12, fig.pos="H"}
-
-grid.arrange(range, sd, ncol = 1)
-
-```
+grid.arrange(max,min,mean,range, sd, ncol = 2)
+
+```
+
+### Wave and wind parameters
 
@@
 
-## MEthod 2 (shows individual legends, still squashed)
+##  (shows individual legends, still squashed)
 
@@
 
-### Wave and wind parameters
-
 Not only do the temperature parameters display seasonal fluctuations, but we see similar observations for wave parameters.
@@
 
-grid.arrange(medianDir1, meanHs1, sdHs1, meanTp1, sdTp1)
+grid.arrange(medianDir1, meanHs1, sdHs1, meanTp1, sdTp1, ncol = 2)
 
@@
 
-grid.arrange(medianDirw1, meanSpw1, sdSpw1)
+grid.arrange(medianDirw1, meanSpw1, sdSpw1, ncol = 2)
 
@@
 # Order site names by location west to east
+
 positionsEck <- c("Betty's Bay", "Kalk Bay", "A-Frame", 

### 2018-10-19 chapter_2.Rmd
house keeping

@@
 output:
-  pdf_document:
+  word_document:
+    fig_caption: yes
+    fig_height: 7
+    fig_width: 7
     toc: yes
@@
     toc: yes
-  word_document:
-    fig_caption: yes
-    fig_height: 7
-    fig_width: 7
+  pdf_document:
     toc: yes
@@
 ```
+
 ### Temperature parameters

### 2018-10-19 chapter_2.Rmd
some more house cleaning

@@
 Outside of these latitudes, they are also found in cool, deep water towards the tropics [@Graham2007]: this deeper tropical water, due to its water clarity, also allows for suitable photon flux and a nitrate concentration [@Luning1990; @Dayton1999; @Zimmerman1984], the two other environmental variables that support kelp productivity.
-Although kelps have a low taxonomic diversity, their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude other turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Dunton1987; @Duggins1989; @Steneck2002], each depending on a wide suite of ecological services provided by the kelp habitat [@Gaines1987; @Bologna1993; @Levin1994; @Anderson1997].
+Although kelps have a low taxonomic diversity, their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude other turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Duggins1989; @Steneck2002], each depending on a wide suite of ecological services provided by the kelp habitat [@Gaines1987; @Bologna1993; @Levin1994; @Anderson1997].
 
@@
 The biogeographic distribution of kelp is limited by seawater temperature [@Bolton2010], where increasing temperature gradients reduce kelp distribution.
-Due to this limiting factor, the two main species of kelps in southern African waters, _Ecklonia maxima_ and _Laminaria pallida_, are distributed along a section of the south coast from De Hoop, extending west around the Cape Peninsula, and thriving north into Namibia [@Molloy1990; @Stegenga1997].
+Due to this limiting factor, the two main species of kelps in southern African waters, _Ecklonia maxima_ and _Laminaria pallida_, are distributed along a section of the south coast from De Hoop, extending west around the Cape Peninsula, and thriving north into Namibia [@Molloy1996; @Stegenga1997].
 This distribution follows a temperate gradient, where sea temperatures increase as one moves south from Namibia, around Cape Point and towards De Hoop.
@@
 Across the majority of the coastline, _Laminaria pallida_ remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 meters [@Field1980b; @Bolton1987; @Molloy1996].
-This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 meters [@Field1980b; @Molloy1990; @Molloy1996; @Stegenga1997].
+This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 meters [@Field1980b; @Molloy1996; @Molloy1996; @Stegenga1997].
 Towards the north along the west coast, from around Hondeklipbaai, _Laminaria pallida_ replaces _Ecklonia maxima_ as the dominant kelp species [@Velimirov1977; @Stegenga1997] and it also occupies increasingly shallower subtidal regions.
@@
 The model is fully spectral in all frequencies and directions (0o -360o) (can solve swell and local wind generated waves simultaneously, propagating in different direction) and was used to model the propagation of short crested swell waves, waves generated by wind, non-linear wave-wave interaction and dissipation and depth induced breaking and refraction.
-Wave energy dissipation due to bottom friction and whitecapping is also accounted for (Deltares, 2014).
+Wave energy dissipation due to bottom friction and whitecapping is also accounted for [@Deltares2014].
 
@@
 Each nearshore output point thus contained a 17-year timeseries of significant wave heights, peak periods and mean wave directions (with both swell and wind waves).
-This data set is available on a 500m resolution around most of the South African coastline and on a 200m resolution with False Bay as provided by the South African Department of Environmental Affairs (DEA) and Council for Scientific and Industrial Research (CSIR)(@heron2014).
-The total 17-year average energy, for each output location, was then computed using linear wave theory, taking into account deep, intermediate and shallow water wave approximations (based on wave length) for the wave energy (Holthuijsen, 2007; Kasiulis, Punys and Kofoed, 2015; Foteinis *et al.*, 2017) which is a function of wave height and period.
+This data set is available on a 500m resolution around most of the South African coastline and on a 200m resolution with False Bay as provided by the South African Department of Environmental Affairs (DEA) and Council for Scientific and Industrial Research (CSIR)(@Theron2014).
+The total 17-year average energy, for each output location, was then computed using linear wave theory, taking into account deep, intermediate and shallow water wave approximations (based on wave length) for the wave energy [@Holthuijsen2007; @Kasiulis2015; @Foteinis2017] which is a function of wave height and period.
 
@@
 The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits.
-It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world [Norderhaug2012; @Liberti2013; @Sundblad2014].
+It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world [@Norderhaug2012; @Liberti2013; @Sundblad2014].
 Kelp and other ecosystems adapt to their typical environmental exposure and thus what is classified as extreme abiotic coastal forcings (temperature, winds, waves and currents) will vary around the world depending on the geographical setting.
@@
 Hindcast data from WWIII span from 1994-2013 at a 3-hour resolution.
-The data are then used to model short –crested waves generated by the wind into the coastal environment, using Simulating Waves in the Nearshore [SWAN; @booij1996].
+The data are then used to model short –crested waves generated by the wind into the coastal environment, using Simulating Waves in the Nearshore SWAN [@booij1996].
 SWAN allows one to extract wave parameters from specific gridded locations in the nearshore.
@@
 ```
+
 #### Waves as a driver of _Laminaria pallida_ morphometrics

### 2018-10-29 chapter_2.Rmd
changed figure label for temperature parameters. deleted chapter2 html output from repo
added kelp comparison script. Writing to follow

@@
 
-```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Morphometric site locations represented by various temperature parameters.
+```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature parameters relative to each morphometric collection site around the Western Cape coast.
 Temperature parameters include minimum, maximum, mean, range and standard deviation (° Celsius).

### 2018-10-29 chapter_2.Rmd
script did not commit in last commit. Re-try. depth morphology data set added

@@
 
+```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Boxplots representing the different _Ecklonia maxima_ morphometrics between deep and shallow water sites measured around the Western Cape coastline, with the X-axis depicting the specific morphology measured, with units provided.
+Boxplots represent the minimum, 25^th^ percentile, median and 75^th^ percentile of the morphometrics measured.
+Interquartile range can be deduced as the different between the 75^th^ and 25^th^ percentiles, and dots represent outliers in the data.
+Sites are ordered sequentially on the Y-axis by location along the coast.
+The top site is St.
+Helena Bay and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.height=8, fig.width=12, fig.pos="H"}
+
+# load data
+comp_df <- read.csv("data/morph_comp.csv", sep = ";")
+
+# remove unnecessary columns
+comp_df <- subset(comp_df, select = -c(3))
+
+# alter dataframe for more friendly plotting
+comp_gather <- gather(comp_df, "Morphology", "Measurement", 3:12, na.rm = TRUE)
+
+# Replace morphology names
+comp_gather <- comp_gather %>%
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "fond_mass", "Frond mass (kg)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "primary_length", "Primary length (cm)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "primary_width", "Primary width (cm)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "frond_length", "Frond length (cm)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "stipe_mass", "Stipe mass (kg)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "stipe_length", "Stipe length (cm)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "stipe_circ", "Stipe circumference (cm)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "tufts", "Number of tufts")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "epi_length", "Epiphyte length (cm)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "total_length", "Total length (cm)"))
+
+# Mean morphology per site
+
+comp_gather_mean <- comp_gather %>%
+  group_by(site, Morphology) %>%
+  summarise(mean = mean(Measurement, na.rm = TRUE)) %>%
+  ungroup()
+
+# Order site names by location west to east
+
+positionsComp <- c("Kalk Bay", 
+                  "Miller's Point", "Buffels Bay", "Soetwater", "Kommetjie")
+
+# Remove Depth and Rocky Bank site
+eck_gather <- subset(eck_gather, !Morphology == "Depth (m)")
+eck_gather <- subset(eck_gather, !site == "Rocky Bank" | !site == "De Hoop")
+
+# Plot morphologies
+
+ggplot(data = comp_gather, aes(x = site, y = Measurement, fill = depth)) +
+  geom_boxplot() +
+  facet_wrap(~Morphology, scales = "free_x", strip.position = "bottom", ncol = 5) +
+  coord_flip() + 
+  scale_x_discrete(limits = positionsComp) +
+  theme(strip.background = element_blank(), strip.placement = "outside", text = element_text(size   = 4.5)) +
+  xlab("Site") +
+  ylab("Morphology measurement") + 
+  theme_bw()
+
+```
+
+### Abiotic correlations
 ```{r, echo= FALSE, message=FALSE, warning=FALSE, results="hide"}
@@
 ```
-
-### Abiotic correlations
 

### 2018-10-29 chapter_2.Rmd
Morphological methods for shallow collections added

@@
 
+Between February 2017 and Novevember 2018, morphological measurements for _Ecklonia maxima_ individuals in shallow water (<1m) at 5 sites along the Western Cape coast of South Africa were collected.
+These sites represent a wave gradient moving south along the peninsula and into False Bay.
+The same morphometric measurements were taken as per the deeper _Ecklonia maxima_, and this allowed comparison between morphological characteristics between deep and shallow individuals.
+
 ```{r, echo= FALSE, message=FALSE, warning=FALSE}

### 2018-10-29 chapter_2.Rmd
results section added for new morphology comparison section

@@
 
+When comparing morphologies bewteen deep and shallow water _Ecklonia maxima_ individuals, clear pattern is seen in certian morphological characteristics.
+These are frond length, frond mass, primary length, primary width and stipe circumference, and all show a similar pattern.
+Significant differences in these morphological characterstics at Kommetjie and Soetwater are evident, however there are no significant differences from Buffel's Bay to Kalk Bay.
+
 ```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Boxplots representing the different _Ecklonia maxima_ morphometrics between deep and shallow water sites measured around the Western Cape coastline, with the X-axis depicting the specific morphology measured, with units provided.

### 2018-11-08 chapter_1.Rmd
minor edit. added in headings to discuss later
some minor edits and changes

@@
 For example a study by @Koehletal2008 showed that transplanted *Nereocystis luetkeana* plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days.
-Another study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and was a combination of *in situ* sampling and transplant of juvenile plants.
+Another study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and through a combination of *in situ* sampling and transplantation of juvenile plants.
 The results showed that morphology differed between wave-sheltered and wave-exposed sites (thin thallus at sheltered sites and a narrow, thick thallus with a thick stipe at exposed sites), and was consistent with previous studies.
-Juveniles transplanted into wave exposed sites under went rapid morphological adaption, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaption.
+Juveniles transplanted into wave exposed sites under went rapid morphological adaptation, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaptation.
 
@@
 The morphology of macroalgae are not fixed gentic traits.
-A study by @??
-showed that *species name* was able to rapidly adapt to the local wave environment when transplanted from a low wave exposure site to a high wave exposure site.
-Advances in genetic techniques and taxonomy have revealed that species delination based on morphology has been inaccurate, and organisms that were once considered two seperate species are actually one species.**insert example here**.
+A study by @Koehletal2008 showed that transplanted *Nereocystis luetkeana* plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days.
+Advances in genetic techniques and taxonomy have revealed that species delination based on morphology has been inaccurate, and organisms that were once considered two seperate species are actually one species.
 The morphological adaption that macroalgae display are therefore driven by site conditions, therefore individuals must be morphologically flexible to persist in stochastic environments.
@@
 In general, flat strap-like blades are common in areas that are exposed to high wave energy, while at proected sites blade morphology is wide and undulated.
-
-@moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed).
-The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid concentrations.
 
@@
 
+# Coastal flow
+
+# Kelp-rafting
+
 #Aims of research

### 2018-11-08 chapter_2.Rmd
addition to intro section

@@
 The resilience of kelps is evident through their distribution, which spans a range of environmental conditions that vary spatially and temporally.
-Wave exposure and temperature are regarded as important environmental drivers with regards to macroalgae, and play a role in the growth, reproduction, and survival of kelp species.
+Wave exposure and temperature are regarded as important environmental drivers with regards to macroalgae, and play a role in the distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of kelps.
 Wave exposure is recognised as the main cause of kelp mortality, which may manifest itself through dislodgment of individuals or breakage of important morphological structures.
-These mechanisms usually occur during times of high wave energy or pulse disturbance events, such as storms.
+These mechanisms usually occur during times of high wave energy or pulse disturbance events, such as storms.Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10- 20 m s-1 with accelerations of 400 m s-2 [@Friedland1995] are the biggest threat to kelp survival.
 Kelps are able to reduce the probability of dislodgment or breakage through morphological adaptation, which may involve several strategies.
-Kelps can either reduce drag, increase attachment strength or increase flexibility in order to survive high wave energy environments.
+Kelps can either reduce drag, increase attachment strength or increase flexibility in order to survive high wave energy environments.The strategy that the algal organism employs is dependent on the species.
+For instance, branched, articulated algae will vary its morphology differently to that of foliose, more flexible macroalgae.
+For example, @moss1948 investigated the anatomy, chemical composition of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed).
+The authors found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations.
+The authors also noted a ‘crumpling effect’ displayed by individuals from exposed sites and inferred that this strategy may reduce overall drag.
+Other studies show that macroalgae in wave exposed environments have morphologies that reduce overall drag, increase strength of attachment or increase flexibility.
+For example, a study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and through a combination of *in situ* sampling and transplantation of juvenile plants.
+The results showed that morphology differed between wave-sheltered and wave-exposed sites (thin thallus at sheltered sites and a narrow, thick thallus with a thick stipe at exposed sites), and was consistent with previous studies.
+Juveniles transplanted into wave exposed sites under went rapid morphological adaptation, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaptation.
+Dislodgement through increased wave exposure is not the only hydrological parameter playing a role in kelp mortality.
+A study by @DeBettignies2013 tested if there is a linked between orbital water velocity representative of wave-swept reefs and the morphology and biomass of *Ecklonia radiata*.
+The results showed that morphological adaptation through drag reduction and strength increase strategies do not reduce drag as efficiently as reduction in overall biomass through reduction in thallus surface area.
+The authors concluded that at peak orbitat water velocities size (biomass) and not morphological adaptation was the main driver of hydrological performance in wave-swept environments.
+There is also evidence that morphological adaptation is driven by currents, and in fact may be driving hydrological performance of macrolgae.
+
+### The South African context
 
@@
 Although the two species occur together for the majority of the coast, their basic morphologies and resource needs vary to a degree.
-The larger species, _Ecklonia maxima_, is distributed from Lüderitz to Cape Agulhas [@Bolton1985a; @Probyn1985; @Bolton1987; @Bolton2012].
-Characterised by a large distal swollen bulb filled with gas, and smooth fronds, this species grows to approximately 10 meters [@Bolton1987].
-There was, however, a 17-meter specimen collected in 2015 off Cape Point (Smit, unpubl.
-data).This species of kelp not only dominate the biomass of the South African nearshore, but plays an important ecological role [@Bustamante1996].
-The estimated productivity of _Ecklonia maxima_ within South Africa varies between 350 and 1500g Cm-2yr-1 [@Mann1982].
-Across the majority of the coastline, _Laminaria pallida_ remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 meters [@Field1980b; @Bolton1987; @Molloy1996].
-This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 meters [@Field1980b; @Molloy1996; @Molloy1996; @Stegenga1997].
-Towards the north along the west coast, from around Hondeklipbaai, _Laminaria pallida_ replaces _Ecklonia maxima_ as the dominant kelp species [@Velimirov1977; @Stegenga1997] and it also occupies increasingly shallower subtidal regions.
-The northern populations also exhibit an increase in stipe hollowness, compared to the solid stipe morphs in the species’ southern distributions [@Molloy1996].
-This variation in morphology was thought to represent two distinct species, with the northern populations formerly described as _Laminaria schinzii_ Foslie [@Molloy1996].
-Genetic work has subsequently shown that the two morphs are in fact the same species [@Rothman2017].
-In southern African waters, the primary production of _Laminaria pallida_ is between 120 and 1900g C m2yr1, similar to that of _Ecklonia maxima_ [@Mann1982].
-Primary production is not the only pathway.
+The larger species, _Ecklonia maxima_, is distributed from Lüderitz to Cape Agulhas (Fig.
+1) [@Bolton1985a; @Probyn1985; @Bolton1987; @Bolton2012].
 
@@
 
+Characterised by a large distal swollen bulb filled with gas, and smooth fronds, this species grows to approximately 10 meters [@Bolton1987].
+There was, however, a 17-meter specimen collected in 2015 off Cape Point (Smit, unpubl.
+data).This species of kelp not only dominate the biomass of the South African nearshore, but plays an important ecological role [@Bustamante1996].
+The estimated productivity of _Ecklonia maxima_ within South Africa varies between 350 and 1500g Cm-2yr-1 [@Mann1982].
+Across the majority of the coastline, _Laminaria pallida_ remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 meters [@Field1980b; @Bolton1987; @Molloy1996].
+This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 meters [@Field1980b; @Molloy1996; @Molloy1996; @Stegenga1997].
+Towards the north along the west coast, from around Hondeklipbaai, _Laminaria pallida_ replaces _Ecklonia maxima_ as the dominant kelp species [@Velimirov1977; @Stegenga1997] and it also occupies increasingly shallower subtidal regions.
+The northern populations also exhibit an increase in stipe hollowness, compared to the solid stipe morphs in the species’ southern distributions [@Molloy1996].
+This variation in morphology was thought to represent two distinct species, with the northern populations formerly described as _Laminaria schinzii_ Foslie [@Molloy1996].
+Genetic work has subsequently shown that the two morphs are in fact the same species [@Rothman2017].
+In southern African waters, the primary production of _Laminaria pallida_ is between 120 and 1900g C m2yr1, similar to that of _Ecklonia maxima_ [@Mann1982].
+Primary production is not the only pathway.
+
 The aim of this study is, therefore, to understand how environmental drivers such as temperature and wave energy can influence morphoplasticity in two species of kelps around South Africa.
@@
 
-In Figure 1 the total and seasonal wave roses of the directional wave buoy just offshore from Cape Town is given form the years 2000 till end 2017 (the period for which directional wave spectra was available).
+In Figure 2 the total and seasonal wave roses of the directional wave buoy just offshore from Cape Town is given form the years 2000 till end 2017 (the period for which directional wave spectra was available).
 It is due to the near consistent South-westerly swell waves and the complex orography around the peninsula, that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.
@@
 
-In Figure 2 the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
+In Figure 3 the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
 These values have been computed using model output from approximately 20 year of simulated nearshore wave conditions.
@@
 
-The directional sheltering effect of the Cape Peninsula, against the dominant swell direction, (given in Figure 1) is clear observed in the wave exposures presented in Figure 2.
+The directional sheltering effect of the Cape Peninsula, against the dominant swell direction, (given in Figure 2) is clear observed in the wave exposures presented in Figure 2.
 The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits.

### 2018-11-08 chapter_2.Rmd
some minor house cleaning and changes

@@
 
+### Background
+
 Kelps are sessile species and have to adapt to local environmental conditions in order to persist and survive.
@@
 In Figure 3 the propagation of a typical offshore wave spectrum is given as produced from a single time-step in SWAN.
-Figure 3 is presented to clarify the averaged wave exposure maps presented in Figure 2.
+Figure 3 is presented to clarify the averaged wave exposure maps presented in Figure 4.
 Tracing the wave height contours into False Bay its clear why this Bay’s western periphery is predominantly sheltered.
@@
 During February and August, the west side of Cape point experiences decreased temperatures (Fig.
-2).
+5).
 This region also experiences lower mean and maximum temperatures when compared to False Bay, a region known as east of Cape Point.
 The range of temperatures within False Bay are larger in winter (August) compared to summer (February) (Fig.
-3).
+5).
 The sites north of Kommetjie on the west side display larger temperatures ranges than sites south of Kommetjie on the west side.
@@
 Median wind direction is northerly in winter (Fig.
-5), and turns southerly in summer.
+6), and turns southerly in summer.
 
@@
 Lamina length for laminaria pallida showed no geographical pattern moving from west to east.
-Kommetjie, Olifantsbos and Batsata Rock showed great variability in lamina length, and both Buffels Bay and Betty’s Bay were visually different to Miller’s Point and Roman Rock, when comparing summary data (Fig.6).
+Kommetjie, Olifantsbos and Batsata Rock showed great variability in lamina length, and both Buffels Bay and Betty’s Bay were visually different to Miller’s Point and Roman Rock, when comparing summary data (Fig.8).
 Lamina thickness showed great variation across sites, with Baboon Rock, Miller’s Point, A-Frame and Roman rock displaying large lamina thickness, significantly different to the rest of the sites.
@@
 Helena Bay to Kommetjie and Soetwater (Fig.
-7).
+9).
 Hout Bay, Kommetjie and Soetwater show similarities in their stipes lengths to Buffels Bay, Batsata Rock and Betty’s Bay, and are significantly different to west of Cape Point counterpart sites such as Oudekraal and Scarborough.
@@
 
-When comparing morphologies bewteen deep and shallow water _Ecklonia maxima_ individuals, clear pattern is seen in certian morphological characteristics.
+When comparing morphologies bewteen deep and shallow water _Ecklonia maxima_ individuals, clear pattern is seen in certian morphological characteristics (Fig.
+10).
 These are frond length, frond mass, primary length, primary width and stipe circumference, and all show a similar pattern.
 Significant differences in these morphological characterstics at Kommetjie and Soetwater are evident, however there are no significant differences from Buffel's Bay to Kalk Bay.
+<!-- I cannot get the black and white them to work on this plot...apologies I will try fix it as soon as possible.
+-->
 
@@
 There are various strong correlations within the temperature parameters, and within the wave parameters on an annual time scale (Fig.
-8).
+11).
 SD Significant wave height (Hs) and mean Hs showed a strong positive correlate(0.92), as well as  SD wave period (Tp) and mean Hs with a strong negative correlation(-0.9) For SD wind speed and mean wind speed a strong positive correlation existed (0.967).
@@
 Median wind direction and mean wind speed are also correlated well (0.954).
-We however see no strong correlations between wave and temperature parameters.
+We however see no strong correlations between wave and temperature parameters (Fig 12).
 
 These correlations vary through the three timescales (Annual, August and February).
-Although there are no strong correlations between temperature and wave parameters, we see interesting differences between the seasonal timescales The mean temperature and mean wind speed are weakly correlated in August (0.253)(Fig.
-9), but are observed to be more strongly negatively correlated in February (-0.788)(Fig.
-10).
+Although there are no strong correlations between temperature and wave parameters, we see interesting differences between the seasonal timescales.
+The mean temperature and mean wind speed are weakly correlated in August (0.253)(Fig.
+12), but are observed to be more strongly negatively correlated in February (-0.788)(Fig.
+13).
 The same is seen for minimum temperature and median wind direction in August (0.129) compared to February (-0.715).
@@
 The morphometrics of _Ecklonia maxima_ were explained more by wave parameters (75%), than temperature parameters (66%), when separate RDA’s were constructed (Fig.
-11).
+14).
 The first two axes for wave parameters driving _Ecklonia maxima_ morphology explained 34% of the variation.
@@
 Although temperature parameters for _Ecklonia maxima_ showed some influence of the morphometric data, the explanation of morphology was not strong, with an R^2^ value of 0.662 and a negative adjusted R^2^ value (-0.138)(Fig.
-12).
+15).
 Adjusted R^2^ values are modified R^2^ values to include the number predictors in a model.
@@
 This is supported by the lack of relationships between environmental (temperature) and response (morphology) variables (Fig.
-12).
+15).
 
@@
 For _Laminaria pallida_, the percentage of explained data by wave parameters was 89%, compared to temperature parameters explaining 82% (Fig.
-13).
+16).
 By using adjusted R^2^ values, the first two axes for wave parameters driving _Laminaria pallida_ morphology explained 57% of the variation of morphology in this species.
@@
 Provided by adjusted R^2^ values, the first two axes of temperature effect on _Laminaria pallida_ morphologies amounted to only 11%, but exhibited a stronger influence than on _Ecklonia maxima_ (Fig.
-14).
+17).
 The mean and minimum temperatures during February were observed to negatively influence _Laminaria pallida_ morphologies such as stipe mass, stipe length and stipe diameter, as well as thallus mass and total length.
@@
 
-Seasonal variations in significant wave height (Hs), wave period and wave direction were observed from the data (Fig.
-4, 5).
+Seasonal variations in significant wave height (Hs), wave period and wave direction were observed from the data.
 The direction of swell swings to the south west in winter, generated by strong low pressures that originate from the southern ocean [@Reason2006].
@@
 The wind speed and wind direction support the presence of upwelling in summer, where southerly winds blow parallel to the coast and trigger upwelling [@Field1980a].
-This is supported by the decreased in temperature in summer (Fig.
-2, 3), specifically along the west side of Cape Point.
+This is supported by the decreased in temperature in summer, specifically along the west side of Cape Point.
 We however see a differentiation of median wind directions in summer, with the western coast of False Bay experiencing south easterly winds, compared to the west side of Cape Point experiencing south westerly winds.
@@
 Although this study did not measure force of attachment, other morphological responses to wave parameters was evident.
-An increase in mean annual wave period saw an increase of lamina length of _Laminaria pallida_, but a decrease in lamina thickness (Fig.
-13).
+An increase in mean annual wave period saw an increase of lamina length of _Laminaria pallida_, but a decrease in lamina thickness.
 By decreasing the thickness of the lamina, and increasing the length of the lamina (directly increasing surface area), a larger more flexible kelp can survive in environments with greater wave period.
@@
 
-Lamina thickness showed positive correlation with annual SD wave period (Fig.
-13).
+Lamina thickness showed positive correlation with annual SD wave period.
 Although this variable may just negatively correlate with annual mean wave period, another reason may exist.
@@
 
-Mean annual significant wave height (Hs) was observed to influence _Ecklonia maxima_ stipe circumference, with increased stipe circumference at larger Hs (Fig.
-11).
+Mean annual significant wave height (Hs) was observed to influence _Ecklonia maxima_ stipe circumference, with increased stipe circumference at larger Hs.
 An increase in Hs saw greater wave actions as waves are larger, as well as greater wave periods that are positively correlated with Hs.
@@
 _Ecklonia maxima_ are thought to be influenced greatly by light attenuation [@Rothman2017], and thus have developed increased (stronger) stipe circumferences to be able to cope in large Hs environments, to access cleaner, nutrient-filled water.
-The same patterns are observed in _Laminaria pallida_ morphologies, that display increased stipe diameters at Soetwater and Kommetjie as well (Fig.
-13).
+The same patterns are observed in _Laminaria pallida_ morphologies, that display increased stipe diameters at Soetwater and Kommetjie as well.
 This increase in stipe diameter correlates with stipe mass of _Laminaria pallida_, providing evidence that this species has developed thicker, strong stipes to persist in this region.
@@
 Therefore adult _Ecklonia maxima_ sporophytes would experience a lower stress ratio compared to adult _Laminaria pallida_ sporophytes in the same environment, where _Laminaria pallida_ is often a subsurface species to _Ecklonia maxima_.
-
-Stipe circumference for _Ecklonia maxima_ were explained by annual wave direction (Fig.
-11).
+Stipe circumference for _Ecklonia maxima_ were explained by annual wave direction.
 For _Ecklonia maxima_, the sporophytes may be producing stipes that have a greater circumference to cope with increased wave energy.
@@
 We however see a lack of upwelling events for sites such as Betty’s Bay and Batsata Rock.
-These two sites display _Ecklonia maxima_ sporophytes with the largest total lengths (Fig.
-7), as well as Betty’s Bay displaying the largest total length of _Laminaria pallida_ sporophytes(Fig.
-6).
+These two sites display _Ecklonia maxima_ sporophytes with the largest total lengths, as well as Betty’s Bay displaying the largest total length of _Laminaria pallida_ sporophytes.
 The effect of depth on stipe length is regarded as insignificant in this study as all kelp were sampled at similar depths.
 Stipe elongation has been shown to be a morphological adaptation in wave exposed environment, as a longer kelp has time to extend back and forth in swell reducing the tension on the holdfast.
-Stipe lengths of _Ecklonia maxima_ from Kommetjie, Hout Bay and Soetwater are similar to Batsata rock and Betty’s Bay, but differs to west side neighbours such as Oudekraal and Scarborough (Fig.
-7).
-Oudekraal and Scarborough experience more variability in wave direction, however they are observed to experience smaller annual mean significant wave heights, when compared to Kommetjie, Hout Bay, Soetwater, as well as Batsata Rock and Betty’s Bay (Fig.
-4).@Friedland1995 found that longer kelps were able to reduce drag and tension as the sinusoidal displacement of most waves would not ‘stretch’ a kelp out long enough before falling.
+Stipe lengths of _Ecklonia maxima_ from Kommetjie, Hout Bay and Soetwater are similar to Batsata rock and Betty’s Bay, but differs to west side neighbours such as Oudekraal and Scarborough.
+Oudekraal and Scarborough experience more variability in wave direction, however they are observed to experience smaller annual mean significant wave heights, when compared to Kommetjie, Hout Bay, Soetwater, as well as Batsata Rock and Betty’s Bay.
+@Friedland1995 found that longer kelps were able to reduce drag and tension as the sinusoidal displacement of most waves would not ‘stretch’ a kelp out long enough before falling.
 The ability to grow longer stipes would provide the kelp with a structure that could surpass most wave heights without the risk of dislodgement.

### 2018-11-08 chapter_2.Rmd
some minor additions to discussion
added to background section and begining of conclusion

@@
 The authors concluded that at peak orbitat water velocities size (biomass) and not morphological adaptation was the main driver of hydrological performance in wave-swept environments.
+
 There is also evidence that morphological adaptation is driven by currents, and in fact may be driving hydrological performance of macrolgae.
+@Duggins2003 examined the direct and indirect flow effects on population dynamics, morphology and biomechanics of several understorey macroalgae species.
+These species included *Costaria costata*, *Agarum fimbriatum*, and *Laminaria complanata* and * Nereocystis luetkeana*.
+The results showed that in wave impacted sites (wave exposed) had higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity.
+The authors concluded that although tidal and current velocity did not play a significant role in determining kelp survival, it did play a role in morphological adaptation.
+The results from this study suggest that high current and tidal stresses are the main driver of kelp morphological adaptation.
+This in turn make those individuals more resilient to dislodgement to wave exposure.
+Wave exposure is stochastic in nature compared to tidal and ocean currents which are more regular in their frequency and magnitude.
+Therefore the regular forces of tidal and ocean currents may make kelp individuals more resilient to mechanical dislodgement over time.
 
@@
 In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an explanation for increased variability of Hs and Tp.
+However, changing winds may also be influencing waves generated within False Bay.
+As seen in Figure 1 the summer and winter swells are relativley constant in terms of direction adn this may be driving the waves locally (I am still looking for the report where Christo cites this).
 The wind speed and wind direction support the presence of upwelling in summer, where southerly winds blow parallel to the coast and trigger upwelling [@Field1980a].
@@
 
+## Conclusion
+
+(More to add here but I wanted to add it so long)
+Mechanical dislodgment through hydrological stress is the main source on mortality for kelp, and previous research has shown that wave exposure is the main driver of kelp morphology.
+Therefore, morphological adaptation is the main strategy to reduce drag and ultimately dislodgement in hydrological stressful environments.
+The results from this study support previous findings, and shows that particular wave parameters (significant wave height and median wave direction) act more strongly in driving kelp morphological plasticity.
+
 # References

### 2018-11-08 chapter_2.Rmd
Edits by AJS.

@@
 title: "The shape of kelp: environmental influences on kelp morphology"
+author:
+- affiliation: University of the Western Cape
+  name: Ross Coppin
+date: "8 November 2018"
 output:
   pdf_document:
-    toc: yes
+    fig_caption: yes
+    fig_height: 6.5
+    fig_width: 6.5
+    highlight: default
+    keep_tex: yes
+    latex_engine: xelatex
+    template: LaTeX/template-article_ajs.tex
   html_notebook:
@@
     toc_float: no
-fontsize: 12pt
+tables: yes
+fontsize: 10pt
+papersize: A4
 geometry: margin=1in
 language: Australian
-mainfont: PT Serif
-monofont: PT Mono
 csl: LaTeX/ecography.csl
-papersize: A4
-sansfont: PT Sans
-tables: yes
-bibliography: LaTex/mycollection.bib
+bibliography: LaTeX/mycollection.bib
 ---
@@
 Outside of these latitudes, they are also found in cool, deep water towards the tropics [@Graham2007]: this deeper tropical water, due to its water clarity, also allows for suitable photon flux and a nitrate concentration [@Luning1990; @Dayton1999; @Zimmerman1984], the two other environmental variables that support kelp productivity.
-Although kelps have a low taxonomic diversity, their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude other turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Duggins1989; @Steneck2002], each depending on a wide suite of ecological services provided by the kelp habitat [@Gaines1987; @Bologna1993; @Levin1994; @Anderson1997].
+Although kelps have a low taxonomic diversity, their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude other turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Duggins1989; @Steneck2002], each depending on a wide suite of ecological services provided by the kelp and the habitat structure they provide [@Gaines1987; @Bologna1993; @Levin1994; @Anderson1997].
 
@@
 
-Kelps are sessile species and have to adapt to local environmental conditions in order to persist and survive.
+Kelps are sessile species and have to adapt to local environmental conditions in order to grow, repruce and persist.
 The resilience of kelps is evident through their distribution, which spans a range of environmental conditions that vary spatially and temporally.
@@
 Wave exposure is recognised as the main cause of kelp mortality, which may manifest itself through dislodgment of individuals or breakage of important morphological structures.
-These mechanisms usually occur during times of high wave energy or pulse disturbance events, such as storms.Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10- 20 m s-1 with accelerations of 400 m s-2 [@Friedland1995] are the biggest threat to kelp survival.
+These mechanisms usually occur during times of high wave energy or pulse disturbance events, such as storms.
+Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10-20 m.s^-1^ with accelerations of 400 m.s^-1^ [@Friedland1995] are the biggest threat to kelp survival.
 Kelps are able to reduce the probability of dislodgment or breakage through morphological adaptation, which may involve several strategies.
-Kelps can either reduce drag, increase attachment strength or increase flexibility in order to survive high wave energy environments.The strategy that the algal organism employs is dependent on the species.
+Kelps can either reduce drag, increase attachment strength, or increase flexibility in order to survive high wave energy environments.
+The strategy that the algal organism employs is dependent on the species.
 For instance, branched, articulated algae will vary its morphology differently to that of foliose, more flexible macroalgae.
-For example, @moss1948 investigated the anatomy, chemical composition of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed).
-The authors found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations.
+@moss1948 investigated the anatomy and chemical composition of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed).
+The author found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations.
 The authors also noted a ‘crumpling effect’ displayed by individuals from exposed sites and inferred that this strategy may reduce overall drag.
 Other studies show that macroalgae in wave exposed environments have morphologies that reduce overall drag, increase strength of attachment or increase flexibility.
-For example, a study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and through a combination of *in situ* sampling and transplantation of juvenile plants.
-The results showed that morphology differed between wave-sheltered and wave-exposed sites (thin thallus at sheltered sites and a narrow, thick thallus with a thick stipe at exposed sites), and was consistent with previous studies.
-Juveniles transplanted into wave exposed sites under went rapid morphological adaptation, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaptation.
-Dislodgement through increased wave exposure is not the only hydrological parameter playing a role in kelp mortality.
-A study by @DeBettignies2013 tested if there is a linked between orbital water velocity representative of wave-swept reefs and the morphology and biomass of *Ecklonia radiata*.
-The results showed that morphological adaptation through drag reduction and strength increase strategies do not reduce drag as efficiently as reduction in overall biomass through reduction in thallus surface area.
-The authors concluded that at peak orbitat water velocities size (biomass) and not morphological adaptation was the main driver of hydrological performance in wave-swept environments.
+For example, a study by @FowlerWalker2006 looked for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites, using a combination of *in situ* sampling and transplantation of juvenile plants.
+The results showed that morphology differed between wave-sheltered and wave-exposed sites, with thin thalli developing at sheltered sites and a narrower, thicker thallus with a thick stipe forming at exposed sites.
+This finding was consistent with previous studies {AJS: name the studies and the species involved here}.
+Juveniles transplanted into wave exposed sites underwent rapid morphological adaptation, whilst the opposite was true for wave-sheltered sites, which showed slower morphological adjustment {refs.}.
+Dislodgement through increased wave exposure is not the only hydrological variable playing a role in kelp mortality.
+A study by @DeBettignies2013 tested if there is a link between orbital water velocity representative of wave-swept reefs and the morphology and biomass of *Ecklonia radiata*.
+The results showed that morphological adaptation through drag reduction and strength increase does not reduce drag as efficiently as reduction in overall biomass through reduction in thallus surface area.
+The authors concluded that at peak orbital water velocities, size (as biomass) and not morphological adaptation was the main determinant of hydrological performance in wave-swept environments.
 
@@
 @Duggins2003 examined the direct and indirect flow effects on population dynamics, morphology and biomechanics of several understorey macroalgae species.
-These species included *Costaria costata*, *Agarum fimbriatum*, and *Laminaria complanata* and * Nereocystis luetkeana*.
-The results showed that in wave impacted sites (wave exposed) had higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity.
+These species included *Costaria costata*, *Agarum fimbriatum*, *Laminaria complanata* and *Nereocystis luetkeana*.
+The results showed that wave impacted sites (wave exposed) exhibited higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity.
 The authors concluded that although tidal and current velocity did not play a significant role in determining kelp survival, it did play a role in morphological adaptation.
@@
 
-```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="A map indicating the sites where morphometric measurements were collected for _Ecklonia maxima_ and _Laminaria pallida_.", fig.height=8, fig.width=12}
+```{r, echo= FALSE, message=FALSE, warning=FALSE, fig.cap="A map indicating the sites where morphometric measurements were collected for _Ecklonia maxima_ and _Laminaria pallida_."}
 
@@
 Characterised by a large distal swollen bulb filled with gas, and smooth fronds, this species grows to approximately 10 meters [@Bolton1987].
-There was, however, a 17-meter specimen collected in 2015 off Cape Point (Smit, unpubl.
-data).This species of kelp not only dominate the biomass of the South African nearshore, but plays an important ecological role [@Bustamante1996].
-The estimated productivity of _Ecklonia maxima_ within South Africa varies between 350 and 1500g Cm-2yr-1 [@Mann1982].
-Across the majority of the coastline, _Laminaria pallida_ remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 meters [@Field1980b; @Bolton1987; @Molloy1996].
-This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 meters [@Field1980b; @Molloy1996; @Molloy1996; @Stegenga1997].
+There was, however, a 17 m specimen collected in 2015 off Cape Point (Smit, unpubl.
+data).
+This species of kelp not only dominate the biomass of the South African nearshore, but plays an important ecological role [@Bustamante1996].
+The estimated productivity of _Ecklonia maxima_ within South Africa varies between 350 and 1500 gC.m^-2^yr^-1^ [@Mann1982].
+Across the majority of the coastline, _Laminaria pallida_ remains a subsurface kelp, dominating the kelp biomass at depths greater than 10 m [@Field1980b; @Bolton1987; @Molloy1996].
+This species is distributed from Danger Point, east of the Cape Peninsula, to Rocky Point in northern Namibia, and reaches depths of greater than 20 m [@Field1980b; @Molloy1996; @Molloy1996; @Stegenga1997].
 Towards the north along the west coast, from around Hondeklipbaai, _Laminaria pallida_ replaces _Ecklonia maxima_ as the dominant kelp species [@Velimirov1977; @Stegenga1997] and it also occupies increasingly shallower subtidal regions.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Coastal wave energy exposure around the Cape Peninsula for the (a) total 17-year nearshore wave record, (b) only winter months and (c) only summer months."}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Coastal wave energy exposure around the Cape Peninsula for the (a) total 17-year nearshore wave record, (b) only winter months and (c) only summer months."}
 
@@
 
-grid.arrange(img_we_total, img_we_winter,img_we_summer, ncol = 2,nrow = 2)
+grid.arrange(img_we_total, img_we_winter, img_we_summer, ncol = 2, nrow = 2)
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="SWAN numerical output for an example typical offshore wave conditions refracting around the Cape Peninsula."}
-
-knitr::include_graphics('figures/regional_waves.png')
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "SWAN numerical output for an example typical offshore wave conditions refracting around the Cape Peninsula."}
+
+thing <- rasterGrob(as.raster(readPNG("figures/regional_waves.png")), interpolate = FALSE)
+grid.arrange(thing, ncol = 1)
+# knitr::include_graphics('figures/regional_waves.png')
 
@@
 The temperature parameters around the Western Cape coast vary among sites and seasons.
-During February and August, the west side of Cape point experiences decreased temperatures (Fig.
+During February and August, the west side of Cape Point experiences decreased temperatures (Fig.
 5).
@@
 Sites are ordered sequentially on the Y-axis by location along the coast.
-The top site is Sea Point and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.height=8, fig.width=12, fig.pos="H"}
+The top site is Sea Point and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.pos="H"}
 
@@
 The top site is St.
-Helena Bay and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.height=8, fig.width=12, fig.pos="H"}
+Helena Bay and is located at the north western boundary, and Betty’s Bay as the bottom site is located at the south eastern boundary, from our sample region.", fig.pos="H"}
 
@@
 Each point of each linear graph represents a collection site.
-The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.height=8, fig.width=12, fig.pos="H"}
+The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.pos="H"}
 # # ggpairs to plot correlation coefficients
@@
 Each point of each linear graph represents a collection site.
-The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.height=8, fig.width=12, fig.pos="H"}
+The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.pos="H"}
 
@@
 Sites are represented in black.
-The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.height=8, fig.width=12, fig.pos="H"}
+The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.pos="H"}
 # Plot scaling = 2
@@
 Sites are represented in black.
-The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.height=8, fig.width=12, fig.pos="H"}
+The top X-axis and right Y-axis represent the explanatory variable axes, while bottom X-axis and left Y-axis represent the response variables axes.", fig.pos="H"}
 
@@
 
-Kelps, in particular brown seaweeds, are robust and resilient organisms that have been shown to adapt their morphology to suit local environmental conditions.
-Wave exposure has been shown to be the main driver of morphological variation in brown seaweeds as a strategy to reduce drag and ultimately dislodgement.
-Although wave exposure is regarded as an important abiotic variable that drives kelp morphology it does not act independently.
+Kelps, in particular brown seaweeds, are robust and resilient organisms that have been shown to adapt their morphology to suit local environmental conditions (refs.).
+Wave exposure has been shown to be the main driver of morphological variation in brown seaweeds as a strategy to reduce drag and ultimately dislodgement (refs.).
+Although wave exposure is regarded as an important abiotic variable that drives kelp morphology, it does not act independently {AJS: of what?}.
 For instance, a study by @Wernberg2005 investigated the effect of wave exposure on the morphology of _Ecklonia radiata_ of six locations along 1100 km of the southwest Australian coastline.
-The authors concluded that wave exposure was the main driver of kelp morphological adaptation in the study, however the authors also noted the variation in morphology between sites was not consistent and concluded that wave parameters may play different roles in determining kelp morphology.
-The results from this current study confirm the idea that kelp morphology is driven by specific wave parameters, particularly significant wave height, wave period and wave direction.
+The authors concluded that wave exposure was the main driver of kelp morphological adaptation in the study; however, they also noted the variation in morphology between sites was not consistent and concluded that wave parameters may play different roles in determining kelp morphology {AJS: What does this mean?}.
+The results from this current study confirm the idea that kelp morphology is driven by specific wave parameters, particularly significant wave height, wave period and wave direction {AJS: what about?}.
 

### 2018-11-08 chapter_2.tex
Edits by AJS.

\documentclass[10pt,a4,]{article}
% \RequirePackage[l2tabu, orthodox]{nag}
\usepackage[a4paper,text={16.5cm,25.2cm},centering,margin=2.6cm]{geometry}
% \usepackage[left=1.0in,top=1.0in,right=1.0in,bottom=1.0in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage{hyperref,amsmath,amssymb,bm,url,enumitem,dcolumn,upquote,framed,alltt,textgreek,xfrac,fixltx2e}
\usepackage[australian]{babel}
\usepackage[compact,small]{titlesec}
\setlength{\parskip}{1.2ex}
\setlength{\parindent}{0em}

\def\tightlist{}

\usepackage{ifxetex}
\ifxetex
  \usepackage{fontspec}
  \defaultfontfeatures{Ligatures=TeX} % To support LaTeX quoting style
 \defaultfontfeatures{Ligatures=TeX,Numbers={OldStyle}}
 \setmainfont{Minion Pro}
 \setsansfont[Scale=MatchLowercase]{Myriad Pro}
 \setmonofont[Scale=MatchLowercase]{Ubuntu Mono}
\else
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{lmodern}
  % \usepackage[full]{textcomp} % directly use the degree (and some other) symbol
\fi

% place after fonts; even better typesetting for improved readability:
\graphicspath{ {figure/} }
\usepackage{tabularx} % for 'tabularx' environment and 'X' column type
\usepackage{ragged2e}  % for '\RaggedRight' macro (allows hyphenation)
\usepackage{siunitx}
    \sisetup{%
        detect-mode,
        group-digits            = false,
        input-symbols           = ( ) [ ] - + < > *,
        table-align-text-post   = false,
        round-mode              = places,
        round-precision         = 3
        }
% \usepackage[font={small, sf}, labelfont=bf]{caption} % tweaking the captions
\usepackage[font={small}, labelfont=bf]{caption} % tweaking the captions
\usepackage[color=yellow, textsize=tiny]{todonotes}
\frenchspacing%
% \usepackage[kerning=false,protrusion=true,expansion=true]{microtype}

\usepackage{abstract}
\renewcommand{\abstractname}{} % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
{{%
\setlength{\leftmargin}{0mm}
\setlength{\rightmargin}{\leftmargin}%
}%
\relax}
{\endlist}

\makeatletter
\def\@maketitle{%
\newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
\let \footnote \thanks
 {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother

\setcounter{secnumdepth}{0}

\usepackage{longtable,booktabs}
\setlength\heavyrulewidth{0.1em}
\setlength\lightrulewidth{0.0625em}

\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\title{The shape of kelp: environmental influences on kelp morphology  }

\author{\Large Ross Coppin\vspace{0.05in} \newline\normalsize\emph{University of the Western Cape}  }

\date{}

% PENALTIES
\widowpenalty=1000
\clubpenalty=1000
\doublehyphendemerits=9999 % Almost no consecutive line hyphens
\brokenpenalty=10000 % No broken words across columns/pages
\interfootnotelinepenalty=9999 % Almost never break footnotes

% SECTION, SUBSECETC.TITLES
\usepackage[compact]{titlesec}
\titleformat{\chapter}
  {\normalfont\LARGE\sffamily\bfseries}
  {\thechapter}
  {1em}
  {}
\titleformat{\section}
  {\normalfont\LARGE\sffamily\bfseries}
  {\thesection}
  {1em}
  {}
\titleformat{\subsection}
  {\normalfont\Large\sffamily\bfseries}
  {\thesubsection}
  {1em}
  {}
\titleformat{\subsubsection}
  {\normalfont\large\sffamily\bfseries\slshape}
  {\thesubsubsection}
  {1em}
  {}
% \titlespacing*{<command>}{<left>}{<before-sep>}{<after-sep>}
\titlespacing*{\chapter}
  {0pt}
  {1.2ex plus 1ex minus .2ex}
  {0.5ex plus .1ex minus .1ex}
\titlespacing*{\section}
  {0pt}
  {1.2ex plus 1ex minus .2ex}
  {0.5ex plus .1ex minus .1ex}
\titlespacing*{\subsection}
  {0pt}
  {1.2ex plus 1ex minus .2ex}
  {0.5ex plus .1ex minus .1ex}
\titlespacing*{\subsubsection}
  {0pt}
  {1.2ex plus 1ex minus .2ex}
  {0.5ex plus .1ex minus .1ex}

\newtheorem{hypothesis}{Hypothesis}
\usepackage{setspace}

\makeatletter

\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[colorlinks=true, citecolor=blue, linkcolor=cyan, pdfborder={0 0 0 }, unicode=true]{hyperref} % place after other packages, but before cleveref
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}

\usepackage{cleveref} % cleverly cross referencing figures and tables; last package to include
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}

% To use for resetting the numbering of Appendix Tables and Figures:
%\setcounter{table}{0}
%\renewcommand{\thetable}{A\arabic{table}}
%\setcounter{figure}{0}
%\renewcommand{\thefigure}{A\arabic{figure}}

\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Ross Coppin (University of the Western Cape)},
             pdfkeywords = {},
            pdftitle={The shape of kelp: environmental influences on kelp morphology},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\begin{document}

% \maketitle

{% \usefont{T1}{pnc}{m}{n}
\setlength{\parindent}{0pt}
\thispagestyle{plain}
{\fontsize{18}{20}\selectfont\raggedright
\maketitle  % title \par
}
{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12}
\textbf{\authorfont Ross Coppin} \hskip 15pt \emph{\small University of the Western Cape}   
}
}

\vskip 6.5pt

\noindent

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Kelps are a group of large seaweeds of the order Laminariales
(Ochrophyta), which despite their relatively low taxonomic diversity of
species in genera (Bolton 2010), nevertheless form the basis of one of
the most productive ecosystems globally (Mann 1973). Kelps generally
have a dependence on cool, temperate and arctic seawater temperatures
(Santelices 2007, Bolton 2010), and dominate the nearshore biomass
within the rocky shallow coasts in both hemispheres (Steneck et al.
2002). Outside of these latitudes, they are also found in cool, deep
water towards the tropics (Graham et al. 2007): this deeper tropical
water, due to its water clarity, also allows for suitable photon flux
and a nitrate concentration (Zimmerman and Kremer 1984, Lüning 1990,
Dayton et al. 1999), the two other environmental variables that support
kelp productivity. Although kelps have a low taxonomic diversity, their
size and complex morphology provide a heterogeneous habitat structure
(Steneck et al. 2002) that accommodate a multitude other turf and
subcanopy seaweed species, and diverse assemblages of sessile and mobile
invertebrates and vertebrates ({\textbf{???}}, Mann 1973, Steneck et al.
2002), each depending on a wide suite of ecological services provided by
the kelp and the habitat structure they provide (Gaines and Roughgarden
1987, Bologna and Steneck 1993, Levin 1994, Anderson et al. 1997).

\hypertarget{background}{%
\subsubsection{Background}\label{background}}

Kelps are sessile species and have to adapt to local environmental
conditions in order to grow, repruce and persist. The resilience of
kelps is evident through their distribution, which spans a range of
environmental conditions that vary spatially and temporally. Wave
exposure and temperature are regarded as important environmental drivers
with regards to macroalgae, and play a role in the distribution,
abundance, diversity, composition, growth (Cousens 1982) and
productivity (Pedersen and Nejrup 2012) of kelps. Wave exposure is
recognised as the main cause of kelp mortality, which may manifest
itself through dislodgment of individuals or breakage of important
morphological structures. These mechanisms usually occur during times of
high wave energy or pulse disturbance events, such as storms. Mechanical
forces generated by the hydrodynamic environment, in the form of sudden
strong ocean currents or storms, between 10-20 m.s\textsuperscript{-1}
with accelerations of 400 m.s\textsuperscript{-1} (Friedland and Denny
1995) are the biggest threat to kelp survival. Kelps are able to reduce
the probability of dislodgment or breakage through morphological
adaptation, which may involve several strategies. Kelps can either
reduce drag, increase attachment strength, or increase flexibility in
order to survive high wave energy environments. The strategy that the
algal organism employs is dependent on the species. For instance,
branched, articulated algae will vary its morphology differently to that
of foliose, more flexible macroalgae. Moss (1948) investigated the
anatomy and chemical composition of \emph{Fucus spiralis} at three sites
that varied in wave exposure (sheltered, medium exposure and exposed).
The author found that individuals in exposed sites showed less branching
of thalli as well as variation physiological components, such as organic
nitrogen, mannitol, laminarian and alginic acid concentrations. The
authors also noted a `crumpling effect' displayed by individuals from
exposed sites and inferred that this strategy may reduce overall drag.
Other studies show that macroalgae in wave exposed environments have
morphologies that reduce overall drag, increase strength of attachment
or increase flexibility. For example, a study by Fowler-Walker et al.
(2006) looked for differences in morphology of \emph{Ecklonia radiata}
between wave-sheltered and wave-exposed sites, using a combination of
\emph{in situ} sampling and transplantation of juvenile plants. The
results showed that morphology differed between wave-sheltered and
wave-exposed sites, with thin thalli developing at sheltered sites and a
narrower, thicker thallus with a thick stipe forming at exposed sites.
This finding was consistent with previous studies \{AJS: name the
studies and the species involved here\}. Juveniles transplanted into
wave exposed sites underwent rapid morphological adaptation, whilst the
opposite was true for wave-sheltered sites, which showed slower
morphological adjustment \{refs.\}. Dislodgement through increased wave
exposure is not the only hydrological variable playing a role in kelp
mortality. A study by Bettignies et al. (2013) tested if there is a link
between orbital water velocity representative of wave-swept reefs and
the morphology and biomass of \emph{Ecklonia radiata}. The results
showed that morphological adaptation through drag reduction and strength
increase does not reduce drag as efficiently as reduction in overall
biomass through reduction in thallus surface area. The authors concluded
that at peak orbital water velocities, size (as biomass) and not
morphological adaptation was the main determinant of hydrological
performance in wave-swept environments.

There is also evidence that morphological adaptation is driven by
currents, and in fact may be driving hydrological performance of
macrolgae. Duggins et al. (2003) examined the direct and indirect flow
effects on population dynamics, morphology and biomechanics of several
understorey macroalgae species. These species included \emph{Costaria
costata}, \emph{Agarum fimbriatum}, \emph{Laminaria complanata} and
\emph{Nereocystis luetkeana}. The results showed that wave impacted
sites (wave exposed) exhibited higher rates of mortality, and no
significance was found between survival of individuals and tidal or
current velocity. The authors concluded that although tidal and current
velocity did not play a significant role in determining kelp survival,
it did play a role in morphological adaptation. The results from this
study suggest that high current and tidal stresses are the main driver
of kelp morphological adaptation. This in turn make those individuals
more resilient to dislodgement to wave exposure. Wave exposure is
stochastic in nature compared to tidal and ocean currents which are more
regular in their frequency and magnitude. Therefore the regular forces
of tidal and ocean currents may make kelp individuals more resilient to
mechanical dislodgement over time.

\hypertarget{the-south-african-context}{%
\subsubsection{The South African
context}\label{the-south-african-context}}

The biogeographic distribution of kelp is limited by seawater
temperature (Bolton 2010), where increasing temperature gradients reduce
kelp distribution. Due to this limiting factor, the two main species of
kelps in southern African waters, \emph{Ecklonia maxima} and
\emph{Laminaria pallida}, are distributed along a section of the south
coast from De Hoop, extending west around the Cape Peninsula, and
thriving north into Namibia (Molloy and Bolton 1996, Stegenga et al.
1997). This distribution follows a temperate gradient, where sea
temperatures increase as one moves south from Namibia, around Cape Point
and towards De Hoop. Although the two species occur together for the
majority of the coast, their basic morphologies and resource needs vary
to a degree. The larger species, \emph{Ecklonia maxima}, is distributed
from Lüderitz to Cape Agulhas (Fig. 1) (Bolton and Levitt 1985, Probyn
and McQuaid 1985, Bolton and Anderson 1987, Bolton et al. 2012).

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{A map indicating the sites where morphometric measurements were
collected for \emph{Ecklonia maxima} and \emph{Laminaria pallida}.}
\end{figure}

Characterised by a large distal swollen bulb filled with gas, and smooth
fronds, this species grows to approximately 10 meters (Bolton and
Anderson 1987). There was, however, a 17 m specimen collected in 2015
off Cape Point (Smit, unpubl. data). This species of kelp not only
dominate the biomass of the South African nearshore, but plays an
important ecological role (Bustamante and Branch 1996). The estimated
productivity of \emph{Ecklonia maxima} within South Africa varies
between 350 and 1500 gC.m\textsuperscript{-2}yr\textsuperscript{-1}
(Mann 1982). Across the majority of the coastline, \emph{Laminaria
pallida} remains a subsurface kelp, dominating the kelp biomass at
depths greater than 10 m (Field et al. 1980a, Bolton and Anderson 1987,
Molloy and Bolton 1996). This species is distributed from Danger Point,
east of the Cape Peninsula, to Rocky Point in northern Namibia, and
reaches depths of greater than 20 m (Field et al. 1980a, Molloy and
Bolton 1996, Stegenga et al. 1997). Towards the north along the west
coast, from around Hondeklipbaai, \emph{Laminaria pallida} replaces
\emph{Ecklonia maxima} as the dominant kelp species (Velimirov et al.
1977, Stegenga et al. 1997) and it also occupies increasingly shallower
subtidal regions. The northern populations also exhibit an increase in
stipe hollowness, compared to the solid stipe morphs in the species'
southern distributions (Molloy and Bolton 1996). This variation in
morphology was thought to represent two distinct species, with the
northern populations formerly described as \emph{Laminaria schinzii}
Foslie (Molloy and Bolton 1996). Genetic work has subsequently shown
that the two morphs are in fact the same species (Rothman et al. 2017).
In southern African waters, the primary production of \emph{Laminaria
pallida} is between 120 and 1900g C m2yr1, similar to that of
\emph{Ecklonia maxima} (Mann 1982). Primary production is not the only
pathway.

The aim of this study is, therefore, to understand how environmental
drivers such as temperature and wave energy can influence
morphoplasticity in two species of kelps around South Africa. This will
be achieved by initially understanding the variation in abiotic
parameters, and morphometrics of \emph{Ecklonia maxima} and
\emph{Laminaria pallida}, around the Western Cape coast. Thereafter we
will look at how the abiotic parameters both correlate and influence
each other in the nearshore environment of our study region. Finally, we
will investigate which abiotic parameters best describe the
morphological variation of the two kelp species, using statistical
tools. It is predicted that higher wave energy environments will exhibit
kelps with reduced undulations, possessing strap-like blades, opposed to
low flow regimes that will show greater undulations in frond morphology.
An increase in temperature is predicted to influence specific
morphological traits that rely on nutrient uptake, as temperature and
nutrients share an inverse relationship. An increase in temperature is
also predicted to reduce the dominance of the colder water kelp,
\emph{Laminaria pallida}, at depth, with a total absence of
\emph{Laminaria pallida} as one approaches the eastern boundary of
sample sites.

\hypertarget{study-area}{%
\subsubsection{Study area}\label{study-area}}

Due to the Cape Peninsula's temperate latitude, winter months bring an
increased frequency of frontal depressions that originate from the
Southern Ocean (Reason et al. 2006). These low pressures are joined by
large swells with increased wave energy. The nearshore environment, with
the accompanied biota, therefore experiences high wave energy events,
with increased frequency in winter (Veitch et al. 2018). The large
peninsula acts as an obstruction for large south westerly swells,
providing decreased wave energy along the west side of False Bay
(Shipley 1964). Conversely, the west coast of Cape Point is battered by
these large swells at full force. Multiple sites, therefore, exist where
kelps grow in diverse temperature and wave energy climates, in close
proximity.

In Figure 2 the total and seasonal wave roses of the directional wave
buoy just offshore from Cape Town is given form the years 2000 till end
2017 (the period for which directional wave spectra was available). It
is due to the near consistent South-westerly swell waves and the complex
orography around the peninsula, that the wave energy distribution around
the Cape Peninsula varies significantly over a small geographical area.
For details on the seasonality of the wave climate around the Cape
peninsula please refer to (Veitch et al. 2018).

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{Wave roses of the Cape Point wave rider buoy situated at
18.29oE, message=FALSE, warning=FALSE, 34.2oS in 70m water depth. (a)
total wave record, (b) winter waves and (c) summer waves.}
\end{figure}

In Figure 3 the total coastal wave exposure of the Cape Peninsula is
given in terms of wave energy (kW per meter wave crest length). These
values have been computed using model output from approximately 20 year
of simulated nearshore wave conditions. Hindcast global wave parameters
were refracted into the nearshore via a spectral phase-averaged
numerical wave modelling code SWAN (Simulating, WAves in the Nearshore -
3rd generation). The model is fully spectral in all frequencies and
directions (0o -360o) (can solve swell and local wind generated waves
simultaneously, propagating in different direction) and was used to
model the propagation of short crested swell waves, waves generated by
wind, non-linear wave-wave interaction and dissipation and depth induced
breaking and refraction. Wave energy dissipation due to bottom friction
and whitecapping is also accounted for ({\textbf{???}}).

Time series output from 1997 till 2013 of this model was then produced
at the 7m and 15m depth contour lines. Each nearshore output point thus
contained a 17-year timeseries of significant wave heights, peak periods
and mean wave directions (with both swell and wind waves). This data set
is available on a 500m resolution around most of the South African
coastline and on a 200m resolution with False Bay as provided by the
South African Department of Environmental Affairs (DEA) and Council for
Scientific and Industrial Research (CSIR)(@Theron2014). The total
17-year average energy, for each output location, was then computed
using linear wave theory, taking into account deep, intermediate and
shallow water wave approximations (based on wave length) for the wave
energy ({\textbf{???}}, {\textbf{???}}, {\textbf{???}}) which is a
function of wave height and period.

The directional sheltering effect of the Cape Peninsula, against the
dominant swell direction, (given in Figure 2) is clear observed in the
wave exposures presented in Figure 2. The classification from fully
sheltered to extremely exposed is based on the total wave energy upper
and lower limits. It should be noted that what is classified as
sheltered around the South African coastline (a high energy coastline)
might be classified as exposed in other regions of the world
({\textbf{???}}, Norderhaug et al. 2012, Sundblad et al. 2014). Kelp and
other ecosystems adapt to their typical environmental exposure and thus
what is classified as extreme abiotic coastal forcings (temperature,
winds, waves and currents) will vary around the world depending on the
geographical setting.

From Figure 2 the western periphery of the Cape peninsula is almost
continuously producing high coastal wave exposures while the Eastern
periphery of the peninsula (western coastline of False Bay) revealed
sheltered wave exposures. Here the marked seasonality, with higher
energy waves during winter, may be clearly observed once more.

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Coastal wave energy exposure around the Cape Peninsula for the
(a) total 17-year nearshore wave record, (b) only winter months and (c)
only summer months.}
\end{figure}

In Figure 3 the propagation of a typical offshore wave spectrum is given
as produced from a single time-step in SWAN. Figure 3 is presented to
clarify the averaged wave exposure maps presented in Figure 4. Tracing
the wave height contours into False Bay its clear why this Bay's western
periphery is predominantly sheltered. It should be mentioned that some
of the annual winter frontal depression systems pass the Cape Peninsula
from the west to east, resulting in wave propagating towards the
continent from much more southerly directions. This results in positive
and negative wave exposure anomalies all around the peninsula.

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{SWAN numerical output for an example typical offshore wave
conditions refracting around the Cape Peninsula.}
\end{figure}

\newpage

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{morphometrics-collection}{%
\subsubsection{Morphometrics
collection}\label{morphometrics-collection}}

Between October 2014 and April 2015, morphological measurements of
\emph{Laminaria pallida} and \emph{Ecklonia maxima} were collected at 18
sites along the Western Cape coast of South Africa (Fig. 1). Eleven
samples were collected per morphology, for each species (Table 1, 2).
These varying morphometrics allowed measurements such as weight, length
and thickness to be compared between sites. Because the macroalgae
differ in morphological features, species-specific morphometrics were
included. These sites span across the majority of the south-west coast,
in varying thermal and wave energy regimes.

Between February 2017 and Novevember 2018, morphological measurements
for \emph{Ecklonia maxima} individuals in shallow water (\textless{}1m)
at 5 sites along the Western Cape coast of South Africa were collected.
These sites represent a wave gradient moving south along the peninsula
and into False Bay. The same morphometric measurements were taken as per
the deeper \emph{Ecklonia maxima}, and this allowed comparison between
morphological characteristics between deep and shallow individuals.

\hypertarget{sites}{%
\subsubsection{Sites}\label{sites}}

Sites were chosen to represent an array of morphological differences
seen within \emph{Ecklonia maxima} and \emph{Laminaria pallida}. Sites
were also chosen to reflect locations that displayed variable wave and
temperature regimes (Fig. 2-5), to allow us to robustly test our
hypothesis of environmental drivers influencing kelp morphology.
St.~Helena Bay and Betty's Bay constituted the north western and south
eastern boundary sites respectively. These sites are roughly 300km
apart, and lie within separate marine provinces, as outlined above. The
Cape Peninsula provides an interesting topographical boundary that
shelters the coast in False Bay. Sites were therefore chosen to
represent an array of environments, from offshore reefs (Batsata Rock),
to sheltered intertidal zones (Miller's Point). West of Cape Point, a
number of sites were chosen to highlight the presence of upwelling
(Oudekraal, Kommetjie), as well as kelps growing in protected bays (Hout
Bay).

\hypertarget{abiotic-parameters}{%
\subsubsection{Abiotic parameters}\label{abiotic-parameters}}

In order to compare abiotic parameters for sites around the coast, large
historical databases for both temperature and wave energy were accessed.
Shallow water temperatures were sourced from the South African Coastal
Temperature Network (SACTN) website
(\url{https://github.com/ajsmit/SACTN}). Seven different organisations
within South Africa contribute to the SACTN, where \emph{in situ}
temperature measurements are made around the South African coast using
either hand-held thermometers or digital temperature recorders
positioned underwater. The mean duration of the 135 daily time series is
19.7 years, and these \emph{in situ} data are preferred over satellite
SST, which have shown to exhibit large biases (Smit et al. 2013). Linear
interpolated SST were calculated for sites where \emph{in situ}
recorders were absent. Wave energy data formed part of the South African
Coastal Vulnerability Assessment, presented to the Department of
Environmental Affairs ({\textbf{???}}). These data are first forecasted
using NOAA Wave Watch III (WWIII), with National Centers for
Environmental Predictions (NCEP) product as the numerical input
({\textbf{???}}). Hindcast data from WWIII span from 1994-2013 at a
3-hour resolution. The data are then used to model short --crested waves
generated by the wind into the coastal environment, using Simulating
Waves in the Nearshore SWAN ({\textbf{???}}). SWAN allows one to extract
wave parameters from specific gridded locations in the nearshore. For
False Bay, a resolution of 200 meters was modelled, at both 7 meter and
15 meter contours. A 200 meter resolution was used as False Bay was
nested within a larger grid area for the research from where the data
were sourced. For Table Bay and east of Cape Hangklip the resolution 500
meters at 7 meters and 15 meters. For this study the 7 meter contours
were used.

\#\#\#Statistical analyses

To compare how kelp morphology varies around the coast, boxplots were
constructed to summarise descriptive statistics for all the
morphometrics of both species.These boxplots highlight five different
descriptive statistics (Minimum, 25\textsuperscript{th} percentile,
median, 75\textsuperscript{th} percentile and maximum), as well as the
interquartile range. These allow us to visually identify variations and
differences in morphology, and to provide evidence that kelp
morphologies vary around the coast.Pairwise correlations were plotted to
compare the abiotic parameters and understand if wave and temperature
parameters correlate with one another around the coast. Therefore
fluctuations such as minimum, maximum, range and standard deviations
were included as temperature parameters, and standard deviations were
included as wave parameters. Median calculations were made for wind and
wave direction, as issues arise when calculating mean and standard
deviation for compass metrics. Redundancy Analyses (RDA) were performed
to understand how kelp morphology is driven by environmental drivers. An
RDA performs multiple linear regressions between explanatory and
response variables. This allows the user to calculate the amount of
variation in response variables explained for by explanatory variables.
Therefore response variables are influenced by explanatory variables.
Response variables were represented by morphology measurements, with
wave and temperature variables selected as explanatory variables.
Temperature and wave parameters were modelled separately, to fully
understand and tease apart which abiotic variables most strongly explain
kelp morphology variation. This was therefore performed for each
species, equating to four RDA's in total.

\newpage

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{temperature-parameters}{%
\subsubsection{Temperature parameters}\label{temperature-parameters}}

The temperature parameters around the Western Cape coast vary among
sites and seasons. During February and August, the west side of Cape
Point experiences decreased temperatures (Fig. 5). This region also
experiences lower mean and maximum temperatures when compared to False
Bay, a region known as east of Cape Point. The range of temperatures
within False Bay are larger in winter (August) compared to summer
(February) (Fig. 5). The sites north of Kommetjie on the west side
display larger temperatures ranges than sites south of Kommetjie on the
west side.

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-12-1.pdf}
\caption{Temperature parameters relative to each morphometric collection
site around the Western Cape coast. Temperature parameters include
minimum, maximum, mean, range and standard deviation (° Celsius). These
site locations are coloured by the temperature statistic relative to the
legends provided. Each temperature parameter is also divided into August
(winter) and February (summer) as well as the annual mean.}
\end{figure}

\hypertarget{wave-and-wind-parameters}{%
\subsubsection{Wave and wind
parameters}\label{wave-and-wind-parameters}}

Not only do the temperature parameters display seasonal fluctuations,
but we see similar observations for wave parameters. The west side of
Cape Point exhibit increased median direction, SD and mean significant
wave height compared to within False Bay (Fig. 4). Sites within False
Bay however display increased standard deviations of wave period. Median
wind direction is northerly in winter (Fig. 6), and turns southerly in
summer.

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-15-1.pdf}
\caption{Wave parameters relative to each morphometric collection site
around the Western Cape coast. These wave parameters include mean and
standard deviation of wave direction (° True north), significant wave
height (Meters) as well as wave period (Seconds). Sites are colour-coded
by the parameter statistic provided by the legend. Each wave parameter
is also divided into August (winter), February (summer) and annual
means, to visualise seasonal differences.}
\end{figure}

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-16-1.pdf}
\caption{Wind parameters relative to each morphometric collection site
around the Western Cape coast. These wind parameters include mean and
standard deviation of wind direction (° True north) as well as wind
speed (Meters/ second). Sites are colour-coded by the parameter
statistic provided by the legend. Each wave parameter is also divided
into August (winter), February (summer) and annual means, to visualise
seasonal differences.}
\end{figure}

\hypertarget{morphologies}{%
\subsubsection{Morphologies}\label{morphologies}}

\hypertarget{laminaria-pallida}{%
\paragraph{\texorpdfstring{\emph{Laminaria
pallida}}{Laminaria pallida}}\label{laminaria-pallida}}

Lamina length for laminaria pallida showed no geographical pattern
moving from west to east. Kommetjie, Olifantsbos and Batsata Rock showed
great variability in lamina length, and both Buffels Bay and Betty's Bay
were visually different to Miller's Point and Roman Rock, when comparing
summary data (Fig.8). Lamina thickness showed great variation across
sites, with Baboon Rock, Miller's Point, A-Frame and Roman rock
displaying large lamina thickness, significantly different to the rest
of the sites. Lamina weight was observed to vary for Baboon Rock and
Betty's Bay. Neighbouring sites A-Frame and Roman Rock also showed
visual difference when comparing boxplot summary statistics. An increase
in the number of digits was observed as one moved from Cape Point north
along the western side of False Bay. This ceased at Batstata Rock, which
exhibited significantly less digits compared to the previous site,
Bordjies reef North. Stipe diameter showed some geographical grouping,
with west of Cape Point sites exhibiting larger stipe diameters compared
to False Bay sites.

Stipe diameter decreased as one rounded the point into False Bay, where
a sudden, significant difference was seen between Bordjies reef North
and Batsata Rock. Greater variation of stipe length was observed for
sites outside of False Bay (Kommetjie, Olifantsbos and Betty's Bay).
Baboon Rock, Miller's Point, A-Frame and Roman Rock were again grouped
together exhibiting the lowest stipe lengths, that were all
significantly different to sites found west of Cape Point. Stipe mass
displays similar patterns to stipe length, with larger stipe lengths
west of Cape Point compared to within False Bay. The thallus mass of
\emph{Laminaria pallida} was observed to be greater for sites around
Cape Point, with an observed difference between Batsata Rock and
Bordjies reef North, similar to the number of digit patterns. Larger
total lengths were observed around Cape Point and Betty's Bay, with the
smallest total lengths at sites that exhibited the greatest lamina
thicknesses (Baboon Rock, Miller's Point, A-Frame and Roman Rock).

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-17-1.pdf}
\caption{Boxplots representing the different \emph{Laminaria pallida}
morphometrics measured around the Western Cape coastline, with the
X-axis depicting the specific morphology measured, with units provided.
Boxplots represent the minimum, 25\textsuperscript{th} percentile,
median and 75\textsuperscript{th} percentile of the morphometrics
measured. Interquartile range can be deduced as the different between
the 75\textsuperscript{th} and 25\textsuperscript{th} percentiles, and
dots represent outliers in the data. Sites are ordered sequentially on
the Y-axis by location along the coast. The top site is Sea Point and is
located at the north western boundary, and Betty's Bay as the bottom
site is located at the south eastern boundary, from our sample region.}
\end{figure}

\hypertarget{ecklonia-maxima}{%
\paragraph{\texorpdfstring{\emph{Ecklonia
maxima}}{Ecklonia maxima}}\label{ecklonia-maxima}}

For frond length, frond mass, stipe circumferences, stipe length and
total length of \emph{Ecklonia maxima} morphometrics, we see a gradual
increase in value as one moves south from St.~Helena Bay to Kommetjie
and Soetwater (Fig. 9). Hout Bay, Kommetjie and Soetwater show
similarities in their stipes lengths to Buffels Bay, Batsata Rock and
Betty's Bay, and are significantly different to west of Cape Point
counterpart sites such as Oudekraal and Scarborough. A difference for
the majority of the morphologies are seen between Soetwater and
Scarborough, with significant differences for epiphyte length, frond
length, frond mass, stipe length, stipe mass and total length. A
separation along the same stretch of coast (west of Cape Point) for
stipe mass shows that Hout Bay, Kommetjie and Soetwater are again
significantly different to neighboring west side sites such as Oudekraal
and Scarborough.

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-18-1.pdf}
\caption{Boxplots representing the different \emph{Ecklonia maxima}
morphometrics measured around the Western Cape coastline, with the
X-axis depicting the specific morphology measured, with units provided.
Boxplots represent the minimum, 25\textsuperscript{th} percentile,
median and 75\textsuperscript{th} percentile of the morphometrics
measured. Interquartile range can be deduced as the different between
the 75\textsuperscript{th} and 25\textsuperscript{th} percentiles, and
dots represent outliers in the data. Sites are ordered sequentially on
the Y-axis by location along the coast. The top site is St.~Helena Bay
and is located at the north western boundary, and Betty's Bay as the
bottom site is located at the south eastern boundary, from our sample
region.}
\end{figure}

When comparing morphologies bewteen deep and shallow water
\emph{Ecklonia maxima} individuals, clear pattern is seen in certian
morphological characteristics (Fig. 10). These are frond length, frond
mass, primary length, primary width and stipe circumference, and all
show a similar pattern. Significant differences in these morphological
characterstics at Kommetjie and Soetwater are evident, however there are
no significant differences from Buffel's Bay to Kalk Bay.

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-19-1.pdf}
\caption{Boxplots representing the different \emph{Ecklonia maxima}
morphometrics between deep and shallow water sites measured around the
Western Cape coastline, with the X-axis depicting the specific
morphology measured, with units provided. Boxplots represent the
minimum, 25\textsuperscript{th} percentile, median and
75\textsuperscript{th} percentile of the morphometrics measured.
Interquartile range can be deduced as the different between the
75\textsuperscript{th} and 25\textsuperscript{th} percentiles, and dots
represent outliers in the data. Sites are ordered sequentially on the
Y-axis by location along the coast. The top site is St.~Helena Bay and
is located at the north western boundary, and Betty's Bay as the bottom
site is located at the south eastern boundary, from our sample region.}
\end{figure}

\hypertarget{abiotic-correlations}{%
\subsubsection{Abiotic correlations}\label{abiotic-correlations}}

There are various strong correlations within the temperature parameters,
and within the wave parameters on an annual time scale (Fig. 11). SD
Significant wave height (Hs) and mean Hs showed a strong positive
correlate(0.92), as well as SD wave period (Tp) and mean Hs with a
strong negative correlation(-0.9) For SD wind speed and mean wind speed
a strong positive correlation existed (0.967). Minimum and mean
temperatures correlate strongly (0.918), as well as SD wind speed and
median wind direction (0.999). A strong correlation exists between SD
temperature and temperature range (0.942). Median wind direction and
mean wind speed are also correlated well (0.954). We however see no
strong correlations between wave and temperature parameters (Fig 12).

These correlations vary through the three timescales (Annual, August and
February). Although there are no strong correlations between temperature
and wave parameters, we see interesting differences between the seasonal
timescales. The mean temperature and mean wind speed are weakly
correlated in August (0.253)(Fig. 12), but are observed to be more
strongly negatively correlated in February (-0.788)(Fig. 13). The same
is seen for minimum temperature and median wind direction in August
(0.129) compared to February (-0.715).

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-21-1.pdf}
\caption{A correlation graph depicting the relationships that various
temperature and wave parameters share, at an annual timescale. The left
Y-axis as well as bottom X-axis represent the measurements for the
parameters, while the top X-axis and right Y-axis depict the abiotic
parameter name. The top right section of the graph represents
correlation coefficients between abiotic parameters. The bottom left
section provides individual linear regressions between each abiotic
parameter, with a fitted line. Each point of each linear graph
represents a collection site. The diagonal density plots represent the
spread of the data for each abiotic relationship.}
\end{figure}

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-22-1.pdf}
\caption{A correlation graph depicting the relationships that various
temperature and wave parameters share, during August. The left Y-axis as
well as bottom X-axis represent the measurements for the parameters,
while the top X-axis and right Y-axis depict the abiotic parameter name.
The top right section of the graph represents correlation coefficients
between abiotic parameters. The bottom left section provides individual
linear regressions between each abiotic parameter, with a fitted line.
Each point of each linear graph represents a collection site. The
diagonal density plots represent the spread of the data for each abiotic
relationship.}
\end{figure}

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-23-1.pdf}
\caption{A correlation graph depicting the relationships that various
temperature and wave parameters share, during February. The left Y-axis
as well as bottom X-axis represent the measurements for the parameters,
while the top X-axis and right Y-axis depict the abiotic parameter name.
The top right section of the graph represents correlation coefficients
between abiotic parameters. The bottom left section provides individual
linear regressions between each abiotic parameter, with a fitted line.
Each point of each linear graph represents a collection site. The
diagonal density plots represent the spread of the data for each abiotic
relationship.}
\end{figure}

\hypertarget{redundancy-analyses}{%
\subsubsection{Redundancy analyses}\label{redundancy-analyses}}

\hypertarget{waves-as-a-driver-of-ecklonia-maxima-morphometrics}{%
\paragraph{\texorpdfstring{Waves as a driver of \emph{Ecklonia maxima}
morphometrics}{Waves as a driver of Ecklonia maxima morphometrics}}\label{waves-as-a-driver-of-ecklonia-maxima-morphometrics}}

The morphometrics of \emph{Ecklonia maxima} were explained more by wave
parameters (75\%), than temperature parameters (66\%), when separate
RDA's were constructed (Fig. 14). The first two axes for wave parameters
driving \emph{Ecklonia maxima} morphology explained 34\% of the
variation. Stipe circumference was positively influenced by both mean
and SD annual significant wave height as well as annual mean wave
direction. There was also a negative influence by annual SD of wave
period and wave direction (which correlate with one another) on stipe
circumference. The primary length of \emph{Ecklonia maxima} was
influenced positively by annual mean wind direction, and negatively
explained by annual SD wind direction.

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-25-1.pdf}
\caption{An RDA, represented by the first two RDA axes, depict the
influence that annual wave parameters have on the morphology of
\emph{Ecklonia maxima} sporophytes. Environmental (explanatory)
variables, in this case annual wave parameters, are represented by blue
vectors extending from the origin. Response variables, in this case
\emph{Ecklonia maxima} morphologies, are representing by red points
ordinated across the plane. Sites are represented in black. The top
X-axis and right Y-axis represent the explanatory variable axes, while
bottom X-axis and left Y-axis represent the response variables axes}
\end{figure}

\hypertarget{temperature-as-a-driver-of-ecklonia-maxima-morphometrics}{%
\paragraph{\texorpdfstring{Temperature as a driver of \emph{Ecklonia
maxima}
morphometrics}{Temperature as a driver of Ecklonia maxima morphometrics}}\label{temperature-as-a-driver-of-ecklonia-maxima-morphometrics}}

Although temperature parameters for \emph{Ecklonia maxima} showed some
influence of the morphometric data, the explanation of morphology was
not strong, with an R\textsuperscript{2} value of 0.662 and a negative
adjusted R\textsuperscript{2} value (-0.138)(Fig. 15). Adjusted
R\textsuperscript{2} values are modified R\textsuperscript{2} values to
include the number predictors in a model. A negative adjusted R squared
values equate to low explanatory power. This is supported by the lack of
relationships between environmental (temperature) and response
(morphology) variables (Fig. 15).

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-27-1.pdf}
\caption{An RDA, represented by the first two RDA axes, depict the
influence that seasonal temperature parameters have on the morphology of
\emph{Ecklonia maxima} sporophytes. Environmental (explanatory)
variables, in this case seasonal temperature parameters, are represented
by blue vectors extending from the origin. Response variables, in this
case \emph{Ecklonia maxima} morphologies, are representing by red points
ordinated across the plane. Sites are represented in black. The top
X-axis and right Y-axis represent the explanatory variable axes, while
bottom X-axis and left Y-axis represent the response variables axes.}
\end{figure}

\hypertarget{waves-as-a-driver-of-laminaria-pallida-morphometrics}{%
\paragraph{\texorpdfstring{Waves as a driver of \emph{Laminaria pallida}
morphometrics}{Waves as a driver of Laminaria pallida morphometrics}}\label{waves-as-a-driver-of-laminaria-pallida-morphometrics}}

For \emph{Laminaria pallida}, the percentage of explained data by wave
parameters was 89\%, compared to temperature parameters explaining 82\%
(Fig. 16). By using adjusted R\textsuperscript{2} values, the first two
axes for wave parameters driving \emph{Laminaria pallida} morphology
explained 57\% of the variation of morphology in this species. An
increase in annual SD of wave period and wave direction was a strong
influence on lamina thickness of \emph{Laminaria pallida}. Annual mean
wave period (which negatively correlates with both annual SD of wave
period and wave direction) saw a strong positive influence on lamina
length, and a negative influence on lamina thickness. Stipe mass of
\emph{Laminaria pallida} was influenced by annual mean wind direction,
as well as annual mean significant wave height, wave direction and SD of
significant wave height. The Total length of \emph{Laminaria pallida}
specimens were similarly explained by annual mean significant wave
height, wave direction and SD of significant wave height.

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-29-1.pdf}
\caption{An RDA, represented by the first two RDA axes, depict the
influence that annual wave parameters have on the morphology of
\emph{Laminaria pallida} sporophytes. Environmental (explanatory)
variables, in this case annual wave parameters, are represented by blue
vectors extending from the origin. Response variables, in this case
\emph{Laminaria pallida} morphologies, are representing by red points
ordinated across the plane. Sites are represented in black. The top
X-axis and right Y-axis represent the explanatory variable axes, while
bottom X-axis and left Y-axis represent the response variables axes.}
\end{figure}

\hypertarget{temperature-as-a-driver-of-laminaria-pallida-morphometrics}{%
\paragraph{\texorpdfstring{Temperature as a driver of \emph{Laminaria
pallida}
morphometrics}{Temperature as a driver of Laminaria pallida morphometrics}}\label{temperature-as-a-driver-of-laminaria-pallida-morphometrics}}

Provided by adjusted R\textsuperscript{2} values, the first two axes of
temperature effect on \emph{Laminaria pallida} morphologies amounted to
only 11\%, but exhibited a stronger influence than on \emph{Ecklonia
maxima} (Fig. 17). The mean and minimum temperatures during February
were observed to negatively influence \emph{Laminaria pallida}
morphologies such as stipe mass, stipe length and stipe diameter, as
well as thallus mass and total length. Maximum august temperatures
however were observed to positively explain stipe diameter of
\emph{Laminaria pallida}.

\begin{figure}
\centering
\includegraphics{chapter_2_files/figure-latex/unnamed-chunk-31-1.pdf}
\caption{An RDA, represented by the first two RDA axes, depict the
influence that seasonal temperature parameters have on the morphology of
\emph{Laminaria pallida} sporophytes. Environmental (explanatory)
variables, in this case seasonal temperature parameters, are represented
by blue vectors extending from the origin. Response variables, in this
case \emph{Laminaria pallida} morphologies, are representing by red
points ordinated across the plane. Sites are represented in black. The
top X-axis and right Y-axis represent the explanatory variable axes,
while bottom X-axis and left Y-axis represent the response variables
axes.}
\end{figure}

\newpage

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

Kelps, in particular brown seaweeds, are robust and resilient organisms
that have been shown to adapt their morphology to suit local
environmental conditions (refs.). Wave exposure has been shown to be the
main driver of morphological variation in brown seaweeds as a strategy
to reduce drag and ultimately dislodgement (refs.). Although wave
exposure is regarded as an important abiotic variable that drives kelp
morphology, it does not act independently \{AJS: of what?\}. For
instance, a study by Wernberg and Thomsen (2005) investigated the effect
of wave exposure on the morphology of \emph{Ecklonia radiata} of six
locations along 1100 km of the southwest Australian coastline. The
authors concluded that wave exposure was the main driver of kelp
morphological adaptation in the study; however, they also noted the
variation in morphology between sites was not consistent and concluded
that wave parameters may play different roles in determining kelp
morphology \{AJS: What does this mean?\}. The results from this current
study confirm the idea that kelp morphology is driven by specific wave
parameters, particularly significant wave height, wave period and wave
direction \{AJS: what about?\}.

Seasonal variations in significant wave height (Hs), wave period and
wave direction were observed from the data. The direction of swell
swings to the south west in winter, generated by strong low pressures
that originate from the southern ocean (Reason et al. 2006). These
swells were found to correlate with increased wave period (Tp), wave
direction and Hs. False Bay is therefore shielded by these south
westerly swells by the Cape Peninsula (Shipley 1964). In summer, these
swells rotate anticlockwise and are able to enter False Bay, providing
an explanation for increased variability of Hs and Tp. However, changing
winds may also be influencing waves generated within False Bay. As seen
in Figure 1 the summer and winter swells are relativley constant in
terms of direction adn this may be driving the waves locally (I am still
looking for the report where Christo cites this). The wind speed and
wind direction support the presence of upwelling in summer, where
southerly winds blow parallel to the coast and trigger upwelling (Field
et al. 1980b). This is supported by the decreased in temperature in
summer, specifically along the west side of Cape Point. We however see a
differentiation of median wind directions in summer, with the western
coast of False Bay experiencing south easterly winds, compared to the
west side of Cape Point experiencing south westerly winds. It is
hypothesised that the topography and elevation along the Cape Peninsula
channels, shields winds along the strip of land. This is however absent
in winter, where strong northerly winds are experiences from St.~Helena
Bay to Betty's Bay (Field et al. 1980b). St.~Helena bay experiences
decreased median wave directions, and is protected by a headland,
similar to the sites found along the western side of False Bay, that are
shielded by the Cape Peninsula. There is however increased variability
in the wave direction from both False Bay sites and St.~Helena Bay,
which is encouraged by refraction of waves.

Morphological adaptation due to water motion may manifest itself in a
number of ways to high wave energy environments. For instance, reduction
of blade thickness, blade elongation, increase of stipe length, and
stipe circumference increase and force of attachment. Although this
study did not measure force of attachment, other morphological responses
to wave parameters was evident. An increase in mean annual wave period
saw an increase of lamina length of \emph{Laminaria pallida}, but a
decrease in lamina thickness. By decreasing the thickness of the lamina,
and increasing the length of the lamina (directly increasing surface
area), a larger more flexible kelp can survive in environments with
greater wave period. Increased thickness of cortical tissues within the
lamina aid photosynthetic ability ({\textbf{???}}), however increased
wave exposure deceases the boundary diffusion layer ({\textbf{???}},
{\textbf{???}}). Therefore \emph{Laminaria pallida} in increased wave
period sites may be able to reduce the need for thick lamina for
photosynthesis, as the increased wave period provide longer wave events,
which decreases the diffusion boundary layer and allow easier nutrient
uptake. Conversely we see that sites with reduced wave height and wave
period, such as Baboon Rock, Miller's Point, A-Frame and Roman Rock
displayed the greatest lamina thickness, with better photosynthetic
ability, for \emph{Laminaria pallida}. Pace ({\textbf{???}}) however
studied the effect of wave exposure on \emph{Macrocystis pyrifera}
morphologies, and found that lamina thickness had a strong positive
relationship with wave exposure. This may be species specific, with
\emph{Macrocystis pyrifera} attaining larger sizes, but with
incomparable stipe and lamina morphologies to \emph{Ecklonia maxima} and
\emph{Laminaria pallida}.

Lamina thickness showed positive correlation with annual SD wave period.
Although this variable may just negatively correlate with annual mean
wave period, another reason may exist. The greatest variation of wave
period and wave direction occurs in August along the western side of
False Bay. Kelps at these sites may have developed thicker lamina, with
a possible strategy of maximising photosynthesis at the expense of maybe
being dislodged through rare increased wave energy events. The
`spreading out' of the lamina may also suggest a `go with the flow'
tactic, to reduce breaking by being less stiff (Friedland and Denny
1995). Denny et al. (1997) explored the stress forces that wave period
played on \emph{Nereocystis luetkeana} by exposing various kelp plant
sizes to a range of wave period that would be found in coastal
environments. Results showed that for shorter kelps (smaller stipe
lengths), the maximum stress on a plant occurs at 10s wave periods, but
only at 5.5s for larger kelps that are similar in length to the water
depth. Similar results were observed for kelps with larger blades, with
maximal stress forces occurring at approximately 8s, compared to a
stress ratio of only 0.4 at 16s wave periods.

Mean annual significant wave height (Hs) was observed to influence
\emph{Ecklonia maxima} stipe circumference, with increased stipe
circumference at larger Hs. An increase in Hs saw greater wave actions
as waves are larger, as well as greater wave periods that are positively
correlated with Hs. An increase in stipe circumference provides
\emph{Ecklonia maxima} with a more rigid structure to withstand the
onslaught of waves. These greater stipe circumferences of \emph{Ecklonia
maxima} were observed at west side sites such as Soetwater and
Kommetjie. While these sites experience increased Hs, they also are
subject to seasonal upwelling events ({\textbf{???}}), that bring with
it cold clean water. \emph{Ecklonia maxima} are thought to be influenced
greatly by light attenuation (Rothman et al. 2017), and thus have
developed increased (stronger) stipe circumferences to be able to cope
in large Hs environments, to access cleaner, nutrient-filled water. The
same patterns are observed in \emph{Laminaria pallida} morphologies,
that display increased stipe diameters at Soetwater and Kommetjie as
well. This increase in stipe diameter correlates with stipe mass of
\emph{Laminaria pallida}, providing evidence that this species has
developed thicker, strong stipes to persist in this region. Denny et
al.~(1997) found that the stress ratio decreased as stipe lengths
increases, in a fixed depth. Therefore adult \emph{Ecklonia maxima}
sporophytes would experience a lower stress ratio compared to adult
\emph{Laminaria pallida} sporophytes in the same environment, where
\emph{Laminaria pallida} is often a subsurface species to \emph{Ecklonia
maxima}. Stipe circumference for \emph{Ecklonia maxima} were explained
by annual wave direction. For \emph{Ecklonia maxima}, the sporophytes
may be producing stipes that have a greater circumference to cope with
increased wave energy. The west side of the peninsula is more exposed to
significant wave heights during winter due to the direction of swell,
while other sites such as Betty's Bay and Basata rock are more protected
due to their location along the coast and the presence of headlands
which refract wave parameters.

Temperature parameters were not found to be a significant driver of kelp
morphology in this study. Although temperature plays an important role
in distribution of kelp and physiological functioning of adults and
gametophytes, there is little evidence that temperature is a driver of
morphological variation in kelps. Although turbidity is not a direct
wave or temperature parameter, turbidity is influence by various abiotic
variables. Increased swells around sandy beaches increase sedimentation
and turbidity ({\textbf{???}}), however wind direction and speed can
encourage upwelling, with brings to the surface cold, clean,
nutrient-rich water ({\textbf{???}}). Increased wind speed at sites
along the west side of the peninsula in a southerly direction trigger
upwelling events ({\textbf{???}}). We however see a lack of upwelling
events for sites such as Betty's Bay and Batsata Rock. These two sites
display \emph{Ecklonia maxima} sporophytes with the largest total
lengths, as well as Betty's Bay displaying the largest total length of
\emph{Laminaria pallida} sporophytes. The effect of depth on stipe
length is regarded as insignificant in this study as all kelp were
sampled at similar depths. Stipe elongation has been shown to be a
morphological adaptation in wave exposed environment, as a longer kelp
has time to extend back and forth in swell reducing the tension on the
holdfast. Stipe lengths of \emph{Ecklonia maxima} from Kommetjie, Hout
Bay and Soetwater are similar to Batsata rock and Betty's Bay, but
differs to west side neighbours such as Oudekraal and Scarborough.
Oudekraal and Scarborough experience more variability in wave direction,
however they are observed to experience smaller annual mean significant
wave heights, when compared to Kommetjie, Hout Bay, Soetwater, as well
as Batsata Rock and Betty's Bay. Friedland and Denny (1995) found that
longer kelps were able to reduce drag and tension as the sinusoidal
displacement of most waves would not `stretch' a kelp out long enough
before falling. The ability to grow longer stipes would provide the kelp
with a structure that could surpass most wave heights without the risk
of dislodgement.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

(More to add here but I wanted to add it so long) Mechanical dislodgment
through hydrological stress is the main source on mortality for kelp,
and previous research has shown that wave exposure is the main driver of
kelp morphology. Therefore, morphological adaptation is the main
strategy to reduce drag and ultimately dislodgement in hydrological
stressful environments. The results from this study support previous
findings, and shows that particular wave parameters (significant wave
height and median wave direction) act more strongly in driving kelp
morphological plasticity.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Anderson1997}{}%
Anderson, R. J. et al. 1997. Holdfasts of adult kelp Ecklonia maxima
provide refuges from grazing for recruitment of juvenile kelps. - Marine
Ecology Progress Series 159: 265--273.

\leavevmode\hypertarget{ref-DeBettignies2013}{}%
Bettignies, T. de et al. 2013. Size, not morphology, determines
hydrodynamic performance of a kelp during peak flow. - Marine Biology in
press.

\leavevmode\hypertarget{ref-Bologna1993}{}%
Bologna, P. and Steneck, R. S. 1993. Kelp beds as habitat for American
lobster Homarus americanus. - Marine \ldots{} in press.

\leavevmode\hypertarget{ref-Bolton2010}{}%
Bolton, J. J. 2010. The biogeography of kelps (Laminariales,
Phaeophyceae): A global analysis with new insights from recent advances
in molecular phylogenetics. - Helgoland Marine Research 64: 263--279.

\leavevmode\hypertarget{ref-Bolton1985a}{}%
Bolton, J. J. and Levitt, G. J. 1985. Light and temperature requirements
for growth and reproduction in gametophytes of Ecklonia maxima
(Alariaceae: Laminariales). - Marine Biology 87: 131--135.

\leavevmode\hypertarget{ref-Bolton1987}{}%
Bolton, J. J. and Anderson, R. J. 1987. Temperature tolerances of two
southern African Ecklonia species (Alariaceae: Laminariales) and of
hybrids between them. - Marine Biology 96: 293--297.

\leavevmode\hypertarget{ref-Bolton2012}{}%
Bolton, J. J. et al. 2012. South African kelp moving eastwards: the
discovery of Ecklonia maxima (Osbeck) Papenfuss at De Hoop Nature
Reserve on the south coast of South Africa. - African Journal of Marine
Science 34: 147--151.

\leavevmode\hypertarget{ref-Bustamante1996}{}%
Bustamante, R. H. and Branch, G. M. 1996. The dependence of intertidal
consumers on kelp-derived organic matter on the west coast of South
Africa. - Journal of Experimental Marine Biology and Ecology 196: 1--28.

\leavevmode\hypertarget{ref-Cousens1982}{}%
Cousens, R. 1982. The Effect of Exposure to Wave Action on the
Morphology and Pigmentation of Ascophyllum nodosum (L.) Le Jolis in
South-Eastern Canada. - Bontanica Marina XXV: 191--195.

\leavevmode\hypertarget{ref-Dayton1999}{}%
Dayton, P. K. et al. 1999. Temporal and spatial patterns of kelp
demography: The role of Oceanographic climate. - Ecological Monographs
69: 219--250.

\leavevmode\hypertarget{ref-Denny1997}{}%
Denny, M. et al. 1997. Flow and Flexibility: II. THE ROLES OF SIZE AND
SHAPE IN DETERMINING WAVE FORCES ON THE BULL KELP NEREOCYSTIS LUETKEANA
MARK. - Journal of Experimental Biology 200: 3165--3183.

\leavevmode\hypertarget{ref-Duggins2003}{}%
Duggins, D. O. et al. 2003. Population, morphometric and biomechanical
studies of three understory kelps along a hydrodynamic gradient. -
Marine Ecology Progress Series 265: 57--76.

\leavevmode\hypertarget{ref-Field1980b}{}%
Field, J. G. et al. 1980a. Variation in Structure and Biomass of Kelp
Communities Along the South-West Cape Coast. - Transactions of the Royal
Society of South Africa 44: 145--203.

\leavevmode\hypertarget{ref-Field1980a}{}%
Field, J. G. et al. 1980b. Upwelling in a nearshore marine ecosystem and
its biological implications. - Estuarine and Coastal Marine Science 11:
133--150.

\leavevmode\hypertarget{ref-FowlerWalker2006}{}%
Fowler-Walker, M. J. et al. 2006. Differences in kelp morphology between
wave sheltered and exposed localities: Morphologically plastic or fixed
traits? - Marine Biology 148: 755--767.

\leavevmode\hypertarget{ref-Friedland1995}{}%
Friedland, M. T. and Denny, M. W. 1995. Surviving hydrodynamic forces in
a wave-swept environment: consequences of morphology in the feather boa
kelp, Egregia menziesii (Turner). - Journal of Experimental Marine
Biology and \ldots{} in press.

\leavevmode\hypertarget{ref-Gaines1987}{}%
Gaines, S. D. and Roughgarden, J. 1987. Fish in offshore kelp forests
affect recruitment to intertidal barnacle populations. - Science 235:
479--481.

\leavevmode\hypertarget{ref-Graham2007}{}%
Graham, M. H. et al. 2007. Global Ecology of the Giant Kelp Macrocystis
: From Ecotypes To Ecosystems. - Oceanography and Marine Biology 45:
39--88.

\leavevmode\hypertarget{ref-Levin1994}{}%
Levin, P. S. 1994. Fine-scale temporal variation in recruitment of a
temperate demersal fish: the importance of settlement versus
post-settlement loss. - Oecologia 97: 124--133.

\leavevmode\hypertarget{ref-Luning1990}{}%
Lüning, K. 1990. Seaweeds: their environment, biogeography, and
ecophysiology (C Yarish and H Kurkman, Eds.). in press.

\leavevmode\hypertarget{ref-Mann1973}{}%
Mann, K. H. 1973. Seaweeds: Their Productivity and Strategy for Growth.
- Science 182: 975--981.

\leavevmode\hypertarget{ref-Mann1982}{}%
Mann, K. H. 1982. Ecology of Coastal Waters: A Systems Approach. in
press.

\leavevmode\hypertarget{ref-Molloy1996}{}%
Molloy, F. J. and Bolton, J. J. 1996. The effects of wave exposure and
depth on the morphology of inshore populations of the Namibian kelp,
Laminaria schinzii Foslie. - Botanica Marina 39: 525--531.

\leavevmode\hypertarget{ref-moss1948}{}%
Moss, B. 1948. Studies in the genus Fucus: I. on the structure and
chemical composition of Fucus vesiculosus from three Scottish
localities. - Annals of Botany 12: 267--279.

\leavevmode\hypertarget{ref-Norderhaug2012}{}%
Norderhaug, K. M. et al. 2012. Does the diversity of kelp forest
macrofauna increase with wave exposure? - Journal of Sea Research 69:
36--42.

\leavevmode\hypertarget{ref-pedersen2012a}{}%
Pedersen, M. F. and Nejrup, L. B. 2012. Effects of wave exposure on
population structure, demography, biomass and productivity of the kelp
Laminaria hyperborea. - Marine Ecology Progress Series 451: 45--60.

\leavevmode\hypertarget{ref-Probyn1985}{}%
Probyn, T. and McQuaid, C. 1985. In-situ measurements of nitrogenous
nutrient uptake by kelp (Ecklonia maxima) and phytoplankton in a
nitrate-rich upwelling environment. - Marine Biology 88: 149--154.

\leavevmode\hypertarget{ref-Reason2006}{}%
Reason, C. J. C. et al. 2006. Seasonal to decadal prediction of southern
African climate and its links with variability of the Atlantic ocean. -
Bulletin of the American Meteorological Society 87: 941--955.

\leavevmode\hypertarget{ref-Rothman2017}{}%
Rothman, M. D. et al. 2017. Geographical variation in morphology of the
two dominant kelp species, Ecklonia maxima and Laminaria pallida
(Phaeophyceae, Laminariales), on the west coast of Southern Africa. -
Journal of Applied Phycology 29: 2627--2639.

\leavevmode\hypertarget{ref-Santelices2007}{}%
Santelices, B. 2007. The discovery of kelp forests in deep-water
habitats of tropical regions. - Proceedings of the National Academy of
Sciences 104: 19163--19164.

\leavevmode\hypertarget{ref-Shipley1964}{}%
Shipley, A. 1964. Some aspects of wave refraction in False Bay. - South
African Journal of Science 60: 115--120.

\leavevmode\hypertarget{ref-Smit2013}{}%
Smit, A. J. et al. 2013. A coastal seawater temperature dataset for
biogeographical studies: Large biases between in situ and
remotely-sensed data sets around the coast of South Africa. - PLoS ONE
in press.

\leavevmode\hypertarget{ref-Stegenga1997}{}%
Stegenga, H. et al. 1997. Seaweeds of the South African west coast. -
Contributions of the Bolus Herbarium 18: 3--637.

\leavevmode\hypertarget{ref-Steneck2002}{}%
Steneck, R. S. et al. 2002. Kelp forest ecosystems: biodiversity,
stability, resilience and future. - Environmental Conservation 29:
436--459.

\leavevmode\hypertarget{ref-Sundblad2014}{}%
Sundblad, G. et al. 2014. Comparing the ecological relevance of four
wave exposure models. - Estuarine, Coastal and Shelf Science 140: 7--13.

\leavevmode\hypertarget{ref-Veitch2018}{}%
Veitch, J. et al. 2018. 'The Cape Point wave record and the role of
large-scale modes of climate variability'. - Reprint submitted to the
Journal of Marine Systems: 131--135.

\leavevmode\hypertarget{ref-Velimirov1977}{}%
Velimirov, B. et al. 1977. The ecology of kelp bed communities in the
Benguela upwelling system - Analysis of biomass and spatial
distribution. - Helgoländer Wissenschaftliche Meeresuntersuchungen 30:
495--518.

\leavevmode\hypertarget{ref-Wernberg2005}{}%
Wernberg, T. and Thomsen, M. S. 2005. The effect of wave exposure on the
morphology of Ecklonia radiata. - Aquatic Botany 83: 61--70.

\leavevmode\hypertarget{ref-Zimmerman1984}{}%
Zimmerman, R. C. and Kremer, J. N. 1984. Episodic nutrient supply to a
kelp forest ecosystem in Southern California. - Journal of Marine
Research 42: 591--604.

\end{document}

### 2018-11-20 chapter_2.Rmd
commit

--- 2018-11-08 chapter_2.Rmd
@@
 
-# fix incorrect
-
 ```
@@
 The diagonal density plots represent the spread of the data for each abiotic relationship.", fig.height=8, fig.width=12, fig.pos="H"}
-# # ggpairs to plot correlation coefficients
-# # Annual
+## ggpairs to plot correlation coefficients
+## Annual
 ggpairs(data = ann_abio_cor, columns = 2:14, 
@@
 
-# # August
+## August
 ggpairs(data = aug_abio_cor, columns = 2:14, 
@@
 
-# # February
+## February
 ggpairs(data = feb_abio_cor, columns = 2:14, 

### 2018-12-14 stats_chapter_2.R
updated shallow morphology data set. include VIF calculation after each rda

@@ -822,11 +822,9 @@ ggpairs(data = feb_abio_cor, columns = 2:14,
   theme(text = element_text(size = 9)) + 
   ggtitle("February")
 
-### Redundancy analyses 
+#### Redundancy analyses 
 #### USing RDA's to see how wave parameters and temperature parameters
-# influence the morphology pf kelp species
-
-###
+#### influence the morphology of kelp species
 
 # rename eck_all to eck-site0
 
@@ -867,6 +865,8 @@ eck_wave_bio <- eck_wave_bio%>%
 # run RDA
 eck_wave_RDA <- rda(eck_wave_bio ~ ., data = eck_wave)
 
+vif.cca(eck_wave_RDA)
+
 # Plot scaling = 2
 plot(eck_wave_RDA, scaling  = 2)
 
@@ -892,6 +892,9 @@ eck_temp_season <- eck_temp_season %>%
 
 # run RDA
 eck_temp_season_RDA <- rda(eck_wave_bio ~ ., data = eck_temp_season)
+
+vif.cca(eck_temp_season_RDA)
+
 
 # Plot scaling = 2
 plot(eck_temp_season_RDA, scaling = 2)
@@ -940,11 +943,14 @@ lam_wave_bio <- lam_wave_bio%>%
 # run RDA
 lam_wave_RDA <- rda(lam_wave_bio ~ ., data = lam_wave)
 
+vif.cca(lam_wave_RDA)
+
 # Plot scaling = 2
 plot(lam_wave_RDA, scaling  = 2)
 
 # Summary
 summary(lam_wave_RDA)
+
 ## Temp
 
 # Subset temps for Lam sites (seasons)
@@ -964,6 +970,8 @@ lam_temp_season <- lam_temp_season %>%
 
 # run RDA
 lam_temp_season_RDA <- rda(lam_wave_bio ~ ., data = lam_temp_season)
+
+vif.cca(lam_temp_season_RDA)
 
 # Plot scaling = 2
 plot(lam_temp_season_RDA, scaling = 2)

### 2018-12-15 stats_chapter_2.R
added more code to work out VIF and forward selection using ordistep for the RDA investigating wave variables and ecklonia morphology.

@@ -840,7 +840,6 @@ eck_wave <- subset(eck_site0, select = c(27:32, 45, 46))
 eck_wave <- eck_wave %>%
   decostand(method = "standardize")
 
-
 # subset ecklonia morpho's
 eck_wave_bio <- subset(eck_site0, select = c(2:11))
 
@@ -855,7 +854,6 @@ eck_wave <- eck_wave %>%
   remove_rownames %>%
   column_to_rownames(var = "site")
 
-
 eck_wave_bio <- cbind(eck_site, eck_wave_bio)
 
 eck_wave_bio <- eck_wave_bio%>%
@@ -865,13 +863,55 @@ eck_wave_bio <- eck_wave_bio%>%
 # run RDA
 eck_wave_RDA <- rda(eck_wave_bio ~ ., data = eck_wave)
 
-vif.cca(eck_wave_RDA)
-
 # Plot scaling = 2
 plot(eck_wave_RDA, scaling  = 2)
 
 # Summary
 summary(eck_wave_RDA)
+
+# Use the "cor" function to calculate Pearson correlation between predictors.
+
+cor_eck_wave <- round(cor(eck_wave, use = "pair"), 2)
+cor_eck_wave
+
+# Calculate "Variance Inflation Factor" using vegan
+
+vif.cca(eck_wave_RDA)
+
+# Global adjusted R^2
+(R2a.all <- RsquareAdj(eck_wave_RDA)$adj.r.squared) 
+
+# Forward selection using vegan's ordistep
+# Note that is function does not use R^2adjusted-based stopping criterion
+
+step.forward <- ordistep(rda(eck_wave_bio ~ 1, data = eck_wave),
+                         scope=formula(eck_wave_RDA), direction = "forward", pstep = 1000)
+
+# Application of stopping criterion, threshold of 0.585435
+# Copy and paste code and the add predictor, rinse and repeat until threshold reached
+
+RsquareAdj(rda(eck_wave_bio ~ Ann_tp_sd, data = eck_wave))$adj.r.squared
+RsquareAdj(rda(eck_wave_bio ~ Ann_tp_sd + Ann_spw_mean, data = eck_wave))$adj.r.squared
+RsquareAdj(rda(eck_wave_bio ~ Ann_tp_sd + Ann_spw_mean + Ann_spw_sd, data = eck_wave))$adj.r.squared
[…]

### 2018-12-17 stats_chapter_2.R
Updated RDAs' and added new RDA for annual temperature variables

@@ -913,7 +913,7 @@ vif.cca(eck_wave_pars)
 # Plot parsimonious RDA
 plot(eck_wave_pars, scaling  = 2)
 
-## Temp
+## Temp: seasonal
 
 # Subset temps for ecklonia sites (for seasons)
 eck_temp_season <- subset(eck_site0, select = c(17:26))
@@ -933,6 +933,9 @@ eck_temp_season <- eck_temp_season %>%
 # run RDA
 eck_temp_season_RDA <- rda(eck_wave_bio ~ ., data = eck_temp_season)
 
+# Plot RDA
+plot(eck_temp_season_RDA, scaling  = 2)
+
 # Use the "cor" function to calculate Pearson correlation between predictors.
 
 cor_eck_temp_season <- round(cor(eck_temp_season, use = "pair"), 2)
@@ -951,11 +954,75 @@ vif.cca(eck_temp_season_RDA)
 step.forward <- ordistep(rda(eck_wave_bio ~ 1, data = eck_temp_season),
                          scope=formula(eck_temp_season_RDA), direction = "forward", pstep = 1000)
 
-# Plot scaling = 2
-plot(eck_temp_season_RDA, scaling = 2)
+# Parsimonious RDA
+
+eck_temp_season_pars <- rda(eck_wave_bio ~ Aug_mean_temp + Feb_mean_temp, data = eck_temp_season)
+eck_temp_season_pars
+anova.cca(eck_temp_season_pars, step = 1000)
+anova.cca(eck_temp_season_pars, step = 1000, by = "axis")
+vif.cca(eck_temp_season_pars)
+(R2a.pars <- RsquareAdj(eck_temp_season_pars)$adj.r.squared)
+
+# Plot parsimonious RDA
+plot(eck_temp_season_pars, scaling  = 2)
 
 # Summary
-summary(eck_temp_season_RDA)
+summary(eck_temp_season_pars)
+
+## Temp: Annual
+
+# Subset temps for ecklonia sites (for seasons)
+eck_temp_annual <- subset(eck_site0, select = c(12:16))
+
+# standardise wave measurements
+eck_temp_annual <- eck_temp_annual %>%
+  decostand(method = "standardize")
+
+# Force site names as column 0
+eck_temp_annual <- cbind(eck_site, eck_temp_annual)
+
+eck_temp_annual <- eck_temp_annual %>%
+  remove_rownames %>%
[…]

### 2019-07-02 paper_morph_new.Rmd
add new paper draft as rmd.file. I have also played around with extra graphs and plots but those scripts are messy and won't be added.

---
title: Temperature, wave exposure and wind as drivers of morphology in two kelp species
  along the coast of South Africa
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera (Bolton 2010), nevertheless form the basis of one of the most productive ecosystems globally (Mann 1973). Kelps generally have a dependence on cool, temperate and arctic seawater temperatures (Bolton 2010; Mohring et al. 2014; Cavanaugh et al. 2011; Dayton 1985), and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres (Steneck et al. 2002). Their size and complex morphology provide a heterogeneous habitat structure (Steneck et al. 2002) that accommodate a multitude of turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates (Mann 1973; Steneck et al. 2002), each depending on a wide suite of ecological services provided by the kelp forests (Gaines & Roughgarden 1987; Paul & Steneck 1993; Levin 1994; Willis & Anderson 2003; Anderson et al. 1997; Dayton 1985). Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in the distribution (Gorman et al. 2013), abundance (Cavanaugh et al. 2011; Dayton et al. 1998), diversity (Wing et al. 2007; Wernberg & Goldberg 2008), composition (Leliaert et al. 2000; Dayton 1985; Norderhaug et al. 2012; Harley et al. 2012), growth (Cousens 1982) and productivity (Pedersen & Nejrup 2012; Dayton et al. 1998; Krumhansl & Scheibling 2012) of kelps.

Wave exposure and temperature have also been shown to be the main drivers of morphological adaptation across various kelp species. Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility (Denny & Gaylord 2002). For example, a study by (Wernberg & Thomsen 2005)) examined the consistency of wave exposure as a driver of *Ecklonia radiata* (C. Agardh) J. Agardh across a broad geographic range, and  showed trends towards drag-reducing (small size, narrow laterals and blades, low spinosity) and increased strength (large holdfast, thick stipe and thick blades and lamina) at high wave exposure sites. The morphological adaptation to wave exposure must be kept in balance with other important processes, such as nutrient assimilation and light absorption, which are dependent on the amount of surface area of the blades. Therefore, there is a trade-off between reducing overall drag and maintaining nutrient and photosynthetic ability. Temperature has also been shown to play a role in driving kelp morphology. For example, a study by (Serisawa et al. 2002) compared the morphology of Ecklonia cava Kjellman growing in warm temperate and cool temperate morphologies. The results showed that wrinkles in the blade seem to be a characteristic of warm temperate regions. However this study did not take into account the interaction with other environmental variables such as wave exposure, which also affect kelp morphology. The morphological adapatibility of kelps is driven by a combination of environmental factors that in turn do not act independently of one another. A study by (Wernberg et al. 2003) investigated the morphology of *E. radiata* in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australasian coast. They found no correlation between spatial distance and morphological similarity and rather the morphology of kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales (Wernberg et al. 2003). Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species. Wave exposure may be greater in shallower environments, or it may be reduced through damping effects of kelp further offshore and natural barriers such as rock outcrops; also, seawater temperature may be higher in shallow water environments due to higher solar irradiance. Wind has also been shown to play a role in influencing stipe length. A study by (Denny et al. 1997) used a combination of field and laboratory tests to model the drag properties on the canopy kelp *Nereocystis luetkeana* (K. Mertens) Postels & Ruprecht. The results from the model suggest that wind may play a role in the extension capabilities of the stipe under certain wave/wind conditions. Furthermore, wind generates surface-gravity waves under certain conditions, and may influence the overall wave climate at a specific location in-time (Holthuijsen 2010).

Comparatively little research has been done on other species of kelp, such as *Laminaria pallida* and *Ecklonia maxima* (Osbeck) Papenfuss  which are important habitat forming seaweeds that exist around the coast of South Africa. Although both these species exist in the subtidal, *L. pallida* dominates deeper waters while *E. maxima* forms dense surface canopies in the shallower waters. Therefore, *E. maxima* is exposed to variations in waved, swell and temperature, while *L. pallida* is exposed to variations in waves and temperature. (Molloy & Bolton 1996) investigated the effect of depth and wave exposure on *L. pallida* at different depths and wave exposure and  showed that depth had a greater effect than wave exposure when considering all the morphological characteristics; when considering individual characteristics, however, wave exposure had the most significant effect on blade thickness. Another study by (Rothman et al. 2017) investigated the changes in morphology in shallow populations of *L. pallida* and *E. maxima* along the South African coastline and into Namibia; *E. maxima* exhibited no morphological changes along the coast but the stipes of *L. pallida* become increasingly hollow further north along the coastline. They suggested that turbidity was the environmental driver responsible for this change. Research using advanced numerical tools is lacking. Previous research investigating drivers of kelp morphological characteristics used qualitative estimates of wave exposure and did not consider other morphological characteristics. Furthermore, only shallow populations have been considered in previous research which ignores the effects of wave damping. No work currently exists where other morphological characters have been considered, quantitative estimates of environmental drivers have been calculated, and the addition of wind as a possible driver of morphological characteristics has been considered.

The aim of this study is, therefore, to understand how temperature, wave exposure and wind can influence the morphology in two species of kelps around South Africa. This will be achieved by initially understanding the variation in temperature and waves and the consequences for morphological characteristics of *E. maxima* and *L. pallida*.

## Methods
### Site selection

Sites were chosen to represent an array of environmental gradients, as indicated in Fig. 1. St. Helena Bay and Betty?s Bay constituted the north western and south eastern boundary sites, respectively.

These sites are roughly 300 km apart, and lie within separate marine provinces[3] and cover a cline in wave exposure and temperature. Study sites  span across the majority of the south-west coast, in varying thermal and wave energy regimes. The region is dominated by kelp communities that persist in contrasting abiotic environments. The west coast region has been termed as ?cool temperate? region, which is defined as a region where mean monthly temperatures are always above 10?C and always below 15?C (Smit et al. 2013). East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone (Smit et al. 2013). The Agulhas Marine Province is characterised by a wide temperature range of up to 7?C difference between mean monthly temperatures between summer and winter (Smit et al. 2013). Due to the Cape Peninsula temperate latitude, winter months bring an increased frequency of frontal depressions that originate from the Southern Ocean (Reason et al. 2006). These low pressures are joined by large swells with increased wave energy. The nearshore environment, with the accompanied biota, therefore experiences high wave energy events, with increased frequency in winter (Veitch et al. 2018). The large peninsula acts as an obstruction for large south westerly swells, providing decreased wave energy along the west side of False Bay (Shipley 1964). Conversely, the west coast of Cape Point is battered by these large swells. Multiple sites, therefore, exist where kelps grow in diverse temperature and wave energy climates, in close proximity.

The topography and elevation along the Cape Peninsula channel and shield winds along False Bay. This is however absent in winter, where strong northerly winds are prevalent from St. Helena Bay to Betty?s Bay (Field et al. 1980; Jury et al. 1985; Andrews & Hutchings 1980; Jury 1980). Patterns are also evident in the wave data. Seasonal variations in significant wave height (Hs), wave period and wave direction are also present. The direction of the dominant swell changes to the south west in winter, generated by strong low pressures that originate from the southern ocean (Reason et al. 2006), which False Bay is sheltered from (Shipley 1964; Atkins 1970; Dufois & Rouault 2012). In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an increased variability of Hs and Tp in this region. It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world (Norderhaug et al. 2012; Leliaert et al. 2000). It is due to the near consistent south-westerly swell and the complex orography around the peninsula that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.

## Study species

The morphological characteristics of both species are presented in Fig. 2. *Laminaria pallida* is characterised by a single smooth blade which is divided longitudinally into sections, and develops from a single meristematic region located at the junction between the blade and the stipe (Dyer 2018). This species has a solid stipe but develops a hollow stipe along the west coast northward and into Namibia (Rothman et al. 2017). *Ecklonia maxima* consists of a single primary blade which develops above a gas-filled float and a hollow stipe below. Secondary blades are produced laterally from the primary blade from several meristematic regions along the margins of the primary blade, known as digits (Dyer 2018). Both species are held to the substrate by finger-like haptera, collectively known as the holdfast.

## Abiotic variables

In order to compare abiotic variables for sites around the coast, large historical databases for temperature, wave energy and wind were accessed. Shallow water temperatures were sourced from The South African Coastal Temperature Network (SACTN) website (https://github.com/ajsmit/SACTN). In terms of nearshore temperature, in situ data are preferred over satellite SST, which have shown to exhibit large biases (Smit et al. 2013). Linear interpolated SST were calculated for sites where in situ recorders were absent. Wave model variable data formed part of the South African Coastal Vulnerability Assessment, presented to the Department of Environmental Affairs (DEA) and produced by the Council for Scientific and Industrial Research (CSIR) (Rautenbach 2015). The South African wave climate was modeled via 20 spectral numerical wave models that simulated the offshore wave climate to the nearshore. The boundary conditions of these models were obtained by using the NOAA Wave Watch III (WWIII) model output, distributed via the National Centers for Environmental Prediction (NCEP) product (Office of the Director 2000; Environmental Modeling Center / Marine Modeling Branch 2005). The particular hindcast product utilised during the DEA-CSIR study spans 1994-2013 at a 3-hour resolution. These data were then used to model swell propagation into the coastal models while wind waves (seas) were generated via stationary computations in the Simulating Waves in the Nearshore SWAN model (Booij et al. 1997). The assumption of stationary computations are acceptable as the model domains were small enough so the temporal variation of the model boundary were slower than the time it takes for that boundary condition to propagate to the coast.  SWAN allows one to extract wave variables from specific gridded locations in the nearshore. For False Bay, a resolution of 200 meters was modelled and output produced at both the 7 meter and 15 meter isobaths. A 200 meter resolution was used as the False Bay computational grid was nested within a larger grid (1 kilometer resolution). This allowed for a computational effective wave resolution of increasing resolution, from the NCEP, low resolution output to nested, high resolution coastal output. For Table Bay and east of Cape Hangklip the resolution was 500 meters and also had output at the 7 meters and 15 meter isobaths. These contour outputs were chosen in the original study by the CSIR as most engineering run-up calculations require wave parameter information at these contour depths and were the main focus of the original study. For this study the 7 meter contours were used. Site-specific summary statistics were calculated for each of the environmental variables considered in this study. The summary statistics calculated for temperature, waves and wind variables were minimum, maximum, mean, range and standard deviation for annual, February[6] (Austral summer) and August (Austral winter) timescales, respectively. In addition, annual and seasonal median wave and wind direction were also calculated; all summary data and their respective abbreviation are presented in Table 3. The median was calculated for wind and wave direction, as issues arise when calculating the mean and standard deviation for compass metrics. Dot and whisker plots were used to present the summary statistics for both temperature and wave variables. Summary statistics for wind was not plotted and instead are discussed, as the data was course relative to the other environmental variables considered in this study.

## Results
### Temperature variables

```{r echo=FALSE}

## Data wrangling and preparation

library(tidyverse)
library(ggpubr)
library(gridExtra)

## Preparing data
# Complete lon/lats
kelp_sites <- read_csv("metadata/sites_complete.csv")

# Ecklonia
eck_data <- read_csv("data/Ecklonia_morphometrics.csv") %>%
  mutate(site = replace(site, site == "Buffels Bay", "Buffelsbaai")) %>%
  mutate(site = replace(site, site == "St. Helena Bay", "St Helena Bay")) %>%
  # Not using Rocky Bank in this study
  filter(site != "Rocky Bank")

# Laminaria
lam_data <- read_csv("data/Laminaria_morphometrics.csv") %>%
  mutate(site = replace(site, site == "Buffels Bay", "Buffelsbaai")) %>%
  # NB: Even though I am changing the name, there will still be no lon/lat values for this site
  # We need to decide how we want to handle this. Give it the same lon/lat as Buffelsbaai?
  mutate(site = replace(site, site == "Buffels Bay South", "Buffelsbaai South")) %>%
  # NB: This may also not be an appropriate change
  mutate(site = replace(site, site == "Bordjies reef North", "Bordjies Deep")) %>%
  # Not using Rocky Bank in this study
  filter(site != "Rocky Bank")

# NB: (RWS) I have removed Roman Rock as this has less than a year of temperature data

# The SACTN monthly data
load("data/SACTNmonthly_v4.1.Rdata")

# The missing site temperatures
load("data/interp_sites.Rdata")

# The patched Roman Rock temperatures
load("data/rr_patch.Rdata")

## Subset temperatures for collection sites
temps <- SACTNmonthly_v4.1 %>% 
  filter(site %in% eck_data$site | site %in% lam_data$site) %>% 
  select(site, date, temp) %>% 
  rbind(., interp_sites) %>% 
  filter(site!= "Roman Rock") %>% 
  rbind(., rr_patch)

## Temperature analyses
# Up next we need to calculate the relevant temperature statistics for each site. These are: #mean, range, and sd, for Annual, February, and August climatologies.
# Monthly clims

temp_feb_aug <- temps %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  group_by(site, date) %>% 
  summarise(mean_temp = mean(temp, na.rm = T),
            min_temp = min(temp, na.rm = T),
            max_temp = max(temp, na.rm = T),
            range_temp = range(temp, na.rm = T)[2]-range(temp, na.rm = T)[1],
            sd_temp = sd(temp, na.rm = T)) %>%
  filter(date %in% c("Feb", "Aug")) %>% 
  ungroup()

# Annual clims
temp_ann <- temps %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  group_by(site, date) %>% 
  summarise(mean_temp = mean(temp, na.rm = T),
            min_temp = min(temp, na.rm = T),
            max_temp = max(temp, na.rm = T),
            range_temp = range(temp, na.rm = T)[2]-range(temp, na.rm = T)[1],
            sd_temp = sd(temp, na.rm = T)) %>% 
  ungroup() %>% 
  select(-date) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  mutate(date = "Ann") %>% 
  select(site, date, everything())

# Combine
temp_clim <- rbind(temp_feb_aug, temp_ann) %>% 
  group_by(site) %>% 
  gather(key = statistic, value = value, 
         mean_temp:sd_temp) %>% 
  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
  select(-date, -statistic) %>% 
  spread(key = clim_stat, value = value)

```

```{r}

### Plots for temperature variables

positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","De Hoop", "St Helena Bay")

# Annual temp
temp_plot_A <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  ylim(8, 22) + 
  geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) + 
  geom_point(data = temp_ann, aes(x = site, y = max_temp, colour = "blue")) + 
  geom_point(data = temp_ann, aes(x = site, y = min_temp, colour = "pink")) + 
  xlab("") + 
  ylab("") + 
  ggtitle("A") +
  theme(plot.title = element_text(hjust = 0)) +
  theme(legend.position = "none") + 
  coord_flip()

# Summer temp
temp_plot_B <- ggplot(data = temp_clim, aes(x = site, y = Feb_mean_temp)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  ylim(8, 22) +
  geom_errorbar(aes(x = site, ymin = Feb_mean_temp - Feb_sd_temp, ymax = Feb_mean_temp + Feb_sd_temp   )) + 
  geom_point(data = temp_clim, aes(x = site, y = Feb_max_temp, colour = "blue")) + 
  geom_point(data = temp_clim, aes(x = site, y = Feb_min_temp, colour = "pink")) + 
  xlab("") + 
  ylab("") + 
  ggtitle("B") +
  theme(legend.position = "none") + 
  coord_flip()

# Winter temp
temp_plot_C <- ggplot(data = temp_clim, aes(x = site, y = Aug_mean_temp)) +
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  ylim(8, 22) + 
  geom_errorbar(aes(x = site, ymin = Aug_mean_temp - Aug_sd_temp, ymax = Aug_mean_temp + Aug_sd_temp   )) + 
  geom_point(data = temp_clim, aes(x = site, y = Aug_max_temp, colour = "blue")) + 
  geom_point(data = temp_clim, aes(x = site, y = Aug_min_temp, colour = "pink")) + 
  xlab("") + 
  ylab("") + 
  ggtitle("C") +
  theme(legend.position = "none") + 
  coord_flip()

```

```{r Temperature plot, echo=FALSE, fig.width = 7, fig.height = 9}

temp_plot_all <- grid.arrange(temp_plot_A, temp_plot_B,temp_plot_C, ncol = 1, nrow = 3 )

ggsave("temperature_plot.png", plot = temp_plot_all, path = "figures/", dpi = 300, width = 5, height = 5, limitsize = TRUE, scale = 2)

```

### Wave variables

```{r echo=FALSE}

## Data wrangling and preparation

# Wave data

# Load the loading functions
source("func/load.wave.R")

# Then load the data
fb_model <- load.model("data/wave_data/FB")
tb_model <- load.model("data/wave_data/TB")
he_model <- load.model("data/wave_data/HE")
sh_model <- load.model("data/wave_data/SH")

# Combine
wave_model <- rbind(fb_model, tb_model, he_model, sh_model)
rm(fb_model, tb_model)

# The site list key
wave_sites <- read_csv("metadata/sites_complete.csv") %>% 
  select(site, wave_7, wave_15)

# Calculate the necessary stats
# Monthly clims
wave_feb_aug <- wave_model %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp'
  group_by(site, date) %>% 
  summarise_all(funs(mean = mean, sd = sd), na.rm = T) %>%
  filter(date %in% c("Feb", "Aug")) %>% 
  ungroup()

# Annual clims
wave_ann <- wave_model %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp'
  group_by(site, date) %>% 
  summarise_all(funs(mean = mean, sd = sd), na.rm = T) %>%
  ungroup() %>% 
  select(-date) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean), na.rm = T) %>% 
  mutate(date = "Ann") %>% 
  select(site, date, everything())

# Subset directions out
dirs <- subset(wave_model, select = c(1, 2, 5, 6))

# Seasons
dirs_feb_aug <- dirs %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  group_by(site, date) %>% 
  summarise(dir_min = min(dir, na.rm = T),
            dir_max = max(dir, na.rm = T),
            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
            dir_median = median(dir, na.rm = T), 
            dirw_min = min(dirw, na.rm = T),
            dirw_max = max(dirw, na.rm = T),
            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
            dirw_median = median(dirw, na.rm = T)) %>%
  filter(date %in% c("Feb", "Aug")) %>% 
  ungroup()

# Annual
dirs_ann <- dirs %>% 
  mutate(date = lubridate::month(date, label = TRUE)) %>% 
  group_by(site, date) %>% 
  summarise(dir_min = min(dir, na.rm = T),
            dir_max = max(dir, na.rm = T),
            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
            dir_median = median(dir, na.rm = T), 
            dirw_min = min(dirw, na.rm = T),
            dirw_max = max(dirw, na.rm = T),
            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
            dirw_median = median(dirw, na.rm = T)) %>% 
  ungroup() %>% 
  select(-date) %>% 
  group_by(site) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  mutate(date = "Ann") %>% 
  select(site, date, everything())

# subset useless columns
dirs_feb_aug <- subset(dirs_feb_aug, select = c(1, 2, 6, 10))
dirs_ann <- subset(dirs_ann, select = c(1, 2, 6, 10))

# combine all direction metrics
dirs_clim <- rbind(dirs_feb_aug, dirs_ann) %>% 
  group_by(site) %>% 
  gather(key = statistic, value = value, 
         3:4) %>% 
  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
  select(-date, -statistic) %>% 
  spread(key = clim_stat, value = value) %>% 
  left_join(wave_sites, by = c("site" = "wave_7")) %>% 
  ungroup() %>% 
  select(-site, wave_15) %>%
  rename(site = site.y) %>% 
  select(site, everything()) %>% 
  na.omit()

# Combine
# NB: This only uses the wave data from the 7 metre contour
wave_clim <- rbind(wave_feb_aug, wave_ann) %>% 
  group_by(site) %>% 
  gather(key = statistic, value = value, 
         hs_mean:spw_sd) %>% 
  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
  select(-date, -statistic) %>% 
  spread(key = clim_stat, value = value) %>% 
  left_join(wave_sites, by = c("site" = "wave_7")) %>% 
  ungroup() %>% 
  select(-site, wave_15) %>%
  rename(site = site.y) %>% 
  select(site, everything()) %>% 
  na.omit()

# combine all wave metrics
wave_clim <- cbind(wave_clim, dirs_clim)

# remove all useless direction metrics from original wave_clim
wave_clim <- subset(wave_clim, select = -c(2,3,4,5,12,13,14,15,22,23,24,25,32,33,40))

```

```{r}

### Plots for wave variables
#Annual wave plots

wave_plot_A <- ggplot(data = wave_clim, aes(x = site, y = Ann_hs_mean)) +
  theme_bw() + 
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Ann_hs_mean - Ann_hs_sd, ymax = Ann_hs_mean + Ann_hs_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  ggtitle("") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  coord_flip()

wave_plot_B <- ggplot(data = wave_clim, aes(x = site, y = Ann_tp_mean)) +
  theme_bw() + 
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Ann_tp_mean - Ann_tp_sd, ymax = Ann_tp_mean + Ann_tp_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  ggtitle("") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  coord_flip()

#Summer wave plots
wave_plot_C <- ggplot(data = wave_clim, aes(x = site, y = Feb_hs_mean)) +
  theme_bw() + 
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Feb_hs_mean - Feb_hs_sd, ymax = Feb_hs_mean + Feb_hs_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  coord_flip()

wave_plot_D <- ggplot(data = wave_clim, aes(x = site, y = Feb_tp_mean)) +
  theme_bw() + 
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Feb_tp_mean - Feb_tp_sd, ymax = Feb_tp_mean + Feb_tp_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  coord_flip()

# Winter summary temp
wave_plot_E <- ggplot(data = wave_clim, aes(x = site, y = Aug_hs_mean)) +
  theme_bw() + 
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Aug_hs_mean - Aug_hs_sd, ymax = Aug_hs_mean + Aug_hs_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  coord_flip()

# Winter temperature range
wave_plot_F <- ggplot(data = wave_clim, aes(x = site, y = Aug_tp_mean)) +
  theme_bw() + 
  geom_point() + 
  scale_x_discrete(limits = positions_temp) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
  geom_errorbar(aes(x = site, ymin = Aug_tp_mean - Aug_tp_sd, ymax = Aug_tp_mean + Aug_tp_sd)) + 
  xlab("") + 
  ylab("") + 
  theme(legend.position = "none") + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  coord_flip()

```

```{r echo=FALSE, fig.height= 8, fig.width= 8}

wave_plot_all <- grid.arrange(wave_plot_A, wave_plot_B, wave_plot_C, wave_plot_D, wave_plot_E, wave_plot_F, ncol = 2, nrow = 3)

ggsave("wave_plot.png", plot = wave_plot_all, path = "figures/", dpi = 300, width = 6, height = 6, limitsize = TRUE, scale = 2)

```

### 2019-07-02 stats_chapter_2.R
commiting older updates on orginal script

@@ -222,8 +222,6 @@ dirs_feb_aug <- dirs %>%
   filter(date %in% c("Feb", "Aug")) %>% 
   ungroup()
 
-
-
 # Annual
 dirs_ann <- dirs %>% 
   mutate(date = lubridate::month(date, label = TRUE)) %>% 

### 2019-07-02 stats_chapter_2_alt.R

@@ -11,10 +11,13 @@ temp_plot_A <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
   geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) + 
   geom_point(data = temp_ann, aes(x = site, y = max_temp, colour = "blue")) + 
   geom_point(data = temp_ann, aes(x = site, y = min_temp, colour = "pink")) + 
-  xlab("") + 
+  xlab("Annual") + 
   ylab("") + 
+  ggtitle("Temperature (\u00B0C)") +
+  theme(plot.title = element_text(hjust = 0.5)) +
   theme(legend.position = "none") + 
   coord_flip() 
+
 # Annual temperature range
 temp_plot_B <- ggplot(data = temp_ann, aes(x = site, y = range_temp)) +
   theme_bw() + 
@@ -24,6 +27,8 @@ temp_plot_B <- ggplot(data = temp_ann, aes(x = site, y = range_temp)) +
   geom_point() + 
   xlab("") + 
   ylab("") + 
+  ggtitle("Temperature range (\u00B0C)") + 
+  theme(plot.title = element_text(hjust = 0.5)) + 
   coord_flip()
 
 # Summer summary temp
@@ -35,7 +40,7 @@ temp_plot_C <- ggplot(data = temp_clim, aes(x = site, y = Feb_mean_temp)) +
   geom_errorbar(aes(x = site, ymin = Feb_mean_temp - Feb_sd_temp, ymax = Feb_mean_temp + Feb_sd_temp   )) + 
   geom_point(data = temp_clim, aes(x = site, y = Feb_max_temp, colour = "blue")) + 
   geom_point(data = temp_clim, aes(x = site, y = Feb_min_temp, colour = "pink")) + 
-  xlab("Sites") + 
+  xlab("Summer") + 
   ylab("") + 
   theme(legend.position = "none") + 
   coord_flip() 
@@ -60,8 +65,8 @@ temp_plot_E <- ggplot(data = temp_clim, aes(x = site, y = Aug_mean_temp)) +
   geom_errorbar(aes(x = site, ymin = Aug_mean_temp - Aug_sd_temp, ymax = Aug_mean_temp + Aug_sd_temp   )) + 
   geom_point(data = temp_clim, aes(x = site, y = Aug_max_temp, colour = "blue")) + 
[…]

### 2019-07-02 doodle.R
working script added. this will be deleted later

@@ -1,0 +1,112 @@
+# Calculate the necessary stats
+# Monthly clims
+wave_feb_aug <- wave_model %>% 
+  mutate(date = lubridate::month(date, label = TRUE)) %>% 
+  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp' 
+  group_by(site, date) %>% 
+  summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T) %>%
+  select(-dir_mean, -dirw_mean, -dir_sd, dirw_sd, -spw_mean,-dirw_sd, -spw_sd, -dir_min, -dirw_min, - spw_min, -dir_max, -dirw_max, -spw_max ) %>% 
+  filter(date %in% c("Feb", "Aug")) %>% 
+  ungroup()
+
+# Annual clims
+wave_ann <- wave_model %>% 
+  mutate(date = lubridate::month(date, label = TRUE)) %>% 
+  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp'
+  group_by(site, date) %>% 
+  summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T) %>%
+  ungroup() %>% 
+  select(-date) %>% 
+  group_by(site) %>% 
+  summarise_all(funs(mean), na.rm = T) %>% 
+  mutate(date = "Ann") %>% 
+  select(site, date, everything())  %>%
+  select(-dir_mean, -dirw_mean, -dir_sd, dirw_sd, -spw_mean,-dirw_sd, -spw_sd, -dir_min, -dirw_min, - spw_min, -dir_max, -dirw_max, -spw_max )
+
+# Subset directions out
+dirs <- subset(wave_model, select = c(1, 2, 5, 6, 7))
+
+# Seasons
+dirs_feb_aug <- dirs %>% 
+  mutate(date = lubridate::month(date, label = TRUE)) %>% 
+  group_by(site, date) %>% 
+  summarise(dir_min = min(dir, na.rm = T),
+            dir_max = max(dir, na.rm = T),
+            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
+            dir_median = median(dir, na.rm = T), 
+            dirw_min = min(dirw, na.rm = T),
+            dirw_max = max(dirw, na.rm = T),
+            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
[…]

### 2019-07-02 morp_graphs_update.R

@@ -1,0 +1,111 @@
+library(tidyverse)
+library(ggpubr)
+library(magrittr)
+
+# Morphologies
+# L. pallida
+
+# load data
+lam <- read.csv("data/Laminaria_morphometrics.csv", sep = ",")
+
+# remove unnecessary columns
+lam <- subset(lam, select = -c(1,2,4))
+
+# alter dataframe for more friendly plotting
+lam_gather <- gather(lam, "Morphology", "Measurement", 2:10, na.rm = TRUE)
+
+# Replace morphology names
+lam_gather <- lam_gather %>%
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "digits", "Number of digits")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "laminaLength", "Lamina length (mm)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "laminaThickness", "Lamina thickness (mm)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "laminaWeight", "Lamina weight (g)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "stipeDiameter", "Stipe diameter (mm)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "stipeLength", "Stipe Length (mm)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "stipeMass", "Stipe mass (g)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "thallusMass", "Thallus mass (g)")) %>% 
+  mutate(Morphology = replace(Morphology, 
+                              Morphology == "totalLength", "Total length (mm)"))
+# Order site names by location west to east
+positionsLam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point", "Baboon Rock", 
+                  "Batsata Rock", "Buffels Bay","Bordjies North","Olifantsbos", "Kommetjie", "Sea Point")
+
+# Visualize: Specify the comparisons you want
[…]

### 2019-07-02 morp_graphs_update.R
deleted extra script

@@ -1,111 +1,0 @@
-library(tidyverse)
-library(ggpubr)
-library(magrittr)
-
-# Morphologies
-# L. pallida
-
-# load data
-lam <- read.csv("data/Laminaria_morphometrics.csv", sep = ",")
-
-# remove unnecessary columns
-lam <- subset(lam, select = -c(1,2,4))
-
-# alter dataframe for more friendly plotting
-lam_gather <- gather(lam, "Morphology", "Measurement", 2:10, na.rm = TRUE)
-
-# Replace morphology names
-lam_gather <- lam_gather %>%
-  mutate(Morphology = replace(Morphology, 
-                              Morphology == "digits", "Number of digits")) %>% 
-  mutate(Morphology = replace(Morphology, 
-                              Morphology == "laminaLength", "Lamina length (mm)")) %>% 
-  mutate(Morphology = replace(Morphology, 
-                              Morphology == "laminaThickness", "Lamina thickness (mm)")) %>% 
-  mutate(Morphology = replace(Morphology, 
-                              Morphology == "laminaWeight", "Lamina weight (g)")) %>% 
-  mutate(Morphology = replace(Morphology, 
-                              Morphology == "stipeDiameter", "Stipe diameter (mm)")) %>% 
-  mutate(Morphology = replace(Morphology, 
-                              Morphology == "stipeLength", "Stipe Length (mm)")) %>% 
-  mutate(Morphology = replace(Morphology, 
-                              Morphology == "stipeMass", "Stipe mass (g)")) %>% 
-  mutate(Morphology = replace(Morphology, 
-                              Morphology == "thallusMass", "Thallus mass (g)")) %>% 
-  mutate(Morphology = replace(Morphology, 
-                              Morphology == "totalLength", "Total length (mm)"))
-# Order site names by location west to east
-positionsLam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point", "Baboon Rock", 
-                  "Batsata Rock", "Buffels Bay","Bordjies North","Olifantsbos", "Kommetjie", "Sea Point")
-
-# Visualize: Specify the comparisons you want
[…]

### 2019-07-02 stats_chapter_2_alt.R

@@ -1,311 +1,0 @@
-### Plots for temperature variables 
-
-positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","De Hoop", "St Helena Bay")
-
-# Annual summary temp
-temp_plot_A <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
-  theme_bw() + 
-  geom_point() + 
-  scale_x_discrete(limits = positions_temp) + 
-  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
-  geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) + 
-  geom_point(data = temp_ann, aes(x = site, y = max_temp, colour = "blue")) + 
-  geom_point(data = temp_ann, aes(x = site, y = min_temp, colour = "pink")) + 
-  xlab("Annual") + 
-  ylab("") + 
-  ggtitle("Temperature (\u00B0C)") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme(legend.position = "none") + 
-  coord_flip() 
-
-# Annual temperature range
-temp_plot_B <- ggplot(data = temp_ann, aes(x = site, y = range_temp)) +
-  theme_bw() + 
-  geom_point() + 
-  scale_x_discrete(limits = positions_temp) + 
-  theme(axis.text.x = element_text(angle = 90, hjust = , vjust = 0.5)) +
-  geom_point() + 
-  xlab("") + 
-  ylab("") + 
-  ggtitle("Temperature range (\u00B0C)") + 
-  theme(plot.title = element_text(hjust = 0.5)) + 
-  coord_flip()
-
-# Summer summary temp
-temp_plot_C <- ggplot(data = temp_clim, aes(x = site, y = Feb_mean_temp)) +
-  theme_bw() + 
-  geom_point() + 
-  scale_x_discrete(limits = positions_temp) + 
-  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
[…]

### 2019-07-02 paper_morph_new.Rmd
figure captions added and small updates to temperature and wave plots. kelp schematic added.

@@
 output:
-  word_document: default
+  word_document: 
+    fig_caption: yes
+tables: yes
+fontsize: 10pt
+papersize: A4
+geometry: margin=1in  
 ---
@@
 
-Sites were chosen to represent an array of environmental gradients, as indicated in Fig. 1. St. Helena Bay and Betty?s Bay constituted the north western and south eastern boundary sites, respectively.
-
+Sites were chosen to represent an array of environmental gradients, as indicated in Fig.
+1.
+St.
+Helena Bay and Betty's Bay constituted the north western and south eastern boundary sites, respectively.
 These sites are roughly 300 km apart, and lie within separate marine provinces[3] and cover a cline in wave exposure and temperature.
@@
 The region is dominated by kelp communities that persist in contrasting abiotic environments.
-The west coast region has been termed as ?cool temperate?
-region, which is defined as a region where mean monthly temperatures are always above 10?C and always below 15?C (Smit et al.
+The west coast region has been termed as a cool temperate region, which is defined as a region where mean monthly temperatures are always above 10°C and always below 15°C (Smit et al.
 2013).
@@
 2013).
-The Agulhas Marine Province is characterised by a wide temperature range of up to 7?C difference between mean monthly temperatures between summer and winter (Smit et al.
+The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter (Smit et al.
 2013).
@@
 This is however absent in winter, where strong northerly winds are prevalent from St.
-Helena Bay to Betty?s Bay (Field et al.
+Helena Bay to Betty's Bay (Field et al.
 1980; Jury et al.
@@
 It is due to the near consistent south-westerly swell and the complex orography around the peninsula that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.
+
+```{r echo=FALSE, fig.cap='Schematic of *E. maxima* on the left and *L. pallida* on the right.REF?'}
+
+library(png)
+library(grid)
+
+img <- readPNG("figures/kelp_diagram.png")
+grid.raster(img)
+
+```
 
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r Temperature plot, echo=FALSE, fig.width = 7, fig.height = 9}
+```{r Temperature plot, echo=FALSE, fig.width = 7, fig.height = 9, fig.cap= 'Temperature variables at the collection sites around the Cape Peninsula.
+Temperature variables include minimum represented by blue dots, maximum represented by red dots, mean represented by black dots and standard deviation by error bars.
+Temperature variables are further categorised by season with "A" representing annual temperature variables, "B" summer temperature variables and "C" winter temperature variables.'}
 
@@
 wave_plot_A <- ggplot(data = wave_clim, aes(x = site, y = Ann_hs_mean)) +
-  theme_bw() + 
   geom_point() + 
@@
   theme(legend.position = "none") + 
-  ggtitle("") + 
-  theme(plot.title = element_text(hjust = 0.5)) + 
+  ggtitle("A") + 
+  theme(plot.title = element_text(hjust = 0)) + 
   coord_flip()
@@
 wave_plot_B <- ggplot(data = wave_clim, aes(x = site, y = Ann_tp_mean)) +
-  theme_bw() + 
   geom_point() + 
@@
   theme(legend.position = "none") + 
-  ggtitle("") + 
-  theme(plot.title = element_text(hjust = 0.5)) + 
+  ggtitle("B") + 
+  theme(plot.title = element_text(hjust = 0)) + 
   theme(axis.title.y=element_blank(),
@@
 wave_plot_C <- ggplot(data = wave_clim, aes(x = site, y = Feb_hs_mean)) +
-  theme_bw() + 
   geom_point() + 
@@
   theme(legend.position = "none") + 
+  ggtitle("C") + 
   coord_flip()
@@
 wave_plot_D <- ggplot(data = wave_clim, aes(x = site, y = Feb_tp_mean)) +
-  theme_bw() + 
   geom_point() + 
@@
   theme(legend.position = "none") + 
+    ggtitle("D") + 
   theme(axis.title.y=element_blank(),
@@
 wave_plot_E <- ggplot(data = wave_clim, aes(x = site, y = Aug_hs_mean)) +
-  theme_bw() + 
   geom_point() + 
@@
   ylab("") + 
+  ggtitle("E") + 
   theme(legend.position = "none") + 
@@
 wave_plot_F <- ggplot(data = wave_clim, aes(x = site, y = Aug_tp_mean)) +
-  theme_bw() + 
   geom_point() + 
@@
   theme(legend.position = "none") + 
+  ggtitle("F") + 
   theme(axis.title.y=element_blank(),
@@
 
-```{r echo=FALSE, fig.height= 8, fig.width= 8}
+```{r echo=FALSE, fig.height= 8, fig.width= 8, fig.cap='Wave variables at each collection site around the Western Cape coast.
+Mean variables are represented by dots and standard deviation by whiskers.
+Plot "A" represents annual significant wave height, plot "B" annual significant wave period, plot "C" winter significant wave height, plot "D" winter significant wave period, "E" summer significant wave height and plot "F" winter significant wave period.'}
 

### 2019-07-07 paper_morph_new.Rmd
added section into the introduction and altered some writing.

@@
 Wave exposure may be greater in shallower environments, or it may be reduced through damping effects of kelp further offshore and natural barriers such as rock outcrops; also, seawater temperature may be higher in shallow water environments due to higher solar irradiance.
-Wind has also been shown to play a role in influencing stipe length.
-A study by (Denny et al.
-1997) used a combination of field and laboratory tests to model the drag properties on the canopy kelp *Nereocystis luetkeana* (K.
-Mertens) Postels & Ruprecht.
-The results from the model suggest that wind may play a role in the extension capabilities of the stipe under certain wave/wind conditions.
-Furthermore, wind generates surface-gravity waves under certain conditions, and may influence the overall wave climate at a specific location in-time (Holthuijsen 2010).
+The kelp species
 
-Comparatively little research has been done on other species of kelp, such as *Laminaria pallida* and *Ecklonia maxima* (Osbeck) Papenfuss  which are important habitat forming seaweeds that exist around the coast of South Africa.
+Other species of kelp, such as *Laminaria pallida* and *Ecklonia maxima* (Osbeck) Papenfuss which are important habitat forming seaweeds that exist around the coast of South Africa, offer a unique opportunity to investigate the drivers of macroalgal morphological characteristics between deep and shallow water environments.
 Although both these species exist in the subtidal, *L.
@@
 Therefore, *E.
-maxima* is exposed to variations in waved, swell and temperature, while *L.
+maxima* is exposed to variations in wave, swell and temperature, while *L.
 pallida* is exposed to variations in waves and temperature.
@@
 
+Measures of wave exposure in ecological studies often incorporate integrative measures of hydrodynamic conditions at a particular site which are based on cartographical models of wave exposure.
+Cartographical models of wave exposure are regarded as 'fetch-based-models' which measure the length of open water associated with a particular site in a straight line, and are regarded as simple measures of wave exposure.
+Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitave, reproducible approach for measuring wave exposure.
+Currently, no ecological studies investigating macroalgal communities have incorporated 3D spectral numerical modelling.
 The aim of this study is, therefore, to understand how temperature, wave exposure and wind can influence the morphology in two species of kelps around South Africa.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
+```{r Temperature data prep, echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r}
+```{r Temperature variable processing}
 
@@
 
-```{r Temperature plot, echo=FALSE, fig.width = 7, fig.height = 9, fig.cap= 'Temperature variables at the collection sites around the Cape Peninsula.
-Temperature variables include minimum represented by blue dots, maximum represented by red dots, mean represented by black dots and standard deviation by error bars.
-Temperature variables are further categorised by season with "A" representing annual temperature variables, "B" summer temperature variables and "C" winter temperature variables.'}
+```{r Temperature variables plot, echo=FALSE, fig.cap='Temperature variables at the collection sites around the Cape Peninsula.
+Temperature variables include minimum represented by blue dots, fig.height=9, fig.width=7}
 
@@
 
-```{r echo=FALSE}
+```{r Wave data prep, echo=FALSE}
 
@@
 
-```{r}
+```{r Wave variable processing, echo=FALSE}
 
@@
 
-```{r echo=FALSE, fig.height= 8, fig.width= 8, fig.cap='Wave variables at each collection site around the Western Cape coast.
+```{r Wave variable plot, echo=FALSE, fig.cap='Wave variables at each collection site around the Western Cape coast.
 Mean variables are represented by dots and standard deviation by whiskers.
-Plot "A" represents annual significant wave height, plot "B" annual significant wave period, plot "C" winter significant wave height, plot "D" winter significant wave period, "E" summer significant wave height and plot "F" winter significant wave period.'}
+Plot "A" represents annual significant wave height, fig.height=8, fig.width=8}
 

### 2019-07-07 paper_morph_new.Rmd
pdf version added, word version updated and general house cleaning.

@@
 author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
+geometry: margin=1in
 output:
+  pdf_document: default
   word_document:
     fig_caption: yes
+papersize: A4
 tables: yes
 fontsize: 10pt
-papersize: A4
-geometry: margin=1in  
 ---
@@
 Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility (Denny & Gaylord 2002).
-For example, a study by (Wernberg & Thomsen 2005)) examined the consistency of wave exposure as a driver of *Ecklonia radiata* (C.
+For example, a study by Wernberg & Thomsen (2005) examined the consistency of wave exposure as a driver of *Ecklonia radiata* (C.
 Agardh) J.
@@
 Temperature has also been shown to play a role in driving kelp morphology.
-For example, a study by (Serisawa et al.
-2002) compared the morphology of Ecklonia cava Kjellman growing in warm temperate and cool temperate morphologies.
+For example, a study by Serisawa et al.
+(2002) compared the morphology of Ecklonia cava Kjellman growing in warm temperate and cool temperate morphologies.
 The results showed that wrinkles in the blade seem to be a characteristic of warm temperate regions.
@@
 The morphological adapatibility of kelps is driven by a combination of environmental factors that in turn do not act independently of one another.
-A study by (Wernberg et al.
-2003) investigated the morphology of *E.
+A study by Wernberg et al.
+(2003) investigated the morphology of *E.
 radiata* in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australasian coast.
@@
 Wave exposure may be greater in shallower environments, or it may be reduced through damping effects of kelp further offshore and natural barriers such as rock outcrops; also, seawater temperature may be higher in shallow water environments due to higher solar irradiance.
-The kelp species
 
@@
 pallida* at different depths and wave exposure and  showed that depth had a greater effect than wave exposure when considering all the morphological characteristics; when considering individual characteristics, however, wave exposure had the most significant effect on blade thickness.
-Another study by (Rothman et al.
-2017) investigated the changes in morphology in shallow populations of *L.
+Another study by Rothman et al.
+(2017) investigated the changes in morphology in shallow populations of *L.
 pallida* and *E.
@@
 They suggested that turbidity was the environmental driver responsible for this change.
-Research using advanced numerical tools is lacking.
-Previous research investigating drivers of kelp morphological characteristics used qualitative estimates of wave exposure and did not consider other morphological characteristics.
-Furthermore, only shallow populations have been considered in previous research which ignores the effects of wave damping.
-No work currently exists where other morphological characters have been considered, quantitative estimates of environmental drivers have been calculated, and the addition of wind as a possible driver of morphological characteristics has been considered.
 
@@
 Currently, no ecological studies investigating macroalgal communities have incorporated 3D spectral numerical modelling.
+Previous research investigating drivers of kelp morphological characteristics used qualitative estimates of wave exposure and did not consider other morphological characteristics.
+Furthermore, only shallow populations have been considered in previous research which ignores the effects of wave damping.
+No work currently exists where other morphological characters have been considered, quantitative estimates of environmental drivers have been calculated, and the addition of wind as a possible driver of morphological characteristics has been considered.
 The aim of this study is, therefore, to understand how temperature, wave exposure and wind can influence the morphology in two species of kelps around South Africa.
@@
 
-```{r echo=FALSE, fig.cap='Schematic of *E. maxima* on the left and *L. pallida* on the right.REF?'}
+```{r echo=FALSE, fig.cap='Schematic of *E. maxima* on the left and *L. pallida* on the right. (Dyer 2018)'}
 
@@
 
-```{r Temperature variable processing}
+```{r Temperature variable processing, message=FALSE, warning=TRUE, include=FALSE}
 
@@
 
-```{r Temperature variables plot, echo=FALSE, fig.cap='Temperature variables at the collection sites around the Cape Peninsula.
-Temperature variables include minimum represented by blue dots, fig.height=9, fig.width=7}
+```{r Temperature variables plot, echo=FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula.
+Temperature variables include minimum represented by blue dots.", fig.height=9, fig.width=7}
 
@@
 
-```{r Wave data prep, echo=FALSE}
+```{r Wave data prep, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-```{r Wave variable processing, echo=FALSE}
+```{r Wave variable processing, include=FALSE}
 
@@
 Mean variables are represented by dots and standard deviation by whiskers.
-Plot "A" represents annual significant wave height, fig.height=8, fig.width=8}
+Plot "A" represents annual significant wave height.', fig.height=8, fig.width=8}
 

### 2019-07-11 paper_morph_new.Rmd
minor edits to rmd script. Data house cleaning

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(include = FALSE)
 ```
@@
 
-```{r Temperature data prep, echo=FALSE, message=FALSE, warning=FALSE}
+```{r Temperature data prep, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-```{r Temperature variables plot, echo=FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula.
+```{r Temperature variables plot, echo=FALSE, include = FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula.
 Temperature variables include minimum represented by blue dots.", fig.height=9, fig.width=7}
@@
 
-```{r Wave variable processing, include=FALSE}
+```{r Wave variable processing, warning=FALSE, include=FALSE, echo = FALSE}
 
@@
 Mean variables are represented by dots and standard deviation by whiskers.
-Plot "A" represents annual significant wave height.', fig.height=8, fig.width=8}
+Plot "A" represents annual significant wave height.', fig.height=8, fig.width=7}
 

### 2019-07-14 paper_morph_new.Rmd
script for deep ecklonia and laminaria added. Final outputs need to be updated. RDA to follow

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(include = FALSE)
-```
-
-## Introduction
+# Introduction
 
@@
 
-## Methods
+# Methods
 ### Site selection
@@
 
-## Results
-### Temperature variables
+# Results
+## Temperature variables
 
@@
 # Complete lon/lats
-kelp_sites <- read_csv("metadata/sites_complete.csv")
+kelp_sites <- read.csv("metadata/site_meta_complete.csv", sep = ";")
 
 # Ecklonia
-eck_data <- read_csv("data/Ecklonia_morphometrics.csv") %>%
-  mutate(site = replace(site, site == "Buffels Bay", "Buffelsbaai")) %>%
-  mutate(site = replace(site, site == "St. Helena Bay", "St Helena Bay")) %>%
-  # Not using Rocky Bank in this study
-  filter(site != "Rocky Bank")
+eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
+  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
+  filter(site!= "De Hoop")
 
 # Laminaria
-lam_data <- read_csv("data/Laminaria_morphometrics.csv") %>%
+lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>%
   mutate(site = replace(site, site == "Buffels Bay", "Buffelsbaai")) %>%
@@
 # The SACTN monthly data
-load("data/SACTNmonthly_v4.1.Rdata")
+load("data/temp_data/SACTNmonthly_v4.1.Rdata")
 
 # The missing site temperatures
-load("data/interp_sites.Rdata")
+load("data/temp_data/interp_sites.Rdata")
 
@@
 
-```{r Temperature variable processing, message=FALSE, warning=TRUE, include=FALSE}
+```{r Temperature variable processing, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-```{r Temperature variables plot, echo=FALSE, include = FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula.
+```{r Temperature variables plot, echo=FALSE, warning=FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula.
 Temperature variables include minimum represented by blue dots.", fig.height=9, fig.width=7}
@@
 
-### Wave variables
+## Wave variables
 
-```{r Wave data prep, message=FALSE, warning=FALSE, include=FALSE}
+```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 # The site list key
-wave_sites <- read_csv("metadata/sites_complete.csv") %>% 
+wave_sites <- read_csv2("metadata/site_meta_complete.csv") %>% 
   select(site, wave_7, wave_15)
@@
 
-```{r Wave variable processing, warning=FALSE, include=FALSE, echo = FALSE}
+```{r wave variable processing, warning=FALSE, include=FALSE, echo = FALSE}
 
@@
 
-```{r Wave variable plot, echo=FALSE, fig.cap='Wave variables at each collection site around the Western Cape coast.
+```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE, fig.cap='Wave variables at each collection site around the Western Cape coast.
 Mean variables are represented by dots and standard deviation by whiskers.
@@
 ```
+
+## Morphology
+### *L. pallida*
+```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE}
+
+lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>% 
+   filter(site!= "Rocky Bank")
+
+# remove unnecessary columns
+lam_df <- subset(lam_data, select = -c(1,2,6))
+
+pos_lam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point", "Baboon Rock", 
+                  "Batsata Rock", "Buffels Bay"," Buffels Bay South","Bordjies North","Olifantsbos", "Kommetjie", "Sea Point")
+pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Sheltered", "Moderatley exposed", "Exposed", "Very exposed")
+
+lamina_weight_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure_lam) + 
+         geom_boxplot(outlier.shape = NA) 
+lamina_length_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure_lam) + 
+         geom_boxplot(outlier.shape = NA) 
+lamina_thickness_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure_lam) + 
+         geom_boxplot(outlier.shape = NA) 
+stipe_mass_lam_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure_lam) + 
+         geom_boxplot(outlier.shape = NA) 
+stipe_length_lam_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure_lam) + 
+         geom_boxplot(outlier.shape = NA) 
+stipe_diameter_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure_lam) + 
+         geom_boxplot(outlier.shape = NA) 
+digits_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure_lam) + 
+         geom_boxplot(outlier.shape = NA) 
+thallus_mass_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure_lam) + 
+         geom_boxplot(outlier.shape = NA) 
+total_length_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure_lam) + 
+         geom_boxplot(outlier.shape = NA) 
+```
+
+```{r laminaria plots, echo=FALSE, message=FALSE, warning=FALSE}
+grid.arrange(lamina_weight_plot,lamina_length_plot,lamina_thickness_plot,stipe_mass_lam_plot,stipe_length_lam_plot,stipe_diameter_plot, digits_plot, thallus_mass_plot,total_length_plot,total_length_plot, ncol = 3, nrow = 4)
+
+```
+
+### *E.maxima*
+
+```{r ecklonia  processing, message=FALSE, warning=FALSE, include=FALSE}
+
+library(ggplot2)
+
+## Preparing data
+# Ecklonia
+eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
+ filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
+ filter(site!= "De Hoop")  
+# group_by(wave_exposure, frond_mass) %>% 
+# summarise(mean = mean(frond_mass:total_length, na.rm = TRUE)) %>% 
+# ungroup()
+
+eck_deep <- eck_data %>% 
+  filter(population == "Deep")
+
+# custom positions for wave exposure categories and sites
+pos_exposure <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Sheltered", "Moderatley exposed", "Exposed", "Very exposed")
+pos_eck<- c("Betty's Bay", "Kalk Bay", "A-Frame",  "Miller's Point", "Baboon Rock", "Batsata Rock", "Buffelsbaai", "Scarborough", "Soetwater","Kommetjie","Hout Bay", "Oudekraal", "St Helena Bay")
+
+frond_mass_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_mass, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure) + 
+         geom_boxplot(outlier.shape = NA) 
+primary_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure) + 
+         geom_boxplot(outlier.shape = NA) 
+primary_width_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure) + 
+         geom_boxplot(outlier.shape = NA) 
+frond_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure) + 
+         geom_boxplot(outlier.shape = NA) 
+stipe_mass_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure) + 
+         geom_boxplot(outlier.shape = NA) 
+stipe_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure) + 
+         geom_boxplot(outlier.shape = NA) 
+stipe_circ_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure) + 
+         geom_boxplot(outlier.shape = NA) 
+tufts_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure) + 
+         geom_boxplot(outlier.shape = NA) 
+epi_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
+         scale_x_discrete(limits = pos_exposure) + 
+         geom_boxplot(outlier.shape = NA) 
+total_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure) + 
+  geom_boxplot(outlier.shape = NA) +
+  ylim(0, 2000)
+
+```
+
+```{r ecklonia plots, echo=FALSE, message=FALSE, warning=FALSE}
+
+grid.arrange(frond_mass_plot,primary_length_plot,primary_width_plot,frond_length_plot,stipe_mass_plot,stipe_length_plot,stipe_circ_plot, tufts_plot, epi_length_plot,total_length_plot, ncol = 3, nrow = 4)
+
+```

### 2019-07-15 paper_morph_new.Rmd
morphology graphs for laminaria

@@
 ---
-title: Temperature, wave exposure and wind as drivers of morphology in two kelp species
+title: Temperature and wave exposure as drivers of morphology in two kelp species
   along the coast of South Africa
@@
 output:
-  pdf_document: default
   word_document:
     fig_caption: yes
+  pdf_document: default
 papersize: A4
@@
 ## Temperature variables
-
-```{r Temperature data prep, message=FALSE, warning=FALSE, include=FALSE}
-
-## Data wrangling and preparation
-
+```{r software packages, message=TRUE, warning=TRUE, include=FALSE, echo=FALSE}
 library(tidyverse)
@@
 library(gridExtra)
+```
 
+```{r Temperature data prep, message=FALSE, warning=FALSE, include=FALSE}
+## Data wrangling and preparation
 ## Preparing data
@@
 ### *L. pallida*
+
 ```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE}
-
 lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>% 
@@
 
-pos_lam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point", "Baboon Rock", 
-                  "Batsata Rock", "Buffels Bay"," Buffels Bay South","Bordjies North","Olifantsbos", "Kommetjie", "Sea Point")
-pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Sheltered", "Moderatley exposed", "Exposed", "Very exposed")
+pos_lam_temp <- c("Cool temperate", "Warm temperate")
 
-lamina_weight_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
+LT1 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_lam_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Lamina weight") +
+  scale_fill_discrete(name="Temperature regime") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+LT2 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_lam_temp)  + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Lamina length") +
+  scale_fill_discrete(name="Temperature regime") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+LT3 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_lam_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Lamina thickness") +
+  scale_fill_discrete(name="Temperature regime") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+LT4 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_lam_temp)  + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Stipe mass (g)") +
+  scale_fill_discrete(name="Temperature regime") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+LT5 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_lam_temp)  + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Stipe length (cm)") +
+  scale_fill_discrete(name="Temperature regime") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+LT6 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_lam_temp)  + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Stipe diameter (cm)") +
+  scale_fill_discrete(name="Temperature regime") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+LT7 <- ggplot(data = lam_df, aes(x = temp_regime, y = digits, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_lam_temp)  + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("No. of digits") +
+  scale_fill_discrete(name="Temperature regime") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+LT8 <- ggplot(data = lam_df, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_lam_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Thallus mass (g)") +
+  scale_fill_discrete(name="Temperature regime") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+LT9 <- ggplot(data = lam_df, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_lam_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Total length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+
+lam_df$wave_exposure <- factor(lam_df$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))
+pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed")
+
+LW1 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam) + 
-         geom_boxplot(outlier.shape = NA) 
-lamina_length_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  ylab("Lamina weight (g)") + 
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))
+
+LW2 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
-         geom_boxplot(outlier.shape = NA) 
-lamina_thickness_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  ylab("Lamina length (cm)") + 
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))
+LW3 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam) + 
-         geom_boxplot(outlier.shape = NA) 
-stipe_mass_lam_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  ylab("Lamina thickness") + 
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
+LW4 <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
-         geom_boxplot(outlier.shape = NA) 
-stipe_length_lam_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  ylab("Stipe mass (g)") + 
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
+LW5 <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
-         geom_boxplot(outlier.shape = NA) 
-stipe_diameter_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  ylab("Stipe length (cm)") + 
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
+LW6 <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
-         geom_boxplot(outlier.shape = NA) 
-digits_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  ylab("Stipe diameter (cm)") + 
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
+LW7 <- ggplot(data = lam_df, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
-         geom_boxplot(outlier.shape = NA) 
-thallus_mass_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  ylab("No. of digits") + 
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
+LW8 <- ggplot(data = lam_df, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam) + 
-         geom_boxplot(outlier.shape = NA) 
-total_length_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  ylab("Thallus mass (kg)") + 
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
+LW9 <- ggplot(data = lam_df, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam) + 
-         geom_boxplot(outlier.shape = NA) 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  ylab("Total length (cm)") + 
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
 ```
 
-```{r laminaria plots, echo=FALSE, message=FALSE, warning=FALSE}
-grid.arrange(lamina_weight_plot,lamina_length_plot,lamina_thickness_plot,stipe_mass_lam_plot,stipe_length_lam_plot,stipe_diameter_plot, digits_plot, thallus_mass_plot,total_length_plot,total_length_plot, ncol = 3, nrow = 4)
+```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE}
+ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom", label = "p.format", ncol = 3, nrow = 3)
+```
 
+```{r laminaria wave plots, echo=FALSE, message=FALSE, warning=FALSE}
+ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
 ```

### 2019-07-16 paper_morph_new.Rmd
update ecklonia deep wave plots

@@
 
-pos_lam_temp <- c("Cool temperate", "Warm temperate")
+pos_temp <- c("Cool temperate", "Warm temperate")
 
 LT1 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
-  scale_x_discrete(limits = pos_lam_temp) + 
+  scale_x_discrete(limits = pos_temp) + 
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
@@
 LT2 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
-  scale_x_discrete(limits = pos_lam_temp)  + 
+  scale_x_discrete(limits = pos_temp)  + 
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
@@
 LT3 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
-  scale_x_discrete(limits = pos_lam_temp) + 
+  scale_x_discrete(limits = pos_temp) + 
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
@@
 LT4 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
-  scale_x_discrete(limits = pos_lam_temp)  + 
+  scale_x_discrete(limits = pos_temp)  + 
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
@@
 LT5 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
-  scale_x_discrete(limits = pos_lam_temp)  + 
+  scale_x_discrete(limits = pos_temp)  + 
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
@@
 LT6 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
-  scale_x_discrete(limits = pos_lam_temp)  + 
+  scale_x_discrete(limits = pos_temp)  + 
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
@@
 LT7 <- ggplot(data = lam_df, aes(x = temp_regime, y = digits, fill = temp_regime)) +
-  scale_x_discrete(limits = pos_lam_temp)  + 
+  scale_x_discrete(limits = pos_temp)  + 
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
@@
 LT8 <- ggplot(data = lam_df, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
-  scale_x_discrete(limits = pos_lam_temp) + 
+  scale_x_discrete(limits = pos_temp) + 
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
@@
 LT9 <- ggplot(data = lam_df, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
-  scale_x_discrete(limits = pos_lam_temp) + 
+  scale_x_discrete(limits = pos_temp) + 
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
@@
 ```{r ecklonia  processing, message=FALSE, warning=FALSE, include=FALSE}
-
 library(ggplot2)
@@
   filter(population == "Deep")
+eck_shall <- eck_data %>% 
+  filter(population == "Shallow")
+
+eck_deep$wave_exposure <- factor(eck_deep$wave_exposure, levels=c("Fully sheltered", "Extremely sheltered", "Very sheltered","Moderately exposed", "Exposed", "Very exposed"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered","Moderately exposed", "Exposed", "Very exposed"))
 
 # custom positions for wave exposure categories and sites
-pos_exposure <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Sheltered", "Moderatley exposed", "Exposed", "Very exposed")
-pos_eck<- c("Betty's Bay", "Kalk Bay", "A-Frame",  "Miller's Point", "Baboon Rock", "Batsata Rock", "Buffelsbaai", "Scarborough", "Soetwater","Kommetjie","Hout Bay", "Oudekraal", "St Helena Bay")
+pos_exposure <- c("Fully sheltered", "Extremely sheltered", "Very sheltered","Moderately exposed", "Exposed", "Very exposed")
 
@@
          scale_x_discrete(limits = pos_exposure) + 
-         geom_boxplot(outlier.shape = NA) 
+         geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Frond mass (g)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
 primary_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
-         geom_boxplot(outlier.shape = NA) 
+         geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Primary length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
 primary_width_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
-         geom_boxplot(outlier.shape = NA) 
+         geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Primary width (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
 frond_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
-         geom_boxplot(outlier.shape = NA) 
+         geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Frond length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
 stipe_mass_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
-         geom_boxplot(outlier.shape = NA) 
+         geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Stipe mass (g)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
 stipe_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
-         geom_boxplot(outlier.shape = NA) 
+         geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Stipe length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
 stipe_circ_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
-         geom_boxplot(outlier.shape = NA) 
+         geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Stipe circumference (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
 tufts_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
-         geom_boxplot(outlier.shape = NA) 
+         geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("No of Tufts") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
 epi_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
-         geom_boxplot(outlier.shape = NA) 
+         geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Epi length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
 total_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure) + 
-  geom_boxplot(outlier.shape = NA) +
-  ylim(0, 2000)
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("Total length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))
 
@@
 
-grid.arrange(frond_mass_plot,primary_length_plot,primary_width_plot,frond_length_plot,stipe_mass_plot,stipe_length_plot,stipe_circ_plot, tufts_plot, epi_length_plot,total_length_plot, ncol = 3, nrow = 4)
+ggarrange(frond_mass_plot,primary_length_plot,primary_width_plot,frond_length_plot,stipe_mass_plot,stipe_length_plot,stipe_circ_plot, tufts_plot, epi_length_plot,total_length_plot, labels = "AUTO", common.legend = TRUE, legend = "bottom")
 

### 2019-07-16 paper_morph_new.Rmd
shallow morphology and significance tests added

@@
 output:
+  pdf_document: 
+    fig_crop: no
   word_document:
     fig_caption: yes
-  pdf_document: default
 papersize: A4
@@
 ## Temperature variables
-```{r software packages, message=TRUE, warning=TRUE, include=FALSE, echo=FALSE}
+
+```{r software packages, message=TRUE, warning=FALSE, include=FALSE, echo=FALSE}
 library(tidyverse)
@@
 ```{r Temperature variables plot, echo=FALSE, warning=FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula.
-Temperature variables include minimum represented by blue dots.", fig.height=9, fig.width=7}
+Temperature variables include minimum represented by blue dots.", fig.height=8, fig.width=8}
 
@@
 
-```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE}
+```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
 
@@
 
-```{r wave variable processing, warning=FALSE, include=FALSE, echo = FALSE}
+```{r wave variable processing, warning=FALSE, include=FALSE, echo = FALSE, message=FALSE}
 
@@
 ```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE, fig.cap='Wave variables at each collection site around the Western Cape coast.
-Mean variables are represented by dots and standard deviation by whiskers.
-Plot "A" represents annual significant wave height.', fig.height=8, fig.width=7}
+Mean variables are represented by dots and standard deviation by whiskers.', fig.height=8, fig.width=8}
 
@@
 
-```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE}
+```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
 lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>% 
@@
 
-pos_temp <- c("Cool temperate", "Warm temperate")
-
+pos_temp <- c("Cool temperate", "Warm temperate")#custom temperature categories position on x-axis
+#temperature and lamanaria  
 LT1 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means() + 
+  stat_compare_means(method = "wilcox.test") + 
   xlab("") + 
@@
 
+#wave exposure and lamanaria
 lam_df$wave_exposure <- factor(lam_df$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))
@@
   scale_x_discrete(limits = pos_exposure_lam) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
+  stat_compare_means(method = "kruskal.test") + 
   theme(axis.title.x=element_blank(),
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-
 LW2 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
+  stat_compare_means(method = "kruskal.test") + 
   theme(axis.title.x=element_blank(),
@@
   scale_x_discrete(limits = pos_exposure_lam)  + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
+  stat_compare_means(method = "kruskal.test") + 
   theme(axis.title.x=element_blank(),
@@
 
-```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE}
+```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=9, fig.width=9, fig.cap='Boxplot of morphological characteristics for *L. pallida* categorised by temperature regime.'}
 ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom", label = "p.format", ncol = 3, nrow = 3)
@@
 
-```{r laminaria wave plots, echo=FALSE, message=FALSE, warning=FALSE}
+```{r laminaria wave plots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=9, fig.width=9, fig.cap ='Boxplot of morphological characteristics for *L. pallida* categorised by wave exposure.'}
 ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
@@
 
-```{r ecklonia  processing, message=FALSE, warning=FALSE, include=FALSE}
+```{r deep ecklonia  processing, message=FALSE, warning=FALSE, include=FALSE}
 library(ggplot2)
-
 ## Preparing data
@@
 # ungroup()
-
 eck_deep <- eck_data %>% 
@@
   filter(population == "Shallow")
+pos_temp <- c("Cool temperate", "Warm temperate") #custom temperature categories position on x-axis
+#temperature plots made seperately for each morphological characteristic
+ET1 <- ggplot(data = eck_deep, aes(x = temp_regime, y = frond_mass, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Frond mass (g)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+ET2 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Primary length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+ET3 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Primary width (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+ET4 <- ggplot(data = eck_deep, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Frond length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+ET5 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Stipe mass (g)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+ET6 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Stipe length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+ET7 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Stipe circ (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+ET8 <- ggplot(data = eck_deep, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("No. of tufts") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+ET9 <- ggplot(data = eck_deep, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Epi length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+ET10 <- ggplot(data = eck_deep, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Total length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
 
@@
 pos_exposure <- c("Fully sheltered", "Extremely sheltered", "Very sheltered","Moderately exposed", "Exposed", "Very exposed")
-
-frond_mass_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_mass, fill = wave_exposure)) +
+#morphology and wave exposure
+EW1 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_mass, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
          geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means() + 
+  stat_compare_means(method = "kruskal.test") + 
   xlab("") + 
@@
   guides(fill=guide_legend(title="Wave exposure level")) 
-primary_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
+EW2 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
          geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means() + 
+  stat_compare_means(method = "kruskal.test") + 
   xlab("") + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))  
-primary_width_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
+EW3 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
          geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means() + 
+  stat_compare_means(method = "kruskal.test") + 
   xlab("") + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))  
-frond_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
+EW4 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
          geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means() + 
+  stat_compare_means(method = "kruskal.test") + 
   xlab("") + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))  
-stipe_mass_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
+EW5 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
          geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means() + 
+  stat_compare_means(method = "kruskal.test") + 
   xlab("") + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))  
-stipe_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
+EW6 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
          geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means() + 
+  stat_compare_means(method = "kruskal.test") + 
   xlab("") + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))  
-stipe_circ_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +
+EW7 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
          geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means() + 
+  stat_compare_means(method = "kruskal.test") + 
   xlab("") + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))  
-tufts_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
+EW8 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
          geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means() + 
+  stat_compare_means(method = "kruskal.test") + 
   xlab("") + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))  
-epi_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
+EW9 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
          scale_x_discrete(limits = pos_exposure) + 
          geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means() + 
+  stat_compare_means(method = "kruskal.test") + 
   xlab("") + 
@@
   guides(fill=guide_legend(title="Wave exposure level")) 
-total_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
+EW10 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure) + 
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means() + 
+  stat_compare_means(method = "kruskal.test") + 
   xlab("") + 
@@
   guides(fill=guide_legend(title="Wave exposure level")) 
-
-```
-
-```{r ecklonia plots, echo=FALSE, message=FALSE, warning=FALSE}
-
-ggarrange(frond_mass_plot,primary_length_plot,primary_width_plot,frond_length_plot,stipe_mass_plot,stipe_length_plot,stipe_circ_plot, tufts_plot, epi_length_plot,total_length_plot, labels = "AUTO", common.legend = TRUE, legend = "bottom")
-
-```
+```
+
+```{r deep ecklonia temp plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=9, fig.width=9, fig.cap ='Boxplot of morphological characteristics for deep population *E.
+maxima* categorised by temperature regime.'}
+ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
+```
+
+```{r deep ecklonia wave plots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=9, fig.width=9, fig.cap = 'Boxplot of morphological characteristics for deep population *E.
+maxima* categorised by wave exposure.'}
+ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8, EW9, EW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")
+```
+
+```{r shallow ecklonia temp processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+EST1 <- ggplot(data = eck_shall, aes(x = temp_regime, y = frond_mass, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Frond mass (g)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+EST2 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Primary length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+EST3 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Primary width (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+EST4 <- ggplot(data = eck_shall, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Frond length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+EST5 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Stipe mass (g)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+EST6 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Stipe length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+EST7 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Stipe circ (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+EST8 <- ggplot(data = eck_shall, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("No. of tufts") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+EST9 <- ggplot(data = eck_shall, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Epi length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+EST10 <- ggplot(data = eck_shall, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
+  scale_x_discrete(limits = pos_temp) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "wilcox.test") + 
+  xlab("") + 
+  ylab("Total length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+```
+
+```{r shallow ecklonia wave processing, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+eck_shall$wave_exposure <- factor(eck_shall$wave_exposure, levels=c("Fully sheltered", "Extremely sheltered", "Sheltered", "Exposed", "Very exposed"), labels=c("Fully sheltered", "Extremely sheltered", "Sheltered", "Exposed", "Very exposed"))
+# custom positions for wave exposure categories and sites
+pos_exposure_shall <- c("Fully sheltered", "Extremely sheltered", "Sheltered", "Exposed", "Very exposed")
+#morphology and wave exposure
+ESW1 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = frond_mass, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure_shall) + 
+  geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means(method = "kruskal.test") + 
+  xlab("") + 
+  ylab("Frond mass (g)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
+ESW2 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure_shall) + 
+  geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means(method = "kruskal.test") + 
+  xlab("") + 
+  ylab("Primary length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
+ESW3 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure_shall) + 
+  geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means(method = "kruskal.test") + 
+  xlab("") + 
+  ylab("Primary width (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
+ESW4 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure_shall) + 
+  geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means(method = "kruskal.test") + 
+  xlab("") + 
+  ylab("Frond length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
+ESW5 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure_shall) + 
+  geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means(method = "kruskal.test") + 
+  xlab("") + 
+  ylab("Stipe mass (g)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
+ESW6 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure_shall) + 
+  geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means(method = "kruskal.test") + 
+  xlab("") + 
+  ylab("Stipe length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
+ESW7 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure_shall) + 
+  geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means(method = "kruskal.test") + 
+  xlab("") + 
+  ylab("Stipe circumference (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
+ESW8 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure_shall) + 
+  geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means(method = "kruskal.test") + 
+  xlab("") + 
+  ylab("No of Tufts") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level"))  
+ESW9 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure_shall) + 
+  geom_boxplot(outlier.shape = NA)  + 
+  stat_compare_means(method = "kruskal.test") + 
+  xlab("") + 
+  ylab("Epi length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
+ESW10 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
+  scale_x_discrete(limits = pos_exposure_shall) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  stat_compare_means(method = "kruskal.test") + 
+  xlab("") + 
+  ylab("Total length (cm)") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) +
+  scale_fill_brewer(palette="OrRd") +
+  guides(fill=guide_legend(title="Wave exposure level")) 
+```
+
+```{r shallow ecklonia temp plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=9, fig.width=9, fig.cap ='Boxplot of morphological characteristics for shallow population *E.
+maxima* categorised by temperature regime.'}
+ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
+```
+
+```{r shallow eck wave plot, message=FALSE,warning=FALSE,include=TRUE, fig.height=9, fig.width=9 ,fig.cap = 'Boxplot of morphological characteristics for shallow population *E.
+maxima* categorised by wave exposure.'}
+ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, ESW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")
+```

### 2019-07-16 doodle.R
housecleaning

@@ -1,112 +1,0 @@
-# Calculate the necessary stats
-# Monthly clims
-wave_feb_aug <- wave_model %>% 
-  mutate(date = lubridate::month(date, label = TRUE)) %>% 
-  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp' 
-  group_by(site, date) %>% 
-  summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T) %>%
-  select(-dir_mean, -dirw_mean, -dir_sd, dirw_sd, -spw_mean,-dirw_sd, -spw_sd, -dir_min, -dirw_min, - spw_min, -dir_max, -dirw_max, -spw_max ) %>% 
-  filter(date %in% c("Feb", "Aug")) %>% 
-  ungroup()
-
-# Annual clims
-wave_ann <- wave_model %>% 
-  mutate(date = lubridate::month(date, label = TRUE)) %>% 
-  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp'
-  group_by(site, date) %>% 
-  summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T) %>%
-  ungroup() %>% 
-  select(-date) %>% 
-  group_by(site) %>% 
-  summarise_all(funs(mean), na.rm = T) %>% 
-  mutate(date = "Ann") %>% 
-  select(site, date, everything())  %>%
-  select(-dir_mean, -dirw_mean, -dir_sd, dirw_sd, -spw_mean,-dirw_sd, -spw_sd, -dir_min, -dirw_min, - spw_min, -dir_max, -dirw_max, -spw_max )
-
-# Subset directions out
-dirs <- subset(wave_model, select = c(1, 2, 5, 6, 7))
-
-# Seasons
-dirs_feb_aug <- dirs %>% 
-  mutate(date = lubridate::month(date, label = TRUE)) %>% 
-  group_by(site, date) %>% 
-  summarise(dir_min = min(dir, na.rm = T),
-            dir_max = max(dir, na.rm = T),
-            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
-            dir_median = median(dir, na.rm = T), 
-            dirw_min = min(dirw, na.rm = T),
-            dirw_max = max(dirw, na.rm = T),
-            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
[…]

### 2019-07-16 stats_chapter_2.R

@@ -1,1089 +1,0 @@
-library(tidyverse)
-library(vegan)
-library(RColorBrewer)
-library(gridExtra)
-library(lubridate)
-library(GGally)
-library(ggpubr)
-library(ggrepel)
-
-## Map of study sites
-# Load map data
-
-load("maps/south_africa_coast.Rdata")
-load("maps/sa_provinces.Rdata")
-load("maps/site_list.Rdata")
-
-# remove De Hoop
-kelp_sites <- read_csv("metadata/sites_complete.csv")
-kelp_sites_map <- subset(kelp_sites, !site == "De Hoop")
-
-# Subset into 4 regions
-west__up_site <- subset(kelp_sites, site == "St Helena Bay" | site == "Sea Point" 
-                        | site == "Oudekraal" | site == "Hout Bay")
-
-west__down_site <- subset(kelp_sites, site == "Soetwater" | site == "Olifantsbos" | site == "Scarborough" | site == "Kommetjie")
-
-east_up_site <- subset(kelp_sites, site == "Roman Rock" | site == "A-Frame" | site == "Miller's Point" | site == "Baboon Rock" | site == "Kalk Bay")
-
-east_down_site <- subset(kelp_sites, site == "Batsata Rock" | site == "Bordjies Deep" | site == "Buffelsbaai South" 
-                         | site == "Buffelsbaai" | site == "Betty's Bay")
-
-# plot map
-
-cape_town_df <- data.frame(lat=c(-33.9249), # making seperate data frame to inlcude Cape Town 
-                lon=c(18.4241))
-
-ggplot(data = kelp_sites_map, aes(x = lon, y = lat)) +
-  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
-  geom_path(data = sa_provinces, aes(group = group)) +
-  geom_point(data = kelp_sites_map)  +
-  coord_fixed(xlim = c(17.65, 19.5), ylim = c(-32.75, -34.6)) +
-  geom_label_repel(data = west__up_site, aes(x = lon, y = lat, label = site), 
-                   size = 3, box.padding = 0.5, nudge_x = -0.5, nudge_y = 0.2, segment.alpha = 0.4) +
-  geom_label_repel(data = west__down_site, aes(x = lon, y = lat, label = site), 
-                   size = 3, box.padding = 0.5, nudge_x = -0.5, nudge_y = -0.2, segment.alpha = 0.4) +
[…]

### 2019-07-19 LaTeX/frontiers_LaTex_templates/preview_paper_morph.Rmd
frontiers template added. latex files and rmd file updated

---
title: "R Notebook"
output:
  word_document: default
  pdf_document: default
  html_notebook:
    css: frontiersSCNS.cls
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

### 2019-07-19 paper_morph_new.Rmd
frontiers template added. latex files and rmd file updated

@@
 output:
-  pdf_document: 
-    fig_crop: no
   word_document: 
     fig_caption: yes
+    reference_docx: frontiers_template.docx
+  pdf_document: 
+    fig_caption: yes
+    fig_crop: no
+    fig_height: 7
+    fig_width: 7
+  html_notebook: default
 papersize: A4
@@
 
-# Methods
-### Site selection
+# Materials and methods
+## Site selection
 
@@
 
-```{r echo=FALSE, fig.cap='Schematic of *E. maxima* on the left and *L. pallida* on the right. (Dyer 2018)'}
-
-library(png)
-library(grid)
-
-img <- readPNG("figures/kelp_diagram.png")
-grid.raster(img)
-
-```
-
-## Study species
-
-The morphological characteristics of both species are presented in Fig.
-2.
-*Laminaria pallida* is characterised by a single smooth blade which is divided longitudinally into sections, and develops from a single meristematic region located at the junction between the blade and the stipe (Dyer 2018).
-This species has a solid stipe but develops a hollow stipe along the west coast northward and into Namibia (Rothman et al.
-2017).
-*Ecklonia maxima* consists of a single primary blade which develops above a gas-filled float and a hollow stipe below.
-Secondary blades are produced laterally from the primary blade from several meristematic regions along the margins of the primary blade, known as digits (Dyer 2018).
-Both species are held to the substrate by finger-like haptera, collectively known as the holdfast.
-
-## Abiotic variables
+## Abiotic environment
+### Temperature
 
@@
 2013).
-Linear interpolated SST were calculated for sites where in situ recorders were absent.
+Linear interpolated SST were calculated for sites where *in situ* recorders were absent.
+
+### wave environment
+
 Wave model variable data formed part of the South African Coastal Vulnerability Assessment, presented to the Department of Environmental Affairs (DEA) and produced by the Council for Scientific and Industrial Research (CSIR) (Rautenbach 2015).
@@
 For this study the 7 meter contours were used.
+
+### Collection of kelp morphological characteristics
+
+The morphological characteristics of both species are presented in Fig.
+2.
+*Laminaria pallida* is characterised by a single smooth blade which is divided longitudinally into sections, and develops from a single meristematic region located at the junction between the blade and the stipe (Dyer 2018).
+This species has a solid stipe but develops a hollow stipe along the west coast northward and into Namibia (Rothman et al.
+2017).
+*Ecklonia maxima* consists of a single primary blade which develops above a gas-filled float and a hollow stipe below.
+Secondary blades are produced laterally from the primary blade from several meristematic regions along the margins of the primary blade, known as digits (Dyer 2018).
+Both species are held to the substrate by finger-like haptera, collectively known as the holdfast.
+
+```{r echo=FALSE, fig.cap='Schematic of *E. maxima* on the left and *L. pallida* on the right. (Dyer 2018)'}
+library(png)
+library(grid)
+
+img <- readPNG("figures/kelp_diagram.png")
+grid.raster(img)
+```
+
+Morphological measurements of *L.
+pallida* and *E.
+maxima* were collected at 18 sites along the Western Cape coast of South Africa (Fig.
+1) between October 2014 and April 2015.
+In each instance only the largest individuals were sampled to ensure that only mature and fully grown sporophytes were measured.
+Thirteen samples were collected for each species at each of the sites.
+The morphological traits named in Tables 1 and 2 were measured for comparison between sites.
+Because the macroalgae differ in morphological features, species-specific morphological characteristics were included.
+This allowed comparison between adjacent and non-adjacent sites around the peninsula.
+Between February 2017 and November 2018, morphological measurements for *E.
+maxima* individuals in shallow water (<1m) at seven sites along the Western Cape coast of South Africa were also collected.
+The same morphological characteristic measurements were taken as for the deeper *E.
+maxima*, and this allowed comparison between morphological characteristics between deep and shallow individuals within sites.
+Measurements were not collected for *L.
+pallida* in the shallow depths, as this species is largely absent from the shallow in this portion of the South African coastline.
+
+## Statisical analyses
+
 Site-specific summary statistics were calculated for each of the environmental variables considered in this study.
-The summary statistics calculated for temperature, waves and wind variables were minimum, maximum, mean, range and standard deviation for annual, February[6] (Austral summer) and August (Austral winter) timescales, respectively.
+The summary statistics calculated for temperature, waves and wind variables were minimum, maximum, mean, range and standard deviation for annual, February (Austral summer) and August (Austral winter) timescales, respectively.
 In addition, annual and seasonal median wave and wind direction were also calculated; all summary data and their respective abbreviation are presented in Table 3.
@@
 
+*insert paragraph about morphology*
+Characterising and comparison of morphological characteristics between deep and shallow *E.
+maxima* was achieved by notched boxplots.
+Distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package (Oksanen et al.
+2016) in R (R Core Team, 2019).
+The role of temperature, waves and wind as drivers of kelp morphology was investigated.
+The abiotic data covered all study sites for both species of kelp.
+Both morphology and abiotic data (temperature, waves and wind) were standardised using the *decostand* function in the *vegan* software package; morphology data were used as response variables and the abiotic data as explanatory variables.
+To determine the explanatory variables that best describe patterns in the response data, a full RDA was performed using a complete set of explanatory variables.
+Forward selection was then used to reduce the number of explanatory variables as well as prevent the inflation of overall type I error.
+To further improve the model, pairwise coefficients and Variance Inflation Factor (VIF) were calculated to identify variables with high multicollinearity.
+The computation of the parsimonious RDA was followed by permutation tests of the adjusted *R*2 to assess significance of constraints.
+
 # Results
-## Temperature variables
+## Abiotic environment
+
+The mean annual coastal water temperature for the study sites located on the western side of the peninsula ranged from 14.5 ± 0.9 ℃ (mean ± SD) at St.
+Helena Bay, the most northern site on the western side of the peninsula, to 14.6 ± 0.3 ℃ at Olifantsbos, the most southern site on the western side of the peninsula.
+For sites located on the eastern side of the peninsula (within False Bay), the annual mean coastal water temperatures ranged from 15.5 ± 0.9 ℃ at Buffelsbaai to 15.0 ± 0.9 ℃ at Betty’s Bay, the most eastern site on the coastline in this study.
+Oudekraal had the lowest annual mean coastal water temperature of 12.3 ± 0.9.
+Annual maximum coastal temperatures for study sites located on the western side of the peninsula ranged from 16.3 ℃ at St.
+Helena Bay to 15.1 ℃ at Olifantsbos.
+For sites located east of the peninsula, annual maximum coastal water temperatures ranged from 16 ℃ at Buffelsbaai to 16.5 ℃ at Betty’s Bay.
+Amongst all sites, Kalk Bay had the highest annual maximum temperature of 19 ℃.
+Annual minimum coastal water temperatures for the study sites on the western side of the peninsula ranged from 12.9 ℃ at St.
+Helena Bay to 14.1 ℃ at Olifantsbos, and on the western side of the peninsula ranged from 15.0 ℃ at Buffelsbaai to 13.6 ℃ at Betty’s Bay.
+The annual range in temperature on the western side on the peninsula was 3.4 at St.
+Helena Bay to 0.9 at Olifantsbos, and on the eastern side on of the peninsula was 1.0 at Buffelsbaai to 2.9 at Betty’s Bay.
+As can be seen in Fig 3, the trend in the range of temperature decreases from St.
+Helena Bay to Miller’s point and then increases from Kalk Bay to Betty’s Bay.
+Also, annual range in temperatures within False Bay are larger in winter (August) compared to summer (February).
 
@@
   spread(key = clim_stat, value = value)
-
 ```
@@
   theme(legend.position = "none") + 
-  coord_flip()
-
+  coord_flip() +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
 # Summer temp
@@
   theme(legend.position = "none") + 
-  coord_flip()
+  coord_flip() +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
 
@@
   theme(legend.position = "none") + 
-  coord_flip()
-
+  coord_flip() +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
 ```
@@
 ```
-
-## Wave variables
 
@@
 wave_clim <- cbind(wave_clim, dirs_clim)
-
 # remove all useless direction metrics from original wave_clim
 wave_clim <- subset(wave_clim, select = -c(2,3,4,5,12,13,14,15,22,23,24,25,32,33,40))
-
 ```
@@
   theme(plot.title = element_text(hjust = 0)) + 
-  coord_flip()
+  coord_flip() +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
 
@@
   ggtitle("C") + 
-  coord_flip()
+  coord_flip() +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
 
@@
         axis.ticks.y=element_blank()) +
-  coord_flip()
+  coord_flip() +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
 
@@
         axis.ticks.y=element_blank()) +
-  coord_flip()
-
+  coord_flip() +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
 ```
@@
 
-## Morphology
-### *L. pallida*
+Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula on both an annual and seasonal scale.
+Annual mean significant wave height ranged from 0.5 ± 0.2 m (mean ± SD) at St.
+Helena Bay to 2.3  ± 0.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side of the peninsula it ranged from 0.9 ± 0.4 m at Buffelsbaai to 1.9 ± 0.6 m at Betty’s Bay (Fig.
+4).
+The annual maximum significant wave height ranged from 1.4 m at St.
+Helena Bay to 4.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side ranged from 2.6 m at Buffelsbaai to 4.2 m at Betty’s Bay.
+These wave heights and their variability are modulated by the relative sheltering against the predominant south-westerly swell direction and the size of the fetch for local wave generation.
+
+Annual mean wave peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St.
+Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay ([Fig.
+4].
+Annual peak wave period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and summer respectively, Miller’s point and had the lowest peak wave period amongst all sites.
+These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side of the peninsula, with Miller’s point showing the highest variation across timescales.
+Annual maximum peak wave period ranged from 18.9 s at St.
+Helena Bay to 18.4 s at Olifantsbos for the western side of the peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty’s Bay on the eastern side of the peninsula.
+Data for median wave direction showed no trend across timescales; however the variation (SD) in wave direction is lower on the western side of the peninsula compared to the eastern side both annually and for winter.
+Summer data were the exception and showed lower variation for the eastern side of the peninsula [Fig.
+4].
+Wave variables divided into Annual (top), Austral summer (middle) and Austral winter (bottom) respectively.
+This phenomenon can be explained by the sheltering effect of the Cape Peninsula, resulting in a wave climate with a narrower directional spreading on the eastern side of the peninsula (refer to Fig.
+7).
+
+**Insert wave exposure maps here**
+
+In Figure 5 the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
+The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in [Fig.
+6].
+The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits.
+The western periphery of the Cape peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures [Fig.
+6].
+THere the marked seasonality, with higher energy waves during winter, may be clearly observed once more.
+To clarify the averaged wave exposure maps presented, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in [Fig.
+7].
+Tracing the wave height contours into False Bay its clear why this bay’s western periphery is predominantly sheltered.
+It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions.
+This results in positive and negative wave exposure anomalies all around the peninsula.
+
+## Drivers of kelp morphological characteristics
 
@@
 
-```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=9, fig.width=9, fig.cap='Boxplot of morphological characteristics for *L. pallida* categorised by temperature regime.'}
-ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom", label = "p.format", ncol = 3, nrow = 3)
-```
-
-```{r laminaria wave plots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=9, fig.width=9, fig.cap ='Boxplot of morphological characteristics for *L. pallida* categorised by wave exposure.'}
+```{r laminaria plots, echo=FALSE, fig.cap='Boxplots of *L.
+pallida morphological characteristics group by temperature regime and wave exposure category.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, include=FALSE}
+
+ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
 ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
 ```
-
-### *E.maxima*
 
@@
 
-```{r deep ecklonia temp plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=9, fig.width=9, fig.cap ='Boxplot of morphological characteristics for deep population *E.
-maxima* categorised by temperature regime.'}
+```{r deep ecklonia  plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=8, fig.width=8, fig.cap ='Boxplot of morphological characteristics for deep population *E.
+maxima* grouped by temperature regime and wave exposure.'}
 ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
-```
-
-```{r deep ecklonia wave plots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=9, fig.width=9, fig.cap = 'Boxplot of morphological characteristics for deep population *E.
-maxima* categorised by wave exposure.'}
 ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8, EW9, EW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")
@@
 
-```{r shallow ecklonia temp processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+```{r shallow ecklonia processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
 EST1 <- ggplot(data = eck_shall, aes(x = temp_regime, y = frond_mass, fill = temp_regime)) +
@@
         panel.background = element_blank())
-```
-
-```{r shallow ecklonia wave processing, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+
+# morphology grouped by wave exposure
 eck_shall$wave_exposure <- factor(eck_shall$wave_exposure, levels=c("Fully sheltered", "Extremely sheltered", "Sheltered", "Exposed", "Very exposed"), labels=c("Fully sheltered", "Extremely sheltered", "Sheltered", "Exposed", "Very exposed"))
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
 ```
 
-```{r shallow ecklonia temp plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=9, fig.width=9, fig.cap ='Boxplot of morphological characteristics for shallow population *E.
-maxima* categorised by temperature regime.'}
+```{r shallow ecklonia temp plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=8, fig.width=8, fig.cap ='Boxplot of morphological characteristics for shallow population *E.
+maxima* grouped by temperature regime and wave exposure.'}
+
 ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
+ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, ESW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")
+
 ```
 
-```{r shallow eck wave plot, message=FALSE,warning=FALSE,include=TRUE, fig.height=9, fig.width=9 ,fig.cap = 'Boxplot of morphological characteristics for shallow population *E.
-maxima* categorised by wave exposure.'}
-ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, ESW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")
-```
+# Discussion
+
+This study considered important drivers of kelp morphology that have been identified in previous research (Fowler-Walker et al.
+2006; Miller et al.
+2011; Fowler-Walker et al.
+2005; Stewart et al.
+2009; Wing et al.
+2007; Molloy & Bolton 1996; Pedersen & Nejrup 2012; Wernberg & Thomsen 2005), namely temperature and wave exposure, around the Cape Peninsula on two prolific kelp species.
+Unlike previous research, this study used a numerical model to provide the various variables of wave exposure.
+We also considered wind in addition to temperature and wave variables, since wind has been shown to influence the extension capabilities of canopy kelp species (Denny et al.
+1997).
+Furthermore, the protrusion of the primary blade of *E.
+maxima* out of the water may result in some direct effects of wind-drag on morphology.
+The results show that wave exposure the main driver of kelp morphology with temperature and wind playing only minor roles.
+Furthermore, the investigation of differences in morphological characteristics between shallow and deep populations of *E.
+maxima *suggests that in low wave exposure environments the role of temperature as a morphological driver increases.
+
+The complex geomorphology of the western Cape coastline creates an ideal natural laboratory for studies of kelps in interaction with their environment.
+This complex coastal environment created by the coastal wind and wave climate creates an ideal opportunity to explore the effects of waves and the thermal regime on the morphological development of kelp.
+The annual and seasonal patterns in both wind influence the thermal regime around the coastline by inducing upwelling.
+Upwelling brings deep, cool nutrient rich water to coastal areas and therefore induces a decrease in temperature (Field et al.
+1980; Andrews & Hutchings 1980; Gill & Clarke 1974; Cram 1970; Rouault et al.
+2010; Blanke et al.
+2002.
+The wind speed and wind direction help drive upwelling in summer on the western side of the Cape peninsula, where southerly winds blow parallel to the coast and trigger upwelling (Field et al.
+1980; Rouault et al.
+2010).
+This is not the case for False Bay which is shielded from the dominating wind and swell the Cape Peninsula is subject to, with the exception being the winter months where upwelling becomes more favourable for the region due to the change in swell and wind direction, causing lower temperatures to dominate during this time (Rouault et al.
+2010; Andrews & Hutchings 1980).
+Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula (Rouault et al.
+2010; Field et al.
+1980).
+We however see a possible lack of upwelling events for sites such as Betty’s Bay and Batsata Rock.
+
+We provide strong support for the idea that variations in the environmental variables, particularly wave exposure variables, are driving kelp morphological characteristics around the Cape Peninsula.
+Morphological adaptation due to water motion may manifest in a number of ways in high wave energy environments.
+For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment (Fowler-Walker et al.
+2006; Wernberg & Thomsen 2005; Friedland & Denny 1995; Denny & Roberson 2002; Bekkby et al.
+2014; Denny et al.
+1997; Denny & Gaylord 2002) have been suggested in previous studies.
+Although this study did not measure force of attachment, other morphological responses to wave variables were evident.

### 2019-07-19 paper_morph_new.tex
frontiers template added. latex files and rmd file updated

\documentclass[10pt,a4paper,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Temperature and wave exposure as drivers of morphology in two kelp species along the coast of South Africa},
            pdfauthor={R. Coppin, C. Rautenbach, T. Ponton and A.J Smit},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Temperature and wave exposure as drivers of morphology in two kelp
species along the coast of South Africa}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{R. Coppin, C. Rautenbach, T. Ponton and A.J Smit}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Kelps are a group of large seaweeds of the order Laminariales
(Ochrophyta), which despite their relatively low taxonomic diversity of
species in genera (Bolton 2010), nevertheless form the basis of one of
the most productive ecosystems globally (Mann 1973). Kelps generally
have a dependence on cool, temperate and arctic seawater temperatures
(Bolton 2010; Mohring et al.~2014; Cavanaugh et al.~2011; Dayton 1985),
and dominate the nearshore biomass within the rocky shallow coasts in
both hemispheres (Steneck et al.~2002). Their size and complex
morphology provide a heterogeneous habitat structure (Steneck et
al.~2002) that accommodate a multitude of turf and subcanopy seaweed
species, and diverse assemblages of sessile and mobile invertebrates and
vertebrates (Mann 1973; Steneck et al.~2002), each depending on a wide
suite of ecological services provided by the kelp forests (Gaines \&
Roughgarden 1987; Paul \& Steneck 1993; Levin 1994; Willis \& Anderson
2003; Anderson et al.~1997; Dayton 1985). Wave exposure and temperature
are regarded as important environmental drivers of kelp forests, and
play a role in the distribution (Gorman et al.~2013), abundance
(Cavanaugh et al.~2011; Dayton et al.~1998), diversity (Wing et
al.~2007; Wernberg \& Goldberg 2008), composition (Leliaert et al.~2000;
Dayton 1985; Norderhaug et al.~2012; Harley et al.~2012), growth
(Cousens 1982) and productivity (Pedersen \& Nejrup 2012; Dayton et
al.~1998; Krumhansl \& Scheibling 2012) of kelps.

Wave exposure and temperature have also been shown to be the main
drivers of morphological adaptation across various kelp species. Past
research has shown that in highly wave exposed areas kelp morphology
tends to take on characteristics which reduce drag, increase strength
and increase flexibility (Denny \& Gaylord 2002). For example, a study
by Wernberg \& Thomsen (2005) examined the consistency of wave exposure
as a driver of \emph{Ecklonia radiata} (C. Agardh) J. Agardh across a
broad geographic range, and showed trends towards drag-reducing (small
size, narrow laterals and blades, low spinosity) and increased strength
(large holdfast, thick stipe and thick blades and lamina) at high wave
exposure sites. The morphological adaptation to wave exposure must be
kept in balance with other important processes, such as nutrient
assimilation and light absorption, which are dependent on the amount of
surface area of the blades. Therefore, there is a trade-off between
reducing overall drag and maintaining nutrient and photosynthetic
ability. Temperature has also been shown to play a role in driving kelp
morphology. For example, a study by Serisawa et al.~(2002) compared the
morphology of Ecklonia cava Kjellman growing in warm temperate and cool
temperate morphologies. The results showed that wrinkles in the blade
seem to be a characteristic of warm temperate regions. However this
study did not take into account the interaction with other environmental
variables such as wave exposure, which also affect kelp morphology. The
morphological adapatibility of kelps is driven by a combination of
environmental factors that in turn do not act independently of one
another. A study by Wernberg et al.~(2003) investigated the morphology
of \emph{E. radiata} in order to quantify the morphological variation
and whether it was dependent on spatial differences along the
Australasian coast. They found no correlation between spatial distance
and morphological similarity and rather the morphology of kelps was
representative of multiple environmental forcings on different
morphological characters at different spatial scales (Wernberg et
al.~2003). Due to the complex effects of environmental drivers on kelp
morphology, one can also expect differences in morphology between deep
and shallow water populations of the same species. Wave exposure may be
greater in shallower environments, or it may be reduced through damping
effects of kelp further offshore and natural barriers such as rock
outcrops; also, seawater temperature may be higher in shallow water
environments due to higher solar irradiance.

Other species of kelp, such as \emph{Laminaria pallida} and
\emph{Ecklonia maxima} (Osbeck) Papenfuss which are important habitat
forming seaweeds that exist around the coast of South Africa, offer a
unique opportunity to investigate the drivers of macroalgal
morphological characteristics between deep and shallow water
environments. Although both these species exist in the subtidal,
\emph{L. pallida} dominates deeper waters while \emph{E. maxima} forms
dense surface canopies in the shallower waters. Therefore, \emph{E.
maxima} is exposed to variations in wave, swell and temperature, while
\emph{L. pallida} is exposed to variations in waves and temperature.
(Molloy \& Bolton 1996) investigated the effect of depth and wave
exposure on \emph{L. pallida} at different depths and wave exposure and
showed that depth had a greater effect than wave exposure when
considering all the morphological characteristics; when considering
individual characteristics, however, wave exposure had the most
significant effect on blade thickness. Another study by Rothman et
al.~(2017) investigated the changes in morphology in shallow populations
of \emph{L. pallida} and \emph{E. maxima} along the South African
coastline and into Namibia; \emph{E. maxima} exhibited no morphological
changes along the coast but the stipes of \emph{L. pallida} become
increasingly hollow further north along the coastline. They suggested
that turbidity was the environmental driver responsible for this change.

Measures of wave exposure in ecological studies often incorporate
integrative measures of hydrodynamic conditions at a particular site
which are based on cartographical models of wave exposure.
Cartographical models of wave exposure are regarded as
`fetch-based-models' which measure the length of open water associated
with a particular site in a straight line, and are regarded as simple
measures of wave exposure. Advances in numerical modelling based on
physical/linear wave theory incorporate more complexity (wind forcing,
wave-wave interactions, wave breaking, diffraction and variation in wave
direction) into the models and allow for a quantitave, reproducible
approach for measuring wave exposure. Currently, no ecological studies
investigating macroalgal communities have incorporated 3D spectral
numerical modelling. Previous research investigating drivers of kelp
morphological characteristics used qualitative estimates of wave
exposure and did not consider other morphological characteristics.
Furthermore, only shallow populations have been considered in previous
research which ignores the effects of wave damping. No work currently
exists where other morphological characters have been considered,
quantitative estimates of environmental drivers have been calculated,
and the addition of wind as a possible driver of morphological
characteristics has been considered. The aim of this study is,
therefore, to understand how temperature, wave exposure and wind can
influence the morphology in two species of kelps around South Africa.
This will be achieved by initially understanding the variation in
temperature and waves and the consequences for morphological
characteristics of \emph{E. maxima} and \emph{L. pallida}.

\hypertarget{materials-and-methods}{%
\section{Materials and methods}\label{materials-and-methods}}

\hypertarget{site-selection}{%
\subsection{Site selection}\label{site-selection}}

Sites were chosen to represent an array of environmental gradients, as
indicated in Fig. 1. St.~Helena Bay and Betty's Bay constituted the
north western and south eastern boundary sites, respectively. These
sites are roughly 300 km apart, and lie within separate marine
provinces{[}3{]} and cover a cline in wave exposure and temperature.
Study sites span across the majority of the south-west coast, in varying
thermal and wave energy regimes. The region is dominated by kelp
communities that persist in contrasting abiotic environments. The west
coast region has been termed as a cool temperate region, which is
defined as a region where mean monthly temperatures are always above
10°C and always below 15°C (Smit et al.~2013). East of Cape Point marks
the beginning of an overlap or transition area, which is also referred
to as the Benguela-Agulhas Transition Zone (Smit et al.~2013). The
Agulhas Marine Province is characterised by a wide temperature range of
up to 7°C difference between mean monthly temperatures between summer
and winter (Smit et al.~2013). Due to the Cape Peninsula temperate
latitude, winter months bring an increased frequency of frontal
depressions that originate from the Southern Ocean (Reason et al.~2006).
These low pressures are joined by large swells with increased wave
energy. The nearshore environment, with the accompanied biota, therefore
experiences high wave energy events, with increased frequency in winter
(Veitch et al.~2018). The large peninsula acts as an obstruction for
large south westerly swells, providing decreased wave energy along the
west side of False Bay (Shipley 1964). Conversely, the west coast of
Cape Point is battered by these large swells. Multiple sites, therefore,
exist where kelps grow in diverse temperature and wave energy climates,
in close proximity.

The topography and elevation along the Cape Peninsula channel and shield
winds along False Bay. This is however absent in winter, where strong
northerly winds are prevalent from St.~Helena Bay to Betty's Bay (Field
et al.~1980; Jury et al.~1985; Andrews \& Hutchings 1980; Jury 1980).
Patterns are also evident in the wave data. Seasonal variations in
significant wave height (Hs), wave period and wave direction are also
present. The direction of the dominant swell changes to the south west
in winter, generated by strong low pressures that originate from the
southern ocean (Reason et al.~2006), which False Bay is sheltered from
(Shipley 1964; Atkins 1970; Dufois \& Rouault 2012). In summer, these
swells rotate anticlockwise and are able to enter False Bay, providing
an increased variability of Hs and Tp in this region. It should be noted
that what is classified as sheltered around the South African coastline
(a high energy coastline) might be classified as exposed in other
regions of the world (Norderhaug et al.~2012; Leliaert et al.~2000). It
is due to the near consistent south-westerly swell and the complex
orography around the peninsula that the wave energy distribution around
the Cape Peninsula varies significantly over a small geographical area.

\hypertarget{abiotic-environment}{%
\subsection{Abiotic environment}\label{abiotic-environment}}

\hypertarget{temperature}{%
\subsubsection{Temperature}\label{temperature}}

In order to compare abiotic variables for sites around the coast, large
historical databases for temperature, wave energy and wind were
accessed. Shallow water temperatures were sourced from The South African
Coastal Temperature Network (SACTN) website
(\url{https://github.com/ajsmit/SACTN}). In terms of nearshore
temperature, in situ data are preferred over satellite SST, which have
shown to exhibit large biases (Smit et al.~2013). Linear interpolated
SST were calculated for sites where \emph{in situ} recorders were
absent.

\hypertarget{wave-environment}{%
\subsubsection{wave environment}\label{wave-environment}}

Wave model variable data formed part of the South African Coastal
Vulnerability Assessment, presented to the Department of Environmental
Affairs (DEA) and produced by the Council for Scientific and Industrial
Research (CSIR) (Rautenbach 2015). The South African wave climate was
modeled via 20 spectral numerical wave models that simulated the
offshore wave climate to the nearshore. The boundary conditions of these
models were obtained by using the NOAA Wave Watch III (WWIII) model
output, distributed via the National Centers for Environmental
Prediction (NCEP) product (Office of the Director 2000; Environmental
Modeling Center / Marine Modeling Branch 2005). The particular hindcast
product utilised during the DEA-CSIR study spans 1994-2013 at a 3-hour
resolution. These data were then used to model swell propagation into
the coastal models while wind waves (seas) were generated via stationary
computations in the Simulating Waves in the Nearshore SWAN model (Booij
et al.~1997). The assumption of stationary computations are acceptable
as the model domains were small enough so the temporal variation of the
model boundary were slower than the time it takes for that boundary
condition to propagate to the coast. SWAN allows one to extract wave
variables from specific gridded locations in the nearshore. For False
Bay, a resolution of 200 meters was modelled and output produced at both
the 7 meter and 15 meter isobaths. A 200 meter resolution was used as
the False Bay computational grid was nested within a larger grid (1
kilometer resolution). This allowed for a computational effective wave
resolution of increasing resolution, from the NCEP, low resolution
output to nested, high resolution coastal output. For Table Bay and east
of Cape Hangklip the resolution was 500 meters and also had output at
the 7 meters and 15 meter isobaths. These contour outputs were chosen in
the original study by the CSIR as most engineering run-up calculations
require wave parameter information at these contour depths and were the
main focus of the original study. For this study the 7 meter contours
were used.

\hypertarget{collection-of-kelp-morphological-characteristics}{%
\subsubsection{Collection of kelp morphological
characteristics}\label{collection-of-kelp-morphological-characteristics}}

The morphological characteristics of both species are presented in Fig.
2. \emph{Laminaria pallida} is characterised by a single smooth blade
which is divided longitudinally into sections, and develops from a
single meristematic region located at the junction between the blade and
the stipe (Dyer 2018). This species has a solid stipe but develops a
hollow stipe along the west coast northward and into Namibia (Rothman et
al.~2017). \emph{Ecklonia maxima} consists of a single primary blade
which develops above a gas-filled float and a hollow stipe below.
Secondary blades are produced laterally from the primary blade from
several meristematic regions along the margins of the primary blade,
known as digits (Dyer 2018). Both species are held to the substrate by
finger-like haptera, collectively known as the holdfast.

\begin{figure}
\centering
\includegraphics{paper_morph_new_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{Schematic of \emph{E. maxima} on the left and \emph{L. pallida}
on the right. (Dyer 2018)}
\end{figure}

Morphological measurements of \emph{L. pallida} and \emph{E. maxima}
were collected at 18 sites along the Western Cape coast of South Africa
(\protect\hyperlink{D2L_fig_ref_Mapux5cux2520ofux5cux2520Capeux5cux2520peninsulaux5cux2520andux5cux2520studyux5cux2520sitesux5cux2520forux5cux2520bothux5cux2520L.ux5cux2520pallidaux5cux2520andux5cux2520E.ux5cux2520maxima.ux5cux2520Theux5cux2520Redux5cux2520dotux5cux2520representsux5cux2520theux5cux2520positionux5cux2520ofux5cux2520Capeux5cux2520Townux5cux2520asux5cux2520aux5cux2520pointux5cux2520ofux5cux2520reference.}{Fig.
1}) between October 2014 and April 2015. In each instance only the
largest individuals were sampled to ensure that only mature and fully
grown sporophytes were measured. Thirteen samples were collected for
each species at each of the sites. The morphological traits named in
Tables 1 and 2 were measured for comparison between sites. Because the
macroalgae differ in morphological features, species-specific
morphological characteristics were included. This allowed comparison
between adjacent and non-adjacent sites around the peninsula. Between
February 2017 and November 2018, morphological measurements for \emph{E.
maxima} individuals in shallow water (\textless{}1m) at seven sites
along the Western Cape coast of South Africa were also collected. The
same morphological characteristic measurements were taken as for the
deeper \emph{E. maxima}, and this allowed comparison between
morphological characteristics between deep and shallow individuals
within sites. Measurements were not collected for \emph{L. pallida} in
the shallow depths, as this species is largely absent from the shallow
in this portion of the South African coastline.

\hypertarget{statisical-analyses}{%
\subsection{Statisical analyses}\label{statisical-analyses}}

Site-specific summary statistics were calculated for each of the
environmental variables considered in this study. The summary statistics
calculated for temperature, waves and wind variables were minimum,
maximum, mean, range and standard deviation for annual, February{[}6{]}
(Austral summer) and August (Austral winter) timescales, respectively.
In addition, annual and seasonal median wave and wind direction were
also calculated; all summary data and their respective abbreviation are
presented in Table 3. The median was calculated for wind and wave
direction, as issues arise when calculating the mean and standard
deviation for compass metrics. Dot and whisker plots were used to
present the summary statistics for both temperature and wave variables.
Summary statistics for wind was not plotted and instead are discussed,
as the data was course relative to the other environmental variables
considered in this study.

\emph{insert paragraph about morphology} Characterising and comparison
of morphological characteristics between deep and shallow \emph{E.
maxima} was achieved by notched boxplots. Distance-based redundancy
analysis was performed using the \emph{rda} function in the \emph{vegan}
software package \href{https://paperpile.com/c/rrvrnI/yQkq}{(Oksanen et
al.~2016)} in R (R Core Team, 2019). The role of temperature, waves and
wind as drivers of kelp morphology was investigated. The abiotic data
covered all study sites for both species of kelp. Both morphology and
abiotic data (temperature, waves and wind) were standardised using the
\emph{decostand} function in the \emph{vegan} software package;
morphology data were used as response variables and the abiotic data as
explanatory variables. To determine the explanatory variables that best
describe patterns in the response data, a full RDA was performed using a
complete set of explanatory variables. Forward selection was then used
to reduce the number of explanatory variables as well as prevent the
inflation of overall type I error
\href{https://paperpile.com/c/rrvrnI/IjDn}{(Borcard et al.~2011)}. To
further improve the model, pairwise coefficients and Variance Inflation
Factor (VIF) were calculated to identify variables with high
multicollinearity. The computation of the parsimonious RDA was followed
by permutation tests of the adjusted \emph{R}2 to assess significance of
constraints.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{abiotic-environment-1}{%
\subsection{Abiotic environment}\label{abiotic-environment-1}}

The mean annual coastal water temperature for the study sites located on
the western side of the peninsula ranged from 14.5 ± 0.9 ℃ (mean ± SD)
at St.~Helena Bay, the most northern site on the western side of the
peninsula, to 14.6 ± 0.3 ℃ at Olifantsbos, the most southern site on the
western side of the peninsula. For sites located on the eastern side of
the peninsula (within False Bay), the annual mean coastal water
temperatures ranged from 15.5 ± 0.9 ℃ at Buffelsbaai to 15.0 ± 0.9 ℃ at
Betty's Bay, the most eastern site on the coastline in this study.
Oudekraal had the lowest annual mean coastal water temperature of 12.3 ±
0.9 ℃
(\protect\hyperlink{D2L_fig_ref_Temperatureux5cux2520variablesux5cux2520relativeux5cux2520toux5cux2520eachux5cux2520morphologicalux5cux2520characteristicux5cux2520collectionux5cux2520siteux5cux2520aroundux5cux2520theux5cux2520Westernux5cux2520Capeux5cux2520coast.ux5cux2520Temperatureux5cux2520parametersux5cux2520includeux5cux2520minimumux2cux5cux2520maximumux2cux5cux2520meanux2cux5cux2520rangeux5cux2520andux5cux2520standardux5cux2520deviationux5cux2520ux28ux2103ux29.ux5cux2520Forux5cux2520figuresux5cux2520inux5cux2520columnux5cux2520Aux2cux5cux2520blueux5cux2520dotsux5cux2520representux5cux2520minimumux5cux2520temperaturesux2cux5cux2520redux5cux2520representsux5cux2520maximumsux5cux2520andux5cux2520blackux5cux2520datesux5cux2520theux5cux2520meanux5cux2520temperature.ux5cux2520Errorux5cux2520barsux5cux2520representux5cux2520theux5cux2520standardux5cux2520deviation.ux5cux2520Forux5cux2520figuresux5cux2520inux5cux2520columnux5cux2520Bux2cux5cux2520blackux5cux2520dotsux5cux2520representux5cux2520temperatureux5cux2520range.ux5cux2520Temperatureux5cux2520variablesux5cux2520areux5cux2520alsoux5cux2520dividedux5cux2520intoux5cux2520Annualux5cux2520ux28topux29ux2cux5cux2520Australux5cux2520summerux5cux2520ux28middleux29ux5cux2520andux5cux2520Australux5cux2520winterux5cux2520ux28bottomux29.}{Fig.
3}), amongst all sites.

Annual maximum coastal temperatures for study sites located on the
western side of the peninsula ranged from 16.3 ℃ at St.~Helena Bay to
15.1 ℃ at Olifantsbos. For sites located east of the peninsula, annual
maximum coastal water temperatures ranged from 16 ℃ at Buffelsbaai to
16.5 ℃ at Betty's Bay. Amongst all sites, Kalk Bay had the highest
annual maximum temperature of 19 ℃. Annual minimum coastal water
temperatures for the study sites on the western side of the peninsula
ranged from 12.9 ℃ at St.~Helena Bay to 14.1 ℃ at Olifantsbos, and on
the western side of the peninsula ranged from 15.0 ℃ at Buffelsbaai to
13.6 ℃ at Betty's Bay. The annual range in temperature on the western
side on the peninsula was 3.4 at St.~Helena Bay to 0.9 at Olifantsbos,
and on the eastern side on of the peninsula was 1.0 at Buffelsbaai to
2.9 at Betty's Bay. As can be seen in Fig 3, the trend in the range of
temperature decreases from St.~Helena Bay to Miller's point and then
increases from Kalk Bay to Betty's Bay. Also, annual range in
temperatures within False Bay are larger in winter (August) compared to
summer (February).

\begin{figure}
\centering
\includegraphics{paper_morph_new_files/figure-latex/Temperature variables plot-1.pdf}
\caption{Temperature variables at the collection sites around the Cape
Peninsula. Temperature variables include minimum represented by blue
dots.}
\end{figure}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_point).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_errorbar).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_point).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_errorbar).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_point).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_errorbar).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_point).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_errorbar).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_point).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_errorbar).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_point).
\end{verbatim}

\begin{verbatim}
## Warning: Removed 3 rows containing missing values (geom_errorbar).
\end{verbatim}

\begin{figure}
\centering
\includegraphics{paper_morph_new_files/figure-latex/wave variable plot-1.pdf}
\caption{Wave variables at each collection site around the Western Cape
coast. Mean variables are represented by dots and standard deviation by
whiskers.}
\end{figure}

Our data show the western side of the peninsula experiences higher
significant wave heights and variation in wave heights compared to the
eastern side of the peninsula on both an annual and seasonal scale.
Annual mean significant wave height ranged from 0.5 ± 0.2 m (mean ± SD)
at St.~Helena Bay to 2.3 ± 0.7 m at Olifantsbos on the western side of
the peninsula, and on the eastern side of the peninsula it ranged from
0.9 ± 0.4 m at Buffelsbaai to 1.9 ± 0.6 m at Betty's Bay
(\protect\hyperlink{D2L_fig_ref_ux5cux2520Waveux5cux2520parametersux5cux2520relativeux5cux2520toux5cux2520eachux5cux2520morphologicalux5cux2520characteristicux5cux2520collectionux5cux2520siteux5cux2520aroundux5cux2520theux5cux2520Westernux5cux2520Capeux5cux2520coast.ux5cux2520Variablesux5cux2520areux5cux2520colourux5cux2520codedux2cux5cux2520plotsux5cux2520withux5cux2520blueux5cux2520dotsux5cux2520representux5cux2520significantux5cux2520waveux5cux2520heightux5cux2520ux28metersux29ux5cux2520andux5cux2520theux5cux2520associatedux5cux2520standardux5cux2520deviationux2cux5cux2520andux5cux2520plotsux5cux2520withux5cux2520redux5cux2520dotsux5cux2520representux5cux2520significantux5cux2520waveux5cux2520periodux5cux2520ux28secondsux29ux5cux2520andux5cux2520theux5cux2520associatedux5cux2520standardux5cux2520deviation.ux5cux2520Plotsux5cux2520withux5cux2520blackux5cux2520dotsux5cux2520representux5cux2520waveux5cux2520directionux5cux2520ux28bearingux29.ux5cux2520Waveux5cux2520variablesux5cux2520dividedux5cux2520intoux5cux2520Annualux5cux2520ux28topux29ux2cux5cux2520Australux5cux2520summerux5cux2520ux28middleux29ux5cux2520andux5cux2520Australux5cux2520winterux5cux2520ux28bottomux29ux5cux2520respectively.}{Fig.
4}). The annual maximum significant wave height ranged from 1.4 m at
St.~Helena Bay to 4.7 m at Olifantsbos on the western side of the
peninsula, and on the eastern side ranged from 2.6 m at Buffelsbaai to
4.2 m at Betty's Bay. These wave heights and their variability are
modulated by the relative sheltering against the predominant
south-westerly swell direction and the size of the fetch for local wave
generation.

Annual mean wave peak period for sites on the western side of the
peninsula ranged from 10.3 ± 3.5 s at St.~Helena Bay to 11.0 ± 2.0 s at
Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s
at Betty's Bay
(\protect\hyperlink{D2L_fig_ref_ux5cux2520Waveux5cux2520parametersux5cux2520relativeux5cux2520toux5cux2520eachux5cux2520morphologicalux5cux2520characteristicux5cux2520collectionux5cux2520siteux5cux2520aroundux5cux2520theux5cux2520Westernux5cux2520Capeux5cux2520coast.ux5cux2520Variablesux5cux2520areux5cux2520colourux5cux2520codedux2cux5cux2520plotsux5cux2520withux5cux2520blueux5cux2520dotsux5cux2520representux5cux2520significantux5cux2520waveux5cux2520heightux5cux2520ux28metersux29ux5cux2520andux5cux2520theux5cux2520associatedux5cux2520standardux5cux2520deviationux2cux5cux2520andux5cux2520plotsux5cux2520withux5cux2520redux5cux2520dotsux5cux2520representux5cux2520significantux5cux2520waveux5cux2520periodux5cux2520ux28secondsux29ux5cux2520andux5cux2520theux5cux2520associatedux5cux2520standardux5cux2520deviation.ux5cux2520Plotsux5cux2520withux5cux2520blackux5cux2520dotsux5cux2520representux5cux2520waveux5cux2520directionux5cux2520ux28bearingux29.ux5cux2520Waveux5cux2520variablesux5cux2520dividedux5cux2520intoux5cux2520Annualux5cux2520ux28topux29ux2cux5cux2520Australux5cux2520summerux5cux2520ux28middleux29ux5cux2520andux5cux2520Australux5cux2520winterux5cux2520ux28bottomux29ux5cux2520respectively.}{Fig.
4}). The data shows no trend in mean peak wave period for the coastline.
Annual peak wave period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and
summer respectively, Miller's point and had the lowest peak wave period
amongst all sites. These data show that the western side of the
peninsula has a lower variation (SD) compared to the eastern side of the
peninsula, with Miller's point showing the highest variation across
timescales. Annual maximum peak wave period ranged from 18.9 s at
St.~Helena Bay to 18.4 s at Olifantsbos for the western side of the
peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty's
Bay on the eastern side of the peninsula. Data for median wave direction
showed no trend across timescales; however the variation (SD) in wave
direction is lower on the western side of the peninsula compared to the
eastern side both annually and for winter. Summer data were the
exception and showed lower variation for the eastern side of the
peninsula
(\protect\hyperlink{D2L_fig_ref_ux5cux2520Waveux5cux2520parametersux5cux2520relativeux5cux2520toux5cux2520eachux5cux2520morphologicalux5cux2520characteristicux5cux2520collectionux5cux2520siteux5cux2520aroundux5cux2520theux5cux2520Westernux5cux2520Capeux5cux2520coast.ux5cux2520Variablesux5cux2520areux5cux2520colourux5cux2520codedux2cux5cux2520plotsux5cux2520withux5cux2520blueux5cux2520dotsux5cux2520representux5cux2520significantux5cux2520waveux5cux2520heightux5cux2520ux28metersux29ux5cux2520andux5cux2520theux5cux2520associatedux5cux2520standardux5cux2520deviationux2cux5cux2520andux5cux2520plotsux5cux2520withux5cux2520redux5cux2520dotsux5cux2520representux5cux2520significantux5cux2520waveux5cux2520periodux5cux2520ux28secondsux29ux5cux2520andux5cux2520theux5cux2520associatedux5cux2520standardux5cux2520deviation.ux5cux2520Plotsux5cux2520withux5cux2520blackux5cux2520dotsux5cux2520representux5cux2520waveux5cux2520directionux5cux2520ux28bearingux29.ux5cux2520Waveux5cux2520variablesux5cux2520dividedux5cux2520intoux5cux2520Annualux5cux2520ux28topux29ux2cux5cux2520Australux5cux2520summerux5cux2520ux28middleux29ux5cux2520andux5cux2520Australux5cux2520winterux5cux2520ux28bottomux29ux5cux2520respectively.}{Fig.
4}). This phenomenon can be explained by the sheltering effect of the
Cape Peninsula, resulting in a wave climate with a narrower directional
spreading on the eastern side of the peninsula (refer to Fig. 7).

In Figure 5 the total coastal wave exposure of the Cape Peninsula is
given in terms of wave energy (kW per meter wave crest length). The
directional sheltering effect of the Cape Peninsula, against the
dominant swell direction is clearly observed in the wave exposures
presented in {[}Fig. 6{]}(\#D2L\_fig\_ref\_The total and seasonal wave
roses of the directional wave buoy just offshore from Cape Town is given
form the years 2000 till end 2017 (the period for which directional wave
spectra was available.). The classification from fully sheltered to
extremely exposed is based on the total wave energy upper and lower
limits. The western periphery of the Cape peninsula is almost
continuously exposed to high wave exposures while the eastern periphery
of the peninsula (western coastline of False Bay) revealed sheltered
wave exposures {[}Fig. 6{]}(\#D2L\_fig\_ref\_The total and seasonal wave
roses of the directional wave buoy just offshore from Cape Town is given
form the years 2000 till end 2017 (the period for which directional wave
spectra was available.). Here the marked seasonality, with higher energy
waves during winter, may be clearly observed once more. To clarify the
averaged wave exposure maps presented, the propagation of a typical
offshore wave spectrum as produced from a single time-step in SWAN is
presented in
\protect\hyperlink{D2L_fig_ref_Theux5cux2520propagationux5cux2520ofux5cux2520aux5cux2520typicalux5cux2520offshoreux5cux2520waveux5cux2520spectrumux5cux2520isux5cux2520givenux5cux2520asux5cux2520producedux5cux2520fromux5cux2520aux5cux2520singleux5cux2520time-stepux5cux2520inux5cux2520SWAN}{Fig.
7}. Tracing the wave height contours into False Bay its clear why this
bay's western periphery is predominantly sheltered. It should be
mentioned that some of the annual winter frontal depression systems pass
the Cape Peninsula from the west to east, resulting in wave propagating
towards the continent from much more southerly directions. This results
in positive and negative wave exposure anomalies all around the
peninsula.

\hypertarget{drivers-of-kelp-morphological-characteristics}{%
\subsection{Drivers of kelp morphological
characteristics}\label{drivers-of-kelp-morphological-characteristics}}

\includegraphics{paper_morph_new_files/figure-latex/laminaria plots-1.pdf}
\includegraphics{paper_morph_new_files/figure-latex/laminaria plots-2.pdf}

\includegraphics{paper_morph_new_files/figure-latex/deep ecklonia plots-1.pdf}
\includegraphics{paper_morph_new_files/figure-latex/deep ecklonia plots-2.pdf}

\includegraphics{paper_morph_new_files/figure-latex/shallow ecklonia temp plots-1.pdf}
\includegraphics{paper_morph_new_files/figure-latex/shallow ecklonia temp plots-2.pdf}

\end{document}

### 2019-07-24 morphology_paper.Rmd
upadted rmd file, renamed. Script reduced. test outputs added.

--- 2019-07-19 paper_morph_new.Rmd
@@
 output:
-  word_document: 
-    fig_caption: yes
-    reference_docx: frontiers_template.docx
   pdf_document:
@@
   html_notebook: default
+  html_document:
+    df_print: paged
+  word_document:
+    reference_docx: frontiers_template.docx
 papersize: A4
@@
 library(gridExtra)
+library(ggplot2)
+library(ggord)
+library(vegan)
 ```
@@
 ```{r Temperature variable processing, message=FALSE, warning=FALSE, include=FALSE}
-
 ### Plots for temperature variables 
-
 positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","De Hoop", "St Helena Bay")
@@
 # Annual temp
-temp_plot_A <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
+temp_plot <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
   geom_point() + 
@@
   ylab("") + 
-  ggtitle("A") +
   theme(plot.title = element_text(hjust = 0)) +
@@
         panel.background = element_blank())
-# Summer temp
-temp_plot_B <- ggplot(data = temp_clim, aes(x = site, y = Feb_mean_temp)) +
-  geom_point() + 
-  scale_x_discrete(limits = positions_temp) + 
-  ylim(8, 22) +
-  geom_errorbar(aes(x = site, ymin = Feb_mean_temp - Feb_sd_temp, ymax = Feb_mean_temp + Feb_sd_temp   )) + 
-  geom_point(data = temp_clim, aes(x = site, y = Feb_max_temp, colour = "blue")) + 
-  geom_point(data = temp_clim, aes(x = site, y = Feb_min_temp, colour = "pink")) + 
-  xlab("") + 
-  ylab("") + 
-  ggtitle("B") +
-  theme(legend.position = "none") + 
-  coord_flip() +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
-
-# Winter temp
-temp_plot_C <- ggplot(data = temp_clim, aes(x = site, y = Aug_mean_temp)) +
-  geom_point() + 
-  scale_x_discrete(limits = positions_temp) + 
-  ylim(8, 22) + 
-  geom_errorbar(aes(x = site, ymin = Aug_mean_temp - Aug_sd_temp, ymax = Aug_mean_temp + Aug_sd_temp   )) + 
-  geom_point(data = temp_clim, aes(x = site, y = Aug_max_temp, colour = "blue")) + 
-  geom_point(data = temp_clim, aes(x = site, y = Aug_min_temp, colour = "pink")) + 
-  xlab("") + 
-  ylab("") + 
-  ggtitle("C") +
-  theme(legend.position = "none") + 
-  coord_flip() +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
-```
-
-```{r Temperature variables plot, echo=FALSE, warning=FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula.
-Temperature variables include minimum represented by blue dots.", fig.height=8, fig.width=8}
-
-temp_plot_all <- grid.arrange(temp_plot_A, temp_plot_B,temp_plot_C, ncol = 1, nrow = 3 )
-
-ggsave("temperature_plot.png", plot = temp_plot_all, path = "figures/", dpi = 300, width = 5, height = 5, limitsize = TRUE, scale = 2)
-
+
+```
+
+```{r Temperature variables plot, echo=FALSE, warning=FALSE, fig.cap='Temperature variables at the collection sites around the Cape Peninsula.
+Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.'}
+temp_plot
 ```
@@
 ```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-
 ## Data wrangling and preparation
-
 # Wave data 
-
 # Load the loading functions
 source("func/load.wave.R")
-
 # Then load the data
@@
   group_by(site, date) %>% 
-  summarise_all(funs(mean = mean, sd = sd), na.rm = T) %>%
+  summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T) %>%
   filter(date %in% c("Feb", "Aug")) %>% 
@@
   group_by(site, date) %>% 
-  summarise_all(funs(mean = mean, sd = sd), na.rm = T) %>%
+  summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T) %>%
   ungroup() %>% 
@@
   mutate(date = "Ann") %>% 
-  select(site, date, everything())
-
-# Subset directions out
-dirs <- subset(wave_model, select = c(1, 2, 5, 6))
-
-# Seasons
-dirs_feb_aug <- dirs %>% 
-  mutate(date = lubridate::month(date, label = TRUE)) %>% 
-  group_by(site, date) %>% 
-  summarise(dir_min = min(dir, na.rm = T),
-            dir_max = max(dir, na.rm = T),
-            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
-            dir_median = median(dir, na.rm = T), 
-            dirw_min = min(dirw, na.rm = T),
-            dirw_max = max(dirw, na.rm = T),
-            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
-            dirw_median = median(dirw, na.rm = T)) %>%
-  filter(date %in% c("Feb", "Aug")) %>% 
-  ungroup()
-
-# Annual
-dirs_ann <- dirs %>% 
-  mutate(date = lubridate::month(date, label = TRUE)) %>% 
-  group_by(site, date) %>% 
-  summarise(dir_min = min(dir, na.rm = T),
-            dir_max = max(dir, na.rm = T),
-            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
-            dir_median = median(dir, na.rm = T), 
-            dirw_min = min(dirw, na.rm = T),
-            dirw_max = max(dirw, na.rm = T),
-            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
-            dirw_median = median(dirw, na.rm = T)) %>% 
-  ungroup() %>% 
-  select(-date) %>% 
+  select(site, date, everything()) %>% 
   group_by(site) %>% 
-  summarise_all(funs(mean(., na.rm = T))) %>% 
-  mutate(date = "Ann") %>% 
-  select(site, date, everything())
-
-# subset useless columns
-dirs_feb_aug <- subset(dirs_feb_aug, select = c(1, 2, 6, 10))
-dirs_ann <- subset(dirs_ann, select = c(1, 2, 6, 10))
-
-# combine all direction metrics
-dirs_clim <- rbind(dirs_feb_aug, dirs_ann) %>% 
-  group_by(site) %>% 
-  gather(key = statistic, value = value, 
-         3:4) %>% 
-  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
-  select(-date, -statistic) %>% 
-  spread(key = clim_stat, value = value) %>% 
   left_join(wave_sites, by = c("site" = "wave_7")) %>% 
@@
 
-# Combine
-# NB: This only uses the wave data from the 7 metre contour
-wave_clim <- rbind(wave_feb_aug, wave_ann) %>% 
-  group_by(site) %>% 
-  gather(key = statistic, value = value, 
-         hs_mean:spw_sd) %>% 
-  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
-  select(-date, -statistic) %>% 
-  spread(key = clim_stat, value = value) %>% 
-  left_join(wave_sites, by = c("site" = "wave_7")) %>% 
-  ungroup() %>% 
-  select(-site, wave_15) %>%
-  rename(site = site.y) %>% 
-  select(site, everything()) %>% 
-  na.omit()
-
-# combine all wave metrics
-wave_clim <- cbind(wave_clim, dirs_clim)
-# remove all useless direction metrics from original wave_clim
-wave_clim <- subset(wave_clim, select = -c(2,3,4,5,12,13,14,15,22,23,24,25,32,33,40))
+wave_ann <- wave_ann[-c(9,15,20,22), -c(2,23)]
+
+#remove useless coloumns
+wave_ann <- wave_ann %>% 
+  select(site, hs_mean, hs_min, hs_max, hs_sd, tp_mean, tp_min, tp_max, tp_sd) 
 ```
@@
 ```{r wave variable processing, warning=FALSE, include=FALSE, echo = FALSE, message=FALSE}
-
 ### Plots for wave variables
@@
 
-wave_plot_A <- ggplot(data = wave_clim, aes(x = site, y = Ann_hs_mean)) +
+wave_plot_A <- ggplot(data = wave_ann, aes(x = site, y = hs_mean)) +
   geom_point() + 
@@
   theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
-  geom_errorbar(aes(x = site, ymin = Ann_hs_mean - Ann_hs_sd, ymax = Ann_hs_mean + Ann_hs_sd)) + 
+  geom_errorbar(aes(x = site, ymin = hs_mean - hs_sd, ymax = hs_mean + hs_sd)) + 
   xlab("") + 
@@
   theme(legend.position = "none") + 
-  ggtitle("A") + 
   theme(plot.title = element_text(hjust = 0)) + 
@@
 
-wave_plot_B <- ggplot(data = wave_clim, aes(x = site, y = Ann_tp_mean)) +
+wave_plot_B <- ggplot(data = wave_ann, aes(x = site, y = tp_mean)) +
   geom_point() + 
@@
   theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
-  geom_errorbar(aes(x = site, ymin = Ann_tp_mean - Ann_tp_sd, ymax = Ann_tp_mean + Ann_tp_sd)) + 
+  geom_errorbar(aes(x = site, ymin = tp_mean - tp_sd, ymax = tp_mean + tp_sd)) + 
   xlab("") + 
@@
   theme(legend.position = "none") + 
-  ggtitle("B") + 
   theme(plot.title = element_text(hjust = 0)) + 
-  theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) +
-  coord_flip()
-
-#Summer wave plots
-wave_plot_C <- ggplot(data = wave_clim, aes(x = site, y = Feb_hs_mean)) +
-  geom_point() + 
-  scale_x_discrete(limits = positions_temp) + 
-  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
-  geom_errorbar(aes(x = site, ymin = Feb_hs_mean - Feb_hs_sd, ymax = Feb_hs_mean + Feb_hs_sd)) + 
-  xlab("") + 
-  ylab("") + 
-  theme(legend.position = "none") + 
-  ggtitle("C") + 
   coord_flip() +
@@
         panel.background = element_blank())  
-
-wave_plot_D <- ggplot(data = wave_clim, aes(x = site, y = Feb_tp_mean)) +
-  geom_point() + 
-  scale_x_discrete(limits = positions_temp) + 
-  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
-  geom_errorbar(aes(x = site, ymin = Feb_tp_mean - Feb_tp_sd, ymax = Feb_tp_mean + Feb_tp_sd)) + 
-  xlab("") + 
-  ylab("") + 
-  theme(legend.position = "none") + 
-    ggtitle("D") + 
-  theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) +
-  coord_flip() +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
-
-# Winter summary temp
-wave_plot_E <- ggplot(data = wave_clim, aes(x = site, y = Aug_hs_mean)) +
-  geom_point() + 
-  scale_x_discrete(limits = positions_temp) + 
-  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
-  geom_errorbar(aes(x = site, ymin = Aug_hs_mean - Aug_hs_sd, ymax = Aug_hs_mean + Aug_hs_sd)) + 
-  xlab("") + 
-  ylab("") + 
-  ggtitle("E") + 
-  theme(legend.position = "none") + 
-  coord_flip() +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
-
-# Winter temperature range
-wave_plot_F <- ggplot(data = wave_clim, aes(x = site, y = Aug_tp_mean)) +
-  geom_point() + 
-  scale_x_discrete(limits = positions_temp) + 
-  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
-  geom_errorbar(aes(x = site, ymin = Aug_tp_mean - Aug_tp_sd, ymax = Aug_tp_mean + Aug_tp_sd)) + 
-  xlab("") + 
-  ylab("") + 
-  theme(legend.position = "none") + 
-  ggtitle("F") + 
-  theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) +
-  coord_flip() +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
 ```
@@
 ```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE, fig.cap='Wave variables at each collection site around the Western Cape coast.
-Mean variables are represented by dots and standard deviation by whiskers.', fig.height=8, fig.width=8}
-
-wave_plot_all <- grid.arrange(wave_plot_A, wave_plot_B, wave_plot_C, wave_plot_D, wave_plot_E, wave_plot_F, ncol = 2, nrow = 3)
-
-ggsave("wave_plot.png", plot = wave_plot_all, path = "figures/", dpi = 300, width = 6, height = 6, limitsize = TRUE, scale = 2)
-
+Mean variables are represented by dots and standard deviation by whiskers.'}
+ggarrange(wave_plot_A, wave_plot_B, ncol = 2, nrow = 1)
 ```
@@
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
   stat_compare_means(method = "wilcox.test") + 
@@
   scale_x_discrete(limits = pos_temp)  + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
   stat_compare_means(method = "wilcox.test") + 
@@
 
-#wave exposure and lamanaria
-lam_df$wave_exposure <- factor(lam_df$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))
-pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed")
+lam_df$wave_exposure <- factor(lam_df$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))#wave exposure levels for lamanaria
+pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed")#custom positions
 
@@
 ```{r laminaria plots, echo=FALSE, fig.cap='Boxplots of *L.
-pallida morphological characteristics group by temperature regime and wave exposure category.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, include=FALSE}
-
-ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
-ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
+pallida* morphological characteristics group by temperature regime and wave exposure category.', message=FALSE, warning=FALSE, include=FALSE}
+ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = FALSE, legend = "bottom")
 ```
@@
 ```{r deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
-library(ggplot2)
 ## Preparing data
@@
 
-```{r deep ecklonia  plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=8, fig.width=8, fig.cap ='Boxplot of morphological characteristics for deep population *E.
-maxima* grouped by temperature regime and wave exposure.'}
-ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
-ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8, EW9, EW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")
+```{r deep ecklonia plot, message=FALSE, warning= FALSE, fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by temperature regime and wave exposure.'}
+ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10,EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
 ```
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+        panel.background = element_blank()) + 
+  ylim(0, 20)
 EST4 <- ggplot(data = eck_shall, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+        panel.background = element_blank()) +
+  ylim(0, 30)
 EST8 <- ggplot(data = eck_shall, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
@@
   scale_fill_brewer(palette="OrRd") +
-  guides(fill=guide_legend(title="Wave exposure level"))  
+  guides(fill=guide_legend(title="Wave exposure level")) + 
+  ylim(0,25)
 ESW4 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
@@
 
-```{r shallow ecklonia temp plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=8, fig.width=8, fig.cap ='Boxplot of morphological characteristics for shallow population *E.
+```{r shallow ecklonia temp plots, echo=FALSE, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E.
 maxima* grouped by temperature regime and wave exposure.'}
-
-ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
-ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, ESW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")
-
+ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = FALSE, legend = "bottom")
+```
+
+```{r echo=FALSE}
+#calculate mean values for lamanaria morphology
+lam_mean <- lam_df %>% 
+  subset(select = -c(1,2)) %>% 
+  group_by(site) %>% 
+  summarise_all(funs(mean), na.rm = T) %>% 
+  ungroup()
+
+# merge temp and wave data to create environmental data-frame
+abio_clim <- merge(temp_ann, wave_ann, by = "site")
+
+#remove date column
+abio_clim <- subset(abio_clim, select = -c(2))
+
+# standardise measurements
+abio_clim <- abio_clim %>%
+  subset(select = -c(1)) %>% 
+  decostand(method = "standardize")
+
+# run RDA
+RDA_all_eck <- rda(eck_bio_deep ~ ., data = eck_abio_deep)
+
+# Plot RDA
+plot(RDA_all_eck, scaling  = 2)
 ```
@@
 Although this study did not measure force of attachment, other morphological responses to wave variables were evident.
+
+# Conclusion
+
+Past research shows that macroalgal morphological characteristics are largely driven by a wave exposure while others have suggested a complex interaction between various environmental variables.
+We confirm that wave variables are the main driver of kelp morphological characteristics in both *E.
+maxima *and *L.
+pallida* populations around the Cape Peninsula, and differences in shallow and deep populations of *E.
+maxima* may reflect small differences in wave and temperature variables as drivers of kelp morphological characteristics.
+
+# Conflict of Interest Statement
+
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest
+
+# Author Contributions
+
+RC conceptualised the scope of the research reported in this paper, collected samples, undertook the majority of numerical and statistical analyses, made the first round of interpretation, and did the majority of the writing.
+AS and CR provided guidance in terms of  writing, statistical analyses, and the final draft of the paper, as well some writing in the relevant sections.
+RS aided in fieldwork and statistical analyses for the temperature data.
+TP performed the initial analyses and interpretation of the data.
+
+# Funding
+
+The research was funded by the South African National Research Foundation ([http://www.nrf.ac.za](http://www.nrf.ac.za)) programme "Thermal characteristics of the South African nearshore:implications for biodiversity" (CPRR14072378735).
+
+Aside from funding provided, the funders had no role in the study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+
+# Acknowledgments
+
+Thank you to AS and CR for their guidance, and the honours class of 2014-2015 that aided in fieldwork. Carla Louise for making the GIS plot etc.

### 2019-07-26 morphology_paper.Rmd
rda outputs updated. figures created and added. project outputs updated

@@
 output:
+  html_notebook: default
   pdf_document:
@@
     fig_width: 7
-  html_notebook: default
   html_document:
@@
 #Annual wave plots
+pos_temp <- c("Cool temperate", "Warm temperate")
 
@@
 ```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>% 
-   filter(site!= "Rocky Bank")
-
-# remove unnecessary columns
-lam_df <- subset(lam_data, select = -c(1,2,6))
-
-pos_temp <- c("Cool temperate", "Warm temperate")#custom temperature categories position on x-axis
+#custom temperature categories position on x-axis
+pos_temp <- c("Cool temperate", "Warm temperate")
+
 #temperature and lamanaria  
-LT1 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
+LT1 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
@@
         panel.background = element_blank())
-LT2 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
+LT2 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp)  + 
@@
         panel.background = element_blank())
-LT3 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
+LT3 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
@@
         panel.background = element_blank())
-LT4 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
+LT4 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp)  + 
@@
         panel.background = element_blank())
-LT5 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
+LT5 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp)  + 
@@
         panel.background = element_blank())
-LT6 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
+LT6 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp)  + 
@@
         panel.background = element_blank())
-LT7 <- ggplot(data = lam_df, aes(x = temp_regime, y = digits, fill = temp_regime)) +
+LT7 <- ggplot(data = lam_data, aes(x = temp_regime, y = digits, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp)  + 
@@
         panel.background = element_blank())
-LT8 <- ggplot(data = lam_df, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
+LT8 <- ggplot(data = lam_data, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
@@
         panel.background = element_blank())
-LT9 <- ggplot(data = lam_df, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
+LT9 <- ggplot(data = lam_data, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
@@
 
-lam_df$wave_exposure <- factor(lam_df$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))#wave exposure levels for lamanaria
+lam_data$wave_exposure <- factor(lam_data$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))#wave exposure levels for lamanaria
 pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed")#custom positions
 
-LW1 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
+LW1 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam) + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-LW2 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
+
+LW2 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-LW3 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
+
+LW3 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam) + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-LW4 <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
+
+LW4 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-LW5 <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
+
+LW5 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-LW6 <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
+
+LW6 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-LW7 <- ggplot(data = lam_df, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
+
+LW7 <- ggplot(data = lam_data, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam)  + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-LW8 <- ggplot(data = lam_df, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
+
+LW8 <- ggplot(data = lam_data, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam) + 
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-LW9 <- ggplot(data = lam_df, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
+
+LW9 <- ggplot(data = lam_data, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
   scale_x_discrete(limits = pos_exposure_lam) + 
@@
 
-```{r echo=FALSE}
+```{r rda analysis lamanaria, include=FALSE}
+
+# remove unnecessary columns
+lam_df <- subset(lam_data, select = -c(1,2,6))
+
 #calculate mean values for lamanaria morphology
@@
 
+#subset sites into seperate object for aligning data sets
+lam_sites <- subset(lam_mean, select = c(1))
+
+#remove Bordjies North from data-frame
+lam_mean <- lam_mean[-c(5),]
+
+lam_meta <- kelp_sites %>% 
+  filter(population == "Deep") %>% 
+  filter(species == "Both")
+
+lam_meta <- merge(lam_sites, kelp_sites, by = "site")
+#drop useless columns
+lam_meta <- lam_meta[-c(1,9,11), ]
+
+#standardise all abiotic measurements
+bio_lam <- lam_mean %>%
+  subset(select = -c(1)) %>% 
+  decostand(method = "standardize")
+
+#subset sites into seperate object for aligning data sets
+lam_sites <- subset(lam_mean, select = c(1))
+
+#bind site column back to dataframe
+bio_lam <- cbind(lam_sites, bio_lam)
+
+bio_lam <- bio_lam %>%
+  remove_rownames %>%
+  column_to_rownames(var = "site")
+
+#subset sites into seperate object for aligning data sets
+lam_sites <- subset(lam_mean, select = c(1))
+
 # merge temp and wave data to create environmental data-frame
@@
 
+#use site lam_site to align appropriate data
+abio_lam <- merge(lam_sites, abio_clim, by = "site")
+
 #remove date column
-abio_clim <- subset(abio_clim, select = -c(2))
-
-# standardise measurements
-abio_clim <- abio_clim %>%
+abio_lam <- abio_lam[ , -c(2)]
+
+# standardise all abiotic measurements
+abio_lam <- abio_lam %>%
   subset(select = -c(1)) %>% 
@@
 
-# run RDA
-RDA_all_eck <- rda(eck_bio_deep ~ ., data = eck_abio_deep)
-
+abio_lam <- cbind(lam_sites, abio_lam)
+
+abio_lam <- abio_lam %>%
+  remove_rownames %>%
+  column_to_rownames(var = "site")
+
+#run RDA for abiotic variables
+rda_all_lam <- rda(bio_lam ~ ., data = abio_lam)
+
+#summary of rda output
+summary(rda_all_lam)
+
+# Forward selection using vegan's ordistep
+# Note that is function does not use R^2adjusted-based stopping criterion
+
+step.forward <- ordistep(rda(bio_lam ~ 1, data = abio_lam),
+                         scope=formula(rda_all_lam), direction = "forward", pstep = 1000)
+#calculate variance inlfation factor
+#vif.cca(rda_all_lam)
+
+#use pearson correlations to determine best abiotic variables for model
+cor_lam_abio <- round(cor(abio_lam, use = "pair"), 2)
+cor_lam_abio
+
+# Parsimonious RDA
+rda_lam <- rda(bio_lam ~ tp_sd + max_temp + hs_mean, data = abio_lam)
+summary(rda_lam)
+
+#visual screenshot of rda axes to help determine which axes explain the most variation
+screeplot(rda_lam)
+
+#retreive adjusted R2 value which measures the unbiased amount of explained variation
+R2adj <- RsquareAdj(rda_lam)$adj.r.squared
+R2adj
+
+```
+
+```{r rda plot lamanaria, echo=FALSE, message=FALSE, warning=FALSE}
 # Plot RDA
-plot(RDA_all_eck, scaling  = 2)
+ggord(rda_lam, lam_meta$region, ellipse = FALSE, hull = FALSE, grp_title = "Region", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9)  
+```
+
+```{r rda ecklonia processing, echo=FALSE, message=FALSE, warning=FALSE}
+
+#filter to select only deep population ecklonia
+eck_df <- eck_data %>% 
+  filter(population == "Deep")
+
+# remove unnecessary columns
+eck_df <- subset(eck_df, select = -c(1,2,3,4,8,9,10))
+
+#calculate mean values for lamanaria morphology
+eck_mean <- eck_df %>% 
+  subset(select = -c(1,2)) %>% 
+  group_by(site) %>% 
+  summarise_all(funs(mean), na.rm = T) %>% 
+  ungroup()
+
+#remove St Helena Bay for now as wave model not available for that site
+eck_mean <- eck_mean[-c(13), ]
+
+#subset sites into seperate object for aligning data sets
+eck_sites <- subset(eck_mean, select = c(1))
+
+#remove Miller's Point site as it is causing issues, also this keeps the amount of western and eastern sites more balanced
+eck_sites <- eck_sites %>% 
+  filter(site != "Miller's Point")
+
+eck_meta <- read.csv("metadata/eck_meta.csv", sep = ";") 
+eck_meta <- eck_meta %>% 
+  filter(population == "Deep") %>% 
+  mutate(site = replace(site, site == "Miller's Point", "Miller's Point"))
+
+eck_meta <- merge(eck_meta, eck_sites,  by = "site")
+
+#standardise all abiotic measurements
+bio_eck<- eck_mean %>%
+  filter(site != "Miller's Point") %>% 
+  subset(select = -c(1)) %>%
+  decostand(method = "standardize")
+
+#bind site column back to dataframe
+bio_eck <- cbind(eck_sites, bio_eck)
+
+bio_eck <- bio_eck %>%
+  remove_rownames %>%
+  column_to_rownames(var = "site")
+
+#use site eck_site to align appropriate data from abio_clim data from the lamanaria processing section
+abio_eck <- merge(eck_sites, abio_clim, by = "site")
+
+#remove date column
+abio_eck <- abio_eck[ , -c(2)]
+
+# standardise all abiotic measurements
+abio_eck <- abio_eck %>%
+  filter(site != "Miller's Point") %>% 
+  subset(select = -c(1)) %>% 
+  decostand(method = "standardize")
+
+abio_eck <- cbind(eck_sites, abio_eck)
+
+abio_eck <- abio_eck %>%
+  remove_rownames %>%
+  column_to_rownames(var = "site")
+
+#run RDA for abiotic variables
+rda_all_eck <- rda(bio_eck ~ ., data = abio_eck)
+
+#summary of rda output
+summary(rda_all_eck)
+
+# Forward selection using vegan's ordistep
+# Note that is function does not use R^2adjusted-based stopping criterion
+
+step.forward <- ordistep(rda(bio_eck ~ 1, data = abio_eck),
+                         scope=formula(rda_all_eck), direction = "forward", pstep = 1000)
+#calculate variance inlfation factor
+vif.cca(rda_all_eck)
+
+#use pearson correlations to determine best abiotic variables for model
+cor_eck_abio <- round(cor(abio_eck, use = "pair"), 2)
+cor_eck_abio
+
+# Parsimonious RDA
+rda_eck <- rda(bio_eck ~ tp_mean + hs_mean + tp_sd + hs_sd + mean_temp, data = abio_eck)
+summary(rda_eck)
+
+#visual screenshot of rda axes to help determine which axes explain the most variation
+screeplot(rda_eck)
+
+#retreive adjusted R2 value which measures the unbiased amount of explained variation
+R2adj <- RsquareAdj(rda_eck)$adj.r.squared
+R2adj
+
+# Plot RDA
+ggord(rda_eck, eck_meta$region, ellipse = FALSE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  # looking at the raw code, this is plotting the 'wa scores', the blue
+```
+
+```{r rda shallow ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
+
+#filter to select only deep population ecklonia and remove Millers B and C for now...
+eck_df_sh <- eck_data %>% 
+  filter(population == "Shallow") %>% 
+  filter(site != "Miller's Point B") %>% 
+  filter(site != "Miller's Point C")
+
+# remove unnecessary columns
+eck_df_sh <- subset(eck_df_sh, select = -c(1,2,3,4,8,9,10))
+
+#calculate mean values for lamanaria morphology
+eck_mean_sh <- eck_df_sh %>% 
+  subset(select = -c(1,2)) %>% 
+  group_by(site) %>% 
+  summarise_all(funs(mean), na.rm = T) %>% 
+  ungroup()
+
+#subset sites into seperate object for aligning data sets
+eck_sites_sh <- subset(eck_mean_sh, select = c(1))
+
+eck_meta_sh <- read.csv("metadata/eck_meta.csv", sep = ";") 
+eck_meta_sh <- eck_meta_sh %>% 
+  filter(population == "Shallow")
+
+eck_meta_sh <- merge(eck_meta_sh, eck_sites_sh,  by = "site")
+
+#standardise all abiotic measurements
+bio_eck_sh<- eck_mean_sh %>%
+  subset(select = -c(1)) %>%
+  decostand(method = "standardize")
+
+#bind site column back to dataframe
+bio_eck_sh <- cbind(eck_sites_sh, bio_eck_sh)
+
+bio_eck_sh <- bio_eck_sh %>%
+  remove_rownames %>%
+  column_to_rownames(var = "site")
+
+#use site eck_site to align appropriate data from abio_clim data from the lamanaria processing section
+abio_eck_sh <- merge(eck_sites_sh, abio_clim, by = "site")
+
+#remove date column
+abio_eck_sh <- abio_eck_sh[ , -c(2)]
+
+# standardise all abiotic measurements
+abio_eck_sh <- abio_eck_sh %>%
+  subset(select = -c(1)) %>% 
+  decostand(method = "standardize")
+
+abio_eck_sh <- cbind(eck_sites_sh, abio_eck_sh)
+
+abio_eck_sh <- abio_eck_sh %>%
+  remove_rownames %>%
+  column_to_rownames(var = "site")
+
+#run RDA for abiotic variables
+rda_all_eck <- rda(bio_eck_sh ~ ., data = abio_eck_sh)
+
+#summary of rda output
+summary(rda_all_eck)
+
+# Forward selection using vegan's ordistep
+# Note that is function does not use R^2adjusted-based stopping criterion
+
+step.forward <- ordistep(rda(bio_eck ~ 1, data = abio_eck),
+                         scope=formula(rda_all_eck), direction = "forward", pstep = 1000)
+#calculate variance inlfation factor
+vif.cca(rda_all_eck)
+
+#use pearson correlations to determine best abiotic variables for model
+cor_eck_abio <- round(cor(abio_eck, use = "pair"), 2)
+cor_eck_abio
+
+# Parsimonious RDA
+rda_eck <- rda(bio_eck ~ tp_mean + hs_mean + tp_sd + hs_sd + mean_temp, data = abio_eck)
+summary(rda_eck)
+
+#visual screenshot of rda axes to help determine which axes explain the most variation
+screeplot(rda_eck)
+
+#retreive adjusted R2 value which measures the unbiased amount of explained variation
+R2adj <- RsquareAdj(rda_eck)$adj.r.squared
+R2adj
+
+# Plot RDA
+ggord(rda_eck, eck_meta$region, ellipse = FALSE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  # looking at the raw code, this is plotting the 'wa scores', the blue
+
 ```

### 2019-07-30 morphology_paper.Rmd
refining outputs

@@
 output:
-  html_notebook: default
+  html_notebook: 
+    css: frontiers.csl
+    fig_caption: yes
+    fig_height: 7
+    fig_width: 7
   pdf_document:
@@
 # Complete lon/lats
-kelp_sites <- read.csv("metadata/site_meta_complete.csv", sep = ";")
 
@@
   # NB: This may also not be an appropriate change
-  mutate(site = replace(site, site == "Bordjies reef North", "Bordjies Deep")) %>%
+  mutate(site = replace(site, site == "Bordjies North", "Bordjies")) %>%
   # Not using Rocky Bank in this study
@@
 load("data/rr_patch.Rdata")
+kelp_sites <- read.csv("metadata/site_meta_complete.csv", sep = ";") %>% 
+  mutate(site = replace(site, site == "Bakoven", "Sea Point"))
 
@@
   rbind(., rr_patch)
-
 ## Temperature analyses
-# Up next we need to calculate the relevant temperature statistics for each site. These are: #mean, range, and sd, for Annual, February, and August climatologies.
-# Monthly clims
-
-temp_feb_aug <- temps %>% 
-  mutate(date = lubridate::month(date, label = TRUE)) %>% 
-  group_by(site, date) %>% 
-  summarise(mean_temp = mean(temp, na.rm = T),
-            min_temp = min(temp, na.rm = T),
-            max_temp = max(temp, na.rm = T),
-            range_temp = range(temp, na.rm = T)[2]-range(temp, na.rm = T)[1],
-            sd_temp = sd(temp, na.rm = T)) %>%
-  filter(date %in% c("Feb", "Aug")) %>% 
-  ungroup()
-
+# Up next we need to calculate the relevant temperature statistics for each site. These are: #mean, range, and sd, for Annual climatology.
 # Annual clims
@@
   select(site, date, everything())
-
-# Combine
-temp_clim <- rbind(temp_feb_aug, temp_ann) %>% 
-  group_by(site) %>% 
-  gather(key = statistic, value = value, 
-         mean_temp:sd_temp) %>% 
-  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
-  select(-date, -statistic) %>% 
-  spread(key = clim_stat, value = value)
 ```
@@
 ### Plots for temperature variables 
-positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","De Hoop", "St Helena Bay")
+positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
 
@@
   scale_x_discrete(limits = positions_temp) + 
-  ylim(8, 22) + 
+  ylim(9, 21) + 
   geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) + 
@@
   theme(legend.position = "none") + 
-  coord_flip() +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
-
+  coord_flip() 
 ```
@@
   select(site, wave_7, wave_15)
-
-# Calculate the necessary stats
-# Monthly clims
-wave_feb_aug <- wave_model %>% 
-  mutate(date = lubridate::month(date, label = TRUE)) %>% 
-  filter(site != "FB025") %>% # FB_25 has issues with 'hs' and 'tp'
-  group_by(site, date) %>% 
-  summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T) %>%
-  filter(date %in% c("Feb", "Aug")) %>% 
-  ungroup()
 
@@
 #Annual wave plots
-pos_temp <- c("Cool temperate", "Warm temperate")
+
+positions_wave <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St.
+Helena Bay")
 
@@
   geom_point() + 
-  scale_x_discrete(limits = positions_temp) + 
+  scale_x_discrete(limits = positions_wave) + 
   theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
@@
   theme(plot.title = element_text(hjust = 0)) + 
-  coord_flip() +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  coord_flip()
 
@@
   theme(plot.title = element_text(hjust = 0)) + 
-  coord_flip() +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())  
+  coord_flip() 
 ```
@@
 eck_shall <- eck_data %>% 
-  filter(population == "Shallow")
+  filter(population == "Shallow") %>% 
+  mutate(site = replace(site, site == "Bakoven", "Sea Point"))
+
 pos_temp <- c("Cool temperate", "Warm temperate") #custom temperature categories position on x-axis
@@
 maxima* grouped by temperature regime and wave exposure.'}
-ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = FALSE, legend = "bottom")
+ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10,ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = FALSE, legend = "bottom")
 ```
@@
   summarise_all(funs(mean), na.rm = T) %>% 
+  filter(site != "Roman Rock") %>% 
   ungroup()
@@
 lam_sites <- subset(lam_mean, select = c(1))
-
-#remove Bordjies North from data-frame
-lam_mean <- lam_mean[-c(5),]
-
-lam_meta <- kelp_sites %>% 
-  filter(population == "Deep") %>% 
-  filter(species == "Both")
-
-lam_meta <- merge(lam_sites, kelp_sites, by = "site")
-#drop useless columns
-lam_meta <- lam_meta[-c(1,9,11), ]
 
@@
 
-#subset sites into seperate object for aligning data sets
-lam_sites <- subset(lam_mean, select = c(1))
-
 #bind site column back to dataframe
@@
   column_to_rownames(var = "site")
-
-#subset sites into seperate object for aligning data sets
-lam_sites <- subset(lam_mean, select = c(1))
 
@@
 
-```
-
-```{r rda plot lamanaria, echo=FALSE, message=FALSE, warning=FALSE}
+lam_meta_rda <- merge(lam_sites, kelp_sites, by = "site")
+lam_meta_rda <- lam_meta_rda[-c(1,9,11,14), ]
+
 # Plot RDA
-ggord(rda_lam, lam_meta$region, ellipse = FALSE, hull = FALSE, grp_title = "Region", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9)  
-```
-
-```{r rda ecklonia processing, echo=FALSE, message=FALSE, warning=FALSE}
+lam_rda <- ggord(rda_lam, lam_meta_rda$region, ellipse = TRUE, hull = FALSE, grp_title = "Region", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9) 
+```
+
+```{r rda deep ecklonia processing, echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 eck_df <- eck_data %>% 
-  filter(population == "Deep")
+  filter(population == "Deep") %>% 
+  filter(site != "Miller's Point")
 
@@
 #remove St Helena Bay for now as wave model not available for that site
-eck_mean <- eck_mean[-c(13), ]
+eck_mean <- eck_mean %>% 
+  filter(site != "St Helena Bay")
 
@@
 
-#remove Miller's Point site as it is causing issues, also this keeps the amount of western and eastern sites more balanced
-eck_sites <- eck_sites %>% 
-  filter(site != "Miller's Point")
-
-eck_meta <- read.csv("metadata/eck_meta.csv", sep = ";") 
-eck_meta <- eck_meta %>% 
-  filter(population == "Deep") %>% 
-  mutate(site = replace(site, site == "Miller's Point", "Miller's Point"))
-
-eck_meta <- merge(eck_meta, eck_sites,  by = "site")
-
 #standardise all abiotic measurements
 bio_eck<- eck_mean %>%
-  filter(site != "Miller's Point") %>% 
   subset(select = -c(1)) %>%
@@
 abio_eck <- abio_eck %>%
-  filter(site != "Miller's Point") %>% 
   subset(select = -c(1)) %>% 
@@
 #calculate variance inlfation factor
-vif.cca(rda_all_eck)
+#vif.cca(rda_all_eck)
 
@@
 
+eck_meta_rda <- merge(eck_sites, kelp_sites, by = "site")
+eck_meta_rda <- eck_meta_rda %>% 
+  filter(population == "Deep")
+
 # Plot RDA
-ggord(rda_eck, eck_meta$region, ellipse = FALSE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  # looking at the raw code, this is plotting the 'wa scores', the blue
+rda_eck_deep <- ggord(rda_eck, ellipse = TRUE, hull = FALSE, ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  # looking at the raw code, this is plotting the 'wa scores', the blue 
 ```
@@
   filter(population == "Shallow") %>% 
-  filter(site != "Miller's Point B") %>% 
-  filter(site != "Miller's Point C")
+  mutate(site = replace(site, site == "Miller's Point A", "Miller's Point")) %>% 
+  mutate(site = replace(site, site == "Bakoven", "Sea Point"))
 
@@
 
-#calculate mean values for lamanaria morphology
+#calculate mean values for shallow ecklonia morphology
 eck_mean_sh <- eck_df_sh %>% 
@@
 
-eck_meta_sh <- read.csv("metadata/eck_meta.csv", sep = ";") 
-eck_meta_sh <- eck_meta_sh %>% 
-  filter(population == "Shallow")
-
-eck_meta_sh <- merge(eck_meta_sh, eck_sites_sh,  by = "site")
-
 #standardise all abiotic measurements
@@
 #use site eck_site to align appropriate data from abio_clim data from the lamanaria processing section
-abio_eck_sh <- merge(eck_sites_sh, abio_clim, by = "site")
+abio_eck_sh <- merge(abio_clim, eck_sites_sh, by = "site")
 
@@
 #run RDA for abiotic variables
-rda_all_eck <- rda(bio_eck_sh ~ ., data = abio_eck_sh)
+rda_all_ecksh <- rda(bio_eck_sh ~ ., data = abio_eck_sh)
 
 #summary of rda output
-summary(rda_all_eck)
+summary(rda_all_ecksh)
 
@@
 
-step.forward <- ordistep(rda(bio_eck ~ 1, data = abio_eck),
+step.forward <- ordistep(rda(bio_eck_sh ~ 1, data = abio_eck_sh),
                          scope=formula(rda_all_eck), direction = "forward", pstep = 1000)
 #calculate variance inlfation factor
-vif.cca(rda_all_eck)
+#vif.cca(rda_all_eck)
 
 #use pearson correlations to determine best abiotic variables for model
-cor_eck_abio <- round(cor(abio_eck, use = "pair"), 2)
+cor_eck_abio <- round(cor(abio_eck_sh, use = "pair"), 2)
 cor_eck_abio
@@
 # Parsimonious RDA
-rda_eck <- rda(bio_eck ~ tp_mean + hs_mean + tp_sd + hs_sd + mean_temp, data = abio_eck)
-summary(rda_eck)
+rda_eck_shallow <- rda(bio_eck_sh ~ hs_max + hs_mean + sd_temp + mean_temp, data = abio_eck_sh)
+summary(rda_eck_shallow)
 
 #visual screenshot of rda axes to help determine which axes explain the most variation
-screeplot(rda_eck)
+screeplot(rda_eck_shallow)
 
 #retreive adjusted R2 value which measures the unbiased amount of explained variation
-R2adj <- RsquareAdj(rda_eck)$adj.r.squared
+R2adj <- RsquareAdj(rda_eck_shallow)$adj.r.squared
 R2adj
@@
 # Plot RDA
-ggord(rda_eck, eck_meta$region, ellipse = FALSE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  # looking at the raw code, this is plotting the 'wa scores', the blue
-
+rda_eck_shallow <- ggord(rda_eck_shallow, eck_meta_sh$region,ellipse = TRUE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9, coord_fix = TRUE)  # looking at the raw code, this is plotting the 'wa scores', the blue
+
+```
+
+```{r rda plots, echo=FALSE}
+ggarrange(lam_rda, rda_eck_deep,rda_eck_shallow, common.legend = TRUE, labels = "AUTO", legend = "bottom")
 ```

### 2019-07-31 data/temp_data/SACTN_v4.2.R
clean up. deleted and moved latex assocaited files. Deleted figures. metadata updated

@@ -1,0 +1,63 @@
+source("pipe/SACTN_products_v4.2.R")
+
+# Central meeting point for product creation.
+# The rest of the pipe is back end to this
+# Source the scripts used for the SACTN pipe
+
+
+################################################################################
+# List sources with new data added
+sources <- c("DAFF", "DEA", "EKZNW", "KZNSB", "SAEON", "SAWS", "UWC")
+
+# Proto
+## NB: Compile and save the proto data with the following line of code
+  ## Then create whichever other products below you would like
+  ## You must include at least one source or this won't run
+  ## Note that these data are not clean or QAd
+system.time(SACTN_proto_v4.2 <- proto_data(sources)) # 43 seconds
+## Or just load the pre-existing proto data without updating them
+SACTN_proto_v4.2 <- load_proto()
+##
+
+# Site list (meta-data)
+system.time(site_data()) # 26 seconds
+
+# Hourly
+  # This is slow because it is averaging millions of rows
+system.time(hourly_data()) # 132 seconds
+
+# Daily
+system.time(daily_data()) # 30 seconds
+
+# Monthly
+system.time(monthly_data()) # 27 seconds
+
+# Yearly
+system.time(yearly_data()) # 27 seconds
+
+# Climatology
+system.time(clim_data()) # 7 seconds
+
+
+################################################################################
+## Run "shiny/SACTN/app.R" to push the new data to the web
+## One may also use the code within the shiny app to load and save the files correctly
+
+
+################################################################################
+## Then run the following line to regenerate all of the meta-data and figures for the 
+## data from the chosen contributors.
+## It is good practice to e-mail any of the sources with the updated reports for which
+## data they have with us as soon as possible after they submit it.
+source("request_data/contributor_meta_data.R")
+system.time(contributor_meta(sources)) # 37 seconds
+
+
+################################################################################
+## Open the SACTN project and run 'startup.R' to populate the new data in the public access directory
+## Remember to push that project to GitHub to update the public access data
+
+
+################################################################################
+## Extract specific time series as desired
+## Still needs work...

### 2019-07-31 morphology_paper.Rmd
added to intro. boxplots for temp and wave exposure seperated. general additions

@@
 output:
-  html_notebook: 
-    css: frontiers.csl
-    fig_caption: yes
-    fig_height: 7
-    fig_width: 7
   pdf_document:
@@
     fig_width: 7
+  html_notebook:
+    fig_caption: yes
+    fig_height: 7
+    fig_width: 7
   html_document:
@@
   word_document:
-    reference_docx: frontiers_template.docx
+    reference_docx: morphology_paper_files/frontiers_template.docx
 papersize: A4
@@
 ---
+
+```{r include=FALSE}
+knitr::opts_chunk$set(cache=FALSE)
+```
 
@@
 
-Wave exposure and temperature have also been shown to be the main drivers of morphological adaptation across various kelp species.
+Temperature and wave exposure have been shown to play a role in driving kelp morphology.
+For example, a study by Serisawa et al.
+(2002) compared the morphology of Ecklonia cava Kjellman growing in warm temperate and cool temperate morphologies.
+The results showed that wrinkles in the blade seem to be a characteristic of warm temperate regions.
+However this study did not take into account the interaction with other environmental variables such as wave exposure.
+Wave exposure has also been shown to be the main drivers of morphological adaptation across various kelp species.
 Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility (Denny & Gaylord 2002).
@@
 Therefore, there is a trade-off between reducing overall drag and maintaining nutrient and photosynthetic ability.
-Temperature has also been shown to play a role in driving kelp morphology.
-For example, a study by Serisawa et al.
-(2002) compared the morphology of Ecklonia cava Kjellman growing in warm temperate and cool temperate morphologies.
-The results showed that wrinkles in the blade seem to be a characteristic of warm temperate regions.
-However this study did not take into account the interaction with other environmental variables such as wave exposure, which also affect kelp morphology.
 The morphological adapatibility of kelps is driven by a combination of environmental factors that in turn do not act independently of one another.
@@
 pallida* dominates deeper waters while *E.
-maxima* forms dense surface canopies in the shallower waters.
+maxima* forms dense surface canopies in from the subtidal to shallower waters.
 Therefore, *E.
@@
 Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitave, reproducible approach for measuring wave exposure.
-Currently, no ecological studies investigating macroalgal communities have incorporated 3D spectral numerical modelling.
+Currently, no ecological studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
 Previous research investigating drivers of kelp morphological characteristics used qualitative estimates of wave exposure and did not consider other morphological characteristics.
-Furthermore, only shallow populations have been considered in previous research which ignores the effects of wave damping.
-No work currently exists where other morphological characters have been considered, quantitative estimates of environmental drivers have been calculated, and the addition of wind as a possible driver of morphological characteristics has been considered.
-The aim of this study is, therefore, to understand how temperature, wave exposure and wind can influence the morphology in two species of kelps around South Africa.
-This will be achieved by initially understanding the variation in temperature and waves and the consequences for morphological characteristics of *E.
+Furthermore, possible differences in morphology between shallow and deep water populations have not been considered in previous research which ignores the effects of wave damping.
+The aim of this study is, therefore, to understand how temperature and wave exposure can influence the morphology in two species of kelps around South Africa, as well as between deep and shallow populations of the same species.
+This will be achieved by initially understanding the variation in temperature and waves using advanced numerical models.
+These numerical models will then be used as a basis for investigating the consequences for morphological chracteristics of *E.
 maxima* and *L.
-pallida*.
+pallida* using statistical tools.
 
@@
 Linear interpolated SST were calculated for sites where *in situ* recorders were absent.
+These data were used to group sites into 'Cool temperate' and 'Warm temperate' categorisations of thermal regime.
 
@@
 For this study the 7 meter contours were used.
+These data were used to group sites according to
 
@@
   # NB: This may also not be an appropriate change
-  mutate(site = replace(site, site == "Bordjies North", "Bordjies")) %>%
+  mutate(site = replace(site, site == "Bordjies North", "Bordjies Deep")) %>%
   # Not using Rocky Bank in this study
@@
 load("data/rr_patch.Rdata")
-kelp_sites <- read.csv("metadata/site_meta_complete.csv", sep = ";") %>% 
-  mutate(site = replace(site, site == "Bakoven", "Sea Point"))
 
@@
   rbind(., rr_patch)
+
 ## Temperature analyses
@@
 
-```{r laminaria plots, echo=FALSE, fig.cap='Boxplots of *L.
-pallida* morphological characteristics group by temperature regime and wave exposure category.', message=FALSE, warning=FALSE, include=FALSE}
-ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = FALSE, legend = "bottom")
+```{r laminaria temp plots, echo=FALSE, warning=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime.', message=FALSE, warning=FALSE, include=FALSE}
+ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = FALSE, legend = "bottom")
+```
+
+```{r lamanaria wave plots, echo=FALSE,warning=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.'}
+ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = FALSE, legend = "bottom")
 ```
@@
 # Ecklonia
-eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
-  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
-  filter(site!= "De Hoop")  
 # group_by(wave_exposure, frond_mass) %>% 
@@
 
-```{r deep ecklonia plot, message=FALSE, warning= FALSE, fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by temperature regime and wave exposure.'}
-ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10,EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
+```{r deep ecklonia temp plot, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by temperature.'}
+ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
+```
+
+```{r deep ecklonia wave plots, warning=FALSE,fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.'}
+ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
 ```
@@
 
-```{r shallow ecklonia temp plots, echo=FALSE, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by temperature regime and wave exposure.'}
-ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10,ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = FALSE, legend = "bottom")
-```
-
-```{r rda analysis lamanaria, include=FALSE}
+```{r shallow ecklonia temp plots, echo=FALSE, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by temperature regime.'}
+ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
+```
+
+```{r echo=FALSE, warning=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by wave exposure.'}
+ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = FALSE, legend = "bottom")
+```
+
+```{r rda analysis lamanaria, eval=FALSE, include=FALSE}
 
@@
   filter(site != "Roman Rock") %>% 
+  filter(site != "Miller's Point") %>%  #the scipt keeps dropping it when combing the site list with the metadata
   ungroup()
@@
 
-#remove date column
-abio_lam <- abio_lam[ , -c(2)]
+#remove date column and extra Betty's Bay
+abio_lam <- abio_lam[-c(5) , -c(2)]
 
@@
 
-```{r rda deep ecklonia processing, echo=FALSE, message=FALSE, warning=FALSE}
+```{r rda deep ecklonia processing, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 eck_df <- eck_data %>% 
-  filter(population == "Deep") %>% 
-  filter(site != "Miller's Point")
+  filter(population == "Deep")
 
@@
 eck_mean <- eck_mean %>% 
-  filter(site != "St Helena Bay")
+  filter(site != "St Helena Bay") %>% 
+  filter(site != "Miller's Point")
 
@@
 
-#remove date column
-abio_eck <- abio_eck[ , -c(2)]
+#remove date column and extra Betty's Bay row
+abio_eck <- abio_eck[-c(5) , -c(2)]
 
@@
 # Plot RDA
-rda_eck_deep <- ggord(rda_eck, ellipse = TRUE, hull = FALSE, ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  # looking at the raw code, this is plotting the 'wa scores', the blue 
-```
-
-```{r rda shallow ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
+rda_eck_deep <- ggord(rda_eck, eck_meta_rda$region ,ellipse = TRUE, hull = FALSE, ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  # looking at the raw code, this is plotting the 'wa scores', the blue
+```
+
+```{r rda shallow ecklonia processing, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 
@@
   filter(population == "Shallow") %>% 
-  mutate(site = replace(site, site == "Miller's Point A", "Miller's Point")) %>% 
-  mutate(site = replace(site, site == "Bakoven", "Sea Point"))
+  mutate(site = replace(site, site == "Buffelsbaai", "Buffelsbaai South")) %>% 
+  mutate(site = replace(site, site == "Miller's Point C", "Miller's Point")) %>% 
+  mutate(site = replace(site, site == "Miller's Point B", "Baboon Rock")) %>% 
+  filter(site != "St James") %>% 
+  filter(site != "Slangkop") %>% 
+  filter(site != "Miller's Point A")
 
@@
   ungroup() 
+eck_mean_sh <- eck_mean_sh %>% 
+  mutate(site = replace(site, site == "Yzerfontein", "St Helena Bay"))
 
@@
 #use site eck_site to align appropriate data from abio_clim data from the lamanaria processing section
+abio_clim_temp <- abio_clim %>% 
+  mutate(site = replace(site, site == "Sea Point", "Bakoven"))
+
 abio_eck_sh <- merge(abio_clim, eck_sites_sh, by = "site")
@@
 # Plot RDA
-rda_eck_shallow <- ggord(rda_eck_shallow, eck_meta_sh$region,ellipse = TRUE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9, coord_fix = TRUE)  # looking at the raw code, this is plotting the 'wa scores', the blue
-
-```
-
-```{r rda plots, echo=FALSE}
+rda_eck_shallow <- ggord(rda_eck_shallow,ellipse = TRUE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9, coord_fix = TRUE)  # looking at the raw code, this is plotting the 'wa scores', the blue
+
+```
+
+```{r rda plots, eval=FALSE, include=FALSE}
 ggarrange(lam_rda, rda_eck_deep,rda_eck_shallow, common.legend = TRUE, labels = "AUTO", legend = "bottom")

### 2019-07-31 morphology_paper.Rmd
updated project options. added in results section for boxplots and rda outputs completed

@@
 output:
+  html_notebook:
+    fig_caption: yes
+    fig_height: 7
+    fig_width: 7
   pdf_document:
@@
     fig_crop: no
-    fig_height: 7
-    fig_width: 7
-  html_notebook:
-    fig_caption: yes
     fig_height: 7
@@
 2013).
-The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter (Smit et al.
+The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm temperate region (Smit et al.
 2013).
@@
 ## Abiotic environment
+
+In order to compare abiotic variables for sites around the coast, large historical databases for temperature, wave energy and wind were accessed.
+
 ### Temperature
 
-In order to compare abiotic variables for sites around the coast, large historical databases for temperature, wave energy and wind were accessed.
 Shallow water temperatures were sourced from The South African Coastal Temperature Network (SACTN) website (https://github.com/ajsmit/SACTN).
-In terms of nearshore temperature, in situ data are preferred over satellite SST, which have shown to exhibit large biases (Smit et al.
+In terms of nearshore temperature, *in situ* data are preferred over satellite SST, which have shown to exhibit large biases (Smit et al.
 2013).
 Linear interpolated SST were calculated for sites where *in situ* recorders were absent.
-These data were used to group sites into 'Cool temperate' and 'Warm temperate' categorisations of thermal regime.
-
-### wave environment
+These data were used to group sites into 'Cool temperate' and 'Warm temperate' categorisations to investigate the possible affect of thermal regime on kelp morphological characteristics.
+
+### Wave environment
 
@@
 For this study the 7 meter contours were used.
-These data were used to group sites according to
+These data were then used to calculate over all wave power (kW/m) as this was considered the best measure of overall wave exposure.
+Annual wave power was plotted and categorised into different wave exposure categories which ranged from fully sheltered to extremely exposed.
 
@@
 # Results
+
+```{r software packages, message=TRUE, warning=FALSE, include=FALSE, echo=FALSE}
+library(tidyverse)
+library(ggpubr)
+library(gridExtra)
+library(ggplot2)
+library(ggord)
+library(vegan)
+```
+
 ## Abiotic environment
@@
 Also, annual range in temperatures within False Bay are larger in winter (August) compared to summer (February).
-
-```{r software packages, message=TRUE, warning=FALSE, include=FALSE, echo=FALSE}
-library(tidyverse)
-library(ggpubr)
-library(gridExtra)
-library(ggplot2)
-library(ggord)
-library(vegan)
-```
 
@@
 ```
+
+Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula on both an annual and seasonal scale.
+Annual mean significant wave height ranged from 0.5 ± 0.2 m (mean ± SD) at St.
+Helena Bay to 2.3  ± 0.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side of the peninsula it ranged from 0.9 ± 0.4 m at Buffelsbaai to 1.9 ± 0.6 m at Betty’s Bay.
+Annual mean wave peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St.
+Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay.
+The data shows no trend in mean peak wave period for the coastline.
+Annual peak wave period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and summer respectively, Miller’s point and had the lowest peak wave period amongst all sites.
+These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side of the peninsula, with Miller’s point showing the highest variation across timescales.
+Annual maximum peak wave period ranged from 18.9 s at St.
+Helena Bay to 18.4 s at Olifantsbos for the western side of the peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty’s Bay on the eastern side of the peninsula.
 
@@
 
-```{r wave variable processing, warning=FALSE, include=FALSE, echo = FALSE, message=FALSE}
+```{r wave variable processing, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 ### Plots for wave variables
@@
 
-positions_wave <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St.
-Helena Bay")
+positions_wave <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
 
@@
 
-**Insert wave exposure maps here**
+```{r wave power map, echo=FALSE, fig.cap='Schematic of *E. maxima* on the left and *L. pallida* on the right. (Dyer 2018)'}
+library(png)
+library(grid)
+
+img_wavepower <- readPNG("figures/CapePen_Total.PNG")
+grid.raster(img_wavepower)
+```
 
@@
 
+Significant differences were found between cool temperate and warm temperate regimes for all *L.
+pallida* morphological differences with lamina length, number of digits and thallus mass the exceptions.
+Certian morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significanlty higher means for the cool temperate regimes when compared to the warm temperate regime.
+Conversely, lamina weight and lamina thickness has significantly lower means for the cool temperate regime compared to the warm temperate regime.
+When *L.
+pallida* morphological characteristics were grouped to wave exposure categories kruskal-wallis tests revealed significant differences amoungst categories for all morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
+The morphological characteristics that were significantly different amoungst sites exibited similar patterns, with increasing mean values from the full sheltered to extremely sheltered categories and then a decrease in mean values for the remainder of the categories.
+Two morphological characteristics, namely lamina thickness and stipe diameter, both exibit higher variations for the exposed categories.
+
 ```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
@@
 
-```{r laminaria temp plots, echo=FALSE, warning=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime.', message=FALSE, warning=FALSE, include=FALSE}
-ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = FALSE, legend = "bottom")
-```
-
-```{r lamanaria wave plots, echo=FALSE,warning=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.'}
-ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = FALSE, legend = "bottom")
-```
+```{r laminaria temp plots, echo=FALSE, warning=FALSE, fig.cap='Boxplots of *L.
+pallida* morphological characteristics group by temperature regime.', message=FALSE, warning=FALSE, include=FALSE, fig.height=8, fig.width=7}
+ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
+```
+
+```{r lamanaria wave plots, echo=FALSE,warning=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.', fig.height=8, fig.width=8}
+ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
+```
+
+All deep *E.
+maxima* morphological characteristics showed differences between cool temperate and warm temperate regimes except for frond mass, frond length and stipe mass.
+Mean values for primary length, primary width and stipe circumference were all significantly higher for the cool temperate regime than the warm temperate regime.
+The remaining morphological characteristics which showed significant differences between regimes exhibited significanlty lower values for the cool temperate regime compared to the warm temperate regime.
+Significant differences amoungst wave exposure categories were found for all morphological characteristics of deep *E.
+maxima* populations.
+Certian morphological characteristics showed a gradual increase in mean value with inncreasing wave exposure.
+The remaining morphological characteristics exhibited decreased values for only the higher exposure categories and then a sharp increase in value for the highest wave exposure category.
+In all cases, mean values for the highest wave exposure level (very exposed) was greatest compared to other levels of exposure.
 
@@
 
-```{r deep ecklonia temp plot, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by temperature.'}
-ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
-```
-
-```{r deep ecklonia wave plots, warning=FALSE,fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.'}
-ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
+```{r deep ecklonia temp plot, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by temperature.',fig.height=8, fig.width=8}
+ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
+```
+
+```{r deep ecklonia wave plots, warning=FALSE,fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.',fig.height=8, fig.width=8}
+ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
 ```

### 2019-08-01 morphology_paper.Rmd
fixed issue when creating abiotic dataset. More writing added in results section. RDA test output html

@@
 output:
-  html_notebook:
-    fig_caption: yes
-    fig_height: 7
-    fig_width: 7
+  html_document:
+    df_print: paged
   pdf_document:
@@
     fig_width: 7
-  html_document:
-    df_print: paged
   word_document:
     reference_docx: morphology_paper_files/frontiers_template.docx
+  html_notebook:
+    fig_caption: yes
+    fig_height: 7
+    fig_width: 7
 papersize: A4
@@
   filter(site!= "De Hoop")
+eck_data$site <- as.character(eck_data$site)
 
@@
 lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>%
-  mutate(site = replace(site, site == "Buffels Bay", "Buffelsbaai")) %>%
-  # NB: Even though I am changing the name, there will still be no lon/lat values for this site
-  # We need to decide how we want to handle this. Give it the same lon/lat as Buffelsbaai?
-  mutate(site = replace(site, site == "Buffels Bay South", "Buffelsbaai South")) %>%
   # NB: This may also not be an appropriate change
-  mutate(site = replace(site, site == "Bordjies North", "Bordjies Deep")) %>%
   # Not using Rocky Bank in this study
   filter(site != "Rocky Bank")
+lam_data$site <- as.character(lam_data$site)
 
@@
 ### Plots for temperature variables 
-positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
+positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock", "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
 
@@
   select(site, wave_7, wave_15)
-
 # Annual clims
@@
 
+#remove useless coloumns
 wave_ann <- wave_ann[-c(9,15,20,22), -c(2,23)]
 
-#remove useless coloumns
 wave_ann <- wave_ann %>% 
   select(site, hs_mean, hs_min, hs_max, hs_sd, tp_mean, tp_min, tp_max, tp_sd)
+
+# Subset directions out
+dirs <- subset(wave_model, select = c(1, 2, 5, 6))
+
+# Annual
+dirs_ann <- dirs %>% 
+  mutate(date = lubridate::month(date, label = TRUE)) %>% 
+  group_by(site, date) %>% 
+  summarise(dir_min = min(dir, na.rm = T),
+            dir_max = max(dir, na.rm = T),
+            dir_range = range(dir, na.rm = T)[2]-range(dir, na.rm = T)[1],
+            dir_median = median(dir, na.rm = T), 
+            dirw_min = min(dirw, na.rm = T),
+            dirw_max = max(dirw, na.rm = T),
+            dirw_range = range(dirw, na.rm = T)[2]-range(dirw, na.rm = T)[1],
+            dirw_median = median(dirw, na.rm = T)) %>% 
+  ungroup() %>% 
+  select(-date) %>% 
+  group_by(site) %>% 
+  summarise_all(funs(mean(., na.rm = T))) %>% 
+  mutate(date = "Ann") %>% 
+  select(site, date, everything())
+
+# combine all direction metrics
+dirs_ann <- dirs_ann %>% 
+  group_by(site) %>% 
+  gather(key = statistic, value = value, 
+         3:4) %>% 
+  mutate(clim_stat = paste(date, statistic, sep = "_")) %>% 
+  select(-date, -statistic) %>% 
+  spread(key = clim_stat, value = value) %>% 
+  left_join(wave_sites, by = c("site" = "wave_7")) %>% 
+  ungroup() %>% 
+  select(-site, wave_15) %>%
+  rename(site = site.y) %>% 
+  select(site, everything()) %>% 
+  na.omit()
+
+# Combine
+#these will be used later for the redudancy analysis
+clim <- merge(wave_ann, dirs_ann, by = "site")  
+clim <- clim[ , -c(12,13,16,17,18)]
+#remove duplicate sites
+clim <- clim[-c(1,6,7,8,12,18,20), ]
+
+# merge temp and wave data to create environmental data-frame
+abio_clim <- merge(temp_ann, clim, by = "site")
 ```
@@
 
-```{r wave power map, echo=FALSE, fig.cap='Schematic of *E. maxima* on the left and *L. pallida* on the right. (Dyer 2018)'}
+```{r wave power map, echo=FALSE, fig.cap='wave power maps'}
 library(png)
@@
 img_wavepower <- readPNG("figures/CapePen_Total.PNG")
-grid.raster(img_wavepower)
+legend_wave_power <- readPNG("figures/CapePen_Legend.PNG")
+grid.raster(img_wavepower,legend_wave_power)
 ```
@@
 Significant differences were found between cool temperate and warm temperate regimes for all *L.
-pallida* morphological differences with lamina length, number of digits and thallus mass the exceptions.
-Certian morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significanlty higher means for the cool temperate regimes when compared to the warm temperate regime.
-Conversely, lamina weight and lamina thickness has significantly lower means for the cool temperate regime compared to the warm temperate regime.
+pallida* morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
+Certian morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significanlty higher means kelps in cool temperate regimes when compared to kelps for the warm temperate regime.
+Conversely, lamina weight and lamina thickness had significantly lower means for kelps in the cool temperate regime compared to kelps in the warm temperate regime.
 When *L.
 pallida* morphological characteristics were grouped to wave exposure categories kruskal-wallis tests revealed significant differences amoungst categories for all morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
-The morphological characteristics that were significantly different amoungst sites exibited similar patterns, with increasing mean values from the full sheltered to extremely sheltered categories and then a decrease in mean values for the remainder of the categories.
-Two morphological characteristics, namely lamina thickness and stipe diameter, both exibit higher variations for the exposed categories.
+The morphological characteristics that were significantly different amoungst sites exibited similar patterns, with increasing mean values from the full sheltered to extremely sheltered categories and a decrease in mean values for the remainder of the categories.
+Two morphological characteristics, namely lamina thickness and stipe diameter, both exibit higher variations for the exposed categories compared to other morphological characteristics.
 
@@
 
-All deep *E.
-maxima* morphological characteristics showed differences between cool temperate and warm temperate regimes except for frond mass, frond length and stipe mass.
-Mean values for primary length, primary width and stipe circumference were all significantly higher for the cool temperate regime than the warm temperate regime.
-The remaining morphological characteristics which showed significant differences between regimes exhibited significanlty lower values for the cool temperate regime compared to the warm temperate regime.
+Six out of the ten deep *E.
+maxima* morphological characteristics showed differences between cool temperate and warm temperate regions except for frond mass, frond length, stipe mass and epiphyte length.
+Mean values for primary length, primary width and stipe circumference were significantly higher for kelps from cool temperate regions than for kelps from warm temperate regions.
+The remaining morphological characteristics, stipe length, number of tufts and total length had significanlty lower values for kelps from cool temperate compared to kelps from the warm temperate regime.
 Significant differences amoungst wave exposure categories were found for all morphological characteristics of deep *E.
 maxima* populations.
-Certian morphological characteristics showed a gradual increase in mean value with inncreasing wave exposure.
-The remaining morphological characteristics exhibited decreased values for only the higher exposure categories and then a sharp increase in value for the highest wave exposure category.
-In all cases, mean values for the highest wave exposure level (very exposed) was greatest compared to other levels of exposure.
+Certian morphological characteristics showed a gradual increase in value with increasing degree of wave exposure.
+The remaining morphological characteristics exhibited decreased values for only the higher wave exposure categories and then a sharp increase in value for the highest wave exposure category.
 
@@
   stat_compare_means(method = "kruskal.test") + 
+ ylim(0,30) +
   xlab("") + 
@@
 
-```{r shallow ecklonia temp plots, echo=FALSE, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by temperature regime.'}
+```{r shallow ecklonia temp plots, echo=FALSE, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E.
+maxima* grouped by temperature regime.',fig.height=8, fig.width=8, include=TRUE}
 ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
@@
 
-```{r echo=FALSE, warning=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by wave exposure.'}
-ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = FALSE, legend = "bottom")
-```
-
-```{r rda analysis lamanaria, eval=FALSE, include=FALSE}
-
+```{r echo=FALSE, warning=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E.
+maxima* grouped by wave exposure.',fig.height=8, fig.width=8, include=TRUE}
+ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom")
+```
+
+```{r rda analysis lamanaria, eval=FALSE, include=TRUE}
+#data processed in the temperature and wave sections will be used here eg. abio_clim
 # remove unnecessary columns
@@
   filter(site != "Roman Rock") %>% 
-  filter(site != "Miller's Point") %>%  #the scipt keeps dropping it when combing the site list with the metadata
   ungroup()
@@
 
-# merge temp and wave data to create environmental data-frame
-abio_clim <- merge(temp_ann, wave_ann, by = "site")
-
 #use site lam_site to align appropriate data
@@
 
-#remove date column and extra Betty's Bay
-abio_lam <- abio_lam[-c(5) , -c(2)]
+#remove date column 
+abio_lam <- abio_lam[ , -c(2)]
 
@@
 
-abio_lam <- cbind(lam_sites, abio_lam)
+#have an issue with Miller's Point not merging so removing for now. creating temp site list for this
+lam_sites_temp <- lam_sites[-c(9), ]
+abio_lam <- cbind(lam_sites_temp, abio_lam)
 
@@
   column_to_rownames(var = "site")
-
+#remove Millers from bio_lam to match up sites as I am still having issues
+bio_lam <- bio_lam[-c(9), ]
 #run RDA for abiotic variables
@@
 # Parsimonious RDA
-rda_lam <- rda(bio_lam ~ tp_sd + max_temp + hs_mean, data = abio_lam)
+rda_lam <- rda(bio_lam ~ tp_sd + hs_mean + mean_temp + dir_median, data = abio_lam)
 summary(rda_lam)
+anova.cca(rda_lam, step = 1000)
+anova.cca(rda_lam, step = 1000, by = "axis")
 
@@
 
-lam_meta_rda <- merge(lam_sites, kelp_sites, by = "site")
-lam_meta_rda <- lam_meta_rda[-c(1,9,11,14), ]
+meta_all <- read.csv("metadata/site_meta_complete.csv", sep = ";")
+meta_lam_rda <- merge(lam_sites, meta_all, by = "site")
+meta_lam_rda <- lam_meta_rda[-c(9), ] #remove Miller's point as it is dropping during processing
 
 # Plot RDA
-lam_rda <- ggord(rda_lam, lam_meta_rda$region, ellipse = TRUE, hull = FALSE, grp_title = "Region", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9) 
+lam_rda <- ggord(rda_lam, meta_lam_rda$region,ellipse = TRUE, hull = FALSE, grp_title = "Region", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9) 
 ```
@@
 
-```{r rda plots, eval=FALSE, include=FALSE}
-ggarrange(lam_rda, rda_eck_deep,rda_eck_shallow, common.legend = TRUE, labels = "AUTO", legend = "bottom")
+```{r rda plots, echo=FALSE, message=FALSE, warning=FALSE}
+lam_rda
 ```

### 2019-08-02 morphology_paper.Rmd
moved fromtier template. updated shallow data sheet as some data was not being read in. Completed RDAs and output with updated sites and accompying data

@@
 
-```{r software packages, message=TRUE, warning=FALSE, include=FALSE, echo=FALSE}
+```{r software packages and intial data prep, message=TRUE, warning=FALSE, include=FALSE, echo=FALSE}
+
 library(tidyverse)
@@
 library(vegan)
+
+# Ecklonia
+eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
+  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
+  filter(site!= "De Hoop")
+eck_data$site <- as.character(eck_data$site)
+
+# Laminaria
+lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>%
+  # NB: This may also not be an appropriate change
+  # Not using Rocky Bank in this study
+  filter(site != "Rocky Bank")
+lam_data$site <- as.character(lam_data$site)
+
+meta_all <- read.csv("metadata/site_meta_complete.csv", sep = ";")
+
 ```
@@
 # Complete lon/lats
-
-# Ecklonia
-eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
-  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
-  filter(site!= "De Hoop")
-eck_data$site <- as.character(eck_data$site)
-
-# Laminaria
-lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>%
-  # NB: This may also not be an appropriate change
-  # Not using Rocky Bank in this study
-  filter(site != "Rocky Bank")
-lam_data$site <- as.character(lam_data$site)
 
@@
 
-```{r wave power map, echo=FALSE, fig.cap='wave power maps'}
+```{r wave power map, eval=FALSE, fig.cap='wave power map', include=FALSE}
 library(png)
@@
 ```
+**insert wave exposure map here**
 
@@
 pallida* morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
-Certian morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significanlty higher means kelps in cool temperate regimes when compared to kelps for the warm temperate regime.
+Certian morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significanlty higher means for kelps in cool temperate regimes when compared to kelps for the warm temperate regime.
 Conversely, lamina weight and lamina thickness had significantly lower means for kelps in the cool temperate regime compared to kelps in the warm temperate regime.
@@
 
-```{r laminaria temp plots, echo=FALSE, warning=FALSE, fig.cap='Boxplots of *L.
-pallida* morphological characteristics group by temperature regime.', message=FALSE, warning=FALSE, include=FALSE, fig.height=8, fig.width=7}
+```{r laminaria temp plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime.', fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
 ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
@@
 
-```{r lamanaria wave plots, echo=FALSE,warning=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.', fig.height=8, fig.width=8}
+```{r lamanaria wave plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.', fig.height=7, fig.width=7, warning=FALSE}
 ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
@@
 Six out of the ten deep *E.
-maxima* morphological characteristics showed differences between cool temperate and warm temperate regions except for frond mass, frond length, stipe mass and epiphyte length.
+maxima* morphological characteristics showed differences between cool temperate and warm temperate regime except for frond mass, frond length, stipe mass and epiphyte length.
 Mean values for primary length, primary width and stipe circumference were significantly higher for kelps from cool temperate regions than for kelps from warm temperate regions.
@@
 
-```{r deep ecklonia temp plot, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by temperature.',fig.height=8, fig.width=8}
+```{r deep ecklonia temp plot, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E.
+maxima* grouped by temperature.', fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
 ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
@@
 
-```{r deep ecklonia wave plots, warning=FALSE,fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.',fig.height=8, fig.width=8}
+```{r deep ecklonia wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.', fig.height=7, fig.width=7, warning=FALSE}
 ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
@@
 
-```{r shallow ecklonia temp plots, echo=FALSE, message=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by temperature regime.',fig.height=8, fig.width=8, include=TRUE}
+```{r shallow ecklonia temp plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E.
+maxima* grouped by temperature regime.', fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
 ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
@@
 
-```{r echo=FALSE, warning=FALSE, warning=FALSE, fig.cap ='Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by wave exposure.',fig.height=8, fig.width=8, include=TRUE}
+```{r shallow rda wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by wave exposure.', fig.height=7, fig.width=7, warning=FALSE}
 ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom")
@@
 
-```{r rda analysis lamanaria, eval=FALSE, include=TRUE}
+```{r rda analysis lamanaria, eval=TRUE, include=FALSE, message=FALSE, warning=FALSE}
 #data processed in the temperature and wave sections will be used here eg. abio_clim
@@
 
-meta_all <- read.csv("metadata/site_meta_complete.csv", sep = ";")
-meta_lam_rda <- merge(lam_sites, meta_all, by = "site")
-meta_lam_rda <- lam_meta_rda[-c(9), ] #remove Miller's point as it is dropping during processing
+meta_lam_rda <- read.csv("metadata/site_meta_complete.csv", sep = ";")
+meta_lam_rda <- merge(lam_sites, meta_lam_rda, by = "site")
+meta_lam_rda <- meta_lam_rda[-c(9), ] #remove Miller's point as it is dropping during processing
 
@@
 eck_mean <- eck_mean %>% 
-  filter(site != "St Helena Bay") %>% 
   filter(site != "Miller's Point")
@@
 #remove date column and extra Betty's Bay row
-abio_eck <- abio_eck[-c(5) , -c(2)]
+abio_eck <- abio_eck[ , -c(2)]
 
@@
 
-#summary of rda output
-summary(rda_all_eck)
-
 # Forward selection using vegan's ordistep
@@
 # Parsimonious RDA
-rda_eck <- rda(bio_eck ~ tp_mean + hs_mean + tp_sd + hs_sd + mean_temp, data = abio_eck)
+rda_eck <- rda(bio_eck ~ tp_mean + hs_mean + tp_sd + min_temp + dir_median, data = abio_eck)
 summary(rda_eck)
+anova.cca(rda_eck, step = 1000)
+anova.cca(rda_eck, step = 1000, by = "axis")
 
@@
 
-eck_meta_rda <- merge(eck_sites, kelp_sites, by = "site")
+eck_meta_rda <- merge(eck_sites, meta_all, by = "site")
 eck_meta_rda <- eck_meta_rda %>% 
@@
 
-```{r rda shallow ecklonia processing, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+```{r rda shallow ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
 
@@
   filter(population == "Shallow") %>% 
-  mutate(site = replace(site, site == "Buffelsbaai", "Buffelsbaai South")) %>% 
-  mutate(site = replace(site, site == "Miller's Point C", "Miller's Point")) %>% 
-  mutate(site = replace(site, site == "Miller's Point B", "Baboon Rock")) %>% 
-  filter(site != "St James") %>% 
-  filter(site != "Slangkop") %>% 
-  filter(site != "Miller's Point A")
+  mutate(site = replace(site, site == "Miller's Point A", "Miller's Point")) %>% 
+  mutate(site = replace(site, site == "Bakoven", "Sea Point")) %>% 
+  filter(site != "Miller's Point B") %>% 
+  filter(site != "Miller's Point C") %>% 
+  filter(site != "Slangkop")
 
@@
 #use site eck_site to align appropriate data from abio_clim data from the lamanaria processing section
-abio_clim_temp <- abio_clim %>% 
-  mutate(site = replace(site, site == "Sea Point", "Bakoven"))
-
-abio_eck_sh <- merge(abio_clim, eck_sites_sh, by = "site")
+abio_clim_v2 <- abio_clim %>% 
+  mutate(site = replace(site, site == "Bakoven", "Sea Point")) %>% 
+  mutate(site = replace(site, site == "Bordjies Deep", "Bordjies"))
+
+abio_eck_sh <- merge(abio_clim_v2, eck_sites_sh, by = "site")
 
@@
 
-#summary of rda output
-summary(rda_all_ecksh)
-
 # Forward selection using vegan's ordistep
@@
 step.forward <- ordistep(rda(bio_eck_sh ~ 1, data = abio_eck_sh),
-                         scope=formula(rda_all_eck), direction = "forward", pstep = 1000)
+                         scope=formula(rda_all_ecksh), direction = "forward", pstep = 1000)
 #calculate variance inlfation factor
-#vif.cca(rda_all_eck)
+#vif.cca(rda_all_ecksh)
 
@@
 # Parsimonious RDA
-rda_eck_shallow <- rda(bio_eck_sh ~ hs_max + hs_mean + sd_temp + mean_temp, data = abio_eck_sh)
+rda_eck_shallow <- rda(bio_eck_sh ~ mean_temp + dirw_median + tp_max + max_temp, data = abio_eck_sh)
 summary(rda_eck_shallow)
+anova.cca(rda_eck_shallow, step = 1000)
+anova.cca(rda_eck_shallow, step = 1000, by = "axis")
 
@@
 
+meta_all$site <- as.character(meta_all$site)
+meta_shallow_rda <- meta_all %>% 
+  mutate(site = replace(site, site == "Bakoven", "Sea Point")) %>%  
+  mutate(site = replace(site, site == "Yzerfontein", "St Helena Bay")) %>% 
+  mutate(site = replace(site, site == "Bordjies Deep", "Bordjies"))
+meta_shallow_rda <- merge(meta_shallow_rda, eck_sites_sh, by = "site")
+
 # Plot RDA
-rda_eck_shallow <- ggord(rda_eck_shallow,ellipse = TRUE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9, coord_fix = TRUE)  # looking at the raw code, this is plotting the 'wa scores', the blue
-
-```
-
-```{r rda plots, echo=FALSE, message=FALSE, warning=FALSE}
-lam_rda
+rda_eck_shallow <- ggord(rda_eck_shallow,meta_shallow_rda$region,ellipse = TRUE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9, coord_fix = TRUE)  # looking at the raw code, this is plotting the 'wa scores', the blue
+
+```
+
+### Abiotic drivers of *L. pallida* morphological characteristics
+
+Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of *L.
+pallida* morphological characteristics with an adjusted *R*2 of 0.73, explaining 85% of the variation (global permutation test on final model: d.f = 4, *F* = 7.18, *p* = 0.001).
+The model consisted of two significant axes: RDA1 and RDA2 explained 66% and 24% of the variation, respectively.
+bla bla
+
+### Abiotic drivers of *E. maxima* morphological characteristics
+
+Forward selection, assessment of VIF and pearson correlation coefficients retained *E.
+maxima* morphological characteristics with an adjusted *R*2 of 0.77, explaining 92% of the variation (global permutation test on final model: d.f = 8, *F* = 6.08, *p* = 0.001).
+bla bla
+
+Forward selection, assessment of VIF and pearson correlation coefficients retained *E.
+maxima* morphological characteristics with an adjusted *R*2 of 0.36, explaining 66% of the variation (global permutation test on final model: d.f = 3, *F* = 2.54, *p* = 0.027).
+bla bla bla
+
+```{r rda plots, echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
+ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE)
 ```

### 2019-08-04 morphology_paper.Rmd
new sections added in results for redundancy analysis and html output updated

@@
   word_document:
-    reference_docx: morphology_paper_files/frontiers_template.docx
+    reference_docx: frontiers_template.docx
   html_notebook:
@@
 
-```{r echo=FALSE, fig.cap='Schematic of *E. maxima* on the left and *L. pallida* on the right. (Dyer 2018)'}
+```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology.'}
 library(png)
@@
 # The patched Roman Rock temperatures
-load("data/rr_patch.Rdata")
+load("data/temp_data/rr_patch.Rdata")
 
@@
 
+```{r wave rose figure, echo=FALSE, eval= FALSE, fig.cap='Wave rose figure'}
+
+img_wave_rose <- readPNG("figures/wave_rose.png")
+grid.raster(wave rose)
+```
+**insert wave roses here**
 Annual mean wave peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St.
@@
 
-```{r wave power map, eval=FALSE, fig.cap='wave power map', include=FALSE}
+```{r wave power map, echo=FALSE, eval= FALSE, fig.cap='Wave power map'}
+
 library(png)
@@
 ```
-**insert wave exposure map here**
-
+**insert wave power map here**
 In Figure 5 the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
@@
 
-```{r rda analysis lamanaria, eval=TRUE, include=FALSE, message=FALSE, warning=FALSE}
+```{r rda analysis lamanaria, message=FALSE, warning=FALSE, include=FALSE}
 #data processed in the temperature and wave sections will be used here eg. abio_clim
@@
 
-```{r rda deep ecklonia processing, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+### Abiotic drivers of *L. pallida* morphological characteristics
+
+Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of *L.
+pallida* morphological characteristics with an adjusted *R*2 of 0.49, explaining 72% of the variation (global permutation test on final model: d.f = 1, *F* = 8.51, *p* = 0.008).
+RDA1 was the only significant axis in the model and explained 48% of the variation while RDA2 was not significant and explained 63% of the variation.
+Total length, stipe length, stipe diameter and stipe mass were postivley influenced (i.e., increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* and negativley influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp*.
+Study sites from the eastern region clustred close together while no clear clustering was evident for the western region.
+
+```{r rda deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 ```
+
+### Abiotic drivers of *E. maxima* morphological characteristics
+
+Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of deep *E.
+maxima* morphological characteristics with an adjusted *R*2 of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, *F* = 2.4, *p* = 0.026).
+RDA1 was the only significant axis in the model (p = 0.03) and explained 72% of the variation.
+The non-significant axis for the model explained 17% of the variation.
+Primary length, primary width and stipe circumference were positivley influenced by *hs_mean*, *tp_mean* and *dir_median* for for kelps from the western region.
+The remaining morphological characteristics were postivley inlfuenced by *min_temp* and negatively influenced by *tp_sd* for kelps ffrom the eastern region.
 
@@
 ```
-
-### Abiotic drivers of *L. pallida* morphological characteristics
-
-Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of *L.
-pallida* morphological characteristics with an adjusted *R*2 of 0.73, explaining 85% of the variation (global permutation test on final model: d.f = 4, *F* = 7.18, *p* = 0.001).
-The model consisted of two significant axes: RDA1 and RDA2 explained 66% and 24% of the variation, respectively.
-bla bla
-
-### Abiotic drivers of *E. maxima* morphological characteristics
-
-Forward selection, assessment of VIF and pearson correlation coefficients retained *E.
-maxima* morphological characteristics with an adjusted *R*2 of 0.77, explaining 92% of the variation (global permutation test on final model: d.f = 8, *F* = 6.08, *p* = 0.001).
-bla bla
 

### 2019-08-04 morphology_paper.Rmd
results section for RDA analysis updated. word output updated.

@@
 output:
-  html_document:
-    df_print: paged
+  word_document:
+    reference_docx: frontiers_template.docx
   pdf_document:
@@
     fig_width: 7
-  word_document:
-    reference_docx: frontiers_template.docx
+  html_document:
+    df_print: paged
   html_notebook:
@@
 
-Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of *L.
+Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of *L.
 pallida* morphological characteristics with an adjusted *R*2 of 0.49, explaining 72% of the variation (global permutation test on final model: d.f = 1, *F* = 8.51, *p* = 0.008).
 RDA1 was the only significant axis in the model and explained 48% of the variation while RDA2 was not significant and explained 63% of the variation.
-Total length, stipe length, stipe diameter and stipe mass were postivley influenced (i.e., increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* and negativley influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp*.
-Study sites from the eastern region clustred close together while no clear clustering was evident for the western region.
+Total length, stipe length, stipe diameter and stipe mass were postivley influenced (i.e., increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* for kelps mostly from the western region and negativley influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp* for kelps from the eastern region.
+Although sites from from the western region did not cluster as closely as eastern region study sites, grouping according to region was still evident.
 
@@
 Primary length, primary width and stipe circumference were positivley influenced by *hs_mean*, *tp_mean* and *dir_median* for for kelps from the western region.
-The remaining morphological characteristics were postivley inlfuenced by *min_temp* and negatively influenced by *tp_sd* for kelps ffrom the eastern region.
+The remaining morphological characteristics were postivley inlfuenced by *min_temp* and negatively influenced by *tp_sd* for kelps from the eastern region.
 
@@
 
-Forward selection, assessment of VIF and pearson correlation coefficients retained *E.
-maxima* morphological characteristics with an adjusted *R*2 of 0.36, explaining 66% of the variation (global permutation test on final model: d.f = 3, *F* = 2.54, *p* = 0.027).
-bla bla bla
+Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retian the most parsimonious descriptors of *E.
+maxima* morphological characteristics with an adjusted *R*2 of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, *F* = 2.41, *p* = 0.013).
+The model consisted of one significant axis, RDA1 (p = 0.02) and explained 62% of the variation.
+Stipe length, stipe mass and stipe circumference were positivley influenced by *max_temp* and *mean_temp* along RDA1, and explained 50% and 55% of the variation respectivley.
+The remaining morphological characteristics were positivley influenced by by *tp_max* and negativley inlfuenced by *dirw_median* along RDA2 and explained 86% and 94% of the variation respectivley.
+There was overlapping of clusters according to region with no clear patterns in sites evident.
 

### 2019-08-04 morphology_paper.Rmd
script for additional graph added to temperature plot

@@
 ---
-title: Temperature and wave exposure as drivers of morphology in two kelp species
-  along the coast of South Africa
+title: Using numerical modelling to investigate temperature and wave exposure as drivers of macroalgal characteristics
 author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
@@
 output:
-  word_document:
-    reference_docx: frontiers_template.docx
+  html_document:
+    df_print: paged
   pdf_document:
@@
     fig_width: 7
-  html_document:
-    df_print: paged
+  word_document:
+    reference_docx: frontiers_template.docx
   html_notebook:
@@
 ## Abiotic environment
+### Temperature
 
@@
   coord_flip()
+
+temp_ann_meta <- merge(meta_all, temp_ann, by = "site")
+temp_ann_meta$site <- as.character(temp_ann_meta$site)
+
+temp_boxplot <- ggplot(data = temp_ann_meta, aes(x = region, y = mean_temp, fill  = region)) + 
+  geom_boxplot() + 
+  scale_x_discrete(limit = c("Western", "Eastern")) + 
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("")
+
 ```
@@
 Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.'}
-temp_plot
-```
+ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 2, nrow = 1, labels = c("AUTO"))
+```
+
+### Wave environment
 
@@
 
-```{r laminaria temp plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime.', fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
+```{r laminaria temp plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime.', fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
 ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
@@
 
-```{r lamanaria wave plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.', fig.height=7, fig.width=7, warning=FALSE}
+```{r lamanaria wave plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.', fig.height=6, fig.width=6, warning=FALSE}
 ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
@@
 ```{r deep ecklonia temp plot, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E.
-maxima* grouped by temperature.', fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
+maxima* grouped by temperature.', fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
 ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
@@
 
-```{r deep ecklonia wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.', fig.height=7, fig.width=7, warning=FALSE}
+```{r deep ecklonia wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.', fig.height=6, fig.width=6, warning=FALSE}
 ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
@@
 ```{r shallow ecklonia temp plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by temperature regime.', fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
+maxima* grouped by temperature regime.', fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
 ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
@@
 
-```{r shallow rda wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by wave exposure.', fig.height=7, fig.width=7, warning=FALSE}
+```{r shallow rda wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by wave exposure.', fig.height=6, fig.width=6, warning=FALSE}
 ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom")
@@
 
-Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retian the most parsimonious descriptors of *E.
+```{r rda plots, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
+ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE)
+```
+
+Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retian the most parsimonious descriptors of shallow *E.
 maxima* morphological characteristics with an adjusted *R*2 of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, *F* = 2.41, *p* = 0.013).
@@
 There was overlapping of clusters according to region with no clear patterns in sites evident.
-
-```{r rda plots, echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
-ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE)
-```
-
 # Discussion

### 2019-08-04 morphology_paper.Rmd
script for extra graph created

@@
   scale_x_discrete(limit = c("Western", "Eastern")) +
+  scale_fill_discrete(name="Region") +
   stat_compare_means() + 
@@
   coord_flip()
+
+wave_ann_meta <- merge(meta_all, wave_ann, by = "site")
+wave_ann_meta$site <- as.character(wave_ann_meta$site)
+
+wave_boxplot_1 <- ggplot(data = wave_ann_meta, aes(x = region, y = hs_mean, fill  = region)) + 
+  geom_boxplot() + 
+  scale_x_discrete(limit = c("Western", "Eastern")) + 
+  scale_fill_discrete(name="Region") +
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("")
+
+wave_boxplot_2 <- ggplot(data = wave_ann_meta, aes(x = region, y = tp_mean, fill  = region)) + 
+  geom_boxplot() + 
+  scale_x_discrete(limit = c("Western", "Eastern")) +
+  scale_fill_discrete(name="Region") +
+  stat_compare_means() + 
+  xlab("") + 
+  ylab("")
 ```
@@
 Mean variables are represented by dots and standard deviation by whiskers.'}
-ggarrange(wave_plot_A, wave_plot_B, ncol = 2, nrow = 1)
+ggarrange(wave_plot_A,wave_boxplot_1, wave_plot_B, wave_boxplot_2, common.legend = TRUE, labels = c("AUTO"))
 ```

### 2019-08-04 morphology_paper.Rmd
methods section updated and added to

@@
 
-Site-specific summary statistics were calculated for each of the environmental variables considered in this study.
-The summary statistics calculated for temperature, waves and wind variables were minimum, maximum, mean, range and standard deviation for annual, February (Austral summer) and August (Austral winter) timescales, respectively.
-In addition, annual and seasonal median wave and wind direction were also calculated; all summary data and their respective abbreviation are presented in Table 3.
+Summary statistics were calculated for each site and regions for each of the environmental variables considered in this study.
+The summary statistics calculated for temperature, waves and wind variables were minimum, maximum, mean, range and standard deviation for annual timescales.
+In addition, annual median wave and wind direction were also calculated; all summary data and their respective abbreviation are presented in Table 3.
 The median was calculated for wind and wave direction, as issues arise when calculating the mean and standard deviation for compass metrics.
@@
 
-*insert paragraph about morphology*
-Characterising and comparison of morphological characteristics between deep and shallow *E.
-maxima* was achieved by notched boxplots.
-Distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package (Oksanen et al.
+Characterisation of morphological characteristics in relation to temperature and wave exposure non-parametric statistical methods were used.
+Morphological characteristics were categorised according to temperature regime and wave exposure category for each each species and population of kelp in this study.
+Significant differences between temperatre regime and wave exposure categories were tested with wilcox tests.
+Notched boxplots allowed comparison between individual categories of wave exposure.
+In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package (Oksanen et al.
 2016) in R (R Core Team, 2019).
-The role of temperature, waves and wind as drivers of kelp morphology was investigated.
 The abiotic data covered all study sites for both species of kelp.

### 2019-08-04 morphology_paper.Rmd
general house cleaning. updated all output types

@@
 output:
-  html_document:
-    df_print: paged
   pdf_document:
@@
     fig_width: 7
+  html_document:
+    df_print: paged
   word_document:
@@
 
-```{r Temperature variables plot, echo=FALSE, warning=FALSE, fig.cap='Temperature variables at the collection sites around the Cape Peninsula.
+```{r Temperature variables plot, echo=FALSE, warning=FALSE,fig.height=7,fig.width=7, fig.cap='Temperature variables at the collection sites around the Cape Peninsula.
 Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.'}
@@
 #Annual wave plots
+#combine with meta data for easier plotting
+wave_ann_meta <- merge(meta_all, wave_ann, by = "site")
+wave_ann_meta$site <- as.character(wave_ann_meta$site)#make character to ensure no issues arise for site and region names
 
@@
 
-wave_ann_meta <- merge(meta_all, wave_ann, by = "site")
-wave_ann_meta$site <- as.character(wave_ann_meta$site)
-
 wave_boxplot_1 <- ggplot(data = wave_ann_meta, aes(x = region, y = hs_mean, fill  = region)) + 
@@
 
-```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE, fig.cap='Wave variables at each collection site around the Western Cape coast.
+```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE,fig.height=7,fig.width=7, fig.cap='Wave variables at each collection site around the Western Cape coast.
 Mean variables are represented by dots and standard deviation by whiskers.'}

### 2019-08-05 morphology_paper.Rmd
added to discussion. general house cleaning of script. and updated word output

@@
 output:
+  word_document:
+    reference_docx: frontiers_template.docx
+  html_document:
+    df_print: paged
   pdf_document:
@@
     fig_width: 7
-  html_document:
-    df_print: paged
-  word_document:
-    reference_docx: frontiers_template.docx
   html_notebook:
@@
   geom_boxplot() + 
-  scale_x_discrete(limit = c("Western", "Eastern")) +
-  scale_fill_discrete(name="Region") +
-  stat_compare_means() + 
+  scale_fill_discrete(breaks=c("Western","Eastern"), name = "Region") +
   xlab("") + 
   ylab("") 
-
-```
-
-```{r Temperature variables plot, echo=FALSE, warning=FALSE,fig.height=7,fig.width=7, fig.cap='Temperature variables at the collection sites around the Cape Peninsula.
+```
+
+```{r Temperature variables plot, echo=FALSE, warning=FALSE,fig.height=5,fig.width=6, fig.cap='Temperature variables at the collection sites around the Cape Peninsula.
 Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.'}
-ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 2, nrow = 1, labels = c("AUTO"))
+ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 1, nrow = 2, labels = c("AUTO"))
 ```
@@
 There was overlapping of clusters according to region with no clear patterns in sites evident.
+
 # Discussion
@@
 The complex geomorphology of the western Cape coastline creates an ideal natural laboratory for studies of kelps in interaction with their environment.
-This complex coastal environment created by the coastal wind and wave climate creates an ideal opportunity to explore the effects of waves and the thermal regime on the morphological development of kelp.
 The annual and seasonal patterns in both wind influence the thermal regime around the coastline by inducing upwelling.
@@
 Although this study did not measure force of attachment, other morphological responses to wave variables were evident.
+Higher temperatures tended to be assocaited with reduced measurements of stipe mass, stipe circumference/diameter and stipe length for the sub-canopy species *L.
+pallida*, which suggests that there is a reduction in overall surface area of the plant in warm temperate environments.
+The ecological consequence of this is unclear...The canopy species *E.
+maxima* exibits increased stipe length and decreased stipe circumference in warm temperate areas for both deep and shallow populations, which suggests a 'slender kelp' tactic and has been shown in previous work.**add examples here**.
+The response of kelp morphological characteristics to wave exposure was evident and both species show a tactics based on the magnitde of wave exposure.
+Strength increasing traits were exibited for lower exposure levels while a 'go-with-the-flow' tactic for moderate levels of wave exposure.
+Denny and Galord ???
+showed that increased flexiblity or a 'go-with-the-flow' tactic has a threshold value, where increased stipe length actaully increase drag under high wave exposure scenarios.
+Therefore, under high wave exposure scenarios size reducing traits, such as reduced surface area of the overall plant should be favoured instead.
+This was evident in the results which showed that stipe length decreases under high wave exposure categories.
+This suggests that how morphological characteristcs manifest themselves is dependent on the magnitude of wave exposure.
+Furthermore, the results suggest that in the absence of wave exposure, temperatre becomes the dominant morphological driver.
+This is evident from the redundancy analysis for each species and deep populations of *E.
+maxima* which show clustering of study sites according to the region; wave variables drive morphological characteristics for kelps from the western region and temperature variables for kelps from the eastern region.
+Clustering of sites according to region overlapped almost completey for the RDA analysis of shallow *E.
+maxima* suggesting that for shallow kelp the effect of wave exposure is less variable.
 

### 2019-08-06 morphology_paper.Rmd
site morphology draft for deep ecklonia added. outputs updated

@@
 Patterns are also evident in the wave data.
-Seasonal variations in significant wave height (Hs), wave period and wave direction are also present.
+Seasonal variations in significant wave height (Hs), peak period and wave direction are also present.
 The direction of the dominant swell changes to the south west in winter, generated by strong low pressures that originate from the southern ocean (Reason et al.
@@
 temp_plot <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
-  geom_point() + 
+  geom_point(show.legend = FALSE) + 
   scale_x_discrete(limits = positions_temp) + 
@@
   geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) + 
-  geom_point(data = temp_ann, aes(x = site, y = max_temp, colour = "blue")) + 
-  geom_point(data = temp_ann, aes(x = site, y = min_temp, colour = "pink")) + 
+  geom_point(data = temp_ann, aes(x = site, y = max_temp, colour = "blue"),show.legend = FALSE) + 
+  geom_point(data = temp_ann, aes(x = site, y = min_temp, colour = "pink"),show.legend = FALSE) + 
   xlab("") + 
@@
   geom_boxplot() + 
-  scale_fill_discrete(breaks=c("Western","Eastern"), name = "Region") +
-  xlab("") + 
-  ylab("") 
+  stat_compare_means(label.y = 13) +
+  xlab("") + 
+  ylab("") +
+  coord_flip()
 ```
@@
 Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.'}
-ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 1, nrow = 2, labels = c("AUTO"))
+ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 1, nrow = 2, labels = c("AUTO"), align = c("h"), legend = c("bottom"))
 ```
@@
 Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay.
-The data shows no trend in mean peak wave period for the coastline.
-Annual peak wave period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and summer respectively, Miller’s point and had the lowest peak wave period amongst all sites.
+The data shows no trend in mean peak peak period for the coastline.
+Annual peak period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and summer respectively, Miller’s point and had the lowest peak period amongst all sites.
 These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side of the peninsula, with Miller’s point showing the highest variation across timescales.
-Annual maximum peak wave period ranged from 18.9 s at St.
+Annual maximum peak period ranged from 18.9 s at St.
 Helena Bay to 18.4 s at Olifantsbos for the western side of the peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty’s Bay on the eastern side of the peninsula.
@@
   geom_boxplot() + 
-  scale_x_discrete(limit = c("Western", "Eastern")) + 
   scale_fill_discrete(name="Region") +
-  stat_compare_means() + 
-  xlab("") + 
-  ylab("")
+  stat_compare_means(label.y =1,label.x = 1.5) + 
+  xlab("") + 
+  ylab("") + 
+  coord_flip()
 
@@
   geom_boxplot() + 
-  scale_x_discrete(limit = c("Western", "Eastern")) +
+  stat_compare_means(label.y = 9, label.x = 1.5) + 
   scale_fill_discrete(name="Region") +
-  stat_compare_means() + 
-  xlab("") + 
-  ylab("")
+  xlab("") + 
+  ylab("") +
+  ylim(8,12) + 
+  coord_flip()
+
 ```
@@
 Mean variables are represented by dots and standard deviation by whiskers.'}
-ggarrange(wave_plot_A,wave_boxplot_1, wave_plot_B, wave_boxplot_2, common.legend = TRUE, labels = c("AUTO"))
+ggarrange(wave_plot_A,wave_boxplot_1, wave_plot_B, wave_boxplot_2, common.legend = TRUE, labels = c("A1", "A2", "B1", "B2"))
 ```
@@
 4].
-Annual peak wave period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and summer respectively, Miller’s point and had the lowest peak wave period amongst all sites.
+Annual peak peak period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and summer respectively, Miller’s point and had the lowest peak peak period amongst all sites.
 These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side of the peninsula, with Miller’s point showing the highest variation across timescales.
-Annual maximum peak wave period ranged from 18.9 s at St.
+Annual maximum peak period ranged from 18.9 s at St.
 Helena Bay to 18.4 s at Olifantsbos for the western side of the peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty’s Bay on the eastern side of the peninsula.
@@
 
-Significant differences were found between cool temperate and warm temperate regimes for all *L.
+Significant differences were found between cool temperate and warm temperate regimes for most *L.
 pallida* morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
@@
 When *L.
-pallida* morphological characteristics were grouped to wave exposure categories kruskal-wallis tests revealed significant differences amoungst categories for all morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
+pallida* morphological characteristics were grouped to wave exposure categories kruskal-wallis tests revealed significant differences amoungst categories for most morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
 The morphological characteristics that were significantly different amoungst sites exibited similar patterns, with increasing mean values from the full sheltered to extremely sheltered categories and a decrease in mean values for the remainder of the categories.
@@
 
-```{r rda plots, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
-ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE)
+```{r rda plots, echo=FALSE, fig.height=9, fig.width=9, message=FALSE, warning=FALSE}
+ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE, ncol = 3, nrow = 1, legend = c("bottom"))
 ```
@@
 Thank you to AS and CR for their guidance, and the honours class of 2014-2015 that aided in fieldwork. Carla Louise for making the GIS plot etc.
+
+# Appendix
+
+```{r appendix figure 1, echo=FALSE, message=FALSE, warning=FALSE}
+positionsLam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point ", "Baboon Rock", 
+                  "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Bordjies Deep","Olifantsbos", "Kommetjie", "Sea Point")
+eck_all_A <- ggplot(data = lam_data, aes(x = site, y = laminaWeight, fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsLam) + 
+           ylab("Lamina Weight (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") +
+        theme(axis.title.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+eck_all_B <- ggplot(data = lam_data, aes(x = site, y = laminaLength, fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsLam) + 
+           ylab("Lamina Length (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") +
+        theme(axis.title.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+eck_all_C <- ggplot(data = lam_data, aes(x = site, y = laminaThickness, fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsLam) + 
+           ylab("Lamina Thickness (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") +
+        theme(axis.title.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+eck_all_D <- ggplot(data = lam_data, aes(x = site, y = stipeMass, fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsLam) + 
+           ylab("Stipe Mass (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") +
+        theme(axis.title.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+eck_all_E <- ggplot(data = lam_data, aes(x = site, y = stipeLength, fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsLam) + 
+           ylab("Stipe Length (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") +
+        theme(axis.title.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+eck_all_F <- ggplot(data = lam_data, aes(x = site, y = stipeDiameter, fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsLam) + 
+           ylab("Stipe Diameter (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") +
+        theme(axis.title.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+eck_all_G <- ggplot(data = lam_data, aes(x = site, y = digits, fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsLam) + 
+           ylab("No. of Digits") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") +
+        theme(axis.title.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank()) 
+eck_all_H <- ggplot(data = lam_data, aes(x = site, y = thallusMass, fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsLam) + 
+           ylab("Thallus Mass (g)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") +
+        theme(axis.title.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+eck_all_I <- ggplot(data = lam_data, aes(x = site, y = totalLength, fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsLam) + 
+           ylab("Total length (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") +
+        theme(axis.title.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank())
+
+ggarrange(eck_all_A,eck_all_B,eck_all_C,eck_all_D,eck_all_E,eck_all_F,eck_all_G,eck_all_H,eck_all_I, ncol = 3, nrow = 3, common.legend = TRUE)
+
+```

### 2019-08-06 morphology_paper.Rmd
next appendix figure added, code changed slighlty. Kept standard theme instead

@@
                   "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Bordjies Deep","Olifantsbos", "Kommetjie", "Sea Point")
-eck_all_A <- ggplot(data = lam_data, aes(x = site, y = laminaWeight, fill = wave_exposure)) +
+lam_all_A <- ggplot(data = lam_data, aes(x = site, y = laminaWeight, fill = wave_exposure)) +
            geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
@@
            guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") +
-        theme(axis.title.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
-eck_all_B <- ggplot(data = lam_data, aes(x = site, y = laminaLength, fill = wave_exposure)) +
+           xlab("")
+lam_all_B <- ggplot(data = lam_data, aes(x = site, y = laminaLength, fill = wave_exposure)) +
            geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
@@
            guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") +
-        theme(axis.title.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
-eck_all_C <- ggplot(data = lam_data, aes(x = site, y = laminaThickness, fill = wave_exposure)) +
+           xlab("")
+lam_all_C <- ggplot(data = lam_data, aes(x = site, y = laminaThickness, fill = wave_exposure)) +
            geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
@@
            guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") +
-        theme(axis.title.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
-eck_all_D <- ggplot(data = lam_data, aes(x = site, y = stipeMass, fill = wave_exposure)) +
+           xlab("") 
+lam_all_D <- ggplot(data = lam_data, aes(x = site, y = stipeMass, fill = wave_exposure)) +
            geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
@@
            guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") +
-        theme(axis.title.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
-eck_all_E <- ggplot(data = lam_data, aes(x = site, y = stipeLength, fill = wave_exposure)) +
+           xlab("") 
+lam_all_E <- ggplot(data = lam_data, aes(x = site, y = stipeLength, fill = wave_exposure)) +
            geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
@@
            guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") +
-        theme(axis.title.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
-eck_all_F <- ggplot(data = lam_data, aes(x = site, y = stipeDiameter, fill = wave_exposure)) +
+           xlab("") 
+lam_all_F <- ggplot(data = lam_data, aes(x = site, y = stipeDiameter, fill = wave_exposure)) +
            geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
@@
            guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") +
-        theme(axis.title.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
-eck_all_G <- ggplot(data = lam_data, aes(x = site, y = digits, fill = wave_exposure)) +
+           xlab("")
+lam_all_G <- ggplot(data = lam_data, aes(x = site, y = digits, fill = wave_exposure)) +
            geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
@@
            guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") +
-        theme(axis.title.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
-eck_all_H <- ggplot(data = lam_data, aes(x = site, y = thallusMass, fill = wave_exposure)) +
+           xlab("") 
+lam_all_H <- ggplot(data = lam_data, aes(x = site, y = thallusMass, fill = wave_exposure)) +
            geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
@@
            guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") +
-        theme(axis.title.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
-eck_all_I <- ggplot(data = lam_data, aes(x = site, y = totalLength, fill = wave_exposure)) +
+           xlab("") 
+lam_all_I <- ggplot(data = lam_data, aes(x = site, y = totalLength, fill = wave_exposure)) +
            geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
@@
            guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") +
-        theme(axis.title.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+           xlab("")
+
+ggarrange(lam_all_A,lam_all_B,lam_all_C,lam_all_D,lam_all_E,lam_all_F,lam_all_G,lam_all_H,lam_all_I, ncol = 3, nrow = 3, common.legend = TRUE)
+
+```
+
+```{r appendix figure 2, echo=FALSE, message=FALSE, warning=FALSE}
+# Order site names by location west to east
+
+positionsEck <- c("Betty's Bay", "Kalk Bay", "A-Frame", 
+                  "Miller's Point", "Baboon Rock", "Batsata Rock", 
+                  "Buffels Bay", "Scarborough", 
+                  "Soetwater","Kommetjie","Hout Bay", 
+                  "Oudekraal", "St. Helena Bay")
+
+eck_all_A <- ggplot(data = eck_data, aes(x = site, y = frond_mass, fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Frond mass (g)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") 
+eck_all_B <- ggplot(data = eck_data, aes(x = site, y = primary_length , fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Primary Length  (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") 
+eck_all_C <- ggplot(data = eck_data, aes(x = site, y = primary_width , fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Primary Width (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") 
+eck_all_D <- ggplot(data = eck_data, aes(x = site, y = frond_length , fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Frond length (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") 
+eck_all_E <- ggplot(data = eck_data, aes(x = site, y = stipe_mass , fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Stipe mass (g)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") 
+eck_all_F <- ggplot(data = eck_data, aes(x = site, y = stipe_length , fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Stipe length (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") 
+eck_all_G <- ggplot(data = eck_data, aes(x = site, y = stipe_circ , fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Stipe circ (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") 
+eck_all_H <- ggplot(data = eck_data, aes(x = site, y = tufts , fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("No. of tufts") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("") 
+eck_all_I <- ggplot(data = eck_data, aes(x = site, y = epi_length , fill = wave_exposure)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Epi Length  (cm)") + 
+           scale_fill_brewer(palette="OrRd") +
+           coord_flip() +
+           guides(fill=guide_legend(title="Wave Exposure")) +
+           xlab("")
 
 ggarrange(eck_all_A,eck_all_B,eck_all_C,eck_all_D,eck_all_E,eck_all_F,eck_all_G,eck_all_H,eck_all_I, ncol = 3, nrow = 3, common.legend = TRUE)
-
-```
+```

### 2019-08-13 morphology_paper.Rmd
map added and word output updated
updated abiotic figures. theme_bw() will be used from now on. Extra details to figure added. All following graphs will be updated accordingly.

@@
     reference_docx: frontiers_template.docx
+  pdf_document: 
   html_document:
     df_print: paged
-  pdf_document:
-    fig_caption: yes
-    fig_crop: no
-    fig_height: 7
-    fig_width: 7
   html_notebook:
@@
 
+```{r map study sites, echo=FALSE, message=FALSE, warning=FALSE}
+library(png)
+library(grid)
+
+map <- readPNG("figures/sa_map.png")
+grid.raster(map)
+```
+
 The topography and elevation along the Cape Peninsula channel and shield winds along False Bay.
@@
 ```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology.'}
-library(png)
-library(grid)
 
@@
   theme(legend.position = "none") + 
-  coord_flip()
+  coord_flip() + 
+  theme_bw()
 
@@
   geom_boxplot() + 
-  stat_compare_means(label.y = 13) +
+  stat_compare_means(label.y = 14.5 ,label.x = 1.5, label = "p.format") +
   xlab("") + 
   ylab("") +
-  coord_flip()
-```
-
-```{r Temperature variables plot, echo=FALSE, warning=FALSE,fig.height=5,fig.width=6, fig.cap='Temperature variables at the collection sites around the Cape Peninsula.
-Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.'}
-ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 1, nrow = 2, labels = c("AUTO"), align = c("h"), legend = c("bottom"))
+  coord_flip() + 
+  theme_bw()
+
+```
+
+```{r Temperature variables plot, echo=FALSE, warning=FALSE,fig.height=5,fig.width=8, fig.cap='Temperature variables at the collection sites around the Cape Peninsula are represented in plot A.
+Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.
+Summary of temperature data for sites grouped by region are represented in plot B.
+See appendix for significance level guide.'}
+ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 2, nrow = 1, labels = c("AUTO"), align = c("h"), legend = FALSE)
 ```
@@
 
-```{r wave variable processing, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+```{r wave variable processing, message=FALSE, warning=FALSE, include=FALSE}
 ### Plots for wave variables
@@
 positions_wave <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
-
-wave_plot_A <- ggplot(data = wave_ann, aes(x = site, y = hs_mean)) +
-  geom_point() + 
+cols <- c("Betty's Bay" = "red", "Roman Rock"= "red","Kalk Bay"= "red", "A-Frame"= "red", "Miller's Point"= "red", "Baboon Rock"= "red",  "Batsata Rock"= "red","Bordjies Deep"= "red","Buffelsbaai"= "red","Buffelsbaai South"= "red","Olifantsbos"= "blue","Scarborough"= "blue", "Soetwater"= "blue", "Kommetjie"= "blue","Hout Bay"= "blue","Oudekraal"= "blue","Sea Point"= "blue","St Helena Bay"= "blue")
+
+wave_plot_A <- ggplot(data = abio_clim, aes(x = site, y = hs_mean)) +
+  geom_point(aes(colour = factor(site)), show.legend = FALSE) + 
   scale_x_discrete(limits = positions_wave) + 
+  scale_colour_manual(values = cols) +
   theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
@@
   theme(plot.title = element_text(hjust = 0)) + 
-  coord_flip()
-
-wave_plot_B <- ggplot(data = wave_ann, aes(x = site, y = tp_mean)) +
-  geom_point() + 
+  coord_flip() + 
+  theme_bw()
+
+wave_plot_B <- ggplot(data = abio_clim, aes(x = site, y = tp_mean)) +
+  geom_point(aes(colour = factor(site)), show.legend = FALSE) + 
   scale_x_discrete(limits = positions_wave) + 
+  scale_colour_manual(values = cols) +
   theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
@@
   theme(plot.title = element_text(hjust = 0)) + 
-  coord_flip()
+  coord_flip() +
+  theme_bw()
 
 wave_boxplot_1 <- ggplot(data = wave_ann_meta, aes(x = region, y = hs_mean, fill  = region)) + 
-  geom_boxplot() + 
-  scale_fill_discrete(name="Region") +
-  stat_compare_means(label.y =1,label.x = 1.5) + 
+  geom_boxplot(show.legend = FALSE) + 
+  stat_compare_means(label.y =1.4,label.x = 1.5, label = "p.format") + 
   xlab("") + 
   ylab("") + 
-  coord_flip()
+  coord_flip() + 
+  theme_bw()
 
 wave_boxplot_2 <- ggplot(data = wave_ann_meta, aes(x = region, y = tp_mean, fill  = region)) + 
-  geom_boxplot() + 
-  stat_compare_means(label.y = 9, label.x = 1.5) + 
-  scale_fill_discrete(name="Region") +
+  geom_boxplot(show.legend = FALSE) + 
+  stat_compare_means(label.y = 10, label.x = 1.5, label = "p.format") + 
   xlab("") + 
@@
   ylim(8,12) + 
-  coord_flip()
-
-```
-
-```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE,fig.height=7,fig.width=7, fig.cap='Wave variables at each collection site around the Western Cape coast.
-Mean variables are represented by dots and standard deviation by whiskers.'}
+  coord_flip() + 
+  theme_bw()
+
+```
+
+```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE,fig.height=7,fig.width=8, fig.cap='In plot A1 and A2 wave variables at each collection site around the Western Cape coast are presented.
+Mean variables are represented by dots and standard deviation by whiskers.
+Blue dots represent sites grouped to the western category and red dots represent sites grouped to the eastern category.'}
 ggarrange(wave_plot_A,wave_boxplot_1, wave_plot_B, wave_boxplot_2, common.legend = TRUE, labels = c("A1", "A2", "B1", "B2"))
@@
 
-```{r wave rose figure, echo=FALSE, eval= FALSE, fig.cap='Wave rose figure'}
+```{r wave rose figure, echo=FALSE, warning = FALSE, eval= FALSE, fig.cap='Wave rose figure'}
 
@@
 
-We provide strong support for the idea that variations in the environmental variables, particularly wave exposure variables, are driving kelp morphological characteristics around the Cape Peninsula.
+We provide strong support that variations in the environmental variables, particularly wave exposure variables, are driving kelp morphological characteristics around the Cape Peninsula.
 Morphological adaptation due to water motion may manifest in a number of ways in high wave energy environments.
@@
 
-```{r appendix figure 1, echo=FALSE, message=FALSE, warning=FALSE}
+```{r appendix figure 1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=7}
 positionsLam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point ", "Baboon Rock", 
@@
 
-```{r appendix figure 2, echo=FALSE, message=FALSE, warning=FALSE}
+```{r appendix figure 2, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=7}
 # Order site names by location west to east
@@
 
-ggarrange(eck_all_A,eck_all_B,eck_all_C,eck_all_D,eck_all_E,eck_all_F,eck_all_G,eck_all_H,eck_all_I, ncol = 3, nrow = 3, common.legend = TRUE)
-```
+ggarrange(eck_all_A,eck_all_B,eck_all_C,eck_all_D,eck_all_E,eck_all_F,eck_all_G,eck_all_H,eck_all_I, ncol = 3, nrow = 3, common.legend = TRUE, legend = FALSE)
+```

### 2019-08-21 morphology_paper.Rmd
literature review added. updates to rmd script added. issues with pushing to repro rectified.

@@
 
-```{r map study sites, echo=FALSE, message=FALSE, warning=FALSE}
+```{r map study sites, echo=FALSE, fig.height=7, fig.width=6, message=FALSE, warning=FALSE}
 library(png)
@@
 
-```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology.'}
+```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology.', fig.height=7, fig.width=7}
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
@@
   scale_fill_discrete(name="Temperature regime") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
+
 LT2 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
@@
   scale_fill_discrete(name="Temperature regime") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
+
 LT3 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
@@
   scale_fill_discrete(name="Temperature regime") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
+
 LT4 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
@@
   scale_fill_discrete(name="Temperature regime") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
+
 LT5 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
@@
   scale_fill_discrete(name="Temperature regime") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
+
 LT6 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
@@
   scale_fill_discrete(name="Temperature regime") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
+
 LT7 <- ggplot(data = lam_data, aes(x = temp_regime, y = digits, fill = temp_regime)) +
@@
   scale_fill_discrete(name="Temperature regime") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
+
 LT8 <- ggplot(data = lam_data, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
@@
   scale_fill_discrete(name="Temperature regime") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
+
 LT9 <- ggplot(data = lam_data, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
   ylab("Total length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test") + 
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Lamina weight (g)") + 
   scale_fill_brewer(palette="OrRd") +
-  guides(fill=guide_legend(title="Wave exposure level"))
+  guides(fill=guide_legend(title="Wave exposure level")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test") + 
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Lamina length (cm)") + 
   scale_fill_brewer(palette="OrRd") +
-  guides(fill=guide_legend(title="Wave exposure level"))
+  guides(fill=guide_legend(title="Wave exposure level")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test") + 
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Lamina thickness") + 
   scale_fill_brewer(palette="OrRd") +
-  guides(fill=guide_legend(title="Wave exposure level"))
+  guides(fill=guide_legend(title="Wave exposure level")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test") + 
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") + 
   scale_fill_brewer(palette="OrRd") +
-  guides(fill=guide_legend(title="Wave exposure level"))
+  guides(fill=guide_legend(title="Wave exposure level")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test") + 
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") + 
   scale_fill_brewer(palette="OrRd") +
-  guides(fill=guide_legend(title="Wave exposure level"))
+  guides(fill=guide_legend(title="Wave exposure level")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test") + 
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe diameter (cm)") + 
   scale_fill_brewer(palette="OrRd") +
-  guides(fill=guide_legend(title="Wave exposure level"))
+  guides(fill=guide_legend(title="Wave exposure level")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test") + 
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("No. of digits") + 
   scale_fill_brewer(palette="OrRd") +
-  guides(fill=guide_legend(title="Wave exposure level"))
+  guides(fill=guide_legend(title="Wave exposure level")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test") + 
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Thallus mass (kg)") + 
   scale_fill_brewer(palette="OrRd") +
-  guides(fill=guide_legend(title="Wave exposure level"))
-
+  guides(fill=guide_legend(title="Wave exposure level")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 LW9 <- ggplot(data = lam_data, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test") + 
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Total length (cm)") + 
   scale_fill_brewer(palette="OrRd") +
-  guides(fill=guide_legend(title="Wave exposure level")) 
-```
-
-```{r laminaria temp plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime.', fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
+  guides(fill=guide_legend(title="Wave exposure level")) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
+```
+
+```{r laminaria temp plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
 ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
@@
 
-```{r lamanaria wave plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.', fig.height=6, fig.width=6, warning=FALSE}
+```{r lamanaria wave plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.', fig.height=7, fig.width=8, warning=FALSE}
 ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
@@
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Frond mass (g)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 ET2 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Primary length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 ET3 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Primary width (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 ET4 <- ggplot(data = eck_deep, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Frond length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 ET5 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 ET6 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 ET7 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe circ (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 ET8 <- ggplot(data = eck_deep, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("No. of tufts") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 ET9 <- ggplot(data = eck_deep, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Epi length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 ET10 <- ggplot(data = eck_deep, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format") + 
   ylab("Total length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
   ylab("Frond mass (g)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
-  xlab("") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Primary length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
   ylab("Primary width (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
   ylab("Frond length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
   ylab("Stipe mass (g)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
   ylab("Stipe length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
   ylab("Stipe circumference (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
   ylab("No of Tufts") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
   ylab("Epi length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "kruskal.test") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   xlab("") + 
   ylab("Total length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
 ```{r deep ecklonia temp plot, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E.
-maxima* grouped by temperature.', fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
-ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
-```
-
-```{r deep ecklonia wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.', fig.height=6, fig.width=6, warning=FALSE}
-ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
+maxima* grouped by temperature.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", legend = FALSE, ncol = 3, nrow = 4)
+```
+
+```{r deep ecklonia wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.', fig.height=8, fig.width=8, warning=FALSE}
+ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3, nrow = 4)
 ```
@@
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Frond mass (g)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 EST2 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Primary length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 EST3 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Primary width (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) + 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") + 
   ylim(0, 20)
@@
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Frond length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 EST5 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 EST6 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 EST7 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe circ (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   ylim(0, 30)
@@
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("No. of tufts") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 EST9 <- ggplot(data = eck_shall, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Epi length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) 
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 EST10 <- ggplot(data = eck_shall, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "wilcox.test") + 
-  xlab("") + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Total length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank())
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("")
 
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
-  xlab("") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Frond mass (g)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
-  xlab("") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Primary width (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
-  xlab("") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Frond length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
-  xlab("") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
-  xlab("") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  ylim(0,30) +
-  xlab("") + 
   ylab("Stipe circumference (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
-  xlab("") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("No of Tufts") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA)  + 
-  stat_compare_means(method = "kruskal.test") + 
-  xlab("") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Epi length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
-  stat_compare_means(method = "kruskal.test") + 
-  xlab("") + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
   ylab("Total length (cm)") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank()) +
+  theme_bw() +
+  theme(axis.text.x = element_blank()) + 
+  xlab("") +
   scale_fill_brewer(palette="OrRd") +
@@
 ```{r shallow ecklonia temp plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by temperature regime.', fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
-ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
-```
-
-```{r shallow rda wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by wave exposure.', fig.height=6, fig.width=6, warning=FALSE}
-ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom")
+maxima* grouped by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
+ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom", ncol = 3, nrow = 4)
+```
+
+```{r shallow rda wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by wave exposure.', fig.height=7, fig.width=8, warning=FALSE}
+ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3,nrow = 3)
 ```
@@
 # Plot RDA
-rda_eck_deep <- ggord(rda_eck, eck_meta_rda$region ,ellipse = TRUE, hull = FALSE, ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  # looking at the raw code, this is plotting the 'wa scores', the blue
+rda_eck_deep <- ggord(rda_eck, eck_meta_rda$region ,ellipse = TRUE, hull = FALSE, ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  
 ```
@@
 # Plot RDA
-rda_eck_shallow <- ggord(rda_eck_shallow,meta_shallow_rda$region,ellipse = TRUE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9, coord_fix = TRUE)  # looking at the raw code, this is plotting the 'wa scores', the blue
-
-```
-
-```{r rda plots, echo=FALSE, fig.height=9, fig.width=9, message=FALSE, warning=FALSE}
-ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE, ncol = 3, nrow = 1, legend = c("bottom"))
+rda_eck_shallow <- ggord(rda_eck_shallow,meta_shallow_rda$region,ellipse = TRUE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9, coord_fix = TRUE)
+
+```
+
+```{r rda plots, echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
+ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, ncol = 1, nrow = 3, label = "AUTO")
+lam_rda
+rda_eck_deep
+rda_eck_shallow
 ```
@@
 Unlike previous research, this study used a numerical model to provide the various variables of wave exposure.
-We also considered wind in addition to temperature and wave variables, since wind has been shown to influence the extension capabilities of canopy kelp species (Denny et al.
-1997).
-Furthermore, the protrusion of the primary blade of *E.
-maxima* out of the water may result in some direct effects of wind-drag on morphology.
-The results show that wave exposure the main driver of kelp morphology with temperature and wind playing only minor roles.
+We also considered wind in addition to temperature and wave variables, since wind is an important component of surface driven waves.
+The results show that specific variables of wave exposure are the main drivers of kelp morphology with temperature and wind playing only minor roles.
 Furthermore, the investigation of differences in morphological characteristics between shallow and deep populations of *E.
-maxima *suggests that in low wave exposure environments the role of temperature as a morphological driver increases.
+maxima * suggests that in low wave exposure environments the role of temperature as a morphological driver increases, such as in False Bay.
 
@@
 1997; Denny & Gaylord 2002) have been suggested in previous studies.
-Although this study did not measure force of attachment, other morphological responses to wave variables were evident.
-Higher temperatures tended to be assocaited with reduced measurements of stipe mass, stipe circumference/diameter and stipe length for the sub-canopy species *L.
-pallida*, which suggests that there is a reduction in overall surface area of the plant in warm temperate environments.
-The ecological consequence of this is unclear...The canopy species *E.
-maxima* exibits increased stipe length and decreased stipe circumference in warm temperate areas for both deep and shallow populations, which suggests a 'slender kelp' tactic and has been shown in previous work.**add examples here**.
+Although this study did not measure force of attachment, other morphological responses to temperature and wave variables were evident.
+Species-specific responses are evident to both wave exposure and temperature.
+In cool-temperate environments *L.
+pallida* tended to show increases in certain morphological characteristics (stipe mass, stipe length and stipe diameter) while in the warm-temperate environments these were significantly lower.
+This was also true for deep *E.
+maxima* populations, which were significantly taller and thinner for kelp in the warm-temperate compared to kelp in the cool-temperate.
+This has been documented previously by ?????, and show that the 'slender' kelp morphology is a response to warmer temperatures.
+In warmer environments kelp metabolism increases, using more nutrients than under cooler conditions causing the kelp to become slender **more needed here**.
+It should be noted however, that from these analyses that in general the cool-temperate region is more exposed to waves than that of the warm-temperate region.
+Therefore, the significantly larger morphological characteristics for kelp from the cool-temperate region may overlap with responses to wave exposure which was exhibited strongly in both species.
+When grouped by wave exposure category, *L.
+pallida* characteristics showed a significant increase in length for the 'Extremely sheltered' category (6-10 kW/m; figure ??) which maybe a 'go-with-the-flow' tactic.
+Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave.
+However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under higher exposure levels.
+This is reflected in Figure ??
+which shows an overall reduction of *L pallida* morphological characteristics suggesting a size reducing tactic to cope with higher levels of wave exposure.
+The morphological characteristics of *E.
+maxima* show similar patterns when grouped to wave exposure.
+The results suggest that deep *E.
+maxima* populations exhibit a different response as well as a higher wave exposure threshold.
+Unlike *L.
+pallida*, *E.
+maxima* exhibits a size reducing tactic with in an increase in exposure level by significantly decreasing stipe mass, stipe length, stipe circumference, total length and frond length.
+This size reduction tactic has been shown before (citations ???) and is regarded as strategy to reduce overall drag on the plant.
+However, *E.
+maxima* morphological characteristics changed significantly as wave exposure increases.
+When grouped to the higher wave exposure levels, *E.
+maxima* morphology exhibited a strength and flexibility increasing trait (see figure ??).
+Furthermore, the morphological response of *E.
+maxima* to wave exposure only occurs at the 'Moderately exposed' (20-30 kW/m) level compared to *L.
+pallida* which exhibits a response at a lower wave exposure level.
+This suggests that *E.
+maxima* can tolerate higher exposure levels before having to exhibit a morphological response.
+The redundancy analysis performed confirms the patterns and responses observed as well as identifying specific temperature and wave variables as drivers of kelp morphology.
+Kelp morphology characteristics are largely wave driven for both species but differ in terms of specific temperature and wave metrics.
+For instance, both species morphological characteristics are determined by swell direction (dir_median), standard deviation in peak period (tp_sd) and mean significant swell height (hs_mean), but *L.
+pallida* temperature driver is mean temperature (mean_temp) and *E.
+maxima* by minimum temperature (min_temp).
+Minimum temperature as an important metric for *E.
+maxima* may be related to nutrient availability.
+Low temperatures are often associated with upwelling events which bring cool, nutrient rich water into the nearshore.
+False Bay has comparativley little upwelling events compared to the western side of the peninsula, and so nutrients may be a limiting factor for *E.
+maxima* populations within False Bay (see figure ??).
+The kelp *L.
+pallida* is a sub-canopy species where more efficent mixing of water within the water column occurs (ref ??).
+The difference in canopy type between the species may also be the reason why *E.
+maxima* is driven by multiple wave metrics compared to *L.
+pallida*.
+Since *E.
+maxima* is a canopy-kelp, it is will be exposed to all components of a wave compared to *L.
+pallida* which occurs deeper in the water column.
+
+Annual maximum and winter mean temperatures influence stipe length and frond length in shallow *E.
+maxima* individuals located at sites within False Bay.
+This was not the case for deeper *E.
+maxima* individuals as well as *L.
+pallida* which were largely driven by wave variables.
+The shallow *E.
+maxima* individuals may not experience the same amount of wave energy as do deeper populations.
+This may be due to wave dampening from deeper populations of *E.
+maxima*, which attenuate the wave energy entering coastal environments (Kobayashi et al.
+1993; Jackson 1984; Mork 1996).
+The morphological variations in characteristics for deep *E.
+maxima* populations are not seen for shallow *E.
+maxima* populations.
+The lack of variation in shallow *E.
+maxima* populations was also found in previous research by (Rothman et al.
+2017), who showed that light was a significant driver.
+This influence of light may suggest photosynthetic ability and nutrient uptake as important drivers in shallow environments.
+Wave exposure may be reduced in shallow environments due to the wave damping effect of deep kelp and seaweed populations (Kobayashi et al.
+1993; Mork 1996; Jackson 1984), or rocky outcrops, thereby reducing the need to adapt morphological characteristics to reduce drag.
+Furthermore, the reduction in differences between deep and shallow *E.
+maxima *at wave exposed sites on the western of the peninsula compared to sites in False Bay provides further evidence that wave exposure is not an important driver of shallow *E.
+maxima *populations.
+Increased temperatures corresponded with a reduction in  stipe circumference, frond length, stipe length, total length and stipe mass resulting in slim or ‘skinny’ kelp individuals, which supports findings from previous research by (Mabin et al.
+2013; Serisawa et al.
+2002).
+
+# Conclusion
+
+Past research shows that macroalgal morphological characteristics are largely driven by a wave exposure while others have suggested a complex interaction between various environmental variables.
+We confirm that wave variables are the main driver of kelp morphological characteristics in both *E.
+maxima *and *L.
+pallida* populations around the Cape Peninsula, and differences in shallow and deep populations of *E.
+maxima* may reflect small differences in wave and temperature variables as drivers of kelp morphological characteristics.
 The response of kelp morphological characteristics to wave exposure was evident and both species show a tactics based on the magnitde of wave exposure.
@@
 maxima* suggesting that for shallow kelp the effect of wave exposure is less variable.
-
-# Conclusion
-
-Past research shows that macroalgal morphological characteristics are largely driven by a wave exposure while others have suggested a complex interaction between various environmental variables.
-We confirm that wave variables are the main driver of kelp morphological characteristics in both *E.
-maxima *and *L.
-pallida* populations around the Cape Peninsula, and differences in shallow and deep populations of *E.
-maxima* may reflect small differences in wave and temperature variables as drivers of kelp morphological characteristics.
+Temperature variables were not a significant driver of kelp morphology in this study.
+Although temperature plays an important role in distribution (Bolton et al.
+2012; Miller et al.
+2011; Rinde et al.
+2014) of kelp and physiological functioning of adults and gametophytes (Mohring et al.
+2014; Steneck et al.
+2002; Gerard 1997; Bearham et al.
+2013; Smale & Moore 2017), there is little evidence that temperature is an important driver of morphological variation.
+However, we suggest that temperature plays a larger role in determining kelp morphology in areas where wave energy does not exert large amounts of impact.
 

### 2019-08-26 morphology_paper.Rmd
minor adjustments to YAML. Cannot get bibtex file to work with pandoc. removing

@@
     reference_docx: frontiers_template.docx
-  pdf_document: 
-  html_document:
-    df_print: paged
-  html_notebook:
-    fig_caption: yes
-    fig_height: 7
-    fig_width: 7
 papersize: A4
@@
 fontsize: 10pt
+bibliography: morphology.bib
 ---
@@
 Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera (Bolton 2010), nevertheless form the basis of one of the most productive ecosystems globally (Mann 1973).
-Kelps generally have a dependence on cool, temperate and arctic seawater temperatures (Bolton 2010; Mohring et al.
+Kelps generally have a dependence on cool, temperate and arctic seawater temperatures (Bolton 2010; @Mohring et al.
 2014; Cavanaugh et al.
-2011; Dayton 1985), and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres (Steneck et al.
-2002).
-Their size and complex morphology provide a heterogeneous habitat structure (Steneck et al.
-2002) that accommodate a multitude of turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates (Mann 1973; Steneck et al.
-2002), each depending on a wide suite of ecological services provided by the kelp forests (Gaines & Roughgarden 1987; Paul & Steneck 1993; Levin 1994; Willis & Anderson 2003; Anderson et al.
-1997; Dayton 1985).
+2011; Dayton 1985), and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
+Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates (Mann 1973; Steneck 2002), each depending on a wide suite of ecological services provided by the kelp forests (Gaines & nRoughgarden 1987; Paul & Steneck 1993; Levin 1994; Willis & Anderson2003; Anderson 1997; Dayton 1985).
 Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in the distribution (Gorman et al.
-2013), abundance (Cavanaugh et al.
-2011; Dayton et al.
-1998), diversity (Wing et al.
-2007; Wernberg & Goldberg 2008), composition (Leliaert et al.
+2013], abundance (Cavanaugh et al.
+2011; Dayton 1998), diversity (Wing 2007; Wernberg & Goldberg 2008), composition (Leliaert et al.
 2000; Dayton 1985; Norderhaug et al.
 2012; Harley et al.
-2012), growth (Cousens 1982) and productivity (Pedersen & Nejrup 2012; Dayton et al.
-1998; Krumhansl & Scheibling 2012) of kelps.
+2012), growth (Cousens 1982) and productivity (Pedersen & Nejrup 2012; Dayton 1998; Krumhansl & Scheibling 2012) of kelps.
 
@@
 
+```{r rda lam plot, echo=FALSE, fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
+lam_rda
+```
+
 ```{r rda deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
@@
 
-```{r rda plots, echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
-ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, ncol = 1, nrow = 3, label = "AUTO")
-lam_rda
-rda_eck_deep
-rda_eck_shallow
-```
-
+```{r rda eck plots, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+ggarrange(rda_eck_deep, rda_eck_shallow, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 1, nrow = 2)
+```
 Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retian the most parsimonious descriptors of shallow *E.
@@
 2013; Smale & Moore 2017), there is little evidence that temperature is an important driver of morphological variation.
-However, we suggest that temperature plays a larger role in determining kelp morphology in areas where wave energy does not exert large amounts of impact.
+However, we suggest that temperature plays a larger role in determining kelp morphology at sites where wave energy is low enough not to elecit a response in morphological characteristics.
 
@@
 Thank you to AS and CR for their guidance, and the honours class of 2014-2015 that aided in fieldwork. Carla Louise for making the GIS plot etc.
+
+# References
 

### 2019-08-27 morphology_paper.Rmd
finally figured out referencing issue. Hopefully no more will arise...deleted old references, beginning to cite via rmarkdown syntax

@@
 fontsize: 10pt
-bibliography: morphology.bib
+bibliography: refs_morphology_paper.bib
 ---
@@
 # Introduction
-
+@Bolton2010
 Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera (Bolton 2010), nevertheless form the basis of one of the most productive ecosystems globally (Mann 1973).
-Kelps generally have a dependence on cool, temperate and arctic seawater temperatures (Bolton 2010; @Mohring et al.
+Kelps generally have a dependence on cool, temperate and arctic seawater temperatures (Bolton 2010; Mohring et al.
 2014; Cavanaugh et al.
-2011; Dayton 1985), and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
-Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates (Mann 1973; Steneck 2002), each depending on a wide suite of ecological services provided by the kelp forests (Gaines & nRoughgarden 1987; Paul & Steneck 1993; Levin 1994; Willis & Anderson2003; Anderson 1997; Dayton 1985).
+2011; Dayton 1985), and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres (Steneck et al.
+2002).
+Their size and complex morphology provide a heterogeneous habitat structure (Steneck et al.
+2002) that accommodate a multitude of turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates (Mann 1973; Steneck 2002), each depending on a wide suite of ecological services provided by the kelp forests (Gaines & nRoughgarden 1987; Paul & Steneck 1993; Levin 1994; Willis & Anderson2003; Anderson 1997; Dayton 1985).
 Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in the distribution (Gorman et al.
@@
 # References
+Anderson, R.J. et al., 1997. Holdfasts of adult kelp Ecklonia maxima provide refuges from grazing for recruitment of juvenile kelps. Marine ecology progress series, 159, pp.265–273.
+
+Andrews, W.R.H. & Hutchings, L., 1980. Upwelling in the Southern Benguela Current. Progress in oceanography, 9(1), pp.1–81.
+Atkins, G.R., 1970. Winds and Current Patterns in False Bay. Transactions of the Royal Society of South Africa, 39(2), pp.139–148.
+
+Bearham, D., Vanderklift, M.A.
+& Gunson, J.R., 2013.
+Temperature and light explain spatial variation in growth and productivity of the kelp Ecklonia radiata.
+Marine ecology progress series.
+Available at: http://www.int-res.com/abstracts/meps/v476/p59-70/.
+
+Bekkby, T. et al., 2014. Length, strength and water flow: Relative importance of wave and current exposure on morphology in kelp Laminaria hyperborea. Marine ecology progress series, 506, pp.61–70.
+
+Blanke, B. et al., 2002. Linking wind and interannual upwelling variability in a regional model of the southern Benguela. Geophysical research letters, 29(24), pp.41–41.
+
+Bolton, J.J.
+et al., 2012.
+South African kelp moving eastwards: the discovery of Ecklonia maxima (Osbeck) Papenfuss at De Hoop Nature Reserve on the south coast of South Africa.
+African journal of marine science, 34(1), pp.147–151.
+
+Bolton, J.J., 2010.
+The biogeography of kelps (Laminariales, Phaeophyceae): A global analysis with new insights from recent advances in molecular phylogenetics.
+Helgoland marine research, 64(4), pp.263–279.
+
+Booij, N., Holthuijsen, L.H. & Ris, R.C., 1997. The“ SWAN” wave model for shallow water. In Coastal Engineering 1996. ascelibrary.org, pp. 668–676.
+
+Borcard, D., Gillet, F. & Legendre, P., 2011. Numerical Ecology with R, Springer.
+Cavanaugh, K.C. et al., 2011. Environmental controls of giant-kelp biomass in the Santa Barbara Channel, California. Marine ecology progress series, 429, pp.1–17.
+
+Cousens, R., 1982. The Effect of Exposure to Wave Action on the Morphology and Pigmentation of Ascophyllum nodosum (L.) Le Jolis in South-Eastern Canada. Bontanica Marina, 25(4), pp.191–195.
+
+Cram, D.L., 1970. A suggested origin for the cold surface water in central False Bay. Transactions of the Royal Society of South Africa, 39(2), pp.129–137.
+
+Dayton, P.K., 1985. Ecology of Kelp Communities. Annual review of ecology and systematics, 16(1), pp.215–245.
+
+Dayton, P.K.
+et al., 1998.
+Sliding baselines, ghosts, and reduced expectations in kelp forest communities.
+Ecological Applications, 8(2), pp.309–322.
+Available at: http://dx.doi.org/10.1890/1051-0761(1998)008[0309:SBGARE]2.0.CO;2.
+
+Denny, M. & Gaylord, B., 2002. The mechanics of wave-swept algae. The Journal of experimental biology, 205(Pt 10), pp.1355–1362.
+
+Denny, M., Gaylord, B.
+& Cowen, E., 1997.
+Flow and Flexibility: II.
+The roles of size and shape In determining wave forces on the bull kelp Nereocystis Luetkeana.
+The Journal of experimental biology, 200, pp.3165–3183.
+Denny, M. & Roberson, L., 2002. Blade motion and nutrient flux to the kelp, Eisenia arborea. The Biological bulletin, 203(1), pp.1–13.
+
+Dufois, F.
+& Rouault, M., 2012.
+Sea surface temperature in False Bay (South Africa): Towards a better understanding of its seasonal and inter-annual variability.
+Continental shelf research, 43, pp.24–35.
+
+Dyer, D.C., 2018. Stable Isotope Ecology of South African Kelp Forests. Doctor of Philosophy. University of Cape Town.
+Environmental Modeling Center / Marine Modeling Branch, 2005. WAVEWATCH III® Model. Available at: https://polar.ncep.noaa.gov/waves/wavewatch/ [Accessed March 8, 2019].
+
+Field, J.G. et al., 1980. Upwelling in a nearshore marine ecosystem and its biological implications. Estuarine and Coastal Marine Science, 11(2), pp.133–150.
+
+Fowler-Walker, M.J., Connell, S.D.
+& Gillanders, B.M., 2005.
+To what extent do geographic and associated environmental variables correlate with kelp morphology across temperate Australia?
+Marine and Freshwater Research, 56(6), pp.877–887.
+Fowler-Walker, M.J., Wernberg, T.
+& Connell, S.D., 2006.
+Differences in kelp morphology between wave sheltered and exposed localities: Morphologically plastic or fixed traits?
+Marine biology, 148(4), pp.755–767.
+
+Friedland, M.T.
+& Denny, M.W., 1995.
+Surviving hydrodynamic forces in a wave-swept environment: consequences of morphology in the feather boa kelp, Egregia menziesii (Turner).
+Journal of experimental marine biology and ecology, 190, pp.109–133.
+Gaines, S.D. & Roughgarden, J., 1987. Fish in Offshore Kelp Forests Affect Recruitment to Intertidal Barnacle Populations. Science, 235(4787), pp.479–481.
+
+Gaylord, B.
+& Denny, M., 1997.
+Flow and flexibility.
+I.
+Effects Of size, shape and stiffness in determining wave forces on the stipitate kelps Eisenia arborea and Pterygophora californica.
+The Journal of experimental biology, 200 (Pt 24), pp.3141–3164.
+
+Gaylord, B., Denny, M.W. & Koehl, M.A.R., 2008. Flow forces on seaweeds: field evidence for roles of wave impingement and organism inertia. The Biological bulletin, 215(3), pp.295–308.
+
+Gerard, V.A., 1997. The role of nitrogen nutrition in high-temperature tolerance of the kelp Laminaria saccarina (Chromophyta). Journal of phycology, 33(5), pp.800–810.
+
+Gill, A.E. & Clarke, A.J., 1974. Wind-induced upwelling, coastal currents and sea-level changes. Deep Sea Research and Oceanographic Abstracts, 21(5), pp.325–345.
+
+Gorman, D. et al., 2013. Modeling kelp forest distribution and biomass along temperate rocky coastlines. Marine biology, 160(2), pp.309–325.
+
+Harley, C. et al., 2012. EFfects Of Climate Change On Global Seaweed Communities. Journal of phycology, 48(5), pp.1064–1078.
+
+Holthuijsen, L.H., 2010. Waves in Oceanic and Coastal Waters, Cambridge University Press.
+
+Jackson, G.A., 1984.
+Internal Wave Attenuation by Coastal Kelp Stands.
+Journal of Physical Oceanography, 14(8), pp.1300–1306.
+Available at: 2.0.CO;2">http://dx.doi.org/10.1175/1520-0485(1984)014<1300:IWABCK>2.0.CO;2.
+
+Jury, M.R., 1980. Characteristics of summer wind fields and air sea interactions over the Cape Peninsula upwelling region. Available at: http://nrfnexus.nrf.ac.za/handle/20.500.11892/52807.
+
+Jury, M.R., Kamstra, F.
+& Taunton-Clark, J., 1985.
+Synoptic summer wind cycles and upwelling off the southern portion of the Cape Peninsula.
+South African Journal of Marine Science/Suid-Afrikaanse Tydskrif vir Seewetenskap, 3(1), pp.33–42.
+
+Kain, J.M., 1982. Morphology and growth of the giant kelp Macrocystis pyrifera in New Zealand and California. Marine biology, 67(2), pp.143–157.
+
+Kobayashi, N., Raichle, A.W. & Asano, T., 1993. Wave attenuation by vegetation. Journal of Waterway, Port, Coastal and Ocean Engineering, 119(1), pp.30–48.
+
+Krumhansl, K.A. & Scheibling, R.E., 2012. Production and fate of kelp detritus. Marine ecology progress series, 467, pp.281–302.
+
+Leliaert, F. et al., 2000. Subtidal Understorey Algal Community Structure in Kelp Beds around the Cape Peninsula (Western Cape, South Africa). Botanica marina, 43(4), pp.359–366.
+
+Levin, P.S., 1994. Fine-scale temporal variation in recruitment of a temperate demersal fish: the importance of settlement versus post-settlement loss. Oecologia, 97(1), pp.124–133.
+
+Mabin, C.J.T.
+et al., 2013.
+Variation in the morphology, reproduction and development of the habitat-forming kelp Ecklonia radiata with changing temperature and nutrients.
+Marine ecology progress series, 483, pp.117–131.
+
+Mann, K.H., 1973. Seaweeds: Their Productivity and Strategy for Growth. Science, 182(4116), pp.975–981.
+
+Miller, S.M., Hurd, C.L.
+& Wing, S.R., 2011.
+Variations in growth, erosion, productivity, and morphology of Ecklonia Radiata (Alarlaceae; Laminarlales) along a fjord in southern New Zealand.
+Journal of phycology, 47(3), pp.505–516.
+
+Mohring, M.B. et al., 2014. Biogeographic variation in temperature drives performance of kelp gametophytes during warming. Marine ecology progress series, 513, pp.85–96.
+
+Molloy, F.J. & Bolton, J.J., 1996. The effects of wave exposure and depth on the morphology of inshore populations of the Namibian kelp, Laminaria schinzii Foslie. Botanica marina, 39, pp.525–531.
+
+Mork, M., 1996. The effect of kelp in wave damping. Sarsia, 80(4), pp.323–327.
+
+Norderhaug, K.M. et al., 2012. Does the diversity of kelp forest macrofauna increase with wave exposure? Journal of sea research, 69, pp.36–42.
+
+Office of the Director, 2000. National Centers for Environmental Prediction. Available at: https://www.ncep.noaa.gov/ [Accessed March 8, 2019].
+
+Oksanen, J. et al., 2016. Vegan: community ecology package. R package 2.3-3, p.Available at: https://cran.r–project.org/web/packa.
+
+Paul, A.X. & Steneck, R., 1993. Kelp Beds as Habitat for American Lobster Homarus americanus. Marine Ecology Progress Series, 100(40545), p.127.
+
+Pedersen, M.F.
+& Nejrup, L.B., 2012.
+Effects of wave exposure on population structure, demography, biomass and productivity of the kelp Laminaria hyperborea.
+Marine ecology progress series, 451, pp.45–60.
+Rautenbach, C., 2015. Southern African Coastal Vulnerability Assessment. In Csir, ed. Southern African Coastal Vulnerability Assessment.
+
+Rinde, E.
+et al., 2014.
+The influence of physical factors on kelp and sea urchin distribution in previously and still grazed areas in the NE Atlantic.
+PloS one, 9(6).
+Available at: http://dx.doi.org/10.1371/journal.pone.0100222.
+
+Rothman, M.D.
+et al., 2017.
+Geographical variation in morphology of the two dominant kelp species, Ecklonia maxima and Laminaria pallida (Phaeophyceae, Laminariales), on the west coast of Southern Africa.
+Journal of applied phycology, 29(5), pp.2627–2639.
+
+Rouault, M., Pohl, B. & Penven, P., 2010. Coastal oceanic climate change and variability from 1982 to 2009 around South Africa. African journal of marine science, 32(2), pp.237–246.
+
+Serisawa, Y. et al., 2002. Morphometric study of Ecklonia cava (Laminariales, Phaeophyta) sporophytes in two localities with different temperature conditions. Phycological research, 50(3), pp.193–199.
+
+Shipley, A.M., 1964. Some aspects of wave refraction in False Bay. South African journal of science, 60(6), pp.115–120.
+Smale, D.A. & Moore, P.J., 2017. Variability in kelp forest structure along a latitudinal gradient in ocean temperature. Journal of experimental marine biology and ecology, 486, pp.255–264.
+
+Smit, A.J.
+et al., 2013.
+A coastal seawater temperature dataset for biogeographical studies: large biases between in situ and remotely-sensed data sets around the coast of South Africa.
+PloS one, 8(12), p.e81944.
+
+Steneck, R.S. et al., 2002. Kelp forest ecosystems: biodiversity, stability, resilience and future. Environmental conservation, 29(04), pp.436–459.
+
+Stewart, H.L.
+et al., 2009.
+Differences in growth, morphology and tissue carbon and nitrogen of Macrocystis pyrifera within and at the outer edge of a giant kelp forest in California, USA.
+Marine ecology progress series, 375, pp.101–112.
+
+Wernberg, T. et al., 2003. Morphology of Ecklonia radiata (Phaeophyta: Laminarales) along its geographic distribution in south-western Australia and Australasia. Marine biology, 143(1), pp.47–55.
+
+Wernberg, T.
+& Goldberg, N., 2008.
+Short-term temporal dynamics of algal species in a subtidal kelp bed in relation to changes in environmental conditions and canopy biomass.
+Estuarine, coastal and shelf science, 76(2), pp.265–272.
+
+Wernberg, T. & Thomsen, M.S., 2005. The effect of wave exposure on the morphology of Ecklonia radiata. Aquatic botany, 83(1), pp.61–70.
+
+Willis, T.J. & Anderson, M.J., 2003. Structure of reef fish assemblages: relationship with habitat characteristics and predator density. Marine ecology progress series, 257(2000), pp.209–221.
+
+Wing, S.R.
+et al., 2007.
+Topographic shading and wave exposure influence morphology and ecophysiology of Ecklonia radiata (C.
+Agardh 1817) in Fiordland, New Zealand.
+Limnology and oceanography, 52(5), pp.1853–1864.
 

### 2019-08-27 morphology_paper.Rmd
adding reference list. final output updated with citations completed thus far

--- 2019-08-26 morphology_paper.Rmd
@@
 
-Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera (Bolton 2010), nevertheless form the basis of one of the most productive ecosystems globally (Mann 1973).
-Kelps generally have a dependence on cool, temperate and arctic seawater temperatures (Bolton 2010; @Mohring et al.
-2014; Cavanaugh et al.
-2011; Dayton 1985), and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
-Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates (Mann 1973; Steneck 2002), each depending on a wide suite of ecological services provided by the kelp forests (Gaines & nRoughgarden 1987; Paul & Steneck 1993; Levin 1994; Willis & Anderson2003; Anderson 1997; Dayton 1985).
-Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in the distribution (Gorman et al.
-2013], abundance (Cavanaugh et al.
-2011; Dayton 1998), diversity (Wing 2007; Wernberg & Goldberg 2008), composition (Leliaert et al.
-2000; Dayton 1985; Norderhaug et al.
-2012; Harley et al.
-2012), growth (Cousens 1982) and productivity (Pedersen & Nejrup 2012; Dayton 1998; Krumhansl & Scheibling 2012) of kelps.
+Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera [@Bolton2010], nevertheless form the basis of one of the most productive ecosystems globally [@Mann1973].
+Kelps generally have a dependence on cool, temperate and arctic seawater temperatures [@Bolton2010; @Mohring2014-bk; @Cavanaugh2011-ib; @Dayton1985], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
+Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002], each depending on a wide suite of ecological services provided by the kelp forests [@Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @Willis2003-xo; @anderson1997; @Dayton1985].
+Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in the distribution [@Gorman2013-nj], abundance [@Cavanaugh2011-ib; @Dayton1998-xw], diversity [@Wing2007-tv; @Wernberg2008], composition [@Leliaert2000a; @Dayton1985; @Norderhaug2012-ks; @Harley2012], growth [@Cousens1982-ck] and productivity [@Pedersen2012-vr; @Dayton1998-xw; @Krumhansl2012] of kelps.
 
 Temperature and wave exposure have been shown to play a role in driving kelp morphology.
-For example, a study by Serisawa et al.
-(2002) compared the morphology of Ecklonia cava Kjellman growing in warm temperate and cool temperate morphologies.
+For example, a study by @Serisawa2002-dc compared the morphology of Ecklonia cava Kjellman growing in warm temperate and cool temperate morphologies.
 The results showed that wrinkles in the blade seem to be a characteristic of warm temperate regions.
@@
 Wave exposure has also been shown to be the main drivers of morphological adaptation across various kelp species.
-Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility (Denny & Gaylord 2002).
-For example, a study by Wernberg & Thomsen (2005) examined the consistency of wave exposure as a driver of *Ecklonia radiata* (C.
+Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002-ob].
+For example, a study by @Wernberg2005c examined the consistency of wave exposure as a driver of *Ecklonia radiata* (C.
 Agardh) J.
@@
 The morphological adapatibility of kelps is driven by a combination of environmental factors that in turn do not act independently of one another.
-A study by Wernberg et al.
-(2003) investigated the morphology of *E.
+A study by @Wernberg2003-mp investigated the morphology of *E.
 radiata* in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australasian coast.
-They found no correlation between spatial distance and morphological similarity and rather the morphology of kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales (Wernberg et al.
-2003).
+They found no correlation between spatial distance and morphological similarity and rather the morphology of kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales [@Wernberg2003-mp].
 Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
@@
 pallida* is exposed to variations in waves and temperature.
-(Molloy & Bolton 1996) investigated the effect of depth and wave exposure on *L.
+@Molloy1996-la investigated the effect of depth and wave exposure on *L.
 pallida* at different depths and wave exposure and showed that depth had a greater effect than wave exposure when considering all the morphological characteristics; when considering individual characteristics, however, wave exposure had the most significant effect on blade thickness.
-Another study by Rothman et al.
-(2017) investigated the changes in morphology in shallow populations of *L.
+Another study by @Rothman2017-pn investigated the changes in morphology in shallow populations of *L.
 pallida* and *E.
@@
 Helena Bay and Betty's Bay constituted the north western and south eastern boundary sites, respectively.
-These sites are roughly 300 km apart, and lie within separate marine provinces[3] and cover a cline in wave exposure and temperature.
+These sites are roughly 300 km apart, and lie within separate marine provinces and cover a cline in wave exposure and temperature.
 Study sites span across the majority of the south-west coast, in varying thermal and wave energy regimes.
 The region is dominated by kelp communities that persist in contrasting abiotic environments.
-The west coast region has been termed as a cool temperate region, which is defined as a region where mean monthly temperatures are always above 10°C and always below 15°C (Smit et al.
-2013).
-East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone (Smit et al.
-2013).
-The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm temperate region (Smit et al.
-2013).
-Due to the Cape Peninsula temperate latitude, winter months bring an increased frequency of frontal depressions that originate from the Southern Ocean (Reason et al.
-2006).
+The west coast region has been termed as a cool temperate region, which is defined as a region where mean monthly temperatures are always above 10°C and always below 15°C [@Smit2013a].
+East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone [@Smit2013a].
+The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm temperate region [@Smit2013a].
+Due to the Cape Peninsula temperate latitude, winter months bring an increased frequency of frontal depressions that originate from the Southern Ocean [@Reason2006].
 These low pressures are joined by large swells with increased wave energy.
-The nearshore environment, with the accompanied biota, therefore experiences high wave energy events, with increased frequency in winter (Veitch et al.
-2018).
-The large peninsula acts as an obstruction for large south westerly swells, providing decreased wave energy along the west side of False Bay (Shipley 1964).
+The nearshore environment, with the accompanied biota, therefore experiences high wave energy events, with increased frequency in winter [@Veitch2019].
+The large peninsula acts as an obstruction for large south westerly swells, providing decreased wave energy along the west side of False Bay [@Shipley1964].
 Conversely, the west coast of Cape Point is battered by these large swells.
@@
 
-```{r map study sites, echo=FALSE, fig.height=7, fig.width=6, message=FALSE, warning=FALSE}
+```{r map study sites, echo=FALSE, fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
 library(png)
@@
 This is however absent in winter, where strong northerly winds are prevalent from St.
-Helena Bay to Betty's Bay (Field et al.
-1980; Jury et al.
-1985; Andrews & Hutchings 1980; Jury 1980).
+Helena Bay to Betty's Bay [@Field1980a; @Jury1985-nr; @Andrews1980; @Jury1980-ov].
 Patterns are also evident in the wave data.
 Seasonal variations in significant wave height (Hs), peak period and wave direction are also present.
-The direction of the dominant swell changes to the south west in winter, generated by strong low pressures that originate from the southern ocean (Reason et al.
-2006), which False Bay is sheltered from (Shipley 1964; Atkins 1970; Dufois & Rouault 2012).
+The direction of the dominant swell changes to the south west in winter, generated by strong low pressures that originate from the southern ocean [@Reason2006], which False Bay is sheltered from [@Shipley1964; @Atkins1970-bx; @Dufois2012-vy].
 In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an increased variability of Hs and Tp in this region.
-It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world (Norderhaug et al.
-2012; Leliaert et al.
-2000).
+It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world [@Norderhaug2012-ks; @Leliaert2000a].
 It is due to the near consistent south-westerly swell and the complex orography around the peninsula that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.
@@
 Shallow water temperatures were sourced from The South African Coastal Temperature Network (SACTN) website (https://github.com/ajsmit/SACTN).
-In terms of nearshore temperature, *in situ* data are preferred over satellite SST, which have shown to exhibit large biases (Smit et al.
-2013).
+In terms of nearshore temperature, *in situ* data are preferred over satellite SST, which have shown to exhibit large biases [@Smit2013a].
 Linear interpolated SST were calculated for sites where *in situ* recorders were absent.
@@
 
-Wave model variable data formed part of the South African Coastal Vulnerability Assessment, presented to the Department of Environmental Affairs (DEA) and produced by the Council for Scientific and Industrial Research (CSIR) (Rautenbach 2015).
+Wave model variable data formed part of the South African Coastal Vulnerability Assessment, presented to the Department of Environmental Affairs (DEA) and produced by the Council for Scientific and Industrial Research (CSIR) [@Rautenbach2015a].
 The South African wave climate was modeled via 20 spectral numerical wave models that simulated the offshore wave climate to the nearshore.
@@
 The particular hindcast product utilised during the DEA-CSIR study spans 1994-2013 at a 3-hour resolution.
-These data were then used to model swell propagation into the coastal models while wind waves (seas) were generated via stationary computations in the Simulating Waves in the Nearshore SWAN model (Booij et al.
-1997).
+These data were then used to model swell propagation into the coastal models while wind waves (seas) were generated via stationary computations in the Simulating Waves in the Nearshore SWAN model [@Booij1997-nx].
 The assumption of stationary computations are acceptable as the model domains were small enough so the temporal variation of the model boundary were slower than the time it takes for that boundary condition to propagate to the coast.
@@
 2.
-*Laminaria pallida* is characterised by a single smooth blade which is divided longitudinally into sections, and develops from a single meristematic region located at the junction between the blade and the stipe (Dyer 2018).
-This species has a solid stipe but develops a hollow stipe along the west coast northward and into Namibia (Rothman et al.
-2017).
+*Laminaria pallida* is characterised by a single smooth blade which is divided longitudinally into sections, and develops from a single meristematic region located at the junction between the blade and the stipe [@Dyer2018-rv].
+This species has a solid stipe but develops a hollow stipe along the west coast northward and into Namibia [@Rothman2017-pn].
 *Ecklonia maxima* consists of a single primary blade which develops above a gas-filled float and a hollow stipe below.
-Secondary blades are produced laterally from the primary blade from several meristematic regions along the margins of the primary blade, known as digits (Dyer 2018).
+Secondary blades are produced laterally from the primary blade from several meristematic regions along the margins of the primary blade, known as digits [@Dyer2018-rv].
 Both species are held to the substrate by finger-like haptera, collectively known as the holdfast.
 
-```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology.', fig.height=7, fig.width=7}
+```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology.', fig.height=6, fig.width=6}
 
@@
 Notched boxplots allowed comparison between individual categories of wave exposure.
-In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package (Oksanen et al.
-2016) in R (R Core Team, 2019).
+In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package [@Oksanen2015-vj] in R (R Core Team, 2019).
 The abiotic data covered all study sites for both species of kelp.
@@
 
-```{r rda lam plot, echo=FALSE, fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
-lam_rda
-```
-
 ```{r rda deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
@@
 
-```{r rda eck plots, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
-ggarrange(rda_eck_deep, rda_eck_shallow, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 1, nrow = 2)
-```
+```{r rda eck plots, echo=FALSE, fig.height=7, fig.width=9, message=FALSE, warning=FALSE}
+ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE, legend = "bottom", labels = "AUTO",ncol = 2, nrow = 2)
+```
+
 Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retian the most parsimonious descriptors of shallow *E.
@@
 
-This study considered important drivers of kelp morphology that have been identified in previous research (Fowler-Walker et al.
-2006; Miller et al.
-2011; Fowler-Walker et al.
-2005; Stewart et al.
-2009; Wing et al.
-2007; Molloy & Bolton 1996; Pedersen & Nejrup 2012; Wernberg & Thomsen 2005), namely temperature and wave exposure, around the Cape Peninsula on two prolific kelp species.
+This study considered important drivers of kelp morphology that have been identified in previous research [@FowlerWalker2006; @Miller2011-pm; @Fowler-Walker2005; @Stewart2009; @Wing2007-tv; @Molloy1996-la; @Pedersen2012-vr; @Wernberg2005c], namely temperature and wave exposure, around the Cape Peninsula on two prolific kelp species.
 Unlike previous research, this study used a numerical model to provide the various variables of wave exposure.
@@
 The annual and seasonal patterns in both wind influence the thermal regime around the coastline by inducing upwelling.
-Upwelling brings deep, cool nutrient rich water to coastal areas and therefore induces a decrease in temperature (Field et al.
-1980; Andrews & Hutchings 1980; Gill & Clarke 1974; Cram 1970; Rouault et al.
-2010; Blanke et al.
-2002.
-The wind speed and wind direction help drive upwelling in summer on the western side of the Cape peninsula, where southerly winds blow parallel to the coast and trigger upwelling (Field et al.
-1980; Rouault et al.
-2010).
-This is not the case for False Bay which is shielded from the dominating wind and swell the Cape Peninsula is subject to, with the exception being the winter months where upwelling becomes more favourable for the region due to the change in swell and wind direction, causing lower temperatures to dominate during this time (Rouault et al.
-2010; Andrews & Hutchings 1980).
-Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula (Rouault et al.
-2010; Field et al.
-1980).
+Upwelling brings deep, cool nutrient rich water to coastal areas and therefore induces a decrease in temperature [@Field1980a; @Andrews1980; @Gill1974-nt; @Cram1970-jw; @Rouault2010-gi; @Blanke2002].
+The wind speed and wind direction help drive upwelling in summer on the western side of the Cape peninsula, where southerly winds blow parallel to the coast and trigger upwelling [@Field1980a; @Rouault2010-gi].
+This is not the case for False Bay which is shielded from the dominating wind and swell the Cape Peninsula is subject to, with the exception being the winter months where upwelling becomes more favourable for the region due to the change in swell and wind direction, causing lower temperatures to dominate during this time [@Rouault2010-gi; @Andrews1980].
+Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula [@Rouault2010-gi; @Field1980a].
 We however see a possible lack of upwelling events for sites such as Betty’s Bay and Batsata Rock.
@@
 Morphological adaptation due to water motion may manifest in a number of ways in high wave energy environments.
-For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment (Fowler-Walker et al.
-2006; Wernberg & Thomsen 2005; Friedland & Denny 1995; Denny & Roberson 2002; Bekkby et al.
-2014; Denny et al.
-1997; Denny & Gaylord 2002) have been suggested in previous studies.
+For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment ([@FowlerWalker2006; @Wernberg2005c; @Friedland1995-vi; @Denny2002; @Bekkby2014; @Denny1997; Denny2002-ob]) have been suggested in previous studies.
 Although this study did not measure force of attachment, other morphological responses to temperature and wave variables were evident.
@@
 This may be due to wave dampening from deeper populations of *E.
-maxima*, which attenuate the wave energy entering coastal environments (Kobayashi et al.
-1993; Jackson 1984; Mork 1996).
+maxima*, which attenuate the wave energy entering coastal environments [@Kobayashi1993-os; @Jackson1984; @Mork1996-md].
 The morphological variations in characteristics for deep *E.
@@
 The lack of variation in shallow *E.
-maxima* populations was also found in previous research by (Rothman et al.
-2017), who showed that light was a significant driver.
+maxima* populations was also found in previous research by [@Rothman2017-pn] who showed that light was a significant driver.
 This influence of light may suggest photosynthetic ability and nutrient uptake as important drivers in shallow environments.
-Wave exposure may be reduced in shallow environments due to the wave damping effect of deep kelp and seaweed populations (Kobayashi et al.
-1993; Mork 1996; Jackson 1984), or rocky outcrops, thereby reducing the need to adapt morphological characteristics to reduce drag.
+Wave exposure may be reduced in shallow environments due to the wave damping effect of deep kelp and seaweed populations [@Kobayashi1993-os; @Mork1996-md; @Jackson1984], or rocky outcrops, thereby reducing the need to adapt morphological characteristics to reduce drag.
 Furthermore, the reduction in differences between deep and shallow *E.
@@
 maxima *populations.
-Increased temperatures corresponded with a reduction in  stipe circumference, frond length, stipe length, total length and stipe mass resulting in slim or ‘skinny’ kelp individuals, which supports findings from previous research by (Mabin et al.
-2013; Serisawa et al.
-2002).
+Increased temperatures corresponded with a reduction in  stipe circumference, frond length, stipe length, total length and stipe mass resulting in slim or ‘skinny’ kelp individuals, which supports findings from previous research by [@Mabin2013-zs; @Serisawa2002-dc].
 
@@
 
-# References
-
 # Appendix
@@
 ```
+
+# References

### 2019-08-27 morphology_paper.Rmd
word output updated. general house cleaning.

@@
 
-```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology.', fig.height=6, fig.width=6}
+```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
 
@@
 
-```{r rda eck plots, echo=FALSE, fig.height=7, fig.width=9, message=FALSE, warning=FALSE}
+```{r rda eck plots, echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
 ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE, legend = "bottom", labels = "AUTO",ncol = 2, nrow = 2)

### 2019-08-27 morphology_paper.Rmd
a bit of house cleaning

@@
 
-```{r wave rose figure, echo=FALSE, warning = FALSE, eval= FALSE, fig.cap='Wave rose figure'}
+```{r wave rose figure, echo=FALSE, warning = FALSE, eval= TRUE, fig.cap='Wave rose figure'}
 
 img_wave_rose <- readPNG("figures/wave_rose.png")
-grid.raster(wave rose)
-```
-**insert wave roses here**
+grid.raster(img_wave_rose)
+```
+
 Annual mean wave peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St.
@@
 ```
-**insert wave power map here**
+
 In Figure 5 the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
@@
 Unlike previous research, this study used a numerical model to provide the various variables of wave exposure.
-We also considered wind in addition to temperature and wave variables, since wind is an important component of surface driven waves.
-The results show that specific variables of wave exposure are the main drivers of kelp morphology with temperature and wind playing only minor roles.
+We also considered wind in addition to temperature and wave variables, since wind is an important component of wind-driven surface waves.
+The results show that specific variables of wave exposure are the main drivers of kelp morphology with temperature only playing a minor role.
 Furthermore, the investigation of differences in morphological characteristics between shallow and deep populations of *E.

### 2019-08-28 morphology_paper.Rmd
references finished. I will review shortly but otherwise discussion is finished.

@@
 ---
-title: Using numerical modelling to investigate temperature and wave exposure as drivers of macroalgal characteristics
+title: Using numerical modelling to investigate temperature and wave exposure as drivers
+  of macroalgal characteristics
 author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
-geometry: margin=1in
+fontsize: 10pt
 output:
@@
     reference_docx: frontiers_template.docx
+  pdf_document: default
+geometry: margin=1in
 papersize: A4
 tables: yes
-fontsize: 10pt
 bibliography: refs_morphology_paper.bib
@@
 Temperature and wave exposure have been shown to play a role in driving kelp morphology.
-For example, a study by @Serisawa2002-dc compared the morphology of Ecklonia cava Kjellman growing in warm temperate and cool temperate morphologies.
-The results showed that wrinkles in the blade seem to be a characteristic of warm temperate regions.
+For example, a study by @Serisawa2002-dc compared the morphology of *Ecklonia cava* Kjellman growing in warm temperate and cool temperate morphologies.
+The results showed that wrinkles in the blade and reduced size to be a characteristic of warm temperate regions.
 However this study did not take into account the interaction with other environmental variables such as wave exposure.
 Wave exposure has also been shown to be the main drivers of morphological adaptation across various kelp species.
-Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002-ob].
+Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002-ob; @Hurd2000; @Fowler-Walker2005, @Thomsen2004-yd].
 For example, a study by @Wernberg2005c examined the consistency of wave exposure as a driver of *Ecklonia radiata* (C.
@@
 Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
-Wave exposure may be greater in shallower environments, or it may be reduced through damping effects of kelp further offshore and natural barriers such as rock outcrops; also, seawater temperature may be higher in shallow water environments due to higher solar irradiance.
-
-Other species of kelp, such as *Laminaria pallida* and *Ecklonia maxima* (Osbeck) Papenfuss which are important habitat forming seaweeds that exist around the coast of South Africa, offer a unique opportunity to investigate the drivers of macroalgal morphological characteristics between deep and shallow water environments.
+Wave exposure may be greater in shallower environments, or it may be reduced through damping effects of kelp and other subtidal species [@Kobayashi1993-os; Mia2013; @Dubi1995].
+Shallower environments are exposed to higher irradiance levels which increases temperatures [@Dellatorre2012; @Dunne2001] as well as an altered hydrodynamic environment which effect vertical mixing in the shallow nearshore.
+
+The kelps *Laminaria pallida* and *Ecklonia maxima* (Osbeck) Papenfuss are important habitat forming seaweeds that exist around the coast of South Africa and offer a unique opportunity to investigate the drivers of macroalgal morphological characteristics between canopy and sub-canopy species, as well as between deep and shallow water environments.
 Although both these species exist in the subtidal, *L.
@@
 Therefore, *E.
-maxima* is exposed to variations in wave, swell and temperature, while *L.
-pallida* is exposed to variations in waves and temperature.
+maxima* is exposed to variations in wind drive surface waves, swell and temperature, while *L.
+pallida* is exposed to variations in swell and temperature.
 @Molloy1996-la investigated the effect of depth and wave exposure on *L.
-pallida* at different depths and wave exposure and showed that depth had a greater effect than wave exposure when considering all the morphological characteristics; when considering individual characteristics, however, wave exposure had the most significant effect on blade thickness.
+pallida* at different depths and wave exposure and showed that depth had a greater effect than wave exposure when considering all the morphological characteristics (wave exposure alters with depth); when considering individual characteristics, however, wave exposure had the most significant effect on blade thickness.
 Another study by @Rothman2017-pn investigated the changes in morphology in shallow populations of *L.
@@
 Currently, no ecological studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
-Previous research investigating drivers of kelp morphological characteristics used qualitative estimates of wave exposure and did not consider other morphological characteristics.
-Furthermore, possible differences in morphology between shallow and deep water populations have not been considered in previous research which ignores the effects of wave damping.
+Previous research investigating drivers of kelp morphological characteristics have often used qualitative estimates of wave exposure and did not consider other morphological characteristics.
+Furthermore, wind is an important component of surface gravity waves [@Holthuijsen2010-vl] and previous work has not differentiated between surface gravity waves and swell waves, which the current work will aim to achieve.
+Other considerations such as possible differences in morphology between sub-canopy and canopy species, shallow and deep water populations have not been considered in previous research which ignores the effects of wave damping.
 The aim of this study is, therefore, to understand how temperature and wave exposure can influence the morphology in two species of kelps around South Africa, as well as between deep and shallow populations of the same species.
-This will be achieved by initially understanding the variation in temperature and waves using advanced numerical models.
-These numerical models will then be used as a basis for investigating the consequences for morphological chracteristics of *E.
+This will be achieved by initially understanding the variation in temperature and waves using an advanced numerical model.
+The numerical model will then be used as a basis for investigating the consequences for morphological chracteristics of *E.
 maxima* and *L.
-pallida* using statistical tools.
+pallida* using multiple statistical tools.
 
@@
 The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm temperate region [@Smit2013a].
+
+```{r map study sites, echo=FALSE, fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
+library(png)
+library(grid)
+
+map <- readPNG("figures/sa_map.png")
+grid.raster(map)
+```
+
 Due to the Cape Peninsula temperate latitude, winter months bring an increased frequency of frontal depressions that originate from the Southern Ocean [@Reason2006].
@@
 Multiple sites, therefore, exist where kelps grow in diverse temperature and wave energy climates, in close proximity.
-
-```{r map study sites, echo=FALSE, fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
-library(png)
-library(grid)
-
-map <- readPNG("figures/sa_map.png")
-grid.raster(map)
-```
-
 The topography and elevation along the Cape Peninsula channel and shield winds along False Bay.
@@
 
-```{r rda eck plots, echo=FALSE, fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
+```{r rda eck plots, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
 ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE, legend = "bottom", labels = "AUTO",ncol = 2, nrow = 2)
@@
 
-This study considered important drivers of kelp morphology that have been identified in previous research [@FowlerWalker2006; @Miller2011-pm; @Fowler-Walker2005; @Stewart2009; @Wing2007-tv; @Molloy1996-la; @Pedersen2012-vr; @Wernberg2005c], namely temperature and wave exposure, around the Cape Peninsula on two prolific kelp species.
+This study considered important drivers of kelp morphology that have been identified in previous research [@FowlerWalker2006; @Miller2011-pm; @Fowler-Walker2005; @Stewart2009; @Wing2007-tv; @Molloy1996-la; @Pedersen2012-vr; @Wernberg2005c; @Thomsen2004-yd], namely temperature and wave exposure, around the Cape Peninsula on two prolific kelp species.
+The complex geomorphology of the western Cape coastline creates an ideal natural laboratory for studies of kelps in interaction with their environment.
 Unlike previous research, this study used a numerical model to provide the various variables of wave exposure.
-We also considered wind in addition to temperature and wave variables, since wind is an important component of wind-driven surface waves.
+We also considered wind in addition to temperature and wave variables, since wind is an important component of wind-driven surface gravity waves [@Holthuijsen2010-vl].
 The results show that specific variables of wave exposure are the main drivers of kelp morphology with temperature only playing a minor role.
@@
 
-The complex geomorphology of the western Cape coastline creates an ideal natural laboratory for studies of kelps in interaction with their environment.
 The annual and seasonal patterns in both wind influence the thermal regime around the coastline by inducing upwelling.
@@
 Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula [@Rouault2010-gi; @Field1980a].
-We however see a possible lack of upwelling events for sites such as Betty’s Bay and Batsata Rock.
 
@@
 Morphological adaptation due to water motion may manifest in a number of ways in high wave energy environments.
-For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment ([@FowlerWalker2006; @Wernberg2005c; @Friedland1995-vi; @Denny2002; @Bekkby2014; @Denny1997; Denny2002-ob]) have been suggested in previous studies.
+For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment [@FowlerWalker2006; @Wernberg2005c; @Friedland1995-vi; @Denny2002; @Bekkby2014; @Denny1997; @Denny2002-ob] have been identified in previous studies.
 Although this study did not measure force of attachment, other morphological responses to temperature and wave variables were evident.
-Species-specific responses are evident to both wave exposure and temperature.
+Species-specific responses are evident in both wave exposure and temperature.
 In cool-temperate environments *L.
@@
 This was also true for deep *E.
-maxima* populations, which were significantly taller and thinner for kelp in the warm-temperate compared to kelp in the cool-temperate.
-This has been documented previously by ?????, and show that the 'slender' kelp morphology is a response to warmer temperatures.
-In warmer environments kelp metabolism increases, using more nutrients than under cooler conditions causing the kelp to become slender **more needed here**.
+maxima* populations, which had longer, thinner stipes for kelp in the warm-temperate compared to kelp in the cool-temperate.
+Reduction in certian morphological charateristics has been attributed to temperature by @Serisawa2002-dc in the kelp *E.
+cava*, which was smaller and shorter in warmer sites compared to cooler sites.
+The reduction in size of adult may be a response to low nutrient conditions, which has been shown to reduce growth rate and overall morphology [@Simonson2015].
+Warmer temperatures are associated with low nutrient concentrations [@Waldron1992] and the low frequency of upwelling conditions in False Bay (low nutrient supply) coupled with warmer temperatures may be a contributing factor.
 It should be noted however, that from these analyses that in general the cool-temperate region is more exposed to waves than that of the warm-temperate region.
@@
 When grouped by wave exposure category, *L.
-pallida* characteristics showed a significant increase in length for the 'Extremely sheltered' category (6-10 kW/m; figure ??) which maybe a 'go-with-the-flow' tactic.
-Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave.
-However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under higher exposure levels.
+pallida* characteristics showed a significant increase in length for the 'Extremely sheltered' category (6-10 kW/m; figure ??) compared to the fully sheltered category, which maybe a 'go-with-the-flow' tactic.
+Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny2002-ob; @Hurd2000].
+However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under higher exposure levels [@Denny2002-ob; @Denny1997].
 This is reflected in Figure ??
@@
 maxima* exhibits a size reducing tactic with in an increase in exposure level by significantly decreasing stipe mass, stipe length, stipe circumference, total length and frond length.
-This size reduction tactic has been shown before (citations ???) and is regarded as strategy to reduce overall drag on the plant.
+A magnitude related reponse has been suggested by @Wernberg2010 who investigated the temporal and spatial variations of various environmental drivers of the kelp *E.
+radiata*.
+The authors identified wave exposure as the most important driver of kelp morphological characteristics and that the type of response elicited is dependent on the magnitude of wave exposure.
+
+A size reduction tactic has been shown before [@FowlerWalker2006; @Denny2002-ob; @Hurd2000; @Fowler-Walker2005] and is regarded as strategy to reduce overall drag on the plant.
 However, *E.
@@
 Kelp morphology characteristics are largely wave driven for both species but differ in terms of specific temperature and wave metrics.
-For instance, both species morphological characteristics are determined by swell direction (dir_median), standard deviation in peak period (tp_sd) and mean significant swell height (hs_mean), but *L.
-pallida* temperature driver is mean temperature (mean_temp) and *E.
-maxima* by minimum temperature (min_temp).
+For instance, both species morphological characteristics are determined by swell direction (*dir_median*), standard deviation in peak period (*tp_sd*) and mean significant swell height (*hs_mean*), but *L.
+pallida* temperature driver is mean temperature (*mean_temp*) and *E.
+maxima* by minimum temperature (*min_temp*).
 Minimum temperature as an important metric for *E.
@@
 False Bay has comparativley little upwelling events compared to the western side of the peninsula, and so nutrients may be a limiting factor for *E.
-maxima* populations within False Bay (see figure ??).
-The kelp *L.
-pallida* is a sub-canopy species where more efficent mixing of water within the water column occurs (ref ??).
+maxima* populations within False Bay (see figure ??) and hence the identification of minimum temperature (*min_temp*) as a driver.
 The difference in canopy type between the species may also be the reason why *E.
@@
 Since *E.
-maxima* is a canopy-kelp, it is will be exposed to all components of a wave compared to *L.
+maxima* is a canopy-kelp, it is  exposed to all components of a wave compared to *L.
 pallida* which occurs deeper in the water column.
+Mean temperature identified as a driver for *L.
+pallida* may be related to diurnal temperature fluctuations in the water column.
+Solar heating of the water surface in combination with wind-driven transport causes fluctuations within the water coloumn, which can occur daily or seasonally [@Kaplan2003].
 
@@
 This influence of light may suggest photosynthetic ability and nutrient uptake as important drivers in shallow environments.
-Wave exposure may be reduced in shallow environments due to the wave damping effect of deep kelp and seaweed populations [@Kobayashi1993-os; @Mork1996-md; @Jackson1984], or rocky outcrops, thereby reducing the need to adapt morphological characteristics to reduce drag.
+Wave exposure may be reduced in shallow environments due to the wave damping effect of deep kelp and seaweed populations [@Kobayashi1993-os; @Mork1996-md; @Jackson1984] thereby reducing the need to adapt morphological characteristics to reduce drag.
 Furthermore, the reduction in differences between deep and shallow *E.
-maxima *at wave exposed sites on the western of the peninsula compared to sites in False Bay provides further evidence that wave exposure is not an important driver of shallow *E.
-maxima *populations.
+maxima * at wave exposed sites on the western of the peninsula compared to sites in False Bay provides further evidence that wave exposure is not an important driver of shallow *E.
+maxima* populations.
 Increased temperatures corresponded with a reduction in stipe circumference, frond length, stipe length, total length and stipe mass resulting in slim or ‘skinny’ kelp individuals, which supports findings from previous research by [@Mabin2013-zs; @Serisawa2002-dc].
+These findings were confirmed by the redundancy analysis which showed that contribution of wave variables as a driver of kelp morphological characterstics was lower compared to temperature variables.
+The sites do cluster by region but almost overlap entirely, suggesting that wave variables are not the main driver.
+Although wave variables were not identified as an important driver is is important to note the wave variables identified in the RDA differ to deep *E.
+maxima*.
+The identification of median wind direction (*dirw_median*) suggests that surface gravity waves as a driver compared to swell waves (*dir_median*) identifiend for *L.
+pallida* and *E.
+maxima*.
+The identification of maximum temperature as a driver of kelp morphological characteristics may be due to depth and irradiance.
+Temperature conditions along the shallow depths of the coastline have been shown to be warmer than deeper waters due to a combination of depth and higher irradiance in shallow coastal environments as water is able to heat up more effectivley than deeper waters.
 

### 2019-08-28 morphology_paper.Rmd
added inshore hydrodynamic section to discussion for the laminiara and mean temp

@@
 ```{r rda eck plots, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
-ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE, legend = "bottom", labels = "AUTO",ncol = 2, nrow = 2)
+ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE, legend = "bottom", labels = "AUTO", ncol = 2, nrow = 2)
 ```
@@
 maxima* populations within False Bay (see figure ??) and hence the identification of minimum temperature (*min_temp*) as a driver.
+Vertical mixing of the water coloumn for inshore regions are driven by several abiotic processes: (1) Turbulence of breaking waves inside or outside of the surf zone (2) convective mixing through a combination of cooling and evaporation (3) wind driven currents and (4) tidal mixing [@Smit2013a].
+These abiotic processes cause effective vertical mixing of surface and deeper water stratifications which leads to a homogeneous thermal environment, which in this study may be interpreted as the mean temperature.
+Therefore, the homogenious thermal regime of inshore regions may be a reason why mean temperature is driver of *L.
+pallida* morphological characteristics.
+
 The difference in canopy type between the species may also be the reason why *E.
@@
 Annual maximum and winter mean temperatures influence stipe length and frond length in shallow *E.
-maxima* individuals located at sites within False Bay.
-This was not the case for deeper *E.
+maxima* individuals located at sites within False Bay; which was not the case for deeper *E.
 maxima* individuals as well as *L.
@@
 Increased temperatures corresponded with a reduction in stipe circumference, frond length, stipe length, total length and stipe mass resulting in slim or ‘skinny’ kelp individuals, which supports findings from previous research by [@Mabin2013-zs; @Serisawa2002-dc].
-These findings were confirmed by the redundancy analysis which showed that contribution of wave variables as a driver of kelp morphological characterstics was lower compared to temperature variables.
-The sites do cluster by region but almost overlap entirely, suggesting that wave variables are not the main driver.
-Although wave variables were not identified as an important driver is is important to note the wave variables identified in the RDA differ to deep *E.
+These findings were confirmed by the redundancy analysis which showed the contribution of wave variables as a driver of kelp morphological characterstics was lower compared to temperature variables.
+The sites do cluster by region but almost overlap entirely, further indicating wave variables are not the main driver.
+Although wave variables were not identified as an important driver, it is important to note the wave variables identified in the RDA differ to deep *E.
 maxima*.
@@
 maxima*.
-The identification of maximum temperature as a driver of kelp morphological characteristics may be due to depth and irradiance.
-Temperature conditions along the shallow depths of the coastline have been shown to be warmer than deeper waters due to a combination of depth and higher irradiance in shallow coastal environments as water is able to heat up more effectivley than deeper waters.
+The identification of maximum temperature as a driver of kelp morphological characteristics may be due heating of shallow coastal waters due to solar heating.
 

### 2019-08-28 morphology_paper.Rmd
a bit extra...
added a bit extra to discussion

@@
 Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula [@Rouault2010-gi; @Field1980a].
+In general, the western side of the peninsula is more exposed than the eastern side (False Bay) and experiences significanlty higher significant swell height and significant peak period.
+This was also reflected in the wave exposure categories which show a trend of decreasing wave exposure (wave power) around the peninsula and into False Bay.
+There are also differences between the types of waves that each side of the peninsula experiences.
+The ????
+analysis shows the western side of the peninsula is not only more exposed but the hydrodynamics are largely swell driven, while on the eastern side most of the wave energy is driven by a combination of swell and wind.
 
@@
 pallida* morphological characteristics.
-
-The difference in canopy type between the species may also be the reason why *E.
+The difference in canopy type between the species may be the reason why *E.
 maxima* is driven by multiple wave metrics compared to *L.

### 2019-08-28 morphology_paper.Rmd
yaml edit

@@
 output:
+  pdf_document: default
   word_document:
     reference_docx: frontiers_template.docx
-  pdf_document: default
 geometry: margin=1in

### 2019-08-29 morphology_paper.Rmd
adding latex cross-referecning syntax. Deleted duplicate paragraphs in wave section.

@@
 
-Sites were chosen to represent an array of environmental gradients, as indicated in Fig.
-1.
+Sites were chosen to represent an array of environmental gradients, as indicated in figure \ref{fig:map study sites}.
 St.
@@
 
-```{r map study sites, echo=FALSE, fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
+```{r map study sites, echo=FALSE, fig.height=5, fig.width=5, message=FALSE, warning=FALSE, fig.cap="\\label{fig:map study sites}Map of the study area and sampling sites.
+The red dot indicates the position of Cape Town for sake of reference."}
+
 library(png)
@@
 
-The morphological characteristics of both species are presented in Fig.
-2.
+The morphological characteristics of both species are presented in figure \ref{fig:schematic}.
 *Laminaria pallida* is characterised by a single smooth blade which is divided longitudinally into sections, and develops from a single meristematic region located at the junction between the blade and the stipe [@Dyer2018-rv].
@@
 
-```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
+```{r schematic, echo=FALSE, fig.cap='\\label{fig:schematic}Schematic of *L. pallida* and *E. maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
 
@@
 The mean annual coastal water temperature for the study sites located on the western side of the peninsula ranged from 14.5 ± 0.9 ℃ (mean ± SD) at St.
-Helena Bay, the most northern site on the western side of the peninsula, to 14.6 ± 0.3 ℃ at Olifantsbos, the most southern site on the western side of the peninsula.
-For sites located on the eastern side of the peninsula (within False Bay), the annual mean coastal water temperatures ranged from 15.5 ± 0.9 ℃ at Buffelsbaai to 15.0 ± 0.9 ℃ at Betty’s Bay, the most eastern site on the coastline in this study.
+Helena Bay, the most northern site on the western side of the peninsula, to 14.6 ± 0.3 ℃ at Olifantsbos, the most southern site on the western side of the peninsula (see figure \ref{fig:Temperature variables plot}A).
+For sites located on the eastern side of the peninsula (within False Bay), the annual mean coastal water temperatures ranged from 15.5 ± 0.9 ℃ at Buffelsbaai to 15.0 ± 0.9 ℃ at Betty’s Bay, the most eastern site on the coastline in this study (see figure \ref{fig:Temperature variables plot}A).
 Oudekraal had the lowest annual mean coastal water temperature of 12.3 ± 0.9.
@@
 Helena Bay to 0.9 at Olifantsbos, and on the eastern side on of the peninsula was 1.0 at Buffelsbaai to 2.9 at Betty’s Bay.
-As can be seen in Fig 3, the trend in the range of temperature decreases from St.
+As can be seen in figure \ref{fig:Temperature variables plot}A, the trend in the range of temperature decreases from St.
 Helena Bay to Miller’s point and then increases from Kalk Bay to Betty’s Bay.
@@
 
-```{r Temperature variables plot, echo=FALSE, warning=FALSE,fig.height=5,fig.width=8, fig.cap='Temperature variables at the collection sites around the Cape Peninsula are represented in plot A.
+```{r Temperature variables plot, echo=FALSE, warning=FALSE,fig.height=5,fig.width=8, fig.cap='\\label{fig:Temperature variables plot}Temperature variables at the collection sites around the Cape Peninsula are represented in plot A.
 Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.
@@
 
-Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula on both an annual and seasonal scale.
-Annual mean significant wave height ranged from 0.5 ± 0.2 m (mean ± SD) at St.
-Helena Bay to 2.3  ± 0.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side of the peninsula it ranged from 0.9 ± 0.4 m at Buffelsbaai to 1.9 ± 0.6 m at Betty’s Bay.
-Annual mean wave peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St.
-Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay.
-The data shows no trend in mean peak peak period for the coastline.
-Annual peak period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and summer respectively, Miller’s point and had the lowest peak period amongst all sites.
+Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula on both an annual and seasonal scale (see figure \ref{fig:wave variable plot}).
+Mean significant wave height ranged from 0.5 ± 0.2 m (mean ± SD) at St.
+Helena Bay to 2.3  ± 0.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side of the peninsula it ranged from 0.9 ± 0.4 m at Buffelsbaai to 1.9 ± 0.6 m at Betty’s Bay (see figure \ref{fig:wave variable plot}A).
+The maximum significant wave height ranged from 1.4 m at St.
+Helena Bay to 4.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side ranged from 2.6 m at Buffelsbaai to 4.2 m at Betty’s Bay.
+Mean peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St.
+Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay (see figure \ref{fig:wave variable plot}B).
+The data shows no trend in mean peak period for the coastline.
 These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side of the peninsula, with Miller’s point showing the highest variation across timescales.
-Annual maximum peak period ranged from 18.9 s at St.
+Maximum peak period ranged from 18.9 s at St.
 Helena Bay to 18.4 s at Olifantsbos for the western side of the peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty’s Bay on the eastern side of the peninsula.
+When wave variables were grouped by region significant differences were found for both mean significant swell height and mean peak period.
+The western region The western region experiences a significantly higher mean significant swell height (p < 0.05; see figure \ref{fig:wave variable plot}C) and mean peak period (p < 0.05; see figure \ref{fig:wave variable plot}D).
+These wave heights and their variability are modulated by the relative sheltering against the predominant south-westerly swell direction and the size of the fetch for local wave generation.
 
@@
 
-```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE,fig.height=7,fig.width=8, fig.cap='In plot A1 and A2 wave variables at each collection site around the Western Cape coast are presented.
+```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE,fig.height=7,fig.width=8, fig.cap='\\label{fig:wave variable plot}In plot A1 and A2 wave variables at each collection site around the Western Cape coast are presented.
 Mean variables are represented by dots and standard deviation by whiskers.
 Blue dots represent sites grouped to the western category and red dots represent sites grouped to the eastern category.'}
-ggarrange(wave_plot_A,wave_boxplot_1, wave_plot_B, wave_boxplot_2, common.legend = TRUE, labels = c("A1", "A2", "B1", "B2"))
-```
-
-Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula on both an annual and seasonal scale.
-Annual mean significant wave height ranged from 0.5 ± 0.2 m (mean ± SD) at St.
-Helena Bay to 2.3  ± 0.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side of the peninsula it ranged from 0.9 ± 0.4 m at Buffelsbaai to 1.9 ± 0.6 m at Betty’s Bay (Fig.
-4).
-The annual maximum significant wave height ranged from 1.4 m at St.
-Helena Bay to 4.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side ranged from 2.6 m at Buffelsbaai to 4.2 m at Betty’s Bay.
-These wave heights and their variability are modulated by the relative sheltering against the predominant south-westerly swell direction and the size of the fetch for local wave generation.
-
-```{r wave rose figure, echo=FALSE, warning = FALSE, eval= TRUE, fig.cap='Wave rose figure'}
+ggarrange(wave_plot_A,wave_boxplot_1, wave_plot_B, wave_boxplot_2, common.legend = TRUE, labels = c("A", "C", "B", "D"))
+```
+
+```{r wave rose, echo=FALSE, warning = FALSE, eval= TRUE, fig.cap='\\label{fig:wave rose}Wave rose figure'}
 
@@
 
-Annual mean wave peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St.
-Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay ([Fig.
-4].
-Annual peak peak period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and summer respectively, Miller’s point and had the lowest peak peak period amongst all sites.
-These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side of the peninsula, with Miller’s point showing the highest variation across timescales.
-Annual maximum peak period ranged from 18.9 s at St.
-Helena Bay to 18.4 s at Olifantsbos for the western side of the peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty’s Bay on the eastern side of the peninsula.
 Data for median wave direction showed no trend across timescales; however the variation (SD) in wave direction is lower on the western side of the peninsula compared to the eastern side both annually and for winter.
-Summer data were the exception and showed lower variation for the eastern side of the peninsula [Fig.
-4].
+Summer data were the exception and showed lower variation for the eastern side of the peninsula (see figure \ref{fig:wave rose}).
 Wave variables divided into Annual (top), Austral summer (middle) and Austral winter (bottom) respectively.
@@
 
-```{r wave power map, echo=FALSE, eval= FALSE, fig.cap='Wave power map'}
+```{r wave power map, echo=FALSE, eval= FALSE, fig.cap='\\label{fig:wave power map}Wave power map'}
 
@@
 In Figure 5 the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
-The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in [Fig.
-6].
 The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits.
-The western periphery of the Cape peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures [Fig.
-6].
-THere the marked seasonality, with higher energy waves during winter, may be clearly observed once more.
+The western periphery of the Cape peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures (see figure \ref{fig:wave power map}).
+There the marked seasonality, with higher energy waves during winter, may be clearly observed once more.
 To clarify the averaged wave exposure maps presented, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in [Fig.
-7].
+???].
 Tracing the wave height contours into False Bay its clear why this bay’s western periphery is predominantly sheltered.
@@
 This is not the case for False Bay which is shielded from the dominating wind and swell the Cape Peninsula is subject to, with the exception being the winter months where upwelling becomes more favourable for the region due to the change in swell and wind direction, causing lower temperatures to dominate during this time [@Rouault2010-gi; @Andrews1980].
+The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in figure \ref{fig:wave power map}.
 Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula [@Rouault2010-gi; @Field1980a].
@@
 The ????
-analysis shows the western side of the peninsula is not only more exposed but the hydrodynamics are largely swell driven, while on the eastern side most of the wave energy is driven by a combination of swell and wind.
+analysis shows the western side of the peninsula is not only more exposed but the hydrodynamics are largely swell driven, while on the eastern side most of the wave energy is driven by a combination of swell and wind with the latter being the predominant driver.
 

### 2019-08-29 morphology_paper.Rmd
cleaned up clumsy sentence in intro

@@
 Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
-Wave exposure may be greater in shallower environments, or it may be reduced through damping effects of kelp and other subtidal species [@Kobayashi1993-os; Mia2013; @Dubi1995].
-Shallower environments are exposed to higher irradiance levels which increases temperatures [@Dellatorre2012; @Dunne2001] as well as an altered hydrodynamic environment which effect vertical mixing in the shallow nearshore.
+Deep and shallow water environments may differ in abiotic processes such as the degree of water coloumn mixing [@Smit2013a], solar heating [@Dellatorre2012; @Dunne2001] and effects of wave dampening [@Kobayashi1993-os; Mia2013; @Dubi1995], all of which do not act idenpendently and instead influence one another.
 
@@
 There the marked seasonality, with higher energy waves during winter, may be clearly observed once more.
-To clarify the averaged wave exposure maps presented, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in [Fig.
-???].
+
+```{r FB spectrum, echo=FALSE, eval= FALSE, fig.cap='\\label{fig:FB spectrum}Wave power map'}
+
+library(png)
+library(grid)
+
+img_fb <- readPNG("figures/falsebay_spectrum.png")
+grid.raster(img_fb)
+```
+
+To clarify the averaged wave exposure maps presented, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in figure \ref{fig:FB spectrum}.
 Tracing the wave height contours into False Bay its clear why this bay’s western periphery is predominantly sheltered.
@@
 Significant differences were found between cool temperate and warm temperate regimes for most *L.
-pallida* morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
+pallida* morphological characteristics with lamina length, number of digits and thallus mass the exceptions (see figure \ref{fig:lamanaria temp plots}).
 Certian morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significanlty higher means for kelps in cool temperate regimes when compared to kelps for the warm temperate regime.
@@
 
-```{r laminaria temp plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
+```{r laminaria temp plots, echo=FALSE, fig.cap='\\label{fig:lamanaria temp plots}Boxplots of *L.
+pallida* morphological characteristics group by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
 ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
@@
 
-```{r lamanaria wave plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.', fig.height=7, fig.width=8, warning=FALSE}
+```{r lamanaria wave plots, echo=FALSE, fig.cap='\\label{fig:lamanaria wave plots}Boxplots of *L.
+pallida* morphological characteristics group by wave exposure category.', fig.height=7, fig.width=8, warning=FALSE}
 ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
@@
 maxima* morphological characteristics showed differences between cool temperate and warm temperate regime except for frond mass, frond length, stipe mass and epiphyte length.
-Mean values for primary length, primary width and stipe circumference were significantly higher for kelps from cool temperate regions than for kelps from warm temperate regions.
+Mean values for primary length, primary width and stipe circumference were significantly higher for kelps from cool temperate regions than for kelps from warm temperate regions (see figure \ref{fig:lamanaria wave plots}.
 The remaining morphological characteristics, stipe length, number of tufts and total length had significanlty lower values for kelps from cool temperate compared to kelps from the warm temperate regime.
@@
 
-```{r deep ecklonia wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.', fig.height=8, fig.width=8, warning=FALSE}
+```{r deep ecklonia wave, echo=FALSE, fig.cap='\\label{fig:deep ecklonia wave}Boxplot of morphological characteristics for deep population *E.
+maxima* grouped by wave exposure.', fig.height=8, fig.width=8, warning=FALSE}
 ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3, nrow = 4)
@@
 
-```{r shallow ecklonia temp plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E.
+```{r shallow ecklonia temp, echo=FALSE, fig.cap='\\label{fig:shallow ecklonia temp}Boxplot of morphological characteristics for shallow population *E.
 maxima* grouped by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
@@
 
-```{r shallow rda wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by wave exposure.', fig.height=7, fig.width=8, warning=FALSE}
+```{r shallow ecklonia wave, echo=FALSE, fig.cap='\\label{fig:shallow ecklonia wave}Boxplot of morphological characteristics for shallow population *E.
+maxima* grouped by wave exposure.', fig.height=7, fig.width=8, warning=FALSE}
 ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3,nrow = 3)
 ```
+
+### Abiotic drivers of *L. pallida* morphological characteristics
+
+Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of *L.
+pallida* morphological characteristics with an adjusted *R*2 of 0.49, explaining 72% of the variation (global permutation test on final model: d.f = 1, *F* = 8.51, *p* = 0.008).
+RDA1 was the only significant axis in the model and explained 48% of the variation while RDA2 was not significant and explained 63% of the variation.
+Total length, stipe length, stipe diameter and stipe mass were postivley influenced (i.e., increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* for kelps mostly from the western region and negativley influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp* for kelps from the eastern region.
+Although sites from from the western region did not cluster as closely as eastern region study sites, grouping according to region was still evident.
+
+### Abiotic drivers of *E. maxima* morphological characteristics
 
@@
 
-### Abiotic drivers of *L. pallida* morphological characteristics
-
-Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of *L.
-pallida* morphological characteristics with an adjusted *R*2 of 0.49, explaining 72% of the variation (global permutation test on final model: d.f = 1, *F* = 8.51, *p* = 0.008).
-RDA1 was the only significant axis in the model and explained 48% of the variation while RDA2 was not significant and explained 63% of the variation.
-Total length, stipe length, stipe diameter and stipe mass were postivley influenced (i.e., increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* for kelps mostly from the western region and negativley influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp* for kelps from the eastern region.
-Although sites from from the western region did not cluster as closely as eastern region study sites, grouping according to region was still evident.
-
 ```{r rda deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
@@
 ```
-
-### Abiotic drivers of *E. maxima* morphological characteristics
-
-Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of deep *E.
-maxima* morphological characteristics with an adjusted *R*2 of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, *F* = 2.4, *p* = 0.026).
-RDA1 was the only significant axis in the model (p = 0.03) and explained 72% of the variation.
-The non-significant axis for the model explained 17% of the variation.
-Primary length, primary width and stipe circumference were positivley influenced by *hs_mean*, *tp_mean* and *dir_median* for for kelps from the western region.
-The remaining morphological characteristics were postivley inlfuenced by *min_temp* and negatively influenced by *tp_sd* for kelps from the eastern region.
 
@@
 
-```{r rda eck plots, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of deep *E.
+maxima* morphological characteristics with an adjusted *R*2 of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, *F* = 2.4, *p* = 0.026).
+RDA1 was the only significant axis in the model (p = 0.03) and explained 72% of the variation.
+The non-significant axis for the model explained 17% of the variation.
+Primary length, primary width and stipe circumference were positivley influenced by *hs_mean*, *tp_mean* and *dir_median* for for kelps from the western region.
+The remaining morphological characteristics were postivley inlfuenced by *min_temp* and negatively influenced by *tp_sd* for kelps from the eastern region.
+
+```{r rda plots, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="\\label{fig:rda plots}RDA output for *L.
+pallida* (A), deep *E.
+maxima* (B) and shallow *E.
+maxima* (C)."}
 ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE, legend = "bottom", labels = "AUTO", ncol = 2, nrow = 2)
@@
 When grouped by wave exposure category, *L.
-pallida* characteristics showed a significant increase in length for the 'Extremely sheltered' category (6-10 kW/m; figure ??) compared to the fully sheltered category, which maybe a 'go-with-the-flow' tactic.
+pallida* characteristics showed a significant increase in length for the 'Extremely sheltered' category (6-10 kW/m; figure \ref{fig:wave power map}) compared to the fully sheltered category (<5 kW/m; see figure \ref{fig:wave power map}), which maybe a 'go-with-the-flow' tactic.
 Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny2002-ob; @Hurd2000].
@@
 When grouped to the higher wave exposure levels, *E.
-maxima* morphology exhibited a strength and flexibility increasing trait (see figure ??).
+maxima* morphology exhibited a strength and flexibility increasing trait (see figure \ref{fig:deep ecklonia wave}).
 Furthermore, the morphological response of *E.
@@
 False Bay has comparativley little upwelling events compared to the western side of the peninsula, and so nutrients may be a limiting factor for *E.
-maxima* populations within False Bay (see figure ??) and hence the identification of minimum temperature (*min_temp*) as a driver.
+maxima* populations within False Bay, hence the identification of minimum temperature (*min_temp*) as a driver.
 Vertical mixing of the water coloumn for inshore regions are driven by several abiotic processes: (1) Turbulence of breaking waves inside or outside of the surf zone (2) convective mixing through a combination of cooling and evaporation (3) wind driven currents and (4) tidal mixing [@Smit2013a].

### 2019-08-29 morphology_paper.Rmd
tiny change in statistical methods section

@@
 Morphological characteristics were categorised according to temperature regime and wave exposure category for each each species and population of kelp in this study.
-Significant differences between temperatre regime and wave exposure categories were tested with wilcox tests.
-Notched boxplots allowed comparison between individual categories of wave exposure.
+Significant differences between temperatre regime and wave exposure categories were tested with wilcox tests and notched boxplots allowed comparison between individual categories of wave exposure.
 In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package [@Oksanen2015-vj] in R (R Core Team, 2019).

### 2019-08-30 morphology_paper.Rmd
commit before testing of inclusion of wave power map
updated pdf output for literature review, commited minor changes to rmd file and added to frontiers of marine science pdf template to the 'markdown' folder

@@
 
-```{r wave power map, echo=FALSE, eval= FALSE, fig.cap='\\label{fig:wave power map}Wave power map'}
+```{r wave power map, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:wave power map}Wave power map'}
 
@@
 
-```{r FB spectrum, echo=FALSE, eval= FALSE, fig.cap='\\label{fig:FB spectrum}Wave power map'}
+```{r FB spectrum, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:FB spectrum}Wave power map'}
 

### 2019-09-04 morphology_paper.Rmd
general house cleaning and output updated

--- 2019-08-29 morphology_paper.Rmd
@@
 output:
-  pdf_document: default
   word_document:
     reference_docx: frontiers_template.docx
+  pdf_document: default
 geometry: margin=1in

### 2019-09-12 morphology_paper.Rmd
added references to literature review. was still not happy with intro for morphology paper so added more and changed it a fair amount. docx output updated

@@
 
-Temperature and wave exposure have been shown to play a role in driving kelp morphology.
-For example, a study by @Serisawa2002-dc compared the morphology of *Ecklonia cava* Kjellman growing in warm temperate and cool temperate morphologies.
-The results showed that wrinkles in the blade and reduced size to be a characteristic of warm temperate regions.
-However this study did not take into account the interaction with other environmental variables such as wave exposure.
-Wave exposure has also been shown to be the main drivers of morphological adaptation across various kelp species.
-Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002-ob; @Hurd2000; @Fowler-Walker2005, @Thomsen2004-yd].
+Given kelps are cool/temperate organisms which are vulernable to dislodgment under high wave exposure scenarios, changes in ocean temperature and an increase in storm frequency and magnitude pose the biggest direct threats to their survival.
+Warmer temperatures reduces the resilience of kelp individuals to perbutations [@wernberg2010], causes fragmentation through weakening tissues [@simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
+Despite these threats and disturbances kelps persist which across a broad range of environments which is largely due to their morphological plasticity [@fowler2006; @reed2003].
+A study by @Wernberg2003-mp investigated the morphology of *E.
+radiata* in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australasian coast.
+They found no correlation between spatial distance and morphological similarity and rather the morphology of kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales [@Wernberg2003-mp].Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
+Deep and shallow water environments may differ in abiotic processes such as the degree of water column mixing [@Smit2013a], solar heating [@Dellatorre2012; @Dunne2001] and effects of wave dampening [@Kobayashi1993-os; Mia2013; @Dubi1995].
+
+Although other factors play a role in kelp morphological adaptation, wave exposure and temperature have been identified as the main drivers of morphological adaptation across various kelp species [@bekkby2014; @wernberg2005; @Denny2002-ob; @Hurd2000; @Fowler-Walker2005; @Thomsen2004-yd; @mabin2013].
+For example, a study by @Serisawa2002-dc compared the morphology of *Ecklonia cava* Kjellman growing at warmer sites and cooler sites, and showed wrinkling of the blade and reduced size to be a characteristic of higher temperature ranges.
+Temperature affects important kelp physiological processes such as photosynthesis and respiration, which ultimately influences growth and productivity resulting in reduced size [@bearham2013; @gerard1997; @gao2013; @zimmerman1986].
+Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002-ob; @Hurd2000; @Fowler-Walker2005; @Thomsen2004-yd].
 For example, a study by @Wernberg2005c examined the consistency of wave exposure as a driver of *Ecklonia radiata*  (C.
 Agardh) J.
-Agardh across a broad geographic range, and  showed trends towards drag-reducing (small size, narrow laterals and blades, low spinosity) and increased strength (large holdfast, thick stipe and thick blades and lamina) at high wave exposure sites.
-The morphological adaptation to wave exposure must be kept in balance with other important processes, such as nutrient assimilation and light absorption, which are dependent on the amount of surface area of the blades.
-Therefore, there is a trade-off between reducing overall drag and maintaining nutrient and photosynthetic ability.
-The morphological adapatibility of kelps is driven by a combination of environmental factors that in turn do not act independently of one another.
-A study by @Wernberg2003-mp investigated the morphology of *E.
-radiata* in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australasian coast.
-They found no correlation between spatial distance and morphological similarity and rather the morphology of kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales [@Wernberg2003-mp].
-Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
-Deep and shallow water environments may differ in abiotic processes such as the degree of water coloumn mixing [@Smit2013a], solar heating [@Dellatorre2012; @Dunne2001] and effects of wave dampening [@Kobayashi1993-os; Mia2013; @Dubi1995], all of which do not act idenpendently and instead influence one another.
+Agardh.
+across a broad geographic range, and showed trends towards drag-reducing (small size, narrow laterals and blades, low spinosity) and increased strength (large holdfast, thick stipe and thick blades and lamina) at high wave exposure sites.
 
@@
 pallida* dominates deeper waters while *E.
-maxima* forms dense surface canopies in from the subtidal to shallower waters.
+maxima* forms dense surface canopies in the subtidal to shallower waters.
 Therefore, *E.
-maxima* is exposed to variations in wind drive surface waves, swell and temperature, while *L.
+maxima* is exposed to variations in wind driven surface waves, swell and temperature, while *L.
 pallida* is exposed to variations in swell and temperature.
@@
 pallida* become increasingly hollow further north along the coastline.
-They suggested that turbidity was the environmental driver responsible for this change.
-
-Measures of wave exposure in ecological studies often incorporate integrative measures of hydrodynamic conditions at a particular site which are based on cartographical models of wave exposure.
-Cartographical models of wave exposure are regarded as 'fetch-based-models' which measure the length of open water associated with a particular site in a straight line, and are regarded as simple measures of wave exposure.
-Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitave, reproducible approach for measuring wave exposure.
-Currently, no ecological studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
-Previous research investigating drivers of kelp morphological characteristics have often used qualitative estimates of wave exposure and did not consider other morphological characteristics.
-Furthermore, wind is an important component of surface gravity waves [@Holthuijsen2010-vl] and previous work has not differentiated between surface gravity waves and swell waves, which the current work will aim to achieve.
-Other considerations such as possible differences in morphology between sub-canopy and canopy species, shallow and deep water populations have not been considered in previous research which ignores the effects of wave damping.
+They suggested that turbidity in relation to light attenuation was the environmental driver responsible for this change.
+
+Measures of wave exposure in ecological studies often incorporate integrative measures of hydrodynamic conditions at a particular site which are based on cartographical models of exposure.
+Cartographical models of wave exposure are 'fetch-based-models' which measure the length of open water associated with a particular site in a straight line, and are regarded as simple measures of wave exposure.
+Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitave, reproducible approach for quantifying wave exposure.
+Although @mabin2013 investigated kelp morphological characteristics using a 2 dimensional WAM model, currently no studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
+Furthermore no current studies have identified specific wave metrics in driving kelp morphological charactersitcs using advanced numerical model, which the current work will aim to achieve.
+Other considerations such as possible differences in morphology between sub-canopy and canopy species, shallow and deep water populations have not been considered using advanced numerical and statisical tools.
 The aim of this study is, therefore, to understand how temperature and wave exposure can influence the morphology in two species of kelps around South Africa, as well as between deep and shallow populations of the same species.
@@
 maxima* and *L.
-pallida* using multiple statistical tools.
+pallida* using advanced statistical tools.
 

### 2019-10-02 morphology_paper.Rmd
general updates and additions. added customised reference list. added in new tables which can be found under the taables subfolder

@@
     reference_docx: frontiers_template.docx
-  pdf_document: default
+  pdf_document: 
+    fig_caption: yes
 geometry: margin=1in
@@
 knitr::opts_chunk$set(cache=FALSE)
+knitr::opts_chunk$set(fig.align = "center")
 ```
@@
 
-Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera [@Bolton2010], nevertheless form the basis of one of the most productive ecosystems globally [@Mann1973].
+Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera [@Bolton2010], form the basis of one of the most productive ecosystems globally [@Mann1973].
 Kelps generally have a dependence on cool, temperate and arctic seawater temperatures [@Bolton2010; @Mohring2014-bk; @Cavanaugh2011-ib; @Dayton1985], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
@@
 Given kelps are cool/temperate organisms which are vulernable to dislodgment under high wave exposure scenarios, changes in ocean temperature and an increase in storm frequency and magnitude pose the biggest direct threats to their survival.
-Warmer temperatures reduces the resilience of kelp individuals to perbutations [@wernberg2010], causes fragmentation through weakening tissues [@simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
-Despite these threats and disturbances kelps persist which across a broad range of environments which is largely due to their morphological plasticity [@fowler2006; @reed2003].
+Warmer temperatures reduces the resilience of kelp individuals to perbutations [@wernberg2010], causes fragmentation through weakening tissues [@Simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
+Despite these threats and disturbances kelps persist across a broad range of environments which is largely due to their morphological plasticity [@FowlerWalker2006; @wernberg2010].
 A study by @Wernberg2003-mp investigated the morphology of *E.
 radiata* in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australasian coast.
-They found no correlation between spatial distance and morphological similarity and rather the morphology of kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales [@Wernberg2003-mp].Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
+They found no correlation between spatial distance and morphological similarity, and rather the morphology of kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales [@Wernberg2003-mp].
+Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
 Deep and shallow water environments may differ in abiotic processes such as the degree of water column mixing [@Smit2013a], solar heating [@Dellatorre2012; @Dunne2001] and effects of wave dampening [@Kobayashi1993-os; Mia2013; @Dubi1995].
 
-Although other factors play a role in kelp morphological adaptation, wave exposure and temperature have been identified as the main drivers of morphological adaptation across various kelp species [@bekkby2014; @wernberg2005; @Denny2002-ob; @Hurd2000; @Fowler-Walker2005; @Thomsen2004-yd; @mabin2013].
-For example, a study by @Serisawa2002-dc compared the morphology of *Ecklonia cava* Kjellman growing at warmer sites and cooler sites, and showed wrinkling of the blade and reduced size to be a characteristic of higher temperature ranges.
+Although other factors play a role in kelp morphological adaptation, wave exposure and temperature have been identified as the main drivers of morphological adaptation across various kelp species [@bekkby2014; @Wernberg2005c; @Denny2002; @Hurd2000; @FowlerWalker2005; @Thomsen2004; @mabin2013].
+For example, a study by @Serisawa2002-dc compared the morphology of *Ecklonia cava* Kjellman growing at warmer and cooler sites, and showed wrinkling of the blade and reduced size to be a characteristic of higher temperature ranges.
 Temperature affects important kelp physiological processes such as photosynthesis and respiration, which ultimately influences growth and productivity resulting in reduced size [@bearham2013; @gerard1997; @gao2013; @zimmerman1986].
-Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002-ob; @Hurd2000; @Fowler-Walker2005; @Thomsen2004-yd].
+Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002; @Hurd2000; @FowlerWalker2005; @Thomsen2004].
 For example, a study by @Wernberg2005c examined the consistency of wave exposure as a driver of *Ecklonia radiata*  (C.
@@
 Cartographical models of wave exposure are 'fetch-based-models' which measure the length of open water associated with a particular site in a straight line, and are regarded as simple measures of wave exposure.
-Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitave, reproducible approach for quantifying wave exposure.
-Although @mabin2013 investigated kelp morphological characteristics using a 2 dimensional WAM model, currently no studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
+Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitave, reproducible approach for characterising the hydrodynamic environment.
+Although @mabin2013 investigated kelp morphological characteristics using a 2D WAM model, currently no studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
 Furthermore no current studies have identified specific wave metrics in driving kelp morphological charactersitcs using advanced numerical model, which the current work will aim to achieve.
@@
 The aim of this study is, therefore, to understand how temperature and wave exposure can influence the morphology in two species of kelps around South Africa, as well as between deep and shallow populations of the same species.
+The current study will also aim to identify specific temperature and wave metrics which drive kelp morphological characteristics.
 This will be achieved by initially understanding the variation in temperature and waves using an advanced numerical model.
@@
 Helena Bay and Betty's Bay constituted the north western and south eastern boundary sites, respectively.
-These sites are roughly 300 km apart, and lie within separate marine provinces and cover a cline in wave exposure and temperature.
-Study sites span across the majority of the south-west coast, in varying thermal and wave energy regimes.
+These sites are roughly 300 km apart, lie within separate marine provinces, and span across the majority of the south-west coast, in varying thermal and wave energy regimes.
 The region is dominated by kelp communities that persist in contrasting abiotic environments.
-The west coast region has been termed as a cool temperate region, which is defined as a region where mean monthly temperatures are always above 10°C and always below 15°C [@Smit2013a].
+The west coast region has been termed as a cool temperate, which is defined as a region where mean monthly temperatures are always above 10°C and always below 15°C [@Smit2013a].
 East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone [@Smit2013a].
@@
 library(grid)
+library(gridExtra)
 
@@
 
+The annual and seasonal patterns in both wind influence the thermal regime around the coastline by inducing upwelling.
+Upwelling brings deep, cool nutrient rich water to coastal areas and therefore induces a decrease in temperature [@Field1980a; @Andrews1980; @Gill1974-nt; @Cram1970-jw; @Rouault2010-gi; @Blanke2002].
+The wind speed and wind direction help drive upwelling in summer on the western side of the Cape peninsula, where southerly winds blow parallel to the coast and trigger upwelling [@Field1980a; @Rouault2010-gi].
+This is not the case for False Bay which is shielded from the dominating wind and swell the Cape Peninsula is subject to, with the exception being the winter months where upwelling becomes more favourable for the region due to the change in swell and wind direction, causing lower temperatures to dominate during this time [@Rouault2010-gi; @Andrews1980].
 Due to the Cape Peninsula temperate latitude, winter months bring an increased frequency of frontal depressions that originate from the Southern Ocean [@Reason2006].
@@
 Helena Bay to Betty's Bay [@Field1980a; @Jury1985-nr; @Andrews1980; @Jury1980-ov].
-Patterns are also evident in the wave data.
-Seasonal variations in significant wave height (Hs), peak period and wave direction are also present.
-The direction of the dominant swell changes to the south west in winter, generated by strong low pressures that originate from the southern ocean [@Reason2006], which False Bay is sheltered from [@Shipley1964; @Atkins1970-bx; @Dufois2012-vy].
-In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an increased variability of Hs and Tp in this region.
-It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world [@Norderhaug2012-ks; @Leliaert2000a].
-It is due to the near consistent south-westerly swell and the complex orography around the peninsula that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.
 
@@
 
-img <- readPNG("figures/kelp_diagram.png")
-grid.raster(img)
+schematic <- readPNG("figures/kelp_diagram.png")
+grid.raster(schematic)
 ```
@@
 
-```{r Temperature data prep, message=FALSE, warning=FALSE, include=FALSE}
+```{r Temperature data prep, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
 ## Data wrangling and preparation
@@
 
-```{r Temperature variable processing, message=FALSE, warning=FALSE, include=FALSE}
+```{r Temperature variable processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
 ### Plots for temperature variables 
@@
   geom_boxplot() + 
-  stat_compare_means(label.y = 14.5 ,label.x = 1.5, label = "p.format") +
+  stat_compare_means(label.x = 2.5, label.y = 12.5, label = "p.format") +
   xlab("") + 
@@
   theme_bw()
-
-```
-
-```{r Temperature variables plot, echo=FALSE, warning=FALSE,fig.height=5,fig.width=8, fig.cap='\\label{fig:Temperature variables plot}Temperature variables at the collection sites around the Cape Peninsula are represented in plot A.
+```
+
+```{r Temperature variables plot, echo=FALSE, warning=FALSE, eval=TRUE,fig.height=5,fig.width=8, fig.cap='\\label{fig:Temperature variables plot}Temperature variables at the collection sites around the Cape Peninsula are represented in plot A.
 Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.
@@
 
-```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, eval=TRUE}
 ## Data wrangling and preparation
@@
 
-```{r wave variable processing, message=FALSE, warning=FALSE, include=FALSE}
+```{r wave variable processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
 ### Plots for wave variables
@@
   geom_boxplot(show.legend = FALSE) + 
-  stat_compare_means(label.y =1.4,label.x = 1.5, label = "p.format") + 
+  stat_compare_means(label.y =0.5 ,label.x = 2.5, label = "p.format") + 
   xlab("") + 
@@
   geom_boxplot(show.legend = FALSE) + 
-  stat_compare_means(label.y = 10, label.x = 1.5, label = "p.format") + 
+  stat_compare_means(label.y = 8.1, label.x = 2.5, label = "p.format") + 
   xlab("") + 
@@
 
-```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE,fig.height=7,fig.width=8, fig.cap='\\label{fig:wave variable plot}In plot A1 and A2 wave variables at each collection site around the Western Cape coast are presented.
+```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE, eval=TRUE,fig.height=7,fig.width=8, fig.cap='\\label{fig:wave variable plot}In plot A1 and A2 wave variables at each collection site around the Western Cape coast are presented.
 Mean variables are represented by dots and standard deviation by whiskers.
@@
 
+library(png)
+library(grid)
+library(gridExtra)
+
 img_wave_rose <- readPNG("figures/wave_rose.png")
 grid.raster(img_wave_rose)
+
 ```
@@
 library(grid)
-
-img_wavepower <- readPNG("figures/CapePen_Total.PNG")
-legend_wave_power <- readPNG("figures/CapePen_Legend.PNG")
-grid.raster(img_wavepower,legend_wave_power)
+library(gridExtra)
+
+wp_1 <-  rasterGrob(as.raster(readPNG("figures/CapePen_Total.PNG")), interpolate = FALSE)
+wp_2 <-  rasterGrob(as.raster(readPNG("figures/CapePen_Legend.PNG")), interpolate = FALSE)
+grid.arrange(wp_1, wp_2, ncol = 2, widths = c(2,1))
+
 ```
@@
 
-```{r FB spectrum, echo=FALSE, eval= FALSE, fig.cap='\\label{fig:FB spectrum}Wave power map'}
+```{r FB spectrum, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:FB spectrum}False Bay wave spectrum'}
 
@@
 library(grid)
-
-img_fb <- readPNG("figures/falsebay_spectrum.png")
+library(jpeg)
+
+img_fb <- readJPEG("figures/falsebay_spectrum.jpg")
 grid.raster(img_fb)
+
 ```
@@
 
-```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, eval=TRUE}
 #custom temperature categories position on x-axis
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   xlab("") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("Lamina weight (g)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("Lamina length (cm)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("Lamina thickness") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("Stipe diameter (cm)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("No. of digits") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("Thallus mass (kg)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("Total length (cm)") + 
@@
 ```{r laminaria temp plots, echo=FALSE, fig.cap='\\label{fig:lamanaria temp plots}Boxplots of *L.
-pallida* morphological characteristics group by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
+pallida* morphological characteristics group by temperature regime.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
 ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
@@
 ```{r lamanaria wave plots, echo=FALSE, fig.cap='\\label{fig:lamanaria wave plots}Boxplots of *L.
-pallida* morphological characteristics group by wave exposure category.', fig.height=7, fig.width=8, warning=FALSE}
+pallida* morphological characteristics group by wave exposure category.', fig.height=8, fig.width=8, warning=FALSE, eval=TRUE}
 ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
@@
 
-```{r deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
+```{r deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
 ## Preparing data
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Frond mass (g)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Primary length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Primary width (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Frond length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe circ (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("No. of tufts") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Epi length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format") + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Total length (cm)") +
@@
 ```{r deep ecklonia temp plot, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E.
-maxima* grouped by temperature.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
+maxima* grouped by temperature.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
 ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", legend = FALSE, ncol = 3, nrow = 4)
@@
 ```{r deep ecklonia wave, echo=FALSE, fig.cap='\\label{fig:deep ecklonia wave}Boxplot of morphological characteristics for deep population *E.
-maxima* grouped by wave exposure.', fig.height=8, fig.width=8, warning=FALSE}
+maxima* grouped by wave exposure.', fig.height=8, fig.width=8, warning=FALSE, eval=TRUE}
 ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3, nrow = 4)
@@
 
-```{r shallow ecklonia processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+```{r shallow ecklonia processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, eval=TRUE}
 EST1 <- ggplot(data = eck_shall, aes(x = temp_regime, y = frond_mass, fill = temp_regime)) +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Frond mass (g)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Primary length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Primary width (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Frond length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe circ (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("No. of tufts") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Epi length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Total length (cm)") +
@@
 ```{r shallow ecklonia temp, echo=FALSE, fig.cap='\\label{fig:shallow ecklonia temp}Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
+maxima* grouped by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
 ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom", ncol = 3, nrow = 4)
@@
 ```{r shallow ecklonia wave, echo=FALSE, fig.cap='\\label{fig:shallow ecklonia wave}Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by wave exposure.', fig.height=7, fig.width=8, warning=FALSE}
+maxima* grouped by wave exposure.', fig.height=7, fig.width=8, warning=FALSE, eval=TRUE}
 ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3,nrow = 3)
@@
 
-```{r rda analysis lamanaria, message=FALSE, warning=FALSE, include=FALSE}
+```{r rda analysis lamanaria, message=FALSE, warning=FALSE, include=FALSE, eval= TRUE}
 #data processed in the temperature and wave sections will be used here eg. abio_clim
@@
 # Plot RDA
-lam_rda <- ggord(rda_lam, meta_lam_rda$region,ellipse = TRUE, hull = FALSE, grp_title = "Region", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9) 
-```
-
-```{r rda deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
+lam_rda <- ggord(rda_lam, meta_lam_rda$region,ellipse = FALSE, hull = TRUE, grp_title = "Region", ptslab = TRUE, parse = TRUE, addsize = 3.5, repel = TRUE, ylims = c(-1.1,1.1), xlims = c(-1.1,1.1)) 
+```
+
+```{r rda deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
 
@@
 # Plot RDA
-rda_eck_deep <- ggord(rda_eck, eck_meta_rda$region ,ellipse = TRUE, hull = FALSE, ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  
-```
-
-```{r rda shallow ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}
+rda_eck_deep <- ggord(rda_eck, eck_meta_rda$region ,ellipse = FALSE, hull = TRUE, ptslab = TRUE, parse = TRUE, addsize = 3.5, repel = TRUE, ylims = c(-1.1,1.1), xlims = c(-1.1,1.1))  
+```
+
+```{r rda shallow ecklonia processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
 
@@
 # Plot RDA
-rda_eck_shallow <- ggord(rda_eck_shallow,meta_shallow_rda$region,ellipse = TRUE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9, coord_fix = TRUE)
-
+rda_eck_shallow <- ggord(rda_eck_shallow,meta_shallow_rda$region,ellipse = FALSE, hull = TRUE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, coord_fix = TRUE, repel = TRUE, facet = FALSE, ylims = c(-1.1,1.1), xlims = c(-1.1,1.1))
 ```
@@
 
-```{r rda plots, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="\\label{fig:rda plots}RDA output for *L.
+```{r rda plots, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE, fig.cap="\\label{fig:rda plots}RDA output for *L.
 pallida* (A), deep *E.
@@
 maxima* (C)."}
-ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE, legend = "bottom", labels = "AUTO", ncol = 2, nrow = 2)
+ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, legend = "none", labels = "AUTO", ncol = 2, nrow = 2)
 ```
@@
 
-This study considered important drivers of kelp morphology that have been identified in previous research [@FowlerWalker2006; @Miller2011-pm; @Fowler-Walker2005; @Stewart2009; @Wing2007-tv; @Molloy1996-la; @Pedersen2012-vr; @Wernberg2005c; @Thomsen2004-yd], namely temperature and wave exposure, around the Cape Peninsula on two prolific kelp species.
+This study considered important drivers of kelp morphology that have been identified in previous research [@FowlerWalker2006; @Miller2011-pm; @FowlerWalker2005; @Stewart2009; @Wing2007-tv; @Molloy1996-la; @Pedersen2012-vr; @Wernberg2005c; @Thomsen2004], namely temperature and wave exposure, around the Cape Peninsula on two prolific kelp species.
 The complex geomorphology of the western Cape coastline creates an ideal natural laboratory for studies of kelps in interaction with their environment.
@@
 
-The annual and seasonal patterns in both wind influence the thermal regime around the coastline by inducing upwelling.
-Upwelling brings deep, cool nutrient rich water to coastal areas and therefore induces a decrease in temperature [@Field1980a; @Andrews1980; @Gill1974-nt; @Cram1970-jw; @Rouault2010-gi; @Blanke2002].
-The wind speed and wind direction help drive upwelling in summer on the western side of the Cape peninsula, where southerly winds blow parallel to the coast and trigger upwelling [@Field1980a; @Rouault2010-gi].
-This is not the case for False Bay which is shielded from the dominating wind and swell the Cape Peninsula is subject to, with the exception being the winter months where upwelling becomes more favourable for the region due to the change in swell and wind direction, causing lower temperatures to dominate during this time [@Rouault2010-gi; @Andrews1980].
+There were clear patterns and clines in both the temperature and wave exposure data.
+The thermal and wave exposure regimes around the Cape Peninsula are driven by a complex interaction between wind, temperature and wave metrics, which do not act independently but instead influence one another.
+The direction of the dominant swell changes to the south west in winter, generated by strong low pressures that originate from the southern ocean [@Reason2006], which False Bay is sheltered from [@Shipley1964; @Atkins1970-bx; @Dufois2012-vy].
+In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an increased variability of Hs and Tp in this region.
+It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world [@Norderhaug2012-ks; @Leliaert2000a].
+It is due to the near consistent south-westerly swell and the complex orography around the peninsula that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.
 The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in figure \ref{fig:wave power map}.
@@
 There are also differences between the types of waves that each side of the peninsula experiences.
-The ????
-analysis shows the western side of the peninsula is not only more exposed but the hydrodynamics are largely swell driven, while on the eastern side most of the wave energy is driven by a combination of swell and wind with the latter being the predominant driver.
+The wave analysis shows the western side of the peninsula is not only more exposed but the hydrodynamics are largely swell driven, while on the eastern side most of the wave energy is driven by a combination of swell and wind with the latter being the predominant driver.
 
@@
 Morphological adaptation due to water motion may manifest in a number of ways in high wave energy environments.
-For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment [@FowlerWalker2006; @Wernberg2005c; @Friedland1995-vi; @Denny2002; @Bekkby2014; @Denny1997; @Denny2002-ob] have been identified in previous studies.
+For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment [@FowlerWalker2006; @Wernberg2005c; @Friedland1995-vi; @Denny2002; @bekkby2014; @Denny1997] have been identified in previous studies.
 Although this study did not measure force of attachment, other morphological responses to temperature and wave variables were evident.
@@
 pallida* characteristics showed a significant increase in length for the 'Extremely sheltered' category (6-10 kW/m; figure \ref{fig:wave power map}) compared to the fully sheltered category (<5 kW/m; see figure \ref{fig:wave power map}), which maybe a 'go-with-the-flow' tactic.
-Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny2002-ob; @Hurd2000].
-However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under higher exposure levels [@Denny2002-ob; @Denny1997].
+Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny2002; @Hurd2000].
+However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under higher exposure levels [@Denny2002; @Denny1997].
 This is reflected in Figure ??
 which shows an overall reduction of *L pallida* morphological characteristics suggesting a size reducing tactic to cope with higher levels of wave exposure.
-The morphological characteristics of *E.
-maxima* show similar patterns when grouped to wave exposure.
 The results suggest that deep *E.
-maxima* populations exhibit a different response as well as a higher wave exposure threshold.
+maxima* populations exhibit a different response, as well as a higher wave exposure threshold.
 Unlike *L.
@@
 maxima* exhibits a size reducing tactic with in an increase in exposure level by significantly decreasing stipe mass, stipe length, stipe circumference, total length and frond length.
-A magnitude related reponse has been suggested by @Wernberg2010 who investigated the temporal and spatial variations of various environmental drivers of the kelp *E.
+A magnitude related reponse has been suggested by @wernberg2010 who investigated the temporal and spatial variations of various environmental drivers of the kelp *E.
 radiata*.
@@
 
-A size reduction tactic has been shown before [@FowlerWalker2006; @Denny2002-ob; @Hurd2000; @Fowler-Walker2005] and is regarded as strategy to reduce overall drag on the plant.
+A size reduction tactic has been shown before [@FowlerWalker2006; @Denny2002; @Hurd2000; @FowlerWalker2005] and is regarded as strategy to reduce overall drag on the plant.
 However, *E.
@@
 maxima* populations.
-Increased temperatures corresponded with a reduction in stipe circumference, frond length, stipe length, total length and stipe mass resulting in slim or ‘skinny’ kelp individuals, which supports findings from previous research by [@Mabin2013-zs; @Serisawa2002-dc].
+Increased temperatures corresponded with a reduction in stipe circumference, frond length, stipe length, total length and stipe mass resulting in slim or ‘skinny’ kelp individuals, which supports findings from previous research by [@mabin2013; @Serisawa2002-dc].
 These findings were confirmed by the redundancy analysis which showed the contribution of wave variables as a driver of kelp morphological characterstics was lower compared to temperature variables.
@@
 This suggests that how morphological characteristcs manifest themselves is dependent on the magnitude of wave exposure.
-Furthermore, the results suggest that in the absence of wave exposure, temperatre becomes the dominant morphological driver.
-This is evident from the redundancy analysis for each species and deep populations of *E.
-maxima* which show clustering of study sites according to the region; wave variables drive morphological characteristics for kelps from the western region and temperature variables for kelps from the eastern region.
-Clustering of sites according to region overlapped almost completey for the RDA analysis of shallow *E.
-maxima* suggesting that for shallow kelp the effect of wave exposure is less variable.
 Temperature variables were not a significant driver of kelp morphology in this study.
@@
 
+```{r}
+
+```
+
 ```{r appendix figure 1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=7}
@@
 lam_all_A <- ggplot(data = lam_data, aes(x = site, y = laminaWeight, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 lam_all_B <- ggplot(data = lam_data, aes(x = site, y = laminaLength, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 lam_all_C <- ggplot(data = lam_data, aes(x = site, y = laminaThickness, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 lam_all_D <- ggplot(data = lam_data, aes(x = site, y = stipeMass, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 lam_all_E <- ggplot(data = lam_data, aes(x = site, y = stipeLength, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 lam_all_F <- ggplot(data = lam_data, aes(x = site, y = stipeDiameter, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 lam_all_G <- ggplot(data = lam_data, aes(x = site, y = digits, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 lam_all_H <- ggplot(data = lam_data, aes(x = site, y = thallusMass, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 lam_all_I <- ggplot(data = lam_data, aes(x = site, y = totalLength, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 eck_all_A <- ggplot(data = eck_data, aes(x = site, y = frond_mass, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 eck_all_B <- ggplot(data = eck_data, aes(x = site, y = primary_length , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 eck_all_C <- ggplot(data = eck_data, aes(x = site, y = primary_width , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 eck_all_D <- ggplot(data = eck_data, aes(x = site, y = frond_length , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 eck_all_E <- ggplot(data = eck_data, aes(x = site, y = stipe_mass , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 eck_all_F <- ggplot(data = eck_data, aes(x = site, y = stipe_length , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 eck_all_G <- ggplot(data = eck_data, aes(x = site, y = stipe_circ , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 eck_all_H <- ggplot(data = eck_data, aes(x = site, y = tufts , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 
@@
 eck_all_I <- ggplot(data = eck_data, aes(x = site, y = epi_length , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
            stat_compare_means(method = "kruskal.test", label.x = 5) + 

### 2019-10-07 morphology_paper.Rmd
updated outputs and uploaded bibtex file for lit review

@@
   geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) + 
-  geom_point(data = temp_ann, aes(x = site, y = max_temp, colour = "blue"),show.legend = FALSE) + 
-  geom_point(data = temp_ann, aes(x = site, y = min_temp, colour = "pink"),show.legend = FALSE) + 
+  geom_point(data = temp_ann, aes(x = site, y = max_temp), show.legend = FALSE, shape = 17, size =2) + 
+  geom_point(data = temp_ann, aes(x = site, y = min_temp), show.legend = FALSE, shape = 6, size =2) + 
   xlab("") + 
@@
 temp_boxplot <- ggplot(data = temp_ann_meta, aes(x = region, y = mean_temp, fill  = region)) + 
-  geom_boxplot() + 
+  geom_boxplot(show.legend = FALSE) + 
   stat_compare_means(label.x = 2.5, label.y = 12.5, label = "p.format") +
@@
   coord_flip() + 
-  theme_bw()
+  theme_bw() +
+  scale_fill_grey(start = 0.4, end = 0.8)
 ```
@@
 ```{r Temperature variables plot, echo=FALSE, warning=FALSE, eval=TRUE,fig.height=5,fig.width=8, fig.cap='\\label{fig:Temperature variables plot}Temperature variables at the collection sites around the Cape Peninsula are represented in plot A.
-Temperature variables include minimum represented by blue dots, mean represented by balc dots, maximum by red dots and whiskers standard deviation.
+Temperature variables include minimum represented by inverted triangles, mean represented by balck dots, maximum by black triangles, and whiskers standard deviation.
 Summary of temperature data for sites grouped by region are represented in plot B.
 See appendix for significance level guide.'}
-ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 2, nrow = 1, labels = c("AUTO"), align = c("h"), legend = FALSE)
+ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 2, nrow = 1, labels = c("AUTO"), align = c("hv"), legend = FALSE)
 ```
@@
 
-positions_wave <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
-cols <- c("Betty's Bay" = "red", "Roman Rock"= "red","Kalk Bay"= "red", "A-Frame"= "red", "Miller's Point"= "red", "Baboon Rock"= "red",  "Batsata Rock"= "red","Bordjies Deep"= "red","Buffelsbaai"= "red","Buffelsbaai South"= "red","Olifantsbos"= "blue","Scarborough"= "blue", "Soetwater"= "blue", "Kommetjie"= "blue","Hout Bay"= "blue","Oudekraal"= "blue","Sea Point"= "blue","St Helena Bay"= "blue")
-
-wave_plot_A <- ggplot(data = abio_clim, aes(x = site, y = hs_mean)) +
-  geom_point(aes(colour = factor(site)), show.legend = FALSE) + 
+positions_wave <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Millers Point", "Baboon Rock",  "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
+#cols <- c("Betty's Bay" = "red", "Roman Rock"= "red","Kalk Bay"= "red", "A-Frame"= "red", "Miller's Point"= "red", "Baboon #Rock"= "red",  "Batsata Rock"= "red","Bordjies Deep"= "red","Buffelsbaai"= "red","Buffelsbaai South"= "red","Olifantsbos"= #"blue","Scarborough"= "blue", "Soetwater"= "blue", "Kommetjie"= "blue","Hout Bay"= "blue","Oudekraal"= "blue","Sea Point"= #"blue","St Helena Bay"= "blue") #  scale_colour_manual(values = cols) #script to assign colours to specific sites
+
+wave_plot_1 <- ggplot(data = abio_clim, aes(x = site, y = hs_mean)) +
+  geom_point(show.legend = FALSE, shape = 17, size = 2) + 
   scale_x_discrete(limits = positions_wave) + 
-  scale_colour_manual(values = cols) +
   theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
@@
 
-wave_plot_B <- ggplot(data = abio_clim, aes(x = site, y = tp_mean)) +
-  geom_point(aes(colour = factor(site)), show.legend = FALSE) + 
+wave_plot_2 <- ggplot(data = abio_clim, aes(x = site, y = tp_mean)) +
+  geom_point(show.legend = FALSE, shape = 16, size = 2) + 
   scale_x_discrete(limits = positions_wave) + 
-  scale_colour_manual(values = cols) +
   theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5)) +
@@
   coord_flip() + 
-  theme_bw()
+  theme_bw() +
+  scale_fill_grey(start = 0.4, end = 0.8)
 
@@
 Blue dots represent sites grouped to the western category and red dots represent sites grouped to the eastern category.'}
-ggarrange(wave_plot_A,wave_boxplot_1, wave_plot_B, wave_boxplot_2, common.legend = TRUE, labels = c("A", "C", "B", "D"))
-```
-
-```{r wave rose, echo=FALSE, warning = FALSE, eval= TRUE, fig.cap='\\label{fig:wave rose}Wave rose figure'}
+ggarrange(wave_plot_1,wave_boxplot_1, wave_plot_2, wave_boxplot_2, common.legend = TRUE, labels = c("A", "B", "C", "D"), align = "hv")
+```
+
+```{r wave rose, echo=FALSE, warning = FALSE, eval= TRUE, fig.align="center",fig.cap='\\label{fig:wave rose}Wave rose figure'}
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.62, label.y.npc = c('top')) + 
   xlab("") + 
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.65, label.y.npc = c('top')) + 
   xlab("") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.85, label.y.npc = c('top')) + 
   ylab("Lamina weight (g)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
   ylab("Stipe diameter (cm)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
   ylab("Thallus mass (kg)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
   ylab("Total length (cm)") + 
@@
 pallida* morphological characteristics group by temperature regime.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
-ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
+ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "none", align ='hv')
 ```
@@
 pallida* morphological characteristics group by wave exposure category.', fig.height=8, fig.width=8, warning=FALSE, eval=TRUE}
-ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
+ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "none", align = "hv")
 ```
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 ET2 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 ET3 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 ET4 <- ggplot(data = eck_deep, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 ET5 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 ET6 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 ET7 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 ET8 <- ggplot(data = eck_deep, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 ET9 <- ggplot(data = eck_deep, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 ET10 <- ggplot(data = eck_deep, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
   scale_x_discrete(limits = pos_temp) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = TRUE) + 
   stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
@@
 maxima* grouped by temperature.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
-ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", legend = FALSE, ncol = 3, nrow = 4)
+ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = c("none"), ncol = 3, nrow = 4)
 ```
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 EST2 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 EST3 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
@@
   xlab("") + 
-  ylim(0, 20)
+  ylim(0, 20) +
+  scale_fill_grey(start = 0.8, end = 0.4)
 EST4 <- ggplot(data = eck_shall, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 EST5 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 EST6 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 EST7 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
@@
   xlab("") +
-  ylim(0, 30)
+  ylim(0, 30) +
+  scale_fill_grey(start = 0.8, end = 0.4)
 EST8 <- ggplot(data = eck_shall, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 EST9 <- ggplot(data = eck_shall, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
@@
   theme(axis.text.x = element_blank()) + 
-  xlab("")
+  xlab("") +
+  scale_fill_grey(start = 0.8, end = 0.4)
 EST10 <- ggplot(data = eck_shall, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
@@
 maxima* (C)."}
-ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, legend = "none", labels = "AUTO", ncol = 2, nrow = 2)
+ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, legend = "none", labels = "AUTO", ncol = 2, nrow = 2, align = "hv")
 ```
@@
 It should be noted however, that from these analyses that in general the cool-temperate region is more exposed to waves than that of the warm-temperate region.
-Therefore, the significantly larger morphological characteristics for kelp from the cool-temperate region may overlap with responses to wave exposure which was exhibited strongly in both species.
+Therefore, the significantly larger morphological characteristics for kelp from the cool-temperate region may overlap with responses to wave exposure.
+The response of kelp morphological characteristics to wave exposure was evident and both species show a tactics based on the magnitde of wave exposure.
+Strength increasing traits were exibited for lower exposure levels while a 'go-with-the-flow' tactic for moderate levels of wave exposure.
+Denny and Galord ???
+showed that increased flexiblity or a 'go-with-the-flow' tactic has a threshold value, where increased stipe length actaully increase drag under high wave exposure scenarios.
+Therefore, under high wave exposure scenarios size reducing traits, such as reduced surface area of the overall plant should be favoured instead.
+This was evident in the results which showed that stipe length decreases under high wave exposure categories.
+This suggests that how morphological characteristcs manifest themselves is dependent on the magnitude of wave exposure.
 When grouped by wave exposure category, *L.
@@
 
-Annual maximum and winter mean temperatures influence stipe length and frond length in shallow *E.
+Temperature variables were not a significant driver of kelp morphology in this study.
+Although temperature plays an important role in distribution (Bolton et al.
+2012; Miller et al.
+2011; Rinde et al.
+2014) of kelp and physiological functioning of adults and gametophytes (Mohring et al.
+2014; Steneck et al.
+2002; Gerard 1997; Bearham et al.
+2013; Smale & Moore 2017), there is little evidence that temperature is an important driver of morphological variation.
+However, we suggest that temperature plays a larger role in determining kelp morphology at sites where wave energy is low enough not to elecit a response in morphological characteristics.
+Temperature metrics influence stipe length and frond length in shallow *E.
 maxima* individuals located at sites within False Bay; which was not the case for deeper *E.
@@
 maxima *and *L.
-pallida* populations around the Cape Peninsula, and differences in shallow and deep populations of *E.
-maxima* may reflect small differences in wave and temperature variables as drivers of kelp morphological characteristics.
-The response of kelp morphological characteristics to wave exposure was evident and both species show a tactics based on the magnitde of wave exposure.
-Strength increasing traits were exibited for lower exposure levels while a 'go-with-the-flow' tactic for moderate levels of wave exposure.
-Denny and Galord ???
-showed that increased flexiblity or a 'go-with-the-flow' tactic has a threshold value, where increased stipe length actaully increase drag under high wave exposure scenarios.
-Therefore, under high wave exposure scenarios size reducing traits, such as reduced surface area of the overall plant should be favoured instead.
-This was evident in the results which showed that stipe length decreases under high wave exposure categories.
-This suggests that how morphological characteristcs manifest themselves is dependent on the magnitude of wave exposure.
-Temperature variables were not a significant driver of kelp morphology in this study.
-Although temperature plays an important role in distribution (Bolton et al.
-2012; Miller et al.
-2011; Rinde et al.
-2014) of kelp and physiological functioning of adults and gametophytes (Mohring et al.
-2014; Steneck et al.
-2002; Gerard 1997; Bearham et al.
-2013; Smale & Moore 2017), there is little evidence that temperature is an important driver of morphological variation.
-However, we suggest that temperature plays a larger role in determining kelp morphology at sites where wave energy is low enough not to elecit a response in morphological characteristics.
+pallida* populations around the Cape Peninsula, and differences in morphological characteristics between shallow and deep populations of *E.
+maxima* may reflect differences in hydrodynamic and temperature regimes.
+This study also confirms the findings of previous research which shows morphological adpatation is dependent on the magnitude of wave exposure, but provides quantitative, reproducible measure of wave exposure while identifying the specific wave metrics and wave energy thresholds invloved.
+Furthermore, the current study highlights the need to understand how kelp populations, which are not exposed to high wave energy and morphological characteristics are largely driven by temperature, are vulnerable to changes in storm frequency and magnitude.
+Kelp populations located in low wave energy environments, such as False Bay, may be more vulnerable to increased storms frequency and magnitude in future climate scenarios.
 

### 2019-10-07 morphology_paper.Rmd
moved the last paragraph to earlier in the intro, after the explanation of environmental variables

@@
 output:
+  pdf_document: 
+    fig_caption: yes
   word_document:
     reference_docx: frontiers_template.docx
-  pdf_document: 
-    fig_caption: yes
 geometry: margin=1in
@@
 Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera [@Bolton2010], form the basis of one of the most productive ecosystems globally [@Mann1973].
-Kelps generally have a dependence on cool, temperate and arctic seawater temperatures [@Bolton2010; @Mohring2014-bk; @Cavanaugh2011-ib; @Dayton1985], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
-Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002], each depending on a wide suite of ecological services provided by the kelp forests [@Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @Willis2003-xo; @anderson1997; @Dayton1985].
+Kelps generally have a dependence on cool-temperate and arctic seawater temperatures [@Bolton2010; @Mohring2014-bk; @Cavanaugh2011-ib; @Dayton1985], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
+Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002]; each depending on a wide suite of ecological services provided by kelp forests [@Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @Willis2003-xo; @anderson1997; @Dayton1985].
 Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in the distribution [@Gorman2013-nj], abundance [@Cavanaugh2011-ib; @Dayton1998-xw], diversity [@Wing2007-tv; @Wernberg2008], composition [@Leliaert2000a; @Dayton1985; @Norderhaug2012-ks; @Harley2012], growth [@Cousens1982-ck] and productivity [@Pedersen2012-vr; @Dayton1998-xw; @Krumhansl2012] of kelps.
 
-Given kelps are cool/temperate organisms which are vulernable to dislodgment under high wave exposure scenarios, changes in ocean temperature and an increase in storm frequency and magnitude pose the biggest direct threats to their survival.
+Given kelps are cool-temperate organisms which are vulernable to dislodgment under high wave exposure scenarios, changes in ocean temperature and an increase in storm frequency and magnitude pose the biggest direct threats to their survival.
 Warmer temperatures reduces the resilience of kelp individuals to perbutations [@wernberg2010], causes fragmentation through weakening tissues [@Simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
@@
 
+Measures of wave exposure in ecological studies often incorporate integrative measures of hydrodynamic conditions at a particular site which are based on cartographical models of exposure.
+Cartographical models of wave exposure are 'fetch-based-models' which measure the length of open water associated with a particular site in a straight line, and are regarded as simple measures of wave exposure.
+Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitave, reproducible approach for characterising the hydrodynamic environment.
+Although @mabin2013 investigated kelp morphological characteristics using a 2D WAM model, currently no studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
+Furthermore no current studies have identified specific wave metrics in driving kelp morphological charactersitcs using advanced numerical model, which the current work will aim to achieve.
+Other considerations such as possible differences in morphology between sub-canopy and canopy species, shallow and deep water populations have not been considered using advanced numerical and statisical tools.
+
 The kelps *Laminaria pallida* and *Ecklonia maxima* (Osbeck) Papenfuss are important habitat forming seaweeds that exist around the coast of South Africa and offer a unique opportunity to investigate the drivers of macroalgal morphological characteristics between canopy and sub-canopy species, as well as between deep and shallow water environments.
@@
 Therefore, *E.
-maxima* is exposed to variations in wind driven surface waves, swell and temperature, while *L.
-pallida* is exposed to variations in swell and temperature.
+maxima* is exposed to variations in wind driven surface waves, swell waves and temperature, while *L.
+pallida* is exposed to variations in swell waves and temperature.
 @Molloy1996-la investigated the effect of depth and wave exposure on *L.
@@
 
-Measures of wave exposure in ecological studies often incorporate integrative measures of hydrodynamic conditions at a particular site which are based on cartographical models of exposure.
-Cartographical models of wave exposure are 'fetch-based-models' which measure the length of open water associated with a particular site in a straight line, and are regarded as simple measures of wave exposure.
-Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitave, reproducible approach for characterising the hydrodynamic environment.
-Although @mabin2013 investigated kelp morphological characteristics using a 2D WAM model, currently no studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
-Furthermore no current studies have identified specific wave metrics in driving kelp morphological charactersitcs using advanced numerical model, which the current work will aim to achieve.
-Other considerations such as possible differences in morphology between sub-canopy and canopy species, shallow and deep water populations have not been considered using advanced numerical and statisical tools.
 The aim of this study is, therefore, to understand how temperature and wave exposure can influence the morphology in two species of kelps around South Africa, as well as between deep and shallow populations of the same species.

### 2019-10-08 morphology_paper.Rmd
a few touch ups

@@
 output:
+  word_document:
+    reference_docx: frontiers_template.docx
   pdf_document: 
     fig_caption: yes
-  word_document:
-    reference_docx: frontiers_template.docx
 geometry: margin=1in
@@
 
-Although other factors play a role in kelp morphological adaptation, wave exposure and temperature have been identified as the main drivers of morphological adaptation across various kelp species [@bekkby2014; @Wernberg2005c; @Denny2002; @Hurd2000; @FowlerWalker2005; @Thomsen2004; @mabin2013].
+Although multiple environmental factors play a role in kelp morphological adaptation, wave exposure and temperature have been identified as the main drivers across various kelp species [@bekkby2014; @Wernberg2005c; @Denny2002; @Hurd2000; @FowlerWalker2005; @Thomsen2004; @mabin2013].
 For example, a study by @Serisawa2002-dc compared the morphology of *Ecklonia cava* Kjellman growing at warmer and cooler sites, and showed wrinkling of the blade and reduced size to be a characteristic of higher temperature ranges.
-Temperature affects important kelp physiological processes such as photosynthesis and respiration, which ultimately influences growth and productivity resulting in reduced size [@bearham2013; @gerard1997; @gao2013; @zimmerman1986].
+Warmer temperature effects important kelp physiological processes such as photosynthesis and respiration, which influences growth and productivity and ultimately leads to a reduction in size [@bearham2013; @gerard1997; @gao2013; @zimmerman1986].
 Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002; @Hurd2000; @FowlerWalker2005; @Thomsen2004].
@@
 Although @mabin2013 investigated kelp morphological characteristics using a 2D WAM model, currently no studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
-Furthermore no current studies have identified specific wave metrics in driving kelp morphological charactersitcs using advanced numerical model, which the current work will aim to achieve.
-Other considerations such as possible differences in morphology between sub-canopy and canopy species, shallow and deep water populations have not been considered using advanced numerical and statisical tools.
+Furthermore no current studies have identified specific wave metrics that drive kelp morphological charactersitcs, which the current work will aim to achieve.
+Other considerations such as possible differences in morphology between sub-canopy and canopy species, and shallow and deep water kelp populations have not been considered using advanced numerical and statisical tools.
 
@@
 Therefore, *E.
-maxima* is exposed to variations in wind driven surface waves, swell waves and temperature, while *L.
+maxima* may be exposed to variations in wind driven surface waves, swell waves and temperature; while *L.
 pallida* is exposed to variations in swell waves and temperature.
@@
 pallida* become increasingly hollow further north along the coastline.
-They suggested that turbidity in relation to light attenuation was the environmental driver responsible for this change.
-
-The aim of this study is, therefore, to understand how temperature and wave exposure can influence the morphology in two species of kelps around South Africa, as well as between deep and shallow populations of the same species.
+The authors suggested that turbidity in relation to light attenuation was the environmental driver responsible for this change.
+
+The aim of this study is, therefore, to understand how temperature and wave exposure can influence the morphology of sub-canopy and canopy kelps, as well as between deep and shallow populations of the same species.
 The current study will also aim to identify specific temperature and wave metrics which drive kelp morphological characteristics.
@@
 
-Sites were chosen to represent an array of environmental gradients, as indicated in figure \ref{fig:map study sites}.
+Sites were chosen to represent an array of temperature and wave gradients, as indicated in figure \ref{fig:map study sites}.
 St.
@@
 The region is dominated by kelp communities that persist in contrasting abiotic environments.
-The west coast region has been termed as a cool temperate, which is defined as a region where mean monthly temperatures are always above 10°C and always below 15°C [@Smit2013a].
+The west coast region has been termed as cool temperate, which is defined as a region where mean monthly temperatures are always above 10°C and always below 15°C [@Smit2013a].
 East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone [@Smit2013a].
-The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm temperate region [@Smit2013a].
+The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm temperate [@Smit2013a].
 
@@
 The annual and seasonal patterns in both wind influence the thermal regime around the coastline by inducing upwelling.
-Upwelling brings deep, cool nutrient rich water to coastal areas and therefore induces a decrease in temperature [@Field1980a; @Andrews1980; @Gill1974-nt; @Cram1970-jw; @Rouault2010-gi; @Blanke2002].
-The wind speed and wind direction help drive upwelling in summer on the western side of the Cape peninsula, where southerly winds blow parallel to the coast and trigger upwelling [@Field1980a; @Rouault2010-gi].
+Upwelling brings deep, cool nutrient rich water to coastal areas which induces a decrease in temperature [@Field1980a; @Andrews1980; @Gill1974-nt; @Cram1970-jw; @Rouault2010-gi; @Blanke2002].
+The wind speed and wind direction help drive upwelling in summer on the western side of the Cape peninsula, where southerly winds blow parallel to the coast and trigger upwelling events [@Field1980a; @Rouault2010-gi].
 This is not the case for False Bay which is shielded from the dominating wind and swell the Cape Peninsula is subject to, with the exception being the winter months where upwelling becomes more favourable for the region due to the change in swell and wind direction, causing lower temperatures to dominate during this time [@Rouault2010-gi; @Andrews1980].
@@
 *Laminaria pallida* is characterised by a single smooth blade which is divided longitudinally into sections, and develops from a single meristematic region located at the junction between the blade and the stipe [@Dyer2018-rv].
-This species has a solid stipe but develops a hollow stipe along the west coast northward and into Namibia [@Rothman2017-pn].
+This species has a solid stipe along the south-west coast but develops a hollow stipe along the west coast northward and into Namibia [@Rothman2017-pn].
 *Ecklonia maxima* consists of a single primary blade which develops above a gas-filled float and a hollow stipe below.
-Secondary blades are produced laterally from the primary blade from several meristematic regions along the margins of the primary blade, known as digits [@Dyer2018-rv].
+Secondary blades are produced laterally from the primary blade from several meristematic regions along the margins of the primary blade [@Dyer2018-rv].
 Both species are held to the substrate by finger-like haptera, collectively known as the holdfast.
@@
 1) between October 2014 and April 2015.
-In each instance only the largest individuals were sampled to ensure that only mature and fully grown sporophytes were measured.
-Thirteen samples were collected for each species at each of the sites.
-The morphological traits named in Tables 1 and 2 were measured for comparison between sites.
-Because the macroalgae differ in morphological features, species-specific morphological characteristics were included.
-This allowed comparison between adjacent and non-adjacent sites around the peninsula.
+In each instance only the thirteen largest individuals were sampled to ensure that only mature and fully grown sporophytes were measured.
+In instances where one of the species was not present a site, the nearest population of the missing species was sampled.
+Because the macroalgae differ in morphological features, species-specific morphological characteristics were included and are presented in Tables 1 and 2.
 Between February 2017 and November 2018, morphological measurements for *E.
@@
 
-Summary statistics were calculated for each site and regions for each of the environmental variables considered in this study.
+Summary statistics were calculated for each site and regions for each of the environmental variables considered in this study, and are presented in Table 3.
 The summary statistics calculated for temperature, waves and wind variables were minimum, maximum, mean, range and standard deviation for annual timescales.
-In addition, annual median wave and wind direction were also calculated; all summary data and their respective abbreviation are presented in Table 3.
+In addition, annual median wave and wind direction were also calculated.
 The median was calculated for wind and wave direction, as issues arise when calculating the mean and standard deviation for compass metrics.
@@
 Morphological characteristics were categorised according to temperature regime and wave exposure category for each each species and population of kelp in this study.
-Significant differences between temperatre regime and wave exposure categories were tested with wilcox tests and notched boxplots allowed comparison between individual categories of wave exposure.
+Significant differences between temperature regimes and wave exposure categories were tested with wilcox tests and notched boxplots allowed comparison between individual categories of wave exposure.
 In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package [@Oksanen2015-vj] in R (R Core Team, 2019).

### 2019-10-09 morphology_paper.Rmd
added extra paragraph to temperature results section

@@
 Helena Bay to 0.9 at Olifantsbos, and on the eastern side on of the peninsula was 1.0 at Buffelsbaai to 2.9 at Betty’s Bay.
-As can be seen in figure \ref{fig:Temperature variables plot}A, the trend in the range of temperature decreases from St.
-Helena Bay to Miller’s point and then increases from Kalk Bay to Betty’s Bay.
-Also, annual range in temperatures within False Bay are larger in winter (August) compared to summer (February).
 
@@
 ```
+
+In general, the mean temperature increases for sites located within False Bay and sites located within False Bay also exibited less variation around the mean temperature compared to sites on the western side of the peninsula(\ref{fig:Temperature variables plot}A).
+This is verified by \ref{fig:Temperature variables plot}B, which shows a significanlty lower (p<0.05) median temperature and less variation in temperature for sites located on the western side of the peninsula compared to sites on the eastern side.
 

### 2019-10-09 morphology_paper.Rmd
moved and deteled some sections from results wave section. spelling checked

--- 2019-10-08 morphology_paper.Rmd
@@
 
+In general, the mean temperature increases for sites located within False Bay and sites located within False Bay also exhibited less variation around the mean temperature compared to sites on the western side of the peninsula(\ref{fig:Temperature variables plot}A).
+This is verified by \ref{fig:Temperature variables plot}B, which shows a significantly lower (p<0.05) median temperature and less variation in temperature for sites located on the western side of the peninsula compared to sites on the eastern side.
+
 ### Wave environment
 
-Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula on both an annual and seasonal scale (see figure \ref{fig:wave variable plot}).
+Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula on both an annual scale (see figure \ref{fig:wave variable plot}).
 Mean significant wave height ranged from 0.5 ± 0.2 m (mean ± SD) at St.
@@
 The western region The western region experiences a significantly higher mean significant swell height (p < 0.05; see figure \ref{fig:wave variable plot}C) and mean peak period (p < 0.05; see figure \ref{fig:wave variable plot}D).
-These wave heights and their variability are modulated by the relative sheltering against the predominant south-westerly swell direction and the size of the fetch for local wave generation.
 
@@
 
-```{r wave rose, echo=FALSE, warning = FALSE, eval= TRUE, fig.align="center",fig.cap='\\label{fig:wave rose}Wave rose figure'}
+```{r wave rose, echo=FALSE, warning = FALSE, eval= TRUE, fig.align="center",fig.cap='\\label{fig:wave rose}Wave rose figure.
+Wave variables divided into Annual (top), Austral summer (middle) and Austral winter (bottom) respectively'}
 
@@
 Summer data were the exception and showed lower variation for the eastern side of the peninsula (see figure \ref{fig:wave rose}).
-Wave variables divided into Annual (top), Austral summer (middle) and Austral winter (bottom) respectively.
-This phenomenon can be explained by the sheltering effect of the Cape Peninsula, resulting in a wave climate with a narrower directional spreading on the eastern side of the peninsula (refer to Fig.
-7).
 
@@
 The thermal and wave exposure regimes around the Cape Peninsula are driven by a complex interaction between wind, temperature and wave metrics, which do not act independently but instead influence one another.
+The hydrodynamic regime is modulated by the relative sheltering against the predominant south-westerly swell direction and the size of the fetch for local wave generation.
 The direction of the dominant swell changes to the south west in winter, generated by strong low pressures that originate from the southern ocean [@Reason2006], which False Bay is sheltered from [@Shipley1964; @Atkins1970-bx; @Dufois2012-vy].

### 2019-10-09 morphology_paper.Rmd
added more detailed caption to figure 4 (wave variables plot)

@@
 
-```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE, eval=TRUE,fig.height=7,fig.width=8, fig.cap='\\label{fig:wave variable plot}In plot A1 and A2 wave variables at each collection site around the Western Cape coast are presented.
-Mean variables are represented by dots and standard deviation by whiskers.
-Blue dots represent sites grouped to the western category and red dots represent sites grouped to the eastern category.'}
+```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE, eval=TRUE,fig.height=7,fig.width=8, fig.cap='\\label{fig:wave variable plot}Mean significant wave height and mean significant peak period across sites and by region.
+Mean significant wave height across sites is presented in plot A and is represented by solid triangles and standard deviation by the whiskers.
+Differences in mean significant wave height between regions are presented in plot B.
+Mean significant peak period across sites is presented in plot C and is represented by solid dots and standard deviation by whiskers.
+Differences in mena significant peak period between regions are presented in plot D.'}
 ggarrange(wave_plot_1,wave_boxplot_1, wave_plot_2, wave_boxplot_2, common.legend = TRUE, labels = c("A", "B", "C", "D"), align = "hv")

### 2019-10-09 morphology_paper.Rmd
added more detail to wave rose caption

@@
 
-```{r wave rose, echo=FALSE, warning = FALSE, eval= TRUE, fig.align="center",fig.cap='\\label{fig:wave rose}Wave rose figure.
-Wave variables divided into Annual (top), Austral summer (middle) and Austral winter (bottom) respectively'}
+```{r wave rose, echo=FALSE, warning = FALSE, eval= TRUE, fig.align="center",fig.cap='\\label{fig:wave rose}The total and seasonal wave roses of the directional wave buoy just offshore from Cape Town is given form the years 2000 till end 2017 (the period for which directional wave spectra was available).
+Annual (top left), Austral summer (top right) and Austral winter (bottom) respectively'}
 

### 2019-10-09 morphology_paper.Rmd
added in seasonal wave power (exposure maps) as per Christo's recommendation.

@@
 wp_1 <- rasterGrob(as.raster(readPNG("figures/CapePen_Total.PNG")), interpolate = FALSE)
-wp_2 <-  rasterGrob(as.raster(readPNG("figures/CapePen_Legend.PNG")), interpolate = FALSE)
-grid.arrange(wp_1, wp_2, ncol = 2, widths = c(2,1))
-
-```
-
-In Figure 5 the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
+wp_2 <- rasterGrob(as.raster(readPNG("figures/CapePen_Summer.PNG")), interpolate = FALSE)
+wp_3 <- rasterGrob(as.raster(readPNG("figures/CapePen_Winter.PNG")), interpolate = FALSE)
+wp_4 <- rasterGrob(as.raster(readPNG("figures/CapePen_Spring.PNG")), interpolate = FALSE)
+wp_5 <- rasterGrob(as.raster(readPNG("figures/CapePen_Autumn.PNG")), interpolate = FALSE)
+wp_6 <- rasterGrob(as.raster(readPNG("figures/CapePen_Legend.PNG")), interpolate = FALSE)
+
+grid.arrange(wp_1, wp_2, wp_3, wp_4, wp_5, wp_6, ncol = 3, nrow = 2)
+
+```
+
+In Figure \ref{fig:wave power map} the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
 The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits.
@@
 
-```{r FB spectrum, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:FB spectrum}False Bay wave spectrum'}
+```{r FB spectrum, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:FB spectrum}The propagation of a typical offshore wave spectrum is given as produced from a single time-step in SWAN'}
 

### 2019-10-09 morphology_paper.Rmd
added detail to wave power (exposure) figure caption

@@
 
-```{r wave power map, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:wave power map}Wave power map'}
+```{r wave power map, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:wave power map}The total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length.
+Annual total (A), summer (B), winter (C), spring (D) and autumn (D).'}
 

### 2019-10-09 morphology_paper.Rmd
moved a sentence from results FB wave spectra to discussion. going to test a word output now.

@@
 
-To clarify the averaged wave exposure maps presented, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in figure \ref{fig:FB spectrum}.
+To clarify the averaged wave exposure maps presented in figure \ref{fig:wave power map}, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in figure \ref{fig:FB spectrum}.
 Tracing the wave height contours into False Bay its clear why this bay’s western periphery is predominantly sheltered.
-It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions.
-This results in positive and negative wave exposure anomalies all around the peninsula.
 
@@
 It is due to the near consistent south-westerly swell and the complex orography around the peninsula that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.
-The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in figure \ref{fig:wave power map}.
+The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in figure \ref{fig:wave power map}.It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions.
+This results in positive and negative wave exposure anomalies all around the peninsula.
 Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula [@Rouault2010-gi; @Field1980a].

### 2019-10-09 morphology_paper.Rmd
noticed lables were slighlty off on temp plot. this has been adjusted.

@@
   ylab("") + 
-  theme(plot.title = element_text(hjust = 0)) +
   theme(legend.position = "none") + 
   coord_flip() + 
-  theme_bw()
+  theme_bw() + 
+  theme(axis.text.y=element_text(vjust=0.3))
 
@@
   theme_bw() +
-  scale_fill_grey(start = 0.4, end = 0.8)
+  scale_fill_grey(start = 0.4, end = 0.8) + 
+  theme(axis.text.y=element_text(vjust=0.3)) 
 ```

### 2019-10-09 morphology_paper.Rmd
similar adjustments made to wave plot as per previous commit. Also adjusted p label position in both wave boxplots.

@@
   theme(legend.position = "none") + 
-  theme(plot.title = element_text(hjust = 0)) + 
-  coord_flip() + 
-  theme_bw()
+  coord_flip() + 
+  theme_bw() + 
+  theme(axis.text.y=element_text(vjust=0.3))
 
@@
   theme(legend.position = "none") + 
-  theme(plot.title = element_text(hjust = 0)) + 
-  coord_flip() +
-  theme_bw()
+  coord_flip() +
+  theme_bw() + 
+  theme(axis.text.y=element_text(vjust=0.3))
 
@@
   geom_boxplot(show.legend = FALSE) + 
-  stat_compare_means(label.y =0.5 ,label.x = 2.5, label = "p.format") + 
+  stat_compare_means(label.y =0.6 ,label.x = 2.5, label = "p.format") + 
   xlab("") + 
@@
   theme_bw() +
-  scale_fill_grey(start = 0.4, end = 0.8)
+  scale_fill_grey(start = 0.4, end = 0.8) + 
+  theme(axis.text.y=element_text(vjust=0.3))
 
@@
   geom_boxplot(show.legend = FALSE) + 
-  stat_compare_means(label.y = 8.1, label.x = 2.5, label = "p.format") + 
+  stat_compare_means(label.y = 8.2, label.x = 2.5, label = "p.format") + 
   xlab("") + 

### 2019-10-11 morphology_paper.Rmd
some general housekeeping. decided to keep notches in wave plots.

@@
 Morphological characteristics were categorised according to temperature regime and wave exposure category for each each species and population of kelp in this study.
-Significant differences between temperature regimes and wave exposure categories were tested with wilcox tests and notched boxplots allowed comparison between individual categories of wave exposure.
+Significant differences between temperature regimes and across wave exposure categories were tested with wilcox tests and kruskal tests respectively, and notched boxplots allowed comparison between individual categories of wave exposure.
 In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package [@Oksanen2015-vj] in R (R Core Team, 2019).
@@
 To determine the explanatory variables that best describe patterns in the response data, a full RDA was performed using a complete set of explanatory variables.
-Forward selection was then used to reduce the number of explanatory variables as well as prevent the inflation of overall type I error.
+Forward selection was then used to reduce the number of explanatory variables, as well as prevent the inflation of overall type I error.
 To further improve the model, pairwise coefficients and Variance Inflation Factor (VIF) were calculated to identify variables with high multicollinearity.
-The computation of the parsimonious RDA was followed by permutation tests of the adjusted *R*2 to assess significance of constraints.
+The computation of the parsimonious RDA was followed by permutation tests of the adjusted *R*^2^ to assess significance of constraints.
 
@@
 temp_boxplot <- ggplot(data = temp_ann_meta, aes(x = region, y = mean_temp, fill  = region)) + 
-  geom_boxplot(show.legend = FALSE) + 
+  geom_boxplot(show.legend = FALSE, width = 0.5) + 
+  geom_jitter(width = 0.2) + 
   stat_compare_means(label.x = 2.5, label.y = 12.5, label = "p.format") +
@@
 
-In general, the mean temperature increases for sites located within False Bay and sites located within False Bay also exhibited less variation around the mean temperature compared to sites on the western side of the peninsula(\ref{fig:Temperature variables plot}A).
-This is verified by \ref{fig:Temperature variables plot}B, which shows a significantly lower (p<0.05) median temperature and less variation in temperature for sites located on the western side of the peninsula compared to sites on the eastern side.
+In general, the mean temperature increases for sites located within False Bay with less variation around the mean temperature compared to sites on the western side of the peninsula(\ref{fig:Temperature variables plot}A).
+This is verified by \ref{fig:Temperature variables plot}B, which shows a significantly lower (p < 0.05, figure \ref{fig:Temperature variables plot}B) mean temperature and less variation in temperature for sites located on the western side of the peninsula compared to sites on the eastern side.
 
@@
   geom_boxplot(show.legend = FALSE) + 
+  geom_jitter(width = 0.2) +
   stat_compare_means(label.y =0.6 ,label.x = 2.5, label = "p.format") + 
@@
   geom_boxplot(show.legend = FALSE) + 
+  geom_jitter(width = 0.2) +
   stat_compare_means(label.y = 8.2, label.x = 2.5, label = "p.format") + 
@@
 pallida* morphological characteristics with lamina length, number of digits and thallus mass the exceptions (see figure \ref{fig:lamanaria temp plots}).
-Certian morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significanlty higher means for kelps in cool temperate regimes when compared to kelps for the warm temperate regime.
-Conversely, lamina weight and lamina thickness had significantly lower means for kelps in the cool temperate regime compared to kelps in the warm temperate regime.
+Certian morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significanlty higher means for kelps from the cool temperate regime when compared to kelps for the warm temperate regime.
+Conversely, lamina weight and lamina thickness had significantly lower means for kelps from the cool temperate regime compared to kelps from the warm temperate regime.
 When *L.
 pallida* morphological characteristics were grouped to wave exposure categories kruskal-wallis tests revealed significant differences amoungst categories for most morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
-The morphological characteristics that were significantly different amoungst sites exibited similar patterns, with increasing mean values from the full sheltered to extremely sheltered categories and a decrease in mean values for the remainder of the categories.
-Two morphological characteristics, namely lamina thickness and stipe diameter, both exibit higher variations for the exposed categories compared to other morphological characteristics.
+The morphological characteristics that were significantly different amoungst sites exibited similar patterns, with increasing mean values from the fully sheltered to extremely sheltered categories and a decrease in mean values for the remainder of the categories.
+Two morphological characteristics, namely lamina thickness and stipe diameter, both exibited higher variations for the exposed categories compared to other morphological characteristics.
 
@@
   scale_x_discrete(limits = pos_exposure_lam) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
   stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.85, label.y.npc = c('top')) + 
@@
   scale_x_discrete(limits = pos_exposure_lam)  + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
   stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
@@
   scale_x_discrete(limits = pos_exposure_lam) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
   stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
@@
   scale_x_discrete(limits = pos_exposure_lam)  + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
   stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
@@
   scale_x_discrete(limits = pos_exposure_lam) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
   stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
@@
 
-```{r laminaria temp plots, echo=FALSE, fig.cap='\\label{fig:lamanaria temp plots}Boxplots of *L.
-pallida* morphological characteristics group by temperature regime.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
+```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.cap='\\label{fig:lamanaria temp plots}Boxplots of *L.
+pallida* morphological characteristics group by temperature regime.', fig.height=8, fig.width=8}
 ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "none", align ='hv')
@@
 
-```{r lamanaria wave plots, echo=FALSE, fig.cap='\\label{fig:lamanaria wave plots}Boxplots of *L.
-pallida* morphological characteristics group by wave exposure category.', fig.height=8, fig.width=8, warning=FALSE, eval=TRUE}
-ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "none", align = "hv")
+```{r lamanaria wave plots, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:lamanaria wave plots}Boxplots of *L.
+pallida* morphological characteristics group by wave exposure category.', fig.height=8, fig.width=8}
+ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom", align = "hv")
 ```
@@
 maxima* morphological characteristics showed differences between cool temperate and warm temperate regime except for frond mass, frond length, stipe mass and epiphyte length.
-Mean values for primary length, primary width and stipe circumference were significantly higher for kelps from cool temperate regions than for kelps from warm temperate regions (see figure \ref{fig:lamanaria wave plots}.
-The remaining morphological characteristics, stipe length, number of tufts and total length had significanlty lower values for kelps from cool temperate compared to kelps from the warm temperate regime.
+Mean values for primary length, primary width and stipe circumference were significantly higher for kelps from the cool temperate regime than for kelps from the warm temperate regime (see figure \ref{fig:lamanaria wave plots}.
+The remaining morphological characteristics, stipe length, number of tufts and total length had significanlty lower means for kelps from cool temperate compared to kelps from the warm temperate regime.
 Significant differences amoungst wave exposure categories were found for all morphological characteristics of deep *E.
 maxima* populations.
-Certian morphological characteristics showed a gradual increase in value with increasing degree of wave exposure.
-The remaining morphological characteristics exhibited decreased values for only the higher wave exposure categories and then a sharp increase in value for the highest wave exposure category.
+Certian morphological characteristics showed a gradual increase in mean value with increasing degree of wave exposure.
+The remaining morphological characteristics exhibited decreased means for only the higher wave exposure categories, followed by a sharp increase in mean value for the highest wave exposure category.
 
@@
   scale_x_discrete(limits = pos_exposure) + 
-  geom_boxplot(outlier.shape = NA)  + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
   stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
@@
   scale_x_discrete(limits = pos_exposure) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) + 
   stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
@@
 
-```{r deep ecklonia wave, echo=FALSE, fig.cap='\\label{fig:deep ecklonia wave}Boxplot of morphological characteristics for deep population *E.
-maxima* grouped by wave exposure.', fig.height=8, fig.width=8, warning=FALSE, eval=TRUE}
+```{r deep ecklonia wave, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:deep ecklonia wave}Boxplot of morphological characteristics for deep population *E.
+maxima* grouped by wave exposure.', fig.height=8, fig.width=8}
 ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3, nrow = 4)
@@
   scale_x_discrete(limits = pos_exposure_shall) + 
-  geom_boxplot(outlier.shape = NA)  + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
   stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
@@
   scale_x_discrete(limits = pos_exposure_shall) + 
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) + 
   stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
@@
 
-```{r shallow ecklonia temp, echo=FALSE, fig.cap='\\label{fig:shallow ecklonia temp}Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
+```{r shallow ecklonia temp, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.cap='\\label{fig:shallow ecklonia temp}Boxplot of morphological characteristics for shallow population *E.
+maxima* grouped by temperature regime.', fig.height=7, fig.width=8}
 ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom", ncol = 3, nrow = 4)
@@
 
-```{r shallow ecklonia wave, echo=FALSE, fig.cap='\\label{fig:shallow ecklonia wave}Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by wave exposure.', fig.height=7, fig.width=8, warning=FALSE, eval=TRUE}
+```{r shallow ecklonia wave, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:shallow ecklonia wave}Boxplot of morphological characteristics for shallow population *E.
+maxima* grouped by wave exposure.', fig.height=7, fig.width=8}
 ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3,nrow = 3)
@@
 RDA1 was the only significant axis in the model (p = 0.03) and explained 72% of the variation.
-The non-significant axis for the model explained 17% of the variation.
 Primary length, primary width and stipe circumference were positivley influenced by *hs_mean*, *tp_mean* and *dir_median* for for kelps from the western region.
@@
 Furthermore, the investigation of differences in morphological characteristics between shallow and deep populations of *E.
-maxima * suggests that in low wave exposure environments the role of temperature as a morphological driver increases, such as in False Bay.
+maxima* suggests that in low wave exposure environments the role of temperature as a morphological driver increases, such as in False Bay.
 
@@
 The direction of the dominant swell changes to the south west in winter, generated by strong low pressures that originate from the southern ocean [@Reason2006], which False Bay is sheltered from [@Shipley1964; @Atkins1970-bx; @Dufois2012-vy].
-In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an increased variability of Hs and Tp in this region.
+In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an increased variability of swell height and peak period in this region.
 It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world [@Norderhaug2012-ks; @Leliaert2000a].
 It is due to the near consistent south-westerly swell and the complex orography around the peninsula that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.
-The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in figure \ref{fig:wave power map}.It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions.
+The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposure maps presented in figure \ref{fig:wave power map}.It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions.
 This results in positive and negative wave exposure anomalies all around the peninsula.
 Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula [@Rouault2010-gi; @Field1980a].
-In general, the western side of the peninsula is more exposed than the eastern side (False Bay) and experiences significanlty higher significant swell height and significant peak period.
+In general, the western side of the peninsula is more exposed than the eastern side (False Bay) and experiences significanlty higher mean significant swell height and mean significant peak period.
 This was also reflected in the wave exposure categories which show a trend of decreasing wave exposure (wave power) around the peninsula and into False Bay.
@@
 Species-specific responses are evident in both wave exposure and temperature.
-In cool-temperate environments *L.
+In cool temperate environments *L.
 pallida* tended to show increases in certain morphological characteristics (stipe mass, stipe length and stipe diameter) while in the warm-temperate environments these were significantly lower.
 This was also true for deep *E.
-maxima* populations, which had longer, thinner stipes for kelp in the warm-temperate compared to kelp in the cool-temperate.
+maxima* populations, which had longer, thinner stipes for kelps from the warm temperate compared to kelps from the cool temperate.
 Reduction in certian morphological charateristics has been attributed to temperature by @Serisawa2002-dc in the kelp *E.
@@
 Warmer temperatures are associated with low nutrient concentrations [@Waldron1992] and the low frequency of upwelling conditions in False Bay (low nutrient supply) coupled with warmer temperatures may be a contributing factor.
-It should be noted however, that from these analyses that in general the cool-temperate region is more exposed to waves than that of the warm-temperate region.
-Therefore, the significantly larger morphological characteristics for kelp from the cool-temperate region may overlap with responses to wave exposure.
+It should be noted however, that from these analyses that in general the cool temperate region is more exposed to waves than that of the warm-temperate region.
+Therefore, the significantly larger morphological characteristics for kelp from the cool temperate region may overlap with responses to wave exposure.
 The response of kelp morphological characteristics to wave exposure was evident and both species show a tactics based on the magnitde of wave exposure.
 Strength increasing traits were exibited for lower exposure levels while a 'go-with-the-flow' tactic for moderate levels of wave exposure.
-Denny and Galord ???
-showed that increased flexiblity or a 'go-with-the-flow' tactic has a threshold value, where increased stipe length actaully increase drag under high wave exposure scenarios.
-Therefore, under high wave exposure scenarios size reducing traits, such as reduced surface area of the overall plant should be favoured instead.
-This was evident in the results which showed that stipe length decreases under high wave exposure categories.
 This suggests that how morphological characteristcs manifest themselves is dependent on the magnitude of wave exposure.
 When grouped by wave exposure category, *L.
-pallida* characteristics showed a significant increase in length for the 'Extremely sheltered' category (6-10 kW/m; figure \ref{fig:wave power map}) compared to the fully sheltered category (<5 kW/m; see figure \ref{fig:wave power map}), which maybe a 'go-with-the-flow' tactic.
+pallida* characteristics showed a significant increase in length for the 'Extremely sheltered' category (6-10 kW/m; figure \ref{fig:wave power map}) compared to the fully sheltered category (<5 kW/m; see figure \ref{fig:wave power map}), which may be a 'go-with-the-flow' tactic.
 Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny2002; @Hurd2000].
-However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under higher exposure levels [@Denny2002; @Denny1997].
-This is reflected in Figure ??
-which shows an overall reduction of *L pallida* morphological characteristics suggesting a size reducing tactic to cope with higher levels of wave exposure.
+However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under high wave exposure [@Denny2002; @Denny1997].
+This is reflected in figure \ref{fig:lamanaria wave plots} which shows an overall reduction of *L pallida* morphological characteristics suggesting a size reducing tactic to cope with higher levels of wave exposure.
 The results suggest that deep *E.
@@
 pallida*, *E.
-maxima* exhibits a size reducing tactic with in an increase in exposure level by significantly decreasing stipe mass, stipe length, stipe circumference, total length and frond length.
+maxima* exhibits a size reducing tactic with increasing wave exposure by  decreasing stipe mass, stipe length, stipe circumference, total length and frond length.
 A magnitude related reponse has been suggested by @wernberg2010 who investigated the temporal and spatial variations of various environmental drivers of the kelp *E.
@@
 maxima* populations within False Bay, hence the identification of minimum temperature (*min_temp*) as a driver.
+Mean temperature identified as a driver for *L.
+pallida* may be related to diurnal temperature fluctuations in the water column.
+Solar heating of the water surface in combination with wind-driven transport causes fluctuations within the water coloumn, which can occur daily or seasonally [@Kaplan2003].
 Vertical mixing of the water coloumn for inshore regions are driven by several abiotic processes: (1) Turbulence of breaking waves inside or outside of the surf zone (2) convective mixing through a combination of cooling and evaporation (3) wind driven currents and (4) tidal mixing [@Smit2013a].
@@
 pallida* which occurs deeper in the water column.
-Mean temperature identified as a driver for *L.
-pallida* may be related to diurnal temperature fluctuations in the water column.
-Solar heating of the water surface in combination with wind-driven transport causes fluctuations within the water coloumn, which can occur daily or seasonally [@Kaplan2003].
 
 Temperature variables were not a significant driver of kelp morphology in this study.
-Although temperature plays an important role in distribution (Bolton et al.
-2012; Miller et al.
-2011; Rinde et al.
-2014) of kelp and physiological functioning of adults and gametophytes (Mohring et al.
-2014; Steneck et al.
-2002; Gerard 1997; Bearham et al.
-2013; Smale & Moore 2017), there is little evidence that temperature is an important driver of morphological variation.
+Although temperature plays an important role in distribution [@Bolton2010; @Miller2011-pm; @Rinde2014] of kelp and physiological functioning of adults and gametophytes [@Mohring2014-bk; @Steneck2002; @gerard1997; @bearham2013; @Smale2017], there is little evidence in this study that temperature is an important driver of morphological variation.
 However, we suggest that temperature plays a larger role in determining kelp morphology at sites where wave energy is low enough not to elecit a response in morphological characteristics.
@@
 The lack of variation in shallow *E.
-maxima* populations was also found in previous research by [@Rothman2017-pn] who showed that light was a significant driver.
+maxima* populations was also found in previous research by [@Rothman2017-pn] who identified light as a significant driver.
 This influence of light may suggest photosynthetic ability and nutrient uptake as important drivers in shallow environments.
@@
 maxima*.
-The identification of maximum temperature as a driver of kelp morphological characteristics may be due heating of shallow coastal waters due to solar heating.
 

### 2019-10-16 chp3_raft_model.Rmd
housecleaning

---
title: "Hydrodynamic and particle dispersion modelling of floating solitary kelp: An experimental approach"
author: "Ross Coppin"
date: "16/10/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2019-10-16 morphology_paper.Rmd
changed .bib file. one reference.bib will be used from now on. updated latest output of word file

@@
 title: Using numerical modelling to investigate temperature and wave exposure as drivers
-  of macroalgal characteristics
+  of macroalgal morphological characteristics
 author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
@@
 tables: yes
-bibliography: refs_morphology_paper.bib
+bibliography: references.bib
 ---
@@
 Differences in mena significant peak period between regions are presented in plot D.'}
-ggarrange(wave_plot_1,wave_boxplot_1, wave_plot_2, wave_boxplot_2, common.legend = TRUE, labels = c("A", "B", "C", "D"), align = "hv")
+ggarrange(wave_plot_1,wave_boxplot_1, wave_plot_2, wave_boxplot_2, common.legend = FALSE, labels = c("A", "B", "C", "D"), align = "hv")
 ```

### 2019-10-19 morphology_paper.Rmd
some minor touch ups. added more figure captions. sorted out minor issues with references. added abstract and output updated. final review and touch ups are next...

@@
 author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
+abstract: "*Abstract*: Macroalgae morphological characteristics are determined by a combination of environmental factors with wave exposure and temperature being identified as the main drivers of morphological adaptation.
+Macroalgae have shown to exhibit different responses to different magnitudes of exposure, such as reduction in overall size and strength increasing traits.
+In terms of temperature, warmer environments have been shown to reduce the overall size of resident and transplanted species.
+However, none of the past studies have identified specific wave and temperature metrics responsible for the morphological adaptation macroalgae exhibit.
+Past research has often used simple or two dimensional models of wave exposure which does not take into account important aspects of the nearshore environment such as wave breaking, refraction and diffraction.
+Furthermore, past studies have often used satellite as sources for temperature data which has been shown to have large bias when measuring the nearshore environment.
+This study used *in situ* temperature data and wave power calculated from a 3D-numerical model to identify specific temperature and wave metrics responsible for morphological adaptation.
+The results identify wave exposure as the main driver of morphological adaptation while identifying specific wave metrics.
+Furthermore, the results show differences in wave metrics between species; and between deep and shallow populations."
 fontsize: 10pt
@@
   geom_boxplot(show.legend = FALSE, width = 0.5) + 
-  geom_jitter(width = 0.2) + 
   stat_compare_means(label.x = 2.5, label.y = 12.5, label = "p.format") +
@@
 ```{r Temperature variables plot, echo=FALSE, warning=FALSE, eval=TRUE,fig.height=5,fig.width=8, fig.cap='\\label{fig:Temperature variables plot}Temperature variables at the collection sites around the Cape Peninsula are represented in plot A.
-Temperature variables include minimum represented by inverted triangles, mean represented by balck dots, maximum by black triangles, and whiskers standard deviation.
+Temperature variables include minimum represented by inverted triangles, mean represented by black dots, maximum by black triangles, and whiskers standard deviation.
 Summary of temperature data for sites grouped by region are represented in plot B.
@@
   geom_boxplot(show.legend = FALSE) + 
-  geom_jitter(width = 0.2) +
   stat_compare_means(label.y =0.6 ,label.x = 2.5, label = "p.format") + 
@@
   geom_boxplot(show.legend = FALSE) + 
-  geom_jitter(width = 0.2) +
   stat_compare_means(label.y = 8.2, label.x = 2.5, label = "p.format") + 
@@
 wp_6 <- rasterGrob(as.raster(readPNG("figures/CapePen_Legend.PNG")), interpolate = FALSE)
-
 grid.arrange(wp_1, wp_2, wp_3, wp_4, wp_5, wp_6, ncol = 3, nrow = 2)
+
+#wave_power_map <- readPNG("figures/wave_power_all_labels.png")
+#grid.raster(wave_power_map)
 
@@
 
-```{r FB spectrum, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:FB spectrum}The propagation of a typical offshore wave spectrum is given as produced from a single time-step in SWAN'}
+```{r FB spectrum, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:FB spectrum}The propagation of a typical offshore wave spectrum is given as produced from a single time-step in SWAN.'}
 
@@
 
-```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.cap='\\label{fig:lamanaria temp plots}Boxplots of *L.
-pallida* morphological characteristics group by temperature regime.', fig.height=8, fig.width=8}
+```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.cap='\\label{fig:lamanaria temp plots}Boxplots representing *L.
+pallida* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+The lower and upper "hinges" correspond to the first and third quartiles.
+The “whiskers” represent the range, solid black lines represent the median and black dots are outliers.
+Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=8, fig.width=8}
 ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "none", align ='hv')
@@
 
-```{r lamanaria wave plots, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:lamanaria wave plots}Boxplots of *L.
-pallida* morphological characteristics group by wave exposure category.', fig.height=8, fig.width=8}
+```{r lamanaria wave plots, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.height=8, fig.width=8, fig.cap='\\label{fig:lamanaria wave plots}Boxplots representing *L.
+pallida* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+The lower and upper hinges correspond to the first and third quartiles.
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.'}
 ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom", align = "hv")
@@
 
-```{r deep ecklonia temp plot, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E.
-maxima* grouped by temperature.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
+```{r deep ecklonia temp plot, echo=FALSE, fig.cap='Boxplots representing deep population *E.
+maxima* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+The lower and upper "hinges" correspond to the first and third quartiles.
+The “whiskers” represent the range, solid black lines represent the median and black dots are outliers.
+Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
 ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = c("none"), ncol = 3, nrow = 4)
@@
 
-```{r deep ecklonia wave, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:deep ecklonia wave}Boxplot of morphological characteristics for deep population *E.
-maxima* grouped by wave exposure.', fig.height=8, fig.width=8}
+```{r deep ecklonia wave, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:deep ecklonia wave}Boxplots representing deep *E.
+maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+The lower and upper "hinges" correspond to the first and third quartiles.
+The “whiskers” represent the range, solid black lines represent the median and black dots are outliers.', fig.height=8, fig.width=8}
 ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3, nrow = 4)
@@
 
-```{r shallow ecklonia temp, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.cap='\\label{fig:shallow ecklonia temp}Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by temperature regime.', fig.height=7, fig.width=8}
+```{r shallow ecklonia temp, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.cap='\\label{fig:shallow ecklonia temp} Boxplots representing shallow *E.
+maxima* morphological characteristics  grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+The lower and upper "hinges" correspond to the first and third quartiles.
+The “whiskers” represent the range, solid black lines represent the median and black dots are outliers.
+Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=7, fig.width=8}
 ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom", ncol = 3, nrow = 4)
@@
 
-```{r shallow ecklonia wave, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:shallow ecklonia wave}Boxplot of morphological characteristics for shallow population *E.
-maxima* grouped by wave exposure.', fig.height=7, fig.width=8}
+```{r shallow ecklonia wave, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:shallow ecklonia wave}Boxplots representing shallow *E.
+maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+The lower and upper "hinges" correspond to the first and third quartiles.
+The “whiskers” represent the range, solid black lines represent the median and black dots are outliers.', fig.height=7, fig.width=8}
 ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3,nrow = 3)
@@
 maxima* (B) and shallow *E.
-maxima* (C)."}
+maxima* (C).
+The blue cluster represnts sites from the western region and the red represents sites from the eastern region."}
 ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, legend = "none", labels = "AUTO", ncol = 2, nrow = 2, align = "hv")
@@
 
-```{r appendix figure 1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=7}
+```{r appendix figure 1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=8}
 positionsLam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point ", "Baboon Rock", 
@@
 
-```{r appendix figure 2, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=7}
+```{r appendix figure 2, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=8}
 # Order site names by location west to east

### 2019-10-22 morphology_paper.Rmd
final spelling and grammer check done.

@@
 author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
-abstract: "*Abstract*: Macroalgae morphological characteristics are determined by a combination of environmental factors with wave exposure and temperature being identified as the main drivers of morphological adaptation.
+abstract: "**Abstract**: Macroalgae morphological characteristics are determined by a combination of environmental factors with wave exposure and temperature being identified as the main drivers of morphological adaptation.
 Macroalgae have shown to exhibit different responses to different magnitudes of exposure, such as reduction in overall size and strength increasing traits.
@@
 Kelps generally have a dependence on cool-temperate and arctic seawater temperatures [@Bolton2010; @Mohring2014-bk; @Cavanaugh2011-ib; @Dayton1985], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
-Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and subcanopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002]; each depending on a wide suite of ecological services provided by kelp forests [@Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @Willis2003-xo; @anderson1997; @Dayton1985].
+Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and sub-canopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002]; each depending on a wide suite of ecological services provided by kelp forests [@Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @Willis2003-xo; @anderson1997; @Dayton1985].
 Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in the distribution [@Gorman2013-nj], abundance [@Cavanaugh2011-ib; @Dayton1998-xw], diversity [@Wing2007-tv; @Wernberg2008], composition [@Leliaert2000a; @Dayton1985; @Norderhaug2012-ks; @Harley2012], growth [@Cousens1982-ck] and productivity [@Pedersen2012-vr; @Dayton1998-xw; @Krumhansl2012] of kelps.
 
-Given kelps are cool-temperate organisms which are vulernable to dislodgment under high wave exposure scenarios, changes in ocean temperature and an increase in storm frequency and magnitude pose the biggest direct threats to their survival.
-Warmer temperatures reduces the resilience of kelp individuals to perbutations [@wernberg2010], causes fragmentation through weakening tissues [@Simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
+Given kelps are cool-temperate organisms which are vulnerable to dislodgement under high wave exposure scenarios, changes in ocean temperature and an increase in storm frequency and magnitude pose the biggest direct threats to their survival.
+Warmer temperatures reduces the resilience of kelp individuals to storm disturbance [@wernberg2010], causes fragmentation through weakening tissues [@Simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
 Despite these threats and disturbances kelps persist across a broad range of environments which is largely due to their morphological plasticity [@FowlerWalker2006; @wernberg2010].
@@
 Cartographical models of wave exposure are 'fetch-based-models' which measure the length of open water associated with a particular site in a straight line, and are regarded as simple measures of wave exposure.
-Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitave, reproducible approach for characterising the hydrodynamic environment.
+Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitative, reproducible approach for characterising the hydrodynamic environment.
 Although @mabin2013 investigated kelp morphological characteristics using a 2D WAM model, currently no studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
-Furthermore no current studies have identified specific wave metrics that drive kelp morphological charactersitcs, which the current work will aim to achieve.
-Other considerations such as possible differences in morphology between sub-canopy and canopy species, and shallow and deep water kelp populations have not been considered using advanced numerical and statisical tools.
+Furthermore no current studies have identified specific wave metrics that drive kelp morphological characteristics, which the current work will aim to achieve.
+Other considerations such as possible differences in morphology between sub-canopy and canopy species, and shallow and deep water kelp populations have not been considered using advanced numerical and statistical tools.
 
 The kelps *Laminaria pallida* and *Ecklonia maxima* (Osbeck) Papenfuss are important habitat forming seaweeds that exist around the coast of South Africa and offer a unique opportunity to investigate the drivers of macroalgal morphological characteristics between canopy and sub-canopy species, as well as between deep and shallow water environments.
-Although both these species exist in the subtidal, *L.
+Although both these species exist in the sub-tidal, *L.
 pallida* dominates deeper waters while *E.
-maxima* forms dense surface canopies in the subtidal to shallower waters.
+maxima* forms dense surface canopies in the sub-tidal to shallower waters.
 Therefore, *E.
@@
 This will be achieved by initially understanding the variation in temperature and waves using an advanced numerical model.
-The numerical model will then be used as a basis for investigating the consequences for morphological chracteristics of *E.
+The numerical model will then be used as a basis for investigating the consequences for morphological characteristics of *E.
 maxima* and *L.
@@
 The region is dominated by kelp communities that persist in contrasting abiotic environments.
-The west coast region has been termed as cool temperate, which is defined as a region where mean monthly temperatures are always above 10°C and always below 15°C [@Smit2013a].
+The west coast region has been termed as cool-temperate, which is defined as a region where mean monthly temperatures are always above 10°C and always below 15°C [@Smit2013a].
 East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone [@Smit2013a].
-The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm temperate [@Smit2013a].
+The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm-temperate [@Smit2013a].
 
@@
 Linear interpolated SST were calculated for sites where *in situ* recorders were absent.
-These data were used to group sites into 'Cool temperate' and 'Warm temperate' categorisations to investigate the possible affect of thermal regime on kelp morphological characteristics.
+These data were used to group sites into 'Cool temperate' and 'Warm temperate' categorisations to investigate the possible effect of thermal regime on kelp morphological characteristics.
 
@@
 
-## Statisical analyses
+## Statistical analyses
 
@@
 
-Characterisation of morphological characteristics in relation to temperature and wave exposure non-parametric statistical methods were used.
-Morphological characteristics were categorised according to temperature regime and wave exposure category for each each species and population of kelp in this study.
+Morphological characteristics in relation to temperature and wave exposure were investigated using non-parametric statistical methods.
+Morphological characteristics were categorised according to temperature regime and wave exposure category for each species and population of kelp in this study.
 Significant differences between temperature regimes and across wave exposure categories were tested with wilcox tests and kruskal tests respectively, and notched boxplots allowed comparison between individual categories of wave exposure.
@@
 Temperature variables include minimum represented by inverted triangles, mean represented by black dots, maximum by black triangles, and whiskers standard deviation.
-Summary of temperature data for sites grouped by region are represented in plot B.
-See appendix for significance level guide.'}
+Summary of temperature data for sites grouped by region are represented in plot B.'}
 ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 2, nrow = 1, labels = c("AUTO"), align = c("hv"), legend = FALSE)
@@
 Mean significant peak period across sites is presented in plot C and is represented by solid dots and standard deviation by whiskers.
-Differences in mena significant peak period between regions are presented in plot D.'}
+Differences in mean significant peak period between regions are presented in plot D.'}
 ggarrange(wave_plot_1,wave_boxplot_1, wave_plot_2, wave_boxplot_2, common.legend = FALSE, labels = c("A", "B", "C", "D"), align = "hv")
@@
 pallida* morphological characteristics with lamina length, number of digits and thallus mass the exceptions (see figure \ref{fig:lamanaria temp plots}).
-Certian morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significanlty higher means for kelps from the cool temperate regime when compared to kelps for the warm temperate regime.
-Conversely, lamina weight and lamina thickness had significantly lower means for kelps from the cool temperate regime compared to kelps from the warm temperate regime.
+Certain morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significantly higher means for kelp from the cool temperate regime when compared to kelp from the warm temperate regime.
+Conversely, lamina weight and lamina thickness had significantly lower means for kelp from the cool temperate regime compared to kelp from the warm temperate regime.
 When *L.
-pallida* morphological characteristics were grouped to wave exposure categories kruskal-wallis tests revealed significant differences amoungst categories for most morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
-The morphological characteristics that were significantly different amoungst sites exibited similar patterns, with increasing mean values from the fully sheltered to extremely sheltered categories and a decrease in mean values for the remainder of the categories.
-Two morphological characteristics, namely lamina thickness and stipe diameter, both exibited higher variations for the exposed categories compared to other morphological characteristics.
+pallida* morphological characteristics were grouped to wave exposure categories kruskal-wallis tests revealed significant differences amongst categories for most morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
+The morphological characteristics that were significantly different amongst sites exhibited similar patterns, with increasing mean values from the fully sheltered to extremely sheltered categories and a decrease in mean values for the remainder of the categories.
+Two morphological characteristics, namely lamina thickness and stipe diameter, both exhibited higher variations for the exposed categories compared to other morphological characteristics.
 
@@
 maxima* morphological characteristics showed differences between cool temperate and warm temperate regime except for frond mass, frond length, stipe mass and epiphyte length.
-Mean values for primary length, primary width and stipe circumference were significantly higher for kelps from the cool temperate regime than for kelps from the warm temperate regime (see figure \ref{fig:lamanaria wave plots}.
-The remaining morphological characteristics, stipe length, number of tufts and total length had significanlty lower means for kelps from cool temperate compared to kelps from the warm temperate regime.
-Significant differences amoungst wave exposure categories were found for all morphological characteristics of deep *E.
+Mean values for primary length, primary width and stipe circumference were significantly higher for kelp from the cool temperate regime than for kelp from the warm temperate regime (see figure \ref{fig:lamanaria wave plots}.
+The remaining morphological characteristics, stipe length, number of tufts and total length had significantly lower means for kelp from cool temperate compared to kelp from the warm temperate regime.
+Significant differences amongst wave exposure categories were found for all morphological characteristics of deep *E.
 maxima* populations.
-Certian morphological characteristics showed a gradual increase in mean value with increasing degree of wave exposure.
+Certain morphological characteristics showed a gradual increase in mean value with increasing degree of wave exposure.
 The remaining morphological characteristics exhibited decreased means for only the higher wave exposure categories, followed by a sharp increase in mean value for the highest wave exposure category.
@@
 RDA1 was the only significant axis in the model and explained 48% of the variation while RDA2 was not significant and explained 63% of the variation.
-Total length, stipe length, stipe diameter and stipe mass were postivley influenced (i.e., increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* for kelps mostly from the western region and negativley influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp* for kelps from the eastern region.
+Total length, stipe length, stipe diameter and stipe mass were positively influenced (i.e., increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* for kelp mostly from the western region and negatively influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp* for kelp from the eastern region.
 Although sites from from the western region did not cluster as closely as eastern region study sites, grouping according to region was still evident.
@@
 RDA1 was the only significant axis in the model (p = 0.03) and explained 72% of the variation.
-Primary length, primary width and stipe circumference were positivley influenced by *hs_mean*, *tp_mean* and *dir_median* for for kelps from the western region.
-The remaining morphological characteristics were postivley inlfuenced by *min_temp* and negatively influenced by *tp_sd* for kelps from the eastern region.
+Primary length, primary width and stipe circumference were positively influenced by *hs_mean*, *tp_mean* and *dir_median* for for kelp from the western region.
+The remaining morphological characteristics were positivist influenced by *min_temp* and negatively influenced by *tp_sd* for kelp from the eastern region.
 
@@
 
-Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retian the most parsimonious descriptors of shallow *E.
+Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of shallow *E.
 maxima* morphological characteristics with an adjusted *R*2 of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, *F* = 2.41, *p* = 0.013).
 The model consisted of one significant axis, RDA1 (p = 0.02) and explained 62% of the variation.
-Stipe length, stipe mass and stipe circumference were positivley influenced by *max_temp* and *mean_temp* along RDA1, and explained 50% and 55% of the variation respectivley.
-The remaining morphological characteristics were positivley influenced by by *tp_max* and negativley inlfuenced by *dirw_median* along RDA2 and explained 86% and 94% of the variation respectivley.
+Stipe length, stipe mass and stipe circumference were positively influenced by *max_temp* and *mean_temp* along RDA1, and explained 50% and 55% of the variation respectively.
+The remaining morphological characteristics were positively influenced by *tp_max* and negatively influenced by *dirw_median* along RDA2 and explained 86% and 94% of the variation respectively.
 There was overlapping of clusters according to region with no clear patterns in sites evident.
@@
 Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula [@Rouault2010-gi; @Field1980a].
-In general, the western side of the peninsula is more exposed than the eastern side (False Bay) and experiences significanlty higher mean significant swell height and mean significant peak period.
+In general, the western side of the peninsula is more exposed than the eastern side (False Bay) and experiences significantly higher mean significant swell height and mean significant peak period.
 This was also reflected in the wave exposure categories which show a trend of decreasing wave exposure (wave power) around the peninsula and into False Bay.
@@
 This was also true for deep *E.
-maxima* populations, which had longer, thinner stipes for kelps from the warm temperate compared to kelps from the cool temperate.
-Reduction in certian morphological charateristics has been attributed to temperature by @Serisawa2002-dc in the kelp *E.
+maxima* populations, which had longer, thinner stipes for kelp from the warm temperate compared to kelp from the cool temperate.
+Reduction in certain morphological characteristics has been attributed to temperature by @Serisawa2002-dc in the kelp *E.
 cava*, which was smaller and shorter in warmer sites compared to cooler sites.
@@
 Therefore, the significantly larger morphological characteristics for kelp from the cool temperate region may overlap with responses to wave exposure.
-The response of kelp morphological characteristics to wave exposure was evident and both species show a tactics based on the magnitde of wave exposure.
-Strength increasing traits were exibited for lower exposure levels while a 'go-with-the-flow' tactic for moderate levels of wave exposure.
-This suggests that how morphological characteristcs manifest themselves is dependent on the magnitude of wave exposure.
+The response of kelp morphological characteristics to wave exposure was evident and both species exhibit tactics based on the magnitude of wave exposure.
+Strength increasing traits were exhibited for lower exposure levels while a 'go-with-the-flow' tactic for moderate levels of wave exposure.
+This suggests that how morphological characteristics manifest themselves is dependent on the magnitude of wave exposure.
 When grouped by wave exposure category, *L.
@@
 maxima* exhibits a size reducing tactic with increasing wave exposure by  decreasing stipe mass, stipe length, stipe circumference, total length and frond length.
-A magnitude related reponse has been suggested by @wernberg2010 who investigated the temporal and spatial variations of various environmental drivers of the kelp *E.
+A magnitude related response has been suggested by @wernberg2010 who investigated the temporal and spatial variations of various environmental drivers of the kelp *E.
 radiata*.
@@
 
-A size reduction tactic has been shown before [@FowlerWalker2006; @Denny2002; @Hurd2000; @FowlerWalker2005] and is regarded as strategy to reduce overall drag on the plant.
+A size reduction tactic has been shown before [@FowlerWalker2006; @Denny2002; @Hurd2000; @FowlerWalker2005] and is regarded as a strategy to reduce overall drag on the plant.
 However, *E.
@@
 Low temperatures are often associated with upwelling events which bring cool, nutrient rich water into the nearshore.
-False Bay has comparativley little upwelling events compared to the western side of the peninsula, and so nutrients may be a limiting factor for *E.
+False Bay has comparatively little upwelling events compared to the western side of the peninsula, and so nutrients may be a limiting factor for *E.
 maxima* populations within False Bay, hence the identification of minimum temperature (*min_temp*) as a driver.
@@
 pallida* may be related to diurnal temperature fluctuations in the water column.
-Solar heating of the water surface in combination with wind-driven transport causes fluctuations within the water coloumn, which can occur daily or seasonally [@Kaplan2003].
-Vertical mixing of the water coloumn for inshore regions are driven by several abiotic processes: (1) Turbulence of breaking waves inside or outside of the surf zone (2) convective mixing through a combination of cooling and evaporation (3) wind driven currents and (4) tidal mixing [@Smit2013a].
-These abiotic processes cause effective vertical mixing of surface and deeper water stratifications which leads to a homogeneous thermal environment, which in this study may be interpreted as the mean temperature.
-Therefore, the homogenious thermal regime of inshore regions may be a reason why mean temperature is driver of *L.
+Solar heating of the water surface in combination with wind-driven transport causes fluctuations within the water column, which can occur daily or seasonally [@Kaplan2003].
+Vertical mixing of the water column for inshore regions are driven by several abiotic processes: (1) Turbulence of breaking waves inside or outside of the surf zone (2) convective mixing through a combination of cooling and evaporation (3) wind driven currents and (4) tidal mixing [@Smit2013a].
+These abiotic processes cause effective vertical mixing of surface and deeper water stratification which leads to a homogeneous thermal environment, which in this study may be interpreted as the mean temperature.
+Therefore, the homogeneous thermal regime of inshore regions may be a reason why mean temperature is driver of *L.
 pallida* morphological characteristics.
@@
 Although temperature plays an important role in distribution [@Bolton2010; @Miller2011-pm; @Rinde2014] of kelp and physiological functioning of adults and gametophytes [@Mohring2014-bk; @Steneck2002; @gerard1997; @bearham2013; @Smale2017], there is little evidence in this study that temperature is an important driver of morphological variation.
-However, we suggest that temperature plays a larger role in determining kelp morphology at sites where wave energy is low enough not to elecit a response in morphological characteristics.
+However, we suggest that temperature plays a larger role in determining kelp morphology at sites where wave energy is low enough not to elicit a response in morphological characteristics.
 Temperature metrics influence stipe length and frond length in shallow *E.
@@
 Increased temperatures corresponded with a reduction in stipe circumference, frond length, stipe length, total length and stipe mass resulting in slim or ‘skinny’ kelp individuals, which supports findings from previous research by [@mabin2013; @Serisawa2002-dc].
-These findings were confirmed by the redundancy analysis which showed the contribution of wave variables as a driver of kelp morphological characterstics was lower compared to temperature variables.
+These findings were confirmed by the redundancy analysis which showed the contribution of wave variables as a driver of kelp morphological characteristics was lower compared to temperature variables.
 The sites do cluster by region but almost overlap entirely, further indicating wave variables are not the main driver.
@@
 maxima*.
-The identification of median wind direction (*dirw_median*) suggests that surface gravity waves as a driver compared to swell waves (*dir_median*) identifiend for *L.
+The identification of median wind direction (*dirw_median*) suggests that surface gravity waves as a driver compared to swell waves (*dir_median*) identified for *L.
 pallida* and *E.
@@
 maxima* may reflect differences in hydrodynamic and temperature regimes.
-This study also confirms the findings of previous research which shows morphological adpatation is dependent on the magnitude of wave exposure, but provides quantitative, reproducible measure of wave exposure while identifying the specific wave metrics and wave energy thresholds invloved.
+This study also confirms the findings of previous research which shows morphological adaptation is dependent on the magnitude of wave exposure, but provides quantitative, reproducible measure of wave exposure while identifying the specific wave metrics and wave energy thresholds involved.
 Furthermore, the current study highlights the need to understand how kelp populations, which are not exposed to high wave energy and morphological characteristics are largely driven by temperature, are vulnerable to changes in storm frequency and magnitude.
-Kelp populations located in low wave energy environments, such as False Bay, may be more vulnerable to increased storms frequency and magnitude in future climate scenarios.
+Kelp populations located in low wave energy environments, such as False Bay, may be more vulnerable to increased storm frequency and magnitude in future climate scenarios.
 

### 2019-10-22 morphology_paper.Rmd
minor adjustments and house cleaning

@@
 When *L.
-pallida* morphological characteristics were grouped to wave exposure categories kruskal-wallis tests revealed significant differences amongst categories for most morphological characteristics with lamina length, number of digits and thallus mass the exceptions.
+pallida* morphological characteristics were grouped to wave exposure categories kruskal-wallis tests revealed significant differences amongst categories for most morphological characteristics with lamina length, number of digits and thallus mass the exceptions (see figure ref\{fig:lamanaria wave plots}).
 The morphological characteristics that were significantly different amongst sites exhibited similar patterns, with increasing mean values from the fully sheltered to extremely sheltered categories and a decrease in mean values for the remainder of the categories.
@@
 maxima* morphological characteristics showed differences between cool temperate and warm temperate regime except for frond mass, frond length, stipe mass and epiphyte length.
-Mean values for primary length, primary width and stipe circumference were significantly higher for kelp from the cool temperate regime than for kelp from the warm temperate regime (see figure \ref{fig:lamanaria wave plots}.
-The remaining morphological characteristics, stipe length, number of tufts and total length had significantly lower means for kelp from cool temperate compared to kelp from the warm temperate regime.
+Mean values for primary length, primary width and stipe circumference were significantly higher for kelp from the cool temperate regime than for kelp from the warm temperate regime (see figure \ref{fig:{deep eck temp plot}}.
+The remaining morphological characteristics, stipe length, number of tufts and total length had significantly lower means for kelp from cool temperate compared to kelp from the warm temperate regime (p < 0.05; see figure \ref{fig:deep eck temp plot}).
 Significant differences amongst wave exposure categories were found for all morphological characteristics of deep *E.
@@
 
-```{r deep ecklonia temp plot, echo=FALSE, fig.cap='Boxplots representing deep population *E.
+```{r deep ecklonia temp plot, echo=FALSE, fig.cap='\\label{fig:deep eck temp plot}Boxplots representing deep population *E.
 maxima* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
@@
 ```
+
+When grouped by temperature regime, shallow *E.
+maxima* exhibited significant differences between temperature regimes for frond mass, primary length, primary width, frond length, stipe mass, stipe length, stipe circumference and total length (p < 0.05; see figure \ref{fig:shallow ecklonia temp}).
+All of the significantly different morphological characteristics, except primary length, had lower values for the western region compared to the eastern region.
+When shallow *E.
+maxima* were grouped by wave exposure category significant difference were found for all morphological characteristics (p < 0.05; see figure \ref{fig:shallow ecklonia wave}).
+In general, values of the various morphological characteristics increase for the lower wave exposure categories, followed by a sharp decrease and a gradual increase in value.
 

### 2019-11-08 morphology_paper.Rmd
added section to lit review, updated outputs, and added references to bib. file

@@
 author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
-abstract: "**Abstract**: Macroalgae morphological characteristics are determined by a combination of environmental factors with wave exposure and temperature being identified as the main drivers of morphological adaptation.
-Macroalgae have shown to exhibit different responses to different magnitudes of exposure, such as reduction in overall size and strength increasing traits.
-In terms of temperature, warmer environments have been shown to reduce the overall size of resident and transplanted species.
-However, none of the past studies have identified specific wave and temperature metrics responsible for the morphological adaptation macroalgae exhibit.
-Past research has often used simple or two dimensional models of wave exposure which does not take into account important aspects of the nearshore environment such as wave breaking, refraction and diffraction.
-Furthermore, past studies have often used satellite as sources for temperature data which has been shown to have large bias when measuring the nearshore environment.
-This study used *in situ* temperature data and wave power calculated from a 3D-numerical model to identify specific temperature and wave metrics responsible for morphological adaptation.
-The results identify wave exposure as the main driver of morphological adaptation while identifying specific wave metrics.
-Furthermore, the results show differences in wave metrics between species; and between deep and shallow populations."
-fontsize: 10pt
+bibliography: references.bib
 output:
+  html_document:
+    df_print: paged
+  pdf_document:
+    fig_caption: yes
   word_document:
     reference_docx: frontiers_template.docx
-  pdf_document: 
-    fig_caption: yes
 geometry: margin=1in
+fontsize: 10pt
 papersize: A4
 tables: yes
-bibliography: references.bib
+abstract: 'Macroalgae morphological characteristics are determined by
+  a combination of environmental factors with wave exposure and temperature being
+  identified as the main drivers of morphological adaptation. Macroalgae have shown
+  to exhibit different responses to different magnitudes of exposure, such as reduction
+  in overall size and strength increasing traits. In terms of temperature, warmer
+  environments have been shown to reduce the overall size of resident and transplanted
+  species. However, none of the past studies have identified specific wave and temperature
+  metrics responsible for the morphological adaptation macroalgae exhibit. Past research
+  has often used simple or two dimensional models of wave exposure which does not
+  take into account important aspects of the nearshore environment such as wave breaking,
+  refraction and diffraction. Furthermore, past studies have often used satellite
+  as sources for temperature data which has been shown to have large bias when measuring
+  the nearshore environment. This study used *in situ* temperature data and wave power
+  calculated from a 3D-numerical model to identify specific temperature and wave metrics
+  responsible for morphological adaptation.  The results identify wave exposure as
+  the main driver of morphological adaptation while identifying specific wave metrics.
+  Furthermore, the results show differences in wave metrics between species; and between
+  deep and shallow populations.'
 ---

### 2019-11-28 morphology_paper.Rmd
grammer and spelling recommendations added. Updated figures and extra paragraph to follow

@@
 tables: yes
-abstract: 'Macroalgae morphological characteristics are determined by
-  a combination of environmental factors with wave exposure and temperature being
-  identified as the main drivers of morphological adaptation. Macroalgae have shown
-  to exhibit different responses to different magnitudes of exposure, such as reduction
+abstract: 'Macroalgal morphological variation is determined to a large extent by
+  a combination of environmental factors, with wave exposure and temperature
+  perhaps the main influences as they are key environmental properties to which a species becomes locally adapted.
+Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction
   in overall size and strength increasing traits. In terms of temperature, warmer
@@
   metrics responsible for the morphological adaptation macroalgae exhibit. Past research
-  has often used simple or two dimensional models of wave exposure which does not
+  has often used simple or two-dimensional models of wave exposure, which do not
   take into account important aspects of the nearshore environment such as wave breaking,
-  refraction and diffraction. Furthermore, past studies have often used satellite
-  as sources for temperature data which has been shown to have large bias when measuring
-  the nearshore environment. This study used *in situ* temperature data and wave power
-  calculated from a 3D-numerical model to identify specific temperature and wave metrics
-  responsible for morphological adaptation.  The results identify wave exposure as
-  the main driver of morphological adaptation while identifying specific wave metrics.
+  refraction and diffraction. Furthermore, past studies have often used satellite-derived
+  datasets as sources for temperature data; however, such data have been shown to have large bias when 
+  applied to the nearshore environment. This study used *in situ* temperature data and wave power
+  metrics calculated from a 3D-numerical model to identify specific temperature and wave metrics
+  responsible for morphological adaptation of the kelp, *Ecklonia maxima* and *Laminaria pallida*.
+The results identify wave      exposure as the main driver of morphological adaptation while identifying specific wave metrics.
   Furthermore, the results show differences in wave metrics between species; and between
@@
 Kelps generally have a dependence on cool-temperate and arctic seawater temperatures [@Bolton2010; @Mohring2014-bk; @Cavanaugh2011-ib; @Dayton1985], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
-Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and sub-canopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002]; each depending on a wide suite of ecological services provided by kelp forests [@Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @Willis2003-xo; @anderson1997; @Dayton1985].
-Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in the distribution [@Gorman2013-nj], abundance [@Cavanaugh2011-ib; @Dayton1998-xw], diversity [@Wing2007-tv; @Wernberg2008], composition [@Leliaert2000a; @Dayton1985; @Norderhaug2012-ks; @Harley2012], growth [@Cousens1982-ck] and productivity [@Pedersen2012-vr; @Dayton1998-xw; @Krumhansl2012] of kelps.
-
-Given kelps are cool-temperate organisms which are vulnerable to dislodgement under high wave exposure scenarios, changes in ocean temperature and an increase in storm frequency and magnitude pose the biggest direct threats to their survival.
+Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and sub-canopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002], each depending on a wide suite of ecological services provided by kelp forests [@Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @Willis2003-xo; @anderson1997; @Dayton1985].
+Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in their distribution [@Gorman2013-nj], abundance [@Cavanaugh2011-ib; @Dayton1998-xw], diversity [@Wing2007-tv; @Wernberg2008], composition [@Leliaert2000a; @Dayton1985; @Norderhaug2012-ks; @Harley2012], growth [@Cousens1982-ck] and productivity [@Pedersen2012-vr; @Dayton1998-xw; @Krumhansl2012].
+
+Given that kelps are cool-temperate organisms which are vulnerable to dislodgement under high wave exposure scenarios, increases in storm frequency and magnitude, and changes in ocean temperature, changes in these influences pose a direct threat to their survival; which is a concern considering that these variables are known to be affected by climate change[@Russo2014; @Reguero2019; @Weisse2010; @Rose2012; @Rummukainen2012].
 Warmer temperatures reduces the resilience of kelp individuals to storm disturbance [@wernberg2010], causes fragmentation through weakening tissues [@Simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
 Despite these threats and disturbances kelps persist across a broad range of environments which is largely due to their morphological plasticity [@FowlerWalker2006; @wernberg2010].
-A study by @Wernberg2003-mp investigated the morphology of *E.
-radiata* in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australasian coast.
-They found no correlation between spatial distance and morphological similarity, and rather the morphology of kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales [@Wernberg2003-mp].
+A study by @Wernberg2003-mp investigated the morphology of *Ecklonia radiata* (C.
+Agardh) J.
+Agardh.
+in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australian coast.
+They found no correlation between spatial distance and morphological similarity; rather the morphology of the kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales [@Wernberg2003-mp].
 Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
-Deep and shallow water environments may differ in abiotic processes such as the degree of water column mixing [@Smit2013a], solar heating [@Dellatorre2012; @Dunne2001] and effects of wave dampening [@Kobayashi1993-os; Mia2013; @Dubi1995].
+Deep and shallow water environments may differ in abiotic processes such as the degree of water column mixing [@Smit2013a], solar heating [@Dellatorre2012; @Dunne2001] and effects of wave dampening [@Kobayashi1993-os; @Mia2013; @Dubi1995].
 
@@
 For example, a study by @Serisawa2002-dc compared the morphology of *Ecklonia cava* Kjellman growing at warmer and cooler sites, and showed wrinkling of the blade and reduced size to be a characteristic of higher temperature ranges.
-Warmer temperature effects important kelp physiological processes such as photosynthesis and respiration, which influences growth and productivity and ultimately leads to a reduction in size [@bearham2013; @gerard1997; @gao2013; @zimmerman1986].
+Warmer temperature affects important kelp physiological processes such as photosynthesis and respiration, which influences growth and productivity and ultimately leads to a reduction in size [@bearham2013; @gerard1997; @gao2013; @zimmerman1986].
 Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002; @Hurd2000; @FowlerWalker2005; @Thomsen2004].
-For example, a study by @Wernberg2005c examined the consistency of wave exposure as a driver of *Ecklonia radiata*  (C.
-Agardh) J.
-Agardh.
-across a broad geographic range, and showed trends towards drag-reducing (small size, narrow laterals and blades, low spinosity) and increased strength (large holdfast, thick stipe and thick blades and lamina) at high wave exposure sites.
+For example, a study by @Wernberg2005c examined the consistency of wave exposure as a driver of *E.
+radiata* across a broad geographic range, and showed trends towards drag-reducing (small size, narrow laterals and blades, low spinosity) and increased strength (large holdfast, thick stipe and thick blades and lamina) at high wave exposure sites.
 
@@
 Although @mabin2013 investigated kelp morphological characteristics using a 2D WAM model, currently no studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
-Furthermore no current studies have identified specific wave metrics that drive kelp morphological characteristics, which the current work will aim to achieve.
+Furthermore no studies have identified specific wave metrics that drive kelp morphological characteristics, which the current work will aim to achieve.
 Other considerations such as possible differences in morphology between sub-canopy and canopy species, and shallow and deep water kelp populations have not been considered using advanced numerical and statistical tools.
 
-The kelps *Laminaria pallida* and *Ecklonia maxima* (Osbeck) Papenfuss are important habitat forming seaweeds that exist around the coast of South Africa and offer a unique opportunity to investigate the drivers of macroalgal morphological characteristics between canopy and sub-canopy species, as well as between deep and shallow water environments.
+The kelps *Laminaria pallida* Greville and *Ecklonia maxima* (Osbeck) Papenfuss are important habitat forming seaweeds that exist around the coast of South Africa and offer a unique opportunity to investigate the drivers of macroalgal morphological characteristics between canopy and sub-canopy species, as well as between deep and shallow water environments.
 Although both these species exist in the sub-tidal, *L.
 pallida* dominates deeper waters while *E.
-maxima* forms dense surface canopies in the sub-tidal to shallower waters.
+maxima* forms dense surface canopies in the shallow sub-littoral.
 Therefore, *E.
-maxima* may be exposed to variations in wind driven surface waves, swell waves and temperature; while *L.
+maxima* may be exposed to variations in wind driven surface waves, swell waves and temperature, while *L.
 pallida* is exposed to variations in swell waves and temperature.
 @Molloy1996-la investigated the effect of depth and wave exposure on *L.
-pallida* at different depths and wave exposure and showed that depth had a greater effect than wave exposure when considering all the morphological characteristics (wave exposure alters with depth); when considering individual characteristics, however, wave exposure had the most significant effect on blade thickness.
+pallida* at different depths and wave exposure and showed that depth had a greater effect than wave exposure when considering all the morphological characteristics; when considering individual characteristics, however, wave exposure had the most significant effect on blade thickness.
 Another study by @Rothman2017-pn investigated the changes in morphology in shallow populations of *L.
@@
 St.
-Helena Bay and Betty's Bay constituted the north western and south eastern boundary sites, respectively.
-These sites are roughly 300 km apart, lie within separate marine provinces, and span across the majority of the south-west coast, in varying thermal and wave energy regimes.
+Helena Bay and Betty's Bay constituted the north-western and south-eastern boundary sites, respectively.
+These sites are roughly 300 km apart, lie within separate marine provinces @Smit2017, and span the majority of the south-west coast, in varying thermal and wave energy regimes.
 The region is dominated by kelp communities that persist in contrasting abiotic environments.
-The west coast region has been termed as cool-temperate, which is defined as a region where mean monthly temperatures are always above 10°C and always below 15°C [@Smit2013a].
+The west coast region has been termed cool-temperate, which is defined as a region where mean monthly temperatures are always above 10°C and below 15°C [@Smit2013a].
 East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone [@Smit2013a].
@@
 The annual and seasonal patterns in both wind influence the thermal regime around the coastline by inducing upwelling.
-Upwelling brings deep, cool nutrient rich water to coastal areas which induces a decrease in temperature [@Field1980a; @Andrews1980; @Gill1974-nt; @Cram1970-jw; @Rouault2010-gi; @Blanke2002].
-The wind speed and wind direction help drive upwelling in summer on the western side of the Cape peninsula, where southerly winds blow parallel to the coast and trigger upwelling events [@Field1980a; @Rouault2010-gi].
-This is not the case for False Bay which is shielded from the dominating wind and swell the Cape Peninsula is subject to, with the exception being the winter months where upwelling becomes more favourable for the region due to the change in swell and wind direction, causing lower temperatures to dominate during this time [@Rouault2010-gi; @Andrews1980].
+Upwelling brings deep, cool nutrient-rich water to coastal areas which causes a decrease in temperature [@Field1980a; @Andrews1980; @Gill1974-nt; @Cram1970-jw; @Rouault2010-gi; @Blanke2002].
+The wind speed and wind direction help drive upwelling in summer on the western side of the Cape Peninsula, where southerly winds blow parallel to the coast and trigger upwelling events [@Field1980a; @Rouault2010-gi].
+This is not the case for False Bay, which is shielded from the dominating wind and swell that the Cape Peninsula is subjected to.
+In the winter months, upwelling becomes more favourable in False Bay due to the change in swell and wind direction, causing lower temperatures to dominate during this time [@Rouault2010-gi; @Andrews1980].
 Due to the Cape Peninsula temperate latitude, winter months bring an increased frequency of frontal depressions that originate from the Southern Ocean [@Reason2006].
@@
 Conversely, the west coast of Cape Point is battered by these large swells.
-Multiple sites, therefore, exist where kelps grow in diverse temperature and wave energy climates, in close proximity.
+Multiple sites, therefore, exist where kelps grow in diverse temperature and wave energy climates, in close proximity and on the same side.
 The topography and elevation along the Cape Peninsula channel and shield winds along False Bay.
-This is however absent in winter, where strong northerly winds are prevalent from St.
+This is, however, absent in winter, when strong northerly winds are prevalent from St.
 Helena Bay to Betty's Bay [@Field1980a; @Jury1985-nr; @Andrews1980; @Jury1980-ov].
@@
 These data were then used to model swell propagation into the coastal models while wind waves (seas) were generated via stationary computations in the Simulating Waves in the Nearshore SWAN model [@Booij1997-nx].
-The assumption of stationary computations are acceptable as the model domains were small enough so the temporal variation of the model boundary were slower than the time it takes for that boundary condition to propagate to the coast.
+The assumption of stationary computations is acceptable as the model domains were small enough so the temporal variation of the model boundary was slower than the time it takes for that boundary condition to propagate to the coast.
 SWAN allows one to extract wave variables from specific gridded locations in the nearshore.
 For False Bay, a resolution of 200 meters was modelled and output produced at both the 7 meter and 15 meter isobaths.
-A 200 meter resolution was used as the False Bay computational grid was nested within a larger grid (1 kilometer resolution).
+A 200 meter resolution was used as the False Bay computational grid was nested within a larger 1 kilometer resolution grid.
 This allowed for a computational effective wave resolution of increasing resolution, from the NCEP, low resolution output to nested, high resolution coastal output.
@@
 
-```{r schematic, echo=FALSE, fig.cap='\\label{fig:schematic}Schematic of *L. pallida* and *E. maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
+```{r schematic, echo=FALSE, fig.cap='\\label{fig:schematic}Schematic of *Laminaria pallida* and *Ecklonia maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
 
@@
 1) between October 2014 and April 2015.
-In each instance only the thirteen largest individuals were sampled to ensure that only mature and fully grown sporophytes were measured.
+In each instance only the 13 largest individuals were sampled to ensure that only mature and fully grown sporophytes were measured.
 In instances where one of the species was not present a site, the nearest population of the missing species was sampled.
@@
 
-Summary statistics were calculated for each site and regions for each of the environmental variables considered in this study, and are presented in Table 3.
-The summary statistics calculated for temperature, waves and wind variables were minimum, maximum, mean, range and standard deviation for annual timescales.
+Summary statistics were calculated for each site and region for each of the environmental variables considered in this study, and are presented in Table 3.
 In addition, annual median wave and wind direction were also calculated.
 The median was calculated for wind and wave direction, as issues arise when calculating the mean and standard deviation for compass metrics.
-Dot and whisker plots were used to present the summary statistics for both temperature and wave variables.
-Summary statistics for wind was not plotted and instead are discussed, as the data was course relative to the other environmental variables considered in this study.
+Summary statistics for wind was not plotted and instead are discussed, as the data were course relative to the other environmental variables considered in this study.
 
@@
 Morphological characteristics were categorised according to temperature regime and wave exposure category for each species and population of kelp in this study.
-Significant differences between temperature regimes and across wave exposure categories were tested with wilcox tests and kruskal tests respectively, and notched boxplots allowed comparison between individual categories of wave exposure.
-In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package [@Oksanen2015-vj] in R (R Core Team, 2019).
+Significant differences between temperature regimes and across wave exposure categories were tested with Wilcox tests and Kruskal tests, respectively, and notched boxplots allowed comparison between individual categories of wave exposure.
+In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis (db-RDA) was performed using the *rda* function in the *vegan* software package [@Oksanen2015-vj] in R (R Core Team, 2019).
 The abiotic data covered all study sites for both species of kelp.
@@
 To determine the explanatory variables that best describe patterns in the response data, a full RDA was performed using a complete set of explanatory variables.
-Forward selection was then used to reduce the number of explanatory variables, as well as prevent the inflation of overall type I error.
+Forward selection was then used to reduce the number of explanatory variables, as well as prevent the inflation of overall Type I error.
 To further improve the model, pairwise coefficients and Variance Inflation Factor (VIF) were calculated to identify variables with high multicollinearity.
@@
 Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay (see figure \ref{fig:wave variable plot}B).
-The data shows no trend in mean peak period for the coastline.
-These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side of the peninsula, with Miller’s point showing the highest variation across timescales.
+The data show no trend in mean peak period for the coastline.
+These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side, with Miller’s Point showing the highest variation across timescales.
 Maximum peak period ranged from 18.9 s at St.
@@
 When wave variables were grouped by region significant differences were found for both mean significant swell height and mean peak period.
-The western region The western region experiences a significantly higher mean significant swell height (p < 0.05; see figure \ref{fig:wave variable plot}C) and mean peak period (p < 0.05; see figure \ref{fig:wave variable plot}D).
+The western region experiences a significantly higher mean significant swell height (p < 0.05; see figure \ref{fig:wave variable plot}C) and mean peak period (*p* < 0.05; see figure \ref{fig:wave variable plot}D).
 
@@
 
-Data for median wave direction showed no trend across timescales; however the variation (SD) in wave direction is lower on the western side of the peninsula compared to the eastern side both annually and for winter.
+Data for median wave direction showed no trend across timescales; however, the variation (SD) in wave direction is lower on the western side of the peninsula compared to the eastern side both annually and for winter.
 Summer data were the exception and showed lower variation for the eastern side of the peninsula (see figure \ref{fig:wave rose}).
@@
 The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits.
-The western periphery of the Cape peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures (see figure \ref{fig:wave power map}).
+The western periphery of the Cape Peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures (see figure \ref{fig:wave power map}).
 There the marked seasonality, with higher energy waves during winter, may be clearly observed once more.
@@
 When *L.
-pallida* morphological characteristics were grouped to wave exposure categories kruskal-wallis tests revealed significant differences amongst categories for most morphological characteristics with lamina length, number of digits and thallus mass the exceptions (see figure ref\{fig:lamanaria wave plots}).
+pallida* morphological characteristics were grouped to wave exposure categories Kruskal-Wallis tests revealed significant differences amongst categories for most morphological characteristics with lamina length, number of digits and thallus mass the exceptions (see figure ref\{fig:lamanaria wave plots}).
 The morphological characteristics that were significantly different amongst sites exhibited similar patterns, with increasing mean values from the fully sheltered to extremely sheltered categories and a decrease in mean values for the remainder of the categories.
@@
 ```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.cap='\\label{fig:lamanaria temp plots}Boxplots representing *L.
-pallida* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+pallida* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
 The lower and upper "hinges" correspond to the first and third quartiles.
-The “whiskers” represent the range, solid black lines represent the median and black dots are outliers.
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.
 Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=8, fig.width=8}
@@
 ```{r lamanaria wave plots, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.height=8, fig.width=8, fig.cap='\\label{fig:lamanaria wave plots}Boxplots representing *L.
-pallida* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+pallida* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
 The lower and upper hinges correspond to the first and third quartiles.
@@
 Mean values for primary length, primary width and stipe circumference were significantly higher for kelp from the cool temperate regime than for kelp from the warm temperate regime (see figure \ref{fig:{deep eck temp plot}}.
-The remaining morphological characteristics, stipe length, number of tufts and total length had significantly lower means for kelp from cool temperate compared to kelp from the warm temperate regime (p < 0.05; see figure \ref{fig:deep eck temp plot}).
+The remaining morphological characteristics, stipe length, number of tufts and total length had significantly lower means for kelp from cool temperate compared to kelp from the warm temperate regime (*p* < 0.05; see figure \ref{fig:deep eck temp plot}).
 Significant differences amongst wave exposure categories were found for all morphological characteristics of deep *E.
@@
 maxima* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
-The lower and upper "hinges" correspond to the first and third quartiles.
-The “whiskers” represent the range, solid black lines represent the median and black dots are outliers.
+The lower and upper hinges correspond to the first and third quartiles.
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.
 Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
@@
 maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
-The lower and upper "hinges" correspond to the first and third quartiles.
-The “whiskers” represent the range, solid black lines represent the median and black dots are outliers.', fig.height=8, fig.width=8}
+The lower and upper hinges correspond to the first and third quartiles.
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.', fig.height=8, fig.width=8}
 ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3, nrow = 4)
@@
 maxima* morphological characteristics  grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
-The lower and upper "hinges" correspond to the first and third quartiles.
-The “whiskers” represent the range, solid black lines represent the median and black dots are outliers.
+The lower and upper hinges correspond to the first and third quartiles.
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.
 Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=7, fig.width=8}
@@
 maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
-The lower and upper "hinges" correspond to the first and third quartiles.
-The “whiskers” represent the range, solid black lines represent the median and black dots are outliers.', fig.height=7, fig.width=8}
+The lower and upper hinges correspond to the first and third quartiles.
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.', fig.height=7, fig.width=8}
 ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3,nrow = 3)
@@
 When grouped by temperature regime, shallow *E.
-maxima* exhibited significant differences between temperature regimes for frond mass, primary length, primary width, frond length, stipe mass, stipe length, stipe circumference and total length (p < 0.05; see figure \ref{fig:shallow ecklonia temp}).
+maxima* exhibited significant differences between temperature regimes for frond mass, primary length, primary width, frond length, stipe mass, stipe length, stipe circumference and total length (*p* < 0.05; see figure \ref{fig:shallow ecklonia temp}).
 All of the significantly different morphological characteristics, except primary length, had lower values for the western region compared to the eastern region.
 When shallow *E.
-maxima* were grouped by wave exposure category significant difference were found for all morphological characteristics (p < 0.05; see figure \ref{fig:shallow ecklonia wave}).
+maxima* were grouped by wave exposure category significant difference were found for all morphological characteristics (*p* < 0.05; see figure \ref{fig:shallow ecklonia wave}).
 In general, values of the various morphological characteristics increase for the lower wave exposure categories, followed by a sharp decrease and a gradual increase in value.
@@
 RDA1 was the only significant axis in the model and explained 48% of the variation while RDA2 was not significant and explained 63% of the variation.
-Total length, stipe length, stipe diameter and stipe mass were positively influenced (i.e., increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* for kelp mostly from the western region and negatively influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp* for kelp from the eastern region.
+Total length, stipe length, stipe diameter and stipe mass were positively influenced (*i.e.* increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* for kelp mostly from the western region and negatively influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp* for kelp from the eastern region.
 Although sites from from the western region did not cluster as closely as eastern region study sites, grouping according to region was still evident.
@@
 Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of deep *E.
-maxima* morphological characteristics with an adjusted *R*2 of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, *F* = 2.4, *p* = 0.026).
-RDA1 was the only significant axis in the model (p = 0.03) and explained 72% of the variation.
+maxima* morphological characteristics with an adjusted *R*^2^ of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, *F* = 2.4, *p* = 0.026).
+RDA1 was the only significant axis in the model (*p* = 0.03) and explained 72% of the variation.
 Primary length, primary width and stipe circumference were positively influenced by *hs_mean*, *tp_mean* and *dir_median* for for kelp from the western region.
@@
 Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of shallow *E.
-maxima* morphological characteristics with an adjusted *R*2 of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, *F* = 2.41, *p* = 0.013).
-The model consisted of one significant axis, RDA1 (p = 0.02) and explained 62% of the variation.
+maxima* morphological characteristics with an adjusted *R*^2^ of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, *F* = 2.41, *p* = 0.013).
+The model consisted of one significant axis, RDA1 (*p* = 0.02) and explained 62% of the variation.
 Stipe length, stipe mass and stipe circumference were positively influenced by *max_temp* and *mean_temp* along RDA1, and explained 50% and 55% of the variation respectively.
@@
 
-This study considered important drivers of kelp morphology that have been identified in previous research [@FowlerWalker2006; @Miller2011-pm; @FowlerWalker2005; @Stewart2009; @Wing2007-tv; @Molloy1996-la; @Pedersen2012-vr; @Wernberg2005c; @Thomsen2004], namely temperature and wave exposure, around the Cape Peninsula on two prolific kelp species.
+This study considered important drivers of the morphology of *Ecklonia maxima* and *Laminaria pallida* arpound tehe Cape Peninsula.
+These abiotic variables, namely temperature and wave exposure, have been identified in previous research [@FowlerWalker2006; @Miller2011-pm; @FowlerWalker2005; @Stewart2009; @Wing2007-tv; @Molloy1996-la; @Pedersen2012-vr; @Wernberg2005c; @Thomsen2004].
 The complex geomorphology of the western Cape coastline creates an ideal natural laboratory for studies of kelps in interaction with their environment.
@@
 We also considered wind in addition to temperature and wave variables, since wind is an important component of wind-driven surface gravity waves [@Holthuijsen2010-vl].
-The results show that specific variables of wave exposure are the main drivers of kelp morphology with temperature only playing a minor role.
+The results show that specific variables of wave exposure are the main drivers of kelp morphology with temperature playing only a minor role.
 Furthermore, the investigation of differences in morphological characteristics between shallow and deep populations of *E.
-maxima* suggests that in low wave exposure environments the role of temperature as a morphological driver increases, such as in False Bay.
+maxima* suggests that in low wave exposure environments the role of temperature as a morphological driver increases; this is particulary the case for False Bay.
 
@@
 It is due to the near consistent south-westerly swell and the complex orography around the peninsula that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.
-The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposure maps presented in figure \ref{fig:wave power map}.It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions.
+The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposure maps.
+It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions.
 This results in positive and negative wave exposure anomalies all around the peninsula.
@@
 When grouped by wave exposure category, *L.
-pallida* characteristics showed a significant increase in length for the 'Extremely sheltered' category (6-10 kW/m; figure \ref{fig:wave power map}) compared to the fully sheltered category (<5 kW/m; see figure \ref{fig:wave power map}), which may be a 'go-with-the-flow' tactic.
+pallida* characteristics showed a significant increase in length for the 'extremely sheltered' category compared to the fully sheltered category, which may be a 'go-with-the-flow' tactic [@Denny1997; @Denny2002; @Hurd2000].
 Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny2002; @Hurd2000].
 However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under high wave exposure [@Denny2002; @Denny1997].
-This is reflected in figure \ref{fig:lamanaria wave plots} which shows an overall reduction of *L pallida* morphological characteristics suggesting a size reducing tactic to cope with higher levels of wave exposure.
+This is reflected in the *L.
+pallida* morphology which shows an overall reduction of morphological characteristics suggesting a size reducing tactic to cope with higher levels of wave exposure.
 The results suggest that deep *E.
@@
 When grouped to the higher wave exposure levels, *E.
-maxima* morphology exhibited a strength and flexibility increasing trait (see figure \ref{fig:deep ecklonia wave}).
+maxima* morphology exhibited a strength and flexibility increasing trait.
 Furthermore, the morphological response of *E.
@@
 Solar heating of the water surface in combination with wind-driven transport causes fluctuations within the water column, which can occur daily or seasonally [@Kaplan2003].
-Vertical mixing of the water column for inshore regions are driven by several abiotic processes: (1) Turbulence of breaking waves inside or outside of the surf zone (2) convective mixing through a combination of cooling and evaporation (3) wind driven currents and (4) tidal mixing [@Smit2013a].
+Vertical mixing of the water column for inshore regions are driven by several abiotic processes: (1) turbulence of breaking waves inside or outside of the surf zone (2) convective mixing through a combination of cooling and evaporation; and (3) wind driven currents and (4) tidal mixing [@Smit2013a].
 These abiotic processes cause effective vertical mixing of surface and deeper water stratification which leads to a homogeneous thermal environment, which in this study may be interpreted as the mean temperature.

### 2019-11-29 morphology_paper.Rmd
checked consistancy of in text figure references. Significance levels on graphs adjusted.

@@
 author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
-bibliography: references.bib
 output:
+  word_document:
+    reference_docx: frontiers_template.docx
+  pdf_document:
+    fig_caption: yes
   html_document:
     df_print: paged
-  pdf_document:
-    fig_caption: yes
-  word_document:
-    reference_docx: frontiers_template.docx
-geometry: margin=1in
-fontsize: 10pt
+bibliography: references.bib
 papersize: A4
@@
 Helena Bay, the most northern site on the western side of the peninsula, to 14.6 ± 0.3 ℃ at Olifantsbos, the most southern site on the western side of the peninsula (see figure \ref{fig:Temperature variables plot}A).
-For sites located on the eastern side of the peninsula (within False Bay), the annual mean coastal water temperatures ranged from 15.5 ± 0.9 ℃ at Buffelsbaai to 15.0 ± 0.9 ℃ at Betty’s Bay, the most eastern site on the coastline in this study (see figure \ref{fig:Temperature variables plot}A).
+For sites located on the eastern side of the peninsula (within False Bay), the annual mean coastal water temperatures ranged from 15.5 ± 0.9 ℃ at Buffelsbaai to 15.0 ± 0.9 ℃ at Betty’s Bay, the most eastern site on the coastline in this study (see figure\ref{fig:Temperature variables plot}A).
 Oudekraal had the lowest annual mean coastal water temperature of 12.3 ± 0.9.
@@
   geom_boxplot(show.legend = FALSE, width = 0.5) + 
-  stat_compare_means(label.x = 2.5, label.y = 12.5, label = "p.format") +
+  stat_compare_means(label.x = 2.5, label.y = 12.5, label = "p.signif") +
   xlab("") + 
@@
   theme(axis.text.y=element_text(vjust=0.3))
+
 ```
@@
 In general, the mean temperature increases for sites located within False Bay with less variation around the mean temperature compared to sites on the western side of the peninsula(\ref{fig:Temperature variables plot}A).
-This is verified by \ref{fig:Temperature variables plot}B, which shows a significantly lower (p < 0.05, figure \ref{fig:Temperature variables plot}B) mean temperature and less variation in temperature for sites located on the western side of the peninsula compared to sites on the eastern side.
+This is verified by figure\ref{fig:Temperature variables plot}B, which shows a significantly lower (*p* < 0.05, figure \ref{fig:Temperature variables plot}B) mean temperature and less variation in temperature for sites located on the western side of the peninsula compared to sites on the eastern side.
 
@@
 When wave variables were grouped by region significant differences were found for both mean significant swell height and mean peak period.
-The western region experiences a significantly higher mean significant swell height (p < 0.05; see figure \ref{fig:wave variable plot}C) and mean peak period (*p* < 0.05; see figure \ref{fig:wave variable plot}D).
+The western region experiences a significantly higher mean significant swell height (*p* < 0.05; see figure \ref{fig:wave variable plot}C) and mean peak period (*p* < 0.05; see figure \ref{fig:wave variable plot}D).
 
@@
   geom_boxplot(show.legend = FALSE) + 
-  stat_compare_means(label.y =0.6 ,label.x = 2.5, label = "p.format") + 
+  stat_compare_means(label.y =0.6 ,label.x = 2.5, label = "p.signif") + 
   xlab("") + 
@@
   geom_boxplot(show.legend = FALSE) + 
-  stat_compare_means(label.y = 8.2, label.x = 2.5, label = "p.format") + 
+  stat_compare_means(label.y = 8.2, label.x = 2.5, label = "p.signif") + 
   xlab("") + 
@@
 
-In Figure \ref{fig:wave power map} the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
+In figure \ref{fig:wave power map} the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
 The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits.
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.62, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.62, label.y.npc = c('top')) + 
   xlab("") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   xlab("") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.65, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.65, label.y.npc = c('top')) + 
   xlab("") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.85, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.85, label.y.npc = c('top')) + 
   ylab("Lamina weight (g)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("Lamina length (cm)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("Lamina thickness") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
   ylab("Stipe diameter (cm)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
   ylab("No. of digits") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
   ylab("Thallus mass (kg)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
   ylab("Total length (cm)") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Frond mass (g)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Primary length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Primary width (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Frond length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe circ (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("No. of tufts") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Epi length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = TRUE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
   ylab("Total length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   xlab("") + 
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   ylab("Primary length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   xlab("") + 
@@
 ```{r deep ecklonia temp plot, echo=FALSE, fig.cap='\\label{fig:deep eck temp plot}Boxplots representing deep population *E.
-maxima* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+maxima* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
 The lower and upper hinges correspond to the first and third quartiles.
@@
 ```{r deep ecklonia wave, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:deep ecklonia wave}Boxplots representing deep *E.
-maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
 The lower and upper hinges correspond to the first and third quartiles.
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Frond mass (g)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Primary length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Primary width (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Frond length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Stipe circ (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("No. of tufts") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Epi length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
-  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 0.6, label.y.npc = c('top')) + 
+  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
   ylab("Total length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   ylab("Frond mass (g)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   ylab("Primary width (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   ylab("Frond length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   ylab("Stipe mass (g)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   ylab("Stipe length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
  ylim(0,30) +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   ylab("No of Tufts") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE)  + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   ylab("Epi length (cm)") +
@@
   geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) + 
-  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
   ylab("Total length (cm)") +
@@
 ```{r shallow ecklonia temp, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.cap='\\label{fig:shallow ecklonia temp} Boxplots representing shallow *E.
-maxima* morphological characteristics  grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+maxima* morphological characteristics  grouped by temperature regime measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
 The lower and upper hinges correspond to the first and third quartiles.
@@
 ```{r shallow ecklonia wave, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:shallow ecklonia wave}Boxplots representing shallow *E.
-maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
 The lower and upper hinges correspond to the first and third quartiles.
@@
 
-```{r}
-
-```
-
 ```{r appendix figure 1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=8}

### 2019-12-01 morphology_paper.Rmd
changes to yaml, added scritp to save plot panels, and then some

@@
 tables: yes
-abstract: 'Macroalgal morphological variation is determined to a large extent by
+abstract:  '**Abstract**: Macroalgal morphological variation is determined to a large extent by
   a combination of environmental factors, with wave exposure and temperature
@@
   deep and shallow populations.'
+
+nocite:  |
+  @Algaebase
 ---
@@
   theme(axis.text.y=element_text(vjust=0.3)) 
+ggsave("hs_plot.jpeg", plot = wave_boxplot_1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
   theme(axis.text.y=element_text(vjust=0.3))
+ggsave("tp_plot.jpeg", plot = wave_boxplot_2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 
+ggsave("fig8A.jpeg", plot = LT1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LT2 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
@@
 
+ggsave("fig8B.jpeg", plot = LT2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LT3 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
@@
 
+ggsave("fig8C.jpeg", plot = LT3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LT4 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
@@
 
+ggsave("fig8D.jpeg", plot = LT4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LT5 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
@@
 
+ggsave("fig8E.jpeg", plot = LT5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LT6 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
@@
   xlab("") +
-  scale_fill_grey(start = 0.8, end = 0.4)
+  scale_fill_grey()
+
+ggsave("fig8F.jpeg", plot = LT6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 
+ggsave("fig8G.jpeg", plot = LT7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LT8 <- ggplot(data = lam_data, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
@@
 
+ggsave("fig8H.jpeg", plot = LT8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LT9 <- ggplot(data = lam_data, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig8I.jpeg", plot = LT9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 
+ggsave("fig9A.jpeg", plot = LW1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LW2 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
@@
 
+ggsave("fig9B.jpeg", plot = LW2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LW3 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
@@
 
+ggsave("fig9C.jpeg", plot = LW3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LW4 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
@@
 
+ggsave("fig9D.jpeg", plot = LW4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LW5 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
@@
 
+ggsave("fig9E.jpeg", plot = LW5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LW6 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
@@
 
+ggsave("fig9F.jpeg", plot = LW6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LW7 <- ggplot(data = lam_data, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
@@
 
+ggsave("fig9G.jpeg", plot = LW7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LW8 <- ggplot(data = lam_data, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
@@
   xlab("")
+
+ggsave("fig9H.jpeg", plot = LW8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 LW9 <- ggplot(data = lam_data, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
@@
   xlab("")
+
+ggsave("fig9I.jpeg", plot = LW9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ```
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig10A.jpeg", plot = ET1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ET2 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig10B.jpeg", plot = ET2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ET3 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig10C.jpeg", plot = ET3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ET4 <- ggplot(data = eck_deep, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig10D.jpeg", plot = ET4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ET5 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig10E.jpeg", plot = ET5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ET6 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig10F.jpeg", plot = ET6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ET7 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig10G.jpeg", plot = ET7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ET8 <- ggplot(data = eck_deep, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig10H.jpeg", plot = ET8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ET9 <- ggplot(data = eck_deep, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig10I.jpeg", plot = ET9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ET10 <- ggplot(data = eck_deep, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig10J.jpeg", plot = ET10, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig11A.jpeg", plot = EW1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EW2 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig11B.jpeg", plot = EW2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EW3 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig11C.jpeg", plot = EW3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EW4 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig11D.jpeg", plot = EW4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EW5 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig11E.jpeg", plot = EW5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EW6 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig11F.jpeg", plot = EW6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EW7 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +
@@
   xlab("") + 
-  ylab("Stipe circumference (cm)") +
+  ylab("Stipe circ (cm)") +
   theme_bw() +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig11G.jpeg", plot = EW7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EW8 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig11H.jpeg", plot = EW8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EW9 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig11I.jpeg", plot = EW9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EW10 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig11J.jpeg", plot = EW10, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ```
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig12A.jpeg", plot = EST1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EST2 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig12B.jpeg", plot = EST2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EST3 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig12C.jpeg", plot = EST3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EST4 <- ggplot(data = eck_shall, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig12D.jpeg", plot = EST4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EST5 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig12E.jpeg", plot = EST5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EST6 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig12F.jpeg", plot = EST6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EST7 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig12G.jpeg", plot = EST7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EST8 <- ggplot(data = eck_shall, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig12H.jpeg", plot = EST8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EST9 <- ggplot(data = eck_shall, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig12I.jpeg", plot = EST9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 EST10 <- ggplot(data = eck_shall, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
+
+ggsave("fig12J.jpeg", plot = EST10, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig13A.jpeg", plot = ESW1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ESW2 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig13B.jpeg", plot = ESW2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ESW3 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
@@
   ylim(0,25)
+
+ggsave("fig13C.jpeg", plot = ESW3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ESW4 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig13D.jpeg", plot = ESW4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ESW5 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig13E.jpeg", plot = ESW5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ESW6 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig13F.jpeg", plot = ESW6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ESW7 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +
@@
  ylim(0,30) +
-  ylab("Stipe circumference (cm)") +
+  ylab("Stipe circ (cm)") +
   theme_bw() +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig13G.jpeg", plot = ESW7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ESW8 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig13H.jpeg", plot = ESW8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ESW9 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig13I.jpeg", plot = ESW9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
+
 ESW10 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
+
+ggsave("fig13J.jpeg", plot = ESW10, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 

### 2019-12-02 morphology_paper.Rmd
Final push for paper. output may be manually edited for some aspects but the changes will be incorportaed into the rmd

@@
 ```
-
 # Introduction
@@
 Given that kelps are cool-temperate organisms which are vulnerable to dislodgement under high wave exposure scenarios, increases in storm frequency and magnitude, and changes in ocean temperature, changes in these influences pose a direct threat to their survival; which is a concern considering that these variables are known to be affected by climate change[@Russo2014; @Reguero2019; @Weisse2010; @Rose2012; @Rummukainen2012].
-Warmer temperatures reduces the resilience of kelp individuals to storm disturbance [@wernberg2010], causes fragmentation through weakening tissues [@Simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
-Despite these threats and disturbances kelps persist across a broad range of environments which is largely due to their morphological plasticity [@FowlerWalker2006; @wernberg2010].
+Warmer temperatures reduces the resilience of kelp individuals to storm disturbance [@Wernberg2010b], causes fragmentation through weakening tissues [@Simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
+Despite these threats and disturbances kelps persist across a broad range of environments which is largely due to their morphological plasticity [@FowlerWalker2006; @Wernberg2010a].
 A study by @Wernberg2003-mp investigated the morphology of *Ecklonia radiata* (C.
@@
   theme(axis.text.y=element_text(vjust=0.3))
-ggsave("hs_plot.jpeg", plot = wave_boxplot_1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
   theme(axis.text.y=element_text(vjust=0.3))
-ggsave("tp_plot.jpeg", plot = wave_boxplot_2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 ```{r wave rose, echo=FALSE, warning = FALSE, eval= TRUE, fig.align="center",fig.cap='\\label{fig:wave rose}The total and seasonal wave roses of the directional wave buoy just offshore from Cape Town is given form the years 2000 till end 2017 (the period for which directional wave spectra was available).
-Annual (top left), Austral summer (top right) and Austral winter (bottom) respectively'}
+Total (A), Austral winter (B) and Austral summer (C) respectively'}
+
+library(png)
+library(jpeg)
+library(grid)
+library(gridExtra)
+
+img_wave_rose_A <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_5A.jpg")))
+img_wave_rose_B <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_5B.jpg")))
+img_wave_rose_C <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_5C.jpg")))
+
+ggarrange(img_wave_rose_A, img_wave_rose_B, img_wave_rose_C, ncol = 2, nrow = 2, labels = "AUTO")
+```
+
+Data for median wave direction showed no trend across timescales; however, the variation (SD) in wave direction is lower on the western side of the peninsula compared to the eastern side both annually and for winter.
+Summer data were the exception and showed lower variation for the eastern side of the peninsula (see figure \ref{fig:wave rose}).
+
+```{r wave power map, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:wave power map}The total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length.
+Total (A), summer (B), winter (C), spring (D) and autumn (E).'}
 
@@
 
-img_wave_rose <- readPNG("figures/wave_rose.png")
-grid.raster(img_wave_rose)
-
-```
-
-Data for median wave direction showed no trend across timescales; however, the variation (SD) in wave direction is lower on the western side of the peninsula compared to the eastern side both annually and for winter.
-Summer data were the exception and showed lower variation for the eastern side of the peninsula (see figure \ref{fig:wave rose}).
-
-```{r wave power map, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:wave power map}The total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length.
-Annual total (A), summer (B), winter (C), spring (D) and autumn (D).'}
-
-library(png)
-library(grid)
-library(gridExtra)
-
-wp_1 <- rasterGrob(as.raster(readPNG("figures/CapePen_Total.PNG")), interpolate = FALSE)
-wp_2 <- rasterGrob(as.raster(readPNG("figures/CapePen_Summer.PNG")), interpolate = FALSE)
-wp_3 <- rasterGrob(as.raster(readPNG("figures/CapePen_Winter.PNG")), interpolate = FALSE)
-wp_4 <- rasterGrob(as.raster(readPNG("figures/CapePen_Spring.PNG")), interpolate = FALSE)
-wp_5 <- rasterGrob(as.raster(readPNG("figures/CapePen_Autumn.PNG")), interpolate = FALSE)
-wp_6 <- rasterGrob(as.raster(readPNG("figures/CapePen_Legend.PNG")), interpolate = FALSE)
-grid.arrange(wp_1, wp_2, wp_3, wp_4, wp_5, wp_6, ncol = 3, nrow = 2)
+wp_1 <- rasterGrob(as.raster(readPNG("figures/CapePen_Total.PNG")))
+wp_2 <- rasterGrob(as.raster(readPNG("figures/CapePen_Summer.PNG")))
+wp_3 <- rasterGrob(as.raster(readPNG("figures/CapePen_Winter.PNG")))
+wp_4 <- rasterGrob(as.raster(readPNG("figures/CapePen_Spring.PNG")))
+wp_5 <- rasterGrob(as.raster(readPNG("figures/CapePen_Autumn.PNG")))
+wp_6 <- rasterGrob(as.raster(readPNG("figures/CapePen_Legend.PNG")))
+
+ggarrange(wp_1, wp_2, wp_3, wp_4, wp_5, wp_6, ncol = 3, nrow = 2, labels = "AUTO", align = "hv")
 
@@
 #grid.raster(wave_power_map)
-
 ```
@@
 
-ggsave("fig8A.jpeg", plot = LT1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LT2 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
@@
 
-ggsave("fig8B.jpeg", plot = LT2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LT3 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
@@
 
-ggsave("fig8C.jpeg", plot = LT3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LT4 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
@@
 
-ggsave("fig8D.jpeg", plot = LT4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LT5 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
@@
 
-ggsave("fig8E.jpeg", plot = LT5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LT6 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
@@
 
-ggsave("fig8F.jpeg", plot = LT6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LT7 <- ggplot(data = lam_data, aes(x = temp_regime, y = digits, fill = temp_regime)) +
@@
 
-ggsave("fig8G.jpeg", plot = LT7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LT8 <- ggplot(data = lam_data, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
@@
 
-ggsave("fig8H.jpeg", plot = LT8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LT9 <- ggplot(data = lam_data, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
-
-ggsave("fig8I.jpeg", plot = LT9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 
-ggsave("fig9A.jpeg", plot = LW1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LW2 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
@@
 
-ggsave("fig9B.jpeg", plot = LW2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LW3 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
@@
 
-ggsave("fig9C.jpeg", plot = LW3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LW4 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
@@
 
-ggsave("fig9D.jpeg", plot = LW4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LW5 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
@@
 
-ggsave("fig9E.jpeg", plot = LW5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LW6 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
@@
 
-ggsave("fig9F.jpeg", plot = LW6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LW7 <- ggplot(data = lam_data, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
@@
 
-ggsave("fig9G.jpeg", plot = LW7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LW8 <- ggplot(data = lam_data, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
@@
 
-ggsave("fig9H.jpeg", plot = LW8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 LW9 <- ggplot(data = lam_data, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
@@
   xlab("")
-
-ggsave("fig9I.jpeg", plot = LW9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 
-ggsave("fig10A.jpeg", plot = ET1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ET2 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
@@
 
-ggsave("fig10B.jpeg", plot = ET2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ET3 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
@@
 
-ggsave("fig10C.jpeg", plot = ET3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ET4 <- ggplot(data = eck_deep, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
@@
 
-ggsave("fig10D.jpeg", plot = ET4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ET5 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
@@
 
-ggsave("fig10E.jpeg", plot = ET5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ET6 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
@@
 
-ggsave("fig10F.jpeg", plot = ET6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ET7 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
@@
 
-ggsave("fig10G.jpeg", plot = ET7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ET8 <- ggplot(data = eck_deep, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
@@
 
-ggsave("fig10H.jpeg", plot = ET8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ET9 <- ggplot(data = eck_deep, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
@@
 
-ggsave("fig10I.jpeg", plot = ET9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ET10 <- ggplot(data = eck_deep, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
-
-ggsave("fig10J.jpeg", plot = ET10, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 
-ggsave("fig11A.jpeg", plot = EW1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EW2 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
@@
 
-ggsave("fig11B.jpeg", plot = EW2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EW3 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
@@
 
-ggsave("fig11C.jpeg", plot = EW3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EW4 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
@@
 
-ggsave("fig11D.jpeg", plot = EW4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EW5 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
@@
 
-ggsave("fig11E.jpeg", plot = EW5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EW6 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
@@
 
-ggsave("fig11F.jpeg", plot = EW6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EW7 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +
@@
 
-ggsave("fig11G.jpeg", plot = EW7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EW8 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
@@
 
-ggsave("fig11H.jpeg", plot = EW8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EW9 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
@@
 
-ggsave("fig11I.jpeg", plot = EW9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EW10 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-
-ggsave("fig11J.jpeg", plot = EW10, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 
-ggsave("fig12A.jpeg", plot = EST1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EST2 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
@@
 
-ggsave("fig12B.jpeg", plot = EST2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EST3 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
@@
 
-ggsave("fig12C.jpeg", plot = EST3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EST4 <- ggplot(data = eck_shall, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
@@
 
-ggsave("fig12D.jpeg", plot = EST4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EST5 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
@@
 
-ggsave("fig12E.jpeg", plot = EST5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EST6 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
@@
 
-ggsave("fig12F.jpeg", plot = EST6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EST7 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
@@
 
-ggsave("fig12G.jpeg", plot = EST7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EST8 <- ggplot(data = eck_shall, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
@@
 
-ggsave("fig12H.jpeg", plot = EST8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EST9 <- ggplot(data = eck_shall, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
@@
 
-ggsave("fig12I.jpeg", plot = EST9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 EST10 <- ggplot(data = eck_shall, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
@@
   scale_fill_grey(start = 0.8, end = 0.4)
-
-ggsave("fig12J.jpeg", plot = EST10, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 
-ggsave("fig13A.jpeg", plot = ESW1, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ESW2 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
@@
 
-ggsave("fig13B.jpeg", plot = ESW2, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ESW3 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
@@
 
-ggsave("fig13C.jpeg", plot = ESW3, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ESW4 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
@@
 
-ggsave("fig13D.jpeg", plot = ESW4, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ESW5 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
@@
 
-ggsave("fig13E.jpeg", plot = ESW5, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ESW6 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-
-ggsave("fig13F.jpeg", plot = ESW6, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 
-ggsave("fig13G.jpeg", plot = ESW7, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ESW8 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
@@
 
-ggsave("fig13H.jpeg", plot = ESW8, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ESW9 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
@@
 
-ggsave("fig13I.jpeg", plot = ESW9, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
-
 ESW10 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
@@
   guides(fill=guide_legend(title="Wave exposure level"))
-
-ggsave("fig13J.jpeg", plot = ESW10, dpi = 300, device = "jpeg", path = "extra/morphology_paper_prep/")
 
@@
 
+Kelps living in warmer environments have been shown to have a lower resilience to purbutation, such as storms [@Wernberg2010b].
+The kelps in False Bay, may therefore be more vulnerable to dislodgment under high wave exposure scenarios compared to kelps on the western side of the Cape Peninsula, which experiences higher wave energy on a regular basis.
+In South Africa, the main contributing factors to the hydrodynamic environment are the Southern Annular Node (SAM) and El Nino Southern Oscillation (ENSO), while the latter is also the main source of variability [@Veitch2019].
+These climate systems are expected to alter storm frequency and magnitude under future climate scenarios, which may increase the overall wave power around the Cape Peninsula.
+The magnitude-specific response of the kelp species to wave exposure in this study suggests there is a limit to morphological adaption in reducing the probability of dislodgement.
+Therefore, if this limit is reached and exceeded in future climate scenarios, kelp populations may risk being reduced over time.
+Furthermore, although recent studies suggest a cooling trend of coastal waters around South Africa [@Rouault2010-gi; @Smit2013a], the impact of marine heat waves may reduce the resilience of kelps to storms and high wave energy.
+Although this study identifies specific wave and temperature metrics that drive morphological adaptation and suggests a possible wave exposure threshold, it does not identify the metrics or magnitude of wave exposure required to dislodge kelp.
+Therefore, future studies should consider using advanced numerical methods to investigate the specific wave power needed to dislodge kelp, and ideally, given a specific set of morphological measures.
+These types of studies in conjunction with forecasting studies can be used to determine the vulnerability of kelp populations to future climate scenarios.
+
 # Conclusion
@@
 RC conceptualised the scope of the research reported in this paper, collected samples, undertook the majority of numerical and statistical analyses, made the first round of interpretation, and did the majority of the writing.
-AS and CR provided guidance in terms of  writing, statistical analyses, and the final draft of the paper, as well some writing in the relevant sections.
-RS aided in fieldwork and statistical analyses for the temperature data.
-TP performed the initial analyses and interpretation of the data.
+AS and CR provided guidance in terms of  writing, statistical analyses, and the final draft of the paper.
 
@@
 
-The research was funded by the South African National Research Foundation ([http://www.nrf.ac.za](http://www.nrf.ac.za)) programme "Thermal characteristics of the South African nearshore:implications for biodiversity" (CPRR14072378735).
+The research was funded by the South African National Research Foundation ([http://www.nrf.ac.za](http://www.nrf.ac.za)). Funding number is CSRP170430229220
 

### 2020-01-06 morphology_paper.Rmd
updated map and word output

@@
 title: Using numerical modelling to investigate temperature and wave exposure as drivers
-  of macroalgal morphological characteristics
+  of morphological characteristics of the kelps, Ecklonia maxima and Laminaria pallida
 author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
@@
 
-```{r map study sites, echo=FALSE, fig.height=5, fig.width=5, message=FALSE, warning=FALSE, fig.cap="\\label{fig:map study sites}Map of the study area and sampling sites.
+```{r map processing, message=FALSE, warning=FALSE, include=FALSE}
+## Load software:
+library(ggplot2)
+library(rgdal)
+library(tidyverse)
+library(viridis)
+library(gridExtra)
+source("map/scaleBarFunc.R")
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+## Read in metadata
+study_sites <- read.csv("map/metadata_map.csv", sep = ";")
+
+study_sites_edit <- study_sites %>%
+  filter(site == "St Helena Bay"|site == "Betty's Bay")
+
+## Make first map of whole study area
+map_A <- ggplot() +
+geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+            size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = study_sites, aes(x = lon, y = lat), color = "red") +
+  geom_text(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = -0.1, nudge_y = -0.01) +
+  annotate("text", label = "WARM\nTEMPERATE", x = 19.0, y = -34.5, size = 4.0, angle = 0, colour = "black") +
+  annotate("text", label = "COOL\nTEMPERATE", x =17.8, y = -33.5, size = 4.0, angle = 0, colour = "black") +
+  annotate("text", label = "Cape\nPeninsula", x = 18.4, y = -34.5, size = 3.0, angle = 0, colour = "black") +
+  scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
+  arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
+  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
+  theme(axis.ticks = element_blank()) + # Remove tick marks
+  xlab("Latitude") +
+  ylab("Longitude")
+ggsave(filename = "map_A.jpg", plot = last_plot(), device = "jpeg", path = "map/")
+## Make 2nd map of just the peninsula
+map_B <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = study_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
+  geom_text(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0.02, nudge_y = 0) +
+  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -34.100)) +
+  theme(axis.ticks = element_blank(), # Remove tick marks
+        axis.text = element_blank(), # Remove lat/ lon numbers
+        axis.title = element_blank()) # Remove lat/ lon labels
+ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")
+
+# Convert the figures to grobs
+grob_A <- ggplotGrob(map_A)
+grob_B <- ggplotGrob(map_B)
+
+## Create combined map
+map_combined <- ggplot() +
+  # First set the x and y axis limits so we know what the ranges are
+  coord_equal(xlim = c(1, 10), ylim = c(1, 10), ratio = 1, expand = TRUE, clip = "on") +
+  # Then we place our facetsover one another using the coordinates we created
+  annotation_custom(grob_A,
+                    xmin = 1, xmax = 10, ymin = 1, ymax = 10) +
+  annotation_custom(grob_B,
+                    xmin = 4, xmax = 10, ymin = 4.5, ymax = 10)
+ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8)
+```
+
+```{r map study sites, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="\\label{fig:map study sites}Map of the study area and sampling sites.
 The red dot indicates the position of Cape Town for sake of reference."}
-
-library(png)
-library(grid)
-library(gridExtra)
-
-map <- readPNG("figures/sa_map.png")
-grid.raster(map)
+map_combined
 ```
@@
 ```{r schematic, echo=FALSE, fig.cap='\\label{fig:schematic}Schematic of *Laminaria pallida* and *Ecklonia maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
-
-schematic <- readPNG("figures/kelp_diagram.png")
+library(jpeg)
+library(grid)
+library(gridExtra)
+
+schematic <- readJPEG("figures/kelp_diagram.jpg")
 grid.raster(schematic)
@@
 
-library(png)
+library(jpeg)
 library(grid)
 library(gridExtra)
-
-wp_1 <- rasterGrob(as.raster(readPNG("figures/CapePen_Total.PNG")))
-wp_2 <- rasterGrob(as.raster(readPNG("figures/CapePen_Summer.PNG")))
-wp_3 <- rasterGrob(as.raster(readPNG("figures/CapePen_Winter.PNG")))
-wp_4 <- rasterGrob(as.raster(readPNG("figures/CapePen_Spring.PNG")))
-wp_5 <- rasterGrob(as.raster(readPNG("figures/CapePen_Autumn.PNG")))
-wp_6 <- rasterGrob(as.raster(readPNG("figures/CapePen_Legend.PNG")))
+library(ggpubr)
+
+wp_1 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6A.jpg")))
+wp_2 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6B.jpg")))
+wp_3 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6C.jpg")))
+wp_4 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6D.jpg")))
+wp_5 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6E.jpg")))
+wp_6 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6_legend.jpg")))
 

### 2020-01-07 thesis_chapter_2.Rmd
major house cleaning and re-organisng

--- 2020-01-06 morphology_paper.Rmd
@@
   word_document:
-    reference_docx: frontiers_template.docx
+    reference_docx: markdown/frontiers_template.docx
+  html_document:
+    df_print: paged
   pdf_document:
     fig_caption: yes
-  html_document:
-    df_print: paged
 bibliography: references.bib
@@
 
-Sites were chosen to represent an array of temperature and wave gradients, as indicated in figure \ref{fig:map study sites}.
+Sites were chosen to represent an array of temperature and wave gradients, as indicated in figure \ref{fig:map}.
 St.
@@
 
-```{r map study sites, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="\\label{fig:map study sites}Map of the study area and sampling sites.
+```{r map, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="\\label{fig:map}Map of the study area and sampling sites.
 The red dot indicates the position of Cape Town for sake of reference."}
@@
 
-```{r schematic, echo=FALSE, fig.cap='\\label{fig:schematic}Schematic of *Laminaria pallida* and *Ecklonia maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
+```{r schematic, echo=FALSE, warning=FALSE, fig.cap='\\label{fig:schematic}Schematic of *Laminaria pallida* and *Ecklonia maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
 library(jpeg)
@@
 
-schematic <- readJPEG("figures/kelp_diagram.jpg")
+schematic <- readJPEG("figures/chapter_2/fig_2.jpg")
 grid.raster(schematic)
@@
 
-img_wave_rose_A <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_5A.jpg")))
-img_wave_rose_B <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_5B.jpg")))
-img_wave_rose_C <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_5C.jpg")))
+img_wave_rose_A <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_5A.jpg")))
+img_wave_rose_B <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_5B.jpg")))
+img_wave_rose_C <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_5C.jpg")))
 
@@
 
-wp_1 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6A.jpg")))
-wp_2 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6B.jpg")))
-wp_3 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6C.jpg")))
-wp_4 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6D.jpg")))
-wp_5 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6E.jpg")))
-wp_6 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6_legend.jpg")))
+wp_1 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6A.jpg")))
+wp_2 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6B.jpg")))
+wp_3 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6C.jpg")))
+wp_4 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6D.jpg")))
+wp_5 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6E.jpg")))
+wp_6 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6_legend.jpg")))
 
@@
 
-library(png)
 library(grid)
@@
 
-img_fb <- readJPEG("figures/falsebay_spectrum.jpg")
+img_fb <- readJPEG("figures/chapter_2/fig_7.jpg")
 grid.raster(img_fb)
@@
 maxima* exhibits a size reducing tactic with increasing wave exposure by  decreasing stipe mass, stipe length, stipe circumference, total length and frond length.
-A magnitude related response has been suggested by @wernberg2010 who investigated the temporal and spatial variations of various environmental drivers of the kelp *E.
+A magnitude related response has been suggested by @Wernberg2010a who investigated the temporal and spatial variations of various environmental drivers of the kelp *E.
 radiata*.

### 2020-01-07 thesis_chapter_3.Rmd
major house cleaning and re-organisng

---
title: 'Hydrodynamic and particle dispersion modelling of floating solitary kelp:
  An experimental approach'
author: "Ross Coppin"
date: "16/10/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Importance of kelp rafting as a dispersal mechanism

Discuss previous rafting studies and the gaps within them
Identify important coeffients

> Bouyancy (Net)
> Drag and resistance
> Effect of wind on surface currents/waves 
> Stokes drift
> Wind and swell direction
> Direct effect of wind on kelp
> Dispersal scale

Highlight the gaps and explain how current study will address gaps

Aims and objectives

# Methods
## Study species
## Study site
## Calculating coefficents
## Dispersal model

# Results
## Coeffients
## Dispersal model

# Discussion

### 2020-01-07 figures/chapter_2/frontiers figure prep.R

@@ -1,0 +1,910 @@
+### Remember to run previous code in the rmd file
+### Plots for temperature variables
+positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock", "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
+
+# Annual temp
+temp_plot <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
+  geom_point(show.legend = FALSE) +
+  scale_x_discrete(limits = positions_temp) +
+  ylim(9, 21) +
+  geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) +
+  geom_point(data = temp_ann, aes(x = site, y = max_temp), show.legend = FALSE, shape = 17, size =2) +
+  geom_point(data = temp_ann, aes(x = site, y = min_temp), show.legend = FALSE, shape = 6, size =2) +
+  xlab("") +
+  ylab("") +
+  theme(legend.position = "none") +
+  coord_flip() +
+  theme_bw() +
+  theme(axis.text.y=element_text(vjust=0.3))
+
+ggsave("fig_3A.jpeg", plot = temp_plot, dpi = 300, device = "jpeg" , path = "figures/")
+
+temp_ann_meta <- merge(meta_all, temp_ann, by = "site")
+temp_ann_meta$site <- as.character(temp_ann_meta$site)
+
+temp_boxplot <- ggplot(data = temp_ann_meta, aes(x = region, y = mean_temp, fill  = region)) +
+  geom_boxplot(show.legend = FALSE, width = 0.5) +
+  stat_compare_means(label.x = 2.5, label.y = 12.5, label = "p.signif") +
+  xlab("") +
+  ylab("") +
+  coord_flip() +
+  theme_bw() +
+  scale_fill_grey(start = 0.4, end = 0.8) +
+  theme(axis.text.y=element_text(vjust=0.3))
+
+ggsave("fig_3B", plot = temp_boxplot, dpi = 300, device = "jpeg", path = "figures/frontiers_fig_prep/")
+
+### Plots for wave variables
+#Annual wave plots
+#combine with meta data for easier plotting
[…]

### 2020-01-07 thesis_chapter_3.Rmd
added in new template for thesis. outputs updated for each chapter

@@
 author: "Ross Coppin"
-date: "16/10/2019"
 output:
+  word_document:
+    reference_docx: markdown/thesis_template.docx
   html_document:
     df_print: paged
+  pdf_document:
+    fig_caption: yes
+bibliography: references.bib
 ---

### 2020-01-08 thesis_chapter_2.Rmd
additional housecleaning

@@
 output:
+  html_document:
+    df_print: paged
   word_document:
     reference_docx: markdown/frontiers_template.docx
-  html_document:
-    df_print: paged
   pdf_document:

### 2020-01-22 thesis_chapter_2.Rmd
added to calculating drift section and updated all outputs

--- 2020-01-07 thesis_chapter_2.Rmd
@@
         axis.title = element_blank()) # Remove lat/ lon labels
-ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")
+#ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")
 
@@
                     xmin = 4, xmax = 10, ymin = 4.5, ymax = 10)
-ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8)
+#ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8)
 ```

### 2020-02-10 thesis_chapter_3.Rmd
Test commit from browser
added to methods section

@@
 ## Study species
-## Study site
-## Calculating coefficents
+## Study sites
+## Experimental work
+### Kelp bouyancy
+
+The aim of this experiment as to determine the epibiont load required to cause a floating kelp individual to become negatively buoyant, i.e to sink.
+Since epibiont load may be in the form of epiphytes and bryozoans, estimated percentage cover for each case was recorded for the whole sporophyte as well as locations along the plant.
+Before the experiments took place, a weighted net bag was suspended by a string from a spring balance so that the bag was submersed in water, and the weight recorded.
+A plant was then inserted into the bag, and the new weight was measured and recorded.
+The difference between weights recorded with and without the plant was considered the plants *net submerged buoyancy*.
+The plant was then placed in a tidal pool and allowed to float freely.
+The percentage of the whole plant, head (primary blade and fronds), stipe and holdfast which was submerged was recorded.
+In addition, photos using a setsquare for scale were also taken for each case.
+Finally, various morphological characteristics were measured (see table ??) based on methods by @rothman2017 and @coppin2020.
+For for each location along the sporophyte, epiphytes were carefully scrapped off into a bag and weighed.
+
+### Kelp drift characteristcs
+
+The aim of this experiment was to determine drift characteristics and drift rate of an *E.
+maxima* individual floating at the surface.
+In each case, a beach-cast or *in situ* individual was collected and the same morphological characteristics as per the buoyancy experiment were collected, however the various mass measurements were recorded after the experiment was complete as this requires the individual to be cut and divided into sections, thereby compromising buoyancy.
+The kelp individual was released from behind the breakwater and allowed to float freely.
+The percentage of the kelp plant submerged for each part of the plant was recorded, and a GPS coordinate recorded.
+The floating kelp was then followed by a small inflatable or kayak and a GPS location of both the kelp and observer was recorded.
+When recording the GPS location of the kelp from a boat, a diver swam to the location of the kelp and recorded its coordinates with a GPS.
+This was done to ensure the boat wake did not interfere with the natural drift characteristics of the kelp individual.
+
+Wind and wave data were collected by means of hind-cast numerical modelling to provide quantitative estimates.
+
+## Calculating coefficients
 ## Dispersal model
@@
 # Results
-## Coeffients
+## Coefficients
 ## Dispersal model

### 2020-02-18 thesis_chapter_3.Rmd
added to intro and materials and methods section

@@
 ---
-title: 'Hydrodynamic and particle dispersion modelling of floating solitary kelp:
+title: 'Hydrodynamic and particle dispersion modelling of floating solitary kelp
   An experimental approach'
@@
 
-Discuss previous rafting studies and the gaps within them
-Identify important coeffients
+The aim of this experiment was to determine the effects of epibiont growth and tissue fragmentation on the longevity of solitary rafting kelp.
+The growth rate of epiphytes has been shown to influence kelp raft longevity, with warmer temperatures favoring increased growth rate of epiphytes [@graiff2016; @macaya2016; @rothäusler2011; @tala2013; @hernández2006; @thiel2005; @hobday2000a] and thereby decreasing raft longevity.
+Therefore, increased sea surface temperatures could reduce the dispersal potential of kelp as the increase of epiphyte biomass reduces the buoyancy of kelp rafts, causing them to sink.
+In addition, tissue fragmentation leads to failure in the structural components of the plant, such as the stipe or pneumatocyst, which become filled with water and sink.
+Other factors such as herbivory and solar radiation have also been identified as important in determining the persistence of kelp rafts in the ocean.
+Finally, previous work has shown epiphytic species differ in their distributions along the kelp plant.
+For example, in South Africa Gelidium vittatum (Linnaeus) Kützing grows on the stipe or on the limpet Cymbula compressa (Linnaeus, 1758) that in turn inhabits kelp stipes, while Carpoblepharis flaccida (J.V.Lamouroux) Kützing grows on the fronds [@Anderson2006].
+This experiment will also help characterise the distribution of epiphytic algae species along E.
+maxima sporophytes.
 
@@
 ## Study sites
-## Experimental work
+## Experimental design
+
+Kelp sporophytes were collected from a kelp forest close to the study site.
+A total of 10 adult sporophytes were sampled, with the holdfast intact, with varying morphological characteristics.
+Before kelp are tethered, various morphological characteristics were measured and all fauna and flora removed if present.
+Individual sporophytes were tethered to separate floats approximately 1-2m apart in a row along the the breadth of the tidal-pool, with the holdfasts facing towards the shore.
+Surgical tubing was used to attach individual kelp to the float to reduce the risk of physical damage over the course of the experiment, and floats were anchored down within the tidal-pool using weights.
+Each kelp will be tagged with an ID relating to the experiment and a UWC ID tag which has contact details engraved on.
+After the first week, the kelp were checked daily to ensure the exact day of sinking was identified.
+Environmental data such as temperature and light was collected by means of HOBO loggers that were attached to a separate float in the study area.
+
+## Epibiont load
+
+The tethered kelp individuals were monitored for epibiont load in the form of epiphytes and bryozoans, which may colonise the kelp, every 7 days over the course of the experiment.
+Three secondary blades were photographed in situ using a floating board with a tape measure for reference.
+Photographs were cataloged and the bryozoan cover was expressed as a percentage total bryozoan cover.
+Total biomass of bryozoan cover was determined indirectly since bryozoans are inseparable from the plant, and so bryozoan biomass was calculated from the percentage cover estimate, using the following equation: $BrM = 4.5516 × BrC0.524$
+
+where $$BrM$$ is the bryozoan proportional mass (%) and $$BrC$$ is the bryozoan cover (%) (cite??).
+Epiphyte cover will also be estimated every 7 days using the same methods as per bryozoan cover, however epiphyte biomass on the day of sinking was directly measured by scraping off the biomass and weighing it using a spring balance.
+Epiphytes were identified until species and the part of the plant where it was removed from was also recorded.
+
+### Morphological characteristics
+
+Before the start of the experiment and day of sinking, multiple morphological characteristics will be measured (see table ??).
+Most of the morphological characteristics will be measured using a tape measure and calipers will be used to measure features of thickness.
+Multiple measurements will be taken for the stipe and pneumatocyst, to determine overall volume.
+To determine the amount of ‘hollowness’, each individual will be cut open and the thickness of the stipe and pneumatocyst will be measured using digital calipers.
+In each case, the thickest part of the plant was measured.
+The volume of hollowness will be subtracted from the overall  sporophyte volume to determine the actual volume of air present in the plant, which in turn will be used to determine buoyancy.
+Biomass for the whole kelp will be measured using a spring scale bothe before and after the experiment, however biomass of the head, stipe and holdfast will only measured after sinking, as this requires cutting the plant into the relevant sections.
+Epiphytes will be scraped off and weighed before the weight of the sporophyte is measured.
+In the case of bryozoans, the previously calculated indirect bryozoan biomass will be subtracted from the sporophyte measured to determine the actual sporophyte biomass.
+
 ### Kelp bouyancy

### 2020-02-18 thesis_chapter_3.Rmd
fixed equation syntax mistake

@@
 Photographs were cataloged and the bryozoan cover was expressed as a percentage total bryozoan cover.
-Total biomass of bryozoan cover was determined indirectly since bryozoans are inseparable from the plant, and so bryozoan biomass was calculated from the percentage cover estimate, using the following equation: $BrM = 4.5516 × BrC0.524$
+Total biomass of bryozoan cover was determined indirectly since bryozoans are inseparable from the plant, and so bryozoan biomass was calculated from the percentage cover estimate, using the following equation: $$BrM = 4.5516 × BrC0.524$$
 
-where $$BrM$$ is the bryozoan proportional mass (%) and $$BrC$$ is the bryozoan cover (%) (cite??).
+where $BrM$ is the bryozoan proportional mass (%) and $BrC$ is the bryozoan cover (%) (cite??).
 Epiphyte cover will also be estimated every 7 days using the same methods as per bryozoan cover, however epiphyte biomass on the day of sinking was directly measured by scraping off the biomass and weighing it using a spring balance.

### 2020-04-23 thesis_chapter_2.Rmd
changes to plots chapter 2

--- 2020-01-07 thesis_chapter_2.Rmd
@@
             size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = study_sites, aes(x = lon, y = lat), color = "red") +
-  geom_text(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = -0.1, nudge_y = -0.01) +
-  annotate("text", label = "WARM\nTEMPERATE", x = 19.0, y = -34.5, size = 4.0, angle = 0, colour = "black") +
-  annotate("text", label = "COOL\nTEMPERATE", x =17.8, y = -33.5, size = 4.0, angle = 0, colour = "black") +
-  annotate("text", label = "Cape\nPeninsula", x = 18.4, y = -34.5, size = 3.0, angle = 0, colour = "black") +
+  geom_point(data = study_sites, aes(x = lon, y = lat), color = "red", nudge_x = 0, nudge_y = 0) +
+  geom_label(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = -0.02, nudge_y = 0.1) +
+  annotate("text", label = "Warm\nTemperate", x = 19.0, y = -34.5, size = 4.0, angle = 0, colour = "black") +
+  annotate("text", label = "Cool\nTemperate", x =17.8, y = -33.5, size = 4.0, angle = 0, colour = "black") +
+  annotate("text", label = "FALSE\nBAY", x =18.65, y = -34.25, size = 4.0, angle = 0, colour = "black") +
   scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
@@
   geom_point(data = study_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
-  geom_text(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0.02, nudge_y = 0) +
+  geom_label(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
   coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -34.100)) +
@@
                     xmin = 4, xmax = 10, ymin = 4.5, ymax = 10)
-ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8)
+#ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8, dpi = 600)
 ```

### 2020-04-23 thesis_chapter_3.Rmd
added to intro for chapter 3

--- 2020-02-10 thesis_chapter_3.Rmd
@@
 ---
-title: 'Hydrodynamic and particle dispersion modelling of floating solitary kelp:
-  An experimental approach'
+title: 'Trajectory model of passive kelp dispersal'
 author: "Ross Coppin"
 output:
+  html_document:
+    df_print: paged
   word_document:
     reference_docx: markdown/thesis_template.docx
-  html_document:
-    df_print: paged
   pdf_document:
@@
 
-Importance of kelp rafting as a dispersal mechanism
+There is a range of objects, both natural and anthropogenic, floating in the ocean but macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
+Floating kelp as a passive dispersal mechanism for a range of marine taxa is well recognised and is sometimes referred to the tumble weed of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
+Although in recent years much of the focus of flotsam studies have shifted towards microplastics, considerable research has been performed investigating macroalgae as a passive dispersal mechanism in the ocean.
+Some macroalgae species are negatively buoyant and sink to the seafloor when detached from the substrate; usually through intense wave power.
+Other species of macroalgae have air-filled pneumatocysts or stipes, which allow the plants to reach the surface where light is more abundant.
+In turn, the positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum.
+The structure and number of pneumatocysts varies between species.
+For example species from the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005b; @graiff2016; @batista2018].
+In some cases the stipe itself is air-filled, such as with *E.
+maxima* and *E.
+radiata*.
+Although there have been reports of floating of Chlorophyta and some Rhodophyta species, Phaeophyceae species are the most commonly reported forms of floating algae.
+This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped inbetween or in the thalli [@dromgoole1982; @back2000].
+The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018] and the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018] have been the focus of much of the research regarding floating macroalgae and the ecological role of rafting.
+Comparatively, little research has been conducted on the dispersal characteristics of other Phaeophyceae species.
 
-Discuss previous rafting studies and the gaps within them
-Identify important coeffients
+The raft-time or float-time, ie.
+the amount of time macroalgae floats at the surface before becoming negatively buoyant and sinks vary across species.
+Raft-time also referred to as raft longevity, is determined by epibiont load, macroalgal growth rate and the loss of kelp biomass over time through fragmentation.
+Epibiont load consists of epiphytic and bryozoan species which tend to grow more rapidly in higher light and temperature environments, such as the ocean surface [@thiel2005b; @thiel2005a; @graiff2016].
+The increase in epibiont load reduces the buoyancy of floating macroalgal rafts over time eventually causing them to sink to the seafloor, while also potentially increasing drag [@hobday2000; @tala2013; @craw2018].
+Fragmentation of kelp biomass over time may also reduce raft times further, while growth may counteract the effects of epibiont load and fragmentation provided the environmental conditions are favorable [@graiff2016; @macaya2016].
+For instance a study by @rothausler2011 investigated the abiotic and biotic factors influencing raft time and dispersal potential of the giant kelp *M.
+pyrifera* by a combination of tethering experiments and field surveys.
+The results showed the physiological performance of kelp declined with increasing epiphyte biomass and that higher temperatures increased growth rate of epibionts and fragmentation of kelp.
+The authors concluded from the results of both the experiments and field surveys that *M.
+pyrifera* dispersal is dependent on low temperature and moderate irradiance conditions, with high temperatures and higher irradiance reducing overall raft-time and hence dispersal potential [@rothausler2011].
 
-> Bouyancy (Net)
-> Drag and resistance
-> Effect of wind on surface currents/waves 
-> Stokes drift
-> Wind and swell direction
-> Direct effect of wind on kelp
-> Dispersal scale
+Many past authors have also taken various approaches to estimate not only the trajectory of macroalgal rafts but also the oceanographic influencers of macroalgae trajectory.
+Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005a, @thiel2005b; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
+Although ocean currents are regarded as the primary influencer, the relative importance of wind versus surface current is still not known, although the role of wind has been recognised as important in several studies.
+For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M.
+pyrifera* in nearshore Monterrey Bay.
+The results showed that kelp rafts with little surface area exposed to the wind were largely driven by a combination of wind and wind waves, although the relative importance of wind and wind waves was not clear.
+In addition, the tagged kelp trajectories were more consistent with the formation of eddies during winter.
+Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems.
+For example the subAntartic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating kelp.
+Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013].
+For instance, a study by @Fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft.
+The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate, while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin.
+The authors suggested that wind and water-movement were the primary influences of trajectory, however, this was only inferred from the genetic results and local climatology data [@fraser2011].
 
-Highlight the gaps and explain how current study will address gaps
+Other aspects such as buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material.
+However, the past research on macroalgal trajectory has not investigated these factors which have been shown to be important aspects of trajectory for other materials such as icebergs, marine craft and microplastics.
+The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in water form drag.
+Furthermore, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as influencers of trajectory, as well as orientation of the object.
+If the length of the object is longer than the significant wave height then waves will be the primary influencer of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].
+There are considerable gaps in the literature with regards to macroalgae trajectory such as buoyancy, air form drag, water form drag, epibiont load, growth, fragmentation, size, shape and orientation which have not been investigated numerically in the past.
 
-Aims and objectives
+In order to determine the coefficients needed for the transport model, a series of experiments are proposed.
+These experiments will aim to determine aspects of buoyancy, epiphyte growth rates and transport rate, all of which have been regarded as important coefficients needed for an accurate model.
+The experiments will be perfomred numerically to determine the main influencers of passive solitary kelp dispersal.
 
 # Methods
-## Study species
-## Study sites
-## Experimental work
-### Kelp bouyancy
+## Model and parameterisation
 
-The aim of this experiment as to determine the epibiont load required to cause a floating kelp individual to become negatively buoyant, i.e to sink.
-Since epibiont load may be in the form of epiphytes and bryozoans, estimated percentage cover for each case was recorded for the whole sporophyte as well as locations along the plant.
-Before the experiments took place, a weighted net bag was suspended by a string from a spring balance so that the bag was submersed in water, and the weight recorded.
-A plant was then inserted into the bag, and the new weight was measured and recorded.
-The difference between weights recorded with and without the plant was considered the plants *net submerged buoyancy*.
-The plant was then placed in a tidal pool and allowed to float freely.
-The percentage of the whole plant, head (primary blade and fronds), stipe and holdfast which was submerged was recorded.
-In addition, photos using a setsquare for scale were also taken for each case.
-Finally, various morphological characteristics were measured (see table ??) based on methods by @rothman2017 and @coppin2020.
-For for each location along the sporophyte, epiphytes were carefully scrapped off into a bag and weighed.
+Numerical calculations of particle trajectory with a Lagrangian method [@delandmeter2019], and be described by the equation
 
-### Kelp drift characteristcs
+$$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
+where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model, $X_b(t)$ is a change in position due to 'behaviour' of the particle.
+This can range from swimming in fish to sinking or beaching.
 
-The aim of this experiment was to determine drift characteristics and drift rate of an *E.
-maxima* individual floating at the surface.
-In each case, a beach-cast or *in situ* individual was collected and the same morphological characteristics as per the buoyancy experiment were collected, however the various mass measurements were recorded after the experiment was complete as this requires the individual to be cut and divided into sections, thereby compromising buoyancy.
-The kelp individual was released from behind the breakwater and allowed to float freely.
-The percentage of the kelp plant submerged for each part of the plant was recorded, and a GPS coordinate recorded.
-The floating kelp was then followed by a small inflatable or kayak and a GPS location of both the kelp and observer was recorded.
-When recording the GPS location of the kelp from a boat, a diver swam to the location of the kelp and recorded its coordinates with a GPS.
-This was done to ensure the boat wake did not interfere with the natural drift characteristics of the kelp individual.
+The model is based on parameterisations that have been used for previous studies investigating iceberg, capsized marine vessels and microplastics.
+It should be noted that most of the previous studies regarding macroalgae trajectory have investigated rafts and not solitary floating individuals.
+A kelp raft will vary in size and can range from a few meters across to large dense mats the size of a sports field.
+Furthermore, some authors suggest that most macroalgae become entangled through ocean currents and not through the hydrodynamic forces that dislodge them.
+In addition, large kelp rafts are not a characteristics around the coast of the Cape Peninsula but rather solitary kelp are (Pers.
+Obs).
+Therefore, the trajectory of solitary *E.
+maxima* individuals will be considered.
+Previous studies have made the assumption that kelp rafts move with the same velocity as the ocean current, and instead model a ensemble of particles which trajectories are determined by a combination of ocean currents and/or the effects of winds on currents.
+In addition, these studies assume the particle is fully buoyant and does not vary with time.
 
-Wind and wave data were collected by means of hind-cast numerical modelling to provide quantitative estimates.
+The the velocity of the kelp sporophyte is a function of mass and time, as the mass of the individual will vary temporarily with changes in growth rate, fragmentation and epibiont load (mass), which in turn will effect buoyancy and thereby the surface area exposed to wind drag and seawater drag.
+A momentum equation can be used to describe how the mass of the kelp individual changes temporarily.
+A previous study by @Andersson2017 who used this approach to describe the change in velocity of an iceberg over time which has been adapted to suite the current project; and taking into account the additional parameterisitions:
+$$m\frac{du_i}{dt} = F_{corr} + F_{a} + F_{c} + F_{r} + F_{e}$$
+where $m$ is the overall kelp mass, $V_{i}$ is the initial kelp velocity, $d$ is the distance, $F_{corr}$ is Coriolis Force, $F_{a}$ is the air drag force, $F_c$ is the seawater drag force, $F_r$ is the effect of waves on the kelp, $F_{e}$ is the epibiont load force.
 
-## Calculating coefficients
-## Dispersal model
+The holdfast, stipe and kelp-head (fronds, primary blade and pneomatocyst) will the the focus of calculating the overall seawater drag force.
+Past studies refer to the surface area of the object exposed to wind forces as the "sail" area/volume, and the surface area exposed to the waves as the "keel" area/volume.
+The volume of the objects is required to calculate the seawater drag force and is expressed as,
+
+$$F_{c} = F_{kelphead} +  F_{holdfast} + F_{stipekeel}$$
+where $F_{c}$ is overall seawater drag force, $F_{kelphead}$ is the estimated drag coefficient for the top portion of the plant (fronds, primary blade, pnuematocyst), $V_{holdfast}$ is the estimated drag coefficient for the holdfast, $V_{stipekeel}$ is the estimated drag coefficient for the volume of the stipe submerged.
+The assumption made is that the kelp-head and holdfast will not change in volume while floating and that the entire holdfast will be submerged.
+Provided the exposed volume of the kelp exposed to the air is know the submerged portion $V_{stipekeel}$ and overall mass can be calculated based on Archimedes law [@Andersson2017]:
+$$V_{stipekeel} = V_{stipesail}\frac{p_{ice}}{p_w-p_{ice}}$$,
+
+$$m = (V_{stipesail}+V_{stipekeel})p_{ice} =(1 + \frac{p_w}{p_w -p{ice}})V_{stipesail}p_{ice}$$
+To calculate the volume of the stipe a idealised cyliderical form is used with height being subsituted for height such that:
+$$V = \pi r^2 l$$
 
 # Results
-## Coefficients
-## Dispersal model
 

### 2020-04-23 thesis_chapter_3.Rmd
alterations to chapter 3 and updated output and template

@@
 output:
+  word_document:
+    reference_docx: markdown/thesis_template.docx
   html_document:
     df_print: paged
-  word_document:
-    reference_docx: markdown/thesis_template.docx
   pdf_document:
@@
 The structure and number of pneumatocysts varies between species.
-For example species from the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005b; @graiff2016; @batista2018].
+For example species from the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018].
 In some cases the stipe itself is air-filled, such as with *E.
@@
 Raft-time also referred to as raft longevity, is determined by epibiont load, macroalgal growth rate and the loss of kelp biomass over time through fragmentation.
-Epibiont load consists of epiphytic and bryozoan species which tend to grow more rapidly in higher light and temperature environments, such as the ocean surface [@thiel2005b; @thiel2005a; @graiff2016].
+Epibiont load consists of epiphytic and bryozoan species which tend to grow more rapidly in higher light and temperature environments, such as the ocean surface [@thiel2005; @thiel2005; @graiff2016].
 The increase in epibiont load reduces the buoyancy of floating macroalgal rafts over time eventually causing them to sink to the seafloor, while also potentially increasing drag [@hobday2000; @tala2013; @craw2018].
@@
 Many past authors have also taken various approaches to estimate not only the trajectory of macroalgal rafts but also the oceanographic influencers of macroalgae trajectory.
-Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005a, @thiel2005b; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
+Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005, @thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
 Although ocean currents are regarded as the primary influencer, the relative importance of wind versus surface current is still not known, although the role of wind has been recognised as important in several studies.
@@
 Provided the exposed volume of the kelp exposed to the air is know the submerged portion $V_{stipekeel}$ and overall mass can be calculated based on Archimedes law [@Andersson2017]:
-$$V_{stipekeel} = V_{stipesail}\frac{p_{ice}}{p_w-p_{ice}}$$,
+$$V_{stipekeel} = V_{stipesail}\frac{p_{ice}}{p_w-p_{ice}}$$
 
+and so,
 $$m = (V_{stipesail}+V_{stipekeel})p_{ice} =(1 + \frac{p_w}{p_w -p{ice}})V_{stipesail}p_{ice}$$

### 2020-06-29 thesis_chapter_2.Rmd
updates to outputs and rmd files for chapters 1,2 and 3. Updated figures to follow and final chpater 2 to follow

--- 2020-01-22 thesis_chapter_2.Rmd
@@
             size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = study_sites, aes(x = lon, y = lat), color = "red") +
-  geom_text(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = -0.1, nudge_y = -0.01) +
-  annotate("text", label = "WARM\nTEMPERATE", x = 19.0, y = -34.5, size = 4.0, angle = 0, colour = "black") +
-  annotate("text", label = "COOL\nTEMPERATE", x =17.8, y = -33.5, size = 4.0, angle = 0, colour = "black") +
-  annotate("text", label = "Cape\nPeninsula", x = 18.4, y = -34.5, size = 3.0, angle = 0, colour = "black") +
+  geom_point(data = study_sites, aes(x = lon, y = lat), color = "red", nudge_x = 0, nudge_y = 0.1) +
+  geom_label(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
+  annotate("text", label = "Warm\nTemperate", x = 18.7, y = -34.5, size = 4.0, angle = 0, colour = "black") +
+  annotate("text", label = "Cool\nTemperate", x = 18.0, y = -34.0, size = 4.0, angle = 0, colour = "black") +
+  annotate("text", label = "FALSE\nBAY", x =18.65, y = -34.25, size = 4.0, angle = 0, colour = "black") +
   scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
@@
   ylab("Longitude")
-ggsave(filename = "map_A.jpg", plot = last_plot(), device = "jpeg", path = "map/")
+#ggsave(filename = "map_A.jpg", plot = last_plot(), device = "jpeg", path = "map/")
+
 ## Make 2nd map of just the peninsula
@@
   geom_point(data = study_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
-  geom_text(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0.02, nudge_y = 0) +
-  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -34.100)) +
+  geom_label(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
+  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -33.915)) +
   theme(axis.ticks = element_blank(), # Remove tick marks
@@
   annotation_custom(grob_B,
-                    xmin = 4, xmax = 10, ymin = 4.5, ymax = 10)
-#ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8)
+                    xmin = 6.2, xmax = 10, ymin = 3.8, ymax = 10)
+#ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8, dpi = 600)
 ```
@@
 ```{r Temperature variable processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
+
 ### Plots for temperature variables
+
+#temp_ann$site <- as.character(temp_ann$site)#make character to ensure no issues arise for site and region names
+#meta_all$site <- as.character(meta_all$site)#make character to ensure no issues arise for site and region names
+
+#temp_ann_meta_v2 <- left_join(temp_ann, meta_all, by = "site", copy = FALSE, keep = TRUE)
+
+region <- c("Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Western", "Eastern","Western", "Eastern", "Western", "Western", "Eastern", "Western", "Western", "Western", "Western", "Western")
+
+temp_ann$region <- region
+temp_ann$region <- as.factor(temp_ann$region)
+
+temp_ann_meta <- merge(meta_all, temp_ann)
+
 positions_temp <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock", "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
@@
 # Annual temp
-temp_plot <- ggplot(data = temp_ann, aes(x = site, y = mean_temp)) +
-  geom_point(show.legend = FALSE) + 
+temp_plot <- ggplot(data = temp_ann_meta, aes(x = site, y = mean_temp)) +
+  geom_point(show.legend = FALSE, aes(fill = region)) +
   scale_x_discrete(limits = positions_temp) +
   ylim(9, 21) +
-  geom_errorbar(aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) + 
+  geom_errorbar(data = temp_ann, aes(x = site, ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp)) +
   geom_point(data = temp_ann, aes(x = site, y = max_temp), show.legend = FALSE, shape = 17, size =2) +
   geom_point(data = temp_ann, aes(x = site, y = min_temp), show.legend = FALSE, shape = 6, size =2) +
+  geom_point(data = temp_ann, aes(x = site, y = min_temp), show.legend = FALSE, shape = 6, size =2) +
+  geom_point(data = temp_ann, aes(x = site, y = mean_temp, colour = region), show.legend = FALSE, size =2) +
   xlab("") +
@@
 
-temp_ann_meta <- merge(meta_all, temp_ann, by = "site")
-temp_ann_meta$site <- as.character(temp_ann_meta$site)
+#ggsave("fig_3A.jpeg", plot = temp_plot, dpi = 300, device = "jpeg" , path = "figures/")
 
@@
 
-```
-
-```{r Temperature variables plot, echo=FALSE, warning=FALSE, eval=TRUE,fig.height=5,fig.width=8, fig.cap='\\label{fig:Temperature variables plot}Temperature variables at the collection sites around the Cape Peninsula are represented in plot A.
-Temperature variables include minimum represented by inverted triangles, mean represented by black dots, maximum by black triangles, and whiskers standard deviation.
-Summary of temperature data for sites grouped by region are represented in plot B.'}
+#ggsave("fig_3B", plot = temp_boxplot, dpi = 300, device = "jpeg", path = "figures/frontiers_fig_prep/")
+
+```
+
+```{r Temperature variables plot, eval=TRUE, echo=FALSE, fig.cap='\\label{fig:Temperature variables plot}Annual temperature variables at the collections sites around the Cape Peninsula are represented by plot A, with sites in order from North to South.
+Temperature variables include minimum temperature represented by inverted triangles, mean represented by black dots, maximum by black triangles, and whiskers standard deviation.
+Summary of temperature data for sites grouped by region and represented in plot B.', fig.height=5, fig.width=8, warning=FALSE}
+
 ggarrange(temp_plot, temp_boxplot, common.legend = FALSE, ncol = 2, nrow = 1, labels = c("AUTO"), align = c("hv"), legend = FALSE)
+
+#ggsave("fig_3", plot = temp_boxplot, dpi = 300, device = "jpeg", path = "figures/")
+
 ```
@@
 #combine with meta data for easier plotting
+
+wave_ann$site <- as.character(wave_ann$site)#make character to ensure no issues arise for site and region names
 wave_ann_meta <- merge(meta_all, wave_ann, by = "site")
+
 wave_ann_meta$site <- as.character(wave_ann_meta$site)#make character to ensure no issues arise for site and region names
-
-positions_wave <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Millers Point", "Baboon Rock",  "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
+wave_ann_meta <- wave_ann_meta %>%
+  mutate(site = replace(site, site == "Miller's Point C", "Miller's Point"))
+
+#wave_ann_meta$site <- as.character(wave_ann_meta$site)#make character to ensure no issues arise for site and region names
+
+positions_wave <- c("Betty's Bay", "Roman Rock","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock",  "Batsata Rock","Bordjies Deep","Buffelsbaai","Buffelsbaai South","Olifantsbos","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","Sea Point","St Helena Bay")
 #cols <- c("Betty's Bay" = "red", "Roman Rock"= "red","Kalk Bay"= "red", "A-Frame"= "red", "Miller's Point"= "red", "Baboon #Rock"= "red",  "Batsata Rock"= "red","Bordjies Deep"= "red","Buffelsbaai"= "red","Buffelsbaai South"= "red","Olifantsbos"= #"blue","Scarborough"= "blue", "Soetwater"= "blue", "Kommetjie"= "blue","Hout Bay"= "blue","Oudekraal"= "blue","Sea Point"= #"blue","St Helena Bay"= "blue") #  scale_colour_manual(values = cols) #script to assign colours to specific sites
 
-wave_plot_1 <- ggplot(data = abio_clim, aes(x = site, y = hs_mean)) +
-  geom_point(show.legend = FALSE, shape = 17, size = 2) + 
+wave_plot_1 <- ggplot(data = wave_ann_meta, aes(x = site, y = hs_mean)) +
+  geom_point(show.legend = FALSE, shape = 17, size = 2, aes(colour = region)) + 
   scale_x_discrete(limits = positions_wave) + 
@@
 
-wave_plot_2 <- ggplot(data = abio_clim, aes(x = site, y = tp_mean)) +
-  geom_point(show.legend = FALSE, shape = 16, size = 2) + 
+wave_plot_2 <- ggplot(data = wave_ann_meta, aes(x = site, y = tp_mean)) +
+  geom_point(show.legend = FALSE, shape = 16, size = 2, aes(colour = region)) + 
   scale_x_discrete(limits = positions_wave) + 
@@
 ggarrange(wave_plot_1,wave_boxplot_1, wave_plot_2, wave_boxplot_2, common.legend = FALSE, labels = c("A", "B", "C", "D"), align = "hv")
+
+ggsave(plot = last_plot(), filename = "Fig_3", path = "figures/chapter_2/", device = "jpeg", units = "cm", width = 7, height = 7)
 ```
@@
   xlab("") + 
-  ylab("Lamina weight") +
+  ylab("Lamina weight (kg)") +
   scale_fill_discrete(name="Temperature regime") +
@@
   xlab("") + 
-  ylab("Lamina length") +
+  ylab("Lamina length (cm)") +
   scale_fill_discrete(name="Temperature regime") +
@@
   xlab("") + 
-  ylab("Lamina thickness") +
+  ylab("Lamina thickness (mm)") +
   scale_fill_discrete(name="Temperature regime") +
@@
   xlab("") + 
-  ylab("Stipe mass (g)") +
+  ylab("Stipe mass (kg)") +
   scale_fill_discrete(name="Temperature regime") +
@@
   xlab("") + 
-  ylab("Stipe diameter (cm)") +
+  ylab("Stipe diameter (mm)") +
   scale_fill_discrete(name="Temperature regime") +
@@
   xlab("") + 
-  ylab("No. of digits") +
+  ylab("No. of digits (count)") +
   scale_fill_discrete(name="Temperature regime") +
@@
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.85, label.y.npc = c('top')) + 
-  ylab("Lamina weight (g)") + 
+  ylab("Lamina weight (kg)") + 
   scale_fill_brewer(palette="OrRd") +
@@
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
-  ylab("Lamina thickness") + 
+  ylab("Lamina thickness (mm)") + 
   scale_fill_brewer(palette="OrRd") +
@@
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
-  ylab("Stipe mass (g)") + 
+  ylab("Stipe mass (kg)") + 
   scale_fill_brewer(palette="OrRd") +
@@
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
-  ylab("Stipe diameter (cm)") + 
+  ylab("Stipe diameter (mm)") + 
   scale_fill_brewer(palette="OrRd") +
@@
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
-  ylab("No. of digits") + 
+  ylab("No. of digits (count)") + 
   scale_fill_brewer(palette="OrRd") +
@@
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
-  ylab("Thallus mass (kg)") + 
+  ylab("Thallus mass (g)") + 
   scale_fill_brewer(palette="OrRd") +
@@
   stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
-  ylab("Frond mass (g)") +
+  ylab("Frond mass (kg)") +
   theme_bw() +
@@
   stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
-  ylab("Stipe mass (g)") +
+  ylab("Stipe mass (kg)") +
   theme_bw() +
@@
   stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
-  ylab("Stipe circ (cm)") +
+  ylab("Stipe circ (mm)") +
   theme_bw() +
@@
   stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
-  ylab("No. of tufts") +
+  ylab("No. of tufts (count)") +
   theme_bw() +
@@
   xlab("") + 
-  ylab("Frond mass (g)") +
+  ylab("Frond mass (kg)") +
   theme_bw() +
@@
   xlab("") + 
-  ylab("Stipe mass (g)") +
+  ylab("Stipe mass (kg)") +
   theme_bw() +
@@
   xlab("") + 
-  ylab("Stipe circ (cm)") +
+  ylab("Stipe circ (mm)") +
   theme_bw() +
@@
   xlab("") + 
-  ylab("No of Tufts") +
+  ylab("No of Tufts (count)") +
   theme_bw() +
@@
 Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
-ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = c("none"), ncol = 3, nrow = 4)
+ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = c("none"), ncol = 4, nrow = 3, vjust = 1.5)
 ```
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.', fig.height=8, fig.width=8}
-ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3, nrow = 4)
+ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 4, nrow = 3, vjust = 1.5)
 ```
@@
   stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
-  ylab("Frond mass (g)") +
+  ylab("Frond mass (kg)") +
   theme_bw() +
@@
   stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
-  ylab("Stipe mass (g)") +
+  ylab("Stipe mass (kg)") +
   theme_bw() +
@@
   stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
-  ylab("Stipe circ (cm)") +
+  ylab("Stipe circ (mm)") +
   theme_bw() +
@@
   stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 0.6, label.y.npc = c('top')) + 
-  ylab("No. of tufts") +
+  ylab("No. of tufts (count)") +
   theme_bw() +
@@
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
-  ylab("Frond mass (g)") +
+  ylab("Frond mass (kg)") +
   theme_bw() +
@@
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
-  ylab("Stipe mass (g)") +
+  ylab("Stipe mass (kg)") +
   theme_bw() +
@@
  ylim(0,30) +
-  ylab("Stipe circ (cm)") +
+  ylab("Stipe circ (mm)") +
   theme_bw() +
@@
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.y.npc = c('top')) + 
-  ylab("No of Tufts") +
+  ylab("No of Tufts (count)") +
   theme_bw() +
@@
 Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=7, fig.width=8}
-ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom", ncol = 3, nrow = 4)
+ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom", ncol = 4, nrow = 3, vjust = 1.5)
 ```
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.', fig.height=7, fig.width=8}
-ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3,nrow = 3)
+ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9,ESW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 4,nrow = 3)
 ```

### 2020-06-29 thesis_chapter_3.Rmd
updates to outputs and rmd files for chapters 1,2 and 3. Updated figures to follow and final chpater 2 to follow

@@
 
-The the velocity of the kelp sporophyte is a function of mass and time, as the mass of the individual will vary temporarily with changes in growth rate, fragmentation and epibiont load (mass), which in turn will effect buoyancy and thereby the surface area exposed to wind drag and seawater drag.
+The the velocity of the kelp sporophyte is a function of mass and time, as the mass of the individual will vary temporarily with changes in growth rate, erosion and epibiont load (mass), which in turn will effect buoyancy and thereby the surface area exposed to wind drag and seawater drag.
+
 A momentum equation can be used to describe how the mass of the kelp individual changes temporarily.
@@
 $$m\frac{du_i}{dt} = F_{corr} + F_{a} + F_{c} + F_{r} + F_{e}$$
-where $m$ is the overall kelp mass, $V_{i}$ is the initial kelp velocity, $d$ is the distance, $F_{corr}$ is Coriolis Force, $F_{a}$ is the air drag force, $F_c$ is the seawater drag force, $F_r$ is the effect of waves on the kelp, $F_{e}$ is the epibiont load force.
+where $m$ is the overall kelp mass, $u_{i}$ is the starting kelp velocity, $d$ is the distance, $F_{corr}$ is Coriolis Force, $F_{a}$ is the air drag force, $F_c$ is the seawater drag force, $F_r$ is the effect of waves on the kelp, $F_{e}$ is the epibiont load force.
 

### 2020-06-29 thesis_chapter_3.Rmd
Update thesis_chapter_3.Rmd

@@
 If the length of the object is longer than the significant wave height then waves will be the primary influencer of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].
+
+Most of the past studies have focused on aspects of raft longevity, with various raft times for a handful of species estimated.
+Although this is an area that requires more research, other factors that are important in the dispersal of macroalgae have been given comparatively little attention in recent years.
+An example is the relative importance of surface currents, winds, and waves in the dispersal of marine algae.
+In addition water- and -air form drag has not been considered before in relation to passive macroalgal dispersal.
+Drag coefficients related to wind and surface current have been shown to be important properties to consider when estimating trajectory and forecasting drift for search and sea rescue operations.
+Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae not only across species but also across environments.
+
+Langrangian trajectory modeling is a useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
+To accurately determine the trajectory of marine macroalgae these aspects need to be taken into account.
+Past research by @plourde1999 and @breivik2008 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
+However, this work does not consider biological material such as macroalgae.
+Although drag estimates do exist, these have been calculated for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
+
+Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulae to calculate drag based on shape.
+
 There are considerable gaps in the literature with regards to macroalgae trajectory such as buoyancy, air form drag, water form drag, epibiont load, growth, fragmentation, size, shape and orientation which have not been investigated numerically in the past.
-
 In order to determine the coefficients needed for the transport model, a series of experiments are proposed.

### 2020-07-07 thesis_chapter_3.Rmd
Added to introduction

@@
 Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulae to calculate drag based on shape.
+Chemical engineers in the field of fluid dynamics which come across similar problems when including drag use an approach known as a 'spherecity factor'.
+A spherecity factor is the ratio of the surface area of a sphere having the volume as the irregular shaped particle (s) to the actualo surface area of the particle (s_d) such that $$\psi=s/s_d$$.
+The values are dependent on the shape of the object and a range of values have been determined based on numerous experimental work.
+A table of various shapes and the associated spherecity factor is provided in the appendix.
 
@@
 # Discussion
+
+# Appendix
+
+Insert table of spherecity factors.

### 2020-07-07 notes.Rmd
adding notes.rmd

---
title: "Notes"
author: "Ross Coppin"
date: "20/03/2020"
output: word_document
---

###

Drift forecasting models have been developed for different applications such as predicting conditions in search areas for search and rescue operations and objects containg harardous material [@Breivik2011; @Hacket2006; @Eide2007; @Breivik2008; @Davidson2009]. These models make use of Monte Carlo (stochasic) techniques to numerically output a probability of trajectories from an ensemble of trajectory models. These models include near-surface currents and  wind field, as well as drift characteristics specific to the object. Near- or surface-currents can be estimated from tidal consistutents or from observations from high-resolution ocean models [@Breivik2011]. Surface drifters and high-frequency radar can also be used for currents bewteen 0.3 and 1.0 m depth, with the latter providing real-time measurements for short-term forecasts [@Breivik2011].

Wind has been shown to be an important influencer on passive drifting objects in the ocean and much research has gone into understanding how to properly predict the effect of wind on an objects drift trajectory. Small objects respond almost linearly to changes in the wind speed [@Breivik2008]. The rule of thumb in trajectory modelling is to assume a linear relation between 10m wind speed and the objects motion relative to the water. In general, an object's motion through the ambient water masses (referred to as its slip, windage or leeway) is roughly inversely proportional to the immersion ratio. The inverse relationship to windage and immersion ration has been shown previously by @Geyer1989 and @oDonnel1997, and has been used in modelling shipping container trajectories. However, objects with complex shapes can be influenced substantially by crosswind motion. In addition, the relation between the crosswind motion and immersion ratio is not well understood and can vary greatly between objects [@Breivik2011].

###

The position of a floating object is computed by numerically by intergrating the total drift velcoity

Inertia is defined here as an object's resistence to changes in velocity. Largrangian objects which are dispersed in the ocean by means of currents can have their trajectory altered by inertial forces, which are depedent on the physical properties of the object. For instance, differences in density can cause objects to become entrained or expelled from eddies [@Brooks; @Beron-Vera2015], and objects less dense than the surrounding water tend to deflect left of the mean flow [@Brooks]. Other factors such as leeway and drag are additional examples. Therefore, incorprating the aforementioned influencers into numerical investigations into *E. maxima* trajectory will shed light on the role these organisms play in marine connectvity as well as ecological function of nearhsore and offshore ecological communities. Macroalgae have been shown to be important vectors of dispersal in the marine environment as well as an important source of organic material for near and distant ecosystems.

Due to the potential changes in overall biomass that the kelp individal will undergo during its trajectory is the underlying physiological model that will be nested within the trajectory model.

The overall density of floating *E. maxima* is a function of mass which is determined by growth rate, erosion rate and epibiont growth. In turn growth rate, erosion rate and epibiont growth are determined by the environmental conditions the floating *E. maxima* individual is exposed to while adrift, namely temperature and light.

There is very little data available within the literature to determine continuous growth rates for both the kelp and epibionts. For this reason, estimates will be used based on the limited literature available. For growth rate, a constant of ??g/day will be used and a constant of ??/day will be used based of work by @

The epibiont component will determine overall float time. The growth of epibionts will increase the overall mass of the kelp individial over time causing it to be come negativley buoyant and sink. The epibiont growth will consist of threshold levels that will determine if growth rate increases or decreases above or below a constant value.

### 2020-07-13 thesis_chapter_3.Rmd
intro and output updated chapter 3

@@
 output:
+  pdf_document:
+    fig_caption: yes
+  html_document:
+    df_print: paged
   word_document:
     reference_docx: markdown/thesis_template.docx
-  html_document:
-    df_print: paged
-  pdf_document:
-    fig_caption: yes
 bibliography: references.bib
@@
 
-There is a range of objects, both natural and anthropogenic, floating in the ocean but macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
-Floating kelp as a passive dispersal mechanism for a range of marine taxa is well recognised and is sometimes referred to the tumble weed of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
-Although in recent years much of the focus of flotsam studies have shifted towards microplastics, considerable research has been performed investigating macroalgae as a passive dispersal mechanism in the ocean.
-Some macroalgae species are negatively buoyant and sink to the seafloor when detached from the substrate; usually through intense wave power.
-Other species of macroalgae have air-filled pneumatocysts or stipes, which allow the plants to reach the surface where light is more abundant.
+There is a range of objects, both natural and anthropogenic, floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
+Floating kelp  acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
+Considerable research has been performed investigating macroalgae as a passive dispersal mechanism in the ocean, however, in recent years much of the focus has shifted towards microplastics.
+
+Some macroalgae species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species of macroalgae have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.
 In turn, the positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum.
@@
 radiata*.
-Although there have been reports of floating of Chlorophyta and some Rhodophyta species, Phaeophyceae species are the most commonly reported forms of floating algae.
+Although there have been reports of floating Chlorophyta species and some Rhodophyta species, Phaeophyceae species are the most commonly reported forms of floating algae.
 This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped inbetween or in the thalli [@dromgoole1982; @back2000].
-The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018] and the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018] have been the focus of much of the research regarding floating macroalgae and the ecological role of rafting.
-Comparatively, little research has been conducted on the dispersal characteristics of other Phaeophyceae species.
+The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018] and the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
 
-The raft-time or float-time, ie.
-the amount of time macroalgae floats at the surface before becoming negatively buoyant and sinks vary across species.
-Raft-time also referred to as raft longevity, is determined by epibiont load, macroalgal growth rate and the loss of kelp biomass over time through fragmentation.
+The raft-time is the amount of time macroalgae floats at the surface before becoming negatively buoyant and begins to sink to the ocean-floor.
+Raft-time, also referred to as raft longevity, is determined by epibiont load, macroalgal growth rate and the loss of kelp biomass over time through fragmentation.
 Epibiont load consists of epiphytic and bryozoan species which tend to grow more rapidly in higher light and temperature environments, such as the ocean surface [@thiel2005; @thiel2005; @graiff2016].
@@
 
-Many past authors have also taken various approaches to estimate not only the trajectory of macroalgal rafts but also the oceanographic influencers of macroalgae trajectory.
+Many past authors have also taken various approaches to estimate not only the raft-time of macroalgal rafts but also the oceanographic influencers of macroalgae trajectory.
 Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005, @thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
-Although ocean currents are regarded as the primary influencer, the relative importance of wind versus surface current is still not known, although the role of wind has been recognised as important in several studies.
+Although ocean currents are regarded as the primary influencer, the relative importance of wind versus surface current is still not known; although the role of wind has been recognised as important in several studies.
 For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M.
 pyrifera* in nearshore Monterrey Bay.
-The results showed that kelp rafts with little surface area exposed to the wind were largely driven by a combination of wind and wind waves, although the relative importance of wind and wind waves was not clear.
+The results showed that kelp rafts with little surface area exposed to the wind were largely driven by a combination of wind and wind waves, however the relative importance of wind and wind waves was not clear.
 In addition, the tagged kelp trajectories were more consistent with the formation of eddies during winter.
@@
 Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013].
-For instance, a study by @Fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft.
-The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate, while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin.
+For example, a study by @Fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft.
+The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate; while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin.
 The authors suggested that wind and water-movement were the primary influences of trajectory, however, this was only inferred from the genetic results and local climatology data [@fraser2011].
@@
 The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in water form drag.
-Furthermore, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as influencers of trajectory, as well as orientation of the object.
+Drag coefficients related to wind and surface current have been shown to be important properties to consider when estimating trajectory and forecasting drift for search and sea rescue operations.
+Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.
+In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as influencers of trajectory, as well as orientation of the object.
 If the length of the object is longer than the significant wave height then waves will be the primary influencer of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].
 
-Most of the past studies have focused on aspects of raft longevity, with various raft times for a handful of species estimated.
-Although this is an area that requires more research, other factors that are important in the dispersal of macroalgae have been given comparatively little attention in recent years.
-An example is the relative importance of surface currents, winds, and waves in the dispersal of marine algae.
-In addition water- and -air form drag has not been considered before in relation to passive macroalgal dispersal.
-Drag coefficients related to wind and surface current have been shown to be important properties to consider when estimating trajectory and forecasting drift for search and sea rescue operations.
-Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae not only across species but also across environments.
+Although various approaches have been used in the past to investigate floating macroalagae trajectory, very few studies have employed the use of Langrangian trajectory modelling.
+Furthermore, non of the existing studies using this modeling apporach have considered macroalgal morphology (i.e.
+shape) as an influencer on drag and ultimately trajectory.
+Currently, work by @Brooks2019 which investigated the effect of inertia and radial size on the trajectory of pelagic *Sargussum* rafts.
+The study used a custome growth model to estimate changes in biomass and ultimately radial size, while a customised Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
+The results showed that trajectory of pelagic *Sargussum* was signficantly influenced by inertia and the radial size of the rafts.
+Langrangian trajectory modeling is a useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
+The radial size included by @Brooks2019 is an improvement, however this approach does not take into account the complex morphology on an individual level into account.
 
-Langrangian trajectory modeling is a useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
 To accurately determine the trajectory of marine macroalgae these aspects need to be taken into account.
@@
 Although drag estimates do exist, these have been calculated for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
-
 Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulae to calculate drag based on shape.
 Chemical engineers in the field of fluid dynamics which come across similar problems when including drag use an approach known as a 'spherecity factor'.
-A spherecity factor is the ratio of the surface area of a sphere having the volume as the irregular shaped particle (s) to the actualo surface area of the particle (s_d) such that $$\psi=s/s_d$$.
+A spherecity factor is the ratio of the surface area of a sphere having the volume as the irregular shaped particle (s) to the actual surface area of the particle (s_d) such that $$\psi=s/s_d$$.
 The values are dependent on the shape of the object and a range of values have been determined based on numerous experimental work.

### 2020-07-23 thesis_chapter_3.Rmd
updated references and added in raft-longevity

@@
 $$V = \pi r^2 l$$
+# Raft-longevity
+
+Raft-longevity refers to the time a macroalgae raft remians afloat at the surface of the ocean after detachment.
+Raft-longevity was estimated from past studies by @graiff2016 and @graiff2013; which found mean raft times of approximately 30 days in moderate conditions.
 

### 2020-08-05 thesis_chapter_3.Rmd
added to methods

--- 2020-07-13 thesis_chapter_3.Rmd
@@
 $$V = \pi r^2 l$$
+# Shape-based drag estimates
+
+The software used to simulate the drifting of *E.
+maxima* individuals assumes the objects are spherical in shape.
+In order to account for variation in drag based on kelp morphology, the drag for the seperate structural components of the plant are calculated and converted to a sphereical object quivalent before each simulation takes place.
+The calculated drag estimates were then incorportaed into each simulaton run and were assumed to remain constant over the duration of the estimated raft-time.
+
+# Raft-time
+
+Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remians afloat at the surface of the ocean after detachment.
+Raft-time was estimated from past studies by @graiff2016 and @graiff2013; which found mean raft times of approximately 30 days in moderate conditions.
 

### 2020-08-11 thesis_chapter_3.Rmd
updated methods section

--- 2020-07-13 thesis_chapter_3.Rmd
@@
 # Methods
-## Model and parameterisation
 
-Numerical calculations of particle trajectory with a Lagrangian method [@delandmeter2019], and be described by the equation
+## Hydrodynamic model
 
-$$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
-where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model, $X_b(t)$ is a change in position due to 'behaviour' of the particle.
-This can range from swimming in fish to sinking or beaching.
+## Particle tracking model
 
@@
 Previous studies have made the assumption that kelp rafts move with the same velocity as the ocean current, and instead model a ensemble of particles which trajectories are determined by a combination of ocean currents and/or the effects of winds on currents.
-In addition, these studies assume the particle is fully buoyant and does not vary with time.
+In addition, these studies assume the particle is fully buoyant.
 
-The the velocity of the kelp sporophyte is a function of mass and time, as the mass of the individual will vary temporarily with changes in growth rate, erosion and epibiont load (mass), which in turn will effect buoyancy and thereby the surface area exposed to wind drag and seawater drag.
+Numerical calculations of particle trajectory with a Lagrangian method [@delandmeter2019], and be described by the equation
 
-A momentum equation can be used to describe how the mass of the kelp individual changes temporarily.
-A previous study by @Andersson2017 who used this approach to describe the change in velocity of an iceberg over time which has been adapted to suite the current project; and taking into account the additional parameterisitions:
-$$m\frac{du_i}{dt} = F_{corr} + F_{a} + F_{c} + F_{r} + F_{e}$$
-where $m$ is the overall kelp mass, $u_{i}$ is the starting kelp velocity, $d$ is the distance, $F_{corr}$ is Coriolis Force, $F_{a}$ is the air drag force, $F_c$ is the seawater drag force, $F_r$ is the effect of waves on the kelp, $F_{e}$ is the epibiont load force.
+$$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
+where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model, $X_b(t)$ is a change in position due to 'behaviour' of the particle.
+This can range from swimming in fish to sinking or beaching.
+IN order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behaviour.
+This study simulated two scenarious; one where the plant is fully submerged and only exposed to seawater drag and the other where parts of the plants are partially exposed resulting in a combination of seawater and air form drag.
+Unlike previous studies, drag based on shape based coefficients was used in the calculations.
 
-The holdfast, stipe and kelp-head (fronds, primary blade and pneomatocyst) will the the focus of calculating the overall seawater drag force.
-Past studies refer to the surface area of the object exposed to wind forces as the "sail" area/volume, and the surface area exposed to the waves as the "keel" area/volume.
-The volume of the objects is required to calculate the seawater drag force and is expressed as,
+## Hydrodynamic drag
 
-$$F_{c} = F_{kelphead} +  F_{holdfast} + F_{stipekeel}$$
-where $F_{c}$ is overall seawater drag force, $F_{kelphead}$ is the estimated drag coefficient for the top portion of the plant (fronds, primary blade, pnuematocyst), $V_{holdfast}$ is the estimated drag coefficient for the holdfast, $V_{stipekeel}$ is the estimated drag coefficient for the volume of the stipe submerged.
-The assumption made is that the kelp-head and holdfast will not change in volume while floating and that the entire holdfast will be submerged.
-Provided the exposed volume of the kelp exposed to the air is know the submerged portion $V_{stipekeel}$ and overall mass can be calculated based on Archimedes law [@Andersson2017]:
-$$V_{stipekeel} = V_{stipesail}\frac{p_{ice}}{p_w-p_{ice}}$$
+## Air-form drag
 
-and so,
-$$m = (V_{stipesail}+V_{stipekeel})p_{ice} =(1 + \frac{p_w}{p_w -p{ice}})V_{stipesail}p_{ice}$$
-To calculate the volume of the stipe a idealised cyliderical form is used with height being subsituted for height such that:
-$$V = \pi r^2 l$$
+## Raft-time
+
+Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remians afloat at the surface of the ocean after detachment.
+Raft-time was estimated from past studies by @graiff2016 and @graiff2013; which found mean raft times of approximately 30 days in moderate conditions.
+
+## Simulations
 

### 2020-08-17 thesis_chapter_3.Rmd
update started

@@
 
-## Air-form drag
+The use of shape coefficients was used as a novel approach to simulating biological drifting debris.
+The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
+The cross-sectional area for each of the virtual kelp parts was determined and used in the calculation of hydrodynamic drag.
 

### 2020-08-17 thesis_chapter_3.Rmd
added in equations to methods section

@@
 
+The energy momemtum equation was used to calculate the drag force exterted on the virtual kelp particle, $$m({d_u/d_t})=F_d + F_w +....$$
+
+The velocity of the virtual kelp float is a combination of the current velocity and the reduction in velocity due to drag such that, $$V = V_c + V_k $$
 ## Hydrodynamic drag
@@
 The cross-sectional area for each of the virtual kelp parts was determined and used in the calculation of hydrodynamic drag.
+The hydrodynamic/form drag can be calculated by the equation: $$ F_d = 0.5 \rho A_c C_d U^2$$
 

### 2020-08-18 thesis_chapter_3.Rmd
updates

@@
 shape) as an influencer on drag and ultimately trajectory.
-Currently, work by @Brooks2019 which investigated the effect of inertia and radial size on the trajectory of pelagic *Sargussum* rafts.
+Currently, work by @brooks2019 which investigated the effect of inertia and radial size on the trajectory of pelagic *Sargussum* rafts.
 The study used a custome growth model to estimate changes in biomass and ultimately radial size, while a customised Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
@@
 Langrangian trajectory modeling is a useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
-The radial size included by @Brooks2019 is an improvement, however this approach does not take into account the complex morphology on an individual level into account.
+The radial size included by @brooks2019 is an improvement, however this approach does not take into account the complex morphology on an individual level into account.
 
 To accurately determine the trajectory of marine macroalgae these aspects need to be taken into account.
-Past research by @plourde1999 and @breivik2008 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
+Past research by @plourde1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
 However, this work does not consider biological material such as macroalgae.
@@
 The velocity of the virtual kelp float is a combination of the current velocity and the reduction in velocity due to drag such that, $$V = V_c + V_k $$
+
 ## Hydrodynamic drag

### 2020-08-18 thesis_chapter_3.Rmd
error in citations fixed. outputs updated.

@@
 output:
+  html_document:
+    df_print: paged
   pdf_document:
     fig_caption: yes
-  html_document:
-    df_print: paged
   word_document:
@@
 Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013].
-For example, a study by @Fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft.
+For example, a study by @fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft.
 The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate; while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin.
@@
 To accurately determine the trajectory of marine macroalgae these aspects need to be taken into account.
-Past research by @plourde1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
+Past research by @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
 However, this work does not consider biological material such as macroalgae.

### 2020-08-18 thesis_chapter_3.Rmd
altered intro, added to methods, html updated

@@
 ---
-title: 'Trajectory model of passive kelp dispersal'
+title: 'Numerical trajectory patterns of floating macroalgae'
 author: "Ross Coppin"
@@
 
-Many past authors have also taken various approaches to estimate not only the raft-time of macroalgal rafts but also the oceanographic influencers of macroalgae trajectory.
+Many past authors have also taken various approaches to estimate not only the raft-time of macroalgal rafts but also the oceanographic influences of macroalgae trajectory.
 Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005, @thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
-Although ocean currents are regarded as the primary influencer, the relative importance of wind versus surface current is still not known; although the role of wind has been recognised as important in several studies.
+Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still not known; although the role of wind has been recognised as important in several studies.
 For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M.
@@
 Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.
-In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as influencers of trajectory, as well as orientation of the object.
-If the length of the object is longer than the significant wave height then waves will be the primary influencer of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].
+In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object.
+If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].
 
 Although various approaches have been used in the past to investigate floating macroalagae trajectory, very few studies have employed the use of Langrangian trajectory modelling.
-Furthermore, non of the existing studies using this modeling apporach have considered macroalgal morphology (i.e.
-shape) as an influencer on drag and ultimately trajectory.
+Furthermore, non of the existing studies using this modeling approach have considered macroalgal morphology (i.e.
+shape) as an aspect of drag and ultimately trajectory.
 Currently, work by @brooks2019 which investigated the effect of inertia and radial size on the trajectory of pelagic *Sargussum* rafts.
-The study used a custome growth model to estimate changes in biomass and ultimately radial size, while a customised Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
-The results showed that trajectory of pelagic *Sargussum* was signficantly influenced by inertia and the radial size of the rafts.
+The study used a custom growth model to estimate changes in biomass and ultimately radial size, while a customised Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
+The results showed that trajectory of pelagic *Sargussum* was significantly influenced by inertia and the radial size of the rafts.
 Langrangian trajectory modeling is a useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
@@
 Although drag estimates do exist, these have been calculated for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
-Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulae to calculate drag based on shape.
+Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulas to calculate drag based on shape.
 Chemical engineers in the field of fluid dynamics which come across similar problems when including drag use an approach known as a 'spherecity factor'.
@@
 There are considerable gaps in the literature with regards to macroalgae trajectory such as buoyancy, air form drag, water form drag, epibiont load, growth, fragmentation, size, shape and orientation which have not been investigated numerically in the past.
-In order to determine the coefficients needed for the transport model, a series of experiments are proposed.
-These experiments will aim to determine aspects of buoyancy, epiphyte growth rates and transport rate, all of which have been regarded as important coefficients needed for an accurate model.
-The experiments will be perfomred numerically to determine the main influencers of passive solitary kelp dispersal.
+In this study, a hypothetical approach was used to shed light on the role of drag on macroalgal drift trajectory by means of Langrangian based trajectory simulations.
+These simulations aimed to determine the role that hydrodynamic and wind drag might play in drift trajectory of macroalgae.
+In terms of wind drag, the direct effect of wind on the kelp was estimated rather than its affect on surface-currents.
 
@@
 
+The characteristics of the underlying numerical flow model will be explained here...
 ## Particle tracking model
@@
 $$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
-where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model, $X_b(t)$ is a change in position due to 'behaviour' of the particle.
+where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model, $X_b(t)$ is a change in position due to 'behavior' of the particle.
 This can range from swimming in fish to sinking or beaching.
-IN order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behaviour.
-This study simulated two scenarious; one where the plant is fully submerged and only exposed to seawater drag and the other where parts of the plants are partially exposed resulting in a combination of seawater and air form drag.
-Unlike previous studies, drag based on shape based coefficients was used in the calculations.
-
-The energy momemtum equation was used to calculate the drag force exterted on the virtual kelp particle, $$m({d_u/d_t})=F_d + F_w +....$$
-
-The velocity of the virtual kelp float is a combination of the current velocity and the reduction in velocity due to drag such that, $$V = V_c + V_k $$
+IN order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.
+This study simulated two scenarios; one where the plant is fully submerged and only exposed to seawater drag and the other where parts of the plants are partially exposed resulting in a combination of seawater and air form drag.
+Unlike previous studies, drag based on shape based coefficients was used in the calculations for determining the reduction in velocity as a result of drag.
+A momentum energy equation was used to calculate the drag forces for the relevant simulation.
+This approach has been employed in modeling iceberg drift trajectory [@andersson2017; @lichey2001; @eik2009] and was adapted to suit this particular study.
+The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was, $$m({d_u/d_t})=F_c + F_a + F_f$$
+where $m$ is kelp mass, $d_u$ the virtual kelp particle velocity, $F_c$ the hydrodynamic drag, $F_a$ is the wind drag, and $F_f$ the surface-current flow.
+The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations.
+Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculated before input into the model.
 
@@
 
-Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remians afloat at the surface of the ocean after detachment.
+Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remains afloat at the surface of the ocean after detachment.
 Raft-time was estimated from past studies by @graiff2016 and @graiff2013; which found mean raft times of approximately 30 days in moderate conditions.

### 2020-08-18 thesis_chapter_3.Rmd
csl file used for citations. outputs updated

@@
 output:
+  word_document:
+    reference_docx: markdown/thesis_template.docx
   html_document:
@@
     fig_caption: yes
-  word_document:
-    reference_docx: markdown/thesis_template.docx
 bibliography: references.bib
+csl: markdown/frontiers.csl.txt
 ---
@@
 Insert table of spherecity factors.
+
+# References

### 2020-08-18 thesis_chapter_3.Rmd
slight adjustments made. html updated

--- 2020-08-18 thesis_chapter_3.Rmd
@@
   html_document:
-    df_print: paged
+  word_document:
+    reference_docx: markdown/thesis_template.docx
   pdf_document:
     fig_caption: yes
-  word_document:
-    reference_docx: markdown/thesis_template.docx
 bibliography: references.bib
+csl: markdown/frontiers.csl.txt
 ---

### 2020-09-19 thesis_chapter_3.Rmd
updated chapter 3 and chpater 1. general comits and temporary track data added

@@
 
+## Study site
+
+```{r map}
+
+## Load software:
+library(ggplot2)
+library(rgdal)
+library(tidyverse)
+library(viridis)
+library(gridExtra)
+source("map/scaleBarFunc.R")
+
+load("map/south_africa_coast_hi_res.Rdata")
+
+map_pen <- ggplot(data = south_africa_coast_hi_res) +
+geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+            size = 1.0, colour = "black", fill = NA) +
+  #geom_point(data = study_sites, aes(x = lon, y = lat), color = "red", nudge_x = 0, nudge_y = 0.1) +
+  #geom_label(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
+  geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.10, xmax=18.50, ymin=-33.70, ymax=-34.50), inherit.aes = FALSE, fill = "blue", alpha = 0.01) +
+  geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.40, xmax=18.90, ymin=-34.00, ymax=-34.50), inherit.aes = FALSE, fill = "red", alpha = 0.01) +
+  #annotate("text", label = "Warm\nTemperate", x = 18.7, y = -34.5, size = 4.0, angle = 0, colour = "black") +
+  #annotate("text", label = "Cool\nTemperate", x = 18.0, y = -34.0, size = 4.0, angle = 0, colour = "black") +
+  annotate("text", label = "FALSE\nBAY", x =18.65, y = -34.25, size = 4.0, angle = 0, colour = "black") +
+  scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
+  arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
+  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
+  #theme(axis.ticks = element_blank()) + # Remove tick marks
+  xlab("Latitude") +
+  ylab("Longitude")
+
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+release_sites <- read.csv("metadata/release_sites.csv", sep = ";")
+
+## Make 2nd map of just the peninsula
+release_map <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = release_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
+  geom_label(data = release_sites, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
+  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -33.600)) +
+  theme(axis.ticks = element_blank(), # Remove tick marks
+        #axis.text = element_blank(), # Remove lat/ lon numbers
+        axis.title = element_blank()) # Remove lat/ lon labels
+#ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")
+
+```
+
+```{r}
+
+library(ggpubr)
+
+ggarrange(map_pen, release_map, labels = c("AUTO"), align = c("hv"), ncol = 2, nrow = 1)
+
+```
+
 ## Hydrodynamic model
@@
 The characteristics of the underlying numerical flow model will be explained here...
+
 ## Particle tracking model
@@
 IN order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.
+
 This study simulated two scenarios; one where the plant is fully submerged and only exposed to seawater drag and the other where parts of the plants are partially exposed resulting in a combination of seawater and air form drag.
@@
 
+## Windage
+
+In order to calculate the direct effect of wind-drag on the kelp the wind velcocity is needed, which can be estimated by the surface-current velocity.
+
+The total wind drag effect was calculated for the hypothetical cross-sectional area of the kelp particle which was exposed above the water surface, which in this study, was estimated as 10% of the total cross-sectional area.
+
 ## Raft-time
@@
 
+```{r}
+
+# Read in data and data wrangling
+
+library(tidync)
+library(tidyverse)
+
+p_particle <- tidync("data/trajectory_data/PassiveFloat.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+
+p_particle_mean <- p_particle %>% 
+   group_by(t) %>% 
+   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+
+p_particle$type <- "passive" # add extra column 'type' to dataframe and assigned 'passive' label
+
+k_particle <- tidync("data/trajectory_data/KelpFloat.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+
+k_particle_mean <- k_particle %>% 
+   group_by(t) %>% 
+   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+
+k_particle$type <- "kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+total_data <- rbind(p_particle, k_particle)
+
+```
+
+## Simulations
+
+```{r}
+
+plot_A <- ggplot(data = p_particle, aes(x = lon, y = lat)) + 
+  geom_point() + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  theme_bw()
+
+plot_B <- ggplot(data = k_particle, aes(x = lon, y = lat)) + 
+  geom_point() + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  theme_bw()
+
+```
+
+```{r}
+
+library(ggpubr)
+ggarrange(plot_A,plot_B, labels = c("A","B"))
+
+```
+
 # Discussion

### 2020-09-19 map/theme.R

@@ -1,0 +1,28 @@
+require(ggplot2)
+require(grid)
+theme_set(theme_bw())
+bw_update <- theme_bw() +
+theme(plot.background = element_blank(),
+      panel.background = element_rect(colour = NA, fill = NA),
+      panel.border = element_rect(colour = "black",
+                                  fill = NA, size = 0.8),
+      panel.grid.minor = element_line(colour = NA),
+      panel.grid.major = element_line(colour = "black",
+                                      size = 0.2,
+                                      linetype = "dotted"),
+      axis.title = element_text(size = 8),
+      axis.text = element_text(size = 8),
+      axis.ticks = element_line(size = 0.5),
+      axis.ticks.length = unit(2, "mm"),
+      #legend.position = "top",
+      #legend.direction = "horizontal",
+      legend.title = element_text(size = 8),
+      legend.text = element_text(size = 8),
+      legend.key = element_blank(),
+      legend.key.height = unit(0.44, "cm"),
+      #legend.key.width = unit(0.30, "cm"),
+      legend.background = element_blank(),
+      plot.title = element_text(size = 8, hjust = 0),
+      strip.background = element_blank(),
+      strip.text = element_text(size = 8))
+theme_set(bw_update)

### 2020-09-21 thesis_chapter_3.Rmd
chapter 3, map and and general additions/modifications

@@
 A spherecity factor is the ratio of the surface area of a sphere having the volume as the irregular shaped particle (s) to the actual surface area of the particle (s_d) such that $$\psi=s/s_d$$.
-The values are dependent on the shape of the object and a range of values have been determined based on numerous experimental work.
-A table of various shapes and the associated spherecity factor is provided in the appendix.
+The values are dependent on the shape of the object and a range of values have been
+determined based on numerous experimental work. A table of various shapes and the associated spherecity factor is provided in the appendix.
 
@@
 
-```{r map}
+```{r map, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-```{r}
+```{r message=FALSE, warning=FALSE}
 
@@
 
+trac_map_passive <- ggplot(data = south_africa_coast_hi_res) +
+geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+            size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = p_particle, aes(x = lon, y = lat), color = "red") +
+  #geom_label(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
+  geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.10, xmax=18.50, ymin=-33.70, ymax=-34.50), inherit.aes = FALSE, fill = "blue", alpha = 0.01) +
+  geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.40, xmax=18.90, ymin=-34.00, ymax=-34.50), inherit.aes = FALSE, fill = "red", alpha = 0.01) +
+  #annotate("text", label = "Warm\nTemperate", x = 18.7, y = -34.5, size = 4.0, angle = 0, colour = "black") +
+  #annotate("text", label = "Cool\nTemperate", x = 18.0, y = -34.0, size = 4.0, angle = 0, colour = "black") +
+  annotate("text", label = "FALSE\nBAY", x =18.65, y = -34.25, size = 4.0, angle = 0, colour = "black") +
+  scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
+  arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
+  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
+  #theme(axis.ticks = element_blank()) + # Remove tick marks
+  xlab("Latitude") +
+  ylab("Longitude")
+
 ```

### 2020-09-23 thesis_chapter_3.Rmd
added morphology graphs and summary stats for area estimation

--- 2020-09-19 thesis_chapter_3.Rmd
@@
 
-```{r}
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
+## Cross-sectional area
+
+In order to incorporate hydrodynamic and wind drag, the cross-sectional area of the kelp was calculated first.
+Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant, for details please see image ???.
+The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin] and estimated from figure \ref{fig:morphology}.
+
+```{r morphology_wrangling, message=FALSE, warning=FALSE, include=FALSE}
+# Investigate morphology data to assist in estimating surface area
+# Load software
+library(tidyverse)
+library(ggpubr)
+library(gridExtra)
+library(ggplot2)
+library(ggord)
+library(vegan)
+# Read in morphology data
+eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
+  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
+  filter(site!= "De Hoop")
+eck_data$site <- as.character(eck_data$site)
+eck_deep <- eck_data %>% 
+  filter(population == "Deep")
+
+eck_long <- gather(eck_deep, morphology, measurement, frond_mass:total_length, factor_key=TRUE)
+# Summary stats to make it easier to refer to later, these will not be used any analysis.
+eck_stats <- eck_long %>% 
+  group_by(site, morphology) %>% 
+  summarise(mean = mean(measurement, na.rm = T),
+            min = min(measurement, na.rm = T),
+            max = max(measurement, na.rm = T),
+            range = range(measurement, na.rm = T)[2]-range(measurement, na.rm = T)[1],
+            sd = sd(measurement, na.rm = T)) %>% 
+  ungroup()
+# Make custom positions
+positionsEck <- c("Betty's Bay", "Kalk Bay", "A-Frame", 
+                  "Miller's Point", "Baboon Rock", "Batsata Rock", 
+                  "Buffelsbaai", "Scarborough", 
+                  "Soetwater","Kommetjie","Hout Bay", 
+                  "Oudekraal", "St Helena Bay")
+# Make plots for each morphology feature and grouped by site.
+frond_mass <- ggplot(data = eck_data, aes(x = site, y = frond_mass, fill = site)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Frond mass (g)") + 
+           coord_flip() +
+           xlab("") + 
+           theme_bw()
+primary_length <- ggplot(data = eck_data, aes(x = site, y = primary_length , fill = site)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Primary Length  (cm)") + 
+           coord_flip() +
+           xlab("") + 
+           theme_bw()
+primary_width <- ggplot(data = eck_data, aes(x = site, y = primary_width , fill = site)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Primary Width (cm)") + 
+           coord_flip() +
+           xlab("") + 
+           theme_bw() 
+frond_length <- ggplot(data = eck_data, aes(x = site, y = frond_length , fill = site)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Frond length (cm)") + 
+           coord_flip() +
+           xlab("") + 
+           theme_bw() 
+stipe_mass <- ggplot(data = eck_data, aes(x = site, y = stipe_mass , fill = site)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Stipe mass (g)") + 
+           coord_flip() +
+           xlab("") + 
+           theme_bw() 
+stipe_length <- ggplot(data = eck_data, aes(x = site, y = stipe_length , fill = site)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Stipe length (cm)") + 
+           coord_flip() +
+           xlab("") + 
+           theme_bw() 
+stipe_circ <- ggplot(data = eck_data, aes(x = site, y = stipe_circ , fill = site)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Stipe circ (cm)") + 
+           coord_flip() +
+           xlab("") + 
+           theme_bw() 
+tufts <- ggplot(data = eck_data, aes(x = site, y = tufts , fill = site)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("No. of tufts") + 
+           coord_flip() +
+           xlab("") + 
+           theme_bw() 
+epi_length <- ggplot(data = eck_data, aes(x = site, y = epi_length , fill = site)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Epi Length  (cm)") + 
+           coord_flip() +
+           xlab("") + 
+           theme_bw()
+
+total_length <- ggplot(data = eck_data, aes(x = site, y = epi_length, fill = site)) +
+           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
+           stat_compare_means(method = "kruskal.test", label.x = 5) + 
+           scale_x_discrete(limits = positionsEck) + 
+           ylab("Total length  (cm)") + 
+           coord_flip() +
+           xlab("") + 
+           theme_bw()
+```
+
+```{r morphology, echo=FALSE, , fig.height=7, fig.width=8, message=FALSE, warning=FALSE, fig.cap='\\label{fig:morphology}Boxplots representing *E.
+pallida* morphological characteristics grouped by site around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured and with units provided.
+The lower and upper "hinges" correspond to the first and third quartiles.
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.'}
+# Plot the morphology data based on site (see previous section)
+ggarrange(frond_length, frond_mass,primary_length, primary_width, stipe_circ, stipe_length, stipe_mass, total_length, tufts, common.legend = TRUE, legend = c("none"), align = c("hv"))
+```
+
 ## Hydrodynamic drag
@@
 The cross-sectional area for each of the virtual kelp parts was determined and used in the calculation of hydrodynamic drag.
-The hydrodynamic/form drag can be calculated by the equation: $$ F_d = 0.5 \rho A_c C_d U^2$$
+The hydrodynamic/form drag can be calculated by: $$ F_d = 0.5 \rho A_c C_d U^2$$
 
@@
 
-p_particle <- tidync("data/trajectory_data/PassiveFloat.nc") %>%
+p_particle <- tidync("OceanParcels/trajectory_data/PassiveFloat.nc") %>%
   hyper_tibble() %>%
@@
 
-k_particle <- tidync("data/trajectory_data/KelpFloat.nc") %>%
+k_particle <- tidync("OceanParcels/trajectory_data/KelpFloat.nc") %>%
   hyper_tibble() %>%

### 2020-09-28 thesis_chapter_3.Rmd
added script for 2dhistogram and heatmaps

@@
 output:
-  html_document:
   word_document:
     reference_docx: markdown/thesis_template.docx
+  html_document:
   pdf_document:
@@
 
+## Wind model
+
+The characteristics of the underlying numerical wind model will be explained here...
+
 ## Particle tracking model
@@
 The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin] and estimated from figure \ref{fig:morphology}.
+
+## Hydrodynamic drag
+
+The use of shape coefficients was used as a novel approach to simulating biological drifting debris.
+The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
+The cross-sectional area for each of the virtual kelp parts was determined and used in the calculation of hydrodynamic drag.
+The hydrodynamic/form drag can be calculated by: $$ F_d = 0.5 \rho A_c C_d U^2$$
+
+## Wind drag
+
+The total wind drag effect was calculated for the hypothetical cross-sectional area of the kelp particle which was exposed above the water surface, which in this study, was estimated as 10% of the total cross-sectional area.
+
+## Raft-time
+
+Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remains afloat at the surface of the ocean after detachment.
+Raft-time was estimated from past studies by @graiff2016 and @graiff2013; which found mean raft times of approximately 30 days in moderate conditions.
+
+## Simulations
+
+# Results
+
+```{r data, message=FALSE, warning=FALSE, include=FALSE}
+
+# Read in data and data wrangling
+
+library(tidync)
+library(tidyverse)
+
+p_RK4 <- tidync("OceanParcels/trajectory_data/PassiveFloat.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+
+p_RK4_mean <- p_RK4 %>% # summarise data based on time 
+   group_by(t) %>% 
+   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+
+p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'passive' label
+
+k_app1 <- tidync("OceanParcels/trajectory_data/KelpFloatApp1.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+
+k_app1_mean <- k_app1 %>% # summarise data based on time 
+   group_by(t) %>% 
+   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+
+k_app1$type <- "Approach 1" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+k_app2 <- tidync("OceanParcels/trajectory_data/KelpFloatApp2.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+
+k_app1_mean <- k_app2 %>% # summarise data based on time 
+   group_by(t) %>% 
+   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+
+k_app2$type <- "Approach 2" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+total_data <- rbind(p_RK4, k_app1, k_app2)
+
+```
+
+```{r plots, message=FALSE, warning=FALSE, include=FALSE}
+
+## 2D histogram plots of number of particles
+
+bin_RK4 <- ggplot(p_RK4, aes(x=lon, y=lat)) +
+  geom_bin2d(bins = 20) +
+  scale_fill_continuous(type = "viridis") +
+  coord_fixed() +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+bin_APP1 <- ggplot(k_app1, aes(x=lon, y=lat)) +
+  geom_bin2d(bins = 20) +
+  scale_fill_continuous(type = "viridis") +
+  coord_fixed() +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+bin_APP2 <- ggplot(k_app2, aes(x=lon, y=lat)) +
+  geom_bin2d(bins = 20) +
+  scale_fill_continuous(type = "viridis") +
+  coord_fixed() +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+## Heatmaps of study domain, use previous data for study site map and histograms
+
+heatmap_RK4 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_APP1 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_APP2 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+## Scatterplots
+
+total_traj <- ggplot(data = total_data, aes(x = lon, y = lat, colour = type)) + 
+  geom_point() + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  theme_bw() +
+  guides(colour=guide_legend(title="Particle type"))
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+library(ggpubr)
+plot(total_traj)
+ggarrange(heatmap_RK4,heatmap_APP1, heatmap_APP2, labels = c("A","B","C"), common.legend = TRUE, legend = c("right"))
+
+```
+
+# Discussion
+
+# Appendix
+
+Insert table of spherecity factors.
 
@@
 
-## Hydrodynamic drag
-
-The use of shape coefficients was used as a novel approach to simulating biological drifting debris.
-The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
-The cross-sectional area for each of the virtual kelp parts was determined and used in the calculation of hydrodynamic drag.
-The hydrodynamic/form drag can be calculated by: $$ F_d = 0.5 \rho A_c C_d U^2$$
-
-## Windage
-
-In order to calculate the direct effect of wind-drag on the kelp the wind velcocity is needed, which can be estimated by the surface-current velocity.
-
-The total wind drag effect was calculated for the hypothetical cross-sectional area of the kelp particle which was exposed above the water surface, which in this study, was estimated as 10% of the total cross-sectional area.
-
-## Raft-time
-
-Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remains afloat at the surface of the ocean after detachment.
-Raft-time was estimated from past studies by @graiff2016 and @graiff2013; which found mean raft times of approximately 30 days in moderate conditions.
-
-## Simulations
-
-# Results
-
-```{r}
-
-# Read in data and data wrangling
-
-library(tidync)
-library(tidyverse)
-
-p_particle <- tidync("OceanParcels/trajectory_data/PassiveFloat.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
-
-p_particle_mean <- p_particle %>% 
-   group_by(t) %>% 
-   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
-
-p_particle$type <- "passive" # add extra column 'type' to dataframe and assigned 'passive' label
-
-k_particle <- tidync("OceanParcels/trajectory_data/KelpFloat.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
-
-k_particle_mean <- k_particle %>% 
-   group_by(t) %>% 
-   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
-
-k_particle$type <- "kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-total_data <- rbind(p_particle, k_particle)
-
-```
-
-## Simulations
-
-```{r}
-
-plot_A <- ggplot(data = p_particle, aes(x = lon, y = lat)) + 
-  geom_point() + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  theme_bw()
-
-plot_B <- ggplot(data = k_particle, aes(x = lon, y = lat)) + 
-  geom_point() + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  theme_bw()
-
-```
-
-```{r}
-
-library(ggpubr)
-ggarrange(plot_A,plot_B, labels = c("A","B"))
-
-```
-
-# Discussion
-
-# Appendix
-
-Insert table of spherecity factors.
-
 # References

### 2020-09-30 thesis_chapter_3.Rmd
script for extra datasets added

@@
 
-p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'passive' label
+p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
+
+p_EE <- tidync("OceanParcels/trajectory_data/PassiveFloat.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+
+p_EE_mean <- p_EE %>% # summarise data based on time 
+   group_by(t) %>% 
+   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+
+p_EE$type <- "EE" # add extra column 'type' to dataframe and assigned 'EE' label
 

### 2020-10-02 thesis_chapter_3.Rmd
updated html. added to figures

@@
 output:
+  html_document:
   word_document:
     reference_docx: markdown/thesis_template.docx
-  html_document:
   pdf_document:
@@
 
-```{r data, message=FALSE, warning=FALSE, include=FALSE}
+```{r data wrangling, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-p_RK4 <- tidync("OceanParcels/trajectory_data/PassiveFloat.nc") %>%
+p_RK4 <- tidync("OceanParcels/trajectory_data/PassiveFloatRK4.nc") %>%
   hyper_tibble() %>%
@@
 
-p_EE <- tidync("OceanParcels/trajectory_data/PassiveFloat.nc") %>%
+p_EE <- tidync("OceanParcels/trajectory_data/PassiveFloatEE.nc") %>%
   hyper_tibble() %>%
@@
 
-k_app1_mean <- k_app2 %>% # summarise data based on time 
+k_app2_mean <- k_app2 %>% # summarise data based on time 
    group_by(t) %>% 
@@
 
-total_data <- rbind(p_RK4, k_app1, k_app2)
+total_data <- rbind(p_RK4, p_EE, k_app1, k_app2)
+
+```
+
+```{r scatter advection, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
+
+total_traj <- ggplot(data = total_data, aes(x = lon, y = lat, colour = type)) + 
+  geom_point(aes(alpha = 0.2)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  theme_bw() +
+  guides(colour=guide_legend(title="Particle type"))
+total_traj
 
@@
 bin_RK4 <- ggplot(p_RK4, aes(x=lon, y=lat)) +
+  geom_bin2d(bins = 20) +
+  scale_fill_continuous(type = "viridis") +
+  coord_fixed() +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+bin_RK4 <- ggplot(p_EE, aes(x=lon, y=lat)) +
   geom_bin2d(bins = 20) +
@@
 
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
 heatmap_RK4 <- ggplot() +
@@
                size = 1.0, colour = "black", fill = NA) +
-  geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_EE <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
@@
   theme_bw() 
-
-## Scatterplots
-
-total_traj <- ggplot(data = total_data, aes(x = lon, y = lat, colour = type)) + 
-  geom_point() + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  theme_bw() +
-  guides(colour=guide_legend(title="Particle type"))
-
 ```
@@
 library(ggpubr)
-plot(total_traj)
-ggarrange(heatmap_RK4,heatmap_APP1, heatmap_APP2, labels = c("A","B","C"), common.legend = TRUE, legend = c("right"))
+ggarrange(heatmap_RK4, heatmap_EE, heatmap_APP1, heatmap_APP2, labels = c("A","B","C","D"), common.legend = TRUE, legend = c("right"))
 

### 2020-10-09 thesis_chapter_3.Rmd
added hexiplot feature to density plots. scatter age plots added. added to site description

@@
 
-```{r map, message=FALSE, warning=FALSE, include=FALSE}
+The region of interest for this study was the Cape Peninsula in the Western Cape, South Africa.
+This was further sub-divided into two domains, the Atlantic domain and False Bay domain (see figure \ref{fig:map}A).
+One release site is located in each of the domains, namely Kommetjie in the Atlantic and Buffelsbaai in False Bay (see figure \ref{fig:map}B).
+The hydrodynamics between the two domains are driven by a complex interaction of wind and hydrodynamic vectors, with wind driven processes primarily occurring in False Bay.
+For a more detailed explanation of the hydrodynamics of these two domains the reader is referred to @coppin2020.
+The hydrodynamic characteristics of these two domains are also influenced by larger oceanic processes, which in South Africa can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC).
+The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly southwestward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean.
+
+```{r map data, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
+```{r map,echo=FALSE, message=FALSE, warning=FALSE, fig.cap='\\label{fig:map}Map of the study area and sampling sites.
+The blue shaded area is the Atlantic region and the red shaded region is the False Bay region.'}
 
@@
 This approach has been employed in modeling iceberg drift trajectory [@andersson2017; @lichey2001; @eik2009] and was adapted to suit this particular study.
-The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was, $$m({d_u/d_t})=F_c + F_a + F_f$$
-where $m$ is kelp mass, $d_u$ the virtual kelp particle velocity, $F_c$ the hydrodynamic drag, $F_a$ is the wind drag, and $F_f$ the surface-current flow.
+The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was, $$m({d_u/d_t})=F_d + F_a + F_f$$
+where $m$ is kelp mass, $d_u$ the virtual kelp particle velocity, $F_d$ the hydrodynamic drag, $F_a$ is the wind drag, and $F_f$ the surface-current flow.
 The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations.
@@
 
-## Cross-sectional area
+## Model inputs
+### Cross-sectional area
 
 In order to incorporate hydrodynamic and wind drag, the cross-sectional area of the kelp was calculated first.
-Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant, for details please see image ???.
-The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin] and estimated from figure \ref{fig:morphology}.
-
-## Hydrodynamic drag
-
-The use of shape coefficients was used as a novel approach to simulating biological drifting debris.
-The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
-The cross-sectional area for each of the virtual kelp parts was determined and used in the calculation of hydrodynamic drag.
-The hydrodynamic/form drag can be calculated by: $$ F_d = 0.5 \rho A_c C_d U^2$$
-
-## Wind drag
-
-The total wind drag effect was calculated for the hypothetical cross-sectional area of the kelp particle which was exposed above the water surface, which in this study, was estimated as 10% of the total cross-sectional area.
-
-## Raft-time
-
-Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remains afloat at the surface of the ocean after detachment.
-Raft-time was estimated from past studies by @graiff2016 and @graiff2013; which found mean raft times of approximately 30 days in moderate conditions.
-
-## Simulations
-
-# Results
-
-```{r data wrangling, message=FALSE, warning=FALSE, include=FALSE}
-
-# Read in data and data wrangling
-
-library(tidync)
-library(tidyverse)
-
-p_RK4 <- tidync("OceanParcels/trajectory_data/PassiveFloatRK4.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
-
-p_RK4_mean <- p_RK4 %>% # summarise data based on time 
-   group_by(t) %>% 
-   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
-
-p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
-
-p_EE <- tidync("OceanParcels/trajectory_data/PassiveFloatEE.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
-
-p_EE_mean <- p_EE %>% # summarise data based on time 
-   group_by(t) %>% 
-   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
-
-p_EE$type <- "EE" # add extra column 'type' to dataframe and assigned 'EE' label
-
-k_app1 <- tidync("OceanParcels/trajectory_data/KelpFloatApp1.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
-
-k_app1_mean <- k_app1 %>% # summarise data based on time 
-   group_by(t) %>% 
-   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
-
-k_app1$type <- "Approach 1" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-k_app2 <- tidync("OceanParcels/trajectory_data/KelpFloatApp2.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
-
-k_app2_mean <- k_app2 %>% # summarise data based on time 
-   group_by(t) %>% 
-   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
-
-k_app2$type <- "Approach 2" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-total_data <- rbind(p_RK4, p_EE, k_app1, k_app2)
-
-```
-
-```{r scatter advection, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
-
-total_traj <- ggplot(data = total_data, aes(x = lon, y = lat, colour = type)) + 
-  geom_point(aes(alpha = 0.2)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  theme_bw() +
-  guides(colour=guide_legend(title="Particle type"))
-total_traj
-
-```
-
-```{r plots, message=FALSE, warning=FALSE, include=FALSE}
-
-## 2D histogram plots of number of particles
-
-bin_RK4 <- ggplot(p_RK4, aes(x=lon, y=lat)) +
-  geom_bin2d(bins = 20) +
-  scale_fill_continuous(type = "viridis") +
-  coord_fixed() +
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
-
-bin_RK4 <- ggplot(p_EE, aes(x=lon, y=lat)) +
-  geom_bin2d(bins = 20) +
-  scale_fill_continuous(type = "viridis") +
-  coord_fixed() +
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
-
-bin_APP1 <- ggplot(k_app1, aes(x=lon, y=lat)) +
-  geom_bin2d(bins = 20) +
-  scale_fill_continuous(type = "viridis") +
-  coord_fixed() +
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
-
-bin_APP2 <- ggplot(k_app2, aes(x=lon, y=lat)) +
-  geom_bin2d(bins = 20) +
-  scale_fill_continuous(type = "viridis") +
-  coord_fixed() +
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
-
-## Heatmaps of study domain, use previous data for study site map and histograms
-
-## Coastline of Southern Africa
-load("map/south_africa_coast_hi_res.Rdata")
-
-heatmap_RK4 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
-
-heatmap_EE <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
-
-heatmap_APP1 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
-
-heatmap_APP2 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() 
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-
-library(ggpubr)
-ggarrange(heatmap_RK4, heatmap_EE, heatmap_APP1, heatmap_APP2, labels = c("A","B","C","D"), common.legend = TRUE, legend = c("right"))
-
-```
-
-# Discussion
-
-# Appendix
-
-Insert table of spherecity factors.
+Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant, for details please see image/table ???.
+The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020] and estimated from figure \ref{fig:morphology}.
 
@@
 
-```{r morphology, echo=FALSE, , fig.height=7, fig.width=8, message=FALSE, warning=FALSE, fig.cap='\\label{fig:morphology}Boxplots representing *E.
-pallida* morphological characteristics grouped by site around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured and with units provided.
+```{r morphology, echo=FALSE, , fig.height=7, fig.width=8, message=FALSE, warning=FALSE, fig.cap='\\label{fig:morphology}Boxplots representing *E.maxima* morphological characteristics by site around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured and with units provided.
 The lower and upper "hinges" correspond to the first and third quartiles.
@@
 
+### Hydrodynamic drag
+
+The use of a sphererical shape coefficient was used in the approach to simulate the trajectory of solitary floating *E.
+maxima*.
+The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
+The cross-sectional area for each of each virtual plant was converted to the equivalent cross-sectional area for that of a sphere in the equation for calculating hyrdrodynamic drag: $$ F_d = 0.5 \rho A_c C_d U^2$$
+
+### Wind drag
+
+The total wind drag effect was calculated for the hypothetical cross-sectional area of the kelp particle which was exposed above the water surface, which in this study, was estimated as 10% of the total cross-sectional area.
+
+### Raft-time
+
+Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remains afloat at the surface of the ocean after detachment before sinking.
+Raft-time was estimated from the results of past studies by @graiff2016 and @graiff2013 for each season.
+For summer the raft time was 25 days, winter was 38 days, and autumn was 41 days respectivley.
+
+## Simulations
+
+In order to determine the influence of morphology on drag components, simulations were performed for the minimum, median and maximum cross-sectional areas for one site on the Atlantic side and one site within False Bay (see figure \ref{fig:map})
+
+## Statisical analysis
+
+# Results
+
+```{r data wrangling, message=FALSE, warning=FALSE, include=FALSE}
+
+# Read in data and data wrangling
+
+library(tidync)
+library(tidyverse)
+
+p_RK4 <- tidync("OceanParcels/trajectory_data/PassiveFloatRK4.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+
+p_RK4_mean <- p_RK4 %>% # summarise data based on time 
+   group_by(t) %>% 
+   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+
+p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
+
+p_EE <- tidync("OceanParcels/trajectory_data/PassiveFloatEE.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+
+p_EE_mean <- p_EE %>% # summarise data based on time 
+   group_by(t) %>% 
+   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+
+p_EE$type <- "EE" # add extra column 'type' to dataframe and assigned 'EE' label
+
+k_app1 <- tidync("OceanParcels/trajectory_data/KelpFloatApp1.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+
+k_app1_mean <- k_app1 %>% # summarise data based on time 
+   group_by(t) %>% 
+   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+
+k_app1$type <- "Approach 1" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+k_app2 <- tidync("OceanParcels/trajectory_data/KelpFloatApp2.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+
+k_app2_mean <- k_app2 %>% # summarise data based on time 
+   group_by(t) %>% 
+   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+
+k_app2$type <- "Approach 2" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+total_data <- rbind(p_RK4, p_EE, k_app1, k_app2)
+
+```
+
+```{r plots, message=FALSE, warning=FALSE, include=FALSE}
+
+library(hexbin)
+
+## Heatmaps of study domain, use previous data for study site map and histograms
+
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+heatmap_RK4 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_EE <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = p_EE, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_APP1 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = k_app1, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_APP2 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Comparison of density of particles within each grid cell for the end run time of each scenario.'}
+
+library(ggpubr)
+ggarrange(heatmap_RK4, heatmap_EE, heatmap_APP1, heatmap_APP2, labels = c("A","B","C","D"), common.legend = TRUE, legend = c("right"), ncol = 2, nrow = 2)
+
+```
+
+```{r scatter plot processing, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+
+age_RK4 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = p_RK4, aes(x = lon, y = lat, alpha = 0.2, color = age)) + 
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Age of particles")) +
+  theme_bw()
+
+age_EE <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = p_EE, aes(x = lon, y = lat, alpha = 0.2, color = age)) + 
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Age of particles")) +
+  theme_bw()
+
+age_APP1 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = k_app1, aes(x = lon, y = lat, alpha = 0.2, color = age)) + 
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Age of particles")) +
+  theme_bw()
+
+age_APP2 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = k_app2, aes(x = lon, y = lat, alpha = 0.2, color = age)) + 
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Age of particles")) +
+  theme_bw()
+
+```
+
+```{r age plots, warning=FALSE, echo=FALSE, include=TRUE, message=FALSE}
+
+ggarrange(age_RK4, age_EE, age_APP1, age_APP2, labels = c("A","B","C","D"), common.legend = TRUE, legend = c("right"), ncol = 2, nrow = 2)
+```
+
+# Discussion
+
+# Appendix
+
+Insert table of spherecity factors.
+
 # References

### 2020-10-16 thesis_chapter_3.Rmd
added to methods section. basic stats table for morphology added to appendix. general tidy.

@@
   html_document:
+  pdf_document:
+    fig_caption: yes
   word_document:
     reference_docx: markdown/thesis_template.docx
-  pdf_document:
-    fig_caption: yes
 bibliography: references.bib
@@
 One release site is located in each of the domains, namely Kommetjie in the Atlantic and Buffelsbaai in False Bay (see figure \ref{fig:map}B).
-The hydrodynamics between the two domains are driven by a complex interaction of wind and hydrodynamic vectors, with wind driven processes primarily occurring in False Bay.
+The hydrodynamics between the two domains are driven by a complex interaction of wind and wave vectors, with wind driven processes primarily occurring in False Bay.
 For a more detailed explanation of the hydrodynamics of these two domains the reader is referred to @coppin2020.
-The hydrodynamic characteristics of these two domains are also influenced by larger oceanic processes, which in South Africa can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC).
+The hydrodynamic characteristics of these regions are also influenced by larger oceanic processes, which in South Africa can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC).
 The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly southwestward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean.
@@
 maxima* individuals will be considered.
-Previous studies have made the assumption that kelp rafts move with the same velocity as the ocean current, and instead model a ensemble of particles which trajectories are determined by a combination of ocean currents and/or the effects of winds on currents.
-In addition, these studies assume the particle is fully buoyant.
+Previous studies have made the assumption that floating macroalage move with the same velocity as the ocean current, and instead model a ensemble of particles which trajectories are determined by a combination of ocean currents and/or the effects of winds on currents.
 
@@
 
-This study simulated two scenarios; one where the plant is fully submerged and only exposed to seawater drag and the other where parts of the plants are partially exposed resulting in a combination of seawater and air form drag.
-Unlike previous studies, drag based on shape based coefficients was used in the calculations for determining the reduction in velocity as a result of drag.
+This study simulated two scenarios; one where the plant is fully submerged and only exposed to seawater drag; and the other where parts of the plants are partially exposed resulting in a combination of seawater and air form drag.
+Drag based on shape based coefficients were used in the calculations for determining the overall kelp velocity vectors.
 A momentum energy equation was used to calculate the drag forces for the relevant simulation.
@@
 The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations.
-Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculated before input into the model.
+Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculatedfor each time step of each simulation.
 
@@
 Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant, for details please see image/table ???.
-The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020] and estimated from figure \ref{fig:morphology}.
+The dimensional data needed was estimated in cases were data was not available for that particular morphological characteristic.
+The bulb/pneumatocyst is a highly variable morphological characteristic and in some cases can appear absent, the same is true for the holdfast.
+Therefore, the cross-sectional area of the bulb/pneumatocyst was not considered in the calculation of overall cross-sectional area.
+In terms of the holdfast, a standard cross-sectional area was used for simulations where the inclusion of the holdfast was considered.
 
-```{r morphology_wrangling, message=FALSE, warning=FALSE, include=FALSE}
-# Investigate morphology data to assist in estimating surface area
-# Load software
-library(tidyverse)
-library(ggpubr)
-library(gridExtra)
-library(ggplot2)
-library(ggord)
-library(vegan)
-# Read in morphology data
-eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
-  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
-  filter(site!= "De Hoop")
-eck_data$site <- as.character(eck_data$site)
-eck_deep <- eck_data %>% 
-  filter(population == "Deep")
-
-eck_long <- gather(eck_deep, morphology, measurement, frond_mass:total_length, factor_key=TRUE)
-# Summary stats to make it easier to refer to later, these will not be used any analysis.
-eck_stats <- eck_long %>% 
-  group_by(site, morphology) %>% 
-  summarise(mean = mean(measurement, na.rm = T),
-            min = min(measurement, na.rm = T),
-            max = max(measurement, na.rm = T),
-            range = range(measurement, na.rm = T)[2]-range(measurement, na.rm = T)[1],
-            sd = sd(measurement, na.rm = T)) %>% 
-  ungroup()
-# Make custom positions
-positionsEck <- c("Betty's Bay", "Kalk Bay", "A-Frame", 
-                  "Miller's Point", "Baboon Rock", "Batsata Rock", 
-                  "Buffelsbaai", "Scarborough", 
-                  "Soetwater","Kommetjie","Hout Bay", 
-                  "Oudekraal", "St Helena Bay")
-# Make plots for each morphology feature and grouped by site.
-frond_mass <- ggplot(data = eck_data, aes(x = site, y = frond_mass, fill = site)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Frond mass (g)") + 
-           coord_flip() +
-           xlab("") + 
-           theme_bw()
-primary_length <- ggplot(data = eck_data, aes(x = site, y = primary_length , fill = site)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Primary Length  (cm)") + 
-           coord_flip() +
-           xlab("") + 
-           theme_bw()
-primary_width <- ggplot(data = eck_data, aes(x = site, y = primary_width , fill = site)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Primary Width (cm)") + 
-           coord_flip() +
-           xlab("") + 
-           theme_bw() 
-frond_length <- ggplot(data = eck_data, aes(x = site, y = frond_length , fill = site)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Frond length (cm)") + 
-           coord_flip() +
-           xlab("") + 
-           theme_bw() 
-stipe_mass <- ggplot(data = eck_data, aes(x = site, y = stipe_mass , fill = site)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Stipe mass (g)") + 
-           coord_flip() +
-           xlab("") + 
-           theme_bw() 
-stipe_length <- ggplot(data = eck_data, aes(x = site, y = stipe_length , fill = site)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Stipe length (cm)") + 
-           coord_flip() +
-           xlab("") + 
-           theme_bw() 
-stipe_circ <- ggplot(data = eck_data, aes(x = site, y = stipe_circ , fill = site)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Stipe circ (cm)") + 
-           coord_flip() +
-           xlab("") + 
-           theme_bw() 
-tufts <- ggplot(data = eck_data, aes(x = site, y = tufts , fill = site)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("No. of tufts") + 
-           coord_flip() +
-           xlab("") + 
-           theme_bw() 
-epi_length <- ggplot(data = eck_data, aes(x = site, y = epi_length , fill = site)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Epi Length  (cm)") + 
-           coord_flip() +
-           xlab("") + 
-           theme_bw()
-
-total_length <- ggplot(data = eck_data, aes(x = site, y = epi_length, fill = site)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Total length  (cm)") + 
-           coord_flip() +
-           xlab("") + 
-           theme_bw()
-```
-
-```{r morphology, echo=FALSE, , fig.height=7, fig.width=8, message=FALSE, warning=FALSE, fig.cap='\\label{fig:morphology}Boxplots representing *E.maxima* morphological characteristics by site around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured and with units provided.
-The lower and upper "hinges" correspond to the first and third quartiles.
-The whiskers represent the range, solid black lines represent the median and black dots are outliers.'}
-# Plot the morphology data based on site (see previous section)
-ggarrange(frond_length, frond_mass,primary_length, primary_width, stipe_circ, stipe_length, stipe_mass, total_length, tufts, common.legend = TRUE, legend = c("none"), align = c("hv"))
-```
+The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
+The minimum, maximum and mean were calculated for each morphological characteristic (see appendix), which were used for calculating the overall minimum, maximum and mean cross-sectional areas needed to run the various simulations.
+The trajectory of these "types" of cross-sectional areas (minimum, maximum and mean overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
 
@@
 
-The use of a sphererical shape coefficient was used in the approach to simulate the trajectory of solitary floating *E.
+The use of a spherical shape coefficient was used in the approach to simulate the trajectory of solitary floating *E.
 maxima*.
 The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
-The cross-sectional area for each of each virtual plant was converted to the equivalent cross-sectional area for that of a sphere in the equation for calculating hyrdrodynamic drag: $$ F_d = 0.5 \rho A_c C_d U^2$$
+The cross-sectional area for each of each virtual plant was converted to the equivalent cross-sectional area for that of a sphere in the equation for calculating hydrodynamic drag: $$ F_d = 0.5 \rho A_c C_d U^2$$
 
@@
 Raft-time was estimated from the results of past studies by @graiff2016 and @graiff2013 for each season.
-For summer the raft time was 25 days, winter was 38 days, and autumn was 41 days respectivley.
+For summer the raft time was 25 days, winter was 38 days, and autumn was 41 days respectively.
 
@@
 
-## Statisical analysis
+## Statistical analysis
 
@@
 
+```{r morphology_wrangling, message=FALSE, warning=FALSE, include=FALSE}
+# Investigate morphology data to assist in estimating surface area
+
+# Load software
+library(tidyverse)
+library(ggpubr)
+library(gridExtra)
+library(ggplot2)
+library(ggord)
+library(vegan)
+library(dplyr)
+
+# Read in morphology data
+eck_df <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
+  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
+  filter(site!= "De Hoop")
+
+eck_df$site <- as.character(eck_df$site)
+eck_df <- eck_df %>% 
+  filter(population == "Deep")
+
+eck_long_df <- gather(eck_df, Morphology, measurement, frond_mass:total_length, factor_key=TRUE)
+# Summary stats to make it easier to refer to later, these will not be used any analysis.
+
+site_list <- c("Kommetjie")
+
+eck_stats <- eck_long_df %>% 
+  group_by(site, Morphology) %>% 
+  summarise(mean = mean(measurement, na.rm = T),
+            minimum = min(measurement, na.rm = T),
+            maximum = max(measurement, na.rm = T)) %>% 
+  ungroup() %>% 
+  filter(site %in% site_list) %>% 
+  rename(Site = site) %>% 
+  rename(Mean = mean) %>% 
+  rename(Minimum = minimum) %>% 
+  rename(Maximum = maximum)
+
+#df <- eck_stats
+#rownames(df)[rownames(df) == "frond_mass"] <- "Frond mass"
+#
+#  eck_stats <- eck_stats %>% 
+#  column_to_rownames(var = "Morphology") %>% 
+#  mutate(Morphology = replace(Morphology, 
+#                              Morphology == "epi_length", "Epiphyte length (cm)")) %>% 
+#  mutate(Morphology = replace(Morphology, 
+#                              Morphology == "frond_length", "Frond length (cm)")) %>% 
+#  mutate(Morphology = replace(Morphology, 
+#                              Morphology == "frond_mass", "Frond mass (kg)")) %>% 
+#  mutate(Morphology = replace(Morphology, 
+#                              Morphology == "primary_length", "Primary length (cm)")) %>% 
+#  mutate(Morphology = replace(Morphology, 
+#                              Morphology == "primary_width", "Primary width (cm)")) %>% 
+#  mutate(Morphology = replace(Morphology, 
+#                              Morphology == "stipe_circ", "Stipe circumference (mm)")) %>% 
+#  mutate(Morphology = replace(Morphology, 
+#                              Morphology == "stipe_length", "Stipe length (cm)")) %>% 
+#  mutate(Morphology = replace(Morphology, 
+#                              Morphology == "stipe_mass", "Stipe mass (kg)")) %>% 
+#  mutate(Morphology = replace(Morphology, 
+#                              Morphology == "total_length", "Total length (cm)")) %>% 
+#  mutate(Morphology = replace(Morphology, 
+#                              Morphology == "tufts", "Number of tufts"))
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+library(knitr)
+library(kableExtra)
+
+knitr::kable(eck_stats, format = "html") %>%
+  kable_styling()
+
+```
+
 Insert table of spherecity factors.

### 2020-10-29 thesis_chapter_3.Rmd
updated rmd and outputs. table added

@@
   word_document:
-    reference_docx: markdown/thesis_template.docx
 bibliography: references.bib
@@
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
 
@@
 Chemical engineers in the field of fluid dynamics which come across similar problems when including drag use an approach known as a 'spherecity factor'.
-A spherecity factor is the ratio of the surface area of a sphere having the volume as the irregular shaped particle (s) to the actual surface area of the particle (s_d) such that $$\psi=s/s_d$$.
-The values are dependent on the shape of the object and a range of values have been
-determined based on numerous experimental work. A table of various shapes and the associated spherecity factor is provided in the appendix.
+A spherecity factor is expressed mathematically by $\psi=s/s_d$, and the ratio of the surface area of a sphere having the volume as the irregular shaped particle ($s$) to the actual surface area of the particle ($s_d$).
+The values are dependent on the shape of the object and a range of values have been determined based on numerous experimental work.
 
-There are considerable gaps in the literature with regards to macroalgae trajectory such as buoyancy, air form drag, water form drag, epibiont load, growth, fragmentation, size, shape and orientation which have not been investigated numerically in the past.
 In this study, a hypothetical approach was used to shed light on the role of drag on macroalgal drift trajectory by means of Langrangian based trajectory simulations.
@@
 
-## Load software:
+## Load software need for entire script
+
 library(ggplot2)
@@
 library(gridExtra)
+library(ggpubr)
+#library(ggord)
+library(vegan)
+library(dplyr)
+library(ggpubr)
+library(maps)
+
 source("map/scaleBarFunc.R")
@@
 
-library(ggpubr)
-
 ggarrange(map_pen, release_map, labels = c("AUTO"), align = c("hv"), ncol = 2, nrow = 1)
@@
 The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations.
-Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculatedfor each time step of each simulation.
+Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculated for each time step of each simulation.
 
@@
 In terms of the holdfast, a standard cross-sectional area was used for simulations where the inclusion of the holdfast was considered.
-
 The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
@@
 The trajectory of these "types" of cross-sectional areas (minimum, maximum and mean overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
+
+```{r sim table, echo=FALSE, message=FALSE, warning=FALSE}
+
+library(flextable)
+library(officer)
+
+sim_table <- read.csv("tables/chapter_3/simulation_table.csv", sep = ";")
+
+sim_table <- flextable(data = sim_table)
+sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
+sim_table <- set_header_labels(sim_table, Particle.behaviour = "Particle behaviour", Cross.sectional.area = "Cross-sectional area", Time.Range = "Time Range", Release.Lon = "Release longitude", Release.Lat = "Release Latitude", Hydrodynamic.model = "Hydrodynamic model", Wind.Model = "Wind model")
+sim_table <- align(x= sim_table, align = "center", part = "all")
+sim_table
+
+```
 
@@
 
-```{r morphology_wrangling, message=FALSE, warning=FALSE, include=FALSE}
+```{r morphology_wrangling, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 # Investigate morphology data to assist in estimating surface area
-
 # Load software
@@
 library(ggplot2)
-library(ggord)
 library(vegan)
 library(dplyr)
-
 # Read in morphology data
@@
   filter(site!= "De Hoop") 
-
 eck_df$site <- as.character(eck_df$site)
@@
   filter(population == "Deep") 
-
 eck_long_df <- gather(eck_df, Morphology, measurement, frond_mass:total_length, factor_key=TRUE)
 # Summary stats to make it easier to refer to later, these will not be used any analysis.
-
 site_list <- c("Kommetjie")
-
 eck_stats <- eck_long_df %>% 
@@
 
-#df <- eck_stats
-#rownames(df)[rownames(df) == "frond_mass"] <- "Frond mass"
-#
-#  eck_stats <- eck_stats %>% 
-#  column_to_rownames(var = "Morphology") %>% 
-#  mutate(Morphology = replace(Morphology, 
-#                              Morphology == "epi_length", "Epiphyte length (cm)")) %>% 
-#  mutate(Morphology = replace(Morphology, 
-#                              Morphology == "frond_length", "Frond length (cm)")) %>% 
-#  mutate(Morphology = replace(Morphology, 
-#                              Morphology == "frond_mass", "Frond mass (kg)")) %>% 
-#  mutate(Morphology = replace(Morphology, 
-#                              Morphology == "primary_length", "Primary length (cm)")) %>% 
-#  mutate(Morphology = replace(Morphology, 
-#                              Morphology == "primary_width", "Primary width (cm)")) %>% 
-#  mutate(Morphology = replace(Morphology, 
-#                              Morphology == "stipe_circ", "Stipe circumference (mm)")) %>% 
-#  mutate(Morphology = replace(Morphology, 
-#                              Morphology == "stipe_length", "Stipe length (cm)")) %>% 
-#  mutate(Morphology = replace(Morphology, 
-#                              Morphology == "stipe_mass", "Stipe mass (kg)")) %>% 
-#  mutate(Morphology = replace(Morphology, 
-#                              Morphology == "total_length", "Total length (cm)")) %>% 
-#  mutate(Morphology = replace(Morphology, 
-#                              Morphology == "tufts", "Number of tufts"))
+eck_stats <- flextable(data = eck_stats)
+eck_stats <- theme_zebra(eck_stats, odd_header = "transparent", even_header = "transparent")
+eck_stats <- align(x= eck_stats, align = "center", part = "all")
+eck_stats
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-
-library(knitr)
-library(kableExtra)
-
-knitr::kable(eck_stats, format = "html") %>%
-  kable_styling()
-
-```
-
-Insert table of spherecity factors.
-
 # References

### 2020-11-02 thesis_chapter_3.Rmd
added in ocean and wind model desciption

@@
 
-The characteristics of the underlying numerical flow model will be explained here...
+Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service.
+The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
+The model has a spatial resolution of 1/12$^\circ$ and is interpolated on a Arakawa C native grid.
 
@@
 
-The characteristics of the underlying numerical wind model will be explained here...
+In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (WIND_L4_NRT_OBSERVATIONS_012_004).
+The outputs from the model is composed of 6-hourly avergaed fields of surface 10m wind velocity vectors with a spatial resolution of 1/4$^\circ$.
 

### 2020-11-02 thesis_chapter_3.Rmd
changed and updated simulation table

@@
 
-```{r sim table, echo=FALSE, message=FALSE, warning=FALSE}
-
-library(flextable)
-library(officer)
-
-sim_table <- read.csv("tables/chapter_3/simulation_table.csv", sep = ";")
-
-sim_table <- flextable(data = sim_table)
-sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
-sim_table <- set_header_labels(sim_table, Particle.behaviour = "Particle behaviour", Cross.sectional.area = "Cross-sectional area", Time.Range = "Time Range", Release.Lon = "Release longitude", Release.Lat = "Release Latitude", Hydrodynamic.model = "Hydrodynamic model", Wind.Model = "Wind model")
-sim_table <- align(x= sim_table, align = "center", part = "all")
-sim_table
-
-```
-
 ### Hydrodynamic drag
@@
 In order to determine the influence of morphology on drag components, simulations were performed for the minimum, median and maximum cross-sectional areas for one site on the Atlantic side and one site within False Bay (see figure \ref{fig:map})
+
+```{r sim table, echo=FALSE, message=FALSE, warning=FALSE}
+
+library(flextable)
+library(officer)
+
+sim_table <- read.csv("tables/chapter_3/simulation_table.csv", sep = ";")
+
+sim_table <- flextable(data = sim_table)
+sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
+sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Time.Range = "Time Range", Drag.Exposure = "Drag Exposure level")
+sim_table <- align(x= sim_table, align = "center", part = "all")
+sim_table
+
+```
 

### 2020-11-04 thesis_chapter_3.Rmd
added surface area table
update

@@
 output:
-  html_document:
   pdf_document:
     fig_caption: yes
+  html_document:
   word_document:
@@
 
+area_table <- read.csv("tables/chapter_3/area_table.csv", sep = ";")
+area_table <- flextable(data = area_table)
+area_table <- theme_zebra(area_table, odd_header = "transparent", even_header = "transparent")
+area_table <- align(x= area_table, align = "center", part = "all")
+area_table
 ```

### 2020-11-05 thesis_chapter_3.Rmd
changes to script and tables updated

--- 2020-11-02 thesis_chapter_3.Rmd
@@
 
-p_RK4 <- tidync("OceanParcels/trajectory_data/PassiveFloatRK4.nc") %>%
+p_RK4 <- tidync("OceanParcels/trajectory_data/RK4_PFloat.nc") %>%
   hyper_tibble() %>%
@@
   dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
-
-p_RK4_mean <- p_RK4 %>% # summarise data based on time 
-   group_by(t) %>% 
-   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
 
@@
 
-p_EE <- tidync("OceanParcels/trajectory_data/PassiveFloatEE.nc") %>%
+Mean_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H100_W00.nc") %>%
   hyper_tibble() %>%
@@
   dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
 
-p_EE_mean <- p_EE %>% # summarise data based on time 
-   group_by(t) %>% 
-   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+Mean_Kfloat_H100_W00$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
 
-p_EE$type <- "EE" # add extra column 'type' to dataframe and assigned 'EE' label
-
-k_app1 <- tidync("OceanParcels/trajectory_data/KelpFloatApp1.nc") %>%
+Mean_KFloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H095_W05.nc") %>%
   hyper_tibble() %>%
@@
   dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
 
-k_app1_mean <- k_app1 %>% # summarise data based on time 
-   group_by(t) %>% 
-   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+Mean_KFloat_H095_W05$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
 
-k_app1$type <- "Approach 1" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-k_app2 <- tidync("OceanParcels/trajectory_data/KelpFloatApp2.nc") %>%
+Mean_KFloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H090_W10.nc") %>%
   hyper_tibble() %>%
@@
   dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
 
-k_app2_mean <- k_app2 %>% # summarise data based on time 
-   group_by(t) %>% 
-   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)
+Mean_KFloat_H090_W10$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
 
-k_app2$type <- "Approach 2" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H085_W15.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
 
-total_data <- rbind(p_RK4, p_EE, k_app1, k_app2)
+Mean_KFloat_H085_W15$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+total_data <- rbind(p_RK4, Mean_Kfloat_H100_W00, Mean_KFloat_H095_W05, Mean_KFloat_H090_W10, Mean_KFloat_H085_W15)
 
@@
 
-heatmap_EE <- ggplot() +
+heatmap_Mean_Kfloat_H100_W00 <- ggplot() +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_EE, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
 
-heatmap_APP1 <- ggplot() +
+heatmap_Mean_KFloat_H095_W05 <- ggplot() +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = k_app1, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
 
-heatmap_APP2 <- ggplot() +
+heatmap_Mean_KFloat_H090_W10 <- ggplot() +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_Mean_KFloat_H085_W15 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
@@
 library(ggpubr)
-ggarrange(heatmap_RK4, heatmap_EE, heatmap_APP1, heatmap_APP2, labels = c("A","B","C","D"), common.legend = TRUE, legend = c("right"), ncol = 2, nrow = 2)
+ggarrange(heatmap_RK4, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, labels = c("A","B","C","D", "E"), common.legend = TRUE, legend = c("right"), ncol = 2, nrow = 3)
 
@@
 
-```{r scatter plot processing, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+```{r scatter plot processing, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-```{r age plots, warning=FALSE, echo=FALSE, include=TRUE, message=FALSE}
+```{r age plots, warning=FALSE, echo=FALSE, include=TRUE, message=FALSE, eval=FALSE}
 
@@
 
-```{r morphology_wrangling, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+```{r morphology_wrangling, eval=TRUE, message=FALSE, warning=FALSE, include=TRUE}
 # Investigate morphology data to assist in estimating surface area

### 2020-11-05 thesis_chapter_3.Rmd
updated script to match new simulations

@@
 
+Min_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H100_W00.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Min_Kfloat_H100_W00$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H095_W05.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Min_Kfloat_H095_W05$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H090_W10.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Min_Kfloat_H090_W10$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H085_W15.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Min_Kfloat_H085_W15$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+
 Mean_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H100_W00.nc") %>%
@@
 
-Mean_KFloat_H085_W15$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
 
-total_data <- rbind(p_RK4, Mean_Kfloat_H100_W00, Mean_KFloat_H095_W05, Mean_KFloat_H090_W10, Mean_KFloat_H085_W15)
+Max_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H100_W00.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Max_Kfloat_H100_W00$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H095_W05.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Max_Kfloat_H095_W05$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H090_W10.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Max_Kfloat_H090_W10$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H085_W15.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Max_Kfloat_H085_W15$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+total_data <- rbind(p_RK4, Min_Kfloat_H100_W00 , Min_Kfloat_H095_W05, Min_Kfloat_H090_W10, Min_Kfloat_H085_W15, Mean_Kfloat_H100_W00, Mean_KFloat_H095_W05, Mean_KFloat_H090_W10, Mean_KFloat_H085_W15, Max_Kfloat_H100_W00, Max_Kfloat_H095_W05, Max_Kfloat_H090_W10, Max_Kfloat_H085_W15)
 
@@
   geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_Min_Kfloat_H100_W00 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_Min_Kfloat_H095_W05 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_Min_Kfloat_H090_W10 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_Min_Kfloat_H085_W15 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
 
+heatmap_Max_KFloat_H100_W00 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_Max_KFloat_H095_W05 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_Max_KFloat_H090_W10 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+heatmap_Max_KFloat_H085_W15 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
 ```
@@
 library(ggpubr)
-ggarrange(heatmap_RK4, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, labels = c("A","B","C","D", "E"), common.legend = TRUE, legend = c("right"), ncol = 2, nrow = 3)
+ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15 , labels = c("AUTO"), common.legend = TRUE, legend = c("right"))
 

### 2020-11-05 thesis_chapter_3.Rmd
removed unneeded script

@@
 
-```{r scatter plot processing, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
-
-age_RK4 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = p_RK4, aes(x = lon, y = lat, alpha = 0.2, color = age)) + 
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Age of particles")) +
-  theme_bw()
-
-age_EE <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = p_EE, aes(x = lon, y = lat, alpha = 0.2, color = age)) + 
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Age of particles")) +
-  theme_bw()
-
-age_APP1 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = k_app1, aes(x = lon, y = lat, alpha = 0.2, color = age)) + 
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Age of particles")) +
-  theme_bw()
-
-age_APP2 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = k_app2, aes(x = lon, y = lat, alpha = 0.2, color = age)) + 
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Age of particles")) +
-  theme_bw()
-
-```
-
-```{r age plots, warning=FALSE, echo=FALSE, include=TRUE, message=FALSE, eval=FALSE}
-
-ggarrange(age_RK4, age_EE, age_APP1, age_APP2, labels = c("A","B","C","D"), common.legend = TRUE, legend = c("right"), ncol = 2, nrow = 2)
-```
-
 # Discussion

### 2020-11-09 thesis_chapter_3.Rmd
added more traj plots

@@
 ---
+
+```{r software, message=FALSE, warning=FALSE, include=FALSE}
+## Load software need for entire script
+
+library(ggplot2)
+library(rgdal)
+library(tidyverse)
+library(viridis)
+library(gridExtra)
+library(ggpubr)
+#library(ggord)
+library(vegan)
+library(dplyr)
+library(ggpubr)
+library(maps)
+library(flextable)
+library(officer)
+library(tidync)
+library(tidyverse)
+library(hexbin)
+library(ggpubr)
+
+```
 
@@
 ```{r map data, message=FALSE, warning=FALSE, include=FALSE}
-
-## Load software need for entire script
-
-library(ggplot2)
-library(rgdal)
-library(tidyverse)
-library(viridis)
-library(gridExtra)
-library(ggpubr)
-#library(ggord)
-library(vegan)
-library(dplyr)
-library(ggpubr)
-library(maps)
 
@@
 
-library(flextable)
-library(officer)
-
 sim_table <- read.csv("tables/chapter_3/simulation_table.csv", sep = ";")
@@
 
-library(tidync)
-library(tidyverse)
-
 p_RK4 <- tidync("OceanParcels/trajectory_data/RK4_PFloat.nc") %>%
   hyper_tibble() %>%
-  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
   dplyr::rename(t = time) %>%
@@
   hyper_tibble() %>%
-  dplyr::select(lon, lat, time, distance, age) %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
   dplyr::rename(t = time) %>%
@@
 
-```{r plots, message=FALSE, warning=FALSE, include=FALSE}
+```{r summary wrangle, message= FALSE, warning = FALSE, echo = FALSE}
 
-library(hexbin)
+p_RK4_mean <- p_RK4 %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H100_W00_mean <- Min_Kfloat_H100_W00 %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H095_W05_mean <- Min_Kfloat_H095_W05 %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H090_W10_mean <- Min_Kfloat_H090_W10 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H085_W15_mean <- Min_Kfloat_H085_W15 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_Kfloat_H100_W00_mean <- Mean_Kfloat_H100_W00 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H095_W05_mean <- Mean_KFloat_H095_W05 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H090_W10_mean <- Mean_KFloat_H090_W10 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H085_W15_mean <- Mean_KFloat_H085_W15 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H100_W00_mean <- Max_Kfloat_H100_W00 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H095_W05_mean <- Max_Kfloat_H095_W05 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H090_W10_mean <- Max_Kfloat_H090_W10 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H085_W15_mean <- Max_Kfloat_H085_W15 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+```
+
+```{r density plots, message=FALSE, warning=FALSE, echo=FALSE}
 
@@
 
-library(ggpubr)
 ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15 , labels = c("AUTO"), common.legend = TRUE, legend = c("right"))
+```
 
+```{r path plots, warning=FALSE, message=FALSE, echo=FALSE}
+
+min_area_path <- ggplot() + 
+  geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'blue'), show.legend = FALSE) +
+  geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'yellow'), show.legend = FALSE) +
+  geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'pink'), show.legend = FALSE) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("Minimum cross-sectional area") + 
+  theme_bw()
+
+mean_area_path <- ggplot() + 
+  geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'blue'), show.legend = FALSE) +
+  geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'yellow'), show.legend = FALSE) +
+  geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'pink'), show.legend = FALSE) +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("Mean cross-sectional area") + 
+  theme_bw()
+
+max_area_path <- ggplot() + 
+  geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'blue'), show.legend = FALSE) +
+  geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'yellow'), show.legend = FALSE) +
+  geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'pink'), show.legend = FALSE) +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("Maximum cross-sectional area") + 
+  theme_bw()
+
+ggarrange(min_area_path, mean_area_path, max_area_path, labels = "AUTO")
 ```
@@
 
-```{r morphology_wrangling, eval=TRUE, message=FALSE, warning=FALSE, include=TRUE}
+```{r morphology wrangling, message=FALSE, warning=FALSE, include=FALSE}
 # Investigate morphology data to assist in estimating surface area
-# Load software
-library(tidyverse)
-library(ggpubr)
-library(gridExtra)
-library(ggplot2)
-library(vegan)
-library(dplyr)
+
 # Read in morphology data
@@
 area_table <- align(x= area_table, align = "center", part = "all")
+
+```
+
+```{r area table, echo=FALSE, message=FALSE, warning=FALSE}
 area_table

### 2020-11-11 thesis_chapter_3.Rmd
tidy up

@@
 output:
+  html_document:
   pdf_document:
     fig_caption: yes
-  html_document:
   word_document:
@@
   ggtitle("Minimum cross-sectional area") + 
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ggtitle("Mean cross-sectional area") + 
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ggtitle("Maximum cross-sectional area") + 
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
 ggarrange(min_area_path, mean_area_path, max_area_path, labels = "AUTO")
+
 ```

### 2020-11-16 thesis_chapter_3.Rmd
added eddy script. script tidy for simulation data

@@
 
-# Read in data and data wrangling
+# load and data wrangle for eddy simulation
+
+p_eddy <- tidync("OceanParcels/trajectory_data/PassiveEddy.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+p_eddy_mean <- p_eddy %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+p_eddy$type <- "Passive Eddy" # add extra column 'type' to dataframe
+
+Min_KFloat_Eddy <- tidync("OceanParcels/trajectory_data/Min_KFloat_Eddy.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+Min_K_Eddy_mean <- Min_KFloat_Eddy %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_KFloat_Eddy$type <- "Min_KFloat_Eddy" # add extra column 'type' to dataframe
+
+Mean_KFloat_Eddy <- tidync("OceanParcels/trajectory_data/Mean_KFloat_Eddy.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+Mean_KFloat_Eddy_mean <- Mean_KFloat_Eddy %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_Eddy$type <- "Mean_KFloat_Eddy" # add extra column 'type' to dataframe
+
+Max_KFloat_Eddy <- tidync("OceanParcels/trajectory_data/Max_KFloat_Eddy.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+Max_KFloat_Eddy_mean <- Max_KFloat_Eddy %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_KFloat_Eddy$type <- "Max_KFloat_Eddy" # add extra column 'type' to dataframe
+
+# Read in simulation data and data wrangling
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+p_RK4_mean <- p_RK4 %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H100_W00_mean <- Min_Kfloat_H100_W00 %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H095_W05_mean <- Min_Kfloat_H095_W05 %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H090_W10_mean <- Min_Kfloat_H090_W10 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H085_W15_mean <- Min_Kfloat_H085_W15 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_Kfloat_H100_W00_mean <- Mean_Kfloat_H100_W00 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H095_W05_mean <- Mean_KFloat_H095_W05 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H090_W10_mean <- Mean_KFloat_H090_W10 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H085_W15_mean <- Mean_KFloat_H085_W15 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H100_W00_mean <- Max_Kfloat_H100_W00 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H095_W05_mean <- Max_Kfloat_H095_W05 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H090_W10_mean <- Max_Kfloat_H090_W10 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H085_W15_mean <- Max_Kfloat_H085_W15 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
 
@@
 
-```{r summary wrangle, message= FALSE, warning = FALSE, echo = FALSE}
+```{r eddy plots, message=FALSE, warning=FALSE, include=FALSE}
 
-p_RK4_mean <- p_RK4 %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
+passive_eddy <- ggplot() + 
+  geom_point(data = p_eddy, aes(x = lon, y = lat)) +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Passive") + 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
-Min_Kfloat_H100_W00_mean <- Min_Kfloat_H100_W00 %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
+min_eddy <- ggplot() + 
+  geom_point(data = Min_K_Eddy, aes(x = lon, y = lat)) +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Minimum Ac") + 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
-Min_Kfloat_H095_W05_mean <- Min_Kfloat_H095_W05 %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
+mean_eddy <- ggplot() + 
+  geom_point(data = Mean_KFloat_Eddy, aes(x = lon, y = lat)) +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Mean Ac") + 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
-Min_Kfloat_H090_W10_mean <- Min_Kfloat_H090_W10 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
+max_eddy <- ggplot() + 
+  geom_point(data = Max_KFloat_Eddy, aes(x = lon, y = lat)) +
+  xlab("Longtitude") + 
+  ylab("Latitude") +
+  ggtitle("Maximum Ac") + 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
-Min_Kfloat_H085_W15_mean <- Min_Kfloat_H085_W15 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
+```
 
-Mean_Kfloat_H100_W00_mean <- Mean_Kfloat_H100_W00 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H095_W05_mean <- Mean_KFloat_H095_W05 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H090_W10_mean <- Mean_KFloat_H090_W10 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H085_W15_mean <- Mean_KFloat_H085_W15 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H100_W00_mean <- Max_Kfloat_H100_W00 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H095_W05_mean <- Max_Kfloat_H095_W05 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H090_W10_mean <- Max_Kfloat_H090_W10 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H085_W15_mean <- Max_Kfloat_H085_W15 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+ggarrange(passive_eddy, min_eddy, mean_eddy, max_eddy)
 ```

### 2020-11-21 thesis_chapter_2.Rmd
chapter 2 updates

@@
 library(ggplot2)
-library(ggord)
+#library(ggord)
 library(vegan)
@@
 
-wp_1 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6A.jpg")))
+wp_1 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6A.jpg"))) #figure missing, copy of fig_6B used
 wp_2 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6B.jpg")))

### 2020-11-21 thesis_chapter_2.Rmd
abstract updated

@@
 tables: yes
-abstract:  '**Abstract**: Macroalgal morphological variation is determined to a large extent by
-  a combination of environmental factors, with wave exposure and temperature
-  perhaps the main influences as they are key environmental properties to which a species becomes locally adapted.
-Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction
-  in overall size and strength increasing traits. In terms of temperature, warmer
-  environments have been shown to reduce the overall size of resident and transplanted
-  species. However, none of the past studies have identified specific wave and temperature
-  metrics responsible for the morphological adaptation macroalgae exhibit. Past research
-  has often used simple or two-dimensional models of wave exposure, which do not
-  take into account important aspects of the nearshore environment such as wave breaking,
-  refraction and diffraction. Furthermore, past studies have often used satellite-derived
-  datasets as sources for temperature data; however, such data have been shown to have large bias when 
-  applied to the nearshore environment. This study used *in situ* temperature data and wave power
-  metrics calculated from a 3D-numerical model to identify specific temperature and wave metrics
-  responsible for morphological adaptation of the kelp, *Ecklonia maxima* and *Laminaria pallida*.
-The results identify wave      exposure as the main driver of morphological adaptation while identifying specific wave metrics.
-  Furthermore, the results show differences in wave metrics between species; and between
-  deep and shallow populations.'
+abstract:  '**Abstract**: Macroalgal morphological variation is determined to a large extent by a combination of environmental factors, with wave exposure and temperature perhaps the main influences, as they are key environmental properties to which a species becomes locally adapted.
+Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction in overall size and strength increasing traits.
+In terms of temperature, warmer environments have been shown to reduce the overall size of resident and transplanted species.
+However, none of the past studies have identified specific wave and temperature metrics responsible for the morphological adaptation macroalgae exhibit.
+Past research has often used simple or two-dimensional models of wave exposure, which do not take into account important aspects of the nearshore environment such as wave breaking, refraction and diffraction.
+Furthermore, past studies have often used satellite-derived datasets as sources for temperature data, however, such data have been shown to have large bias when applied to the nearshore environment.
+This study used in situ temperature data and wave power metrics calculated from a 3D-numerical model to identify specific temperature and wave metrics responsible for morphological adaptation of the kelp, Ecklonia maxima and Laminaria pallida.
+Between temperature and wave exposure, the results identify wave exposure as the main influencer of morphological adaptation while identifying specific wave metrics.
+Furthermore, the results show differences in wave metrics between species; and between deep and shallow populations.'
 

### 2020-11-21 thesis_chapter_2.Rmd
intro updated to latest

@@
 
-Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera [@Bolton2010], form the basis of one of the most productive ecosystems globally [@Mann1973].
+Kelps are a group of large seaweeds of the order Laminariales (Ochrophyta), which despite their relatively low taxonomic diversity of species in genera [@Bolton2010], form the basis of one of the most productive ecosystems globally [@Mann1973; @Krumhansl2012].
 Kelps generally have a dependence on cool-temperate and arctic seawater temperatures [@Bolton2010; @Mohring2014-bk; @Cavanaugh2011-ib; @Dayton1985], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
-Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and sub-canopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002], each depending on a wide suite of ecological services provided by kelp forests [@Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @Willis2003-xo; @anderson1997; @Dayton1985].
-Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in their distribution [@Gorman2013-nj], abundance [@Cavanaugh2011-ib; @Dayton1998-xw], diversity [@Wing2007-tv; @Wernberg2008], composition [@Leliaert2000a; @Dayton1985; @Norderhaug2012-ks; @Harley2012], growth [@Cousens1982-ck] and productivity [@Pedersen2012-vr; @Dayton1998-xw; @Krumhansl2012].
-
-Given that kelps are cool-temperate organisms which are vulnerable to dislodgement under high wave exposure scenarios, increases in storm frequency and magnitude, and changes in ocean temperature, changes in these influences pose a direct threat to their survival; which is a concern considering that these variables are known to be affected by climate change[@Russo2014; @Reguero2019; @Weisse2010; @Rose2012; @Rummukainen2012].
+Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and sub-canopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002], each depending on a wide suite of ecological services provided by kelp forests [@Dayton1985; @Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @Anderson1997; @Willis2003-xo].
+Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in their distribution [@Gorman2013-nj], abundance [@Dayton1998; @Cavanaugh2011-ib], diversity [@Wing2007-tv; @Wernberg2008], composition [@Dayton1985; @Leliaert2000a; @Harley2012; @Norderhaug2012-ks], growth [@Cousens1982-ck] and productivity [@Dayton1998;@Krumhansl2012; @Pedersen2012].
+
+Given that kelps are cool-temperate organisms which are vulnerable to dislodgement under high wave exposure scenarios, increases in storm frequency and magnitude, and changes in ocean temperature, pose a direct threat to their survival; which is a concern considering that these variables are known to be affected by climate change [@Russo2014; @Reguero2019; @Weisse2010; @Rose2012; @Rummukainen2012].
 Warmer temperatures reduces the resilience of kelp individuals to storm disturbance [@Wernberg2010b], causes fragmentation through weakening tissues [@Simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
 Despite these threats and disturbances kelps persist across a broad range of environments which is largely due to their morphological plasticity [@FowlerWalker2006; @Wernberg2010a].
-A study by @Wernberg2003-mp investigated the morphology of *Ecklonia radiata* (C.
-Agardh) J.
-Agardh.
-in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australian coast.
+A study by [@Wernberg2003-mp] investigated the morphology of Ecklonia radiata (C.Agardh) J.Agardh in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australian coast.
 They found no correlation between spatial distance and morphological similarity; rather the morphology of the kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales [@Wernberg2003-mp].
-Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
-Deep and shallow water environments may differ in abiotic processes such as the degree of water column mixing [@Smit2013a], solar heating [@Dellatorre2012; @Dunne2001] and effects of wave dampening [@Kobayashi1993-os; @Mia2013; @Dubi1995].
-
-Although multiple environmental factors play a role in kelp morphological adaptation, wave exposure and temperature have been identified as the main drivers across various kelp species [@bekkby2014; @Wernberg2005c; @Denny2002; @Hurd2000; @FowlerWalker2005; @Thomsen2004; @mabin2013].
-For example, a study by @Serisawa2002-dc compared the morphology of *Ecklonia cava* Kjellman growing at warmer and cooler sites, and showed wrinkling of the blade and reduced size to be a characteristic of higher temperature ranges.
+
+Although multiple environmental factors play a role in kelp morphological adaptation, wave exposure and temperature have been identified as the main influencers across various kelp species [@Hurd2000;@Denny2002;@Thomsen2004;@FowlerWalker2005;@Wernberg2005c;@mabin2013;@Bekkby2014].
+For example, a study by  @Serisawa2002-d compared the morphology of Ecklonia cava Kjellman growing at warmer and cooler sites, and showed wrinkling of the blade and reduced size to be a characteristic of higher temperature ranges.
 Warmer temperature affects important kelp physiological processes such as photosynthesis and respiration, which influences growth and productivity and ultimately leads to a reduction in size [@bearham2013; @gerard1997; @gao2013; @zimmerman1986].
 Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002; @Hurd2000; @FowlerWalker2005; @Thomsen2004].
-For example, a study by @Wernberg2005c examined the consistency of wave exposure as a driver of *E.
-radiata* across a broad geographic range, and showed trends towards drag-reducing (small size, narrow laterals and blades, low spinosity) and increased strength (large holdfast, thick stipe and thick blades and lamina) at high wave exposure sites.
+For example, a study by @Wernberg2005c examined the consistency of wave exposure as a driver of E.
+radiata across a broad geographic range, and showed trends toward drag-reducing (small size, narrow laterals and blades, low spinosity) and increased strength (large holdfast, thick stipe and thick blades and lamina) at high wave exposure sites.
 
 Measures of wave exposure in ecological studies often incorporate integrative measures of hydrodynamic conditions at a particular site which are based on cartographical models of exposure.
-Cartographical models of wave exposure are 'fetch-based-models' which measure the length of open water associated with a particular site in a straight line, and are regarded as simple measures of wave exposure.
-Advances in numerical modelling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitative, reproducible approach for characterising the hydrodynamic environment.
-Although @mabin2013 investigated kelp morphological characteristics using a 2D WAM model, currently no studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modelling.
+Cartographical models of wave exposure are “fetch-based-models” which measure the length of open water associated with a particular site in a straight line, and are regarded as simple measures of wave exposure.
+Advances in numerical modeling based on physical/linear wave theory incorporate more complexity (wind forcing, wave-wave interactions, wave breaking, diffraction and variation in wave direction) into the models and allow for a quantitative, reproducible approach for characterizing the hydrodynamic environment.
+Although @mabin2013 investigated kelp morphological characteristics using a 2D WAM model, currently no studies investigating macroalgal morphological characteristics in relation to the wave environment have incorporated 3D spectral numerical modeling.
 Furthermore no studies have identified specific wave metrics that drive kelp morphological characteristics, which the current work will aim to achieve.
 Other considerations such as possible differences in morphology between sub-canopy and canopy species, and shallow and deep water kelp populations have not been considered using advanced numerical and statistical tools.
+Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
+Deep and shallow water environments may differ in abiotic processes such as the degree of water column mixing [@Smit2013a], solar heating [@Dellatorre2012; @Dunne2001], and effects of wave dampening [@Kobayashi1993-os; @Mia2013; @Dubi1995].
 
@@
 maxima* forms dense surface canopies in the shallow sub-littoral.
-Therefore, *E.
-maxima* may be exposed to variations in wind driven surface waves, swell waves and temperature, while *L.
-pallida* is exposed to variations in swell waves and temperature.
+Therefore, E.
+maxima may be exposed to variations in wind driven surface waves, swell waves and temperature, while *L.
+pallida* is exposed to variations in swell waves, wave exposure and temperature due to differences in depth.
 @Molloy1996-la investigated the effect of depth and wave exposure on *L.
-pallida* at different depths and wave exposure and showed that depth had a greater effect than wave exposure when considering all the morphological characteristics; when considering individual characteristics, however, wave exposure had the most significant effect on blade thickness.
+pallida* and showed that depth had a greater effect than wave exposure when considering all the morphological characteristics; when considering individual characteristics, however, wave exposure had the most significant effect on blade thickness.
 Another study by @Rothman2017-pn investigated the changes in morphology in shallow populations of *L.
@@
 The current study will also aim to identify specific temperature and wave metrics which drive kelp morphological characteristics.
-This will be achieved by initially understanding the variation in temperature and waves using an advanced numerical model.
+This will be achieved by initially understanding the variation in temperature and waves using a numerical model.
 The numerical model will then be used as a basis for investigating the consequences for morphological characteristics of *E.

### 2020-11-21 thesis_chapter_2.Rmd
abstract error

--- 2020-11-21 thesis_chapter_2.Rmd
@@
 tables: yes
-abstract:  '**Abstract**: Macroalgal morphological variation is determined to a large extent by a combination of environmental factors, with wave exposure and temperature perhaps the main influences, as they are key environmental properties to which a species becomes locally adapted.
+abstract: Macroalgal morphological variation is determined to a large extent by a combination of environmental factors, with wave exposure and temperature perhaps the main influences, as they are key environmental properties to which a species becomes locally adapted.
 Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction in overall size and strength increasing traits.

### 2020-11-21 thesis_chapter_2.Rmd
abstract error on github

--- 2020-11-21 thesis_chapter_2.Rmd
@@
 tables: yes
-abstract:  '**Abstract**: Macroalgal morphological variation is determined to a large extent by a combination of environmental factors, with wave exposure and temperature perhaps the main influences, as they are key environmental properties to which a species becomes locally adapted.
+abstract:  'Macroalgal morphological variation is determined to a large extent by a combination of environmental factors, with wave exposure and temperature perhaps the main influences, as they are key environmental properties to which a species becomes locally adapted.
 Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction in overall size and strength increasing traits.

### 2020-11-21 doodle.R
commiting all

@@ -1,0 +1,29 @@
+library(gganimate)
+library(tidync)
+library(tidyverse)
+library(ncdf4)
+
+
+climate_filepath <- paste0("OceanParcels/trajectory_data/Mean_KFloat_H100_W00.nc")
+climate_output <- nc_open(climate_filepath)
+
+
+p_eddy <- tidync("OceanParcels/trajectory_data/PassiveEddy.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+p_eddy_mean <- p_eddy %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+p_eddy$type <- "Passive Eddy" # add extra column 'type' to dataframe
+
+
+
+
+
+
+
+
+

### 2020-11-21 thesis_chapter_2.Rmd
intro updated

@@
 tables: yes
-abstract:  'Macroalgal morphological variation is determined to a large extent by a combination of environmental factors, with wave exposure and temperature perhaps the main influences, as they are key environmental properties to which a species becomes locally adapted.
+abstract: Macroalgal morphological variation is determined to a large extent by a combination of environmental factors, with wave exposure and temperature perhaps the main influences, as they are key environmental properties to which a species becomes locally adapted.
 Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction in overall size and strength increasing traits.

### 2020-11-21 thesis_chapter_2.Rmd
methods updated

@@
 
-Sites were chosen to represent an array of temperature and wave gradients, as indicated in figure \ref{fig:map}.
+Sites were chosen to represent an array of temperature and wave gradients, as indicated in Figure \ref{fig:map}.
 St.
-Helena Bay and Betty's Bay constituted the north-western and south-eastern boundary sites, respectively.
+Helena Bay and Betty’s Bay constituted the north-western and south-eastern boundary sites, respectively.
 These sites are roughly 300 km apart, lie within separate marine provinces  @Smit2017, and span the majority of the south-west coast, in varying thermal and wave energy regimes.
@@
 East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone [@Smit2013a].
-The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm-temperate [@Smit2013a].
+The Agulhas Marine Province is characterized by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm-temperate [@Smit2013a].
 
@@
 
-The annual and seasonal patterns in both wind influence the thermal regime around the coastline by inducing upwelling.
+The annual and seasonal patterns in both wind and swell direction influence the thermal regime around the coastline by inducing upwelling.
 Upwelling brings deep, cool nutrient-rich water to coastal areas which causes a decrease in temperature [@Field1980a; @Andrews1980; @Gill1974-nt; @Cram1970-jw; @Rouault2010-gi; @Blanke2002].
@@
 This is not the case for False Bay, which is shielded from the dominating wind and swell that the Cape Peninsula is subjected to.
-In the winter months, upwelling becomes more favourable in False Bay due to the change in swell and wind direction, causing lower temperatures to dominate during this time [@Rouault2010-gi; @Andrews1980].
+In the summer months, upwelling becomes more favorable in False Bay due to the change in swell and wind direction, causing lower temperatures to dominate during this time [@Rouault2010-gi; @Andrews1980].
 Due to the Cape Peninsula temperate latitude, winter months bring an increased frequency of frontal depressions that originate from the Southern Ocean [@Reason2006].
@@
 This is, however, absent in winter, when strong northerly winds are prevalent from St.
-Helena Bay to Betty's Bay [@Field1980a; @Jury1985-nr; @Andrews1980; @Jury1980-ov].
+Helena Bay to Betty’s Bay [@Field1980a; @Jury1985-nr; @Andrews1980; @Jury1980-ov].
 
@@
 
-Shallow water temperatures were sourced from The South African Coastal Temperature Network (SACTN) website (https://github.com/ajsmit/SACTN).
-In terms of nearshore temperature, *in situ* data are preferred over satellite SST, which have shown to exhibit large biases [@Smit2013a].
-Linear interpolated SST were calculated for sites where *in situ* recorders were absent.
-These data were used to group sites into 'Cool temperate' and 'Warm temperate' categorisations to investigate the possible effect of thermal regime on kelp morphological characteristics.
+Shallow water temperatures were sourced from The South African Coastal Temperature Network (SACTN) website1.
+In terms of nearshore temperature, in situ data are preferred over satellite sea surface temperature (SST), which have shown to exhibit large biases [@Smit2013a].
+The underwater temperature recorders (UTRs) are attached to concrete mooring blocks or railway line sections deployed in the shallow at ∼3 m depth.
+The UTRs were comprised of Starmon Mini recorders (Star-Oddi, Reykjavik, Iceland/Onset Hobo® U22-001) accurate to ±0.2°C.
+All in situ temperature data were reduced to a time-coded, continuous series of monthly means over various time-scales [@Smit2013a].
+It was decided that the annual time-scale would be used for analysis.
+Linear interpolated SST were calculated for sites where in situ recorders were absent.
+These data were used to group sites into “Cool temperate” and “Warm temperate” categorizations to investigate the possible effect of thermal regime on kelp morphological characteristics.
 
@@
 The South African wave climate was modeled via 20 spectral numerical wave models that simulated the offshore wave climate to the nearshore.
-The boundary conditions of these models were obtained by using the NOAA Wave Watch III (WWIII) model output, distributed via the National Centers for Environmental Prediction (NCEP) product (Office of the Director 2000; Environmental Modeling Center / Marine Modeling Branch 2005).
-The particular hindcast product utilised during the DEA-CSIR study spans 1994-2013 at a 3-hour resolution.
+The boundary conditions of these models were obtained by using the NOAA Wave Watch III (WWIII) model output, distributed via the National Centers for Environmental Prediction (NCEP) product (Office of the Director 2000; Environmental Modeling Center/Marine Modeling Branch 2005).
+The particular hindcast product utilized during the DEA-CSIR study spans 1994–2013 at a 3-hour resolution.
 These data were then used to model swell propagation into the coastal models while wind waves (seas) were generated via stationary computations in the Simulating Waves in the Nearshore SWAN model [@Booij1997-nx].
@@
 SWAN allows one to extract wave variables from specific gridded locations in the nearshore.
-For False Bay, a resolution of 200 meters was modelled and output produced at both the 7 meter and 15 meter isobaths.
-A 200 meter resolution was used as the False Bay computational grid was nested within a larger 1 kilometer resolution grid.
+For False Bay, a resolution of 200 meters was modeled and output produced at both the 7 m and 15 m isobaths.
+A 200 m resolution was used as the False Bay computational grid was nested within a larger 1 km resolution grid.
 This allowed for a computational effective wave resolution of increasing resolution, from the NCEP, low resolution output to nested, high resolution coastal output.
-For Table Bay and east of Cape Hangklip the resolution was 500 meters and also had output at the 7 meters and 15 meter isobaths.
+For Table Bay and east of Cape Hangklip the resolution was 500 m and also had output at the 7 and 15 m isobaths.
 These contour outputs were chosen in the original study by the CSIR as most engineering run-up calculations require wave parameter information at these contour depths and were the main focus of the original study.
-For this study the 7 meter contours were used.
+For this study the 7 m contours were used.
 These data were then used to calculate over all wave power (kW/m) as this was considered the best measure of overall wave exposure.
-Annual wave power was plotted and categorised into different wave exposure categories which ranged from fully sheltered to extremely exposed.
+Annual wave power was plotted and categorized into different wave exposure categories which ranged from fully sheltered to extremely exposed.
 
@@
 The morphological characteristics of both species are presented in figure \ref{fig:schematic}.
-*Laminaria pallida* is characterised by a single smooth blade which is divided longitudinally into sections, and develops from a single meristematic region located at the junction between the blade and the stipe [@Dyer2018-rv].
+*Laminaria pallida* is characterized by a single smooth blade which is divided longitudinally into sections, and develops from a single meristematic region located at the junction between the blade and the stipe [@Dyer2018-rv].
 This species has a solid stipe along the south-west coast but develops a hollow stipe along the west coast northward and into Namibia [@Rothman2017-pn].
@@
 
-Morphological measurements of *L.
-pallida* and *E.
-maxima* were collected at 18 sites along the Western Cape coast of South Africa (Fig.
-1) between October 2014 and April 2015.
+Morphological measurements of L.
+pallida and E.
+maxima were collected at 18 sites along the Western Cape coast of South Africa (Figure \ref{fig:map}) between October 2014 and April 2015.
+AlgaeBase was used to confirm the latest taxonomic associations for each species [@Guiry2019].
+Divers sampled kelp individuals by means of SCUBA for deeper populations and freediving for shallower populations.
+The sampled individuals were then taken ashore and the relevant morphological measurements taken.
 In each instance only the 13 largest individuals were sampled to ensure that only mature and fully grown sporophytes were measured.
+The age of *E.
+maxima* was determined by size, presence at the surface as well as a darker coloration of the stipe; as this occurs with age.
+In the case of *L.
+pallida*, the individuals with the longest stipe were selected which is considered a reasonable estimate of age.
+In each case the age of the individuals was estimated visually before the individual was sampled.
 In instances where one of the species was not present a site, the nearest population of the missing species was sampled.
-Because the macroalgae differ in morphological features, species-specific morphological characteristics were included and are presented in Tables 1 and 2.
-Between February 2017 and November 2018, morphological measurements for *E.
-maxima* individuals in shallow water (<1m) at seven sites along the Western Cape coast of South Africa were also collected.
-The same morphological characteristic measurements were taken as for the deeper *E.
-maxima*, and this allowed comparison between morphological characteristics between deep and shallow individuals within sites.
-Measurements were not collected for *L.
-pallida* in the shallow depths, as this species is largely absent from the shallow in this portion of the South African coastline.
+Because the macroalgae differ in morphological features, species-specific morphological characteristics were included and are presented in Table 1.
+The stipe length for *E.
+maxima* was considered to be the portion of the plant just below the pneumatocyst and just above the holdfast.
+Stipe circumference was measured just above the holdfast for both species.
+Between February 2017 and November 2018, morphological measurements for E.
+maxima individuals in shallow water (<1 m) at seven sites along the Western Cape coast of South Africa were also collected.
+The same morphological characteristic measurements were taken as for the deeper E.
+maxima, and this allowed comparison between morphological characteristics between deep and shallow individuals within sites.
+Measurements were not collected for L.
+pallida in the shallow depths, as this species is largely absent from the shallow in this portion of the South African coastline.
 
@@
 
-Summary statistics were calculated for each site and region for each of the environmental variables considered in this study, and are presented in Table 3.
+Summary statistics were calculated for each site and region for each of the environmental variables considered in this study, and are presented in Table 2.
+The associated abbreviations used are also presented in Table 2.
 In addition, annual median wave and wind direction were also calculated.
@@
 Morphological characteristics in relation to temperature and wave exposure were investigated using non-parametric statistical methods.
-Morphological characteristics were categorised according to temperature regime and wave exposure category for each species and population of kelp in this study.
-Significant differences between temperature regimes and across wave exposure categories were tested with Wilcox tests and Kruskal tests, respectively, and notched boxplots allowed comparison between individual categories of wave exposure.
-In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis (db-RDA) was performed using the *rda* function in the *vegan* software package [@Oksanen2015-vj] in R (R Core Team, 2019).
+Morphological characteristics were placed into categories according to temperature regime and wave exposure class for each species and population of kelp in this study.
+The appropriate category was determined by kelp’s sampling location along the coastline and the according wave exposure and temperature category assigned to that species and/or population.
+The R software package ggplot2 was used for graphical output and non-parametric analyses.
+Significant differences between temperature regimes and across wave exposure categories were tested with Wilcox tests and Kruskal -Wallis tests, respectively, and notched boxplots allowed comparison between individual categories of wave exposure.
+The alpha level used to define significance was p = 0.05.
+Boxplots were created using the software package ggplot2 [@Wickham2009].
+In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis (db-RDA) was performed using the rda function in the vegan software package [@Oksanen2015-vj] in R (R Core Team, 2014; RStudio Team, 2020).
 The abiotic data covered all study sites for both species of kelp.
-Both morphology and abiotic data (temperature, waves and wind) were standardised using the *decostand* function in the *vegan* software package; morphology data were used as response variables and the abiotic data as explanatory variables.
+Both morphology and abiotic data (temperature, waves and wind) were standardized using the decostand function in the vegan software package; morphology data were used as response variables and the abiotic data as explanatory variables.
+A list of the abbreviations used and the associated metrics are presented in Table 2.
 To determine the explanatory variables that best describe patterns in the response data, a full RDA was performed using a complete set of explanatory variables.
@@
 To further improve the model, pairwise coefficients and Variance Inflation Factor (VIF) were calculated to identify variables with high multicollinearity.
-The computation of the parsimonious RDA was followed by permutation tests of the adjusted *R*^2^ to assess significance of constraints.
+The computation of the parsimonious RDA was followed by permutation tests of the adjusted R2 to assess significance of constraints.
 

### 2020-11-21 thesis_chapter_2.Rmd
italics error

@@
 maxima individuals in shallow water (<1 m) at seven sites along the Western Cape coast of South Africa were also collected.
-The same morphological characteristic measurements were taken as for the deeper E.
-maxima, and this allowed comparison between morphological characteristics between deep and shallow individuals within sites.
-Measurements were not collected for L.
-pallida in the shallow depths, as this species is largely absent from the shallow in this portion of the South African coastline.
+The same morphological characteristic measurements were taken as for the deeper *E.
+maxima*, and this allowed comparison between morphological characteristics between deep and shallow individuals within sites.
+Measurements were not collected for *L.
+pallida* in the shallow depths, as this species is largely absent from the shallow in this portion of the South African coastline.
 

### 2020-11-21 thesis_chapter_2.Rmd
results updated

@@
 
-The mean annual coastal water temperature for the study sites located on the western side of the peninsula ranged from 14.5 ± 0.9 ℃ (mean ± SD) at St.
-Helena Bay, the most northern site on the western side of the peninsula, to 14.6 ± 0.3 ℃ at Olifantsbos, the most southern site on the western side of the peninsula (see figure \ref{fig:Temperature variables plot}A).
-For sites located on the eastern side of the peninsula (within False Bay), the annual mean coastal water temperatures ranged from 15.5 ± 0.9 ℃ at Buffelsbaai to 15.0 ± 0.9 ℃ at Betty’s Bay, the most eastern site on the coastline in this study (see figure\ref{fig:Temperature variables plot}A).
-Oudekraal had the lowest annual mean coastal water temperature of 12.3 ± 0.9.
-Annual maximum coastal temperatures for study sites located on the western side of the peninsula ranged from 16.3 ℃ at St.
-Helena Bay to 15.1 ℃ at Olifantsbos.
-For sites located east of the peninsula, annual maximum coastal water temperatures ranged from 16 ℃ at Buffelsbaai to 16.5 ℃ at Betty’s Bay.
-Amongst all sites, Kalk Bay had the highest annual maximum temperature of 19 ℃.
-Annual minimum coastal water temperatures for the study sites on the western side of the peninsula ranged from 12.9 ℃ at St.
-Helena Bay to 14.1 ℃ at Olifantsbos, and on the western side of the peninsula ranged from 15.0 ℃ at Buffelsbaai to 13.6 ℃ at Betty’s Bay.
-The annual range in temperature on the western side on the peninsula was 3.4 at St.
-Helena Bay to 0.9 at Olifantsbos, and on the eastern side on of the peninsula was 1.0 at Buffelsbaai to 2.9 at Betty’s Bay.
+The mean annual coastal water temperature for the study sites located on the western side of the peninsula ranged from 14.5 ± 0.9°C (mean ± SD) at St.
+Helena Bay, the most northern site on the western side of the peninsula, to 14.6 ± 0.3°C at Olifantsbos, the most southern site on the western side of the peninsula (see Figure \ref{fig:Temperature variables plot}A).
+For sites located on the eastern side of the peninsula (within False Bay), the annual mean coastal water temperatures ranged from 15.5 ± 0.9°C at Buffelsbaai to 15.0 ± 0.9°C at Betty’s Bay, the most eastern site on the coastline in this study (see Figure \ref{fig:Temperature variables plot}A).
 
@@
 ```
-
-In general, the mean temperature increases for sites located within False Bay with less variation around the mean temperature compared to sites on the western side of the peninsula(\ref{fig:Temperature variables plot}A).
-This is verified by figure\ref{fig:Temperature variables plot}B, which shows a significantly lower (*p* < 0.05, figure \ref{fig:Temperature variables plot}B) mean temperature and less variation in temperature for sites located on the western side of the peninsula compared to sites on the eastern side.
+In general, the mean temperature increases for sites located within False Bay with less variation around the mean temperature compared to sites on the western side of the peninsula (see Figure \ref{fig:Temperature variables plot}A).
+This is verified by statistical difference based on region, which shows a significantly lower (p < 0.05, Figure \ref{fig:Temperature variables plot}B) mean temperature and less variation in temperature for sites located on the western side of the peninsula compared to sites on the eastern side.
 
@@
 
-Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula on both an annual scale (see figure \ref{fig:wave variable plot}).
+Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula (see Figure \ref{fig:wave variable plot})A).
 Mean significant wave height ranged from 0.5 ± 0.2 m (mean ± SD) at St.
-Helena Bay to 2.3  ± 0.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side of the peninsula it ranged from 0.9 ± 0.4 m at Buffelsbaai to 1.9 ± 0.6 m at Betty’s Bay (see figure \ref{fig:wave variable plot}A).
+Helena Bay to 2.3 ± 0.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side of the peninsula it ranged from 0.9 ± 0.4 m at Buffelsbaai to 1.9 ± 0.6 m at Betty’s Bay (see Figure \ref{fig:wave variable plot})A).
 The maximum significant wave height ranged from 1.4 m at St.
@@
 Mean peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St.
-Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay (see figure \ref{fig:wave variable plot}B).
+Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay (see Figure \ref{fig:wave variable plot})B).
 The data show no trend in mean peak period for the coastline.
@@
 When wave variables were grouped by region significant differences were found for both mean significant swell height and mean peak period.
-The western region experiences a significantly higher mean significant swell height (*p* < 0.05; see figure \ref{fig:wave variable plot}C) and mean peak period (*p* < 0.05; see figure \ref{fig:wave variable plot}D).
+The western region experiences a significantly higher mean significant swell height (p < 0.05; see Figure \ref{fig:wave variable plot})C) and mean peak period (p < 0.05; see Figure \ref{fig:wave variable plot})D).
+
+Data for median wave direction showed no trend across timescales, however, the variation (SD) in wave direction is lower on the western side of the peninsula compared to the eastern side both annually and for winter.
+Summer data were the exception and showed lower variation for the eastern side of the peninsula (see Appendix).
+
+In Figure 5A the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
+The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits (see Figure \ref{fig:wave power map}B).
+The western periphery of the Cape Peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures (see Figure \ref{fig:wave power map}).
+There the marked seasonality, with higher energy waves during winter, may be clearly observed once more.
 
@@
 ```
-
-In figure \ref{fig:wave power map} the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length).
-The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits.
-The western periphery of the Cape Peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures (see figure \ref{fig:wave power map}).
-There the marked seasonality, with higher energy waves during winter, may be clearly observed once more.
+To clarify the averaged wave exposure map presented in Figure \ref{fig:wave power map}A, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in Figure \ref{fig:wave power map}C.
+Tracing the wave height contours into False Bay its clear why this bay’s western periphery is predominantly sheltered.
 
@@
 ```
-
-To clarify the averaged wave exposure maps presented in figure \ref{fig:wave power map}, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in figure \ref{fig:FB spectrum}.
-Tracing the wave height contours into False Bay its clear why this bay’s western periphery is predominantly sheltered.
-
 ## Drivers of kelp morphological characteristics
@@
 Significant differences were found between cool temperate and warm temperate regimes for most *L.
-pallida* morphological characteristics with lamina length, number of digits and thallus mass the exceptions (see figure \ref{fig:lamanaria temp plots}).
-Certain morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significantly higher means for kelp from the cool temperate regime when compared to kelp from the warm temperate regime.
+pallida* morphological characteristics with lamina length, number of digits and thallus mass the exceptions (see Figure 6).
+Certain morphological characteristics such as stipe mass, stipe length, total length and stipe diameter had significantly higher means for kelp from the cool temperate regime when compared to kelp from the warm temperate regime (p < 0.05; see Figure \ref{fig:lamanaria temp plots}).
 Conversely, lamina weight and lamina thickness had significantly lower means for kelp from the cool temperate regime compared to kelp from the warm temperate regime.
 When *L.
-pallida* morphological characteristics were grouped to wave exposure categories Kruskal-Wallis tests revealed significant differences amongst categories for most morphological characteristics with lamina length, number of digits and thallus mass the exceptions (see figure ref\{fig:lamanaria wave plots}).
+pallida* morphological characteristics were grouped to wave exposure categories Kruskal-Wallis tests revealed significant differences amongst categories for most morphological characteristics with lamina length, number of digits and thallus mass the exceptions (see Figure ref\{fig:lamanaria wave plots}).
 The morphological characteristics that were significantly different amongst sites exhibited similar patterns, with increasing mean values from the fully sheltered to extremely sheltered categories and a decrease in mean values for the remainder of the categories.
@@
 maxima* morphological characteristics showed differences between cool temperate and warm temperate regime except for frond mass, frond length, stipe mass and epiphyte length.
-Mean values for primary length, primary width and stipe circumference were significantly higher for kelp from the cool temperate regime than for kelp from the warm temperate regime (see figure \ref{fig:{deep eck temp plot}}.
-The remaining morphological characteristics, stipe length, number of tufts and total length had significantly lower means for kelp from cool temperate compared to kelp from the warm temperate regime (*p* < 0.05; see figure \ref{fig:deep eck temp plot}).
-Significant differences amongst wave exposure categories were found for all morphological characteristics of deep *E.
-maxima* populations.
+Mean values for primary length, primary width and stipe circumference were significantly higher for kelp from the cool temperate regime than for kelp from the warm temperate regime (see Figure \ref{fig:{deep eck temp plot}}).
+The remaining morphological characteristics, stipe length, number of tufts and total length had significantly lower means for kelp from cool temperate compared to kelp from the warm temperate regime (p < 0.05; see Figure 8).
+Significant differences amongst wave exposure categories were found for all morphological characteristics of deep E.
+maxima populations (see Figure \ref{fig:deep eck temp plot}
+).
 Certain morphological characteristics showed a gradual increase in mean value with increasing degree of wave exposure.
@@
 ```
-
 When grouped by temperature regime, shallow *E.
-maxima* exhibited significant differences between temperature regimes for frond mass, primary length, primary width, frond length, stipe mass, stipe length, stipe circumference and total length (*p* < 0.05; see figure \ref{fig:shallow ecklonia temp}).
+maxima* exhibited significant differences between temperature regimes for frond mass, primary length, primary width, frond length, stipe mass, stipe length, stipe circumference and total length (p < 0.05; see Figure \ref{fig:shallow ecklonia temp}).
 All of the significantly different morphological characteristics, except primary length, had lower values for the western region compared to the eastern region.
-When shallow *E.
-maxima* were grouped by wave exposure category significant difference were found for all morphological characteristics (*p* < 0.05; see figure \ref{fig:shallow ecklonia wave}).
+When shallow E.
+maxima were grouped by wave exposure category significant difference were found for all morphological characteristics (p < 0.05; see Figure \ref{fig:shallow ecklonia wave}).
 In general, values of the various morphological characteristics increase for the lower wave exposure categories, followed by a sharp decrease and a gradual increase in value.
@@
 
-Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of *L.
-pallida* morphological characteristics with an adjusted *R*2 of 0.49, explaining 72% of the variation (global permutation test on final model: d.f = 1, *F* = 8.51, *p* = 0.008).
+Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of L.
+pallida morphological characteristics with an adjusted R2 of 0.49, explaining 72% of the variation (global permutation test on final model: d.f = 1, F = 8.51, p = 0.008; Figure \ref{fig:rda plots}A).
 RDA1 was the only significant axis in the model and explained 48% of the variation while RDA2 was not significant and explained 63% of the variation.
-Total length, stipe length, stipe diameter and stipe mass were positively influenced (*i.e.* increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* for kelp mostly from the western region and negatively influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp* for kelp from the eastern region.
-Although sites from from the western region did not cluster as closely as eastern region study sites, grouping according to region was still evident.
+Total length, stipe length, stipe diameter and stipe mass were positively influenced (i.e., increased size corresponding with the environmental driver) by mean significant wave height and median swell direction for kelp mostly from the western region and negatively influenced (i.e., decreased size corresponding with the environmental driver) by peak period standard deviation and mean temperature for kelp from the eastern region.
+Although sites from the western region did not cluster as closely as eastern region study sites, grouping according to region was still evident.
 
@@
 ```
-
-Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of deep *E.
-maxima* morphological characteristics with an adjusted *R*^2^ of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, *F* = 2.4, *p* = 0.026).
-RDA1 was the only significant axis in the model (*p* = 0.03) and explained 72% of the variation.
-Primary length, primary width and stipe circumference were positively influenced by *hs_mean*, *tp_mean* and *dir_median* for for kelp from the western region.
-The remaining morphological characteristics were positivist influenced by *min_temp* and negatively influenced by *tp_sd* for kelp from the eastern region.
+The same approaches to identify the descriptors for L.
+pallida were employed for both populations of E.
+maxima.
+Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of deep E.
+maxima morphological characteristics with an adjusted R2 of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, F = 2.4, p = 0.026; Figure \ref{fig:rda plots}B).
+RDA1 was the only significant axis in the model (p = 0.03) and explained 72% of the variation.
+Primary length, primary width and stipe circumference were positively influenced by mean significant wave height, mean peak period and median swell direction for kelp from the western region.
+The remaining morphological characteristics were positively influenced by minimum temperature and negatively influenced by peak period standard deviation for kelp from the eastern region.
 
@@
 ```
-
-Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of shallow *E.
-maxima* morphological characteristics with an adjusted *R*^2^ of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, *F* = 2.41, *p* = 0.013).
-The model consisted of one significant axis, RDA1 (*p* = 0.02) and explained 62% of the variation.
-Stipe length, stipe mass and stipe circumference were positively influenced by *max_temp* and *mean_temp* along RDA1, and explained 50% and 55% of the variation respectively.
-The remaining morphological characteristics were positively influenced by *tp_max* and negatively influenced by *dirw_median* along RDA2 and explained 86% and 94% of the variation respectively.
+Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of shallow E.
+maxima morphological characteristics with an adjusted R2 of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, F = 2.41, p = 0.013, Figure \ref{fig:rda plots}C).
+The model consisted of one significant axis, RDA1 (p = 0.02) and explained 62% of the variation.
+Stipe length, stipe mass and stipe circumference were positively influenced by maximum temperature and mean temperature along RDA1, and explained 50 and 55% of the variation, respectively.
+The remaining morphological characteristics were positively influenced by maximum peak period and negatively influenced by median wind direction along RDA2 and explained 86 and 94% of the variation, respectively.
 There was overlapping of clusters according to region with no clear patterns in sites evident.

### 2020-11-21 thesis_chapter_2.Rmd
error in results section corrected

@@
 ```
-
-Data for median wave direction showed no trend across timescales; however, the variation (SD) in wave direction is lower on the western side of the peninsula compared to the eastern side both annually and for winter.
-Summer data were the exception and showed lower variation for the eastern side of the peninsula (see figure \ref{fig:wave rose}).
 

### 2020-11-21 thesis_chapter_2.Rmd
discussion and conclusion updated. figure to follow...

@@
 
-This study considered important drivers of the morphology of *Ecklonia maxima* and *Laminaria pallida* arpound tehe Cape Peninsula.
-These abiotic variables, namely temperature and wave exposure, have been identified in previous research [@FowlerWalker2006; @Miller2011-pm; @FowlerWalker2005; @Stewart2009; @Wing2007-tv; @Molloy1996-la; @Pedersen2012-vr; @Wernberg2005c; @Thomsen2004].
-The complex geomorphology of the western Cape coastline creates an ideal natural laboratory for studies of kelps in interaction with their environment.
-Unlike previous research, this study used a numerical model to provide the various variables of wave exposure.
+This study considered important drivers of the morphology of Ecklonia maxima and Laminaria pallida around the Cape Peninsula.
+These abiotic variables, namely temperature and wave exposure, have been identified in previous research on other brown algae species such as E.
+radiata and M.
+pyrifera [@Thomsen2004; @FowlerWalker2005;@FowlerWalker2006; @Wernberg2005c; @Wing2007-tv; @Stewart2009; @Miller2011-pm; @Pedersen2012-vr].
+The complex geomorphology of the Western Cape coastline creates an ideal natural laboratory for studies of kelps in interaction with their environment.
+Unlike previous research, this study used a complex numerical model to provide the various variables of wave exposure, which is based on linear wave theory to calculate overall wave exposure.
 We also considered wind in addition to temperature and wave variables, since wind is an important component of wind-driven surface gravity waves [@Holthuijsen2010-vl].
 The results show that specific variables of wave exposure are the main drivers of kelp morphology with temperature playing only a minor role.
-Furthermore, the investigation of differences in morphological characteristics between shallow and deep populations of *E.
-maxima* suggests that in low wave exposure environments the role of temperature as a morphological driver increases; this is particulary the case for False Bay.
+The morphological adaptation was also shown to be associated with the magnitude of wave exposure which had only been inferred in previous studies.
+Finally, the investigation of differences in morphological characteristics between shallow and deep populations of E.
+maxima suggests that in low wave exposure environments the role of temperature as a morphological driver increases; this is particularly the case for False Bay.
 
@@
 The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposure maps.
-It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions.
+It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating toward the continent from much more southerly directions.
 This results in positive and negative wave exposure anomalies all around the peninsula.
@@
 There are also differences between the types of waves that each side of the peninsula experiences.
-The wave analysis shows the western side of the peninsula is not only more exposed but the hydrodynamics are largely swell driven, while on the eastern side most of the wave energy is driven by a combination of swell and wind with the latter being the predominant driver.
-
 We provide strong support that variations in the environmental variables, particularly wave exposure variables, are driving kelp morphological characteristics around the Cape Peninsula.
@@
 maxima* populations, which had longer, thinner stipes for kelp from the warm temperate compared to kelp from the cool temperate.
-Reduction in certain morphological characteristics has been attributed to temperature by @Serisawa2002-dc in the kelp *E.
-cava*, which was smaller and shorter in warmer sites compared to cooler sites.
+Reduction in certain morphological characteristics has been attributed to temperature by @Serisawa2002-dc in the kelp E.
+cava, which was smaller and shorter in warmer sites compared to cooler sites.
 The reduction in size of adult may be a response to low nutrient conditions, which has been shown to reduce growth rate and overall morphology [@Simonson2015].
 Warmer temperatures are associated with low nutrient concentrations [@Waldron1992] and the low frequency of upwelling conditions in False Bay (low nutrient supply) coupled with warmer temperatures may be a contributing factor.
-It should be noted however, that from these analyses that in general the cool temperate region is more exposed to waves than that of the warm-temperate region.
+It should be noted, however, that from these analyses that in general the cool temperate region is more exposed to waves than that of the warm-temperate region.
 Therefore, the significantly larger morphological characteristics for kelp from the cool temperate region may overlap with responses to wave exposure.
 The response of kelp morphological characteristics to wave exposure was evident and both species exhibit tactics based on the magnitude of wave exposure.
-Strength increasing traits were exhibited for lower exposure levels while a 'go-with-the-flow' tactic for moderate levels of wave exposure.
+Strength increasing traits were exhibited for lower exposure levels while a “go-with-the-flow” tactic for moderate levels of wave exposure.
 This suggests that how morphological characteristics manifest themselves is dependent on the magnitude of wave exposure.
 When grouped by wave exposure category, *L.
-pallida* characteristics showed a significant increase in length for the 'extremely sheltered' category compared to the fully sheltered category, which may be a 'go-with-the-flow' tactic [@Denny1997; @Denny2002; @Hurd2000].
-Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny2002; @Hurd2000].
-However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under high wave exposure [@Denny2002; @Denny1997].
+pallida* characteristics showed a significant increase in length for the “extremely sheltered” category compared to the fully sheltered category, which may be a “go-with-the-flow” tactic [@Denny1997; @Denny2002; @Hurd2000].
+Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny1997; @Denny2002; @Hurd2000].
+However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under high wave exposure [@Denny1997; @Denny2002; @Hurd2000].
 This is reflected in the *L.
@@
 maxima* exhibits a size reducing tactic with increasing wave exposure by decreasing stipe mass, stipe length, stipe circumference, total length and frond length.
-A magnitude related response has been suggested by @Wernberg2010a who investigated the temporal and spatial variations of various environmental drivers of the kelp *E.
+A magnitude related response has been suggested by Wernberg and Vanderklift (2010) who investigated the temporal and spatial variations of various environmental drivers of the kelp *E.
 radiata*.
@@
 Furthermore, the morphological response of *E.
-maxima* to wave exposure only occurs at the 'Moderately exposed' (20-30 kW/m) level compared to *L.
+maxima* to wave exposure only occurs at the “Moderately exposed” level compared to *L.
 pallida* which exhibits a response at a lower wave exposure level.
@@
 Kelp morphology characteristics are largely wave driven for both species but differ in terms of specific temperature and wave metrics.
-For instance, both species morphological characteristics are determined by swell direction (*dir_median*), standard deviation in peak period (*tp_sd*) and mean significant swell height (*hs_mean*), but *L.
-pallida* temperature driver is mean temperature (*mean_temp*) and *E.
-maxima* by minimum temperature (*min_temp*).
-Minimum temperature as an important metric for *E.
-maxima* may be related to nutrient availability.
+
+The wave metrics identified in this study are important components of determining overall wave power, which will vary seasonally in the region based on the swell direction.
+The differences in temperature metrics between *E.
+maxima* and *L.
+pallida* point to underlying differences in nutrient availability.
 Low temperatures are often associated with upwelling events which bring cool, nutrient rich water into the nearshore.
 False Bay has comparatively little upwelling events compared to the western side of the peninsula, and so nutrients may be a limiting factor for *E.
-maxima* populations within False Bay, hence the identification of minimum temperature (*min_temp*) as a driver.
+maxima* populations within False Bay, hence the identification of minimum temperature as a influencer.
 Mean temperature identified as a driver for *L.
@@
 Solar heating of the water surface in combination with wind-driven transport causes fluctuations within the water column, which can occur daily or seasonally [@Kaplan2003].
-Vertical mixing of the water column for inshore regions are driven by several abiotic processes: (1) turbulence of breaking waves inside or outside of the surf zone (2) convective mixing through a combination of cooling and evaporation; and (3) wind driven currents and (4) tidal mixing [@Smit2013a].
+Vertical mixing of the water column for inshore regions are driven by several abiotic processes: (1) turbulence of breaking waves inside or outside of the surf zone; (2) convective mixing through a combination of cooling and evaporation; (3) wind driven currents; and (4) tidal mixing [@Smit2013a].
 These abiotic processes cause effective vertical mixing of surface and deeper water stratification which leads to a homogeneous thermal environment, which in this study may be interpreted as the mean temperature.
@@
 Since *E.
-maxima* is a canopy-kelp, it is  exposed to all components of a wave compared to *L.
+maxima* are a canopy-kelp, it is exposed to all components of a wave compared to *L.
 pallida* which occurs deeper in the water column.
 
-Temperature variables were not a significant driver of kelp morphology in this study.
 Although temperature plays an important role in distribution [@Bolton2010; @Miller2011-pm; @Rinde2014] of kelp and physiological functioning of adults and gametophytes [@Mohring2014-bk; @Steneck2002; @gerard1997; @bearham2013; @Smale2017], there is little evidence in this study that temperature is an important driver of morphological variation.
 However, we suggest that temperature plays a larger role in determining kelp morphology at sites where wave energy is low enough not to elicit a response in morphological characteristics.
-Temperature metrics influence stipe length and frond length in shallow *E.
-maxima* individuals located at sites within False Bay; which was not the case for deeper *E.
-maxima* individuals as well as *L.
+Temperature metrics influence stipe length and frond length in shallow E.
+maxima individuals located at sites within False Bay; which was not the case for deeper E.
+maxima individuals as well as *L.
 pallida* which were largely driven by wave variables.
@@
 The morphological variations in characteristics for deep *E.
-maxima* populations are not seen for shallow *E.
-maxima* populations.
-The lack of variation in shallow *E.
+maxima* populations are not seen for shallow E.
+maxima populations.
+The lack of variation in shallow *.
 maxima* populations was also found in previous research by [@Rothman2017-pn] who identified light as a significant driver.
@@
 Wave exposure may be reduced in shallow environments due to the wave damping effect of deep kelp and seaweed populations [@Kobayashi1993-os; @Mork1996-md; @Jackson1984] thereby reducing the need to adapt morphological characteristics to reduce drag.
-Furthermore, the reduction in differences between deep and shallow *E.
-maxima * at wave exposed sites on the western of the peninsula compared to sites in False Bay provides further evidence that wave exposure is not an important driver of shallow *E.
-maxima* populations.
-Increased temperatures corresponded with a reduction in stipe circumference, frond length, stipe length, total length and stipe mass resulting in slim or ‘skinny’ kelp individuals, which supports findings from previous research by [@mabin2013; @Serisawa2002-dc].
+Furthermore, the reduction in differences between deep and shallow E.
+maxima at wave exposed sites on the western of the peninsula compared to sites in False Bay provides further evidence that wave exposure may not be an important driver of shallow *E.
+maxima* populations in this study.
+Increased temperatures corresponded with a reduction in stipe circumference, frond length, stipe length, total length and stipe mass resulting in slim or “skinny” kelp individuals, which supports findings from previous research by @mabin2013, @Serisawa2002-dc.
 These findings were confirmed by the redundancy analysis which showed the contribution of wave variables as a driver of kelp morphological characteristics was lower compared to temperature variables.
-The sites do cluster by region but almost overlap entirely, further indicating wave variables are not the main driver.
-Although wave variables were not identified as an important driver, it is important to note the wave variables identified in the RDA differ to deep *E.
-maxima*.
-The identification of median wind direction (*dirw_median*) suggests that surface gravity waves as a driver compared to swell waves (*dir_median*) identified for *L.
+The sites do cluster by region but almost overlap entirely; further indicating wave variables are not the main driver.
+Although wave variables were not identified as an important driver, it is important to note the wave variables identified in the RDA differ to deep E.
+maxima.
+The identification of median wind direction suggests that surface gravity waves as a driver compared to swell waves identified for *L.
 pallida* and *E.
@@
 
-Kelps living in warmer environments have been shown to have a lower resilience to purbutation, such as storms [@Wernberg2010b].
-The kelps in False Bay, may therefore be more vulnerable to dislodgment under high wave exposure scenarios compared to kelps on the western side of the Cape Peninsula, which experiences higher wave energy on a regular basis.
-In South Africa, the main contributing factors to the hydrodynamic environment are the Southern Annular Node (SAM) and El Nino Southern Oscillation (ENSO), while the latter is also the main source of variability [@Veitch2019].
+Kelps living in warmer environments have been shown to have a lower resilience to disturbance, such as storms [@Wernberg2010b].
+The kelps in False Bay may therefore be more vulnerable to dislodgment under high wave exposure scenarios compared to kelps on the western side of the Cape Peninsula, which experiences higher wave energy on a regular basis.
+In South Africa, the main contributing factors to the hydrodynamic environment are the Southern Annular Node (SAM) and El Niño Southern Oscillation (ENSO), while the latter is also the main source of variability [@Veitch2019].
 These climate systems are expected to alter storm frequency and magnitude under future climate scenarios, which may increase the overall wave power around the Cape Peninsula.
-The magnitude-specific response of the kelp species to wave exposure in this study suggests there is a limit to morphological adaption in reducing the probability of dislodgement.
+The magnitude-specific response of the kelp species to wave exposure in this study suggests there is a limit to morphological adaptation in reducing the probability of dislodgement.
 Therefore, if this limit is reached and exceeded in future climate scenarios, kelp populations may risk being reduced over time.
@@
 We confirm that wave variables are the main driver of kelp morphological characteristics in both *E.
-maxima *and *L.
+maxima* and *L.
 pallida* populations around the Cape Peninsula, and differences in morphological characteristics between shallow and deep populations of *E.
 maxima* may reflect differences in hydrodynamic and temperature regimes.
-This study also confirms the findings of previous research which shows morphological adaptation is dependent on the magnitude of wave exposure, but provides quantitative, reproducible measure of wave exposure while identifying the specific wave metrics and wave energy thresholds involved.
+This study also confirms the findings of previous research which shows morphological adaptation is dependent on the magnitude of wave exposure, but also provides quantitative, reproducible measure of wave exposure while identifying the specific wave metrics and wave energy thresholds involved.
 Furthermore, the current study highlights the need to understand how kelp populations, which are not exposed to high wave energy and morphological characteristics are largely driven by temperature, are vulnerable to changes in storm frequency and magnitude.
@@
 
-The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest
+CR was employed by the company MetOcean Solutions Ltd.
+
+The remaining authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
 
@@
 
-RC conceptualised the scope of the research reported in this paper, collected samples, undertook the majority of numerical and statistical analyses, made the first round of interpretation, and did the majority of the writing.
-AS and CR provided guidance in terms of  writing, statistical analyses, and the final draft of the paper.
+RC conceptualized the scope of the research reported in this manuscript, collected samples, undertook the majority of numerical and statistical analyses, made the first round of interpretation, and did the majority of the writing.
+AS and CR provided the guidance in terms of writing, statistical analyses, and overall presentation of the manuscript.
+All authors contributed to the article and approved the submitted version.
 
@@
 
-The research was funded by the South African National Research Foundation ([http://www.nrf.ac.za](http://www.nrf.ac.za)). Funding number is CSRP170430229220
-
+This research was funded by the South African National Research Foundation (http://www.nrf.ac.za).
+Funding number is CSRP170430229220.
 Aside from funding provided, the funders had no role in the study design, data collection and analysis, decision to publish, or preparation of the manuscript.
@@
 
-Thank you to AS and CR for their guidance, and the honours class of 2014-2015 that aided in fieldwork. Carla Louise for making the GIS plot etc.
+We thank A.
+J Smit and Christo Rautenbach for their guidance on this project.
+We also thank Robert Schlegel and Robert Williamson for their help coordinating the honors class of 2014–2015, in particular Jesse Smith, Janine Greuel, Tevya Lotriet, and Rudy Philander.
+The work would have not been possible without the dataset from the CSIR report 2014 where the processed data is available at the following DOIs: 10.15493/DEFF.10000004 and 10.15493/DEFF.10000003.
+Finally, we thank the South African Coastal Temperature Network (SACTN) for access the temperature data used in this study.
+
+# Supplementary Material
+
+The Supplementary Material for this article can be found online at: https://www.frontiersin.org/articles/10.3389/fmars.2020.00567/full#supplementary-material
 

### 2020-11-21 thesis_chapter_2.Rmd
housecleaning word output updated

@@
 output:
+  pdf_document: 
+    fig_caption: yes
+    latex_engine: xelatex
+  html_document:
+    df_print: paged
   word_document:
     reference_docx: markdown/frontiers_template.docx
-  html_document:
-    df_print: paged
-  pdf_document:
-    fig_caption: yes
 bibliography: references.bib
@@
 
-```{r rda analysis lamanaria, message=FALSE, warning=FALSE, include=FALSE, eval= TRUE}
+```{r rda analysis lamanaria, message=FALSE, warning=FALSE, include=FALSE, eval= FALSE}
 #data processed in the temperature and wave sections will be used here eg. abio_clim
@@
 
-```{r rda deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
+```{r rda deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
 
@@
 
-```{r rda shallow ecklonia processing, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
+```{r rda shallow ecklonia processing, message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
 
@@
 ```
-The same approaches to identify the descriptors for L.
-pallida were employed for both populations of E.
-maxima.
-Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of deep E.
-maxima morphological characteristics with an adjusted R2 of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, F = 2.4, p = 0.026; Figure \ref{fig:rda plots}B).
+The same approaches to identify the descriptors for *L.
+pallida* were employed for both populations of *E.
+maxima*.
+Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of deep *E.
+maxima* morphological characteristics with an adjusted R2 of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, F = 2.4, p = 0.026; Figure \ref{fig:rda plots}B).
 RDA1 was the only significant axis in the model (p = 0.03) and explained 72% of the variation.
@@
 The blue cluster represnts sites from the western region and the red represents sites from the eastern region."}
-ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, legend = "none", labels = "AUTO", ncol = 2, nrow = 2, align = "hv")
-```
-Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of shallow E.
-maxima morphological characteristics with an adjusted R2 of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, F = 2.41, p = 0.013, Figure \ref{fig:rda plots}C).
+#ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, legend = "none", labels = "AUTO", ncol = 2, nrow = 2, align = "hv")
+
+rda_1 <- rasterGrob(as.raster(readJPEG("thesis_chapter_2_files/figures_frontiers/fig_12A.jpeg"))) #figure missing, copy of fig_6B used
+rda_2 <- rasterGrob(as.raster(readJPEG("thesis_chapter_2_files/figures_frontiers/fig_12B.jpeg")))
+rda_3 <- rasterGrob(as.raster(readJPEG("thesis_chapter_2_files/figures_frontiers/fig_12C.jpeg")))
+
+ggarrange(rda_1, rda_2, rda_3, labels = "AUTO", align = "hv")
+```
+Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of shallow *E.
+maxima* morphological characteristics with an adjusted R2 of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, F = 2.41, p = 0.013, Figure \ref{fig:rda plots}C).
 The model consisted of one significant axis, RDA1 (p = 0.02) and explained 62% of the variation.
@@
 This study considered important drivers of the morphology of Ecklonia maxima and Laminaria pallida around the Cape Peninsula.
-These abiotic variables, namely temperature and wave exposure, have been identified in previous research on other brown algae species such as E.
-radiata and M.
-pyrifera [@Thomsen2004; @FowlerWalker2005;@FowlerWalker2006; @Wernberg2005c; @Wing2007-tv; @Stewart2009; @Miller2011-pm; @Pedersen2012-vr].
+These abiotic variables, namely temperature and wave exposure, have been identified in previous research on other brown algae species such as *E.
+radiata* and *M.
+pyrifera* [@Thomsen2004; @FowlerWalker2005;@FowlerWalker2006; @Wernberg2005c; @Wing2007-tv; @Stewart2009; @Miller2011-pm; @Pedersen2012-vr].
 The complex geomorphology of the Western Cape coastline creates an ideal natural laboratory for studies of kelps in interaction with their environment.

### 2020-11-24 thesis_chapter_3.Rmd
saved
deleted duplicate in software load in

@@
 library(hexbin)
-library(ggpubr)
-
+s
 ```

### 2020-11-25 thesis_chapter_3.Rmd
added script for lat and lon analysis

@@
 library(hexbin)
-s
+
 ```
@@
 
-```{r eddy plots, message=FALSE, warning=FALSE, include=FALSE}
+```{r eddy plots, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
 
-passive_eddy <- ggplot() + 
-  geom_point(data = p_eddy, aes(x = lon, y = lat)) +
+total_eddy <- rbind(p_eddy, Min_KFloat_Eddy, Mean_KFloat_Eddy, Max_KFloat_Eddy)
+
+ggplot() + 
+  geom_point(data = total_eddy, aes(x = lon, y = lat, color = type)) +
   xlab("Longtitude") +
   ylab("Latitude") +
-  ggtitle("Passive") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
 
-min_eddy <- ggplot() + 
-  geom_point(data = Min_K_Eddy, aes(x = lon, y = lat)) +
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("Minimum Ac") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-mean_eddy <- ggplot() + 
-  geom_point(data = Mean_KFloat_Eddy, aes(x = lon, y = lat)) +
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("Mean Ac") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-max_eddy <- ggplot() + 
-  geom_point(data = Max_KFloat_Eddy, aes(x = lon, y = lat)) +
-  xlab("Longtitude") + 
-  ylab("Latitude") +
-  ggtitle("Maximum Ac") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-```
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-ggarrange(passive_eddy, min_eddy, mean_eddy, max_eddy)
 ```
@@
 
-```{r path plots, warning=FALSE, message=FALSE, echo=FALSE}
+```{r path plots, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
 
@@
 ggarrange(min_area_path, mean_area_path, max_area_path, labels = "AUTO")
+
+```
+
+```{r trajectory plots, echo=FALSE, message=FALSE, warning=FALSE}
+
+total_mean_traj <- total_data %>% #calculate mean trajectory 
+  group_by(t, type) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+lon_plot <- ggplot() +
+  geom_point(data = total_mean_traj, aes(x = t, y = lon, color = type)) +
+  xlab("Time") +
+  ylab("Longitude") +
+  theme_bw() +
+  guides(fill=guide_legend(title="Type"))
+
+lat_plot <- ggplot() +
+  geom_point(data = total_mean_traj, aes(x = t, y = lat, color = type)) +
+  xlab("Time") +
+  ylab("Latitude") +
+  theme_bw() +
+  guides(fill=guide_legend(title="Type"))
+
+ggarrange(lon_plot, lat_plot, labels = "AUTO", common.legend = TRUE, ncol = 1, nrow = 2, legend = "right")
 

### 2020-11-30 thesis_chapter_3.Rmd
outputs updated

@@
 # References
+
+```{r}
+
+p_single <- tidync("OceanParcels/trajectory_data/RK4_Single.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Min_Single_H100<- tidync("OceanParcels/trajectory_data/Min_Single_H100_W00.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Min_Single_H095_H05 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H095_W05.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Min_Single_H090_H10 <- tidync("OceanParcels/trajectory_data/Min_Single_H090_W10.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+Min_Single_H085_H15 <- tidync("OceanParcels/trajectory_data/Min_Single_H085_W15.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, time, distance) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+```

### 2020-12-11 thesis_chapter_3.Rmd
updated outputs. added to methods. alterations to intro

--- 2020-11-25 thesis_chapter_3.Rmd
@@
 output:
+  html_document:
+  word_document:
   pdf_document:
     fig_caption: yes
-  html_document:
-  word_document:
 bibliography: references.bib
@@
 
+#library(computeMSD)
 library(ggplot2)
@@
 
-The raft-time is the amount of time macroalgae floats at the surface before becoming negatively buoyant and begins to sink to the ocean-floor.
-Raft-time, also referred to as raft longevity, is determined by epibiont load, macroalgal growth rate and the loss of kelp biomass over time through fragmentation.
-Epibiont load consists of epiphytic and bryozoan species which tend to grow more rapidly in higher light and temperature environments, such as the ocean surface [@thiel2005; @thiel2005; @graiff2016].
-The increase in epibiont load reduces the buoyancy of floating macroalgal rafts over time eventually causing them to sink to the seafloor, while also potentially increasing drag [@hobday2000; @tala2013; @craw2018].
-Fragmentation of kelp biomass over time may also reduce raft times further, while growth may counteract the effects of epibiont load and fragmentation provided the environmental conditions are favorable [@graiff2016; @macaya2016].
-For instance a study by @rothausler2011 investigated the abiotic and biotic factors influencing raft time and dispersal potential of the giant kelp *M.
-pyrifera* by a combination of tethering experiments and field surveys.
-The results showed the physiological performance of kelp declined with increasing epiphyte biomass and that higher temperatures increased growth rate of epibionts and fragmentation of kelp.
-The authors concluded from the results of both the experiments and field surveys that *M.
-pyrifera* dispersal is dependent on low temperature and moderate irradiance conditions, with high temperatures and higher irradiance reducing overall raft-time and hence dispersal potential [@rothausler2011].
-
-Many past authors have also taken various approaches to estimate not only the raft-time of macroalgal rafts but also the oceanographic influences of macroalgae trajectory.
 Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005, @thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
@@
 
+To accurately determine the trajectory of marine macroalgae these aspects need to be taken into account.
+Past research by @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
+However, this work does not consider biological material such as macroalgae.
+Although drag estimates do exist, these have been calculated for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
+Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulas to calculate drag based on shape.
+Chemical engineers in the field of fluid dynamics which come across similar problems when including drag use an approach known as a 'spherecity factor'.
+A spherecity factor is expressed mathematically by $\psi=s/s_d$, and the ratio of the surface area of a sphere having the volume as the irregular shaped particle ($s$) to the actual surface area of the particle ($s_d$).
+The values are dependent on the shape of the object and a range of values have been determined based on numerous experimental work.
+
 Although various approaches have been used in the past to investigate floating macroalagae trajectory, very few studies have employed the use of Langrangian trajectory modelling.
@@
 
-To accurately determine the trajectory of marine macroalgae these aspects need to be taken into account.
-Past research by @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
-However, this work does not consider biological material such as macroalgae.
-Although drag estimates do exist, these have been calculated for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
-Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulas to calculate drag based on shape.
-Chemical engineers in the field of fluid dynamics which come across similar problems when including drag use an approach known as a 'spherecity factor'.
-A spherecity factor is expressed mathematically by $\psi=s/s_d$, and the ratio of the surface area of a sphere having the volume as the irregular shaped particle ($s$) to the actual surface area of the particle ($s_d$).
-The values are dependent on the shape of the object and a range of values have been determined based on numerous experimental work.
-
-In this study, a hypothetical approach was used to shed light on the role of drag on macroalgal drift trajectory by means of Langrangian based trajectory simulations.
-These simulations aimed to determine the role that hydrodynamic and wind drag might play in drift trajectory of macroalgae.
-In terms of wind drag, the direct effect of wind on the kelp was estimated rather than its affect on surface-currents.
+In this study, a numerical approach is used to shed light on the role of water and direct wind drag on macroalgal drift trajectory using lagrangian based trajectory simulations.
+This will be achieved by comparing the trajectory outputs of virtual 'passive' particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of water and wind drag exposure.
+The different degrees of wave and wind exposure are meant to reflect the different buoyancy characteristics, as well as identify the role of magnitude of the drag components on overall trajectory.
 
@@
 
-## Hydrodynamic model
+## Hydrodynamic and wind model
 
@@
 The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
-The model has a spatial resolution of 1/12$^\circ$ and is interpolated on a Arakawa C native grid.
-
-## Wind model
+The model has a spatial resolution of 0.08$^\circ$ and is interpolated on a Arakawa C native grid.
 
 In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (WIND_L4_NRT_OBSERVATIONS_012_004).
-The outputs from the model is composed of 6-hourly avergaed fields of surface 10m wind velocity vectors with a spatial resolution of 1/4$^\circ$.
+The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
+Both averaged ocean current and wind data are stored as meridonal and zonal velocity vectors.
 
@@
 
-### Hydrodynamic drag
+```{r morphology wrangling, message=FALSE, warning=FALSE, include=FALSE}
+# Investigate morphology data to assist in estimating surface area
+
+# Read in morphology data
+eck_df <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
+  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
+  filter(site!= "De Hoop") 
+eck_df$site <- as.character(eck_df$site)
+eck_df <- eck_df %>% 
+  filter(population == "Deep") 
+eck_long_df <- gather(eck_df, Morphology, measurement, frond_mass:total_length, factor_key=TRUE)
+# Summary stats to make it easier to refer to later, these will not be used any analysis.
+site_list <- c("Kommetjie")
+eck_stats <- eck_long_df %>% 
+  group_by(site, Morphology) %>% 
+  summarise(mean = mean(measurement, na.rm = T),
+            minimum = min(measurement, na.rm = T),
+            maximum = max(measurement, na.rm = T)) %>% 
+  ungroup() %>% 
+  filter(site %in% site_list) %>% 
+  rename(Site = site) %>% 
+  rename(Mean = mean) %>% 
+  rename(Minimum = minimum) %>% 
+  rename(Maximum = maximum)
+
+eck_stats <- flextable(data = eck_stats)
+eck_stats <- theme_zebra(eck_stats, odd_header = "transparent", even_header = "transparent")
+eck_stats <- align(x= eck_stats, align = "center", part = "all")
+eck_stats
+
+area_table <- read.csv("tables/chapter_3/area_table.csv", sep = ";")
+area_table <- flextable(data = area_table)
+area_table <- theme_zebra(area_table, odd_header = "transparent", even_header = "transparent")
+area_table <- align(x= area_table, align = "center", part = "all")
+
+```
+
+```{r area table, echo=FALSE, message=FALSE, warning=FALSE}
+area_table
+```
+
+### Hydrodynamic and wind drag
 
@@
 The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
-The cross-sectional area for each of each virtual plant was converted to the equivalent cross-sectional area for that of a sphere in the equation for calculating hydrodynamic drag: $$ F_d = 0.5 \rho A_c C_d U^2$$
+The cross-sectional area for each of each virtual plant was converted to the equivalent cross-sectional area for that of a sphere in the equation for calculating hydrodynamic drag: $ F_d = 0.5 \rho A_c C_d U^2$.
+Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step.
+A similar approach was used for the wind drag force.
+The same equation that was used to calculate hydrodynamic drag, was used to calculate the wind drag the virtual particle.
+Instead of estimating an ad hoc percentage of wind drag, the wind velocity vectors were interpolated for each time-step and were used in the overall calculation for wind drag force.
 
-### Wind drag
-
-The total wind drag effect was calculated for the hypothetical cross-sectional area of the kelp particle which was exposed above the water surface, which in this study, was estimated as 10% of the total cross-sectional area.
-
-### Raft-time
-
-Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remains afloat at the surface of the ocean after detachment before sinking.
-Raft-time was estimated from the results of past studies by @graiff2016 and @graiff2013 for each season.
-For summer the raft time was 25 days, winter was 38 days, and autumn was 41 days respectively.
+To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios.
+These scenarios were meant to act as proxy for bouyancy, which determines the surface area exposed to hydrodynamic and wind drag.
+These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation.
+For example, if 80% of the plant is submerged and 20% of above water, then the overall hydrodynamic and wind drag forces would be 80% and 20% of the total drag forces respectively.
+The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for water drag; 0 %, 5%, 10%, and 15% for wind.
 
@@
 
-In order to determine the influence of morphology on drag components, simulations were performed for the minimum, median and maximum cross-sectional areas for one site on the Atlantic side and one site within False Bay (see figure \ref{fig:map})
+In order to determine the influence of morphology on drag components, simulations were performed for the minimum, mean and maximum cross-sectional areas for one site on the Atlantic side and one site within False Bay (see figure \ref{fig:map})
 
@@
 
-# load and data wrangle for eddy simulation
-
-p_eddy <- tidync("OceanParcels/trajectory_data/PassiveEddy.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-p_eddy_mean <- p_eddy %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-p_eddy$type <- "Passive Eddy" # add extra column 'type' to dataframe
-
-Min_KFloat_Eddy <- tidync("OceanParcels/trajectory_data/Min_KFloat_Eddy.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-Min_K_Eddy_mean <- Min_KFloat_Eddy %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_KFloat_Eddy$type <- "Min_KFloat_Eddy" # add extra column 'type' to dataframe
-
-Mean_KFloat_Eddy <- tidync("OceanParcels/trajectory_data/Mean_KFloat_Eddy.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-Mean_KFloat_Eddy_mean <- Mean_KFloat_Eddy %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_Eddy$type <- "Mean_KFloat_Eddy" # add extra column 'type' to dataframe
-
-Max_KFloat_Eddy <- tidync("OceanParcels/trajectory_data/Max_KFloat_Eddy.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-Max_KFloat_Eddy_mean <- Max_KFloat_Eddy %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_KFloat_Eddy$type <- "Max_KFloat_Eddy" # add extra column 'type' to dataframe
-
 # Read in simulation data and data wrangling
@@
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::select(trajectory, lon, lat, time, distance, uwind, vwind) %>%
   dplyr::rename(t = time) %>%
@@
 total_data <- rbind(p_RK4, Min_Kfloat_H100_W00 , Min_Kfloat_H095_W05, Min_Kfloat_H090_W10, Min_Kfloat_H085_W15, Mean_Kfloat_H100_W00, Mean_KFloat_H095_W05, Mean_KFloat_H090_W10, Mean_KFloat_H085_W15, Max_Kfloat_H100_W00, Max_Kfloat_H095_W05, Max_Kfloat_H090_W10, Max_Kfloat_H085_W15)
-
-```
-
-```{r eddy plots, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
-
-total_eddy <- rbind(p_eddy, Min_KFloat_Eddy, Mean_KFloat_Eddy, Max_KFloat_Eddy)
-
-ggplot() + 
-  geom_point(data = total_eddy, aes(x = lon, y = lat, color = type)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
 
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
+  #guides(fill=guide_legend(title="Number of particles")) +
   theme_bw()
@@
 
-ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15 , labels = c("AUTO"), common.legend = TRUE, legend = c("right"))
+a <- ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15 , labels = c("AUTO"), common.legend = TRUE, legend = c("none"))
+
+ggarrange(heatmap_RK4, a, common.legend = TRUE, legend = c("right"), ncol = 1, nrow = 2)
+
 ```
@@
 
+```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE}
+pos <- c("RK4",
+         "Min_Kfloat_H100_W00",
+         "Min_Kfloat_H095_W05",
+         "Min_Kfloat_H090_W10",
+         "Min_Kfloat_H085_W15",
+         "Mean_Kfloat_H100_W00",
+         "Mean_Kfloat_H095_W05",
+         "Mean_Kfloat_H090_W10",
+         "Mean_Kfloat_H085_W15",
+         "Max_Kfloat_H100_W00",
+         "Max_Kfloat_H095_W05",
+         "Max_Kfloat_H090_W10",
+         "Max_Kfloat_H085_W15")
+
+dis_plot <- ggplot(data = total_data, aes(x = type, y = distance, fill = type, show.legend = FALSE))+
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(limits = pos)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif")+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Type")+
+  coord_flip()
+
+```
+
+```{r distance plot, message=FALSE, warning=FALSE}
+dis_plot
+```
+
 # Discussion
@@
 
-```{r morphology wrangling, message=FALSE, warning=FALSE, include=FALSE}
-# Investigate morphology data to assist in estimating surface area
-
-# Read in morphology data
-eck_df <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
-  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
-  filter(site!= "De Hoop") 
-eck_df$site <- as.character(eck_df$site)
-eck_df <- eck_df %>% 
-  filter(population == "Deep") 
-eck_long_df <- gather(eck_df, Morphology, measurement, frond_mass:total_length, factor_key=TRUE)
-# Summary stats to make it easier to refer to later, these will not be used any analysis.
-site_list <- c("Kommetjie")
-eck_stats <- eck_long_df %>% 
-  group_by(site, Morphology) %>% 
-  summarise(mean = mean(measurement, na.rm = T),
-            minimum = min(measurement, na.rm = T),
-            maximum = max(measurement, na.rm = T)) %>% 
-  ungroup() %>% 
-  filter(site %in% site_list) %>% 
-  rename(Site = site) %>% 
-  rename(Mean = mean) %>% 
-  rename(Minimum = minimum) %>% 
-  rename(Maximum = maximum)
-
-eck_stats <- flextable(data = eck_stats)
-eck_stats <- theme_zebra(eck_stats, odd_header = "transparent", even_header = "transparent")
-eck_stats <- align(x= eck_stats, align = "center", part = "all")
-eck_stats
-
-area_table <- read.csv("tables/chapter_3/area_table.csv", sep = ";")
-area_table <- flextable(data = area_table)
-area_table <- theme_zebra(area_table, odd_header = "transparent", even_header = "transparent")
-area_table <- align(x= area_table, align = "center", part = "all")
-
-```
-
-```{r area table, echo=FALSE, message=FALSE, warning=FALSE}
-area_table
-```
-
 # References

### 2020-12-14 thesis_chapter_3.Rmd
deleted mean traj syntax and added to heatmaps. html updated

@@
 
-## Hydrodynamic and wind model
-
-Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service.
-The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
-The model has a spatial resolution of 0.08$^\circ$ and is interpolated on a Arakawa C native grid.
-
-In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (WIND_L4_NRT_OBSERVATIONS_012_004).
-The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
-Both averaged ocean current and wind data are stored as meridonal and zonal velocity vectors.
-
 ## Particle tracking model
@@
 Obs).
-Therefore, the trajectory of solitary *E.
+Therefore, the trajectory of only olitary *E.
 maxima* individuals will be considered.
@@
 
+## Hydrodynamic and wind model
+
+Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service.
+The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
+The model has a spatial resolution of 0.08$^\circ$ and is interpolated on a Arakawa C native grid.
+
+In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (WIND_L4_NRT_OBSERVATIONS_012_004).
+The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
+Both averaged ocean current and wind data are stored as meridonal and zonal velocity vectors.
+
+## Simulations
+
+Virtual particles were released with drag or wind effects from the study site in order to act as a basis for comparison to simulations with varying degrees of hydrodynamic and wind drag.
+All simulations has the same release location and run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme.
+Brownian motion was also incorporated in order to account for any stochasticity not resolved by the model.
+These are standard practices for Lagrangian trajectory studies.
+In order to determine the influence of cross-sectional area on drag components, simulations were performed for the minimum, mean and maximum cross-sectional areas for Kommetjie kelp forest population (see figure \ref{fig:map}).
+In order to determine the direct effect of wind on trajectory, simulations were run for varying magnitudes of wind exposure for each type of cross-sectional area.
+
+```{r sim table, echo=FALSE, message=FALSE, warning=FALSE}
+
+sim_table <- read.csv("tables/chapter_3/simulation_table.csv", sep = ";")
+
+sim_table <- flextable(data = sim_table)
+sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
+sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Time.Range = "Time Range", Drag.Exposure = "Drag Exposure level")
+sim_table <- align(x= sim_table, align = "center", part = "all")
+sim_table
+
+```
+
 ## Model inputs
@@
 The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
-The minimum, maximum and mean were calculated for each morphological characteristic (see appendix), which were used for calculating the overall minimum, maximum and mean cross-sectional areas needed to run the various simulations.
-The trajectory of these "types" of cross-sectional areas (minimum, maximum and mean overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
+The minimum, mean and maximum were calculated for each morphological characteristic (see appendix), which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations.
+The trajectory of these "types" of cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
 
@@
 The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
-The cross-sectional area for each of each virtual plant was converted to the equivalent cross-sectional area for that of a sphere in the equation for calculating hydrodynamic drag: $ F_d = 0.5 \rho A_c C_d U^2$.
+The cross-sectional area for each of each virtual plant was converted to the equivalent cross-sectional area for that of a sphere in the equation for calculating hydrodynamic drag: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to water or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
+Water and wind drag components were calculated separately for both the meridonal and zonal velocity vectors.
 Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step.
@@
 The same equation that was used to calculate hydrodynamic drag, was used to calculate the wind drag the virtual particle.
-Instead of estimating an ad hoc percentage of wind drag, the wind velocity vectors were interpolated for each time-step and were used in the overall calculation for wind drag force.
 
 To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios.
-These scenarios were meant to act as proxy for bouyancy, which determines the surface area exposed to hydrodynamic and wind drag.
+These scenarios were meant to act as proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag.
 These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation.
@@
 The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for water drag; 0 %, 5%, 10%, and 15% for wind.
-
-## Simulations
-
-In order to determine the influence of morphology on drag components, simulations were performed for the minimum, mean and maximum cross-sectional areas for one site on the Atlantic side and one site within False Bay (see figure \ref{fig:map})
-
-```{r sim table, echo=FALSE, message=FALSE, warning=FALSE}
-
-sim_table <- read.csv("tables/chapter_3/simulation_table.csv", sep = ";")
-
-sim_table <- flextable(data = sim_table)
-sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
-sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Time.Range = "Time Range", Drag.Exposure = "Drag Exposure level")
-sim_table <- align(x= sim_table, align = "center", part = "all")
-sim_table
-
-```
 
@@
   geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 13000)) +
+  geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
 
-a <- ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15 , labels = c("AUTO"), common.legend = TRUE, legend = c("none"))
+ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15 , labels = c("AUTO"), common.legend = TRUE, legend = c("none"))
 
-ggarrange(heatmap_RK4, a, common.legend = TRUE, legend = c("right"), ncol = 1, nrow = 2)
+#ggarrange(heatmap_RK4, a, common.legend = TRUE, legend = c("right"), ncol = 1, nrow = 2)
 
 ```
-
-```{r path plots, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
-
-min_area_path <- ggplot() + 
-  geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'blue'), show.legend = FALSE) +
-  geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'yellow'), show.legend = FALSE) +
-  geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'pink'), show.legend = FALSE) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("Minimum cross-sectional area") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-mean_area_path <- ggplot() + 
-  geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'blue'), show.legend = FALSE) +
-  geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'yellow'), show.legend = FALSE) +
-  geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'pink'), show.legend = FALSE) +
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("Mean cross-sectional area") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-max_area_path <- ggplot() + 
-  geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'blue'), show.legend = FALSE) +
-  geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'yellow'), show.legend = FALSE) +
-  geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'pink'), show.legend = FALSE) +
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("Maximum cross-sectional area") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-ggarrange(min_area_path, mean_area_path, max_area_path, labels = "AUTO")
-
-```
-
 ```{r trajectory plots, echo=FALSE, message=FALSE, warning=FALSE}

### 2020-12-15 thesis_chapter_3.Rmd
added pca. output updated

@@
 
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+library(ggfortify)
+
+total_end <- total_data %>%
+             filter(t == "2018-01-30 00:00:00")
+
+df <- total_end[2:3]
+pca_res <- prcomp(df, scale. = TRUE)
+
+autoplot(pca_res, data = total_end, colour = 'type', loadings = TRUE, loading.label = TRUE, frame = TRUE, frame.type = 'norm')
+```
+
 ```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE}
@@
 
-```{r distance plot, message=FALSE, warning=FALSE}
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE}
 dis_plot

### 2020-12-15 chapter_4.Rmd
added in template for chapter 4

---
title: "Chapter 4"
author: "Ross Coppin"
date: "11/12/2020"
output: html_document
---

# Introduction

The raft-time is the amount of time macroalgae floats at the surface before becoming negatively buoyant and begins to sink to the ocean-floor. Raft-time, also referred to as raft longevity, is determined by epibiont load, macroalgal growth rate and the loss of kelp biomass over time through fragmentation. Epibiont load consists of epiphytic and bryozoan species which tend to grow more rapidly in higher light and temperature environments, such as the ocean surface [@thiel2005; @thiel2005; @graiff2016]. The increase in epibiont load reduces the buoyancy of floating macroalgal rafts over time eventually causing them to sink to the seafloor, while also potentially increasing drag [@hobday2000; @tala2013; @craw2018]. Fragmentation of kelp biomass over time may also reduce raft times further, while growth may counteract the effects of epibiont load and fragmentation provided the environmental conditions are favorable [@graiff2016; @macaya2016]. For instance a study by @rothausler2011 investigated the abiotic and biotic factors influencing raft time and dispersal potential of the giant kelp *M. pyrifera* by a combination of tethering experiments and field surveys. The results showed the physiological performance of kelp declined with increasing epiphyte biomass and that higher temperatures increased growth rate of epibionts and fragmentation of kelp. The authors concluded from the results of both the experiments and field surveys that *M. pyrifera* dispersal is dependent on low temperature and moderate irradiance conditions, with high temperatures and higher irradiance reducing overall raft-time and hence dispersal potential [@rothausler2011].

# Methods

## Raft-time

Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remains afloat at the surface of the ocean after detachment before sinking. Raft-time was estimated from the results of past studies by @graiff2016 and @graiff2013 for each season. For summer the raft time was 25 days, winter was 38 days, and autumn was 41 days respectively.

### 2020-12-16 thesis_chapter_3.Rmd
fixed tables. added analyses methods

--- 2020-12-14 thesis_chapter_3.Rmd
@@
 output:
+  word_document:
+    reference_docx: markdown/frontiers_template.docx
   html_document:
-  word_document:
   pdf_document:
@@
 library(tidyverse)
-library(viridis)
 library(gridExtra)
@@
 maxima* individuals will be considered.
-Previous studies have made the assumption that floating macroalage move with the same velocity as the ocean current, and instead model a ensemble of particles which trajectories are determined by a combination of ocean currents and/or the effects of winds on currents.
-
 Numerical calculations of particle trajectory with a Lagrangian method [@delandmeter2019], and be described by the equation
@@
 The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations.
-Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculated for each time step of each simulation.
+Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculated for each time step for each simulation.
 
@@
 
-Virtual particles were released with drag or wind effects from the study site in order to act as a basis for comparison to simulations with varying degrees of hydrodynamic and wind drag.
-All simulations has the same release location and run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme.
+Virtual particles were released with no drag or wind effects from the study site in order to act as a basis for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are refferred to hereafter as 'passive particles'.
+All simulations have the same release location and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme.
 Brownian motion was also incorporated in order to account for any stochasticity not resolved by the model.
 These are standard practices for Lagrangian trajectory studies.
-In order to determine the influence of cross-sectional area on drag components, simulations were performed for the minimum, mean and maximum cross-sectional areas for Kommetjie kelp forest population (see figure \ref{fig:map}).
-In order to determine the direct effect of wind on trajectory, simulations were run for varying magnitudes of wind exposure for each type of cross-sectional area.
 
-```{r sim table, echo=FALSE, message=FALSE, warning=FALSE}
+```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Details for each simulation run for the study"}
 
@@
 sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
-sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Time.Range = "Time Range", Drag.Exposure = "Drag Exposure level")
+sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Drag.Exposure = "Drag Exposure level")
 sim_table <- align(x= sim_table, align = "center", part = "all")
+sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
 sim_table
@@
 ```
+
+In order to determine the influence of cross-sectional area on drag components, simulations were performed for the minimum, mean and maximum cross-sectional areas for Kommetjie kelp forest population (see figure \ref{fig:map}).
+In order to determine the direct effect of wind on trajectory, simulations were run for varying magnitudes of wind exposure for each type of cross-sectional area.
+The virtual particles released with the hydrodynamic or wind drag components are referred to as 'kelp particles'.
 
@@
 
-```{r morphology wrangling, message=FALSE, warning=FALSE, include=FALSE}
+```{r morphology wrangling, message=FALSE, warning=FALSE, include=FALSE, fig.cap="Table summarising the atrributes used and estimated to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation."}
 # Investigate morphology data to assist in estimating surface area
@@
 eck_stats <- align(x= eck_stats, align = "center", part = "all")
-eck_stats
+eck_stats <- set_table_properties(sim_table, width = 1, layout = "autofit")
 
@@
 area_table <- align(x= area_table, align = "center", part = "all")
+area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
+area_table <- set_header_labels(area_table, Plant.characteristic = "Plant characteristic", Approximate.shape = "Approximate shape", Equation.used = "Equation Used", Plant.dimensions = "Plant dimensions")
+area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.maximum = "Ac Maximum")
 
@@
 
-## Statistical analysis
+## Analysis
+
+In order to compare trajectories between cross-sectional area types and wind exposure scenarios, density distribution maps were produced.
+This approach is an effective for assessing pathways and ocean connectivity [@sebille2018].
+In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs.
+The distance that particles traveled was also measured and used to produce boxplots and scatterplots.
+Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@sebille2018].
+If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; @putman2018; @sebille2018].
 
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 13000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 13000)) +
+  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
   geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
   geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
   geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
   geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
   geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
   geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
 ```
+
 ```{r trajectory plots, echo=FALSE, message=FALSE, warning=FALSE}
@@
 lon_plot <- ggplot() +
-  geom_point(data = total_mean_traj, aes(x = t, y = lon, color = type)) +
+  geom_line(data = total_mean_traj, aes(x = t, y = lon, color = type), shape = "type") +
   xlab("Time") +
@@
 lat_plot <- ggplot() +
-  geom_point(data = total_mean_traj, aes(x = t, y = lat, color = type)) +
+  geom_line(data = total_mean_traj, aes(x = t, y = lat, color = type)) +
   xlab("Time") +
@@
 ```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE}
+
 pos <- c("RK4",

### 2020-12-16 thesis_chapter_3.Rmd
figure captions added

@@
 area_table <- set_header_labels(area_table, Plant.characteristic = "Plant characteristic", Approximate.shape = "Approximate shape", Equation.used = "Equation Used", Plant.dimensions = "Plant dimensions")
-area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.maximum = "Ac Maximum")
+area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
 
@@
 In order to compare trajectories between cross-sectional area types and wind exposure scenarios, density distribution maps were produced.
-This approach is an effective for assessing pathways and ocean connectivity [@sebille2018].
+This approach is an effective for assessing pathways and ocean connectivity [@Sebille2018].
 In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs.
 The distance that particles traveled was also measured and used to produce boxplots and scatterplots.
-Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@sebille2018].
-If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; @putman2018; @sebille2018].
+Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@Sebille2018].
+If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@Putman2020; @Putman2018; @Sebille2018].
 
@@
 
-```{r trajectory plots, echo=FALSE, message=FALSE, warning=FALSE}
+```{r trajectory plots, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean trajectory paths of each exposure scenario over the entire simulation run.
+Trajectory along the latitudinal and longitudal paths are plotted over time seperatley.
+Plot A is the longitudinal and plot B the latitudinal trajectories respectively."}
 
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE}
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Boxplots of total distance from source over the course of the simulation for each exposure scenario."}
 dis_plot

### 2020-12-16 thesis_chapter_3.Rmd
outputs updated

@@
 output:
+  html_document:
   word_document:
     reference_docx: markdown/frontiers_template.docx
-  html_document:
   pdf_document:

### 2021-01-05 thesis_chapter_3.Rmd
comparison graphs added. html updated

@@
   html_document:
+  pdf_document:
+    fig_caption: yes
   word_document:
     reference_docx: markdown/frontiers_template.docx
-  pdf_document:
-    fig_caption: yes
 bibliography: references.bib
@@
 library(ggpubr)
-#library(ggord)
+library(ggord)
 library(vegan)
@@
 library(hexbin)
+library(lubridate)
 
@@
 geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-            size = 1.0, colour = "black", fill = NA) +
+            size = 1.0, colour = "black", fill = "green") +
   #geom_point(data = study_sites, aes(x = lon, y = lat), color = "red", nudge_x = 0, nudge_y = 0.1) +
@@
   geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.10, xmax=18.50, ymin=-33.70, ymax=-34.50), inherit.aes = FALSE, fill = "blue", alpha = 0.01) +
-  geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.40, xmax=18.90, ymin=-34.00, ymax=-34.50), inherit.aes = FALSE, fill = "red", alpha = 0.01) +
+  #geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.40, xmax=18.90, ymin=-34.00, ymax=-34.50), inherit.aes = FALSE, fill = "red", alpha = 0.01) +
   #annotate("text", label = "Warm\nTemperate", x = 18.7, y = -34.5, size = 4.0, angle = 0, colour = "black") +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+               size = 1.0, colour = "black", fill = "green") +
   geom_point(data = release_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
@@
   theme(axis.ticks = element_blank(), # Remove tick marks
-        #axis.text = element_blank(), # Remove lat/ lon numbers
+        axis.text = element_blank(), # Remove lat/ lon numbers
         axis.title = element_blank()) # Remove lat/ lon labels
@@
 area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
+area_table
 
-```
-
-```{r area table, echo=FALSE, message=FALSE, warning=FALSE}
-area_table
 ```
@@
 In order to compare trajectories between cross-sectional area types and wind exposure scenarios, density distribution maps were produced.
-This approach is an effective for assessing pathways and ocean connectivity [@Sebille2018].
+Density distribution maps are an effective method for assessing pathways and ocean connectivity [@Sebille2018].
 In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs.
@@
 
+## Domain features
+
+```{r, message=FALSE, warning = FALSE, include= TRUE, echo = FALSE, fig.height=8, fig.width=8}
+library(jpeg)
+library(grid)
+library(gridExtra)
+
+fieldUend <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/fieldUend.jpg")))
+fieldVend <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/fieldVend.jpg")))
+windUend <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/windUend.jpg")))
+windUend <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/windVend.jpg")))
+
+ggarrange(fieldUend, fieldVend, windUend, windUend, ncol = 2, nrow = 2, labels = "AUTO")
+```
+
 ```{r data wrangling, message=FALSE, warning=FALSE, include=FALSE}
@@
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, uwind, vwind) %>%
+  dplyr::select(trajectory, lon, lat, time) %>%
   dplyr::rename(t = time) %>%
@@
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, uwind, vwind) %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
   dplyr::rename(t = time) %>%
@@
 
-total_data <- rbind(p_RK4, Min_Kfloat_H100_W00 , Min_Kfloat_H095_W05, Min_Kfloat_H090_W10, Min_Kfloat_H085_W15, Mean_Kfloat_H100_W00, Mean_KFloat_H095_W05, Mean_KFloat_H090_W10, Mean_KFloat_H085_W15, Max_Kfloat_H100_W00, Max_Kfloat_H095_W05, Max_Kfloat_H090_W10, Max_Kfloat_H085_W15)
+total_data <- rbind(Min_Kfloat_H100_W00 , Min_Kfloat_H095_W05, Min_Kfloat_H090_W10, Min_Kfloat_H085_W15, Mean_Kfloat_H100_W00, Mean_KFloat_H095_W05, Mean_KFloat_H090_W10, Mean_KFloat_H085_W15, Max_Kfloat_H100_W00, Max_Kfloat_H095_W05, Max_Kfloat_H090_W10, Max_Kfloat_H085_W15)
 
 ```
+
+## Trajectory comparison
+
+```{r message=FALSE, warning=FALSE, include=FALSE}
+waterdrag_plot <- ggplot()+
+  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Max_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+wwdrag_plot <- ggplot()+
+  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = Min_Kfloat_H095_W05_mean, aes(x = lon, y = lat), color = 'snow4')+
+  geom_path(data = Min_Kfloat_H090_W10_mean, aes(x = lon, y = lat), color = 'green')+
+  geom_path(data = Min_Kfloat_H085_W15_mean, aes(x = lon, y = lat), color = 'pink')+
+  geom_path(data = Mean_KFloat_H095_W05_mean, aes(x = lon, y = lat), color = 'yellow')+
+  geom_path(data = Mean_KFloat_H090_W10_mean, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Mean_KFloat_H085_W15_mean, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = Max_Kfloat_H095_W05_mean, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = Max_Kfloat_H090_W10_mean, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Max_Kfloat_H085_W15_mean, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() 
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
+ggarrange(waterdrag_plot, wwdrag_plot)
+```
+
+## Comparison of density distributions
 
@@
   geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Comparison of density of particles within each grid cell for the end run time of each scenario.'}
-
-ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15 , labels = c("AUTO"), common.legend = TRUE, legend = c("none"))
-
-#ggarrange(heatmap_RK4, a, common.legend = TRUE, legend = c("right"), ncol = 1, nrow = 2)
-
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Comparison of density of particles within each grid cell for the end run time of each scenario.', fig.height=8, fig.width=8}
+ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, heatmap_RK4, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"))
 ```
 
-```{r trajectory plots, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean trajectory paths of each exposure scenario over the entire simulation run.
-Trajectory along the latitudinal and longitudal paths are plotted over time seperatley.
-Plot A is the longitudinal and plot B the latitudinal trajectories respectively."}
-
-total_mean_traj <- total_data %>% #calculate mean trajectory 
-  group_by(t, type) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-lon_plot <- ggplot() +
-  geom_line(data = total_mean_traj, aes(x = t, y = lon, color = type), shape = "type") +
-  xlab("Time") +
-  ylab("Longitude") +
-  theme_bw() +
-  guides(fill=guide_legend(title="Type"))
-
-lat_plot <- ggplot() +
-  geom_line(data = total_mean_traj, aes(x = t, y = lat, color = type)) +
-  xlab("Time") +
-  ylab("Latitude") +
-  theme_bw() +
-  guides(fill=guide_legend(title="Type"))
-
-ggarrange(lon_plot, lat_plot, labels = "AUTO", common.legend = TRUE, ncol = 1, nrow = 2, legend = "right")
-
-```
-
-```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE}
+## Comparison of distances 
+```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE, fig.height=3, fig.width=4}
 
@@
 
+```{r shape data, message=FALSE, echo=FALSE, warning=FALSE}
+
+Max_ShapeFloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_ShapeFloat_H100_W00_mean <- Max_ShapeFloat_H100_W00 %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_ShapeFloat_H100_W00$type <- "Max_ShapeFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+# Make mean trajectory plot
+shape_plot <- ggplot()+
+  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = Max_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = Max_ShapeFloat_H100_W00_mean, aes(x = lon, y = lat), color = 'orange')+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+# make density map
+heatmap_Max_ShapeFloat_H100_W00 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_ShapeFloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw()
+
+shape_data_total <- rbind(Max_Kfloat_H100_W00, Max_ShapeFloat_H100_W00)
+
+shape_dis_plot <- ggplot(data = shape_data_total, aes(x = type, y = distance, fill = type, show.legend = FALSE))+
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(limits = pos)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif")+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Type")
+
+```
+
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Comparison of trajectory data between a kelp sphere and kelp cylinder.", fig.height=8, fig.width=8}
+
+ggarrange(shape_plot, shape_dis_plot, heatmap_Max_KFloat_H100_W00, heatmap_Max_ShapeFloat_H100_W00, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 2)
+
+```
+
 # Discussion

### 2021-01-06 thesis_chapter_3.Rmd
added to results section

@@
 
-Although various approaches have been used in the past to investigate floating macroalagae trajectory, very few studies have employed the use of Langrangian trajectory modelling.
+Although various approaches have been used in the past to investigate floating macroalagae trajectory, very few studies have employed the use of Langrangian trajectory modeling.
 Furthermore, non of the existing studies using this modeling approach have considered macroalgal morphology (i.e.
@@
 Obs).
-Therefore, the trajectory of only olitary *E.
+Therefore, the trajectory of only solitary *E.
 maxima* individuals will be considered.
@@
 
-Virtual particles were released with no drag or wind effects from the study site in order to act as a basis for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are refferred to hereafter as 'passive particles'.
+Virtual particles were released with no drag or wind effects from the study site in order to act as a basis for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'.
 All simulations have the same release location and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme.
@@
 Therefore, the cross-sectional area of the bulb/pneumatocyst was not considered in the calculation of overall cross-sectional area.
-In terms of the holdfast, a standard cross-sectional area was used for simulations where the inclusion of the holdfast was considered.
+In terms of the holdfast, a standard cross-sectional area was used for all simulations.
 The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
@@
 
-The use of a spherical shape coefficient was used in the approach to simulate the trajectory of solitary floating *E.
-maxima*.
 The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
-The cross-sectional area for each of each virtual plant was converted to the equivalent cross-sectional area for that of a sphere in the equation for calculating hydrodynamic drag: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to water or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
+The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to water or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
 Water and wind drag components were calculated separately for both the meridonal and zonal velocity vectors.
@@
 These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation.
-For example, if 80% of the plant is submerged and 20% of above water, then the overall hydrodynamic and wind drag forces would be 80% and 20% of the total drag forces respectively.
+For example, if 80% of the plant is submerged and 20% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively.
 The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for water drag; 0 %, 5%, 10%, and 15% for wind.
@@
 
-fieldUend <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/fieldUend.jpg")))
-fieldVend <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/fieldVend.jpg")))
-windUend <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/windUend.jpg")))
-windUend <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/windVend.jpg")))
+vector_2 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day2.jpg")))
+vector_5 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day5.jpg")))
+vector_10 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day10.jpg")))
+vector_15 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day15.jpg")))
+vector_20 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day20.jpg")))
+vector_25 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day25.jpg")))
+vector_30 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day30.jpg")))
 
-ggarrange(fieldUend, fieldVend, windUend, windUend, ncol = 2, nrow = 2, labels = "AUTO")
+ggarrange(vector_2, vector_5, vector_10, vector_15, vector_20, vector_25, vector_30, labels = "AUTO")
 ```
@@
 ## Trajectory comparison
+
+The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles.
+When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen with only slight variation.
+The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation.
+The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well as being further from the release site.
+When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories is evident.
 

### 2021-01-06 thesis_chapter_3.Rmd
added more to results section

@@
 Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step.
-A similar approach was used for the wind drag force.
-The same equation that was used to calculate hydrodynamic drag, was used to calculate the wind drag the virtual particle.
+The same approach was used for the wind drag force.
+The equation that was used to calculate hydrodynamic drag, was also used to calculate the wind drag the virtual particle.
 
@@
 The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well as being further from the release site.
-When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories is evident.
+When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories was evident.
+In all simulations particles flowed in a north-westward direction.
 
@@
 ## Comparison of density distributions
+
+Across all simulations particles flowed in a north-westward direction.
+The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has a almost even density of particles across grids.
+In addition the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticy while particles in the passive simulation deflected away.
+Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.
 
@@
 ## Comparison of distances
+
+There were significant differences among the median distances travelled for all particles in each simulation (see figure ??; p < 0.05).
+The simulations that considered for 100% hydrodynamic drag travelled less distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
+Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations amongst simulations.
+
 ```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE, fig.height=3, fig.width=4}

### 2021-01-07 thesis_chapter_3.Rmd
added to results section

@@
 output:
-  html_document:
+  word_document:
+    reference_docx: markdown/frontiers_template.docx
   pdf_document:
     fig_caption: yes
-  word_document:
-    reference_docx: markdown/frontiers_template.docx
+  html_document:
 bibliography: references.bib
@@
 Although drag estimates do exist, these have been calculated for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
-Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulas to calculate drag based on shape.
-Chemical engineers in the field of fluid dynamics which come across similar problems when including drag use an approach known as a 'spherecity factor'.
-A spherecity factor is expressed mathematically by $\psi=s/s_d$, and the ratio of the surface area of a sphere having the volume as the irregular shaped particle ($s$) to the actual surface area of the particle ($s_d$).
-The values are dependent on the shape of the object and a range of values have been determined based on numerous experimental work.
-
 Although various approaches have been used in the past to investigate floating macroalagae trajectory, very few studies have employed the use of Langrangian trajectory modeling.
@@
 
-In order to compare trajectories between cross-sectional area types and wind exposure scenarios, density distribution maps were produced.
+In order to compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated and density distribution maps were produced.
 Density distribution maps are an effective method for assessing pathways and ocean connectivity [@Sebille2018].
 In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs.
-The distance that particles traveled was also measured and used to produce boxplots and scatterplots.
+The distance that particles traveled was also measured and used to produce boxplots.
 Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@Sebille2018].
@@
 
-```{r message=FALSE, warning=FALSE, include=FALSE}
+```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
 waterdrag_plot <- ggplot()+
@@
   ylab("Latitude") + 
-  #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  ggtitle("Comparison of hydrodynamic drag simulations") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ylab("Latitude") + 
+  ggtitle("Comparison of hydrodynamic and wind drag simulations")
   #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
 ```
@@
   ylab("Latitude") + 
-  #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  ggtitle("Passive") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ylab("Latitude") + 
+  ggtitle("H100W100") +
   guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  ggtitle("H095W05") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ylab("Latitude") + 
+  ggtitle("H090W10") + 
   guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ylab("Latitude") + 
+  ggtitle("H085W15") + 
   guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ylab("Latitude") + 
+  ggtitle("H100W00") +
   guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ylab("Latitude") + 
+  ggtitle("H095W05") +
   guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ylab("Latitude") + 
+  ggtitle("H090W10") +
   guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ylab("Latitude") + 
+  ggtitle("H085W15") +
   guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ylab("Latitude") + 
+  ggtitle("H095W05") +
+  theme(plot.title = element_text(hjust = 0.5)) + 
   guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   xlab("Type")+
-  coord_flip()
+  coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   ylab("Latitude") +
+  ggtitle("Mean trajectory comparison") +
   #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw()
+  ggtitle("Cylinder") + 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# make another density map for the maximum so that title can be manipulated
+
+heatmap_Max_Sphere_H100_W00 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Sphere") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
   stat_compare_means(method = "kruskal.test", label = "p.signif")+
-  theme_bw()+
   ylab("Distance (km)")+
-  xlab("Type")
+  xlab("Type") + 
+  ggtitle("Comparison of distance travelled") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
 
-ggarrange(shape_plot, shape_dis_plot, heatmap_Max_KFloat_H100_W00, heatmap_Max_ShapeFloat_H100_W00, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 2)
+ggarrange(shape_plot, shape_dis_plot, heatmap_Max_Sphere_H100_W00, heatmap_Max_ShapeFloat_H100_W00, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 2)
 

### 2021-01-09 thesis_chapter_3.Rmd
updated discussion added. more to come

@@
 
+The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices.
+Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics, however current research has focused on large macroalgae-rafts that can vary greatly in size and surface area.
+None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae.
+This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag into account.
+The passive simulations deviated from passive particles trajectories when including varing levels of hydrodynamic and wind drag scenarios into account.
+The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns and as well our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.
+
+The general flow in the study domain is north-westward and is characterised by sub-mesoscale processes such as vortices and sub-mesoscale eddies.
+
+Past research has inferred from direct and indirect techniques that drifting macroalgae tend to follow general flow patterns in the region in question. For example...
+
+The results from this numerical experiment confirm that drifting macroalgae do follow the surface current in the study region when considering only hydrodynamic drag.
+In addition, not only are the trajectories similar but so are the end points at the end of simulation.
+However, when ad-hoc wind drag (windage) is considered in combination with hydrodynamic drag the trajectories differ greatly when compared to purely langrangian particle trajectories.
+The inclusion of wind drag causes the particles to flow further away from the coast and with a greater distance travelled from the release location.
+The difference in trajectory and distance covered can be attributed to the inclusion of wind drag.
+Wind drag adds to the effects of hydrodynamic drag when the wind direction and current direction are opposite to each other.
+However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object.
+This causes the particles to travel further and to be exposed to different time varying flows.
+insert putman study here.
+This is also reflected in the analysis comparing distances travelled by all the particles in each simulation.
+The particles in the simulations that included wind drag travlled significantly greater distances compared to the passive or hydrodynamic only drag.
+
+The addition of drag forces into the simulation also causes the particles to cluster together along the mean trajectory compared to that of the passive particles which are more evenly dispersed.
+This is most likely due to the shared drag characteristics among the particles which included the plants weight and surface area into the calculation.
+Other studies investigating flotsam trajectory characteristics have found similar results.
+A study by @miron2020 conducted field experiments using a range of objects such as spheres, cubes, cuboids to compare to investigate the effect of inertia on particles dynamics.
+The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia.
+Although drag forces were used in this study and not inertia, this study provides evidence to account for the higher density of particles along the mean trajectory.
+
+The ocean is made up of varying time-varying flows and changes in an objects velocity can cause the object to be exposed to different flow patterns over time.
+The lack of dissimilarity between different cross-sectional area types (minimum, mean and maximum) was an unexpected result.
+A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region.
+The results suggested that the hatchlings trajectories are mostly influenced by the ocean currents in the first year of hatching due to low swimming speeds of juveniles.
+Therefore, the differences in cross-sectional area types may not be significant enough in order to have a significant effect on the relevant velocity vectors (zonal and meridonal).
+Another possible reason is the resolution of the ocean model used.
+The resolution of the underlying ocean model is an important aspect of lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications.
+A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study.
+The authors found that the inclusion of brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications.
+Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.
+
+The same is also true when comparing different hydrodynamic and wind drag exposure scenarios.
+These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag.
+The similarities between the different combined drag scenarios suggests that higher magnitudes of wind exposure do not significantly alter trajectories.
+Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns.
+All the simulations that included any form of drag resulted in a fairly high density of particles entrained in a vortex.
+The entrainment and expulsion from eddies and vortices is an important characteristic of floating macroalgae.
+Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real sargassum rafts.
+However, the aforementioned studies investigated large rafts and not floating individuals.
+Inertia is based on the size/weight of the raft, while hydrodynamic and wind drag are based on current velocity and surface area.
+Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag should be considered.
+The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.
+
+??the study by putman which used different ad-hoc windage paramterisations showed that higher windsage resulted in greater accuracy of entrainment in eddies??.
+
 # Appendix

### 2021-01-10 thesis_chapter_3.Rmd
chapter 3 discussion updated

@@
 
-The general flow in the study domain is north-westward and is characterised by sub-mesoscale processes such as vortices and sub-mesoscale eddies.
-
-Past research has inferred from direct and indirect techniques that drifting macroalgae tend to follow general flow patterns in the region in question. For example...
-
+Past research has inferred from direct and indirect techniques that drifting macroalgae tend to follow the prevailing surface currents [@thiel2005, @thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
 The results from this numerical experiment confirm that drifting macroalgae do follow the surface current in the study region when considering only hydrodynamic drag.
 In addition, not only are the trajectories similar but so are the end points at the end of simulation.
-However, when ad-hoc wind drag (windage) is considered in combination with hydrodynamic drag the trajectories differ greatly when compared to purely langrangian particle trajectories.
+However, when ad-hoc wind drag (windage) is considered in combination with hydrodynamic drag the trajectories differ greatly when compared to purely lagrangian particle trajectories.
 The inclusion of wind drag causes the particles to flow further away from the coast and with a greater distance travelled from the release location.
@@
 However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object.
-This causes the particles to travel further and to be exposed to different time varying flows.
-insert putman study here.
+This causes the particles to travel further as well as exposed to different time varying flows.
+Past studies have shown this when including windage as a function of inertia [@putman2018;@brooks2019; @putman2020].
+A study by @putman2020 used ad-hoc windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic Sargassum.
+The results showed that including a ad-hoc windage factor in proving virtual trajectories to tracked Sargassum mats which was partly due to the inclusion of momemtum energy transfer from the wind.
 This is also reflected in the analysis comparing distances travelled by all the particles in each simulation.
-The particles in the simulations that included wind drag travlled significantly greater distances compared to the passive or hydrodynamic only drag.
+The particles in the simulations that included wind drag traveled significantly greater distances compared to the passive or hydrodynamic only drag.
 
@@
 This is most likely due to the shared drag characteristics among the particles which included the plants weight and surface area into the calculation.
-Other studies investigating flotsam trajectory characteristics have found similar results.
+Other studies investigating flotsam trajectory characteristics have found similar results [@breivik2011; @olascoaga2020; @miron2020].
 A study by @miron2020 conducted field experiments using a range of objects such as spheres, cubes, cuboids to compare to investigate the effect of inertia on particles dynamics.
 The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia.
-Although drag forces were used in this study and not inertia, this study provides evidence to account for the higher density of particles along the mean trajectory.
+Although drag forces were used in this study and not inertia, the same conclusion applies, drag is an important characteristic to consider when investigating or predicting solitary drifting macrolagae.
 
@@
 The entrainment and expulsion from eddies and vortices is an important characteristic of floating macroalgae.
-Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real sargassum rafts.
+Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts.
 However, the aforementioned studies investigated large rafts and not floating individuals.
@@
 
-??the study by putman which used different ad-hoc windage paramterisations showed that higher windsage resulted in greater accuracy of entrainment in eddies??.
+The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory.
+Instead, the inclusion of a differnt shape coefficient increased to the density of particles entrained in an eddy field.
+This suggest that inclusion of different shape coefficients play a role in further improving estmiates and hindcasts of macroalgal disperal patterns.
+
+# Conclusion
+
+Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape, however no research exists attempting to clarify the drifting characteristics of solitary macroalgae.
+Understanding how to accurately model the distribution of solitary kelp can lead to be understanding of accumulation zones and sinks and ultimately the ecological links pertaining to those processes.
+The findings from this study indicate that solitary floating virtual *E.
+maxima* particles do tend to follow the prevailing surface currents but that the inclusion of wind causes the trajectories to differ greatly from that of a purely lagrangian particle.
+Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies.
+Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are neglible when investigating dispersal patterns of drifting *E.maxima*.
+In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory but also causes higher density of particles to become entrained in an eddy.
+Overall the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not.
+This study also provides an approach which can be adapted to model floating solitary macroalgae, provided the surface area can be estimated accurately.
+However, more gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal as well as how these change seasonally.
+The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates and raft-times.
 

### 2021-01-11 thesis_chapter_3.Rmd
minor alterations. table captions issue resolved

@@
 output:
+  html_document:
+  pdf_document:
+    fig_caption: yes
   word_document:
     reference_docx: markdown/frontiers_template.docx
-  pdf_document:
-    fig_caption: yes
-  html_document:
 bibliography: references.bib
@@
 However, the past research on macroalgal trajectory has not investigated these factors which have been shown to be important aspects of trajectory for other materials such as icebergs, marine craft and microplastics.
-The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in water form drag.
+The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag.
 Drag coefficients related to wind and surface current have been shown to be important properties to consider when estimating trajectory and forecasting drift for search and sea rescue operations.
@@
 
-In this study, a numerical approach is used to shed light on the role of water and direct wind drag on macroalgal drift trajectory using lagrangian based trajectory simulations.
-This will be achieved by comparing the trajectory outputs of virtual 'passive' particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of water and wind drag exposure.
+In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using lagrangian based trajectory simulations.
+This will be achieved by comparing the trajectory outputs of virtual 'passive' particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure.
 The different degrees of wave and wind exposure are meant to reflect the different buoyancy characteristics, as well as identify the role of magnitude of the drag components on overall trajectory.
@@
 
-```{r map data, message=FALSE, warning=FALSE, include=FALSE}
+```{r map, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Map of the study domain. This is a temporary map until issues with the map script are resolved."}
+library(jpeg)
+library(grid)
+library(gridExtra)
 
-source("map/scaleBarFunc.R")
-
-load("map/south_africa_coast_hi_res.Rdata")
-
-map_pen <- ggplot(data = south_africa_coast_hi_res) +
-geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-            size = 1.0, colour = "black", fill = "green") +
-  #geom_point(data = study_sites, aes(x = lon, y = lat), color = "red", nudge_x = 0, nudge_y = 0.1) +
-  #geom_label(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
-  geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.10, xmax=18.50, ymin=-33.70, ymax=-34.50), inherit.aes = FALSE, fill = "blue", alpha = 0.01) +
-  #geom_rect(data = south_africa_coast_hi_res,aes(xmin=18.40, xmax=18.90, ymin=-34.00, ymax=-34.50), inherit.aes = FALSE, fill = "red", alpha = 0.01) +
-  #annotate("text", label = "Warm\nTemperate", x = 18.7, y = -34.5, size = 4.0, angle = 0, colour = "black") +
-  #annotate("text", label = "Cool\nTemperate", x = 18.0, y = -34.0, size = 4.0, angle = 0, colour = "black") +
-  annotate("text", label = "FALSE\nBAY", x =18.65, y = -34.25, size = 4.0, angle = 0, colour = "black") +
-  scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
-  arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
-  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
-  #theme(axis.ticks = element_blank()) + # Remove tick marks
-  xlab("Latitude") +
-  ylab("Longitude")
-
-## Coastline of Southern Africa
-load("map/south_africa_coast_hi_res.Rdata")
-
-release_sites <- read.csv("metadata/release_sites.csv", sep = ";")
-
-## Make 2nd map of just the peninsula
-release_map <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = "green") +
-  geom_point(data = release_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
-  geom_label(data = release_sites, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
-  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -33.600)) +
-  theme(axis.ticks = element_blank(), # Remove tick marks
-        axis.text = element_blank(), # Remove lat/ lon numbers
-        axis.title = element_blank()) # Remove lat/ lon labels
-#ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")
-
-```
-
-```{r map,echo=FALSE, message=FALSE, warning=FALSE, fig.cap='\\label{fig:map}Map of the study area and sampling sites.
-The blue shaded area is the Atlantic region and the red shaded region is the False Bay region.'}
-
-ggarrange(map_pen, release_map, labels = c("AUTO"), align = c("hv"), ncol = 2, nrow = 1)
-
+map_chp3 <- rasterGrob(as.raster(readJPEG("map/tempmap_chapter3.jpg")))
+ggarrange(map_chp3)
 ```
@@
 
-This study simulated two scenarios; one where the plant is fully submerged and only exposed to seawater drag; and the other where parts of the plants are partially exposed resulting in a combination of seawater and air form drag.
+This study simulated two scenarios; one where the plant is fully submerged and only exposed to hydrodynamic drag; and the other where parts of the plants are partially exposed resulting in a combination of hydrodynamic and air form drag.
 Drag based on shape based coefficients were used in the calculations for determining the overall kelp velocity vectors.
@@
 sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
+sim_table <- set_caption(sim_table, "Table 1:Details for each simulation executed for the study.")
 sim_table
-
 ```
@@
 In order to incorporate hydrodynamic and wind drag, the cross-sectional area of the kelp was calculated first.
-Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant, for details please see image/table ???.
+Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant, for details please refer to table 2.
 The dimensional data needed was estimated in cases were data was not available for that particular morphological characteristic.
@@
 
-```{r morphology wrangling, message=FALSE, warning=FALSE, include=FALSE, fig.cap="Table summarising the atrributes used and estimated to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation."}
+```{r morphology wrangling, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
 # Investigate morphology data to assist in estimating surface area
@@
 
-eck_stats <- flextable(data = eck_stats)
-eck_stats <- theme_zebra(eck_stats, odd_header = "transparent", even_header = "transparent")
-eck_stats <- align(x= eck_stats, align = "center", part = "all")
-eck_stats <- set_table_properties(sim_table, width = 1, layout = "autofit")
+#eck_stats <- flextable(data = eck_stats)
+#eck_stats <- theme_zebra(eck_stats, odd_header = "transparent", even_header = "transparent")
+#eck_stats <- align(x= eck_stats, align = "center", part = "all")
+#eck_stats <- set_table_properties(eck_stats, width = 1, layout = "autofit")
 
+```
+
+```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation."}
 area_table <- read.csv("tables/chapter_3/area_table.csv", sep = ";")
@@
 area_table <- align(x= area_table, align = "center", part = "all")
-area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
 area_table <- set_header_labels(area_table, Plant.characteristic = "Plant characteristic", Approximate.shape = "Approximate shape", Equation.used = "Equation Used", Plant.dimensions = "Plant dimensions")
 area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
+area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
+area_table <- set_caption(sim_table, "Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.")
 area_table
-
 ```
@@
 The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
-The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to water or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
-Water and wind drag components were calculated separately for both the meridonal and zonal velocity vectors.
+The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
+hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors.
 Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step.
@@
 The equation that was used to calculate hydrodynamic drag, was also used to calculate the wind drag the virtual particle.
-
 To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios.
@@
 For example, if 80% of the plant is submerged and 20% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively.
-The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for water drag; 0 %, 5%, 10%, and 15% for wind.
+The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1).
 
@@
 # Results
-
-## Domain features
-
-```{r, message=FALSE, warning = FALSE, include= TRUE, echo = FALSE, fig.height=8, fig.width=8}
-library(jpeg)
-library(grid)
-library(gridExtra)
-
-vector_2 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day2.jpg")))
-vector_5 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day5.jpg")))
-vector_10 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day10.jpg")))
-vector_15 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day15.jpg")))
-vector_20 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day20.jpg")))
-vector_25 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day25.jpg")))
-vector_30 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day30.jpg")))
-
-ggarrange(vector_2, vector_5, vector_10, vector_15, vector_20, vector_25, vector_30, labels = "AUTO")
-```
 
@@
 
-The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles.
+The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (\ref{fig:meantrajA}.
 When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen with only slight variation.
-The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation.
+The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation(\ref{fig:meantrajB}.
 The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well as being further from the release site.
@@
 
-```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
+```{r meantraj, message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
+
 waterdrag_plot <- ggplot()+
@@
   ylab("Latitude") + 
-  ggtitle("Comparison of hydrodynamic drag simulations") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
 
@@
   ylab("Latitude") + 
-  ggtitle("Comparison of hydrodynamic and wind drag simulations")
-  #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
-ggarrange(waterdrag_plot, wwdrag_plot)
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Comparison of mean trajectories.
+Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag."}
+ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO")
 ```
@@
   ylab("Latitude") + 
-  ggtitle("H100W00") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") + 
-  ggtitle("H095W05") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") + 
-  ggtitle("H090W10") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") + 
-  ggtitle("H085W15") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") + 
-  ggtitle("H095W05") +
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
   theme(plot.title = element_text(hjust = 0.5))
-
 ```
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Comparison of density of particles within each grid cell for the end run time of each scenario.', fig.height=8, fig.width=8}
-ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, heatmap_RK4, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"))
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Comparison of density of particles within each grid cell for the end run time of each scenario.
+Plots A-D repressent the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
+The bottom plot is the density plot for the passive simulation.', fig.height=10.2, fig.width=7.5}
+
+figure_a <- ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, labels = c("AUTO"), common.legend = TRUE, legend = c("none"))
+
+figure_b <- ggarrange(heatmap_RK4, legend = c("right"))
+
+ggarrange(figure_a, figure_b, ncol = 1, nrow = 2)
+
 ```
@@
 
-There were significant differences among the median distances travelled for all particles in each simulation (see figure ??; p < 0.05).
+There were significant differences among the median distances travelled for all particles in each simulation (see figure (\ref{fig:distance plot}; p < 0.05).
 The simulations that considered for 100% hydrodynamic drag travelled less distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
@@
 
-dis_plot <- ggplot(data = total_data, aes(x = type, y = distance, fill = type, show.legend = FALSE))+
+dis_plot <- ggplot(data = total_data, aes(x = type, y = distance, show.legend = FALSE))+
   geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Boxplots of total distance from source over the course of the simulation for each exposure scenario."}
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Boxplots of total distance of all particles from the release site over the course of the simulation for each exposure scenario."}
 dis_plot
 ```
+
+## Comparison of coefficents
+
+The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no dissimilarity.
+This was also reflected when comparing distance travelled by particles and particle density plots.
+The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient.
 
@@
 
-shape_dis_plot <- ggplot(data = shape_data_total, aes(x = type, y = distance, fill = type, show.legend = FALSE))+
+shape_dis_plot <- ggplot(data = shape_data_total, aes(x = type, y = distance, show.legend = FALSE))+
   geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
@@
 
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Comparison of trajectory data between a kelp sphere and kelp cylinder.", fig.height=8, fig.width=8}
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Comparison of trajectory data between a kelp sphere and kelp cylinder.
+Plot A is the comparison of mean trajectories, plot B is the comparison of distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.", fig.height=8, fig.width=8}
 
@@
 
-The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices.
-Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics, however current research has focused on large macroalgae-rafts that can vary greatly in size and surface area.
+The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @Sebille2018; @miron2020].
+Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], however current research has focused on large macroalgae-rafts that can vary greatly in size and surface area.
 None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae.
@@
 Wind drag adds to the effects of hydrodynamic drag when the wind direction and current direction are opposite to each other.
-However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object.
+However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object [@hackett2006; @putman2020].
 This causes the particles to travel further as well as exposed to different time varying flows.
@@
 The ocean is made up of varying time-varying flows and changes in an objects velocity can cause the object to be exposed to different flow patterns over time.
-The lack of dissimilarity between different cross-sectional area types (minimum, mean and maximum) was an unexpected result.
+The lack of dissimilarity between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible.
 A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region.
@@
 The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.
-
 The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory.
@@
 
+## Domain features
+
+```{r, message=FALSE, warning = FALSE, include= TRUE, echo = FALSE, fig.height=8, fig.width=8}
+library(jpeg)
+library(grid)
+library(gridExtra)
+
+vector_2 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day2.jpg")))
+vector_5 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day5.jpg")))
+vector_10 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day10.jpg")))
+vector_15 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day15.jpg")))
+vector_20 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day20.jpg")))
+vector_25 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day25.jpg")))
+vector_30 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day30.jpg")))
+
+ggarrange(vector_2, vector_5, vector_10, vector_15, vector_20, vector_25, vector_30, labels = "AUTO")
+```
+
 # References

### 2021-01-11 thesis_chapter_3.Rmd
chnages made to plots

@@
 ```{r density plots, message=FALSE, warning=FALSE, echo=FALSE}
-
 ## Heatmaps of study domain, use previous data for study site map and histograms
@@
 
-Max_ShapeFloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00.nc") %>%
+sphere <- tidync("OceanParcels/trajectory_data/Max_KFloat_H100_W00.nc") %>%
   hyper_tibble() %>%
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_ShapeFloat_H100_W00_mean <- Max_ShapeFloat_H100_W00 %>% #calculate mean trajectory 
+sphere_mean <- sphere %>% #calculate mean trajectory 
   group_by(t) %>% 
@@
 
-Max_ShapeFloat_H100_W00$type <- "Max_ShapeFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+sphere$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+sphere_mean$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+# read in data again for one max simulation, as cannot sort issue out with rename df...
+cyl <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+cyl_mean <- cyl %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+cyl$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+cyl_mean$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
   geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Max_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Max_ShapeFloat_H100_W00_mean, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = sphere_mean, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = cyl_mean, aes(x = lon, y = lat), color = 'orange')+
   xlab("Longtitude") + 
@@
 # make density map
-heatmap_Max_ShapeFloat_H100_W00 <- ggplot() +
+heatmap_sphere <- ggplot() +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_ShapeFloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = sphere, bins = 80, aes(x = lon, y = lat)) +
   scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  geom_path(data = sphere_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
 
-# make another density map for the maximum so that title can be manipulated
-
-heatmap_Max_Sphere_H100_W00 <- ggplot() +
+heatmap_cyl <- ggplot() +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = cyl, bins = 80, aes(x = lon, y = lat)) +
   scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  geom_path(data = cyl_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
 
-shape_data_total <- rbind(Max_Kfloat_H100_W00, Max_ShapeFloat_H100_W00)
+shape_data_total <- rbind(sphere, cyl)
 
@@
 
-ggarrange(shape_plot, shape_dis_plot, heatmap_Max_Sphere_H100_W00, heatmap_Max_ShapeFloat_H100_W00, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 2)
+ggarrange(shape_plot, shape_dis_plot, heatmap_sphere, heatmap_cyl, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 2)
 

### 2021-01-11 thesis_chapter_3.Rmd
cross-references added

@@
 The region of interest for this study was the Cape Peninsula in the Western Cape, South Africa.
-This was further sub-divided into two domains, the Atlantic domain and False Bay domain (see figure \ref{fig:map}A).
-One release site is located in each of the domains, namely Kommetjie in the Atlantic and Buffelsbaai in False Bay (see figure \ref{fig:map}B).
+This was further sub-divided into two domains, the Atlantic domain and False Bay domain (Figure 1A).
+One release site is located in each of the domains, namely Kommetjie in the Atlantic and Buffelsbaai in False Bay (Figure 2B).
 The hydrodynamics between the two domains are driven by a complex interaction of wind and wave vectors, with wind driven processes primarily occurring in False Bay.
@@
 
-```{r map, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Map of the study domain. This is a temporary map until issues with the map script are resolved."}
+```{r map, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 1:Map of the study domain. This is a temporary map until issues with the map script are resolved."}
 library(jpeg)
@@
 
-In order to determine the influence of cross-sectional area on drag components, simulations were performed for the minimum, mean and maximum cross-sectional areas for Kommetjie kelp forest population (see figure \ref{fig:map}).
+In order to determine the influence of cross-sectional area on drag components, simulations were performed for the minimum, mean and maximum cross-sectional areas for Kommetjie kelp forest population (Figure 1).
 In order to determine the direct effect of wind on trajectory, simulations were run for varying magnitudes of wind exposure for each type of cross-sectional area.
@@
 
-The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (\ref{fig:meantrajA}.
+The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (Figure 1A).
 When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen with only slight variation.
-The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation(\ref{fig:meantrajB}.
+The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation (Figure 1B).
 The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well as being further from the release site.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Figure 2:Comparison of mean trajectories.
 Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag."}
@@
 
-Across all simulations particles flowed in a north-westward direction.
-The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has a almost even density of particles across grids.
-In addition the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticy while particles in the passive simulation deflected away.
+Across all simulations particles flowed in a north-westward direction (Figure 3).
+The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has a almost even density of particles across grids (Figure 3A, 3E, 3I).
+In addition the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticy while particles in the passive simulation deflected away (Figure 3).
 Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Comparison of density of particles within each grid cell for the end run time of each scenario.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3:Comparison of density of particles within each grid cell for the end run time of each scenario.
 Plots A-D repressent the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
@@
 
-There were significant differences among the median distances travelled for all particles in each simulation (see figure (\ref{fig:distance plot}; p < 0.05).
+There were significant differences among the median distances travelled for all particles in each simulation (Figure 4; p < 0.05).
 The simulations that considered for 100% hydrodynamic drag travelled less distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Boxplots of total distance of all particles from the release site over the course of the simulation for each exposure scenario."}
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4:Boxplots of total distance of all particles from the release site over the course of the simulation for each exposure scenario."}
 dis_plot
@@
 
-The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no dissimilarity.
-This was also reflected when comparing distance travelled by particles and particle density plots.
-The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient.
+The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no dissimilarity (Figure 5A).
+This was also reflected when comparing distance travelled by particles and particle density plots (Figure 5B).
+The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (Figure 5C-D).
 
@@
 
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Comparison of trajectory data between a kelp sphere and kelp cylinder.
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5:Comparison of trajectory data between a kelp sphere and kelp cylinder.
 Plot A is the comparison of mean trajectories, plot B is the comparison of distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.", fig.height=8, fig.width=8}

### 2021-01-11 thesis_chapter_3.Rmd
updating rmd file and output.

@@
 In order to compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated and density distribution maps were produced.
-Density distribution maps are an effective method for assessing pathways and ocean connectivity [@Sebille2018].
+Density distribution maps are an effective method for assessing pathways and ocean connectivity [@sebille2018].
 In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs.
 The distance that particles traveled was also measured and used to produce boxplots.
-Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@Sebille2018].
-If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@Putman2020; @Putman2018; @Sebille2018].
+Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@sebille2018].
+If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; putman2018; @sebille2018].
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Figure 2:Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Figure 2: Comparison of mean trajectories.
 Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag."}
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3:Comparison of density of particles within each grid cell for the end run time of each scenario.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density of particles within each grid cell for the end run time of each scenario.
 Plots A-D repressent the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
@@
 The simulations that considered for 100% hydrodynamic drag travelled less distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
-Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations amongst simulations.
+Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations.
 
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4:Boxplots of total distance of all particles from the release site over the course of the simulation for each exposure scenario."}
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance of all particles from the release site over the course of the simulation for each exposure scenario."}
 dis_plot
@@
 
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5:Comparison of trajectory data between a kelp sphere and kelp cylinder.
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5: Comparison of trajectory data between a kelp sphere and kelp cylinder.
 Plot A is the comparison of mean trajectories, plot B is the comparison of distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.", fig.height=8, fig.width=8}
@@
 
-The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @Sebille2018; @miron2020].
+The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @sebille2018; @miron2020].
 Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], however current research has focused on large macroalgae-rafts that can vary greatly in size and surface area.

### 2021-01-11 thesis_chapter_3.Rmd
references added and minor errors sorted

@@
 Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
-Considerable research has been performed investigating macroalgae as a passive dispersal mechanism in the ocean, however, in recent years much of the focus has shifted towards microplastics.
-
 Some macroalgae species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species of macroalgae have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.
@@
 However, this work does not consider biological material such as macroalgae.
-Although drag estimates do exist, these have been calculated for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
+Although drag estimates do exist, these have been applied for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
 Although various approaches have been used in the past to investigate floating macroalagae trajectory, very few studies have employed the use of Langrangian trajectory modeling.
-Furthermore, non of the existing studies using this modeling approach have considered macroalgal morphology (i.e.
+Furthermore, non of the existing studies using this modeling approach have considered macroalgal morphology cross-sectional area (i.e.
 shape) as an aspect of drag and ultimately trajectory.
-Currently, work by @brooks2019 which investigated the effect of inertia and radial size on the trajectory of pelagic *Sargussum* rafts.
-The study used a custom growth model to estimate changes in biomass and ultimately radial size, while a customised Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
+Studies by @brooks2019, @putman2018, @putman2020 used lagrangian approaches to investigate the effects of inertia, raft-size, and windage on macroalgae trajectory.
+@brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customised Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
 The results showed that trajectory of pelagic *Sargussum* was significantly influenced by inertia and the radial size of the rafts.
+Other work by @putman2020 investigated the effect of including windage in macroalgae trajectory simulations.
+The authors showed that including ad-hoc windage factors greatly improved accuracy regarding entrainment or expulsion from eddies when compared to tracked sargassum mats in the ocean.
 Langrangian trajectory modeling is a useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
@@
 
-The region of interest for this study was the Cape Peninsula in the Western Cape, South Africa.
-This was further sub-divided into two domains, the Atlantic domain and False Bay domain (Figure 1A).
-One release site is located in each of the domains, namely Kommetjie in the Atlantic and Buffelsbaai in False Bay (Figure 2B).
-The hydrodynamics between the two domains are driven by a complex interaction of wind and wave vectors, with wind driven processes primarily occurring in False Bay.
-For a more detailed explanation of the hydrodynamics of these two domains the reader is referred to @coppin2020.
-The hydrodynamic characteristics of these regions are also influenced by larger oceanic processes, which in South Africa can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC).
-The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly southwestward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean.
+The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A).
+A release site was chosen for this study, Kommetjie, which is located on the Atlantic side of the Cape Peninsula (Figure 2B).
+The hydrodynamics in the study region are driven by a complex interaction of wind and waves.
+which are also influenced by larger oceanic processes In South Africa, the larger processes can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC).
+The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean.
 
@@
 The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
-hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors.
+Hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors.
 Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step.
 The same approach was used for the wind drag force.
-The equation that was used to calculate hydrodynamic drag, was also used to calculate the wind drag the virtual particle.
+The equation that was used to calculate hydrodynamic drag, was also used to calculate wind.
 To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios.
@@
 In order to compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated and density distribution maps were produced.
-Density distribution maps are an effective method for assessing pathways and ocean connectivity [@sebille2018].
+Density distribution maps are an effective method for assessing pathways and ocean connectivity [@vanSebille2018].
 In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs.
 The distance that particles traveled was also measured and used to produce boxplots.
-Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@sebille2018].
-If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; putman2018; @sebille2018].
+Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@VanSebille2018].
+If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; @putman2018; @vanSebille2018].
 
@@
 Across all simulations particles flowed in a north-westward direction (Figure 3).
-The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has a almost even density of particles across grids (Figure 3A, 3E, 3I).
+The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids (Figure 3A, 3E, 3I).
 In addition the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticy while particles in the passive simulation deflected away (Figure 3).
@@
 
-The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @sebille2018; @miron2020].
-Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], however current research has focused on large macroalgae-rafts that can vary greatly in size and surface area.
+The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @VanSebille2018; @miron2020].
+Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], current research has focused on large macroalgae-rafts which can vary greatly in size and surface area.
 None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae.
-This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag into account.
-The passive simulations deviated from passive particles trajectories when including varing levels of hydrodynamic and wind drag scenarios into account.
-The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns and as well our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.
+This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag.
+The mean virtual kelp trajectory deviated from the mean passive  trajectory when including varying levels of hydrodynamic and wind drag scenarios.
+The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns, and also our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.
 
-Past research has inferred from direct and indirect techniques that drifting macroalgae tend to follow the prevailing surface currents [@thiel2005, @thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
+Past research has inferred from direct and indirect techniques that drifting macroalgae tend to follow the prevailing surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
 The results from this numerical experiment confirm that drifting macroalgae do follow the surface current in the study region when considering only hydrodynamic drag.
@@
 A study by @putman2020 used ad-hoc windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic Sargassum.
-The results showed that including a ad-hoc windage factor in proving virtual trajectories to tracked Sargassum mats which was partly due to the inclusion of momemtum energy transfer from the wind.
+The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked Sargassum mats which was partly due to the inclusion of momemtum energy transfer from wind.
 This is also reflected in the analysis comparing distances travelled by all the particles in each simulation.
@@
 The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia.
-Although drag forces were used in this study and not inertia, the same conclusion applies, drag is an important characteristic to consider when investigating or predicting solitary drifting macrolagae.
+Although drag forces were used in this study and not inertia, the same conclusion applies, drag is an important characteristic to consider when investigating or predicting solitary drifting macroalgae.
 
@@
 The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory.
-Instead, the inclusion of a differnt shape coefficient increased to the density of particles entrained in an eddy field.
-This suggest that inclusion of different shape coefficients play a role in further improving estmiates and hindcasts of macroalgal disperal patterns.
+Instead, the inclusion of a different shape coefficient increased to the density of particles entrained in an eddy field.
+This suggest that inclusion of different shape coefficients play a role in further improving estimates and hindcasts of macroalgal dispersal patterns.
 
@@
 The findings from this study indicate that solitary floating virtual *E.
-maxima* particles do tend to follow the prevailing surface currents but that the inclusion of wind causes the trajectories to differ greatly from that of a purely lagrangian particle.
+maxima* particles tend to follow the prevailing surface currents and the inclusion of wind causes the trajectories to differ greatly from that of a purely lagrangian particle.
 Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies.
@@
 Overall the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not.
-This study also provides an approach which can be adapted to model floating solitary macroalgae, provided the surface area can be estimated accurately.
-However, more gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal as well as how these change seasonally.
+This study also provides an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately.
+Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal as well as how these vary seasonally.
 The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates and raft-times.

### 2021-01-11 thesis_chapter_3.Rmd
one error

@@
 Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies.
-Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are neglible when investigating dispersal patterns of drifting *E.maxima*.
+Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E.maxima*.
 In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory but also causes higher density of particles to become entrained in an eddy.

### 2021-01-12 thesis_chapter_3.Rmd
draft version ready

@@
 The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean.
+The region where the retroflection occurs is known as the Agulhas retroflection and consists of some of the highest levels of mesoscale variability [@garzoli1996].
+The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988; @lutjeharms2006; @@lutjeharms2007].
+The Agulhas Rings are the main contributors to the Agulhas leakage which is the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015].
+The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009].
+The Benguela region is characterised by strong coastal upwelling cells which a primarily driven by equatorward winds [@shannon1996; @hardman2003], and is known as the Benguela upwelling region.
+The Benguela upwelling region starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996].
+The region is also characterised by high amounts of mesoscale variability in the form of eddies and filaments[@blanke2002; @blanke2005; @rubio2009].
+The Benguela inshore region is dominated by a northwestward flow as a result of topographical steering, wind stress and interactions with passing Agulhas rings and eddies [@veitch2010].
 
@@
 area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
-area_table <- set_caption(sim_table, "Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.")
+area_table <- set_caption(area_table, "Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.")
 area_table
@@
 
-The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @VanSebille2018; @miron2020].
+The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @miron2020].
 Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], current research has focused on large macroalgae-rafts which can vary greatly in size and surface area.

### 2021-01-15 animations.R
animation script and outputs added

@@ -1,0 +1,417 @@
+# use chapter 3 script to load the data, packages specific to the animation will be loaded here
+
+#require(xtractomatic)
+require(tidyverse)
+require(oce)
+require(lubridate)
+require(gganimate)
+require(sf)
+require(transformr)
+
+# Animations for the density plots
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+ani_heatmap_RK4 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Passive") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  transition_time(t)
+
+ani_heatmap_Min_Kfloat_H100_W00 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("H100W100") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  transition_time(t)
+
+
+ani_heatmap_Min_Kfloat_H095_W05 <- ggplot() +
[…]

### 2021-01-15 thesis_chapter_3.Rmd
minor alterations

@@
 
-The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (Figure 1A).
+The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (Figure 2A).
 When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen with only slight variation.
-The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation (Figure 1B).
+The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation (Figure 2B).
 The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well as being further from the release site.

### 2021-01-20 thesis_chapter_3.Rmd
added in velocity column data wrangling

@@
 title: 'Numerical trajectory patterns of floating macroalgae'
-author: "Ross Coppin"
+author:
+- R Coppin^[University of the Western Cape]
+- C Rautenbach^[South African Weather Service]
+- AJ Smit^[University of the Western Cape]
 output:
@@
 Min_Kfloat_H100_W00$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00$velocity <- with(Min_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Min_Kfloat_H095_W05$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05$velocity <- with(Min_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Min_Kfloat_H090_W10$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10$velocity <- with(Min_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Min_Kfloat_H085_W15$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15$velocity <- with(Min_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Mean_Kfloat_H100_W00$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00$velocity <- with(Mean_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Mean_KFloat_H095_W05$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05$velocity <- with(Mean_KFloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Mean_KFloat_H090_W10$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10$velocity <- with(Mean_KFloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Mean_KFloat_H085_W15$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15$velocity <- with(Mean_KFloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Max_Kfloat_H100_W00$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00$velocity <- with(Max_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Max_Kfloat_H095_W05$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05$velocity <- with(Max_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Max_Kfloat_H090_W10$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10$velocity <- with(Max_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 Max_Kfloat_H085_W15$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15$velocity <- with(Max_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
 

### 2021-01-25 thesis_chapter_2.Rmd
minor and general alterations

@@
   arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
-  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
+  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-34.55, -32.50)) +
   theme(axis.ticks = element_blank()) + # Remove tick marks
@@
   geom_label(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0, nudge_y = 0) +
-  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -33.915)) +
+  coord_cartesian(xlim = c(18.275, 18.525), ylim = c( -34.375, -33.915)) +
   theme(axis.ticks = element_blank(), # Remove tick marks

### 2021-01-25 thesis_chapter_3.Rmd
map problem sorted

@@
 output:
+  html_document:
   word_document:
@@
     fig_caption: yes
-  html_document:
+    latex_engine: xelatex
 bibliography: references.bib
@@
 library(ggord)
+library(mapproj)
 library(vegan)
@@
 
-```{r map, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 1:Map of the study domain. This is a temporary map until issues with the map script are resolved."}
-library(jpeg)
+```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1:Map of the study domain. This is a temporary map until issues with the map script are resolved."}
+library(png)
 library(grid)
@@
 
-map_chp3 <- rasterGrob(as.raster(readJPEG("map/tempmap_chapter3.jpg")))
-ggarrange(map_chp3)
+map_chp3A <- rasterGrob(as.raster(readPNG("map/chp3_mapA.png")))
+map_chp3B <- rasterGrob(as.raster(readPNG("map/chp3_mapB.png")))
+
+ggarrange(map_chp3A, map_chp3B, ncol = 1, nrow = 2, labels = "AUTO")
 ```
@@
 ```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Details for each simulation run for the study"}
+set_flextable_defaults(fonts_ignore=TRUE)
 
@@
 ```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation."}
+
+set_flextable_defaults(fonts_ignore=TRUE)
+
 area_table <- read.csv("tables/chapter_3/area_table.csv", sep = ";")
@@
 ```{r, message=FALSE, warning = FALSE, include= TRUE, echo = FALSE, fig.height=8, fig.width=8}
+
 library(jpeg)

### 2021-01-25 animations.R
updated animations to include velocity

@@ -1,14 +1,27 @@
 # use chapter 3 script to load the data, packages specific to the animation will be loaded here
 
 #require(xtractomatic)
-require(tidyverse)
-require(oce)
-require(lubridate)
-require(gganimate)
-require(sf)
-require(transformr)
+library(tidyverse)
+library(oce)
+library(lubridate)
+library(gganimate)
+library(sf)
+library(transformr)
 
-# Animations for the density plots
+waterdrag_plot <- ggplot()+
+  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black') +
+  geom_point(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black') +
+  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'blue') +
+  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'orange') +
+  geom_path(data = Max_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'red') +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  transition_time(t)
+
+######################################## Animations for the density plots ################################################
 ## Coastline of Southern Africa
 load("map/south_africa_coast_hi_res.Rdata")
 
@@ -17,29 +30,32 @@ ani_heatmap_RK4 <- ggplot() +
                size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_viridis_c(name = "Density", limits = c(0, 1000)) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
   xlab("Longtitude") +
   ylab("Latitude") +
   ggtitle("Passive") +
-  guides(fill=guide_legend(title="Number of particles")) +
   theme_bw() +
[…]

### 2021-01-25 thesis_chapter_4.Rmd
preliminary table added to chapter 4

--- 2020-12-15 chapter_4.Rmd
@@
 ---
+
+```{r message=FALSE, warning=FALSE, include=FALSE}
+
+library(tidyverse)
+library(flextable)
+
+```
 
@@
 For summer the raft time was 25 days, winter was 38 days, and autumn was 41 days respectively.
+
+```{r morphology surface area table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="TBD."}
+
+# Read in morphology data
+
+eck_df <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
+  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
+  filter(site!= "De Hoop") 
+eck_df$site <- as.character(eck_df$site)
+eck_df <- eck_df %>% 
+  filter(population == "Deep") 
+eck_long_df <- gather(eck_df, Morphology, measurement, frond_mass:total_length, factor_key=TRUE)
+
+# Summary stats to make it easier to refer to later, these will not be used any analysis.
+#site_list <- c("Kommetjie")
+
+morph_data <- eck_long_df %>% 
+  group_by(site, Morphology) %>% 
+  summarise(maximum = max(measurement, na.rm = T)) %>% 
+  ungroup() %>% 
+  rename(Site = site) %>% 
+  rename(Maximum = maximum) %>% 
+  spread(Morphology, Maximum)
+
+# Add in and/calculate estimates
+morph_data$stipe_radius <- with(morph_data, stipe_circ/(pi*2))
+morph_data$frond_width <- with(morph_data, 5)
+morph_data$frond_thickness <- with(morph_data, 0.2)
+morph_data$bulb_length <- with(morph_data, 10)
+morph_data$bulb_base_radius <- with(morph_data, 5)
+morph_data$top_radius <- with(morph_data, 3)
+morph_data$bottom_radius <- with(morph_data, 10)
+morph_data$stipe_radius <- with(morph_data, stipe_circ/(pi*2))
+morph_data$height <- with(morph_data, 5)
+
+# Define the value of pi for ease
+pi <- 3.14
+
+# Calculate the total cross-sectional areas for each site from all morphological characteristics
+morph_data$total_Ac <- with(morph_data, ((2*frond_length*frond_width) + (2*frond_length*frond_thickness) + (2*frond_width*frond_thickness) + (primary_length * primary_width)/2 + (4*pi*bulb_base_radius*2) + (2*pi*bulb_base_radius*bulb_length) + 2*pi*stipe_radius*(stipe_radius*stipe_length) +  (pi*(2*top_radius + 2*bottom_radius) + pi*(top_radius + bottom_radius)*pi*(top_radius - bottom_radius)*2 + height*2))/1000)
+
+# Make table based on new entries
+site_surface_area <- morph_data
+site_surface_area <- flextable(data = site_surface_area)
+site_surface_area <- theme_zebra(site_surface_area, odd_header = "transparent", even_header = "transparent")
+site_surface_area <- align(x= site_surface_area, align = "center", part = "all")
+#area_table <- set_header_labels(area_table, Plant.characteristic = "Plant characteristic", Approximate.shape = "Approximate shape", Equation.used = "Equation Used", Plant.dimensions = "Plant dimensions")
+#area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
+site_surface_area <- set_table_properties(site_surface_area, width = 1, layout = "autofit")
+#site_surface_area <- set_caption(site_surface_area, "Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.")
+site_surface_area
+```

### 2021-02-16 thesis_chapter_3.Rmd
CR comments added. Updated map and morph fig added. more to come...

@@
 output:
-  html_document:
   word_document:
     reference_docx: markdown/frontiers_template.docx
+  html_document:
   pdf_document:
@@
 library(rgdal)
-library(tidyverse)
 library(gridExtra)
@@
 
-There is a range of objects, both natural and anthropogenic, floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
+There are a range of objects, both natural and anthropogenic, floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
 Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
@@
 
-Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005, @thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
-Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still not known; although the role of wind has been recognised as important in several studies.
+Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
+Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown although the role of wind has been recognised as important in several studies.
 For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M.
 pyrifera* in nearshore Monterrey Bay.
-The results showed that kelp rafts with little surface area exposed to the wind were largely driven by a combination of wind and wind waves, however the relative importance of wind and wind waves was not clear.
+The results showed that kelp rafts, with little surface, area exposed to the wind were largely driven by a combination of wind and wind waves, however the relative importance of wind and wind waves was not clear.
 In addition, the tagged kelp trajectories were more consistent with the formation of eddies during winter.
 Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems.
-For example the subAntartic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating kelp.
+For example, the subAntartic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating kelp.
 Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013].
@@
 Although various approaches have been used in the past to investigate floating macroalagae trajectory, very few studies have employed the use of Langrangian trajectory modeling.
-Furthermore, non of the existing studies using this modeling approach have considered macroalgal morphology cross-sectional area (i.e.
+Furthermore, none of the existing studies using this modeling approach have considered macroalgal morphology cross-sectional area (i.e.
 shape) as an aspect of drag and ultimately trajectory.
@@
 # Methods
-
 ## Study site
@@
 A release site was chosen for this study, Kommetjie, which is located on the Atlantic side of the Cape Peninsula (Figure 2B).
-The hydrodynamics in the study region are driven by a complex interaction of wind and waves.
-which are also influenced by larger oceanic processes In South Africa, the larger processes can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC).
-The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean.
+The hydrodynamics in the study region are driven by a complex interaction of wind and waves, which are also influenced by larger oceanic processes.
+In South Africa, the larger processes can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC)[@lutjeharms2000;@lutjeharms2007;@garzoli1996;@Hutchings2009;@rubio2009].
+The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@@lutjeharms2000;@lutjeharms2007].
 The region where the retroflection occurs is known as the Agulhas retroflection and consists of some of the highest levels of mesoscale variability [@garzoli1996].
-The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988; @lutjeharms2006; @@lutjeharms2007].
-The Agulhas Rings are the main contributors to the Agulhas leakage which is the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015].
+The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988; @lutjeharms2006; @lutjeharms2007].
+The Agulhas Rings are the main contributors to the Agulhas leakage which are the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015].
 The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009].
-The Benguela region is characterised by strong coastal upwelling cells which a primarily driven by equatorward winds [@shannon1996; @hardman2003], and is known as the Benguela upwelling region.
+The Benguela region is characterised by strong coastal upwelling cells which are primarily driven by equatorward winds [@shannon1996; @hardman2003], and is known as the Benguela upwelling region.
 The Benguela upwelling region starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996].
-The region is also characterised by high amounts of mesoscale variability in the form of eddies and filaments[@blanke2002; @blanke2005; @rubio2009].
+The region is also characterised by high amounts of mesoscale variability in the form of eddies and filaments [@blanke2002; @blanke2005; @rubio2009].
 The Benguela inshore region is dominated by a northwestward flow as a result of topographical steering, wind stress and interactions with passing Agulhas rings and eddies [@veitch2010].
@@
 
-map_chp3A <- rasterGrob(as.raster(readPNG("map/chp3_mapA.png")))
-map_chp3B <- rasterGrob(as.raster(readPNG("map/chp3_mapB.png")))
-
-ggarrange(map_chp3A, map_chp3B, ncol = 1, nrow = 2, labels = "AUTO")
+#map_chp3A <- rasterGrob(as.raster(readPNG("map/chp3_mapA.png")))
+#map_chp3B <- rasterGrob(as.raster(readPNG("map/chp3_mapB.png")))
+#ggarrange(map_chp3A, map_chp3B, ncol = 2, nrow = 1, labels = "AUTO", align = c("h"))
+map_chp3_combined <- rasterGrob(as.raster(readPNG("map/chp3_map_combined_hz.png")))
+ggarrange(map_chp3_combined)
 ```
@@
 $$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
-where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model, $X_b(t)$ is a change in position due to 'behavior' of the particle.
+where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model at particle point in time, $t$ is time, $X_b(t)$ is a change in position due to 'behavior' of the particle at a particular point in time.
 This can range from swimming in fish to sinking or beaching.
-IN order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.
+In order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.
 
 This study simulated two scenarios; one where the plant is fully submerged and only exposed to hydrodynamic drag; and the other where parts of the plants are partially exposed resulting in a combination of hydrodynamic and air form drag.
-Drag based on shape based coefficients were used in the calculations for determining the overall kelp velocity vectors.
+Drag based on shape-based coefficients were used in the calculations for determining the overall kelp velocity vectors.
 A momentum energy equation was used to calculate the drag forces for the relevant simulation.
@@
 The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was, $$m({d_u/d_t})=F_d + F_a + F_f$$
-where $m$ is kelp mass, $d_u$ the virtual kelp particle velocity, $F_d$ the hydrodynamic drag, $F_a$ is the wind drag, and $F_f$ the surface-current flow.
+where $m$ is kelp mass, $d_u$ the virtual kelp particle velocity, $F_d$ the hydrodynamic drag, $F_a$ is the wind drag, $F_f$ the surface-current flow, and $d_t$ is the time step.
 The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations.
@@
 
-## Hydrodynamic and wind model
-
-Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service.
-The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
-The model has a spatial resolution of 0.08$^\circ$ and is interpolated on a Arakawa C native grid.
-
-In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (WIND_L4_NRT_OBSERVATIONS_012_004).
-The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
-Both averaged ocean current and wind data are stored as meridonal and zonal velocity vectors.
-
 ## Simulations
 
-Virtual particles were released with no drag or wind effects from the study site in order to act as a basis for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'.
-All simulations have the same release location and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme.
+Virtual particles were released with no drag or wind effects from the study site in order to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'.
+All simulations have the same release location and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018].
 Brownian motion was also incorporated in order to account for any stochasticity not resolved by the model.
-These are standard practices for Lagrangian trajectory studies.
+These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018].
+To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios.
+These scenarios were meant to act as proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag.
+These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation.
+For example, if 85% of the plant is submerged and 15% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively.
+The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1).
 
@@
 In order to determine the direct effect of wind on trajectory, simulations were run for varying magnitudes of wind exposure for each type of cross-sectional area.
+The magnitudes of exposure where based on personal observations of drifting *E.
+maxima* from previous studies and fieldwork.
 The virtual particles released with the hydrodynamic or wind drag components are referred to as 'kelp particles'.
@@
 ## Model inputs
+### Ocean and wind model
+
+Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service.
+The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
+The model has a spatial resolution of 0.08$^\circ$ and is interpolated on an Arakawa C native grid.
+
+In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (WIND_L4_NRT_OBSERVATIONS_012_004).
+The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
+Both averaged ocean current and wind data are stored as meridonal and zonal velocity vectors.
+
+### Hydrodynamic and wind drag
+
+The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
+The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
+Hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors.
+Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step.
+The same approach was used for the wind drag force.
+The equation that was used to calculate hydrodynamic drag, was also used to calculate wind.
+
 ### Cross-sectional area
@@
 In order to incorporate hydrodynamic and wind drag, the cross-sectional area of the kelp was calculated first.
-Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant, for details please refer to table 2.
+Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant, for details please refer to figure 2.
 The dimensional data needed was estimated in cases were data was not available for that particular morphological characteristic.
@@
 In terms of the holdfast, a standard cross-sectional area was used for all simulations.
+
+```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2:Diagram of the *E.maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic."}
+
+kelp_shape_figure <- rasterGrob(as.raster(readPNG("figures/chapter_3/chp3_shapedrawing.png")))
+ggarrange(kelp_shape_figure)
+
+```
+
 The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
-The minimum, mean and maximum were calculated for each morphological characteristic (see appendix), which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations.
+The minimum, mean and maximum were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (see table 1).
 The trajectory of these "types" of cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
@@
 ```
-
-### Hydrodynamic and wind drag
-
-The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
-The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
-Hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors.
-Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step.
-The same approach was used for the wind drag force.
-The equation that was used to calculate hydrodynamic drag, was also used to calculate wind.
-To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios.
-These scenarios were meant to act as proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag.
-These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation.
-For example, if 80% of the plant is submerged and 20% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively.
-The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1).
-
 ## Analysis
@@
   geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Min_Kfloat_H095_W05_mean, aes(x = lon, y = lat), color = 'snow4')+
+  geom_path(data = Min_Kfloat_H095_W05_mean, aes(x = lon, y = lat), color = 'grey')+
   geom_path(data = Min_Kfloat_H090_W10_mean, aes(x = lon, y = lat), color = 'green')+
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Figure 2: Comparison of mean trajectories.
-Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag."}
+Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
+In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10.
+In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13."}
 ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO")
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density of particles within each grid cell for the end run time of each scenario.
-Plots A-D repressent the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
+Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
 The bottom plot is the density plot for the passive simulation.', fig.height=10.2, fig.width=7.5}

### 2021-02-16 thesis_chapter_3.Rmd
figures updated according to CR comments

@@
 The bulb/pneumatocyst is a highly variable morphological characteristic and in some cases can appear absent, the same is true for the holdfast.
-Therefore, the cross-sectional area of the bulb/pneumatocyst was not considered in the calculation of overall cross-sectional area.
-In terms of the holdfast, a standard cross-sectional area was used for all simulations.
+Therefore, a standard cross-sectional area was used for both the bulb and holdfast, for all simulations.
 
@@
   ylab("Latitude") + 
-  ggtitle("Passive") +
+  ggtitle("1") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") + 
-  ggtitle("H100W100") +
+  #ggtitle("H100W100") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("H095W05") +
+  #ggtitle("H095W05") +
   theme_bw() +
@@
   ylab("Latitude") + 
-  ggtitle("H090W10") + 
+  #ggtitle("H090W10") + 
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") + 
-  ggtitle("H085W15") + 
+  #ggtitle("H085W15") + 
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-figure_a <- ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, labels = c("AUTO"), common.legend = TRUE, legend = c("none"))
+figure_a <- ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, labels = c("2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"))
 
@@
   xlab("Type")+
-  coord_flip() +
+  #coord_flip() +
   theme_bw() +
   theme(plot.title = element_text(hjust = 0.5))
+
+custom_labs <- c("1","2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot
+
+dis_plot <- dis_plot +
+  scale_x_discrete(labels= custom_labs)
 

### 2021-02-16 thesis_chapter_3.Rmd
changed label

@@
   ylab("Distance (km)")+
-  xlab("Type")+
+  xlab("Scenario")+
   #coord_flip() +

### 2021-02-19 thesis_chapter_3.Rmd
chapter CR additions and extra additions

@@
 This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped inbetween or in the thalli [@dromgoole1982; @back2000].
-The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018] and the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
+The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @putman2018; @brooks2018] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
 
@@
 The authors suggested that wind and water-movement were the primary influences of trajectory, however, this was only inferred from the genetic results and local climatology data [@fraser2011].
-
-Other aspects such as buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material.
-However, the past research on macroalgal trajectory has not investigated these factors which have been shown to be important aspects of trajectory for other materials such as icebergs, marine craft and microplastics.
-The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag.
-Drag coefficients related to wind and surface current have been shown to be important properties to consider when estimating trajectory and forecasting drift for search and sea rescue operations.
-Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.
+Past research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
+The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag [@breivik2011; @griffin2017].
 In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object.
 If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].
+Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.
 
@@
 The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@@lutjeharms2000;@lutjeharms2007].
-The region where the retroflection occurs is known as the Agulhas retroflection and consists of some of the highest levels of mesoscale variability [@garzoli1996].
+The region where the retroflection occurs is known as the Agulhas retroflection and consists of high levels of mesoscale variability [@garzoli1996].
 The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988; @lutjeharms2006; @lutjeharms2007].
@@
 ```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1:Map of the study domain. This is a temporary map until issues with the map script are resolved."}
+
 library(png)
@@
 ggarrange(map_chp3_combined)
+
 ```
+
+The study domain falls within the southern Benguela upwelling ecosystem, which is a region of high mesoscale variability and activity.
+These occur mostly in the form of eddies, filaments and vortices which are generated from the Agulhas retroflection [@veitch2018; @veitch2017];  generally flowing northward along the coastline and begin to move offshore near Cape Columbine.
+The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, eckman transport and geostrophic currents [@stramma1989, @veitch2017; @veitch2018; @hutchings2009].
+The stability of the jet varies seasonally with more stable flows in winter and splitting of the jet during summer and spring.
+In January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2009] which results into a weak narrow inshore branch which flows along the shelf edge and a strong flowing north-westwards offshore branch.
+The area is also an area of high upwelling activity driven by strong westerly winds [@hutchings2009].
+These winds are most intense during summer (December-March) which favours upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet; which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
+In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
 
@@
 For example, if 85% of the plant is submerged and 15% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively.
+This approach has been used in past research investigating pelagic Sargassum rafts [@putman2020;@putman2018].
 The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1).
+Each scenario consisted of a particle set with a size of 1000 where released from Kommetjie at the start of simulation.
+The particles sets were advected with a Runge-Kutte fourth order scheme for 30 days and incorporated brownian motion.
+Since the aim of this investigation was to investigate the relative surface behaviour of virtual kelp particles, only the surface layer was used for simulations.
+It should also be noted that the Copernicus model used places velocity vectors over some of the land covering the coastal domain.
+This was not considered a concern, as once again, the aim was to investigate surface drift behavior.
+The time period chosen to run the simulations was 01-01-2018 un 30-01-2018.
+This particular time period was chosen as the splitting of the Benguela Jet occurs in Janurary, which also conincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity [@rubio2009; @veitch2018; @ragoasha2019].
 
 ```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Details for each simulation run for the study"}
+
 set_flextable_defaults(fonts_ignore=TRUE)
@@
 sim_table
+
 ```
@@
 The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
-Hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors.
+Hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors, as per @furnans2008.
 Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step.
 The same approach was used for the wind drag force.
-The equation that was used to calculate hydrodynamic drag, was also used to calculate wind.
+The equation that was used to calculate hydrodynamic drag, was also used to calculate wind drag.
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Figure 2: Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Comparison of mean trajectories.
 Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
@@
 The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @miron2020].
-Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], current research has focused on large macroalgae-rafts which can vary greatly in size and surface area.
+Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], past research has focused on large macroalgae-rafts which can vary greatly in size and surface area.
 None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae.
@@
 
-The addition of drag forces into the simulation also causes the particles to cluster together along the mean trajectory compared to that of the passive particles which are more evenly dispersed.
+The addition of drag forces causes the particles to cluster together along the mean trajectory compared to that of the passive particles which are more evenly dispersed.
 This is most likely due to the shared drag characteristics among the particles which included the plants weight and surface area into the calculation.
@@
 The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia.
-Although drag forces were used in this study and not inertia, the same conclusion applies, drag is an important characteristic to consider when investigating or predicting solitary drifting macroalgae.
+Although drag forces were used in this study and not inertia, drag is an important characteristic to consider when investigating or predicting solitary drifting macroalgae.
 
@@
 The results suggested that the hatchlings trajectories are mostly influenced by the ocean currents in the first year of hatching due to low swimming speeds of juveniles.
-Therefore, the differences in cross-sectional area types may not be significant enough in order to have a significant effect on the relevant velocity vectors (zonal and meridonal).
+Therefore, the differences in cross-sectional area types may not be significant enough in order to have an effect on the relevant velocity vectors (zonal and meridonal).
 Another possible reason is the resolution of the ocean model used.
@@
 Inertia is based on the size/weight of the raft, while hydrodynamic and wind drag are based on current velocity and surface area.
-Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag should be considered.
+Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered.
 The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.

### 2021-02-21 thesis_chapter_3.Rmd
more detail added to table and figure captions

@@
 
-```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1:Map of the study domain. This is a temporary map until issues with the map script are resolved."}
+```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1:Map of the study domain and the main features pertaining to transport in the region.
+In panel A, the map of the entire South African coastline and the main currents presented.
+The resulting eddies and filaments from the Agulhas retroflection and study area are presented.
+In panel B, the splitting of the Benguela Jet, eddie features and general flow of the study domain are presented."}
 
@@
 
-```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Details for each simulation run for the study"}
+```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Details for each simulation run for the study.
+The table cosists of parameterisations for each scenario.
+The ID coloumn refers to the syntax used in the relevant OceanParcels script.
+Type refers to either a passive particle with no drag behaviour or a kelp particle that has been advected with any drag exposure.
+The drag exposure column refers to the parameterisations regarding the level of water and wind drag simulated."}
 
@@
 
-```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation."}
+```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.
+Approximate shape refers to the dimensions of the geometric shape used to represent a particular morpholgical characteristic.
+Equations used are the equations relating to the dimensions used for the geometric shape estimate.
+Plant dimensions are the measurements used, collected in previous work by @coppin2020, used to calculate the associated surface areas.
+The total surface areas for each plant characteristic for each surface area type make up the remaining columns."}
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density of particles within each grid cell for the end run time of each scenario.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time of each scenario.
+The plots are regular hexagons within the density of particles is calculated.
 Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
-The bottom plot is the density plot for the passive simulation.', fig.height=10.2, fig.width=7.5}
+The bottom plot is the density plot for the passive simulation.
+The mean trajectories for each scenario are also plotted for comparison.', fig.height=10.2, fig.width=7.5}
 
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance of all particles from the release site over the course of the simulation for each exposure scenario."}
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario.
+The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.
+The lower and upper "hinges" correspond to the first and third quartiles.
+The whiskers represent the range, solid black lines represent the median and black dots are outliers."}
 dis_plot
@@
 ```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5: Comparison of trajectory data between a kelp sphere and kelp cylinder.
-Plot A is the comparison of mean trajectories, plot B is the comparison of distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.", fig.height=8, fig.width=8}
+Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.", fig.height=8, fig.width=8}
 

### 2021-02-21 thesis_chapter_3.Rmd
refer to table...added to captions

@@
 Virtual particles were released with no drag or wind effects from the study site in order to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'.
-All simulations have the same release location and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018].
+All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018].
 Brownian motion was also incorporated in order to account for any stochasticity not resolved by the model.
@@
 In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10.
-In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13."}
+In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
 ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO")
@@
 The bottom plot is the density plot for the passive simulation.
-The mean trajectories for each scenario are also plotted for comparison.', fig.height=10.2, fig.width=7.5}
+The mean trajectories for each scenario are also plotted for comparison.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
 
@@
 The lower and upper "hinges" correspond to the first and third quartiles.
-The whiskers represent the range, solid black lines represent the median and black dots are outliers."}
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
 dis_plot
@@
 ```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5: Comparison of trajectory data between a kelp sphere and kelp cylinder.
-Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.", fig.height=8, fig.width=8}
+Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=8}
 

### 2021-03-08 thesis_chapter_3.Rmd
additions to chapter 3

@@
 csl: markdown/frontiers.csl.txt
+abstract: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
+Macroalage are one of the most important natural dispersal mechanisms in the ocean, and are regarded as the tumble-weed of the ocean.
+Despite the many studies that exist, the relative role of wind and surface currents in macrolage trajectory is still unknown.
+Past studies also tend to focus on kelp-rafts which can vary greatly in size and making it difficult to form a concensus as to the role of wind and surface currents.
+In addition past studies have not conisidered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam.
+The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectory of flotsam in the ocean.
+This study aimed to shed light on the relative role of wind versus surface currents, as well as the role of drag, in macrolagae trajectory and accumulation.
+This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals.
+This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley lagrangian particles.
+The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scnearios causes accumulation in eddies and vortices within the study domain.
+The results also point to the importance of sub-mesoscale features, such as eddies and vortices, as important surface accumlation zones for solitary floating macroalage.
 ---
@@
 
+In addition, past studies focus on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddies and vortices in the accumulation of kelp at the surface.
+Gyres, eddies and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018].
+Past studies investigating the dispersal of marine debris have also included the possible effects of inertia in accumulation of ocean drifter instruments in mesoscale features.
+For example, a study by @beron2016 used a modeling approach to investigated the role of eckman transport and inertia to explain the observed behaviour of undrogued drifters in the ocean gyres.
+Undrogued drifter trajectory is affected by wind unlike the drogue counter parts which makes the effects of wind neglible.
+The results showed that the combined effect inertia, not eckman transport, played a significant role in the observed drifter behaviour [@beron2016].
+The authors also noted that the presence of other flotsam along with the undrogued drifters in ocean gyres also suggests that aspects inertia also play a role in accumulation of other marine debris [@beron2016].
+The approach by @beron2016 has also been improved upon by @beron2019 but also shows the significant role of inertia in flotsam trajectory and accumulation.
+
 In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using lagrangian based trajectory simulations.
+It should be noted that this study does not attempt to characterise the dispersal of drifting macroalage within the study domain.
+The authors argue that too little is know about the drift characteristics of solitary floating macroalgae to characterise dispersal effectivley.
+Instead, this study will aim to shed light on the 'behaviour' of surface drifting macroalgae from a coastal environment into the open ocean, which will inform future characterisation studies.
 This will be achieved by comparing the trajectory outputs of virtual 'passive' particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure.
 The different degrees of wave and wind exposure are meant to reflect the different buoyancy characteristics, as well as identify the role of magnitude of the drag components on overall trajectory.
+In addition, the modeling of single kelp offers the opportunity to study the role of eddies and vortices in accumulation points on the surface.
+Therefore, separate simulations will be run to compare the entrained characteristics of passive particle with a kelp particle.
 
@@
 The study domain falls within the southern Benguela upwelling ecosystem, which is a region of high mesoscale variability and activity.
-These occur mostly in the form of eddies, filaments and vortices which are generated from the Agulhas retroflection [@veitch2018; @veitch2017];  generally flowing northward along the coastline and begin to move offshore near Cape Columbine.
-The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, eckman transport and geostrophic currents [@stramma1989, @veitch2017; @veitch2018; @hutchings2009].
+These occur mostly in the form of eddies, filaments and vortices which are generated from the Agulhas retroflection [@veitch2018; @veitch2017], generally flowing northward along the coastline and begin to move offshore near Cape Columbine.
+The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, eckman transport and geostrophic currents [@stramma1989; @veitch2017; @veitch2018; @hutchings2009].
 The stability of the jet varies seasonally with more stable flows in winter and splitting of the jet during summer and spring.
@@
 In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
-
 ## Particle tracking model
@@
 Virtual particles were released with no drag or wind effects from the study site in order to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'.
+All particles that have any form of drag exposure are hereafter referred to as 'kelp' particles.
 All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018].
@@
 The table cosists of parameterisations for each scenario.
-The ID coloumn refers to the syntax used in the relevant OceanParcels script.
+The ID coloumn refers to the syntax used in the relevant OceanParcels script (see additional material for link to script).
 Type refers to either a passive particle with no drag behaviour or a kelp particle that has been advected with any drag exposure.
@@
 
-In order to determine the influence of cross-sectional area on drag components, simulations were performed for the minimum, mean and maximum cross-sectional areas for Kommetjie kelp forest population (Figure 1).
-In order to determine the direct effect of wind on trajectory, simulations were run for varying magnitudes of wind exposure for each type of cross-sectional area.
-The magnitudes of exposure where based on personal observations of drifting *E.
-maxima* from previous studies and fieldwork.
-The virtual particles released with the hydrodynamic or wind drag components are referred to as 'kelp particles'.
+To determine if vortices and eddies play a role in kelp accumulation on the surface, a separate set of simulations were performed.
+Unlike the previous simulations these simulation released a particle every hour.
+These were run over the same time period for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme and incorporated Brownian motion.
+In order to determine if the shape co-efficient used plays a role in model used, a simulation of a kelp with the maximum surface area and drag exposure was run using the shape coefficient for a cylinder to compare to a simulation using a sphere, and was run for 30 days using a fourth order Runge-Kutte advection scheme with brownian motion and a particle set of size 1000.
 
@@
 The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.
-The lower and upper "hinges" correspond to the first and third quartiles.
+The lower and upper hinges correspond to the first and third quartiles.
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.
@@
 ```
+## Accumulation
+
+```{r message=FALSE, echo=FALSE, warning=FALSE}
+
+kelp_acc <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+kelp_acc_mean <- kelp_acc %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+kelp_acc$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_mean$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+pass_acc <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+pass_acc_mean <- pass_acc %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+pass_acc$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_mean$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+total_acc_data <- rbind(pass_acc, kelp_acc)
+total_acc_mean <- rbind(pass_acc_mean, kelp_acc_mean)
+
+mean_traj_acc <- ggplot() +
+  geom_path(data = total_acc_mean, aes(x = lon, y = lat, colour = type)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+heatmap_RK4_acc <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Passive particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_kelp_acc <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Kelp particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_RK4_acc <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = pass_acc, aes(x = lon, y = lat, colour = distance)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_kelp_acc <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Age")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_kelp_acc <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  #scale_color_viridis_c(name = "Age") +
+  guides(fill=guide_legend(title="Age")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+dis_plot_acc <- ggplot(data = total_acc_data, aes(x = type, y = distance, show.legend = FALSE)) +
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(limits = pos)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = "right") +
+  theme_bw() +
+  ylab("Distance (km)") +
+  xlab("Scenario") +
+  #coord_flip() +
+  theme_bw()
+  #theme(plot.title = element_text(hjust = 1))
+
+```
+
+```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 6: Comparison of trajectory data between passive and kelp accumulation scenarios.
+Panel A are a comparison of density plots  and panel B is a comparison of age.", fig.height=8, fig.width=8}
+
+plot_A <- ggarrange(heatmap_RK4_acc, heatmap_kelp_acc, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 1)
+plot_B <- ggarrange(point_RK4_acc, point_kelp_acc, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 1)
+
+ggarrange(plot_A, plot_B, labels = c("AUTO"), ncol = 1, nrow = 2)
+
+```
+
+The role of eddies that exist within the study domain over the time period play a clear role in the accumulation and dispersal of kelp particles.
+Passive particles tend to remain on the outer edge of the eddy and then are expelled on the other side and continue to follow the ocean current (see Figure 5A).
+The opposite is true for kelp particles which have the highest exposure level of hydrodynamic and wind drag.
+The kelp particles are entrained in the eddy (see Figure 5A).
+This is also reflected when comparing the age of particles at the end of the simulation.
+In Figure 5B, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
 
@@
 
-The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no dissimilarity (Figure 5A).
-This was also reflected when comparing distance travelled by particles and particle density plots (Figure 5B).
-The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (Figure 5C-D).
+The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no dissimilarity (Figure 6A).
+This was also reflected when comparing distance travelled by particles and particle density plots (Figure 6B).
+The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (Figure 6C-D).
 
@@
 
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5: Comparison of trajectory data between a kelp sphere and kelp cylinder.
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 6: Comparison of trajectory data between a kelp sphere and kelp cylinder.
 Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.
@@
 In addition, not only are the trajectories similar but so are the end points at the end of simulation.
-However, when ad-hoc wind drag (windage) is considered in combination with hydrodynamic drag the trajectories differ greatly when compared to purely lagrangian particle trajectories.
+However, when ad-hoc wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely lagrangian particle trajectories.
 The inclusion of wind drag causes the particles to flow further away from the coast and with a greater distance travelled from the release location.
@@
 The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia.
-Although drag forces were used in this study and not inertia, drag is an important characteristic to consider when investigating or predicting solitary drifting macroalgae.
 
 The ocean is made up of varying time-varying flows and changes in an objects velocity can cause the object to be exposed to different flow patterns over time.
-The lack of dissimilarity between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible.
+The lack of dissimilarity in trajectory between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible.
 A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region.
-The results suggested that the hatchlings trajectories are mostly influenced by the ocean currents in the first year of hatching due to low swimming speeds of juveniles.
+The results suggested that the hatchlings trajectories are mostly influenced by the ocean currents in the first year of hatching due to  as juveniles do are not swim fast enough to alter thier overall trajectory.
 Therefore, the differences in cross-sectional area types may not be significant enough in order to have an effect on the relevant velocity vectors (zonal and meridonal).
@@
 The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.
+
+The role of eddies and vortices in the entrainment of kelp is further supported by the accumulation simulations.
+These results cleary show that the inclusion of drag greatly increases the number of particles entrained in the eddy within the study domain.
+Eddies and vortices may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in area below the eddy.
+In a study by @woodborne1989 references a site offshore approximately 120km from Cape Columbine where dredged kelp was found.
+The area referenced by @woodborne1989 coincides with the area where virtual kelp are entrained by a vorticy.
 The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory.
@@
 
-```{r, message=FALSE, warning = FALSE, include= TRUE, echo = FALSE, fig.height=8, fig.width=8}
+```{r eval=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=TRUE}
 

### 2021-03-08 thesis_chapter_4.Rmd
chapter 4 added and started

@@
 library(flextable)
+library(tidync)
+library(ncdf4)
+library(lubridate)
+library(gridExtra)
+library(ggpubr)
 
@@
 
-# Define the value of pi for ease
+# Define the value of pi 
 pi <- 3.14
@@
 
+# Calculate the total mass of kelp individual based on site morphological characteristics
+
+morph_data$total_mass <- with(morph_data, (frond_mass*6) + (stipe_mass) )
+
 # Make table based on new entries
@@
 site_surface_area <- flextable(data = site_surface_area)
-site_surface_area <- theme_zebra(site_surface_area, odd_header = "transparent", even_header = "transparent")
+#site_surface_area <- theme_zebra(site_surface_area, odd_header = "transparent", even_header = "transparent")
 site_surface_area <- align(x= site_surface_area, align = "center", part = "all")
@@
 ```
+
+```{r, warning=FALSE, message=FALSE, echo = FALSE}
+
+houtbay <- tidync("OceanParcels/trajectory_data/chapter_4/houtbay.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+
+houtbay_mean <- houtbay %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+houtbay$site <- "Hout Bay" 
+houtbay_mean$site <- "Hout Bay"
+
+kommetjie <- tidync("OceanParcels/trajectory_data/chapter_4/kommetjie.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+
+kommetjie_mean <- kommetjie %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+kommetjie$site <- "kommetjie" 
+kommetjie_mean$site <- "kommetjie"
+
+soetwater <- tidync("OceanParcels/trajectory_data/chapter_4/soetwater.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+
+soetwater_mean <- soetwater %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+soetwater$site <- "Soetwater" 
+soetwater_mean$site <- "Soetwater"
+
+millers <- tidync("OceanParcels/trajectory_data/chapter_4/millers.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+
+millers_mean <- millers %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+millers$site <- "millers" 
+millers_mean$site <- "millers"
+
+total_data <- rbind(houtbay, soetwater, kommetjie, millers)
+total_data_mean <- rbind(houtbay_mean, soetwater_mean,kommetjie_mean, millers_mean)
+```
+
+## Mean trajectory
+
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
+
+library(ggplot2)
+
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+mean_traj_site <- ggplot() + 
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_path(data = total_data_mean, aes(x = lon, y = lat, colour = site)) +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+mean_traj_site
+
+```
+
+## Density comparisons
+
+```{r, message=FALSE, echo=FALSE, warning=FALSE}
+
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+heatmap_houtbay <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = houtbay, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 130000)) +
+  geom_path(data = houtbay_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("HoutBay") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_komm <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = kommetjie, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 130000)) +
+  geom_path(data = kommetjie_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Kommetjie") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_soetwater <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = soetwater, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 130000)) +
+  geom_path(data = soetwater_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Soetwater") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_soetwater <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = soetwater, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 130000)) +
+  geom_path(data = soetwater_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Kommetjie") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_millers <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = millers, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 130000)) +
+  geom_path(data = millers_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Millers") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+ggarrange(heatmap_houtbay, heatmap_komm, heatmap_soetwater, heatmap_millers, labels = "AUTO", common.legend = TRUE)
+
+```

### 2021-03-08 map/map_chapter3.R
additional files chapter 3

@@ -1,0 +1,41 @@
+# remember to load software from parent script
+
+load("map/south_africa_coast_hi_res.Rdata")
+source("map/scaleBarFunc.R")
+release_sites_chp3 <- read.csv("metadata/release_sites_chp3.csv", sep = ";")
+
+
+map_sa <- ggplot(data = south_africa_coast_hi_res) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = "gray") +
+  geom_rect(data = south_africa_coast_hi_res,aes(xmin=14, xmax=20, ymin=-30, ymax=-35), inherit.aes = FALSE, fill = NA, colour =    "black", size = 1.5) +
+  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
+  #coord_map("mercator", clip = "on") +
+  #coord_cartesian(xlim = c(15, 20), ylim = c(-30, -35)) +
+  #scale_x_continuous(trans = "reverse") +
+  #theme(axis.ticks = element_blank()) + # Remove tick marks
+  theme_bw() +
+  xlab("Latitude") +
+  ylab("Longitude")
+map_sa
+
+ggsave(filename = "map_sa.jpg", plot = last_plot(), device = "jpeg", path = "map/", dpi = 320, width = 10, height = 10)
+
+## Make 2nd map of just the peninsula
+map_pen <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
+  coord_map("mercator", clip = "on") +
+  scaleBar(lon = 15, lat = -34.9, distanceLon = 20, distanceLat = 20, distanceLegend = -5, dist.unit = "km",
+           arrow.length = 40, arrow.distance = 20, arrow.North.size = 5) +
+  geom_point(data = release_sites_chp3, aes(x = lon, y = lat), color = "red", nudge_x = 0, nudge_y = 0.1) +
+  theme_bw() +
+  xlab("Latitude") +
+  ylab("Longitude") +
+  coord_cartesian(xlim = c(15, 20), ylim = c(-35, -30))
+  #theme(axis.ticks = element_blank(), # Remove tick marks
+  #     axis.text = element_blank(), # Remove lat/ lon numbers
[…]

### 2021-03-11 thesis_chapter_3.Rmd
added July data set and figures. more to follow

@@
 ```{r density plots, message=FALSE, warning=FALSE, echo=FALSE}
-## Heatmaps of study domain, use previous data for study site map and histograms
-
-## Coastline of Southern Africa
+## Heatmaps of study domain, use previous data for study site map and histograms## Coastline of Southern Africa
 load("map/south_africa_coast_hi_res.Rdata")
@@
 
+## Additional simulations
+
+```{r additional sims, message=FALSE, echo=FALSE, warning=FALSE}
+
+# Read in simulation data and data wrangling
+
+## July
+
+p_RK4_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+p_RK4_mean_July <- p_RK4_July %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+p_RK4_July$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
+
+Min_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H100_W00_mean_July <- Min_Kfloat_H100_W00_July %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H100_W00_July$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_July$velocity <- with(Min_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Min_Kfloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H095_W05_mean_July <- Min_Kfloat_H095_W05_July %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H095_W05_July$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_July$velocity <- with(Min_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Min_Kfloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H090_W10_mean_July <- Min_Kfloat_H090_W10_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H090_W10_July$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_July$velocity <- with(Min_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Min_Kfloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H085_W15_mean_July <- Min_Kfloat_H085_W15_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H085_W15_July$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_July$velocity <- with(Min_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_Kfloat_H100_W00_mean_July <- Mean_Kfloat_H100_W00_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_Kfloat_H100_W00_July$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_July$velocity <- with(Mean_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_KFloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H095_W05_mean_July <- Mean_KFloat_H095_W05_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H095_W05_July$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_July$velocity <- with(Mean_KFloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_KFloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H090_W10_mean_July <- Mean_KFloat_H090_W10_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H090_W10_July$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_July$velocity <- with(Mean_KFloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_KFloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H085_W15_mean_July <- Mean_KFloat_H085_W15_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H085_W15_July$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_July$velocity <- with(Mean_KFloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H100_W00_mean_July <- Max_Kfloat_H100_W00_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H100_W00_July$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_July$velocity <- with(Max_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H095_W05_mean_July <- Max_Kfloat_H095_W05_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H095_W05_July$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_July$velocity <- with(Max_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H090_W10_mean_July <- Max_Kfloat_H090_W10_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H090_W10_July$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_July$velocity <- with(Max_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H085_W15_mean_July <- Max_Kfloat_H085_W15_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H085_W15_July$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_July$velocity <- with(Max_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+total_data_July <- rbind(Min_Kfloat_H100_W00_July, Min_Kfloat_H095_W05_July, Min_Kfloat_H090_W10_July, Min_Kfloat_H085_W15_July, Mean_Kfloat_H100_W00_July, Mean_KFloat_H095_W05_July, Mean_KFloat_H090_W10_July, Mean_KFloat_H085_W15_July, Max_Kfloat_H100_W00_July, Max_Kfloat_H095_W05_July, Max_Kfloat_H090_W10_July, Max_Kfloat_H085_W15_July)
+
+## October
+
+p_RK4_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+p_RK4_mean_Oct <- p_RK4_Oct %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+p_RK4_Oct$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
+
+Min_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H100_W00_mean_Oct <- Min_Kfloat_H100_W00_Oct %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H100_W00_Oct$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_Octvelocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Min_Kfloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H095_W05_mean_Oct <- Min_Kfloat_H095_W05_Oct %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H095_W05_Oct$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Min_Kfloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H090_W10_mean_Oct <- Min_Kfloat_H090_W10_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H090_W10_Oct$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Min_Kfloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H085_W15_mean_Oct <- Min_Kfloat_H085_W15_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H085_W15_Oct$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_Kfloat_H100_W00_mean_Oct <- Mean_Kfloat_H100_W00_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_Kfloat_H100_W00_Oct$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_KFloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H095_W05_mean_Oct <- Mean_KFloat_H095_W05_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H095_W05_Oct$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_KFloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H090_W10_mean_Oct <- Mean_KFloat_H090_W10_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H090_W10_Oct$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_KFloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H085_W15_mean_Oct <- Mean_KFloat_H085_W15_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H085_W15_Oct$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H100_W00_mean_Oct <- Max_Kfloat_H100_W00_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H100_W00_Oct$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H095_W05_mean_Oct <- Max_Kfloat_H095_W05_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H095_W05_Oct$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H090_W10_mean_Oct <- Max_Kfloat_H090_W10_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H090_W10_Oct$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H085_W15_mean_Oct <- Max_Kfloat_H085_W15_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H085_W15_Oct$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+
+total_data_Oct <- rbind(Min_Kfloat_H100_W00_Oct, Min_Kfloat_H095_W05_Oct, Min_Kfloat_H090_W10_Oct, Min_Kfloat_H085_W15_Oct, Mean_Kfloat_H100_W00_Oct, Mean_KFloat_H095_W05_Oct, Mean_KFloat_H090_W10_Oct, Mean_KFloat_H085_W15_Oct, Max_Kfloat_H100_W00_Oct, Max_Kfloat_H095_W05_Oct, Max_Kfloat_H090_W10_Oct, Max_Kfloat_H085_W15_Oct)
+
+```
+
+```{r, message=FALSE, echo=FALSE, warning=FALSE}
+
+waterdrag_plot_July <- ggplot()+
+  geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = Mean_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = Mean_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Max_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+wwdrag_plot_July <- ggplot()+
+  geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = Min_Kfloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'grey')+
+  geom_path(data = Min_Kfloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'green')+
+  geom_path(data = Min_Kfloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'pink')+
+  geom_path(data = Mean_KFloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'yellow')+
+  geom_path(data = Mean_KFloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = Max_Kfloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = Max_Kfloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Max_Kfloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Comparison of mean trajectories.
+Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
+In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10.
+In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
+
+ggarrange(waterdrag_plot_July, wwdrag_plot_July, labels = "AUTO")
+```
+```{r density plots, message=FALSE, warning=FALSE, echo=FALSE}
+## Heatmaps of study domain, use previous data for study site map and histograms## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+heatmap_RK4_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = p_RK4_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H100_W00_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H100W100") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H095_W05_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  #ggtitle("H095W05") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H090_W10_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H090W10") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H085_W15_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H085W15") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_Kfloat_H100_W00_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_path(data = Mean_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H095_W05_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H090_W10_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H085_W15_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H100_W00_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H100_W00_mean_July ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H095_W05_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  theme(plot.title = element_text(hjust = 0.5)) + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H090_W10_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H085_W15_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time of each scenario.
+The plots are regular hexagons within the density of particles is calculated.
+Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
+The bottom plot is the density plot for the passive simulation.
+The mean trajectories for each scenario are also plotted for comparison.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
+
+figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"))
+
+figure_b_July <- ggarrange(heatmap_RK4_July, legend = c("right"))
+
+ggarrange(figure_a_July, figure_b_July, ncol = 1, nrow = 2)
+
+```
+
 # References

### 2021-03-12 thesis_chapter_3.Rmd
updating chapter 3. additional sims

@@
 ```
+
 ## Accumulation
@@
 Min_Kfloat_H100_W00_Oct$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H100_W00_Octvelocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Comparison of mean trajectories.
+waterdrag_plot_Oct <- ggplot()+
+  geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Max_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+wwdrag_plot_Oct <- ggplot()+
+  geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = Min_Kfloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'grey')+
+  geom_path(data = Min_Kfloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'green')+
+  geom_path(data = Min_Kfloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'pink')+
+  geom_path(data = Mean_KFloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'yellow')+
+  geom_path(data = Mean_KFloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Mean_KFloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = Max_Kfloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = Max_Kfloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Max_Kfloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure ??: Comparison of mean trajectories.
 Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
@@
 
-ggarrange(waterdrag_plot_July, wwdrag_plot_July, labels = "AUTO")
-```
+ggarrange(waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, labels = "AUTO")
+
+```
+
 ```{r density plots, message=FALSE, warning=FALSE, echo=FALSE}
@@
 load("map/south_africa_coast_hi_res.Rdata")
+
+## July
 
@@
   theme(plot.title = element_text(hjust = 0.5))
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time of each scenario.
+
+## October
+
+heatmap_RK4_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = p_RK4_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H100_W00_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H100W100") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H095_W05_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  #ggtitle("H095W05") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H090_W10_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H090W10") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H085_W15_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H085W15") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_Kfloat_H100_W00_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H095_W05_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H090_W10 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H085_W15_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H100_W00_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H095_W05_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  theme(plot.title = element_text(hjust = 0.5)) + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H090_W10_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H085_W15_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+```
+
+## Density comparison
+### July
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure ??: Comparison of density plots at the end run time of each scenario for July 2018 (winter).
 The plots are regular hexagons within the density of particles is calculated.
@@
 ```
+### October
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure ??: Comparison of density plots at the end run time of each scenario for October 2018 (spring).
+The plots are regular hexagons within the density of particles is calculated.
+Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
+The bottom plot is the density plot for the passive simulation.
+The mean trajectories for each scenario are also plotted for comparison.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
+
+figure_a_Oct <- ggarrange(heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"))
+
+figure_b_Oct <- ggarrange(heatmap_RK4_Oct, legend = c("right"))
+
+ggarrange(figure_a_Oct, figure_b_Oct, ncol = 1, nrow = 2)
+
+```
+## Comparison of distances
+
+```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE, fig.height=3, fig.width=4}
+## July
+
+pos <- c("RK4_July",
+         "Min_Kfloat_H100_W00_July",
+         "Min_Kfloat_H095_W05_July",
+         "Min_Kfloat_H090_W10_July",
+         "Min_Kfloat_H085_W15_July",
+         "Mean_Kfloat_H100_W00_July",
+         "Mean_Kfloat_H095_W05_July",
+         "Mean_Kfloat_H090_W10_July",
+         "Mean_Kfloat_H085_W15_July",
+         "Max_Kfloat_H100_W00_July",
+         "Max_Kfloat_H095_W05_July",
+         "Max_Kfloat_H090_W10_July",
+         "Max_Kfloat_H085_W15_July")
+
+dis_plot_July <- ggplot(data = total_data_July, aes(x = type, y = distance, show.legend = FALSE))+
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(limits = pos)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif")+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Scenario")+
+  #coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+custom_labs_July <- c("1","2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot
+
+dis_plot_July <- dis_plot_July +
+  scale_x_discrete(labels= custom_labs_July)
+
+## Oct
+
+pos <- c("RK4_Oct",
+         "Min_Kfloat_H100_W00_Oct",
+         "Min_Kfloat_H095_W05_Oct",
+         "Min_Kfloat_H090_W10_Oct",
+         "Min_Kfloat_H085_W15_Oct",
+         "Mean_Kfloat_H100_W00_Oct",
+         "Mean_Kfloat_H095_W05_Oct",
+         "Mean_Kfloat_H090_W10_Oct",
+         "Mean_Kfloat_H085_W15_Oct",
+         "Max_Kfloat_H100_W00_Oct",
+         "Max_Kfloat_H095_W05_Oct",
+         "Max_Kfloat_H090_W10_Oct",
+         "Max_Kfloat_H085_W15_Oct")
+
+dis_plot_Oct <- ggplot(data = total_data_Oct, aes(x = type, y = distance, show.legend = FALSE))+
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(limits = pos)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif")+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Scenario")+
+  #coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+custom_labs_Oct <- c("1","2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot
+
+dis_plot_Oct <- dis_plot_Oct +
+  scale_x_discrete(labels= custom_labs_Oct)
+
+```
+
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring).
+The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.
+The lower and upper hinges correspond to the first and third quartiles.
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
+
+ggarrange(dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 2)
+
+```
 

### 2021-03-12 thesis_chapter_2.tex
adding support files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  a4paper,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Using numerical modelling to investigate temperature and wave exposure as drivers of morphological characteristics of the kelps, Ecklonia maxima and Laminaria pallida},
  pdfauthor={R. Coppin, C. Rautenbach, T. Ponton and A.J Smit},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{Using numerical modelling to investigate temperature and wave exposure
as drivers of morphological characteristics of the kelps, Ecklonia
maxima and Laminaria pallida}
\author{R. Coppin, C. Rautenbach, T. Ponton and A.J Smit}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Macroalgal morphological variation is determined to a large extent by a
combination of environmental factors, with wave exposure and temperature
perhaps the main influences, as they are key environmental properties to
which a species becomes locally adapted. Macroalgae have shown to
exhibit different responses to different magnitudes of exposure to
waves, such as reduction in overall size and strength increasing traits.
In terms of temperature, warmer environments have been shown to reduce
the overall size of resident and transplanted species. However, none of
the past studies have identified specific wave and temperature metrics
responsible for the morphological adaptation macroalgae exhibit. Past
research has often used simple or two-dimensional models of wave
exposure, which do not take into account important aspects of the
nearshore environment such as wave breaking, refraction and diffraction.
Furthermore, past studies have often used satellite-derived datasets as
sources for temperature data, however, such data have been shown to have
large bias when applied to the nearshore environment. This study used in
situ temperature data and wave power metrics calculated from a
3D-numerical model to identify specific temperature and wave metrics
responsible for morphological adaptation of the kelp, Ecklonia maxima
and Laminaria pallida. Between temperature and wave exposure, the
results identify wave exposure as the main influencer of morphological
adaptation while identifying specific wave metrics. Furthermore, the
results show differences in wave metrics between species; and between
deep and shallow populations.'
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Kelps are a group of large seaweeds of the order Laminariales
(Ochrophyta), which despite their relatively low taxonomic diversity of
species in genera (Bolton 2010), form the basis of one of the most
productive ecosystems globally (Mann 1973; Krumhansl and Scheibling
2012). Kelps generally have a dependence on cool-temperate and arctic
seawater temperatures (Bolton 2010; Mohring et al. 2014; Cavanaugh et
al. 2011; Dayton 1985), and dominate the nearshore biomass within the
rocky shallow coasts in both hemispheres (Steneck et al. 2002). Their
size and complex morphology provide a heterogeneous habitat structure
(Steneck et al. 2002) that accommodate a multitude of turf and
sub-canopy seaweed species, and diverse assemblages of sessile and
mobile invertebrates and vertebrates (Mann 1973; Steneck et al. 2002),
each depending on a wide suite of ecological services provided by kelp
forests (Dayton 1985; Gaines and Roughgarden 1987; Paul and Steneck
1993; Levin 1994; {\textbf{???}}; Willis and Anderson 2003). Wave
exposure and temperature are regarded as important environmental drivers
of kelp forests, and play a role in their distribution (Gorman et al.
2013), abundance ({\textbf{???}}; Cavanaugh et al. 2011), diversity
(Wing et al. 2007; Wernberg and Goldberg 2008), composition (Dayton
1985; Leliaert et al. 2000; Harley et al. 2012; Norderhaug et al. 2012),
growth (Cousens 1982) and productivity ({\textbf{???}}; {\textbf{???}};
Krumhansl and Scheibling 2012).

Given that kelps are cool-temperate organisms which are vulnerable to
dislodgement under high wave exposure scenarios, increases in storm
frequency and magnitude, and changes in ocean temperature, pose a direct
threat to their survival; which is a concern considering that these
variables are known to be affected by climate change (Russo et al. 2014;
Reguero, Losada, and Méndez 2019; Weisse 2010; Rose, Smale, and Botting
2012; Rummukainen 2012). Warmer temperatures reduces the resilience of
kelp individuals to storm disturbance (T. Wernberg et al. 2010), causes
fragmentation through weakening tissues (Simonson, Scheibling, and
Metaxas 2015) and reduces growth and productivity (Bearham, Vanderklift,
and Gunson 2013; Gerard 1997; Gao et al. 2013; Zimmerman and Kremer
1986); while storms and high wave energy dislodge and break kelps
(Byrnes et al. 2011; Graham 2004; Seymour et al. 1989). Despite these
threats and disturbances kelps persist across a broad range of
environments which is largely due to their morphological plasticity
(Fowler-Walker, Wernberg, and Connell 2006; T. Wernberg and Vanderklift
2010). A study by (Wernberg et al. 2003) investigated the morphology of
Ecklonia radiata (C.Agardh) J.Agardh in order to quantify the
morphological variation and whether it was dependent on spatial
differences along the Australian coast. They found no correlation
between spatial distance and morphological similarity; rather the
morphology of the kelps was representative of multiple environmental
forcings on different morphological characters at different spatial
scales (Wernberg et al. 2003).

Although multiple environmental factors play a role in kelp
morphological adaptation, wave exposure and temperature have been
identified as the main influencers across various kelp species (Hurd
2000; Denny and Gaylord 2002; Thomsen, Wernberg, and Kendrick 2004;
Fowler-Walker, Connell, and Gillanders 2005; Wernberg and Thomsen 2005;
Mabin, Gribben, and Fischer 2013; {\textbf{???}}). For example, a study
by ({\textbf{???}}) compared the morphology of Ecklonia cava Kjellman
growing at warmer and cooler sites, and showed wrinkling of the blade
and reduced size to be a characteristic of higher temperature ranges.
Warmer temperature affects important kelp physiological processes such
as photosynthesis and respiration, which influences growth and
productivity and ultimately leads to a reduction in size (Bearham,
Vanderklift, and Gunson 2013; Gerard 1997; Gao et al. 2013; Zimmerman
and Kremer 1986). Past research has shown that in highly wave exposed
areas kelp morphology tends to take on characteristics which reduce
drag, increase strength and increase flexibility (Denny and Gaylord
2002; Hurd 2000; Fowler-Walker, Connell, and Gillanders 2005; Thomsen,
Wernberg, and Kendrick 2004). For example, a study by Wernberg and
Thomsen (2005) examined the consistency of wave exposure as a driver of
E. radiata across a broad geographic range, and showed trends toward
drag-reducing (small size, narrow laterals and blades, low spinosity)
and increased strength (large holdfast, thick stipe and thick blades and
lamina) at high wave exposure sites.

Measures of wave exposure in ecological studies often incorporate
integrative measures of hydrodynamic conditions at a particular site
which are based on cartographical models of exposure. Cartographical
models of wave exposure are ``fetch-based-models'' which measure the
length of open water associated with a particular site in a straight
line, and are regarded as simple measures of wave exposure. Advances in
numerical modeling based on physical/linear wave theory incorporate more
complexity (wind forcing, wave-wave interactions, wave breaking,
diffraction and variation in wave direction) into the models and allow
for a quantitative, reproducible approach for characterizing the
hydrodynamic environment. Although Mabin, Gribben, and Fischer (2013)
investigated kelp morphological characteristics using a 2D WAM model,
currently no studies investigating macroalgal morphological
characteristics in relation to the wave environment have incorporated 3D
spectral numerical modeling. Furthermore no studies have identified
specific wave metrics that drive kelp morphological characteristics,
which the current work will aim to achieve. Other considerations such as
possible differences in morphology between sub-canopy and canopy
species, and shallow and deep water kelp populations have not been
considered using advanced numerical and statistical tools. Due to the
complex effects of environmental drivers on kelp morphology, one can
also expect differences in morphology between deep and shallow water
populations of the same species. Deep and shallow water environments may
differ in abiotic processes such as the degree of water column mixing
(Smit et al. 2013), solar heating (Dellatorre et al. 2012; Dunne and
Brown 2001), and effects of wave dampening (Kobayashi, Raichle, and
Asano 1993; Ma et al. 2013; Dubi and Tørum 1995).

The kelps \emph{Laminaria pallida} Greville and \emph{Ecklonia maxima}
(Osbeck) Papenfuss are important habitat forming seaweeds that exist
around the coast of South Africa and offer a unique opportunity to
investigate the drivers of macroalgal morphological characteristics
between canopy and sub-canopy species, as well as between deep and
shallow water environments. Although both these species exist in the
sub-tidal, \emph{L. pallida} dominates deeper waters while \emph{E.
maxima} forms dense surface canopies in the shallow sub-littoral.
Therefore, E. maxima may be exposed to variations in wind driven surface
waves, swell waves and temperature, while \emph{L. pallida} is exposed
to variations in swell waves, wave exposure and temperature due to
differences in depth. Molloy and Bolton (1996) investigated the effect
of depth and wave exposure on \emph{L. pallida} and showed that depth
had a greater effect than wave exposure when considering all the
morphological characteristics; when considering individual
characteristics, however, wave exposure had the most significant effect
on blade thickness. Another study by Rothman et al. (2017) investigated
the changes in morphology in shallow populations of \emph{L. pallida}
and \emph{E. maxima} along the South African coastline and into Namibia;
\emph{E. maxima} exhibited no morphological changes along the coast but
the stipes of \emph{L. pallida} become increasingly hollow further north
along the coastline. The authors suggested that turbidity in relation to
light attenuation was the environmental driver responsible for this
change.

The aim of this study is, therefore, to understand how temperature and
wave exposure can influence the morphology of sub-canopy and canopy
kelps, as well as between deep and shallow populations of the same
species. The current study will also aim to identify specific
temperature and wave metrics which drive kelp morphological
characteristics. This will be achieved by initially understanding the
variation in temperature and waves using a numerical model. The
numerical model will then be used as a basis for investigating the
consequences for morphological characteristics of \emph{E. maxima} and
\emph{L. pallida} using advanced statistical tools.

\hypertarget{materials-and-methods}{%
\section{Materials and methods}\label{materials-and-methods}}

\hypertarget{site-selection}{%
\subsection{Site selection}\label{site-selection}}

Sites were chosen to represent an array of temperature and wave
gradients, as indicated in Figure \ref{fig:map}. St.~Helena Bay and
Betty's Bay constituted the north-western and south-eastern boundary
sites, respectively. These sites are roughly 300 km apart, lie within
separate marine provinces Smit, Bolton, and Anderson (2017), and span
the majority of the south-west coast, in varying thermal and wave energy
regimes. The region is dominated by kelp communities that persist in
contrasting abiotic environments. The west coast region has been termed
cool-temperate, which is defined as a region where mean monthly
temperatures are always above 10°C and below 15°C (Smit et al. 2013).
East of Cape Point marks the beginning of an overlap or transition area,
which is also referred to as the Benguela-Agulhas Transition Zone (Smit
et al. 2013). The Agulhas Marine Province is characterized by a wide
temperature range of up to 7°C difference between mean monthly
temperatures between summer and winter and is classified as
warm-temperate (Smit et al. 2013).

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/map-1}

}

\caption{\label{fig:map}Map of the study area and sampling sites. The red dot indicates the position of Cape Town for sake of reference.}\label{fig:map}
\end{figure}

The annual and seasonal patterns in both wind and swell direction
influence the thermal regime around the coastline by inducing upwelling.
Upwelling brings deep, cool nutrient-rich water to coastal areas which
causes a decrease in temperature (Field et al. 1980; Andrews and
Hutchings 1980; Gill and Clarke 1974; Cram 1970; Rouault, Pohl, and
Penven 2010; Blanke et al. 2002). The wind speed and wind direction help
drive upwelling in summer on the western side of the Cape Peninsula,
where southerly winds blow parallel to the coast and trigger upwelling
events (Field et al. 1980; Rouault, Pohl, and Penven 2010). This is not
the case for False Bay, which is shielded from the dominating wind and
swell that the Cape Peninsula is subjected to. In the summer months,
upwelling becomes more favorable in False Bay due to the change in swell
and wind direction, causing lower temperatures to dominate during this
time (Rouault, Pohl, and Penven 2010; Andrews and Hutchings 1980). Due
to the Cape Peninsula temperate latitude, winter months bring an
increased frequency of frontal depressions that originate from the
Southern Ocean (Reason, Landman, and Tennant 2006). These low pressures
are joined by large swells with increased wave energy. The nearshore
environment, with the accompanied biota, therefore experiences high wave
energy events, with increased frequency in winter (Veitch et al. 2019).
The large peninsula acts as an obstruction for large south westerly
swells, providing decreased wave energy along the west side of False Bay
(Shipley 1964). Conversely, the west coast of Cape Point is battered by
these large swells. Multiple sites, therefore, exist where kelps grow in
diverse temperature and wave energy climates, in close proximity and on
the same side. The topography and elevation along the Cape Peninsula
channel and shield winds along False Bay. This is, however, absent in
winter, when strong northerly winds are prevalent from St.~Helena Bay to
Betty's Bay (Field et al. 1980; Jury, Kamstra, and Taunton-Clark 1985;
Andrews and Hutchings 1980; Jury 1980).

\hypertarget{abiotic-environment}{%
\subsection{Abiotic environment}\label{abiotic-environment}}

In order to compare abiotic variables for sites around the coast, large
historical databases for temperature, wave energy and wind were
accessed.

\hypertarget{temperature}{%
\subsubsection{Temperature}\label{temperature}}

Shallow water temperatures were sourced from The South African Coastal
Temperature Network (SACTN) website1. In terms of nearshore temperature,
in situ data are preferred over satellite sea surface temperature (SST),
which have shown to exhibit large biases (Smit et al. 2013). The
underwater temperature recorders (UTRs) are attached to concrete mooring
blocks or railway line sections deployed in the shallow at ∼3 m depth.
The UTRs were comprised of Starmon Mini recorders (Star-Oddi, Reykjavik,
Iceland/Onset Hobo® U22-001) accurate to ±0.2°C. All in situ temperature
data were reduced to a time-coded, continuous series of monthly means
over various time-scales (Smit et al. 2013). It was decided that the
annual time-scale would be used for analysis. Linear interpolated SST
were calculated for sites where in situ recorders were absent. These
data were used to group sites into ``Cool temperate'' and ``Warm
temperate'' categorizations to investigate the possible effect of
thermal regime on kelp morphological characteristics.

\hypertarget{wave-environment}{%
\subsubsection{Wave environment}\label{wave-environment}}

Wave model variable data formed part of the South African Coastal
Vulnerability Assessment, presented to the Department of Environmental
Affairs (DEA) and produced by the Council for Scientific and Industrial
Research (CSIR) (Rautenbach 2015). The South African wave climate was
modeled via 20 spectral numerical wave models that simulated the
offshore wave climate to the nearshore. The boundary conditions of these
models were obtained by using the NOAA Wave Watch III (WWIII) model
output, distributed via the National Centers for Environmental
Prediction (NCEP) product (Office of the Director 2000; Environmental
Modeling Center/Marine Modeling Branch 2005). The particular hindcast
product utilized during the DEA-CSIR study spans 1994--2013 at a 3-hour
resolution. These data were then used to model swell propagation into
the coastal models while wind waves (seas) were generated via stationary
computations in the Simulating Waves in the Nearshore SWAN model (Booij,
Holthuijsen, and Ris 1997). The assumption of stationary computations is
acceptable as the model domains were small enough so the temporal
variation of the model boundary was slower than the time it takes for
that boundary condition to propagate to the coast. SWAN allows one to
extract wave variables from specific gridded locations in the nearshore.
For False Bay, a resolution of 200 meters was modeled and output
produced at both the 7 m and 15 m isobaths. A 200 m resolution was used
as the False Bay computational grid was nested within a larger 1 km
resolution grid. This allowed for a computational effective wave
resolution of increasing resolution, from the NCEP, low resolution
output to nested, high resolution coastal output. For Table Bay and east
of Cape Hangklip the resolution was 500 m and also had output at the 7
and 15 m isobaths. These contour outputs were chosen in the original
study by the CSIR as most engineering run-up calculations require wave
parameter information at these contour depths and were the main focus of
the original study. For this study the 7 m contours were used. These
data were then used to calculate over all wave power (kW/m) as this was
considered the best measure of overall wave exposure. Annual wave power
was plotted and categorized into different wave exposure categories
which ranged from fully sheltered to extremely exposed.

\hypertarget{collection-of-kelp-morphological-characteristics}{%
\subsubsection{Collection of kelp morphological
characteristics}\label{collection-of-kelp-morphological-characteristics}}

The morphological characteristics of both species are presented in
figure \ref{fig:schematic}. \emph{Laminaria pallida} is characterized by
a single smooth blade which is divided longitudinally into sections, and
develops from a single meristematic region located at the junction
between the blade and the stipe (Dyer 2018). This species has a solid
stipe along the south-west coast but develops a hollow stipe along the
west coast northward and into Namibia (Rothman et al. 2017).
\emph{Ecklonia maxima} consists of a single primary blade which develops
above a gas-filled float and a hollow stipe below. Secondary blades are
produced laterally from the primary blade from several meristematic
regions along the margins of the primary blade (Dyer 2018). Both species
are held to the substrate by finger-like haptera, collectively known as
the holdfast.

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/schematic-1}

}

\caption{\label{fig:schematic}Schematic of *Laminaria pallida* and *Ecklonia maxima* morphology [@Dyer2018-rv].}\label{fig:schematic}
\end{figure}

Morphological measurements of L. pallida and E. maxima were collected at
18 sites along the Western Cape coast of South Africa (Figure
\ref{fig:map}) between October 2014 and April 2015. AlgaeBase was used
to confirm the latest taxonomic associations for each species
({\textbf{???}}). Divers sampled kelp individuals by means of SCUBA for
deeper populations and freediving for shallower populations. The sampled
individuals were then taken ashore and the relevant morphological
measurements taken. In each instance only the 13 largest individuals
were sampled to ensure that only mature and fully grown sporophytes were
measured. The age of \emph{E. maxima} was determined by size, presence
at the surface as well as a darker coloration of the stipe; as this
occurs with age. In the case of \emph{L. pallida}, the individuals with
the longest stipe were selected which is considered a reasonable
estimate of age. In each case the age of the individuals was estimated
visually before the individual was sampled. In instances where one of
the species was not present a site, the nearest population of the
missing species was sampled. Because the macroalgae differ in
morphological features, species-specific morphological characteristics
were included and are presented in Table 1. The stipe length for
\emph{E. maxima} was considered to be the portion of the plant just
below the pneumatocyst and just above the holdfast. Stipe circumference
was measured just above the holdfast for both species. Between February
2017 and November 2018, morphological measurements for E. maxima
individuals in shallow water (\textless1 m) at seven sites along the
Western Cape coast of South Africa were also collected. The same
morphological characteristic measurements were taken as for the deeper
\emph{E. maxima}, and this allowed comparison between morphological
characteristics between deep and shallow individuals within sites.
Measurements were not collected for \emph{L. pallida} in the shallow
depths, as this species is largely absent from the shallow in this
portion of the South African coastline.

\hypertarget{statistical-analyses}{%
\subsection{Statistical analyses}\label{statistical-analyses}}

Summary statistics were calculated for each site and region for each of
the environmental variables considered in this study, and are presented
in Table 2. The associated abbreviations used are also presented in
Table 2. In addition, annual median wave and wind direction were also
calculated. The median was calculated for wind and wave direction, as
issues arise when calculating the mean and standard deviation for
compass metrics. Summary statistics for wind was not plotted and instead
are discussed, as the data were course relative to the other
environmental variables considered in this study.

Morphological characteristics in relation to temperature and wave
exposure were investigated using non-parametric statistical methods.
Morphological characteristics were placed into categories according to
temperature regime and wave exposure class for each species and
population of kelp in this study. The appropriate category was
determined by kelp's sampling location along the coastline and the
according wave exposure and temperature category assigned to that
species and/or population. The R software package ggplot2 was used for
graphical output and non-parametric analyses. Significant differences
between temperature regimes and across wave exposure categories were
tested with Wilcox tests and Kruskal -Wallis tests, respectively, and
notched boxplots allowed comparison between individual categories of
wave exposure. The alpha level used to define significance was p = 0.05.
Boxplots were created using the software package ggplot2
({\textbf{???}}). In order to investigate specific drivers of
morphological characteristics a distance-based redundancy analysis
(db-RDA) was performed using the rda function in the vegan software
package (Oksanen et al. 2015) in R (R Core Team, 2014; RStudio Team,
2020). The abiotic data covered all study sites for both species of
kelp. Both morphology and abiotic data (temperature, waves and wind)
were standardized using the decostand function in the vegan software
package; morphology data were used as response variables and the abiotic
data as explanatory variables. A list of the abbreviations used and the
associated metrics are presented in Table 2. To determine the
explanatory variables that best describe patterns in the response data,
a full RDA was performed using a complete set of explanatory variables.
Forward selection was then used to reduce the number of explanatory
variables, as well as prevent the inflation of overall Type I error. To
further improve the model, pairwise coefficients and Variance Inflation
Factor (VIF) were calculated to identify variables with high
multicollinearity. The computation of the parsimonious RDA was followed
by permutation tests of the adjusted R2 to assess significance of
constraints.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{abiotic-environment-1}{%
\subsection{Abiotic environment}\label{abiotic-environment-1}}

\hypertarget{temperature-1}{%
\subsubsection{Temperature}\label{temperature-1}}

The mean annual coastal water temperature for the study sites located on
the western side of the peninsula ranged from 14.5 ± 0.9°C (mean ± SD)
at St.~Helena Bay, the most northern site on the western side of the
peninsula, to 14.6 ± 0.3°C at Olifantsbos, the most southern site on the
western side of the peninsula (see Figure
\ref{fig:Temperature variables plot}A). For sites located on the eastern
side of the peninsula (within False Bay), the annual mean coastal water
temperatures ranged from 15.5 ± 0.9°C at Buffelsbaai to 15.0 ± 0.9°C at
Betty's Bay, the most eastern site on the coastline in this study (see
Figure \ref{fig:Temperature variables plot}A).

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/Temperature variables plot-1}

}

\caption{\label{fig:Temperature variables plot}Annual temperature variables at the collections sites around the Cape Peninsula are represented by plot A, with sites in order from North to South. Temperature variables include minimum temperature represented by inverted triangles, mean represented by black dots, maximum by black triangles, and whiskers standard deviation. Summary of temperature data for sites grouped by region and represented in plot B.}\label{fig:Temperature variables plot}
\end{figure}

In general, the mean temperature increases for sites located within
False Bay with less variation around the mean temperature compared to
sites on the western side of the peninsula (see Figure
\ref{fig:Temperature variables plot}A). This is verified by statistical
difference based on region, which shows a significantly lower (p
\textless{} 0.05, Figure \ref{fig:Temperature variables plot}B) mean
temperature and less variation in temperature for sites located on the
western side of the peninsula compared to sites on the eastern side.

\hypertarget{wave-environment-1}{%
\subsubsection{Wave environment}\label{wave-environment-1}}

Our data show the western side of the peninsula experiences higher
significant wave heights and variation in wave heights compared to the
eastern side of the peninsula (see Figure
\ref{fig:wave variable plot})A). Mean significant wave height ranged
from 0.5 ± 0.2 m (mean ± SD) at St.~Helena Bay to 2.3 ± 0.7 m at
Olifantsbos on the western side of the peninsula, and on the eastern
side of the peninsula it ranged from 0.9 ± 0.4 m at Buffelsbaai to 1.9 ±
0.6 m at Betty's Bay (see Figure \ref{fig:wave variable plot})A). The
maximum significant wave height ranged from 1.4 m at St.~Helena Bay to
4.7 m at Olifantsbos on the western side of the peninsula, and on the
eastern side ranged from 2.6 m at Buffelsbaai to 4.2 m at Betty's Bay.
Mean peak period for sites on the western side of the peninsula ranged
from 10.3 ± 3.5 s at St.~Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and
ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty's Bay
(see Figure \ref{fig:wave variable plot})B). The data show no trend in
mean peak period for the coastline. These data show that the western
side of the peninsula has a lower variation (SD) compared to the eastern
side, with Miller's Point showing the highest variation across
timescales. Maximum peak period ranged from 18.9 s at St.~Helena Bay to
18.4 s at Olifantsbos for the western side of the peninsula, and ranged
from 18.0 s at Buffelsbaai to 18.0 s at Betty's Bay on the eastern side
of the peninsula. When wave variables were grouped by region significant
differences were found for both mean significant swell height and mean
peak period. The western region experiences a significantly higher mean
significant swell height (p \textless{} 0.05; see Figure
\ref{fig:wave variable plot})C) and mean peak period (p \textless{}
0.05; see Figure \ref{fig:wave variable plot})D).

Data for median wave direction showed no trend across timescales,
however, the variation (SD) in wave direction is lower on the western
side of the peninsula compared to the eastern side both annually and for
winter. Summer data were the exception and showed lower variation for
the eastern side of the peninsula (see Appendix).

In Figure 5A the total coastal wave exposure of the Cape Peninsula is
given in terms of wave energy (kW per meter wave crest length). The
classification from fully sheltered to extremely exposed is based on the
total wave energy upper and lower limits (see Figure
\ref{fig:wave power map}B). The western periphery of the Cape Peninsula
is almost continuously exposed to high wave exposures while the eastern
periphery of the peninsula (western coastline of False Bay) revealed
sheltered wave exposures (see Figure \ref{fig:wave power map}). There
the marked seasonality, with higher energy waves during winter, may be
clearly observed once more.

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/wave variable plot-1}

}

\caption{\label{fig:wave variable plot}Mean significant wave height and mean significant peak period across sites and by region. Mean significant wave height across sites is presented in plot A and is represented by solid triangles and standard deviation by the whiskers. Differences in mean significant wave height between regions are presented in plot B. Mean significant peak period across sites is presented in plot C and is represented by solid dots and standard deviation by whiskers. Differences in mean significant peak period between regions are presented in plot D.}\label{fig:wave variable plot}
\end{figure}

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/wave rose-1}

}

\caption{\label{fig:wave rose}The total and seasonal wave roses of the directional wave buoy just offshore from Cape Town is given form the years 2000 till end 2017 (the period for which directional wave spectra was available). Total (A), Austral winter (B) and Austral summer (C) respectively}\label{fig:wave rose}
\end{figure}

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/wave power map-1}

}

\caption{\label{fig:wave power map}The total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length. Total (A), summer (B), winter (C), spring (D) and autumn (E).}\label{fig:wave power map}
\end{figure}

To clarify the averaged wave exposure map presented in Figure
\ref{fig:wave power map}A, the propagation of a typical offshore wave
spectrum as produced from a single time-step in SWAN is presented in
Figure \ref{fig:wave power map}C. Tracing the wave height contours into
False Bay its clear why this bay's western periphery is predominantly
sheltered.

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/FB spectrum-1}

}

\caption{\label{fig:FB spectrum}The propagation of a typical offshore wave spectrum is given as produced from a single time-step in SWAN.}\label{fig:FB spectrum}
\end{figure}

\hypertarget{drivers-of-kelp-morphological-characteristics}{%
\subsection{Drivers of kelp morphological
characteristics}\label{drivers-of-kelp-morphological-characteristics}}

Significant differences were found between cool temperate and warm
temperate regimes for most \emph{L. pallida} morphological
characteristics with lamina length, number of digits and thallus mass
the exceptions (see Figure 6). Certain morphological characteristics
such as stipe mass, stipe length, total length and stipe diameter had
significantly higher means for kelp from the cool temperate regime when
compared to kelp from the warm temperate regime (p \textless{} 0.05; see
Figure \ref{fig:lamanaria temp plots}). Conversely, lamina weight and
lamina thickness had significantly lower means for kelp from the cool
temperate regime compared to kelp from the warm temperate regime. When
\emph{L. pallida} morphological characteristics were grouped to wave
exposure categories Kruskal-Wallis tests revealed significant
differences amongst categories for most morphological characteristics
with lamina length, number of digits and thallus mass the exceptions
(see Figure ref\{fig:lamanaria wave plots\}). The morphological
characteristics that were significantly different amongst sites
exhibited similar patterns, with increasing mean values from the fully
sheltered to extremely sheltered categories and a decrease in mean
values for the remainder of the categories. Two morphological
characteristics, namely lamina thickness and stipe diameter, both
exhibited higher variations for the exposed categories compared to other
morphological characteristics.

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/laminaria temp plots-1}

}

\caption{\label{fig:lamanaria temp plots}Boxplots representing *L. pallida* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided. The lower and upper "hinges" correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers. Light grey boxes represent cool-temperate and dark grey warm-temperate.}\label{fig:laminaria temp plots}
\end{figure}

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/lamanaria wave plots-1}

}

\caption{\label{fig:lamanaria wave plots}Boxplots representing *L. pallida* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers.}\label{fig:lamanaria wave plots}
\end{figure}

Six out of the ten deep \emph{E. maxima} morphological characteristics
showed differences between cool temperate and warm temperate regime
except for frond mass, frond length, stipe mass and epiphyte length.
Mean values for primary length, primary width and stipe circumference
were significantly higher for kelp from the cool temperate regime than
for kelp from the warm temperate regime (see Figure
\ref{fig:{deep eck temp plot}}). The remaining morphological
characteristics, stipe length, number of tufts and total length had
significantly lower means for kelp from cool temperate compared to kelp
from the warm temperate regime (p \textless{} 0.05; see Figure 8).
Significant differences amongst wave exposure categories were found for
all morphological characteristics of deep E. maxima populations (see
Figure \ref{fig:deep eck temp plot} ). Certain morphological
characteristics showed a gradual increase in mean value with increasing
degree of wave exposure. The remaining morphological characteristics
exhibited decreased means for only the higher wave exposure categories,
followed by a sharp increase in mean value for the highest wave exposure
category.

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/deep ecklonia temp plot-1}

}

\caption{\label{fig:deep eck temp plot}Boxplots representing deep population *E. maxima* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers. Light grey boxes represent cool-temperate and dark grey warm-temperate.}\label{fig:deep ecklonia temp plot}
\end{figure}

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/deep ecklonia wave-1}

}

\caption{\label{fig:deep ecklonia wave}Boxplots representing deep *E. maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers.}\label{fig:deep ecklonia wave}
\end{figure}

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/shallow ecklonia temp-1}

}

\caption{\label{fig:shallow ecklonia temp} Boxplots representing shallow *E. maxima* morphological characteristics  grouped by temperature regime measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers. Light grey boxes represent cool-temperate and dark grey warm-temperate.}\label{fig:shallow ecklonia temp}
\end{figure}

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/shallow ecklonia wave-1}

}

\caption{\label{fig:shallow ecklonia wave}Boxplots representing shallow *E. maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers.}\label{fig:shallow ecklonia wave}
\end{figure}

When grouped by temperature regime, shallow \emph{E. maxima} exhibited
significant differences between temperature regimes for frond mass,
primary length, primary width, frond length, stipe mass, stipe length,
stipe circumference and total length (p \textless{} 0.05; see Figure
\ref{fig:shallow ecklonia temp}). All of the significantly different
morphological characteristics, except primary length, had lower values
for the western region compared to the eastern region. When shallow E.
maxima were grouped by wave exposure category significant difference
were found for all morphological characteristics (p \textless{} 0.05;
see Figure \ref{fig:shallow ecklonia wave}). In general, values of the
various morphological characteristics increase for the lower wave
exposure categories, followed by a sharp decrease and a gradual increase
in value.

\hypertarget{abiotic-drivers-of-l.-pallida-morphological-characteristics}{%
\subsubsection{\texorpdfstring{Abiotic drivers of \emph{L. pallida}
morphological
characteristics}{Abiotic drivers of L. pallida morphological characteristics}}\label{abiotic-drivers-of-l.-pallida-morphological-characteristics}}

Forward selection, assessment of VIF and an examination of pairwise
Pearson correlation coefficients allowed us to retain the most
parsimonious descriptors of L. pallida morphological characteristics
with an adjusted R2 of 0.49, explaining 72\% of the variation (global
permutation test on final model: d.f = 1, F = 8.51, p = 0.008; Figure
\ref{fig:rda plots}A). RDA1 was the only significant axis in the model
and explained 48\% of the variation while RDA2 was not significant and
explained 63\% of the variation. Total length, stipe length, stipe
diameter and stipe mass were positively influenced (i.e., increased size
corresponding with the environmental driver) by mean significant wave
height and median swell direction for kelp mostly from the western
region and negatively influenced (i.e., decreased size corresponding
with the environmental driver) by peak period standard deviation and
mean temperature for kelp from the eastern region. Although sites from
the western region did not cluster as closely as eastern region study
sites, grouping according to region was still evident.

\hypertarget{abiotic-drivers-of-e.-maxima-morphological-characteristics}{%
\subsubsection{\texorpdfstring{Abiotic drivers of \emph{E. maxima}
morphological
characteristics}{Abiotic drivers of E. maxima morphological characteristics}}\label{abiotic-drivers-of-e.-maxima-morphological-characteristics}}

The same approaches to identify the descriptors for \emph{L. pallida}
were employed for both populations of \emph{E. maxima}. Forward
selection, assessment of VIF and Pearson correlation coefficients
allowed us to retained as the most parsimonious descriptors of deep
\emph{E. maxima} morphological characteristics with an adjusted R2 of
0.39, explaining 67\% of the variation (global permutation test on final
model: d.f = 5, F = 2.4, p = 0.026; Figure \ref{fig:rda plots}B). RDA1
was the only significant axis in the model (p = 0.03) and explained 72\%
of the variation. Primary length, primary width and stipe circumference
were positively influenced by mean significant wave height, mean peak
period and median swell direction for kelp from the western region. The
remaining morphological characteristics were positively influenced by
minimum temperature and negatively influenced by peak period standard
deviation for kelp from the eastern region.

\begin{figure}

{\centering \includegraphics{thesis_chapter_2_files/figure-latex/rda plots-1}

}

\caption{\label{fig:rda plots}RDA output for *L. pallida* (A), deep *E. maxima* (B) and shallow *E. maxima* (C). The blue cluster represnts sites from the western region and the red represents sites from the eastern region.}\label{fig:rda plots}
\end{figure}

Forward selection, assessment of VIF and Pearson correlation
coefficients allowed us to retain the most parsimonious descriptors of
shallow \emph{E. maxima} morphological characteristics with an adjusted
R2 of 0.36, explaining 61\% of the variation (global permutation test on
final model: d.f = 4, F = 2.41, p = 0.013, Figure \ref{fig:rda plots}C).
The model consisted of one significant axis, RDA1 (p = 0.02) and
explained 62\% of the variation. Stipe length, stipe mass and stipe
circumference were positively influenced by maximum temperature and mean
temperature along RDA1, and explained 50 and 55\% of the variation,
respectively. The remaining morphological characteristics were
positively influenced by maximum peak period and negatively influenced
by median wind direction along RDA2 and explained 86 and 94\% of the
variation, respectively. There was overlapping of clusters according to
region with no clear patterns in sites evident.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

This study considered important drivers of the morphology of Ecklonia
maxima and Laminaria pallida around the Cape Peninsula. These abiotic
variables, namely temperature and wave exposure, have been identified in
previous research on other brown algae species such as E. radiata and M.
pyrifera (Thomsen, Wernberg, and Kendrick 2004; Fowler-Walker, Connell,
and Gillanders 2005; Fowler-Walker, Wernberg, and Connell 2006; Wernberg
and Thomsen 2005; Wing et al. 2007; Stewart et al. 2009; Miller, Hurd,
and Wing 2011; Pedersen and Nejrup 2012). The complex geomorphology of
the Western Cape coastline creates an ideal natural laboratory for
studies of kelps in interaction with their environment. Unlike previous
research, this study used a complex numerical model to provide the
various variables of wave exposure, which is based on linear wave theory
to calculate overall wave exposure. We also considered wind in addition
to temperature and wave variables, since wind is an important component
of wind-driven surface gravity waves (Holthuijsen 2010). The results
show that specific variables of wave exposure are the main drivers of
kelp morphology with temperature playing only a minor role. The
morphological adaptation was also shown to be associated with the
magnitude of wave exposure which had only been inferred in previous
studies. Finally, the investigation of differences in morphological
characteristics between shallow and deep populations of E. maxima
suggests that in low wave exposure environments the role of temperature
as a morphological driver increases; this is particularly the case for
False Bay.

There were clear patterns and clines in both the temperature and wave
exposure data. The thermal and wave exposure regimes around the Cape
Peninsula are driven by a complex interaction between wind, temperature
and wave metrics, which do not act independently but instead influence
one another. The hydrodynamic regime is modulated by the relative
sheltering against the predominant south-westerly swell direction and
the size of the fetch for local wave generation. The direction of the
dominant swell changes to the south west in winter, generated by strong
low pressures that originate from the southern ocean (Reason, Landman,
and Tennant 2006), which False Bay is sheltered from (Shipley 1964;
Atkins 1970; Dufois and Rouault 2012). In summer, these swells rotate
anticlockwise and are able to enter False Bay, providing an increased
variability of swell height and peak period in this region. It should be
noted that what is classified as sheltered around the South African
coastline (a high energy coastline) might be classified as exposed in
other regions of the world (Norderhaug et al. 2012; Leliaert et al.
2000). It is due to the near consistent south-westerly swell and the
complex orography around the peninsula that the wave energy distribution
around the Cape Peninsula varies significantly over a small geographical
area. The directional sheltering effect of the Cape Peninsula, against
the dominant swell direction is clearly observed in the wave exposure
maps. It should be mentioned that some of the annual winter frontal
depression systems pass the Cape Peninsula from the west to east,
resulting in wave propagating toward the continent from much more
southerly directions. This results in positive and negative wave
exposure anomalies all around the peninsula. Increased wind speed at
sites along the west side of the peninsula in a southerly direction
trigger upwelling events along the western side of the peninsula
(Rouault, Pohl, and Penven 2010; Field et al. 1980). In general, the
western side of the peninsula is more exposed than the eastern side
(False Bay) and experiences significantly higher mean significant swell
height and mean significant peak period. This was also reflected in the
wave exposure categories which show a trend of decreasing wave exposure
(wave power) around the peninsula and into False Bay. There are also
differences between the types of waves that each side of the peninsula
experiences. We provide strong support that variations in the
environmental variables, particularly wave exposure variables, are
driving kelp morphological characteristics around the Cape Peninsula.
Morphological adaptation due to water motion may manifest in a number of
ways in high wave energy environments. For instance, reduction of blade
thickness, blade elongation, increase of stipe length, increase in stipe
circumference and force of attachment (Fowler-Walker, Wernberg, and
Connell 2006; Wernberg and Thomsen 2005; Friedland and Denny 1995; Denny
and Gaylord 2002; Bekkby et al., n.d.; Denny, Gaylord, and Cowen 1997)
have been identified in previous studies. Although this study did not
measure force of attachment, other morphological responses to
temperature and wave variables were evident. Species-specific responses
are evident in both wave exposure and temperature. In cool temperate
environments \emph{L. pallida} tended to show increases in certain
morphological characteristics (stipe mass, stipe length and stipe
diameter) while in the warm-temperate environments these were
significantly lower. This was also true for deep \emph{E. maxima}
populations, which had longer, thinner stipes for kelp from the warm
temperate compared to kelp from the cool temperate. Reduction in certain
morphological characteristics has been attributed to temperature by
Serisawa et al. (2002) in the kelp E. cava, which was smaller and
shorter in warmer sites compared to cooler sites. The reduction in size
of adult may be a response to low nutrient conditions, which has been
shown to reduce growth rate and overall morphology (Simonson,
Scheibling, and Metaxas 2015). Warmer temperatures are associated with
low nutrient concentrations (Waldron and Probyn 1992) and the low
frequency of upwelling conditions in False Bay (low nutrient supply)
coupled with warmer temperatures may be a contributing factor. It should
be noted, however, that from these analyses that in general the cool
temperate region is more exposed to waves than that of the
warm-temperate region. Therefore, the significantly larger morphological
characteristics for kelp from the cool temperate region may overlap with
responses to wave exposure. The response of kelp morphological
characteristics to wave exposure was evident and both species exhibit
tactics based on the magnitude of wave exposure. Strength increasing
traits were exhibited for lower exposure levels while a
``go-with-the-flow'' tactic for moderate levels of wave exposure. This
suggests that how morphological characteristics manifest themselves is
dependent on the magnitude of wave exposure. When grouped by wave
exposure category, \emph{L. pallida} characteristics showed a
significant increase in length for the ``extremely sheltered'' category
compared to the fully sheltered category, which may be a
``go-with-the-flow'' tactic (Denny, Gaylord, and Cowen 1997; Denny and
Gaylord 2002; Hurd 2000). Kelp are able to increase flexibility by
increasing stipe length which increases the extension capabilities of
kelp to a passing wave (Denny, Gaylord, and Cowen 1997; Denny and
Gaylord 2002; Hurd 2000). However, increasing stipe length is only
beneficial under lower exposure levels as a long stipe actually
increases overall drag on the plant under high wave exposure (Denny,
Gaylord, and Cowen 1997; Denny and Gaylord 2002; Hurd 2000). This is
reflected in the \emph{L. pallida} morphology which shows an overall
reduction of morphological characteristics suggesting a size reducing
tactic to cope with higher levels of wave exposure. The results suggest
that deep \emph{E. maxima} populations exhibit a different response, as
well as a higher wave exposure threshold. Unlike \emph{L. pallida},
\emph{E. maxima} exhibits a size reducing tactic with increasing wave
exposure by decreasing stipe mass, stipe length, stipe circumference,
total length and frond length. A magnitude related response has been
suggested by Wernberg and Vanderklift (2010) who investigated the
temporal and spatial variations of various environmental drivers of the
kelp \emph{E. radiata}. The authors identified wave exposure as the most
important driver of kelp morphological characteristics and that the type
of response elicited is dependent on the magnitude of wave exposure.

A size reduction tactic has been shown before (Fowler-Walker, Wernberg,
and Connell 2006; Denny and Gaylord 2002; Hurd 2000; Fowler-Walker,
Connell, and Gillanders 2005) and is regarded as a strategy to reduce
overall drag on the plant. However, \emph{E. maxima} morphological
characteristics changed significantly as wave exposure increases. When
grouped to the higher wave exposure levels, \emph{E. maxima} morphology
exhibited a strength and flexibility increasing trait. Furthermore, the
morphological response of \emph{E. maxima} to wave exposure only occurs
at the ``Moderately exposed'' level compared to \emph{L. pallida} which
exhibits a response at a lower wave exposure level. This suggests that
\emph{E. maxima} can tolerate higher exposure levels before having to
exhibit a morphological response. The redundancy analysis performed
confirms the patterns and responses observed as well as identifying
specific temperature and wave variables as drivers of kelp morphology.
Kelp morphology characteristics are largely wave driven for both species
but differ in terms of specific temperature and wave metrics.

The wave metrics identified in this study are important components of
determining overall wave power, which will vary seasonally in the region
based on the swell direction. The differences in temperature metrics
between \emph{E. maxima} and \emph{L. pallida} point to underlying
differences in nutrient availability. Low temperatures are often
associated with upwelling events which bring cool, nutrient rich water
into the nearshore. False Bay has comparatively little upwelling events
compared to the western side of the peninsula, and so nutrients may be a
limiting factor for \emph{E. maxima} populations within False Bay, hence
the identification of minimum temperature as a influencer. Mean
temperature identified as a driver for \emph{L. pallida} may be related
to diurnal temperature fluctuations in the water column. Solar heating
of the water surface in combination with wind-driven transport causes
fluctuations within the water column, which can occur daily or
seasonally (Kaplan et al. 2003). Vertical mixing of the water column for
inshore regions are driven by several abiotic processes: (1) turbulence
of breaking waves inside or outside of the surf zone; (2) convective
mixing through a combination of cooling and evaporation; (3) wind driven
currents; and (4) tidal mixing (Smit et al. 2013). These abiotic
processes cause effective vertical mixing of surface and deeper water
stratification which leads to a homogeneous thermal environment, which
in this study may be interpreted as the mean temperature. Therefore, the
homogeneous thermal regime of inshore regions may be a reason why mean
temperature is driver of \emph{L. pallida} morphological
characteristics. The difference in canopy type between the species may
be the reason why \emph{E. maxima} is driven by multiple wave metrics
compared to \emph{L. pallida}. Since \emph{E. maxima} are a canopy-kelp,
it is exposed to all components of a wave compared to \emph{L. pallida}
which occurs deeper in the water column.

Although temperature plays an important role in distribution (Bolton
2010; Miller, Hurd, and Wing 2011; Rinde et al. 2014) of kelp and
physiological functioning of adults and gametophytes (Mohring et al.
2014; Steneck et al. 2002; Gerard 1997; Bearham, Vanderklift, and Gunson
2013; Smale and Moore 2017), there is little evidence in this study that
temperature is an important driver of morphological variation. However,
we suggest that temperature plays a larger role in determining kelp
morphology at sites where wave energy is low enough not to elicit a
response in morphological characteristics. Temperature metrics influence
stipe length and frond length in shallow E. maxima individuals located
at sites within False Bay; which was not the case for deeper E. maxima
individuals as well as \emph{L. pallida} which were largely driven by
wave variables. The shallow \emph{E. maxima} individuals may not
experience the same amount of wave energy as do deeper populations. This
may be due to wave dampening from deeper populations of \emph{E.
maxima}, which attenuate the wave energy entering coastal environments
(Kobayashi, Raichle, and Asano 1993; Jackson 1984; Mork 1996). The
morphological variations in characteristics for deep \emph{E. maxima}
populations are not seen for shallow E. maxima populations. The lack of
variation in shallow \emph{. maxima} populations was also found in
previous research by (Rothman et al. 2017) who identified light as a
significant driver. This influence of light may suggest photosynthetic
ability and nutrient uptake as important drivers in shallow
environments. Wave exposure may be reduced in shallow environments due
to the wave damping effect of deep kelp and seaweed populations
(Kobayashi, Raichle, and Asano 1993; Mork 1996; Jackson 1984) thereby
reducing the need to adapt morphological characteristics to reduce drag.
Furthermore, the reduction in differences between deep and shallow E.
maxima at wave exposed sites on the western of the peninsula compared to
sites in False Bay provides further evidence that wave exposure may not
be an important driver of shallow \emph{E. maxima} populations in this
study. Increased temperatures corresponded with a reduction in stipe
circumference, frond length, stipe length, total length and stipe mass
resulting in slim or ``skinny'' kelp individuals, which supports
findings from previous research by Mabin, Gribben, and Fischer (2013),
Serisawa et al. (2002). These findings were confirmed by the redundancy
analysis which showed the contribution of wave variables as a driver of
kelp morphological characteristics was lower compared to temperature
variables. The sites do cluster by region but almost overlap entirely;
further indicating wave variables are not the main driver. Although wave
variables were not identified as an important driver, it is important to
note the wave variables identified in the RDA differ to deep E. maxima.
The identification of median wind direction suggests that surface
gravity waves as a driver compared to swell waves identified for
\emph{L. pallida} and \emph{E. maxima}.

Kelps living in warmer environments have been shown to have a lower
resilience to disturbance, such as storms (T. Wernberg et al. 2010). The
kelps in False Bay may therefore be more vulnerable to dislodgment under
high wave exposure scenarios compared to kelps on the western side of
the Cape Peninsula, which experiences higher wave energy on a regular
basis. In South Africa, the main contributing factors to the
hydrodynamic environment are the Southern Annular Node (SAM) and El Niño
Southern Oscillation (ENSO), while the latter is also the main source of
variability (Veitch et al. 2019). These climate systems are expected to
alter storm frequency and magnitude under future climate scenarios,
which may increase the overall wave power around the Cape Peninsula. The
magnitude-specific response of the kelp species to wave exposure in this
study suggests there is a limit to morphological adaptation in reducing
the probability of dislodgement. Therefore, if this limit is reached and
exceeded in future climate scenarios, kelp populations may risk being
reduced over time. Furthermore, although recent studies suggest a
cooling trend of coastal waters around South Africa (Rouault, Pohl, and
Penven 2010; Smit et al. 2013), the impact of marine heat waves may
reduce the resilience of kelps to storms and high wave energy. Although
this study identifies specific wave and temperature metrics that drive
morphological adaptation and suggests a possible wave exposure
threshold, it does not identify the metrics or magnitude of wave
exposure required to dislodge kelp. Therefore, future studies should
consider using advanced numerical methods to investigate the specific
wave power needed to dislodge kelp, and ideally, given a specific set of
morphological measures. These types of studies in conjunction with
forecasting studies can be used to determine the vulnerability of kelp
populations to future climate scenarios.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Past research shows that macroalgal morphological characteristics are
largely driven by a wave exposure while others have suggested a complex
interaction between various environmental variables. We confirm that
wave variables are the main driver of kelp morphological characteristics
in both \emph{E. maxima} and \emph{L. pallida} populations around the
Cape Peninsula, and differences in morphological characteristics between
shallow and deep populations of \emph{E. maxima} may reflect differences
in hydrodynamic and temperature regimes. This study also confirms the
findings of previous research which shows morphological adaptation is
dependent on the magnitude of wave exposure, but also provides
quantitative, reproducible measure of wave exposure while identifying
the specific wave metrics and wave energy thresholds involved.
Furthermore, the current study highlights the need to understand how
kelp populations, which are not exposed to high wave energy and
morphological characteristics are largely driven by temperature, are
vulnerable to changes in storm frequency and magnitude. Kelp populations
located in low wave energy environments, such as False Bay, may be more
vulnerable to increased storm frequency and magnitude in future climate
scenarios.

\hypertarget{conflict-of-interest-statement}{%
\section{Conflict of Interest
Statement}\label{conflict-of-interest-statement}}

CR was employed by the company MetOcean Solutions Ltd.

The remaining authors declare that the research was conducted in the
absence of any commercial or financial relationships that could be
construed as a potential conflict of interest.

\hypertarget{author-contributions}{%
\section{Author Contributions}\label{author-contributions}}

RC conceptualized the scope of the research reported in this manuscript,
collected samples, undertook the majority of numerical and statistical
analyses, made the first round of interpretation, and did the majority
of the writing. AS and CR provided the guidance in terms of writing,
statistical analyses, and overall presentation of the manuscript. All
authors contributed to the article and approved the submitted version.

\hypertarget{funding}{%
\section{Funding}\label{funding}}

This research was funded by the South African National Research
Foundation (\url{http://www.nrf.ac.za}). Funding number is
CSRP170430229220. Aside from funding provided, the funders had no role
in the study design, data collection and analysis, decision to publish,
or preparation of the manuscript.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

We thank A. J Smit and Christo Rautenbach for their guidance on this
project. We also thank Robert Schlegel and Robert Williamson for their
help coordinating the honors class of 2014--2015, in particular Jesse
Smith, Janine Greuel, Tevya Lotriet, and Rudy Philander. The work would
have not been possible without the dataset from the CSIR report 2014
where the processed data is available at the following DOIs:
10.15493/DEFF.10000004 and 10.15493/DEFF.10000003. Finally, we thank the
South African Coastal Temperature Network (SACTN) for access the
temperature data used in this study.

\hypertarget{supplementary-material}{%
\section{Supplementary Material}\label{supplementary-material}}

The Supplementary Material for this article can be found online at:
\url{https://www.frontiersin.org/articles/10.3389/fmars.2020.00567/full\#supplementary-material}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\begin{center}\includegraphics{thesis_chapter_2_files/figure-latex/appendix figure 1-1} \end{center}

\begin{center}\includegraphics{thesis_chapter_2_files/figure-latex/appendix figure 2-1} \end{center}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Andrews1980}{}%
Andrews, WRH, and L Hutchings. 1980. ``Upwelling in the Southern
Benguela Current.'' \emph{Progress in Oceanography} 9 (1): 1--81.

\leavevmode\hypertarget{ref-Atkins1970-bx}{}%
Atkins, G R. 1970. ``Winds and Current Patterns in False Bay.''
\emph{Transactions of the Royal Society of South Africa} 39 (2):
139--48.

\leavevmode\hypertarget{ref-bearham2013}{}%
Bearham, Douglas, Mathew A Vanderklift, and Jim R Gunson. 2013.
``Temperature and Light Explain Spatial Variation in Growth and
Productivity of the Kelp \emph{Ecklonia Radiata}.'' \emph{Marine Ecology
Progress Series} 476: 59--70.

\leavevmode\hypertarget{ref-bekkby2014}{}%
Bekkby, Trine, Eli Rinde, Hege Gundersen, Kjell Magnus Norderhaug, Janne
K. Gitmark, and Hartvig Christie. n.d. ``Length, strength and water
flow: Relative importance of wave and current exposure on morphology in
kelp hyperborea.'' \emph{Marine Ecology Progress Series}, 61--70.
\url{https://doi.org/10.3354/meps10778}.

\leavevmode\hypertarget{ref-Blanke2002}{}%
Blanke, Bruno, Claude Roy, Pierrick Penven, Sabrina Speich, James
McWilliams, and Greville Nelson. 2002. ``Linking Wind and Interannual
Upwelling Variability in a Regional Model of the Southern Benguela.''
\emph{Geophysical Research Letters} 29 (24): 41--41.

\leavevmode\hypertarget{ref-Bolton2010}{}%
Bolton, J J. 2010. ``The biogeography of kelps (Laminariales,
Phaeophyceae): A global analysis with new insights from recent advances
in molecular phylogenetics.'' \emph{Helgoland Marine Research} 64 (4):
263--79. \url{https://doi.org/10.1007/s10152-010-0211-6}.

\leavevmode\hypertarget{ref-Booij1997-nx}{}%
Booij, N, L H Holthuijsen, and R C Ris. 1997. ``The "SWAN'' wave model
for shallow water.'' In \emph{Coastal Engineering 1996}, 668--76.
ascelibrary.org.

\leavevmode\hypertarget{ref-byrnes2011}{}%
Byrnes, Jarrett E, Daniel C Reed, Bradley J Cardinale, Kyle C Cavanaugh,
Sally J Holbrook, and Russell J Schmitt. 2011. ``Climate-Driven
Increases in Storm Frequency Simplify Kelp Forest Food Webs.''
\emph{Global Change Biology} 17 (8): 2513--24.

\leavevmode\hypertarget{ref-Cavanaugh2011-ib}{}%
Cavanaugh, K C, D A Siegel, D C Reed, and P E Dennison. 2011.
``Environmental controls of giant-kelp biomass in the Santa Barbara
Channel, California.'' \emph{Marine Ecology Progress Series} 429: 1--17.

\leavevmode\hypertarget{ref-Cousens1982-ck}{}%
Cousens, R. 1982. ``The Effect of Exposure to Wave Action on the
Morphology and Pigmentation of \emph{Ascophyllum nodosum} (L.) Le Jolis
in South-Eastern Canada.'' \emph{Bontanica Marina} 25 (4): 191--95.

\leavevmode\hypertarget{ref-Cram1970-jw}{}%
Cram, D L. 1970. ``A suggested origin for the cold surface water in
central False Bay.'' \emph{Transactions of the Royal Society of South
Africa} 39 (2): 129--37.

\leavevmode\hypertarget{ref-Dayton1985}{}%
Dayton, P K. 1985. ``Ecology of Kelp Communities.'' \emph{Annual Review
of Ecology and Systematics} 16 (1): 215--45.
\url{https://doi.org/10.1146/annurev.es.16.110185.001243}.

\leavevmode\hypertarget{ref-Dellatorre2012}{}%
Dellatorre, Fernando G, Juan P Pisoni, Pedro J Barón, and Andrés L
Rivas. 2012. ``Tide and Wind Forced Nearshore Dynamics in Nuevo Gulf
(Northern Patagonia, Argentina): Potential Implications for Cross-Shore
Transport.'' \emph{Journal of Marine Systems} 96: 82--89.

\leavevmode\hypertarget{ref-Denny2002}{}%
Denny, Mark, and Brian Gaylord. 2002. ``The mechanics of wave-swept
algae.'' \emph{Journal of Experimental Biology} 205 (Pt 10): 1355--62.

\leavevmode\hypertarget{ref-Denny1997}{}%
Denny, M, B Gaylord, and E Cowen. 1997. ``Flow and Flexibility: II. The
roles of size and shape in determining wave forces on the bull kelp
Nereocystis luetkeana.'' \emph{Journal of Experimental Biology} 200:
3165--83.

\leavevmode\hypertarget{ref-Dubi1995}{}%
Dubi, Alfonse, and Alf Tørum. 1995. ``Wave Damping by Kelp Vegetation.''
In \emph{Coastal Engineering 1994}, 142--56.

\leavevmode\hypertarget{ref-Dufois2012-vy}{}%
Dufois, François, and Mathieu Rouault. 2012. ``Sea surface temperature
in False Bay (South Africa): Towards a better understanding of its
seasonal and inter-annual variability.'' \emph{Continental Shelf
Research} 43 (July): 24--35.

\leavevmode\hypertarget{ref-Dunne2001}{}%
Dunne, R, and B Brown. 2001. ``The Influence of Solar Radiation on
Bleaching of Shallow Water Reef Corals in the Andaman Sea, 1993--1998.''
\emph{Coral Reefs} 20 (3): 201--10.

\leavevmode\hypertarget{ref-Dyer2018-rv}{}%
Dyer, David Clifford. 2018. ``Stable Isotope Ecology of South African
Kelp Forests.'' Edited by John Bolton, Robert Anderson, and A J Smit.
PhD thesis, University of Cape Town.

\leavevmode\hypertarget{ref-Field1980a}{}%
Field, J G, C L Griffiths, E A Linley, R A Carter, and P Zoutendyk.
1980. ``Upwelling in a nearshore marine ecosystem and its biological
implications.'' \emph{Estuarine and Coastal Marine Science} 11 (2):
133--50. \url{https://doi.org/10.1016/S0302-3524(80)80037-5}.

\leavevmode\hypertarget{ref-FowlerWalker2005}{}%
Fowler-Walker, M J, S D Connell, and B M Gillanders. 2005. ``To what
extent do geographic and associated environmental variables correlate
with kelp morphology across temperate Australia?'' \emph{Marine and
Freshwater Research} 56 (6): 877--87.
\url{https://doi.org/10.1071/MF05042}.

\leavevmode\hypertarget{ref-FowlerWalker2006}{}%
Fowler-Walker, M J, T Wernberg, and S D Connell. 2006. ``Differences in
kelp morphology between wave sheltered and exposed localities:
Morphologically plastic or fixed traits?'' \emph{Marine Biology} 148
(4): 755--67. \url{https://doi.org/10.1007/s00227-005-0125-z}.

\leavevmode\hypertarget{ref-Friedland1995-vi}{}%
Friedland, M T, and M W Denny. 1995. ``Surviving hydrodynamic forces in
a wave-swept environment: consequences of morphology in the feather boa
kelp, \emph{Egregia menziesii} (Turner).'' \emph{Journal of Experimental
Biology} 190: 109--33.

\leavevmode\hypertarget{ref-Gaines1987-bg}{}%
Gaines, S D, and J Roughgarden. 1987. ``Fish in Offshore Kelp Forests
Affect Recruitment to Intertidal Barnacle Populations.'' \emph{Science}
235 (4787): 479--81.

\leavevmode\hypertarget{ref-gao2013}{}%
Gao, Xu, Hikaru Endo, Kazuya Taniguchi, and Yukio Agatsuma. 2013.
``Combined Effects of Seawater Temperature and Nutrient Condition on
Growth and Survival of Juvenile Sporophytes of the Kelp \emph{Undaria
Pinnatifida} (Laminariales; Phaeophyta) Cultivated in Northern Honshu,
Japan.'' \emph{Journal of Applied Phycology} 25 (1): 269--75.

\leavevmode\hypertarget{ref-gerard1997}{}%
Gerard, Valrie A. 1997. ``The Role of Nitrogen Nutrition in
High-Temperature Tolerance of the Kelp, \emph{Laminaria Saccharina}
(Chlorophyta).'' \emph{Journal of Phycology} 33 (5): 800--810.

\leavevmode\hypertarget{ref-Gill1974-nt}{}%
Gill, A E, and A J Clarke. 1974. ``Wind-induced upwelling, coastal
currents and sea-level changes.'' \emph{Deep Sea Research and
Oceanographic Abstracts} 21 (5): 325--45.

\leavevmode\hypertarget{ref-Gorman2013-nj}{}%
Gorman, Daniel, Touria Bajjouk, Jacques Populus, Mickaël Vasquez, and
Axel Ehrhold. 2013. ``Modeling kelp forest distribution and biomass
along temperate rocky coastlines.'' \emph{Marine Biology} 160 (2):
309--25.

\leavevmode\hypertarget{ref-graham2004}{}%
Graham, Michael H. 2004. ``Effects of Local Deforestation on the
Diversity and Structure of Southern California Giant Kelp Forest Food
Webs.'' \emph{Ecosystems} 7 (4): 341--57.

\leavevmode\hypertarget{ref-Algaebase}{}%
Guiry, MD, and GM Guiry. 2019. \emph{AlgaeBase: World-Wide Electronic
Publication}. National University of Ireland, Galway.

\leavevmode\hypertarget{ref-Harley2012}{}%
Harley, C, K M Anderson, K W Demes, J P Jorve, R L Kordas, T Coyle, and
M H Graham. 2012. ``EFfects of Climate Change on Global Seaweed
Communities.'' \emph{Journal of Phycology} 48 (5): 1064--78.
\url{https://doi.org/10.1111/j.1529-8817.2012.01224.x}.

\leavevmode\hypertarget{ref-Holthuijsen2010-vl}{}%
Holthuijsen, Leo H. 2010. \emph{Waves in Oceanic and Coastal Waters}.
Cambridge University Press.

\leavevmode\hypertarget{ref-Hurd2000}{}%
Hurd, Catriona L. 2000. ``Water Motion, Marine Macroalgal Physiology,
and Production.'' \emph{Journal of Phycology} 36 (3): 453--72.

\leavevmode\hypertarget{ref-Jackson1984}{}%
Jackson, G A. 1984. ``Internal Wave Attenuation by Coastal Kelp
Stands.'' \emph{Journal of Physical Oceanography}.
\url{https://doi.org/10.1175/1520-0485(1984)014\%3C1300:IWABCK\%3E2.0.CO;2}.

\leavevmode\hypertarget{ref-Jury1980-ov}{}%
Jury, Mark R. 1980. ``Characteristics of summer wind fields and air sea
interactions over the Cape Peninsula upwelling region.''

\leavevmode\hypertarget{ref-Jury1985-nr}{}%
Jury, M R, F Kamstra, and J Taunton-Clark. 1985. ``Synoptic summer wind
cycles and upwelling off the southern portion of the Cape Peninsula.''
\emph{South African Journal of Marine Science} 3 (1): 33--42.

\leavevmode\hypertarget{ref-Kaplan2003}{}%
Kaplan, David M, John L Largier, Sergio Navarrete, Ricardo Guiñez, and
Juan Carlos Castilla. 2003. ``Large Diurnal Temperature Fluctuations in
the Nearshore Water Column.'' \emph{Estuarine, Coastal and Shelf
Science} 57 (3): 385--98.

\leavevmode\hypertarget{ref-Kobayashi1993-os}{}%
Kobayashi, Nobuhisa, Andrew W Raichle, and Toshiyuki Asano. 1993. ``Wave
attenuation by vegetation.'' \emph{J. Waterway Port Coast. Ocean Eng.}
119 (1): 30--48.

\leavevmode\hypertarget{ref-Krumhansl2012}{}%
Krumhansl, Kira A., and Robert E. Scheibling. 2012. ``Production and
fate of kelp detritus.'' \emph{Marine Ecology Progress Series} 467:
281--302. \url{https://doi.org/10.3354/meps09940}.

\leavevmode\hypertarget{ref-Leliaert2000a}{}%
Leliaert, F, R J Anderson, J Bolton, and E Coppejans. 2000. ``Subtidal
Understorey Algal Community Structure in Kelp Beds around the Cape
Peninsula (Western Cape, South Africa).'' \emph{Botanica Marina} 43 (4):
359--66. \url{https://doi.org/10.1515/BOT.2000.036}.

\leavevmode\hypertarget{ref-Levin1994-qf}{}%
Levin, P S. 1994. ``Fine-scale temporal variation in recruitment of a
temperate demersal fish: the importance of settlement versus
post-settlement loss.'' \emph{Oecologia} 97 (1): 124--33.

\leavevmode\hypertarget{ref-Mia2013}{}%
Ma, Gangfeng, James T Kirby, Shih-Feng Su, Jens Figlus, and Fengyan Shi.
2013. ``Numerical Study of Turbulence and Wave Damping Induced by
Vegetation Canopies.'' \emph{Coastal Engineering} 80: 68--78.

\leavevmode\hypertarget{ref-mabin2013}{}%
Mabin, CJT, PE Gribben, and A Fischer. 2013. ``Variation in the
morphology, reproduction and development of the habitat-forming kelp
Ecklonia radiata with changing temperature and nutrients.'' \emph{Marine
Ecology Progress}.
\url{http://www.int-res.com/abstracts/meps/v483/p117-131}.

\leavevmode\hypertarget{ref-Mann1973}{}%
Mann, K H. 1973. ``Seaweeds: Their Productivity and Strategy for
Growth.'' \emph{Science} 182 (4116): 975--81.

\leavevmode\hypertarget{ref-Miller2011-pm}{}%
Miller, Sheryl M, Catriona L Hurd, and Stephen R Wing. 2011.
``Variations in Growth, Erosion, Productivity, and Morphology of
\emph{Ecklonia Radiata} (Alariaceae; Laminariales) Along a Fjord in
Southern New Zealand.'' \emph{Journal of Phycology} 47 (3): 505--16.

\leavevmode\hypertarget{ref-Mohring2014-bk}{}%
Mohring, M B, T Wernberg, J T Wright, S D Connell, and B D Russell.
2014. ``Biogeographic variation in temperature drives performance of
kelp gametophytes during warming.'' \emph{Marine Ecology Progress
Series} 513: 85--96.

\leavevmode\hypertarget{ref-Molloy1996-la}{}%
Molloy, F J, and J J Bolton. 1996. ``The effects of wave exposure and
depth on the morphology of inshore populations of the Namibian kelp,
\emph{Laminaria schinzii} Foslie.'' \emph{Botanica Marina} 39: 525--31.

\leavevmode\hypertarget{ref-Mork1996-md}{}%
Mork, Martin. 1996. ``The effect of kelp in wave damping.''
\emph{Sarsia} 80 (4): 323--27.

\leavevmode\hypertarget{ref-Norderhaug2012-ks}{}%
Norderhaug, K M, H Christie, G S Andersen, and T Bekkby. 2012. ``Does
the diversity of kelp forest macrofauna increase with wave exposure?''
\emph{Journal of Sea Research} 69: 36--42.

\leavevmode\hypertarget{ref-Oksanen2015-vj}{}%
Oksanen, Jari, F Guillaume Blanchet, Roeland Kindt, Pierre Legendre,
Peter R Minchin, R B O'hara, Gavin L Simpson, et al. 2015. ``Package
`Vegan'.'' \emph{Community Ecology Package, Version} 2 (9).

\leavevmode\hypertarget{ref-Paul1993-bm}{}%
Paul, A X, and Robert Steneck. 1993. ``Kelp Beds as Habitat for American
Lobster \emph{Homarus americanus}.'' \emph{Marine Ecology Progress
Series} 100 (40545): 127.

\leavevmode\hypertarget{ref-Pedersen2012-vr}{}%
Pedersen, M F, and L B Nejrup. 2012. ``Effects of wave exposure on
population structure, demography, biomass and productivity of the kelp
\emph{Laminaria hyperborea}.'' \emph{Marine Ecology Progress Series}
451: 45--60.

\leavevmode\hypertarget{ref-Rautenbach2015a}{}%
Rautenbach, C. 2015. ``Southern African Coastal Vulnerability
Assessment.'' In \emph{Southern African Coastal Vulnerability
Assessment}, edited by CSIR.

\leavevmode\hypertarget{ref-Reason2006}{}%
Reason, CJC, W Landman, and W Tennant. 2006. ``Seasonal to Decadal
Prediction of Southern African Climate and Its Links with Variability of
the Atlantic Ocean.'' \emph{Bulletin of the American Meteorological
Society} 87 (7): 941--56.

\leavevmode\hypertarget{ref-Reguero2019}{}%
Reguero, Borja G, Iñigo J Losada, and Fernando J Méndez. 2019. ``A
Recent Increase in Global Wave Power as a Consequence of Oceanic
Warming.'' \emph{Nature Communications} 10.

\leavevmode\hypertarget{ref-Rinde2014}{}%
Rinde, E, H Christie, C W Fagerli, T Bekkby, H Gundersen, K M
Norderhaug, and D Hjermann. 2014. ``The influence of physical factors on
kelp and sea urchin distribution in previously and still grazed areas in
the NE Atlantic.'' \emph{PLoS ONE} 9 (6).
\url{https://doi.org/10.1371/journal.pone.0100222}.

\leavevmode\hypertarget{ref-Rose2012}{}%
Rose, TH, DA Smale, and G Botting. 2012. ``The 2011 Marine Heat Wave in
Cockburn Sound, Southwest Australia.'' \emph{Ocean Science} 8 (4):
545--50.

\leavevmode\hypertarget{ref-Rothman2017-pn}{}%
Rothman, M D, J J Bolton, M S Stekoll, C J T Boothroyd, F A Kemp, and R
J Anderson. 2017. ``Geographical variation in morphology of the two
dominant kelp species, \emph{Ecklonia maxima} and \emph{Laminaria
pallida} (Phaeophyceae, Laminariales), on the west coast of Southern
Africa.'' \emph{Journal of Applied Phycology} 29 (5): 2627--39.

\leavevmode\hypertarget{ref-Rouault2010-gi}{}%
Rouault, M, B Pohl, and P Penven. 2010. ``Coastal oceanic climate change
and variability from 1982 to 2009 around South Africa.'' \emph{African
Journal of Marine Science} 32 (2): 237--46.

\leavevmode\hypertarget{ref-Rummukainen2012}{}%
Rummukainen, Markku. 2012. ``Changes in Climate and Weather Extremes in
the 21st Century.'' \emph{Wiley Interdisciplinary Reviews: Climate
Change} 3 (2): 115--29.

\leavevmode\hypertarget{ref-Russo2014}{}%
Russo, Simone, Alessandro Dosio, Rune G Graversen, Jana Sillmann, Hugo
Carrao, Martha B Dunbar, Andrew Singleton, Paolo Montagna, Paulo
Barbola, and Jürgen V Vogt. 2014. ``Magnitude of Extreme Heat Waves in
Present Climate and Their Projection in a Warming World.'' \emph{Journal
of Geophysical Research: Atmospheres} 119 (22): 12--500.

\leavevmode\hypertarget{ref-Serisawa2002-dc}{}%
Serisawa, Yukihiko, Hideki Akino, Kazuyo Matsuyama, Masao Ohno, Jiro
Tanaka, and Yasutsugu Yokohama. 2002. ``Morphometric study of
\emph{Ecklonia cava} (Laminariales, Phaeophyta) sporophytes in two
localities with different temperature conditions.'' \emph{Phycological
Research} 50 (3): 193--99.

\leavevmode\hypertarget{ref-seymour1989}{}%
Seymour, RJ, MJ Tegner, PK Dayton, and PE Parnell. 1989. ``Storm Wave
Induced Mortality of Giant Kelp, Macrocystis Pyrifera, in Southern
California.'' \emph{Estuarine, Coastal and Shelf Science} 28: 277--92.

\leavevmode\hypertarget{ref-Shipley1964}{}%
Shipley, A. M. 1964. ``Some aspects of wave refraction in False Bay.''
\emph{South African Journal of Science} 60 (6): 115--20.

\leavevmode\hypertarget{ref-Simonson2015}{}%
Simonson, EJ, RE Scheibling, and A Metaxas. 2015. ``Kelp in Hot Water:
I. Warming Seawater Temperature Induces Weakening and Loss of Kelp
Tissue.'' \emph{Marine Ecology Progress Series} 537: 89--104.

\leavevmode\hypertarget{ref-Smale2017}{}%
Smale, D A, and P J Moore. 2017. ``Variability in kelp forest structure
along a latitudinal gradient in ocean temperature.'' \emph{Journal of
Experimental Marine Biology and Ecology} 486: 255--64.
\url{https://doi.org/10.1016/j.jembe.2016.10.023}.

\leavevmode\hypertarget{ref-Smit2013a}{}%
Smit, A J, M Roberts, R J Anderson, F Dufois, S Dudley, T Bornman, J
Olbers, and J J Bolton. 2013. ``A coastal seawater temperature dataset
for biogeographical studies: Large biases between in situ and
remotely-sensed data sets around the coast of South Africa.'' \emph{PLoS
ONE} 8 (12). \url{https://doi.org/10.1371/journal.pone.0081944}.

\leavevmode\hypertarget{ref-Smit2017}{}%
Smit, Albertus J, John J Bolton, and Robert J Anderson. 2017. ``Seaweeds
in Two Oceans: Beta-Diversity.'' \emph{Frontiers in Marine Science} 4:
404.

\leavevmode\hypertarget{ref-Steneck2002}{}%
Steneck, R S, M H Graham, B J Bourque, D Corbett, J M Erlandson, J A
Estes, and M J Tegner. 2002. ``Kelp forest ecosystems: biodiversity,
stability, resilience and future.'' \emph{Environmental Conservation} 29
(04): 436--59. \url{https://doi.org/10.1017/S0376892902000322}.

\leavevmode\hypertarget{ref-Stewart2009}{}%
Stewart, H L, J P Fram, D C Reed, S L Williams, M A Brzezinski, S
Maclntyreb, and B Gaylord. 2009. ``Differences in growth, morphology and
tissue carbon and nitrogen of \emph{Macrocystis pyrifera} within and at
the outer edge of a giant kelp forest in California, USA.'' \emph{Marine
Ecology Progress Series} 375: 101--12.
\url{https://doi.org/10.3354/meps07752}.

\leavevmode\hypertarget{ref-Thomsen2004}{}%
Thomsen, M S, T Wernberg, and G A Kendrick. 2004. ``The effect of
thallus size, life stage, aggregation, wave exposure and substratum
conditions on the forces required to break or dislodge the small kelp
\emph{Ecklonia radiata}.'' \emph{Botanica Marina}.

\leavevmode\hypertarget{ref-Veitch2019}{}%
Veitch, Jennifer, Christo Rautenbach, Juliet Hermes, and Chris Reason.
2019. ``The Cape Point Wave Record, Extreme Events and the Role of
Large-Scale Modes of Climate Variability.'' \emph{Journal of Marine
Systems}, 103185.

\leavevmode\hypertarget{ref-Waldron1992}{}%
Waldron, HN, and TA Probyn. 1992. ``Nitrate Supply and Potential New
Production in the Benguela Upwelling System.'' \emph{South African
Journal of Marine Science} 12 (1): 29--39.

\leavevmode\hypertarget{ref-Weisse2010}{}%
Weisse, Ralf. 2010. \emph{Marine Climate and Climate Change: Storms,
Wind Waves and Storm Surges}. Springer Science \& Business Media.

\leavevmode\hypertarget{ref-Wernberg2003-mp}{}%
Wernberg, T, M Coleman, A Fairhead, S Miller, and M Thomsen. 2003.
``Morphology of \emph{Ecklonia radiata} (Phaeophyta: Laminarales) along
its geographic distribution in south-western Australia and
Australasia.'' \emph{Marine Biology} 143 (1): 47--55.

\leavevmode\hypertarget{ref-Wernberg2008}{}%
Wernberg, T, and N Goldberg. 2008. ``Short-term temporal dynamics of
algal species in a subtidal kelp bed in relation to changes in
environmental conditions and canopy biomass.'' \emph{Estuarine, Coastal
and Shelf Science} 76 (2): 265--72.
\url{https://doi.org/10.1016/j.ecss.2007.07.008}.

\leavevmode\hypertarget{ref-Wernberg2010b}{}%
Wernberg, Thomas, Mads S Thomsen, Fernando Tuya, Gary A Kendrick, Peter
A Staehr, and Benjamin D Toohey. 2010. ``Decreasing Resilience of Kelp
Beds Along a Latitudinal Temperature Gradient: Potential Implications
for a Warmer Future.'' \emph{Ecology Letters} 13 (6): 685--94.

\leavevmode\hypertarget{ref-Wernberg2010a}{}%
Wernberg, Thomas, and Mathew A Vanderklift. 2010. ``Contribution of
Temporal and Spatial Components to Morphological Variation in the Kelp
\emph{Ecklonia} (Laminariales).'' \emph{Journal of Phycology} 46 (1):
153--61.

\leavevmode\hypertarget{ref-Wernberg2005c}{}%
Wernberg, T, and M S Thomsen. 2005. ``The effect of wave exposure on the
morphology of \emph{Ecklonia radiata}.'' \emph{Aquatic Botany} 83 (1):
61--70. \url{https://doi.org/10.1016/j.aquabot.2005.05.007}.

\leavevmode\hypertarget{ref-Willis2003-xo}{}%
Willis, T J, and M J Anderson. 2003. ``Structure of cryptic reef fish
assemblages: relationships with habitat characteristics and predator
density.'' \emph{Marine Ecology Progress Series} 257: 209--21.

\leavevmode\hypertarget{ref-Wing2007-tv}{}%
Wing, S R, J J Leichter, C Perrin, S M Rutger, M H Bowman, and C D
Cornelisen. 2007. ``Topographic shading and wave exposure influence
morphology and ecophysiology of \emph{Ecklonia radiata} (C. Agardh 1817)
in Fiordland, New Zealand.'' \emph{Limnology and Oceanography} 52 (5):
1853--64.

\leavevmode\hypertarget{ref-zimmerman1986}{}%
Zimmerman, Richard C, and James N Kremer. 1986. ``In Situ Growth and
Chemical Composition of the Giant Kelp, Macrocystis Pyrifera: Response
to Temporal Changes in Ambient Nutrient Availability.'' \emph{Marine
Ecology Progress Series} 27 (2): 277--85.

\end{document}

### 2021-03-12 thesis_chapter_3.tex
adding support files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Numerical trajectory patterns of floating macroalgae},
  pdfauthor={R Coppin; C Rautenbach; AJ Smit},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}

\title{Numerical trajectory patterns of floating macroalgae}
\author{R Coppin\footnote{University of the Western Cape} \and C Rautenbach\footnote{South African Weather Service} \and AJ Smit\footnote{University of the Western Cape}}
\date{}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

There is a range of objects, both natural and anthropogenic, floating in
the ocean of which macroalgae are regarded as one of the most important
passive dispersal mechanisms of marine taxa. Floating kelp acts as a
passive dispersal mechanism for a range of marine taxa and is sometimes
referred to as the `tumble-weed' of the ocean (Edgar, 1987; Norton,
1992; Bushing, 1994; Helmuth et al., 1994; Holmquist, 1994; Smith, 2002;
McCormick et al., 2008). Some macroalgae species are negatively buoyant
and sink to the seafloor when detached from the substrate, while other
species of macroalgae have air-filled pneumatocysts or stipes which
allow the plants to reach the surface where light is more abundant. In
turn, the positively buoyant pneumatocysts cause plants to float to the
surface when dislodged from the substratum. The structure and number of
pneumatocysts varies between species. For example species from the
genera \emph{Macrocystis}, \emph{Sargassum}, \emph{Ascophyllum}, and
\emph{Fucus} have thalli with many small pneumatocysts, while other kelp
species such as \emph{Nereocystis luetkeana}, \emph{Pelagophycus porra},
\emph{Ecklonia radiata}, \emph{Ecklonia maxima} have a single, large
pneumatocysts (Dayton, 1985; Smith, 2002; Thiel and Gutow, 2005; Graiff
et al., 2016; Batista et al., 2018). In some cases the stipe itself is
air-filled, such as with \emph{E. maxima} and \emph{E. radiata}.
Although there have been reports of floating Chlorophyta species and
some Rhodophyta species, Phaeophyceae species are the most commonly
reported forms of floating algae. This is most likely because the green
and red species reported floating are not actually positively buoyant,
but instead are kept at the surface by gas trapped inbetween or in the
thalli (Dromgoole, 1982; Bäck et al., 2000). The giant kelp
\emph{Macrocystis pyrifera} (Helmuth et al., 1994; Kingsford, 1995;
Hobday, 2000; Macaya et al., 2005; Graiff et al., 2016; Batista et al.,
2018) and the bull kelp \emph{Durvillaea antarctica} (Smith, 2002;
Collins et al., 2010; Wichmann et al., 2012; Tala et al., 2013 , 2017;
Saunders, 2014; Batista et al., 2018) have been the focus of much of the
research regarding spatial and temporal dispersal patterns, the
ecological role of rafting, marine connectivity and raft-time.

Past research points to macroalgae trajectory being largely determined
by prevailing wind conditions and surface currents (Hobday, 2000; Thiel
and Gutow, 2005, @thiel2005; Fraser et al., 2011; Rothäusler et al.,
2011, 2011). Although ocean currents are regarded as the primary
influence, the relative importance of wind versus surface current is
still not known; although the role of wind has been recognised as
important in several studies. For example, a study by Harrold and Lisin
(1989) investigated the seasonal trajectories of radio-transmitter
tagged \emph{M. pyrifera} in nearshore Monterrey Bay. The results showed
that kelp rafts with little surface area exposed to the wind were
largely driven by a combination of wind and wind waves, however the
relative importance of wind and wind waves was not clear. In addition,
the tagged kelp trajectories were more consistent with the formation of
eddies during winter. Previous studies have identified wind as an
important mechanism of dispersal in wind dominated ocean systems. For
example the subAntartic latitudes the West Wind Drift causes continuous
unidirectional surface flow and is regarded as an important potential
mechanism for dispersal of floating kelp. Other studies have used
genetic approaches to determine macroalgae raft trajectory
characteristics by inferring source location from genetically similar
populations (Nikula et al., 2013). For example, a study by Fraser et al.
(2011) on the rafting capabilities of \emph{Durvillaea antarctica} used
a combination of population genetics and relative age estimate of
`goosebarnacles' attached to the raft. The presence of goosebarnacles
suggests a long raft time as these species have a slow growth rate;
while the genetic analyses showed these species are able to raft up to
\(\sim\) 390km from their local origin. The authors suggested that wind
and water-movement were the primary influences of trajectory, however,
this was only inferred from the genetic results and local climatology
data (Fraser et al., 2011).

Other aspects such as buoyancy and drag also play a role in determining
the trajectory and rate of transport for surface floating material.
However, the past research on macroalgal trajectory has not investigated
these factors which have been shown to be important aspects of
trajectory for other materials such as icebergs, marine craft and
microplastics. The ``sail'' area of an object floating at sea is the
surface area exposed to the wind which results in air form drag, while
the area of the object below the surface of the water is exposed to
surface currents which result in hydrodynamic form drag. Drag
coefficients related to wind and surface current have been shown to be
important properties to consider when estimating trajectory and
forecasting drift for search and sea rescue operations. Drag is
ultimately determined by the size and shape of the object which are
properties that vary considerably with macroalgae species. In addition,
past research conducted within the maritime industry has shown that the
size and shape of a vessel determine the relative importance of waves or
wind as drivers of trajectory, as well as orientation of the object. If
the length of the object is longer than the significant wave height then
waves will be the primary driver of trajectory while the opposite is
true for smaller objects where the effect of waves is regarded as
negligible (Breivik et al., 2011; Griffin et al., 2017).

To accurately determine the trajectory of marine macroalgae these
aspects need to be taken into account. Past research by Allen and
Plourde (1999) and Breivik et al. (2011) have provided estimates of drag
for various objects based on experimental work, which consisted of both
direct and indirect methods of the Leeway model. However, this work does
not consider biological material such as macroalgae. Although drag
estimates do exist, these have been applied for macroalgae not detached
from the substratum and are regarded as fixed-point estimates. Although
various approaches have been used in the past to investigate floating
macroalagae trajectory, very few studies have employed the use of
Langrangian trajectory modeling. Furthermore, non of the existing
studies using this modeling approach have considered macroalgal
morphology cross-sectional area (i.e.~shape) as an aspect of drag and
ultimately trajectory. Studies by Brooks et al. (2019), Putman et al.
(2018), Putman et al. (2020) used lagrangian approaches to investigate
the effects of inertia, raft-size, and windage on macroalgae trajectory.
Brooks et al. (2019) used a custom growth model to estimate changes in
biomass and ultimately radial size, while a customised Hybrid Coordinate
Ocean Model (HYCOM) for the trajectory simulations was used. The results
showed that trajectory of pelagic \emph{Sargussum} was significantly
influenced by inertia and the radial size of the rafts. Other work by
Putman et al. (2020) investigated the effect of including windage in
macroalgae trajectory simulations. The authors showed that including
ad-hoc windage factors greatly improved accuracy regarding entrainment
or expulsion from eddies when compared to tracked sargassum mats in the
ocean. Langrangian trajectory modeling is a useful tool, however, the
various models available often assume the object to be a spherical
particle, which is obviously not the case for both biological and
anthropogenic forms of marine debris in nature. The radial size included
by Brooks et al. (2019) is an improvement, however this approach does
not take into account the complex morphology on an individual level into
account.

In this study, a numerical approach is used to shed light on the role of
hydrodynamic and direct wind drag on macroalgal drift trajectory using
lagrangian based trajectory simulations. This will be achieved by
comparing the trajectory outputs of virtual `passive' particles (no drag
characteristics added) with kelp particles (drag characteristics added)
of varying degrees of hydrodynamic and wind drag exposure. The different
degrees of wave and wind exposure are meant to reflect the different
buoyancy characteristics, as well as identify the role of magnitude of
the drag components on overall trajectory.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{study-site}{%
\subsection{Study site}\label{study-site}}

The region of interest for this study was the Cape Peninsula in the
Western Cape (Figure 1A). A release site was chosen for this study,
Kommetjie, which is located on the Atlantic side of the Cape Peninsula
(Figure 2B). The hydrodynamics in the study region are driven by a
complex interaction of wind and waves. which are also influenced by
larger oceanic processes In South Africa, the larger processes can
largely be attributed to the Agulhas Current (AC) and the Benguela
Current (BC). The AC is a component of the South-west Indian Ocean
sub-gyre, flowing predominantly south-westward following the continental
shelf edge and eventually retroflecting eastward back into the South
Indian Ocean. The region where the retroflection occurs is known as the
Agulhas retroflection and consists of some of the highest levels of
mesoscale variability (Garzoli et al., 1996). The variability results in
shedding of rings and filaments known as the Agulhas Rings and have an
average diameter of \textasciitilde324km (Lutjeharms and Van
Ballegooyen, 1988; Lutjeharms, 2006, @ 2007). The Agulhas Rings are the
main contributors to the Agulhas leakage which is the transport of warm
water from the Indian Ocean to the South Atlantic Ocean (Garzoli et al.,
1996; Lutjeharms, 2006; Beal et al., 2015). The Agulhas leakage is also
a source of meso-scale variability within the Benguela region as
anticyclonic rings are shed and drift northwestward along the southern
Benguela slope region (Rubio et al., 2009). The Benguela region is
characterised by strong coastal upwelling cells which a primarily driven
by equatorward winds (Shannon and Nelson, 1996; Hardman-Mountford et
al., 2003), and is known as the Benguela upwelling region. The Benguela
upwelling region starts at 27\(^\circ\)S and extends southward to
35\(^\circ\)S (Shannon and Nelson, 1996). The region is also
characterised by high amounts of mesoscale variability in the form of
eddies and filaments(Blanke et al., 2002, 2005; Rubio et al., 2009). The
Benguela inshore region is dominated by a northwestward flow as a result
of topographical steering, wind stress and interactions with passing
Agulhas rings and eddies (Veitch et al., 2010).

\begin{figure}
\centering
\includegraphics{thesis_chapter_3_files/figure-latex/map-1.pdf}
\caption{Figure 1:Map of the study domain. This is a temporary map until
issues with the map script are resolved.}
\end{figure}

\hypertarget{particle-tracking-model}{%
\subsection{Particle tracking model}\label{particle-tracking-model}}

The model is based on parameterisations that have been used for previous
studies investigating iceberg, capsized marine vessels and
microplastics. It should be noted that most of the previous studies
regarding macroalgae trajectory have investigated rafts and not solitary
floating individuals. A kelp raft will vary in size and can range from a
few meters across to large dense mats the size of a sports field.
Furthermore, some authors suggest that most macroalgae become entangled
through ocean currents and not through the hydrodynamic forces that
dislodge them. In addition, large kelp rafts are not a characteristics
around the coast of the Cape Peninsula but rather solitary kelp are
(Pers. Obs). Therefore, the trajectory of only solitary \emph{E. maxima}
individuals will be considered. Numerical calculations of particle
trajectory with a Lagrangian method (Delandmeter and Van Sebille, 2019),
and be described by the equation

\[X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)\]
where \(X\) is the three-dimensional position of a particle, \(v(x,t)\)
is the three-dimensional velocity field at the location in the ocean
model, \(X_b(t)\) is a change in position due to `behavior' of the
particle. This can range from swimming in fish to sinking or beaching.
IN order to investigate the effects of drag components of trajectory of
floating macroalgae, this study incorporated drag as the custom
behavior.

This study simulated two scenarios; one where the plant is fully
submerged and only exposed to hydrodynamic drag; and the other where
parts of the plants are partially exposed resulting in a combination of
hydrodynamic and air form drag. Drag based on shape based coefficients
were used in the calculations for determining the overall kelp velocity
vectors. A momentum energy equation was used to calculate the drag
forces for the relevant simulation. This approach has been employed in
modeling iceberg drift trajectory (Lichey and Hellmer, 2001; Eik, 2009;
Andersson et al., 2017) and was adapted to suit this particular study.
The energy momentum equation used to calculate the drag force exerted on
the virtual kelp particle was, \[m({d_u/d_t})=F_d + F_a + F_f\] where
\(m\) is kelp mass, \(d_u\) the virtual kelp particle velocity, \(F_d\)
the hydrodynamic drag, \(F_a\) is the wind drag, and \(F_f\) the
surface-current flow. The assumption made was that the mass of the kelp
did not change for each virtual kelp particle over the course of the
simulations. Hydrodynamic drag, wind drag and surface current-flow are
two-dimensional vector quantities and were calculated for each time step
for each simulation.

\hypertarget{hydrodynamic-and-wind-model}{%
\subsection{Hydrodynamic and wind
model}\label{hydrodynamic-and-wind-model}}

Both passive and kelp particles were simulated around the South African
coastline within hind-cast outputs from the Copernicus Marine
Environment Monitoring Service. The Copernicus outputs used in this
study contains 3D daily current information from the top layer to the
bottom (Global Analysis Forecast PHY\_001\_024). The model has a spatial
resolution of 0.08\(^\circ\) and is interpolated on a Arakawa C native
grid.

In order to incorporate the effects of direct wind drag within the
simulations, the Copernicus Marine Environment Monitoring Service global
wind product was used (WIND\_L4\_NRT\_OBSERVATIONS\_012\_004). The
outputs from the model is composed of 6-hourly averaged fields of
surface 10m wind velocity vectors with a spatial resolution of
0.25\(^\circ\). Both averaged ocean current and wind data are stored as
meridonal and zonal velocity vectors.

\hypertarget{simulations}{%
\subsection{Simulations}\label{simulations}}

Virtual particles were released with no drag or wind effects from the
study site in order to act as a basis for comparison to simulations with
varying degrees of hydrodynamic and wind drag, and are referred to
hereafter as `passive particles'. All simulations have the same release
location and were run for 30 days with hourly outputs using a fourth
order Runge-Kutte advection scheme. Brownian motion was also
incorporated in order to account for any stochasticity not resolved by
the model. These are standard practices for Lagrangian trajectory
studies.

\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}

\setlength{\tabcolsep}{8pt}

\renewcommand*{\arraystretch}{1.5}

\begin{table}

\centering

\begin{longtable}{ccccc}

\caption{Table 1:Details for each simulation executed for the study.}\\

\hhline{~~~~~}

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Scenario}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{ID}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Type}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Cross-sectional area}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Drag Exposure level}}}} \\

\endfirsthead

\hhline{~~~~~}

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Scenario}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{ID}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Type}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Cross-sectional area}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Drag Exposure level}}}} \\

\endhead

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{RK4\_Pfloat}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Passive}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{N/A}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{N/A}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Min\_Kfloat\_H100W00}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Minimum}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag only}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{3}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Min\_Kfloat\_H095W005}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Minimum}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag 95\%, wind drag 5\%}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Min\_Kfloat\_H090W10}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Minimum}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag 90\%, wind drag 10\%}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{5}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Min\_Kfloat\_H085W15}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Minimum}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag 85\%, wind drag 15\%}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{6}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Mean\_Kfloat\_H100W00}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Mean}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag only}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Mean\_Kfloat\_H095W005}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Mean}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag 95\%, wind drag 5\%}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Mean\_Kfloat\_H090W10}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Mean}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag 90\%, wind drag 10\%}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{9}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Mean\_Kfloat\_H085W15}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Mean}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag 85\%, wind drag 15\%}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{10}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Max\_Kfloat\_H100W00}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Maximum}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag only}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{11}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Max\_Kfloat\_H095W005}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Maximum}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag 95\%, wind drag 5\%}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{12}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Max\_Kfloat\_H090W10}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Maximum}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag 90\%, wind drag 10\%}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{13}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Max\_Kfloat\_H085W15}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Maximum}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag 85\%, wind drag 15\%}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{14}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Max\_ShapeFloat\_H085W15}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Kelp}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Maximum area, cylinder shape}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Hydrodynamic drag only}}} \\

\end{longtable}

\end{table}

In order to determine the influence of cross-sectional area on drag
components, simulations were performed for the minimum, mean and maximum
cross-sectional areas for Kommetjie kelp forest population (Figure 1).
In order to determine the direct effect of wind on trajectory,
simulations were run for varying magnitudes of wind exposure for each
type of cross-sectional area. The virtual particles released with the
hydrodynamic or wind drag components are referred to as `kelp
particles'.

\hypertarget{model-inputs}{%
\subsection{Model inputs}\label{model-inputs}}

\hypertarget{cross-sectional-area}{%
\subsubsection{Cross-sectional area}\label{cross-sectional-area}}

In order to incorporate hydrodynamic and wind drag, the cross-sectional
area of the kelp was calculated first. Known geometric shapes reflecting
the relevant plant sections were used to estimate the surface area for
various parts of the plant, for details please refer to table 2. The
dimensional data needed was estimated in cases were data was not
available for that particular morphological characteristic. The
bulb/pneumatocyst is a highly variable morphological characteristic and
in some cases can appear absent, the same is true for the holdfast.
Therefore, the cross-sectional area of the bulb/pneumatocyst was not
considered in the calculation of overall cross-sectional area. In terms
of the holdfast, a standard cross-sectional area was used for all
simulations. The cross-sectional area calculated was site specific and
the dimensions needed were garnered from morphology data from a previous
study (Coppin et al., 2020). The minimum, mean and maximum were
calculated for each morphological characteristic (see appendix), which
were used for calculating the overall minimum, mean and maximum
cross-sectional areas needed to run the various simulations. The
trajectory of these ``types'' of cross-sectional areas (minimum, mean
and maximum overall cross-sectional area) were used to determine the
influence of drag, both hydrodynamic and wind, on overall trajectory.

\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}

\setlength{\tabcolsep}{8pt}

\renewcommand*{\arraystretch}{1.5}

\begin{table}

\centering

\begin{longtable}{ccccccc}

\caption{Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.}\\

\hhline{~~~~~~~}

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Plant characteristic}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Approximate shape}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Equation Used}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Plant dimensions}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Ac Minimum}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Ac Mean}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Ac Maximum}}}} \\

\endfirsthead

\hhline{~~~~~~~}

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Plant characteristic}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Approximate shape}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Equation Used}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Plant dimensions}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Ac Minimum}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Ac Mean}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{\textbf{Ac Maximum}}}} \\

\endhead

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Secondary blade}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Rectangle}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Ac = 2lw + 2lh + 2wh}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{frond length, frond width*,frond thickness*}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{1,136}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{1,993}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{2,550}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Primary blade}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Rhombus}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Ac = (lxb)/2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{primary length, primary width}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{96}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{262}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{536}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Bulb/pneomatocyst}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Capsule}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Ac = 4?r2 + 2?rh}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{bulb length*, bulb base radius*}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{628}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{628}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{628}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Stipe}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Cylinder}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Ac = 2?r(r + h)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{stipe radius from stipe circumference, stipe length}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{9,797}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{19,338}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{34,273}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Holdfast area}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Conical frustrum}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Ac = ?(R2 + r2) + ?(R+r)?(R-r)2 + h2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{top radius*, bottom radius*, height*}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{640}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{640}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{640}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Total Area (centimeters)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{12,297}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{22,861}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{38,628}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Total area (meters)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{123}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{229}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{386}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Total weight}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{N/A}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{mass (kg) = stipe mass + (frond mass x 6)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{Total weight (kg)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{17}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{34}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{}c!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{13}\selectfont{\textcolor[HTML]{000000}{49}}} \\

\end{longtable}

\end{table}

\hypertarget{hydrodynamic-and-wind-drag}{%
\subsubsection{Hydrodynamic and wind
drag}\label{hydrodynamic-and-wind-drag}}

The drag coefficient for a spherical particle was used in the
calculation of hydrodynamic drag as the trajectory model assumes the
particles are spherical. The drag equation used was:
\(F_d = 0.5 \rho A_c C_d U^2\), where \(\rho\) is the density of
seawater or air, \(A_c\) is the cross-sectional area exposed to
hydrodynamic or wind drag, \(C_d\) is the drag coefficient for a
spherical particle and \(U\) is the surface velocity vector of the
flow-/wind- field. Hydrodynamic and wind drag components were calculated
separately for both the meridonal and zonal velocity vectors. Since drag
force is dependent on the velocity vectors which vary with time, the
meridonal and zonal velocities were interpolated and used in the drag
force calculation for each time-step. The same approach was used for the
wind drag force. The equation that was used to calculate hydrodynamic
drag, was also used to calculate wind. To test the different magnitudes
that hydrodynamic and wind drag might play in the overall trajectory,
simulations were run with varying degrees of drag exposure scenarios.
These scenarios were meant to act as proxy for buoyancy, which
determines the surface area exposed to hydrodynamic and wind drag. These
estimates were expressed as percentages which were used to calculate
hydrodynamic and wind drag for applicable simulation. For example, if
80\% of the plant is submerged and 20\% of above water, then the overall
hydrodynamic and wind drag forces would be 85\% and 15\% of the total
drag forces respectively. The exposure scenarios used in this study were
100\%, 95\%, 90\%, and 85\% for hydrodynamic drag; 0 \%, 5\%, 10\%, and
15\% for wind (see table 1).

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

In order to compare trajectories between cross-sectional area types and
wind exposure scenarios, mean trajectories were calculated and density
distribution maps were produced. Density distribution maps are an
effective method for assessing pathways and ocean connectivity
({\textbf{???}}). In addition, mean trajectory was calculated and
included in the density distribution maps to further aid in comparing
simulation outputs. The distance that particles traveled was also
measured and used to produce boxplots. Distance traveled from the
release site can help reveal topographic steering and momentum energy
transfer from the wind field to kelp particle ({\textbf{???}}). If the
wind direction is in the same direction as the current the virtual
particles will cover larger distances due to an increase in the velocity
as a result of momentum energy transfer ({\textbf{???}}; Putman et al.,
2018, 2020).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{trajectory-comparison}{%
\subsection{Trajectory comparison}\label{trajectory-comparison}}

The mean trajectories of the different cross-sectional areas are similar
to the mean trajectory of the passive particles (Figure 2A). When
comparing mean trajectories between the different cross-sectional areas
for the 100\% hydrodynamic drag simulations, no differences can be seen
with only slight variation. The results from the simulations including
different hydrodynamic and wind drag exposures and cross-sectional area
types show differences in mean trajectory when compared to the passive
simulation (Figure 2B). The end points of the simulations including
hydrodynamic and wind drag were different compared to the passive
simulations, as well as being further from the release site. When
comparing the different hydrodynamic and wind drag simulations to each
other, only slight variation in mean trajectories was evident. In all
simulations particles flowed in a north-westward direction.

\begin{figure}
\centering
\includegraphics{thesis_chapter_3_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{Figure 2: Comparison of mean trajectories. Plot A is the
comparison of mean trajectories for only hydrodynamic drag and plot B is
the comparison of mean trajectories for simulations with varying degrees
of hydrodynamic and wind drag.}
\end{figure}

\hypertarget{comparison-of-density-distributions}{%
\subsection{Comparison of density
distributions}\label{comparison-of-density-distributions}}

Across all simulations particles flowed in a north-westward direction
(Figure 3). The density plots for simulations that considered any form
of drag showed a higher density of particles along the mean trajectory
path compared to the passive simulation which has an almost even density
of particles across grids (Figure 3A, 3E, 3I). In addition the density
plots show that the particles in simulations that considered any form of
drag got entrained in a vorticy while particles in the passive
simulation deflected away (Figure 3). Comparison of cross-sectional
areas and different drag exposure scenarios show no differences in mean
trajectory and spatial patterns in density, with only slight variations
across types.

\begin{figure}
\centering
\includegraphics{thesis_chapter_3_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{Figure 3: Comparison of density of particles within each grid
cell for the end run time of each scenario. Plots A-D repressent the
minimum, plots E-H are the mean and plots I-L the maximum
cross-sectional areas. The bottom plot is the density plot for the
passive simulation.}
\end{figure}

\hypertarget{comparison-of-distances}{%
\subsection{Comparison of distances}\label{comparison-of-distances}}

There were significant differences among the median distances travelled
for all particles in each simulation (Figure 4; p \textless{} 0.05). The
simulations that considered for 100\% hydrodynamic drag travelled less
distance than particles for simulations which considered a combination
of hydrodynamic and wind drag scenarios. Comparison between simulations
that considered hydrodynamic and wind drag show no significant
differences with only slight variations among simulations.

\begin{figure}
\centering
\includegraphics{thesis_chapter_3_files/figure-latex/distance plot-1.pdf}
\caption{Figure 4: Boxplots of total distance of all particles from the
release site over the course of the simulation for each exposure
scenario.}
\end{figure}

\hypertarget{comparison-of-coefficents}{%
\subsection{Comparison of coefficents}\label{comparison-of-coefficents}}

The comparison of mean trajectories between the sphere and cylindrical
shape coefficients revealed no dissimilarity (Figure 5A). This was also
reflected when comparing distance travelled by particles and particle
density plots (Figure 5B). The simulation which used a cylindrical shape
coefficient showed higher density in an eddy compared to the simulation
which used a sphere for the shape coefficient (Figure 5C-D).

\begin{figure}
\centering
\includegraphics{thesis_chapter_3_files/figure-latex/shape data outputs-1.pdf}
\caption{Figure 5: Comparison of trajectory data between a kelp sphere
and kelp cylinder. Plot A is the comparison of mean trajectories, plot B
is the comparison of distance travelled by all particles, and plot C and
D are the density plots for simulations using a sphere and cylinderical
shape coefficient.}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The treatment of drifting macroalgae as purely lagrangian will not be
able to accurately determine patterns of passive dispersal, such as
entrainment in eddies or vortices (Fossette et al., 2012; Miron et al.,
2020). Although past research has shown the inclusion of inertia and
windage greatly increase observed macroalgae trajectory and entrainment
characteristics (Brooks et al., 2018, 2019; Putman et al., 2018, 2020),
current research has focused on large macroalgae-rafts which can vary
greatly in size and surface area. None of the past research has taken
advantage of the advancement in numerical ocean models and lagrangian
trajectory modeling to investigate how aspects of drag (hydrodynamic and
wind) affect the trajectory patterns of solitary drifting macroalgae.
This study compared particles trajectories with simulations that
included no forms of drag with simulations that included various forms
of both hydrodynamic and wind drag. The mean virtual kelp trajectory
deviated from the mean passive trajectory when including varying levels
of hydrodynamic and wind drag scenarios. The results from this study
show the inclusion of hydrodynamic and wind drag improves entrainment
patterns, and also our understanding of how surface area plays a role in
solitary drifting macroalgae trajectory, in this case \emph{E.maxima}.

Past research has inferred from direct and indirect techniques that
drifting macroalgae tend to follow the prevailing surface currents
(Hobday, 2000; Thiel and Gutow, 2005; Fraser et al., 2011; Rothäusler et
al., 2011, 2011). The results from this numerical experiment confirm
that drifting macroalgae do follow the surface current in the study
region when considering only hydrodynamic drag. In addition, not only
are the trajectories similar but so are the end points at the end of
simulation. However, when ad-hoc wind drag (windage) is considered in
combination with hydrodynamic drag the trajectories differ greatly when
compared to purely lagrangian particle trajectories. The inclusion of
wind drag causes the particles to flow further away from the coast and
with a greater distance travelled from the release location. The
difference in trajectory and distance covered can be attributed to the
inclusion of wind drag. Wind drag adds to the effects of hydrodynamic
drag when the wind direction and current direction are opposite to each
other. However, when the wind direction is in the same or similar
direction to that of the current, momentum energy from the wind is
transferred to the object (Hackett et al., 2006; Putman et al., 2020).
This causes the particles to travel further as well as exposed to
different time varying flows. Past studies have shown this when
including windage as a function of inertia (Putman et al., 2018, 2020;
Brooks et al., 2019). A study by Putman et al. (2020) used ad-hoc
windage factors that was based on the surface current velocity to assess
the approach for improving transport predictions of pelagic Sargassum.
The results showed that including ad-hoc windage factors improved
virtual trajectories compared to tracked Sargassum mats which was partly
due to the inclusion of momemtum energy transfer from wind. This is also
reflected in the analysis comparing distances travelled by all the
particles in each simulation. The particles in the simulations that
included wind drag traveled significantly greater distances compared to
the passive or hydrodynamic only drag.

The addition of drag forces into the simulation also causes the
particles to cluster together along the mean trajectory compared to that
of the passive particles which are more evenly dispersed. This is most
likely due to the shared drag characteristics among the particles which
included the plants weight and surface area into the calculation. Other
studies investigating flotsam trajectory characteristics have found
similar results (Breivik et al., 2011; Miron et al., 2020; Olascoaga et
al., 2020). A study by Miron et al. (2020) conducted field experiments
using a range of objects such as spheres, cubes, cuboids to compare to
investigate the effect of inertia on particles dynamics. The results
from the study showed that objects tended to cluster according to the
shape of the object which the authors attributed to shared
characteristics of inertia. Although drag forces were used in this study
and not inertia, the same conclusion applies, drag is an important
characteristic to consider when investigating or predicting solitary
drifting macroalgae.

The ocean is made up of varying time-varying flows and changes in an
objects velocity can cause the object to be exposed to different flow
patterns over time. The lack of dissimilarity between different
cross-sectional area types (minimum, mean and maximum) suggests that
reduction in velocity in relation to cross-sectional area is negligible.
A similar result was found by Le Gouvello et al. (2020) who investigated
the effects of swimming behavior on sea turtle hatchling dispersal in
the Agulhas region. The results suggested that the hatchlings
trajectories are mostly influenced by the ocean currents in the first
year of hatching due to low swimming speeds of juveniles. Therefore, the
differences in cross-sectional area types may not be significant enough
in order to have a significant effect on the relevant velocity vectors
(zonal and meridonal). Another possible reason is the resolution of the
ocean model used. The resolution of the underlying ocean model is an
important aspect of lagrangian ocean modeling and must be able to
resolve sub-grid scale ocean processes for finer scale applications. A
study by Hart-Davis et al. (2018) assessed the inclusion of stochastic
motion, wind and currents into forecasting for search and rescue using
the same model ocean model used in this study. The authors found that
the inclusion of brownian motion greatly increased the accuracy in
representing sub-grid scale processes and that the inclusion of wind,
currents and stochastic motion greatly improved forecasting
applications. Therefore, the authors argue that the lack of
dissimilarity between simulations of different cross-sectional area and
drag exposure scenarios is not as a result of the resolution of the
model.

The same is also true when comparing different hydrodynamic and wind
drag exposure scenarios. These scenarios reflect different buoyancy
situations which result in different amounts of surface area exposed to
hydrodynamic and/or wind drag. The similarities between the different
combined drag scenarios suggests that higher magnitudes of wind exposure
do not significantly alter trajectories. Instead, the results from this
study suggest that the inclusion of drag forces in simulating macroalgal
trajectory may result in improving accuracy of entrainment patterns. All
the simulations that included any form of drag resulted in a fairly high
density of particles entrained in a vortex. The entrainment and
expulsion from eddies and vortices is an important characteristic of
floating macroalgae. Accuracy in predicting entrainment and expulsion of
macroalgae-rafts from eddies has been investigated previously by Putman
et al. (2018) and Putman et al. (2020), who show that including windage
greatly improves entrainment patterns when comparing virtual and real
Sargassum rafts. However, the aforementioned studies investigated large
rafts and not floating individuals. Inertia is based on the size/weight
of the raft, while hydrodynamic and wind drag are based on current
velocity and surface area. Therefore, the role of inertia may be
negligible for a solitary individual and rather other forms of drag
should be considered. The results from this study suggest that the
inclusion of hydrodynamic and/or wind drag is an important component
when simulating virtual kelp particles will greatly increase the
accuracy of entrainment and expulsion from sub-grid mesoscale
oceanographic features, such as eddies and vortices. The use of a
cylindrical shape coefficient instead of a sphere in the calculation of
drag force also had no effect on the mean trajectory. Instead, the
inclusion of a different shape coefficient increased to the density of
particles entrained in an eddy field. This suggest that inclusion of
different shape coefficients play a role in further improving estimates
and hindcasts of macroalgal dispersal patterns.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Most of the past research has focused on large macroalgae rafts which
can vary greatly in size and shape, however no research exists
attempting to clarify the drifting characteristics of solitary
macroalgae. Understanding how to accurately model the distribution of
solitary kelp can lead to be understanding of accumulation zones and
sinks and ultimately the ecological links pertaining to those processes.
The findings from this study indicate that solitary floating virtual
\emph{E. maxima} particles tend to follow the prevailing surface
currents and the inclusion of wind causes the trajectories to differ
greatly from that of a purely lagrangian particle. Furthermore, the
inclusion of both hydrodynamic and wind drag cause clustering of
particles along the trajectory as well causing particles to become
entrained in eddies. Different cross-sectional areas exposed to
hydrodynamic and wind drag had no effect on overall trajectory
suggesting that those differences are negligible when investigating
dispersal patterns of drifting \emph{E.maxima}. In addition, the use of
a cylindrical shape coefficient has no effect on mean trajectory but
also causes higher density of particles to become entrained in an eddy.
Overall the inclusion of drag forces is an important aspect to consider
when investigating the dispersal patterns of solitary macroalgae, while
differences in cross-sectional areas are not. This study also provides
an approach which can be adapted to model any floating solitary
macroalgae, provided the surface area can be estimated accurately. Gaps
in the research exist when considering what oceanographic processes play
a role in solitary macroalgal dispersal as well as how these vary
seasonally. The identification of biological and physical factors that
play a role in accumulation zones is also needed, as well as better
estimates for sinking rates and raft-times.

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\hypertarget{domain-features}{%
\subsection{Domain features}\label{domain-features}}

\includegraphics{thesis_chapter_3_files/figure-latex/unnamed-chunk-3-1.pdf}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-allen1999}{}%
Allen, A., and Plourde, J. (1999). Review of leeway: Field experiments
and implementation. US coast guard rep.

\leavevmode\hypertarget{ref-andersson2017}{}%
Andersson, L. E., Scibilia, F., and Imsland, L. (2017). A study on an
iceberg drift trajectory. 8.
doi:\href{https://doi.org/10.1115/OMAE2017-62159}{10.1115/OMAE2017-62159}.

\leavevmode\hypertarget{ref-batista2018}{}%
Batista, M. B., Anderson, A. B., Sanches, P. F., Polito, P. S.,
Silveira, T. C. L., Velez-Rubio, G. M., et al. (2018). Kelps'
long-distance dispersal: role of ecological/oceanographic processes and
implications to marine forest conservation. \emph{Diversity} 10.
doi:\href{https://doi.org/10.3390/d10010011}{10.3390/d10010011}.

\leavevmode\hypertarget{ref-back2000}{}%
Bäck, S., Lehvo, A., and Blomster, J. (2000). Mass occurrence of
unattached \emph{enteromorpha intestinalis} on the finnish baltic sea
coast. 155--161.

\leavevmode\hypertarget{ref-beal2015}{}%
Beal, L. M., Elipot, S., Houk, A., and Leber, G. M. (2015). Capturing
the transport variability of a western boundary jet: Results from the
agulhas current time-series experiment (act). \emph{Journal of Physical
Oceanography} 45, 1302--1324.

\leavevmode\hypertarget{ref-blanke2002}{}%
Blanke, B., Roy, C., Penven, P., Speich, S., McWilliams, J., and Nelson,
G. (2002). Linking wind and interannual upwelling variability in a
regional model of the southern benguela. \emph{Geophysical Research
Letters} 29, 41--1.

\leavevmode\hypertarget{ref-blanke2005}{}%
Blanke, B., Speich, S., Bentamy, A., Roy, C., and Sow, B. (2005).
Modeling the structure and variability of the southern benguela
upwelling using quikscat wind forcing. \emph{Journal of Geophysical
Research: Oceans} 110.

\leavevmode\hypertarget{ref-breivik2011}{}%
Breivik, Ø., Allen, A. A., Maisondieu, C., and Roth, J. C. (2011).
Wind-induced drift of objects at sea: The leeway field method.
\emph{Applied Ocean Research} 33, 100--109.

\leavevmode\hypertarget{ref-brooks2019}{}%
Brooks, M. T., Coles, V. J., and Coles, W. C. (2019). Inertia influences
pelagic sargassum advection and distribution. \emph{Geophysical Research
Letters} 46, 2610--2618.

\leavevmode\hypertarget{ref-brooks2018}{}%
Brooks, M. T., Coles, V. J., Hood, R. R., and Gower, J. F. (2018).
Factors controlling the seasonal distribution of pelagic sargassum.
\emph{Marine Ecology Progress Series} 599, 1--18.

\leavevmode\hypertarget{ref-bushing1994}{}%
Bushing, W. W. (1994). Biogeographic and ecological implications of kelp
rafting as a dispersal vector for marine invertebrates. 22.

\leavevmode\hypertarget{ref-collins2010}{}%
Collins, C. J., Fraser, C. I., Ashcroft, A., and Waters, J. M. (2010).
Asymmetric dispersal of southern bull-kelp (\emph{Durvillaea
antarctica}) adults in coastal New Zealand: testing an oceanographic
hypothesis. \emph{Mol. Ecol.} 19, 4572--4580.

\leavevmode\hypertarget{ref-coppin2020}{}%
Coppin, R., Rautenbach, C., Ponton, T. J., and Smit, A. (2020).
Investigating waves and temperature as drivers of kelp morphology.
\emph{Frontiers in Marine Science} 7, 567.

\leavevmode\hypertarget{ref-dayton1985}{}%
Dayton, P. K. (1985). Ecology of Kelp Communities. \emph{Annual Review
of Ecology and Systematics} 16, 215--245.
doi:\href{https://doi.org/10.1146/annurev.es.16.110185.001243}{10.1146/annurev.es.16.110185.001243}.

\leavevmode\hypertarget{ref-delandmeter2019}{}%
Delandmeter, P., and Van Sebille, E. (2019). The parcels v2. 0
lagrangian framework: New field interpolation schemes.
\emph{Geoscientific Model Development} 12, 3571--3584.

\leavevmode\hypertarget{ref-dromgoole1982}{}%
Dromgoole, F. (1982). The buoyant properties of codium. \emph{Botanica
Marina} 25, 391--398.

\leavevmode\hypertarget{ref-edgar1987}{}%
Edgar, G. (1987). Dispersal of faunal and floral propagules associated
with drifting \emph{macrocystis pyrifera} plants. \emph{Marine Biology}
95, 599--610.

\leavevmode\hypertarget{ref-eik2009}{}%
Eik, K. (2009). Iceberg drift modelling and validation of applied
metocean hindcast data. \emph{Cold Regions Science and Technology} 57,
67--90.

\leavevmode\hypertarget{ref-fossette2012}{}%
Fossette, S., Putman, N. F., Lohmann, K. J., Marsh, R., and Hays, G. C.
(2012). A biologist's guide to assessing ocean currents: A review.
\emph{Marine Ecology Progress Series} 457, 285--301.

\leavevmode\hypertarget{ref-fraser2011}{}%
Fraser, C. I., Nikula, R., and Waters, J. M. (2011). Oceanic rafting by
a coastal community. \emph{Proceedings of the Royal Society B:
Biological Sciences} 278, 649--655.

\leavevmode\hypertarget{ref-garzoli1996}{}%
Garzoli, S. L., Gordon, A. L., Kamenkovich, V., Pillsbury, D., and
Duncombe-Rae, C. (1996). Variability and sources of the southeastern
atlantic circulation. \emph{Journal of Marine Research} 54, 1039--1071.

\leavevmode\hypertarget{ref-graiff2016}{}%
Graiff, A., Pantoja, J. F., Tala, F., and Thiel, M. (2016). Epibiont
load causes sinking of viable kelp rafts: Seasonal variation in floating
persistence of giant kelp \emph{macrocystis pyrifera}. \emph{Marine
biology} 163, 191.

\leavevmode\hypertarget{ref-griffin2017}{}%
Griffin, D., Oke, P., and Jones, E. (2017). \emph{The search for mh370
and ocean surface drift}. Commonwealth Scientific; Industrial Research
Organisation.

\leavevmode\hypertarget{ref-hackett2006}{}%
Hackett, B., Breivik, Ø., and Wettre, C. (2006). Forecasting the drift
of objects and substances in the ocean. 507--523.

\leavevmode\hypertarget{ref-hardman2003}{}%
Hardman-Mountford, N., Richardson, A., Agenbag, J., Hagen, E., Nykjaer,
L., Shillington, F., et al. (2003). Ocean climate of the south east
atlantic observed from satellite data and wind models. \emph{Progress in
Oceanography} 59, 181--221.

\leavevmode\hypertarget{ref-harrold1989}{}%
Harrold, C., and Lisin, S. (1989). Radio-tracking rafts of giant kelp:
Local production and regional transport. \emph{Journal of Experimental
Marine Biology and Ecology} 130, 237--251.

\leavevmode\hypertarget{ref-hart2018}{}%
Hart-Davis, M. G., Backeberg, B. C., and Bakhoday-Paskyabi, M. (2018).
An assessment of the importance of combining wind, ocean currents and
stochastic motions in a particle trajectory model for search and rescue
applications.

\leavevmode\hypertarget{ref-helmuth1994}{}%
Helmuth, B., Veit, R. R., and Holberton, R. (1994). Long-distance
dispersal of a subantarctic brooding bivalve \emph{(Gaimardia
trapesina)} by kelp-rafting. \emph{Marine biology} 120, 421--426.
doi:\href{https://doi.org/10.1007/BF00680216}{10.1007/BF00680216}.

\leavevmode\hypertarget{ref-hobday2000}{}%
Hobday, A. J. (2000). Age of drifting \emph{Macrocystis pyrifera} (L.)
C. Agardh rafts in the Southern California Bight. \emph{Journal of
experimental marine biology and ecology} 253, 97--114.
doi:\href{https://doi.org/10.1016/S0022-0981(00)00255-0}{10.1016/S0022-0981(00)00255-0}.

\leavevmode\hypertarget{ref-holmquist1994}{}%
Holmquist, J. (1994). Benthic macroalgae as a dispersal mechanism for
fauna: Influence of a marine tumbleweed. \emph{Journal of Experimental
Marine Biology and Ecology} 180, 235--251.

\leavevmode\hypertarget{ref-kingsford1995}{}%
Kingsford, M. J. (1995). Drift algae: A contribution to near-shore
habitat complexity in the pelagic environment and an attractant for
fish. \emph{Marine ecology progress series. Oldendorf} 116, 297--301.

\leavevmode\hypertarget{ref-gouvello2020}{}%
Le Gouvello, D. Z., Hart-Davis, M. G., Backeberg, B. C., and Nel, R.
(2020). Effects of swimming behaviour and oceanography on sea turtle
hatchling dispersal at the intersection of two ocean current systems.
\emph{Ecological Modelling} 431, 109130.

\leavevmode\hypertarget{ref-lichey2001}{}%
Lichey, C., and Hellmer, H. H. (2001). Modeling giant-iceberg drift
under the influence of sea ice in the weddell sea, antarctica.
\emph{Journal of Glaciology} 47, 452--460.

\leavevmode\hypertarget{ref-lutjeharms2007}{}%
Lutjeharms, J. (2007). Three decades of research on the greater agulhas
current.

\leavevmode\hypertarget{ref-lutjeharms2006}{}%
Lutjeharms, J. R. (2006). The agulhas current. 5.

\leavevmode\hypertarget{ref-lutjeharms1988}{}%
Lutjeharms, J., and Van Ballegooyen, R. (1988). The retroflection of the
agulhas current. \emph{Journal of Physical Oceanography} 18, 1570--1583.

\leavevmode\hypertarget{ref-macaya2005}{}%
Macaya, E. C., Boltana, S., Hinojosa, I. A., Macchiavello, J. E.,
Valdivia, N. A., Vasquez, N. R., et al. (2005). PRESENCE of sporophylls
in floating kelp rafts of macrocystis spp.(PHAEOPHYCEAE) along the
chilean pacific coast 1. \emph{Journal of Phycology} 41, 913--922.

\leavevmode\hypertarget{ref-McCormick2008}{}%
McCormick, T. B., Buckley, L. M., Brogan, J., and Perry, L. M. (2008).
Drift macroalgae as a potential dispersal mechanism for the white
abalone haliotis sorenseni. \emph{Marine Ecology Progress Series} 362,
225--232.

\leavevmode\hypertarget{ref-miron2020}{}%
Miron, P., Olascoaga, M., Beron-Vera, F., Putman, N., Triñanes, J.,
Lumpkin, R., et al. (2020). Clustering of marine-debris-and
sargassum-like drifters explained by inertial particle dynamics.
\emph{Geophysical Research Letters} 47, e2020GL089874.

\leavevmode\hypertarget{ref-nikula2013}{}%
Nikula, Spencer, and Waters (2013). Passive rafting is a powerful driver
of transoceanic gene flow. \emph{Biology letters} 9, 20120821.
doi:\href{https://doi.org/10.1098/rsbl.2012.0821}{10.1098/rsbl.2012.0821}.

\leavevmode\hypertarget{ref-norton1992}{}%
Norton, T. (1992). Dispersal by macroalgae. \emph{British Phycological
Journal} 27, 293--301.

\leavevmode\hypertarget{ref-olascoaga2020}{}%
Olascoaga, M. J., Beron-Vera, F. J., Miron, P., Triñanes, J., Putman,
N., Lumpkin, R., et al. (2020). Observation and quantification of
inertial effects on the drift of floating objects at the ocean surface.
\emph{Physics of Fluids} 32, 026601.

\leavevmode\hypertarget{ref-putman2018}{}%
Putman, N. F., Goni, G. J., Gramer, L. J., Hu, C., Johns, E. M.,
Trinanes, J., et al. (2018). Simulating transport pathways of pelagic
sargassum from the equatorial atlantic into the caribbean sea.
\emph{Progress in Oceanography} 165, 205--214.

\leavevmode\hypertarget{ref-putman2020}{}%
Putman, N. F., Lumpkin, R., Olascoaga, M. J., Trinanes, J., and Goni, G.
J. (2020). Improving transport predictions of pelagic sargassum.
\emph{Journal of Experimental Marine Biology and Ecology} 529, 151398.

\leavevmode\hypertarget{ref-rothausler2011}{}%
Rothäusler, E., Gómez, I., Hinojosa, I. A., Karsten, U., Miranda, L.,
Tala, F., et al. (2011). Kelp rafts in the humboldt current: Interplay
of abiotic and biotic factors limit their floating persistence and
dispersal potential. \emph{Limnology and oceanography} 56, 1751--1763.

\leavevmode\hypertarget{ref-rubio2009}{}%
Rubio, A., Blanke, B., Speich, S., Grima, N., and Roy, C. (2009).
Mesoscale eddy activity in the southern benguela upwelling system from
satellite altimetry and model data. \emph{Progress in Oceanography} 83,
288--295.

\leavevmode\hypertarget{ref-saunders2014}{}%
Saunders, G. W. (2014). Long distance kelp rafting impacts seaweed
biogeography in the Northeast Pacific: The kelp conveyor hypothesis.
\emph{Journal of phycology} 50, 968--974.
doi:\href{https://doi.org/10.1111/jpy.12237}{10.1111/jpy.12237}.

\leavevmode\hypertarget{ref-shannon1996}{}%
Shannon, L., and Nelson, G. (1996). The benguela: Large scale features
and processes and system variability. 163--210.

\leavevmode\hypertarget{ref-smith2002}{}%
Smith, S. D. A. (2002). Kelp rafts in the Southern Ocean. \emph{Global
ecology and biogeography: a journal of macroecology} 11, 67--69.
doi:\href{https://doi.org/10.1046/j.1466-822X.2001.00259.x}{10.1046/j.1466-822X.2001.00259.x}.

\leavevmode\hypertarget{ref-tala2013}{}%
Tala, F., Gómez, I., Luna-Jorquera, G., and Thiel, M. (2013).
Morphological, physiological and reproductive conditions of rafting bull
kelp \emph{(Durvillaea antarctica)} in northern-central Chile (30°S).
\emph{Marine biology} 160, 1339--1351.
doi:\href{https://doi.org/10.1007/s00227-013-2186-8}{10.1007/s00227-013-2186-8}.

\leavevmode\hypertarget{ref-tala2017}{}%
Tala, F., Penna-Díaz, M. A., Luna-Jorquera, G., Rothäusler, E., and
Thiel, M. (2017). Daily and seasonal changes of photobiological
responses in floating bull kelp \emph{Durvillaea antarctica} (Chamisso)
Hariot (Fucales: Phaeophyceae). \emph{Phycologia} 56, 271--283.

\leavevmode\hypertarget{ref-thiel2005}{}%
Thiel, M., and Gutow, L. (2005). The Ecology of rafting in the marine
environment. II. The rafting organisms and community. \emph{Oceanography
and Marine Biology: An Annual Review} 43, 279--418.

\leavevmode\hypertarget{ref-veitch2010}{}%
Veitch, J., Penven, P., and Shillington, F. (2010). Modeling equilibrium
dynamics of the benguela current system. \emph{Journal of Physical
Oceanography} 40, 1942--1964.

\leavevmode\hypertarget{ref-wichmann2012}{}%
Wichmann, C.-S., Hinojosa, I. A., and Thiel, M. (2012). Floating kelps
in Patagonian Fjords: an important vehicle for rafting invertebrates and
its relevance for biogeography. \emph{Mar. Biol.} 159, 2035--2049.
doi:\href{https://doi.org/10.1007/s00227-012-1990-x}{10.1007/s00227-012-1990-x}.

\end{document}

### 2021-03-12 thesis_chapter_2.Rmd
updating chapter2 rmd to match pub

@@
 
-ggarrange(img_wave_rose_A, img_wave_rose_B, img_wave_rose_C, ncol = 2, nrow = 2, labels = "AUTO")
-```
-
-```{r wave power map, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:wave power map}The total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length.
-Total (A), summer (B), winter (C), spring (D) and autumn (E).'}
-
-library(jpeg)
-library(grid)
-library(gridExtra)
-library(ggpubr)
-
-wp_1 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6A.jpg"))) #figure missing, copy of fig_6B used
-wp_2 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6B.jpg")))
-wp_3 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6C.jpg")))
-wp_4 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6D.jpg")))
-wp_5 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6E.jpg")))
-wp_6 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6_legend.jpg")))
-
-ggarrange(wp_1, wp_2, wp_3, wp_4, wp_5, wp_6, ncol = 3, nrow = 2, labels = "AUTO", align = "hv")
-
-#wave_power_map <- readPNG("figures/wave_power_all_labels.png")
-#grid.raster(wave_power_map)
-```
+ggarrange(img_wave_rose_A, img_wave_rose_B, img_wave_rose_C, ncol = 2, nrow = 2, labels = c("A", "C", "B"))
+```
+
 To clarify the averaged wave exposure map presented in Figure \ref{fig:wave power map}A, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in Figure \ref{fig:wave power map}C.
@@
 
-```{r FB spectrum, echo=FALSE, eval= TRUE, fig.cap='\\label{fig:FB spectrum}The propagation of a typical offshore wave spectrum is given as produced from a single time-step in SWAN.'}
-
-library(grid)
-library(jpeg)
-
-img_fb <- readJPEG("figures/chapter_2/fig_7.jpg")
-grid.raster(img_fb)
-
-```
 ## Drivers of kelp morphological characteristics
@@
 Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=8, fig.width=8}
-ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "none", align ='hv')
+ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, common.legend = TRUE, legend = "none", align ='hv')
 ```
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.'}
-ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom", align = "hv")
+ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, common.legend = TRUE, legend = "bottom", align = "hv")
 ```
@@
 Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
-ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = c("none"), ncol = 4, nrow = 3, vjust = 1.5)
+ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, common.legend = TRUE, legend = c("none"), ncol = 4, nrow = 3, vjust = 1.5)
 ```
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.', fig.height=8, fig.width=8}
-ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 4, nrow = 3, vjust = 1.5)
+ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, common.legend = TRUE, legend = "bottom", ncol = 4, nrow = 3, vjust = 1.5)
 ```
@@
 Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=7, fig.width=8}
-ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom", ncol = 4, nrow = 3, vjust = 1.5)
+ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, common.legend = FALSE, legend = "bottom", ncol = 4, nrow = 3, vjust = 1.5)
 ```
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.', fig.height=7, fig.width=8}
-ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9,ESW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 4,nrow = 3)
+ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9,ESW10, common.legend = TRUE, legend = "bottom", ncol = 4,nrow = 3)
 ```
@@
 #ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, legend = "none", labels = "AUTO", ncol = 2, nrow = 2, align = "hv")
-
-rda_1 <- rasterGrob(as.raster(readJPEG("thesis_chapter_2_files/figures_frontiers/fig_12A.jpeg"))) #figure missing, copy of fig_6B used
-rda_2 <- rasterGrob(as.raster(readJPEG("thesis_chapter_2_files/figures_frontiers/fig_12B.jpeg")))
-rda_3 <- rasterGrob(as.raster(readJPEG("thesis_chapter_2_files/figures_frontiers/fig_12C.jpeg")))
+library(jpeg)
+
+rda_1 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_14A.jpeg"))) #figure missing, copy of fig_6B used
+rda_2 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_14B.jpeg")))
+rda_3 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_14C.jpeg")))
 
@@
 
-```{r appendix figure 1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=8}
-positionsLam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point ", "Baboon Rock", 
-                  "Batsata Rock", "Buffelsbaai","Buffelsbaai South","Bordjies Deep","Olifantsbos", "Kommetjie", "Sea Point")
-lam_all_A <- ggplot(data = lam_data, aes(x = site, y = laminaWeight, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsLam) + 
-           ylab("Lamina Weight (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("")
-lam_all_B <- ggplot(data = lam_data, aes(x = site, y = laminaLength, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsLam) + 
-           ylab("Lamina Length (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("")
-lam_all_C <- ggplot(data = lam_data, aes(x = site, y = laminaThickness, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsLam) + 
-           ylab("Lamina Thickness (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-lam_all_D <- ggplot(data = lam_data, aes(x = site, y = stipeMass, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsLam) + 
-           ylab("Stipe Mass (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-lam_all_E <- ggplot(data = lam_data, aes(x = site, y = stipeLength, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsLam) + 
-           ylab("Stipe Length (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-lam_all_F <- ggplot(data = lam_data, aes(x = site, y = stipeDiameter, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsLam) + 
-           ylab("Stipe Diameter (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("")
-lam_all_G <- ggplot(data = lam_data, aes(x = site, y = digits, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsLam) + 
-           ylab("No. of Digits") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-lam_all_H <- ggplot(data = lam_data, aes(x = site, y = thallusMass, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsLam) + 
-           ylab("Thallus Mass (g)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-lam_all_I <- ggplot(data = lam_data, aes(x = site, y = totalLength, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsLam) + 
-           ylab("Total length (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("")
-
-ggarrange(lam_all_A,lam_all_B,lam_all_C,lam_all_D,lam_all_E,lam_all_F,lam_all_G,lam_all_H,lam_all_I, ncol = 3, nrow = 3, common.legend = TRUE)
-
-```
-
-```{r appendix figure 2, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=8}
-# Order site names by location west to east
-
-positionsEck <- c("Betty's Bay", "Kalk Bay", "A-Frame", 
-                  "Miller's Point", "Baboon Rock", "Batsata Rock", 
-                  "Buffels Bay", "Scarborough", 
-                  "Soetwater","Kommetjie","Hout Bay", 
-                  "Oudekraal", "St. Helena Bay")
-
-eck_all_A <- ggplot(data = eck_data, aes(x = site, y = frond_mass, fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Frond mass (g)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-eck_all_B <- ggplot(data = eck_data, aes(x = site, y = primary_length , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Primary Length  (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-eck_all_C <- ggplot(data = eck_data, aes(x = site, y = primary_width , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Primary Width (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-eck_all_D <- ggplot(data = eck_data, aes(x = site, y = frond_length , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Frond length (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-eck_all_E <- ggplot(data = eck_data, aes(x = site, y = stipe_mass , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Stipe mass (g)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-eck_all_F <- ggplot(data = eck_data, aes(x = site, y = stipe_length , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Stipe length (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-eck_all_G <- ggplot(data = eck_data, aes(x = site, y = stipe_circ , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Stipe circ (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-eck_all_H <- ggplot(data = eck_data, aes(x = site, y = tufts , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("No. of tufts") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("") 
-eck_all_I <- ggplot(data = eck_data, aes(x = site, y = epi_length , fill = wave_exposure)) +
-           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
-           stat_compare_means(method = "kruskal.test", label.x = 5) + 
-           scale_x_discrete(limits = positionsEck) + 
-           ylab("Epi Length  (cm)") + 
-           scale_fill_brewer(palette="OrRd") +
-           coord_flip() +
-           guides(fill=guide_legend(title="Wave Exposure")) +
-           xlab("")
-
-ggarrange(eck_all_A,eck_all_B,eck_all_C,eck_all_D,eck_all_E,eck_all_F,eck_all_G,eck_all_H,eck_all_I, ncol = 3, nrow = 3, common.legend = TRUE, legend = FALSE)
+```{r appendix figure, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=8}
+
+app_A <- rasterGrob(as.raster(readJPEG("figures/chapter_2/appendix_A.jpg"))) #figure missing, copy of fig_6B used
+app_B <- rasterGrob(as.raster(readJPEG("figures/chapter_2/appendix_B.jpg")))
+app_C <- rasterGrob(as.raster(readJPEG("figures/chapter_2/appendix_C.jpg")))
+
+app_plt_A <- ggarrange(app_B, app_C, align = "hv", ncol = 2, nrow = 1, labels = c("B", "C"))
+ggarrange(app_A, app_plt_A, align = "hv", ncol = 1, nrow = 2, labels = c("A", "", ""))
+
 ```

### 2021-03-12 thesis_chapter_2.Rmd
chapter 2 figure caption updated
figure captions updated

@@
 ```{r map, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="\\label{fig:map}Map of the study area and sampling sites.
-The red dot indicates the position of Cape Town for sake of reference."}
+The red dots represent study sites around the Cape Peninsula."}
 map_combined
@@
 
-```{r Temperature variables plot, eval=TRUE, echo=FALSE, fig.cap='\\label{fig:Temperature variables plot}Annual temperature variables at the collections sites around the Cape Peninsula are represented by plot A, with sites in order from North to South.
-Temperature variables include minimum temperature represented by inverted triangles, mean represented by black dots, maximum by black triangles, and whiskers standard deviation.
-Summary of temperature data for sites grouped by region and represented in plot B.', fig.height=5, fig.width=8, warning=FALSE}
+```{r Temperature variables plot, eval=TRUE, echo=FALSE, fig.cap='\\label{fig:Temperature variables plot}Annual temperature variables at the collection sites around the Cape Peninsula are represented in panel (A), with sites in order from north to south.
+Temperature variables include minimum represented by inverted triangles, mean represented by blue and red dots which represent the western and eastern side of the peninsula, respectively, maximum by black triangles, and whiskers standard deviation.
+Summary of temperature data for sites grouped by region are represented in panel (B).
+Significance levels are as follows: * represents ≤0.05, ** represents ≤0.01, *** represents ≤0.001, **** represents ≤0.0001.', fig.height=5, fig.width=8, warning=FALSE}
 
@@
 
-```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE, eval=TRUE,fig.height=7,fig.width=8, fig.cap='\\label{fig:wave variable plot}Mean significant wave height and mean significant peak period across sites and by region.
-Mean significant wave height across sites is presented in plot A and is represented by solid triangles and standard deviation by the whiskers.
-Differences in mean significant wave height between regions are presented in plot B.
-Mean significant peak period across sites is presented in plot C and is represented by solid dots and standard deviation by whiskers.
-Differences in mean significant peak period between regions are presented in plot D.'}
+```{r wave variable plot, warning = FALSE, echo=FALSE, message = FALSE, include= TRUE, eval=TRUE,fig.height=7,fig.width=8, fig.cap='\\label{fig:wave variable plot}Mean significant wave height (meters) and mean significant peak period (seconds) across sites and by region.
+Annual mean significant wave height across sites is presented in panel (A) and is represented by solid triangles and standard deviation by the whiskers.
+Differences in annual mean significant wave height between regions are presented in panel (B).
+Annual mean significant peak period across sites is presented in panel (C) and is represented by solid dots and standard deviation by whiskers.
+Differences in annual mean significant peak period between regions are presented in panel (D).
+Significance levels are as follows: * represents ≤0.05, ** represents ≤0.01, *** represents ≤0.001, **** represents ≤0.0001.'}
 ggarrange(wave_plot_1,wave_boxplot_1, wave_plot_2, wave_boxplot_2, common.legend = FALSE, labels = c("A", "B", "C", "D"), align = "hv")
@@
 
-```{r wave rose, echo=FALSE, warning = FALSE, eval= TRUE, fig.align="center",fig.cap='\\label{fig:wave rose}The total and seasonal wave roses of the directional wave buoy just offshore from Cape Town is given form the years 2000 till end 2017 (the period for which directional wave spectra was available).
-Total (A), Austral winter (B) and Austral summer (C) respectively'}
+```{r wave rose, echo=FALSE, warning = FALSE, eval= TRUE, fig.align="center",fig.cap='\\label{fig:wave rose}The total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length) is presented in panel (A) and the associated legend in panel (B).
+The propagation of a typical offshore wave spectrum is given as produced from a single time-step in SWAN, is represented in panel (C)'}
 
@@
 ```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.cap='\\label{fig:lamanaria temp plots}Boxplots representing *L.
-pallida* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
-The lower and upper "hinges" correspond to the first and third quartiles.
+pallida* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
+The lower and upper “hinges” correspond to the first and third quartiles.
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=8, fig.width=8}
+Light gray boxes represent cool-temperate and dark gray warm-temperate.
+Significance levels are as follows: * represents ≤0.05, ** represents ≤0.01, *** represents ≤0.001, **** represents ≤0.0001.', fig.height=8, fig.width=8}
 ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, common.legend = TRUE, legend = "none", align ='hv')
@@
 ```{r lamanaria wave plots, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.height=8, fig.width=8, fig.cap='\\label{fig:lamanaria wave plots}Boxplots representing *L.
-pallida* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
+pallida* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
 The lower and upper hinges correspond to the first and third quartiles.
-The whiskers represent the range, solid black lines represent the median and black dots are outliers.'}
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.
+Significance levels are as follows: * represents ≤0.05, ** represents ≤0.01, *** represents ≤0.001, **** represents ≤0.0001.'}
 ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, common.legend = TRUE, legend = "bottom", align = "hv")
@@
 ```{r deep ecklonia temp plot, echo=FALSE, fig.cap='\\label{fig:deep eck temp plot}Boxplots representing deep population *E.
-maxima* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
+maxima* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
 The lower and upper hinges correspond to the first and third quartiles.
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
+Light gray boxes represent cool-temperate and dark gray warm-temperate.
+Significance levels are as follows: * represents ≤0.05, ** represents ≤0.01, *** represents ≤0.001, **** represents ≤0.0001.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE, eval=TRUE}
 ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, common.legend = TRUE, legend = c("none"), ncol = 4, nrow = 3, vjust = 1.5)
@@
 ```{r deep ecklonia wave, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:deep ecklonia wave}Boxplots representing deep *E.
-maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
+maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
 The lower and upper hinges correspond to the first and third quartiles.
-The whiskers represent the range, solid black lines represent the median and black dots are outliers.', fig.height=8, fig.width=8}
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.
+Significance levels are as follows: * represents ≤0.05, ** represents ≤0.01, *** represents ≤0.001, **** represents ≤0.0001.', fig.height=8, fig.width=8}
 ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, common.legend = TRUE, legend = "bottom", ncol = 4, nrow = 3, vjust = 1.5)
@@
 ```{r shallow ecklonia temp, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.cap='\\label{fig:shallow ecklonia temp} Boxplots representing shallow *E.
-maxima* morphological characteristics  grouped by temperature regime measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
+maxima* morphological characteristics grouped by temperature regime measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
 The lower and upper hinges correspond to the first and third quartiles.
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-Light grey boxes represent cool-temperate and dark grey warm-temperate.', fig.height=7, fig.width=8}
+Light gray boxes represent cool-temperate and dark gray warm-temperate.
+Significance levels are as follows: * represents ≤0.05, ** represents ≤0.01, *** represents ≤0.001, **** represents ≤0.0001.', fig.height=7, fig.width=8}
 ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, common.legend = FALSE, legend = "bottom", ncol = 4, nrow = 3, vjust = 1.5)
@@
 ```{r shallow ecklonia wave, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE, fig.cap='\\label{fig:shallow ecklonia wave}Boxplots representing shallow *E.
-maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured, with units provided.
+maxima* morphological characteristics grouped by wave exposure category measured around the Western Cape coastline, with the y-axis depicting the specific morphology measured, with units provided.
 The lower and upper hinges correspond to the first and third quartiles.
-The whiskers represent the range, solid black lines represent the median and black dots are outliers.', fig.height=7, fig.width=8}
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.
+Significance levels are as follows: * represents ≤0.05, ** represents ≤0.01, *** represents ≤0.001, **** represents ≤0.0001.', fig.height=7, fig.width=8}
 ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9,ESW10, common.legend = TRUE, legend = "bottom", ncol = 4,nrow = 3)
@@
 maxima* (C).
-The blue cluster represnts sites from the western region and the red represents sites from the eastern region."}
+The blue cluster represents sites from the western region and the red represents sites from the eastern region."}
 #ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, legend = "none", labels = "AUTO", ncol = 2, nrow = 2, align = "hv")
@@
 
-```{r appendix figure, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=8}
+```{r appendix figure, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7,fig.width=8, fig.cap = "The total and seasonal wave roses of the directional wave buoy just offshore from Cape Town is given for years 2000 till end of 2017 (the period for which the directional wave spectra was available).
+The total (A), Austral winter (B), and Austral summer (C) are presented, respectively."}
 

### 2021-03-12 thesis_chapter_2.Rmd
chapter 2 updated, tables to come
tables added. issue with table one header though

@@
 ```
+
 # Introduction
@@
 
+```{r table 2.1, message= FALSE, warning=FALSE, echo=FALSE}
+
+library(flextable)
+
+morph_table <- read.csv("tables/chapter_2/table_1.csv", sep = ";")
+morph_table <- flextable(data = morph_table)
+morph_table <- theme_zebra(morph_table, odd_header = "transparent", even_header = "transparent")
+morph_table <- set_header_labels(morph_table, Morphological.Characteristic = "Morphological Characteristic", Unit.measurement = "Unit.measurement")
+morph_table <- align(x= morph_table, align = "center", part = "all")
+morph_table <- set_table_properties(morph_table, width = 1, layout = "autofit")
+morph_table
+
+```
+
 ## Statistical analyses
@@
 Summary statistics for wind was not plotted and instead are discussed, as the data were course relative to the other environmental variables considered in this study.
+
+```{r table 2.2, message= FALSE, warning=FALSE, echo=FALSE}
+
+var_table <- read.csv("tables/chapter_2/table_2.csv", sep = ";")
+var_table <- flextable(data = var_table)
+var_table <- theme_zebra(var_table, odd_header = "transparent", even_header = "transparent")
+var_table <- align(x= var_table, align = "center", part = "all")
+var_table <- set_table_properties(var_table, width = 1, layout = "autofit")
+var_table
+```
 

### 2021-03-12 thesis_chapter_2.Rmd
outputs updated. minor issues still

@@
 output:
+  word_document:
+    reference_docx: markdown/frontiers_template.docx
+  html_document:
+    df_print: paged
   pdf_document: 
@@
     latex_engine: xelatex
-  html_document:
-    df_print: paged
-  word_document:
-    reference_docx: markdown/frontiers_template.docx
 bibliography: references.bib

### 2021-03-12 doodle.R
file delete

@@ -1,29 +1,0 @@
-library(gganimate)
-library(tidync)
-library(tidyverse)
-library(ncdf4)
-
-
-climate_filepath <- paste0("OceanParcels/trajectory_data/Mean_KFloat_H100_W00.nc")
-climate_output <- nc_open(climate_filepath)
-
-
-p_eddy <- tidync("OceanParcels/trajectory_data/PassiveEddy.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(lon, lat, time) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-p_eddy_mean <- p_eddy %>% # calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-p_eddy$type <- "Passive Eddy" # add extra column 'type' to dataframe
-
-
-
-
-
-
-
-
-

### 2021-03-12 figure_prep.R
delete

@@ -1,11 +1,0 @@
-library(jpeg)
-library(grid)
-library(gridExtra)
-library(ggpubr)
-
-wp_1 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6A.jpg")))
-img_fb <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_7.jpg")))
-wp_6 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6_legend.jpg")))
-
-ggarrange(wp_1,img_fb, wp_6, ncol = 2, nrow = 2, labels = c("A", "C", "B"), widths = c(1,2), align = "hv")
-

### 2021-03-15 thesis_chapter_3.Rmd
added in sentence to first paragraph

@@
 There are a range of objects, both natural and anthropogenic, floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
-Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
+Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008]; as well as input of allochthonous materials to near and distant ecosystems [@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].
+
 Some macroalgae species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species of macroalgae have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.

### 2021-03-15 thesis_chapter_3.Rmd
AJ comments made to chapter 3 (round 1)

@@
 
-Some macroalgae species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species of macroalgae have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.
+Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species of macroalgae have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.
 In turn, the positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum.
@@
 radiata*.
-Although there have been reports of floating Chlorophyta species and some Rhodophyta species, Phaeophyceae species are the most commonly reported forms of floating algae.
-This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped inbetween or in the thalli [@dromgoole1982; @back2000].
-The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @putman2018; @brooks2018] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
+Although there have been reports of floating Chlorophyta species and some Rhodophyta species, Phaeophyceae species are the most reported forms of floating algae.
+This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in the thalli [@dromgoole1982; @back2000].
+The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @putman2018; @brooks2018; ] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
 
@@
 pyrifera* in nearshore Monterrey Bay.
-The results showed that kelp rafts, with little surface, area exposed to the wind were largely driven by a combination of wind and wind waves, however the relative importance of wind and wind waves was not clear.
-In addition, the tagged kelp trajectories were more consistent with the formation of eddies during winter.
+The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents, however the relative importance of wind and was not clear.
+In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter.
 Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems.
-For example, the subAntartic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating kelp.
+For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating bull kelp *Durvillaea antarctica*.
 Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013].
-For example, a study by @fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft.
-The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate; while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin.
-The authors suggested that wind and water-movement were the primary influences of trajectory, however, this was only inferred from the genetic results and local climatology data [@fraser2011].
+For example, a study by @fraser2011 on the rafting capabilities of *D.
+antarctica* used a combination of population genetics and relative age estimate of goosebarnacles attached to the raft.
+The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate.
+Genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin.
+The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011].
 Past research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
-The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag [@breivik2011; @griffin2017].
+The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag [@breivik2011; @griffin2017].
 In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object.
@@
 
-To accurately determine the trajectory of marine macroalgae these aspects need to be taken into account.
+To accurately determine the trajectory of marine macroalgae water and wind drag need to be taken into account.
 Past research by @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
@@
 Although drag estimates do exist, these have been applied for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
-Although various approaches have been used in the past to investigate floating macroalagae trajectory, very few studies have employed the use of Langrangian trajectory modeling.
-Furthermore, none of the existing studies using this modeling approach have considered macroalgal morphology cross-sectional area (i.e.
-shape) as an aspect of drag and ultimately trajectory.
-Studies by @brooks2019, @putman2018, @putman2020 used lagrangian approaches to investigate the effects of inertia, raft-size, and windage on macroalgae trajectory.
+Although various approaches have been used in the past to investigate floating macroalgal trajectories, very few studies have employed the use of Langrangian trajectory modeling.
+Furthermore, none of the existing studies using this modeling approach have considered macroalgal morphology cross-sectional area (**i.e.** shape) as an aspect of drag and ultimately trajectory.
+Studies by @brooks2019, @putman2018, and @putman2020 used Lagrangian approaches to investigate the effects of inertia, raft-size, and windage on macroalgae trajectory.
 @brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customised Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
@@
 Other work by @putman2020 investigated the effect of including windage in macroalgae trajectory simulations.
-The authors showed that including ad-hoc windage factors greatly improved accuracy regarding entrainment or expulsion from eddies when compared to tracked sargassum mats in the ocean.
+The authors showed that including **ad-hoc** windage factors greatly improved accuracy regarding entrainment or expulsion from eddies when compared to tracked sargassum mats in the ocean.
 Langrangian trajectory modeling is a useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
-The radial size included by @brooks2019 is an improvement, however this approach does not take into account the complex morphology on an individual level into account.
+The radial size included by @brooks2019 is an improvement, however this approach does not consider the complex morphology on an individual level into account.
 
@@
 Past studies investigating the dispersal of marine debris have also included the possible effects of inertia in accumulation of ocean drifter instruments in mesoscale features.
-For example, a study by @beron2016 used a modeling approach to investigated the role of eckman transport and inertia to explain the observed behaviour of undrogued drifters in the ocean gyres.
-Undrogued drifter trajectory is affected by wind unlike the drogue counter parts which makes the effects of wind neglible.
-The results showed that the combined effect inertia, not eckman transport, played a significant role in the observed drifter behaviour [@beron2016].
+For example, a study by @beron2016 used a modeling approach to investigated the role of Eckman transport and inertia to explain the observed behavior of undrogued drifters in the ocean gyres.
+Undrogued drifter trajectory is affected by wind unlike the drogue counter parts which makes the effects of wind negligible.
+The results showed that the combined effect inertia, not Eckman transport, played a significant role in the observed drifter behavior [@beron2016].
 The authors also noted that the presence of other flotsam along with the undrogued drifters in ocean gyres also suggests that aspects inertia also play a role in accumulation of other marine debris [@beron2016].
@@
 
-In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using lagrangian based trajectory simulations.
+In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations.
 It should be noted that this study does not attempt to characterise the dispersal of drifting macroalage within the study domain.

### 2021-03-15 thesis_chapter_3.Rmd
study site updated

@@
 The hydrodynamics in the study region are driven by a complex interaction of wind and waves, which are also influenced by larger oceanic processes.
-In South Africa, the larger processes can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC)[@lutjeharms2000;@lutjeharms2007;@garzoli1996;@Hutchings2009;@rubio2009].
-The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@@lutjeharms2000;@lutjeharms2007].
-The region where the retroflection occurs is known as the Agulhas retroflection and consists of high levels of mesoscale variability [@garzoli1996].
+In South Africa, the larger processes can be attributed to the Agulhas Current (AC) and the Benguela Current (BC)[@lutjeharms2000;@lutjeharms2007;@garzoli1996;@Hutchings2009;@rubio2009].
+The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].
+The region where the retroflection occurs is known as the Agulhas Retroflection and consists of high levels of mesoscale variability [@garzoli1996].
 The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988; @lutjeharms2006; @lutjeharms2007].
@@
 The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009].
-The Benguela region is characterised by strong coastal upwelling cells which are primarily driven by equatorward winds [@shannon1996; @hardman2003], and is known as the Benguela upwelling region.
-The Benguela upwelling region starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996].
+The Benguela region is characterised by strong coastal upwelling cells which are primarily driven by equatorward winds [@shannon1996; @hardman2003].
+It is known as the Benguela Upwelling System (BUS) The BUS starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996].
 The region is also characterised by high amounts of mesoscale variability in the form of eddies and filaments [@blanke2002; @blanke2005; @rubio2009].
-The Benguela inshore region is dominated by a northwestward flow as a result of topographical steering, wind stress and interactions with passing Agulhas rings and eddies [@veitch2010].
+The Benguela inshore region is dominated by a northwestward flow because of topographical steering, wind stress and interactions with passing Agulhas Rings and eddies [@veitch2010].
 

### 2021-03-15 thesis_chapter_3.Rmd
more corrections made

@@
 The stability of the jet varies seasonally with more stable flows in winter and splitting of the jet during summer and spring.
-In January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2009] which results into a weak narrow inshore branch which flows along the shelf edge and a strong flowing north-westwards offshore branch.
+In January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results into a weak narrow inshore branch which flows along the shelf edge and a strong flowing north-westwards offshore branch.
 The area is also an area of high upwelling activity driven by strong westerly winds [@hutchings2009].
@@
 It should be noted that most of the previous studies regarding macroalgae trajectory have investigated rafts and not solitary floating individuals.
-A kelp raft will vary in size and can range from a few meters across to large dense mats the size of a sports field.
+A kelp raft will vary in size and can range from a few meters across to large dense mats the size of a rugby field.
 Furthermore, some authors suggest that most macroalgae become entangled through ocean currents and not through the hydrodynamic forces that dislodge them.
-In addition, large kelp rafts are not a characteristics around the coast of the Cape Peninsula but rather solitary kelp are (Pers.
-Obs).
+In addition, large kelp rafts are not a feature around the coast of the Cape Peninsula; rather, solitary kelp are (pers.
+obs).
 Therefore, the trajectory of only solitary *E.
 maxima* individuals will be considered.
-Numerical calculations of particle trajectory with a Lagrangian method [@delandmeter2019], and be described by the equation
+Numerical calculations of particle trajectory were performed with a Lagrangian method [@delandmeter2019], and be described by the equation
 
@@
 A momentum energy equation was used to calculate the drag forces for the relevant simulation.
-This approach has been employed in modeling iceberg drift trajectory [@andersson2017; @lichey2001; @eik2009] and was adapted to suit this particular study.
+This approach has been employed in modeling iceberg drift trajectory [@andersson2017; @lichey2001; @eik2009] and was adapted to suit this study.
 The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was, $$m({d_u/d_t})=F_d + F_a + F_f$$

### 2021-03-15 thesis_chapter_3.Rmd
more corrections chapter 3

@@
 This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals.
-This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley lagrangian particles.
+This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley Lagrangian particles.
 The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scnearios causes accumulation in eddies and vortices within the study domain.
@@
 
-Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service.
+Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service (Copernicus, http://marine.copernicus.com).
+The Copernicus model is based on the modelling framework known as Nucleus for European Modelling of the ocean.
 The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
-The model has a spatial resolution of 0.08$^\circ$ and is interpolated on an Arakawa C native grid.
-
-In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (WIND_L4_NRT_OBSERVATIONS_012_004).
+The model is forcedhas a spatial resolution of 0.08$^\circ$ and is interpolated on an Arakawa C native grid.
+In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (Global Ocean Wind L4_NRT_012_004).
 The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
@@
 
-In order to incorporate hydrodynamic and wind drag, the cross-sectional area of the kelp was calculated first.
-Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant, for details please refer to figure 2.
-The dimensional data needed was estimated in cases were data was not available for that particular morphological characteristic.
-The bulb/pneumatocyst is a highly variable morphological characteristic and in some cases can appear absent, the same is true for the holdfast.
+To incorporate hydrodynamic and wind drag, the cross-sectional area of the kelp was calculated first.
+Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant.
+For details please refer to figure 2.
+The dimensional data needed were estimated in cases where data was not available for that particular morphological characteristic.
+The bulb/pneumatocyst is a highly variable morphological characteristic and in some cases can appear absent; the same is true for the holdfast.
 Therefore, a standard cross-sectional area was used for both the bulb and holdfast, for all simulations.
@@
 The minimum, mean and maximum were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (see table 1).
-The trajectory of these "types" of cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
+The trajectory of these 'types' of cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
 
@@
 The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (Figure 2A).
-When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen with only slight variation.
+When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen; rather, there is only a slight variation.
 The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation (Figure 2B).
-The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well as being further from the release site.
+The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well the final location of the kelp particles being further from the release site.
 When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories was evident.
@@
 The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids (Figure 3A, 3E, 3I).
-In addition the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticy while particles in the passive simulation deflected away (Figure 3).
+In addition the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 3).
 Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.
@@
 There were significant differences among the median distances travelled for all particles in each simulation (Figure 4; p < 0.05).
-The simulations that considered for 100% hydrodynamic drag travelled less distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
+The simulations that considered for 100% hydrodynamic drag travelled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
 Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations.
@@
 
-The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no dissimilarity (Figure 6A).
+The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (Figure 6A).
 This was also reflected when comparing distance travelled by particles and particle density plots (Figure 6B).
@@
 
-The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @miron2020].
-Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], past research has focused on large macroalgae-rafts which can vary greatly in size and surface area.
-None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae.
+The treatment of drifting macroalgae as purely Lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @miron2020].
+Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], past research has focused on large macroalgal-rafts which can vary greatly in size and surface area.
+None of the past research has taken advantage of the advancement in numerical ocean models and Lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae.
 This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag.
 The mean virtual kelp trajectory deviated from the mean passive  trajectory when including varying levels of hydrodynamic and wind drag scenarios.
-The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns, and also our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.
+The results from this study show the inclusion of hydrodynamic and wind drag results in observable entrainment patterns, and broadens our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case that of *E.
+maxima*.
 
@@
 In addition, not only are the trajectories similar but so are the end points at the end of simulation.
-However, when ad-hoc wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely lagrangian particle trajectories.
+However, when **ad-hoc** wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely Lagrangian particle trajectories.
 The inclusion of wind drag causes the particles to flow further away from the coast and with a greater distance travelled from the release location.
@@
 Another possible reason is the resolution of the ocean model used.
-The resolution of the underlying ocean model is an important aspect of lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications.
+The resolution of the underlying ocean model is an important aspect of Lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications.
 A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study.
@@
 The findings from this study indicate that solitary floating virtual *E.
-maxima* particles tend to follow the prevailing surface currents and the inclusion of wind causes the trajectories to differ greatly from that of a purely lagrangian particle.
+maxima* particles tend to follow the prevailing surface currents and the inclusion of wind causes the trajectories to differ greatly from that of a purely Lagrangian particle.
 Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies.

### 2021-03-15 thesis_chapter_3.Rmd
round 1 additions to chapter 3 writing done

@@
 In addition, not only are the trajectories similar but so are the end points at the end of simulation.
-However, when **ad-hoc** wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely Lagrangian particle trajectories.
-The inclusion of wind drag causes the particles to flow further away from the coast and with a greater distance travelled from the release location.
+However, when *ad-hoc* wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely Lagrangian particle trajectories.
+The inclusion of wind drag causes the kelp particles to flow further away from the coast and with a greater distance travelled from the release location.
 The difference in trajectory and distance covered can be attributed to the inclusion of wind drag.
@@
 However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object [@hackett2006; @putman2020].
-This causes the particles to travel further as well as exposed to different time varying flows.
+This causes the kelp particles to travel further as well as exposed to different time varying flows.
 Past studies have shown this when including windage as a function of inertia [@putman2018;@brooks2019; @putman2020].
-A study by @putman2020 used ad-hoc windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic Sargassum.
-The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked Sargassum mats which was partly due to the inclusion of momemtum energy transfer from wind.
+A study by @putman2020 used *ad-hoc* windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic *Sargassum sp*.
+The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked *Sargassum sp* mats which was partly due to the inclusion of momemtum energy transfer from wind.
 This is also reflected in the analysis comparing distances travelled by all the particles in each simulation.
@@
 
-The addition of drag forces causes the particles to cluster together along the mean trajectory compared to that of the passive particles which are more evenly dispersed.
-This is most likely due to the shared drag characteristics among the particles which included the plants weight and surface area into the calculation.
+The addition of drag forces causes the kelp particles to cluster together along the mean trajectory compared to that of the passive particles which are more evenly dispersed.
+This is most likely due to the shared drag characteristics among the kelp particles which included the plants weight and surface area into the calculation.
 Other studies investigating flotsam trajectory characteristics have found similar results [@breivik2011; @olascoaga2020; @miron2020].
-A study by @miron2020 conducted field experiments using a range of objects such as spheres, cubes, cuboids to compare to investigate the effect of inertia on particles dynamics.
+A study by @miron2020 conducted field experiments using a range of objects such as spheres, cubes, and cuboids to compare to investigate the effect of inertia on particles dynamics.
 The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia.
-
-The ocean is made up of varying time-varying flows and changes in an objects velocity can cause the object to be exposed to different flow patterns over time.
+Although drag forces and not inertia were used in this study, the same conclusion applies; drag is therefore an important characteristic to consider when investigating or predicting solitary drifting macroalgae trajectory.
+
+The ocean is comprised of time-varying flows and changes in an objects velocity that can cause the object to be exposed to different flow patterns over time.
 The lack of dissimilarity in trajectory between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible.
 A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region.
-The results suggested that the hatchlings trajectories are mostly influenced by the ocean currents in the first year of hatching due to  as juveniles do are not swim fast enough to alter thier overall trajectory.
-Therefore, the differences in cross-sectional area types may not be significant enough in order to have an effect on the relevant velocity vectors (zonal and meridonal).
+The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due to  as juveniles do are not swim fast enough to alter thier overall trajectory.
+Therefore, the differences in cross-sectional area types may not be significant enough to have an effect on the relevant zonal and meridonal velocity vectors.
 Another possible reason is the resolution of the ocean model used.
@@
 A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study.
-The authors found that the inclusion of brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications.
+The authors found that the inclusion of Brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications.
 Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.
@@
 These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag.
-The similarities between the different combined drag scenarios suggests that higher magnitudes of wind exposure do not significantly alter trajectories.
+The similarities between the different combined drag scenarios suggest that higher magnitudes of wind exposure do not significantly alter trajectories.
 Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns.
-All the simulations that included any form of drag resulted in a fairly high density of particles entrained in a vortex.
-The entrainment and expulsion from eddies and vortices is an important characteristic of floating macroalgae.
+All the simulations that included any form of drag resulted in a  high density of particles entrained in a eddy.
+The entrainment and expulsion from eddies and vorticities are an important characteristic of floating macroalgae.
 Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts.
-However, the aforementioned studies investigated large rafts and not floating individuals.
-Inertia is based on the size/weight of the raft, while hydrodynamic and wind drag are based on current velocity and surface area.
+However, these studies investigated large rafts and not floating individuals.
+Inertia is based on the size/mass of the raft, while hydrodynamic and wind drag are based on current velocity and surface area.
 Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered.
-The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.
+The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles, and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.
 
@@
 Instead, the inclusion of a different shape coefficient increased to the density of particles entrained in an eddy field.
-This suggest that inclusion of different shape coefficients play a role in further improving estimates and hindcasts of macroalgal dispersal patterns.
+Therefore, more accurately parameterizing shape will therefore allow for further improvements in models of macroalgal dispersal patterns.
 
@@
 
-Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape, however no research exists attempting to clarify the drifting characteristics of solitary macroalgae.
+Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape; however, no research exists to clarify the drifting characteristics of solitary macroalgae.
 Understanding how to accurately model the distribution of solitary kelp can lead to be understanding of accumulation zones and sinks and ultimately the ecological links pertaining to those processes.
@@
 Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E.maxima*.
-In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory but also causes higher density of particles to become entrained in an eddy.
-Overall the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not.
+In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory, but it does cause higher density of particles to become entrained in an eddy.
+Overall, the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not.
 This study also provides an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately.

### 2021-03-15 thesis_chapter_3.Rmd
output updated. script updated

@@
 This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in the thalli [@dromgoole1982; @back2000].
-The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @putman2018; @brooks2018; ] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
+The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @putman2018; @brooks2018; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
 
@@
 ```{r density plots, message=FALSE, warning=FALSE, echo=FALSE}
-## Heatmaps of study domain, use previous data for study site map and histograms## Coastline of Southern Africa
+
+## Heatmaps of study domain, use previous data for study site map and histograms
+## Coastline of Southern Africa
+
 load("map/south_africa_coast_hi_res.Rdata")
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
   geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
+  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  
+  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  
+  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
 
-```{r eval=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=TRUE}
+```{r eval=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
 
@@
 
-```{r density plots, message=FALSE, warning=FALSE, echo=FALSE}
+```{r density plots additional, message=FALSE, warning=FALSE, echo=FALSE}
 ## Heatmaps of study domain, use previous data for study site map and histograms## Coastline of Southern Africa
@@
 
-```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE, fig.height=3, fig.width=4}
+```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE, fig.height=3, fig.width=4}
 ## July
 
-pos <- c("RK4_July",
+pos_July <- c("RK4_July",
          "Min_Kfloat_H100_W00_July",
@@
 
-pos <- c("RK4_Oct",
+pos_Oct <- c("RK4_Oct",
          "Min_Kfloat_H100_W00_Oct",
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring).
+```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring).
 The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.

### 2021-03-16 thesis_chapter_3.Rmd
additional corrections made.

@@
 author:
-- R Coppin^[University of the Western Cape]
-- C Rautenbach^[South African Weather Service]
-- AJ Smit^[University of the Western Cape]
+- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape]
 output:
@@
 csl: markdown/frontiers.csl.txt
-abstract: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
+abstract: "*Abstract* The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
 Macroalage are one of the most important natural dispersal mechanisms in the ocean, and are regarded as the tumble-weed of the ocean.
@@
 The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scnearios causes accumulation in eddies and vortices within the study domain.
-The results also point to the importance of sub-mesoscale features, such as eddies and vortices, as important surface accumlation zones for solitary floating macroalage.
+The results also point to the importance of sub-mesoscale features, such as eddies and vortices, as important surface accumlation zones for solitary floating macroalage."
 ---
@@
 The hydrodynamics in the study region are driven by a complex interaction of wind and waves, which are also influenced by larger oceanic processes.
-In South Africa, the larger processes can be attributed to the Agulhas Current (AC) and the Benguela Current (BC)[@lutjeharms2000;@lutjeharms2007;@garzoli1996;@Hutchings2009;@rubio2009].
+In South Africa, the larger processes can be attributed to the Agulhas Current (AC) and the Benguela Current (BC)[@lutjeharms2000; @lutjeharms2007 ;@garzoli1996 ;@hutchings2009; @rubio2009].
 The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].
@@
 In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
+
 ## Particle tracking model
@@
 
+load("map/south_africa_coast_hi_res.Rdata")
+
 waterdrag_plot <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
@@
 wwdrag_plot <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
@@
 
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 6: Comparison of trajectory data between passive and kelp accumulation scenarios.
+```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5: Comparison of trajectory data between passive and kelp accumulation scenarios.
 Panel A are a comparison of density plots  and panel B is a comparison of age.", fig.height=8, fig.width=8}
@@
 
-## Additional simulations
+## Additional comparisons
 
@@
 
+### Comparison of mean trajectories
+
 ```{r, message=FALSE, echo=FALSE, warning=FALSE}
@@
 waterdrag_plot_July <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
@@
 wwdrag_plot_July <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
@@
 waterdrag_plot_Oct <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
@@
 wwdrag_plot_Oct <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure ??: Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 7: Comparison of mean trajectories.
 Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
@@
 
+### Comparison of density
+
 ```{r density plots additional, message=FALSE, warning=FALSE, echo=FALSE}
@@
   ylab("Latitude") + 
-  ggtitle("1") +
+  ggtitle("30") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") + 
-  ggtitle("1") +
+  ggtitle("43") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-## Density comparison
-### July
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure ??: Comparison of density plots at the end run time of each scenario for July 2018 (winter).
+## July
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time of each scenario for July 2018 (winter).
 The plots are regular hexagons within the density of particles is calculated.
@@
 
-figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"))
+figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("32","33","34","35","36","37","338","39","40","41","42","43"), common.legend = TRUE, legend = c("none"))
 
@@
 ```
-### October
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure ??: Comparison of density plots at the end run time of each scenario for October 2018 (spring).
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time of each scenario for October 2018 (spring).
 The plots are regular hexagons within the density of particles is calculated.
@@
 
-figure_a_Oct <- ggarrange(heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"))
+figure_a_Oct <- ggarrange(heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("none"))
 
@@
 
-```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE, fig.height=3, fig.width=4}
+```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE}
 ## July
@@
 
-custom_labs_July <- c("1","2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot
+custom_labs_July <- c("32","33","34","35","36","37","338","39","40","41","42","43") #create custom labels and add them to dis_plot
 
@@
 
-custom_labs_Oct <- c("1","2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot
+custom_labs_Oct <- c("44","45","46","47","48","49","50","51","52","53","54","55") #create custom labels and add them to dis_plot
 
@@
 
-```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring).
+```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 10: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring).
 The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.

### 2021-03-16 thesis_chapter_3.Rmd
map alterations chapter 3

@@
 ---
-title: 'Numerical trajectory patterns of floating macroalgae'
+title: 'A numerical experiment investigating the influence of drag in trajectory patterns of floating macroalgae'
 author: 
@@
 csl: markdown/frontiers.csl.txt
-abstract: "*Abstract* The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
+abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
 Macroalage are one of the most important natural dispersal mechanisms in the ocean, and are regarded as the tumble-weed of the ocean.
@@
 #ggarrange(map_chp3A, map_chp3B, ncol = 2, nrow = 1, labels = "AUTO", align = c("h"))
-map_chp3_combined <- rasterGrob(as.raster(readPNG("map/chp3_map_combined_hz.png")))
+map_chp3_combined <- rasterGrob(as.raster(readPNG("map/ch3_map_final.png")))
 ggarrange(map_chp3_combined)
@@
 waterdrag_plot <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
   geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'blue')+
+  geom_point(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), shape = 20)+
   geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'orange')+
@@
 wwdrag_plot <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
@@
 waterdrag_plot_July <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
@@
 wwdrag_plot_July <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
@@
 waterdrag_plot_Oct <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
@@
 wwdrag_plot_Oct <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
   geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+

### 2021-03-16 map/map_chapter3.R

@@ -24,10 +24,10 @@ ggsave(filename = "map_sa.jpg", plot = last_plot(), device = "jpeg", path = "map
 map_pen <- ggplot() +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 1.0, colour = "black", fill = "gray") +
-  coord_map("mercator", clip = "on") +
+  #coord_map("mercator", clip = "on") +
   scaleBar(lon = 15, lat = -34.9, distanceLon = 20, distanceLat = 20, distanceLegend = -5, dist.unit = "km",
            arrow.length = 40, arrow.distance = 20, arrow.North.size = 5) +
-  geom_point(data = release_sites_chp3, aes(x = lon, y = lat), color = "red", nudge_x = 0, nudge_y = 0.1) +
+  geom_point(data = release_sites_chp3, aes(x = lon, y = lat), color = "black", nudge_x = 0, nudge_y = 0.1, size = 5, shape = 17) +
   theme_bw() +
   xlab("Latitude") +
   ylab("Longitude") +
@@ -38,4 +38,4 @@ map_pen <- ggplot() +
 
 ggsave(filename = "map_pen.jpg", plot = last_plot(), device = "jpeg", path = "map/", dpi = 320, width = 10, height = 10)
 
-ggarrange(map_sa, map_pen, ncol = 1, nrow = 2)
+ggarrange(map_sa, map_pen, ncol = 2, nrow = 1, align = c("hv"))

### 2021-03-17 thesis_chapter_3.Rmd
chapter 3 files updated

@@
 abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
-Macroalage are one of the most important natural dispersal mechanisms in the ocean, and are regarded as the tumble-weed of the ocean.
+Macroalage are one of the most important natural dispersal mechanisms, and are regarded as the tumble-weed of the ocean.
 Despite the many studies that exist, the relative role of wind and surface currents in macrolage trajectory is still unknown.
-Past studies also tend to focus on kelp-rafts which can vary greatly in size and making it difficult to form a concensus as to the role of wind and surface currents.
-In addition past studies have not conisidered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam.
+Past studies also focus on kelp-rafts, which can vary greatly in size making it difficult to form a concensus as to the relative role of wind versus surface currents.
+In addition, past studies have not conisidered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam.
 The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectory of flotsam in the ocean.
-This study aimed to shed light on the relative role of wind versus surface currents, as well as the role of drag, in macrolagae trajectory and accumulation.
+This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagae trajectory and accumulation.
 This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals.
 This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley Lagrangian particles.
-The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scnearios causes accumulation in eddies and vortices within the study domain.
-The results also point to the importance of sub-mesoscale features, such as eddies and vortices, as important surface accumlation zones for solitary floating macroalage."
+The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain.
+The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumlation zones for solitary floating macroalage."
 ---
@@
 
-```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2:Diagram of the *E.maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic."}
+```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Diagram of the *E.maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic."}
 
@@
 
-p_RK4_July$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_July$type <- "RK4_July" # add extra column 'type' to dataframe and assigned 'RK4' label
 
@@
 
-p_RK4_Oct$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_Oct$type <- "RK4_Oct" # add extra column 'type' to dataframe and assigned 'RK4' label
 
@@
 
-figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("32","33","34","35","36","37","338","39","40","41","42","43"), common.legend = TRUE, legend = c("none"))
+figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"))
 
@@
 ```
+
 ## Comparison of distances
@@
 ```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE}
-## July
-
-pos_July <- c("RK4_July",
+
+pos_July <- c("RK4_July",             ## July
          "Min_Kfloat_H100_W00_July",
@@
 
-custom_labs_July <- c("32","33","34","35","36","37","338","39","40","41","42","43") #create custom labels and add them to dis_plot
+custom_labs_July <- c("30","31","32","33","34","35","36","37","38","39","40","41", "42") #create custom labels and add them to dis_plot
 
@@
 
-## Oct
-
-pos_Oct <- c("RK4_Oct",
+pos_Oct <- c("RK4_Oct",              ## Oct
          "Min_Kfloat_H100_W00_Oct",
@@
 
-custom_labs_Oct <- c("44","45","46","47","48","49","50","51","52","53","54","55") #create custom labels and add them to dis_plot
+custom_labs_Oct <- c("43","44","45","46","47","48","49","50","51","52","53","54","55") #create custom labels and add them to dis_plot
 
@@
 
-```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 10: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring).
+```{r distance plot additional, include = FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 10: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring).
 The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.

### 2021-03-17 thesis_chapter_3.Rmd
appendix updated

@@
 
-# Appendix
-
 ## Domain features
@@
 
+## April
+
+p_RK4_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+p_RK4_mean_April <- p_RK4_April %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+p_RK4_July$type <- "RK4_April" # add extra column 'type' to dataframe and assigned 'RK4' label
+
+Min_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H100_W00_mean_April <- Min_Kfloat_H100_W00_April %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H100_W00_April$type <- "Min_Kfloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_April$velocity <- with(Min_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Min_Kfloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H095_W05_mean_April <- Min_Kfloat_H095_W05_April %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H095_W05_April$type <- "Min_Kfloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_April$velocity <- with(Min_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Min_Kfloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H090_W10_mean_April <- Min_Kfloat_H090_W10_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H090_W10_April$type <- "Min_Kfloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_April$velocity <- with(Min_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Min_Kfloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H085_W15_mean_April <- Min_Kfloat_H085_W15_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H085_W15_April$type <- "Min_Kfloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_April$velocity <- with(Min_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_Kfloat_H100_W00_mean_April <- Mean_Kfloat_H100_W00_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_Kfloat_H100_W00_April$type <- "Mean_Kfloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_April$velocity <- with(Mean_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_KFloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H095_W05_mean_April <- Mean_KFloat_H095_W05_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H095_W05_April$type <- "Mean_KFloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_April$velocity <- with(Mean_KFloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_KFloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H090_W10_mean_April <- Mean_KFloat_H090_W10_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H090_W10_April$type <- "Mean_KFloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_April$velocity <- with(Mean_KFloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Mean_KFloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H085_W15_mean_April <- Mean_KFloat_H085_W15_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H085_W15_April$type <- "Mean_KFloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_April$velocity <- with(Mean_KFloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H100_W00_mean_April <- Max_Kfloat_H100_W00_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H100_W00_April$type <- "Max_KFloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_April$velocity <- with(Max_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H095_W05_mean_April <- Max_Kfloat_H095_W05_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H095_W05_April$type <- "Max_KFloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_April$velocity <- with(Max_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H090_W10_mean_April <- Max_Kfloat_H090_W10_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H090_W10_April$type <- "Max_KFloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_April$velocity <- with(Max_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+Max_Kfloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H085_W15_mean_April <- Max_Kfloat_H085_W15_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H085_W15_April$type <- "Max_KFloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_April$velocity <- with(Max_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
+
+total_data_April <- rbind(Min_Kfloat_H100_W00_April, Min_Kfloat_H095_W05_April, Min_Kfloat_H090_W10_April, Min_Kfloat_H085_W15_April, Mean_Kfloat_H100_W00_April, Mean_KFloat_H095_W05_April, Mean_KFloat_H090_W10_April, Mean_KFloat_H085_W15_April, Max_Kfloat_H100_W00_April, Max_Kfloat_H095_W05_April, Max_Kfloat_H090_W10_April, Max_Kfloat_H085_W15_April)
+
 ## July
@@
 
-Min_Kfloat_H100_W00_July$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_July$type <- "Min_Kfloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Min_Kfloat_H100_W00_July$velocity <- with(Min_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Min_Kfloat_H095_W05_July$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_July$type <- "Min_Kfloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Min_Kfloat_H095_W05_July$velocity <- with(Min_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Min_Kfloat_H090_W10_July$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_July$type <- "Min_Kfloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Min_Kfloat_H090_W10_July$velocity <- with(Min_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Min_Kfloat_H085_W15_July$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_July$type <- "Min_Kfloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Min_Kfloat_H085_W15_July$velocity <- with(Min_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Mean_Kfloat_H100_W00_July$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_July$type <- "Mean_Kfloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Mean_Kfloat_H100_W00_July$velocity <- with(Mean_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Mean_KFloat_H095_W05_July$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_July$type <- "Mean_KFloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Mean_KFloat_H095_W05_July$velocity <- with(Mean_KFloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Mean_KFloat_H090_W10_July$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_July$type <- "Mean_KFloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Mean_KFloat_H090_W10_July$velocity <- with(Mean_KFloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Mean_KFloat_H085_W15_July$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_July$type <- "Mean_KFloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Mean_KFloat_H085_W15_July$velocity <- with(Mean_KFloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Max_Kfloat_H100_W00_July$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_July$type <- "Max_KFloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Max_Kfloat_H100_W00_July$velocity <- with(Max_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Max_Kfloat_H095_W05_July$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_July$type <- "Max_KFloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Max_Kfloat_H095_W05_July$velocity <- with(Max_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Max_Kfloat_H090_W10_July$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_July$type <- "Max_KFloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Max_Kfloat_H090_W10_July$velocity <- with(Max_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Max_Kfloat_H085_W15_July$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_July$type <- "Max_KFloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
 Max_Kfloat_H085_W15_July$velocity <- with(Max_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Min_Kfloat_H100_W00_Oct$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_Oct$type <- "Min_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Min_Kfloat_H095_W05_Oct$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_Oct$type <- "Min_Kfloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Min_Kfloat_H090_W10_Oct$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_Oct$type <- "Min_Kfloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Min_Kfloat_H085_W15_Oct$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_Oct$type <- "Min_Kfloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Mean_Kfloat_H100_W00_Oct$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_Oct$type <- "Mean_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Mean_KFloat_H095_W05_Oct$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_Oct$type <- "Mean_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Mean_KFloat_H090_W10_Oct$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_Oct$type <- "Mean_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Mean_KFloat_H085_W15_Oct$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_Oct$type <- "Mean_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Max_Kfloat_H100_W00_Oct$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_Oct$type <- "Max_KFloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Max_Kfloat_H095_W05_Oct$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_Oct$type <- "Max_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Max_Kfloat_H090_W10_Oct$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_Oct$type <- "Max_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 
-Max_Kfloat_H085_W15_Oct$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_Oct$type <- "Max_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
 Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
@@
 ```{r, message=FALSE, echo=FALSE, warning=FALSE}
+waterdrag_plot_April <- ggplot()+
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = Mean_Kfloat_H100_W00_mean_April, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = Mean_Kfloat_H100_W00_mean_April, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Max_Kfloat_H100_W00_mean_April, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+wwdrag_plot_April <- ggplot()+
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat), color = 'grey')+
+  geom_path(data = Min_Kfloat_H090_W10_mean_April, aes(x = lon, y = lat), color = 'green')+
+  geom_path(data = Min_Kfloat_H085_W15_mean_April, aes(x = lon, y = lat), color = 'pink')+
+  geom_path(data = Mean_KFloat_H095_W05_mean_April, aes(x = lon, y = lat), color = 'yellow')+
+  geom_path(data = Mean_KFloat_H090_W10_mean_April, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = Max_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = Max_Kfloat_H090_W10_mean_April, aes(x = lon, y = lat), color = 'orange')+
+  geom_path(data = Max_Kfloat_H085_W15_mean_April, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
 
@@
 
-ggarrange(waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, labels = "AUTO")
-
-```
+ggarrange(waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, labels = "AUTO", ncol = 2, nrow = 3)
+
+```
+In the April simulations, both passive and kelp particles followed the general north-westward flow of the region with clear entrainment of kelp particles in sub-mesoscale features.
+Kelp particles follow the north
 
@@
 
+## April
+
+heatmap_RK4_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = p_RK4_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("30") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H100_W00_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H100W100") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H095_W05_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  #ggtitle("H095W05") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H090_W10_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H090W10") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H085_W15_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_path(data = Min_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H085W15") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_Kfloat_H100_W00_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_path(data = Mean_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H095_W05_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H090_W10_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H085_W15_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H100_W00_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H100_W00_mean_April ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H095_W05_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  theme(plot.title = element_text(hjust = 0.5)) + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H090_W10_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H085_W15_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_path(data = Max_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
 ## July
@@
 
-## July
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time of each scenario for July 2018 (winter).
+## April
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time of each scenario for April 2018 (winter).
 The plots are regular hexagons within the density of particles is calculated.
@@
 
+figure_a_April <- ggarrange(heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"))
+
+figure_b_April <- ggarrange(heatmap_RK4_April, legend = c("right"))
+
+ggarrange(figure_a_April, figure_b_April, ncol = 1, nrow = 2)
+
+```
+
+## July
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time of each scenario for July 2018 (winter).
+The plots are regular hexagons within the density of particles is calculated.
+Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
+The bottom plot is the density plot for the passive simulation.
+The mean trajectories for each scenario are also plotted for comparison.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
+
 figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"))
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time of each scenario for October 2018 (spring).
+## October
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 10: Comparison of density plots at the end run time of each scenario for October 2018 (spring).
 The plots are regular hexagons within the density of particles is calculated.
@@
 
-```{r distance plot additional, include = FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 10: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring).
+```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 11: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for April (Autumn), July (winter), and October (spring).
 The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.

### 2021-03-18 appendix_chapter_3.Rmd
moving appendix for chapter 3 to seperate rmd

--- 2021-03-17 thesis_chapter_3.Rmd
@@
 ---
-title: 'A numerical experiment investigating the influence of drag in trajectory patterns of floating macroalgae'
-author: 
-- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape]
+title: 'Chapter 3: Appendix'
 output:
@@
 csl: markdown/frontiers.csl.txt
-abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
-Macroalage are one of the most important natural dispersal mechanisms, and are regarded as the tumble-weed of the ocean.
-Despite the many studies that exist, the relative role of wind and surface currents in macrolage trajectory is still unknown.
-Past studies also focus on kelp-rafts, which can vary greatly in size making it difficult to form a concensus as to the relative role of wind versus surface currents.
-In addition, past studies have not conisidered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam.
-The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectory of flotsam in the ocean.
-This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagae trajectory and accumulation.
-This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals.
-This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley Lagrangian particles.
-The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain.
-The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumlation zones for solitary floating macroalage."
 ---
-
 ```{r software, message=FALSE, warning=FALSE, include=FALSE}
@@
 
-# Introduction
-
-There are a range of objects, both natural and anthropogenic, floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
-Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008]; as well as input of allochthonous materials to near and distant ecosystems [@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].
-
-Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species of macroalgae have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.
-In turn, the positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum.
-The structure and number of pneumatocysts varies between species.
-For example species from the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018].
-In some cases the stipe itself is air-filled, such as with *E.
-maxima* and *E.
-radiata*.
-Although there have been reports of floating Chlorophyta species and some Rhodophyta species, Phaeophyceae species are the most reported forms of floating algae.
-This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in the thalli [@dromgoole1982; @back2000].
-The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @putman2018; @brooks2018; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
-
-Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
-Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown although the role of wind has been recognised as important in several studies.
-For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M.
-pyrifera* in nearshore Monterrey Bay.
-The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents, however the relative importance of wind and was not clear.
-In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter.
-Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems.
-For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating bull kelp *Durvillaea antarctica*.
-Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013].
-For example, a study by @fraser2011 on the rafting capabilities of *D.
-antarctica* used a combination of population genetics and relative age estimate of goosebarnacles attached to the raft.
-The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate.
-Genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin.
-The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011].
-Past research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
-The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag [@breivik2011; @griffin2017].
-In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object.
-If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].
-Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.
-
-To accurately determine the trajectory of marine macroalgae water and wind drag need to be taken into account.
-Past research by @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
-However, this work does not consider biological material such as macroalgae.
-Although drag estimates do exist, these have been applied for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
-Although various approaches have been used in the past to investigate floating macroalgal trajectories, very few studies have employed the use of Langrangian trajectory modeling.
-Furthermore, none of the existing studies using this modeling approach have considered macroalgal morphology cross-sectional area (**i.e.** shape) as an aspect of drag and ultimately trajectory.
-Studies by @brooks2019, @putman2018, and @putman2020 used Lagrangian approaches to investigate the effects of inertia, raft-size, and windage on macroalgae trajectory.
-@brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customised Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
-The results showed that trajectory of pelagic *Sargussum* was significantly influenced by inertia and the radial size of the rafts.
-Other work by @putman2020 investigated the effect of including windage in macroalgae trajectory simulations.
-The authors showed that including **ad-hoc** windage factors greatly improved accuracy regarding entrainment or expulsion from eddies when compared to tracked sargassum mats in the ocean.
-Langrangian trajectory modeling is a useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
-The radial size included by @brooks2019 is an improvement, however this approach does not consider the complex morphology on an individual level into account.
-
-In addition, past studies focus on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddies and vortices in the accumulation of kelp at the surface.
-Gyres, eddies and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018].
-Past studies investigating the dispersal of marine debris have also included the possible effects of inertia in accumulation of ocean drifter instruments in mesoscale features.
-For example, a study by @beron2016 used a modeling approach to investigated the role of Eckman transport and inertia to explain the observed behavior of undrogued drifters in the ocean gyres.
-Undrogued drifter trajectory is affected by wind unlike the drogue counter parts which makes the effects of wind negligible.
-The results showed that the combined effect inertia, not Eckman transport, played a significant role in the observed drifter behavior [@beron2016].
-The authors also noted that the presence of other flotsam along with the undrogued drifters in ocean gyres also suggests that aspects inertia also play a role in accumulation of other marine debris [@beron2016].
-The approach by @beron2016 has also been improved upon by @beron2019 but also shows the significant role of inertia in flotsam trajectory and accumulation.
-
-In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations.
-It should be noted that this study does not attempt to characterise the dispersal of drifting macroalage within the study domain.
-The authors argue that too little is know about the drift characteristics of solitary floating macroalgae to characterise dispersal effectivley.
-Instead, this study will aim to shed light on the 'behaviour' of surface drifting macroalgae from a coastal environment into the open ocean, which will inform future characterisation studies.
-This will be achieved by comparing the trajectory outputs of virtual 'passive' particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure.
-The different degrees of wave and wind exposure are meant to reflect the different buoyancy characteristics, as well as identify the role of magnitude of the drag components on overall trajectory.
-In addition, the modeling of single kelp offers the opportunity to study the role of eddies and vortices in accumulation points on the surface.
-Therefore, separate simulations will be run to compare the entrained characteristics of passive particle with a kelp particle.
-
-# Methods
-## Study site
-
-The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A).
-A release site was chosen for this study, Kommetjie, which is located on the Atlantic side of the Cape Peninsula (Figure 2B).
-The hydrodynamics in the study region are driven by a complex interaction of wind and waves, which are also influenced by larger oceanic processes.
-In South Africa, the larger processes can be attributed to the Agulhas Current (AC) and the Benguela Current (BC)[@lutjeharms2000; @lutjeharms2007 ;@garzoli1996 ;@hutchings2009; @rubio2009].
-The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].
-The region where the retroflection occurs is known as the Agulhas Retroflection and consists of high levels of mesoscale variability [@garzoli1996].
-The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988; @lutjeharms2006; @lutjeharms2007].
-The Agulhas Rings are the main contributors to the Agulhas leakage which are the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015].
-The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009].
-The Benguela region is characterised by strong coastal upwelling cells which are primarily driven by equatorward winds [@shannon1996; @hardman2003].
-It is known as the Benguela Upwelling System (BUS) The BUS starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996].
-The region is also characterised by high amounts of mesoscale variability in the form of eddies and filaments [@blanke2002; @blanke2005; @rubio2009].
-The Benguela inshore region is dominated by a northwestward flow because of topographical steering, wind stress and interactions with passing Agulhas Rings and eddies [@veitch2010].
-
-```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1:Map of the study domain and the main features pertaining to transport in the region.
-In panel A, the map of the entire South African coastline and the main currents presented.
-The resulting eddies and filaments from the Agulhas retroflection and study area are presented.
-In panel B, the splitting of the Benguela Jet, eddie features and general flow of the study domain are presented."}
-
-library(png)
-library(grid)
-library(gridExtra)
-
-#map_chp3A <- rasterGrob(as.raster(readPNG("map/chp3_mapA.png")))
-#map_chp3B <- rasterGrob(as.raster(readPNG("map/chp3_mapB.png")))
-#ggarrange(map_chp3A, map_chp3B, ncol = 2, nrow = 1, labels = "AUTO", align = c("h"))
-map_chp3_combined <- rasterGrob(as.raster(readPNG("map/ch3_map_final.png")))
-ggarrange(map_chp3_combined)
-
-```
-
-The study domain falls within the southern Benguela upwelling ecosystem, which is a region of high mesoscale variability and activity.
-These occur mostly in the form of eddies, filaments and vortices which are generated from the Agulhas retroflection [@veitch2018; @veitch2017], generally flowing northward along the coastline and begin to move offshore near Cape Columbine.
-The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, eckman transport and geostrophic currents [@stramma1989; @veitch2017; @veitch2018; @hutchings2009].
-The stability of the jet varies seasonally with more stable flows in winter and splitting of the jet during summer and spring.
-In January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results into a weak narrow inshore branch which flows along the shelf edge and a strong flowing north-westwards offshore branch.
-The area is also an area of high upwelling activity driven by strong westerly winds [@hutchings2009].
-These winds are most intense during summer (December-March) which favours upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet; which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
-In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
-
-## Particle tracking model
-
-The model is based on parameterisations that have been used for previous studies investigating iceberg, capsized marine vessels and microplastics.
-It should be noted that most of the previous studies regarding macroalgae trajectory have investigated rafts and not solitary floating individuals.
-A kelp raft will vary in size and can range from a few meters across to large dense mats the size of a rugby field.
-Furthermore, some authors suggest that most macroalgae become entangled through ocean currents and not through the hydrodynamic forces that dislodge them.
-In addition, large kelp rafts are not a feature around the coast of the Cape Peninsula; rather, solitary kelp are (pers.
-obs).
-Therefore, the trajectory of only solitary *E.
-maxima* individuals will be considered.
-Numerical calculations of particle trajectory were performed with a Lagrangian method [@delandmeter2019], and be described by the equation
-
-$$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
-where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model at particle point in time, $t$ is time, $X_b(t)$ is a change in position due to 'behavior' of the particle at a particular point in time.
-This can range from swimming in fish to sinking or beaching.
-In order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.
-
-This study simulated two scenarios; one where the plant is fully submerged and only exposed to hydrodynamic drag; and the other where parts of the plants are partially exposed resulting in a combination of hydrodynamic and air form drag.
-Drag based on shape-based coefficients were used in the calculations for determining the overall kelp velocity vectors.
-A momentum energy equation was used to calculate the drag forces for the relevant simulation.
-This approach has been employed in modeling iceberg drift trajectory [@andersson2017; @lichey2001; @eik2009] and was adapted to suit this study.
-The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was, $$m({d_u/d_t})=F_d + F_a + F_f$$
-where $m$ is kelp mass, $d_u$ the virtual kelp particle velocity, $F_d$ the hydrodynamic drag, $F_a$ is the wind drag, $F_f$ the surface-current flow, and $d_t$ is the time step.
-The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations.
-Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculated for each time step for each simulation.
-
-## Simulations
-
-Virtual particles were released with no drag or wind effects from the study site in order to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'.
-All particles that have any form of drag exposure are hereafter referred to as 'kelp' particles.
-All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018].
-Brownian motion was also incorporated in order to account for any stochasticity not resolved by the model.
-These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018].
-To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios.
-These scenarios were meant to act as proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag.
-These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation.
-For example, if 85% of the plant is submerged and 15% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively.
-This approach has been used in past research investigating pelagic Sargassum rafts [@putman2020;@putman2018].
-The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1).
-Each scenario consisted of a particle set with a size of 1000 where released from Kommetjie at the start of simulation.
-The particles sets were advected with a Runge-Kutte fourth order scheme for 30 days and incorporated brownian motion.
-Since the aim of this investigation was to investigate the relative surface behaviour of virtual kelp particles, only the surface layer was used for simulations.
-It should also be noted that the Copernicus model used places velocity vectors over some of the land covering the coastal domain.
-This was not considered a concern, as once again, the aim was to investigate surface drift behavior.
-The time period chosen to run the simulations was 01-01-2018 un 30-01-2018.
-This particular time period was chosen as the splitting of the Benguela Jet occurs in Janurary, which also conincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity [@rubio2009; @veitch2018; @ragoasha2019].
-
-```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Details for each simulation run for the study.
-The table cosists of parameterisations for each scenario.
-The ID coloumn refers to the syntax used in the relevant OceanParcels script (see additional material for link to script).
-Type refers to either a passive particle with no drag behaviour or a kelp particle that has been advected with any drag exposure.
-The drag exposure column refers to the parameterisations regarding the level of water and wind drag simulated."}
-
-set_flextable_defaults(fonts_ignore=TRUE)
-
-sim_table <- read.csv("tables/chapter_3/simulation_table.csv", sep = ";")
-
-sim_table <- flextable(data = sim_table)
-sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
-sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Drag.Exposure = "Drag Exposure level")
-sim_table <- align(x= sim_table, align = "center", part = "all")
-sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
-sim_table <- set_caption(sim_table, "Table 1:Details for each simulation executed for the study.")
-sim_table
-
-```
-
-To determine if vortices and eddies play a role in kelp accumulation on the surface, a separate set of simulations were performed.
-Unlike the previous simulations these simulation released a particle every hour.
-These were run over the same time period for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme and incorporated Brownian motion.
-In order to determine if the shape co-efficient used plays a role in model used, a simulation of a kelp with the maximum surface area and drag exposure was run using the shape coefficient for a cylinder to compare to a simulation using a sphere, and was run for 30 days using a fourth order Runge-Kutte advection scheme with brownian motion and a particle set of size 1000.
-
-## Model inputs
-### Ocean and wind model
-
-Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service (Copernicus, http://marine.copernicus.com).
-The Copernicus model is based on the modelling framework known as Nucleus for European Modelling of the ocean.
-The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
-The model is forcedhas a spatial resolution of 0.08$^\circ$ and is interpolated on an Arakawa C native grid.
-In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (Global Ocean Wind L4_NRT_012_004).
-The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
-Both averaged ocean current and wind data are stored as meridonal and zonal velocity vectors.
-
-### Hydrodynamic and wind drag
-
-The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
-The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
-Hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors, as per @furnans2008.
-Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step.
-The same approach was used for the wind drag force.
-The equation that was used to calculate hydrodynamic drag, was also used to calculate wind drag.
-
-### Cross-sectional area
-
-To incorporate hydrodynamic and wind drag, the cross-sectional area of the kelp was calculated first.
-Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant.
-For details please refer to figure 2.
-The dimensional data needed were estimated in cases where data was not available for that particular morphological characteristic.
-The bulb/pneumatocyst is a highly variable morphological characteristic and in some cases can appear absent; the same is true for the holdfast.
-Therefore, a standard cross-sectional area was used for both the bulb and holdfast, for all simulations.
-
-```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Diagram of the *E.maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic."}
-
-kelp_shape_figure <- rasterGrob(as.raster(readPNG("figures/chapter_3/chp3_shapedrawing.png")))
-ggarrange(kelp_shape_figure)
-
-```
-
-The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
-The minimum, mean and maximum were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (see table 1).
-The trajectory of these 'types' of cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
-
-```{r morphology wrangling, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-# Investigate morphology data to assist in estimating surface area
-
-# Read in morphology data
-eck_df <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
-  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
-  filter(site!= "De Hoop") 
-eck_df$site <- as.character(eck_df$site)
-eck_df <- eck_df %>% 
-  filter(population == "Deep") 
-eck_long_df <- gather(eck_df, Morphology, measurement, frond_mass:total_length, factor_key=TRUE)
-# Summary stats to make it easier to refer to later, these will not be used any analysis.
-site_list <- c("Kommetjie")
-eck_stats <- eck_long_df %>% 
-  group_by(site, Morphology) %>% 
-  summarise(mean = mean(measurement, na.rm = T),
-            minimum = min(measurement, na.rm = T),
-            maximum = max(measurement, na.rm = T)) %>% 
-  ungroup() %>% 
-  filter(site %in% site_list) %>% 
-  rename(Site = site) %>% 
-  rename(Mean = mean) %>% 
-  rename(Minimum = minimum) %>% 
-  rename(Maximum = maximum)
-
-#eck_stats <- flextable(data = eck_stats)
-#eck_stats <- theme_zebra(eck_stats, odd_header = "transparent", even_header = "transparent")
-#eck_stats <- align(x= eck_stats, align = "center", part = "all")
-#eck_stats <- set_table_properties(eck_stats, width = 1, layout = "autofit")
-
-```
-
-```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.
-Approximate shape refers to the dimensions of the geometric shape used to represent a particular morpholgical characteristic.
-Equations used are the equations relating to the dimensions used for the geometric shape estimate.
-Plant dimensions are the measurements used, collected in previous work by @coppin2020, used to calculate the associated surface areas.
-The total surface areas for each plant characteristic for each surface area type make up the remaining columns."}
-
-set_flextable_defaults(fonts_ignore=TRUE)
-
-area_table <- read.csv("tables/chapter_3/area_table.csv", sep = ";")
-area_table <- flextable(data = area_table)
-area_table <- theme_zebra(area_table, odd_header = "transparent", even_header = "transparent")
-area_table <- align(x= area_table, align = "center", part = "all")
-area_table <- set_header_labels(area_table, Plant.characteristic = "Plant characteristic", Approximate.shape = "Approximate shape", Equation.used = "Equation Used", Plant.dimensions = "Plant dimensions")
-area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
-area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
-area_table <- set_caption(area_table, "Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.")
-area_table
-```
-## Analysis
-
-In order to compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated and density distribution maps were produced.
-Density distribution maps are an effective method for assessing pathways and ocean connectivity [@vanSebille2018].
-In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs.
-The distance that particles traveled was also measured and used to produce boxplots.
-Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@VanSebille2018].
-If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; @putman2018; @vanSebille2018].
-
-# Results
-
-```{r data wrangling, message=FALSE, warning=FALSE, include=FALSE}
-
-# Read in simulation data and data wrangling
-
-p_RK4 <- tidync("OceanParcels/trajectory_data/RK4_PFloat.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-p_RK4_mean <- p_RK4 %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
-
-Min_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H100_W00.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H100_W00_mean <- Min_Kfloat_H100_W00 %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H100_W00$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H100_W00$velocity <- with(Min_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Min_Kfloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H095_W05.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H095_W05_mean <- Min_Kfloat_H095_W05 %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H095_W05$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H095_W05$velocity <- with(Min_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Min_Kfloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H090_W10.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H090_W10_mean <- Min_Kfloat_H090_W10 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H090_W10$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H090_W10$velocity <- with(Min_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Min_Kfloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H085_W15.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H085_W15_mean <- Min_Kfloat_H085_W15 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H085_W15$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H085_W15$velocity <- with(Min_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H100_W00.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_Kfloat_H100_W00_mean <- Mean_Kfloat_H100_W00 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_Kfloat_H100_W00$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_Kfloat_H100_W00$velocity <- with(Mean_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_KFloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H095_W05.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H095_W05_mean <- Mean_KFloat_H095_W05 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H095_W05$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H095_W05$velocity <- with(Mean_KFloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_KFloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H090_W10.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H090_W10_mean <- Mean_KFloat_H090_W10 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H090_W10$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H090_W10$velocity <- with(Mean_KFloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_KFloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H085_W15.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H085_W15_mean <- Mean_KFloat_H085_W15 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H085_W15$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H085_W15$velocity <- with(Mean_KFloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H100_W00.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H100_W00_mean <- Max_Kfloat_H100_W00 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H100_W00$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H100_W00$velocity <- with(Max_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H095_W05.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H095_W05_mean <- Max_Kfloat_H095_W05 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H095_W05$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H095_W05$velocity <- with(Max_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H090_W10.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H090_W10_mean <- Max_Kfloat_H090_W10 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H090_W10$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H090_W10$velocity <- with(Max_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H085_W15.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H085_W15_mean <- Max_Kfloat_H085_W15 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H085_W15$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H085_W15$velocity <- with(Max_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
-
-total_data <- rbind(Min_Kfloat_H100_W00 , Min_Kfloat_H095_W05, Min_Kfloat_H090_W10, Min_Kfloat_H085_W15, Mean_Kfloat_H100_W00, Mean_KFloat_H095_W05, Mean_KFloat_H090_W10, Mean_KFloat_H085_W15, Max_Kfloat_H100_W00, Max_Kfloat_H095_W05, Max_Kfloat_H090_W10, Max_Kfloat_H085_W15)
-
-```
-
-## Trajectory comparison
-
-The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (Figure 2A).
-When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen; rather, there is only a slight variation.
-The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation (Figure 2B).
-The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well the final location of the kelp particles being further from the release site.
-When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories was evident.
-In all simulations particles flowed in a north-westward direction.
-
-```{r meantraj, message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
-
-load("map/south_africa_coast_hi_res.Rdata")
-
-waterdrag_plot <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'blue')+
-  geom_point(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), shape = 20)+
-  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'red')+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-wwdrag_plot <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Min_Kfloat_H095_W05_mean, aes(x = lon, y = lat), color = 'grey')+
-  geom_path(data = Min_Kfloat_H090_W10_mean, aes(x = lon, y = lat), color = 'green')+
-  geom_path(data = Min_Kfloat_H085_W15_mean, aes(x = lon, y = lat), color = 'pink')+
-  geom_path(data = Mean_KFloat_H095_W05_mean, aes(x = lon, y = lat), color = 'yellow')+
-  geom_path(data = Mean_KFloat_H090_W10_mean, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Mean_KFloat_H085_W15_mean, aes(x = lon, y = lat), color = 'red')+
-  geom_path(data = Max_Kfloat_H095_W05_mean, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Max_Kfloat_H090_W10_mean, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H085_W15_mean, aes(x = lon, y = lat), color = 'red')+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Comparison of mean trajectories.
-Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
-In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10.
-In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
-ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO")
-```
-
-## Comparison of density distributions
-
-Across all simulations particles flowed in a north-westward direction (Figure 3).
-The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids (Figure 3A, 3E, 3I).
-In addition the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 3).
-Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.
-
-```{r density plots, message=FALSE, warning=FALSE, echo=FALSE}
-
-## Heatmaps of study domain, use previous data for study site map and histograms
-## Coastline of Southern Africa
-
-load("map/south_africa_coast_hi_res.Rdata")
-
-heatmap_RK4 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("1") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H100_W00 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H095_W05 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H090_W10 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H085_W15 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_Kfloat_H100_W00 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_KFloat_H095_W05 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_KFloat_H090_W10 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_KFloat_H085_W15 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H100_W00 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H095_W05 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  theme(plot.title = element_text(hjust = 0.5)) + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H090_W10 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H085_W15 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time of each scenario.
-The plots are regular hexagons within the density of particles is calculated.
-Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
-The bottom plot is the density plot for the passive simulation.
-The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
-
-figure_a <- ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, labels = c("2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"))
-
-figure_b <- ggarrange(heatmap_RK4, legend = c("right"))
-
-ggarrange(figure_a, figure_b, ncol = 1, nrow = 2)
-
-```
-
-## Comparison of distances
-
-There were significant differences among the median distances travelled for all particles in each simulation (Figure 4; p < 0.05).
-The simulations that considered for 100% hydrodynamic drag travelled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
-Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations.
-
-```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE, fig.height=3, fig.width=4}
-
-pos <- c("RK4",
-         "Min_Kfloat_H100_W00",
-         "Min_Kfloat_H095_W05",
-         "Min_Kfloat_H090_W10",
-         "Min_Kfloat_H085_W15",
-         "Mean_Kfloat_H100_W00",
-         "Mean_Kfloat_H095_W05",
-         "Mean_Kfloat_H090_W10",
-         "Mean_Kfloat_H085_W15",
-         "Max_Kfloat_H100_W00",
-         "Max_Kfloat_H095_W05",
-         "Max_Kfloat_H090_W10",
-         "Max_Kfloat_H085_W15")
-
-dis_plot <- ggplot(data = total_data, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
-  stat_compare_means(method = "kruskal.test", label = "p.signif")+
-  theme_bw()+
-  ylab("Distance (km)")+
-  xlab("Scenario")+
-  #coord_flip() +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-custom_labs <- c("1","2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot
-
-dis_plot <- dis_plot +
-  scale_x_discrete(labels= custom_labs)
-
-```
-
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario.
-The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.
-The lower and upper hinges correspond to the first and third quartiles.
-The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
-dis_plot
-```
-
-## Accumulation
-
-```{r message=FALSE, echo=FALSE, warning=FALSE}
-
-kelp_acc <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-kelp_acc_mean <- kelp_acc %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-kelp_acc$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_mean$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-pass_acc <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-pass_acc_mean <- pass_acc %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-pass_acc$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_mean$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-total_acc_data <- rbind(pass_acc, kelp_acc)
-total_acc_mean <- rbind(pass_acc_mean, kelp_acc_mean)
-
-mean_traj_acc <- ggplot() +
-  geom_path(data = total_acc_mean, aes(x = lon, y = lat, colour = type)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-## Coastline of Southern Africa
-load("map/south_africa_coast_hi_res.Rdata")
-
-heatmap_RK4_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Passive particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_kelp_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Kelp particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_RK4_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = pass_acc, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_kelp_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Age")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_kelp_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  #scale_color_viridis_c(name = "Age") +
-  guides(fill=guide_legend(title="Age")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-dis_plot_acc <- ggplot(data = total_acc_data, aes(x = type, y = distance, show.legend = FALSE)) +
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
-  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = "right") +
-  theme_bw() +
-  ylab("Distance (km)") +
-  xlab("Scenario") +
-  #coord_flip() +
-  theme_bw()
-  #theme(plot.title = element_text(hjust = 1))
-
-```
-
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5: Comparison of trajectory data between passive and kelp accumulation scenarios.
-Panel A are a comparison of density plots  and panel B is a comparison of age.", fig.height=8, fig.width=8}
-
-plot_A <- ggarrange(heatmap_RK4_acc, heatmap_kelp_acc, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 1)
-plot_B <- ggarrange(point_RK4_acc, point_kelp_acc, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 1)
-
-ggarrange(plot_A, plot_B, labels = c("AUTO"), ncol = 1, nrow = 2)
-
-```
-
-The role of eddies that exist within the study domain over the time period play a clear role in the accumulation and dispersal of kelp particles.
-Passive particles tend to remain on the outer edge of the eddy and then are expelled on the other side and continue to follow the ocean current (see Figure 5A).
-The opposite is true for kelp particles which have the highest exposure level of hydrodynamic and wind drag.
-The kelp particles are entrained in the eddy (see Figure 5A).
-This is also reflected when comparing the age of particles at the end of the simulation.
-In Figure 5B, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
-
-## Comparison of coefficents
-
-The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (Figure 6A).
-This was also reflected when comparing distance travelled by particles and particle density plots (Figure 6B).
-The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (Figure 6C-D).
-
-```{r shape data, message=FALSE, echo=FALSE, warning=FALSE}
-
-sphere <- tidync("OceanParcels/trajectory_data/Max_KFloat_H100_W00.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-sphere_mean <- sphere %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-sphere$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
-sphere_mean$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-# read in data again for one max simulation, as cannot sort issue out with rename df...
-cyl <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-cyl_mean <- cyl %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-cyl$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
-cyl_mean$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-# Make mean trajectory plot
-shape_plot <- ggplot()+
-  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = sphere_mean, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = cyl_mean, aes(x = lon, y = lat), color = 'orange')+
-  xlab("Longtitude") + 
-  ylab("Latitude") +
-  ggtitle("Mean trajectory comparison") +
-  #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-# make density map
-heatmap_sphere <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = sphere, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = sphere_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("Cylinder") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_cyl <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = cyl, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = cyl_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("Sphere") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-shape_data_total <- rbind(sphere, cyl)
-
-shape_dis_plot <- ggplot(data = shape_data_total, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
-  stat_compare_means(method = "kruskal.test", label = "p.signif")+
-  ylab("Distance (km)")+
-  xlab("Type") + 
-  ggtitle("Comparison of distance travelled") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-```
-
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 6: Comparison of trajectory data between a kelp sphere and kelp cylinder.
-Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=8}
-
-ggarrange(shape_plot, shape_dis_plot, heatmap_sphere, heatmap_cyl, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 2)
-
-```
-
-# Discussion
-
-The treatment of drifting macroalgae as purely Lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @miron2020].
-Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], past research has focused on large macroalgal-rafts which can vary greatly in size and surface area.
-None of the past research has taken advantage of the advancement in numerical ocean models and Lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae.
-This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag.
-The mean virtual kelp trajectory deviated from the mean passive  trajectory when including varying levels of hydrodynamic and wind drag scenarios.
-The results from this study show the inclusion of hydrodynamic and wind drag results in observable entrainment patterns, and broadens our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case that of *E.
-maxima*.
-
-Past research has inferred from direct and indirect techniques that drifting macroalgae tend to follow the prevailing surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
-The results from this numerical experiment confirm that drifting macroalgae do follow the surface current in the study region when considering only hydrodynamic drag.
-In addition, not only are the trajectories similar but so are the end points at the end of simulation.
-However, when *ad-hoc* wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely Lagrangian particle trajectories.
-The inclusion of wind drag causes the kelp particles to flow further away from the coast and with a greater distance travelled from the release location.
-The difference in trajectory and distance covered can be attributed to the inclusion of wind drag.
-Wind drag adds to the effects of hydrodynamic drag when the wind direction and current direction are opposite to each other.
-However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object [@hackett2006; @putman2020].
-This causes the kelp particles to travel further as well as exposed to different time varying flows.
-Past studies have shown this when including windage as a function of inertia [@putman2018;@brooks2019; @putman2020].
-A study by @putman2020 used *ad-hoc* windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic *Sargassum sp*.
-The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked *Sargassum sp* mats which was partly due to the inclusion of momemtum energy transfer from wind.
-This is also reflected in the analysis comparing distances travelled by all the particles in each simulation.
-The particles in the simulations that included wind drag traveled significantly greater distances compared to the passive or hydrodynamic only drag.
-
-The addition of drag forces causes the kelp particles to cluster together along the mean trajectory compared to that of the passive particles which are more evenly dispersed.
-This is most likely due to the shared drag characteristics among the kelp particles which included the plants weight and surface area into the calculation.
-Other studies investigating flotsam trajectory characteristics have found similar results [@breivik2011; @olascoaga2020; @miron2020].
-A study by @miron2020 conducted field experiments using a range of objects such as spheres, cubes, and cuboids to compare to investigate the effect of inertia on particles dynamics.
-The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia.
-Although drag forces and not inertia were used in this study, the same conclusion applies; drag is therefore an important characteristic to consider when investigating or predicting solitary drifting macroalgae trajectory.
-
-The ocean is comprised of time-varying flows and changes in an objects velocity that can cause the object to be exposed to different flow patterns over time.
-The lack of dissimilarity in trajectory between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible.
-A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region.
-The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due to  as juveniles do are not swim fast enough to alter thier overall trajectory.
-Therefore, the differences in cross-sectional area types may not be significant enough to have an effect on the relevant zonal and meridonal velocity vectors.
-Another possible reason is the resolution of the ocean model used.
-The resolution of the underlying ocean model is an important aspect of Lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications.
-A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study.
-The authors found that the inclusion of Brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications.
-Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.
-
-The same is also true when comparing different hydrodynamic and wind drag exposure scenarios.
-These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag.
-The similarities between the different combined drag scenarios suggest that higher magnitudes of wind exposure do not significantly alter trajectories.
-Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns.
-All the simulations that included any form of drag resulted in a  high density of particles entrained in a eddy.
-The entrainment and expulsion from eddies and vorticities are an important characteristic of floating macroalgae.
-Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts.
-However, these studies investigated large rafts and not floating individuals.
-Inertia is based on the size/mass of the raft, while hydrodynamic and wind drag are based on current velocity and surface area.
-Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered.
-The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles, and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.
-
-The role of eddies and vortices in the entrainment of kelp is further supported by the accumulation simulations.
-These results cleary show that the inclusion of drag greatly increases the number of particles entrained in the eddy within the study domain.
-Eddies and vortices may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in area below the eddy.
-In a study by @woodborne1989 references a site offshore approximately 120km from Cape Columbine where dredged kelp was found.
-The area referenced by @woodborne1989 coincides with the area where virtual kelp are entrained by a vorticy.
-The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory.
-Instead, the inclusion of a different shape coefficient increased to the density of particles entrained in an eddy field.
-Therefore, more accurately parameterizing shape will therefore allow for further improvements in models of macroalgal dispersal patterns.
-
-# Conclusion
-
-Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape; however, no research exists to clarify the drifting characteristics of solitary macroalgae.
-Understanding how to accurately model the distribution of solitary kelp can lead to be understanding of accumulation zones and sinks and ultimately the ecological links pertaining to those processes.
-The findings from this study indicate that solitary floating virtual *E.
-maxima* particles tend to follow the prevailing surface currents and the inclusion of wind causes the trajectories to differ greatly from that of a purely Lagrangian particle.
-Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies.
-Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E.maxima*.
-In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory, but it does cause higher density of particles to become entrained in an eddy.
-Overall, the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not.
-This study also provides an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately.
-Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal as well as how these vary seasonally.
-The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates and raft-times.
+# Appendix
 
@@
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time) %>%
+  dplyr::select(trajectory, lon, lat, time, distance) %>%
   dplyr::rename(t = time) %>%
@@
 
-p_RK4_July$type <- "RK4_April" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_April$type <- "RK4_April" # add extra column 'type' to dataframe and assigned 'RK4' label
 
@@
 Min_Kfloat_H100_W00_April$type <- "Min_Kfloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H100_W00_April$velocity <- with(Min_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H100_W00_April$velocity <- with(Min_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 7: Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=8, fig.cap="Figure 7: Comparison of mean trajectories.
 Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
@@
   ylab("Latitude") + 
-  ggtitle("30") +
+  ggtitle("17") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time of each scenario for April 2018 (winter).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time for scenarios 17 - 29, for April 2018 (winter).
 The plots are regular hexagons within the density of particles is calculated.
-Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
-The bottom plot is the density plot for the passive simulation.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 
-figure_a_April <- ggarrange(heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"))
+figure_a_April <- ggarrange(heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April, labels = c("18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("none"))
 
@@
 ```
-
 ## July
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time of each scenario for July 2018 (winter).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time for scenarios 30 - 42, for July 2018 (winter).
 The plots are regular hexagons within the density of particles is calculated.
-Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
-The bottom plot is the density plot for the passive simulation.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 10: Comparison of density plots at the end run time of each scenario for October 2018 (spring).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 10: Comparison of density plots at the end run time for scenarios 43 - 55, for October 2018 (spring).
 The plots are regular hexagons within the density of particles is calculated.
-Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
-The bottom plot is the density plot for the passive simulation.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 ```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE}
+
+pos_April <- c("RK4_April",             ## April
+         "Min_Kfloat_H100_W00_April",
+         "Min_Kfloat_H095_W05_April",
+         "Min_Kfloat_H090_W10_April",
+         "Min_Kfloat_H085_W15_April",
+         "Mean_Kfloat_H100_W00_April",
+         "Mean_Kfloat_H095_W05_April",
+         "Mean_Kfloat_H090_W10_April",
+         "Mean_Kfloat_H085_W15_April",
+         "Max_Kfloat_H100_W00_April",
+         "Max_Kfloat_H095_W05_April",
+         "Max_Kfloat_H090_W10_April",
+         "Max_Kfloat_H085_W15_April")
+
+dis_plot_April <- ggplot(data = total_data_April, aes(x = type, y = distance, show.legend = FALSE))+
+  geom_boxplot(data = p_RK4_April, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(limits = pos)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif")+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Scenario")+
+  #coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+custom_labs_April <- c("17","18","19","20","21","22","23","24","25","26","27","28", "29") #create custom labels and add them to dis_plot
+
+dis_plot_April <- dis_plot_April +
+  scale_x_discrete(labels= custom_labs_April)
 
@@
 dis_plot_July <- ggplot(data = total_data_July, aes(x = type, y = distance, show.legend = FALSE))+
+  geom_boxplot(data = p_RK4_July, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
   geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
@@
 dis_plot_Oct <- ggplot(data = total_data_Oct, aes(x = type, y = distance, show.legend = FALSE))+
+  geom_boxplot(data = p_RK4_Oct, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
   geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
@@
 
-```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 11: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for April (Autumn), July (winter), and October (spring).
+```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8, fig.width=6, fig.cap="Figure 11: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for April (Autumn), July (winter), and October (spring).
 The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.
@@
 
-ggarrange(dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 2)
+ggarrange(dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 3)
 
 ```
-
-# References

### 2021-03-18 thesis_chapter_3.Rmd
moving appendix for chapter 3 to seperate rmd

--- 2021-03-17 thesis_chapter_3.Rmd
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time of each scenario.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time for scenarios 1 - 13, for January 2018 (summer).
 The plots are regular hexagons within the density of particles is calculated.
-Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
-The bottom plot is the density plot for the passive simulation.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 dis_plot <- ggplot(data = total_data, aes(x = type, y = distance, show.legend = FALSE))+
+  geom_boxplot(data = p_RK4, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
   geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario.
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for scenarios 1 - 13, for January 2018 (summer).
 The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.
@@
 
-# Appendix
-
-## Domain features
-
-```{r eval=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
-
-library(jpeg)
-library(grid)
-library(gridExtra)
-
-vector_2 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day2.jpg")))
-vector_5 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day5.jpg")))
-vector_10 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day10.jpg")))
-vector_15 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day15.jpg")))
-vector_20 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day20.jpg")))
-vector_25 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day25.jpg")))
-vector_30 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day30.jpg")))
-
-ggarrange(vector_2, vector_5, vector_10, vector_15, vector_20, vector_25, vector_30, labels = "AUTO")
-```
-
-## Additional comparisons
-
-```{r additional sims, message=FALSE, echo=FALSE, warning=FALSE}
-
-# Read in simulation data and data wrangling
-
-## July
-
-p_RK4_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-p_RK4_mean_July <- p_RK4_July %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-p_RK4_July$type <- "RK4_July" # add extra column 'type' to dataframe and assigned 'RK4' label
-
-Min_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H100_W00_mean_July <- Min_Kfloat_H100_W00_July %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H100_W00_July$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H100_W00_July$velocity <- with(Min_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Min_Kfloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H095_W05_mean_July <- Min_Kfloat_H095_W05_July %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H095_W05_July$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H095_W05_July$velocity <- with(Min_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Min_Kfloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H090_W10_mean_July <- Min_Kfloat_H090_W10_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H090_W10_July$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H090_W10_July$velocity <- with(Min_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Min_Kfloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H085_W15_mean_July <- Min_Kfloat_H085_W15_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H085_W15_July$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H085_W15_July$velocity <- with(Min_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_Kfloat_H100_W00_mean_July <- Mean_Kfloat_H100_W00_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_Kfloat_H100_W00_July$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_Kfloat_H100_W00_July$velocity <- with(Mean_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_KFloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H095_W05_mean_July <- Mean_KFloat_H095_W05_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H095_W05_July$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H095_W05_July$velocity <- with(Mean_KFloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_KFloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H090_W10_mean_July <- Mean_KFloat_H090_W10_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H090_W10_July$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H090_W10_July$velocity <- with(Mean_KFloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_KFloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H085_W15_mean_July <- Mean_KFloat_H085_W15_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H085_W15_July$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H085_W15_July$velocity <- with(Mean_KFloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H100_W00_mean_July <- Max_Kfloat_H100_W00_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H100_W00_July$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H100_W00_July$velocity <- with(Max_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H095_W05_mean_July <- Max_Kfloat_H095_W05_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H095_W05_July$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H095_W05_July$velocity <- with(Max_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H090_W10_mean_July <- Max_Kfloat_H090_W10_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H090_W10_July$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H090_W10_July$velocity <- with(Max_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H085_W15_mean_July <- Max_Kfloat_H085_W15_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H085_W15_July$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H085_W15_July$velocity <- with(Max_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
-
-total_data_July <- rbind(Min_Kfloat_H100_W00_July, Min_Kfloat_H095_W05_July, Min_Kfloat_H090_W10_July, Min_Kfloat_H085_W15_July, Mean_Kfloat_H100_W00_July, Mean_KFloat_H095_W05_July, Mean_KFloat_H090_W10_July, Mean_KFloat_H085_W15_July, Max_Kfloat_H100_W00_July, Max_Kfloat_H095_W05_July, Max_Kfloat_H090_W10_July, Max_Kfloat_H085_W15_July)
-
-## October
-
-p_RK4_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-p_RK4_mean_Oct <- p_RK4_Oct %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-p_RK4_Oct$type <- "RK4_Oct" # add extra column 'type' to dataframe and assigned 'RK4' label
-
-Min_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H100_W00_mean_Oct <- Min_Kfloat_H100_W00_Oct %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H100_W00_Oct$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Min_Kfloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H095_W05_mean_Oct <- Min_Kfloat_H095_W05_Oct %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H095_W05_Oct$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Min_Kfloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H090_W10_mean_Oct <- Min_Kfloat_H090_W10_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H090_W10_Oct$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Min_Kfloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H085_W15_mean_Oct <- Min_Kfloat_H085_W15_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H085_W15_Oct$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_Kfloat_H100_W00_mean_Oct <- Mean_Kfloat_H100_W00_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_Kfloat_H100_W00_Oct$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_KFloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H095_W05_mean_Oct <- Mean_KFloat_H095_W05_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H095_W05_Oct$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_KFloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H090_W10_mean_Oct <- Mean_KFloat_H090_W10_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H090_W10_Oct$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Mean_KFloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H085_W15_mean_Oct <- Mean_KFloat_H085_W15_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H085_W15_Oct$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H100_W00_mean_Oct <- Max_Kfloat_H100_W00_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H100_W00_Oct$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H095_W05_mean_Oct <- Max_Kfloat_H095_W05_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H095_W05_Oct$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H090_W10_mean_Oct <- Max_Kfloat_H090_W10_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H090_W10_Oct$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-Max_Kfloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H085_W15_mean_Oct <- Max_Kfloat_H085_W15_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H085_W15_Oct$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-
-total_data_Oct <- rbind(Min_Kfloat_H100_W00_Oct, Min_Kfloat_H095_W05_Oct, Min_Kfloat_H090_W10_Oct, Min_Kfloat_H085_W15_Oct, Mean_Kfloat_H100_W00_Oct, Mean_KFloat_H095_W05_Oct, Mean_KFloat_H090_W10_Oct, Mean_KFloat_H085_W15_Oct, Max_Kfloat_H100_W00_Oct, Max_Kfloat_H095_W05_Oct, Max_Kfloat_H090_W10_Oct, Max_Kfloat_H085_W15_Oct)
-
-```
-
-### Comparison of mean trajectories
-
-```{r, message=FALSE, echo=FALSE, warning=FALSE}
-
-waterdrag_plot_July <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'red')+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-wwdrag_plot_July <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Min_Kfloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'grey')+
-  geom_path(data = Min_Kfloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'green')+
-  geom_path(data = Min_Kfloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'pink')+
-  geom_path(data = Mean_KFloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'yellow')+
-  geom_path(data = Mean_KFloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'red')+
-  geom_path(data = Max_Kfloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Max_Kfloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'red')+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-waterdrag_plot_Oct <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'red')+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-wwdrag_plot_Oct <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Min_Kfloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'grey')+
-  geom_path(data = Min_Kfloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'green')+
-  geom_path(data = Min_Kfloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'pink')+
-  geom_path(data = Mean_KFloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'yellow')+
-  geom_path(data = Mean_KFloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Mean_KFloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'red')+
-  geom_path(data = Max_Kfloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Max_Kfloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'red')+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 7: Comparison of mean trajectories.
-Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
-In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10.
-In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
-
-ggarrange(waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, labels = "AUTO")
-
-```
-
-### Comparison of density
-
-```{r density plots additional, message=FALSE, warning=FALSE, echo=FALSE}
-## Heatmaps of study domain, use previous data for study site map and histograms## Coastline of Southern Africa
-load("map/south_africa_coast_hi_res.Rdata")
-
-## July
-
-heatmap_RK4_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = p_RK4_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("30") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H100_W00_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H095_W05_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H090_W10_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H085_W15_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_Kfloat_H100_W00_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
-  geom_path(data = Mean_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_KFloat_H095_W05_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
-  geom_path(data = Mean_KFloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_KFloat_H090_W10_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Mean_KFloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_KFloat_H085_W15_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H100_W00_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H100_W00_mean_July ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H095_W05_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  theme(plot.title = element_text(hjust = 0.5)) + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H090_W10_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H085_W15_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-## October
-
-heatmap_RK4_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = p_RK4_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("43") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H100_W00_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H095_W05_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H090_W10_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Min_Kfloat_H085_W15_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
-  geom_path(data = Min_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_Kfloat_H100_W00_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
-  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_KFloat_H095_W05_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
-  geom_path(data = Mean_KFloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_KFloat_H090_W10 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Mean_KFloat_H085_W15_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Mean_KFloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H100_W00_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H095_W05_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  theme(plot.title = element_text(hjust = 0.5)) + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H090_W10_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_Max_KFloat_H085_W15_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = Max_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-```
-
-## July
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time of each scenario for July 2018 (winter).
-The plots are regular hexagons within the density of particles is calculated.
-Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
-The bottom plot is the density plot for the passive simulation.
-The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
-
-figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"))
-
-figure_b_July <- ggarrange(heatmap_RK4_July, legend = c("right"))
-
-ggarrange(figure_a_July, figure_b_July, ncol = 1, nrow = 2)
-
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time of each scenario for October 2018 (spring).
-The plots are regular hexagons within the density of particles is calculated.
-Plots A-D represents the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas.
-The bottom plot is the density plot for the passive simulation.
-The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
-
-figure_a_Oct <- ggarrange(heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("none"))
-
-figure_b_Oct <- ggarrange(heatmap_RK4_Oct, legend = c("right"))
-
-ggarrange(figure_a_Oct, figure_b_Oct, ncol = 1, nrow = 2)
-
-```
-
-## Comparison of distances
-
-```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE}
-
-pos_July <- c("RK4_July",             ## July
-         "Min_Kfloat_H100_W00_July",
-         "Min_Kfloat_H095_W05_July",
-         "Min_Kfloat_H090_W10_July",
-         "Min_Kfloat_H085_W15_July",
-         "Mean_Kfloat_H100_W00_July",
-         "Mean_Kfloat_H095_W05_July",
-         "Mean_Kfloat_H090_W10_July",
-         "Mean_Kfloat_H085_W15_July",
-         "Max_Kfloat_H100_W00_July",
-         "Max_Kfloat_H095_W05_July",
-         "Max_Kfloat_H090_W10_July",
-         "Max_Kfloat_H085_W15_July")
-
-dis_plot_July <- ggplot(data = total_data_July, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
-  stat_compare_means(method = "kruskal.test", label = "p.signif")+
-  theme_bw()+
-  ylab("Distance (km)")+
-  xlab("Scenario")+
-  #coord_flip() +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-custom_labs_July <- c("30","31","32","33","34","35","36","37","38","39","40","41", "42") #create custom labels and add them to dis_plot
-
-dis_plot_July <- dis_plot_July +
-  scale_x_discrete(labels= custom_labs_July)
-
-pos_Oct <- c("RK4_Oct",              ## Oct
-         "Min_Kfloat_H100_W00_Oct",
-         "Min_Kfloat_H095_W05_Oct",
-         "Min_Kfloat_H090_W10_Oct",
-         "Min_Kfloat_H085_W15_Oct",
-         "Mean_Kfloat_H100_W00_Oct",
-         "Mean_Kfloat_H095_W05_Oct",
-         "Mean_Kfloat_H090_W10_Oct",
-         "Mean_Kfloat_H085_W15_Oct",
-         "Max_Kfloat_H100_W00_Oct",
-         "Max_Kfloat_H095_W05_Oct",
-         "Max_Kfloat_H090_W10_Oct",
-         "Max_Kfloat_H085_W15_Oct")
-
-dis_plot_Oct <- ggplot(data = total_data_Oct, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
-  stat_compare_means(method = "kruskal.test", label = "p.signif")+
-  theme_bw()+
-  ylab("Distance (km)")+
-  xlab("Scenario")+
-  #coord_flip() +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-custom_labs_Oct <- c("43","44","45","46","47","48","49","50","51","52","53","54","55") #create custom labels and add them to dis_plot
-
-dis_plot_Oct <- dis_plot_Oct +
-  scale_x_discrete(labels= custom_labs_Oct)
-
-```
-
-```{r distance plot additional, include = FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 10: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring).
-The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.
-The lower and upper hinges correspond to the first and third quartiles.
-The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
-
-ggarrange(dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 2)
-
-```
-
 # References

### 2021-03-18 appendix_chapter_3.Rmd
additional updates

@@
   ylab("Latitude") + 
-  ggtitle("17") +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") + 
-  ggtitle("30") +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") + 
-  ggtitle("43") +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-figure_a_April <- ggarrange(heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April, labels = c("18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("none"))
-
-figure_b_April <- ggarrange(heatmap_RK4_April, legend = c("right"))
-
-ggarrange(figure_a_April, figure_b_April, ncol = 1, nrow = 2)
+figure_a_April <- ggarrange(heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April, labels = c("18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
+
+figure_b_April <- ggarrange(heatmap_RK4_April, legend = c("right"), labels = c("17"), font.label = list(size = 11, face = "bold"))
+
+ggarrange(figure_b_April, figure_a_April, ncol = 1, nrow = 2)
 
@@
 
-figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"))
-
-figure_b_July <- ggarrange(heatmap_RK4_July, legend = c("right"))
-
-ggarrange(figure_a_July, figure_b_July, ncol = 1, nrow = 2)
+figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
+
+figure_b_July <- ggarrange(heatmap_RK4_July, legend = c("right"), labels = c("30"), font.label = list(size = 11, face = "bold"))
+
+ggarrange(figure_b_July, figure_a_July, ncol = 1, nrow = 2)
 
@@
 
-figure_a_Oct <- ggarrange(heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("none"))
-
-figure_b_Oct <- ggarrange(heatmap_RK4_Oct, legend = c("right"))
-
-ggarrange(figure_a_Oct, figure_b_Oct, ncol = 1, nrow = 2)
+figure_a_Oct <- ggarrange(heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
+
+figure_b_Oct <- ggarrange(heatmap_RK4_Oct, legend = c("right"), labels = c("43"), font.label = list(size = 11, face = "bold"))
+
+ggarrange(figure_b_Oct, figure_a_Oct, ncol = 1, nrow = 2)
 
@@
 
-```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE}
-
-pos_April <- c("RK4_April",             ## April
+```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+
+pos_April <- c("RK4_April",             
          "Min_Kfloat_H100_W00_April",
@@
   #scale_x_discrete(limits = pos)+
-  stat_compare_means(method = "kruskal.test", label = "p.signif")+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
   theme_bw()+
@@
 
-pos_July <- c("RK4_July",             ## July
+pos_July <- c("RK4_July",             
          "Min_Kfloat_H100_W00_July",
@@
 
-pos_Oct <- c("RK4_Oct",              ## Oct
+pos_Oct <- c("RK4_Oct",              
          "Min_Kfloat_H100_W00_Oct",

### 2021-03-18 thesis_chapter_3.Rmd
additional updates

@@
 The resulting eddies and filaments from the Agulhas retroflection and study area are presented.
-In panel B, the splitting of the Benguela Jet, eddie features and general flow of the study domain are presented."}
+In panel B, the splitting of the Benguela Jet, eddie features and general flow of the study domain are presented.
+The release site is represented by a blue star."}
 
@@
 
-```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Details for each simulation run for the study.
+```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=" Details for each simulation run for the study.
 The table cosists of parameterisations for each scenario.
@@
 
-```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.
+```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap=" Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.
 Approximate shape refers to the dimensions of the geometric shape used to represent a particular morpholgical characteristic.
@@
   ylab("Latitude") + 
-  ggtitle("1") +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-figure_a <- ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, labels = c("2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"))
-
-figure_b <- ggarrange(heatmap_RK4, legend = c("right"))
-
-ggarrange(figure_a, figure_b, ncol = 1, nrow = 2)
+figure_a <- ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, labels = c("2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
+
+figure_b <- ggarrange(heatmap_RK4, legend = c("right"), labels = c("1"), font.label = list(size = 11, face = "bold"))
+
+ggarrange(figure_b, figure_a, ncol = 1, nrow = 2)
 

### 2021-03-18 appendix_chapter_3.Rmd
added some analysis

@@
 ---
-```{r software, message=FALSE, warning=FALSE, include=FALSE}
+```{r software, message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
 ## Load software need for entire script
@@
 
-```{r eval=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
+```{r eval=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, eval=FALSE}
 
@@
 
+The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (autumn).
+Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag travelled further northward compared to the passive particles (see Figure 7A-B).
+The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag travelled further north-westard compared to the passive particles.
+Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles travelled less distance compared to kelp particles with both wind and hydrodynamic drag.
+Differences in mean trajectories between kelp and passive particles was also evident in the July (winter) simulations (see Figure 7C-D).
+Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially travel in a southerly direction and then moved westward offshore; resulting in entrainment within an eddy.
+Similar behaviour is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag.
+In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles travelled initially southward and then westward offshore and also becoming entrained in an eddy.
+In the October (Spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 7E-F).
+Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag ended a greater distance from the release site compared to the passive particles.
+This was also evident when comparing kelp particles with different cross-sectional areas and exposure to wind and hydrodynamic drag with passive particles; with the end location a further from the release location.
+In both the April and October simulations all particles flowed north-westward, while in winter the mean trajectory resulted in entrainment within an eddy.
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=8, fig.cap="Figure 7: Comparison of mean trajectories.
@@
 ```
-In the April simulations, both passive and kelp particles followed the general north-westward flow of the region with clear entrainment of kelp particles in sub-mesoscale features.
-Kelp particles follow the north
 
@@
 
-## April
+In April and across all simulations, particles flowed in a north-westward direction (see Figure 8).
+The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation.
+In addition, simulatons that considered any form of drag got entrained within an eddy further north-westward from end location of passive particles.
+Only slight variations across cross-sectional areas and drag exposure scenarios are evident.
+Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in a eddy located westward of the release site compared to passive particles (see Figure 9).
+Variation in density along the mean trajectory and eddy for all kelp particles is evident.
+Kelp particles with only hydrodynamic drag across cross-sectional areas showed lower density of particles along the mean trajectory compared to kelp particles with both wind and hydrodynamic drag.
+October simulations showed no differences in density of particles along the mean trajectory when comparing passive and kelp particles.
+Only slight variations in density and trajectory can be seen with no entrainment (see Figure 10).
+Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.
 
@@
 ```
-## July
 
@@
 ```
-
-## October
 
@@
 
-```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8, fig.width=6, fig.cap="Figure 11: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for April (Autumn), July (winter), and October (spring).
+```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.height=6, fig.width=5, fig.cap="Figure 11: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for April (Autumn), July (winter), and October (spring).
 The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.

### 2021-03-18 appendix_chapter_3.Rmd
appendix done

@@
 
+When comparing distances between passive and kelp particles for April, significant differences are evident.
+(Figure 11, simulations 17-29; p < 0.05).
+Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas travelled less distance across compared to other kelp particles.
+The lowest distance travelled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
+No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; p > 0.05).
+In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; p < 0.05).
+Kelp particles with only 90% hydrodynamic drag across cross-sectional areas travelled less distance compared to all other particles.
+As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure travelled the least distance compared to all other particles.
+
 ```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.height=6, fig.width=5, fig.cap="Figure 11: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for April (Autumn), July (winter), and October (spring).

### 2021-03-19 appendix_chapter_3.Rmd
updates to rmd files

@@
 ---
-```{r software, message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
+```{r software, message=FALSE, warning=FALSE, include=FALSE, eval=TRUE}
 ## Load software need for entire script
@@
 Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag travelled further northward compared to the passive particles (see Figure 7A-B).
-The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag travelled further north-westard compared to the passive particles.
+The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles.
 Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles travelled less distance compared to kelp particles with both wind and hydrodynamic drag.
-Differences in mean trajectories between kelp and passive particles was also evident in the July (winter) simulations (see Figure 7C-D).
-Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially travel in a southerly direction and then moved westward offshore; resulting in entrainment within an eddy.
+Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 7C-D).
+Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore; resulting in entrainment within an eddy.
 Similar behaviour is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=8, fig.cap="Figure 7: Comparison of mean trajectories.
-Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
-In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10.
-In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=8, fig.cap="Figure 7: Comparison of mean trajectories for only hydrodynamic drag and mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
+Panel (A) represents hydrodynamic drag only simulations and Panel (B) represents simulations for both wind and hydrodynamic drag, for April.
+Panel (C) represents hydrodynamic drag only simulations and Panel (D) represents simulations for both wind and hydrodynamic drag, for July.
+Panel (E) represents hydrodynamic drag only simulations and Panel (F) represents simulations for both wind and hydrodynamic drag, for October.
+In all panels the passive simulations are represented by black lines."}
 
@@
 
-### Comparison of density
+### Comparison of density distributions
 
@@
 The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation.
-In addition, simulatons that considered any form of drag got entrained within an eddy further north-westward from end location of passive particles.
+In addition, simulations that considered any form of drag got entrained within an eddy further north-westward from the end location of passive particles.
 Only slight variations across cross-sectional areas and drag exposure scenarios are evident.
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time for scenarios 17 - 29, for April 2018 (winter).
-The plots are regular hexagons within the density of particles is calculated.
+The plots are regular hexagons within the density of particles was calculated.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time for scenarios 30 - 42, for July 2018 (winter).
-The plots are regular hexagons within the density of particles is calculated.
+The plots are regular hexagons within the density of particles was calculated.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 10: Comparison of density plots at the end run time for scenarios 43 - 55, for October 2018 (spring).
-The plots are regular hexagons within the density of particles is calculated.
+The plots are regular hexagons within the density of particles was calculated.
 The mean trajectories for each scenario are also plotted for comparison.

### 2021-03-19 thesis_chapter_3.Rmd
updates to rmd files

@@
 Past studies also focus on kelp-rafts, which can vary greatly in size making it difficult to form a concensus as to the relative role of wind versus surface currents.
-In addition, past studies have not conisidered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam.
-The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectory of flotsam in the ocean.
+In addition, past studies have not considered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam.
+The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectories of flotsam in the ocean.
 This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagae trajectory and accumulation.
@@
 The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain.
-The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumlation zones for solitary floating macroalage."
+The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumulation zones for solitary floating macroalage."
 ---
@@
 It should be noted that this study does not attempt to characterise the dispersal of drifting macroalage within the study domain.
-The authors argue that too little is know about the drift characteristics of solitary floating macroalgae to characterise dispersal effectivley.
+The authors argue that too little is know about the drift characteristics of solitary floating macroalgae to characterise dispersal effectively.
 Instead, this study will aim to shed light on the 'behaviour' of surface drifting macroalgae from a coastal environment into the open ocean, which will inform future characterisation studies.
@@
 These occur mostly in the form of eddies, filaments and vortices which are generated from the Agulhas retroflection [@veitch2018; @veitch2017], generally flowing northward along the coastline and begin to move offshore near Cape Columbine.
-The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, eckman transport and geostrophic currents [@stramma1989; @veitch2017; @veitch2018; @hutchings2009].
+The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, Eckman transport and geostrophic currents [@stramma1989; @veitch2017; @veitch2018; @hutchings2009].
 The stability of the jet varies seasonally with more stable flows in winter and splitting of the jet during summer and spring.
@@
 These winds are most intense during summer (December-March) which favours upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet; which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
-In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
+In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
 
@@
 Each scenario consisted of a particle set with a size of 1000 where released from Kommetjie at the start of simulation.
-The particles sets were advected with a Runge-Kutte fourth order scheme for 30 days and incorporated brownian motion.
+The particle sets were advected with a Runge-Kutte fourth order scheme for 30 days and incorporated brownian motion.
 Since the aim of this investigation was to investigate the relative surface behaviour of virtual kelp particles, only the surface layer was used for simulations.
@@
 The time period chosen to run the simulations was 01-01-2018 un 30-01-2018.
-This particular time period was chosen as the splitting of the Benguela Jet occurs in Janurary, which also conincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity [@rubio2009; @veitch2018; @ragoasha2019].
-
-```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=" Details for each simulation run for the study.
-The table cosists of parameterisations for each scenario.
-The ID coloumn refers to the syntax used in the relevant OceanParcels script (see additional material for link to script).
-Type refers to either a passive particle with no drag behaviour or a kelp particle that has been advected with any drag exposure.
-The drag exposure column refers to the parameterisations regarding the level of water and wind drag simulated."}
+This particular time period was chosen as the splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity [@rubio2009; @veitch2018; @ragoasha2019].
+A summary of the parameters for each simulation are provided in table 1.
+
+```{r sim table, echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
-sim_table <- set_caption(sim_table, "Table 1:Details for each simulation executed for the study.")
+sim_table <- set_caption(sim_table, "Table 1: Details for each simulation run for the study.
+Type refers to either a passive particle with no drag behaviour or a kelp particle that has been advected with any drag exposure.
+The drag exposure column refers to the parameterisations regarding the level of water and wind drag simulated.")
 sim_table
@@
 In order to determine if the shape co-efficient used plays a role in model used, a simulation of a kelp with the maximum surface area and drag exposure was run using the shape coefficient for a cylinder to compare to a simulation using a sphere, and was run for 30 days using a fourth order Runge-Kutte advection scheme with brownian motion and a particle set of size 1000.
+
+Additional simulations were done for months in 2018, which were chosen based on seasonality.
+These simulations had the same parameters used for January and were performed for April (Autumn), July (winter), and October (Spring) in order to assess if any patterns seen in January translate into general patterns for times periods throughout 2018.
 
@@
 The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
-Both averaged ocean current and wind data are stored as meridonal and zonal velocity vectors.
+Both averaged ocean current and wind data are stored as meridional and zonal velocity vectors.
 
@@
 The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
-Hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors, as per @furnans2008.
-Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step.
+Hydrodynamic and wind drag components were calculated separately for both the meridional and zonal velocity vectors, as per @furnans2008.
+Since drag force is dependent on the velocity vectors which vary with time, the meridional and zonal velocities were interpolated and used in the drag force calculation for each time-step.
 The same approach was used for the wind drag force.
@@
 The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
-The minimum, mean and maximum were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (see table 1).
+The minimum, mean and maximum were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (see table 2).
 The trajectory of these 'types' of cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
@@
 
-```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap=" Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.
-Approximate shape refers to the dimensions of the geometric shape used to represent a particular morpholgical characteristic.
-Equations used are the equations relating to the dimensions used for the geometric shape estimate.
-Plant dimensions are the measurements used, collected in previous work by @coppin2020, used to calculate the associated surface areas.
-The total surface areas for each plant characteristic for each surface area type make up the remaining columns."}
+```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
 
@@
 area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
-area_table <- set_caption(area_table, "Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.")
+area_table <- set_caption(area_table, "Table 2: Table summarising the attributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.
+Approximate shape refers to the dimensions of the geometric shape used to represent a particular morphological characteristic.
+Equations used are the equations relating to the dimensions used for the geometric shape estimate.
+Plant dimensions are the measurements used, collected in previous work by @coppin2020, used to calculate the associated surface areas.
+The total surface areas for each plant characteristic for each surface area type make up the remaining columns.")
 area_table
@@
 The distance that particles traveled was also measured and used to produce boxplots.
-Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@VanSebille2018].
+Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particles [@vanSebille2018].
 If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; @putman2018; @vanSebille2018].
@@
 In all simulations particles flowed in a north-westward direction.
+The patterns of north-westward flow and entrainment in eddies is also evident in the additional simulations, however clear differences in mean trajectory between kelp and passive particles is clear (see appendix).
 
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time for scenarios 1 - 13, for January 2018 (summer).
-The plots are regular hexagons within the density of particles is calculated.
+The plots are regular hexagons within the density of particles was calculated.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 ```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 6: Comparison of trajectory data between a kelp sphere and kelp cylinder.
-Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.
+Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylindrical shape coefficient.
 The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=8}
@@
 A study by @putman2020 used *ad-hoc* windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic *Sargassum sp*.
-The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked *Sargassum sp* mats which was partly due to the inclusion of momemtum energy transfer from wind.
+The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked *Sargassum sp* mats which was partly due to the inclusion of momentum energy transfer from wind.
 This is also reflected in the analysis comparing distances travelled by all the particles in each simulation.
 The particles in the simulations that included wind drag traveled significantly greater distances compared to the passive or hydrodynamic only drag.
+Similar patterns are seen in the additional simulations run for April (autumn), July (winter), and October (spring), however even kelp particles with only hydrodynamic drag flow further compared to passive particles in these times of the year.
+This may be a result of the kelp particles exposed to different time-varying flows, which results in the kelp particles following a different surface current vector compared to passive particles.
+This also results in different end-points and therefore these results further confirm that drag is an important factor to account for when investigating macroalgal drift patterns.
 
@@
 A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region.
-The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due to  as juveniles do are not swim fast enough to alter thier overall trajectory.
-Therefore, the differences in cross-sectional area types may not be significant enough to have an effect on the relevant zonal and meridonal velocity vectors.
+The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due to  as juveniles do are not swim fast enough to alter their overall trajectory.
+Therefore, the differences in cross-sectional area types may not be significant enough to have an effect on the relevant zonal and meridional velocity vectors.
 Another possible reason is the resolution of the ocean model used.
@@
 All the simulations that included any form of drag resulted in a  high density of particles entrained in a eddy.
-The entrainment and expulsion from eddies and vorticities are an important characteristic of floating macroalgae.
+The entrainment and expulsion from eddies and vortices are an important characteristic of floating macroalgae.
 Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts.
@@
 The role of eddies and vortices in the entrainment of kelp is further supported by the accumulation simulations.
-These results cleary show that the inclusion of drag greatly increases the number of particles entrained in the eddy within the study domain.
+These results clearly show that the inclusion of drag greatly increases the number of particles entrained in the eddy within the study domain.
 Eddies and vortices may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in area below the eddy.
 In a study by @woodborne1989 references a site offshore approximately 120km from Cape Columbine where dredged kelp was found.
-The area referenced by @woodborne1989 coincides with the area where virtual kelp are entrained by a vorticy.
+The area referenced by @woodborne1989 coincides with the area where virtual kelp are entrained by a eddy.
 The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory.

### 2021-03-19 thesis_chapter_3.Rmd
title updated

@@
 ---
-title: 'A numerical experiment investigating the influence of drag in trajectory patterns of floating macroalgae'
+title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
 author: 

### 2021-03-23 appendix_chapter_3.Rmd
changed mean traj plots

@@
 # Appendix
-
-## Domain features
-
-```{r eval=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, eval=FALSE}
-
-library(jpeg)
-library(grid)
-library(gridExtra)
-
-vector_2 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day2.jpg")))
-vector_5 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day5.jpg")))
-vector_10 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day10.jpg")))
-vector_15 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day15.jpg")))
-vector_20 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day20.jpg")))
-vector_25 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day25.jpg")))
-vector_30 <- rasterGrob(as.raster(readJPEG("OceanParcels/trajectory_data/vector_outputs/jpeg/vector_day30.jpg")))
-
-ggarrange(vector_2, vector_5, vector_10, vector_15, vector_20, vector_25, vector_30, labels = "AUTO")
-```
-
 ## Additional comparisons
@@
 p_RK4_April$type <- "RK4_April" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_mean_April$type <- "17" # add extra column 'type' to dataframe and assigned 'RK4' label
 
@@
 Min_Kfloat_H100_W00_April$velocity <- with(Min_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H100_W00_mean_April$type <- "18" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H095_W05_April$velocity <- with(Min_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H095_W05_mean_April$type <- "19" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H090_W10_April$velocity <- with(Min_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H090_W10_mean_April$type <- "20" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H085_W15_April$velocity <- with(Min_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H085_W15_mean_April$type <- "21" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_Kfloat_H100_W00_April$velocity <- with(Mean_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_Kfloat_H100_W00_mean_April$type <- "22" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H095_W05_April$velocity <- with(Mean_KFloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H095_W05_mean_April$type <- "23" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H090_W10_April$velocity <- with(Mean_KFloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H090_W10_mean_April$type <- "24" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H085_W15_April$velocity <- with(Mean_KFloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H085_W15_mean_April$type <- "25" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H100_W00_April$velocity <- with(Max_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H100_W00_mean_April$type <- "26" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H095_W05_April$velocity <- with(Max_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H095_W05_mean_April$type <- "27" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H090_W10_April$velocity <- with(Max_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H090_W10_mean_April$type <- "28" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H085_W15_April$velocity <- with(Max_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H085_W15_mean_April$type <- "29" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 p_RK4_July$type <- "RK4_July" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_mean_July$type <- "30" # add extra column 'type' to dataframe and assigned 'RK4' label
 
@@
 Min_Kfloat_H100_W00_July$velocity <- with(Min_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H100_W00_mean_July$type <- "31" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H095_W05_July$velocity <- with(Min_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H095_W05_mean_July$type <- "32" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H090_W10_July$velocity <- with(Min_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H090_W10_mean_July$type <- "33" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H085_W15_July$velocity <- with(Min_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H085_W15_mean_July$type <- "34" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_Kfloat_H100_W00_July$velocity <- with(Mean_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_Kfloat_H100_W00_mean_July$type <- "35" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H095_W05_July$velocity <- with(Mean_KFloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H095_W05_mean_July$type <- "36" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H090_W10_July$velocity <- with(Mean_KFloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H090_W10_mean_July$type <- "37" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H085_W15_July$velocity <- with(Mean_KFloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H085_W15_mean_July$type <- "38" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H100_W00_July$velocity <- with(Max_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H100_W00_mean_July$type <- "39" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H095_W05_July$velocity <- with(Max_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H095_W05_mean_July$type <- "40" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H090_W10_July$velocity <- with(Max_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H090_W10_mean_July$type <- "41" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H085_W15_July$velocity <- with(Max_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H085_W15_mean_July$type <- "42" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 p_RK4_Oct$type <- "RK4_Oct" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_mean_Oct$type <- "43" # add extra column 'type' to dataframe and assigned 'RK4' label
 
@@
 Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H100_W00_mean_Oct$type <- "44" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H095_W05_mean_Oct$type <- "45" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H090_W10_mean_Oct$type <- "46" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H085_W15_mean_Oct$type <- "47" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_Kfloat_H100_W00_mean_Oct$type <- "48" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H095_W05_mean_Oct$type <- "49" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H090_W10_mean_Oct$type <- "50" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H085_W15_mean_Oct$type <- "51" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H100_W00_mean_Oct$type <- "52" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H095_W05_mean_Oct$type <- "53" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H090_W10_mean_Oct$type <- "54" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H085_W15_mean_Oct$type <- "55" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 ```{r, message=FALSE, echo=FALSE, warning=FALSE}
+
+total_mean_water_April <- rbind(p_RK4_mean_April, Min_Kfloat_H100_W00_mean_April, Mean_Kfloat_H100_W00_mean_April, Max_Kfloat_H100_W00_mean_April)
+total_mean_ww_April <- rbind(p_RK4_mean_April, Min_Kfloat_H095_W05_mean_April, Min_Kfloat_H090_W10_mean_April, Min_Kfloat_H085_W15_mean_April, Mean_KFloat_H095_W05_mean_April, Mean_KFloat_H090_W10_mean_April, Mean_KFloat_H085_W15_mean_April, Max_Kfloat_H095_W05_mean_April, Max_Kfloat_H090_W10_mean_April, Max_Kfloat_H085_W15_mean_April)
+
+total_mean_water_April$type <- factor(total_mean_water_April$type, levels=c("17", "18", "22", "26"))
+total_mean_ww_April$type <- factor(total_mean_ww_April$type, levels=c("17", "19", "20", "21", "23", "24", "25", "27", "28", "29"))
+
 waterdrag_plot_April <- ggplot()+
   #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean_April, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean_April, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H100_W00_mean_April, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = total_mean_water_April, aes(x = lon, y = lat, colour = type))+
   xlab("Longtitude") + 
@@
   ggtitle("") +
+  scale_colour_discrete("Type") +  
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat), color = 'grey')+
-  geom_path(data = Min_Kfloat_H090_W10_mean_April, aes(x = lon, y = lat), color = 'green')+
-  geom_path(data = Min_Kfloat_H085_W15_mean_April, aes(x = lon, y = lat), color = 'pink')+
-  geom_path(data = Mean_KFloat_H095_W05_mean_April, aes(x = lon, y = lat), color = 'yellow')+
-  geom_path(data = Mean_KFloat_H090_W10_mean_April, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat), color = 'red')+
-  geom_path(data = Max_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Max_Kfloat_H090_W10_mean_April, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H085_W15_mean_April, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, colour = type))+
   xlab("Longtitude") + 
@@
 
+total_mean_water_July <- rbind(p_RK4_mean_July, Min_Kfloat_H100_W00_mean_July, Mean_Kfloat_H100_W00_mean_July, Max_Kfloat_H100_W00_mean_July)
+total_mean_ww_July <- rbind(p_RK4_mean_July, Min_Kfloat_H095_W05_mean_July, Min_Kfloat_H090_W10_mean_July, Min_Kfloat_H085_W15_mean_July, Mean_KFloat_H095_W05_mean_July, Mean_KFloat_H090_W10_mean_July, Mean_KFloat_H085_W15_mean_July, Max_Kfloat_H095_W05_mean_July, Max_Kfloat_H090_W10_mean_July, Max_Kfloat_H085_W15_mean_July)
+
+total_mean_water_July$type <- factor(total_mean_water_July$type, levels=c("30", "31", "35", "39"))
+total_mean_ww_July$type <- factor(total_mean_ww_July$type, levels=c("30", "32", "33", "34", "36", "37", "38", "40", "41", "42"))
+
 waterdrag_plot_July <- ggplot()+
   #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H100_W00_mean_July, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = total_mean_water_July, aes(x = lon, y = lat, colour = type))+
   xlab("Longtitude") + 
@@
   #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Min_Kfloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'grey')+
-  geom_path(data = Min_Kfloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'green')+
-  geom_path(data = Min_Kfloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'pink')+
-  geom_path(data = Mean_KFloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'yellow')+
-  geom_path(data = Mean_KFloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'red')+
-  geom_path(data = Max_Kfloat_H095_W05_mean_July, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Max_Kfloat_H090_W10_mean_July, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H085_W15_mean_July, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, colour = type))+
   xlab("Longtitude") + 
@@
 
+total_mean_water_Oct <- rbind(p_RK4_mean_Oct, Min_Kfloat_H100_W00_mean_Oct, Mean_Kfloat_H100_W00_mean_Oct, Max_Kfloat_H100_W00_mean_Oct)
+total_mean_ww_Oct <- rbind(p_RK4_mean_Oct, Min_Kfloat_H095_W05_mean_Oct, Min_Kfloat_H090_W10_mean_Oct, Min_Kfloat_H085_W15_mean_Oct, Mean_KFloat_H095_W05_mean_Oct, Mean_KFloat_H090_W10_mean_Oct, Mean_KFloat_H085_W15_mean_Oct, Max_Kfloat_H095_W05_mean_Oct, Max_Kfloat_H090_W10_mean_Oct, Max_Kfloat_H085_W15_mean_Oct)
+
+total_mean_water_Oct$type <- factor(total_mean_water_Oct$type, levels=c("43", "44", "48", "52"))
+total_mean_ww_Oct$type <- factor(total_mean_ww_Oct$type, levels=c("43", "45", "46", "47", "49", "50", "51", "53", "54", "55"))
+
 waterdrag_plot_Oct <- ggplot()+
   #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H100_W00_mean_Oct, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, colour = type))+
   xlab("Longtitude") + 
@@
   #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Min_Kfloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'grey')+
-  geom_path(data = Min_Kfloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'green')+
-  geom_path(data = Min_Kfloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'pink')+
-  geom_path(data = Mean_KFloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'yellow')+
-  geom_path(data = Mean_KFloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Mean_KFloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'red')+
-  geom_path(data = Max_Kfloat_H095_W05_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Max_Kfloat_H090_W10_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H085_W15_mean_Oct, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, colour = type))+
   xlab("Longtitude") + 
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = p_RK4_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H100_W00_mean_April ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = p_RK4_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H100_W00_mean_July ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = p_RK4_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 13000)) +
+  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 13000)) +
+  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time for scenarios 17 - 29, for April 2018 (winter).
-The plots are regular hexagons within the density of particles was calculated.
+The plots are regular hexagons within the density of particles were calculated.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time for scenarios 30 - 42, for July 2018 (winter).
-The plots are regular hexagons within the density of particles was calculated.
+The plots are regular hexagons within the density of particles were calculated.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 10: Comparison of density plots at the end run time for scenarios 43 - 55, for October 2018 (spring).
-The plots are regular hexagons within the density of particles was calculated.
+The plots are regular hexagons within the density of particles were calculated.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 (Figure 11, simulations 17-29; p < 0.05).
-Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas travelled less distance across compared to other kelp particles.
-The lowest distance travelled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
+Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas travelled shorter distance across compared to other kelp particles.
+The shortest distance travelled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
 No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; p > 0.05).
 In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; p < 0.05).
-Kelp particles with only 90% hydrodynamic drag across cross-sectional areas travelled less distance compared to all other particles.
-As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure travelled the least distance compared to all other particles.
+Kelp particles with only 90% hydrodynamic drag across cross-sectional areas travelled shorter distance compared to all other particles.
+As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure travelled the shortest distance compared to all other particles.
 

### 2021-03-23 thesis_chapter_3.Rmd
grammer error in 3.5 title
changed mean traj plots

@@
 abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
-Macroalage are one of the most important natural dispersal mechanisms, and are regarded as the tumble-weed of the ocean.
-Despite the many studies that exist, the relative role of wind and surface currents in macrolage trajectory is still unknown.
+Macroalgae are one of the most important natural dispersal mechanisms, and are regarded as the tumble-weed of the ocean.
+Despite the many studies that exist, the relative role of wind and surface currents in macroalgae trajectory is still unknown.
 Past studies also focus on kelp-rafts, which can vary greatly in size making it difficult to form a concensus as to the relative role of wind versus surface currents.
@@
 The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011].
-Past research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
+Past research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
 The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag [@breivik2011; @griffin2017].
@@
 p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_mean$type <- "1" # add extra column 'type' to dataframe and assigned 'RK4' label
 
@@
 Min_Kfloat_H100_W00$velocity <- with(Min_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H100_W00_mean$type <- "2" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H095_W05$velocity <- with(Min_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H095_W05_mean$type <- "3" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H090_W10$velocity <- with(Min_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H090_W10_mean$type <- "4" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H085_W15$velocity <- with(Min_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H085_W15_mean$type <- "5" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_Kfloat_H100_W00$velocity <- with(Mean_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_Kfloat_H100_W00_mean$type <- "6" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H095_W05$velocity <- with(Mean_KFloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H095_W05_mean$type <- "7" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H090_W10$velocity <- with(Mean_KFloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H090_W10_mean$type <- "8" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H085_W15$velocity <- with(Mean_KFloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H085_W15_mean$type <- "9" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H100_W00$velocity <- with(Max_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H100_W00_mean$type <- "10" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H095_W05$velocity <- with(Max_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H095_W05_mean$type <- "11" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H090_W10$velocity <- with(Max_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H090_W10_mean$type <- "12" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H085_W15$velocity <- with(Max_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H085_W15_mean$type <- "13" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 
+total_mean_water <- rbind(p_RK4_mean, Min_Kfloat_H100_W00_mean, Mean_Kfloat_H100_W00_mean, Max_Kfloat_H100_W00_mean)
+total_mean_ww <- rbind(p_RK4_mean, Min_Kfloat_H095_W05_mean, Min_Kfloat_H090_W10_mean, Min_Kfloat_H085_W15_mean, Mean_KFloat_H095_W05_mean, Mean_KFloat_H090_W10_mean, Mean_KFloat_H085_W15_mean, Max_Kfloat_H095_W05_mean, Max_Kfloat_H090_W10_mean, Max_Kfloat_H085_W15_mean)
+
+total_mean_water$type <- factor(total_mean_water$type, levels=c("1", "2", "6", "10"))
+total_mean_ww$type <- factor(total_mean_ww$type, levels=c("1", "3", "4", "5", "7", "8", "9", "11", "12", "13"))
+
 waterdrag_plot <- ggplot()+
   #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'blue')+
-  geom_point(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), shape = 20)+
-  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = total_mean_water, aes(x = lon, y = lat, colour = type))+
   xlab("Longtitude") + 
@@
   #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = Min_Kfloat_H095_W05_mean, aes(x = lon, y = lat), color = 'grey')+
-  geom_path(data = Min_Kfloat_H090_W10_mean, aes(x = lon, y = lat), color = 'green')+
-  geom_path(data = Min_Kfloat_H085_W15_mean, aes(x = lon, y = lat), color = 'pink')+
-  geom_path(data = Mean_KFloat_H095_W05_mean, aes(x = lon, y = lat), color = 'yellow')+
-  geom_path(data = Mean_KFloat_H090_W10_mean, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Mean_KFloat_H085_W15_mean, aes(x = lon, y = lat), color = 'red')+
-  geom_path(data = Max_Kfloat_H095_W05_mean, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = Max_Kfloat_H090_W10_mean, aes(x = lon, y = lat), color = 'orange')+
-  geom_path(data = Max_Kfloat_H085_W15_mean, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = total_mean_ww, aes(x = lon, y = lat, colour = type))+
   xlab("Longtitude") + 
@@
 Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
-In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10.
-In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13.
 The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
-ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO")
+
+ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO", ncol = 1, nrow = 2)
+
 ```
@@
   geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient",limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient",limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient",limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time for scenarios 1 - 13, for January 2018 (summer).
-The plots are regular hexagons within the density of particles was calculated.
+The plots are regular hexagons within the density of particles were calculated.
 The mean trajectories for each scenario are also plotted for comparison.
@@
 
-## Comparison of coefficents
+## Comparison of coefficients
 

### 2021-03-23 thesis_chapter_3.Rmd
chapter 3 updates

@@
   geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
   geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient",limits = c(1, 1100)) +
   geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient",limits = c(1, 1100)) +
   geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient",limits = c(1, 1100)) +
   geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
   geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
   geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
   geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
   geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
   geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
   geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
   geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
   geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +

### 2021-03-24 thesis_chapter_3.Rmd
updates

@@
 
+data_dis <- total_data %>% 
+  select(!(kelpvelU)) %>%
+  select(!(kelpvelV)) %>%
+  select(!(velocity))
+
+total_dis_data <- rbind(p_RK4, data_dis)
+
 pos <- c("RK4",
@@
 
-dis_plot <- ggplot(data = total_data, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(data = p_RK4, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+dis_plot <- ggplot(data = total_dis_data, aes(x = type, y = distance, show.legend = FALSE))+
   geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
@@
 
-custom_labs <- c("1","2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot
+custom_labs <- c("1","2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot
 
@@
 The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
+
 dis_plot
+ggsave(filename = "figure_4.jpg", plot = last_plot(), device = "jpeg", path = "figures/chapter_3/", width = 8, height = 5, dpi = 600)
+
 ```
@@
 kelp_acc_mean$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc$age <- with(kelp_acc, age/3600)
 
@@
 pass_acc_mean$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc$age <- with(pass_acc, age/3600)
 
@@
   geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
+  #scale_fill_continuous(type = "gradient", limits = c(0, 30)) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
@@
   ggtitle("") +
-  guides(fill=guide_legend(title="Age")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_kelp_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  #scale_color_viridis_c(name = "Age") +
   guides(fill=guide_legend(title="Age")) +
@@
 ggarrange(plot_A, plot_B, labels = c("AUTO"), ncol = 1, nrow = 2)
+ggsave(filename = "figure_5.jpg", plot = last_plot(), device = "jpeg", path = "figures/chapter_3/", width = 7, height = 7, dpi = 600)
 
@@
 A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region.
-The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due to  as juveniles do are not swim fast enough to alter their overall trajectory.
+The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due as juveniles do not swim fast enough to alter their overall trajectory.
 Therefore, the differences in cross-sectional area types may not be significant enough to have an effect on the relevant zonal and meridional velocity vectors.
@@
 Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns.
-All the simulations that included any form of drag resulted in a  high density of particles entrained in a eddy.
+All the simulations that included any form of drag resulted in a high density of particles entrained in an eddy.
 The entrainment and expulsion from eddies and vortices are an important characteristic of floating macroalgae.

### 2021-03-30 thesis_chapter_3.Rmd
figure 1 caption update

@@
 
-```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1:Map of the study domain and the main features pertaining to transport in the region.
-In panel A, the map of the entire South African coastline and the main currents presented.
-The resulting eddies and filaments from the Agulhas retroflection and study area are presented.
-In panel B, the splitting of the Benguela Jet, eddie features and general flow of the study domain are presented.
-The release site is represented by a blue star."}
+```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1: Map of the study domain and the main features pertaining to transport in the region.
+In panel A, the map of the entire South African coastline and the main currents is presented.
+The resulting eddies and filaments from the Agulhas retroflection are indicated, and the position of the study area is drawn in as a black rectangle.
+Detail of the study area is in panel B.
+Here the splitting of the Benguela Jet, eddie features, and the general flow of the study domain are presented.
+The release site of simulated kelp ‘particles’ is represented by a blue star."}
 

### 2021-03-30 thesis_chapter_3.Rmd
corrections...

--- 2021-03-24 thesis_chapter_3.Rmd
@@
 csl: markdown/frontiers.csl.txt
-abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
-Macroalgae are one of the most important natural dispersal mechanisms, and are regarded as the tumble-weed of the ocean.
-Despite the many studies that exist, the relative role of wind and surface currents in macroalgae trajectory is still unknown.
-Past studies also focus on kelp-rafts, which can vary greatly in size making it difficult to form a concensus as to the relative role of wind versus surface currents.
-In addition, past studies have not considered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam.
-The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectories of flotsam in the ocean.
-This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagae trajectory and accumulation.
-This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals.
-This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley Lagrangian particles.
+abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
+Macroalgae are one of the most important natural dispersal vectors, and are regarded as the tumble-weed of the ocean.
+Despite many studies that exist on the topic, the relative role of wind and surface currents in macroalgae trajectory is still unknown.
+Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a concensus as to the relative role of wind versus surface currents.
+In addition, past studies have not considered surface-area in relation to drag characterisics of floating macroalgae which have been shown to play a role in the trajectory and accumulation of other flotsam.
+Advances in Lagrangian trajectory modeling have been exponential in recent years, and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean.
+This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagal trajectory and accumulation.
+We did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals.
+This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles.
 The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain.
@@
 
-There are a range of objects, both natural and anthropogenic, floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
-Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008]; as well as input of allochthonous materials to near and distant ecosystems [@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].
-
-Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species of macroalgae have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.
+There is a range of objects --- both natural and anthropogenic --- floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
+Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
+Kelp also function as input of allochthonous materials to near and distant ecosystems and in this manner support the ecological function in a range of adjacent habitats [@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].
+
+Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.
 In turn, the positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum.
 The structure and number of pneumatocysts varies between species.
-For example species from the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018].
-In some cases the stipe itself is air-filled, such as with *E.
+For example species belonging to the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018].
+In some cases the stipe itself may be air-filled, such as with *E.
 maxima* and *E.
@@
 pyrifera* in nearshore Monterrey Bay.
-The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents, however the relative importance of wind and was not clear.
+The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents; however the relative importance of wind and was not clear.
 In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter.
-Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems.
+Previous studies have also identified wind as an important mechanism of dispersal in wind dominated ocean systems.
 For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating bull kelp *Durvillaea antarctica*.
@@
 antarctica* used a combination of population genetics and relative age estimate of goosebarnacles attached to the raft.
-The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate.
-Genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin.
+The presence of goosebarnacles suggests a long rafting, time as these species have a slow growth rate.
+Genetic analyses showed these species are able to raft up to $\sim$ 390 km from their local origin.
 The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011].
 Past research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
-The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag [@breivik2011; @griffin2017].
+The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic-form drag [@breivik2011; @griffin2017].
 In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object.
@@
 
-To accurately determine the trajectory of marine macroalgae water and wind drag need to be taken into account.
+To accurately determine the trajectory of marine macroalgae, water and wind drag need to be taken into account.
 Past research by @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
@@
 Although various approaches have been used in the past to investigate floating macroalgal trajectories, very few studies have employed the use of Langrangian trajectory modeling.
-Furthermore, none of the existing studies using this modeling approach have considered macroalgal morphology cross-sectional area (**i.e.** shape) as an aspect of drag and ultimately trajectory.
+Furthermore, none of the existing studies using this modeling approach have considered macroalgal morphology cross-sectional area (*i.e.* shape) as an aspect of drag and ultimately trajectory.
 Studies by @brooks2019, @putman2018, and @putman2020 used Lagrangian approaches to investigate the effects of inertia, raft-size, and windage on macroalgae trajectory.
 @brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customised Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
-The results showed that trajectory of pelagic *Sargussum* was significantly influenced by inertia and the radial size of the rafts.
+The results showed that the trajectory of pelagic *Sargussum* was significantly influenced by inertia and the radial size of the rafts.
 Other work by @putman2020 investigated the effect of including windage in macroalgae trajectory simulations.
-The authors showed that including **ad-hoc** windage factors greatly improved accuracy regarding entrainment or expulsion from eddies when compared to tracked sargassum mats in the ocean.
-Langrangian trajectory modeling is a useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
+The authors showed that including *ad hoc* windage factors greatly improved accuracy regarding entrainment or expulsion from eddies when compared to tracked *Sargassum* mats in the ocean.
+Langrangian trajectory modeling is clearly; useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
 The radial size included by @brooks2019 is an improvement, however this approach does not consider the complex morphology on an individual level into account.
 
-In addition, past studies focus on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddies and vortices in the accumulation of kelp at the surface.
+Past studies have focused on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddies and vortices in the accumulation of individual kelp plants at the surface.
 Gyres, eddies and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018].
@@
 In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations.
-It should be noted that this study does not attempt to characterise the dispersal of drifting macroalage within the study domain.
-The authors argue that too little is know about the drift characteristics of solitary floating macroalgae to characterise dispersal effectively.
-Instead, this study will aim to shed light on the 'behaviour' of surface drifting macroalgae from a coastal environment into the open ocean, which will inform future characterisation studies.
-This will be achieved by comparing the trajectory outputs of virtual 'passive' particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure.
+We aim to shed light on the 'behaviour' of surface drifting macroalgae from a coastal environment into the open ocean, which will inform future characterisation studies.
+We will achieve this by comparing the trajectory outputs of virtual passive particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure.
 The different degrees of wave and wind exposure are meant to reflect the different buoyancy characteristics, as well as identify the role of magnitude of the drag components on overall trajectory.
-In addition, the modeling of single kelp offers the opportunity to study the role of eddies and vortices in accumulation points on the surface.
-Therefore, separate simulations will be run to compare the entrained characteristics of passive particle with a kelp particle.
+The modeling of single kelp offers the opportunity to study the role of eddies and vortices in accumulation points on the surface.
 
@@
 A release site was chosen for this study, Kommetjie, which is located on the Atlantic side of the Cape Peninsula (Figure 2B).
-The hydrodynamics in the study region are driven by a complex interaction of wind and waves, which are also influenced by larger oceanic processes.
-In South Africa, the larger processes can be attributed to the Agulhas Current (AC) and the Benguela Current (BC)[@lutjeharms2000; @lutjeharms2007 ;@garzoli1996 ;@hutchings2009; @rubio2009].
+The hydrodynamics in the study region are driven by a complex interaction of wind and waves, and are also influenced by larger oceanic processes.
+In South Africa, the larger processes can be attributed to the Agulhas Current (AC) and the Benguela Current (BC) [@lutjeharms2000;@lutjeharms2007;@garzoli1996;@hutchings2009;@rubio2009].
 The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].
 The region where the retroflection occurs is known as the Agulhas Retroflection and consists of high levels of mesoscale variability [@garzoli1996].
-The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988; @lutjeharms2006; @lutjeharms2007].
-The Agulhas Rings are the main contributors to the Agulhas leakage which are the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015].
+The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324 km [@lutjeharms1988;@lutjeharms2006; @lutjeharms2007].
+Agulhas Rings are the main contributors to the Agulhas leakage which is the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015].
 The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009].
-The Benguela region is characterised by strong coastal upwelling cells which are primarily driven by equatorward winds [@shannon1996; @hardman2003].
-It is known as the Benguela Upwelling System (BUS) The BUS starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996].
+The Benguela region is characterised by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003].
+It is known as the Benguela Upwelling System (BUS).
+The BUS starts at 27 S and extends southward to 35 S [@shannon1996].
 The region is also characterised by high amounts of mesoscale variability in the form of eddies and filaments [@blanke2002; @blanke2005; @rubio2009].
@@
 
-```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1:Map of the study domain and the main features pertaining to transport in the region.
+```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1: Map of the study domain and the main features pertaining to transport in the region.
 In panel A, the map of the entire South African coastline and the main currents presented.

### 2021-03-30 thesis_chapter_3.Rmd
figure 1 caption update

@@
 ```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1: Map of the study domain and the main features pertaining to transport in the region.
-In panel A, the map of the entire South African coastline and the main currents presented.
-The resulting eddies and filaments from the Agulhas retroflection and study area are presented.
-In panel B, the splitting of the Benguela Jet, eddie features and general flow of the study domain are presented.
-The release site is represented by a blue star."}
+In panel A, the map of the entire South African coastline and the main currents is presented.
+The resulting eddies and filaments from the Agulhas retroflection are indicated, and the position of the study area is drawn in as a black rectangle.
+Detail of the study area is in panel B.
+Here the splitting of the Benguela Jet, eddie features, and the general flow of the study domain are presented.
+The release site of simulated kelp ‘particles’ is represented by a blue star."}
 

### 2021-03-30 thesis_chapter_3.Rmd
corrects 2....

@@
 
-The study domain falls within the southern Benguela upwelling ecosystem, which is a region of high mesoscale variability and activity.
-These occur mostly in the form of eddies, filaments and vortices which are generated from the Agulhas retroflection [@veitch2018; @veitch2017], generally flowing northward along the coastline and begin to move offshore near Cape Columbine.
+The study domain falls within the southern Benguela upwelling ecosystem, which is a region of high mesoscale variability and activity in the form of eddies, filaments and vortices generated from the Agulhas retroflection [@veitch2018; @veitch2017] they generally flow northward along the coastline and begin to move offshore near Cape Columbine.
 The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, Eckman transport and geostrophic currents [@stramma1989; @veitch2017; @veitch2018; @hutchings2009].
-The stability of the jet varies seasonally with more stable flows in winter and splitting of the jet during summer and spring.
-In January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results into a weak narrow inshore branch which flows along the shelf edge and a strong flowing north-westwards offshore branch.
-The area is also an area of high upwelling activity driven by strong westerly winds [@hutchings2009].
-These winds are most intense during summer (December-March) which favours upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet; which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
+The stability of the jet varies seasonally with more stable flows in winter; and in January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results in a weak narrow inshore branch that flows along the shelf edge and a strong flowing north-westwards offshore branch.
+The area is also characterized by high upwelling activity driven by strong westerly winds [@hutchings2009].
+These winds are most intense during summer (December-March) which favors upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
 In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
@@
 
-The model is based on parameterisations that have been used for previous studies investigating iceberg, capsized marine vessels and microplastics.
+The model is based on parameterizations that have been used for previous studies investigating iceberg, capsized marine vessels and microplastics.
 It should be noted that most of the previous studies regarding macroalgae trajectory have investigated rafts and not solitary floating individuals.
 A kelp raft will vary in size and can range from a few meters across to large dense mats the size of a rugby field.
-Furthermore, some authors suggest that most macroalgae become entangled through ocean currents and not through the hydrodynamic forces that dislodge them.
-In addition, large kelp rafts are not a feature around the coast of the Cape Peninsula; rather, solitary kelp are (pers.
+Some authors suggest that most macroalgae become entangled through ocean currents and not through the hydrodynamic forces that dislodge them.
+Large kelp rafts are not a feature around the coast of the Cape Peninsula; rather, solitary kelp are (pers.
 obs).
@@
 maxima* individuals will be considered.
-Numerical calculations of particle trajectory were performed with a Lagrangian method [@delandmeter2019], and be described by the equation
+We performed numerical calculations of particle trajectory with a Lagrangian method [@delandmeter2019], and be described by the equation
 
 $$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
-where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model at particle point in time, $t$ is time, $X_b(t)$ is a change in position due to 'behavior' of the particle at a particular point in time.
-This can range from swimming in fish to sinking or beaching.
+where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model at particle point in time, $t$ is time $X_b(t)$ is a change in position due to 'behavior' of the particle at a particular point in time.
 In order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.
@@
 The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations.
-Hydrodynamic drag, wind drag and surface current-flow are two-dimensional vector quantities and were calculated for each time step for each simulation.
+Hydrodynamic drag, wind drag, and surface current-flow are two-dimensional vector quantities and were calculated for each time step for each simulation.
 
@@
 Virtual particles were released with no drag or wind effects from the study site in order to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'.
-All particles that have any form of drag exposure are hereafter referred to as 'kelp' particles.
+All particles that have any form of drag exposure are hereafter referred to as 'kelp particles'.
 All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018].
@@
 For example, if 85% of the plant is submerged and 15% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively.
-This approach has been used in past research investigating pelagic Sargassum rafts [@putman2020;@putman2018].
+This approach has been used in past research investigating pelagic *Sargassum* rafts [@putman2020;@putman2018].
 The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1).
-Each scenario consisted of a particle set with a size of 1000 where released from Kommetjie at the start of simulation.
-The particle sets were advected with a Runge-Kutte fourth order scheme for 30 days and incorporated brownian motion.
-Since the aim of this investigation was to investigate the relative surface behaviour of virtual kelp particles, only the surface layer was used for simulations.
+Each scenario consisted of a particle set with a size of 1000 were released from Kommetjie at the start of simulation.
+Since the aim of this investigation was to investigate the relative surface behavior of virtual kelp particles, only the surface layer was used for simulations.
 It should also be noted that the Copernicus model used places velocity vectors over some of the land covering the coastal domain.

### 2021-03-30 thesis_chapter_3.Rmd
table 1 caption

@@
 sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
-sim_table <- set_caption(sim_table, "Table 1: Details for each simulation run for the study.
+sim_table <- set_caption(sim_table, "Details for each simulation run for the study.
 Type refers to either a passive particle with no drag behaviour or a kelp particle that has been advected with any drag exposure.
-The drag exposure column refers to the parameterisations regarding the level of water and wind drag simulated.")
+The drag exposure column refers to the parameterisations regarding the level of water and wind drag simulated.
+All simulations were run with a Runge-Kutte (4th) advection scheme, Brownian motion, 30 days, and included 1000 particles for a 2018 hindcast.")
 sim_table

### 2021-03-30 thesis_chapter_3.Rmd
moved appendix simulations and writeup to main section

@@
 Past studies have focused on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddies and vortices in the accumulation of individual kelp plants at the surface.
-Gyres, eddies and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018].
+Gyres, eddies, and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018].
 Past studies investigating the dispersal of marine debris have also included the possible effects of inertia in accumulation of ocean drifter instruments in mesoscale features.
@@
 This particular time period was chosen as the splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity [@rubio2009; @veitch2018; @ragoasha2019].
+Additional simulations were done for months in 2018, which were chosen based on seasonality.
+These simulations had the same parameters used for January and were performed for April (Autumn), July (winter), and October (Spring) in order to assess if any patterns seen in January translate into general patterns for times periods throughout 2018.
 A summary of the parameters for each simulation are provided in table 1.
@@
 To determine if vortices and eddies play a role in kelp accumulation on the surface, a separate set of simulations were performed.
-Unlike the previous simulations these simulation released a particle every hour.
-These were run over the same time period for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme and incorporated Brownian motion.
+Unlike the previous simulations these simulation released a particle every hour over the same time periods.
 In order to determine if the shape co-efficient used plays a role in model used, a simulation of a kelp with the maximum surface area and drag exposure was run using the shape coefficient for a cylinder to compare to a simulation using a sphere, and was run for 30 days using a fourth order Runge-Kutte advection scheme with brownian motion and a particle set of size 1000.
-
-Additional simulations were done for months in 2018, which were chosen based on seasonality.
-These simulations had the same parameters used for January and were performed for April (Autumn), July (winter), and October (Spring) in order to assess if any patterns seen in January translate into general patterns for times periods throughout 2018.
 
@@
 
-Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service (Copernicus, http://marine.copernicus.com).
+Both passive and kelp particles were simulated around the South African coastline within hindcast outputs from the Copernicus Marine Environment Monitoring Service (Copernicus, http://marine.copernicus.com).
 The Copernicus model is based on the modelling framework known as Nucleus for European Modelling of the ocean.
-The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
-The model is forcedhas a spatial resolution of 0.08$^\circ$ and is interpolated on an Arakawa C native grid.
+The Copernicus outputs used in this study contain 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
+The model has a spatial resolution of 0.08$^\circ$ and is interpolated on an Arakawa C native grid.
 In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (Global Ocean Wind L4_NRT_012_004).
-The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
-Both averaged ocean current and wind data are stored as meridional and zonal velocity vectors.
+The outputs from the model are composed of 6-hourly averaged fields of surface 10 m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
+Averaged ocean current and wind data are stored as meridional and zonal velocity vectors.
 
@@
 Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant.
-For details please refer to figure 2.
+For details refer to figure 2.
 The dimensional data needed were estimated in cases where data was not available for that particular morphological characteristic.
@@
 
-```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Diagram of the *E.maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic."}
+```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Diagram of the *E.
+maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic."}
 
@@
 area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
-area_table <- set_caption(area_table, "Table 2: Table summarising the attributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.
+area_table <- set_caption(area_table, "Table 2: Table summarising the attributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulations.
 Approximate shape refers to the dimensions of the geometric shape used to represent a particular morphological characteristic.
@@
 
+## April
+
+p_RK4_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+p_RK4_mean_April <- p_RK4_April %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+p_RK4_April$type <- "RK4_April" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_mean_April$type <- "17" # add extra column 'type' to dataframe and assigned 'RK4' label
+
+Min_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H100_W00_mean_April <- Min_Kfloat_H100_W00_April %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H100_W00_April$type <- "Min_Kfloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_April$velocity <- with(Min_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H100_W00_mean_April$type <- "18" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H095_W05_mean_April <- Min_Kfloat_H095_W05_April %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H095_W05_April$type <- "Min_Kfloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_April$velocity <- with(Min_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H095_W05_mean_April$type <- "19" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H090_W10_mean_April <- Min_Kfloat_H090_W10_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H090_W10_April$type <- "Min_Kfloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_April$velocity <- with(Min_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H090_W10_mean_April$type <- "20" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H085_W15_mean_April <- Min_Kfloat_H085_W15_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H085_W15_April$type <- "Min_Kfloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_April$velocity <- with(Min_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H085_W15_mean_April$type <- "21" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_Kfloat_H100_W00_mean_April <- Mean_Kfloat_H100_W00_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_Kfloat_H100_W00_April$type <- "Mean_Kfloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_April$velocity <- with(Mean_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_Kfloat_H100_W00_mean_April$type <- "22" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_KFloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H095_W05_mean_April <- Mean_KFloat_H095_W05_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H095_W05_April$type <- "Mean_KFloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_April$velocity <- with(Mean_KFloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H095_W05_mean_April$type <- "23" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_KFloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H090_W10_mean_April <- Mean_KFloat_H090_W10_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H090_W10_April$type <- "Mean_KFloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_April$velocity <- with(Mean_KFloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H090_W10_mean_April$type <- "24" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_KFloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H085_W15_mean_April <- Mean_KFloat_H085_W15_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H085_W15_April$type <- "Mean_KFloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_April$velocity <- with(Mean_KFloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H085_W15_mean_April$type <- "25" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H100_W00_mean_April <- Max_Kfloat_H100_W00_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H100_W00_April$type <- "Max_KFloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_April$velocity <- with(Max_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H100_W00_mean_April$type <- "26" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H095_W05_mean_April <- Max_Kfloat_H095_W05_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H095_W05_April$type <- "Max_KFloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_April$velocity <- with(Max_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H095_W05_mean_April$type <- "27" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H090_W10_mean_April <- Max_Kfloat_H090_W10_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H090_W10_April$type <- "Max_KFloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_April$velocity <- with(Max_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H090_W10_mean_April$type <- "28" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H085_W15_mean_April <- Max_Kfloat_H085_W15_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H085_W15_April$type <- "Max_KFloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_April$velocity <- with(Max_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H085_W15_mean_April$type <- "29" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+total_data_April <- rbind(Min_Kfloat_H100_W00_April, Min_Kfloat_H095_W05_April, Min_Kfloat_H090_W10_April, Min_Kfloat_H085_W15_April, Mean_Kfloat_H100_W00_April, Mean_KFloat_H095_W05_April, Mean_KFloat_H090_W10_April, Mean_KFloat_H085_W15_April, Max_Kfloat_H100_W00_April, Max_Kfloat_H095_W05_April, Max_Kfloat_H090_W10_April, Max_Kfloat_H085_W15_April)
+
+## July
+
+p_RK4_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+p_RK4_mean_July <- p_RK4_July %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+p_RK4_July$type <- "RK4_July" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_mean_July$type <- "30" # add extra column 'type' to dataframe and assigned 'RK4' label
+
+Min_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H100_W00_mean_July <- Min_Kfloat_H100_W00_July %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H100_W00_July$type <- "Min_Kfloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_July$velocity <- with(Min_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H100_W00_mean_July$type <- "31" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H095_W05_mean_July <- Min_Kfloat_H095_W05_July %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H095_W05_July$type <- "Min_Kfloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_July$velocity <- with(Min_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H095_W05_mean_July$type <- "32" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H090_W10_mean_July <- Min_Kfloat_H090_W10_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H090_W10_July$type <- "Min_Kfloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_July$velocity <- with(Min_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H090_W10_mean_July$type <- "33" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H085_W15_mean_July <- Min_Kfloat_H085_W15_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H085_W15_July$type <- "Min_Kfloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_July$velocity <- with(Min_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H085_W15_mean_July$type <- "34" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_Kfloat_H100_W00_mean_July <- Mean_Kfloat_H100_W00_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_Kfloat_H100_W00_July$type <- "Mean_Kfloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_July$velocity <- with(Mean_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_Kfloat_H100_W00_mean_July$type <- "35" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_KFloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H095_W05_mean_July <- Mean_KFloat_H095_W05_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H095_W05_July$type <- "Mean_KFloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_July$velocity <- with(Mean_KFloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H095_W05_mean_July$type <- "36" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_KFloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H090_W10_mean_July <- Mean_KFloat_H090_W10_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H090_W10_July$type <- "Mean_KFloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_July$velocity <- with(Mean_KFloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H090_W10_mean_July$type <- "37" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_KFloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H085_W15_mean_July <- Mean_KFloat_H085_W15_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H085_W15_July$type <- "Mean_KFloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_July$velocity <- with(Mean_KFloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H085_W15_mean_July$type <- "38" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H100_W00_mean_July <- Max_Kfloat_H100_W00_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H100_W00_July$type <- "Max_KFloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_July$velocity <- with(Max_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H100_W00_mean_July$type <- "39" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H095_W05_mean_July <- Max_Kfloat_H095_W05_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H095_W05_July$type <- "Max_KFloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_July$velocity <- with(Max_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H095_W05_mean_July$type <- "40" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H090_W10_mean_July <- Max_Kfloat_H090_W10_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H090_W10_July$type <- "Max_KFloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_July$velocity <- with(Max_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H090_W10_mean_July$type <- "41" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H085_W15_mean_July <- Max_Kfloat_H085_W15_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H085_W15_July$type <- "Max_KFloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_July$velocity <- with(Max_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H085_W15_mean_July$type <- "42" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+total_data_July <- rbind(Min_Kfloat_H100_W00_July, Min_Kfloat_H095_W05_July, Min_Kfloat_H090_W10_July, Min_Kfloat_H085_W15_July, Mean_Kfloat_H100_W00_July, Mean_KFloat_H095_W05_July, Mean_KFloat_H090_W10_July, Mean_KFloat_H085_W15_July, Max_Kfloat_H100_W00_July, Max_Kfloat_H095_W05_July, Max_Kfloat_H090_W10_July, Max_Kfloat_H085_W15_July)
+
+## October
+
+p_RK4_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+p_RK4_mean_Oct <- p_RK4_Oct %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+p_RK4_Oct$type <- "RK4_Oct" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_mean_Oct$type <- "43" # add extra column 'type' to dataframe and assigned 'RK4' label
+
+Min_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H100_W00_mean_Oct <- Min_Kfloat_H100_W00_Oct %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H100_W00_Oct$type <- "Min_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H100_W00_mean_Oct$type <- "44" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H095_W05_mean_Oct <- Min_Kfloat_H095_W05_Oct %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H095_W05_Oct$type <- "Min_Kfloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H095_W05_mean_Oct$type <- "45" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H090_W10_mean_Oct <- Min_Kfloat_H090_W10_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H090_W10_Oct$type <- "Min_Kfloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H090_W10_mean_Oct$type <- "46" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Min_Kfloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Min_Kfloat_H085_W15_mean_Oct <- Min_Kfloat_H085_W15_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Min_Kfloat_H085_W15_Oct$type <- "Min_Kfloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Min_Kfloat_H085_W15_mean_Oct$type <- "47" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_Kfloat_H100_W00_mean_Oct <- Mean_Kfloat_H100_W00_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_Kfloat_H100_W00_Oct$type <- "Mean_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_Kfloat_H100_W00_mean_Oct$type <- "48" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_KFloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H095_W05_mean_Oct <- Mean_KFloat_H095_W05_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H095_W05_Oct$type <- "Mean_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H095_W05_mean_Oct$type <- "49" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_KFloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H090_W10_mean_Oct <- Mean_KFloat_H090_W10_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H090_W10_Oct$type <- "Mean_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H090_W10_mean_Oct$type <- "50" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Mean_KFloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Mean_KFloat_H085_W15_mean_Oct <- Mean_KFloat_H085_W15_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Mean_KFloat_H085_W15_Oct$type <- "Mean_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Mean_KFloat_H085_W15_mean_Oct$type <- "51" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H100_W00_mean_Oct <- Max_Kfloat_H100_W00_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H100_W00_Oct$type <- "Max_KFloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H100_W00_mean_Oct$type <- "52" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H095_W05_mean_Oct <- Max_Kfloat_H095_W05_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H095_W05_Oct$type <- "Max_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H095_W05_mean_Oct$type <- "53" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H090_W10_mean_Oct <- Max_Kfloat_H090_W10_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H090_W10_Oct$type <- "Max_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H090_W10_mean_Oct$type <- "54" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+Max_Kfloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+Max_Kfloat_H085_W15_mean_Oct <- Max_Kfloat_H085_W15_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+Max_Kfloat_H085_W15_Oct$type <- "Max_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
+Max_Kfloat_H085_W15_mean_Oct$type <- "55" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+total_data_Oct <- rbind(Min_Kfloat_H100_W00_Oct, Min_Kfloat_H095_W05_Oct, Min_Kfloat_H090_W10_Oct, Min_Kfloat_H085_W15_Oct, Mean_Kfloat_H100_W00_Oct, Mean_KFloat_H095_W05_Oct, Mean_KFloat_H090_W10_Oct, Mean_KFloat_H085_W15_Oct, Max_Kfloat_H100_W00_Oct, Max_Kfloat_H095_W05_Oct, Max_Kfloat_H090_W10_Oct, Max_Kfloat_H085_W15_Oct)
+
 ```
@@
 In all simulations particles flowed in a north-westward direction.
-The patterns of north-westward flow and entrainment in eddies is also evident in the additional simulations, however clear differences in mean trajectory between kelp and passive particles is clear (see appendix).
+
+The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (autumn).
+Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag travelled further northward compared to the passive particles (see Figure 7A-B).
+The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles.
+Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles travelled less distance compared to kelp particles with both wind and hydrodynamic drag.
+Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 7C-D).
+Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore; resulting in entrainment within an eddy.
+Similar behaviour is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag.
+In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles travelled initially southward and then westward offshore and also becoming entrained in an eddy.
+In the October (Spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 7E-F).
+Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag ended a greater distance from the release site compared to the passive particles.
+This was also evident when comparing kelp particles with different cross-sectional areas and exposure to wind and hydrodynamic drag with passive particles; with the end location a further from the release location.
+In both the April and October simulations all particles flowed north-westward, while in winter the mean trajectory resulted in entrainment within an eddy.
 
@@
 
+total_mean_water_April <- rbind(p_RK4_mean_April, Min_Kfloat_H100_W00_mean_April, Mean_Kfloat_H100_W00_mean_April, Max_Kfloat_H100_W00_mean_April)
+total_mean_ww_April <- rbind(p_RK4_mean_April, Min_Kfloat_H095_W05_mean_April, Min_Kfloat_H090_W10_mean_April, Min_Kfloat_H085_W15_mean_April, Mean_KFloat_H095_W05_mean_April, Mean_KFloat_H090_W10_mean_April, Mean_KFloat_H085_W15_mean_April, Max_Kfloat_H095_W05_mean_April, Max_Kfloat_H090_W10_mean_April, Max_Kfloat_H085_W15_mean_April)
+
+total_mean_water_April$type <- factor(total_mean_water_April$type, levels=c("17", "18", "22", "26"))
+total_mean_ww_April$type <- factor(total_mean_ww_April$type, levels=c("17", "19", "20", "21", "23", "24", "25", "27", "28", "29"))
+
+waterdrag_plot_April <- ggplot()+
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_path(data = total_mean_water_April, aes(x = lon, y = lat, colour = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  scale_colour_discrete("Type") +  
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+wwdrag_plot_April <- ggplot()+
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, colour = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  scale_colour_discrete("Type") +  
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+total_mean_water_July <- rbind(p_RK4_mean_July, Min_Kfloat_H100_W00_mean_July, Mean_Kfloat_H100_W00_mean_July, Max_Kfloat_H100_W00_mean_July)
+total_mean_ww_July <- rbind(p_RK4_mean_July, Min_Kfloat_H095_W05_mean_July, Min_Kfloat_H090_W10_mean_July, Min_Kfloat_H085_W15_mean_July, Mean_KFloat_H095_W05_mean_July, Mean_KFloat_H090_W10_mean_July, Mean_KFloat_H085_W15_mean_July, Max_Kfloat_H095_W05_mean_July, Max_Kfloat_H090_W10_mean_July, Max_Kfloat_H085_W15_mean_July)
+
+total_mean_water_July$type <- factor(total_mean_water_July$type, levels=c("30", "31", "35", "39"))
+total_mean_ww_July$type <- factor(total_mean_ww_July$type, levels=c("30", "32", "33", "34", "36", "37", "38", "40", "41", "42"))
+
+waterdrag_plot_July <- ggplot()+
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_path(data = total_mean_water_July, aes(x = lon, y = lat, colour = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  scale_colour_discrete("Type") +  
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+wwdrag_plot_July <- ggplot()+
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, colour = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  scale_colour_discrete("Type") +  
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+total_mean_water_Oct <- rbind(p_RK4_mean_Oct, Min_Kfloat_H100_W00_mean_Oct, Mean_Kfloat_H100_W00_mean_Oct, Max_Kfloat_H100_W00_mean_Oct)
+total_mean_ww_Oct <- rbind(p_RK4_mean_Oct, Min_Kfloat_H095_W05_mean_Oct, Min_Kfloat_H090_W10_mean_Oct, Min_Kfloat_H085_W15_mean_Oct, Mean_KFloat_H095_W05_mean_Oct, Mean_KFloat_H090_W10_mean_Oct, Mean_KFloat_H085_W15_mean_Oct, Max_Kfloat_H095_W05_mean_Oct, Max_Kfloat_H090_W10_mean_Oct, Max_Kfloat_H085_W15_mean_Oct)
+
+total_mean_water_Oct$type <- factor(total_mean_water_Oct$type, levels=c("43", "44", "48", "52"))
+total_mean_ww_Oct$type <- factor(total_mean_ww_Oct$type, levels=c("43", "45", "46", "47", "49", "50", "51", "53", "54", "55"))
+
+waterdrag_plot_Oct <- ggplot()+
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, colour = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  scale_colour_discrete("Type") +  
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+wwdrag_plot_Oct <- ggplot()+
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, colour = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  scale_colour_discrete("Type") +  
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
 ```
@@
 
-ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO", ncol = 1, nrow = 2)
+ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, labels = "AUTO", ncol = 2, nrow = 4)
 
@@
 
+In April and across all simulations, particles flowed in a north-westward direction (see Figure 8).
+The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation.
+In addition, simulations that considered any form of drag got entrained within an eddy further north-westward from the end location of passive particles.
+Only slight variations across cross-sectional areas and drag exposure scenarios are evident.
+Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in a eddy located westward of the release site compared to passive particles (see Figure 9).
+Variation in density along the mean trajectory and eddy for all kelp particles is evident.
+Kelp particles with only hydrodynamic drag across cross-sectional areas showed lower density of particles along the mean trajectory compared to kelp particles with both wind and hydrodynamic drag.
+October simulations showed no differences in density of particles along the mean trajectory when comparing passive and kelp particles.
+Only slight variations in density and trajectory can be seen with no entrainment (see Figure 10).
+Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.
+
 ```{r density plots, message=FALSE, warning=FALSE, echo=FALSE}
@@
   theme(plot.title = element_text(hjust = 0.5))
+
+## April
+
+heatmap_RK4_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = p_RK4_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H100_W00_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H100W100") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H095_W05_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  #ggtitle("H095W05") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H090_W10_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H090W10") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H085_W15_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H085W15") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_Kfloat_H100_W00_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_path(data = Mean_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H095_W05_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_path(data = Mean_KFloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H090_W10_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Mean_KFloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H085_W15_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H100_W00_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H100_W00_mean_April ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H095_W05_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  theme(plot.title = element_text(hjust = 0.5)) + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H090_W10_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H085_W15_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+## July
+
+heatmap_RK4_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = p_RK4_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H100_W00_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H100W100") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H095_W05_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  #ggtitle("H095W05") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H090_W10_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H090W10") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H085_W15_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H085W15") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_Kfloat_H100_W00_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_path(data = Mean_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H095_W05_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_path(data = Mean_KFloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H090_W10_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Mean_KFloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H085_W15_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H100_W00_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H100_W00_mean_July ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H095_W05_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  theme(plot.title = element_text(hjust = 0.5)) + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H090_W10_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H085_W15_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+## October
+
+heatmap_RK4_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = p_RK4_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H100_W00_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H100W100") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H095_W05_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  #ggtitle("H095W05") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H090_W10_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H090W10") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Min_Kfloat_H085_W15_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_path(data = Min_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  #ggtitle("H085W15") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_Kfloat_H100_W00_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H095_W05_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_path(data = Mean_KFloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H090_W10 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Mean_KFloat_H085_W15_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Mean_KFloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H100_W00_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H095_W05_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  theme(plot.title = element_text(hjust = 0.5)) + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H090_W10_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_Max_KFloat_H085_W15_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_path(data = Max_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
 ```
@@
 
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time for scenarios 17 - 29, for April 2018 (winter).
+The plots are regular hexagons within the density of particles were calculated.
+The mean trajectories for each scenario are also plotted for comparison.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
+
+figure_a_April <- ggarrange(heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April, labels = c("18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
+
+figure_b_April <- ggarrange(heatmap_RK4_April, legend = c("right"), labels = c("17"), font.label = list(size = 11, face = "bold"))
+
+ggarrange(figure_b_April, figure_a_April, ncol = 1, nrow = 2)
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time for scenarios 30 - 42, for July 2018 (winter).
+The plots are regular hexagons within the density of particles were calculated.
+The mean trajectories for each scenario are also plotted for comparison.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
+
+figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
+
+figure_b_July <- ggarrange(heatmap_RK4_July, legend = c("right"), labels = c("30"), font.label = list(size = 11, face = "bold"))
+
+ggarrange(figure_b_July, figure_a_July, ncol = 1, nrow = 2)
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 10: Comparison of density plots at the end run time for scenarios 43 - 55, for October 2018 (spring).
+The plots are regular hexagons within the density of particles were calculated.
+The mean trajectories for each scenario are also plotted for comparison.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
+
+figure_a_Oct <- ggarrange(heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
+
+figure_b_Oct <- ggarrange(heatmap_RK4_Oct, legend = c("right"), labels = c("43"), font.label = list(size = 11, face = "bold"))
+
+ggarrange(figure_b_Oct, figure_a_Oct, ncol = 1, nrow = 2)
+
+```
+
 ## Comparison of distances
@@
 Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations.
+When comparing distances between passive and kelp particles for April, significant differences are evident.
+(Figure 11, simulations 17-29; p < 0.05).
+Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas travelled shorter distance across compared to other kelp particles.
+The shortest distance travelled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
+No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; p > 0.05).
+In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; p < 0.05).
+Kelp particles with only 90% hydrodynamic drag across cross-sectional areas travelled shorter distance compared to all other particles.
+As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure travelled the shortest distance compared to all other particles.
 
@@
 
+pos_April <- c("RK4_April",             
+         "Min_Kfloat_H100_W00_April",
+         "Min_Kfloat_H095_W05_April",
+         "Min_Kfloat_H090_W10_April",
+         "Min_Kfloat_H085_W15_April",
+         "Mean_Kfloat_H100_W00_April",
+         "Mean_Kfloat_H095_W05_April",
+         "Mean_Kfloat_H090_W10_April",
+         "Mean_Kfloat_H085_W15_April",
+         "Max_Kfloat_H100_W00_April",
+         "Max_Kfloat_H095_W05_April",
+         "Max_Kfloat_H090_W10_April",
+         "Max_Kfloat_H085_W15_April")
+
+dis_plot_April <- ggplot(data = total_data_April, aes(x = type, y = distance, show.legend = FALSE))+
+  geom_boxplot(data = p_RK4_April, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(limits = pos)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Scenario")+
+  #coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+custom_labs_April <- c("17","18","19","20","21","22","23","24","25","26","27","28", "29") #create custom labels and add them to dis_plot
+
+dis_plot_April <- dis_plot_April +
+  scale_x_discrete(labels= custom_labs_April)
+
+pos_July <- c("RK4_July",             
+         "Min_Kfloat_H100_W00_July",
+         "Min_Kfloat_H095_W05_July",
+         "Min_Kfloat_H090_W10_July",
+         "Min_Kfloat_H085_W15_July",
+         "Mean_Kfloat_H100_W00_July",
+         "Mean_Kfloat_H095_W05_July",
+         "Mean_Kfloat_H090_W10_July",
+         "Mean_Kfloat_H085_W15_July",
+         "Max_Kfloat_H100_W00_July",
+         "Max_Kfloat_H095_W05_July",
+         "Max_Kfloat_H090_W10_July",
+         "Max_Kfloat_H085_W15_July")
+
+dis_plot_July <- ggplot(data = total_data_July, aes(x = type, y = distance, show.legend = FALSE))+
+  geom_boxplot(data = p_RK4_July, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(limits = pos)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Scenario")+
+  #coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+custom_labs_July <- c("30","31","32","33","34","35","36","37","38","39","40","41", "42") #create custom labels and add them to dis_plot
+
+dis_plot_July <- dis_plot_July +
+  scale_x_discrete(labels= custom_labs_July)
+
+pos_Oct <- c("RK4_Oct",              
+         "Min_Kfloat_H100_W00_Oct",
+         "Min_Kfloat_H095_W05_Oct",
+         "Min_Kfloat_H090_W10_Oct",
+         "Min_Kfloat_H085_W15_Oct",
+         "Mean_Kfloat_H100_W00_Oct",
+         "Mean_Kfloat_H095_W05_Oct",
+         "Mean_Kfloat_H090_W10_Oct",
+         "Mean_Kfloat_H085_W15_Oct",
+         "Max_Kfloat_H100_W00_Oct",
+         "Max_Kfloat_H095_W05_Oct",
+         "Max_Kfloat_H090_W10_Oct",
+         "Max_Kfloat_H085_W15_Oct")
+
+dis_plot_Oct <- ggplot(data = total_data_Oct, aes(x = type, y = distance, show.legend = FALSE))+
+  geom_boxplot(data = p_RK4_Oct, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(limits = pos)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Scenario")+
+  #coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+custom_labs_Oct <- c("43","44","45","46","47","48","49","50","51","52","53","54","55") #create custom labels and add them to dis_plot
+
+dis_plot_Oct <- dis_plot_Oct +
+  scale_x_discrete(labels= custom_labs_Oct)
+
 ```
@@
 
-dis_plot
+ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 3)
+
 ggsave(filename = "figure_4.jpg", plot = last_plot(), device = "jpeg", path = "figures/chapter_3/", width = 8, height = 5, dpi = 600)

### 2021-03-30 thesis_chapter_3.Rmd
corrections 3....

@@
 The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (autumn).
-Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag travelled further northward compared to the passive particles (see Figure 7A-B).
+Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag travelled further northward compared to the passive particles (see Figure 2C-D).
 The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles.
 Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles travelled less distance compared to kelp particles with both wind and hydrodynamic drag.
-Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 7C-D).
+Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 2E-F).
 Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore; resulting in entrainment within an eddy.
@@
 In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles travelled initially southward and then westward offshore and also becoming entrained in an eddy.
-In the October (Spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 7E-F).
+In the October (Spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 7F-G).
 Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag ended a greater distance from the release site compared to the passive particles.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Comparison of mean trajectories.
-Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=6, fig.cap="Figure 2: Comparison of mean trajectories.
+Panel (A) represents hydrodynamic drag only simulations and Panel (B) represents simulations for both wind and hydrodynamic drag, for January, Panel (C) represents hydrodynamic drag only simulations and Panel (D) represents simulations for both wind and hydrodynamic drag, for April.
+Panel (E) represents hydrodynamic drag only simulations and Panel (F) represents simulations for both wind and hydrodynamic drag, for July.
+Panel (G) represents hydrodynamic drag only simulations and Panel (H) represents simulations for both wind and hydrodynamic drag, for October.
+In all panels the passive simulations are represented by black lines."}
 
@@
 Across all simulations particles flowed in a north-westward direction (Figure 3).
-The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids (Figure 3A, 3E, 3I).
+The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids.
 In addition the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 3).
@@
 
-In April and across all simulations, particles flowed in a north-westward direction (see Figure 8).
+In April and across all simulations, particles flowed in a north-westward direction (see Figure 4).
 The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation.
@@
 Only slight variations across cross-sectional areas and drag exposure scenarios are evident.
-Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in a eddy located westward of the release site compared to passive particles (see Figure 9).
+Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in a eddy located westward of the release site compared to passive particles (see Figure 5).
 Variation in density along the mean trajectory and eddy for all kelp particles is evident.
@@
 October simulations showed no differences in density of particles along the mean trajectory when comparing passive and kelp particles.
-Only slight variations in density and trajectory can be seen with no entrainment (see Figure 10).
+Only slight variations in density and trajectory can be seen with no entrainment (see Figure 6).
 Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 8: Comparison of density plots at the end run time for scenarios 17 - 29, for April 2018 (winter).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for scenarios 17 - 29, for April 2018 (winter).
 The plots are regular hexagons within the density of particles were calculated.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 9: Comparison of density plots at the end run time for scenarios 30 - 42, for July 2018 (winter).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end run time for scenarios 30 - 42, for July 2018 (winter).
 The plots are regular hexagons within the density of particles were calculated.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 10: Comparison of density plots at the end run time for scenarios 43 - 55, for October 2018 (spring).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end run time for scenarios 43 - 55, for October 2018 (spring).
 The plots are regular hexagons within the density of particles were calculated.
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for scenarios 1 - 13, for January 2018 (summer).
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 7: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for scenarios 1 - 13, for January 2018 (summer).
 The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.
@@
 
-ggsave(filename = "figure_4.jpg", plot = last_plot(), device = "jpeg", path = "figures/chapter_3/", width = 8, height = 5, dpi = 600)
+ggsave(filename = "figure_7.jpg", plot = last_plot(), device = "jpeg", path = "figures/chapter_3/", width = 8, height = 5, dpi = 600)
 
@@
 
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5: Comparison of trajectory data between passive and kelp accumulation scenarios.
+```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of trajectory data between passive and kelp accumulation scenarios.
 Panel A are a comparison of density plots  and panel B is a comparison of age.", fig.height=8, fig.width=8}
@@
 The role of eddies that exist within the study domain over the time period play a clear role in the accumulation and dispersal of kelp particles.
-Passive particles tend to remain on the outer edge of the eddy and then are expelled on the other side and continue to follow the ocean current (see Figure 5A).
+Passive particles tend to remain on the outer edge of the eddy and then are expelled on the other side and continue to follow the ocean current (see Figure 8A).
 The opposite is true for kelp particles which have the highest exposure level of hydrodynamic and wind drag.
-The kelp particles are entrained in the eddy (see Figure 5A).
+The kelp particles are entrained in the eddy (see Figure 8A).
 This is also reflected when comparing the age of particles at the end of the simulation.
-In Figure 5B, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
+In Figure 8B, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
 
@@
 
-The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (Figure 6A).
-This was also reflected when comparing distance travelled by particles and particle density plots (Figure 6B).
-The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (Figure 6C-D).
+The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (Figure 9A).
+This was also reflected when comparing distance travelled by particles and particle density plots (Figure 9B).
+The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (Figure 9C-D).
 
@@
 
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 6: Comparison of trajectory data between a kelp sphere and kelp cylinder.
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of trajectory data between a kelp sphere and kelp cylinder.
 Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylindrical shape coefficient.
@@
 In addition, not only are the trajectories similar but so are the end points at the end of simulation.
-However, when *ad-hoc* wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely Lagrangian particle trajectories.
+However, when *ad hoc* wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely Lagrangian particle trajectories.
 The inclusion of wind drag causes the kelp particles to flow further away from the coast and with a greater distance travelled from the release location.
@@
 Past studies have shown this when including windage as a function of inertia [@putman2018;@brooks2019; @putman2020].
-A study by @putman2020 used *ad-hoc* windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic *Sargassum sp*.
+A study by @putman2020 used *ad hoc* windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic *Sargassum sp*.
 The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked *Sargassum sp* mats which was partly due to the inclusion of momentum energy transfer from wind.
@@
 The particles in the simulations that included wind drag traveled significantly greater distances compared to the passive or hydrodynamic only drag.
-Similar patterns are seen in the additional simulations run for April (autumn), July (winter), and October (spring), however even kelp particles with only hydrodynamic drag flow further compared to passive particles in these times of the year.
+Similar patterns are seen in the additional simulations run for April (autumn), July (winter), and October (spring); however even kelp particles with only hydrodynamic drag flow further compared to passive particles in these times of the year.
 This may be a result of the kelp particles exposed to different time-varying flows, which results in the kelp particles following a different surface current vector compared to passive particles.
@@
 The entrainment and expulsion from eddies and vortices are an important characteristic of floating macroalgae.
-Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts.
+Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real *Sargassum* rafts.
 However, these studies investigated large rafts and not floating individuals.
@@
 Eddies and vortices may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in area below the eddy.
-In a study by @woodborne1989 references a site offshore approximately 120km from Cape Columbine where dredged kelp was found.
-The area referenced by @woodborne1989 coincides with the area where virtual kelp are entrained by a eddy.
+In a study by @woodborne1989 references a site offshore approximately 120 km from Cape Columbine where dredged kelp was found.
+The area referenced by @woodborne1989 coincides with the area where virtual kelp are entrained by an eddy.
 The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory.
@@
 Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies.
-Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E.maxima*.
+Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E.
+maxima*.
 In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory, but it does cause higher density of particles to become entrained in an eddy.

### 2021-03-30 appendix_chapter_3.Rmd
appendix corrections

@@
 When comparing distances between passive and kelp particles for April, significant differences are evident.
-(Figure 11, simulations 17-29; p < 0.05).
+(Figure 11, simulations 17-29; *p* < 0.05).
 Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas travelled shorter distance across compared to other kelp particles.
 The shortest distance travelled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
-No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; p > 0.05).
-In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; p < 0.05).
+No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; *p* > 0.05).
+In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; *p* < 0.05).
 Kelp particles with only 90% hydrodynamic drag across cross-sectional areas travelled shorter distance compared to all other particles.

### 2021-04-06 thesis_chapter_3.Rmd
accumulation scripted updated

@@
 
+# January
+
 kelp_acc <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc.nc") %>%
@@
 
-total_acc_data <- rbind(pass_acc, kelp_acc)
-total_acc_mean <- rbind(pass_acc_mean, kelp_acc_mean)
-
 mean_traj_acc <- ggplot() +
@@
 
+# April
+
+kelp_acc_April <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+kelp_acc_mean_April <- kelp_acc_April %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+kelp_acc_April$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_mean_April$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_April$age <- with(kelp_acc_April, age/3600)
+
+pass_acc_April <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+pass_acc_mean_April <- pass_acc_April %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+pass_acc_April$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_mean_April$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_April$age <- with(pass_acc, age/3600)
+
+total_acc_data_April <- rbind(pass_acc_April, kelp_acc_April)
+total_acc_mean_April <- rbind(pass_acc_mean_April, kelp_acc_mean_April)
+
+mean_traj_acc_April <- ggplot() +
+  geom_path(data = total_acc_mean_April, aes(x = lon, y = lat, colour = type)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+# July
+
+kelp_acc_July <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+kelp_acc_mean_July <- kelp_acc_July %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+kelp_acc_July$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_mean_July$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_July$age <- with(kelp_acc_July, age/3600)
+
+pass_acc_July <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+pass_acc_mean_July <- pass_acc_July %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+pass_acc_July$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_mean_July$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_July$age <- with(pass_acc_July, age/3600)
+
+total_acc_data_July <- rbind(pass_acc_July, kelp_acc_July)
+total_acc_mean_July <- rbind(pass_acc_mean_July, kelp_acc_mean_July)
+
+mean_traj_acc_July <- ggplot() +
+  geom_path(data = total_acc_mean_July, aes(x = lon, y = lat, colour = type)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+# October
+
+kelp_acc_Oct <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+kelp_acc_mean_Oct <- kelp_acc_Oct %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+kelp_acc_Oct$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_mean_Oct$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_Oct$age <- with(kelp_acc_Oct, age/3600)
+
+pass_acc_Oct <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+pass_acc_mean_Oct <- pass_acc_Oct %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+pass_acc_Oct$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_mean_Oct$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_Oct$age <- with(pass_acc_Oct, age/3600)
+
+total_acc_data_Oct <- rbind(pass_acc_Oct, kelp_acc_Oct)
+total_acc_mean_Oct <- rbind(pass_acc_mean_Oct, kelp_acc_mean_Oct)
+
+mean_traj_acc_Oct <- ggplot() +
+  geom_path(data = total_acc_mean_Oct, aes(x = lon, y = lat, colour = type)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
 ## Coastline of Southern Africa
@@
 
+# January
+
 heatmap_RK4_acc <- ggplot() +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
   geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
   geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
-  #scale_fill_continuous(type = "gradient", limits = c(0, 30)) +
+  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
@@
   theme(plot.title = element_text(hjust = 0.5))
+
+# April
+
+heatmap_RK4_acc_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = pass_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = pass_acc_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Passive particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_kelp_acc_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = kelp_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = kelp_acc_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Kelp particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_RK4_acc_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = pass_acc_April, aes(x = lon, y = lat, colour = distance)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_kelp_acc_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = kelp_acc_April, aes(x = lon, y = lat, colour = age)) +
+  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Age")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# July
+
+heatmap_RK4_acc_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = pass_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = pass_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Passive particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_kelp_acc_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = kelp_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = kelp_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Kelp particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_RK4_acc_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = pass_acc_July, aes(x = lon, y = lat, colour = distance)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_kelp_acc_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = kelp_acc_July, aes(x = lon, y = lat, colour = age)) +
+  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Age")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# October
+
+heatmap_RK4_acc_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = pass_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = pass_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Passive particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_kelp_acc_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = kelp_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = kelp_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Kelp particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_RK4_acc_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = pass_acc_Oct, aes(x = lon, y = lat, colour = distance)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_kelp_acc_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = kelp_acc_Oct, aes(x = lon, y = lat, colour = age)) +
+  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Age")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+total_acc_data <- rbind(pass_acc, kelp_acc, pass_acc_April, kelp_acc_April,pass_acc_July, kelp_acc_July,pass_acc_Oct, kelp_acc_Oct)
+total_acc_mean <- rbind(pass_acc_mean, kelp_acc_mean, pass_acc_mean_April, kelp_acc_mean_April, pass_acc_mean_July, kelp_acc_mean_July, pass_acc_mean_Oct, kelp_acc_mean_Oct)
+
+# Distance
 
@@
 
-plot_A <- ggarrange(heatmap_RK4_acc, heatmap_kelp_acc, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 1)
-plot_B <- ggarrange(point_RK4_acc, point_kelp_acc, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 1)
-
-ggarrange(plot_A, plot_B, labels = c("AUTO"), ncol = 1, nrow = 2)
-ggsave(filename = "figure_5.jpg", plot = last_plot(), device = "jpeg", path = "figures/chapter_3/", width = 7, height = 7, dpi = 600)
+ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4)
 
@@
 
+```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of trajectory data between passive and kelp accumulation scenarios.
+Panel A are a comparison of density plots  and panel B is a comparison of age.", fig.height=8, fig.width=8}
+
+ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4)
+
+```
+
 The role of eddies that exist within the study domain over the time period play a clear role in the accumulation and dispersal of kelp particles.
-Passive particles tend to remain on the outer edge of the eddy and then are expelled on the other side and continue to follow the ocean current (see Figure 8A).
+Passive particles tend to remain on the outer edge of the eddy and then are expelled on the other side and continue to follow the ocean current (see Figure 8).
 The opposite is true for kelp particles which have the highest exposure level of hydrodynamic and wind drag.
@@
 This is also reflected when comparing the age of particles at the end of the simulation.
-In Figure 8B, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
+In Figure 9, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
 
@@
 
-The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (Figure 9A).
-This was also reflected when comparing distance travelled by particles and particle density plots (Figure 9B).
-The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (Figure 9C-D).
+The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (Figure 10).
+This was also reflected when comparing distance travelled by particles and particle density plots (Figure 10).
+The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (Figure 10).
 

### 2021-04-07 thesis_chapter_3.Rmd
updates

@@
 output:
-  word_document:
-    reference_docx: markdown/frontiers_template.docx
   html_document:
@@
     latex_engine: xelatex
+  word_document:
+    reference_docx: markdown/frontiers_template.docx
 bibliography: references.bib
@@
 
-#library(computeMSD)
 library(ggplot2)
@@
 library(ggpubr)
-library(ggord)
 library(mapproj)
@@
 library(maps)
-library(flextable)
+#library(flextable)
 library(officer)
@@
 These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018].
+The time period used in all simulations were January (summer), April (Autumn), July (winter), and October (Spring) of 2018.
+These time periods were chosen based on variable mesoscale activities and stability of the Bengulea Jet that occurs through the year.
+For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity ; while in winter the Benguela jet begins to stablize until summer when the begins to split again [@rubio2009; @veitch2018; @ragoasha2019].
+In order to determine if the shape co-efficient used plays a role in model used, a simulation of a kelp with the maximum surface area and drag exposure was run using the shape coefficient for a cylinder to compare to a simulation using a sphere, and was run for 30 days using a fourth order Runge-Kutte advection scheme with brownian motion and a particle set of size 1000.
+
 To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios.
@@
 This was not considered a concern, as once again, the aim was to investigate surface drift behavior.
-The time period chosen to run the simulations was 01-01-2018 un 30-01-2018.
-This particular time period was chosen as the splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity [@rubio2009; @veitch2018; @ragoasha2019].
-Additional simulations were done for months in 2018, which were chosen based on seasonality.
-These simulations had the same parameters used for January and were performed for April (Autumn), July (winter), and October (Spring) in order to assess if any patterns seen in January translate into general patterns for times periods throughout 2018.
-A summary of the parameters for each simulation are provided in table 1.
-
-```{r sim table, echo=FALSE, message=FALSE, warning=FALSE}
+A summary of the parameters for each simulation are provided in table 1.To determine if vortices and eddies play a role in kelp accumulation on the surface, a separate set of simulations were performed.
+Unlike the previous simulations these simulation released a particle every hour over the same time periods.
+
+```{r sim table, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 ```
-
-To determine if vortices and eddies play a role in kelp accumulation on the surface, a separate set of simulations were performed.
-Unlike the previous simulations these simulation released a particle every hour over the same time periods.
-In order to determine if the shape co-efficient used plays a role in model used, a simulation of a kelp with the maximum surface area and drag exposure was run using the shape coefficient for a cylinder to compare to a simulation using a sphere, and was run for 30 days using a fourth order Runge-Kutte advection scheme with brownian motion and a particle set of size 1000.
 
@@
 
-```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
+```{r morphology table, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, eval=FALSE}
 
@@
 # Results
+## Comparison of coefficients
+
+The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (see appendix).
+This was also reflected when comparing distance travelled by particles and particle density plots (see appendix).
+The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (see appendix).
 
@@
   geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
   geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
   geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient",limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
   geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
+  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1100)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = p_RK4_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H100_W00_mean_April ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = p_RK4_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H100_W00_mean_July ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = p_RK4_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = Max_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
 The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
-
-figure_a <- ggarrange(heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, labels = c("2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
-
-figure_b <- ggarrange(heatmap_RK4, legend = c("right"), labels = c("1"), font.label = list(size = 11, face = "bold"))
-
-ggarrange(figure_b, figure_a, ncol = 1, nrow = 2)
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=7}
+
+ggarrange(heatmap_RK4, heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, labels = c("1","2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("bottom"), font.label = list(size = 11, face = "bold"))
 
@@
 The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
-
-figure_a_April <- ggarrange(heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April, labels = c("18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
-
-figure_b_April <- ggarrange(heatmap_RK4_April, legend = c("right"), labels = c("17"), font.label = list(size = 11, face = "bold"))
-
-ggarrange(figure_b_April, figure_a_April, ncol = 1, nrow = 2)
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=7}
+
+ggarrange(heatmap_RK4_April,heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April, labels = c("17", "18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("bottom"), font.label = list(size = 11, face = "bold"))
 
@@
 
-figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
-
-figure_b_July <- ggarrange(heatmap_RK4_July, legend = c("right"), labels = c("30"), font.label = list(size = 11, face = "bold"))
-
-ggarrange(figure_b_July, figure_a_July, ncol = 1, nrow = 2)
+ggarrange(heatmap_RK4_July,heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("30","31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("bottom"), font.label = list(size = 11, face = "bold"))
 
@@
 The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
-
-figure_a_Oct <- ggarrange(heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
-
-figure_b_Oct <- ggarrange(heatmap_RK4_Oct, legend = c("right"), labels = c("43"), font.label = list(size = 11, face = "bold"))
-
-ggarrange(figure_b_Oct, figure_a_Oct, ncol = 1, nrow = 2)
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=7}
+
+ggarrange(heatmap_RK4_Oct, heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("43","44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("bottom"), font.label = list(size = 11, face = "bold"))
 
@@
 
-ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 3)
-
-ggsave(filename = "figure_7.jpg", plot = last_plot(), device = "jpeg", path = "figures/chapter_3/", width = 8, height = 5, dpi = 600)
+ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 4)
 
@@
 
-mean_traj_acc <- ggplot() +
-  geom_path(data = total_acc_mean, aes(x = lon, y = lat, colour = type)) +
+# April
+
+kelp_acc_April <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+kelp_acc_mean_April <- kelp_acc_April %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+kelp_acc_April$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_mean_April$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_April$age <- with(kelp_acc_April, age/3600)
+
+pass_acc_April <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+pass_acc_mean_April <- pass_acc_April %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+pass_acc_April$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_mean_April$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_April$age <- with(pass_acc, age/3600)
+
+total_acc_data_April <- rbind(pass_acc_April, kelp_acc_April)
+total_acc_mean_April <- rbind(pass_acc_mean_April, kelp_acc_mean_April)
+
+mean_traj_acc_April <- ggplot() +
+  geom_path(data = total_acc_mean_April, aes(x = lon, y = lat, colour = type)) +
   xlab("Longtitude") +
@@
 
-# April
-
-kelp_acc_April <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_April.nc") %>%
+# July
+
+kelp_acc_July <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_July.nc") %>%
   hyper_tibble() %>%
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-kelp_acc_mean_April <- kelp_acc_April %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-kelp_acc_April$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_mean_April$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_April$age <- with(kelp_acc_April, age/3600)
-
-pass_acc_April <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_April.nc") %>%
+kelp_acc_mean_July <- kelp_acc_July %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+kelp_acc_July$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_mean_July$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_July$age <- with(kelp_acc_July, age/3600)
+
+pass_acc_July <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_July.nc") %>%
   hyper_tibble() %>%
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-pass_acc_mean_April <- pass_acc_April %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-pass_acc_April$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_mean_April$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_April$age <- with(pass_acc, age/3600)
-
-total_acc_data_April <- rbind(pass_acc_April, kelp_acc_April)
-total_acc_mean_April <- rbind(pass_acc_mean_April, kelp_acc_mean_April)
-
-mean_traj_acc_April <- ggplot() +
-  geom_path(data = total_acc_mean_April, aes(x = lon, y = lat, colour = type)) +
+pass_acc_mean_July <- pass_acc_July %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+pass_acc_July$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_mean_July$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_July$age <- with(pass_acc_July, age/3600)
+
+total_acc_data_July <- rbind(pass_acc_July, kelp_acc_July)
+total_acc_mean_July <- rbind(pass_acc_mean_July, kelp_acc_mean_July)
+
+mean_traj_acc_July <- ggplot() +
+  geom_path(data = total_acc_mean_July, aes(x = lon, y = lat, colour = type)) +
   xlab("Longtitude") +
@@
 
-# July
-
-kelp_acc_July <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_July.nc") %>%
+# October
+
+kelp_acc_Oct <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_Oct.nc") %>%
   hyper_tibble() %>%
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-kelp_acc_mean_July <- kelp_acc_July %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-kelp_acc_July$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_mean_July$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_July$age <- with(kelp_acc_July, age/3600)
-
-pass_acc_July <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_July.nc") %>%
+kelp_acc_mean_Oct <- kelp_acc_Oct %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+kelp_acc_Oct$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_mean_Oct$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_acc_Oct$age <- with(kelp_acc_Oct, age/3600)
+
+pass_acc_Oct <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_Oct.nc") %>%
   hyper_tibble() %>%
@@
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-pass_acc_mean_July <- pass_acc_July %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-pass_acc_July$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_mean_July$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_July$age <- with(pass_acc_July, age/3600)
-
-total_acc_data_July <- rbind(pass_acc_July, kelp_acc_July)
-total_acc_mean_July <- rbind(pass_acc_mean_July, kelp_acc_mean_July)
-
-mean_traj_acc_July <- ggplot() +
-  geom_path(data = total_acc_mean_July, aes(x = lon, y = lat, colour = type)) +
+pass_acc_mean_Oct <- pass_acc_Oct %>% #calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+pass_acc_Oct$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_mean_Oct$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
+pass_acc_Oct$age <- with(pass_acc_Oct, age/3600)
+
+total_acc_data_Oct <- rbind(pass_acc_Oct, kelp_acc_Oct)
+total_acc_mean_Oct <- rbind(pass_acc_mean_Oct, kelp_acc_mean_Oct)
+
+mean_traj_acc_Oct <- ggplot() +
+  geom_path(data = total_acc_mean_Oct, aes(x = lon, y = lat, colour = type)) +
   xlab("Longtitude") +
@@
 
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+# January
+
+heatmap_RK4_acc <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Passive particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_kelp_acc <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Kelp particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_RK4_acc <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = pass_acc, aes(x = lon, y = lat, colour = distance)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_kelp_acc <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
+  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Age")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# April
+
+heatmap_RK4_acc_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = pass_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = pass_acc_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Passive particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_kelp_acc_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = kelp_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = kelp_acc_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Kelp particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_RK4_acc_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = pass_acc_April, aes(x = lon, y = lat, colour = distance)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_kelp_acc_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = kelp_acc_April, aes(x = lon, y = lat, colour = age)) +
+  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Age")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# July
+
+heatmap_RK4_acc_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = pass_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = pass_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Passive particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_kelp_acc_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = kelp_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = kelp_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Kelp particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_RK4_acc_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = pass_acc_July, aes(x = lon, y = lat, colour = distance)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_kelp_acc_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = kelp_acc_July, aes(x = lon, y = lat, colour = age)) +
+  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Age")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
 # October
 
-kelp_acc_Oct <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-kelp_acc_mean_Oct <- kelp_acc_Oct %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-kelp_acc_Oct$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_mean_Oct$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_Oct$age <- with(kelp_acc_Oct, age/3600)
-
-pass_acc_Oct <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-pass_acc_mean_Oct <- pass_acc_Oct %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-pass_acc_Oct$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_mean_Oct$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_Oct$age <- with(pass_acc_Oct, age/3600)
-
-total_acc_data_Oct <- rbind(pass_acc_Oct, kelp_acc_Oct)
-total_acc_mean_Oct <- rbind(pass_acc_mean_Oct, kelp_acc_mean_Oct)
-
-mean_traj_acc_Oct <- ggplot() +
-  geom_path(data = total_acc_mean_Oct, aes(x = lon, y = lat, colour = type)) +
+heatmap_RK4_acc_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = pass_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = pass_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Passive particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_kelp_acc_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = kelp_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = kelp_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Kelp particles") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_RK4_acc_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = pass_acc_Oct, aes(x = lon, y = lat, colour = distance)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+point_kelp_acc_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_point(data = kelp_acc_Oct, aes(x = lon, y = lat, colour = age)) +
+  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Age")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+total_acc_data <- rbind(pass_acc, kelp_acc, pass_acc_April, kelp_acc_April,pass_acc_July, kelp_acc_July,pass_acc_Oct, kelp_acc_Oct)
+total_acc_mean <- rbind(pass_acc_mean, kelp_acc_mean, pass_acc_mean_April, kelp_acc_mean_April, pass_acc_mean_July, kelp_acc_mean_July, pass_acc_mean_Oct, kelp_acc_mean_Oct)
+
+mean_traj_acc <- ggplot() +
+  geom_path(data = total_acc_mean, aes(x = lon, y = lat, colour = type)) +
   xlab("Longtitude") +
@@
 
-## Coastline of Southern Africa
-load("map/south_africa_coast_hi_res.Rdata")
-
-# January
-
-heatmap_RK4_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Passive particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_kelp_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Kelp particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_RK4_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = pass_acc, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_kelp_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
-  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Age")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-# April
-
-heatmap_RK4_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_path(data = pass_acc_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Passive particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_kelp_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_path(data = kelp_acc_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Kelp particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_RK4_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = pass_acc_April, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_kelp_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = kelp_acc_April, aes(x = lon, y = lat, colour = age)) +
-  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Age")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-# July
-
-heatmap_RK4_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_path(data = pass_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Passive particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_kelp_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_path(data = kelp_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Kelp particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_RK4_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = pass_acc_July, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_kelp_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = kelp_acc_July, aes(x = lon, y = lat, colour = age)) +
-  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Age")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-# October
-
-heatmap_RK4_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_path(data = pass_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Passive particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_kelp_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_path(data = kelp_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Kelp particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_RK4_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = pass_acc_Oct, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-point_kelp_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = kelp_acc_Oct, aes(x = lon, y = lat, colour = age)) +
-  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Age")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-total_acc_data <- rbind(pass_acc, kelp_acc, pass_acc_April, kelp_acc_April,pass_acc_July, kelp_acc_July,pass_acc_Oct, kelp_acc_Oct)
-total_acc_mean <- rbind(pass_acc_mean, kelp_acc_mean, pass_acc_mean_April, kelp_acc_mean_April, pass_acc_mean_July, kelp_acc_mean_July, pass_acc_mean_Oct, kelp_acc_mean_Oct)
-
 # Distance
 
-dis_plot_acc <- ggplot(data = total_acc_data, aes(x = type, y = distance, show.legend = FALSE)) +
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+#dis_plot_acc <- ggplot(data = total_acc_data, aes(x = type, y = distance, show.legend = FALSE)) +
+#  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
   #scale_x_discrete(limits = pos)+
-  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = "right") +
-  theme_bw() +
-  ylab("Distance (km)") +
-  xlab("Scenario") +
+#  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = "right") +
+#  theme_bw() +
+#  ylab("Distance (km)") +
+#  xlab("Scenario") +
   #coord_flip() +
-  theme_bw()
+#  theme_bw()
   #theme(plot.title = element_text(hjust = 1))
@@
 
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of trajectory data between passive and kelp accumulation scenarios.
-Panel A are a comparison of density plots  and panel B is a comparison of age.", fig.height=8, fig.width=8}
-
-ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4)
+```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation scenarios.", fig.height=9, fig.width=7}
+
+ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 11, face = "bold"), align = "hv")
 
@@
 
-```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of trajectory data between passive and kelp accumulation scenarios.
-Panel A are a comparison of density plots  and panel B is a comparison of age.", fig.height=8, fig.width=8}
-
-ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4)
+```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of age between passive and kelp accumulation scenarios.", fig.height=9, fig.width=7}
+
+ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 11, face = "bold"), align = "hv")
 
@@
 The opposite is true for kelp particles which have the highest exposure level of hydrodynamic and wind drag.
-The kelp particles are entrained in the eddy (see Figure 8A).
+The kelp particles are entrained in the eddy (see Figure 8).
 This is also reflected when comparing the age of particles at the end of the simulation.
 In Figure 9, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
-
-## Comparison of coefficients
-
-The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (Figure 10).
-This was also reflected when comparing distance travelled by particles and particle density plots (Figure 10).
-The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (Figure 10).
-
-```{r shape data, message=FALSE, echo=FALSE, warning=FALSE}
-
-sphere <- tidync("OceanParcels/trajectory_data/Max_KFloat_H100_W00.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-sphere_mean <- sphere %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-sphere$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
-sphere_mean$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-# read in data again for one max simulation, as cannot sort issue out with rename df...
-cyl <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-cyl_mean <- cyl %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-cyl$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
-cyl_mean$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-# Make mean trajectory plot
-shape_plot <- ggplot()+
-  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = sphere_mean, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = cyl_mean, aes(x = lon, y = lat), color = 'orange')+
-  xlab("Longtitude") + 
-  ylab("Latitude") +
-  ggtitle("Mean trajectory comparison") +
-  #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-# make density map
-heatmap_sphere <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = sphere, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = sphere_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("Cylinder") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_cyl <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = cyl, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
-  geom_path(data = cyl_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("Sphere") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-shape_data_total <- rbind(sphere, cyl)
-
-shape_dis_plot <- ggplot(data = shape_data_total, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
-  stat_compare_means(method = "kruskal.test", label = "p.signif")+
-  ylab("Distance (km)")+
-  xlab("Type") + 
-  ggtitle("Comparison of distance travelled") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-```
-
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of trajectory data between a kelp sphere and kelp cylinder.
-Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylindrical shape coefficient.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=8}
-
-ggarrange(shape_plot, shape_dis_plot, heatmap_sphere, heatmap_cyl, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 2)
-
-```
 
@@
 # References
+
+# Appendix
+
+```{r shape data, message=FALSE, echo=FALSE, warning=FALSE}
+
+# January
+
+sphere <- tidync("OceanParcels/trajectory_data/Max_KFloat_H100_W00.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+sphere_mean <- sphere %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+sphere$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+sphere_mean$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+# read in data again for one max simulation, as cannot sort issue out with rename df...
+cyl <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+cyl_mean <- cyl %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+cyl$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+cyl_mean$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+# Make mean trajectory plot
+shape_plot <- ggplot()+
+  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = sphere_mean, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = cyl_mean, aes(x = lon, y = lat), color = 'orange')+
+  xlab("Longtitude") + 
+  ylab("Latitude") +
+  ggtitle("Mean trajectory comparison") +
+  #guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# make density map
+heatmap_sphere <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = sphere, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = sphere_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("Cylinder") + 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_cyl <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = cyl, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = cyl_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Sphere") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# April
+
+sphere_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+sphere_mean_April <- sphere_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+sphere_April$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+sphere_mean_April$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+# read in data again for one max simulation, as cannot sort issue out with rename df...
+cyl_April <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00_April.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+cyl_mean_April <- cyl_April %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+cyl_April$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+cyl_mean_April$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+# Make mean trajectory plot
+shape_plot_April <- ggplot()+
+  geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = sphere_mean_April, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = cyl_mean_April, aes(x = lon, y = lat), color = 'orange')+
+  xlab("Longtitude") + 
+  ylab("Latitude") +
+  ggtitle("Mean trajectory comparison") +
+  #guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# make density map
+heatmap_sphere_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = sphere_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = sphere_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("Cylinder") + 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_cyl_April <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = cyl_April, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = cyl_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Sphere") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# July
+
+sphere_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+sphere_mean_July <- sphere_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+sphere_July$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+sphere_mean_July$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+# read in data again for one max simulation, as cannot sort issue out with rename df...
+cyl_July <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00_July.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+cyl_mean_July <- cyl_July %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+cyl_July$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+cyl_mean_July$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+# Make mean trajectory plot
+shape_plot_July <- ggplot()+
+  geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = sphere_mean_July, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = cyl_mean_July, aes(x = lon, y = lat), color = 'orange')+
+  xlab("Longtitude") + 
+  ylab("Latitude") +
+  ggtitle("Mean trajectory comparison") +
+  #guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# make density map
+heatmap_sphere_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = sphere_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = sphere_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("Cylinder") + 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_cyl_July <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = cyl_July, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = cyl_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Sphere") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# October
+
+sphere_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+sphere_mean_Oct <- sphere_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+sphere_Oct$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+sphere_mean_Oct$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+# read in data again for one max simulation, as cannot sort issue out with rename df...
+cyl_Oct <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00_Oct.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+cyl_mean_Oct <- cyl_Oct %>% #calculate mean trajectory 
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+cyl_Oct$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+cyl_mean_Oct$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+# Make mean trajectory plot
+shape_plot_Oct <- ggplot()+
+  geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = sphere_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
+  geom_path(data = cyl_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
+  xlab("Longtitude") + 
+  ylab("Latitude") +
+  ggtitle("Mean trajectory comparison") +
+  #guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# make density map
+heatmap_sphere_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = sphere_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = sphere_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("Cylinder") + 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_cyl_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = cyl_Oct, bins = 80, aes(x = lon, y = lat)) +
+  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_path(data = cyl_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Sphere") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# Distance plot
+
+shape_data_total <- rbind(sphere, cyl, sphere_April, cyl_April, sphere_July, cyl_July, sphere_Oct, cyl_Oct)
+
+shape_dis_plot <- ggplot(data = shape_data_total, aes(x = type, y = distance, show.legend = FALSE))+
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(limits = pos)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif")+
+  ylab("Distance (km)")+
+  xlab("Type") + 
+  ggtitle("Comparison of distance travelled") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of trajectory data between a kelp sphere and kelp cylinder.
+Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylindrical shape coefficient.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=8}
+
+shape_plot_A <- ggarrange(shape_plot, shape_plot_April, shape_plot_July, shape_plot_Oct, heatmap_sphere, heatmap_cyl,heatmap_sphere_April, heatmap_cyl_April, heatmap_sphere_July, heatmap_cyl_July, heatmap_sphere_Oct, heatmap_cyl_Oct, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 3)
+
+ggarrange(shape_plot_A, shape_dis_plot, ncol = 1, nrow = 2)
+
+```

### 2021-04-07 thesis_chapter_3.Rmd
chapter 3 minor update

@@
 output:
-  html_document:
+  word_document:
+    reference_docx: markdown/frontiers_template.docx
   pdf_document:
@@
     latex_engine: xelatex
-  word_document:
-    reference_docx: markdown/frontiers_template.docx
+  html_document:
 bibliography: references.bib

### 2021-04-09 thesis_chapter_3.Rmd
chapter 3 rmd and output updated

@@
 waterdrag_plot <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
   geom_path(data = total_mean_water, aes(x = lon, y = lat, colour = type))+
@@
 wwdrag_plot <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
   geom_path(data = total_mean_ww, aes(x = lon, y = lat, colour = type))+
@@
 waterdrag_plot_April <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
   geom_path(data = total_mean_water_April, aes(x = lon, y = lat, colour = type))+
@@
 wwdrag_plot_April <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill ="grey") +
+  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
   geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, colour = type))+
@@
 waterdrag_plot_July <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
   geom_path(data = total_mean_water_July, aes(x = lon, y = lat, colour = type))+
@@
 wwdrag_plot_July <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
   geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, colour = type))+
@@
 waterdrag_plot_Oct <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
   geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, colour = type))+
@@
 
-wwdrag_plot_Oct <- ggplot()+
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
+wwdrag_plot_Oct <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
   geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, colour = type))+
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=6, fig.cap="Figure 2: Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=5, fig.cap="Figure 2: Comparison of mean trajectories.
 Panel (A) represents hydrodynamic drag only simulations and Panel (B) represents simulations for both wind and hydrodynamic drag, for January, Panel (C) represents hydrodynamic drag only simulations and Panel (D) represents simulations for both wind and hydrodynamic drag, for April.
@@
 
-ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, labels = "AUTO", ncol = 2, nrow = 4)
+ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, labels = "AUTO", ncol = 2, nrow = 4, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("right"))
 

### 2021-04-09 thesis_chapter_3.Rmd
more polishing...

@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
+  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
+  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  
+  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  
+  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  
+  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
+  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
+  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
   geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
+  geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = p_RK4_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
+  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = p_RK4_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
+  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
   geom_path(data = p_RK4_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
+  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000)) +
   geom_path(data = Min_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
+  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
   geom_path(data = Mean_KFloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000)) +
@@
 The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=7}
-
-ggarrange(heatmap_RK4, heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15, labels = c("1","2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("bottom"), font.label = list(size = 11, face = "bold"))
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}
+
+leg_Jan <- get_legend(heatmap_RK4)
+
+ggarrange(heatmap_RK4, heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15,as_ggplot(leg_Jan), labels = c("1","2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"),font.label = list(size = 11, face = "bold"))
 
@@
 The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=7}
-
-ggarrange(heatmap_RK4_April,heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April, labels = c("17", "18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("bottom"), font.label = list(size = 11, face = "bold"))
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}
+
+leg_April <- get_legend(heatmap_RK4_April)
+
+ggarrange(heatmap_RK4_April,heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April,as_ggplot(leg_April), labels = c("17", "18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
 
@@
 The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}
-
-ggarrange(heatmap_RK4_July,heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("30","31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("bottom"), font.label = list(size = 11, face = "bold"))
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}
+
+leg_July <- get_legend(heatmap_RK4_July)
+
+ggarrange(heatmap_RK4_July,heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, as_ggplot(leg_July),labels = c("30","31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
 
@@
 The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=7}
-
-ggarrange(heatmap_RK4_Oct, heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, labels = c("43","44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("bottom"), font.label = list(size = 11, face = "bold"))
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}
+
+leg_Oct <- get_legend(heatmap_RK4_Oct)
+
+ggarrange(heatmap_RK4_Oct, heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, as_ggplot(leg_Oct),labels = c("43","44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
 
@@
 
-```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE, fig.height=3, fig.width=4}
+```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE}
 
@@
   scale_x_discrete(labels= custom_labs_Oct)
-
 ```
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=7, fig.width=5}
 
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
   geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
   geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = pass_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
   geom_path(data = pass_acc_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = kelp_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
   geom_path(data = kelp_acc_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = pass_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
   geom_path(data = pass_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = kelp_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
   geom_path(data = kelp_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = pass_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
   geom_path(data = pass_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_hex(data = kelp_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
   geom_path(data = kelp_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
 
-sphere$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
-sphere_mean$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+sphere$type <- "Sphere" 
+sphere$Month <- "January" 
+sphere_mean$type <- "Sphere"
 
@@
 
-cyl$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
-cyl_mean$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+cyl$type <- "Cylinder" 
+cyl$Month <- "January" 
+cyl_mean$type <- "Cylinder"
 
@@
 
-sphere_April$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
-sphere_mean_April$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+sphere_April$type <- "Sphere" 
+sphere_April$Month <- "April" 
+sphere_mean_April$type <- "Sphere"
 
@@
 
-cyl_April$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
-cyl_mean_April$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+cyl_April$type <- "Cylinder" 
+cyl_April$Month <- "April" 
+cyl_mean_April$type <- "Cylinder"
 
@@
 
-sphere_July$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
-sphere_mean_July$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+sphere_July$type <- "Sphere" 
+sphere_July$Month <- "July" 
+sphere_mean_July$type <- "Sphere"
 
@@
 cyl_July$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+cyl_July$Month <- "July" # add extra column 'type' to dataframe and assigned 'kelp' label
 cyl_mean_July$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
@@
 
-sphere_Oct$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
-sphere_mean_Oct$type <- "Sphere" # add extra column 'type' to dataframe and assigned 'kelp' label
+sphere_Oct$type <- "Sphere" 
+sphere_Oct$Month <- "October" 
+sphere_mean_Oct$type <- "Sphere"
 
@@
 
-cyl_Oct$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
-cyl_mean_Oct$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+cyl_Oct$type <- "Cylinder" 
+cyl_Oct$Month <- "October"
+cyl_mean_Oct$type <- "Cylinder"
 
@@
 shape_data_total <- rbind(sphere, cyl, sphere_April, cyl_April, sphere_July, cyl_July, sphere_Oct, cyl_Oct)
-
-shape_dis_plot <- ggplot(data = shape_data_total, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
+shape_pos <- c("January", "April", "July", "October")
+shape_dis_plot <- ggplot(data = shape_data_total, aes(x = Month, y = distance, show.legend = TRUE, fill = type))+
+  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = TRUE, stat = "boxplot") +
+  scale_x_discrete(limits = shape_pos)+
   stat_compare_means(method = "kruskal.test", label = "p.signif")+
   ylab("Distance (km)")+
-  xlab("Type") + 
+  xlab("Month") + 
   ggtitle("Comparison of distance travelled") +
@@
 
-shape_plot_A <- ggarrange(shape_plot, shape_plot_April, shape_plot_July, shape_plot_Oct, heatmap_sphere, heatmap_cyl,heatmap_sphere_April, heatmap_cyl_April, heatmap_sphere_July, heatmap_cyl_July, heatmap_sphere_Oct, heatmap_cyl_Oct, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 3)
+shape_plot_A <- ggarrange(shape_plot, shape_plot_April, shape_plot_July, shape_plot_Oct, heatmap_sphere, heatmap_sphere_April,heatmap_sphere_July, heatmap_sphere_Oct, heatmap_cyl, heatmap_cyl_April, heatmap_cyl_July, heatmap_cyl_Oct, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 3)
 

### 2021-04-12 thesis_chapter_3.Rmd
small corrections...

@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=5, fig.cap="Figure 2: Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=5, fig.cap="Figure 3: Comparison of mean trajectories.
 Panel (A) represents hydrodynamic drag only simulations and Panel (B) represents simulations for both wind and hydrodynamic drag, for January, Panel (C) represents hydrodynamic drag only simulations and Panel (D) represents simulations for both wind and hydrodynamic drag, for April.
@@
 
-Across all simulations particles flowed in a north-westward direction (Figure 3).
+Across all simulations particles flowed in a north-westward direction.
 The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids.

### 2021-04-13 thesis_chapter_3.Rmd
touchups..

@@
 library(maps)
-#library(flextable)
+library(flextable)
 library(officer)
@@
 
-```{r sim table, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+```{r sim table, eval=TRUE, message=FALSE, warning=FALSE, include=TRUE}
 
@@
 
-```{r morphology table, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, eval=FALSE}
+```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, eval=TRUE}
 
@@
   scale_colour_discrete("Type") +  
+  guides(fill=guide_legend(title="Number of particles")) +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=5, fig.cap="Figure 3: Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=11, fig.width=8, fig.cap="Figure 3: Comparison of mean trajectories.
 Panel (A) represents hydrodynamic drag only simulations and Panel (B) represents simulations for both wind and hydrodynamic drag, for January, Panel (C) represents hydrodynamic drag only simulations and Panel (D) represents simulations for both wind and hydrodynamic drag, for April.
@@
 
-ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, labels = "AUTO", ncol = 2, nrow = 4, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("right"))
+ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, ncol = 2, nrow = 4, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("right"), labels = "AUTO", align = "hv")
 
@@
 
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation scenarios.", fig.height=9, fig.width=7}
+```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation scenarios for each season.For each month, passive particles are plotted on the left colum", fig.height=9, fig.width=7}
 
@@
 
+load("map/south_africa_coast_hi_res.Rdata")
+
 # January
@@
 shape_plot <- ggplot()+
-  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
+  #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
   geom_path(data = sphere_mean, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = cyl_mean, aes(x = lon, y = lat), color = 'orange')+
-  xlab("Longtitude") + 
-  ylab("Latitude") +
-  ggtitle("Mean trajectory comparison") +
+  geom_path(data = cyl_mean, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") +
+  ggtitle("January") +
   #guides(fill=guide_legend(title="Number of particles")) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("Cylinder") + 
+  ggtitle("") + 
   theme_bw() +
@@
   ylab("Latitude") + 
-  ggtitle("Sphere") +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 shape_plot_April <- ggplot()+
-  geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
+  #geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
   geom_path(data = sphere_mean_April, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = cyl_mean_April, aes(x = lon, y = lat), color = 'orange')+
-  xlab("Longtitude") + 
-  ylab("Latitude") +
-  ggtitle("Mean trajectory comparison") +
+  geom_path(data = cyl_mean_April, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") +
+  ggtitle("April") +
   #guides(fill=guide_legend(title="Number of particles")) +
@@
 shape_plot_July <- ggplot()+
-  geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
+  #geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
   geom_path(data = sphere_mean_July, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = cyl_mean_July, aes(x = lon, y = lat), color = 'orange')+
-  xlab("Longtitude") + 
-  ylab("Latitude") +
-  ggtitle("Mean trajectory comparison") +
+  geom_path(data = cyl_mean_July, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") +
+  ggtitle("July") +
   #guides(fill=guide_legend(title="Number of particles")) +
@@
 shape_plot_Oct <- ggplot()+
-  geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
+  #geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
   geom_path(data = sphere_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = cyl_mean_Oct, aes(x = lon, y = lat), color = 'orange')+
-  xlab("Longtitude") + 
-  ylab("Latitude") +
-  ggtitle("Mean trajectory comparison") +
+  geom_path(data = cyl_mean_Oct, aes(x = lon, y = lat), color = 'red')+
+  xlab("Longtitude") + 
+  ylab("Latitude") +
+  ggtitle("October") +
   #guides(fill=guide_legend(title="Number of particles")) +
@@
   ggtitle("Comparison of distance travelled") +
+  guides(fill=guide_legend(title="Coefficient")) +
   theme_bw() +
@@
 
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of trajectory data between a kelp sphere and kelp cylinder.
-Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylindrical shape coefficient.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=8}
-
-shape_plot_A <- ggarrange(shape_plot, shape_plot_April, shape_plot_July, shape_plot_Oct, heatmap_sphere, heatmap_sphere_April,heatmap_sphere_July, heatmap_sphere_Oct, heatmap_cyl, heatmap_cyl_April, heatmap_cyl_July, heatmap_cyl_Oct, labels = c("AUTO"), common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 3)
-
-ggarrange(shape_plot_A, shape_dis_plot, ncol = 1, nrow = 2)
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of trajectory data between spehere and cylinder shape coeficients.The mean trajectory comparison between passive and kel The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=8}
+
+shape_plot_A <- ggarrange(shape_plot, shape_plot_April, shape_plot_July, shape_plot_Oct, heatmap_sphere, heatmap_sphere_April,heatmap_sphere_July, heatmap_sphere_Oct, heatmap_cyl, heatmap_cyl_April, heatmap_cyl_July, heatmap_cyl_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 3, align = "hv", labels = c("A", "", "", "", "B", "", "", "", "C"), font.label = list(size = 11, face = "bold"))
+
+ggarrange(shape_plot_A, shape_dis_plot, ncol = 1, nrow = 2, legend = c("right"), labels = c("", "D") , font.label = list(size = 11, face = "bold"))
 

### 2021-04-13 thesis_chapter_3.Rmd
simulation table updated

@@
 For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity ; while in winter the Benguela jet begins to stablize until summer when the begins to split again [@rubio2009; @veitch2018; @ragoasha2019].
-In order to determine if the shape co-efficient used plays a role in model used, a simulation of a kelp with the maximum surface area and drag exposure was run using the shape coefficient for a cylinder to compare to a simulation using a sphere, and was run for 30 days using a fourth order Runge-Kutte advection scheme with brownian motion and a particle set of size 1000.
 
@@
 
-```{r sim table, eval=TRUE, message=FALSE, warning=FALSE, include=TRUE}
+In order to determine if the shape co-efficient used plays a role in model used and to validate the use of a sphereical shape coefficent, simulations of kelps with the maximum surface area and hydrodynamic drag exposure was run using the shape coefficient for a cylinder to compare to a simulations using a spheres.
+Simulations were run for 30 days using a fourth order Runge-Kutte advection scheme with brownian motion and a particle set of size 1000 for the same seasons as per previous simulations.
+
+```{r sim table, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 # Results
-## Comparison of coefficients
-
-The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (see appendix).
-This was also reflected when comparing distance travelled by particles and particle density plots (see appendix).
-The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (see appendix).
 
@@
 p_RK4_April$type <- "RK4_April" # add extra column 'type' to dataframe and assigned 'RK4' label
-p_RK4_mean_April$type <- "17" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_mean_April$type <- "14" # add extra column 'type' to dataframe and assigned 'RK4' label
 
@@
 Min_Kfloat_H100_W00_April$velocity <- with(Min_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H100_W00_mean_April$type <- "18" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_mean_April$type <- "15" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H095_W05_April$velocity <- with(Min_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H095_W05_mean_April$type <- "19" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_mean_April$type <- "16" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H090_W10_April$velocity <- with(Min_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H090_W10_mean_April$type <- "20" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_mean_April$type <- "17" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H085_W15_April$velocity <- with(Min_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H085_W15_mean_April$type <- "21" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_mean_April$type <- "18" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_Kfloat_H100_W00_April$velocity <- with(Mean_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_Kfloat_H100_W00_mean_April$type <- "22" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_mean_April$type <- "19" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H095_W05_April$velocity <- with(Mean_KFloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H095_W05_mean_April$type <- "23" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_mean_April$type <- "20" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H090_W10_April$velocity <- with(Mean_KFloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H090_W10_mean_April$type <- "24" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_mean_April$type <- "21" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H085_W15_April$velocity <- with(Mean_KFloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H085_W15_mean_April$type <- "25" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_mean_April$type <- "22" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H100_W00_April$velocity <- with(Max_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H100_W00_mean_April$type <- "26" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_mean_April$type <- "23" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H095_W05_April$velocity <- with(Max_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H095_W05_mean_April$type <- "27" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_mean_April$type <- "24" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H090_W10_April$velocity <- with(Max_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H090_W10_mean_April$type <- "28" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_mean_April$type <- "25" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H085_W15_April$velocity <- with(Max_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H085_W15_mean_April$type <- "29" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_mean_April$type <- "26" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 p_RK4_July$type <- "RK4_July" # add extra column 'type' to dataframe and assigned 'RK4' label
-p_RK4_mean_July$type <- "30" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_mean_July$type <- "27" # add extra column 'type' to dataframe and assigned 'RK4' label
 
@@
 Min_Kfloat_H100_W00_July$velocity <- with(Min_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H100_W00_mean_July$type <- "31" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_mean_July$type <- "28" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H095_W05_July$velocity <- with(Min_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H095_W05_mean_July$type <- "32" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_mean_July$type <- "29" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H090_W10_July$velocity <- with(Min_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H090_W10_mean_July$type <- "33" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_mean_July$type <- "30" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H085_W15_July$velocity <- with(Min_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H085_W15_mean_July$type <- "34" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_mean_July$type <- "31" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_Kfloat_H100_W00_July$velocity <- with(Mean_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_Kfloat_H100_W00_mean_July$type <- "35" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_mean_July$type <- "32" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H095_W05_July$velocity <- with(Mean_KFloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H095_W05_mean_July$type <- "36" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_mean_July$type <- "33" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H090_W10_July$velocity <- with(Mean_KFloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H090_W10_mean_July$type <- "37" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_mean_July$type <- "34" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H085_W15_July$velocity <- with(Mean_KFloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H085_W15_mean_July$type <- "38" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_mean_July$type <- "35" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H100_W00_July$velocity <- with(Max_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H100_W00_mean_July$type <- "39" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_mean_July$type <- "36" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H095_W05_July$velocity <- with(Max_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H095_W05_mean_July$type <- "40" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_mean_July$type <- "37" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H090_W10_July$velocity <- with(Max_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H090_W10_mean_July$type <- "41" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_mean_July$type <- "38" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H085_W15_July$velocity <- with(Max_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H085_W15_mean_July$type <- "42" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_mean_July$type <- "39" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H100_W00_mean_Oct$type <- "44" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_mean_Oct$type <- "40" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H095_W05_mean_Oct$type <- "45" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_mean_Oct$type <- "41" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H090_W10_mean_Oct$type <- "46" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_mean_Oct$type <- "42" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H085_W15_mean_Oct$type <- "47" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_mean_Oct$type <- "43" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_Kfloat_H100_W00_mean_Oct$type <- "48" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_mean_Oct$type <- "44" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H095_W05_mean_Oct$type <- "49" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_mean_Oct$type <- "45" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H090_W10_mean_Oct$type <- "50" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_mean_Oct$type <- "46" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H085_W15_mean_Oct$type <- "51" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_mean_Oct$type <- "47" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H100_W00_mean_Oct$type <- "52" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_mean_Oct$type <- "48" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H095_W05_mean_Oct$type <- "53" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_mean_Oct$type <- "49" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H090_W10_mean_Oct$type <- "54" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_mean_Oct$type <- "50" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H085_W15_mean_Oct$type <- "55" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_mean_Oct$type <- "51" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 
-total_mean_water_April$type <- factor(total_mean_water_April$type, levels=c("17", "18", "22", "26"))
-total_mean_ww_April$type <- factor(total_mean_ww_April$type, levels=c("17", "19", "20", "21", "23", "24", "25", "27", "28", "29"))
+total_mean_water_April$type <- factor(total_mean_water_April$type, levels=c("14", "15", "19", "23"))
+total_mean_ww_April$type <- factor(total_mean_ww_April$type, levels=c("14", "16", "17", "18", "20", "21", "22", "24", "25", "26"))
 
@@
 
-total_mean_water_July$type <- factor(total_mean_water_July$type, levels=c("30", "31", "35", "39"))
-total_mean_ww_July$type <- factor(total_mean_ww_July$type, levels=c("30", "32", "33", "34", "36", "37", "38", "40", "41", "42"))
+total_mean_water_July$type <- factor(total_mean_water_July$type, levels=c("27", "28", "32", "36"))
+total_mean_ww_July$type <- factor(total_mean_ww_July$type, levels=c("27", "29", "30", "31", "33", "34", "35", "37", "38", "39"))
 
@@
 
-total_mean_water_Oct$type <- factor(total_mean_water_Oct$type, levels=c("43", "44", "48", "52"))
-total_mean_ww_Oct$type <- factor(total_mean_ww_Oct$type, levels=c("43", "45", "46", "47", "49", "50", "51", "53", "54", "55"))
+total_mean_water_Oct$type <- factor(total_mean_water_Oct$type, levels=c("40", "41", "45", "49"))
+total_mean_ww_Oct$type <- factor(total_mean_ww_Oct$type, levels=c("40", "42", "43", "44", "46", "47", "48", "50", "51", "52"))
 
@@
 
-custom_labs_April <- c("17","18","19","20","21","22","23","24","25","26","27","28", "29") #create custom labels and add them to dis_plot
+custom_labs_April <- c("14","15","16","17","18","19","20","21","22","23","24","25", "26") #create custom labels and add them to dis_plot
 
@@
 
-custom_labs_July <- c("30","31","32","33","34","35","36","37","38","39","40","41", "42") #create custom labels and add them to dis_plot
+custom_labs_July <- c("27","28","29","30","31","32","33","34","35","36","37","38", "39") #create custom labels and add them to dis_plot
 
@@
 
-custom_labs_Oct <- c("43","44","45","46","47","48","49","50","51","52","53","54","55") #create custom labels and add them to dis_plot
+custom_labs_Oct <- c("40","41","42","43","44","45","46","47","48","49","50","51","52") #create custom labels and add them to dis_plot
 
@@
 
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation scenarios for each season.For each month, passive particles are plotted on the left colum", fig.height=9, fig.width=7}
+```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation for scenarios 53-60.
+For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=9, fig.width=7}
 
@@
 
-```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of age between passive and kelp accumulation scenarios.", fig.height=9, fig.width=7}
+```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of age between passive and kelp accumulation scenarios 53-60.
+For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=9, fig.width=7}
 
@@
 In Figure 9, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
+
+## Comparison of coefficients
+
+The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (see appendix).
+This was also reflected when comparing distance travelled by particles and particle density plots (see appendix).
+The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (see appendix).
 
@@
 
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of trajectory data between spehere and cylinder shape coeficients.The mean trajectory comparison between passive and kel The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=8}
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of trajectory data between spehere and cylinder shape coeficients for scenarios 61-64.The mean trajectory comparison between passive and kel The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=8}
 

### 2021-04-13 ocean_map.R
script for vector map

@@ -1,0 +1,35 @@
+library(tidyverse)
+library(oce)
+library(lubridate)
+library(gganimate)
+library(ggplot2)
+
+# Read in simulation data and data wrangling
+
+ocean_vector <- tidync("OceanParcels/global-analysis-forecast-phy-001-024_1603882856077.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(longitude, latitude, time, uo, vo) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+ #filter(between(t, as.POSIXct("2018-01-01"), as.POSIXct("2018-01-05")))
+ #mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))  %>% # calculate mean trajectory
+
+ocean_vector$velocity <- with(ocean_vector, sqrt(uo^2 + vo^2))
+
+ocean_summary = ocean_vector %>%
+  group_by(longitude, latitude, t) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
+load("map/south_africa_coast_hi_res.Rdata")
+
+test_plot = ggplot() +
+  geom_segment(data = ocean_summary,
+               aes(x = longitude, xend = longitude+uo/60, y = latitude,
+               yend = latitude+vo/60), arrow = arrow(length = unit(0.25, "cm"))) +
+  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",fill = "grey") +
+  #coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
+  theme_bw() +
+  theme(axis.text = element_text(size = 11, colour = 1)) +
+  labs(x = NULL, y = NULL)

### 2021-04-16 vector_map_v2.R
vector map script

@@ -1,0 +1,89 @@
+library(tidyverse)
+library(oce)
+library(lubridate)
+library(gganimate)
+library(ggplot2)
+
+# Read in simulation data and data wrangling
+
+ocean_vector <- tidync("OceanParcels/global-analysis-forecast-phy-001-024_1603882856077.nc") %>%
+  hyper_tibble() %>%
+  #dplyr::select(longitude, latitude, time, uo, vo) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) %>%
+  filter(t >= as.Date("2018-01-01 00:00:00") & t <= as.Date("2018-01-30 00:00:00"))
+
+
+
+############################################
+
+
+ocean_vector <- tidync("OceanParcels/global-analysis-forecast-phy-001-024_1603882856077.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, longitude, latitude, uo, vo)
+
+ocean_vector$velocity <- with(ocean_vector, sqrt(uo^2 + vo^2))
+
+ocean_vector <- ocean_vector %>%
+  group_by(longitude, latitude) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
+wind_vector <- tidync("OceanParcels/CERSAT-GLO-BLENDED_WIND_L4-V6-OBS_FULL_TIME_SERIE_1603883929498.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(lon, lat, eastward_wind, northward_wind)
+
+wind_vector$velocity <- with(wind_vector, sqrt(eastward_wind^2 + northward_wind^2))
+
+wind_vector = wind_vector %>%
+  group_by(lon, lat) %>%
+  summarise(eastward_wind = median(eastward_wind, na.rm = TRUE),
+            northward_wind = median(northward_wind, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
+load("map/south_africa_coast_hi_res.Rdata")
+
+ocean_plot <-  ggplot() +
+  geom_tile(data = ocean_vector, aes(fill = velocity, x = longitude, y = latitude)) +
+  geom_segment(data = ocean_vector,
+               aes(x = longitude, xend = longitude+uo, y = latitude,
[…]

### 2021-04-18 thesis_chapter3_vector_map.R
file rename

@@ -108,7 +108,6 @@ ocean_plot_oct <-  ggplot() +
   theme(axis.text = element_text(size = 11, colour = 1)) +
   labs(x = NULL, y = NULL)
 
-ggarrange(ocean_plot, wind_plot, font.label = list(size = 11, face = "bold"), ncol = 2, nrow = 1)
 
 ggsave("ocean_plot.png", dpi = 320, width = 8, height = 8, path = "figures/chapter_3/", units = "in")
 
@@ -228,10 +227,21 @@ wind_plot_oct <- ggplot() +
   theme(axis.text = element_text(size = 11, colour = 1)) +
   labs(x = NULL, y = NULL)
 
-ggarrange(ocean_plot_jan, wind_plot_jan, ocean_plot_apr, wind_plot_apr, ocean_plot_july, wind_plot_july, ocean_plot_oct, wind_plot_oct,
-          font.label = list(size = 11, face = "bold"), ncol = 2, nrow = 3, labels = c("January", "", "April", "", "July", "", "October"))
+ocean <- ggarrange(ocean_plot_jan, ocean_plot_apr, ocean_plot_july, ocean_plot_oct,
+                   font.label = list(size = 12, face = "bold"), labels = c("January", "April", "July","October"),
+                   ncol = 2, nrow = 2, common.legend = TRUE, legend = c("bottom"))
 
-ggsave("vector_combined.png", dpi = 320, width = 8, height = 8, path = "figures/chapter_3/", units = "in")
+wind <- ggarrange(wind_plot_jan, wind_plot_apr, wind_plot_july, wind_plot_oct,
+                  font.label = list(size = 12, face = "bold"),labels = c("January", "April", "July","October"),
+                  ncol = 2, nrow = 2, common.legend = TRUE, legend = c("bottom"))
+
+combined <- ggarrange(ocean, wind, align = c("h"), ncol = 1, nrow = 2)
+
+ggarrange(ocean_plot_jan, wind_plot_jan, ocean_plot_apr, wind_plot_apr, ocean_plot_july, wind_plot_july, ocean_plot_oct, wind_plot_oct,
+          font.label = list(size = 11, face = "bold"), labels = c("January", "", "April", "","July", "", "October"), ncol = 2, nrow = 4, align = "h",
+          common.legend = TRUE, legend = c("bottom"))
+
+ggsave("vector_combined.jpeg", dpi = 320, width = 8, height = 8, path = "figures/chapter_3/", units = "in")
 
 ##########################################################################################################################################################
 # wind rose

### 2021-04-21 thesis_chapter_3.Rmd
version 3 of draft, vector map script updated

@@
 author: 
-- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape]
+- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape South African Observation Network]
 output:
   word_document:
-    reference_docx: markdown/frontiers_template.docx
+    reference_docx: markdown/template.docx
   pdf_document:
@@
 
-```{r sim table, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}
+```{r sim table, eval=TRUE, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
 
@@
 p_RK4_Oct$type <- "RK4_Oct" # add extra column 'type' to dataframe and assigned 'RK4' label
-p_RK4_mean_Oct$type <- "43" # add extra column 'type' to dataframe and assigned 'RK4' label
+p_RK4_mean_Oct$type <- "40" # add extra column 'type' to dataframe and assigned 'RK4' label
 
@@
 Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H100_W00_mean_Oct$type <- "40" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H100_W00_mean_Oct$type <- "41" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H095_W05_mean_Oct$type <- "41" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H095_W05_mean_Oct$type <- "42" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H090_W10_mean_Oct$type <- "42" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H090_W10_mean_Oct$type <- "43" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H085_W15_mean_Oct$type <- "43" # add extra column 'type' to dataframe and assigned 'kelp' label
+Min_Kfloat_H085_W15_mean_Oct$type <- "44" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_Kfloat_H100_W00_mean_Oct$type <- "44" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_Kfloat_H100_W00_mean_Oct$type <- "45" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H095_W05_mean_Oct$type <- "45" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H095_W05_mean_Oct$type <- "46" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H090_W10_mean_Oct$type <- "46" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H090_W10_mean_Oct$type <- "47" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H085_W15_mean_Oct$type <- "47" # add extra column 'type' to dataframe and assigned 'kelp' label
+Mean_KFloat_H085_W15_mean_Oct$type <- "48" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H100_W00_mean_Oct$type <- "48" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H100_W00_mean_Oct$type <- "49" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H095_W05_mean_Oct$type <- "49" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H095_W05_mean_Oct$type <- "50" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H090_W10_mean_Oct$type <- "50" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H090_W10_mean_Oct$type <- "51" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H085_W15_mean_Oct$type <- "51" # add extra column 'type' to dataframe and assigned 'kelp' label
+Max_Kfloat_H085_W15_mean_Oct$type <- "52" # add extra column 'type' to dataframe and assigned 'kelp' label
 
@@
 The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (autumn).
-Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag travelled further northward compared to the passive particles (see Figure 2C-D).
+Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag traveled further northward compared to the passive particles (see Figure 2C-D).
 The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles.
-Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles travelled less distance compared to kelp particles with both wind and hydrodynamic drag.
+Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles traveled less distance compared to kelp particles with both wind and hydrodynamic drag.
 Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 2E-F).
 Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore; resulting in entrainment within an eddy.
-Similar behaviour is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag.
-In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles travelled initially southward and then westward offshore and also becoming entrained in an eddy.
+Similar behavior is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag.
+In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles traveled initially southward and then westward offshore and also becoming entrained in an eddy.
 In the October (Spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 7F-G).
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=11, fig.width=8, fig.cap="Figure 3: Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=7.5, fig.cap="Figure 3: Comparison of mean trajectories.
 Panel (A) represents hydrodynamic drag only simulations and Panel (B) represents simulations for both wind and hydrodynamic drag, for January, Panel (C) represents hydrodynamic drag only simulations and Panel (D) represents simulations for both wind and hydrodynamic drag, for April.
@@
 
-ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, ncol = 2, nrow = 4, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("right"), labels = "AUTO", align = "hv")
+ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, ncol = 2, nrow = 4, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("bottom"), labels = "AUTO", align = "hv")
 
@@
 
-heatmap_Mean_KFloat_H090_W10 <- ggplot() +
+heatmap_Mean_KFloat_H090_W100 <- ggplot() +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
 
-ggarrange(heatmap_RK4_April,heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April,as_ggplot(leg_April), labels = c("17", "18","19","20","21","22","23","24","25","26","27","28", "29"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
+ggarrange(heatmap_RK4_April,heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April,as_ggplot(leg_April), labels = c("14", "15","16","17","18","19","20","21","22","23","24","25", "26"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
 
@@
 
-ggarrange(heatmap_RK4_July,heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, as_ggplot(leg_July),labels = c("30","31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
+ggarrange(heatmap_RK4_July,heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, as_ggplot(leg_July),labels = c("27","28","29","30","31","32","33","34","35","36","37","38","39"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
 
@@
 
-leg_Oct <- get_legend(heatmap_RK4_Oct)
-
-ggarrange(heatmap_RK4_Oct, heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, as_ggplot(leg_Oct),labels = c("43","44","45","46","47","48","49","50","51","52","53","54","55"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
+leg_Oct <- get_legend(heatmap_Min_Kfloat_H100_W00_Oct)
+
+ggarrange(heatmap_RK4_Oct, heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Mean_KFloat_H090_W100, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, as_ggplot(leg_Oct),labels = c("40","41","42","43","44","45","46","47","48","49","50","51","52"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
 
@@
 
-There were significant differences among the median distances travelled for all particles in each simulation (Figure 4; p < 0.05).
-The simulations that considered for 100% hydrodynamic drag travelled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
+There were significant differences among the median distances traveled for all particles in each simulation (Figure 4; p < 0.05).
+The simulations that considered for 100% hydrodynamic drag traveled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
 Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations.
@@
 (Figure 11, simulations 17-29; p < 0.05).
-Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas travelled shorter distance across compared to other kelp particles.
-The shortest distance travelled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
+Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas traveled shorter distance across compared to other kelp particles.
+The shortest distance traveled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
 No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; p > 0.05).
 In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; p < 0.05).
-Kelp particles with only 90% hydrodynamic drag across cross-sectional areas travelled shorter distance compared to all other particles.
-As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure travelled the shortest distance compared to all other particles.
+Kelp particles with only 90% hydrodynamic drag across cross-sectional areas traveled shorter distance compared to all other particles.
+As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure traveled the shortest distance compared to all other particles.
 
@@
 
-ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 4)
+ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "auto", ncol = 1, nrow = 4)
 
@@
 The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (see appendix).
-This was also reflected when comparing distance travelled by particles and particle density plots (see appendix).
+This was also reflected when comparing distance traveled by particles and particle density plots (see appendix).
 The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (see appendix).
@@
 This is also reflected in the analysis comparing distances travelled by all the particles in each simulation.
-The particles in the simulations that included wind drag traveled significantly greater distances compared to the passive or hydrodynamic only drag.
-Similar patterns are seen in the additional simulations run for April (autumn), July (winter), and October (spring); however even kelp particles with only hydrodynamic drag flow further compared to passive particles in these times of the year.
+The particles in the simulations that included wind drag traveled significantly greater distances compared to the passive or hydrodynamic only drag simulations.
+Similar patterns are seen in the simulations run for April (autumn), July (winter), and October (spring); however unlike in January (summer) where only kelp particles with combined drag (hydrodynamic and wind) flowed further, even kelp particles with only hydrodynamic drag flow further compared to passive particles in these times of the year.
 This may be a result of the kelp particles exposed to different time-varying flows, which results in the kelp particles following a different surface current vector compared to passive particles.
 This also results in different end-points and therefore these results further confirm that drag is an important factor to account for when investigating macroalgal drift patterns.
+The different trajectory patterns between seasons reflects the underlying seasonal variation in regional oceanography.
+Transport in the study region is driven by a combination d variation of along-shore jets and wind induced Ekman transport of surface water [@rubio2009; @veitch2018; @ragoasha2019].
+A study by @ragoasha2019 investigated the success of Lagrangian particles reaching St.
+Helena Bay from Cape point using Lagrangian trajectory modelling and inclusion of regional oceanographic processes.
+The authors showed that transport success is higher in spring due to the single coherent Benguela Jet, which concentrates over the 300m isobath [@ragoasha2019].
+During upwelling season (December - March) transport success is less efficient as particles tend to flow offshore.
+The move offshore is a result of the bifurification of the Benguela Jet, which splits as a result of shoaling, and causes particles to be exposed to higher amounts of mesoscale activity.
+In the winter months, the Benguela Jet is still split but the flow weakens which results in lower transport success [@ragoasha2019].
+Similar patterns were found in this study, suggesting that the differences in seasonal patterns reflects the regional oceanographic features, particularly the Benguela Jet.
 
@@
 This study also provides an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately.
+In addition, the results from this study highlights the importance of regional oceanographic process and how these vary seasonally to better understand how kelp flotsam is dispersed in the ocean.
+In particular, the current study highlights the role of the Benguela Jet and mesoscale processes and how these vary seasonally; ultimately determining transport success around the coast.
 Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal as well as how these vary seasonally.
@@
   geom_path(data = cyl_mean_April, aes(x = lon, y = lat), color = 'red')+
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
   xlab("Longtitude") + 
@@
   geom_path(data = cyl_mean_July, aes(x = lon, y = lat), color = 'red')+
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
   xlab("Longtitude") + 
@@
   geom_path(data = cyl_mean_Oct, aes(x = lon, y = lat), color = 'red')+
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
   xlab("Longtitude") + 
@@
 
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of trajectory data between spehere and cylinder shape coeficients for scenarios 61-64.The mean trajectory comparison between passive and kel The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=8}
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 10: Comparison of trajectory data between sphere and cylinder shape coefficients for scenarios 61-64.The mean trajectory comparison between passive and kelp.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=8}
 
@@
 ```
+
+```{r vector maps, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 11: The average ocean (left column) and wind vector maps (right column) for the respective seasons are presented.
+The velocity vectors were averaged over the 30 day simulation times for respective seasons (summer, autumn, winter, spring).
+A represents summer, B autumn, C winter and D spring", fig.height=10, fig.width=8}
+
+library(jpeg)
+library(grid)
+library(gridExtra)
+
+vector_maps <- readJPEG("figures/chapter_3/vector_combined.jpeg")
+grid.raster(vector_maps)
+```

### 2021-04-21 thesis_chapter3_vector_map.R

@@ -46,17 +46,20 @@ ocean_oct_summary <- ocean_oct %>%
             vo = median(vo, na.rm = TRUE),
             velocity = median(velocity, na.rm = TRUE))
 
+#ocean_jan_summary[seq(5, nrow(ocean_jan_summary), 4), ] # every 4th row
+
+
 # January ocean vector map
 
 load("map/south_africa_coast_hi_res.Rdata")
 
 ocean_plot_jan <-  ggplot() +
   geom_tile(data = ocean_jan_summary, aes(fill = velocity, x = longitude, y = latitude)) +
-  geom_segment(data = ocean_jan_summary,
-               aes(x = longitude, xend = longitude+uo, y = latitude,
-                   yend = latitude+vo), arrow = arrow(length = unit(0.30, "cm"))) +
+  geom_segment(data = ocean_jan_summary[seq(5, nrow(ocean_jan_summary),2), ],
+               aes(x = longitude, xend = longitude+uo/30, y = latitude,
+                   yend = latitude+vo/30), arrow = arrow(length = unit(0.30, "cm"))) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",fill = "grey") +
-  scale_fill_continuous(type = "viridis", limits = c(0, 0.6)) +
+  scale_fill_continuous(type = "viridis", limits = c(0, 0.8)) +
   coord_cartesian(xlim = c(15, 20), ylim = c(-35, -30)) +
   guides(fill=guide_legend(title="Velocity")) +
   theme_bw() +
@@ -67,11 +70,11 @@ ocean_plot_jan <-  ggplot() +
 
 ocean_plot_apr <-  ggplot() +
   geom_tile(data = ocean_apr_summary, aes(fill = velocity, x = longitude, y = latitude)) +
-  geom_segment(data = ocean_apr_summary,
-               aes(x = longitude, xend = longitude+uo, y = latitude,
-                   yend = latitude+vo), arrow = arrow(length = unit(0.30, "cm"))) +
+  geom_segment(data = ocean_apr_summary[seq(5, nrow(ocean_jan_summary), 2), ],
+               aes(x = longitude, xend = longitude+uo/30, y = latitude,
+                   yend = latitude+vo/30), arrow = arrow(length = unit(0.30, "cm"))) +
[…]

### 2021-04-22 thesis_chapter_3.Rmd
updates

@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for scenarios 17 - 29, for April 2018 (winter).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for scenarios 14 - 26, for April 2018 (autumn).
 The plots are regular hexagons within the density of particles were calculated.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end run time for scenarios 30 - 42, for July 2018 (winter).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end run time for scenarios 27 - 39, for July 2018 (winter).
 The plots are regular hexagons within the density of particles were calculated.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end run time for scenarios 43 - 55, for October 2018 (spring).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end run time for scenarios 40 - 52, for October 2018 (spring).
 The plots are regular hexagons within the density of particles were calculated.

### 2021-04-22 vector_map_animations.R
animations for vector maps

@@ -1,0 +1,94 @@
+library(gganimate)
+library(plotly)
+
+wind_vector <- tidync("OceanParcels/CERSAT-GLO-BLENDED_WIND_L4-V6-OBS_FULL_TIME_SERIE_1603883929498.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, lon, lat, eastward_wind, northward_wind) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2018-01-01 00:00:00"),
+                     "GMT",
+                     "%Y-%m-%d"))
+
+wind_jan <-  wind_vector %>%
+  filter(t >= ("2136-01-01 00:00:00") & t <= ("2136-01-30 00:00:00"))
+wind_apr <-  wind_vector %>%
+  filter(t >= ("2136-04-01 00:00:00") & t <= ("2136-04-30 00:00:00"))
+wind_july <-  wind_vector %>%
+  filter(t >= ("2136-07-01 00:00:00") & t <= ("2136-07-30 00:00:00"))
+wind_oct <-  wind_vector %>%
+  filter(t >= ("2136-10-01 00:00:00") & t <= ("2136-10-30 00:00:00"))
+
+wind_jan$velocity <- with(wind_jan, sqrt(eastward_wind^2 + northward_wind^2))
+wind_apr$velocity <- with(wind_apr, sqrt(eastward_wind^2 + northward_wind^2))
+wind_july$velocity <- with(wind_july, sqrt(eastward_wind^2 + northward_wind^2))
+wind_oct$velocity <- with(wind_oct, sqrt(eastward_wind^2 + northward_wind^2))
+
+wind_jan_summary_ani <- wind_jan %>%
+  group_by(t, lon, lat) %>%
+  summarise(eastward_wind = median(eastward_wind, na.rm = TRUE),
+            northward_wind = median(northward_wind, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
+
+# January wind vector map
+
+load("map/south_africa_coast_hi_res.Rdata")
+
+wind_ani_jan <- ggplot() +
+  geom_tile(data = wind_jan_summary_ani, aes(fill = velocity, x = lon, y = lat)) +
+  #geom_raster(data = wind_vector, aes(x = lon, y = lat, fill = velocity))+
+  geom_segment(data = wind_jan_summary_ani,
+               aes(x = lon, xend = lon+eastward_wind/30, y = lat,
[…]

### 2021-04-22 vector_map_animations.R
animations for vector maps finished

@@ -1,6 +1,133 @@
 library(gganimate)
 library(plotly)
 
+ocean_vector <- tidync("OceanParcels/global-analysis-forecast-phy-001-024_1603882856077.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, longitude, latitude, uo, vo) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2018-01-01 00:00:00"),
+                     "GMT",
+                     "%Y-%m-%d"))
+
+ocean_jan <-  ocean_vector %>%
+  filter(t >= ("2086-01-01 00:00:00") & t <= ("2086-01-30 00:00:00"))
+ocean_apr <-  ocean_vector %>%
+  filter(t >= ("2086-04-01 00:00:00") & t <= ("2086-04-30 00:00:00"))
+ocean_july <-  ocean_vector %>%
+  filter(t >= ("2086-07-01 00:00:00") & t <= ("2086-07-30 00:00:00"))
+ocean_oct <-  ocean_vector %>%
+  filter(t >= ("2086-10-01 00:00:00") & t <= ("2086-10-30 00:00:00"))
+
+ocean_jan$velocity <- with(ocean_jan, sqrt(uo^2 + vo^2))
+ocean_apr$velocity <- with(ocean_apr, sqrt(uo^2 + vo^2))
+ocean_july$velocity <- with(ocean_july, sqrt(uo^2 + vo^2))
+ocean_oct$velocity <- with(ocean_oct, sqrt(uo^2 + vo^2))
+
+ocean_jan_summary_ani <- ocean_jan %>%
+  group_by(t, longitude, latitude) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
+ocean_apr_summary_ani <- ocean_apr %>%
+  group_by(t, longitude, latitude) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
+ocean_july_summary_ani <- ocean_july %>%
+  group_by(t, longitude, latitude) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
[…]

### 2021-04-26 thesis_chapter_4.Rmd
updates
chapter 4 script

--- 2021-01-25 thesis_chapter_4.Rmd
@@
 ```
+
+```{r, warning=FALSE, message=FALSE, echo = FALSE}
+
+houtbay <- tidync("OceanParcels/trajectory_data/chapter_4_output/houtbay.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+
+houtbay_mean <- houtbay %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+houtbay$site <- "Hout Bay" 
+houtbay_mean$site <- "Hout Bay"
+
+kommetjie <- tidync("OceanParcels/trajectory_data/chapter_4_output/kommetjie.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+
+kommetjie_mean <- kommetjie %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+kommetjie$site <- "kommetjie" 
+kommetjie_mean$site <- "kommetjie"
+
+soetwater <- tidync("OceanParcels/trajectory_data/chapter_4_output/soetwater.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+
+soetwater_mean <- soetwater %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+soetwater$site <- "Soetwater" 
+soetwater_mean$site <- "Soetwater"
+
+millers <- tidync("OceanParcels/trajectory_data/chapter_4_output/millers.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+
+millers_mean <- millers %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+millers$site <- "millers" 
+millers_mean$site <- "millers"
+
+total_data <- rbind(houtbay, soetwater, kommetjie, millers)
+total_data_mean <- rbind(houtbay_mean, soetwater_mean,kommetjie_mean, millers_mean)
+```
+
+## Mean trajectory
+
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
+
+library(ggplot2)
+
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+mean_traj_site <- ggplot() + 
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  geom_path(data = total_data_mean, aes(x = lon, y = lat, colour = site)) +
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw()
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+mean_traj_site
+
+```
+
+## Density comparisons
+
+```{r, message=FALSE, echo=FALSE, warning=FALSE}
+
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+heatmap_houtbay <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size = 1.0, colour = "black", fill = NA) +
+  geom_hex(data = houtbay, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000000)) +
+  geom_path(data = houtbay_mean,aes(x = lon, y = lat, color = 'red'),             show.legend = FALSE) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("HoutBay") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_komm <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = kommetjie, bins = 80, aes(x = lon, y = lat), legend = FALSE) +   scale_fill_continuous(type = "gradient", limits = c(1, 1000000)) +
+  geom_path(data = kommetjie_mean,aes(x = lon, y = lat, color = 'red'),           show.legend = FALSE) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Kommetjie") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_soetwater <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = soetwater, bins = 80, aes(x = lon, y = lat), legend = FALSE) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000000)) +
+  geom_path(data = soetwater_mean,aes(x = lon, y = lat, color = 'red'),           show.legend = FALSE) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Soetwater") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_soetwater <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = soetwater, bins = 80, aes(x = lon, y = lat), legend = FALSE) +
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000000)) +
+  geom_path(data = soetwater_mean,aes(x = lon, y = lat, color = 'red'),           show.legend = FALSE) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Kommetjie") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+heatmap_millers <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = millers, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 1000000)) +
+  geom_path(data = millers_mean,aes(x = lon, y = lat, color = 'red'),             show.legend = FALSE) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Millers") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+ggarrange(heatmap_houtbay, heatmap_soetwater, heatmap_millers, ncol = 1, nrow = 3, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("bottom"), labels = "AUTO", align = "hv")
+```
+
+```{r total heatmap, echo=FALSE, message=FALSE, warning=FALSE}
+
+heatmap_total <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size =    1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = total_data, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
+  scale_fill_continuous(type = "gradient", limits = c(1, 19000000)) +
+  geom_path(data = total_data_mean,aes(x = lon, y = lat, color = 'red'),show.legend = FALSE) +
+  geom_rect(data = south_africa_coast_hi_res,aes(xmin=10, xmax=18.4, ymin=-33, ymax=-33.5),           inherit.aes = FALSE, fill = NA, colour =    "red", size = 1.5) +
+  geom_rect(data = south_africa_coast_hi_res,aes(xmin=10, xmax=18.4, ymin=-33.5, ymax=-34.0),           inherit.aes = FALSE, fill = NA, colour =    "green", size = 1.5) +
+  geom_rect(data = south_africa_coast_hi_res,aes(xmin=10, xmax=18.4, ymin=-34.0, ymax=-34.5),           inherit.aes = FALSE, fill = NA, colour =    "black", size = 1.5) +
+  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  ggtitle("Millers") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+heatmap_total
+```

### 2021-05-02 thesis_chapter_3.Rmd
abstract AJ comments

@@
 abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
-Macroalgae are one of the most important natural dispersal vectors, and are regarded as the tumble-weed of the ocean.
-Despite many studies that exist on the topic, the relative role of wind and surface currents in macroalgae trajectory is still unknown.
-Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a concensus as to the relative role of wind versus surface currents.
-In addition, past studies have not considered surface-area in relation to drag characterisics of floating macroalgae which have been shown to play a role in the trajectory and accumulation of other flotsam.
-Advances in Lagrangian trajectory modeling have been exponential in recent years, and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean.
+Macroalgae are one of the most important natural dispersal vectors and are regarded as the 'tumble-weed' of the ocean.
+Despite many studies that exist on the topic, the relative role of wind and surface currents in influencing the trajectory of macroalgal dispersal is still unknown.
+Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a consenses as to the relative role of wind versus surface currents.
+In addition, past studies have not considered surface-area in relation to drag characteristics of floating macroalgae, both of which have been shown to play a role in the trajectory and accumulation of other flotsam.
+Advances in Lagrangian trajectory modeling have been rapid in recent years and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean.
 This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagal trajectory and accumulation.
-We did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals.
+We focused on solitary kelp individuals.
 This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles.
@@
 library(ggplot2)
-library(rgdal)
+#library(rgdal)
 library(gridExtra)

### 2021-05-02 thesis_chapter_3.Rmd
Intro AJ comments

@@
 
-There is a range of objects --- both natural and anthropogenic --- floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
+There is a range of objects---both natural and anthropogenic---floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
 Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
-Kelp also function as input of allochthonous materials to near and distant ecosystems and in this manner support the ecological function in a range of adjacent habitats [@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].
+Kelps also function as input of allochthonous materials to near and distant ecosystems and in this manner support the ecological function in a range of adjacent habitats [@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].
 
 Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.
-In turn, the positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum.
+The positively buoyant pneumatocysts also cause plants to float to the surface when dislodged from the substratum.
 The structure and number of pneumatocysts varies between species.
-For example species belonging to the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018].
-In some cases the stipe itself may be air-filled, such as with *E.
+For example, species belonging to the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, and *Ecklonia maxima* have single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018].
+In some cases, the stipe itself may be air-filled, such as with *E.
 maxima* and *E.
 radiata*.
-Although there have been reports of floating Chlorophyta species and some Rhodophyta species, Phaeophyceae species are the most reported forms of floating algae.
-This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in the thalli [@dromgoole1982; @back2000].
+Although there have been reports of floating species of Chlorophyta and Rhodophyta, Phaeophyceae species are the most reported forms of floating algae.
+This is most likely because the green and red species reported to float are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in their thalli [@dromgoole1982; @back2000].
 The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @putman2018; @brooks2018; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
 
-Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
-Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown although the role of wind has been recognised as important in several studies.
+Past research points to macroalgal trajectories being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
+Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown although the role of wind has been recognized as important in several studies.
 For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M.
-pyrifera* in nearshore Monterrey Bay.
-The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents; however the relative importance of wind and was not clear.
+pyrifera* in nearshore Monterey Bay.
+The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents; however, the relative importance of wind and was not clear.
 In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter.
@@
 For example, a study by @fraser2011 on the rafting capabilities of *D.
-antarctica* used a combination of population genetics and relative age estimate of goosebarnacles attached to the raft.
-The presence of goosebarnacles suggests a long rafting, time as these species have a slow growth rate.
-Genetic analyses showed these species are able to raft up to $\sim$ 390 km from their local origin.
+antarctica* used a combination of population genetics and relative age estimate of goose barnacles attached to the raft.
+The presence of goose barnacles suggests a long rafting time as these species have a slow growth rate.
+Genetic analyses showed these species can raft up to $\sim$ 390 km from their local origin.
 The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011].
@@
 The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic-form drag [@breivik2011; @griffin2017].
-In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object.
+In addition, past research conducted by the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object.
 If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].
@@
 
-To accurately determine the trajectory of marine macroalgae, water and wind drag need to be taken into account.
+To accurately determine the trajectory of marine macroalgae, both water and wind drag need to be considered.
 Past research by @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
@@
 Studies by @brooks2019, @putman2018, and @putman2020 used Lagrangian approaches to investigate the effects of inertia, raft-size, and windage on macroalgae trajectory.
-@brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customised Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
+@brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customized Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
 The results showed that the trajectory of pelagic *Sargussum* was significantly influenced by inertia and the radial size of the rafts.
@@
 The authors showed that including *ad hoc* windage factors greatly improved accuracy regarding entrainment or expulsion from eddies when compared to tracked *Sargassum* mats in the ocean.
-Langrangian trajectory modeling is clearly; useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
+Langrangian trajectory modeling is clearly a useful tool; however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
 The radial size included by @brooks2019 is an improvement, however this approach does not consider the complex morphology on an individual level into account.
@@
 Gyres, eddies, and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018].
-Past studies investigating the dispersal of marine debris have also included the possible effects of inertia in accumulation of ocean drifter instruments in mesoscale features.
-For example, a study by @beron2016 used a modeling approach to investigated the role of Eckman transport and inertia to explain the observed behavior of undrogued drifters in the ocean gyres.
-Undrogued drifter trajectory is affected by wind unlike the drogue counter parts which makes the effects of wind negligible.
+Investigations into  the dispersal of marine debris have also included the possible effects of inertia in accumulation of ocean drifter instruments in mesoscale features.
+For example, a study by @beron2016 used a modeling approach to investigate the role of Eckman transport and inertia to explain the observed behavior of undrogued drifters in the ocean gyres.
+Undrogued drifter trajectory is affected by wind, unlike the drogue counter parts which makes the effects of wind negligible.
 The results showed that the combined effect inertia, not Eckman transport, played a significant role in the observed drifter behavior [@beron2016].
-The authors also noted that the presence of other flotsam along with the undrogued drifters in ocean gyres also suggests that aspects inertia also play a role in accumulation of other marine debris [@beron2016].
+@beron2016 also noted that the presence of other flotsam, along with the undrogued drifters in ocean gyres, suggests that aspects of inertia also play a role in accumulation of other marine debris [@beron2016].
 The approach by @beron2016 has also been improved upon by @beron2019 but also shows the significant role of inertia in flotsam trajectory and accumulation.
@@
 In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations.
-We aim to shed light on the 'behaviour' of surface drifting macroalgae from a coastal environment into the open ocean, which will inform future characterisation studies.
+We aim to shed light on the 'behavior' of surface drifting macroalgae from a coastal environment into the open ocean, which will inform future characterization studies.
 We will achieve this by comparing the trajectory outputs of virtual passive particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure.

### 2021-05-02 thesis_chapter_3.Rmd
Methods AJ comments

@@
 This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles.
-The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain.
-The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumulation zones for solitary floating macroalage."
+The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddys within the study domain.
+The results also point to the importance of sub-mesoscale features, such as eddys, as important surface accumulation zones for solitary floating macroalage."
 ---
@@
 The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents; however, the relative importance of wind and was not clear.
-In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter.
+In addition, the trajectories of tagged kelp were more consistent with the formation of eddys during winter.
 Previous studies have also identified wind as an important mechanism of dispersal in wind dominated ocean systems.
@@
 Other work by @putman2020 investigated the effect of including windage in macroalgae trajectory simulations.
-The authors showed that including *ad hoc* windage factors greatly improved accuracy regarding entrainment or expulsion from eddies when compared to tracked *Sargassum* mats in the ocean.
+The authors showed that including *ad hoc* windage factors greatly improved accuracy regarding entrainment or expulsion from eddys when compared to tracked *Sargassum* mats in the ocean.
 Langrangian trajectory modeling is clearly a useful tool; however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
@@
 
-Past studies have focused on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddies and vortices in the accumulation of individual kelp plants at the surface.
-Gyres, eddies, and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018].
+Past studies have focused on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddys and vortices in the accumulation of individual kelp plants at the surface.
+Gyres, eddys, and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018].
 Investigations into  the dispersal of marine debris have also included the possible effects of inertia in accumulation of ocean drifter instruments in mesoscale features.
@@
 The different degrees of wave and wind exposure are meant to reflect the different buoyancy characteristics, as well as identify the role of magnitude of the drag components on overall trajectory.
-The modeling of single kelp offers the opportunity to study the role of eddies and vortices in accumulation points on the surface.
+The modeling of single kelp offers the opportunity to study the role of eddys and vortices in accumulation points on the surface.
 
@@
 The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A).
-A release site was chosen for this study, Kommetjie, which is located on the Atlantic side of the Cape Peninsula (Figure 2B).
-The hydrodynamics in the study region are driven by a complex interaction of wind and waves, and are also influenced by larger oceanic processes.
+Kommetjie, located on the Atlantic side of the Cape Peninsula (Figure 2B), was chosen as a release site.
+The hydrodynamics in the study region are characterized by a complex interaction of wind, sea waves and ocean swell, and are also influenced by larger-scale regional oceanic processes.
 In South Africa, the larger processes can be attributed to the Agulhas Current (AC) and the Benguela Current (BC) [@lutjeharms2000;@lutjeharms2007;@garzoli1996;@hutchings2009;@rubio2009].
 The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].
+
 The region where the retroflection occurs is known as the Agulhas Retroflection and consists of high levels of mesoscale variability [@garzoli1996].
@@
 The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009].
-The Benguela region is characterised by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003].
+The Benguela region is characterized by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003;@veitch2018; @veitch2017].
 It is known as the Benguela Upwelling System (BUS).
-The BUS starts at 27 S and extends southward to 35 S [@shannon1996].
-The region is also characterised by high amounts of mesoscale variability in the form of eddies and filaments [@blanke2002; @blanke2005; @rubio2009].
-The Benguela inshore region is dominated by a northwestward flow because of topographical steering, wind stress and interactions with passing Agulhas Rings and eddies [@veitch2010].
+The BUS starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996].
+The region is also characterized by large amounts of mesoscale variability in the form of eddys and filaments [@blanke2002; @blanke2005; @rubio2009].
+The Benguela inshore region is dominated by a northwestward flow because of topographical steering, wind stress and interactions with passing Agulhas Rings and eddys, and is known as the Benguela Jet [@stramma1989; @veitch2010; @veitch2017; @veitch2018; @hutchings2009].
+The stability of the jet varies seasonally with more stable flows in winter; in January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results in a weak narrow inshore branch that flows along the shelf edge and a strong flowing north-westwards offshore branch.
+The area is also characterized by high upwelling activity driven by strong westerly winds [@hutchings2009].
+These winds are most intense during summer (December-March) which favors upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
+In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela Jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
 
 ```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1: Map of the study domain and the main features pertaining to transport in the region.
-In panel A, the map of the entire South African coastline and the main currents is presented.
-The resulting eddies and filaments from the Agulhas retroflection are indicated, and the position of the study area is drawn in as a black rectangle.
+In panel A, the map of the western portion of South African coastline and the main currents is presented.
+The resulting eddys and filaments froming from the Agulhas retroflection are indicated, and the position of the study area is drawn in as a black rectangle.
 Detail of the study area is in panel B.
-Here the splitting of the Benguela Jet, eddie features, and the general flow of the study domain are presented.
+Here the splitting of the Benguela Jet, eddy features, and the general flow within the study domain are presented.
 The release site of simulated kelp ‘particles’ is represented by a blue star."}
@@
 
-The study domain falls within the southern Benguela upwelling ecosystem, which is a region of high mesoscale variability and activity in the form of eddies, filaments and vortices generated from the Agulhas retroflection [@veitch2018; @veitch2017] they generally flow northward along the coastline and begin to move offshore near Cape Columbine.
-The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, Eckman transport and geostrophic currents [@stramma1989; @veitch2017; @veitch2018; @hutchings2009].
-The stability of the jet varies seasonally with more stable flows in winter; and in January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results in a weak narrow inshore branch that flows along the shelf edge and a strong flowing north-westwards offshore branch.
-The area is also characterized by high upwelling activity driven by strong westerly winds [@hutchings2009].
-These winds are most intense during summer (December-March) which favors upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
-In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
-
 ## Particle tracking model
@@
 Some authors suggest that most macroalgae become entangled through ocean currents and not through the hydrodynamic forces that dislodge them.
-Large kelp rafts are not a feature around the coast of the Cape Peninsula; rather, solitary kelp are (pers.
+Large kelp rafts are not a feature around the coast of the Cape Peninsula; rather, solitary kelp is (pers.
 obs).
@@
 maxima* individuals will be considered.
-We performed numerical calculations of particle trajectory with a Lagrangian method [@delandmeter2019], and be described by the equation
+We performed numerical calculations of particle trajectory with a Lagrangian method [@delandmeter2019], which can be described by the equation:
 
 $$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
-where $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model at particle point in time, $t$ is time $X_b(t)$ is a change in position due to 'behavior' of the particle at a particular point in time.
-In order to investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.
+Here, $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model at particle point in time, $t$ is time $X_b(t)$ is a change in position due to 'behavior' of the particle at a particular point in time.
+To investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.
 
@@
 
-Virtual particles were released with no drag or wind effects from the study site in order to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'.
+Virtual particles were released with no drag or wind effects from the study site to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag and are referred to hereafter as 'passive particles'.
 All particles that have any form of drag exposure are hereafter referred to as 'kelp particles'.
 All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018].
-Brownian motion was also incorporated in order to account for any stochasticity not resolved by the model.
+Brownian motion was also incorporated to account for any stochasticity not resolved by the model.
 These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018].
@@
 These time periods were chosen based on variable mesoscale activities and stability of the Bengulea Jet that occurs through the year.
-For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity ; while in winter the Benguela jet begins to stablize until summer when the begins to split again [@rubio2009; @veitch2018; @ragoasha2019].
+For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity while in winter the Benguela jet begins to stabilize until summer when the begins to split again [@rubio2009; @veitch2018; @ragoasha2019].
 
@@
 Since the aim of this investigation was to investigate the relative surface behavior of virtual kelp particles, only the surface layer was used for simulations.
-It should also be noted that the Copernicus model used places velocity vectors over some of the land covering the coastal domain.
+It should also be noted that the Copernicus model (see below) used places velocity vectors over some of the land covering the coastal domain.
 This was not considered a concern, as once again, the aim was to investigate surface drift behavior.
-A summary of the parameters for each simulation are provided in table 1.To determine if vortices and eddies play a role in kelp accumulation on the surface, a separate set of simulations were performed.
-Unlike the previous simulations these simulation released a particle every hour over the same time periods.
-
-In order to determine if the shape co-efficient used plays a role in model used and to validate the use of a sphereical shape coefficent, simulations of kelps with the maximum surface area and hydrodynamic drag exposure was run using the shape coefficient for a cylinder to compare to a simulations using a spheres.
-Simulations were run for 30 days using a fourth order Runge-Kutte advection scheme with brownian motion and a particle set of size 1000 for the same seasons as per previous simulations.
+A summary of the parameters for each simulation are provided in table 1.
+To determine if vortices and eddys play a role in kelp accumulation on the surface, a separate set of simulations were performed.
+Unlike the previous simulations, these simulations released a particle every hour over the same time periods.
+
+In order to determine if the shape coefficient used plays a role in model used and to validate the use of a spherical shape coefficient, simulations of kelps with the maximum surface area and hydrodynamic drag exposure was run using the shape coefficient for a cylinder to compare to a simulations using a sphere.
+Simulations were run for 30 days using a fourth order Runge-Kutte advection scheme with Brownian motion and a particle set of size 1000 for the same seasons as per previous simulations.
 
@@
 sim_table <- set_caption(sim_table, "Details for each simulation run for the study.
-Type refers to either a passive particle with no drag behaviour or a kelp particle that has been advected with any drag exposure.
-The drag exposure column refers to the parameterisations regarding the level of water and wind drag simulated.
+Type refers to either a passive particle with no drag behavior or a kelp particle that has been advected with any drag exposure.
+The drag exposure column refers to the parameterizations regarding the level of water and wind drag simulated.
 All simulations were run with a Runge-Kutte (4th) advection scheme, Brownian motion, 30 days, and included 1000 particles for a 2018 hindcast.")
@@
 The model has a spatial resolution of 0.08$^\circ$ and is interpolated on an Arakawa C native grid.
-In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (Global Ocean Wind L4_NRT_012_004).
+To incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (Global Ocean Wind L4_NRT_012_004).
 The outputs from the model are composed of 6-hourly averaged fields of surface 10 m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
@@
 For details refer to figure 2.
-The dimensional data needed were estimated in cases where data was not available for that particular morphological characteristic.
+The dimensional data needed were estimated in cases where data was not available for that morphological characteristic.
 The bulb/pneumatocyst is a highly variable morphological characteristic and in some cases can appear absent; the same is true for the holdfast.
@@
 
-In order to compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated and density distribution maps were produced.
+To compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated, and density distribution maps were produced.
 Density distribution maps are an effective method for assessing pathways and ocean connectivity [@vanSebille2018].
@@
 Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particles [@vanSebille2018].
-If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; @putman2018; @vanSebille2018].
+If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity because of momentum energy transfer [@putman2020; @putman2018; @vanSebille2018].
 
@@
 
-The role of eddies that exist within the study domain over the time period play a clear role in the accumulation and dispersal of kelp particles.
+The role of eddys that exist within the study domain over the time period play a clear role in the accumulation and dispersal of kelp particles.
 Passive particles tend to remain on the outer edge of the eddy and then are expelled on the other side and continue to follow the ocean current (see Figure 8).
@@
 
-The treatment of drifting macroalgae as purely Lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @miron2020].
+The treatment of drifting macroalgae as purely Lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddys or vortices [@fossette2012; @miron2020].
 Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], past research has focused on large macroalgal-rafts which can vary greatly in size and surface area.
@@
 All the simulations that included any form of drag resulted in a high density of particles entrained in an eddy.
-The entrainment and expulsion from eddies and vortices are an important characteristic of floating macroalgae.
-Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real *Sargassum* rafts.
+The entrainment and expulsion from eddys and vortices are an important characteristic of floating macroalgae.
+Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddys has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real *Sargassum* rafts.
 However, these studies investigated large rafts and not floating individuals.
@@
 Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered.
-The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles, and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.
-
-The role of eddies and vortices in the entrainment of kelp is further supported by the accumulation simulations.
+The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles, and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddys and vortices.
+
+The role of eddys and vortices in the entrainment of kelp is further supported by the accumulation simulations.
 These results clearly show that the inclusion of drag greatly increases the number of particles entrained in the eddy within the study domain.
-Eddies and vortices may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in area below the eddy.
+eddys and vortices may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in area below the eddy.
 In a study by @woodborne1989 references a site offshore approximately 120 km from Cape Columbine where dredged kelp was found.
@@
 maxima* particles tend to follow the prevailing surface currents and the inclusion of wind causes the trajectories to differ greatly from that of a purely Lagrangian particle.
-Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies.
+Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddys.
 Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E.

### 2021-05-02 thesis_chapter_3.Rmd
results AJ comments

@@
 When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories was evident.
-In all simulations particles flowed in a north-westward direction.
+In all simulations, particles flowed in a north-westward direction.
 
@@
 Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 2E-F).
-Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore; resulting in entrainment within an eddy.
+Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore, resulting in entrainment within an eddy.
 Similar behavior is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag.
@@
 Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag ended a greater distance from the release site compared to the passive particles.
-This was also evident when comparing kelp particles with different cross-sectional areas and exposure to wind and hydrodynamic drag with passive particles; with the end location a further from the release location.
+This was also evident when comparing kelp particles with different cross-sectional areas and exposure to wind and hydrodynamic drag with passive particles, with the end location a further from the release location.
 In both the April and October simulations all particles flowed north-westward, while in winter the mean trajectory resulted in entrainment within an eddy.
@@
 
-Across all simulations particles flowed in a north-westward direction.
+Across all simulations, particles flowed in a north-westward direction.
 The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids.
-In addition the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 3).
+In addition, the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 3).
 Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.
@@
 Only slight variations across cross-sectional areas and drag exposure scenarios are evident.
-Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in a eddy located westward of the release site compared to passive particles (see Figure 5).
+Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in an eddy located westward of the release site compared to passive particles (see Figure 5).
 Variation in density along the mean trajectory and eddy for all kelp particles is evident.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for scenarios 14 - 26, for April 2018 (autumn).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end of the run for scenarios 14 - 26, for April 2018 (autumn).
 The plots are regular hexagons within the density of particles were calculated.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end run time for scenarios 27 - 39, for July 2018 (winter).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end of the run for scenarios 27 - 39, for July 2018 (winter).
 The plots are regular hexagons within the density of particles were calculated.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end run time for scenarios 40 - 52, for October 2018 (spring).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end of the run for scenarios 40 - 52, for October 2018 (spring).
 The plots are regular hexagons within the density of particles were calculated.
@@
 
-There were significant differences among the median distances traveled for all particles in each simulation (Figure 4; p < 0.05).
+There were significant differences among the median distances traveled for all particles in each simulation (Figure 4; *p* < 0.05).
 The simulations that considered for 100% hydrodynamic drag traveled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
@@
 When comparing distances between passive and kelp particles for April, significant differences are evident.
-(Figure 11, simulations 17-29; p < 0.05).
+(Figure 11, simulations 17-29; *p* < 0.05).
 Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas traveled shorter distance across compared to other kelp particles.
 The shortest distance traveled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
-No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; p > 0.05).
-In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; p < 0.05).
+No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; *p* > 0.05).
+In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; *p* < 0.05).
 Kelp particles with only 90% hydrodynamic drag across cross-sectional areas traveled shorter distance compared to all other particles.
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 7: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for scenarios 1 - 13, for January 2018 (summer).
-The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers.
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 7: Boxplots of total distance travelled by all particles from the release site over the course of the simulations for scenarios 1 - 52.
+The *x*-axis represents the scenario number, and the *y*-axis is the distance travelled in kilometers.
 The lower and upper hinges correspond to the first and third quartiles.

### 2021-05-02 thesis_chapter_3.Rmd
AJ comments

@@
 The mean virtual kelp trajectory deviated from the mean passive trajectory when including varying levels of hydrodynamic and wind drag scenarios.
-The results from this study show the inclusion of hydrodynamic and wind drag results in observable entrainment patterns, and broadens our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case that of *E.
+The results from this study show the inclusion of hydrodynamic and wind drag results in observable entrainment patterns and broadens our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case that of *E.
 maxima*.
@@
 A study by @putman2020 used *ad hoc* windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic *Sargassum sp*.
-The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked *Sargassum sp* mats which was partly due to the inclusion of momentum energy transfer from wind.
+The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked *Sargassum sp* mats, which was partly due to the inclusion of momentum energy transfer from wind.
 This is also reflected in the analysis comparing distances travelled by all the particles in each simulation.
@@
 The different trajectory patterns between seasons reflects the underlying seasonal variation in regional oceanography.
-Transport in the study region is driven by a combination d variation of along-shore jets and wind induced Ekman transport of surface water [@rubio2009; @veitch2018; @ragoasha2019].
+Transport in the study region is driven by a combination d variation of alongshore jets and wind induced Ekman transport of surface water [@rubio2009; @veitch2018; @ragoasha2019].
 A study by @ragoasha2019 investigated the success of Lagrangian particles reaching St.
@@
 During upwelling season (December - March) transport success is less efficient as particles tend to flow offshore.
-The move offshore is a result of the bifurification of the Benguela Jet, which splits as a result of shoaling, and causes particles to be exposed to higher amounts of mesoscale activity.
+The move offshore is a result of the bifurification of the Benguela Jet, which splits because of shoaling and causes particles to be exposed to higher amounts of mesoscale activity.
 In the winter months, the Benguela Jet is still split but the flow weakens which results in lower transport success [@ragoasha2019].
-Similar patterns were found in this study, suggesting that the differences in seasonal patterns reflects the regional oceanographic features, particularly the Benguela Jet.
+Similar patterns were found in this study, suggesting that the difference in seasonal patterns reflects the regional oceanographic features, particularly the Benguela Jet.
 
@@
 The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia.
-Although drag forces and not inertia were used in this study, the same conclusion applies; drag is therefore an important characteristic to consider when investigating or predicting solitary drifting macroalgae trajectory.
+Although drag force and not inertia was used in this study, the same conclusion applies; drag is therefore an important characteristic to consider when investigating or predicting solitary drifting macroalgae trajectory.
 
@@
 The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due as juveniles do not swim fast enough to alter their overall trajectory.
-Therefore, the differences in cross-sectional area types may not be significant enough to have an effect on the relevant zonal and meridional velocity vectors.
+Therefore, the differences in cross-sectional area types may not be significant enough to influence the relevant zonal and meridional velocity vectors.
 Another possible reason is the resolution of the ocean model used.
 The resolution of the underlying ocean model is an important aspect of Lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications.
-A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study.
-The authors found that the inclusion of Brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications.
-Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.
+A study by @hart2018 assessed the inclusion of stochastic motion, wind, and currents into forecasting for search and rescue using the same model ocean model used in this study.
+The authors found that the inclusion of Brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents, and stochastic motion greatly improved forecasting applications.
+Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not because of the resolution of the model.
 
@@
 Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered.
-The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles, and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddys and vortices.
+The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddys and vortices.
 
@@
 In a study by @woodborne1989 references a site offshore approximately 120 km from Cape Columbine where dredged kelp was found.
-The area referenced by @woodborne1989 coincides with the area where virtual kelp are entrained by an eddy.
+The area referenced by @woodborne1989 coincides with the area where virtual kelp is entrained by an eddy.
 The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory.

### 2021-05-02 thesis_chapter_3.Rmd
AJ almost final chp 3

@@
 The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
-The minimum, mean and maximum were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (see table 2).
-The trajectory of these 'types' of cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
+The minimum, mean and maximum s-sectional areas (Ac, see table 2) were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (see table 2).
+The trajectory of the different cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
 
@@
 
-The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (Figure 2A).
-When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen; rather, there is only a slight variation.
-The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation (Figure 2B).
+The mean trajectories of the virtual kelps with different cross-sectional areas are similar to those of the passive particles during the month of January (Austral summer; Figure 3A).
+When comparing mean trajectories between virtual kelp with different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen; rather, there is only a slight variation in the overall trajectory.
+The results from the simulations including different hydrodynamic and wind drag exposures and kelp cross-sectional areas show differences in mean trajectory when compared to the passive simulation (Figure 3B).
 The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well the final location of the kelp particles being further from the release site.
@@
 
-The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (autumn).
-Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag traveled further northward compared to the passive particles (see Figure 2C-D).
-The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles.
+The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (Austral autumn).
+Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag traveled further northward compared to the passive particles (see Figure 3C-D).
+The mean trajectories of virtual kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles.
 Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles traveled less distance compared to kelp particles with both wind and hydrodynamic drag.
-Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 2E-F).
+Differences in mean trajectories between kelp and passive particles were also evident in the July (Austral winter) simulations (see Figure 3E-F).
 Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore, resulting in entrainment within an eddy.
@@
 In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles traveled initially southward and then westward offshore and also becoming entrained in an eddy.
-In the October (Spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 7F-G).
-Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag ended a greater distance from the release site compared to the passive particles.
+In the October (Austral spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 3F-G).
+Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag stop a greater distance from the release site compared to the passive particles.
 This was also evident when comparing kelp particles with different cross-sectional areas and exposure to wind and hydrodynamic drag with passive particles, with the end location a further from the release location.
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=7.5, fig.cap="Figure 3: Comparison of mean trajectories.
-Panel (A) represents hydrodynamic drag only simulations and Panel (B) represents simulations for both wind and hydrodynamic drag, for January, Panel (C) represents hydrodynamic drag only simulations and Panel (D) represents simulations for both wind and hydrodynamic drag, for April.
-Panel (E) represents hydrodynamic drag only simulations and Panel (F) represents simulations for both wind and hydrodynamic drag, for July.
-Panel (G) represents hydrodynamic drag only simulations and Panel (H) represents simulations for both wind and hydrodynamic drag, for October.
+Simulations of hydrodynamic drag are presented in Panels A, C, E, G for January, April, July, and October, respectively.
+Results of simulations including both hydrodynamic drag and wind are in Panels B, D, F, H for January, April, July, and October, respectively.
 In all panels the passive simulations are represented by black lines."}
@@
 The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids.
-In addition, the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 3).
+In addition, the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 4).
 Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.
@@
 Kelp particles with only hydrodynamic drag across cross-sectional areas showed lower density of particles along the mean trajectory compared to kelp particles with both wind and hydrodynamic drag.
-October simulations showed no differences in density of particles along the mean trajectory when comparing passive and kelp particles.
+Simulations run for October show no difference in density of particles along the mean trajectory when comparing passive and kelp particles.
 Only slight variations in density and trajectory can be seen with no entrainment (see Figure 6).
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time for scenarios 1 - 13, for January 2018 (summer).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for scenarios 1 - 13, for January 2018 (summer).
 The plots are regular hexagons within the density of particles were calculated.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end of the run for scenarios 14 - 26, for April 2018 (autumn).
-The plots are regular hexagons within the density of particles were calculated.
-The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end of the run for scenarios 14 - 26, for April 2018 (autumn).
+See Figure 4 for further details.', fig.height=8, fig.width=8}
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end of the run for scenarios 27 - 39, for July 2018 (winter).
-The plots are regular hexagons within the density of particles were calculated.
-The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end of the run for scenarios 27 - 39, for July 2018 (winter).
+See Figure 4 for further details.', fig.height=8, fig.width=8}
 
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end of the run for scenarios 40 - 52, for October 2018 (spring).
-The plots are regular hexagons within the density of particles were calculated.
-The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}
+See Figure 4 for further details.', fig.height=8, fig.width=8}
 
@@
 
-```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of age between passive and kelp accumulation scenarios 53-60.
+```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of particle age between passive and kelp accumulation scenarios 53-60.
 For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=9, fig.width=7}
@@
 The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (see appendix).
-This was also reflected when comparing distance traveled by particles and particle density plots (see appendix).
-The simulation which used a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (see appendix).
+This is also seen when comparing distance traveled by particles and particle density plots (see appendix).
+The simulation employing a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (see appendix).
 
@@
 
-The treatment of drifting macroalgae as purely Lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddys or vortices [@fossette2012; @miron2020].
+The treatment of drifting macroalgae as purely Lagrangian does not accurately determine patterns of passive dispersal, such as entrainment in eddys or vortices [@fossette2012; @miron2020].
 Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], past research has focused on large macroalgal-rafts which can vary greatly in size and surface area.
@@
 However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object [@hackett2006; @putman2020].
-This causes the kelp particles to travel further as well as exposed to different time varying flows.
+This causes the kelp particles to travel further as well as become exposed to different time varying flows.
 Past studies have shown this when including windage as a function of inertia [@putman2018;@brooks2019; @putman2020].
@@
 The particles in the simulations that included wind drag traveled significantly greater distances compared to the passive or hydrodynamic only drag simulations.
-Similar patterns are seen in the simulations run for April (autumn), July (winter), and October (spring); however unlike in January (summer) where only kelp particles with combined drag (hydrodynamic and wind) flowed further, even kelp particles with only hydrodynamic drag flow further compared to passive particles in these times of the year.
+Similar patterns are seen in the simulations run for April (Austral autumn), July (Austral winter), and October (Austral spring); however unlike in January (Austral summer) where only kelp particles with combined drag (hydrodynamic and wind) flowed further, even kelp particles with only hydrodynamic drag flow further compared to passive particles in these times of the year.
 This may be a result of the kelp particles exposed to different time-varying flows, which results in the kelp particles following a different surface current vector compared to passive particles.
@@
 
-Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape; however, no research exists to clarify the drifting characteristics of solitary macroalgae.
+Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape; however, prior to our study no research existed to clarify the drifting characteristics of solitary macroalgae.
 Understanding how to accurately model the distribution of solitary kelp can lead to be understanding of accumulation zones and sinks and ultimately the ecological links pertaining to those processes.
-The findings from this study indicate that solitary floating virtual *E.
-maxima* particles tend to follow the prevailing surface currents and the inclusion of wind causes the trajectories to differ greatly from that of a purely Lagrangian particle.
-Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddys.
+The findings from this study indicated that solitary floating virtual *E.
+maxima* particles tend to follow the prevailing surface currents and the inclusion of wind caused the trajectories to differ greatly from that of a purely Lagrangian particle.
+Furthermore, the inclusion of both hydrodynamic and wind drag caused clustering of particles along the trajectory as well causing particles to become entrained in eddys.
 Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E.
 maxima*.
-In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory, but it does cause higher density of particles to become entrained in an eddy.
+In addition, the use of a cylindrical shape coefficient had no effect on mean trajectory, but it does cause higher density of particles to become entrained in an eddy.
 Overall, the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not.
-This study also provides an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately.
-In addition, the results from this study highlights the importance of regional oceanographic process and how these vary seasonally to better understand how kelp flotsam is dispersed in the ocean.
+This study also provided an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately.
+Our results from this study highlights the importance of regional oceanographic processes to better understand how kelp flotsam is dispersed in the ocean.
 In particular, the current study highlights the role of the Benguela Jet and mesoscale processes and how these vary seasonally; ultimately determining transport success around the coast.
-Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal as well as how these vary seasonally.
+Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal.
 The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates and raft-times.

### 2021-05-03 thesis_chapter_3.Rmd
final comments AJ. new output

@@
 This approach has been used in past research investigating pelagic *Sargassum* rafts [@putman2020;@putman2018].
-The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1).
+The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1, Appendix).
 Each scenario consisted of a particle set with a size of 1000 were released from Kommetjie at the start of simulation.
@@
 Simulations were run for 30 days using a fourth order Runge-Kutte advection scheme with Brownian motion and a particle set of size 1000 for the same seasons as per previous simulations.
-
-```{r sim table, eval=TRUE, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
-
-set_flextable_defaults(fonts_ignore=TRUE)
-
-sim_table <- read.csv("tables/chapter_3/simulation_table.csv", sep = ";")
-
-sim_table <- flextable(data = sim_table)
-sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
-sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Drag.Exposure = "Drag Exposure level")
-sim_table <- align(x= sim_table, align = "center", part = "all")
-sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
-sim_table <- set_caption(sim_table, "Details for each simulation run for the study.
-Type refers to either a passive particle with no drag behavior or a kelp particle that has been advected with any drag exposure.
-The drag exposure column refers to the parameterizations regarding the level of water and wind drag simulated.
-All simulations were run with a Runge-Kutte (4th) advection scheme, Brownian motion, 30 days, and included 1000 particles for a 2018 hindcast.")
-sim_table
-
-```
 
@@
 area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
+area_table <- fontsize(area_table, size = 8, part = "body")
+area_table <- fontsize(area_table, size = 8, part = "header")
 area_table <- set_caption(area_table, "Table 2: Table summarising the attributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulations.
@@
   scale_colour_discrete("Type") +  
+  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   scale_colour_discrete("Type") + 
+  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   theme(plot.title = element_text(hjust = 0.5)) +
+  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
   theme_bw()
@@
   guides(fill=guide_legend(title="Number of particles")) +
+  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
   theme_bw() +
@@
   guides(fill=guide_legend(title="Age")) +
+  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
   theme_bw() +
@@
 
+```{r sim table, eval=TRUE, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
+
+set_flextable_defaults(fonts_ignore=TRUE)
+
+sim_table <- read.csv("tables/chapter_3/simulation_table.csv", sep = ";")
+
+sim_table <- flextable(data = sim_table)
+sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
+sim_table <- fontsize(sim_table, size = 8, part = "body")
+sim_table <- fontsize(sim_table, size = 8, part = "header")
+sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Drag.Exposure = "Drag Exposure level")
+sim_table <- align(x= sim_table, align = "center", part = "all")
+sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
+sim_table <- set_caption(sim_table, "Details for each simulation run for the study.
+Type refers to either a passive particle with no drag behavior or a kelp particle that has been advected with any drag exposure.
+The drag exposure column refers to the parameterizations regarding the level of water and wind drag simulated.
+All simulations were run with a Runge-Kutte (4th) advection scheme, Brownian motion, 30 days, and included 1000 particles for a 2018 hindcast.")
+sim_table
+
+```
+
 ```{r shape data, message=FALSE, echo=FALSE, warning=FALSE}
@@
 shape_plot <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
   #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
@@
   #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  coord_fixed(ratio=1/1) # fix the aspect ratio of plot
 
@@
 shape_plot_April <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
   #geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5)) +
+  coord_fixed(ratio=1/1)  # fix the aspect ratio of plot
 
@@
 shape_plot_July <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
   #geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
@@
 shape_plot_Oct <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
   #geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+

### 2021-05-04 thesis_chapter_3.Rmd
final corrections for chapter 3

@@
 author: 
-- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape South African Observation Network]
+- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, 41 Market Place Viaduct Harbour, 1010 Auckland, Private Bag 99940, Newmarket, Auckland 1149, New Zealand], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
 output:
   word_document:
-    reference_docx: markdown/template.docx
+    reference_docx: markdown/wiley_template.docx
   pdf_document:
@@
 bibliography: references.bib
-csl: markdown/frontiers.csl.txt
+csl: markdown/apa.csl.txt
 abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_water, aes(x = lon, y = lat, colour = type))+
@@
   scale_colour_discrete("Type") +  
-  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  #coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_ww, aes(x = lon, y = lat, colour = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_water_April, aes(x = lon, y = lat, colour = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill ="grey") +
-  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, colour = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_water_July, aes(x = lon, y = lat, colour = type))+
@@
   scale_colour_discrete("Type") + 
-  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  #coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, colour = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, colour = type))+
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  #coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
   theme_bw()
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, colour = type))+
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  #coord_fixed(ratio=1/1, clip = "on") + # fix the aspect ratio of plot
   theme_bw()
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=7.5, fig.cap="Figure 3: Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=12, fig.width=7, fig.cap="Figure 3: Comparison of mean trajectories.
 Simulations of hydrodynamic drag are presented in Panels A, C, E, G for January, April, July, and October, respectively.
@@
   geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  #coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
   theme_bw() +
@@
   geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_point(data = pass_acc, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  #coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
   theme_bw() +
@@
   geom_path(data = pass_acc_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_path(data = kelp_acc_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_point(data = pass_acc_April, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_path(data = pass_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_path(data = kelp_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_point(data = pass_acc_July, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_path(data = pass_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_path(data = kelp_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_point(data = pass_acc_Oct, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
 
+The findings from this study broaden our understanding of macroalgal trajectory patterns and behavior in the ocean.
+In particular, the results show that consideration of drag and surface area are important when investigating macroalgal trajectory patterns.
 The treatment of drifting macroalgae as purely Lagrangian does not accurately determine patterns of passive dispersal, such as entrainment in eddys or vortices [@fossette2012; @miron2020].
 Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], past research has focused on large macroalgal-rafts which can vary greatly in size and surface area.
-None of the past research has taken advantage of the advancement in numerical ocean models and Lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae.
+None of the extant research has taken advantage of the advancement in numerical ocean models and Lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae.
 This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag.
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
   #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
@@
   #guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  coord_fixed(ratio=1/1) # fix the aspect ratio of plot
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5)) 
+  #coord_fixed(ratio=1/1) # fix the aspect ratio of plot
 
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  coord_fixed(ratio=1/1)  # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) 
+  #coord_fixed(ratio=1/1)  # fix the aspect ratio of plot
 

### 2021-05-05 thesis_chapter_3.Rmd
update 1

@@
 
-kelp_shape_figure <- rasterGrob(as.raster(readPNG("figures/chapter_3/chp3_shapedrawing.png")))
+kelp_shape_figure <- rasterGrob(as.raster(readPNG("figures/chapter_3/figure_2.png")))
 ggarrange(kelp_shape_figure)
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_water, aes(x = lon, y = lat, colour = type))+
@@
   scale_colour_discrete("Type") +  
-  #coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_ww, aes(x = lon, y = lat, colour = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_water_April, aes(x = lon, y = lat, colour = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill ="grey") +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, colour = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_water_July, aes(x = lon, y = lat, colour = type))+
@@
   scale_colour_discrete("Type") + 
-  #coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, colour = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, colour = type))+
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  #coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
   theme_bw()
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, colour = type))+
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  #coord_fixed(ratio=1/1, clip = "on") + # fix the aspect ratio of plot
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
   theme_bw()
@@
 ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, ncol = 2, nrow = 4, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("bottom"), labels = "AUTO", align = "hv")
-
 ```
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end of the run for scenarios 40 - 52, for October 2018 (spring).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 7: Comparison of density plots at the end of the run for scenarios 40 - 52, for October 2018 (spring).
 See Figure 4 for further details.', fig.height=8, fig.width=8}
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 7: Boxplots of total distance travelled by all particles from the release site over the course of the simulations for scenarios 1 - 52.
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 8: Boxplots of total distance travelled by all particles from the release site over the course of the simulations for scenarios 1 - 52.
 The *x*-axis represents the scenario number, and the *y*-axis is the distance travelled in kilometers.
@@
   geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  #coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
   theme_bw() +
@@
   geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_point(data = pass_acc, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   guides(fill=guide_legend(title="Age")) +
-  #coord_fixed(ratio=1/1) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
   theme_bw() +
@@
   geom_path(data = pass_acc_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_path(data = kelp_acc_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_point(data = pass_acc_April, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_path(data = pass_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_path(data = kelp_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_point(data = pass_acc_July, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_point(data = kelp_acc_July, aes(x = lon, y = lat, colour = age)) +
-  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
   coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
@@
   geom_path(data = pass_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_path(data = kelp_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
   geom_point(data = pass_acc_Oct, aes(x = lon, y = lat, colour = distance)) +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
   xlab("Longtitude") +
@@
 sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
-sim_table <- set_caption(sim_table, "Details for each simulation run for the study.
+sim_table <- set_caption(sim_table, "Table 1: Details for each simulation run for the study.
 Type refers to either a passive particle with no drag behavior or a kelp particle that has been advected with any drag exposure.

### 2021-05-05 thesis_chapter_3.Rmd
intro corrections...

@@
 abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
-Macroalgae are one of the most important natural dispersal vectors and are regarded as the 'tumble-weed' of the ocean.
+Macroalgae are one of the most important natural dispersal vectors  of marine organisms and are regarded as the 'tumble-weed of the ocean'.
 Despite many studies that exist on the topic, the relative role of wind and surface currents in influencing the trajectory of macroalgal dispersal is still unknown.
 Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a consenses as to the relative role of wind versus surface currents.
-In addition, past studies have not considered surface-area in relation to drag characteristics of floating macroalgae, both of which have been shown to play a role in the trajectory and accumulation of other flotsam.
+In addition, these studies have not considered surface-area in relation to drag characteristics of the macroalgae, both of which have been shown to play a role in the trajectory and accumulation of other flotsam.
 Advances in Lagrangian trajectory modeling have been rapid in recent years and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean.
@@
 
-There is a range of objects---both natural and anthropogenic---floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa.
-Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
+A range of natural and anthropogenic objects floats in the ocean, and macroalgae are regarded as one of the most important of these passive dispersal mechanisms of marine taxa.
+Floating kelps act as passive dispersal mechanisms for a range of marine taxa and is sometimes referred to as the ‘tumble-weed of the ocean’ [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
 Kelps also function as input of allochthonous materials to near and distant ecosystems and in this manner support the ecological function in a range of adjacent habitats [@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].
@@
 The positively buoyant pneumatocysts also cause plants to float to the surface when dislodged from the substratum.
-The structure and number of pneumatocysts varies between species.
+The structure and number of pneumatocysts varies among species.
 For example, species belonging to the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, and *Ecklonia maxima* have single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018].
@@
 In addition, the trajectories of tagged kelp were more consistent with the formation of eddys during winter.
-Previous studies have also identified wind as an important mechanism of dispersal in wind dominated ocean systems.
+Studies have also identified wind as an important mechanism of dispersal in wind dominated ocean systems.
 For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating bull kelp *Durvillaea antarctica*.
@@
 The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011].
-Past research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
+Research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
 The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic-form drag [@breivik2011; @griffin2017].
-In addition, past research conducted by the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object.
+Work by the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object.
 If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].
@@
 To accurately determine the trajectory of marine macroalgae, both water and wind drag need to be considered.
-Past research by @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
-However, this work does not consider biological material such as macroalgae.
-Although drag estimates do exist, these have been applied for macroalgae not detached from the substratum and are regarded as fixed-point estimates.
+@allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
+However, this work did not consider biological material such as macroalgae.
+Although drag estimates do exist, these have been applied for macroalgae not detached from the substratum and are regarded as fixed-point estimates [@Denny1997; @Denny2002a; @Denny2002b; @Gaylord1997-ow; @Gaylord2008-qv].
 Although various approaches have been used in the past to investigate floating macroalgal trajectories, very few studies have employed the use of Langrangian trajectory modeling.
-Furthermore, none of the existing studies using this modeling approach have considered macroalgal morphology cross-sectional area (*i.e.* shape) as an aspect of drag and ultimately trajectory.
-Studies by @brooks2019, @putman2018, and @putman2020 used Lagrangian approaches to investigate the effects of inertia, raft-size, and windage on macroalgae trajectory.
-@brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customized Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
-The results showed that the trajectory of pelagic *Sargussum* was significantly influenced by inertia and the radial size of the rafts.
-Other work by @putman2020 investigated the effect of including windage in macroalgae trajectory simulations.
-The authors showed that including *ad hoc* windage factors greatly improved accuracy regarding entrainment or expulsion from eddys when compared to tracked *Sargassum* mats in the ocean.
+Furthermore, none of these studies using this modeling approach have considered macroalgal morphology cross-sectional area (*i.e.* shape) as an aspect of drag and ultimately trajectory.
+@brooks2019, @putman2018, and @putman2020 used Lagrangian approaches to investigate the effects of inertia, raft-size, and windage on macroalgae trajectory.
+For example, @brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customized Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
+Their results showed that the trajectory of pelagic *Sargussum* was significantly influenced by inertia and the radial size of the rafts.
+@putman2020 investigated the effect of including windage in macroalgae trajectory simulations and showed that including *ad hoc* windage factors greatly improved accuracy regarding entrainment or expulsion from eddys when compared to tracked *Sargassum* mats in the ocean.
 Langrangian trajectory modeling is clearly a useful tool; however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
-The radial size included by @brooks2019 is an improvement, however this approach does not consider the complex morphology on an individual level into account.
-
-Past studies have focused on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddys and vortices in the accumulation of individual kelp plants at the surface.
+The radial size included by @brooks2019 is an improvement but this approach does not consider the complex morphology on an individual level into account.
+
+Studies have focused on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddys and vortices in the accumulation of individual kelp plants at the surface.
 Gyres, eddys, and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018].
 Investigations into  the dispersal of marine debris have also included the possible effects of inertia in accumulation of ocean drifter instruments in mesoscale features.
-For example, a study by @beron2016 used a modeling approach to investigate the role of Eckman transport and inertia to explain the observed behavior of undrogued drifters in the ocean gyres.
+For example, @beron2016 used a modeling approach to investigate the role of Eckman transport and inertia to explain the observed behavior of undrogued drifters in the ocean gyres.
 Undrogued drifter trajectory is affected by wind, unlike the drogue counter parts which makes the effects of wind negligible.
 The results showed that the combined effect inertia, not Eckman transport, played a significant role in the observed drifter behavior [@beron2016].
-@beron2016 also noted that the presence of other flotsam, along with the undrogued drifters in ocean gyres, suggests that aspects of inertia also play a role in accumulation of other marine debris [@beron2016].
-The approach by @beron2016 has also been improved upon by @beron2019 but also shows the significant role of inertia in flotsam trajectory and accumulation.
+These authors also noted that the presence of other flotsam, along with the undrogued drifters in ocean gyres, suggests that aspects of inertia also play a role in accumulation of other marine debris [@beron2016].
+The approach by @beron2016 has also been improved upon by @beron2019 but they also highlight the significant role of inertia in flotsam trajectory and accumulation.
 
 In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations.
-We aim to shed light on the 'behavior' of surface drifting macroalgae from a coastal environment into the open ocean, which will inform future characterization studies.
+We aim to determine the 'behavior' of surface drifting macroalgae from a coastal environment to the open ocean, which will inform future characterization studies.
 We will achieve this by comparing the trajectory outputs of virtual passive particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure.
@@
 
+<div id="refs"></div>
+
 # Appendix

### 2021-05-05 thesis_chapter_3.Rmd
chapter 3 updates

@@
 bibliography: references.bib
-csl: markdown/apa.csl.txt
+csl: markdown/limnologyandoceanography.csl
 abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
@@
 The hydrodynamics in the study region are characterized by a complex interaction of wind, sea waves and ocean swell, and are also influenced by larger-scale regional oceanic processes.
-In South Africa, the larger processes can be attributed to the Agulhas Current (AC) and the Benguela Current (BC) [@lutjeharms2000;@lutjeharms2007;@garzoli1996;@hutchings2009;@rubio2009].
+The larger processes in the region can be attributed to the Agulhas Current (AC) and the Benguela Current (BC) [@lutjeharms2000;@lutjeharms2007;@garzoli1996;@hutchings2009;@rubio2009].
 The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].
@@
 The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009].
-The Benguela region is characterized by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003;@veitch2018; @veitch2017].
-It is known as the Benguela Upwelling System (BUS).
+The Benguela region ---known as the Benguela Upwelling System (BUS)--- is characterized by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003;@veitch2018; @veitch2017].
 The BUS starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996].
 The region is also characterized by large amounts of mesoscale variability in the form of eddys and filaments [@blanke2002; @blanke2005; @rubio2009].
-The Benguela inshore region is dominated by a northwestward flow because of topographical steering, wind stress and interactions with passing Agulhas Rings and eddys, and is known as the Benguela Jet [@stramma1989; @veitch2010; @veitch2017; @veitch2018; @hutchings2009].
+Inshore, the BUS is dominated by a northwestward flow, known as the Benguela Jet, because of topographical steering, wind stress and interactions with passing Agulhas Rings and eddys [@stramma1989; @veitch2010; @veitch2017; @veitch2018; @hutchings2009].
 The stability of the jet varies seasonally with more stable flows in winter; in January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results in a weak narrow inshore branch that flows along the shelf edge and a strong flowing north-westwards offshore branch.
 The area is also characterized by high upwelling activity driven by strong westerly winds [@hutchings2009].
-These winds are most intense during summer (December-March) which favors upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
+These winds are most intense during Austral summer (December-March) which favors upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
 In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela Jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
@@
 #ggarrange(map_chp3A, map_chp3B, ncol = 2, nrow = 1, labels = "AUTO", align = c("h"))
-map_chp3_combined <- rasterGrob(as.raster(readPNG("map/ch3_map_final.png")))
+map_chp3_combined <- rasterGrob(as.raster(readPNG("map/chp3_map_final.png")))
 ggarrange(map_chp3_combined)
@@
 The model is based on parameterizations that have been used for previous studies investigating iceberg, capsized marine vessels and microplastics.
-It should be noted that most of the previous studies regarding macroalgae trajectory have investigated rafts and not solitary floating individuals.
-A kelp raft will vary in size and can range from a few meters across to large dense mats the size of a rugby field.
+It should be noted that most of the previous studies regarding macroalgal trajectories have investigated rafts and not solitary floating individuals.
+A kelp raft will vary in size and can range from a few meters across to large dense mats the size of a sports field.
 Some authors suggest that most macroalgae become entangled through ocean currents and not through the hydrodynamic forces that dislodge them.
@@
 
-This study simulated two scenarios; one where the plant is fully submerged and only exposed to hydrodynamic drag; and the other where parts of the plants are partially exposed resulting in a combination of hydrodynamic and air form drag.
+This study simulated two scenarios: one where the plant is fully submerged and only exposed to hydrodynamic drag, and the other where parts of the plants are partially exposed resulting in a combination of hydrodynamic and air form drag.
 Drag based on shape-based coefficients were used in the calculations for determining the overall kelp velocity vectors.
@@
 This approach has been employed in modeling iceberg drift trajectory [@andersson2017; @lichey2001; @eik2009] and was adapted to suit this study.
-The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was, $$m({d_u/d_t})=F_d + F_a + F_f$$
+The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was $$m({d_u/d_t})=F_d + F_a + F_f$$
 where $m$ is kelp mass, $d_u$ the virtual kelp particle velocity, $F_d$ the hydrodynamic drag, $F_a$ is the wind drag, $F_f$ the surface-current flow, and $d_t$ is the time step.
@@
 These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018].
-The time period used in all simulations were January (summer), April (Autumn), July (winter), and October (Spring) of 2018.
+The time period used in all simulations were January (Austral summer), April (Austral autumn), July (Austral winter), and October (Austral spring) of 2018.
 These time periods were chosen based on variable mesoscale activities and stability of the Bengulea Jet that occurs through the year.
-For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity while in winter the Benguela jet begins to stabilize until summer when the begins to split again [@rubio2009; @veitch2018; @ragoasha2019].
+For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity while in winter the Benguela Jet begins to stabilize until summer when the begins to split again [@rubio2009; @veitch2018; @ragoasha2019].
 
@@
 The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
-The minimum, mean and maximum s-sectional areas (Ac, see table 2) were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (see table 2).
+The minimum, mean and maximum cross-sectional areas (Ac, table 2) were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (table 2).
 The trajectory of the different cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
@@
 Plant dimensions are the measurements used, collected in previous work by @coppin2020, used to calculate the associated surface areas.
+Ac refers to the cross-sectional area calculated.
 The total surface areas for each plant characteristic for each surface area type make up the remaining columns.")
@@
 The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (Austral autumn).
-Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag traveled further northward compared to the passive particles (see Figure 3C-D).
+Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag traveled further northward compared to the passive particles (Figure 3C-D).
 The mean trajectories of virtual kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles.
@@
 In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles traveled initially southward and then westward offshore and also becoming entrained in an eddy.
-In the October (Austral spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 3F-G).
+In the October (Austral spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (Figure 3F-G).
 Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag stop a greater distance from the release site compared to the passive particles.
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  geom_path(data = total_mean_water, aes(x = lon, y = lat, colour = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  scale_colour_discrete("Type") +  
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
+  geom_path(data = total_mean_water, aes(x = lon, y = lat, linetype = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(linetype=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  geom_path(data = total_mean_ww, aes(x = lon, y = lat, colour = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  scale_colour_discrete("Type") +  
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
+  geom_path(data = total_mean_ww, aes(x = lon, y = lat, linetype = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(linetype=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  geom_path(data = total_mean_water_April, aes(x = lon, y = lat, colour = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  scale_colour_discrete("Type") +  
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
+  geom_path(data = total_mean_water_April, aes(x = lon, y = lat, linetype = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(linetype=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill ="grey") +
-  geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, colour = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  scale_colour_discrete("Type") +  
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
+  geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, linetype = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(linetype=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  geom_path(data = total_mean_water_July, aes(x = lon, y = lat, colour = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  scale_colour_discrete("Type") + 
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
+  geom_path(data = total_mean_water_July, aes(x = lon, y = lat, linetype = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(linetype=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
-  geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, colour = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  scale_colour_discrete("Type") +  
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
+  geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, linetype = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(linetype=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
-  geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, colour = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  scale_colour_discrete("Type") +  
+  geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, linetype = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(linetype=guide_legend(title="Simulation")) +
   theme(plot.title = element_text(hjust = 0.5)) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
   theme_bw()
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
-  geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, colour = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  scale_colour_discrete("Type") +  
-  guides(fill=guide_legend(title="Number of particles")) +
+  geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, linetype = type))+
+  xlab("Longtitude") + 
+  ylab("Latitude") + 
+  guides(linetype=guide_legend(title="Simulation")) +
   theme(plot.title = element_text(hjust = 0.5)) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
   theme_bw()
@@
 Simulations of hydrodynamic drag are presented in Panels A, C, E, G for January, April, July, and October, respectively.
-Results of simulations including both hydrodynamic drag and wind are in Panels B, D, F, H for January, April, July, and October, respectively.
-In all panels the passive simulations are represented by black lines."}
+Results of simulations including both hydrodynamic drag and wind are in Panels B, D, F, H for January, April, July, and October, respectively."}
 
@@
 
-In April and across all simulations, particles flowed in a north-westward direction (see Figure 4).
+In April and across all simulations, particles flowed in a north-westward direction (Figure 4).
 The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation.
@@
 Only slight variations across cross-sectional areas and drag exposure scenarios are evident.
-Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in an eddy located westward of the release site compared to passive particles (see Figure 5).
+Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in an eddy located westward of the release site compared to passive particles (Figure 5).
 Variation in density along the mean trajectory and eddy for all kelp particles is evident.
@@
 Simulations run for October show no difference in density of particles along the mean trajectory when comparing passive and kelp particles.
-Only slight variations in density and trajectory can be seen with no entrainment (see Figure 6).
+Only slight variations in density and trajectory can be seen with no entrainment (Figure 6).
 Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.

### 2021-05-06 thesis_chapter_3.Rmd
chp3 graphs touchups

@@
 
+load("map/south_africa_coast_hi_res.Rdata")
+load("map/sa_bathy.RData")
+
 library(ggplot2)
@@
 Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles traveled less distance compared to kelp particles with both wind and hydrodynamic drag.
-Differences in mean trajectories between kelp and passive particles were also evident in the July (Austral winter) simulations (see Figure 3E-F).
+Differences in mean trajectories between kelp and passive particles were also evident in the July (Austral winter) simulations (Figure 3E-F).
 Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore, resulting in entrainment within an eddy.
@@
 
-load("map/south_africa_coast_hi_res.Rdata")
-
 total_mean_water <- rbind(p_RK4_mean, Min_Kfloat_H100_W00_mean, Mean_Kfloat_H100_W00_mean, Max_Kfloat_H100_W00_mean)
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_water, aes(x = lon, y = lat, linetype = type))+
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_ww, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_water_April, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill ="grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_water_July, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, linetype = type))+
@@
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme_bw()
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, linetype = type))+
@@
   guides(linetype=guide_legend(title="Simulation")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme_bw()
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 
-ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, ncol = 2, nrow = 4, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("bottom"), labels = "AUTO", align = "hv")
+ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, ncol = 2, nrow = 4, font.label = list(size = 12, face = "bold"), common.legend = FALSE, legend = c("bottom"), labels = c("A","B","C","D","E","F","G", "H"), align = "hv")
+
+ggsave(filename = "figure_3.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, dpi = 300, scale = 2.0)
 ```
@@
 load("map/south_africa_coast_hi_res.Rdata")
+load("map/sa_bathy.RData")
 
 heatmap_RK4 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H100_W00 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis",limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("2") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H095_W05 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("3") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H090_W10 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("4") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H085_W15 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("5") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_Kfloat_H100_W00 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("6") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H095_W05 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("7") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H090_W10 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("8") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H085_W15 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("9") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H100_W00 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("10") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H095_W05 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") +
+  ggtitle("11") + 
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H090_W10 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("12") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H085_W15 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("13") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 heatmap_RK4_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = p_RK4_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("14") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H100_W00_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("15") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H095_W05_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("16") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H090_W10_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("17") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H085_W15_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("18") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_Kfloat_H100_W00_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("19") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H095_W05_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("20") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H090_W10_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("21") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H085_W15_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("22") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H100_W00_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H100_W00_mean_April ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("23") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H095_W05_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
+  ggtitle("24") +
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H090_W10_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("25") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H085_W15_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("26") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 heatmap_RK4_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = p_RK4_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("27") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H100_W00_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("28") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H095_W05_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("29") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H090_W10_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("30") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H085_W15_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("31") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_Kfloat_H100_W00_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("32") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H095_W05_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("33") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H090_W10_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("34") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H085_W15_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("35") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H100_W00_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H100_W00_mean_July ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("36") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H095_W05_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
+  ggtitle("37") +
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H090_W10_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("38") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H085_W15_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("39") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 heatmap_RK4_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = p_RK4_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("40") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H100_W00_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("41") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H095_W05_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("42") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H090_W10_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("43") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H085_W15_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("44") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_Kfloat_H100_W00_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("45") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H095_W05_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("46") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H090_W100 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("47") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H085_W15_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("48") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H100_W00_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("49") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H095_W05_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
+  ggtitle("50") +
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H090_W10_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("51") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H085_W15_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  ggtitle("52") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 
-ggarrange(heatmap_RK4, heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15,as_ggplot(leg_Jan), labels = c("1","2","3","4","5","6","7","8","9","10","11","12","13"), common.legend = TRUE, legend = c("none"),font.label = list(size = 11, face = "bold"))
-
+ggarrange(heatmap_RK4, heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15,as_ggplot(leg_Jan), labels = c("A","B","C","D","E","F","G","H","I","J","K","L","M"), common.legend = TRUE, legend = c("none"),font.label = list(size = 12, face = "bold"))
+
+ggsave(filename = "figure_4.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 ```
@@
 
-ggarrange(heatmap_RK4_April,heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April,as_ggplot(leg_April), labels = c("14", "15","16","17","18","19","20","21","22","23","24","25", "26"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
-
+ggarrange(heatmap_RK4_April,heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April,as_ggplot(leg_April), labels = c("A","B","C","D","E","F","G","H","I","J","K","L", "M"), common.legend = TRUE, legend = c("none"), font.label = list(size = 12, face = "bold"))
+
+ggsave(filename = "figure_5.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 ```
@@
 
-ggarrange(heatmap_RK4_July,heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, as_ggplot(leg_July),labels = c("27","28","29","30","31","32","33","34","35","36","37","38","39"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
-
+ggarrange(heatmap_RK4_July,heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, as_ggplot(leg_July),labels = c("A","B","C","D","E","F","G","H","I","J","K","L","M"), common.legend = TRUE, legend = c("none"), font.label = list(size = 12, face = "bold"))
+
+ggsave(filename = "figure_6.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 ```
@@
 
-ggarrange(heatmap_RK4_Oct, heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Mean_KFloat_H090_W100, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, as_ggplot(leg_Oct),labels = c("40","41","42","43","44","45","46","47","48","49","50","51","52"), common.legend = TRUE, legend = c("none"), font.label = list(size = 11, face = "bold"))
-
+ggarrange(heatmap_RK4_Oct, heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Mean_KFloat_H090_W100, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, as_ggplot(leg_Oct),labels = c("A","B","C","D","E","F","G","H","I","J","K","L","M"), common.legend = TRUE, legend = c("none"), font.label = list(size = 12, face = "bold"))
+
+ggsave(filename = "figure_7.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 ```
@@
 heatmap_RK4_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_kelp_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_RK4_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = pass_acc, aes(x = lon, y = lat, colour = distance)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   xlab("Longtitude") +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_kelp_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
@@
   guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 heatmap_RK4_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = pass_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   geom_path(data = pass_acc_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_kelp_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = kelp_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   geom_path(data = kelp_acc_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_RK4_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = pass_acc_April, aes(x = lon, y = lat, colour = distance))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   xlab("Longtitude") +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_kelp_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = kelp_acc_April, aes(x = lon, y = lat, colour = age))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
@@
 heatmap_RK4_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = pass_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   geom_path(data = pass_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_kelp_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = kelp_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   geom_path(data = kelp_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_RK4_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = pass_acc_July, aes(x = lon, y = lat, colour = distance))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   xlab("Longtitude") +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_kelp_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = kelp_acc_July, aes(x = lon, y = lat, colour = age))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
@@
 heatmap_RK4_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = pass_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   geom_path(data = pass_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_kelp_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = kelp_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   geom_path(data = kelp_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_RK4_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = pass_acc_Oct, aes(x = lon, y = lat, colour = distance))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   xlab("Longtitude") +
@@
   guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_kelp_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = kelp_acc_Oct, aes(x = lon, y = lat, colour = age)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = "grey") +
   #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
+  theme_bw()+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 
-ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 11, face = "bold"), align = "hv")
+ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 12, face = "bold"), align = "hv")
 
@@
 
-ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 11, face = "bold"), align = "hv")
+ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 12, face = "bold"), align = "hv")
 
@@
 
-The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (see appendix).
-This is also seen when comparing distance traveled by particles and particle density plots (see appendix).
-The simulation employing a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (see appendix).
+The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (appendix).
+This is also seen when comparing distance traveled by particles and particle density plots (appendix).
+The simulation employing a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (appendix).
 

### 2021-05-06 thesis_chapter_3.Rmd
little touchups

@@
 
-There were significant differences among the median distances traveled for all particles in each simulation (Figure 4; *p* < 0.05).
+There were significant differences among the median distances traveled for all particles in each simulation (Figure 8; *p* < 0.05).
 The simulations that considered for 100% hydrodynamic drag traveled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
@@
 When comparing distances between passive and kelp particles for April, significant differences are evident.
-(Figure 11, simulations 17-29; *p* < 0.05).
+(Figure 8, simulations 17-29; *p* < 0.05).
 Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas traveled shorter distance across compared to other kelp particles.
 The shortest distance traveled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
-No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; *p* > 0.05).
-In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; *p* < 0.05).
+No significant differences are evident for the July simulations, with only slight variation across simulations (Figure 8, simulations 30-42; *p* > 0.05).
+In October, the plot shows significant differences when comparing both passive and kelp particles (Figure 8, simulations 43-55; *p* < 0.05).
 Kelp particles with only 90% hydrodynamic drag across cross-sectional areas traveled shorter distance compared to all other particles.
@@
 
-```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE}
+```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
 
@@
 
-ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "auto", ncol = 1, nrow = 4)
-
+ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 4)
+ggsave(filename = "figure_8.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 ```
@@
   ylab("Latitude") +
-  ggtitle("Passive particles") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") +
-  ggtitle("Kelp particles") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") +
-  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   ylab("Latitude") +
-  ggtitle("") +
   guides(fill=guide_legend(title="Age")) +
@@
   ylab("Latitude") +
-  ggtitle("") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
         axis.title=element_text(size=12)) 
-
-# Distance
-
-#dis_plot_acc <- ggplot(data = total_acc_data, aes(x = type, y = distance, show.legend = FALSE)) +
-#  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
-#  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = "right") +
-#  theme_bw() +
-#  ylab("Distance (km)") +
-#  xlab("Scenario") +
-  #coord_flip() +
-#  theme_bw()
-  #theme(plot.title = element_text(hjust = 1))
-
 ```
 
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation for scenarios 53-60.
+```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of density between passive and kelp accumulation for scenarios 53-60.
 For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=9, fig.width=7}
@@
 ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 12, face = "bold"), align = "hv")
-
+ggsave(filename = "figure_9.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 ```
 
-```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of particle age between passive and kelp accumulation scenarios 53-60.
+```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 10: Comparison of particle age between passive and kelp accumulation scenarios 53-60.
 For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=9, fig.width=7}
@@
 ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 12, face = "bold"), align = "hv")
-
+ggsave(filename = "figure_10.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 ```
@@
 The role of eddys that exist within the study domain over the time period play a clear role in the accumulation and dispersal of kelp particles.
-Passive particles tend to remain on the outer edge of the eddy and then are expelled on the other side and continue to follow the ocean current (see Figure 8).
+Passive particles tend to remain on the outer edge of the eddy and then are expelled on the other side and continue to follow the ocean current (Figure 9).
 The opposite is true for kelp particles which have the highest exposure level of hydrodynamic and wind drag.
-The kelp particles are entrained in the eddy (see Figure 8).
+The kelp particles are entrained in the eddy (Figure 9).
 This is also reflected when comparing the age of particles at the end of the simulation.
-In Figure 9, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
+In Figure 10, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
 

### 2021-05-10 thesis_chapter_3.Rmd
new corrections and ref error sorted

--- 2021-05-05 thesis_chapter_3.Rmd
@@
 author: 
-- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, 41 Market Place Viaduct Harbour, 1010 Auckland, Private Bag 99940, Newmarket, Auckland 1149, New Zealand], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
+- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
 output:
@@
 These winds are most intense during Austral summer (December-March) which favors upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
-In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela Jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
+In Austral winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela Jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
 
@@
 All particles that have any form of drag exposure are hereafter referred to as 'kelp particles'.
-All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018].
+All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @vanSebille2018].
 Brownian motion was also incorporated to account for any stochasticity not resolved by the model.
-These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018].
+These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019;@vanSebille2018].
 The time period used in all simulations were January (Austral summer), April (Austral autumn), July (Austral winter), and October (Austral spring) of 2018.
@@
 ## Model inputs
-### Ocean and wind model
+## Ocean and wind model
 
@@
 
-### Hydrodynamic and wind drag
+## Hydrodynamic and wind drag
 
@@
 waterdrag_plot <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_water, aes(x = lon, y = lat, linetype = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(linetype=guide_legend(title="Simulation")) +
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 wwdrag_plot <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_ww, aes(x = lon, y = lat, linetype = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(linetype=guide_legend(title="Simulation")) +
@@
 waterdrag_plot_April <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_water_April, aes(x = lon, y = lat, linetype = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(linetype=guide_legend(title="Simulation")) +
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 wwdrag_plot_April <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill ="grey") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill ="grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, linetype = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(linetype=guide_legend(title="Simulation")) +
@@
 waterdrag_plot_July <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_water_July, aes(x = lon, y = lat, linetype = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(linetype=guide_legend(title="Simulation")) +
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 wwdrag_plot_July <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",     fill = "grey") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, linetype = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(linetype=guide_legend(title="Simulation")) +
@@
 waterdrag_plot_Oct <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, linetype = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(linetype=guide_legend(title="Simulation")) +
@@
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme_bw()
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 wwdrag_plot_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, linetype = type))+
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(linetype=guide_legend(title="Simulation")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme_bw()
-
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
 ```
@@
 
-Across all simulations, particles flowed in a north-westward direction.
-The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids.
+Across all simulations for January (Austral summer), particles flowed in a north-westward direction.
+The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids (Figure 4).
 In addition, the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 4).
@@
 
-In April and across all simulations, particles flowed in a north-westward direction (Figure 4).
+In April and across all simulations, particles flowed in a north-westward direction (Figure 5).
 The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation.
@@
 Only slight variations across cross-sectional areas and drag exposure scenarios are evident.
-Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in an eddy located westward of the release site compared to passive particles (Figure 5).
+Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in an eddy located westward of the release site compared to passive particles (Figure 6).
 Variation in density along the mean trajectory and eddy for all kelp particles is evident.
@@
 Simulations run for October show no difference in density of particles along the mean trajectory when comparing passive and kelp particles.
-Only slight variations in density and trajectory can be seen with no entrainment (Figure 6).
+Only slight variations in density and trajectory can be seen with no entrainment (Figure 7).
 Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.
@@
 load("map/south_africa_coast_hi_res.Rdata")
+load("map/sa_bathy.RData")
 
 heatmap_RK4 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H100_W00 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis",limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("2") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H095_W05 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("3") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H090_W10 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("4") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H085_W15 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("5") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_Kfloat_H100_W00 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("6") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H095_W05 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("7") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H090_W10 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("8") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H085_W15 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("9") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H100_W00 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("10") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H095_W05 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("11") + 
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H090_W10 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("12") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H085_W15 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("13") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 heatmap_RK4_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = p_RK4_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("14") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H100_W00_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("15") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H095_W05_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("16") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H090_W10_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("17") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H085_W15_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("18") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_Kfloat_H100_W00_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("19") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H095_W05_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("20") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H090_W10_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("21") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H085_W15_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("22") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H100_W00_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H100_W00_mean_April ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("23") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H095_W05_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("24") +
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H090_W10_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("25") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H085_W15_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("26") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 heatmap_RK4_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = p_RK4_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("27") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H100_W00_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("28") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H095_W05_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("29") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H090_W10_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) + 
+  ggtitle("30") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H085_W15_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("31") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_Kfloat_H100_W00_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("32") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H095_W05_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("33") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H090_W10_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("34") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H085_W15_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("35") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H100_W00_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H100_W00_mean_July ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("36") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H095_W05_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("37") +
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H090_W10_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) + 
+  ggtitle("38") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H085_W15_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) + 
+  ggtitle("39") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 heatmap_RK4_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = p_RK4_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("40") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H100_W00_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H100W100") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) + 
+  ggtitle("41") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H095_W05_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  #ggtitle("H095W05") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("42") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H090_W10_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H090W10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) + 
+  ggtitle("43") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Min_Kfloat_H085_W15_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Min_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  #ggtitle("H085W15") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) + 
+  ggtitle("44") + 
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_Kfloat_H100_W00_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("45") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H095_W05_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) + 
+  ggtitle("46") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H090_W100 <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("47") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Mean_KFloat_H085_W15_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Mean_KFloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("48") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H100_W00_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("49") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H095_W05_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) + 
+  ggtitle("50") +
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H090_W10_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("51") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_Max_KFloat_H085_W15_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
   scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
   geom_path(data = Max_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("52") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 
-There were significant differences among the median distances traveled for all particles in each simulation (Figure 4; *p* < 0.05).
-The simulations that considered for 100% hydrodynamic drag traveled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios.
+There were significant differences among the median distances traveled for all particles in each simulation (Figure 8; *p* < 0.05).
+The simulations that considered for 100% hydrodynamic drag traveled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios (Figure 8; *p* < 0.05).
 Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations.
 When comparing distances between passive and kelp particles for April, significant differences are evident.
-(Figure 11, simulations 17-29; *p* < 0.05).
+(Figure 8, simulations 14-26; *p* < 0.05).
 Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas traveled shorter distance across compared to other kelp particles.
 The shortest distance traveled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
-No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; *p* > 0.05).
-In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; *p* < 0.05).
+No significant differences are evident for the July simulations, with only slight variation across simulations (Figure 8, simulations 27-39; *p* > 0.05).
+In October, the plot shows significant differences when comparing both passive and kelp particles (Figure 8, simulations 40-52; *p* < 0.05).
 Kelp particles with only 90% hydrodynamic drag across cross-sectional areas traveled shorter distance compared to all other particles.
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=7, fig.width=5}
-
-ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "auto", ncol = 1, nrow = 4)
-
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=6, fig.width=6}
+
+ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 4)
+ggsave(filename = "figure_8.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 ```
@@
 
-```{r message=FALSE, echo=FALSE, warning=FALSE}
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=FALSE}
 
@@
 heatmap_RK4_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Passive particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Passive") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
   theme_bw() +
   theme(plot.title = element_text(hjust = 0.5))
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_kelp_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Kelp particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Kelp") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
   theme_bw() +
   theme(plot.title = element_text(hjust = 0.5))
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_RK4_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = pass_acc, aes(x = lon, y = lat, colour = distance)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_point(data = pass_acc, aes(x = lon, y = lat, colour = age)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Age")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_kelp_acc <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 heatmap_RK4_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = pass_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = pass_acc_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Passive particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Number of particles")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_kelp_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = kelp_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = kelp_acc_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Kelp particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Number of particles")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_RK4_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = pass_acc_April, aes(x = lon, y = lat, colour = distance)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_point(data = pass_acc_April, aes(x = lon, y = lat, colour = age))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Age")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_kelp_acc_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = kelp_acc_April, aes(x = lon, y = lat, colour = age))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 heatmap_RK4_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = pass_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = pass_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Passive particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Number of particles")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_kelp_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = kelp_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = kelp_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Kelp particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Number of particles")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_RK4_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = pass_acc_July, aes(x = lon, y = lat, colour = distance)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_point(data = pass_acc_July, aes(x = lon, y = lat, colour = age))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Age")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_kelp_acc_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = kelp_acc_July, aes(x = lon, y = lat, colour = age))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
 heatmap_RK4_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = pass_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = pass_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Passive particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Number of particles")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 heatmap_kelp_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = kelp_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = kelp_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("Kelp particles") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Number of particles")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_RK4_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_point(data = pass_acc_Oct, aes(x = lon, y = lat, colour = distance)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_point(data = pass_acc_Oct, aes(x = lon, y = lat, colour = age))  +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(fill=guide_legend(title="Age")) +
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
 point_kelp_acc_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_point(data = kelp_acc_Oct, aes(x = lon, y = lat, colour = age)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -26)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
 
@@
   geom_path(data = total_acc_mean, aes(x = lon, y = lat, colour = type)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-# Distance
-
-#dis_plot_acc <- ggplot(data = total_acc_data, aes(x = type, y = distance, show.legend = FALSE)) +
-#  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
-#  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = "right") +
-#  theme_bw() +
-#  ylab("Distance (km)") +
-#  xlab("Scenario") +
-  #coord_flip() +
-#  theme_bw()
-  #theme(plot.title = element_text(hjust = 1))
-
+  theme_bw()+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
 ```
 
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation for scenarios 53-60.
-For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=9, fig.width=7}
-
-ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 11, face = "bold"), align = "hv")
-
+```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of density between passive and kelp accumulation for scenarios 53-60.
+For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=8, fig.width=8}
+
+ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January","", "April","","July","","October"), font.label = list(size = 12, face = "bold"), align = "hv")
+
+ggsave(filename = "figure_9.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 ```
 
-```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of particle age between passive and kelp accumulation scenarios 53-60.
-For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=9, fig.width=7}
-
-ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 11, face = "bold"), align = "hv")
-
+```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 10: Comparison of particle age between passive and kelp accumulation scenarios 53-60.
+For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=10, fig.width=10}
+
+ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 12, face = "bold"), align = "hv", heights = c(0.5,0.5, 0.5,0.5, 0.5,0.5, 0.5,0.5))
+ggsave(filename = "figure_10.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 ```
@@
 The different trajectory patterns between seasons reflects the underlying seasonal variation in regional oceanography.
-Transport in the study region is driven by a combination d variation of alongshore jets and wind induced Ekman transport of surface water [@rubio2009; @veitch2018; @ragoasha2019].
-A study by @ragoasha2019 investigated the success of Lagrangian particles reaching St.
+Transport in the study region is driven by a combination of variation of alongshore jets and wind induced Ekman transport of surface water [@rubio2009; @veitch2018; @ragoasha2019].
+@ragoasha2019 investigated the success of Lagrangian particles reaching St.
 Helena Bay from Cape point using Lagrangian trajectory modelling and inclusion of regional oceanographic processes.
@@
   geom_path(data = cyl_mean, aes(x = lon, y = lat), color = 'red')+
-  xlab("Longtitude") + 
-  ylab("Latitude") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   ggtitle("January") +
@@
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   ggtitle("") +
@@
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   ggtitle("April") +
@@
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   ggtitle("July") +
@@
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
   ggtitle("October") +
@@
 
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 10: Comparison of trajectory data between sphere and cylinder shape coefficients for scenarios 61-64.The mean trajectory comparison between passive and kelp.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=8}
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 11: Comparison of trajectory data between sphere and cylinder shape coefficients for scenarios 61-64.The mean trajectory comparison between passive and kelp.
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=6}
 
@@
 
-ggarrange(shape_plot_A, shape_dis_plot, ncol = 1, nrow = 2, legend = c("right"), labels = c("", "D") , font.label = list(size = 11, face = "bold"))
+ggarrange(shape_plot_A, shape_dis_plot, ncol = 1, nrow = 2, legend = c("right"), labels = c("", "D") , font.label = list(size = 11, face = "bold"), heights = c(2, 1))
+
+ggsave(filename = "figure_11.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 
@@
 The velocity vectors were averaged over the 30 day simulation times for respective seasons (summer, autumn, winter, spring).
-A represents summer, B autumn, C winter and D spring", fig.height=10, fig.width=8}
+A represents summer, B autumn, C winter and D spring", fig.height=8, fig.width=8}
 

### 2021-05-10 thesis_chapter_3.Rmd
touchups

@@
 ## Model inputs
-## Ocean and wind model
+### Ocean and wind model
 
@@
 
-## Hydrodynamic and wind drag
+### Hydrodynamic and wind drag
 
@@
 ```
+
 ## Analysis
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
+  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000)) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12),
+        plot.title = element_text(hjust = 0.5),
+        element_blank())
 
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720)) +
+  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000)) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
+        axis.title=element_text(size=12),
+        element_blank())
 
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_hex(data = pass_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_hex(data = kelp_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_hex(data = pass_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_hex(data = kelp_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_hex(data = pass_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 720))  +
+  geom_hex(data = kelp_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
 
-ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January","", "April","","July","","October"), font.label = list(size = 12, face = "bold"), align = "hv")
-
-ggsave(filename = "figure_9.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
+ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January","", "April","","July","","October"), font.label = list(size = 12, face = "bold"), align = "hv", label.x = 0.2)
+
+ggsave(filename = "figure_9.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 2, dpi = 300)
 ```
@@
 ```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 10: Comparison of particle age between passive and kelp accumulation scenarios 53-60.
-For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=10, fig.width=10}
-
-ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 12, face = "bold"), align = "hv", heights = c(0.5,0.5, 0.5,0.5, 0.5,0.5, 0.5,0.5))
-ggsave(filename = "figure_10.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
+Age is represented in hours since release.
+For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=8, fig.width=8}
+
+ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 12, face = "bold"), align = "hv", label.x = 0.2)
+ggsave(filename = "figure_10.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 2, dpi = 300)
 ```
@@
 sphere$Month <- "January" 
-sphere_mean$type <- "Sphere"
+sphere_mean$Type <- "Sphere"
 
@@
 cyl$Month <- "January" 
-cyl_mean$type <- "Cylinder"
+cyl_mean$Type <- "Cylinder"
+
+shape_data_jan <- rbind(sphere_mean, cyl_mean)
 
@@
 shape_plot <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  coord_fixed(ratio=1, xlim = c(14, 20), ylim = c(-35, -30)) + 
   #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = sphere_mean, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = cyl_mean, aes(x = lon, y = lat), color = 'red')+
+  geom_path(data = shape_data_jan, aes(x = lon, y = lat, linetype = Type)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   labs(x = expression(Longitude*degree*S)) +
@@
   ggtitle("January") +
-  #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5)) 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5),
+        element_blank()) 
   #coord_fixed(ratio=1/1) # fix the aspect ratio of plot
@@
 heatmap_sphere <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = sphere_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5), 
+        element_blank())
 
 heatmap_cyl <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = cyl_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5), 
+        element_blank())
 
@@
 sphere_April$Month <- "April" 
-sphere_mean_April$type <- "Sphere"
+sphere_mean_April$Type <- "Sphere"
 
@@
 cyl_April$Month <- "April" 
-cyl_mean_April$type <- "Cylinder"
+cyl_mean_April$Type <- "Cylinder"
+
+shape_data_apr <- rbind(sphere_mean_April, cyl_mean_April)
 
@@
 shape_plot_April <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
-  #geom_path(data = p_RK4_mean_April, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = sphere_mean_April, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = cyl_mean_April, aes(x = lon, y = lat), color = 'red')+
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  coord_fixed(ratio=1, xlim = c(14, 20), ylim = c(-35, -30)) + 
+  #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = shape_data_apr, aes(x = lon, y = lat, linetype = Type)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   labs(x = expression(Longitude*degree*S)) +
@@
   ggtitle("April") +
-  #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5)) 
-  #coord_fixed(ratio=1/1)  # fix the aspect ratio of plot
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5), 
+        element_blank())
 
@@
 heatmap_sphere_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = sphere_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5),
+        element_blank())
 
 heatmap_cyl_April <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = cyl_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5),
+        element_blank())
 
@@
 sphere_July$Month <- "July" 
-sphere_mean_July$type <- "Sphere"
+sphere_mean_July$Type <- "Sphere"
 
@@
 cyl_July$Month <- "July" # add extra column 'type' to dataframe and assigned 'kelp' label
-cyl_mean_July$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+cyl_mean_July$Type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+shape_data_july <- rbind(sphere_mean_July, cyl_mean_July)
 
@@
 shape_plot_July <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
-  #geom_path(data = p_RK4_mean_July, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = sphere_mean_July, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = cyl_mean_July, aes(x = lon, y = lat), color = 'red')+
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  coord_fixed(ratio=1, xlim = c(14, 20), ylim = c(-35, -30)) + 
+  #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
+  geom_path(data = shape_data_apr, aes(x = lon, y = lat, linetype = Type)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   labs(x = expression(Longitude*degree*S)) +
@@
   ggtitle("July") +
-  #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5)) 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5),
+        element_blank()) 
   #coord_fixed(ratio=1/1) # fix the aspect ratio of plot
@@
 heatmap_sphere_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = sphere_July, bins = 80, aes(x = lon, y = lat)) +
+    geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5),
+        element_blank())
 
 heatmap_cyl_July <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = cyl_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
 sphere_Oct$Month <- "October" 
-sphere_mean_Oct$type <- "Sphere"
+sphere_mean_Oct$Type <- "Sphere"
 
@@
 cyl_Oct$Month <- "October"
-cyl_mean_Oct$type <- "Cylinder"
+cyl_mean_Oct$Type <- "Cylinder"
+
+shape_data_oct <- rbind(sphere_mean_Oct, cyl_mean_Oct)
 
@@
 shape_plot_Oct <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0,     colour = "black",     fill = "grey") +
-  #geom_path(data = p_RK4_mean_Oct, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = sphere_mean_Oct, aes(x = lon, y = lat), color = 'blue')+
-  geom_path(data = cyl_mean_Oct, aes(x = lon, y = lat), color = 'red')+
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  coord_fixed(ratio=1, xlim = c(14, 20), ylim = c(-35, -30)) + 
+  #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
+  geom_path(data = shape_data_oct, aes(x = lon, y = lat, linetype = Type)) +
   labs(x = expression(Longitude*degree*S)) +
@@
   ggtitle("October") +
-  #guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5)) 
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5),
+        element_blank()) 
   #coord_fixed(ratio=1/1) # fix the aspect ratio of plot
@@
 heatmap_sphere_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
@@
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   geom_path(data = sphere_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
   theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5),
+        element_blank())
 
 heatmap_cyl_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
   geom_hex(data = cyl_Oct, bins = 80, aes(x = lon, y = lat)) +
+    geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "grey") +
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
@@
   scale_x_discrete(limits = shape_pos)+
-  stat_compare_means(method = "kruskal.test", label = "p.signif")+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", aes(group=Month), hide.ns = FALSE)+
   ylab("Distance (km)")+
@@
 
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 11: Comparison of trajectory data between sphere and cylinder shape coefficients for scenarios 61-64.The mean trajectory comparison between passive and kelp.
+```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 11: Comparison of trajectory data between sphere and cylinder shape coefficients for scenarios 61-64.
+Row A represents the mean trajectory comparison for each of the respective months (seasons) simulated.
 The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=6}
 
-shape_plot_A <- ggarrange(shape_plot, shape_plot_April, shape_plot_July, shape_plot_Oct, heatmap_sphere, heatmap_sphere_April,heatmap_sphere_July, heatmap_sphere_Oct, heatmap_cyl, heatmap_cyl_April, heatmap_cyl_July, heatmap_cyl_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 3, align = "hv", labels = c("A", "", "", "", "B", "", "", "", "C"), font.label = list(size = 11, face = "bold"))
-
-ggarrange(shape_plot_A, shape_dis_plot, ncol = 1, nrow = 2, legend = c("right"), labels = c("", "D") , font.label = list(size = 11, face = "bold"), heights = c(2, 1))
-
-ggsave(filename = "figure_11.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
+shape_plot_A <- ggarrange(shape_plot, shape_plot_April, shape_plot_July, shape_plot_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 1, labels = c("A", "", "", ""), font.label = list(size = 12, face = "bold"), align = "hv")
+
+shape_plot_B <- ggarrange(heatmap_sphere, heatmap_sphere_April,heatmap_sphere_July, heatmap_sphere_Oct, heatmap_cyl, heatmap_cyl_April, heatmap_cyl_July, heatmap_cyl_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 2, align = "hv", labels = c("B", "", "", "", "C", "", "", ""), font.label = list(size = 12, face = "bold"))
+
+ggarrange(shape_plot_A, shape_plot_B, shape_dis_plot, ncol = 1, nrow = 3, legend = c("right"), labels = c("", "D") , font.label = list(size = 11, face = "bold"), heights = c(2, 2, 1))
+
+#ggsave(filename = "figure_11.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
 

### 2021-05-11 thesis_chapter_3.Rmd
file rename. updates to chp3. vector map updated

@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=12, fig.width=7, fig.cap="Figure 3: Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=8, fig.cap="Figure 3: Comparison of mean trajectories.
 Simulations of hydrodynamic drag are presented in Panels A, C, E, G for January, April, July, and October, respectively.
@@
 ```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 11: Comparison of trajectory data between sphere and cylinder shape coefficients for scenarios 61-64.
-Row A represents the mean trajectory comparison for each of the respective months (seasons) simulated.
+Row A represents the mean trajectory comparison, row B-C are density comparison plots, and row D is comparison of total distance travelled from the release site for each of the respective months (seasons) simulated.
 The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=6}

### 2021-05-11 thesis_chapter3_vector_map.R

@@ -55,65 +55,69 @@ load("map/south_africa_coast_hi_res.Rdata")
 
 ocean_plot_jan <-  ggplot() +
   geom_tile(data = ocean_jan_summary, aes(fill = velocity, x = longitude, y = latitude)) +
-  geom_segment(data = ocean_jan_summary[seq(5, nrow(ocean_jan_summary),2), ],
-               aes(x = longitude, xend = longitude+uo/30, y = latitude,
-                   yend = latitude+vo/30), arrow = arrow(length = unit(0.30, "cm"))) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",fill = "grey") +
+  #geom_segment(data = ocean_jan_summary[seq(5, nrow(ocean_jan_summary),5), ],
+               #aes(x = longitude, xend = longitude+uo/30, y = latitude,
+                  # yend = latitude+vo/30), arrow = arrow(length = unit(0.25, "cm"))) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",fill = "grey") +
   scale_fill_continuous(type = "viridis", limits = c(0, 0.8)) +
-  coord_cartesian(xlim = c(15, 20), ylim = c(-35, -30)) +
+  coord_fixed(ratio=1, xlim = c(15, 20), ylim = c(-35, -30)) +
   guides(fill=guide_legend(title="Velocity")) +
   theme_bw() +
-  theme(axis.text = element_text(size = 11, colour = 1)) +
-  labs(x = NULL, y = NULL)
+  theme(axis.text = element_text(size = 11, colour = 1),
+        element_blank()) +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E))
 
 # April ocean vector map
 
 ocean_plot_apr <-  ggplot() +
   geom_tile(data = ocean_apr_summary, aes(fill = velocity, x = longitude, y = latitude)) +
-  geom_segment(data = ocean_apr_summary[seq(5, nrow(ocean_jan_summary), 2), ],
-               aes(x = longitude, xend = longitude+uo/30, y = latitude,
-                   yend = latitude+vo/30), arrow = arrow(length = unit(0.30, "cm"))) +
[…]

### 2021-05-11 thesis_chapter3_vector_map_animations.R

@@ -1,0 +1,258 @@
+library(gganimate)
+library(plotly)
+
+ocean_vector <- tidync("OceanParcels/global-analysis-forecast-phy-001-024_1603882856077.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, longitude, latitude, uo, vo) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2018-01-01 00:00:00"),
+                     "GMT",
+                     "%Y-%m-%d"))
+
+ocean_jan <-  ocean_vector %>%
+  filter(t >= ("2086-01-01 00:00:00") & t <= ("2086-01-30 00:00:00"))
+ocean_apr <-  ocean_vector %>%
+  filter(t >= ("2086-04-01 00:00:00") & t <= ("2086-04-30 00:00:00"))
+ocean_july <-  ocean_vector %>%
+  filter(t >= ("2086-07-01 00:00:00") & t <= ("2086-07-30 00:00:00"))
+ocean_oct <-  ocean_vector %>%
+  filter(t >= ("2086-10-01 00:00:00") & t <= ("2086-10-30 00:00:00"))
+
+ocean_jan$velocity <- with(ocean_jan, sqrt(uo^2 + vo^2))
+ocean_apr$velocity <- with(ocean_apr, sqrt(uo^2 + vo^2))
+ocean_july$velocity <- with(ocean_july, sqrt(uo^2 + vo^2))
+ocean_oct$velocity <- with(ocean_oct, sqrt(uo^2 + vo^2))
+
+ocean_jan_summary_ani <- ocean_jan %>%
+  group_by(t, longitude, latitude) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
+ocean_apr_summary_ani <- ocean_apr %>%
+  group_by(t, longitude, latitude) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
+ocean_july_summary_ani <- ocean_july %>%
+  group_by(t, longitude, latitude) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
[…]

### 2021-05-11 ocean_map.R
deleted files

@@ -1,35 +1,0 @@
-library(tidyverse)
-library(oce)
-library(lubridate)
-library(gganimate)
-library(ggplot2)
-
-# Read in simulation data and data wrangling
-
-ocean_vector <- tidync("OceanParcels/global-analysis-forecast-phy-001-024_1603882856077.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(longitude, latitude, time, uo, vo) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
- #filter(between(t, as.POSIXct("2018-01-01"), as.POSIXct("2018-01-05")))
- #mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))  %>% # calculate mean trajectory
-
-ocean_vector$velocity <- with(ocean_vector, sqrt(uo^2 + vo^2))
-
-ocean_summary = ocean_vector %>%
-  group_by(longitude, latitude, t) %>%
-  summarise(uo = median(uo, na.rm = TRUE),
-            vo = median(vo, na.rm = TRUE),
-            velocity = median(velocity, na.rm = TRUE))
-
-load("map/south_africa_coast_hi_res.Rdata")
-
-test_plot = ggplot() +
-  geom_segment(data = ocean_summary,
-               aes(x = longitude, xend = longitude+uo/60, y = latitude,
-               yend = latitude+vo/60), arrow = arrow(length = unit(0.25, "cm"))) +
-  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black",fill = "grey") +
-  #coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
-  theme_bw() +
-  theme(axis.text = element_text(size = 11, colour = 1)) +
-  labs(x = NULL, y = NULL)

### 2021-05-11 vector_map_animations.R

@@ -1,258 +1,0 @@
-library(gganimate)
-library(plotly)
-
-ocean_vector <- tidync("OceanParcels/global-analysis-forecast-phy-001-024_1603882856077.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, longitude, latitude, uo, vo) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2018-01-01 00:00:00"),
-                     "GMT",
-                     "%Y-%m-%d"))
-
-ocean_jan <-  ocean_vector %>%
-  filter(t >= ("2086-01-01 00:00:00") & t <= ("2086-01-30 00:00:00"))
-ocean_apr <-  ocean_vector %>%
-  filter(t >= ("2086-04-01 00:00:00") & t <= ("2086-04-30 00:00:00"))
-ocean_july <-  ocean_vector %>%
-  filter(t >= ("2086-07-01 00:00:00") & t <= ("2086-07-30 00:00:00"))
-ocean_oct <-  ocean_vector %>%
-  filter(t >= ("2086-10-01 00:00:00") & t <= ("2086-10-30 00:00:00"))
-
-ocean_jan$velocity <- with(ocean_jan, sqrt(uo^2 + vo^2))
-ocean_apr$velocity <- with(ocean_apr, sqrt(uo^2 + vo^2))
-ocean_july$velocity <- with(ocean_july, sqrt(uo^2 + vo^2))
-ocean_oct$velocity <- with(ocean_oct, sqrt(uo^2 + vo^2))
-
-ocean_jan_summary_ani <- ocean_jan %>%
-  group_by(t, longitude, latitude) %>%
-  summarise(uo = median(uo, na.rm = TRUE),
-            vo = median(vo, na.rm = TRUE),
-            velocity = median(velocity, na.rm = TRUE))
-
-ocean_apr_summary_ani <- ocean_apr %>%
-  group_by(t, longitude, latitude) %>%
-  summarise(uo = median(uo, na.rm = TRUE),
-            vo = median(vo, na.rm = TRUE),
-            velocity = median(velocity, na.rm = TRUE))
-
-ocean_july_summary_ani <- ocean_july %>%
-  group_by(t, longitude, latitude) %>%
-  summarise(uo = median(uo, na.rm = TRUE),
-            vo = median(vo, na.rm = TRUE),
[…]

### 2021-05-11 map/map_chapter3.R
updated map files

@@ -5,12 +5,11 @@ source("map/scaleBarFunc.R")
 release_sites_chp3 <- read.csv("metadata/release_sites_chp3.csv", sep = ";")
 
 
-map_sa <- ggplot(data = south_africa_coast_hi_res) +
+map_A <- ggplot(data = south_africa_coast_hi_res) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = "gray") +
   geom_rect(data = south_africa_coast_hi_res,aes(xmin=14, xmax=20, ymin=-30, ymax=-35), inherit.aes = FALSE, fill = NA, colour =    "black", size = 1.5) +
-  coord_cartesian(xlim = c(8, 28), ylim = c(-38, -26)) +
-  #coord_map("mercator", clip = "on") +
-  #coord_cartesian(xlim = c(15, 20), ylim = c(-30, -35)) +
+  #coord_fixed(ratio=1, clip = "on", xlim = c(8, 28), ylim = c(-38, -27)) +
+  coord_fixed(ratio = 1, clip = "on", xlim = c(8, 28), ylim = c(-38, -27)) +
   #scale_x_continuous(trans = "reverse") +
   #theme(axis.ticks = element_blank()) + # Remove tick marks
   theme_bw() +
@@ -18,24 +17,27 @@ map_sa <- ggplot(data = south_africa_coast_hi_res) +
   ylab("Longitude")
 map_sa
 
-ggsave(filename = "map_sa.jpg", plot = last_plot(), device = "jpeg", path = "map/", dpi = 320, width = 10, height = 10)
+ggsave(filename = "map_A.jpg", plot = last_plot(), device = "jpeg", path = "map/", dpi = 320, width = 10, height = 10)
 
 ## Make 2nd map of just the peninsula
-map_pen <- ggplot() +
+map_B <- ggplot() +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 1.0, colour = "black", fill = "gray") +
-  #coord_map("mercator", clip = "on") +
-  scaleBar(lon = 15, lat = -34.9, distanceLon = 20, distanceLat = 20, distanceLegend = -5, dist.unit = "km",
+  coord_map(ratio = 1, clip = "on", xlim = c(15, 20), ylim = c(-35, -30)) +
[…]

### 2021-05-11 thesis_chapter_3.Rmd
updated.

@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_path(data = total_mean_water, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_path(data = total_mean_ww, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_path(data = total_mean_water_April, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill ="grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_path(data = total_mean_water_July, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, linetype = type))+
@@
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, linetype = type))+
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=8, fig.cap="Figure 3: Comparison of mean trajectories.
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=6, fig.cap="Figure 3: Comparison of mean trajectories.
 Simulations of hydrodynamic drag are presented in Panels A, C, E, G for January, April, July, and October, respectively.
@@
 
-ggsave(filename = "figure_3.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, dpi = 300, scale = 2.0)
+ggsave(filename = "figure_3.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 6, dpi = 300, scale = 1.2)
 ```
@@
 
-Across all simulations for January (Austral summer), particles flowed in a north-westward direction.
-The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids (Figure 4).
-In addition, the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 4).
+Density distributions summarizing the conclusion of sumlating Scenarios 1-13 (Table 1) are presented in Figures 3-6.
+Across all simulations for January (Austral summer), particles flow in a north-westward direction.
+The density plots for simulations that consider any form of drag show a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids (Figure 4).
+In addition, the density plots show that the particles in simulations that considered any form of drag become entrained in a vorticity while particles in the passive simulation deflected away (Figure 4).
 Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.
-
-In April and across all simulations, particles flowed in a north-westward direction (Figure 5).
-The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation.
-In addition, simulations that considered any form of drag got entrained within an eddy further north-westward from the end location of passive particles.
+In April (Figure 5) across all simulations, particles flow in a north-westward direction.
+The density plots that consider any form of drag show a narrow, higher density along the mean trajectory compared to passive simulation.
+In addition, simulations that consider any form of drag got entrained within an eddy further north-westward from the end location of passive particles.
 Only slight variations across cross-sectional areas and drag exposure scenarios are evident.
-Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in an eddy located westward of the release site compared to passive particles (Figure 6).
+
+Simulations for July (Figure 6) show that the flow is also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show a high density of particles entrained in an eddy located westward of the release site compared to passive particles.
 Variation in density along the mean trajectory and eddy for all kelp particles is evident.
 Kelp particles with only hydrodynamic drag across cross-sectional areas showed lower density of particles along the mean trajectory compared to kelp particles with both wind and hydrodynamic drag.
-Simulations run for October show no difference in density of particles along the mean trajectory when comparing passive and kelp particles.
-Only slight variations in density and trajectory can be seen with no entrainment (Figure 7).
+Simulations run for October (Figure 7) show no difference in density of particles along the mean trajectory when comparing passive and kelp particles.
+Only slight variations in density and trajectory can be seen with no entrainment.
 Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
@@
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
@@
   #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
   #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for scenarios 1 - 13, for January 2018 (summer).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for simulations 1 - 13, for January 2018 (summer).
 The plots are regular hexagons within the density of particles were calculated.
-The mean trajectories for each scenario are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}
+The mean trajectories for each simulation are also plotted for comparison.
+The reader is referred to tables 1 and 2 for details pertaining to each simulation', fig.height=8, fig.width=6}
 
@@
 
-ggsave(filename = "figure_4.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
+ggsave(filename = "figure_4.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 8, scale = 1, dpi = 300)
 ```
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end of the run for scenarios 14 - 26, for April 2018 (autumn).
-See Figure 4 for further details.', fig.height=8, fig.width=8}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end of the run for simulations 14 - 26, for April 2018 (autumn).
+See Figure 4 for further details.', fig.height=8, fig.width=6}
 
@@
 
-ggsave(filename = "figure_5.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
+ggsave(filename = "figure_5.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 8, scale = 1, dpi = 300)
 ```
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end of the run for scenarios 27 - 39, for July 2018 (winter).
-See Figure 4 for further details.', fig.height=8, fig.width=8}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end of the run for simulations 27 - 39, for July 2018 (winter).
+See Figure 4 for further details.', fig.height=8, fig.width=6}
 
@@
 
-ggsave(filename = "figure_6.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
+ggsave(filename = "figure_6.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 8, scale = 1, dpi = 300)
 ```
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 7: Comparison of density plots at the end of the run for scenarios 40 - 52, for October 2018 (spring).
-See Figure 4 for further details.', fig.height=8, fig.width=8}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 7: Comparison of density plots at the end of the run for simulations 40 - 52, for October 2018 (spring).
+See Figure 4 for further details.', fig.height=8, fig.width=6}
 
@@
 
-ggsave(filename = "figure_7.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
+ggsave(filename = "figure_7.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 8, scale = 1, dpi = 300)
 ```
@@
 
-There were significant differences among the median distances traveled for all particles in each simulation (Figure 8; *p* < 0.05).
-The simulations that considered for 100% hydrodynamic drag traveled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios (Figure 8; *p* < 0.05).
+There are significant differences among the median distances traveled for all particles in each simulation (Figure 8; *p* < 0.05) made for January conditions.
+Simulations that considered 100% hydrodynamic drag travel a shorter distance than particles for simulations with a combination of hydrodynamic and wind drag scenarios (Figure 8; *p* < 0.05).
 Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations.
-When comparing distances between passive and kelp particles for April, significant differences are evident.
+When comparing distances between passive and kelp particles during April conditions, significant differences are evident.
 (Figure 8, simulations 14-26; *p* < 0.05).
-Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas traveled shorter distance across compared to other kelp particles.
-The shortest distance traveled was the scenario with the maximum cross-sectional area and drag exposure scenarios.
-No significant differences are evident for the July simulations, with only slight variation across simulations (Figure 8, simulations 27-39; *p* > 0.05).
-In October, the plot shows significant differences when comparing both passive and kelp particles (Figure 8, simulations 40-52; *p* < 0.05).
-Kelp particles with only 90% hydrodynamic drag across cross-sectional areas traveled shorter distance compared to all other particles.
-As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure traveled the shortest distance compared to all other particles.
+Kelp particles with 90% hydrodynamic and 10% wind drag across cross-sectional areas travel a shorter distance across compared to other kelp particles.
+The shortest distance traveled was the simulation with the maximum cross-sectional area and drag exposure scenarios.
+No significant differences are evident for the July simulations, with only a slight variation across simulations (Figure 8, simulations 27-39; *p* > 0.05).
+The plot with results of the simulations made for October conditions shows significant differences when comparing both passive and kelp particles (Figure 8, simulations 40-52; *p* < 0.05).
+Kelp particles with only 90% hydrodynamic drag across cross-sectional areas travel a shorter distance compared to all other particles.
+As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure travel the shortest distance compared to all other particles.
 
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=6, fig.width=6}
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=6}
 
 ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 4)
-ggsave(filename = "figure_8.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
+ggsave(filename = "figure_8.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 7, scale = 1, dpi = 300)
 ```
@@
 heatmap_RK4_acc <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
@@
 heatmap_kelp_acc <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
@@
 point_RK4_acc <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_point(data = pass_acc, aes(x = lon, y = lat, colour = age)) +
@@
 point_kelp_acc <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
@@
 heatmap_RK4_acc_April <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
@@
 heatmap_kelp_acc_April <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
@@
 point_RK4_acc_April <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_point(data = pass_acc_April, aes(x = lon, y = lat, colour = age))  +
@@
 point_kelp_acc_April <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_point(data = kelp_acc_April, aes(x = lon, y = lat, colour = age))  +
@@
 heatmap_RK4_acc_July <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
@@
 heatmap_kelp_acc_July <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
@@
 point_RK4_acc_July <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_point(data = pass_acc_July, aes(x = lon, y = lat, colour = age))  +
@@
 point_kelp_acc_July <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_point(data = kelp_acc_July, aes(x = lon, y = lat, colour = age))  +
@@
 heatmap_RK4_acc_Oct <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
@@
 heatmap_kelp_acc_Oct <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
@@
 point_RK4_acc_Oct <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_point(data = pass_acc_Oct, aes(x = lon, y = lat, colour = age))  +
@@
 point_kelp_acc_Oct <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
   geom_point(data = kelp_acc_Oct, aes(x = lon, y = lat, colour = age)) +
@@
 ```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of density between passive and kelp accumulation for scenarios 53-60.
-For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=8, fig.width=8}
-
-ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January","", "April","","July","","October"), font.label = list(size = 12, face = "bold"), align = "hv", label.x = 0.2)
-
-ggsave(filename = "figure_9.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 2, dpi = 300)
+For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=8, fig.width=6}
+
+ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January","", "April","","July","","October"), font.label = list(size = 12, face = "bold"), align = "hv")
+
+ggsave(filename = "figure_9.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 6, scale = 1, dpi = 300)
 ```
@@
 Age is represented in hours since release.
-For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=8, fig.width=8}
-
-ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 12, face = "bold"), align = "hv", label.x = 0.2)
-ggsave(filename = "figure_10.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 2, dpi = 300)
+For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=8, fig.width=6}
+
+ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 12, face = "bold"), align = "hv", hjust = -0.3)
+
+ggsave(filename = "figure_10.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 6, scale = 1, dpi = 300)
 ```
 
-The role of eddys that exist within the study domain over the time period play a clear role in the accumulation and dispersal of kelp particles.
-Passive particles tend to remain on the outer edge of the eddy and then are expelled on the other side and continue to follow the ocean current (Figure 9).
+The role of eddys that exist within the study domain during the study period play a clear role in the accumulation and dispersal of kelp particles.
+Passive particles tend to remain on the outer edge of the eddy and are then expelled on the other side and continue to follow the ocean current (Figure 9).
 The opposite is true for kelp particles which have the highest exposure level of hydrodynamic and wind drag.
@@
 This is also reflected when comparing the age of particles at the end of the simulation.
-In Figure 10, younger particles are near the release site and older particles further away from shore, while older kelp particles are retained in the eddy.
+In Figure 10, younger particles are near the release site and older particles further away from the shore, while older kelp particles are retained within the eddy.
 
@@
 The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (appendix).
-This is also seen when comparing distance traveled by particles and particle density plots (appendix).
+This was also seen when comparing distance traveled by particles and particle density plots (appendix).
 The simulation employing a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (appendix).
@@
 @ragoasha2019 investigated the success of Lagrangian particles reaching St.
-Helena Bay from Cape point using Lagrangian trajectory modelling and inclusion of regional oceanographic processes.
-The authors showed that transport success is higher in spring due to the single coherent Benguela Jet, which concentrates over the 300m isobath [@ragoasha2019].
-During upwelling season (December - March) transport success is less efficient as particles tend to flow offshore.
-The move offshore is a result of the bifurification of the Benguela Jet, which splits because of shoaling and causes particles to be exposed to higher amounts of mesoscale activity.
-In the winter months, the Benguela Jet is still split but the flow weakens which results in lower transport success [@ragoasha2019].
+Helena Bay from Cape Point using Lagrangian trajectory modelling and inclusion of regional oceanographic processes.
+The authors showed that transport success was higher in spring due to the single coherent Benguela Jet, which is most intense over the 300m isobath [@ragoasha2019].
+During upwelling season (December - March) transport success was less efficient as particles tend to flow offshore.
+The move offshore was a result of the bifurification of the Benguela Jet, which splits because of shoaling and causes particles to be exposed to higher amounts of mesoscale activity.
+In the winter months, the Benguela Jet was still split but the flow weakens which resulted in lower transport success [@ragoasha2019].
 Similar patterns were found in this study, suggesting that the difference in seasonal patterns reflects the regional oceanographic features, particularly the Benguela Jet.
 
-The addition of drag forces causes the kelp particles to cluster together along the mean trajectory compared to that of the passive particles which are more evenly dispersed.
-This is most likely due to the shared drag characteristics among the kelp particles which included the plants weight and surface area into the calculation.
+The addition of drag forces caused the kelp particles to cluster together along the mean trajectory compared to that of the passive particles which were more evenly dispersed.
+This was most likely due to the shared drag characteristics among the kelp particles which included the plants' mass and surface area into the calculation.
 Other studies investigating flotsam trajectory characteristics have found similar results [@breivik2011; @olascoaga2020; @miron2020].
@@
 The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia.
-Although drag force and not inertia was used in this study, the same conclusion applies; drag is therefore an important characteristic to consider when investigating or predicting solitary drifting macroalgae trajectory.
+Although drag force and not inertia was used in this study, the same conclusion applies---drag is therefore an important characteristic to consider when investigating or predicting solitary drifting macroalgal trajectories.
 
 The ocean is comprised of time-varying flows and changes in an objects velocity that can cause the object to be exposed to different flow patterns over time.
-The lack of dissimilarity in trajectory between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible.
+The lack of dissimilarity in trajectory between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area was negligible.
 A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region.
-The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due as juveniles do not swim fast enough to alter their overall trajectory.
+The results suggested that the hatchling trajectories were mostly influenced by the ocean currents in the first year of hatching due as juveniles do not swim fast enough to alter their overall trajectory.
 Therefore, the differences in cross-sectional area types may not be significant enough to influence the relevant zonal and meridional velocity vectors.
-Another possible reason is the resolution of the ocean model used.
+Another possible reason was the resolution of the ocean model used.
 The resolution of the underlying ocean model is an important aspect of Lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications.
@@
 
-The same is also true when comparing different hydrodynamic and wind drag exposure scenarios.
-These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag.
-The similarities between the different combined drag scenarios suggest that higher magnitudes of wind exposure do not significantly alter trajectories.
-Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns.
+The same is true when comparing different hydrodynamic and wind drag exposure scenarios.
+These scenarios reflected different buoyancy situations which resulted in different amounts of surface area exposed to hydrodynamic and/or wind drag.
+The similarities between the different combined drag scenarios suggest that higher magnitudes of wind exposure did not significantly alter trajectories.
+Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of modeling entrainment patterns.
 All the simulations that included any form of drag resulted in a high density of particles entrained in an eddy.
 The entrainment and expulsion from eddys and vortices are an important characteristic of floating macroalgae.
-Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddys has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real *Sargassum* rafts.
+Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddys has been investigated previously by @putman2018 and @putman2020, who showed that including windage greatly improved entrainment patterns when comparing virtual and real *Sargassum* rafts.
 However, these studies investigated large rafts and not floating individuals.
-Inertia is based on the size/mass of the raft, while hydrodynamic and wind drag are based on current velocity and surface area.
+Inertia was based on the size/mass of the raft, while hydrodynamic and wind drag are based on current velocity and surface area.
 Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered.
-The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddys and vortices.
+The results from this study suggest that the inclusion of hydrodynamic and/or wind drag was an important component when simulating virtual kelp particles and greatly increased the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddys and vortices.
 
@@
 eddys and vortices may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in area below the eddy.
-In a study by @woodborne1989 references a site offshore approximately 120 km from Cape Columbine where dredged kelp was found.
-The area referenced by @woodborne1989 coincides with the area where virtual kelp is entrained by an eddy.
+A study by @woodborne1989 references a site offshore approximately 120 km from Cape Columbine where dredged kelp was found.
+The area coincides with the region where virtual kelp was entrained by an eddy.
 The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory.
@@
 
-Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape; however, prior to our study no research existed to clarify the drifting characteristics of solitary macroalgae.
-Understanding how to accurately model the distribution of solitary kelp can lead to be understanding of accumulation zones and sinks and ultimately the ecological links pertaining to those processes.
+Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape; however, prior to our study, no research existed to clarify the drifting characteristics of solitary macroalgae.
+Understanding how to accurately model the distribution of solitary kelp will lead to better understanding of accumulation zones and sinks, and ultimately the ecological links pertaining to those processes.
+The ecological links characterized will highlight ecosystems that are tightly dependent on kelp for proper ecological functioning, and these areas might also have a unique biodiversity compared to surrounding areas where less detrital input is present.
 The findings from this study indicated that solitary floating virtual *E.
-maxima* particles tend to follow the prevailing surface currents and the inclusion of wind caused the trajectories to differ greatly from that of a purely Lagrangian particle.
-Furthermore, the inclusion of both hydrodynamic and wind drag caused clustering of particles along the trajectory as well causing particles to become entrained in eddys.
+maxima* particles tended to follow the prevailing surface currents and the inclusion of wind caused the trajectories to differ greatly from that of a purely Lagrangian particle.
+Furthermore, the inclusion of both hydrodynamic and wind drag caused clustering of particles along the trajectory as well caused particles to become entrained in eddys.
 Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E.
 maxima*.
-In addition, the use of a cylindrical shape coefficient had no effect on mean trajectory, but it does cause higher density of particles to become entrained in an eddy.
+In addition, the use of a cylindrical shape coefficient had no effect on mean trajectory, but did cause higher density of particles to become entrained in an eddy.
 Overall, the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not.
 This study also provided an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately.
-Our results from this study highlights the importance of regional oceanographic processes to better understand how kelp flotsam is dispersed in the ocean.
-In particular, the current study highlights the role of the Benguela Jet and mesoscale processes and how these vary seasonally; ultimately determining transport success around the coast.
+Our results further highlight the importance of regional oceanographic processes to better understand how kelp flotsam is dispersed in the ocean.
+In particular, we highlight the role of the Benguela Jet and mesoscale processes, which ultimately determine the transport success around the coast.
 Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal.
-The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates and raft-times.
+The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates.
 
@@
 sim_table <- fontsize(sim_table, size = 8, part = "header")
-sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Drag.Exposure = "Drag Exposure level")
+sim_table <- set_header_labels(sim_table, Scenario = "Simulation", Cross.sectional.area = "Cross-sectional area", Drag.Exposure = "Drag Exposure level")
 sim_table <- align(x= sim_table, align = "center", part = "all")
@@
                size = 0.4, colour = "black", fill = "grey")  +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   labs(x = expression(Longitude*degree*S)) +
@@
                size = 0.4, colour = "black", fill = "grey")  +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = sphere_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
                size = 0.4, colour = "black", fill = "grey")  +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = cyl_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
                size = 0.4, colour = "black", fill = "grey")  +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = sphere_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
                size = 0.4, colour = "black", fill = "grey")  +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = cyl_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
                size = 0.4, colour = "black", fill = "grey")  +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
@@
                size = 0.4, colour = "black", fill = "grey")  +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = cyl_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
                size = 0.4, colour = "black", fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = shape_data_oct, aes(x = lon, y = lat, linetype = Type)) +
@@
                size = 0.4, colour = "black", fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   geom_path(data = sphere_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
@@
                size = 0.4, colour = "black", fill = "grey") +
+    geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
   scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
@@
 Row A represents the mean trajectory comparison, row B-C are density comparison plots, and row D is comparison of total distance travelled from the release site for each of the respective months (seasons) simulated.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=6}
-
-shape_plot_A <- ggarrange(shape_plot, shape_plot_April, shape_plot_July, shape_plot_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 1, labels = c("A", "", "", ""), font.label = list(size = 12, face = "bold"), align = "hv")
+The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=6}
+
+shape_plot_A <- ggarrange(shape_plot, shape_plot_April, shape_plot_July, shape_plot_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 1, font.label = list(size = 12, face = "bold"), align = "hv", labels = c("A","","",""))
 
@@
 
-ggarrange(shape_plot_A, shape_plot_B, shape_dis_plot, ncol = 1, nrow = 3, legend = c("right"), labels = c("", "D") , font.label = list(size = 11, face = "bold"), heights = c(2, 2, 1))
-
-#ggsave(filename = "figure_11.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 10, scale = 1, dpi = 300)
+ggarrange(shape_plot_A, shape_plot_B, shape_dis_plot, ncol = 1, nrow = 3, legend = c("right"), labels = c("","","","D"), font.label = list(size = 11, face = "bold"), heights = c(1,2,2), widths = c(1,2,2))
+
+ggsave(filename = "figure_11.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 7, scale = 1, dpi = 300)
 
@@
 
-```{r vector maps, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 11: The average ocean (left column) and wind vector maps (right column) for the respective seasons are presented.
-The velocity vectors were averaged over the 30 day simulation times for respective seasons (summer, autumn, winter, spring).
-A represents summer, B autumn, C winter and D spring", fig.height=8, fig.width=8}
+# Supplementary material
+
+```{r vector maps, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 12: The average ocean (left column) and wind vector maps (right column) for the respective seasons are presented.
+The velocity vectors were averaged over the 30 day simulation times for respective seasons (summer, autumn, winter, spring), for both ocean and wind data.
+In the case of wind, direction is also indicated by means of arrows.
+January ocean and wind data are presented in panels A-B, April in panels C-D, July in panels E-F, and October in panels G-H." ,fig.height=8, fig.width=6}
 

### 2021-05-11 thesis_chapter3_vector_map.R
map update

@@ -228,10 +228,10 @@ wind_plot_july <- ggplot() +
 
 wind_plot_oct <- ggplot() +
   geom_tile(data = wind_oct_summary, aes(fill = velocity, x = lon, y = lat)) +
-  #geom_raster(data = wind_vector, aes(x = lon, y = lat, fill = velocity))+
+  geom_raster(data = wind_vector, aes(x = lon, y = lat, fill = velocity))+
   geom_segment(data = wind_oct_summary,
-               aes(x = lon, xend = lon+eastward_wind/30, y = lat,
-                   yend = lat+northward_wind/30), arrow = arrow(length = unit(0.25, "cm"))) +
+              aes(x = lon, xend = lon+eastward_wind/60, y = lat,
+              yend = lat+northward_wind/60), arrow = arrow(length = unit(0.30, "cm"))) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",fill = "grey") +
   scale_fill_continuous(type = "viridis", limits = c(0, 11)) +
   coord_fixed(ratio=1, xlim = c(15, 20), ylim = c(-35, -30)) +
@@ -249,26 +249,36 @@ ocean <- ggarrange(ocean_plot_jan, ocean_plot_apr, ocean_plot_july, ocean_plot_o
                    font.label = list(size = 12, face = "bold"), labels = c("A", "B", "C","D"),
                    ncol = 2, nrow = 2, common.legend = TRUE, legend = c("bottom"))
 
-ocean_lng <- ggarrange(ocean_plot_jan, ocean_plot_apr, ocean_plot_july, ocean_plot_oct,
-                   font.label = list(size = 12, face = "bold"), labels = c("A", "C", "E","G"),
-                   ncol = 1, nrow = 4, common.legend = TRUE, legend = c("bottom"))
-
 wind <- ggarrange(wind_plot_jan, wind_plot_apr, wind_plot_july, wind_plot_oct,
                   font.label = list(size = 12, face = "bold"),labels = c("B", "D", "F","H"),
                   ncol = 2, nrow = 2, common.legend = TRUE, legend = c("bottom"))
 
+ocean_lng <- ggarrange(ocean_plot_jan, ocean_plot_apr, ocean_plot_july, ocean_plot_oct,
+                   font.label = list(size = 11, face = "bold"), labels = c("January","April","July","October"),
[…]

### 2021-05-12 thesis_chapter_3.Rmd
touchups

@@
 Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant.
-For details refer to figure 2.
+For details refer to Figure 2.
 The dimensional data needed were estimated in cases where data was not available for that morphological characteristic.
@@
 The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
-The minimum, mean and maximum cross-sectional areas (Ac, table 2) were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (table 2).
+The minimum, mean and maximum cross-sectional areas (Ac, Table 2) were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (Table 2).
 The trajectory of the different cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
@@
 Simulations of hydrodynamic drag are presented in Panels A, C, E, G for January, April, July, and October, respectively.
-Results of simulations including both hydrodynamic drag and wind are in Panels B, D, F, H for January, April, July, and October, respectively."}
+Results of simulations including both hydrodynamic drag and wind are in Panels B, D, F, H for January, April, July, and October, respectively.
+Significant amounts of overplotting prevents the different line types to be discernable."}
 
@@
 
-Density distributions summarizing the conclusion of sumlating Scenarios 1-13 (Table 1) are presented in Figures 3-6.
+Density distributions summarizing the conclusion of sumlating scenarios 1-13 (Table 1) are presented in Figures 3-6.
 Across all simulations for January (Austral summer), particles flow in a north-westward direction.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for simulations 1 - 13, for January 2018 (summer).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for simulations 1 - 13, for January 2018 (Austral summer).
 The plots are regular hexagons within the density of particles were calculated.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end of the run for simulations 14 - 26, for April 2018 (autumn).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end of the run for simulations 14 - 26, for April 2018 (Austral autumn).
 See Figure 4 for further details.', fig.height=8, fig.width=6}
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end of the run for simulations 27 - 39, for July 2018 (winter).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end of the run for simulations 27 - 39, for July 2018 (Austral winter).
 See Figure 4 for further details.', fig.height=8, fig.width=6}
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 7: Comparison of density plots at the end of the run for simulations 40 - 52, for October 2018 (spring).
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 7: Comparison of density plots at the end of the run for simulations 40 - 52, for October 2018 (Austral spring).
 See Figure 4 for further details.', fig.height=8, fig.width=6}
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=6}
+The reader is referred to Tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=6}
 
@@
 ```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 11: Comparison of trajectory data between sphere and cylinder shape coefficients for scenarios 61-64.
-Row A represents the mean trajectory comparison, row B-C are density comparison plots, and row D is comparison of total distance travelled from the release site for each of the respective months (seasons) simulated.
+Row A represents the mean trajectory comparison, row B-C are density comparison plots, and the bottom boxplot is comparison of shape coefficients total distance travelled from the release site for each of the respective months (seasons) simulated.
 The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=6}

### 2021-07-29 thesis_chapter_4.Rmd
methods addition chp4

@@
 date: "11/12/2020"
-output: html_document
+output:
+  html_document: default
+  pdf_document: default
 ---
@@
 ## Raft-time
-
 Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remains afloat at the surface of the ocean after detachment before sinking.
 Raft-time was estimated from the results of past studies by @graiff2016 and @graiff2013 for each season.
-For summer the raft time was 25 days, winter was 38 days, and autumn was 41 days respectively.
+For summer the raft time was 22 days, autumn was 41 days, winter was 38 days, and spring was 33 days.
 
@@
 
+# Calculate the volume fro each site for all morphological characteristics (meters^3)
+
+morph_data$total_volume <- with(morph_data, (frond_thickness*frond_length*frond_width) + (primary_length * primary_width * frond_thickness) + (pi*(stipe_radius*stipe_radius)*stipe_length) + (pi*height*0.3333*((top_radius*top_radius) + (top_radius * bottom_radius) + (bottom_radius*bottom_radius))))
+
 # Make table based on new entries
@@
 
-```{r, warning=FALSE, message=FALSE, echo = FALSE}
+## Simulations
+
+Simulations used daily outputs of velocity vectors (meridonal and zonal) and salinity (psu) for varying depths for the study region.
+To account for the role of morphological characteristics in influencing drift trajectory in terms of hydrodynamic and wind drag [@coppin2021] simulations were run separately for each site per season.
+Morphological characteristics of macroalgae has been shown to be based on the local wave and temperature environment, hence the need for site specific parameters with regards to morphological characteristics.
+Running simulations per season allowed incorporation of seasonal raft-times identified in pervious work [@graiff2016 and @graiff2013].
+The outputs from each simulation (site) was pooled based on season.
+
+Simulations were run using the OceanParcels software (available from https://oceanparcels.org/).
+All simulations were run using a Runge-Kutte fourth order advection scheme, while custom kernels simulated the effects of hydrodynamic and wind drag on overall trajectory, seasonal raft-times, sinking and settling on seafloor of virtual kelp particles; all important processes in the determination of surface and bottom sinks.
+The simulations were run for 2019, which was divided into various seasons (see table??).
+For incorporating drag, the 'maximum exposure scenario' proposed by @coppin2021 was used for all simulations (85% hydrodynamic and 15% wind drag) while virtual kelp particles were on the surface and then would convert to full hydrodynamic drag exposure (100%) once the virtual kelp started sinking; as wind would no longer act as a direct influencer of trajectory once below the ocean surface.
+
+## Analysis
+
+```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
+```{r eval = FALSE, message=FALSE, warning=FALSE, echo=FALSE}
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
+```{r eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r, message=FALSE, echo=FALSE, warning=FALSE}
+```{r eval = FALSE, message=FALSE, echo=FALSE, warning=FALSE}
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
+```{r , eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r total heatmap, echo=FALSE, message=FALSE, warning=FALSE}
+```{r total heatmap, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
 

### 2021-07-29 chapter_4_test_script.R
dummy file to test repo

@@ -1,0 +1,56 @@
+## Load software need for entire script
+
+load("map/south_africa_coast_hi_res.Rdata")
+load("map/sa_bathy.RData")
+
+library(ggplot2)
+library(gridExtra)
+library(ggpubr)
+library(mapproj)
+library(vegan)
+library(dplyr)
+library(ggpubr)
+library(maps)
+library(flextable)
+library(officer)
+library(hexbin)
+library(lubridate)
+library(tidyverse)
+library(tidync)
+library(ncdf4)
+
+options(scipen = 999)
+
+deVoss_filepath <- paste0("OceanParcels/coppin_extract_01.nc")
+deVoss_output <- nc_open(deVoss_filepath)
+
+coper_filepath <- paste0("OceanParcels/global-analysis-forecast-phy-001-024_1623936804092.nc")
+coper_output <- nc_open(coper_filepath)
+
+test <- tidync("OceanParcels/test.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, Salinity, age, KelpDepth, SinkSpeed, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+
+test_2 <- tidync("OceanParcels/test.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, Salinity, age, SinkSpeed, KelpDepth, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+
+test <- test_kom %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:KelpDepth, mean, na.rm= TRUE))
+
+test_3 <- tidync("OceanParcels/test.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, Salinity, age, SinkSpeed, KelpDepth, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
[…]

### 2021-08-04 thesis_chapter_4.Rmd
started to read data in

@@
 
-houtbay <- tidync("OceanParcels/trajectory_data/chapter_4_output/houtbay.nc") %>%
+houtbay_summer <- tidync("OceanParcels/chapter_4_output/hb_summer.nc") %>%
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
   dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+houtbay_summer$season <- "Summer" 
+houtbay_summer$site <- "Hout Bay" 
+houtbay_summer_mean <- houtbay_summer %>% # calculate mean trajectory
+ group_by(t) %>% 
+ summarise(across(lon:bathy, mean, na.rm= TRUE))
+houtbay_summer_mean$season <- "Summer" 
+houtbay_summer_mean$site <- "Hout Bay"
 
-houtbay_mean <- houtbay %>% # calculate mean trajectory
+houtbay_autumn <- tidync("OceanParcels/chapter_4_output/hb_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+houtbay_autumn$season <- "Autumn" 
+houtbay_autumn$site <- "Hout Bay" 
+houtbay_autumn_mean <- houtbay_autumn %>% # calculate mean trajectory
  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
+ summarise(across(lon:bathy, mean, na.rm= TRUE))
+houtbay_autumn_mean$season <- "Autumn" 
+houtbay_autumn_mean$site <- "Hout Bay"
 
-houtbay$site <- "Hout Bay" 
-houtbay_mean$site <- "Hout Bay"
+houtbay_winter <- tidync("OceanParcels/chapter_4_output/hb_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+houtbay_winter$season <- "Winter" 
+houtbay_winter$site <- "Hout Bay" 
+houtbay_winter_mean <- houtbay_winter %>% # calculate mean trajectory
+ group_by(t) %>% 
+ summarise(across(lon:bathy, mean, na.rm= TRUE))
+houtbay_winter_mean$season <- "Winter" 
+houtbay_winter_mean$site <- "Hout Bay"
 
-kommetjie <- tidync("OceanParcels/trajectory_data/chapter_4_output/kommetjie.nc") %>%
+houtbay_spring <- tidync("OceanParcels/chapter_4_output/hb_spring.nc") %>%
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
   dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+houtbay_spring$season <- "Spring" 
+houtbay_spring$site <- "Hout Bay" 
+houtbay_spring_mean <- houtbay_spring %>% # calculate mean trajectory
+ group_by(t) %>% 
+ summarise(across(lon:bathy, mean, na.rm= TRUE))
+houtbay_spring_mean$season <- "Spring" 
+houtbay_spring_mean$site <- "Hout Bay"
 
-kommetjie_mean <- kommetjie %>% # calculate mean trajectory
+kom_summer <- tidync("OceanParcels/chapter_4_output/kom_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+kom_summer$season <- "Summer" 
+kom_summer$site <- "Kommetjie" 
+kom_summer_mean <- kom_summer %>% # calculate mean trajectory
   group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+kom_summer_mean$season <- "Summer" 
+kom_summer_mean$site <- "Kommetjie"
 
-kommetjie$site <- "kommetjie" 
-kommetjie_mean$site <- "kommetjie"
+kom_autumn <- tidync("OceanParcels/chapter_4_output/kom_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+kom_autumn$season <- "Autumn" 
+kom_autumn$site <- "Kommetjie" 
+kom_autumn_mean <- kom_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+kom_autumn_mean$season <- "Autumn" 
+kom_autumn_mean$site <- "Kommetjie"
 
-soetwater <- tidync("OceanParcels/trajectory_data/chapter_4_output/soetwater.nc") %>%
+kom_winter <- tidync("OceanParcels/chapter_4_output/kom_winter.nc") %>%
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
   dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+kom_winter$season <- "Winter" 
+kom_winter$site <- "Kommetjie" 
+kom_winter_mean <- kom_winter %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+kom_winter_mean$season <- "Winter" 
+kom_winter_mean$site <- "Kommetjie"
 
-soetwater_mean <- soetwater %>% # calculate mean trajectory
+kom_spring <- tidync("OceanParcels/chapter_4_output/kom_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+kom_spring$season <- "Spring" 
+kom_spring$site <- "Kommetjie" 
+kom_spring_mean <- kom_spring %>% # calculate mean trajectory
   group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+kom_spring_mean$season <- "Spring" 
+kom_spring_mean$site <- "Kommetjie"
 
-soetwater$site <- "Soetwater" 
-soetwater_mean$site <- "Soetwater"
+scar_summer <- tidync("OceanParcels/chapter_4_output/scar_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+scar_summer$season <- "Summer" 
+scar_summer$site <- "Scarborough" 
+scar_summer_mean <- scar_summer %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+scar_summer_mean$season <- "Summer" 
+scar_summer_mean$site <- "Scarborough"
 
-millers <- tidync("OceanParcels/trajectory_data/chapter_4_output/millers.nc") %>%
+scar_autumn <- tidync("OceanParcels/chapter_4_output/scar_autumn.nc") %>%
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
   dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2020-01-01 00:00:00"))
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+scar_autumn$season <- "Autumn" 
+scar_autumn$site <- "Scarborough" 
+scar_autumn_mean <- scar_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+scar_autumn_mean$season <- "Autumn" 
+scar_autumn_mean$site <- "Scarborough"
 
-millers_mean <- millers %>% # calculate mean trajectory
+scar_winter <- tidync("OceanParcels/chapter_4_output/scar_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+scar_winter$season <- "Winter" 
+scar_winter$site <- "Scarborough" 
+scar_winter_mean <- scar_winter %>% # calculate mean trajectory
   group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+scar_winter_mean$season <- "Winter" 
+scar_winter_mean$site <- "Scarborough"
 
-millers$site <- "millers" 
-millers_mean$site <- "millers"
+scar_spring <- tidync("OceanParcels/chapter_4_output/scar_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+scar_spring$season <- "Spring" 
+scar_spring$site <- "Scarborough" 
+scar_spring_mean <- scar_spring %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+scar_spring_mean$season <- "Spring" 
+scar_spring_mean$site <- "Scarborough"
 
-total_data <- rbind(houtbay, soetwater, kommetjie, millers)
-total_data_mean <- rbind(houtbay_mean, soetwater_mean,kommetjie_mean, millers_mean)
+oude_summer <- tidync("OceanParcels/chapter_4_output/oudekraal_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+oude_summer$season <- "Summer" 
+oude_summer$site <- "Oudekraal" 
+oude_summer_mean <- oude_summer %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+oude_summer_mean$season <- "Summer" 
+oude_summer_mean$site <- "Oudekraal"
+
+oude_autumn <- tidync("OceanParcels/chapter_4_output/oudekraal_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+oude_autumn$season <- "Autumn" 
+oude_autumn$site <- "Oudekraal" 
+oude_autumn_mean <- oude_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+oude_autumn_mean$season <- "Autumn" 
+oude_autumn_mean$site <- "Oudekraal"
+
+oude_winter <- tidync("OceanParcels/chapter_4_output/oudekraal_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+oude_winter$season <- "Winter" 
+oude_winter$site <- "Oudekraal" 
+oude_wintermean <- oude_winter %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+oude_winter_mean$season <- "Winter" 
+oude_winter_mean$site <- "Oudekraal"
+
+oude_spring <- tidync("OceanParcels/chapter_4_output/oudekraal_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+oude_spring$season <- "Spring" 
+oude_spring$site <- "Oudekraal" 
+oude_spring <- scar_spring %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+oude_spring_mean$season <- "Spring" 
+oude_spring_mean$site <- "Oudekraal"
+
 ```
@@
 ## Mean trajectory
-
-```{r eval = FALSE, message=FALSE, warning=FALSE, echo=FALSE}
-
-library(ggplot2)
-
-## Coastline of Southern Africa
-load("map/south_africa_coast_hi_res.Rdata")
-
-mean_traj_site <- ggplot() + 
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 1.0, colour = "black", fill = NA) +
-  geom_path(data = total_data_mean, aes(x = lon, y = lat, colour = site)) +
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-```
-
-```{r eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-
-mean_traj_site
-
-```
-
-## Density comparisons
-
-```{r eval = FALSE, message=FALSE, echo=FALSE, warning=FALSE}
-
-## Coastline of Southern Africa
-load("map/south_africa_coast_hi_res.Rdata")
-
-heatmap_houtbay <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size = 1.0, colour = "black", fill = NA) +
-  geom_hex(data = houtbay, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000000)) +
-  geom_path(data = houtbay_mean,aes(x = lon, y = lat, color = 'red'),             show.legend = FALSE) +
-  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("HoutBay") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_komm <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kommetjie, bins = 80, aes(x = lon, y = lat), legend = FALSE) +   scale_fill_continuous(type = "gradient", limits = c(1, 1000000)) +
-  geom_path(data = kommetjie_mean,aes(x = lon, y = lat, color = 'red'),           show.legend = FALSE) +
-  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("Kommetjie") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_soetwater <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = soetwater, bins = 80, aes(x = lon, y = lat), legend = FALSE) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000000)) +
-  geom_path(data = soetwater_mean,aes(x = lon, y = lat, color = 'red'),           show.legend = FALSE) +
-  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("Soetwater") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_soetwater <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = soetwater, bins = 80, aes(x = lon, y = lat), legend = FALSE) +
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000000)) +
-  geom_path(data = soetwater_mean,aes(x = lon, y = lat, color = 'red'),           show.legend = FALSE) +
-  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("Kommetjie") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-heatmap_millers <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size = 1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = millers, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 1000000)) +
-  geom_path(data = millers_mean,aes(x = lon, y = lat, color = 'red'),             show.legend = FALSE) +
-  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("Millers") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-```
-
-```{r , eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-
-ggarrange(heatmap_houtbay, heatmap_soetwater, heatmap_millers, ncol = 1, nrow = 3, font.label = list(size = 11, face = "bold"), common.legend = FALSE, legend = c("bottom"), labels = "AUTO", align = "hv")
-```
-
-```{r total heatmap, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
-
-heatmap_total <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group =    PID), size =    1.0, colour = "black", fill = NA) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = total_data, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
-  scale_fill_continuous(type = "gradient", limits = c(1, 19000000)) +
-  geom_path(data = total_data_mean,aes(x = lon, y = lat, color = 'red'),show.legend = FALSE) +
-  geom_rect(data = south_africa_coast_hi_res,aes(xmin=10, xmax=18.4, ymin=-33, ymax=-33.5),           inherit.aes = FALSE, fill = NA, colour =    "red", size = 1.5) +
-  geom_rect(data = south_africa_coast_hi_res,aes(xmin=10, xmax=18.4, ymin=-33.5, ymax=-34.0),           inherit.aes = FALSE, fill = NA, colour =    "green", size = 1.5) +
-  geom_rect(data = south_africa_coast_hi_res,aes(xmin=10, xmax=18.4, ymin=-34.0, ymax=-34.5),           inherit.aes = FALSE, fill = NA, colour =    "black", size = 1.5) +
-  #coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  xlab("Longtitude") + 
-  ylab("Latitude") + 
-  ggtitle("Millers") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-heatmap_total
-```

### 2021-08-04 thesis_chapter_4.Rmd
more sites read-in

@@
 oude_winter$site <- "Oudekraal" 
-oude_wintermean <- oude_winter %>% # calculate mean trajectory
+oude_winter_mean <- oude_winter %>% # calculate mean trajectory
   group_by(t) %>% 
@@
 oude_spring$site <- "Oudekraal" 
-oude_spring <- scar_spring %>% # calculate mean trajectory
+oude_spring <- oude_spring %>% # calculate mean trajectory
   group_by(t) %>% 
@@
 
+betty_summer <- tidync("OceanParcels/chapter_4_output/betty_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+betty_summer$season <- "Summer" 
+betty_summer$site <- "Betty's Bay" 
+betty_summer_mean <- betty_summer %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+betty_summer_mean$season <- "Summer" 
+betty_summer_mean$site <- "Betty's Bay"
+
+betty_autumn <- tidync("OceanParcels/chapter_4_output/betty_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+betty_autumn$season <- "Autumn" 
+betty_autumn$site <- "Betty's Bay" 
+betty_autumn_mean <- betty_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+betty_autumn_mean$season <- "Autumn" 
+betty_autumn_mean$site <- "Betty's Bay"
+
+betty_winter <- tidync("OceanParcels/chapter_4_output/betty_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+betty_winter$season <- "Winter" 
+betty_winter$site <- "Betty's Bay" 
+betty_winter_mean <- betty_winter %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+betty_winter_mean$season <- "Winter" 
+betty_winter_mean$site <- "Betty's Bay"
+
+betty_spring <- tidync("OceanParcels/chapter_4_output/betty_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+betty_spring$season <- "Spring" 
+betty_spring$site <- "Betty's Bay" 
+betty_spring_mean <- betty_spring %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+betty_spring_mean$season <- "Spring" 
+betty_spring_mean$site <- "Betty's Bay"
+
+kalk_summer <- tidync("OceanParcels/chapter_4_output/kalk_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+kalk_summer$season <- "Summer" 
+kalk_summer$site <- "Kalk Bay" 
+kalk_summer_mean <- kalk_summer %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+kalk_summer_mean$season <- "Summer" 
+kalk_summer_mean$site <- "Kalk Bay"
+
+kalk_autumn <- tidync("OceanParcels/chapter_4_output/kalk_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+kalk_autumn$season <- "Autumn" 
+kalk_autumn$site <- "Kalk Bay" 
+kalk_autumn_mean <- kalk_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+kalk_autumn_mean$season <- "Autumn" 
+kalk_autumn_mean$site <- "Kalk Bay"
+
+kalk_winter <- tidync("OceanParcels/chapter_4_output/kalk_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+kalk_winter$season <- "Winter" 
+kalk_winter$site <- "Kalk Bay" 
+kalk_winter_mean <- kalk_winter %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+kalk_winter_mean$season <- "Winter" 
+kalk_winter_mean$site <- "Kalk Bay"
+
+kalk_spring <- tidync("OceanParcels/chapter_4_output/kalk_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+kalk_spring$season <- "Spring" 
+kalk_spring$site <- "Kalk Bay" 
+kalk_spring_mean <- kalk_spring %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+kalk_spring_mean$season <- "Spring" 
+kalk_spring_mean$site <- "Kalk Bay"
+
+aframe_summer <- tidync("OceanParcels/chapter_4_output/aframe_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+aframe_summer$season <- "Summer" 
+aframe_summer$site <- "A-Frame" 
+aframe_summer_mean <- aframe_summer %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+aframe_summer_mean$season <- "Summer" 
+aframe_summer_mean$site <- "A-Frame"
+
+aframe_autumn <- tidync("OceanParcels/chapter_4_output/aframe_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+aframe_autumn$season <- "Autumn" 
+aframe_autumn$site <- "A-Frame" 
+aframe_autumn_mean <- aframe_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+aframe_autumn_mean$season <- "Autumn" 
+aframe_autumn_mean$site <- "A-Frame"
+
+aframe_winter <- tidync("OceanParcels/chapter_4_output/aframe_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+aframe_winter$season <- "Winter" 
+aframe_winter$site <- "A-Frame" 
+aframe_winter_mean <- aframe_winter %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+aframe_winter_mean$season <- "Winter" 
+aframe_winter_mean$site <- "A-Frame"
+
+aframe_spring <- tidync("OceanParcels/chapter_4_output/aframe_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+aframe_spring$season <- "Spring" 
+aframe_spring$site <- "A-Frame" 
+aframe_spring_mean <- aframe_spring %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+aframe_spring_mean$season <- "Spring" 
+aframe_spring_mean$site <- "A-Frame"
+
+sthelena_summer <- tidync("OceanParcels/chapter_4_output/sthelena_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+sthelena_summer$season <- "Summer" 
+sthelena_summer$site <- "St. Helena Bay" 
+sthelena_summer_mean <- sthelena_summer %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+sthelena_summer_mean$season <- "Summer" 
+sthelena_summer_mean$site <- "St. Helena Bay"
+
+sthelena_autumn <- tidync("OceanParcels/chapter_4_output/sthelena_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+sthelena_autumn$season <- "Autumn" 
+sthelena_autumn$site <- "St. Helena Bay" 
+sthelena_autumn_mean <- sthelena_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+sthelena_autumn_mean$season <- "Autumn" 
+sthelena_autumn_mean$site <- "St. Helena Bay"
+
+sthelena_winter <- tidync("OceanParcels/chapter_4_output/sthelena_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+sthelena_winter$season <- "Winter" 
+sthelena_winter$site <- "St. Helena Bay" 
+sthelena_winter_mean <- sthelena_winter %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+sthelena_winter_mean$season <- "Winter" 
+sthelena_winter_mean$site <- "St. Helena Bay"
+
+sthelena_spring <- tidync("OceanParcels/chapter_4_output/sthelena_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+sthelena_spring$season <- "Spring" 
+sthelena_spring$site <- "St. Helena Bay" 
+sthelena_spring_mean <- sthelena_spring %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+sthelena_spring_mean$season <- "Spring" 
+sthelena_spring_mean$site <- "St. Helena Bay"
+
 ```

### 2021-08-04 thesis_chapter_4.Rmd
even more read-in sites

@@
 
+millers_summer <- tidync("OceanParcels/chapter_4_output/millers_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+millers_summer$season <- "Summer" 
+millers_summer$site <- "Miller's Point" 
+millers_summer_mean <- millers_summer %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+millers_summer_mean$season <- "Summer" 
+millers_summer_mean$site <- "Miller's Point"
+
+millers_autumn <- tidync("OceanParcels/chapter_4_output/millers_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+millers_autumn$season <- "Autumn" 
+millers_autumn$site <- "Miller's Point" 
+millers_autumn_mean <- millers_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+millers_autumn_mean$season <- "Autumn" 
+millers_autumn_mean$site <- "Miller's Point"
+
+millers_winter <- tidync("OceanParcels/chapter_4_output/millers_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+millers_winter$season <- "Winter" 
+millers_winter$site <- "Miller's Point" 
+millers_winter_mean <- millers_winter %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+millers_winter_mean$season <- "Winter" 
+millers_winter_mean$site <- "Miller's Point"
+
+millers_spring <- tidync("OceanParcels/chapter_4_output/") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+millers_spring$season <- "Spring" 
+millers_spring$site <- "Miller's Point" 
+millers_spring_mean <- millers_spring %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+millers_spring_mean$season <- "Spring" 
+millers_spring_mean$site <- "Miller's Point"
+
+baboon_summer <- tidync("OceanParcels/chapter_4_output/baboon_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+baboon_summer$season <- "Summer" 
+baboon_summer$site <- "Baboon Rock" 
+baboon_summer_mean <- baboon_summer %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+baboon_summer_mean$season <- "Summer" 
+baboon_summer_mean$site <- "Baboon Rock"
+
+baboon_autumn <- tidync("OceanParcels/chapter_4_output/baboon_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+baboon_autumn$season <- "Autumn" 
+baboon_autumn$site <- "Baboon Rock" 
+baboon_autumn_mean <- baboon_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+baboon_autumn_mean$season <- "Autumn" 
+baboon_autumn_mean$site <- "Baboon Rock"
+
+baboon_winter <- tidync("OceanParcels/chapter_4_output/baboon_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+baboon_winter$season <- "Winter" 
+baboon_winter$site <- "Baboon Rock" 
+baboon_winter_mean <- baboon_winter %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+baboon_winter_mean$season <- "Winter" 
+baboon_winter_mean$site <- "Baboon Rock"
+
+baboon_spring <- tidync("OceanParcels/chapter_4_output/baboon_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+baboon_spring$season <- "Spring" 
+baboon_spring$site <- "Baboon Rock" 
+baboon_spring_mean <- baboon_spring %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+baboon_spring_mean$season <- "Spring" 
+baboon_spring_mean$site <- "Baboon Rock"
+
+basata_summer <- tidync("OceanParcels/chapter_4_output/basata_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+basata_summer$season <- "Summer" 
+basata_summer$site <- "Basata Rock" 
+basata_summer_mean <- basata_summer %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+basata_summer_mean$season <- "Summer" 
+basata_summer_mean$site <- "Basata Rock"
+
+basata_autumn <- tidync("OceanParcels/chapter_4_output/basata_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+basata_autumn$season <- "Autumn" 
+basata_autumn$site <- "Basata Rock" 
+basata_autumn_mean <- basata_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+basata_autumn_mean$season <- "Autumn" 
+basata_autumn_mean$site <- "Basata Rock"
+
+basata_winter <- tidync("OceanParcels/chapter_4_output/basata_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+basata_winter$season <- "Winter" 
+basata_winter$site <- "Basata Rock" 
+basata_winter_mean <- basata_winter %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+basata_winter_mean$season <- "Winter" 
+basata_winter_mean$site <- "Basata Rock"
+
+basata_spring <- tidync("OceanParcels/chapter_4_output/basata_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+basata_spring$season <- "Spring" 
+basata_spring$site <- "Basata Rock" 
+basata_spring_mean <- basata_spring %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+basata_spring_mean$season <- "Spring" 
+basata_spring_mean$site <- "Basata Rock"
+
 ```

### 2021-08-04 thesis_chapter_4.Rmd
inital plot code

@@
 oude_spring$site <- "Oudekraal" 
-oude_spring <- oude_spring %>% # calculate mean trajectory
+oude_spring_mean <- oude_spring %>% # calculate mean trajectory
   group_by(t) %>% 
@@
 
-millers_spring <- tidync("OceanParcels/chapter_4_output/") %>%
+millers_spring <- tidync("OceanParcels/chapter_4_output/millers_spring.nc") %>%
   hyper_tibble() %>%
@@
 
+buffels_summer <- tidync("OceanParcels/chapter_4_output/buffels_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+buffels_summer$season <- "Summer" 
+buffels_summer$site <- "Buffelsbaai" 
+buffels_summer_mean <- buffels_summer %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+buffels_summer_mean$season <- "Summer" 
+buffels_summer_mean$site <- "Buffelsbaai"
+
+buffels_autumn <- tidync("OceanParcels/chapter_4_output/buffels_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+buffels_autumn$season <- "Autumn" 
+buffels_autumn$site <- "Buffelsbaai" 
+buffels_autumn_mean <- buffels_autumn %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+buffels_autumn_mean$season <- "Autumn" 
+buffels_autumn_mean$site <- "Buffelsbaai"
+
+buffels_winter <- tidync("OceanParcels/chapter_4_output/buffels_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+buffels_winter$season <- "Winter" 
+buffels_winter$site <- "Buffelsbaai" 
+buffels_winter_mean <- buffels_winter %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+buffels_winter_mean$season <- "Winter" 
+buffels_winter_mean$site <- "Buffelsbaai"
+
+buffels_spring <- tidync("OceanParcels/chapter_4_output/buffels_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
+buffels_spring$season <- "Spring" 
+buffels_spring$site <- "Buffelsbaai" 
+buffels_spring_mean <- buffels_spring %>% # calculate mean trajectory
+  group_by(t) %>% 
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+buffels_spring_mean$season <- "Spring" 
+buffels_spring_mean$site <- "Buffelsbaai"
+
+soetwater_summer <- tidync("OceanParcels/chapter_4_output/soetwater_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+soetwater_summer$season <- "Summer"
+soetwater_summer$site <- "Soetwater"
+soetwater_summer_mean <- soetwater_summer %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+soetwater_summer_mean$season <- "Summer"
+soetwater_summer_mean$site <- "Soetwater"
+
+soetwater_autumn <- tidync("OceanParcels/chapter_4_output/soetwater_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+soetwater_autumn$season <- "Autumn"
+soetwater_autumn$site <- "Soetwater"
+soetwater_autumn_mean <- soetwater_autumn %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+soetwater_autumn_mean$season <- "Autumn"
+soetwater_autumn_mean$site <- "Soetwater"
+
+soetwater_winter <- tidync("OceanParcels/chapter_4_output/soetwater_winter.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+soetwater_winter$season <- "Winter"
+soetwater_winter$site <- "Soetwater"
+soetwater_winter_mean <- soetwater_winter %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+soetwater_winter_mean$season <- "Winter"
+soetwater_winter_mean$site <- "Soetwater"
+
+soetwater_spring <- tidync("OceanParcels/chapter_4_output/soetwater_spring.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+soetwater_spring$season <- "Spring"
+soetwater_spring$site <- "Soetwater"
+soetwater_spring_mean <- soetwater_spring %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+soetwater_spring_mean$season <- "Spring"
+soetwater_spring_mean$site <- "Soetwater"
+
 ```
@@
 ## Mean trajectory
+
+```{r message=FALSE, warning=FALSE}
+# Summer
+total_summer <- rbind(kom_summer_mean, houtbay_summer_mean, scar_summer_mean, oude_summer_mean, betty_summer_mean, kalk_summer_mean, aframe_summer_mean, sthelena_summer_mean, millers_summer_mean, baboon_summer_mean, basata_summer_mean, buffels_summer_mean, soetwater_summer_mean)
+
+summer_mean <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_path(data = total_summer, aes(x = lon, y = lat, linetype = site))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(linetype=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# Autumn
+total_autumn <- rbind(kom_autumn_mean, houtbay_autumn_mean, scar_autumn_mean, oude_autumn_mean, betty_autumn_mean, kalk_autumn_mean, aframe_autumn_mean, sthelena_autumn_mean, millers_autumn_mean, baboon_autumn_mean, basata_autumn_mean, buffels_autumn_mean, soetwater_autumn_mean)
+
+autumn_mean <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_path(data = total_autumn, aes(x = lon, y = lat, linetype = site))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(linetype=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# Winter
+total_winter <- rbind(kom_winter_mean, houtbay_winter_mean, scar_winter_mean, oude_winter_mean, betty_winter_mean, kalk_winter_mean, aframe_winter_mean, sthelena_winter_mean, millers_winter_mean, baboon_winter_mean, basata_winter_mean, buffels_winter_mean, soetwater_winter_mean)
+
+winter_mean <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_path(data = total_winter, aes(x = lon, y = lat, linetype = site))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(linetype=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+# Spring
+total_spring <- rbind(kom_spring_mean, houtbay_spring_mean, scar_spring_mean, oude_spring_mean, betty_spring_mean, kalk_spring_mean, aframe_spring_mean, sthelena_spring_mean, millers_spring_mean, baboon_spring_mean, basata_spring_mean, buffels_spring_mean, soetwater_spring_mean)
+
+spring_mean <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_path(data = total_spring, aes(x = lon, y = lat, linetype = site))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(linetype=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+ggarrange(summer_mean, autumn_mean, winter_mean, common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"))
+
+```
+
+## Heatmaps
+
+```{r message=FALSE, warning=FALSE}
+
+## Load software need for entire script
+
+load("map/south_africa_coast_hi_res.Rdata")
+load("map/sa_bathy.RData")
+
+aframe_summer_heatmap <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = aframe_summer, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis", limits = c(1, 68000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  geom_path(data = aframe_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+```

### 2021-08-04 chapter_4_test_script.R
test script used and updated

@@ -54,3 +54,65 @@ bathy_output <- nc_open(bathy_filepath)
 
 wind_filepath <- paste0("OceanParcels/copernicus/CERSAT-GLO-BLENDED_WIND_L4-V6-OBS_FULL_TIME_SERIE_1624029428851.nc")
 wind_output <- nc_open(wind_filepath)
+
+
+##################################################
+
+soetwater_summer <- tidync("OceanParcels/chapter_4_output/soetwater_summer.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+soetwater_summer$season <- "Summer"
+soetwater_summer$site <- "Soetwater"
+soetwater_summer_mean <- soetwater_summer %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+soetwater_summer_mean$season <- "Summer"
+soetwater_summer_mean$site <- "Soetwater"
+
+soetwater_autumn <- tidync("OceanParcels/chapter_4_output/soetwater_autumn.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+soetwater_autumn$season <- "Autumn"
+soetwater_autumn$site <- "Soetwater"
+soetwater_autumn_mean <- soetwater_autumn %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+soetwater_autumn_mean$season <- "Autumn"
+soetwater_autumn_mean$site <- "Soetwater"
+
+soetwater_winter <- tidync("OceanParcels/chapter_4_output/soetwater_winter.nc") %>%
+  hyper_tibble() %>%
[…]

### 2021-08-04 thesis_chapter_4.Rmd
yaml
added ocean vector code, adjust mean traj figures

@@
 output:
+  word_document: default
+  pdf_document: default
   html_document: default
-  pdf_document: default
 ---
@@
 ## Analysis
+
+# Results
+
+## Ocean analysis
+
+```{r eval=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+
+ocean_vector_chp4 <- tidync("OceanParcels/chapter_4_oceanmodel/copernicus_phy001_030/global-reanalysis-phy-001-030-daily_1625758460342.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, longitude, latitude, uo, vo) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2019-01-01 00:00:00"),
+                     "GMT",
+                     "%Y-%m-%d-%h"))
+
+ocean_summer <-  ocean_vector_chp4 %>%
+  filter(t >= ("2086-01-01 00:00:00") & t <= ("2086-02-28 00:00:00"))
+ocean_autumn <-  ocean_vector_chp4 %>%
+  filter(t >= ("2086-03-01 00:00:00") & t <= ("2086-05-31 00:00:00"))
+ocean_winter <-  ocean_vector_chp4 %>%
+  filter(t >= ("2086-06-01 00:00:00") & t <= ("2086-08-30 00:00:00"))
+ocean_spring <-  ocean_vector_chp4 %>%
+  filter(t >= ("2086-09-01 00:00:00") & t <= ("2086-11-30 00:00:00"))
+
+ocean_summer$velocity <- with(ocean_jan, sqrt(uo^2 + vo^2))
+ocean_autumn$velocity <- with(ocean_apr, sqrt(uo^2 + vo^2))
+ocean_winter$velocity <- with(ocean_july, sqrt(uo^2 + vo^2))
+ocean_spring$velocity <- with(ocean_oct, sqrt(uo^2 + vo^2))
+
+```
 
@@
 ```{r message=FALSE, warning=FALSE}
+
+load("map/south_africa_coast_hi_res.Rdata")
+load("map/sa_bathy.RData")
+
 # Summer
@@
   labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Summer") +
   guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1, clip = "on", xlim = c(15, 20), ylim = c(-35, -31)) + # fix the aspect ratio of plot
+  theme_bw() +
   theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
   theme(axis.text=element_text(size=12),
@@
   labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Autumn") +
   guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1, clip = "on", xlim = c(15, 20), ylim = c(-35, -31)) + # fix the aspect ratio of plot
+  theme_bw() +
   theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
   theme(axis.text=element_text(size=12),
@@
   labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Winter") +
   guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1, clip = "on", xlim = c(15, 20), ylim = c(-35, -31)) + # fix the aspect ratio of plot
+  theme_bw() +
   theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
   theme(axis.text=element_text(size=12),
@@
   labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Spring") +
   guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  coord_fixed(ratio=1, clip = "on", xlim = c(15, 20), ylim = c(-35, -31)) + # fix the aspect ratio of plot
+  theme_bw() +
   theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
   theme(axis.text=element_text(size=12),
@@
 
-ggarrange(summer_mean, autumn_mean, winter_mean, common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"))
+ggarrange(summer_mean, autumn_mean, winter_mean, spring_mean,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 2)
 

### 2021-08-04 thesis_chapter_4.Rmd
ocean vector script added and heatmaps for each season

@@
 # Summer
-total_summer <- rbind(kom_summer_mean, houtbay_summer_mean, scar_summer_mean, oude_summer_mean, betty_summer_mean, kalk_summer_mean, aframe_summer_mean, sthelena_summer_mean, millers_summer_mean, baboon_summer_mean, basata_summer_mean, buffels_summer_mean, soetwater_summer_mean)
+total_summer_mean <- rbind(kom_summer_mean, houtbay_summer_mean, scar_summer_mean, oude_summer_mean, betty_summer_mean, kalk_summer_mean, aframe_summer_mean, sthelena_summer_mean, millers_summer_mean, baboon_summer_mean, basata_summer_mean, buffels_summer_mean, soetwater_summer_mean)
 
@@
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_summer, aes(x = lon, y = lat, linetype = site))+
+  geom_path(data = total_summer_mean, aes(x = lon, y = lat, linetype = site))+
   labs(x = expression(Longitude*degree*S)) +
@@
 # Autumn
-total_autumn <- rbind(kom_autumn_mean, houtbay_autumn_mean, scar_autumn_mean, oude_autumn_mean, betty_autumn_mean, kalk_autumn_mean, aframe_autumn_mean, sthelena_autumn_mean, millers_autumn_mean, baboon_autumn_mean, basata_autumn_mean, buffels_autumn_mean, soetwater_autumn_mean)
+total_autumn_mean <- rbind(kom_autumn_mean, houtbay_autumn_mean, scar_autumn_mean, oude_autumn_mean, betty_autumn_mean, kalk_autumn_mean, aframe_autumn_mean, sthelena_autumn_mean, millers_autumn_mean, baboon_autumn_mean, basata_autumn_mean, buffels_autumn_mean, soetwater_autumn_mean)
 
@@
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_autumn, aes(x = lon, y = lat, linetype = site))+
+  geom_path(data = total_autumn_mean, aes(x = lon, y = lat, linetype = site))+
   labs(x = expression(Longitude*degree*S)) +
@@
 # Winter
-total_winter <- rbind(kom_winter_mean, houtbay_winter_mean, scar_winter_mean, oude_winter_mean, betty_winter_mean, kalk_winter_mean, aframe_winter_mean, sthelena_winter_mean, millers_winter_mean, baboon_winter_mean, basata_winter_mean, buffels_winter_mean, soetwater_winter_mean)
+total_winter_mean <- rbind(kom_winter_mean, houtbay_winter_mean, scar_winter_mean, oude_winter_mean, betty_winter_mean, kalk_winter_mean, aframe_winter_mean, sthelena_winter_mean, millers_winter_mean, baboon_winter_mean, basata_winter_mean, buffels_winter_mean, soetwater_winter_mean)
 
@@
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_winter, aes(x = lon, y = lat, linetype = site))+
+  geom_path(data = total_winter_mean, aes(x = lon, y = lat, linetype = site))+
   labs(x = expression(Longitude*degree*S)) +
@@
 # Spring
-total_spring <- rbind(kom_spring_mean, houtbay_spring_mean, scar_spring_mean, oude_spring_mean, betty_spring_mean, kalk_spring_mean, aframe_spring_mean, sthelena_spring_mean, millers_spring_mean, baboon_spring_mean, basata_spring_mean, buffels_spring_mean, soetwater_spring_mean)
+total_spring_mean <- rbind(kom_spring_mean, houtbay_spring_mean, scar_spring_mean, oude_spring_mean, betty_spring_mean, kalk_spring_mean, aframe_spring_mean, sthelena_spring_mean, millers_spring_mean, baboon_spring_mean, basata_spring_mean, buffels_spring_mean, soetwater_spring_mean)
 
@@
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_spring, aes(x = lon, y = lat, linetype = site))+
+  geom_path(data = total_spring_mean, aes(x = lon, y = lat, linetype = site))+
   labs(x = expression(Longitude*degree*S)) +
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-aframe_summer_heatmap <- ggplot() +
+# Summer
+total_summer <- rbind(kom_summer, houtbay_summer, scar_summer, oude_summer, betty_summer, kalk_summer, aframe_summer, sthelena_summer, millers_summer, baboon_summer, basata_summer, buffels_summer, soetwater_summer)
+
+summer_heatmap <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_hex(data = aframe_summer, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis", limits = c(1, 68000)) +
+  geom_hex(data = total_summer, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis", limits = c(1, 100000)) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = aframe_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
 
+# Autumn
+
+total_autumn <- rbind(kom_autumn, houtbay_autumn, scar_autumn, oude_autumn, betty_autumn, kalk_autumn, aframe_autumn, sthelena_autumn, millers_autumn, baboon_autumn, basata_autumn, buffels_autumn, soetwater_autumn)
+
+autumn_heatmap <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_autumn, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# Winter
+
+total_winter <- rbind(kom_winter, houtbay_winter, scar_winter, oude_winter, betty_winter, kalk_winter, aframe_winter, sthelena_winter, millers_winter, baboon_winter, basata_winter, buffels_winter, soetwater_winter)
+
+autumn_heatmap <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_winter, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+# Spring
+
+total_spring <- rbind(kom_spring, houtbay_spring, scar_spring, oude_spring, betty_spring, kalk_spring, aframe_spring, sthelena_spring, millers_spring, baboon_spring, basata_spring, buffels_spring, soetwater_spring)
+
+autumn_heatmap <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
 ```

### 2021-08-05 thesis_chapter_4.Rmd
added more plots (depth plots)

@@
 
-```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+```{r data wrangle, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
+# Summer
+total_summer <- rbind(kom_summer, houtbay_summer, scar_summer, oude_summer, betty_summer, kalk_summer, aframe_summer, sthelena_summer, millers_summer, baboon_summer, basata_summer, buffels_summer, soetwater_summer)
+
+# Autumn
+
+total_autumn <- rbind(kom_autumn, houtbay_autumn, scar_autumn, oude_autumn, betty_autumn, kalk_autumn, aframe_autumn, sthelena_autumn, millers_autumn, baboon_autumn, basata_autumn, buffels_autumn, soetwater_autumn)
+
+# Winter
+
+total_winter <- rbind(kom_winter, houtbay_winter, scar_winter, oude_winter, betty_winter, kalk_winter, aframe_winter, sthelena_winter, millers_winter, baboon_winter, basata_winter, buffels_winter, soetwater_winter)
+
+# Spring
+
+total_spring <- rbind(kom_spring, houtbay_spring, scar_spring, oude_spring, betty_spring, kalk_spring, aframe_spring, sthelena_spring, millers_spring, baboon_spring, basata_spring, buffels_spring, soetwater_spring)
+
+# Summer
+total_summer_mean <- rbind(kom_summer_mean, houtbay_summer_mean, scar_summer_mean, oude_summer_mean, betty_summer_mean, kalk_summer_mean, aframe_summer_mean, sthelena_summer_mean, millers_summer_mean, baboon_summer_mean, basata_summer_mean, buffels_summer_mean, soetwater_summer_mean)
+
+# Autumn
+total_autumn_mean <- rbind(kom_autumn_mean, houtbay_autumn_mean, scar_autumn_mean, oude_autumn_mean, betty_autumn_mean, kalk_autumn_mean, aframe_autumn_mean, sthelena_autumn_mean, millers_autumn_mean, baboon_autumn_mean, basata_autumn_mean, buffels_autumn_mean, soetwater_autumn_mean)
+
+# Winter
+total_winter_mean <- rbind(kom_winter_mean, houtbay_winter_mean, scar_winter_mean, oude_winter_mean, betty_winter_mean, kalk_winter_mean, aframe_winter_mean, sthelena_winter_mean, millers_winter_mean, baboon_winter_mean, basata_winter_mean, buffels_winter_mean, soetwater_winter_mean)
+
+# Spring
+total_spring_mean <- rbind(kom_spring_mean, houtbay_spring_mean, scar_spring_mean, oude_spring_mean, betty_spring_mean, kalk_spring_mean, aframe_spring_mean, sthelena_spring_mean, millers_spring_mean, baboon_spring_mean, basata_spring_mean, buffels_spring_mean, soetwater_spring_mean)
+
 ```
@@
 load("map/sa_bathy.RData")
-
-# Summer
-total_summer_mean <- rbind(kom_summer_mean, houtbay_summer_mean, scar_summer_mean, oude_summer_mean, betty_summer_mean, kalk_summer_mean, aframe_summer_mean, sthelena_summer_mean, millers_summer_mean, baboon_summer_mean, basata_summer_mean, buffels_summer_mean, soetwater_summer_mean)
 
@@
 
-# Autumn
-total_autumn_mean <- rbind(kom_autumn_mean, houtbay_autumn_mean, scar_autumn_mean, oude_autumn_mean, betty_autumn_mean, kalk_autumn_mean, aframe_autumn_mean, sthelena_autumn_mean, millers_autumn_mean, baboon_autumn_mean, basata_autumn_mean, buffels_autumn_mean, soetwater_autumn_mean)
-
 autumn_mean <- ggplot()+
@@
 
-# Winter
-total_winter_mean <- rbind(kom_winter_mean, houtbay_winter_mean, scar_winter_mean, oude_winter_mean, betty_winter_mean, kalk_winter_mean, aframe_winter_mean, sthelena_winter_mean, millers_winter_mean, baboon_winter_mean, basata_winter_mean, buffels_winter_mean, soetwater_winter_mean)
-
 winter_mean <- ggplot()+
@@
         axis.title=element_text(size=12))
-# Spring
-total_spring_mean <- rbind(kom_spring_mean, houtbay_spring_mean, scar_spring_mean, oude_spring_mean, betty_spring_mean, kalk_spring_mean, aframe_spring_mean, sthelena_spring_mean, millers_spring_mean, baboon_spring_mean, basata_spring_mean, buffels_spring_mean, soetwater_spring_mean)
 
@@
 load("map/sa_bathy.RData")
-
-# Summer
-total_summer <- rbind(kom_summer, houtbay_summer, scar_summer, oude_summer, betty_summer, kalk_summer, aframe_summer, sthelena_summer, millers_summer, baboon_summer, basata_summer, buffels_summer, soetwater_summer)
 
@@
 
-# Autumn
-
-total_autumn <- rbind(kom_autumn, houtbay_autumn, scar_autumn, oude_autumn, betty_autumn, kalk_autumn, aframe_autumn, sthelena_autumn, millers_autumn, baboon_autumn, basata_autumn, buffels_autumn, soetwater_autumn)
-
 autumn_heatmap <- ggplot() +
@@
 
-# Winter
-
-total_winter <- rbind(kom_winter, houtbay_winter, scar_winter, oude_winter, betty_winter, kalk_winter, aframe_winter, sthelena_winter, millers_winter, baboon_winter, basata_winter, buffels_winter, soetwater_winter)
-
 autumn_heatmap <- ggplot() +
@@
         axis.title=element_text(size=12))
-# Spring
-
-total_spring <- rbind(kom_spring, houtbay_spring, scar_spring, oude_spring, betty_spring, kalk_spring, aframe_spring, sthelena_spring, millers_spring, baboon_spring, basata_spring, buffels_spring, soetwater_spring)
 
@@
 ```
+
+# Point plots
+
+```{r message=FALSE, echo=FALSE, warning=FALSE}
+# summer
+
+total_summer_surface <- total_summer %>% 
+filter(KelpDepth == 0)
+total_summer_sink <- total_summer %>% 
+filter(KelpDepth > 0)
+
+summer_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_summer_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Summer") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# autumn
+
+total_autumn_surface <- total_autumn %>% 
+filter(KelpDepth == 0)
+total_autumn_sink <- total_autumn %>% 
+filter(KelpDepth > 0)
+
+autumn_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_autumn_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Autumn") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+# winter
+
+total_winter_surface <- total_winter %>% 
+filter(KelpDepth == 0)
+total_winter_sink <- total_winter %>% 
+filter(KelpDepth > 0)
+
+winter_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_winter_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Winter") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# spring
+
+total_spring_surface <- total_spring %>% 
+filter(KelpDepth == 0)
+total_spring_sink <- total_spring %>% 
+filter(KelpDepth > 0)
+
+spring_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_spring_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(1, 500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Spring") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  #theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE}
+
+ggarrange(summer_depth_sink, autumn_depth_sink, winter_depth_sink, spring_depth_sink,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 2)
+
+```

### 2021-08-05 thesis_chapter_4.Rmd
updated chapter 4 heatmaps

--- 2021-08-04 thesis_chapter_4.Rmd
@@
 output:
+  html_document: default
   word_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 
-```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+```{r data wrangle, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
+total_summer_surface <- total_summer %>% 
+filter(KelpDepth == 0)
+total_summer_sink <- total_summer %>% 
+filter(KelpDepth > 0)
+total_autumn_surface <- total_autumn %>% 
+filter(KelpDepth == 0)
+total_autumn_sink <- total_autumn %>% 
+filter(KelpDepth > 0)
+total_winter_surface <- total_winter %>% 
+filter(KelpDepth == 0)
+total_winter_sink <- total_winter %>% 
+filter(KelpDepth > 0)
+total_spring_surface <- total_spring %>% 
+filter(KelpDepth == 0)
+total_spring_sink <- total_spring %>% 
+filter(KelpDepth > 0)
+
 ## Load software need for entire script
@@
 
-# Summer
-total_summer <- rbind(kom_summer, houtbay_summer, scar_summer, oude_summer, betty_summer, kalk_summer, aframe_summer, sthelena_summer, millers_summer, baboon_summer, basata_summer, buffels_summer, soetwater_summer)
-
-summer_heatmap <- ggplot() +
+# summer
+summer_heatmap_surface <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_hex(data = total_summer, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  geom_hex(data = total_summer_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 100000)) +
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("1") +
+  ggtitle("Surface") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-# Autumn
-
-total_autumn <- rbind(kom_autumn, houtbay_autumn, scar_autumn, oude_autumn, betty_autumn, kalk_autumn, aframe_autumn, sthelena_autumn, millers_autumn, baboon_autumn, basata_autumn, buffels_autumn, soetwater_autumn)
-
-autumn_heatmap <- ggplot() +
+summer_heatmap_sink <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_hex(data = total_autumn, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  geom_hex(data = total_summer_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis", limits = c(1, 100000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Sinking") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+# autumn
+autumn_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_autumn_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("1") +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-# Winter
-
-total_winter <- rbind(kom_winter, houtbay_winter, scar_winter, oude_winter, betty_winter, kalk_winter, aframe_winter, sthelena_winter, millers_winter, baboon_winter, basata_winter, buffels_winter, soetwater_winter)
-
-autumn_heatmap <- ggplot() +
+autumn_heatmap_sink <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_hex(data = total_winter, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  geom_hex(data = total_autumn_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# winter
+winter_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_winter_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
@@
         axis.title=element_text(size=12))
-# Spring
-
-total_spring <- rbind(kom_spring, houtbay_spring, scar_spring, oude_spring, betty_spring, kalk_spring, aframe_spring, sthelena_spring, millers_spring, baboon_spring, basata_spring, buffels_spring, soetwater_spring)
-
-autumn_heatmap <- ggplot() +
+
+winter_heatmap_sink <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_hex(data = total_spring, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  geom_hex(data = total_winter_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# spring
+spring_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
   scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
@@
         axis.title=element_text(size=12))
+
+spring_heatmap_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
 ```
+
+```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=8, fig.width=6}
+
+ggarrange(summer_heatmap_surface, summer_heatmap_sink, autumn_heatmap_surface, autumn_heatmap_sink, winter_heatmap_surface, winter_heatmap_sink, autumn_heatmap_surface, spring_heatmap_sink,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 4)
+
+```
+
+# Point plots
+## Depth and surface distribution
+```{r message=FALSE, echo=FALSE, warning=FALSE}
+# summer
+summer_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_summer_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Summer") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+# autumn
+autumn_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_autumn_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Autumn") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+# winter
+winter_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_winter_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Winter") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+# spring
+spring_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_spring_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(1, 500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Spring") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  #theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE}
+
+ggarrange(summer_depth_sink, autumn_depth_sink, winter_depth_sink, spring_depth_sink,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 2)
+
+```

### 2021-08-05 thesis_chapter_4.Rmd
final additions for the day

--- 2021-08-04 thesis_chapter_4.Rmd
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-
-ggarrange(summer_mean, autumn_mean, winter_mean, spring_mean,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 2)
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=6, fig.width=6}
+
+ggarrange(summer_mean, autumn_mean, winter_mean, spring_mean,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 2, labels = c("AUTO"))
 
@@
 ```{r message=FALSE, warning=FALSE, include=FALSE}
+
+total_summer_surface <- total_summer %>% 
+filter(KelpDepth == 0)
+total_summer_sink <- total_summer %>% 
+filter(KelpDepth > 0)
+total_autumn_surface <- total_autumn %>% 
+filter(KelpDepth == 0)
+total_autumn_sink <- total_autumn %>% 
+filter(KelpDepth > 0)
+total_winter_surface <- total_winter %>% 
+filter(KelpDepth == 0)
+total_winter_sink <- total_winter %>% 
+filter(KelpDepth > 0)
+total_spring_surface <- total_spring %>% 
+filter(KelpDepth == 0)
+total_spring_sink <- total_spring %>% 
+filter(KelpDepth > 0)
 
@@
 
-# Summer
-total_summer <- rbind(kom_summer, houtbay_summer, scar_summer, oude_summer, betty_summer, kalk_summer, aframe_summer, sthelena_summer, millers_summer, baboon_summer, basata_summer, buffels_summer, soetwater_summer)
-
-summer_heatmap <- ggplot() +
+# summer
+summer_heatmap_surface <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_hex(data = total_summer, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis", limits = c(1, 100000)) +
+  geom_hex(data = total_summer_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C", limits = c(1, 110000)) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
 
-# Autumn
-
-total_autumn <- rbind(kom_autumn, houtbay_autumn, scar_autumn, oude_autumn, betty_autumn, kalk_autumn, aframe_autumn, sthelena_autumn, millers_autumn, baboon_autumn, basata_autumn, buffels_autumn, soetwater_autumn)
-
-autumn_heatmap <- ggplot() +
+summer_heatmap_sink <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_hex(data = total_autumn, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
+  geom_hex(data = total_summer_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 1000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Sub-surface") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+# autumn
+autumn_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_autumn_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C", limits = c(1, 110000)) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
 
-# Winter
-
-total_winter <- rbind(kom_winter, houtbay_winter, scar_winter, oude_winter, betty_winter, kalk_winter, aframe_winter, sthelena_winter, millers_winter, baboon_winter, basata_winter, buffels_winter, soetwater_winter)
-
-autumn_heatmap <- ggplot() +
+autumn_heatmap_sink <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_hex(data = total_winter, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
+  geom_hex(data = total_autumn_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 1000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# winter
+winter_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_winter_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C", limits = c(1, 110000)) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
         axis.title=element_text(size=12))
-# Spring
-
-total_spring <- rbind(kom_spring, houtbay_spring, scar_spring, oude_spring, betty_spring, kalk_spring, aframe_spring, sthelena_spring, millers_spring, baboon_spring, basata_spring, buffels_spring, soetwater_spring)
-
-autumn_heatmap <- ggplot() +
+
+winter_heatmap_sink <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_hex(data = total_spring, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis", limits = c(1, 110000)) +
+  geom_hex(data = total_winter_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 1000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# spring
+spring_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C", limits = c(1, 11000)) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
         axis.title=element_text(size=12))
+
+spring_heatmap_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 1000)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
 ```
+
+```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=10, fig.width=6}
+leg_heatmap_surface <- get_legend(summer_heatmap_surface)
+leg_heatmap_sink <- get_legend(summer_heatmap_sink)
+
+ggarrange(summer_heatmap_surface, summer_heatmap_sink, autumn_heatmap_surface, autumn_heatmap_sink, winter_heatmap_surface, winter_heatmap_sink, autumn_heatmap_surface, spring_heatmap_sink,as_ggplot(leg_heatmap_surface), as_ggplot(leg_heatmap_sink), font.label = list(size = 12, face = "bold"), legend = c("none"), ncol = 2, nrow = 5, labels = c("AUTO"))
+
+```
+
+# Point plots
+## Depth distribution
+```{r message=FALSE, echo=FALSE, warning=FALSE}
+# summer
+summer_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_summer_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Summer") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+# autumn
+autumn_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_autumn_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Autumn") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+# winter
+winter_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_winter_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Winter") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+# spring
+spring_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_spring_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
+  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(1, 500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Spring") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  #theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.height=6, fig.width=6}
+
+ggarrange(summer_depth_sink, autumn_depth_sink, winter_depth_sink, spring_depth_sink,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 2)
+
+```

### 2021-08-06 thesis_chapter_4.Rmd
update to rmd file, age plots included

@@
 total_summer <- rbind(kom_summer, houtbay_summer, scar_summer, oude_summer, betty_summer, kalk_summer, aframe_summer, sthelena_summer, millers_summer, baboon_summer, basata_summer, buffels_summer, soetwater_summer)
+total_summer$age <- with(total_summer, age/3600)
 
@@
 total_autumn <- rbind(kom_autumn, houtbay_autumn, scar_autumn, oude_autumn, betty_autumn, kalk_autumn, aframe_autumn, sthelena_autumn, millers_autumn, baboon_autumn, basata_autumn, buffels_autumn, soetwater_autumn)
+total_autumn$age <- with(total_autumn, age/3600)
 
@@
 total_winter <- rbind(kom_winter, houtbay_winter, scar_winter, oude_winter, betty_winter, kalk_winter, aframe_winter, sthelena_winter, millers_winter, baboon_winter, basata_winter, buffels_winter, soetwater_winter)
+total_winter$age <- with(total_winter, age/3600)
 
@@
 total_spring <- rbind(kom_spring, houtbay_spring, scar_spring, oude_spring, betty_spring, kalk_spring, aframe_spring, sthelena_spring, millers_spring, baboon_spring, basata_spring, buffels_spring, soetwater_spring)
+total_spring$age <- with(total_spring, age/3600)
 
@@
 
-```
-
-## Mean trajectory
-
-```{r message=FALSE, warning=FALSE, include=FALSE}
-
 load("map/south_africa_coast_hi_res.Rdata")
 load("map/sa_bathy.RData")
+
+```
+
+## Mean trajectory
+
+```{r message=FALSE, warning=FALSE, include=FALSE}
 
@@
 ```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=FALSE}
+
+# summer
+summer_age <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_summer, aes(x = lon, y = lat, colour = age)) +
+  scale_colour_viridis_b(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Summer") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+# autumn
+autumn_age <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_autumn, aes(x = lon, y = lat, colour = age)) +
+  scale_colour_viridis_b(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Autumn") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+# winter
+winter_age <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_winter, aes(x = lon, y = lat, colour = age)) +
+  scale_colour_viridis_b(aesthetics = "colour", option = "C", limits = c(0, 1300)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Winter") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+# spring
+spring_age <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_spring, aes(x = lon, y = lat, colour = age)) +
+  scale_colour_viridis_b(aesthetics = "colour", option = "C", limits = c(1, 1500)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Spring") +
+  #theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.height=6, fig.width=6}
+
+ggarrange(summer_age, autumn_age, winter_age, spring_age,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 2)
+
+```

### 2021-08-06 chapter_4_test_script.R
test script adjusted

@@ -24,7 +24,7 @@ options(scipen = 999)
 deVoss_filepath <- paste0("OceanParcels/coppin_extract_01.nc")
 deVoss_output <- nc_open(deVoss_filepath)
 
-coper_filepath <- paste0("OceanParcels/global-analysis-forecast-phy-001-024_1623936804092.nc")
+coper_filepath <- paste0("OceanParcels/chapter_4_oceanmodel/copernicus_phy001_030/global-reanalysis-phy-001-030-daily_1625764682025.nc")
 coper_output <- nc_open(coper_filepath)
 
 test <- tidync("OceanParcels/test.nc") %>%
@@ -110,9 +110,8 @@ soetwater_spring_mean <- soetwater_spring %>% # calculate mean trajectory
 soetwater_spring_mean$season <- "Spring"
 soetwater_spring_mean$site <- "Soetwater"
 
-
-
-
+total_summer_end <- total_summer %>%
+  filter(t =="2019-02-28 00:00:00")
 
 
 

### 2021-08-06 thesis_chapter_3_vectormaps.R
added additional associated files. error in naming corrected for chp3 file

@@ -1,0 +1,301 @@
+# intialise packages from main script
+
+ocean_vector <- tidync("OceanParcels/global-analysis-forecast-phy-001-024_1603882856077.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, longitude, latitude, uo, vo) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2018-01-01 00:00:00"),
+                     "GMT",
+                     "%Y-%m-%d"))
+
+ocean_jan <-  ocean_vector %>%
+  filter(t >= ("2086-01-01 00:00:00") & t <= ("2086-01-30 00:00:00"))
+ocean_apr <-  ocean_vector %>%
+  filter(t >= ("2086-04-01 00:00:00") & t <= ("2086-04-30 00:00:00"))
+ocean_july <-  ocean_vector %>%
+  filter(t >= ("2086-07-01 00:00:00") & t <= ("2086-07-30 00:00:00"))
+ocean_oct <-  ocean_vector %>%
+  filter(t >= ("2086-10-01 00:00:00") & t <= ("2086-10-30 00:00:00"))
+
+ocean_jan$velocity <- with(ocean_jan, sqrt(uo^2 + vo^2))
+ocean_apr$velocity <- with(ocean_apr, sqrt(uo^2 + vo^2))
+ocean_july$velocity <- with(ocean_july, sqrt(uo^2 + vo^2))
+ocean_oct$velocity <- with(ocean_oct, sqrt(uo^2 + vo^2))
+
+ocean_jan_summary <- ocean_jan %>%
+  group_by(longitude, latitude) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
+ocean_apr_summary <- ocean_apr %>%
+  group_by(longitude, latitude) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
+ocean_july_summary <- ocean_july %>%
+  group_by(longitude, latitude) %>%
+  summarise(uo = median(uo, na.rm = TRUE),
+            vo = median(vo, na.rm = TRUE),
+            velocity = median(velocity, na.rm = TRUE))
+
[…]

### 2021-08-06 thesis_chapter_3_vectormap.R
deleted un-needed files

@@ -1,301 +1,0 @@
-# intialise packages from main script
-
-ocean_vector <- tidync("OceanParcels/global-analysis-forecast-phy-001-024_1603882856077.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, longitude, latitude, uo, vo) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2018-01-01 00:00:00"),
-                     "GMT",
-                     "%Y-%m-%d"))
-
-ocean_jan <-  ocean_vector %>%
-  filter(t >= ("2086-01-01 00:00:00") & t <= ("2086-01-30 00:00:00"))
-ocean_apr <-  ocean_vector %>%
-  filter(t >= ("2086-04-01 00:00:00") & t <= ("2086-04-30 00:00:00"))
-ocean_july <-  ocean_vector %>%
-  filter(t >= ("2086-07-01 00:00:00") & t <= ("2086-07-30 00:00:00"))
-ocean_oct <-  ocean_vector %>%
-  filter(t >= ("2086-10-01 00:00:00") & t <= ("2086-10-30 00:00:00"))
-
-ocean_jan$velocity <- with(ocean_jan, sqrt(uo^2 + vo^2))
-ocean_apr$velocity <- with(ocean_apr, sqrt(uo^2 + vo^2))
-ocean_july$velocity <- with(ocean_july, sqrt(uo^2 + vo^2))
-ocean_oct$velocity <- with(ocean_oct, sqrt(uo^2 + vo^2))
-
-ocean_jan_summary <- ocean_jan %>%
-  group_by(longitude, latitude) %>%
-  summarise(uo = median(uo, na.rm = TRUE),
-            vo = median(vo, na.rm = TRUE),
-            velocity = median(velocity, na.rm = TRUE))
-
-ocean_apr_summary <- ocean_apr %>%
-  group_by(longitude, latitude) %>%
-  summarise(uo = median(uo, na.rm = TRUE),
-            vo = median(vo, na.rm = TRUE),
-            velocity = median(velocity, na.rm = TRUE))
-
-ocean_july_summary <- ocean_july %>%
-  group_by(longitude, latitude) %>%
-  summarise(uo = median(uo, na.rm = TRUE),
-            vo = median(vo, na.rm = TRUE),
-            velocity = median(velocity, na.rm = TRUE))
-
[…]

### 2021-08-06 thesis_chapter_4.Rmd
slight adjustments

@@
 # Results
-
 ## Ocean analysis
@@
 
-ocean_vector_chp4 <- tidync("OceanParcels/chapter_4_oceanmodel/copernicus_phy001_030/global-reanalysis-phy-001-030-daily_1625758460342.nc") %>%
+ocean_summer <- tidync("OceanParcels/chapter_4_oceanmodel/copernicus_phy001_030/global-reanalysis-phy-001-030-daily_1625758460342.nc") %>%
   hyper_tibble() %>%
@@
                      "%Y-%m-%d-%h"))
-
-ocean_summer <-  ocean_vector_chp4 %>%
-  filter(t >= ("2086-01-01 00:00:00") & t <= ("2086-02-28 00:00:00"))
-ocean_autumn <-  ocean_vector_chp4 %>%
-  filter(t >= ("2086-03-01 00:00:00") & t <= ("2086-05-31 00:00:00"))
-ocean_winter <-  ocean_vector_chp4 %>%
-  filter(t >= ("2086-06-01 00:00:00") & t <= ("2086-08-30 00:00:00"))
-ocean_spring <-  ocean_vector_chp4 %>%
-  filter(t >= ("2086-09-01 00:00:00") & t <= ("2086-11-30 00:00:00"))
 

### 2021-08-06 thesis_chapter_4.Rmd
chapter4 and vectormap updated

@@
 ```{r eval=FALSE, message=FALSE, echo=FALSE, include=FALSE}
-
-ocean_summer <- tidync("OceanParcels/chapter_4_oceanmodel/copernicus_phy001_030/global-reanalysis-phy-001-030-daily_1625758460342.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, longitude, latitude, uo, vo) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2019-01-01 00:00:00"),
-                     "GMT",
-                     "%Y-%m-%d-%h"))
-
-ocean_summer$velocity <- with(ocean_jan, sqrt(uo^2 + vo^2))
-ocean_autumn$velocity <- with(ocean_apr, sqrt(uo^2 + vo^2))
-ocean_winter$velocity <- with(ocean_july, sqrt(uo^2 + vo^2))
-ocean_spring$velocity <- with(ocean_oct, sqrt(uo^2 + vo^2))
 

### 2021-08-06 thesis_chapter_4_vectormaps.R

@@ -1,0 +1,132 @@
+load("map/south_africa_coast_hi_res.Rdata")
+#ggsave("fig_3A.jpeg", plot = temp_plot, dpi = 300, device = "jpeg" , path = "figures/")
+
+ocean_summer_2018 <- tidync("OceanParcels/chapter_4_oceanmodel/copernicus_phy001_030/global-reanalysis-phy-001-030-daily_1625758460342.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, longitude, latitude, uo, vo) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2019-01-01 00:00:00"),
+                     "GMT",
+                     "%Y-%m-%d-%h"))
+
+ocean_autumn_2018 <- tidync("OceanParcels/chapter_4_oceanmodel/copernicus_phy001_030/global-reanalysis-phy-001-030-daily_1625759550563.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, longitude, latitude, uo, vo) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2019-01-01 00:00:00"),
+                     "GMT",
+                     "%Y-%m-%d-%h"))
+
+ocean_winter_2018 <- tidync("OceanParcels/chapter_4_oceanmodel/copernicus_phy001_030/global-reanalysis-phy-001-030-daily_1625760938909.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, longitude, latitude, uo, vo) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2019-01-01 00:00:00"),
+                     "GMT",
+                     "%Y-%m-%d-%h"))
+
+ocean_spring_2018 <- tidync("OceanParcels/chapter_4_oceanmodel/copernicus_phy001_030/global-reanalysis-phy-001-030-daily_1625761875172.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, longitude, latitude, uo, vo) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.Date(as.POSIXct(t * 3600, origin = "2019-01-01 00:00:00"),
+                     "GMT",
+                     "%Y-%m-%d-%h"))
+
[…]

### 2021-08-07 thesis_chapter_4.Rmd
added initial vector maps

@@
 library(ggpubr)
+library(jpeg)
+library(gridExtra)
 
@@
 ```{r eval=FALSE, message=FALSE, echo=FALSE, include=FALSE}
+
+map_chp4A <- rasterGrob(as.raster(readJPEG("figures/chapter_4/chapter_4_vectormap_summer.jpeg")))
+map_chp4B <- rasterGrob(as.raster(readJPEG("figures/chapter_4/chapter_4_vectormap_autumn.jpeg")))
+map_chp4C <- rasterGrob(as.raster(readJPEG("figures/chapter_4/chapter_4_vectormap_winter.jpeg")))
+map_chp4D <- rasterGrob(as.raster(readJPEG("figures/chapter_4/chapter_4_vectormap_spring.jpeg")))
+
+ggarrange(map_chp4A, map_chp4B, map_chp4C, map_chp4D)
 

### 2021-10-31 thesis_chapter_3.Rmd
new chapter 3 files

@@
 output:
-  word_document:
-    reference_docx: markdown/wiley_template.docx
+  html_document:
   pdf_document:
@@
     latex_engine: xelatex
-  html_document:
+  word_document:
+    reference_docx: markdown/wiley_template.docx
 bibliography: references.bib
@@
 
-library(ggplot2)
-#library(rgdal)
+library(tidyverse)
+library(flextable)
+library(tidync)
+library(ncdf4)
+library(lubridate)
 library(gridExtra)
 library(ggpubr)
-library(mapproj)
-library(vegan)
-library(dplyr)
-library(ggpubr)
-library(maps)
-library(flextable)
-library(officer)
-library(tidync)
-library(tidyverse)
-library(hexbin)
-library(lubridate)
+library(jpeg)
+library(gridExtra)
+library(grid)
 
@@
 
-A range of natural and anthropogenic objects floats in the ocean, and macroalgae are regarded as one of the most important of these passive dispersal mechanisms of marine taxa.
-Floating kelps act as passive dispersal mechanisms for a range of marine taxa and is sometimes referred to as the ‘tumble-weed of the ocean’ [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
-Kelps also function as input of allochthonous materials to near and distant ecosystems and in this manner support the ecological function in a range of adjacent habitats [@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].
+Passive dispersal mechanisms of marine taxa include a range of natural and anthropogenic objects that float in the ocean, and macroalgae are regarded as one of the most important of these.
+Floating kelps, sometimes referred to as the ‘tumble-weed of the ocean,’ are very important in this regard [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
+Kelps also function as allochthonous materials to near and distant ecosystems and in this manner support the ecological function in a range of adjacent habitats[@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].
 
 Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.
-The positively buoyant pneumatocysts also cause plants to float to the surface when dislodged from the substratum.
-The structure and number of pneumatocysts varies among species.
-For example, species belonging to the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, and *Ecklonia maxima* have single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018].
+The positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum.
+The structure and number of pneumatocysts vary among species.
+For example, species belonging to the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, and *Ecklonia maxima* have single, large pneumatocysts[@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018].
 In some cases, the stipe itself may be air-filled, such as with *E.
@@
 This is most likely because the green and red species reported to float are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in their thalli [@dromgoole1982; @back2000].
-The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @putman2018; @brooks2018; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
+The giant kelp *Macrocystis pyrifera* [@hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@tala2017; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @brooks2018; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
 
 Past research points to macroalgal trajectories being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
-Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown although the role of wind has been recognized as important in several studies.
+Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown, although the role of wind has been recognized as important in several studies.
 For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M.
 pyrifera* in nearshore Monterey Bay.
-The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents; however, the relative importance of wind and was not clear.
-In addition, the trajectories of tagged kelp were more consistent with the formation of eddys during winter.
+The results showed that kelp rafts, with little surface area exposed to the wind, were largely driven by a combination of wind and surface currents; however, the relative importance of wind and was not clear.
+In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter.
 Studies have also identified wind as an important mechanism of dispersal in wind dominated ocean systems.
-For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating bull kelp *Durvillaea antarctica*.
+For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important mechanism for dispersal of floating bull kelp *Durvillaea antarctica* [@waters2008; @smith2002; @nikula2010].
 Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013].
@@
 #ggarrange(map_chp3A, map_chp3B, ncol = 2, nrow = 1, labels = "AUTO", align = c("h"))
-map_chp3_combined <- rasterGrob(as.raster(readPNG("map/chp3_map_final.png")))
+map_chp3_combined <- rasterGrob(as.raster(readPNG("map/chp3_map_final_v2.png")))
 ggarrange(map_chp3_combined)
@@
 ```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Diagram of the *E.
-maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic."}
+maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic.
+Schemetaic was adapted from @Dyer2018-rv."}
 
@@
 
-```{r data wrangling, message=FALSE, warning=FALSE, include=FALSE}
-
-# Read in simulation data and data wrangling
-
-p_RK4 <- tidync("OceanParcels/trajectory_data/RK4_PFloat.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-p_RK4_mean <- p_RK4 %>% # calculate mean trajectory
-  group_by(t) %>% 
+```{r data readin, message=FALSE, warning=FALSE, include=FALSE}
+
+total_summer <- read.csv(file = "data/chapter_3/total_summer.csv")
+total_summer_mean <- read.csv(file = "data/chapter_3/total_summer_mean.csv")
+
+total_autumn <- read.csv(file = "data/chapter_3/total_autumn.csv")
+total_autumn_mean <- read.csv(file = "data/chapter_3/total_autumn_mean.csv")
+
+total_winter <- read.csv(file = "data/chapter_3/total_winter.csv")
+total_winter_mean <- read.csv(file = "data/chapter_3/total_winter_mean.csv")
+
+total_spring <- read.csv(file = "data/chapter_3/total_spring.csv")
+total_spring_mean <- read.csv(file = "data/chapter_3/total_spring_mean.csv")
+
+total_passive_summer <- read.csv(file = "data/chapter_3/total_passive_summer.csv")
+total_passive_summer_mean <- total_passive_summer %>% 
+  group_by(t, type) %>%
   summarise(across(lon:lat, mean, na.rm= TRUE))
-
-p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
-p_RK4_mean$type <- "1" # add extra column 'type' to dataframe and assigned 'RK4' label
-
-Min_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H100_W00.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H100_W00_mean <- Min_Kfloat_H100_W00 %>% # calculate mean trajectory
-  group_by(t) %>% 
+total_passive_autumn <- read.csv(file = "data/chapter_3/total_passive_autumn.csv")
+total_passive_autumn_mean <- total_passive_autumn %>% 
+  group_by(t, type) %>%
   summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H100_W00$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H100_W00$velocity <- with(Min_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H100_W00_mean$type <- "2" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H095_W05.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H095_W05_mean <- Min_Kfloat_H095_W05 %>% # calculate mean trajectory
-  group_by(t) %>% 
+total_passive_winter <- read.csv(file = "data/chapter_3/total_passive_winter.csv")
+total_passive_winter_mean <- total_passive_winter %>% 
+  group_by(t, type) %>%
   summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H095_W05$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H095_W05$velocity <- with(Min_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H095_W05_mean$type <- "3" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H090_W10.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H090_W10_mean <- Min_Kfloat_H090_W10 %>% #calculate mean trajectory 
-  group_by(t) %>% 
+total_passive_spring <- read.csv(file = "data/chapter_3/total_passive_spring.csv")
+total_passive_spring_mean <- total_passive_spring %>% 
+  group_by(t, type) %>%
   summarise(across(lon:lat, mean, na.rm= TRUE))
 
-Min_Kfloat_H090_W10$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H090_W10$velocity <- with(Min_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H090_W10_mean$type <- "4" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H085_W15.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H085_W15_mean <- Min_Kfloat_H085_W15 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H085_W15$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H085_W15$velocity <- with(Min_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H085_W15_mean$type <- "5" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H100_W00.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_Kfloat_H100_W00_mean <- Mean_Kfloat_H100_W00 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_Kfloat_H100_W00$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_Kfloat_H100_W00$velocity <- with(Mean_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_Kfloat_H100_W00_mean$type <- "6" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H095_W05.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H095_W05_mean <- Mean_KFloat_H095_W05 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H095_W05$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H095_W05$velocity <- with(Mean_KFloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H095_W05_mean$type <- "7" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H090_W10.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H090_W10_mean <- Mean_KFloat_H090_W10 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H090_W10$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H090_W10$velocity <- with(Mean_KFloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H090_W10_mean$type <- "8" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H085_W15.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H085_W15_mean <- Mean_KFloat_H085_W15 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H085_W15$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H085_W15$velocity <- with(Mean_KFloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H085_W15_mean$type <- "9" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H100_W00.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H100_W00_mean <- Max_Kfloat_H100_W00 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H100_W00$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H100_W00$velocity <- with(Max_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H100_W00_mean$type <- "10" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H095_W05.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H095_W05_mean <- Max_Kfloat_H095_W05 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H095_W05$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H095_W05$velocity <- with(Max_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H095_W05_mean$type <- "11" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H090_W10.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H090_W10_mean <- Max_Kfloat_H090_W10 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H090_W10$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H090_W10$velocity <- with(Max_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H090_W10_mean$type <- "12" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H085_W15.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H085_W15_mean <- Max_Kfloat_H085_W15 %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H085_W15$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H085_W15$velocity <- with(Max_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H085_W15_mean$type <- "13" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-total_data <- rbind(Min_Kfloat_H100_W00 , Min_Kfloat_H095_W05, Min_Kfloat_H090_W10, Min_Kfloat_H085_W15, Mean_Kfloat_H100_W00, Mean_KFloat_H095_W05, Mean_KFloat_H090_W10, Mean_KFloat_H085_W15, Max_Kfloat_H100_W00, Max_Kfloat_H095_W05, Max_Kfloat_H090_W10, Max_Kfloat_H085_W15)
-
-## April
-
-p_RK4_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-p_RK4_mean_April <- p_RK4_April %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-p_RK4_April$type <- "RK4_April" # add extra column 'type' to dataframe and assigned 'RK4' label
-p_RK4_mean_April$type <- "14" # add extra column 'type' to dataframe and assigned 'RK4' label
-
-Min_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H100_W00_mean_April <- Min_Kfloat_H100_W00_April %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H100_W00_April$type <- "Min_Kfloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H100_W00_April$velocity <- with(Min_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H100_W00_mean_April$type <- "15" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H095_W05_mean_April <- Min_Kfloat_H095_W05_April %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H095_W05_April$type <- "Min_Kfloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H095_W05_April$velocity <- with(Min_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H095_W05_mean_April$type <- "16" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H090_W10_mean_April <- Min_Kfloat_H090_W10_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H090_W10_April$type <- "Min_Kfloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H090_W10_April$velocity <- with(Min_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H090_W10_mean_April$type <- "17" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H085_W15_mean_April <- Min_Kfloat_H085_W15_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H085_W15_April$type <- "Min_Kfloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H085_W15_April$velocity <- with(Min_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H085_W15_mean_April$type <- "18" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_Kfloat_H100_W00_mean_April <- Mean_Kfloat_H100_W00_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_Kfloat_H100_W00_April$type <- "Mean_Kfloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_Kfloat_H100_W00_April$velocity <- with(Mean_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_Kfloat_H100_W00_mean_April$type <- "19" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H095_W05_mean_April <- Mean_KFloat_H095_W05_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H095_W05_April$type <- "Mean_KFloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H095_W05_April$velocity <- with(Mean_KFloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H095_W05_mean_April$type <- "20" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H090_W10_mean_April <- Mean_KFloat_H090_W10_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H090_W10_April$type <- "Mean_KFloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H090_W10_April$velocity <- with(Mean_KFloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H090_W10_mean_April$type <- "21" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H085_W15_mean_April <- Mean_KFloat_H085_W15_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H085_W15_April$type <- "Mean_KFloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H085_W15_April$velocity <- with(Mean_KFloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H085_W15_mean_April$type <- "22" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H100_W00_mean_April <- Max_Kfloat_H100_W00_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H100_W00_April$type <- "Max_KFloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H100_W00_April$velocity <- with(Max_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H100_W00_mean_April$type <- "23" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H095_W05_mean_April <- Max_Kfloat_H095_W05_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H095_W05_April$type <- "Max_KFloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H095_W05_April$velocity <- with(Max_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H095_W05_mean_April$type <- "24" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H090_W10_mean_April <- Max_Kfloat_H090_W10_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H090_W10_April$type <- "Max_KFloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H090_W10_April$velocity <- with(Max_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H090_W10_mean_April$type <- "25" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H085_W15_mean_April <- Max_Kfloat_H085_W15_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H085_W15_April$type <- "Max_KFloat_H085_W15_April" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H085_W15_April$velocity <- with(Max_Kfloat_H085_W15_April, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H085_W15_mean_April$type <- "26" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-total_data_April <- rbind(Min_Kfloat_H100_W00_April, Min_Kfloat_H095_W05_April, Min_Kfloat_H090_W10_April, Min_Kfloat_H085_W15_April, Mean_Kfloat_H100_W00_April, Mean_KFloat_H095_W05_April, Mean_KFloat_H090_W10_April, Mean_KFloat_H085_W15_April, Max_Kfloat_H100_W00_April, Max_Kfloat_H095_W05_April, Max_Kfloat_H090_W10_April, Max_Kfloat_H085_W15_April)
-
-## July
-
-p_RK4_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-p_RK4_mean_July <- p_RK4_July %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-p_RK4_July$type <- "RK4_July" # add extra column 'type' to dataframe and assigned 'RK4' label
-p_RK4_mean_July$type <- "27" # add extra column 'type' to dataframe and assigned 'RK4' label
-
-Min_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H100_W00_mean_July <- Min_Kfloat_H100_W00_July %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H100_W00_July$type <- "Min_Kfloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H100_W00_July$velocity <- with(Min_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H100_W00_mean_July$type <- "28" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H095_W05_mean_July <- Min_Kfloat_H095_W05_July %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H095_W05_July$type <- "Min_Kfloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H095_W05_July$velocity <- with(Min_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H095_W05_mean_July$type <- "29" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H090_W10_mean_July <- Min_Kfloat_H090_W10_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H090_W10_July$type <- "Min_Kfloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H090_W10_July$velocity <- with(Min_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H090_W10_mean_July$type <- "30" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H085_W15_mean_July <- Min_Kfloat_H085_W15_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H085_W15_July$type <- "Min_Kfloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H085_W15_July$velocity <- with(Min_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H085_W15_mean_July$type <- "31" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_Kfloat_H100_W00_mean_July <- Mean_Kfloat_H100_W00_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_Kfloat_H100_W00_July$type <- "Mean_Kfloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_Kfloat_H100_W00_July$velocity <- with(Mean_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_Kfloat_H100_W00_mean_July$type <- "32" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H095_W05_mean_July <- Mean_KFloat_H095_W05_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H095_W05_July$type <- "Mean_KFloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H095_W05_July$velocity <- with(Mean_KFloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H095_W05_mean_July$type <- "33" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H090_W10_mean_July <- Mean_KFloat_H090_W10_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H090_W10_July$type <- "Mean_KFloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H090_W10_July$velocity <- with(Mean_KFloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H090_W10_mean_July$type <- "34" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H085_W15_mean_July <- Mean_KFloat_H085_W15_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H085_W15_July$type <- "Mean_KFloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H085_W15_July$velocity <- with(Mean_KFloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H085_W15_mean_July$type <- "35" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H100_W00_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H100_W00_mean_July <- Max_Kfloat_H100_W00_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H100_W00_July$type <- "Max_KFloat_H100_W00_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H100_W00_July$velocity <- with(Max_Kfloat_H100_W00_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H100_W00_mean_July$type <- "36" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H095_W05_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H095_W05_mean_July <- Max_Kfloat_H095_W05_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H095_W05_July$type <- "Max_KFloat_H095_W05_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H095_W05_July$velocity <- with(Max_Kfloat_H095_W05_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H095_W05_mean_July$type <- "37" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H090_W10_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H090_W10_mean_July <- Max_Kfloat_H090_W10_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H090_W10_July$type <- "Max_KFloat_H090_W10_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H090_W10_July$velocity <- with(Max_Kfloat_H090_W10_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H090_W10_mean_July$type <- "38" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H085_W15_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H085_W15_mean_July <- Max_Kfloat_H085_W15_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H085_W15_July$type <- "Max_KFloat_H085_W15_July" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H085_W15_July$velocity <- with(Max_Kfloat_H085_W15_July, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H085_W15_mean_July$type <- "39" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-total_data_July <- rbind(Min_Kfloat_H100_W00_July, Min_Kfloat_H095_W05_July, Min_Kfloat_H090_W10_July, Min_Kfloat_H085_W15_July, Mean_Kfloat_H100_W00_July, Mean_KFloat_H095_W05_July, Mean_KFloat_H090_W10_July, Mean_KFloat_H085_W15_July, Max_Kfloat_H100_W00_July, Max_Kfloat_H095_W05_July, Max_Kfloat_H090_W10_July, Max_Kfloat_H085_W15_July)
-
-## October
-
-p_RK4_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-p_RK4_mean_Oct <- p_RK4_Oct %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-p_RK4_Oct$type <- "RK4_Oct" # add extra column 'type' to dataframe and assigned 'RK4' label
-p_RK4_mean_Oct$type <- "40" # add extra column 'type' to dataframe and assigned 'RK4' label
-
-Min_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H100_W00_mean_Oct <- Min_Kfloat_H100_W00_Oct %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H100_W00_Oct$type <- "Min_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H100_W00_Oct$velocity <- with(Min_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H100_W00_mean_Oct$type <- "41" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H095_W05_mean_Oct <- Min_Kfloat_H095_W05_Oct %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H095_W05_Oct$type <- "Min_Kfloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H095_W05_Oct$velocity <- with(Min_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H095_W05_mean_Oct$type <- "42" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H090_W10_mean_Oct <- Min_Kfloat_H090_W10_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H090_W10_Oct$type <- "Min_Kfloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H090_W10_Oct$velocity <- with(Min_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H090_W10_mean_Oct$type <- "43" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Min_Kfloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Min_Kfloat_H085_W15_mean_Oct <- Min_Kfloat_H085_W15_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Min_Kfloat_H085_W15_Oct$type <- "Min_Kfloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Min_Kfloat_H085_W15_Oct$velocity <- with(Min_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Min_Kfloat_H085_W15_mean_Oct$type <- "44" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H100_W00_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_Kfloat_H100_W00_mean_Oct <- Mean_Kfloat_H100_W00_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_Kfloat_H100_W00_Oct$type <- "Mean_Kfloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_Kfloat_H100_W00_Oct$velocity <- with(Mean_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_Kfloat_H100_W00_mean_Oct$type <- "45" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H095_W05_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H095_W05_mean_Oct <- Mean_KFloat_H095_W05_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H095_W05_Oct$type <- "Mean_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H095_W05_Oct$velocity <- with(Mean_KFloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H095_W05_mean_Oct$type <- "46" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H090_W10_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H090_W10_mean_Oct <- Mean_KFloat_H090_W10_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H090_W10_Oct$type <- "Mean_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H090_W10_Oct$velocity <- with(Mean_KFloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H090_W10_mean_Oct$type <- "47" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Mean_KFloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Mean_KFloat_H085_W15_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Mean_KFloat_H085_W15_mean_Oct <- Mean_KFloat_H085_W15_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Mean_KFloat_H085_W15_Oct$type <- "Mean_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Mean_KFloat_H085_W15_Oct$velocity <- with(Mean_KFloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Mean_KFloat_H085_W15_mean_Oct$type <- "48" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H100_W00_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H100_W00_mean_Oct <- Max_Kfloat_H100_W00_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H100_W00_Oct$type <- "Max_KFloat_H100_W00_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H100_W00_Oct$velocity <- with(Max_Kfloat_H100_W00_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H100_W00_mean_Oct$type <- "49" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H095_W05_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H095_W05_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H095_W05_mean_Oct <- Max_Kfloat_H095_W05_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H095_W05_Oct$type <- "Max_KFloat_H095_W05_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H095_W05_Oct$velocity <- with(Max_Kfloat_H095_W05_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H095_W05_mean_Oct$type <- "50" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H090_W10_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H090_W10_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H090_W10_mean_Oct <- Max_Kfloat_H090_W10_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H090_W10_Oct$type <- "Max_KFloat_H090_W10_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H090_W10_Oct$velocity <- with(Max_Kfloat_H090_W10_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H090_W10_mean_Oct$type <- "51" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-Max_Kfloat_H085_W15_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H085_W15_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-Max_Kfloat_H085_W15_mean_Oct <- Max_Kfloat_H085_W15_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-Max_Kfloat_H085_W15_Oct$type <- "Max_KFloat_H085_W15_Oct" # add extra column 'type' to dataframe and assigned 'kelp' label
-Max_Kfloat_H085_W15_Oct$velocity <- with(Max_Kfloat_H085_W15_Oct, sqrt(kelpvelU^2 + kelpvelV^2))
-Max_Kfloat_H085_W15_mean_Oct$type <- "52" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-total_data_Oct <- rbind(Min_Kfloat_H100_W00_Oct, Min_Kfloat_H095_W05_Oct, Min_Kfloat_H090_W10_Oct, Min_Kfloat_H085_W15_Oct, Mean_Kfloat_H100_W00_Oct, Mean_KFloat_H095_W05_Oct, Mean_KFloat_H090_W10_Oct, Mean_KFloat_H085_W15_Oct, Max_Kfloat_H100_W00_Oct, Max_Kfloat_H095_W05_Oct, Max_Kfloat_H090_W10_Oct, Max_Kfloat_H085_W15_Oct)
+total_eddy_hydro <- read.csv(file = "data/chapter_3/total_eddy_hydro.csv")
+total_eddy_hydrostksdrag <- read.csv(file = "data/chapter_3/total_eddy_hydrostksdrag.csv")
+total_eddy_hydrowinddrag <- read.csv(file = "data/chapter_3/total_eddy_hydrowinddrag.csv")
+total_eddy_hydrowindstksdrag <- read.csv(file = "data/chapter_3/total_eddy_hydrowindstksdrag.csv")
+
+total_eddy_passive_curr <- read.csv(file = "data/chapter_3/total_eddy_passive_curr.csv")
+total_eddy_passive_currstks <- read.csv(file = "data/chapter_3/total_eddy_passive_currstks.csv")
+total_eddy_passive_currwind <- read.csv(file = "data/chapter_3/total_eddy_passive_currwind.csv")
+total_eddy_passive_currwindstks <- read.csv(file = "data/chapter_3/total_eddy_passive_currwindstks.csv")
+
+load("map/south_africa_coast_hi_res.Rdata")
+load("map/sa_bathy.RData")
 
@@
 
-total_mean_water <- rbind(p_RK4_mean, Min_Kfloat_H100_W00_mean, Mean_Kfloat_H100_W00_mean, Max_Kfloat_H100_W00_mean)
-total_mean_ww <- rbind(p_RK4_mean, Min_Kfloat_H095_W05_mean, Min_Kfloat_H090_W10_mean, Min_Kfloat_H085_W15_mean, Mean_KFloat_H095_W05_mean, Mean_KFloat_H090_W10_mean, Mean_KFloat_H085_W15_mean, Max_Kfloat_H095_W05_mean, Max_Kfloat_H090_W10_mean, Max_Kfloat_H085_W15_mean)
-
-total_mean_water$type <- factor(total_mean_water$type, levels=c("1", "2", "6", "10"))
-total_mean_ww$type <- factor(total_mean_ww$type, levels=c("1", "3", "4", "5", "7", "8", "9", "11", "12", "13"))
-
-waterdrag_plot <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_mean_water, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(linetype=guide_legend(title="Simulation")) +
+total_traj_summer <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_passive_summer_mean, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
+  geom_path(data = total_summer_mean, aes(x = lon, y = lat, linetype = type))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Passive Particles")) +
+  guides(linetype=guide_legend(title="Kelp Particles")) +
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
@@
 
-wwdrag_plot <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_mean_ww, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(linetype=guide_legend(title="Simulation")) +
+total_traj_autumn <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_passive_autumn_mean, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
+  geom_path(data = total_autumn_mean, aes(x = lon, y = lat, linetype = type))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Passive Particles")) +
+  guides(linetype=guide_legend(title="Kelp Particles")) +
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
@@
 
-total_mean_water_April <- rbind(p_RK4_mean_April, Min_Kfloat_H100_W00_mean_April, Mean_Kfloat_H100_W00_mean_April, Max_Kfloat_H100_W00_mean_April)
-total_mean_ww_April <- rbind(p_RK4_mean_April, Min_Kfloat_H095_W05_mean_April, Min_Kfloat_H090_W10_mean_April, Min_Kfloat_H085_W15_mean_April, Mean_KFloat_H095_W05_mean_April, Mean_KFloat_H090_W10_mean_April, Mean_KFloat_H085_W15_mean_April, Max_Kfloat_H095_W05_mean_April, Max_Kfloat_H090_W10_mean_April, Max_Kfloat_H085_W15_mean_April)
-
-total_mean_water_April$type <- factor(total_mean_water_April$type, levels=c("14", "15", "19", "23"))
-total_mean_ww_April$type <- factor(total_mean_ww_April$type, levels=c("14", "16", "17", "18", "20", "21", "22", "24", "25", "26"))
-
-waterdrag_plot_April <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_mean_water_April, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(linetype=guide_legend(title="Simulation")) +
+total_traj_winter <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_passive_winter_mean, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
+  geom_path(data = total_winter_mean, aes(x = lon, y = lat, linetype = type))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Passive Particles")) +
+  guides(linetype=guide_legend(title="Kelp Particles")) +
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
@@
 
-wwdrag_plot_April <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill ="grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_mean_ww_April, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(linetype=guide_legend(title="Simulation")) +
+total_traj_spring <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_passive_spring_mean, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
+  geom_path(data = total_spring_mean, aes(x = lon, y = lat, linetype = type))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Passive Particles")) +
+  guides(linetype=guide_legend(title="Kelp Particles")) +
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
@@
         axis.title=element_text(size=12)) 
-
-total_mean_water_July <- rbind(p_RK4_mean_July, Min_Kfloat_H100_W00_mean_July, Mean_Kfloat_H100_W00_mean_July, Max_Kfloat_H100_W00_mean_July)
-total_mean_ww_July <- rbind(p_RK4_mean_July, Min_Kfloat_H095_W05_mean_July, Min_Kfloat_H090_W10_mean_July, Min_Kfloat_H085_W15_mean_July, Mean_KFloat_H095_W05_mean_July, Mean_KFloat_H090_W10_mean_July, Mean_KFloat_H085_W15_mean_July, Max_Kfloat_H095_W05_mean_July, Max_Kfloat_H090_W10_mean_July, Max_Kfloat_H085_W15_mean_July)
-
-total_mean_water_July$type <- factor(total_mean_water_July$type, levels=c("27", "28", "32", "36"))
-total_mean_ww_July$type <- factor(total_mean_ww_July$type, levels=c("27", "29", "30", "31", "33", "34", "35", "37", "38", "39"))
-
-waterdrag_plot_July <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_mean_water_July, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-wwdrag_plot_July <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_mean_ww_July, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-total_mean_water_Oct <- rbind(p_RK4_mean_Oct, Min_Kfloat_H100_W00_mean_Oct, Mean_Kfloat_H100_W00_mean_Oct, Max_Kfloat_H100_W00_mean_Oct)
-total_mean_ww_Oct <- rbind(p_RK4_mean_Oct, Min_Kfloat_H095_W05_mean_Oct, Min_Kfloat_H090_W10_mean_Oct, Min_Kfloat_H085_W15_mean_Oct, Mean_KFloat_H095_W05_mean_Oct, Mean_KFloat_H090_W10_mean_Oct, Mean_KFloat_H085_W15_mean_Oct, Max_Kfloat_H095_W05_mean_Oct, Max_Kfloat_H090_W10_mean_Oct, Max_Kfloat_H085_W15_mean_Oct)
-
-total_mean_water_Oct$type <- factor(total_mean_water_Oct$type, levels=c("40", "41", "45", "49"))
-total_mean_ww_Oct$type <- factor(total_mean_ww_Oct$type, levels=c("40", "42", "43", "44", "46", "47", "48", "50", "51", "52"))
-
-waterdrag_plot_Oct <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_mean_water_Oct, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(linetype=guide_legend(title="Simulation")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-wwdrag_plot_Oct <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_mean_ww_Oct, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12)) 
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=6, fig.cap="Figure 3: Comparison of mean trajectories.
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=6, fig.width=6, fig.cap="Figure 3: Comparison of mean trajectories.
 Simulations of hydrodynamic drag are presented in Panels A, C, E, G for January, April, July, and October, respectively.
@@
 
-ggarrange(waterdrag_plot, wwdrag_plot, waterdrag_plot_April, wwdrag_plot_April, waterdrag_plot_July, wwdrag_plot_July, waterdrag_plot_Oct, wwdrag_plot_Oct, ncol = 2, nrow = 4, font.label = list(size = 12, face = "bold"), common.legend = FALSE, legend = c("bottom"), labels = c("A","B","C","D","E","F","G", "H"), align = "hv")
-
-ggsave(filename = "figure_3.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 6, dpi = 300, scale = 1.2)
+ggarrange(total_traj_summer, total_traj_autumn, total_traj_winter, total_traj_spring ,ncol = 2, nrow = 2, font.label = list(size = 12, face = "bold"), common.legend = TRUE, legend = c("bottom"), labels = c("A","B","C","D"), align = "hv")
+
+ggsave(filename = "figure_3.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 10, dpi = 300, scale = 1.5)
 ```
@@
 ## Heatmaps of study domain, use previous data for study site map and histograms
-## Coastline of Southern Africa
-
-load("map/south_africa_coast_hi_res.Rdata")
-load("map/sa_bathy.RData")
-
-heatmap_RK4 <- ggplot() +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis", limits = c(1, 1100)) +
+
+heatmap_summer <- ggplot() +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_hex(data = total_summer, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis",limits = c(1, 5000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("1") +
+  facet_wrap("type") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-35, -28)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Kelp Particles") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-heatmap_Min_Kfloat_H100_W00 <- ggplot() +
+heatmap_autumn <- ggplot() +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis",limits = c(1, 1100))  +
+  geom_hex(data = total_autumn, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis",limits = c(1, 5000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("2") +
+  facet_wrap("type") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-35, -28)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-heatmap_Min_Kfloat_H095_W05 <- ggplot() +
+heatmap_winter <- ggplot() +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
+  geom_hex(data = total_winter, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis",limits = c(1, 5000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
+  facet_wrap("type") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-35, -28)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Kelp Particles") +
   guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("3") +
   theme_bw() +
@@
 
-heatmap_Min_Kfloat_H090_W10 <- ggplot() +
+heatmap_spring <- ggplot() +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis",limits = c(1, 1100)) +
+  geom_hex(data = total_spring, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis",limits = c(1, 5000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("4") + 
+  facet_wrap("type") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-35, -28)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-heatmap_Min_Kfloat_H085_W15 <- ggplot() +
+### Passive passive
+
+heatmap_summer_passive <- ggplot() +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_hex(data = total_passive_summer, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis",limits = c(1, 5000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("5") + 
+  facet_wrap("type") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-35, -28)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Passive Particles") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-heatmap_Mean_Kfloat_H100_W00 <- ggplot() +
+heatmap_autumn_passive <- ggplot() +
   #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+  geom_hex(data = total_passive_autumn, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis",limits = c(1, 5000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("6") + 
+  facet_wrap("type") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-35, -28)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-heatmap_Mean_KFloat_H095_W05 <- ggplot() +
-  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+heatmap_winter_passive <- ggplot() +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_hex(data = total_passive_winter, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis",limits = c(1, 5000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("7") + 
+  facet_wrap("type") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-35, -28)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Passive Particles") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 
-heatmap_Mean_KFloat_H090_W10 <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
+heatmap_spring_passive <- ggplot() +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_hex(data = total_passive_spring, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis",limits = c(1, 5000))  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
                size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("8") + 
+  facet_wrap("type") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(11, 20), ylim = c(-35, -28)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
         axis.title=element_text(size=12)) 
-
-heatmap_Mean_KFloat_H085_W15 <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("9") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H100_W00 <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("10") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H095_W05 <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("11") + 
-  theme(plot.title = element_text(hjust = 0.5)) + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H090_W10 <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("12") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H085_W15 <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1100))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("13") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-## April
-
-heatmap_RK4_April <- ggplot() +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = p_RK4_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = p_RK4_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("14") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H100_W00_April <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("15") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H095_W05_April <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H095_W05_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("16") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H090_W10_April <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("17") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H085_W15_April <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("18") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_Kfloat_H100_W00_April <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_Kfloat_H100_W00_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_Kfloat_H100_W00_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("19") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_KFloat_H095_W05_April <- ggplot() +
-  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("20") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_KFloat_H090_W10_April <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("21") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_KFloat_H085_W15_April <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H085_W15_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("22") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H100_W00_April <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H100_W00_April , bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H100_W00_mean_April ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("23") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H095_W05_April <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H095_W05_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H095_W05_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("24") +
-  theme(plot.title = element_text(hjust = 0.5)) + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H090_W10_April <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H090_W10_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H090_W10_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("25") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H085_W15_April <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H085_W15_April, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H085_W15_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("26") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-## July
-
-heatmap_RK4_July <- ggplot() +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = p_RK4_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = p_RK4_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("27") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H100_W00_July <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("28") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H095_W05_July <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("29") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H090_W10_July <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) + 
-  ggtitle("30") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H085_W15_July <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("31") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_Kfloat_H100_W00_July <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_Kfloat_H100_W00_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_Kfloat_H100_W00_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("32") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_KFloat_H095_W05_July <- ggplot() +
-  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("33") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_KFloat_H090_W10_July <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("34") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_KFloat_H085_W15_July <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H085_W15_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("35") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H100_W00_July <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H100_W00_July , bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H100_W00_mean_July ,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("36") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H095_W05_July <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H095_W05_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H095_W05_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("37") +
-  theme(plot.title = element_text(hjust = 0.5)) + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H090_W10_July <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H090_W10_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H090_W10_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) + 
-  ggtitle("38") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H085_W15_July <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H085_W15_July, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H085_W15_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) + 
-  ggtitle("39") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-## October
-
-heatmap_RK4_Oct <- ggplot() +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = p_RK4_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = p_RK4_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("40") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H100_W00_Oct <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) + 
-  ggtitle("41") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H095_W05_Oct <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("42") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H090_W10_Oct <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) + 
-  ggtitle("43") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Min_Kfloat_H085_W15_Oct <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Min_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",        limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Min_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) + 
-  ggtitle("44") + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_Kfloat_H100_W00_Oct <- ggplot() +
-  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("45") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_KFloat_H095_W05_Oct <- ggplot() +
-  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  scale_fill_continuous(type = "viridis",       limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) + 
-  ggtitle("46") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_KFloat_H090_W100 <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("47") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Mean_KFloat_H085_W15_Oct <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Mean_KFloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Mean_KFloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("48") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H100_W00_Oct <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H100_W00_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H100_W00_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("49") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H095_W05_Oct <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H095_W05_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H095_W05_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) + 
-  ggtitle("50") +
-  theme(plot.title = element_text(hjust = 0.5)) + 
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H090_W10_Oct <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H090_W10_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H090_W10_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("51") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-heatmap_Max_KFloat_H085_W15_Oct <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_hex(data = Max_Kfloat_H085_W15_Oct, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 1000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  geom_path(data = Max_Kfloat_H085_W15_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("52") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
 ```
@@
 The mean trajectories for each simulation are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each simulation', fig.height=8, fig.width=6}
-
-leg_Jan <- get_legend(heatmap_RK4)
-
-ggarrange(heatmap_RK4, heatmap_Min_Kfloat_H100_W00, heatmap_Min_Kfloat_H095_W05,heatmap_Min_Kfloat_H090_W10,heatmap_Min_Kfloat_H085_W15, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, heatmap_Max_KFloat_H100_W00,heatmap_Max_KFloat_H095_W05, heatmap_Max_KFloat_H090_W10, heatmap_Max_KFloat_H085_W15,as_ggplot(leg_Jan), labels = c("A","B","C","D","E","F","G","H","I","J","K","L","M"), common.legend = TRUE, legend = c("none"),font.label = list(size = 12, face = "bold"))
-
-ggsave(filename = "figure_4.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 8, scale = 1, dpi = 300)
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end of the run for simulations 14 - 26, for April 2018 (Austral autumn).
-See Figure 4 for further details.', fig.height=8, fig.width=6}
-
-leg_April <- get_legend(heatmap_RK4_April)
-
-ggarrange(heatmap_RK4_April,heatmap_Min_Kfloat_H100_W00_April, heatmap_Min_Kfloat_H095_W05_April,heatmap_Min_Kfloat_H090_W10_April,heatmap_Min_Kfloat_H085_W15_April, heatmap_Mean_Kfloat_H100_W00_April, heatmap_Mean_KFloat_H095_W05_April, heatmap_Mean_KFloat_H090_W10_April, heatmap_Mean_KFloat_H085_W15_April, heatmap_Max_KFloat_H100_W00_April,heatmap_Max_KFloat_H095_W05_April, heatmap_Max_KFloat_H090_W10_April, heatmap_Max_KFloat_H085_W15_April,as_ggplot(leg_April), labels = c("A","B","C","D","E","F","G","H","I","J","K","L", "M"), common.legend = TRUE, legend = c("none"), font.label = list(size = 12, face = "bold"))
-
-ggsave(filename = "figure_5.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 8, scale = 1, dpi = 300)
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end of the run for simulations 27 - 39, for July 2018 (Austral winter).
-See Figure 4 for further details.', fig.height=8, fig.width=6}
-
-leg_July <- get_legend(heatmap_RK4_July)
-
-ggarrange(heatmap_RK4_July,heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, as_ggplot(leg_July),labels = c("A","B","C","D","E","F","G","H","I","J","K","L","M"), common.legend = TRUE, legend = c("none"), font.label = list(size = 12, face = "bold"))
-
-ggsave(filename = "figure_6.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 8, scale = 1, dpi = 300)
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 7: Comparison of density plots at the end of the run for simulations 40 - 52, for October 2018 (Austral spring).
-See Figure 4 for further details.', fig.height=8, fig.width=6}
-
-leg_Oct <- get_legend(heatmap_Min_Kfloat_H100_W00_Oct)
-
-ggarrange(heatmap_RK4_Oct, heatmap_Min_Kfloat_H100_W00_Oct, heatmap_Min_Kfloat_H095_W05_Oct,heatmap_Min_Kfloat_H090_W10_Oct,heatmap_Min_Kfloat_H085_W15_Oct, heatmap_Mean_Kfloat_H100_W00_Oct, heatmap_Mean_KFloat_H095_W05_Oct, heatmap_Mean_KFloat_H090_W100, heatmap_Mean_KFloat_H085_W15_Oct, heatmap_Max_KFloat_H100_W00_Oct,heatmap_Max_KFloat_H095_W05_Oct, heatmap_Max_KFloat_H090_W10_Oct, heatmap_Max_KFloat_H085_W15_Oct, as_ggplot(leg_Oct),labels = c("A","B","C","D","E","F","G","H","I","J","K","L","M"), common.legend = TRUE, legend = c("none"), font.label = list(size = 12, face = "bold"))
-
-ggsave(filename = "figure_7.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 8, scale = 1, dpi = 300)
+The reader is referred to tables 1 and 2 for details pertaining to each simulation', fig.height=7, fig.width=6}
+
+ggarrange(heatmap_summer, heatmap_summer_passive, heatmap_autumn, heatmap_autumn_passive, labels = c("AUTO"), nrow = 2, ncol = 2, common.legend = TRUE, legend = c("bottom"),font.label = list(size = 12, face = "bold"))
+
+ggsave(filename = "figure_4.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 9, scale = 1, dpi = 300)
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end run time for simulations 1 - 13, for January 2018 (Austral summer).
+The plots are regular hexagons within the density of particles were calculated.
+The mean trajectories for each simulation are also plotted for comparison.
+The reader is referred to tables 1 and 2 for details pertaining to each simulation', fig.height=10, fig.width=9}
+
+ggarrange(heatmap_winter,heatmap_winter_passive, heatmap_spring, heatmap_spring_passive, labels = c("AUTO"), nrow = 2, ncol = 2, common.legend = TRUE, legend = c("bottom"),font.label = list(size = 12, face = "bold"))
+
+ggsave(filename = "figure_5.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 9, scale = 1, dpi = 300)
+```
+
+## Point distribution
+
+```{r point plots wrangle, message=FALSE, echo=FALSE, warning=FALSE}
+
+total_summer_point <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  #geom_path(data = total_summer %>% 
+              #filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
+  geom_point(data = total_summer, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
+  facet_wrap("type") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme_bw() +
+  #ggtitle("Kelp Particles")
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+total_autumn_point <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  #geom_path(data = total_autumn %>% 
+              #filter(!type =='Passive'), aes(x = lon, y = lat, linetype = type))+
+  geom_point(data = total_autumn, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
+  facet_wrap("type") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+total_winter_point <- ggplot()+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  #geom_path(data = total_winter %>% 
+              #filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
+  geom_point(data = total_winter, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
+  facet_wrap("type") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme_bw() +
+  #ggtitle("Kelp Particles")
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+total_spring_point <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  #geom_path(data = total_spring %>% 
+             # filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
+  geom_point(data = total_spring, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
+  facet_wrap('type') +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# Passive particles
+
+total_summer_point_pass <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  #geom_path(data = total_spring %>% 
+             # filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
+  geom_point(data = total_passive_summer, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
+  facet_wrap('type') +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme_bw() +
+  #ggtitle("Passive Particles")
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+total_autumn_point_pass <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  #geom_path(data = total_spring %>% 
+             # filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
+  geom_point(data = total_passive_autumn, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
+  facet_wrap('type') +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+total_winter_point_pass <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  #geom_path(data = total_spring %>% 
+             # filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
+  geom_point(data = total_passive_winter, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
+  facet_wrap('type') +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme_bw() +
+  ggtitle("Passive Particles")
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+total_spring_point_pass <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  #geom_path(data = total_spring %>% 
+             # filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
+  geom_point(data = total_passive_spring, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
+  facet_wrap('type') +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Simulation")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+```
+
+```{r point plot, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE, fig.cap="Figure 6: Point plots shoing the age of all passive and kelp particles at the end of each simulation run."}
+
+ggarrange(total_summer_point, total_summer_point_pass, total_autumn_point, total_autumn_point_pass,total_winter_point,total_winter_point_pass, total_spring_point,total_spring_point_pass, labels = c("AUTO"), nrow = 2, ncol = 4, common.legend = TRUE, legend = c("bottom"),font.label = list(size = 12, face = "bold"))
+
+ggsave(filename = "figure_age.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 6, width = 6, scale = 1, dpi = 300)
 ```
@@
 
-There are significant differences among the median distances traveled for all particles in each simulation (Figure 8; *p* < 0.05) made for January conditions.
-Simulations that considered 100% hydrodynamic drag travel a shorter distance than particles for simulations with a combination of hydrodynamic and wind drag scenarios (Figure 8; *p* < 0.05).
-Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations.
-When comparing distances between passive and kelp particles during April conditions, significant differences are evident.
-(Figure 8, simulations 14-26; *p* < 0.05).
-Kelp particles with 90% hydrodynamic and 10% wind drag across cross-sectional areas travel a shorter distance across compared to other kelp particles.
-The shortest distance traveled was the simulation with the maximum cross-sectional area and drag exposure scenarios.
-No significant differences are evident for the July simulations, with only a slight variation across simulations (Figure 8, simulations 27-39; *p* > 0.05).
-The plot with results of the simulations made for October conditions shows significant differences when comparing both passive and kelp particles (Figure 8, simulations 40-52; *p* < 0.05).
-Kelp particles with only 90% hydrodynamic drag across cross-sectional areas travel a shorter distance compared to all other particles.
-As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure travel the shortest distance compared to all other particles.
-
 ```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-
-data_dis <- total_data %>% 
-  select(!(kelpvelU)) %>%
-  select(!(kelpvelV)) %>%
-  select(!(velocity))
-
-total_dis_data <- rbind(p_RK4, data_dis)
-
-pos <- c("RK4",
-         "Min_Kfloat_H100_W00",
-         "Min_Kfloat_H095_W05",
-         "Min_Kfloat_H090_W10",
-         "Min_Kfloat_H085_W15",
-         "Mean_Kfloat_H100_W00",
-         "Mean_Kfloat_H095_W05",
-         "Mean_Kfloat_H090_W10",
-         "Mean_Kfloat_H085_W15",
-         "Max_Kfloat_H100_W00",
-         "Max_Kfloat_H095_W05",
-         "Max_Kfloat_H090_W10",
-         "Max_Kfloat_H085_W15")
-
-dis_plot <- ggplot(data = total_dis_data, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
+# Kelp particles
+dis_plot_summer <- ggplot()+
+  #geom_boxplot(data = passive_summer, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(data = total_summer, aes(x = type, y = distance, show.legend = FALSE), notch = TRUE,   notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(labels= pos_dist_plot) +
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
@@
   ylab("Distance (km)")+
-  xlab("Scenario")+
+  xlab("Type")+
   #coord_flip() +
@@
 
-custom_labs <- c("1","2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot
-
-dis_plot <- dis_plot +
-  scale_x_discrete(labels= custom_labs)
-
-pos_April <- c("RK4_April",             
-         "Min_Kfloat_H100_W00_April",
-         "Min_Kfloat_H095_W05_April",
-         "Min_Kfloat_H090_W10_April",
-         "Min_Kfloat_H085_W15_April",
-         "Mean_Kfloat_H100_W00_April",
-         "Mean_Kfloat_H095_W05_April",
-         "Mean_Kfloat_H090_W10_April",
-         "Mean_Kfloat_H085_W15_April",
-         "Max_Kfloat_H100_W00_April",
-         "Max_Kfloat_H095_W05_April",
-         "Max_Kfloat_H090_W10_April",
-         "Max_Kfloat_H085_W15_April")
-
-dis_plot_April <- ggplot(data = total_data_April, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(data = p_RK4_April, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
+dis_plot_autumn <- ggplot()+
+  #geom_boxplot(data = passive_summer, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(data = total_autumn, aes(x = type, y = distance, show.legend = FALSE), notch = TRUE,   notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(labels= pos_dist_plot) +
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
@@
 
-custom_labs_April <- c("14","15","16","17","18","19","20","21","22","23","24","25", "26") #create custom labels and add them to dis_plot
-
-dis_plot_April <- dis_plot_April +
-  scale_x_discrete(labels= custom_labs_April)
-
-pos_July <- c("RK4_July",             
-         "Min_Kfloat_H100_W00_July",
-         "Min_Kfloat_H095_W05_July",
-         "Min_Kfloat_H090_W10_July",
-         "Min_Kfloat_H085_W15_July",
-         "Mean_Kfloat_H100_W00_July",
-         "Mean_Kfloat_H095_W05_July",
-         "Mean_Kfloat_H090_W10_July",
-         "Mean_Kfloat_H085_W15_July",
-         "Max_Kfloat_H100_W00_July",
-         "Max_Kfloat_H095_W05_July",
-         "Max_Kfloat_H090_W10_July",
-         "Max_Kfloat_H085_W15_July")
-
-dis_plot_July <- ggplot(data = total_data_July, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(data = p_RK4_July, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
+dis_plot_winter <- ggplot()+
+  #geom_boxplot(data = passive_summer, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(data = total_winter, aes(x = type, y = distance, show.legend = FALSE), notch = TRUE,   notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(labels= pos_dist_plot) +
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
@@
 
-custom_labs_July <- c("27","28","29","30","31","32","33","34","35","36","37","38", "39") #create custom labels and add them to dis_plot
-
-dis_plot_July <- dis_plot_July +
-  scale_x_discrete(labels= custom_labs_July)
-
-pos_Oct <- c("RK4_Oct",              
-         "Min_Kfloat_H100_W00_Oct",
-         "Min_Kfloat_H095_W05_Oct",
-         "Min_Kfloat_H090_W10_Oct",
-         "Min_Kfloat_H085_W15_Oct",
-         "Mean_Kfloat_H100_W00_Oct",
-         "Mean_Kfloat_H095_W05_Oct",
-         "Mean_Kfloat_H090_W10_Oct",
-         "Mean_Kfloat_H085_W15_Oct",
-         "Max_Kfloat_H100_W00_Oct",
-         "Max_Kfloat_H095_W05_Oct",
-         "Max_Kfloat_H090_W10_Oct",
-         "Max_Kfloat_H085_W15_Oct")
-
-dis_plot_Oct <- ggplot(data = total_data_Oct, aes(x = type, y = distance, show.legend = FALSE))+
-  geom_boxplot(data = p_RK4_Oct, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
-  #scale_x_discrete(limits = pos)+
+dis_plot_spring <- ggplot()+
+  #geom_boxplot(data = passive_summer, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(data = total_spring, aes(x = type, y = distance, show.legend = FALSE), notch = TRUE,   notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(labels= pos_dist_plot) +
   stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
@@
 
-custom_labs_Oct <- c("40","41","42","43","44","45","46","47","48","49","50","51","52") #create custom labels and add them to dis_plot
-
-dis_plot_Oct <- dis_plot_Oct +
-  scale_x_discrete(labels= custom_labs_Oct)
-```
-
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 8: Boxplots of total distance travelled by all particles from the release site over the course of the simulations for scenarios 1 - 52.
+#Passive particles
+
+dis_plot_summer_pass <- ggplot()+
+  #geom_boxplot(data = passive_summer, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(data = total_passive_summer, aes(x = type, y = distance, show.legend = FALSE), notch = TRUE,   notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(labels= pos_dist_plot) +
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Type")+
+  #coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+dis_plot_autumn_pass <- ggplot()+
+  #geom_boxplot(data = passive_summer, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(data = total_passive_autumn, aes(x = type, y = distance, show.legend = FALSE), notch = TRUE,   notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(labels= pos_dist_plot) +
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Type")+
+  #coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+dis_plot_winter_pass <- ggplot()+
+  #geom_boxplot(data = passive_summer, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(data = total_passive_winter, aes(x = type, y = distance, show.legend = FALSE), notch = TRUE,   notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(labels= pos_dist_plot) +
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Type")+
+  #coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+dis_plot_spring_pass <- ggplot()+
+  #geom_boxplot(data = passive_summer, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  geom_boxplot(data = total_passive_spring, aes(x = type, y = distance, show.legend = FALSE), notch = TRUE,   notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
+  #scale_x_discrete(labels= pos_dist_plot) +
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Type")+
+  #coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+```
+
+```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, include=TRUE, fig.cap="Figure 6: Boxplots of total distance travelled by all particles from the release site over the course of the simulations for scenarios 1 - 52.
 The *x*-axis represents the scenario number, and the *y*-axis is the distance travelled in kilometers.
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-The reader is referred to Tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=6}
-
-ggarrange(dis_plot, dis_plot_April, dis_plot_July,dis_plot_Oct, label = "AUTO", ncol = 1, nrow = 4)
-ggsave(filename = "figure_8.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 7, scale = 1, dpi = 300)
-```
-
-## Accumulation
+The reader is referred to Tables 1 and 2 for details pertaining to each scenario.", fig.height=7, fig.width=7}
+
+ggarrange(dis_plot_summer, dis_plot_summer_pass, dis_plot_autumn, dis_plot_autumn_pass, dis_plot_winter, dis_plot_winter_pass,dis_plot_spring, dis_plot_spring_pass, label = "AUTO", ncol = 2, nrow = 4, labels = "AUTO")
+ggsave(filename = "figure_7.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 8, scale = 1, dpi = 300)
+```
+
+## Eddy/Line experiment
 
@@
 
-# January
-
-kelp_acc <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-kelp_acc_mean <- kelp_acc %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-kelp_acc$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_mean$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc$age <- with(kelp_acc, age/3600)
-
-pass_acc <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-pass_acc_mean <- pass_acc %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-pass_acc$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_mean$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc$age <- with(pass_acc, age/3600)
-
-# April
-
-kelp_acc_April <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-kelp_acc_mean_April <- kelp_acc_April %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-kelp_acc_April$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_mean_April$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_April$age <- with(kelp_acc_April, age/3600)
-
-pass_acc_April <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_April.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-pass_acc_mean_April <- pass_acc_April %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-pass_acc_April$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_mean_April$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_April$age <- with(pass_acc, age/3600)
-
-total_acc_data_April <- rbind(pass_acc_April, kelp_acc_April)
-total_acc_mean_April <- rbind(pass_acc_mean_April, kelp_acc_mean_April)
-
-mean_traj_acc_April <- ggplot() +
-  geom_path(data = total_acc_mean_April, aes(x = lon, y = lat, colour = type)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-# July
-
-kelp_acc_July <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-kelp_acc_mean_July <- kelp_acc_July %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-kelp_acc_July$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_mean_July$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_July$age <- with(kelp_acc_July, age/3600)
-
-pass_acc_July <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-pass_acc_mean_July <- pass_acc_July %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-pass_acc_July$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_mean_July$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_July$age <- with(pass_acc_July, age/3600)
-
-total_acc_data_July <- rbind(pass_acc_July, kelp_acc_July)
-total_acc_mean_July <- rbind(pass_acc_mean_July, kelp_acc_mean_July)
-
-mean_traj_acc_July <- ggplot() +
-  geom_path(data = total_acc_mean_July, aes(x = lon, y = lat, colour = type)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-# October
-
-kelp_acc_Oct <- tidync("OceanParcels/trajectory_data/chap3_kelpsim_acc_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-kelp_acc_mean_Oct <- kelp_acc_Oct %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-kelp_acc_Oct$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_mean_Oct$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
-kelp_acc_Oct$age <- with(kelp_acc_Oct, age/3600)
-
-pass_acc_Oct <- tidync("OceanParcels/trajectory_data/chap3_passivesim_acc_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-pass_acc_mean_Oct <- pass_acc_Oct %>% #calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-pass_acc_Oct$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_mean_Oct$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
-pass_acc_Oct$age <- with(pass_acc_Oct, age/3600)
-
-total_acc_data_Oct <- rbind(pass_acc_Oct, kelp_acc_Oct)
-total_acc_mean_Oct <- rbind(pass_acc_mean_Oct, kelp_acc_mean_Oct)
-
-mean_traj_acc_Oct <- ggplot() +
-  geom_path(data = total_acc_mean_Oct, aes(x = lon, y = lat, colour = type)) +
-  xlab("Longtitude") +
-  ylab("Latitude") +
-  ggtitle("") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()
-
-## Coastline of Southern Africa
-load("map/south_africa_coast_hi_res.Rdata")
-
-# January
-
-heatmap_RK4_acc <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  geom_path(data = pass_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
+# summer
+
+eddy_hydro_plot <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_eddy_hydro, aes(x = lon, y = lat, colour = trajectory), show.legend = FALSE) +
+  scale_colour_continuous(type = "viridis",limits = c(0, 9))  +
+  facet_wrap("Season") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  #guides(colour=guide_legend(title="Passive Particles")) +
+  #guides(linetype=guide_legend(title="Kelp Particles")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(7, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme_bw() +
+  ggtitle("Kelp") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+eddy_hydrostksdrag_plot <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_eddy_hydrostksdrag, aes(x = lon, y = lat, colour = trajectory), show.legend = FALSE) +
+  scale_colour_continuous(type = "viridis",limits = c(0, 9))  +
+  facet_wrap("Season") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  #guides(colour=guide_legend(title="Passive Particles")) +
+  #guides(linetype=guide_legend(title="Kelp Particles")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(7, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+eddy_hydrowinddrag_plot <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_eddy_hydrowinddrag, aes(x = lon, y = lat, colour = trajectory), show.legend = FALSE) +
+  scale_colour_continuous(type = "viridis",limits = c(0, 9))  +
+  facet_wrap("Season") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  #guides(colour=guide_legend(title="Passive Particles")) +
+  #guides(linetype=guide_legend(title="Kelp Particles")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(7, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme_bw() +
+  ggtitle("Kelp")
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+eddy_hydrowindstksdrag_plot <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_eddy_hydrowindstksdrag, aes(x = lon, y = lat, colour = trajectory), show.legend = FALSE) +
+  scale_colour_continuous(type = "viridis",limits = c(0, 9))  +
+  facet_wrap("Season") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  #guides(colour=guide_legend(title="Passive Particles")) +
+  #guides(linetype=guide_legend(title="Kelp Particles")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(7, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# Passive particles
+
+eddy_passive_plot_curr <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_eddy_passive_curr, aes(x = lon, y = lat, colour = trajectory), show.legend = FALSE) +
+  scale_colour_continuous(type = "viridis",limits = c(0, 9))  +
+  facet_wrap("Season") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  #guides(colour=guide_legend(title="Passive Particles")) +
+  #guides(linetype=guide_legend(title="Kelp Particles")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(7, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme_bw() +
   ggtitle("Passive") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        plot.title = element_text(hjust = 0.5),
-        element_blank())
-
-heatmap_kelp_acc <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  geom_path(data = kelp_acc_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Kelp") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        plot.title = element_text(hjust = 0.5),
-        element_blank())
-
-point_RK4_acc <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = pass_acc, aes(x = lon, y = lat, colour = age)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-point_kelp_acc <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-# April
-
-heatmap_RK4_acc_April <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  geom_path(data = pass_acc_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-heatmap_kelp_acc_April <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_April, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  geom_path(data = kelp_acc_mean_April,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-point_RK4_acc_April <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = pass_acc_April, aes(x = lon, y = lat, colour = age))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-point_kelp_acc_April <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = kelp_acc_April, aes(x = lon, y = lat, colour = age))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-# July
-
-heatmap_RK4_acc_July <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  geom_path(data = pass_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-heatmap_kelp_acc_July <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_July, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  geom_path(data = kelp_acc_mean_July,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-point_RK4_acc_July <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = pass_acc_July, aes(x = lon, y = lat, colour = age))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-point_kelp_acc_July <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = kelp_acc_July, aes(x = lon, y = lat, colour = age))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  coord_cartesian(xlim = c(11, 20), ylim = c(-36, -26)) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-# October
-
-heatmap_RK4_acc_Oct <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = pass_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  geom_path(data = pass_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-heatmap_kelp_acc_Oct <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = kelp_acc_Oct, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "viridis", limits = c(1, 3000))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  geom_path(data = kelp_acc_mean_Oct, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-point_RK4_acc_Oct <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = pass_acc_Oct, aes(x = lon, y = lat, colour = age))  +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-point_kelp_acc_Oct <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = kelp_acc_Oct, aes(x = lon, y = lat, colour = age)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  #scale_fill_continuous(type = "viridis", limits = c(0, 30)) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Age")) +
-  coord_fixed(ratio=1/1, xlim = c(11, 20), ylim = c(-36, -27)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12),
-        element_blank())
-
-total_acc_data <- rbind(pass_acc, kelp_acc, pass_acc_April, kelp_acc_April,pass_acc_July, kelp_acc_July,pass_acc_Oct, kelp_acc_Oct)
-total_acc_mean <- rbind(pass_acc_mean, kelp_acc_mean, pass_acc_mean_April, kelp_acc_mean_April, pass_acc_mean_July, kelp_acc_mean_July, pass_acc_mean_Oct, kelp_acc_mean_Oct)
-
-mean_traj_acc <- ggplot() +
-  geom_path(data = total_acc_mean, aes(x = lon, y = lat, colour = type)) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw()+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12)) 
-```
-
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 9: Comparison of density between passive and kelp accumulation for scenarios 53-60.
-For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=8, fig.width=6}
-
-ggarrange(heatmap_RK4_acc, heatmap_kelp_acc,heatmap_RK4_acc_April, heatmap_kelp_acc_April, heatmap_RK4_acc_July, heatmap_kelp_acc_July, heatmap_RK4_acc_Oct, heatmap_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January","", "April","","July","","October"), font.label = list(size = 12, face = "bold"), align = "hv")
-
-ggsave(filename = "figure_9.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 6, scale = 1, dpi = 300)
-```
-
-```{r age data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 10: Comparison of particle age between passive and kelp accumulation scenarios 53-60.
-Age is represented in hours since release.
-For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=8, fig.width=6}
-
-ggarrange(point_RK4_acc, point_kelp_acc, point_RK4_acc_April, point_kelp_acc_April, point_RK4_acc_July, point_kelp_acc_July, point_RK4_acc_Oct, point_kelp_acc_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 2, nrow = 4, labels = c("January", "", "April", "","July", "", "October"), font.label = list(size = 12, face = "bold"), align = "hv", hjust = -0.3)
-
-ggsave(filename = "figure_10.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 6, scale = 1, dpi = 300)
-```
-
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+eddy_passive_plot_currstks <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_eddy_passive_currstks, aes(x = lon, y = lat, colour = trajectory), show.legend = FALSE) +
+  scale_colour_continuous(type = "viridis",limits = c(0, 9))  +
+  facet_wrap("Season") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  #guides(colour=guide_legend(title="Passive Particles")) +
+  #guides(linetype=guide_legend(title="Kelp Particles")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(7, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+eddy_passive_plot_currwind <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_eddy_passive_currwind, aes(x = lon, y = lat, colour = trajectory), show.legend = FALSE) +
+  scale_colour_continuous(type = "viridis",limits = c(0, 9))  +
+  facet_wrap("Season") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  #guides(colour=guide_legend(title="Passive Particles")) +
+  #guides(linetype=guide_legend(title="Kelp Particles")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(7, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme_bw() +
+  ggtitle("Passive")
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+eddy_passive_plot_currwindstks <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_point(data = total_eddy_passive_currwindstks, aes(x = lon, y = lat, colour = trajectory), show.legend = FALSE) +
+  scale_colour_continuous(type = "viridis",limits = c(0, 9))  +
+  facet_wrap("Season") +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  #guides(colour=guide_legend(title="Passive Particles")) +
+  #guides(linetype=guide_legend(title="Kelp Particles")) +
+  coord_fixed(ratio=1, clip = "on", xlim = c(7, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+```
+
+```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation for scenarios 53-60.
+For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=6, fig.width=8}
+
+ggarrange(eddy_hydro_plot, eddy_passive_plot_curr, eddy_hydrostksdrag_plot,eddy_passive_plot_currstks, common.legend = TRUE, legend = c("none"), ncol = 2, nrow = 2, labels = c("AUTO"), font.label = list(size = 12, face = "bold"), align = "hv")
+
+ggsave(filename = "figure_8.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 6, width = 8, scale = 2, dpi = 300)
+```
+
+```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation for scenarios 53-60.
+For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=6, fig.width=8}
+
+ggarrange(eddy_hydro_plot, eddy_passive_plot_curr, eddy_hydrostksdrag_plot,eddy_passive_plot_currstks, common.legend = TRUE, legend = c("none"), ncol = 2, nrow = 2, labels = c("AUTO"), font.label = list(size = 12, face = "bold"), align = "hv")
+
+ggsave(filename = "figure_8.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 6, width = 8, scale = 2, dpi = 300)
+```
 The role of eddys that exist within the study domain during the study period play a clear role in the accumulation and dispersal of kelp particles.
@@
 
-```{r shape data, message=FALSE, echo=FALSE, warning=FALSE}
-
-load("map/south_africa_coast_hi_res.Rdata")
-
-# January
-
-sphere <- tidync("OceanParcels/trajectory_data/Max_KFloat_H100_W00.nc") %>%
+```{r test data wrangle, warning = FALSE, echo=FALSE, message=FALSE, eval=FALSE}
+
+kelp_hydro_A <- tidync("OceanParcels/chapter_3_output/kelp_hydro_testA.nc") %>% #1hr
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::select(time, trajectory, lon, lat, distance, age, Phase, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, waterdragU, waterdragV) %>%
   dplyr::rename(t = time) %>%
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-sphere_mean <- sphere %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-sphere$type <- "Sphere" 
-sphere$Month <- "January" 
-sphere_mean$Type <- "Sphere"
-
-# read in data again for one max simulation, as cannot sort issue out with rename df...
-cyl <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00.nc") %>%
+
+kelp_hydro_B <- tidync("OceanParcels/chapter_3_output/kelp_hydro_testB.nc") %>% #3hr
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::select(time, trajectory, lon, lat, distance, age, Phase, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, waterdragU, waterdragV) %>%
   dplyr::rename(t = time) %>%
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-cyl_mean <- cyl %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-cyl$type <- "Cylinder" 
-cyl$Month <- "January" 
-cyl_mean$Type <- "Cylinder"
-
-shape_data_jan <- rbind(sphere_mean, cyl_mean)
-
-# Make mean trajectory plot
-shape_plot <- ggplot()+
-  coord_fixed(ratio=1, xlim = c(14, 20), ylim = c(-35, -30)) + 
-  #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = shape_data_jan, aes(x = lon, y = lat, linetype = Type)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey")  +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("January") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5),
-        element_blank()) 
-  #coord_fixed(ratio=1/1) # fix the aspect ratio of plot
-
-# make density map
-heatmap_sphere <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = sphere, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey")  +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  geom_path(data = sphere_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5), 
-        element_blank())
-
-heatmap_cyl <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = cyl, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey")  +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  geom_path(data = cyl_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5), 
-        element_blank())
-
-# April
-
-sphere_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_April.nc") %>%
+
+kelp_hydro_C <- tidync("OceanParcels/chapter_3_output/kelp_hydro_testC.nc") %>% #6hr
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::select(time, trajectory, lon, lat, distance, age, Phase, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, waterdragU, waterdragV) %>%
   dplyr::rename(t = time) %>%
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-sphere_mean_April <- sphere_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-sphere_April$type <- "Sphere" 
-sphere_April$Month <- "April" 
-sphere_mean_April$Type <- "Sphere"
-
-# read in data again for one max simulation, as cannot sort issue out with rename df...
-cyl_April <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00_April.nc") %>%
+
+kelp_hydro_D <- tidync("OceanParcels/chapter_3_output/kelp_hydro_testD.nc") %>% #12hr
   hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
+  dplyr::select(time, trajectory, lon, lat, distance, age, Phase, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, waterdragU, waterdragV) %>%
   dplyr::rename(t = time) %>%
   mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-cyl_mean_April <- cyl_April %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-cyl_April$type <- "Cylinder" 
-cyl_April$Month <- "April" 
-cyl_mean_April$Type <- "Cylinder"
-
-shape_data_apr <- rbind(sphere_mean_April, cyl_mean_April)
-
-# Make mean trajectory plot
-shape_plot_April <- ggplot()+
-  coord_fixed(ratio=1, xlim = c(14, 20), ylim = c(-35, -30)) + 
-  #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = shape_data_apr, aes(x = lon, y = lat, linetype = Type)) +
+
+heatmap_1hr <- ggplot() +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_hex(data = kelp_hydro_A, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis")  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey")  +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("April") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5), 
-        element_blank())
-
-# make density map
-heatmap_sphere_April <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = sphere_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("1 Hr release") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+heatmap_3hr <- ggplot() +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_hex(data = kelp_hydro_B, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis")  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey")  +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  geom_path(data = sphere_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   labs(y = expression(Latitude*degree*E)) +
+  ggtitle("3 Hr release") +
   guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5),
-        element_blank())
-
-heatmap_cyl_April <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = cyl_April, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+heatmap_6hr <- ggplot() +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_hex(data = kelp_hydro_C, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis")  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey")  +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  geom_path(data = cyl_mean_April, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  ggtitle("6 Hr release") +
   guides(fill=guide_legend(title="Number of particles")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5),
-        element_blank())
-
-# July
-
-sphere_July <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-sphere_mean_July <- sphere_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-sphere_July$type <- "Sphere" 
-sphere_July$Month <- "July" 
-sphere_mean_July$Type <- "Sphere"
-
-# read in data again for one max simulation, as cannot sort issue out with rename df...
-cyl_July <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00_July.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-cyl_mean_July <- cyl_July %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-cyl_July$type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
-cyl_July$Month <- "July" # add extra column 'type' to dataframe and assigned 'kelp' label
-cyl_mean_July$Type <- "Cylinder" # add extra column 'type' to dataframe and assigned 'kelp' label
-
-shape_data_july <- rbind(sphere_mean_July, cyl_mean_July)
-
-# Make mean trajectory plot
-shape_plot_July <- ggplot()+
-  coord_fixed(ratio=1, xlim = c(14, 20), ylim = c(-35, -30)) + 
-  #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_path(data = shape_data_apr, aes(x = lon, y = lat, linetype = Type)) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+heatmap_12hr <- ggplot() +
+  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
+  geom_hex(data = kelp_hydro_D, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_continuous(type = "viridis")  +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey")  +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("July") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5),
-        element_blank()) 
-  #coord_fixed(ratio=1/1) # fix the aspect ratio of plot
-
-# make density map
-heatmap_sphere_July <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = sphere_July, bins = 80, aes(x = lon, y = lat)) +
-    geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey")  +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_path(data = sphere_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   labs(y = expression(Latitude*degree*E)) +
+  ggtitle("12 Hr release") +
   guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5),
-        element_blank())
-
-heatmap_cyl_July <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = cyl_July, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey")  +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  geom_path(data = cyl_mean_July, aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5),
-        element_blank())
-
-# October
-
-sphere_Oct <- tidync("OceanParcels/trajectory_data/chapter3_additional/Max_KFloat_H100_W00_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-sphere_mean_Oct <- sphere_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-sphere_Oct$type <- "Sphere" 
-sphere_Oct$Month <- "October" 
-sphere_mean_Oct$Type <- "Sphere"
-
-# read in data again for one max simulation, as cannot sort issue out with rename df...
-cyl_Oct <- tidync("OceanParcels/trajectory_data/Max_ShapeFloat_H100_W00_Oct.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-cyl_mean_Oct <- cyl_Oct %>% #calculate mean trajectory 
-  group_by(t) %>% 
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-cyl_Oct$type <- "Cylinder" 
-cyl_Oct$Month <- "October"
-cyl_mean_Oct$Type <- "Cylinder"
-
-shape_data_oct <- rbind(sphere_mean_Oct, cyl_mean_Oct)
-
-# Make mean trajectory plot
-shape_plot_Oct <- ggplot()+
-  coord_fixed(ratio=1, xlim = c(14, 20), ylim = c(-35, -30)) + 
-  #geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  geom_path(data = shape_data_oct, aes(x = lon, y = lat, linetype = Type)) +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("October") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5),
-        element_blank()) 
-  #coord_fixed(ratio=1/1) # fix the aspect ratio of plot
-
-# make density map
-heatmap_sphere_Oct <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = sphere_Oct, bins = 80, aes(x = lon, y = lat)) +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  geom_path(data = sphere_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(fill=guide_legend(title="Number of particles")) +
-  ggtitle("") + 
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5),
-        element_blank())
-
-heatmap_cyl_Oct <- ggplot() +
-  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
-  geom_hex(data = cyl_Oct, bins = 80, aes(x = lon, y = lat)) +
-    geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "grey") +
-    geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey") +
-  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
-  geom_path(data = cyl_mean_Oct,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5),
-        element_blank())
-
-# Distance plot
-
-shape_data_total <- rbind(sphere, cyl, sphere_April, cyl_April, sphere_July, cyl_July, sphere_Oct, cyl_Oct)
-shape_pos <- c("January", "April", "July", "October")
-shape_dis_plot <- ggplot(data = shape_data_total, aes(x = Month, y = distance, show.legend = TRUE, fill = type))+
-  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = TRUE, stat = "boxplot") +
-  scale_x_discrete(limits = shape_pos)+
-  stat_compare_means(method = "kruskal.test", label = "p.signif", aes(group=Month), hide.ns = FALSE)+
-  ylab("Distance (km)")+
-  xlab("Month") + 
-  ggtitle("Comparison of distance travelled") +
-  guides(fill=guide_legend(title="Coefficient")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5),
-        element_blank())
-
-```
-
-```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 11: Comparison of trajectory data between sphere and cylinder shape coefficients for scenarios 61-64.
-Row A represents the mean trajectory comparison, row B-C are density comparison plots, and the bottom boxplot is comparison of shape coefficients total distance travelled from the release site for each of the respective months (seasons) simulated.
-The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=10, fig.width=6}
-
-shape_plot_A <- ggarrange(shape_plot, shape_plot_April, shape_plot_July, shape_plot_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 1, font.label = list(size = 12, face = "bold"), align = "hv", labels = c("A","","",""))
-
-shape_plot_B <- ggarrange(heatmap_sphere, heatmap_sphere_April,heatmap_sphere_July, heatmap_sphere_Oct, heatmap_cyl, heatmap_cyl_April, heatmap_cyl_July, heatmap_cyl_Oct, common.legend = TRUE, legend = c("bottom"), ncol = 4, nrow = 2, align = "hv", labels = c("B", "", "", "", "C", "", "", ""), font.label = list(size = 12, face = "bold"))
-
-ggarrange(shape_plot_A, shape_plot_B, shape_dis_plot, ncol = 1, nrow = 3, legend = c("right"), labels = c("","","","D"), font.label = list(size = 11, face = "bold"), heights = c(1,2,2), widths = c(1,2,2))
-
-ggsave(filename = "figure_11.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 7, scale = 1, dpi = 300)
-
-```
-
-# Supplementary material
-
-```{r vector maps, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 12: The average ocean (left column) and wind vector maps (right column) for the respective seasons are presented.
-The velocity vectors were averaged over the 30 day simulation times for respective seasons (summer, autumn, winter, spring), for both ocean and wind data.
-In the case of wind, direction is also indicated by means of arrows.
-January ocean and wind data are presented in panels A-B, April in panels C-D, July in panels E-F, and October in panels G-H." ,fig.height=8, fig.width=6}
-
-library(jpeg)
-library(grid)
-library(gridExtra)
-
-vector_maps <- readJPEG("figures/chapter_3/vector_combined.jpeg")
-grid.raster(vector_maps)
-```
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+```
+
+```{r test data outputs, warning=FALSE, echo=FALSE, message=FALSE,eval=FALSE, fig.cap="Heatmaps comparing densities and patterns across time-release options.", fig.height=6, fig.width=5}
+
+ggarrange(heatmap_1hr, heatmap_3hr, heatmap_6hr, heatmap_12hr,common.legend = TRUE, legend = c("none"), ncol = 2, nrow = 2, labels = c("AUTO"), font.label = list(size = 12, face = "bold"), align = "hv")
+
+#ggsave(filename = "figure_9.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 6, scale = 1, dpi = 300)
+```

### 2021-10-31 thesis_chapter_3_datawrangle.R

@@ -1,0 +1,875 @@
+##-- New passive simulations
+# Passive with Ocean currents only
+
+passive_summer_curr <- tidync("OceanParcels/chapter_3_output/passive_summer_curr.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2017-12-01 00:00:00"))
+passive_summer_curr_mean <- passive_summer_curr %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+passive_summer_curr$type <- "Currents" # add extra column 'type' to dataframe and assigned 'kelp' label
+#passive_summer_curr$velocity <- with(passive_summer_curr, sqrt(velU^2 + velV^2))
+passive_summer_curr_mean$type <- "Currents" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+passive_autumn_curr <- tidync("OceanParcels/chapter_3_output/passive_autumn_curr.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2017-12-01 00:00:00"))
+passive_autumn_curr_mean <- passive_autumn_curr %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+passive_autumn_curr$type <- "Currents" # add extra column 'type' to dataframe and assigned 'kelp' label
+#passive_autumn_curr$velocity <- with(passive_autumn_curr, sqrt(velU^2 + velV^2))
+passive_autumn_curr_mean$type <- "Currents" # add extra column 'type' to dataframe and assigned 'kelp' label
+
+passive_winter_curr <- tidync("OceanParcels/chapter_3_output/passive_winter_curr.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(trajectory, lon, lat, time, distance, age) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2017-12-01 00:00:00"))
+passive_winter_curr_mean <- passive_winter_curr %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
[…]

### 2021-10-31 thesis_chapter_4.Rmd
updated chapter 4 files

--- 2021-08-06 thesis_chapter_4.Rmd
@@
 ---
-title: "Chapter 4"
-author: "Ross Coppin"
-date: "11/12/2020"
+title: "Individual-based numerical experiment to describe the distribution of floating kelp within the Southern Benguela Upwelling System."
+author: 
+- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
 output:
+  word_document: 
+    reference_docx: markdown/wiley_template.docx
   html_document: default
-  word_document: default
   pdf_document: default
+bibliography: references.bib
+csl: markdown/limnologyandoceanography.csl
 ---
@@
 ```{r message=FALSE, warning=FALSE, include=FALSE}
-
 library(tidyverse)
@@
 library(ggpubr)
-
+library(jpeg)
+library(gridExtra)
+library(grid)
 ```
@@
 
-The raft-time is the amount of time macroalgae floats at the surface before becoming negatively buoyant and begins to sink to the ocean-floor.
+The inter- to sub-tidal habitats of temperate shorelines are usually dominated by species of macroalgae, specifically brown seaweeds of the order Laminariales; commonly referred to as kelp [@Steneck2002;@Schiel2015-vr].
+Kelps are highly productive organisms [@Mann1973;@Miller2011-pm; @Dayton1998-xw;@Krumhansl2012;@Pedersen2012-vr] which provide a 3-dimensional habitat for various fauna and flora [@anderson1997;@Leliaert2000a;@Steneck2002; @Norderhaug2012-ks; @Krumhansl2016].
+Kelp are also capable of altering the local wave and light environment by wave dampening [@Mork1996-md; @Mia2013] and shading of the understorey habitat [@duggins1990; @watanabe1992; @clark2004], as well as providing important organic subsidy through erosion of fronds and fragmentation of plant structures [@Krumhansl2012; @Filbee-Dexter2018-wp; @DeBettignies2013; @Simonson2015; @ramirez2021].
+Due to kelps’ ecological role in the nearshore, they are regarded as ecological engineers as well as important influencer of biodiversity along temperate shorelines [@FowlerWalker2005;@graham2004;@Harley2012;@Smale2017; @Wing2007-tv;@Wernberg2008].
+Kelps are also important from a socio-economic perspective, as they provide habitats for economically important species such as abalone and rock lobster [@vasquez2014; @blamey2018], and are harvested for uses ranging from feed for the aquaculture sector [@troell2006] to pharmaceuticals [@smit2004].
+
+Kelps’ ecological and socio-economic contributions are not limited to their local environment.
+During times of high wave energy kelps can become dislodged from the substrate  [@seymour1989; @Thomsen2004] and are transported via ocean currents, eventually deposited on beaches or offshore areas [@hobday2000a].
+The kelp deposited on beaches provides important organic subsidies for coastal invertebrate communities, as well as a resource which can be harvested for economic purposes [@kirkman1997;@troell2006; @holden2018].
+Along with the organic input are 'hitchhikers', which are organisms that remain attached during offshore transport once dislodged [@highsmith1985; @holmquist1994; @macaya2016; @thiel2005].
+The transport of other organisms attached to kelp is regarded as an important processes in invasive species intrusion [@lewis2005; @McCormick2008; @nikula2010; @batista2018] and marine connectivity [@norton1992, @ingolfsson1995; @hobday2000a; @fraser2011].
+The formation of kelp-rafts can also occur, providing habitat complexity in the pelagic environment for various fish species [@kingsford1992; @kingsford1995], Kelps which are transported offshore can become entangled to form a kelp-raft, but there is no clear mechanism for kelp-raft formation.
+The structural integrity of individual kelp or rafts will degrade over time through a combination of biotic and abiotic factors which ultimately results in sinking.
+Once settled on the seafloor, the kelp will degrade further over time resulting in organic input into offshore areas; which may be transported back towards the coast through upwelling processes [@Krumhansl2012; @Filbee-Dexter2018-wp; @ramirez2021].
+
+Understanding the distribution patterns of macroalgae in the ocean and the underlying oceanographic processes influencing trajectory has received much attention in the literature.
+Approaches such as radio-tagged kelp rafts [@harrold1989], geological forms in holdfasts [@waters2018], genetic approaches [@craw2018; @waters2018; @nikula2013], the presence of region-specific epibionts on beach-cast kelp [@fraser2011], and Lagrangian approaches [@putman2018; @brooks2019; @putman2020] have helped to shed light on the processes involved in macroalgae transport.
+Ocean currents are the main influencer of trajectory as well as drag and windage, which is the direct effect of wind on a floating object [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
+More recent work by @coppin2021 has highlighted the importance of both hydrodynamic and wind drag in entrainment patterns within oceanographic mesoscale processes, more specifically eddies.
+
+Although the past research concerning the transport of kelp has greatly improved our understanding of such processes, limitations and gaps still exist.
+For example, kelps are highly morphologically diverse which may influence buoyancy and overall drag characteristics.
+Morphological characteristics, sea surface temperatures, and light also influence the time which kelps remain afloat at the surface; known as raft-times.
+Macroalgae rafts also vary greatly in size, therefore studies investigating a specific raft size limits our understanding of that particular scenario as the findings cannot be inferred for other sizes of kelp rafts.
+Finally, it has been well established that morphological characteristics of macroalgae are influenced by local abiotic factors, which can vary greatly along coastlines.
+Although factors such as light, turbidity, and nutrients have been shown to influence macroalgae morphology; temperature and waves are regarded as the main influencers [@Hurd2000;@Denny2002a;@Thomsen2004;@FowlerWalker2005;@Wernberg2005c;@mabin2013;@bekkby2014; @coppin2021].
+How the variable response in morphological characteristics influences sink rates, buoyancy, and drag is also yet to be established.
+There is also no current work where rafts may form on the surface and the role of eddies in this regard.
+It is most likely that eddies might play a role in this regard as they have shown to cause collection of other flotsam at the surface via entrainment @onink2019; @dasaro2018; @brach2018.
+For example, a study by @beron2019 showed that the inclusion of inertia in numerical experiments resulted in entrainment patterns of flotsam in eddies which more closely reflected observational data.
+
+A better understanding of the distribution of floating kelp and ocean processes involved will aid in understanding the broader ecological role that kelp plays in the marine environment but will also be useful for coastal management and harvesting activities.
+The current work aims to shed light on the distribution patterns of floating kelp by a series of Lagrangian individual-particle tracking numerical experiments.
+This approach will allow for the morphological characteristics that vary with the local temperature and wave environment to be accounted for.
+In addition, since individual kelp has a higher probability of dislodgement compared to kelps in which holdfasts are clustered together, an individual-based approach will reflect real-world scenarios [@Thomsen2004].
+The novel approach of including hydrodynamic and wind drag based on site-specific morphological characteristics by @coppin2021 will be used and the current work will build on this approach by the inclusion of morphological specific buoyancy and sinking rates.
+
+# Methods
+
+## Morphological Characteristics
+
+Morphological characteristics of kelps are influenced by local wave and temperature conditions, which in turn will influence buoyancy, drag, and sinking parameters.
+Kelp morphological data was garnered from a previous study by @coppin2020.
+The assumption made is that the largest kelps (largest surface area) are the most likely to be dislodged from the substrate, therefore only maximum surface areas were garnered from the data.
+In instances where the morphological data needed were not available, reasonable estimates were used based on previous studies and experience.
+To calculate the surface area and volume for kelps from each site, geometric shapes were used which most closely matched the shape of the relevant kelp morphological characteristics as per @coppin2021.
+The shapes assigned to each of the morphological features were used in the calculation of buoyancy, drag, and sinking rate.
+
+## Buoyancy
+
+To effectively model the trajectory of a floating object, buoyancy needs to be calculated.
+The buoyancy will determine the amount of exposure to, and ultimately magnitude, of hydrodynamic and wind drag which in turn influences the trajectory of the object to the variable current and wind fields.
+The portion of an object above and below the surface of the water is determined by the net effect of forces acting on the object, namely buoyancy force (BF) and the weight of the object.
+The BF is the force pushing the object out of the water and the weight is the force pulling the object down.
+In order to calculate the buoyancy force (BF) the density and volume of the object must be known, such that: $$BF = p_fV_sg$$
+
+where $p_f$ is the density of the liquid, $V_s$ is the volume of the object submerged, and $g$ is the gravitational acceleration (m/s).
+
+The force resulting from the weight of an object can be calculated by $W_f = p_oV_og$. Substitution of mass for density and volume alters the equation to: $$W_f = p_oV_og$$
+where $p_o$ is the density of the object, $V_o$ is the total volume of the object and $g$ is gravitational acceleration.
+An object will only float if the forces acting on it ($BF$ and $W_f$) are equal to one another, such that:
+$$BF = W_f$$
+then substituting  density and volume results in:
+$$p_fV_sg = p_oV_og$$
+which can also be further expressed as:
+$$V_s/V_o = p_o/p_f$$
+
+The solution of equation (3) results in a unitless fraction that can be expressed as a percentage of the relevant portion submerged.
+This in turn can be used to calculate the percentage of the object above the surface of the water (exposed to wind).
+The percentage of the object below and above the water surface was used in the calculation of hydrodynamic and wind drag when the virtual kelp is at the surface.
+Since morphological characteristics are influenced by the local wave and temperature environment, the buoyancy of virtual kelp particles may vary from site to site.
+
+## Sink Rate
+
+Kelp floating on the surface of the ocean will begin to sink once the structural integrity of air-filled structures of the plants fails and becomes flooded.
+It is also possible for sinking to begin due to an increase in epibiont biomass which overwhelms the buoyancy of kelp, however, this is beyond the scope of this study.
+The target species of this study, *E.
+maxima*, contains a hollow stipe and a single air-filled pneumatocyst which serve to keep the plant upright in the water column, allowing the light-saturated surface waters to be reached.
+These structures also cause the plant to float if dislodged and therefore was taken into account when calculating the overall volume of the plant; which in turn was used to calculate other important parameters (buoyancy, drag, and sink rate).
+
+## Vertical Velocity
+
+The force responsible for the kelp to begin to sink once the structural integrity has been compromised is the weight of the overall plant.
+The calculation of the force resulting from the weight was based on previous work by @dietrich1982, @lobelle2021 and @kooi2017, such that
+
+$$F_w = (p_s - p_l)gV$$
+
+Where $p_s$ is the density of the particle, $p__l$ the density of the medium, $g$ is the gravitational acceleration constant, and $V$ is the total volume of the particle.
+In order to accurately calculate the density of the kelp based on volume and mass, the volume of the hollow section of the stipe was subtracted from the overall volume.
+The target species used for this experiment (*E.
+maxima*) has an air-filled stipe and therefore the hollowness needs to be accounted for when calculating the overall volume of the stipe.
+The volume of the kelp is:
+$$Volume = V_t - V_h$$
+where $V_t$ is the total volume of the stipe if now hollowness is taken into account and $V_h$ is the hollowness volume of the stipe.
+The other component of the vertical velocity is the drag force in the opposite direction to the downward force of the weight.
+The vertical drag component is calculated by the same equation for drag in the horizontal, such that:
+
+$$F_d = 1/2pv^2C_DA_c$$
+where $p$ is the density of the medium, $v$ is the velocity at a particular time-step, $C_d$ is the shape drag-coefficient, and $A_c$ is the cross-sectional area of the object.
+However, unlike surface drag where the velocity is interpolated from the underlying ocean model, the velocity component for the vertical drag force is calculated from the same equation used for hydrodynamic and wind drag.
+The approach for modeling the 'behavior' of kelp was used in this paper allows for the incorporation of site-specific morphological characteristics.
+
+## Simulations
+
+Simulations used daily outputs of velocity vectors (meridonal and zonal) and salinity (psu) for varying depths for the study region.
+To account for the role of morphological characteristics in influencing drift trajectory in terms of hydrodynamic and wind drag [@coppin2021] simulations were run separately for each site per season.
+Morphological characteristics of macroalgae has been shown to be based on the local wave and temperature environment, hence the need for site specific parameters with regards to morphological characteristics.
+Running simulations per season allowed incorporation of seasonal raft-times identified in pervious work [@graiff2016 and @graiff2013].
+The outputs from each simulation (site) was pooled based on season.
+
+Simulations were run using the OceanParcels software (available from https://oceanparcels.org/).
+All simulations were run using a Runge-Kutte fourth order advection scheme, while custom kernels simulated the effects of hydrodynamic and wind drag on overall trajectory, seasonal raft-times, sinking and settling on seafloor of virtual kelp particles; all important processes in the determination of surface and bottom sinks.
+The simulations were run for 2019, which was divided into various seasons (see table 2).
+For incorporating drag, the site specific buoyancy calculations were used to determine the magnitude of hydrodynamic (portion below surface) and wind (portion above the surface) while virtual kelp particles were on the surface and then converted to full hydrodynamic drag exposure (100%) once the virtual kelp started sinking; as wind would no longer act as a direct influencer of trajectory once below the ocean surface.
+
+## Raft-time
+
+Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remains afloat at the surface of the ocean after detachment before sinking.
 Raft-time, also referred to as raft longevity, is determined by epibiont load, macroalgal growth rate and the loss of kelp biomass over time through fragmentation.
@@
 pyrifera* dispersal is dependent on low temperature and moderate irradiance conditions, with high temperatures and higher irradiance reducing overall raft-time and hence dispersal potential [@rothausler2011].
-
-# Methods
-
-## Raft-time
-Raft-time or sometimes referred to as raft-longevity is the time a macroalgae raft remains afloat at the surface of the ocean after detachment before sinking.
 Raft-time was estimated from the results of past studies by @graiff2016 and @graiff2013 for each season.
@@
 
-```{r morphology surface area table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="TBD."}
+```{r fieldworkdata and table wrangle, message = FALSE, echo=FALSE, warning=FALSE, include=FALSE}
+### data for stipe hollowness
+hollow_df <- read.csv(file = "data/chapter_4/hollowness_data.csv", sep = ";")
+hollow_df$inner_radius <- with(hollow_df, (outer_diameter/2)) # addinmg difference between measurements
+hollow_df$outer_radius <- with(hollow_df, (inner_diameter/2)) # addinmg difference between measurements
+hollow_df$difference <- with(hollow_df, (outer_diameter - inner_diameter)) # addinmg difference between measurements
+hollow_df$percentage <- with(hollow_df, (inner_diameter/outer_diameter)) # inner as percentage of outer
+hollow_df_mean <- hollow_df %>% # calculate mean trajectory
+  group_by(site) %>%
+  summarise(across(outer_diameter:percentage, mean, na.rm= TRUE))
+
+ggplot(data = hollow_df, aes(x = outer_diameter, y = inner_diameter)) +
+  geom_point() +
+  stat_smooth(method=lm)
+
+ggplot(data = hollow_df, aes(x = no, y = percentage)) +
+  geom_point() +
+  geom_smooth()
+
+ggplot(data = hollow_df, aes(x = no, y = difference)) +
+  geom_point() +
+  geom_smooth()
+
+# testing for correlation
+outer_df <- hollow_df$outer_diameter
+inner_df <- hollow_df$inner_diameter
+
+outer_rad <- hollow_df$outer_radius
+inner_rad <- hollow_df$inner_radius
+
+cor(outer_df, inner_df, method = c("pearson"))
+cor.test(outer_df, inner_df, method = c("pearson"))
+cor(outer_rad, inner_rad, method = c("pearson"))
+cor.test(outer_rad, inner_rad, method = c("pearson"))
+
+cor(outer_df, inner_df, method = c("spearman"))
+cor.test(outer_df, inner_df, method = c("spearman"))
+cor(outer_rad, inner_rad, method = c("spearman"))
+cor.test(outer_rad, inner_rad, method = c("spearman"))
 
@@
 morph_data$bottom_radius <- with(morph_data, 10)
-morph_data$stipe_radius <- with(morph_data, stipe_circ/(pi*2))
+morph_data$stipe_radius <- with(morph_data, ((stipe_circ/pi)/2))
+morph_data$hollow_radius <- with(morph_data, (0.61*stipe_radius))
+morph_data$hollow_length <- with(morph_data, (0.90*stipe_length))
 morph_data$height <- with(morph_data, 5)
+morph_data$hollow_volume <- with(morph_data, pi*(hollow_radius*hollow_radius)*hollow_length)
 
@@
 morph_data$total_volume <- with(morph_data, (frond_thickness*frond_length*frond_width) + (primary_length * primary_width * frond_thickness) + (pi*(stipe_radius*stipe_radius)*stipe_length) + (pi*height*0.3333*((top_radius*top_radius) + (top_radius * bottom_radius) + (bottom_radius*bottom_radius))))
+
+# Calculate actual (True) kelp volume (hollowness taken into account)
+
+morph_data$kelp_volume <- with(morph_data, (total_volume - hollow_volume))
+
+# Calculate the density of each kelp
+
+morph_data$kelp_density <- with(morph_data, total_mass/kelp_volume)
 
@@
 site_surface_area <- flextable(data = site_surface_area)
-#site_surface_area <- theme_zebra(site_surface_area, odd_header = "transparent", even_header = "transparent")
+site_surface_area <- theme_zebra(site_surface_area, odd_header = "transparent", even_header = "transparent")
 site_surface_area <- align(x= site_surface_area, align = "center", part = "all")
-#area_table <- set_header_labels(area_table, Plant.characteristic = "Plant characteristic", Approximate.shape = "Approximate shape", Equation.used = "Equation Used", Plant.dimensions = "Plant dimensions")
-#area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
-site_surface_area <- set_table_properties(site_surface_area, width = 1, layout = "autofit")
-#site_surface_area <- set_caption(site_surface_area, "Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.")
+site_surface_area <- theme_vanilla(site_surface_area)
+site_surface_area <- fontsize(site_surface_area, size = 8, part = "body")
+site_surface_area <- fontsize(site_surface_area, size = 8, part = "header")
+site_surface_area <- colformat_num(site_surface_area,digits = 1)
+site_surface_area <- set_header_labels(site_surface_area, frond_mass = "Frond mass", primary_length = "Primary blade length", primary_width = "Primary blade width", frond_length = "Frond length", stipe_mass = "Stipe mass", stipe_length = "Stipe length", stipe_circ = "Stipe circumference", tufts = "No.
+of Tufts", epi_length = "Epiphyte length", total_length = "Total length", stipe_radius = "Stipe radius", frond_width = "Frond width", frond_thickness = "Frond thickness", bulb_length = "Bulb length", bulb_base_radius = "Bulb base radius", top_radius = "Top radius", bottom_radius = "Bottom radius", hollow_radius = "Hollow radius", hollow_length = "Hollow length", height = "Height", hollow_volume = "Hollow Volume", total_Ac = "Total Ac", total_mass = "Total mass", total_volume = "Total volume", kelp_volume = "Kelp Volume", kelp_density = "Kelp density")
+site_surface_area <- set_table_properties(site_surface_area, layout = "fixed")
+```
+
+```{r morphology surface area table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
+site_surface_area <- set_caption(site_surface_area, "Table 1:Summary of the attributes and estimates used to calculate the overall site-specific surface area of kelp individuals used in the various simulations.")
+#site_surface_area <- set_table_properties(site_surface_area, layout = "autofit")
 site_surface_area
+
 ```
@@
 All simulations were run using a Runge-Kutte fourth order advection scheme, while custom kernels simulated the effects of hydrodynamic and wind drag on overall trajectory, seasonal raft-times, sinking and settling on seafloor of virtual kelp particles; all important processes in the determination of surface and bottom sinks.
-The simulations were run for 2019, which was divided into various seasons (see table??).
-For incorporating drag, the 'maximum exposure scenario' proposed by @coppin2021 was used for all simulations (85% hydrodynamic and 15% wind drag) while virtual kelp particles were on the surface and then would convert to full hydrodynamic drag exposure (100%) once the virtual kelp started sinking; as wind would no longer act as a direct influencer of trajectory once below the ocean surface.
+The simulations were run for two years, which were divided into the relevant seasons seasons (see table 2).
+
+```{r simulation table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="TBD."}
+
+chapter4_table2 <- read_csv2(file = "tables/chapter_4/simulation_table.csv")
+chapter4_table2 <- flextable(data = chapter4_table2)
+chapter4_table2 <- theme_zebra(chapter4_table2, odd_header = "transparent", even_header = "transparent")
+chapter4_table2 <- align(x= chapter4_table2, align = "center", part = "all")
+chapter4_table2 <- theme_vanilla(chapter4_table2)
+chapter4_table2 <- fontsize(chapter4_table2, size = 8, part = "body")
+chapter4_table2 <- fontsize(chapter4_table2, size = 8, part = "header")
+chapter4_table2 <- colformat_num(chapter4_table2,digits = 1)
+chapter4_table2 <- set_table_properties(chapter4_table2, layout = "autofit")
+chapter4_table2 <- set_caption(chapter4_table2, "Table 2: Summary of the simulations run for the study.
+Simulations were run for each site over the course of 2 years and then pooled based on season for each year.
+The details pertaining are provided in the below.
+The start of each year was the beginning of Austral summer.")
+chapter4_table2
+```
 
@@
 
+To compare virtual kelp trajectories between sites the mean trajectories were calculated, and density distribution maps were produced.
+Density distribution maps are an effective method for assessing pathways and ocean connectivity [@vanSebille2018].
+The distance particles traveled was also measured and used to produce boxplots.
+Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particles [@vanSebille2018].
+If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity because of momentum energy transfer [@putman2020; @putman2018; @vanSebille2018].
+To aid in seasonal comparison of trajectories, proportions of each kelp phase at the end of each simulation run was calculated and compared.
+The different phases were surface, subsurface, beached, sunk and out of bounds.
+
 # Results
-## Ocean analysis
-
-```{r eval=FALSE, message=FALSE, echo=FALSE, include=FALSE}
-
-```
-
-```{r data wrangle, message=FALSE, warning=FALSE, include=FALSE}
-
-houtbay_summer <- tidync("OceanParcels/chapter_4_output/hb_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-houtbay_summer$season <- "Summer" 
-houtbay_summer$site <- "Hout Bay" 
-houtbay_summer_mean <- houtbay_summer %>% # calculate mean trajectory
- group_by(t) %>% 
- summarise(across(lon:bathy, mean, na.rm= TRUE))
-houtbay_summer_mean$season <- "Summer" 
-houtbay_summer_mean$site <- "Hout Bay"
-
-houtbay_autumn <- tidync("OceanParcels/chapter_4_output/hb_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-houtbay_autumn$season <- "Autumn" 
-houtbay_autumn$site <- "Hout Bay" 
-houtbay_autumn_mean <- houtbay_autumn %>% # calculate mean trajectory
- group_by(t) %>% 
- summarise(across(lon:bathy, mean, na.rm= TRUE))
-houtbay_autumn_mean$season <- "Autumn" 
-houtbay_autumn_mean$site <- "Hout Bay"
-
-houtbay_winter <- tidync("OceanParcels/chapter_4_output/hb_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-houtbay_winter$season <- "Winter" 
-houtbay_winter$site <- "Hout Bay" 
-houtbay_winter_mean <- houtbay_winter %>% # calculate mean trajectory
- group_by(t) %>% 
- summarise(across(lon:bathy, mean, na.rm= TRUE))
-houtbay_winter_mean$season <- "Winter" 
-houtbay_winter_mean$site <- "Hout Bay"
-
-houtbay_spring <- tidync("OceanParcels/chapter_4_output/hb_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-houtbay_spring$season <- "Spring" 
-houtbay_spring$site <- "Hout Bay" 
-houtbay_spring_mean <- houtbay_spring %>% # calculate mean trajectory
- group_by(t) %>% 
- summarise(across(lon:bathy, mean, na.rm= TRUE))
-houtbay_spring_mean$season <- "Spring" 
-houtbay_spring_mean$site <- "Hout Bay"
-
-kom_summer <- tidync("OceanParcels/chapter_4_output/kom_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-kom_summer$season <- "Summer" 
-kom_summer$site <- "Kommetjie" 
-kom_summer_mean <- kom_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-kom_summer_mean$season <- "Summer" 
-kom_summer_mean$site <- "Kommetjie"
-
-kom_autumn <- tidync("OceanParcels/chapter_4_output/kom_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-kom_autumn$season <- "Autumn" 
-kom_autumn$site <- "Kommetjie" 
-kom_autumn_mean <- kom_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-kom_autumn_mean$season <- "Autumn" 
-kom_autumn_mean$site <- "Kommetjie"
-
-kom_winter <- tidync("OceanParcels/chapter_4_output/kom_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-kom_winter$season <- "Winter" 
-kom_winter$site <- "Kommetjie" 
-kom_winter_mean <- kom_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-kom_winter_mean$season <- "Winter" 
-kom_winter_mean$site <- "Kommetjie"
-
-kom_spring <- tidync("OceanParcels/chapter_4_output/kom_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-kom_spring$season <- "Spring" 
-kom_spring$site <- "Kommetjie" 
-kom_spring_mean <- kom_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-kom_spring_mean$season <- "Spring" 
-kom_spring_mean$site <- "Kommetjie"
-
-scar_summer <- tidync("OceanParcels/chapter_4_output/scar_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-scar_summer$season <- "Summer" 
-scar_summer$site <- "Scarborough" 
-scar_summer_mean <- scar_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-scar_summer_mean$season <- "Summer" 
-scar_summer_mean$site <- "Scarborough"
-
-scar_autumn <- tidync("OceanParcels/chapter_4_output/scar_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-scar_autumn$season <- "Autumn" 
-scar_autumn$site <- "Scarborough" 
-scar_autumn_mean <- scar_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-scar_autumn_mean$season <- "Autumn" 
-scar_autumn_mean$site <- "Scarborough"
-
-scar_winter <- tidync("OceanParcels/chapter_4_output/scar_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-scar_winter$season <- "Winter" 
-scar_winter$site <- "Scarborough" 
-scar_winter_mean <- scar_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-scar_winter_mean$season <- "Winter" 
-scar_winter_mean$site <- "Scarborough"
-
-scar_spring <- tidync("OceanParcels/chapter_4_output/scar_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-scar_spring$season <- "Spring" 
-scar_spring$site <- "Scarborough" 
-scar_spring_mean <- scar_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-scar_spring_mean$season <- "Spring" 
-scar_spring_mean$site <- "Scarborough"
-
-oude_summer <- tidync("OceanParcels/chapter_4_output/oudekraal_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-oude_summer$season <- "Summer" 
-oude_summer$site <- "Oudekraal" 
-oude_summer_mean <- oude_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-oude_summer_mean$season <- "Summer" 
-oude_summer_mean$site <- "Oudekraal"
-
-oude_autumn <- tidync("OceanParcels/chapter_4_output/oudekraal_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-oude_autumn$season <- "Autumn" 
-oude_autumn$site <- "Oudekraal" 
-oude_autumn_mean <- oude_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-oude_autumn_mean$season <- "Autumn" 
-oude_autumn_mean$site <- "Oudekraal"
-
-oude_winter <- tidync("OceanParcels/chapter_4_output/oudekraal_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-oude_winter$season <- "Winter" 
-oude_winter$site <- "Oudekraal" 
-oude_winter_mean <- oude_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-oude_winter_mean$season <- "Winter" 
-oude_winter_mean$site <- "Oudekraal"
-
-oude_spring <- tidync("OceanParcels/chapter_4_output/oudekraal_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-oude_spring$season <- "Spring" 
-oude_spring$site <- "Oudekraal" 
-oude_spring_mean <- oude_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-oude_spring_mean$season <- "Spring" 
-oude_spring_mean$site <- "Oudekraal"
-
-betty_summer <- tidync("OceanParcels/chapter_4_output/betty_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-betty_summer$season <- "Summer" 
-betty_summer$site <- "Betty's Bay" 
-betty_summer_mean <- betty_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-betty_summer_mean$season <- "Summer" 
-betty_summer_mean$site <- "Betty's Bay"
-
-betty_autumn <- tidync("OceanParcels/chapter_4_output/betty_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-betty_autumn$season <- "Autumn" 
-betty_autumn$site <- "Betty's Bay" 
-betty_autumn_mean <- betty_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-betty_autumn_mean$season <- "Autumn" 
-betty_autumn_mean$site <- "Betty's Bay"
-
-betty_winter <- tidync("OceanParcels/chapter_4_output/betty_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-betty_winter$season <- "Winter" 
-betty_winter$site <- "Betty's Bay" 
-betty_winter_mean <- betty_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-betty_winter_mean$season <- "Winter" 
-betty_winter_mean$site <- "Betty's Bay"
-
-betty_spring <- tidync("OceanParcels/chapter_4_output/betty_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-betty_spring$season <- "Spring" 
-betty_spring$site <- "Betty's Bay" 
-betty_spring_mean <- betty_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-betty_spring_mean$season <- "Spring" 
-betty_spring_mean$site <- "Betty's Bay"
-
-kalk_summer <- tidync("OceanParcels/chapter_4_output/kalk_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-kalk_summer$season <- "Summer" 
-kalk_summer$site <- "Kalk Bay" 
-kalk_summer_mean <- kalk_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-kalk_summer_mean$season <- "Summer" 
-kalk_summer_mean$site <- "Kalk Bay"
-
-kalk_autumn <- tidync("OceanParcels/chapter_4_output/kalk_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-kalk_autumn$season <- "Autumn" 
-kalk_autumn$site <- "Kalk Bay" 
-kalk_autumn_mean <- kalk_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-kalk_autumn_mean$season <- "Autumn" 
-kalk_autumn_mean$site <- "Kalk Bay"
-
-kalk_winter <- tidync("OceanParcels/chapter_4_output/kalk_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-kalk_winter$season <- "Winter" 
-kalk_winter$site <- "Kalk Bay" 
-kalk_winter_mean <- kalk_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-kalk_winter_mean$season <- "Winter" 
-kalk_winter_mean$site <- "Kalk Bay"
-
-kalk_spring <- tidync("OceanParcels/chapter_4_output/kalk_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-kalk_spring$season <- "Spring" 
-kalk_spring$site <- "Kalk Bay" 
-kalk_spring_mean <- kalk_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-kalk_spring_mean$season <- "Spring" 
-kalk_spring_mean$site <- "Kalk Bay"
-
-aframe_summer <- tidync("OceanParcels/chapter_4_output/aframe_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-aframe_summer$season <- "Summer" 
-aframe_summer$site <- "A-Frame" 
-aframe_summer_mean <- aframe_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-aframe_summer_mean$season <- "Summer" 
-aframe_summer_mean$site <- "A-Frame"
-
-aframe_autumn <- tidync("OceanParcels/chapter_4_output/aframe_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-aframe_autumn$season <- "Autumn" 
-aframe_autumn$site <- "A-Frame" 
-aframe_autumn_mean <- aframe_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-aframe_autumn_mean$season <- "Autumn" 
-aframe_autumn_mean$site <- "A-Frame"
-
-aframe_winter <- tidync("OceanParcels/chapter_4_output/aframe_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-aframe_winter$season <- "Winter" 
-aframe_winter$site <- "A-Frame" 
-aframe_winter_mean <- aframe_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-aframe_winter_mean$season <- "Winter" 
-aframe_winter_mean$site <- "A-Frame"
-
-aframe_spring <- tidync("OceanParcels/chapter_4_output/aframe_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-aframe_spring$season <- "Spring" 
-aframe_spring$site <- "A-Frame" 
-aframe_spring_mean <- aframe_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-aframe_spring_mean$season <- "Spring" 
-aframe_spring_mean$site <- "A-Frame"
-
-sthelena_summer <- tidync("OceanParcels/chapter_4_output/sthelena_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-sthelena_summer$season <- "Summer" 
-sthelena_summer$site <- "St. Helena Bay" 
-sthelena_summer_mean <- sthelena_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-sthelena_summer_mean$season <- "Summer" 
-sthelena_summer_mean$site <- "St. Helena Bay"
-
-sthelena_autumn <- tidync("OceanParcels/chapter_4_output/sthelena_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-sthelena_autumn$season <- "Autumn" 
-sthelena_autumn$site <- "St. Helena Bay" 
-sthelena_autumn_mean <- sthelena_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-sthelena_autumn_mean$season <- "Autumn" 
-sthelena_autumn_mean$site <- "St. Helena Bay"
-
-sthelena_winter <- tidync("OceanParcels/chapter_4_output/sthelena_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-sthelena_winter$season <- "Winter" 
-sthelena_winter$site <- "St. Helena Bay" 
-sthelena_winter_mean <- sthelena_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-sthelena_winter_mean$season <- "Winter" 
-sthelena_winter_mean$site <- "St. Helena Bay"
-
-sthelena_spring <- tidync("OceanParcels/chapter_4_output/sthelena_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-sthelena_spring$season <- "Spring" 
-sthelena_spring$site <- "St. Helena Bay" 
-sthelena_spring_mean <- sthelena_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-sthelena_spring_mean$season <- "Spring" 
-sthelena_spring_mean$site <- "St. Helena Bay"
-
-millers_summer <- tidync("OceanParcels/chapter_4_output/millers_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-millers_summer$season <- "Summer" 
-millers_summer$site <- "Miller's Point" 
-millers_summer_mean <- millers_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-millers_summer_mean$season <- "Summer" 
-millers_summer_mean$site <- "Miller's Point"
-
-millers_autumn <- tidync("OceanParcels/chapter_4_output/millers_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-millers_autumn$season <- "Autumn" 
-millers_autumn$site <- "Miller's Point" 
-millers_autumn_mean <- millers_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-millers_autumn_mean$season <- "Autumn" 
-millers_autumn_mean$site <- "Miller's Point"
-
-millers_winter <- tidync("OceanParcels/chapter_4_output/millers_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-millers_winter$season <- "Winter" 
-millers_winter$site <- "Miller's Point" 
-millers_winter_mean <- millers_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-millers_winter_mean$season <- "Winter" 
-millers_winter_mean$site <- "Miller's Point"
-
-millers_spring <- tidync("OceanParcels/chapter_4_output/millers_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-millers_spring$season <- "Spring" 
-millers_spring$site <- "Miller's Point" 
-millers_spring_mean <- millers_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-millers_spring_mean$season <- "Spring" 
-millers_spring_mean$site <- "Miller's Point"
-
-baboon_summer <- tidync("OceanParcels/chapter_4_output/baboon_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-baboon_summer$season <- "Summer" 
-baboon_summer$site <- "Baboon Rock" 
-baboon_summer_mean <- baboon_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-baboon_summer_mean$season <- "Summer" 
-baboon_summer_mean$site <- "Baboon Rock"
-
-baboon_autumn <- tidync("OceanParcels/chapter_4_output/baboon_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-baboon_autumn$season <- "Autumn" 
-baboon_autumn$site <- "Baboon Rock" 
-baboon_autumn_mean <- baboon_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-baboon_autumn_mean$season <- "Autumn" 
-baboon_autumn_mean$site <- "Baboon Rock"
-
-baboon_winter <- tidync("OceanParcels/chapter_4_output/baboon_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-baboon_winter$season <- "Winter" 
-baboon_winter$site <- "Baboon Rock" 
-baboon_winter_mean <- baboon_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-baboon_winter_mean$season <- "Winter" 
-baboon_winter_mean$site <- "Baboon Rock"
-
-baboon_spring <- tidync("OceanParcels/chapter_4_output/baboon_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-baboon_spring$season <- "Spring" 
-baboon_spring$site <- "Baboon Rock" 
-baboon_spring_mean <- baboon_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-baboon_spring_mean$season <- "Spring" 
-baboon_spring_mean$site <- "Baboon Rock"
-
-basata_summer <- tidync("OceanParcels/chapter_4_output/basata_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-basata_summer$season <- "Summer" 
-basata_summer$site <- "Basata Rock" 
-basata_summer_mean <- basata_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-basata_summer_mean$season <- "Summer" 
-basata_summer_mean$site <- "Basata Rock"
-
-basata_autumn <- tidync("OceanParcels/chapter_4_output/basata_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-basata_autumn$season <- "Autumn" 
-basata_autumn$site <- "Basata Rock" 
-basata_autumn_mean <- basata_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-basata_autumn_mean$season <- "Autumn" 
-basata_autumn_mean$site <- "Basata Rock"
-
-basata_winter <- tidync("OceanParcels/chapter_4_output/basata_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-basata_winter$season <- "Winter" 
-basata_winter$site <- "Basata Rock" 
-basata_winter_mean <- basata_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-basata_winter_mean$season <- "Winter" 
-basata_winter_mean$site <- "Basata Rock"
-
-basata_spring <- tidync("OceanParcels/chapter_4_output/basata_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-basata_spring$season <- "Spring" 
-basata_spring$site <- "Basata Rock" 
-basata_spring_mean <- basata_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-basata_spring_mean$season <- "Spring" 
-basata_spring_mean$site <- "Basata Rock"
-
-buffels_summer <- tidync("OceanParcels/chapter_4_output/buffels_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-buffels_summer$season <- "Summer" 
-buffels_summer$site <- "Buffelsbaai" 
-buffels_summer_mean <- buffels_summer %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-buffels_summer_mean$season <- "Summer" 
-buffels_summer_mean$site <- "Buffelsbaai"
-
-buffels_autumn <- tidync("OceanParcels/chapter_4_output/buffels_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-buffels_autumn$season <- "Autumn" 
-buffels_autumn$site <- "Buffelsbaai" 
-buffels_autumn_mean <- buffels_autumn %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-buffels_autumn_mean$season <- "Autumn" 
-buffels_autumn_mean$site <- "Buffelsbaai"
-
-buffels_winter <- tidync("OceanParcels/chapter_4_output/buffels_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-buffels_winter$season <- "Winter" 
-buffels_winter$site <- "Buffelsbaai" 
-buffels_winter_mean <- buffels_winter %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-buffels_winter_mean$season <- "Winter" 
-buffels_winter_mean$site <- "Buffelsbaai"
-
-buffels_spring <- tidync("OceanParcels/chapter_4_output/buffels_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
-buffels_spring$season <- "Spring" 
-buffels_spring$site <- "Buffelsbaai" 
-buffels_spring_mean <- buffels_spring %>% # calculate mean trajectory
-  group_by(t) %>% 
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-buffels_spring_mean$season <- "Spring" 
-buffels_spring_mean$site <- "Buffelsbaai"
-
-soetwater_summer <- tidync("OceanParcels/chapter_4_output/soetwater_summer.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-soetwater_summer$season <- "Summer"
-soetwater_summer$site <- "Soetwater"
-soetwater_summer_mean <- soetwater_summer %>% # calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-soetwater_summer_mean$season <- "Summer"
-soetwater_summer_mean$site <- "Soetwater"
-
-soetwater_autumn <- tidync("OceanParcels/chapter_4_output/soetwater_autumn.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-soetwater_autumn$season <- "Autumn"
-soetwater_autumn$site <- "Soetwater"
-soetwater_autumn_mean <- soetwater_autumn %>% # calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-soetwater_autumn_mean$season <- "Autumn"
-soetwater_autumn_mean$site <- "Soetwater"
-
-soetwater_winter <- tidync("OceanParcels/chapter_4_output/soetwater_winter.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-soetwater_winter$season <- "Winter"
-soetwater_winter$site <- "Soetwater"
-soetwater_winter_mean <- soetwater_winter %>% # calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-soetwater_winter_mean$season <- "Winter"
-soetwater_winter_mean$site <- "Soetwater"
-
-soetwater_spring <- tidync("OceanParcels/chapter_4_output/soetwater_spring.nc") %>%
-  hyper_tibble() %>%
-  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, KelpDepth, SinkSpeed,       KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
-  dplyr::rename(t = time) %>%
-  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
-soetwater_spring$season <- "Spring"
-soetwater_spring$site <- "Soetwater"
-soetwater_spring_mean <- soetwater_spring %>% # calculate mean trajectory
-  group_by(t) %>%
-  summarise(across(lon:bathy, mean, na.rm= TRUE))
-soetwater_spring_mean$season <- "Spring"
-soetwater_spring_mean$site <- "Soetwater"
-
-# Summer
-total_summer <- rbind(kom_summer, houtbay_summer, scar_summer, oude_summer, betty_summer, kalk_summer, aframe_summer, sthelena_summer, millers_summer, baboon_summer, basata_summer, buffels_summer, soetwater_summer)
-total_summer$age <- with(total_summer, age/3600)
-
-# Autumn
-
-total_autumn <- rbind(kom_autumn, houtbay_autumn, scar_autumn, oude_autumn, betty_autumn, kalk_autumn, aframe_autumn, sthelena_autumn, millers_autumn, baboon_autumn, basata_autumn, buffels_autumn, soetwater_autumn)
-total_autumn$age <- with(total_autumn, age/3600)
-
-# Winter
-
-total_winter <- rbind(kom_winter, houtbay_winter, scar_winter, oude_winter, betty_winter, kalk_winter, aframe_winter, sthelena_winter, millers_winter, baboon_winter, basata_winter, buffels_winter, soetwater_winter)
-total_winter$age <- with(total_winter, age/3600)
-
-# Spring
-
-total_spring <- rbind(kom_spring, houtbay_spring, scar_spring, oude_spring, betty_spring, kalk_spring, aframe_spring, sthelena_spring, millers_spring, baboon_spring, basata_spring, buffels_spring, soetwater_spring)
-total_spring$age <- with(total_spring, age/3600)
-
-# Summer
-total_summer_mean <- rbind(kom_summer_mean, houtbay_summer_mean, scar_summer_mean, oude_summer_mean, betty_summer_mean, kalk_summer_mean, aframe_summer_mean, sthelena_summer_mean, millers_summer_mean, baboon_summer_mean, basata_summer_mean, buffels_summer_mean, soetwater_summer_mean)
-
-# Autumn
-total_autumn_mean <- rbind(kom_autumn_mean, houtbay_autumn_mean, scar_autumn_mean, oude_autumn_mean, betty_autumn_mean, kalk_autumn_mean, aframe_autumn_mean, sthelena_autumn_mean, millers_autumn_mean, baboon_autumn_mean, basata_autumn_mean, buffels_autumn_mean, soetwater_autumn_mean)
-
-# Winter
-total_winter_mean <- rbind(kom_winter_mean, houtbay_winter_mean, scar_winter_mean, oude_winter_mean, betty_winter_mean, kalk_winter_mean, aframe_winter_mean, sthelena_winter_mean, millers_winter_mean, baboon_winter_mean, basata_winter_mean, buffels_winter_mean, soetwater_winter_mean)
-
-# Spring
-total_spring_mean <- rbind(kom_spring_mean, houtbay_spring_mean, scar_spring_mean, oude_spring_mean, betty_spring_mean, kalk_spring_mean, aframe_spring_mean, sthelena_spring_mean, millers_spring_mean, baboon_spring_mean, basata_spring_mean, buffels_spring_mean, soetwater_spring_mean)
+
+```{r data, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-```
-
-## Mean trajectory
+total_summer <- read.csv(file = "OceanParcels/chapter_4_data/df_total_summer.csv") 
+total_summer2 <- read.csv(file = "OceanParcels/chapter_4_data/df_total_summer2.csv")
+total_autumn <- read.csv(file = "OceanParcels/chapter_4_data/df_total_autumn.csv") 
+total_autumn2 <- read.csv(file = "OceanParcels/chapter_4_data/df_total_autumn2.csv")
+total_winter <- read.csv(file = "OceanParcels/chapter_4_data/df_total_winter.csv") 
+total_winter2 <- read.csv(file = "OceanParcels/chapter_4_data/df_total_winter2.csv")
+total_spring <- read.csv(file = "OceanParcels/chapter_4_data/df_total_spring.csv") 
+total_spring2 <- read.csv(file = "OceanParcels/chapter_4_data/df_total_spring2.csv")
+
+# calculate and add overall velocity vector of ocean current sampled at virtual kelp location
+total_summer$velocity <- with(total_summer, sqrt(VelU^2 + VelV^2))
+total_summer2$velocity <- with(total_summer2, sqrt(VelU^2 + VelV^2))
+total_autumn$velocity <- with(total_autumn, sqrt(VelU^2 + VelV^2))
+total_autumn2$velocity <- with(total_autumn2, sqrt(VelU^2 + VelV^2))
+total_winter$velocity <- with(total_winter, sqrt(VelU^2 + VelV^2))
+total_winter2$velocity <- with(total_winter2, sqrt(VelU^2 + VelV^2))
+total_spring$velocity <- with(total_spring, sqrt(VelU^2 + VelV^2))
+total_spring2$velocity <- with(total_spring2, sqrt(VelU^2 + VelV^2))
+# calculate and add overall kelp velocity vector sampled at virtual kelp location
+total_summer$kelpvelocity <- with(total_summer, sqrt(KelpVelU^2 + KelpVelV^2))
+total_summer2$kelpvelocity <- with(total_summer2, sqrt(KelpVelU^2 + KelpVelV^2))
+total_autumn$kelpvelocity <- with(total_autumn, sqrt(KelpVelU^2 + KelpVelV^2))
+total_autumn2$kelpvelocity <- with(total_autumn2, sqrt(KelpVelU^2 + KelpVelV^2))
+total_winter$kelpvelocity <- with(total_winter, sqrt(KelpVelU^2 + KelpVelV^2))
+total_winter2$kelpvelocity <- with(total_winter2, sqrt(KelpVelU^2 + KelpVelV^2))
+total_spring$kelpvelocity <- with(total_spring, sqrt(KelpVelU^2 + KelpVelV^2))
+total_spring2$kelpvelocity <- with(total_spring2, sqrt(KelpVelU^2 + KelpVelV^2))
+
+total_summer_mean <- total_summer %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_summer2_mean <- total_summer2 %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+total_autumn_mean <- total_autumn %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_autumn2_mean <- total_autumn2 %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+total_winter_mean <- total_winter %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_winter2_mean <- total_winter2 %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+total_spring_mean <- total_spring %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_spring2_mean <- total_spring2 %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+#total_summer_mean <- read.csv(file = "OceanParcels/chapter_4_data/df_total_summer_mean.csv") 
+#total_summer2_mean <- read.csv(file = "OceanParcels/chapter_4_data/df_total_summer2_mean.csv")
+#total_autumn_mean <- read.csv(file = "OceanParcels/chapter_4_data/df_total_autumn_mean.csv") 
+#total_autumn2_mean <- read.csv(file = "OceanParcels/chapter_4_data/df_total_autumn2_mean.csv")
+#total_winter_mean <- read.csv(file = "OceanParcels/chapter_4_data/df_total_winter_mean.csv") 
+#total_winter2_mean <- read.csv(file = "OceanParcels/chapter_4_data/df_total_winter2_mean.csv")
+#total_spring_mean <- read.csv(file = "OceanParcels/chapter_4_data/df_total_spring_mean.csv") 
+#total_spring2_mean <- read.csv(file = "OceanParcels/chapter_4_data/df_total_spring2_mean.csv")
+
+total_summer_surface <- total_summer %>% 
+filter(Phase == 1)
+total_summer_subsurface <- total_summer %>% 
+filter(Phase == 2)
+total_summer_sink <- total_summer %>% 
+filter(Phase == 3)
+total_summer_beached <- total_summer %>% 
+filter(Phase == 4)
+total_summer_boundry <- total_summer %>% 
+filter(Phase == 5)
+total_autumn_surface <- total_autumn %>% 
+filter(Phase == 1)
+total_autumn_subsurface <- total_autumn %>% 
+filter(Phase == 2)
+total_autumn_sink <- total_autumn %>% 
+filter(Phase == 3)
+total_autumn_beached <- total_autumn %>% 
+filter(Phase == 4)
+total_autumn_boundry <- total_autumn %>% 
+filter(Phase == 5)
+total_winter_surface <- total_winter %>% 
+filter(Phase == 1)
+total_winter_subsurface <- total_winter %>% 
+filter(Phase == 2)
+total_winter_sink <- total_winter %>% 
+filter(Phase == 3)
+total_winter_beached <- total_winter %>% 
+filter(Phase == 4)
+total_winter_boundry <- total_winter %>% 
+filter(Phase == 5)
+total_spring_surface <- total_spring %>% 
+filter(Phase == 1)
+total_spring_subsurface <- total_spring %>% 
+filter(Phase == 2)
+total_spring_sink <- total_spring %>% 
+filter(Phase == 3)
+total_spring_beached <- total_spring %>% 
+filter(Phase == 4)
+total_spring_boundry <- total_spring %>% 
+filter(Phase == 5)
+
+total_summer2_surface <- total_summer2 %>% 
+filter(Phase == 1)
+total_summer2_subsurface <- total_summer2 %>% 
+filter(Phase == 2)
+total_summer2_sink <- total_summer2 %>% 
+filter(Phase == 3)
+total_summer2_beached <- total_summer2 %>% 
+filter(Phase == 4)
+total_summer2_boundry <- total_summer2 %>% 
+filter(Phase == 5)
+total_autumn2_surface <- total_autumn2 %>% 
+filter(Phase == 1)
+total_autumn2_subsurface <- total_autumn2 %>% 
+filter(Phase == 2)
+total_autumn2_sink <- total_autumn2 %>% 
+filter(Phase == 3)
+total_autumn2_beached <- total_autumn2 %>% 
+filter(Phase == 4)
+total_autumn2_boundry <- total_autumn2 %>% 
+filter(Phase == 5)
+total_winter2_surface <- total_winter2 %>% 
+filter(Phase == 1)
+total_winter2_subsurface <- total_winter2 %>% 
+filter(Phase == 2)
+total_winter2_sink <- total_winter2 %>% 
+filter(Phase == 3)
+total_winter2_beached <- total_winter2 %>% 
+filter(Phase == 4)
+total_winter2_boundry <- total_winter2 %>% 
+filter(Phase == 5)
+total_spring2_surface <- total_spring2 %>% 
+filter(Phase == 1)
+total_spring2_subsurface <- total_spring2 %>% 
+filter(Phase == 2)
+total_spring2_sink <- total_spring2 %>% 
+filter(Phase == 3)
+total_spring2_beached <- total_spring2 %>% 
+filter(Phase == 4)
+total_spring2_boundry <- total_spring2 %>% 
+filter(Phase == 5)
+
+total_summer_surface_mean <- total_summer_surface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_summer_subsurface_mean <- total_summer_subsurface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+total_summer2_surface_mean <- total_summer2_surface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_summer2_subsurface_mean <- total_summer2_subsurface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+total_autumn_surface_mean <- total_autumn_surface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_autumn_subsurface_mean <- total_autumn_subsurface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+total_autumn2_surface_mean <- total_autumn2_surface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_autumn2_subsurface_mean <- total_autumn2_subsurface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+total_winter_surface_mean <- total_winter_surface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_winter_subsurface_mean <- total_winter_subsurface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+total_winter2_surface_mean <- total_winter2_surface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_winter2_subsurface_mean <- total_winter2_subsurface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+total_spring_surface_mean <- total_spring_surface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_spring_subsurface_mean <- total_spring_subsurface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+total_spring2_surface_mean <- total_spring2_surface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+total_spring2_subsurface_mean <- total_spring2_subsurface %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:kelpvelocity, mean, na.rm= TRUE))
+
+# eddy experiments
+
+total_kelp_eddy <- read.csv(file = "OceanParcels/chapter_4_data/df_total_kelp_eddy.csv")
+
+# passive experiments
+
+total_passive1 <- read.csv(file = "OceanParcels/chapter_4_data/df_total_passive_year1.csv") 
+total_passive2 <- read.csv(file = "OceanParcels/chapter_4_data/df_total_passive_year2.csv")
+
+total_passive1_surface <- total_passive1 %>% 
+filter(Year == 1) %>% 
+filter(Phase == 1)
+total_passive1_subsurface <- total_passive1 %>% 
+filter(Year == 1) %>% 
+filter(Phase == 2)
+total_passive1_sink <- total_passive1 %>%
+filter(Year == 1) %>% 
+filter(Phase == 3)
+total_passive1_beached <- total_passive1 %>% 
+filter(Year == 1) %>% 
+filter(Phase == 4)
+total_passive1_boundry <- total_passive1 %>%
+filter(Year == 1) %>% 
+filter(Phase == 5)
+
+total_passive2_surface <- total_passive2 %>% 
+filter(Year == 2) %>% 
+filter(Phase == 1)
+total_passive2_subsurface <- total_passive2 %>% 
+filter(Year == 2) %>% 
+filter(Phase == 2)
+total_passive2_sink <- total_passive2 %>%
+filter(Year == 2) %>% 
+filter(Phase == 3)
+total_passive2_beached <- total_passive2 %>% 
+filter(Year == 2) %>% 
+filter(Phase == 4)
+total_passive2_boundry <- total_passive2 %>%
+filter(Year == 2) %>% 
+filter(Phase == 5)
+
+```
 
 ```{r message=FALSE, warning=FALSE, include=FALSE}
-
-summer_mean <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_summer_mean, aes(x = lon, y = lat, linetype = site))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Summer") +
-  guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(15, 20), ylim = c(-35, -31)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-autumn_mean <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_autumn_mean, aes(x = lon, y = lat, linetype = site))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Autumn") +
-  guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(15, 20), ylim = c(-35, -31)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-winter_mean <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_winter_mean, aes(x = lon, y = lat, linetype = site))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Winter") +
-  guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(15, 20), ylim = c(-35, -31)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-spring_mean <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4, colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_path(data = total_spring_mean, aes(x = lon, y = lat, linetype = site))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Spring") +
-  guides(linetype=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(15, 20), ylim = c(-35, -31)) + # fix the aspect ratio of plot
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12)) 
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=6, fig.width=6}
-
-ggarrange(summer_mean, autumn_mean, winter_mean, spring_mean,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 2, labels = c("AUTO"))
-
-```
-
-## Heatmaps
-
-```{r message=FALSE, warning=FALSE, include=FALSE}
-
-total_summer_surface <- total_summer %>% 
-filter(KelpDepth == 0)
-total_summer_sink <- total_summer %>% 
-filter(KelpDepth > 0)
-total_autumn_surface <- total_autumn %>% 
-filter(KelpDepth == 0)
-total_autumn_sink <- total_autumn %>% 
-filter(KelpDepth > 0)
-total_winter_surface <- total_winter %>% 
-filter(KelpDepth == 0)
-total_winter_sink <- total_winter %>% 
-filter(KelpDepth > 0)
-total_spring_surface <- total_spring %>% 
-filter(KelpDepth == 0)
-total_spring_sink <- total_spring %>% 
-filter(KelpDepth > 0)
-
-## Load software need for entire script
-
-load("map/south_africa_coast_hi_res.Rdata")
-load("map/sa_bathy.RData")
-
 # summer
@@
   geom_hex(data = total_summer_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) + 
-  scale_fill_viridis_c(aesthetics = "fill", option = "C", limits = c(1, 110000)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Surface") +
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+summer2_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_summer2_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) + 
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+summer_heatmap_subsurface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_summer_subsurface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+summer2_heatmap_subsurface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_summer2_subsurface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   geom_hex(data = total_summer_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 1000)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Sub-surface") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12)) 
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 25)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+summer2_heatmap_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_summer2_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 25)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
 # autumn
@@
   geom_hex(data = total_autumn_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_viridis_c(aesthetics = "fill", option = "C", limits = c(1, 110000)) +
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+autumn2_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_autumn2_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+autumn_heatmap_subsurface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_autumn_subsurface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+autumn2_heatmap_subsurface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_autumn2_subsurface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
   labs(x = expression(Longitude*degree*S)) +
@@
   geom_hex(data = total_autumn_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 1000)) +
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 25)) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+autumn2_heatmap_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_autumn2_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 25)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
   labs(x = expression(Longitude*degree*S)) +
@@
   geom_hex(data = total_winter_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_viridis_c(aesthetics = "fill", option = "C", limits = c(1, 110000)) +
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+winter2_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_winter2_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+winter_heatmap_subsurface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_winter_subsurface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+winter2_heatmap_subsurface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_winter2_subsurface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
   labs(x = expression(Longitude*degree*S)) +
@@
   geom_hex(data = total_winter_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 1000)) +
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 25)) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+winter2_heatmap_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_winter2_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 25)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
   labs(x = expression(Longitude*degree*S)) +
@@
   geom_hex(data = total_spring_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_viridis_c(aesthetics = "fill", option = "C", limits = c(1, 11000)) +
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+spring2_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring2_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+spring_heatmap_subsurface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring_subsurface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+spring2_heatmap_subsurface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring2_subsurface, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+spring_heatmap_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 25)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+spring2_heatmap_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring2_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 25)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+```
+
+## Surface distribution
+
+The general direction of particles was north-westward with the exception of winter where direction changed to predominately southward for both years.
+Entrainment within eddies was evident during times of high mesoscale activity, which was summer and winter.
+
+The overall particle direction during summer for both years was north-westward.
+The only visible differences between years 1 and 2 was entrainment of particles in an eddy in year 1, as well as a higher concentration of particles closer to the coast (figure 2).
+The concentration of particles in the north-westward direction was lower during  year 1 of autumn compared to year 2, as well as no entrainment patterns.
+During spring of year 1 there was a high concentration of particles flowing in the north-westward direction, and no signs of entrainment.
+During spring of year 2, the particles were almost evenly spread across the southern portion of the study domain.
+There were also two visible concentrations of particles flowing in the eastward direction.
+The first starts at approximately 18$^\circ$S32.8$^\circ$E and the other at 18$^\circ$S34$^\circ$E.
+
+```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=10, fig.width=6, fig.cap="Figure 2: Hexagonal heatmaps for all virtual surface kelp for both years and all seasons.
+The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon."}
+#leg_heatmap_surface <- get_legend(summer_heatmap_surface)
+#leg_heatmap_sink <- get_legend(summer_heatmap_sink)
+
+ggarrange(summer_heatmap_surface, summer2_heatmap_surface, autumn_heatmap_surface, autumn2_heatmap_surface, winter_heatmap_surface, winter2_heatmap_surface, spring_heatmap_surface, spring2_heatmap_surface, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_2.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+
+```
+
+## Subsurface distribution
+
+Differences in concentration patterns for both years and across seasons were evident when comparing subsurface virtual kelp particle distributions.
+In general, particles were located near the north-western part of the study domain with only slight variation across seasons and between years, with the exception of winter and autumn of year 2; where particles were located in the highest concentrations in the south-western portion of the study domain.
+
+During the summer of year 1, particles were located between the coast and where the continental shelf edge ends while in year 2 particles were located further east of the shelf edge and closer to the coast.
+Clear entrainment of particles was visible in year 1, with no entrainment during the summer of year 2.
+No differences in particle concentrations was evident for Autumn of both years, with only slight variation in the spread of particles.
+During Autumn of year 1, particles were located close to the coast and spread eastwards with an even distribution.
+In Autumn of year 2, particles were located further from the coast compared to year 1 and limited to a narrow band of particles.
+The overall direction of particles and entrainment patterns were different for winter of both years compared to other seasons, with particle direction towards the south-west of the study domain.
+During winter of year 1, particles were initially concentrated in the north-western and south-western portions of the study domain, the latter with higher concentrations.
+The high concentration of particles in the south-western portion also reflected eddy entrainment patterns.
+The concentration of particles which exited the eddy on the northern side spread evenly.
+For winter of year 2, particles were located south of the peninsula in highly concentrated, limited patches.
+
+There were clear differences between particle distribution patterns between years for Spring.
+For Spring of year 1, particles flowed in a north-westward direction evenly and become concentrated in the most north-western portion of the study domain.
+The even distribution occurred over the wider portion of the underlying bathymetry.
+In year 2, particles were highly concentrated in the most south-eastern portion of the study domain.
+
+```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=10, fig.width=6,fig.cap="Figure 3: Hexagonal heatmaps for all virtual sub-surface kelp for both years and all seasons.
+The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon."}
+#leg_heatmap_surface <- get_legend(summer_heatmap_surface)
+#leg_heatmap_sink <- get_legend(summer_heatmap_sink)
+
+ggarrange(summer_heatmap_subsurface, summer2_heatmap_subsurface, autumn_heatmap_subsurface, autumn2_heatmap_subsurface, winter_heatmap_subsurface, winter2_heatmap_subsurface, spring_heatmap_subsurface, spring2_heatmap_subsurface, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_3.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+```
+
+## Sunken distribution
+
+Differences in sink locations and concentration of particles was evident between years and across seasons.
+For summer of year 1, particles sank along the bathymetry in a north-westward direction, while in year 2 particles were concentrated to a patch of particles in the most northern part of the study domain.
+During Autumn of year 1, particles sunk close to the coastline with the highest concentration of particles to the north of the study domain.
+For Autumn of year 2, particles were located further offshore with lower concentrations to the north of the study domain.
+For winter of year 1, particles sank close to the coast in the northern part of the study domain with a lower concentration of particles extending southwards along the coast.
+For winter of year 2, particles were concentrated along the eastern part of the coastline and the study domain.
+Particles sunk in the most north-western part of the study domain for Spring of year 1.
+For year 2, particles were spread out to the south-western portion of the study domain, with no clear differences inconcentrations evident.
+
+```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=10, fig.width=6, fig.cap="Figure 4: Hexagonal heatmaps for all virtual sunken kelp for both years and all seasons.
+The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon."}
+#leg_heatmap_surface <- get_legend(summer_heatmap_surface)
+#leg_heatmap_sink <- get_legend(summer_heatmap_sink)
+
+ggarrange(summer_heatmap_sink, summer2_heatmap_sink, autumn_heatmap_sink, autumn2_heatmap_sink, winter_heatmap_sink, winter2_heatmap_sink, spring_heatmap_sink, spring2_heatmap_sink, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_4.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+```
+
+## Beached distribution
+
+No clear differences, only variation, was evident for beaching locations and concentrations across all seasons and years, with the exception of winter.
+The slight variation was between the summer of year 1 and year 2, where fewer concentrations of particles located to the north of St.
+Helena.
+In year 2 particles, a higher concentration of particles were located north of St.
+Helena Bay compared to year 1.
+For winter, no differences were evident in distributions between years, with the highest concentration of particles located around the southern tip of the Cape Peninsula and within False Bay.
+
+```{r echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
+
+# summer
+summer_heatmap_beached <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_summer_beached, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 300)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(17, 20), ylim = c(-36, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+summer2_heatmap_beached <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_summer2_beached, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 300)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(17, 20), ylim = c(-36, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
+  guides(fill=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# autumn
+autumn_heatmap_beached <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_autumn_beached, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 300)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(17, 20), ylim = c(-36, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+autumn2_heatmap_beached <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_autumn2_beached, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 300)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(17, 20), ylim = c(-36, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# winter
+winter_heatmap_beached <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_winter2_beached, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 300)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(17, 20), ylim = c(-36, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+winter2_heatmap_beached <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_winter2_beached, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 300)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(17, 20), ylim = c(-36, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# spring
+spring_heatmap_beached <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring_beached, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 300)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(17, 20), ylim = c(-36, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Site")) +
+  #theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+spring2_heatmap_beached <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_spring2_beached, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
+  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 300)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(17, 20), ylim = c(-36, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Site")) +
+  #theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+```
+
+```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=10, fig.width=6, fig.cap="Figure 5: Hexagonal heatmaps for all virtual beached kelp for both years and all seasons.
+The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon."}
+
+#leg_heatmap_surface <- get_legend(summer_heatmap_surface)
+#leg_heatmap_sink <- get_legend(summer_heatmap_sink)
+
+ggarrange(summer_heatmap_beached, summer2_heatmap_beached, autumn_heatmap_beached, autumn2_heatmap_beached, winter_heatmap_beached, winter2_heatmap_beached, spring_heatmap_beached, spring2_heatmap_beached, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_5.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+```
+
+## Surface and sink convergence
+
+In both years and across seasons, site particle trajectories were all directed north-westward with the exception of winter.
+In winter for both years, the general direction of site trajectories was south-eastwards and closer to the peninsula.
+For both years and across seasons, the mean trajectory for St.
+Helena Bay flowed separately and did not overlap with the other site trajectories.
+With the exception of Buffelsbaai and Batsata Rock, sites within False Bay do not tend to travel outside of False Bay.
+Instead, these particles generally beached near release locations, except for winter where particles traveled and beached on the eastern side of  False Bay.
+During summer for both years, particle trajectories were more evenly spread out in the north-western direction, with little overlap between trajectories.
+In Autumn and Spring for both years, trajectories were concentrated along a narrow band that starts approximately at -34$^\circ$E.
+The most overlap of particles occurs during winter for both years, where the flow of trajectories was chaotic and directed towards the south-western portion of the study domain and eastwards along the coastline towards Betty’s Bay.
+
+```{r point distrubtion wrangle, message=FALSE, echo=FALSE, warning=FALSE}
+# summer
+
+summer_point <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  #geom_point(data = total_spring2_surface_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  geom_path(data = total_summer_surface_mean, aes(x = lon, y = lat, linetype = site)) +  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30.5)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+summer2_point <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  #geom_point(data = total_spring2_surface_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  geom_path(data = total_summer2_surface_mean, aes(x = lon, y = lat, linetype = site)) +  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30.5)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# autumn
+
+autumn_point <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  #geom_point(data = total_spring2_surface_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  geom_path(data = total_autumn_surface_mean, aes(x = lon, y = lat, linetype = site)) +  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30.5)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+autumn2_point <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  #geom_point(data = total_spring2_surface_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  geom_path(data = total_autumn2_surface_mean, aes(x = lon, y = lat, linetype = site)) +  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30.5)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# winter
+
+winter_point <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  #geom_point(data = total_spring2_surface_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  geom_path(data = total_winter_surface_mean, aes(x = lon, y = lat, linetype = site)) +  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30.5)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+winter2_point <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  #geom_point(data = total_spring2_surface_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  geom_path(data = total_winter2_surface_mean, aes(x = lon, y = lat, linetype = site)) +                   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30.5)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# spring
+
+spring_point <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  #geom_point(data = total_spring2_surface_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  geom_path(data = total_spring_surface_mean, aes(x = lon, y = lat, linetype = site)) +  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30.5)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+spring2_point <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  #geom_point(data = total_spring2_surface_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  geom_path(data = total_spring2_surface_mean, aes(x = lon, y = lat, linetype = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30.5)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.height=10,fig.width=6,fig.cap="Figure 6: The mean point distrubtion for depth of all surface virtual kelp for both years and all seasons.
+The surface kelp distributions are grouped per site and are represented by varying shapes."}
+
+ggarrange(summer_point, summer2_point, autumn_point, autumn2_point, winter_point, winter2_point, spring_point, spring2_point, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_6.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+```
+
+The distribution and overlap of sites for surface and sunken virtual kelp particles differs across seasons and varies between years.
+For all seasons and years particles released from the various sites generally sank in the same location, with Autumn of year 2 to the exception as particles were dispersed over the south-western portion of the continental shelf.
+Some sites from False Bay (Baboon rock, Miller’s point, A-frame and Kalk Bay) were absent from sink locations.
+
+Comparison of summer between years shows no differences in sink sites mostly located along the north-western part of the study domain, with only variation visible.
+In summer of year 1, particles sink in a narrow band, east of the continental shelf with more sites converging further north.
+In year 2, particles were concentrated further northward and not along the continental shelf edge.
+In both years for summer, Buffelsbaai settles near St.
+Helena and close to the coast.
+Variation in sink sites was also evident for autumn between years.
+In Autumn of year 1, most sites sank in similar locations in a narrow band close to most northern part of the coastline in the study domain.
+In year 2, particles were still located in the northern domain and were spread to the east of the coastline.
+In winter of year 1 particles sank towards the north-western part of the study domain.
+The virtual kelp from St.
+Helena Bay sank to the north and separated from the other sites.
+During winter of year 1, more sites were also present at the sink locations.
+In winter of year 2, the only kelp that sank in the study domain were from St.
+Helena Bay, with all other sites absent.
+In year 1 of spring sites sink further north-westward compared to summer of both years.
+Kelp from release sites were clustered together in the most north-western part of the study domain.
+Sites from False Bay such as Baboon rock, Miller’s point, Kalk Bay, and A-frame were absent from sink locations.
+In Autumn of year 2, kelp from the release sites were spread across the south-western portion of the study domain.
+The sunken kelp from a particular site also sank in the same location, for example St.
+Helena Bay towards the most eastern portion of the study domain, and Hout Bay near the continental shelf edge.
+
+```{r mean sink locations, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}
+
+total_summer_sink_mean <- total_summer_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_summer2_sink_mean <- total_summer2_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_autumn_sink_mean <- total_autumn_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_autumn2_sink_mean <- total_autumn2_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_winter_sink_mean <- total_winter_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_winter2_sink_mean <- total_winter2_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_spring_sink_mean <- total_spring_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_spring2_sink_mean <- total_spring2_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+# summer
+
+summer_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_summer_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-spring_heatmap_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_hex(data = total_spring_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) +  
-  scale_fill_viridis_c(aesthetics = "fill", option = "B", limits = c(1, 1000)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  ggtitle("Year 1") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+summer2_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_summer2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
   coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12)) 
-```
-
-```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=10, fig.width=6}
-leg_heatmap_surface <- get_legend(summer_heatmap_surface)
-leg_heatmap_sink <- get_legend(summer_heatmap_sink)
-
-ggarrange(summer_heatmap_surface, summer_heatmap_sink, autumn_heatmap_surface, autumn_heatmap_sink, winter_heatmap_surface, winter_heatmap_sink, autumn_heatmap_surface, spring_heatmap_sink,as_ggplot(leg_heatmap_surface), as_ggplot(leg_heatmap_sink), font.label = list(size = 12, face = "bold"), legend = c("none"), ncol = 2, nrow = 5, labels = c("AUTO"))
-
-```
-
-# Point plots
-## Depth distribution
-```{r message=FALSE, echo=FALSE, warning=FALSE}
-# summer
-summer_depth_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_summer_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
-  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  ggtitle("Year 2") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# autumn
+
+autumn_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_autumn_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Summer") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12)) 
-# autumn
-autumn_depth_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_autumn_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
-  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+autumn2_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_autumn2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Autumn") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
 # winter
+
 winter_depth_sink <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_winter_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
-  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+  geom_point(data = total_winter_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Winter") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12)) 
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+winter2_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_winter2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
 # spring
+
 spring_depth_sink <- ggplot() +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_spring_sink, aes(x = lon, y = lat, colour = KelpDepth)) +
-  scale_colour_viridis_c(aesthetics = "colour", option = "C", limits = c(1, 500)) +
+  geom_point(data = total_spring_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Spring") +
+  ggtitle("") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
         axis.title=element_text(size=12))
-```
-
-```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.height=6, fig.width=6}
-
-ggarrange(summer_depth_sink, autumn_depth_sink, winter_depth_sink, spring_depth_sink,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 2)
-
-```
-
-```{r message=FALSE, echo=FALSE, warning=FALSE, include=FALSE}
-
-# summer
-summer_age <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_summer, aes(x = lon, y = lat, colour = age)) +
-  scale_colour_viridis_b(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
+
+spring2_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_spring2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Summer") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12)) 
-# autumn
-autumn_age <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_autumn, aes(x = lon, y = lat, colour = age)) +
-  scale_colour_viridis_b(aesthetics = "colour", option = "C", limits = c(0, 1500)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Autumn") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-# winter
-winter_age <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_winter, aes(x = lon, y = lat, colour = age)) +
-  scale_colour_viridis_b(aesthetics = "colour", option = "C", limits = c(0, 1300)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Winter") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12)) 
-# spring
-spring_age <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_spring, aes(x = lon, y = lat, colour = age)) +
-  scale_colour_viridis_b(aesthetics = "colour", option = "C", limits = c(1, 1500)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Spring") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
   #theme_bw() +
@@
 
-```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.height=6, fig.width=6}
-
-ggarrange(summer_age, autumn_age, winter_age, spring_age,common.legend = TRUE, font.label = list(size = 12, face = "bold"), legend = c("bottom"), ncol = 2, nrow = 2)
-
-```
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.height=10, fig.width=6,fig.cap="Figure 7: Point distrubtion of all sunken virtual kelp for both years and all seasons.The sunken kelp distributions are grouped per site and are represented by varying shapes."}
+
+ggarrange(summer_depth_sink, summer2_depth_sink, autumn_depth_sink, autumn2_depth_sink, winter_depth_sink, winter2_depth_sink, spring_depth_sink, spring2_depth_sink, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_7.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+
+```
+
+There are clear differences between the amount of kelp that eventually sank, beached or traveled outside the study domain, between seasons with variation between years.
+During summer months, no particles traveled outside the study domain.
+The highest proportion of particles that traveled outside the domain occurred during spring months and to a lesser magnitude during autumn and winter months.
+The highest proportion of kelp particles that beached occurred during winter months, in particular year 2, where almost all the particles beached.
+Kelps released from Buffelsbaai were the only population to travel outside of False Bay during all times of the year, while the remaining sites (Baboon rock, Bachata rock, Miller’s point, and Kalk Bay) only traveled outside the bay during winter of year 2.
+Kelps released from Betty’s Bay (southern site extreme) usually become beached, while kelps from St.
+Helena Bay (northern extreme) had similar proportions of beached and sunken kelp across seasons and years.
+
+```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+
+positions_dotplot <- c("Betty's Bay","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock", "Basata Rock","Buffelsbaai","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","St.
+Helena Bay")
+
+dot_phase_summer <- ggplot(data = total_summer, aes(x = site, y = Phase, size = X)) +
+  #geom_bar(stat = "identity", position = "stack") +
+  #geom_boxplot() +
+  geom_point() +
+  scale_x_discrete(limits = positions_dotplot) +
+  xlab("Site") +
+  ylab("Phase") +
+  coord_flip() +
+  ggtitle("Year 1") +
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))+
+  labs(size = "Observations")
+
+dot_phase_autumn <- ggplot(data = total_autumn, aes(x = site, y = Phase, size = X)) +
+  #geom_bar(stat = "identity", position = "stack") +
+  #geom_boxplot() +
+  geom_point() +
+  scale_x_discrete(limits = positions_dotplot) +
+  xlab("Site") +
+  ylab("Phase") +
+  coord_flip()+
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))+
+  labs(size = "Observations")
+
+dot_phase_winter <- ggplot(data = total_winter, aes(x = site, y = Phase, size = X)) +
+  #geom_bar(stat = "identity", position = "stack") +
+  #geom_boxplot() +
+  geom_point() +
+  scale_x_discrete(limits = positions_dotplot) +
+  xlab("Site") +
+  ylab("Phase") +
+  coord_flip() +
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))+
+  labs(size = "Observations")
+
+dot_phase_spring <- ggplot(data = total_spring, aes(x = site, y = Phase, size = X)) +
+  #geom_bar(stat = "identity", position = "stack") +
+  #geom_boxplot() +
+  geom_point() +
+  scale_x_discrete(limits = positions_dotplot) +
+  xlab("Site") +
+  ylab("Phase") +
+  coord_flip() +
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))+
+  labs(size = "Observations")
+
+dot_phase_summer2 <- ggplot(data = total_summer2, aes(x = site, y = Phase, size = X)) +
+  #geom_bar(stat = "identity", position = "stack") +
+  #geom_boxplot() +
+  geom_point() +
+  scale_x_discrete(limits = positions_dotplot) +
+  xlab("Site") +
+  ylab("Phase") +
+  coord_flip() +
+  ggtitle("Year 2") +
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))+
+  labs(size = "Observations")
+
+dot_phase_autumn2 <- ggplot(data = total_autumn2, aes(x = site, y = Phase, size = X)) +
+  #geom_bar(stat = "identity", position = "stack") +
+  #geom_boxplot() +
+  geom_point() +
+  scale_x_discrete(limits = positions_dotplot) +
+  xlab("Site") +
+  ylab("Phase") +
+  coord_flip() +
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))+
+  labs(size = "Observations")
+
+dot_phase_winter2 <- ggplot(data = total_winter2, aes(x = site, y = Phase, size = X)) +
+  #geom_bar(stat = "identity", position = "stack") +
+  #geom_boxplot() +
+  geom_point() +
+  scale_x_discrete(limits = positions_dotplot) +
+  xlab("Site") +
+  ylab("Phase") +
+  coord_flip() +
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))+
+  labs(size = "Observations")
+
+dot_phase_spring2 <- ggplot(data = total_spring2, aes(x = site, y = Phase, size = X)) +
+  #geom_bar(stat = "identity", position = "stack") +
+  #geom_boxplot() +
+  geom_point() +
+  scale_x_discrete(limits = positions_dotplot) +
+  xlab("Site") +
+  ylab("Phase") +
+  coord_flip() +
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) +
+  labs(size = "Observations")
+```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.height=10, fig.width=8, fig.cap="Figure 8: Porportion of particles for each phase at the end of the simulation run for each respective season and year.
+The size of the point is related to the amount of particles within that life cycle phase at the end of each simulation run."}
+
+ggarrange(dot_phase_summer, dot_phase_summer2, dot_phase_autumn,dot_phase_autumn2, dot_phase_winter, dot_phase_winter2, dot_phase_spring, dot_phase_spring2, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_8.jpeg", plot = last_plot(), scale = 2, dpi = 300)
+```
+
+# Discussion
+
+The findings from this study deepens our understanding the role mesoscale features play in influencing macroalgal trajectory patterns and distribution.
+The importance of including drag components in relation to surface area as shown by @coppin2021 was confirmed, however, the study did not find the site-specific morphological characteristics influenced mean trajectory of virtual kelp particles.
+Instead, the inclusion of site-specific morphological characteristics were important in describing entrainment patterns of virtual kelp particles.
+In addition, the study provides evidence to support the site-specific approach used, which the authors argue should be incorporated into future studies on the subject.
+
+Previous work investigating drifting macroalgal trajectory patterns show that macroalgae tend to follow the ocean currents [@hobday2000; @fraser2011, @coppin2021] and drag components are important for characterising entrainment patterns within mesoscale oceanographic features [@coppin2021;@brooks2019; @putman2020; @beron2016; @beron2019].
+This was evident from the analysis of surface and subsurface kelp distributions, which show high concentrations of virtual kelp particles within mesoscale features during times of high activity (Austral summer and winter).
+Eddies have been identified as important surface accumulation zones for other types of flotsam [@brach2018; @dasaro2020; @dasaro2018; @onink2019].
+For example, @beron2016 used a Maxey-Riley approach to include the effects on inertia on accumulation patterns.
+The authors found that the effects of inertia was responsible for accumulation patterns within mesoscale features, and also noted the presence of other flotsam within these features during field validation work.
+Other work by @brooks2019 investigating distribution of *Sargassum* rafts in relation to radial size and windage.
+The authors found that eddies were important surface accumulation zones, and that the inclusion of windage based on radial raft size greatly increased accuracy of numerical simulations.
+
+Virtual kelp particles remain entrained when sinking which was as a result of the incorporated drag components.
+@coppin2021 found that purely Lagrangian particles remained on the boundary of mesoscale features with very little evidence of entrainment, while virtual kelp exhibited high entrainment patterns and focused towards the centre.
+This trend has been reported in previous work by @beron2019; @miron2020).
+For example, a study by @pichel2012 used a combination of satellite and aircraft observations to detect and describe ‘ghost gear’ distribution within the Gulf of Alaska.
+The authors found that eddies, particularly in the central cores, can collect and retain ghost gear and highlighted the role eddies play in flotsam accumulation on the surface.
+In addition to acting as accumulation zones, eddies also disperse the kelp particles over the study domain and towards the coastline when ‘leaking’ of virtual kelp from an eddy occured.
+Some of the entrained virtual kelp particles frequently exit on the eastern side of mesoscale features.
+This resulted in kelp particles becoming entrained within the Benguela Jet to travel northwest or to become entrained to other passing mesoscale activity, and or eventually beaching.
+Although the leaking of particles was less common than continued entrainment, particle leakage from mesoscale features was a significant influencer of virtual kelp particle dispersal over the study domain.
+The leaking of particles is most likely a result of the underlying bathymetry.
+A study by @rubio2009 which aimed to characterise and validate mesoscale activity in the Southern Benguela Upwelling region found similar trends with regards to volume transport of nutrients.
+The authors showed that eddies (which extend vertically through the water column) interacted with the underlying bathymetry on the coastal side of the feature which results in weaker dynamics.
+These weaker dynamics on one side of the mesoscale feature resulted in volumes of seawater particles exiting on the weaker dynamic side.
+Although the study by @rubio2009 investigated fluid parcels of water and not biological flotsam, the mechanism of weaker dynamics explains the leaking of virtual kelp particles from mesoscale features within the study domain.
+
+Eddies have also been shown to act as dynamic upwelling and downwelling of volume transport in the open ocean.
+Downwelling occurs on the periphery of mesoscale features.
+These dynamic ocean features are regarded as important mechanisms of volume transport.
+Similar relationships between virtual kelp particles and mesoscale features was evident from the analyses of surface and subsurface patterns and distributions within the study domain.
+Virtual kelp particles located towards the centre of an eddy continue to remain entrained while sinking.
+Since eddies are structured vertically through the water column, this aided in the vertical transport and distribution towards the seafloor.
+Work by @mcgillicuddy2003 using an eddy-resolving model to investigate the role of mesoscale features as sources and sinks of nutrients in the North Atlantic.
+The authors showed that the mesoscale features has a significant influence of nutrients within the euphotic zone.
+In fact, the downwelling of mesoscale features is of such high magnitudes that it can counter-balance Ekman driven upwelling in the subpolar gyre.
+Although virtual kelp particles and nutrients, similar patterns are seen in the results for the sub-surface and sunken kelp particles.
+High concentrations of kelp particles are located in areas on the seafloor which coincide with subsurface mesoscale entrainment patterns.
+Generally, kelp particles settle on the seafloor in the northern part of the study domain inshore of the continental shelf; the exception being spring of year 2.
+The distribution of sunken virtual kelp coincides with area where kelp has been found previously.
+@woodborne1989 characterised kelp-rafted geological clasts along the West Coast cites previous work by @gentle1987 who described a possible kelp sink location.
+@ Gentle describes kelp-rafted clasts that were recovered from dredging an area approximately 120km northwest of  Cape Columbine at an approximate depth of 350m deep.
+The dredged samples consisted of sand and pebbles held together by a loose mat of seaweed.
+Unfortunately, @gentle1987 did not describe the species of seaweed but simply as kelp, but based on the most common species of kelp in South Africa, the species could be *E.
+maxima*, *Ecklonia radiata* , *Laminaria pallida* and *Macrocystis Pyrifera*.
+Mesoscale features within the study domain also influence the proportions of virtual kelp particles which beach by acting as barriers to offshore transport.
+Eddy-filaments which originated offshore and occasionally travel shorewards entrain nearby kelp particles which trajectoires become largely influenced by the eddy-filament.
+The trajectory of the kelp particle is therefore determined by the trajectory of the eddy-filament which was either north-westward or shoreward causing higher proportions of beaching during the winter months of both years.
+Mesoscale and sub-mesoscale features have been shown to act as barriers to offshore transport of other flotsam.
+
+## Conclusion
+
+Macro-algae are prone to entrainment within mesoscale features, such as eddies and the associated eddy-filaments, if present but otherwise follow the ocean currents.
+The inclusion of site-specific morphological characteristics allowed for a quantitative approach to estimating windage, hydrodynamic drag, buoyancy and sink rates, all of which are important parameters for accurately characterising the distribution of macroalgae in the ocean.
+In this experiment, the results verify findings from other studies investigating eddies as convergence zones for flotsam but highlights a possible barrier to macralagal transport that eddy associated filaments and fronts may pose.
+Furthermore, the current work highlights the role of kelp in the overall carbon/nutrient cycle within the oceans by identifying possible ‘kelp’ sinks’ in the north-western and south-western areas of the Benguela Upwelling region.
+
+# Appendix
+
+```{r chp4 figure 9, message=FALSE, echo=FALSE, warning=FALSE, include=FALSE}
+## Year 1
+vector_summer_surface <- ggplot() +
+  geom_line(data = total_summer_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_summer_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)") 
+vector_summer_subsurface <- ggplot() +
+  geom_line(data = total_summer_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_summer_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)")
+
+vector_autumn_surface <- ggplot() +
+  geom_line(data = total_autumn_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_autumn_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)") 
+vector_autumn_subsurface <- ggplot() +
+  geom_line(data = total_autumn_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_autumn_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)")
+
+vector_winter_surface <- ggplot() +
+  geom_line(data = total_winter_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_winter_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)") 
+vector_winter_subsurface <- ggplot() +
+  geom_line(data = total_winter_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_winter_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)")
+
+vector_spring_surface <- ggplot() +
+  geom_line(data = total_spring_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_spring_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)") 
+vector_spring_subsurface <- ggplot() +
+  geom_line(data = total_spring_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_spring_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)")
+
+## Year 2
+
+vector_summer2_surface <- ggplot() +
+  geom_line(data = total_summer2_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_summer2_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)") 
+vector_summer2_subsurface <- ggplot() +
+  geom_line(data = total_summer2_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_summer2_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)")
+
+vector_autumn2_surface <- ggplot() +
+  geom_line(data = total_autumn2_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_autumn2_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)") 
+vector_autumn2_subsurface <- ggplot() +
+  geom_line(data = total_autumn2_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_autumn2_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)")
+
+vector_winter2_surface <- ggplot() +
+  geom_line(data = total_winter2_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_winter2_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)") 
+vector_winter2_subsurface <- ggplot() +
+  geom_line(data = total_winter2_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_winter2_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)")
+
+vector_spring2_surface <- ggplot() +
+  geom_line(data = total_spring2_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_spring2_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)") 
+vector_spring2_subsurface <- ggplot() +
+  geom_line(data = total_spring2_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
+  geom_point(data = total_spring2_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  guides(shape=guide_legend(title="Sites")) +
+  xlab("Date") +
+  ylab("Kelp Velocity (m/s)")
+
+```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Figure 9: Time-series of the mean velocity of surface kelp particles grouped by site for each season and year.", fig.height=10, fig.width=8}
+
+ggarrange(vector_summer_surface, vector_summer2_surface, vector_autumn_surface, vector_autumn2_surface, vector_winter_surface, vector_winter2_surface, vector_spring_surface, vector_spring2_surface, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("AUTO"))
+
+ggsave(filename = "figures/chapter_4/figure_9.jpeg", plot = last_plot(), scale = 0.8, dpi = 300)
+```
+
+```{r chp4 figure 10 wrangle, message=FALSE, echo=FALSE, warning=FALSE, include=FALSE}
+
+positions_dotplot <- c("Betty's Bay","Kalk Bay", "A-Frame", "Miller's Point", "Baboon Rock", "Basata Rock","Buffelsbaai","Scarborough", "Soetwater", "Kommetjie","Hout Bay","Oudekraal","St.
+Helena Bay")
+
+dis_plot_summer <- ggplot()+
+  geom_boxplot(data = total_summer, aes(x = site, y = distance, show.legend = FALSE), notch = TRUE,        notchwidth = 1, stat = "boxplot") +
+  scale_x_discrete(limits = positions_dotplot)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Sites")+
+  coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5)) 
+dis_plot_summer2 <- ggplot()+
+  geom_boxplot(data = total_summer2, aes(x = site, y = distance, show.legend = FALSE), notch = TRUE,        notchwidth = 1, stat = "boxplot") +
+  scale_x_discrete(limits = positions_dotplot)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Sites")+
+  coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+dis_plot_autumn <- ggplot()+
+  geom_boxplot(data = total_autumn, aes(x = site, y = distance, show.legend = FALSE), notch = TRUE,        notchwidth = 1, stat = "boxplot") +
+  scale_x_discrete(limits = positions_dotplot)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Sites")+
+  coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5)) 
+dis_plot_autumn2 <- ggplot()+
+  geom_boxplot(data = total_autumn2, aes(x = site, y = distance, show.legend = FALSE), notch = TRUE,        notchwidth = 1, stat = "boxplot") +
+  scale_x_discrete(limits = positions_dotplot)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Sites")+
+  coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+dis_plot_winter <- ggplot()+
+  geom_boxplot(data = total_winter, aes(x = site, y = distance, show.legend = FALSE), notch = TRUE,        notchwidth = 1, stat = "boxplot") +
+  scale_x_discrete(limits = positions_dotplot)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Sites")+
+  coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5)) 
+dis_plot_winter2 <- ggplot()+
+  geom_boxplot(data = total_winter2, aes(x = site, y = distance, show.legend = FALSE), notch = TRUE,        notchwidth = 1, stat = "boxplot") +
+  scale_x_discrete(limits = positions_dotplot)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Sites")+
+  coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+dis_plot_spring <- ggplot()+
+  geom_boxplot(data = total_spring, aes(x = site, y = distance, show.legend = FALSE), notch = TRUE,        notchwidth = 1, stat = "boxplot") +
+  scale_x_discrete(limits = positions_dotplot)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Sites")+
+  coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5)) 
+dis_plot_spring2 <- ggplot()+
+  geom_boxplot(data = total_spring2, aes(x = site, y = distance, show.legend = FALSE), notch = TRUE,        notchwidth = 1, stat = "boxplot") +
+  scale_x_discrete(limits = positions_dotplot)+
+  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
+  theme_bw()+
+  ylab("Distance (km)")+
+  xlab("Sites")+
+  coord_flip() +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+```{r chp4 figure 10, message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Figure 11: Boxplots of the total distance travelled by all particles for both years and all seasons.
+The lower and upper hinges correspond to the first and third quartiles.
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.", fig.height=9, fig.width=8}
+
+ggarrange(dis_plot_summer, dis_plot_summer2, dis_plot_autumn, dis_plot_autumn2, dis_plot_winter, dis_plot_winter2, dis_plot_spring, dis_plot_spring2, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_10.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+```
+
+```{r chp4 figure 11 wrangle, message=FALSE, echo=FALSE, warning=FALSE, include=FALSE}
+
+# year 1
+passive_heatmap_surface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_passive1_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) + 
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  facet_wrap("Season") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+passive_heatmap_subsurface <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_passive1_subsurface, bins = 80, aes(x = lon, y = lat), legend = TRUE) + 
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  facet_wrap("Season") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+passive_heatmap_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_passive1_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) + 
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  facet_wrap("Season") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+passive_heatmap_beached <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_passive1_beached, bins = 80, aes(x = lon, y = lat), legend = TRUE) + 
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  facet_wrap("Season") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# year 2
+passive_heatmap_surface2 <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_passive2_surface, bins = 80, aes(x = lon, y = lat), legend = TRUE) + 
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  facet_wrap("Season") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+passive_heatmap_subsurface2 <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_passive2_subsurface, bins = 80, aes(x = lon, y = lat), legend = TRUE) + 
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  facet_wrap("Season") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+passive_heatmap_sink2 <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_passive2_sink, bins = 80, aes(x = lon, y = lat), legend = TRUE) + 
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  facet_wrap("Season") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+passive_heatmap_beached2 <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_hex(data = total_passive2_beached, bins = 80, aes(x = lon, y = lat), legend = TRUE) + 
+  scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  facet_wrap("Season") +
+  ggtitle("") +
+  guides(fill=guide_legend(title="Number of particles")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+```
+
+```{r chp4 figure 11, message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Figure 11: Hexagonal heatmaps for all  passive surface particles for year 1 and all seasons.
+The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon.", fig.height=10, fig.width=8}
+
+ggarrange(passive_heatmap_surface, passive_heatmap_subsurface, passive_heatmap_sink, passive_heatmap_beached, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = FALSE, ncol = 2, nrow = 2, labels = c("Surface","Subsurface","Sunk","Beached"))
+
+ggsave(filename = "figures/chapter_4/figure_11.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+```
+
+```{r chp4 figure 12, message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Figure 13: Hexagonal heatmaps for all  phases passive surface particles for year 2 and all seasons.
+The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon.", fig.height=10, fig.width=8}
+
+ggarrange(passive_heatmap_surface2, passive_heatmap_subsurface2, passive_heatmap_sink2, passive_heatmap_beached2, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = FALSE, ncol = 2, nrow = 2, labels = c("Surface","Subsurface","Sunk","Beached"))
+
+ggsave(filename = "figures/chapter_4/figure_12.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+```
+
+```{r vector maps, eval=FALSE, message=FALSE, echo=FALSE, include=TRUE}
+
+chapter_4_sitemap <- rasterGrob(as.raster(readJPEG("figures/chapter_4/chapter_4_vectormap_v2.jpeg")))
+
+ggarrange(chapter_4_sitemap)
+
+```

### 2021-10-31 thesis_chapter_4_datawrangle.R

@@ -1,0 +1,1600 @@
+# script for initial data wrangle for chapter 4. this script will be used to create seperate datasets
+#that will be used for the main script ("thesis_chapter_4.rmd"
+
+houtbay_summer <- tidync("OceanParcels/chapter_4_output/hb_summer1.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, Phase, KelpDepth, SinkSpeed, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2017-12-01 00:00:00"))
+houtbay_summer$season <- "Summer"
+houtbay_summer$site <- "Hout Bay"
+houtbay_summer_mean <- houtbay_summer %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+houtbay_summer_mean$season <- "Summer"
+houtbay_summer_mean$site <- "Hout Bay"
+
+houtbay_autumn <- tidync("OceanParcels/chapter_4_output/hb_autumn1.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age, Phase,KelpDepth, SinkSpeed, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2017-12-01 00:00:00"))
+houtbay_autumn$season <- "Autumn"
+houtbay_autumn$site <- "Hout Bay"
+houtbay_autumn_mean <- houtbay_autumn %>% # calculate mean trajectory
+  group_by(t) %>%
+  summarise(across(lon:bathy, mean, na.rm= TRUE))
+houtbay_autumn_mean$season <- "Autumn"
+houtbay_autumn_mean$site <- "Hout Bay"
+
+houtbay_winter <- tidync("OceanParcels/chapter_4_output/hb_winter1.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, Salinity, age,Phase,KelpDepth, SinkSpeed,KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, bathy) %>%
[…]

### 2021-10-31 chapter_3_animations.R
animations

@@ -1,0 +1,652 @@
+# use chapter 3 script to load the data, packages specific to the animation will be loaded here
+
+library(ggplot2)
+library(gganimate)
+library(gifski)
+
+
+waterdrag_plot <- ggplot()+
+  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black') +
+  geom_point(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black') +
+  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'blue') +
+  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'orange') +
+  geom_path(data = Max_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'red') +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_bw() +
+  transition_time(t)
+
+######################################## Animations for the density plots ################################################
+## Coastline of Southern Africa
+load("map/south_africa_coast_hi_res.Rdata")
+
+ani_heatmap_RK4 <- ggplot() +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 1.0, colour = "black", fill = NA) +
+  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
+  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
+  scale_fill_viridis_c(name = "Density", limits = c(0, 1000)) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  xlab("Longtitude") +
+  ylab("Latitude") +
+  ggtitle("Passive") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  transition_time(t)
+
+anim_save("OceanParcels/animations/ani_heatmap_RK4.gif", ani_heatmap_RK4)
+
+ani_heatmap_Min_Kfloat_H100_W00 <- ggplot() +
[…]

### 2021-10-31 chapter_4_animations.R

@@ -1,0 +1,238 @@
+# use chapter 4 script to load the data, packages specific to the animation will be loaded here
+library(ggplot2)
+library(gganimate)
+library(gifski)
+
+load("map/south_africa_coast_hi_res.Rdata")
+load("map/sa_bathy.RData")
+
+total_summer <- total_summer %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+summer_kelp_ani <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  #geom_hex(data = total_summer, bins = 80, aes(x = lon, y = lat), legend = TRUE) +
+  geom_point(data = total_summer, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  #scale_fill_viridis_c(aesthetics = "fill", option = "C") +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  labs(title = 'Hours: {frame_time}') +
+  guides(shape=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) +
+  theme(legend.position = "bottom") +
+  transition_time(age)
+
+animate(summer_kelp_ani, device = 'png', duration = 60)
+anim_save("OceanParcels/animations/chapter_4/summer_kelp_ani.gif", summer_kelp_ani)
+
+
+total_summer2 <- total_summer2 %>%
+  mutate(t = as.Date.POSIXct(t, origin = "2017-12-01 00:00:00", format ="%Y-%m-%d %H:%M:%OS"))
+
+summer2_kelp_ani <- ggplot() +
[…]

### 2022-08-31 thesis_chapter_2.Rmd
uploaded updated analysis files

--- 2021-03-12 thesis_chapter_2.Rmd
@@
 ---
-title: Using numerical modelling to investigate temperature and wave exposure as drivers
-  of morphological characteristics of the kelps, Ecklonia maxima and Laminaria pallida
+title: 'Investigating waves and temperature as drivers of kelp morphology'
 author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
@@
     fig_caption: yes
-    latex_engine: xelatex
+    latex_engine: tectonic
   html_document:
@@
   word_document:
-    reference_docx: markdown/frontiers_template.docx
+    reference_docx: markdown/thesis_template.docx
 bibliography: references.bib
@@
 tables: yes
-abstract: Macroalgal morphological variation is determined to a large extent by a combination of environmental factors, with wave exposure and temperature perhaps the main influences, as they are key environmental properties to which a species becomes locally adapted.
+abstract: 'Macroalgal morphological variation is determined to a large extent by a combination of environmental factors, with wave exposure and temperature perhaps the main influences, as they are key environmental properties to which a species becomes locally adapted.
 Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction in overall size and strength increasing traits.
@@
 Kelps generally have a dependence on cool-temperate and arctic seawater temperatures [@Bolton2010; @Mohring2014-bk; @Cavanaugh2011-ib; @Dayton1985], and dominate the nearshore biomass within the rocky shallow coasts in both hemispheres [@Steneck2002].
-Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and sub-canopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002], each depending on a wide suite of ecological services provided by kelp forests [@Dayton1985; @Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @Anderson1997; @Willis2003-xo].
-Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in their distribution [@Gorman2013-nj], abundance [@Dayton1998; @Cavanaugh2011-ib], diversity [@Wing2007-tv; @Wernberg2008], composition [@Dayton1985; @Leliaert2000a; @Harley2012; @Norderhaug2012-ks], growth [@Cousens1982-ck] and productivity [@Dayton1998;@Krumhansl2012; @Pedersen2012].
-
-Given that kelps are cool-temperate organisms which are vulnerable to dislodgement under high wave exposure scenarios, increases in storm frequency and magnitude, and changes in ocean temperature, pose a direct threat to their survival; which is a concern considering that these variables are known to be affected by climate change [@Russo2014; @Reguero2019; @Weisse2010; @Rose2012; @Rummukainen2012].
-Warmer temperatures reduces the resilience of kelp individuals to storm disturbance [@Wernberg2010b], causes fragmentation through weakening tissues [@Simonson2015] and reduces growth and productivity [@bearham2013; @gerard1997; @gao2013; @zimmerman1986]; while storms and high wave energy dislodge and break kelps [@byrnes2011; @graham2004; @seymour1989].
+Their size and complex morphology provide a heterogeneous habitat structure [@Steneck2002] that accommodate a multitude of turf and sub-canopy seaweed species, and diverse assemblages of sessile and mobile invertebrates and vertebrates [@Mann1973; @Steneck2002], each depending on a wide suite of ecological services provided by kelp forests [@Dayton1985; @Gaines1987-bg; @Paul1993-bm; @Levin1994-qf; @anderson1997; @Willis2003-xo].
+Wave exposure and temperature are regarded as important environmental drivers of kelp forests, and play a role in their distribution [@Gorman2013-nj], abundance [@Dayton1998-xw; @Cavanaugh2011-ib], diversity [@Wing2007-tv; @Wernberg2008], composition [@Dayton1985; @Leliaert2000a; @Harley2012; @Norderhaug2012-ks], growth [@Cousens1982-ck] and productivity [@Dayton1998-xw;@Krumhansl2012; @Pedersen2012a].
+
+Given that kelps are cool-temperate organisms which are vulnerable to dislodgement under high wave exposure scenarios, increases in storm frequency and magnitude, and changes in ocean temperature, pose a direct threat to their survival; which is a concern considering that these variables are known to be affected by climate change [@Weisse2010;@Rose2012; @Rummukainen2012;@Russo2014; @Reguero2019].
+Warmer temperatures reduces the resilience of kelp individuals to storm disturbance [@Wernberg2010b], causes fragmentation through weakening tissues [@Simonson2015] and reduces growth and productivity [@zimmerman1986; @gerard1997; @bearham2013; @gao2013]; while storms and high wave energy dislodge and break kelps[@seymour1989; @graham2004; @byrnes2011].
 Despite these threats and disturbances kelps persist across a broad range of environments which is largely due to their morphological plasticity [@FowlerWalker2006; @Wernberg2010a].
-A study by [@Wernberg2003-mp] investigated the morphology of Ecklonia radiata (C.Agardh) J.Agardh in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australian coast.
+A study by @Wernberg2003-mp investigated the morphology of *Ecklonia radiata* (C.Agardh) J.Agardh in order to quantify the morphological variation and whether it was dependent on spatial differences along the Australian coast.
 They found no correlation between spatial distance and morphological similarity; rather the morphology of the kelps was representative of multiple environmental forcings on different morphological characters at different spatial scales [@Wernberg2003-mp].
 
-Although multiple environmental factors play a role in kelp morphological adaptation, wave exposure and temperature have been identified as the main influencers across various kelp species [@Hurd2000;@Denny2002;@Thomsen2004;@FowlerWalker2005;@Wernberg2005c;@mabin2013;@Bekkby2014].
-For example, a study by  @Serisawa2002-d compared the morphology of Ecklonia cava Kjellman growing at warmer and cooler sites, and showed wrinkling of the blade and reduced size to be a characteristic of higher temperature ranges.
-Warmer temperature affects important kelp physiological processes such as photosynthesis and respiration, which influences growth and productivity and ultimately leads to a reduction in size [@bearham2013; @gerard1997; @gao2013; @zimmerman1986].
-Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility [@Denny2002; @Hurd2000; @FowlerWalker2005; @Thomsen2004].
-For example, a study by @Wernberg2005c examined the consistency of wave exposure as a driver of E.
-radiata across a broad geographic range, and showed trends toward drag-reducing (small size, narrow laterals and blades, low spinosity) and increased strength (large holdfast, thick stipe and thick blades and lamina) at high wave exposure sites.
-
+Although multiple environmental factors play a role in kelp morphological adaptation, wave exposure and temperature have been identified as the main influencers across various kelp species [Hurd2000; @Denny2002a; @Thomsen2004; @FowlerWalker2005; @Wernberg2005c; @mabin2013; @bekkby2014].
+For example, a study by @Serisawa2002-dc compared the morphology of *Ecklonia cava* Kjellman growing at warmer and cooler sites, and showed wrinkling of the blade and reduced size to be a characteristic of higher temperature ranges.
+Warmer temperature affects important kelp physiological processes such as photosynthesis and respiration, which influences growth and productivity and ultimately leads to a reduction in size [@zimmerman1986; @gerard1997; @bearham2013; @gao2013].
+Past research has shown that in highly wave exposed areas kelp morphology tends to take on characteristics which reduce drag, increase strength and increase flexibility (@Hurd2000; @Denny2002a; @Thomsen2004; @FowlerWalker2005].
+For example, a study by @Wernberg2005c examined the consistency of wave exposure as a driver of *E.
+radiata* across a broad geographic range, and showed trends toward drag- reducing (small size, narrow laterals and blades, low spinosity) and increased strength (large holdfast, thick stipe and thick blades and lamina) at high wave exposure sites.
 Measures of wave exposure in ecological studies often incorporate integrative measures of hydrodynamic conditions at a particular site which are based on cartographical models of exposure.
@@
 Due to the complex effects of environmental drivers on kelp morphology, one can also expect differences in morphology between deep and shallow water populations of the same species.
-Deep and shallow water environments may differ in abiotic processes such as the degree of water column mixing [@Smit2013a], solar heating [@Dellatorre2012; @Dunne2001], and effects of wave dampening [@Kobayashi1993-os; @Mia2013; @Dubi1995].
-
+Deep and shallow water environments may differ in abiotic processes such as the degree of water column mixing [@Smit2013a], solar heating [@Dunne2001; @Dellatorre2012], and effects of wave dampening [@Kobayashi1993-os; @Dubi1995; @ma2013].
 The kelps *Laminaria pallida* Greville and *Ecklonia maxima* (Osbeck) Papenfuss are important habitat forming seaweeds that exist around the coast of South Africa and offer a unique opportunity to investigate the drivers of macroalgal morphological characteristics between canopy and sub-canopy species, as well as between deep and shallow water environments.
@@
 maxima* forms dense surface canopies in the shallow sub-littoral.
-Therefore, E.
-maxima may be exposed to variations in wind driven surface waves, swell waves and temperature, while *L.
+Therefore, *E.
+maxima* may be exposed to variations in wind driven surface waves, swell waves and temperature, while *L.
 pallida* is exposed to variations in swell waves, wave exposure and temperature due to differences in depth.
@@
 pallida* and *E.
-maxima* along the South African coastline and into Namibia; *E.
-maxima* exhibited no morphological changes along the coast but the stipes of *L.
-pallida* become increasingly hollow further north along the coastline.
+maxima* along the South African coastline and into Namibia; _E.
+maxima _exhibited no morphological changes along the coast but the stipes of  *L.pallida* become increasingly hollow further north along the coastline.
 The authors suggested that turbidity in relation to light attenuation was the environmental driver responsible for this change.
@@
 The numerical model will then be used as a basis for investigating the consequences for morphological characteristics of *E.
-maxima* and *L.
-pallida* using advanced statistical tools.
+maxima* and *L.pallida* using advanced statistical tools.
 
@@
 
-Sites were chosen to represent an array of temperature and wave gradients, as indicated in Figure \ref{fig:map}.
+Sites were chosen to represent an array of temperature and wave gradients, as indicated in Figure 1.
 St.
 Helena Bay and Betty’s Bay constituted the north-western and south-eastern boundary sites, respectively.
-These sites are roughly 300 km apart, lie within separate marine provinces  @Smit2017, and span the majority of the south-west coast, in varying thermal and wave energy regimes.
+These sites are roughly 300 km apart, lie within separate marine provinces [@Smit2017], and span the majority of the south-west coast, in varying thermal and wave energy regimes.
 The region is dominated by kelp communities that persist in contrasting abiotic environments.
-The west coast region has been termed cool-temperate, which is defined as a region where mean monthly temperatures are always above 10°C and below 15°C [@Smit2013a].
-East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone [@Smit2013a].
-The Agulhas Marine Province is characterized by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm-temperate [@Smit2013a].
+The west coast region has been termed cool-temperate, which is defined as a region where mean monthly temperatures are always above 10◦C and below 15◦C  [@Smit2013a.
+East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone [@Smit2013a.
+The Agulhas Marine Province is characterized by a wide temperature range of up to 7◦C difference between mean monthly temperatures between summer and winter and is classified as warm-temperate  [@Smit2013a.
 
@@
 library(ggplot2)
-library(rgdal)
 library(tidyverse)
@@
 The annual and seasonal patterns in both wind and swell direction influence the thermal regime around the coastline by inducing upwelling.
-Upwelling brings deep, cool nutrient-rich water to coastal areas which causes a decrease in temperature [@Field1980a; @Andrews1980; @Gill1974-nt; @Cram1970-jw; @Rouault2010-gi; @Blanke2002].
+Upwelling brings deep, cool nutrient-rich water to coastal areas which causes a decrease in temperature [@Field1980a; @Andrews1980; @Gill1974-nt; @Cram1970-jw; @Rouault2010-gi; @blanke2002].
 The wind speed and wind direction help drive upwelling in summer on the western side of the Cape Peninsula, where southerly winds blow parallel to the coast and trigger upwelling events [@Field1980a; @Rouault2010-gi].
@@
 Shallow water temperatures were sourced from The South African Coastal Temperature Network (SACTN) website1.
-In terms of nearshore temperature, in situ data are preferred over satellite sea surface temperature (SST), which have shown to exhibit large biases [@Smit2013a].
+In terms of nearshore temperature, *in situ* data are preferred over satellite sea surface temperature (SST), which have shown to exhibit large biases [@Smit2013a].
 The underwater temperature recorders (UTRs) are attached to concrete mooring blocks or railway line sections deployed in the shallow at ∼3 m depth.
-The UTRs were comprised of Starmon Mini recorders (Star-Oddi, Reykjavik, Iceland/Onset Hobo® U22-001) accurate to ±0.2°C.
-All in situ temperature data were reduced to a time-coded, continuous series of monthly means over various time-scales [@Smit2013a].
+The UTRs were comprised of *Starmon Mini* recorders (_Star-Oddi_, Reykjavik, Iceland/_Onset Hobo_⃝R U22-001) accurate to ±0.2◦C.
+All *in situ* temperature data were reduced to a time-coded, continuous series of monthly means over various time-scales [@Smit2013a].
 It was decided that the annual time-scale would be used for analysis.
-Linear interpolated SST were calculated for sites where in situ recorders were absent.
+Linear interpolated SST were calculated for sites where *in situ* recorders were absent.
 These data were used to group sites into “Cool temperate” and “Warm temperate” categorizations to investigate the possible effect of thermal regime on kelp morphological characteristics.
@@
 Wave model variable data formed part of the South African Coastal Vulnerability Assessment, presented to the Department of Environmental Affairs (DEA) and produced by the Council for Scientific and Industrial Research (CSIR) [@Rautenbach2015a].
-The South African wave climate was modeled via 20 spectral numerical wave models that simulated the offshore wave climate to the nearshore.
+The South African wave climate was modeled _via _20 spectral numerical wave models that simulated the offshore wave climate to the nearshore.
 The boundary conditions of these models were obtained by using the NOAA Wave Watch III (WWIII) model output, distributed via the National Centers for Environmental Prediction (NCEP) product (Office of the Director 2000; Environmental Modeling Center/Marine Modeling Branch 2005).
@@
 This allowed for a computational effective wave resolution of increasing resolution, from the NCEP, low resolution output to nested, high resolution coastal output.
+
 For Table Bay and east of Cape Hangklip the resolution was 500 m and also had output at the 7 and 15 m isobaths.
@@
 
-The morphological characteristics of both species are presented in figure \ref{fig:schematic}.
+The morphological characteristics of both species are presented in Figure 2.
 *Laminaria pallida* is characterized by a single smooth blade which is divided longitudinally into sections, and develops from a single meristematic region located at the junction between the blade and the stipe [@Dyer2018-rv].
@@
 
-Morphological measurements of L.
-pallida and E.
-maxima were collected at 18 sites along the Western Cape coast of South Africa (Figure \ref{fig:map}) between October 2014 and April 2015.
-AlgaeBase was used to confirm the latest taxonomic associations for each species [@Guiry2019].
+Morphological measurements of *L.pallida* and _E.
+maxima_ were collected at 18 sites along the Western Cape coast of South Africa (Figure 1) between October 2014 and April 2015.
+AlgaeBase was used to confirm the latest taxonomic associations for each species (Guiry and Guiry, 2019).
 Divers sampled kelp individuals by means of SCUBA for deeper populations and freediving for shallower populations.
@@
 maxima* was determined by size, presence at the surface as well as a darker coloration of the stipe; as this occurs with age.
-In the case of *L.
-pallida*, the individuals with the longest stipe were selected which is considered a reasonable estimate of age.
+In the case of *L.pallida*, the individuals with the longest stipe were selected which is considered a reasonable estimate of age.
 In each case the age of the individuals was estimated visually before the individual was sampled.
-In instances where one of the species was not present a site, the nearest population of the missing species was sampled.
+In instances where one of the species was not present at a site, the nearest population of the missing species was sampled.
 Because the macroalgae differ in morphological features, species-specific morphological characteristics were included and are presented in Table 1.
@@
 Stipe circumference was measured just above the holdfast for both species.
-Between February 2017 and November 2018, morphological measurements for E.
-maxima individuals in shallow water (<1 m) at seven sites along the Western Cape coast of South Africa were also collected.
+Between February 2017 and November 2018, morphological measurements for *E.
+maxima* individuals in shallow water (<1m) at seven sites along the Western Cape coast of South Africa were also collected.
 The same morphological characteristic measurements were taken as for the deeper *E.
 maxima*, and this allowed comparison between morphological characteristics between deep and shallow individuals within sites.
-Measurements were not collected for *L.
-pallida* in the shallow depths, as this species is largely absent from the shallow in this portion of the South African coastline.
+Measurements were not collected for *L.pallida* in the shallow depths, as this species is largely absent from the shallow in this portion of the South African coastline.
+
+```{r table 2.1, message=FALSE, warning=FALSE, echo=FALSE}
+
+library(flextable)
+
+morph_table <- read.csv("tables/chapter_2/table_1.csv", sep = ";")
+morph_table <- flextable(data = morph_table)
+morph_table <- theme_zebra(morph_table, odd_header = "transparent", even_header = "transparent")
+morph_table <- set_header_labels(morph_table, Morphological.Characteristic = "Morphological Characteristic", Unit.measurement = "Unit.measurement")
+morph_table <- align(x= morph_table, align = "center", part = "all")
+morph_table <- set_table_properties(morph_table, width = 1, layout = "autofit")
+morph_table <- set_caption(morph_table, "A list of measured environmental variables (temperature, waves and wind) and their associated abbreviations.")
+morph_table
+
+```
 
@@
 
+```{r table 2.2, message= FALSE, warning=FALSE, echo=FALSE}
+
+var_table <- read.csv("tables/chapter_2/table_2.csv", sep = ";")
+var_table <- flextable(data = var_table)
+var_table <- theme_zebra(var_table, odd_header = "transparent", even_header = "transparent")
+var_table <- align(x= var_table, align = "center", part = "all")
+var_table <- set_table_properties(var_table, width = 1, layout = "autofit")
+var_table <- set_caption(var_table, " A list of measured environmental variables (temperature, waves and wind) and their associated abbreviations.")
+var_table
+```
+
 Morphological characteristics in relation to temperature and wave exposure were investigated using non-parametric statistical methods.
@@
 Significant differences between temperature regimes and across wave exposure categories were tested with Wilcox tests and Kruskal -Wallis tests, respectively, and notched boxplots allowed comparison between individual categories of wave exposure.
-The alpha level used to define significance was p = 0.05.
-Boxplots were created using the software package ggplot2 [@Wickham2009].
-In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis (db-RDA) was performed using the rda function in the vegan software package [@Oksanen2015-vj] in R (R Core Team, 2014; RStudio Team, 2020).
+The alpha level used to define significance was *p*= 0.05.
+Boxplots were created using the software package ggplot2 [@oksanen2013].
+In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis (db-RDA) was performed using the *rda* function in the *vegan* software package [@Oksanen2015-vj] in R [@rstudio].
 The abiotic data covered all study sites for both species of kelp.
-Both morphology and abiotic data (temperature, waves and wind) were standardized using the decostand function in the vegan software package; morphology data were used as response variables and the abiotic data as explanatory variables.
+Both morphology and abiotic data (temperature, waves and wind) were standardized using the _decostand _function in the _vegan _software package; morphology data were used as response variables and the abiotic data as explanatory variables.
 A list of the abbreviations used and the associated metrics are presented in Table 2.
@@
 To further improve the model, pairwise coefficients and Variance Inflation Factor (VIF) were calculated to identify variables with high multicollinearity.
-The computation of the parsimonious RDA was followed by permutation tests of the adjusted R2 to assess significance of constraints.
+The computation of the parsimonious RDA was followed by permutation tests of the adjusted _R_2 to assess significance of constraints.
 
@@
 ```
+
 In general, the mean temperature increases for sites located within False Bay with less variation around the mean temperature compared to sites on the western side of the peninsula (see Figure \ref{fig:Temperature variables plot}A).
@@
 The remaining morphological characteristics, stipe length, number of tufts and total length had significantly lower means for kelp from cool temperate compared to kelp from the warm temperate regime (p < 0.05; see Figure 8).
-Significant differences amongst wave exposure categories were found for all morphological characteristics of deep E.
-maxima populations (see Figure \ref{fig:deep eck temp plot}
+Significant differences amongst wave exposure categories were found for all morphological characteristics of deep *E.
+maxima* populations (see Figure \ref{fig:deep eck temp plot}
 ).
@@
 ```
+
 When grouped by temperature regime, shallow *E.
-maxima* exhibited significant differences between temperature regimes for frond mass, primary length, primary width, frond length, stipe mass, stipe length, stipe circumference and total length (p < 0.05; see Figure \ref{fig:shallow ecklonia temp}).
+maxima* exhibited significant differences between temperature regimes for frond mass, primary length, primary width, frond length, stipe mass, stipe length, stipe circumference and total length (*p* < 0.05; see Figure \ref{fig:shallow ecklonia temp}).
 All of the significantly different morphological characteristics, except primary length, had lower values for the western region compared to the eastern region.
 When shallow E.
-maxima were grouped by wave exposure category significant difference were found for all morphological characteristics (p < 0.05; see Figure \ref{fig:shallow ecklonia wave}).
+maxima were grouped by wave exposure category significant difference were found for all morphological characteristics (*p* < 0.05; see Figure \ref{fig:shallow ecklonia wave}).
 In general, values of the various morphological characteristics increase for the lower wave exposure categories, followed by a sharp decrease and a gradual increase in value.
@@
 Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of L.
-pallida morphological characteristics with an adjusted R2 of 0.49, explaining 72% of the variation (global permutation test on final model: d.f = 1, F = 8.51, p = 0.008; Figure \ref{fig:rda plots}A).
+pallida morphological characteristics with an adjusted R2 of 0.49, explaining 72% of the variation (global permutation test on final model: d.f = 1, F = 8.51, *p* = 0.008; Figure \ref{fig:rda plots}A).
 RDA1 was the only significant axis in the model and explained 48% of the variation while RDA2 was not significant and explained 63% of the variation.
@@
 ```
+
 The same approaches to identify the descriptors for *L.
@@
 Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of deep *E.
-maxima* morphological characteristics with an adjusted R2 of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, F = 2.4, p = 0.026; Figure \ref{fig:rda plots}B).
-RDA1 was the only significant axis in the model (p = 0.03) and explained 72% of the variation.
+maxima* morphological characteristics with an adjusted R2 of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, F = 2.4, *p* = 0.026; Figure \ref{fig:rda plots}B).
+RDA1 was the only significant axis in the model (*p* = 0.03) and explained 72% of the variation.
 Primary length, primary width and stipe circumference were positively influenced by mean significant wave height, mean peak period and median swell direction for kelp from the western region.
@@
 ```
+
 Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retain the most parsimonious descriptors of shallow *E.
-maxima* morphological characteristics with an adjusted R2 of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, F = 2.41, p = 0.013, Figure \ref{fig:rda plots}C).
-The model consisted of one significant axis, RDA1 (p = 0.02) and explained 62% of the variation.
+maxima* morphological characteristics with an adjusted R2 of 0.36, explaining 61% of the variation (global permutation test on final model: d.f = 4, F = 2.41, *p* = 0.013, Figure \ref{fig:rda plots}C).
+The model consisted of one significant axis, RDA1 (*p* = 0.02) and explained 62% of the variation.
 Stipe length, stipe mass and stipe circumference were positively influenced by maximum temperature and mean temperature along RDA1, and explained 50 and 55% of the variation, respectively.
@@
 radiata* and *M.
-pyrifera* [@Thomsen2004; @FowlerWalker2005;@FowlerWalker2006; @Wernberg2005c; @Wing2007-tv; @Stewart2009; @Miller2011-pm; @Pedersen2012-vr].
+pyrifera* [@Thomsen2004; @FowlerWalker2005;@FowlerWalker2006; @Wernberg2005c; @Wing2007-tv; @Stewart2009; @Miller2011-pm; @Pedersen2012a].
 The complex geomorphology of the Western Cape coastline creates an ideal natural laboratory for studies of kelps in interaction with their environment.
 Unlike previous research, this study used a complex numerical model to provide the various variables of wave exposure, which is based on linear wave theory to calculate overall wave exposure.
-We also considered wind in addition to temperature and wave variables, since wind is an important component of wind-driven surface gravity waves [@Holthuijsen2010-vl].
+We also considered wind in addition to temperature and wave variables, since wind is an important component of wind-driven surface gravity waves [@Holthuijsen2010].
 The results show that specific variables of wave exposure are the main drivers of kelp morphology with temperature playing only a minor role.
@@
 Morphological adaptation due to water motion may manifest in a number of ways in high wave energy environments.
-For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment [@FowlerWalker2006; @Wernberg2005c; @Friedland1995-vi; @Denny2002; @bekkby2014; @Denny1997] have been identified in previous studies.
+For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment [@FowlerWalker2006; @Wernberg2005c; @Friedland1995-vi; @Denny2002a; @bekkby2014; @Denny1997] have been identified in previous studies.
 Although this study did not measure force of attachment, other morphological responses to temperature and wave variables were evident.
@@
 When grouped by wave exposure category, *L.
-pallida* characteristics showed a significant increase in length for the “extremely sheltered” category compared to the fully sheltered category, which may be a “go-with-the-flow” tactic [@Denny1997; @Denny2002; @Hurd2000].
-Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny1997; @Denny2002; @Hurd2000].
-However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under high wave exposure [@Denny1997; @Denny2002; @Hurd2000].
+pallida* characteristics showed a significant increase in length for the “extremely sheltered” category compared to the fully sheltered category, which may be a “go-with-the-flow” tactic [@Denny1997; @Denny2002a; @Hurd2000].
+Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny1997; @Denny2002a; @Hurd2000].
+However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under high wave exposure [@Denny1997; @Denny2002a; @Hurd2000].
 This is reflected in the *L.
@@
 
-A size reduction tactic has been shown before [@FowlerWalker2006; @Denny2002; @Hurd2000; @FowlerWalker2005] and is regarded as a strategy to reduce overall drag on the plant.
+A size reduction tactic has been shown before [@FowlerWalker2006; @Denny2002a; @Hurd2000; @FowlerWalker2005] and is regarded as a strategy to reduce overall drag on the plant.
 However, *E.
@@
 
+# References
+
 # Appendix
@@
 ```
-
-# References

### 2022-08-31 thesis_chapter_3.Rmd
uploaded updated analysis files

@@
 title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
-author: 
-- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
+author: "R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216], AJ Smit^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa, South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]"
 output:
+  pdf_document:
+    latex_engine: xelatex 
+    toc: no
   html_document:
-  pdf_document:
-    fig_caption: yes
-    latex_engine: xelatex
   word_document:
-    reference_docx: markdown/wiley_template.docx
+    reference_docx: markdown/frontiers_template.docx
 bibliography: references.bib
-csl: markdown/limnologyandoceanography.csl
-abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic.
-Macroalgae are one of the most important natural dispersal vectors  of marine organisms and are regarded as the 'tumble-weed of the ocean'.
-Despite many studies that exist on the topic, the relative role of wind and surface currents in influencing the trajectory of macroalgal dispersal is still unknown.
-Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a consenses as to the relative role of wind versus surface currents.
-In addition, these studies have not considered surface-area in relation to drag characteristics of the macroalgae, both of which have been shown to play a role in the trajectory and accumulation of other flotsam.
+csl: markdown/elsevier-harvard2.csl
+abstract: "Ocean currents act as an important dispersal mechanism of natural and anthropogenic material on the ocean surface.
+Macroalgae are one of the most important natural dispersal vectors of marine organisms and are regarded as the ‘tumble-weed of the ocean.’ Despite many studies on the topic, the relative role of wind and surface currents in influencing the trajectory of macroalgal dispersal is still uncertain.
+Past studies focused on kelp-rafts, which can vary greatly in size, making it difficult to form a consensus as to the relative role of wind versus surface currents.
+In addition, these studies have not considered surface-area in relation to drag characteristics of the macroalgae, both of which have been shown to play a role in the trajectory and accumulation of flotsam.
 Advances in Lagrangian trajectory modeling have been rapid in recent years and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean.
-This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagal trajectory and accumulation.
-We focused on solitary kelp individuals.
-This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles.
-The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddys within the study domain.
-The results also point to the importance of sub-mesoscale features, such as eddys, as important surface accumulation zones for solitary floating macroalage."
+This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macroalgal trajectory and accumulation.
+We focused on solitary kelp plants.
+This was achieved by comparing simulations of virtual kelp ‘particles’ which incorporate drag with that of purely Lagrangian particles (no drag).
+Geometric representations for each morphological characteristic was used to calculate the overall surface area used in the calculation of the relevant drag components (hydrodynamic and wind) for each scenario (submerged and partially submerged) and flow-field combinations (currents only, currents with wind, currents with Stokes drift, currents with wind and Stokes drift).
+The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain."
+nocite:  |
+  @Algaebase
+  @Worms
+header-includes:
+  - \usepackage[left]{lineno}
+  - \linenumbers
 ---
@@
 
-Passive dispersal mechanisms of marine taxa include a range of natural and anthropogenic objects that float in the ocean, and macroalgae are regarded as one of the most important of these.
-Floating kelps, sometimes referred to as the ‘tumble-weed of the ocean,’ are very important in this regard [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008].
-Kelps also function as allochthonous materials to near and distant ecosystems and in this manner support the ecological function in a range of adjacent habitats[@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].
-
-Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant.
-The positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum.
+Passive dispersal vectors of marine taxa include a range of natural and anthropogenic objects that float in the ocean, and kelp are regarded as one of the most important of these [@edgar1987;@norton1992;@bushing1994;@holmquist1994; @helmuth1994;@smith2002; @McCormick2008].
+Floating macroalgae are regarded as important transport mechanisms for marine invertebrates which remain attached to the plant when dislodged from the substratum [@edgar1987;@norton1992;@bushing1994;@holmquist1994; @helmuth1994;@smith2002; @McCormick2008].
+For example a study by @holmquist1994 investigated the long-distance dispersal of the bivalve *Gaimardia trapesina* Lamarck via drifting kelp.
+The authors used observational techniques to quantify kelp-rafts along with haphazard sampling of the kelp-rafts observed.
+Genetic samples were taken from the target species for each of the haphazardly sampled kelp-rafts for comparison amongst populations.
+The results suggest that kelp-rafting, along with the local wind patterns, most likely played a key role in the dispersal of  *G.
+trapesina* between genetically similar but spatially disconnected populations.
+Kelps ecological role in terms of organic subsidy is not limited to the local environment, and instead plays an important role in organic input into deep, offshore ecosystems [@bustamante1995; @colombini2003;@vanderklift2008;@Krumhansl2012; @filbee2018;@zuercher2019;@cartraud2021].
+For example, a study by @filbee2018, investigated the allochthonous input of kelp detritus into deep fjords within northern Norway.
+The authors used a combination of video surveys, biomass estimates and tagging of detrital material to characterise kelp detritus within shallow, sub-tidal and deep fjord systems.
+The results showed that the organic subsidy into the different systems was linked to loss of biomass from kelp forests during times of high wave exposure.
+Furthermore, the results showed that kelp detritus is an important organic subsidy for distant ecosystems, suggesting resource coupling between nearshore kelp forests and offshore systems [@filbee2018].
+The formation of kelp rafts also act as floating refuges for marine organisms as well as natural fish aggregation devices.
+Work by @witherington2012 showed the reliance of young sea turtles on drifting pelagic Sargassum rafts as a source of food and refuge from marine predators, and work by @kingsford1992 showed the reliance of some fish larvae (*Paralabrax clathratus* Girard, *Chromis punctipinnis* Cooper and Sebastes spp) on kelp-rafts for habitat complexity and transport to offshore areas.
+
+Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species have air-filled pneumatocysts or stipes which allow the plants to remain at the surface where light is more abundant, and in turn cause plants to float to the surface when detached.
 The structure and number of pneumatocysts vary among species.
-For example, species belonging to the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, and *Ecklonia maxima* have single, large pneumatocysts[@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018].
+For example, species belonging to the genera Macrocystis, Sargassum, Ascophyllum, and Fucus have thalli with many small pneumatocysts, while other species such as *Nereocystis luetkeana* (K.Mertens) Postels & Ruprecht, *Pelagophycus porra* (Léman) Setchell, *Ecklonia radiata* (C.Agardh) J.Agardh, and *Ecklonia maxima* (Osbeck) Papenfuss have single, large pneumatocysts [@dayton1985; @smith2002;@thiel2005; @graiff2016; @batista2018].
 In some cases, the stipe itself may be air-filled, such as with *E.
@@
 This is most likely because the green and red species reported to float are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in their thalli [@dromgoole1982; @back2000].
-The giant kelp *Macrocystis pyrifera* [@hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@tala2017; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @brooks2018; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
-
-Past research points to macroalgal trajectories being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011].
+The giant kelp *Macrocystis pyrifera* (Linnaeus) C.Agardh [@helmuth1994; @kingsford1995;@hobday2000;@macaya2005; @graiff2016;@batista2018], the bull kelp *Durvillaea antarctica* (Chamisso) Hariot [@smith2002; @collins2010; @wichmann2012; @tala2013; @tala2017; @saunders2014; @batista2018], and pelagic Sargassum species [@brooks2018; @brooks2019; @putman2018; @putman2020; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.
+
+Past research points to macroalgal trajectories being largely determined by  a combination of prevailing wind conditions and surface currents [@hobday2000; @thiel2005; @rothausler2011;@fraser2011].
 Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown, although the role of wind has been recognized as important in several studies.
@@
 pyrifera* in nearshore Monterey Bay.
-The results showed that kelp rafts, with little surface area exposed to the wind, were largely driven by a combination of wind and surface currents; however, the relative importance of wind and was not clear.
-In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter.
-Studies have also identified wind as an important mechanism of dispersal in wind dominated ocean systems.
-For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important mechanism for dispersal of floating bull kelp *Durvillaea antarctica* [@waters2008; @smith2002; @nikula2010].
-Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013].
-For example, a study by @fraser2011 on the rafting capabilities of *D.
-antarctica* used a combination of population genetics and relative age estimate of goose barnacles attached to the raft.
+The results showed that kelp rafts, with little surface area exposed to the wind, were largely driven by a combination of wind and surface currents; however, the relative importance of wind was not clear.
+This work also showed that the trajectories of tagged kelp were congruent with the formation of eddies during winter.
+Other studies have used genetic approaches to determine macroalgal raft trajectory characteristics by inferring source location from genetically similar populations.
+For example a study by @nikula2013 used genetic markers (mtDNA and AFLP) of invertebrate species to show that in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important mechanism for dispersal of floating bull kelp *D.
+antarctica*.
+Another study by @fraser2011 on the rafting capabilities of *D.
+antarctica* used a combination of population genetics and relative age estimate of goose barnacles attached to kelp-rafts.
 The presence of goose barnacles suggests a long rafting time as these species have a slow growth rate.
-Genetic analyses showed these species can raft up to $\sim$ 390 km from their local origin.
-The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011].
-Research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
-The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic-form drag [@breivik2011; @griffin2017].
-Work by the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object.
-If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].
-Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.
-
-To accurately determine the trajectory of marine macroalgae, both water and wind drag need to be considered.
-@allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model.
-However, this work did not consider biological material such as macroalgae.
-Although drag estimates do exist, these have been applied for macroalgae not detached from the substratum and are regarded as fixed-point estimates [@Denny1997; @Denny2002a; @Denny2002b; @Gaylord1997-ow; @Gaylord2008-qv].
+Genetic analyses showed these species can raft up to 390 km from their local origin.
+The authors suggested that wind and surface currents were the primary influences of trajectory; however, this was inferred from the genetic analyses and local climatology data [@fraser2011].
+
+Besides wind and surface currents, waves are also an important ocean feature and work by the maritime industry has shown that the size, shape, and orientation of the object of a vessel determine the relative importance of waves or wind as drivers of trajectory [@breivik2011; @griffin2016].
+In general, if the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for shorter objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2016].
+Stokes drift is a wave-driven process and previous work has identified this as an aspect that needs to be considered when investigating dispersal of macroalgal species.
+Stokes drift is the net drift velocity of a particle floating at the ocean surface in the direction of wave propagation due to the differences between the velocity of the fluid particle and flow-field [@bremer2018].
+More generally, the particle floating at the surface moves slightly faster in the direction of the wave as a result of the oscillatory motion of fluid particles within a wave.
+Stokes drift can significantly influence the trajectory of a particle or object floating at the surface of the ocean.
+For example, work by @fraser2018 included Stokes drift into the advection scheme for the Lagrangian model used to investigate the dispersal potential of the southern bull kelp *D.
+antarctica* to reach Antarctica.
+The results showed that the inclusion of wave-driven Stokes drift and mesoscale eddies suggest frequent incursions into Antarctica, which the authors attributed to storm-forced surface waves, further highlighting the role of wave-driven processes in the open ocean.
+
+Research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@furnans2008; @beron2016; @trinanes2016; @dasaro2018; @gates2018; @onink2019; @shen2019].
+Therefore, to  accurately determine the trajectory of marine kelp, both water and wind drag need to be considered.
+Drag is the force an object experiences when moving through a fluid which is proportional to the velocity of the object [@Vogel2020].
+The fluid viscosity and the friction from the surface of the object results in the overall drag component, as well as the surface area and shape of the object [@Vogel2020].
+If the object is partially submerged, as a result of buoyancy forces, then drag will be a combination of drag from the air and fluid [@Vogel2020].
+The ‘sail’ area of an object floating at sea is the surface area exposed to the wind (air) which results in air-form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic-form drag [@breivik2011; @griffin2017].
+@allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work using both direct and indirect methods of the Leeway model, However, this work did not consider biological material such as macroalgae.
+Therefore, no drag estimates exist for macroalgae floating at the surface of the ocean and alternative approaches should be used to account for the drag components.
+Although drag estimates do exist, these have been applied for kelp not detached from the substratum and are regarded as fixed-point estimates [@Denny1997; @Gaylord1997-ow; @Denny2002a; @Denny2002b;@Gaylord1997-ow].
+
 Although various approaches have been used in the past to investigate floating macroalgal trajectories, very few studies have employed the use of Langrangian trajectory modeling.
-Furthermore, none of these studies using this modeling approach have considered macroalgal morphology cross-sectional area (*i.e.* shape) as an aspect of drag and ultimately trajectory.
-@brooks2019, @putman2018, and @putman2020 used Lagrangian approaches to investigate the effects of inertia, raft-size, and windage on macroalgae trajectory.
+Works by @brooks2019, @putman2018 and @putman2020, used Lagrangian approaches to investigate the effects of inertia, raft-size, and windage (drag from the wind) on kelp trajectory.
 For example, @brooks2019 used a custom growth model to estimate changes in biomass and ultimately radial size, while a customized Hybrid Coordinate Ocean Model (HYCOM) for the trajectory simulations was used.
-Their results showed that the trajectory of pelagic *Sargussum* was significantly influenced by inertia and the radial size of the rafts.
-@putman2020 investigated the effect of including windage in macroalgae trajectory simulations and showed that including *ad hoc* windage factors greatly improved accuracy regarding entrainment or expulsion from eddys when compared to tracked *Sargassum* mats in the ocean.
-Langrangian trajectory modeling is clearly a useful tool; however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
-The radial size included by @brooks2019 is an improvement but this approach does not consider the complex morphology on an individual level into account.
-
-Studies have focused on kelp-rafts, not solitary floating kelp, and therefore are not able to investigate the role of eddys and vortices in the accumulation of individual kelp plants at the surface.
-Gyres, eddys, and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@onink2019; @dasaro2018; @brach2018].
-Investigations into  the dispersal of marine debris have also included the possible effects of inertia in accumulation of ocean drifter instruments in mesoscale features.
-For example, @beron2016 used a modeling approach to investigate the role of Eckman transport and inertia to explain the observed behavior of undrogued drifters in the ocean gyres.
-Undrogued drifter trajectory is affected by wind, unlike the drogue counter parts which makes the effects of wind negligible.
-The results showed that the combined effect inertia, not Eckman transport, played a significant role in the observed drifter behavior [@beron2016].
+Their results showed that the trajectory of pelagic Sargassum was significantly influenced by inertia and the radial size of the rafts.
+In the inclusion of wind into studies has also been performed on an *ad-hoc* basis, where a fraction of wind velocity field is simply added to the ocean velocity field in terms of the velocity vectors (meridonal and zonal); for example 1%, 2%, 5% of the ocean velocity field has been used previously [@putman2018;@putman2020].
+@putman2020 investigated the effect of including windage in kelp trajectory simulations and showed that including ad hoc windage factors greatly improved accuracy regarding entrainment or expulsion from eddies when compared to tracked Sargassum mats in the ocean.
+Langrangian trajectory modeling is clearly a useful tool; however, the various models available often assume the object to either be purely Lagrangian (no drag) or to be only influenced by windage (drag from the wind), which is obviously not the case for both biological and anthropogenic forms of marine debris in nature.
+The radial size in relation to windage included by @brooks2019 is an improvement but this approach does not consider hydrodynamic drag.
+Another example is work by @zhong2012 describing Sargassum distribution in relation to mesoscale and submesoscale eddies by simulating passive tracers.
+The authors used two model configurations, one with an idealised eddy-flow and another that simulates flow in the western Gulf of Mexico to advect passive tracers as a proxy for Sargassum.
+The authors showed that mesoscale and submesoscale features both play a role in accumulation zones for the virtual Sargassum and were able to explain some of the observed accumulation of Sargassum in the western Gulf of Mexico.
+Treating the virtual Sargassum particles as passive tracers may be the reason for the unexplained patterns during comparison of simulated and observed distributions.
+The authors do mention running simulations which included Stokes drag and inertia using a Maxey-Riley approach, and state that it made no significant difference to the overall results.
+However, the authors only used the density as a parameter for calculating the relevant forces by making the particles 10% lighter than the surrounding fluid.
+Additionally, the authors did not consider surface area or take hydrodynamic drag and windage into account, all of which are important influencers of trajectory.
+
+Gyres, eddies, and converging surface currents have been identified as important in the accumulation and dispersal of plastics and other marine debris [@brach2018; @dasaro2018; @onink2019].
+Investigations into the dispersal of marine debris have also included the possible effects of inertia in the accumulation of ocean drifter instruments in mesoscale features.
+For example, @beron2016 used a modelling approach to investigate the role of Ekman transport and inertia to explain the observed behavior of undrogued drifters in the ocean gyres.
+Undrogued drifter trajectory is affected by wind, unlike the drogue counterparts which makes the effects of wind negligible.
+The results showed that inertia, not Ekman transport, played a significant role in the observed drifter behaviour [@beron2016].
 These authors also noted that the presence of other flotsam, along with the undrogued drifters in ocean gyres, suggests that aspects of inertia also play a role in accumulation of other marine debris [@beron2016].
@@
 
-In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations.
-We aim to determine the 'behavior' of surface drifting macroalgae from a coastal environment to the open ocean, which will inform future characterization studies.
-We will achieve this by comparing the trajectory outputs of virtual passive particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure.
-The different degrees of wave and wind exposure are meant to reflect the different buoyancy characteristics, as well as identify the role of magnitude of the drag components on overall trajectory.
-The modeling of single kelp offers the opportunity to study the role of eddys and vortices in accumulation points on the surface.
+This study aims to use a numerical approach to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations.
+The ‘behaviour’ of surface drifting kelp is investigated with specific focus on trajectories from a coastal environment to the open ocean.
+We achieved this by comparing the trajectory outputs of virtual passive particles (purely Lagrangian i.e no drag characteristics added, as a reference) with kelp particles (drag characteristics added) with surface area exposed to the various drag components.
 
 # Methods
+
 ## Study site
@@
 The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A).
-Kommetjie, located on the Atlantic side of the Cape Peninsula (Figure 2B), was chosen as a release site.
+Kommetjie, located on the Atlantic side of the Cape Peninsula (Figure 1B), was chosen as a release site.
 The hydrodynamics in the study region are characterized by a complex interaction of wind, sea waves and ocean swell, and are also influenced by larger-scale regional oceanic processes.
 The larger processes in the region can be attributed to the Agulhas Current (AC) and the Benguela Current (BC) [@lutjeharms2000;@lutjeharms2007;@garzoli1996;@hutchings2009;@rubio2009].
-The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].
+The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroreflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].
 
 The region where the retroflection occurs is known as the Agulhas Retroflection and consists of high levels of mesoscale variability [@garzoli1996].
-The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324 km [@lutjeharms1988;@lutjeharms2006; @lutjeharms2007].
+The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988;@lutjeharms2006; @lutjeharms2007].
 Agulhas Rings are the main contributors to the Agulhas leakage which is the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015].
-The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009].
+The Agulhas leakage is also a source of mesoscale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009].
 The Benguela region ---known as the Benguela Upwelling System (BUS)--- is characterized by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003;@veitch2018; @veitch2017].
@@
 The region is also characterized by large amounts of mesoscale variability in the form of eddys and filaments [@blanke2002; @blanke2005; @rubio2009].
-Inshore, the BUS is dominated by a northwestward flow, known as the Benguela Jet, because of topographical steering, wind stress and interactions with passing Agulhas Rings and eddys [@stramma1989; @veitch2010; @veitch2017; @veitch2018; @hutchings2009].
+Inshore, the BUS is dominated by a northwestward flow, known as the Benguela Jet, due to topographical steering, wind stress and interactions with passing Agulhas Rings and eddys [@stramma1989; @veitch2010; @veitch2017; @veitch2018; @hutchings2009].
 The stability of the jet varies seasonally with more stable flows in winter; in January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results in a weak narrow inshore branch that flows along the shelf edge and a strong flowing north-westwards offshore branch.
 The area is also characterized by high upwelling activity driven by strong westerly winds [@hutchings2009].
-These winds are most intense during Austral summer (December-March) which favors upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
+These winds are most intense during Austral summer (December-March) which favours upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which favours upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
 In Austral winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela Jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
 
-```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1: Map of the study domain and the main features pertaining to transport in the region.
-In panel A, the map of the western portion of South African coastline and the main currents is presented.
-The resulting eddys and filaments froming from the Agulhas retroflection are indicated, and the position of the study area is drawn in as a black rectangle.
-Detail of the study area is in panel B.
+```{r map chp3, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1: Map of the study domain and the main features pertaining to transport in the region.
+In panel A, the map of the western portion of the South African coastline and the main currents is presented.
+The resulting eddies and filaments forming from the Agulhas retroflection are indicated, and the position of the study area is drawn in as a black rectangle.
+Details of the study area are given in panel B.
 Here the splitting of the Benguela Jet, eddy features, and the general flow within the study domain are presented.
-The release site of simulated kelp ‘particles’ is represented by a blue star."}
+The release site of simulated kelp ‘particles’ is represented by a black triangle."}
 
@@
 
+## Numerical models
+
+Virtual particles were released with no drag effects from the study site to act as a reference for comparison to simulations with varying combinations of hydrodynamic and wind drag and are referred to hereafter as ‘passive particles.’ All particles that have any form of drag exposure are hereafter referred to as ‘kelp particles.’ All simulations have the same release location.
+The release site, Kommetjie, was in an area of a known kelp population, which a previous study by @coppin2020 collected morphological data from.
+Due to the resolution of the underlying ocean model (see Ocean and wind model section) nearshore processes are not resolved well.
+Therefore, the release site was placed in the adjacent grid cell in order to better represent offshore transport and to avoid model limitations with regards to nearshore transport.
+Since the aim of this investigation was to investigate the relative surface behaviour of virtual kelp particles, only the surface layer of the underpinning hydrodynamic models were used.
+It should also be noted that the Copernicus model (see Ocean and wind model section) used velocity vectors over some of the land covering the coastal domain.
+
+The time period used in all simulations were Austral summer, autumn, winter, and spring of 2018.
+These time periods were chosen based on variable mesoscale activities and stability of the Benguela Jet that occurs throughout the year.
+For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity while in winter the Benguela Jet begins to stabilize until summer [@rubio2009; @Veitch2019; @ragoasha2019].
+Simulations were executed for the duration of each season, with 3-hourly timesteps and 3-hourly outputs using a fourth-order Runge-Kutta advection scheme for both passive and kelp particles.
+Drag components for kelp particles were incorporated by the modifiable component of the OceanParcels advection scheme [@North2009; @vanSebille2018].
+
+To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with degrees of drag exposure scenarios based on personal observations of drifting kelp.
+This acts as a proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag.
+These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for the applicable simulation.
+The scenarios used in this study were 100%  and 85% for hydrodynamic and Stokes drag and 0% and 15% for wind.
+Simulations were separated into fully submerged (100% hydrodynamic drag and 0% wind drag, see Figure 2A) and partially submerged (85% hydrodynamic drag, and 15% wind drag, See Figure 2C-D) categories.
+For each of the categories, particles were advected with different combinations of flow-fields (currents, currents-Stokes drift, currents-wind, currents-wind-Stokes drift).
+A similar approach has been used in past research investigating the role of windage in distribution of pelagic Sargassum rafts by @putman2018 and @putman2020 in the form of ad hoc windage estimates; although the authors did not include hydrodynamic drag unlike the current study.
+
+To determine if vortices and eddies play a role in kelp accumulation on the surface, a separate set of simulations were performed.
+Unlike the previous simulations, these simulations released 10 particles over the latitudinal extent of the study domain from the study site, equidistant from each other.
+A summary of the simulations executed are provided in table 2.
+
 ## Particle tracking model
 
-The model is based on parameterizations that have been used for previous studies investigating iceberg, capsized marine vessels and microplastics.
-It should be noted that most of the previous studies regarding macroalgal trajectories have investigated rafts and not solitary floating individuals.
-A kelp raft will vary in size and can range from a few meters across to large dense mats the size of a sports field.
-Some authors suggest that most macroalgae become entangled through ocean currents and not through the hydrodynamic forces that dislodge them.
-Large kelp rafts are not a feature around the coast of the Cape Peninsula; rather, solitary kelp is (pers.
+The model is based on parameterizations that have been used for previous studies investigating the dispersal patterns of various flotsam by implementing custom parameterisations of Lagrangian trajectory models [@furnans2008; @beron2016; @hart2018; @brooks2019; @putman2020; @jutzeler2020].
+Once dislodged, kelp can accumulate on the surface and become entangled to form a dense mat known as a kelp-raft [@hobday2000a; @smith2002].
+Kelp rafts are not a common feature around the coast of the Cape Peninsula; rather, solitary kelp is (pers.
 obs).
@@
 maxima* individuals will be considered.
-We performed numerical calculations of particle trajectory with a Lagrangian method [@delandmeter2019], which can be described by the equation:
+All simulations in this study used the OceanParcels software package (version 2.0), which is an open source, customizable Lagrangian Trajectory tool [@delandmeter2019].
+OceanParcels calculates the advection of particle trajectory with a Lagrangian method [@delandmeter2019], which can be described by the equation:
 
 $$X(t+\Delta t) = X(t) + \int_{t}^{t+\Delta t} v(x, \tau)d \tau + \Delta X_b(t)$$
-Here, $X$ is the three-dimensional position of a particle, $v(x,t)$ is the three-dimensional velocity field at the location in the ocean model at particle point in time, $t$ is time $X_b(t)$ is a change in position due to 'behavior' of the particle at a particular point in time.
-To investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behavior.
+
+Here, ${X}_t$ is the position of the particle at time $t$, ${X}(t + \Delta t)$ is the position after a timestep $\Delta t$, ${v}({x},t)$ is the ocean velocity field, $dt$ is the time-step and $\Delta{X_b}(t)$ is a change in position due to 'behavior' or some kind of effect.
+To determine the new position of a particle, the integral has to be calculated.
+To investigate the effects of drag components of trajectory of floating macroalgae, this study incorporated drag as the custom behaviour for virtual kelp particles and zero effect for passive particles.
 
 This study simulated two scenarios: one where the plant is fully submerged and only exposed to hydrodynamic drag, and the other where parts of the plants are partially exposed resulting in a combination of hydrodynamic and air form drag.
-Drag based on shape-based coefficients were used in the calculations for determining the overall kelp velocity vectors.
-A momentum energy equation was used to calculate the drag forces for the relevant simulation.
-This approach has been employed in modeling iceberg drift trajectory [@andersson2017; @lichey2001; @eik2009] and was adapted to suit this study.
-The energy momentum equation used to calculate the drag force exerted on the virtual kelp particle was $$m({d_u/d_t})=F_d + F_a + F_f$$
-where $m$ is kelp mass, $d_u$ the virtual kelp particle velocity, $F_d$ the hydrodynamic drag, $F_a$ is the wind drag, $F_f$ the surface-current flow, and $d_t$ is the time step.
+Drag based on spherical shape-based coefficients were used in the calculations for determining the overall kelp velocity vectors.
 The assumption made was that the mass of the kelp did not change for each virtual kelp particle over the course of the simulations.
-Hydrodynamic drag, wind drag, and surface current-flow are two-dimensional vector quantities and were calculated for each time step for each simulation.
-
-## Simulations
-
-Virtual particles were released with no drag or wind effects from the study site to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag and are referred to hereafter as 'passive particles'.
-All particles that have any form of drag exposure are hereafter referred to as 'kelp particles'.
-All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @vanSebille2018].
-Brownian motion was also incorporated to account for any stochasticity not resolved by the model.
-These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019;@vanSebille2018].
-The time period used in all simulations were January (Austral summer), April (Austral autumn), July (Austral winter), and October (Austral spring) of 2018.
-These time periods were chosen based on variable mesoscale activities and stability of the Bengulea Jet that occurs through the year.
-For example, splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity while in winter the Benguela Jet begins to stabilize until summer when the begins to split again [@rubio2009; @veitch2018; @ragoasha2019].
-
-To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios.
-These scenarios were meant to act as proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag.
-These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation.
-For example, if 85% of the plant is submerged and 15% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively.
-This approach has been used in past research investigating pelagic *Sargassum* rafts [@putman2020;@putman2018].
-The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1, Appendix).
-Each scenario consisted of a particle set with a size of 1000 were released from Kommetjie at the start of simulation.
-Since the aim of this investigation was to investigate the relative surface behavior of virtual kelp particles, only the surface layer was used for simulations.
-It should also be noted that the Copernicus model (see below) used places velocity vectors over some of the land covering the coastal domain.
-This was not considered a concern, as once again, the aim was to investigate surface drift behavior.
-A summary of the parameters for each simulation are provided in table 1.
-To determine if vortices and eddys play a role in kelp accumulation on the surface, a separate set of simulations were performed.
-Unlike the previous simulations, these simulations released a particle every hour over the same time periods.
-
-In order to determine if the shape coefficient used plays a role in model used and to validate the use of a spherical shape coefficient, simulations of kelps with the maximum surface area and hydrodynamic drag exposure was run using the shape coefficient for a cylinder to compare to a simulations using a sphere.
-Simulations were run for 30 days using a fourth order Runge-Kutte advection scheme with Brownian motion and a particle set of size 1000 for the same seasons as per previous simulations.
+Hydrodynamic drag, wind drag, Stokes drag, and surface current-flow are two-dimensional vector quantities and were calculated for each time step for each simulation.
 
 ## Model inputs
-### Ocean and wind model
-
-Both passive and kelp particles were simulated around the South African coastline within hindcast outputs from the Copernicus Marine Environment Monitoring Service (Copernicus, http://marine.copernicus.com).
-The Copernicus model is based on the modelling framework known as Nucleus for European Modelling of the ocean.
-The Copernicus outputs used in this study contain 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024).
-The model has a spatial resolution of 0.08$^\circ$ and is interpolated on an Arakawa C native grid.
-To incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (Global Ocean Wind L4_NRT_012_004).
-The outputs from the model are composed of 6-hourly averaged fields of surface 10 m wind velocity vectors with a spatial resolution of 0.25$^\circ$.
-Averaged ocean current and wind data are stored as meridional and zonal velocity vectors.
-
-### Hydrodynamic and wind drag
-
-The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical.
-The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field.
-Hydrodynamic and wind drag components were calculated separately for both the meridional and zonal velocity vectors, as per @furnans2008.
-Since drag force is dependent on the velocity vectors which vary with time, the meridional and zonal velocities were interpolated and used in the drag force calculation for each time-step.
-The same approach was used for the wind drag force.
-The equation that was used to calculate hydrodynamic drag, was also used to calculate wind drag.
-
-### Cross-sectional area
+
+### Cross-sectional area and simulation scenarios
 
 To incorporate hydrodynamic and wind drag, the cross-sectional area of the kelp was calculated first.
+AlgaeBase was used to confirm the latest taxonomic associations for the study species [@Algaebase].
 Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant.
-For details refer to Figure 2.
+The structure of *E.
+maxima* consist of a single primary blade that develops above a gas-filled float and a hollow stipe below.
+Secondary blades are produced laterally from the primary blade from several meristematic regions along the margins of the primary blade and is held to the substrate by finger-like haptera known collectively as the holdfast [@Dyer2018-rv].
+For details refer to Figure 2A.
+
 The dimensional data needed were estimated in cases where data was not available for that morphological characteristic.
@@
 
-```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Diagram of the *E.
+```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Panel A is a diagram of the *E.
 maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic.
-Schemetaic was adapted from @Dyer2018-rv."}
-
-kelp_shape_figure <- rasterGrob(as.raster(readPNG("figures/chapter_3/figure_2.png")))
+Panel B represents the full submerged scenario, panel C represents the combined drag scenario when kelp is partially submerged and panel D represents a partially submerged kelp is exposed to hydrodynamic drag and momentum transfer from the wind when in the same direction of the current.
+The kelp Schematic was adapted from @Dyer2018-rv."}
+
+kelp_shape_figure <- rasterGrob(as.raster(readPNG("figures/chapter_3/figure_2_v2.png")))
 ggarrange(kelp_shape_figure)
@@
 
-The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020].
-The minimum, mean and maximum cross-sectional areas (Ac, Table 2) were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (Table 2).
-The trajectory of the different cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.
+The cross-sectional area calculated was site-specific and the dimensions needed were garnered from morphology data from a previous study by @coppin2020.
+The minimum, mean and maximum cross-sectional areas (Ac, Table 1) were calculated for each morphological characteristic, which were used for calculating the cross-sectional area needed to run the various simulations.
+Initially, experiments included minimum and mean surface areas but no differences could be seen.
+Therefore, the minimum and mean cross-sectional areas were ignored and simulations only considered the maximum cross-sectional area.
+Different combinations of flow fields were used for both passive and kelp particles, in order to determine the influence of drag (hydrodynamic, Stokes drift and wind) on macroalgal trajectory.
+The different combination of flow-fields and the relative associated drag components are hereafter referred to as ‘scenarios’.
 
@@
 
-set_flextable_defaults(fonts_ignore=TRUE)
+set_flextable_defaults(fonts_ignore=FALSE)
 
@@
 area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
-area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
+#area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
 area_table <- fontsize(area_table, size = 8, part = "body")
 area_table <- fontsize(area_table, size = 8, part = "header")
-area_table <- set_caption(area_table, "Table 2: Table summarising the attributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulations.
+area_table <- set_caption(area_table, " Summary of the attributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulations.
 Approximate shape refers to the dimensions of the geometric shape used to represent a particular morphological characteristic.
@@
 Ac refers to the cross-sectional area calculated.
-The total surface areas for each plant characteristic for each surface area type make up the remaining columns.")
+The total surface areas for each plant characteristic for each surface area type make up the remaining columns.
+Where applicable r is the radius, h is the heigh, l is the length, w is the width, and h is the height.")
 area_table
@@
 
+### Flow-models
+
+Both passive and kelp particles were simulated around the South African coastline within hindcast outputs from the Copernicus Marine Environment Monitoring Service (Copernicus, http://marine.copernicus.com).
+The Copernicus model is based on the modelling framework known as Nucleus for European Modelling of the Ocean (NEMO).
+The Copernicus outputs used in this study contain 3D daily current information from the surface layer to the bottom (Global Analysis Forecast PHY_001_030).
+The model has a spatial resolution of 0.08 geographical degrees and is interpolated on a rectangular grid.
+To incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (Global Ocean Wind L4_NRT_012_004).
+The outputs from the model are composed of 6-hourly averaged fields of surface 10 m wind velocity vectors with a spatial resolution of 0.25 geographical degrees.
+Finally, to test the direct effects of Stokes-drift, the Copernicus Marine Environment Monitoring Service global wave product was used (Global Analysis Forecast WAVE_001_027) which are composed of 3-hourly instantaneous velocity vectors on a standard grid at 0.08 geographical degrees.
+Averaged ocean current, wind and Stokes drift data are stored as meridional and zonal velocity vectors.
+
+### Drag components
+
+Drag components calculated was incoporated into simulations by means of the modifiable custom behavior component of the underling OceanParcels advection scheme.
+The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag, as the trajectory model assumes the particles are spherical.
+The drag equation used was: $F_d = (0.5 \rho A_c C_d U^2)/m$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to the relevant drag component, $C_d$ is the drag coefficient for a spherical particle, $m$ is the total mas of the plant, and $U$ is the surface velocity vector of the flow-/wind- field.
+Dividing the calculated force by the mass of the plant transforms it to the appropriate units (m/s) for use in the trajectory modelling script.
+Hydrodynamic, wind drag and Stoke's drift components were calculated separately for both the meridional and zonal velocity vectors [@furnans2008; @putman2018; @putman2020].
+The same equation and approach used to calculate hydrodynamic drag, was also used to calculate wind and Stokes drift drag components.
+Since drag force is dependent on the velocity vectors which vary with time, the meridional and zonal velocities were interpolated and used in the drag force calculation for each time-step.
+The velocity of the resultant drag component is then added to the particle velocity, which is based on the underlying Euerlian flow-field, and determines the overall velocity of the particle and grid node position.
+
 ## Analysis
 
-To compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated, and density distribution maps were produced.
-Density distribution maps are an effective method for assessing pathways and ocean connectivity [@vanSebille2018].
-In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs.
+To compare trajectories between scenarios and seasons, trajectories were calculated, and density distribution maps were produced.
+Single trajectories allow comparison of pathways between scenarios, and density distribution maps are an effective method for assessing pathways and ocean connectivity [@vanSebille2018].
 The distance that particles traveled was also measured and used to produce boxplots.
@@
 If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity because of momentum energy transfer [@putman2020; @putman2018; @vanSebille2018].
+Topographic steering occurs when ocean currents make contact with underlying bathymetery features, such as sea mounts and ridges.
+Under topographic steering currents will tend to follow lines of constant potential vorticity [@davis1991;@vanSebille2018].
 
@@
 
+Overall, the results showed that the inclusion of drag parameters influenced trajectory patterns and entrainment patterns of virtual kelp particles.
+Single trajectories for each of the scenarios will be presented, followed by the comparison of density distributions and distance travelled by particles.
+Finally, the results from the separate set of experiments advecting single particles across the study domain will be presented.
+
 ```{r data readin, message=FALSE, warning=FALSE, include=FALSE}
 
-total_summer <- read.csv(file = "data/chapter_3/total_summer.csv")
+## single release
+
+total_summer_single <- read.csv(file = "data/chapter_3/total_summer_single.csv")
+total_autumn_single <- read.csv(file = "data/chapter_3/total_autumn_single.csv")
+total_winter_single <- read.csv(file = "data/chapter_3/total_winter_single.csv")
+total_spring_single <- read.csv(file = "data/chapter_3/total_spring_single.csv")
+
+total_passive_summer_single <- read.csv(file = "data/chapter_3/total_passive_summer_single.csv")
+total_passive_autumn_single <- read.csv(file = "data/chapter_3/total_passive_autumn_single.csv")
+total_passive_winter_single <- read.csv(file = "data/chapter_3/total_passive_winter_single.csv")
+total_passive_spring_single <- read.csv(file = "data/chapter_3/total_passive_spring_single.csv")
+
+total_passive_summer_single <- total_passive_summer_single %>% 
+  filter(type!= "Currents and Wind") %>% 
+  filter(type!= "Currents, Wind and Stokes")
+
+total_passive_winter_single <- total_passive_winter_single %>% 
+  filter(type!= "Currents and Wind") %>% 
+  filter(type!= "Currents and Stokes") %>% 
+  filter(type!= "Currents, Wind and Stokes")
+
+total_passive_spring_single <- total_passive_spring_single %>% 
+  filter(type!= "Currents and Wind") %>% 
+  filter(type!= "Currents and Stokes") %>% 
+  filter(type!= "Currents, Wind and Stokes")
+
+## continued release
+
+total_summer <- read.csv(file = "data/chapter_3/total_summer_noscale.csv")
 total_summer_mean <- read.csv(file = "data/chapter_3/total_summer_mean.csv")
 
-total_autumn <- read.csv(file = "data/chapter_3/total_autumn.csv")
+total_autumn <- read.csv(file = "data/chapter_3/total_autumn_noscale.csv")
 total_autumn_mean <- read.csv(file = "data/chapter_3/total_autumn_mean.csv")
 
-total_winter <- read.csv(file = "data/chapter_3/total_winter.csv")
+total_winter <- read.csv(file = "data/chapter_3/total_winter_noscale.csv")
 total_winter_mean <- read.csv(file = "data/chapter_3/total_winter_mean.csv")
 
-total_spring <- read.csv(file = "data/chapter_3/total_spring.csv")
+total_spring <- read.csv(file = "data/chapter_3/total_spring_noscale.csv")
 total_spring_mean <- read.csv(file = "data/chapter_3/total_spring_mean.csv")
@@
 
+# Line/Eddy experiment
+
 total_eddy_hydro <- read.csv(file = "data/chapter_3/total_eddy_hydro.csv")
@@
 
-The mean trajectories of the virtual kelps with different cross-sectional areas are similar to those of the passive particles during the month of January (Austral summer; Figure 3A).
-When comparing mean trajectories between virtual kelp with different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen; rather, there is only a slight variation in the overall trajectory.
-The results from the simulations including different hydrodynamic and wind drag exposures and kelp cross-sectional areas show differences in mean trajectory when compared to the passive simulation (Figure 3B).
-The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well the final location of the kelp particles being further from the release site.
-When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories was evident.
-In all simulations, particles flowed in a north-westward direction.
-
-The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (Austral autumn).
-Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag traveled further northward compared to the passive particles (Figure 3C-D).
-The mean trajectories of virtual kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles.
-Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles traveled less distance compared to kelp particles with both wind and hydrodynamic drag.
-Differences in mean trajectories between kelp and passive particles were also evident in the July (Austral winter) simulations (Figure 3E-F).
-Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore, resulting in entrainment within an eddy.
-Similar behavior is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag.
-In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles traveled initially southward and then westward offshore and also becoming entrained in an eddy.
-In the October (Austral spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (Figure 3F-G).
-Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag stop a greater distance from the release site compared to the passive particles.
-This was also evident when comparing kelp particles with different cross-sectional areas and exposure to wind and hydrodynamic drag with passive particles, with the end location a further from the release location.
-In both the April and October simulations all particles flowed north-westward, while in winter the mean trajectory resulted in entrainment within an eddy.
+The trajectories of the virtual kelps under different drag scenarios during Austral summer show no differences with those of passive particles advected by currents but do differ with passive particles advected by the different combinations of currents, Stokes drift and wind fields (Austral summer; Figure 3A).
+The results from the simulations of virtual kelp particles show no differences between scenarios.
+Any passive particles advected with Stokes drift travelled further than those that did not, while the combination of wind and currents for passive particles travelled the least distance compared to the other simulations for the same time period (Figure 3A, blue line).
+In all simulations for Austral summer, particles flowed-north-westward except for the currents and wind combination for passive particles which initially travelled shoreward.
+
+The trajectories of kelp particles across all drag scenarios showed no difference with passive particles advected by currents, but differences between the other passive scenarios and kelp particles was evident for Austral Autumn (Figure 3B).
+When comparing passive particles, only slight variations were evident for any simulation including Stokes drift, while currents and currents-wind showed slight variations between each other; with passive particles advected by currents travelling further offshore.
+The results from the simulations show that any scenario including Stokes drift for the passive particles results in the  trajectories flowing closer to shore (Figure 3B).
+Particles flowed in a north-westward direction and then slightly eastward near the end of the simulation run, across all simulations for both passive and kelp particles.
+
+No differences were evident across drag scenarios, and between passive particles advected by currents, and between kelp particles of different drag scenarios for Austral Winter (Figure 3C).
+There was slight variation between passive trajectories with the only variation the further distance travelled by passive particles for the currents-wind scenario.
+Unlike the previous time-periods, the overall direction for both passive and kelp particles was shoreward (eastwards) and then south along the eastern part of the coastline.
+
+In Austral Spring, no differences between kelp particles with different drag scenarios was evident, and no difference was evident between all kelp particle scenarios and passive particles advected by currents (Figure 3D).
+When comparing passive particles clear differences between current and current-wind scenarios is evident, with passive particles advected by currents displaced further offshore compared to passive particles with current-wind scenario.
+No difference in trajectories between passive particles advected by currents-Stokes drift and currents-wind-Stokes drift was evident with only a slight variation in distance travelled.
+Particles travelled in a north-westward direction across all scenarios for both passive and kelp particles.
+Overall, all kelp particles regardless of the drag scenario followed the same trajectory as passive particles advected by currents across all seasons and drag scenarios.
 
@@
 
-total_traj_summer <- ggplot()+
+traj_summer <- ggplot()+
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = total_passive_summer_mean, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
-  geom_path(data = total_summer_mean, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Passive Particles")) +
-  guides(linetype=guide_legend(title="Kelp Particles")) +
+  geom_point(data = total_passive_summer_single, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
+  geom_path(data = total_summer_single, aes(x = lon, y = lat, linetype = type))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Passive Particles",nrow=2,byrow=TRUE)) +
+  guides(linetype=guide_legend(title="Kelp Particles",nrow=2,byrow=TRUE)) +
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
@@
 
-total_traj_autumn <- ggplot()+
+traj_autumn <- ggplot()+
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = total_passive_autumn_mean, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
-  geom_path(data = total_autumn_mean, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Passive Particles")) +
-  guides(linetype=guide_legend(title="Kelp Particles")) +
+  geom_point(data = total_passive_autumn_single, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
+  geom_path(data = total_autumn_single, aes(x = lon, y = lat, linetype = type))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Passive Particles",nrow=2,byrow=TRUE)) +
+  guides(linetype=guide_legend(title="Kelp Particles",nrow=2,byrow=TRUE)) +
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
@@
 
-total_traj_winter <- ggplot()+
+traj_winter <- ggplot()+
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = total_passive_winter_mean, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
-  geom_path(data = total_winter_mean, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Passive Particles")) +
-  guides(linetype=guide_legend(title="Kelp Particles")) +
+  geom_point(data = total_passive_winter_single, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
+  geom_path(data = total_winter_single, aes(x = lon, y = lat, linetype = type))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Passive Particles",nrow=2,byrow=TRUE)) +
+  guides(linetype=guide_legend(title="Kelp Particles",nrow=2,byrow=TRUE)) +
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
@@
 
-total_traj_spring <- ggplot() +
+traj_spring <- ggplot()+
   geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = total_passive_spring_mean, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
-  geom_path(data = total_spring_mean, aes(x = lon, y = lat, linetype = type))+
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Passive Particles")) +
-  guides(linetype=guide_legend(title="Kelp Particles")) +
+  geom_point(data = total_passive_spring_single, aes(x = lon, y = lat, colour = type), alpha = 0.4) +
+  geom_path(data = total_spring_single, aes(x = lon, y = lat, linetype = type))+
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  guides(colour=guide_legend(title="Passive Particles",nrow=2,byrow=TRUE)) +
+  guides(linetype=guide_legend(title="Kelp Particles", nrow=2,byrow=TRUE)) +
   coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
@@
         axis.title=element_text(size=12))
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=6, fig.width=6, fig.cap="Figure 3: Comparison of mean trajectories.
-Simulations of hydrodynamic drag are presented in Panels A, C, E, G for January, April, July, and October, respectively.
-Results of simulations including both hydrodynamic drag and wind are in Panels B, D, F, H for January, April, July, and October, respectively.
-Significant amounts of overplotting prevents the different line types to be discernable."}
-
-ggarrange(total_traj_summer, total_traj_autumn, total_traj_winter, total_traj_spring ,ncol = 2, nrow = 2, font.label = list(size = 12, face = "bold"), common.legend = TRUE, legend = c("bottom"), labels = c("A","B","C","D"), align = "hv")
-
-ggsave(filename = "figure_3.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 9, width = 10, dpi = 300, scale = 1.5)
+
+```
+
+```{r Figure 3, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=8, fig.cap="Figure 3: Comparison of single trajectories.
+Simulations 1,5,9,13,17,21,25,29 are represented in panel A (Austral summer simulations), 2,6,10,14,18,22,26,30 in panel B (Austal aumtumn simulations), 3,7,11,15,19,23,27,31 in panel C (Austral winter simulations), and 4,8,12,16,20,24,28,32 in panel D (Austral spring simulations).
+Overlapping of linetypes will be represented by a solid black line for kelp particles."}
+
+# Remember data for this section has been edited for this section
+ggarrange(traj_summer, traj_autumn, traj_winter, traj_spring, ncol = 2, nrow = 2, font.label = list(size = 12, face = "bold"), common.legend = FALSE, legend = c("bottom"), labels = c("A","B","C","D"), align = "hv")
+
+#ggsave(filename = "figure_3v2.jpg", plot = last_plot(), path = "figures/chapter_3/", units = c("cm"), width = 40, height = 40, scale = 1.2)
+
 ```
@@
 
-Density distributions summarizing the conclusion of sumlating scenarios 1-13 (Table 1) are presented in Figures 3-6.
-Across all simulations for January (Austral summer), particles flow in a north-westward direction.
-The density plots for simulations that consider any form of drag show a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids (Figure 4).
-In addition, the density plots show that the particles in simulations that considered any form of drag become entrained in a vorticity while particles in the passive simulation deflected away (Figure 4).
-Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.
-In April (Figure 5) across all simulations, particles flow in a north-westward direction.
-The density plots that consider any form of drag show a narrow, higher density along the mean trajectory compared to passive simulation.
-In addition, simulations that consider any form of drag got entrained within an eddy further north-westward from the end location of passive particles.
-Only slight variations across cross-sectional areas and drag exposure scenarios are evident.
-
-Simulations for July (Figure 6) show that the flow is also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show a high density of particles entrained in an eddy located westward of the release site compared to passive particles.
-Variation in density along the mean trajectory and eddy for all kelp particles is evident.
-Kelp particles with only hydrodynamic drag across cross-sectional areas showed lower density of particles along the mean trajectory compared to kelp particles with both wind and hydrodynamic drag.
-Simulations run for October (Figure 7) show no difference in density of particles along the mean trajectory when comparing passive and kelp particles.
-Only slight variations in density and trajectory can be seen with no entrainment.
-Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.
-
+Density distributions summarizing the conclusion of simulating all passive and kelp scenarios (Table 2) are presented in Figure 4A-B.
+Across all simulations for Austral summer, particles flow in a north-westward direction.
+The only passive particle scenario that resulted in entrainment were particles advected by currents with particles concentrated on the boundary of the mesoscale feature (Figure 4A).
+The passive particles that were advected with Stokes drift resulted in particles concentrating closer to shore, and less dispersion of particles off the continental shelf (Figure 4A).
+Passive particles advected by currents-wind showed greater dispersion over the continental shelf and closer to shore compared to the other passive particle scenarios.
+The density plots that consider any form of drag showed entrainment patterns, with kelp particles towards the centre of the mesoscale feature (Figure 4B).
+No differences were evident between different drag scenarios for kelp particles.
+
+In Austral Autumn (Figure 4C-D) across all simulations, particles flowed in a north-westward direction.
+Passive particles advected by currents showed similar density distribution to kelp particles, but the other passive particles scenarios showed no entrainment patterns (Figure 4C).
+The density plots for kelp particles with any drag scenario show slight entrainment patterns with no differences in density evident between kelp particle drag scenarios (Figure 4D).
+When comparing kelp particles to passive particles, clear differences in density were evident.
+Kelp particles are less concentrated along the overall trajectory compared to passive particles which show a higher concentration of particles.
+
+Simulations for Austral winter (Figure 5A-B) showed that the flow was also north-westward with only slight variation in density distribution for kelp particles and clear differences between passive particle scenarios.
+Clear differences are evident between passive particle scenarios (Figure 5A).
+Passive particles advected by currents show a similar density distribution to that of kelp particles for the same time-period (Austral Winter, Figure 5A-B).
+Passive particles advected with Stokes drag showed distribution of particles closer to shore compared to the other passive scenarios.
+No differences or variation was evident when comparing kelp particle drag scenarios, only showing slight entrainment patterns in the north-westward portion of the study domain (Figure 5B).
+The currents-wind scenario differed from all the kelp and passive scenarios for the same time-period (Austral Winter, Figure 5B), and showed large spread of particles over the study domain and surrounding mesoscale features offshore.
+
+Simulations executed for Austral Spring (Figure 5C-D) showed no differences in density distribution between kelp particles.
+Passive particles for the same time-period (Austral Spring, Figure 5C) are concentrated along the overall trajectory and are less spread out.
+Passive particles that are close to mesoscale activity follow the boundary of the feature and do not spread out or become entrained (Figure 5C).
+Passive particles advected with Stokes drift are less distributed offshore and are instead concentrated along the overall trajectory and towards shore.
+Across all drag scenarios for kelp particles, entrainment patterns were evident, and particles were spread westwards across the study domain Figure 5B and 5D).
 ```{r density plots, message=FALSE, warning=FALSE, echo=FALSE}
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for simulations 1 - 13, for January 2018 (Austral summer).
-The plots are regular hexagons within the density of particles were calculated.
-The mean trajectories for each simulation are also plotted for comparison.
-The reader is referred to tables 1 and 2 for details pertaining to each simulation', fig.height=7, fig.width=6}
-
-ggarrange(heatmap_summer, heatmap_summer_passive, heatmap_autumn, heatmap_autumn_passive, labels = c("AUTO"), nrow = 2, ncol = 2, common.legend = TRUE, legend = c("bottom"),font.label = list(size = 12, face = "bold"))
+```{r figure 4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for both passive and kelp scenarios, for Austral summer (panel row A) and autumn (panel row B).
+Simulations 33,37,41,45 are represented in panel A, 49,53,57,61 are represented in panel B, simulations 34,38,42,46 are represente in panel C, and simulations 50,54,58,62 are represented in panel D.
+The plots are regular hexagons within which the density of particles were calculated.
+The reader is referred to tables 1 and 2 for details pertaining to each simulation.', fig.height=10, fig.width=9}
+
+ggarrange(heatmap_summer_passive, heatmap_summer, heatmap_autumn_passive, heatmap_autumn, labels = c("AUTO"), nrow = 2, ncol = 2, common.legend = TRUE, legend = c("bottom"),font.label = list(size = 12, face = "bold"))
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end run time for simulations 1 - 13, for January 2018 (Austral summer).
-The plots are regular hexagons within the density of particles were calculated.
-The mean trajectories for each simulation are also plotted for comparison.
+```{r r figure 5, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end run time for both passive and kelp scenarios, for Austral winter (panel row C) and spring (panel row D).
+Simulations 35,39,43,47 are represented in panel A, 51,55,59,63 are represented in panel B, simulations 36,40,44,48 are represente in panel C, and simulations 52,56,60,64 are represented in panel D.
+The plots are regular hexagons within which the density of particles were calculated.
 The reader is referred to tables 1 and 2 for details pertaining to each simulation', fig.height=10, fig.width=9}
 
-ggarrange(heatmap_winter,heatmap_winter_passive, heatmap_spring, heatmap_spring_passive, labels = c("AUTO"), nrow = 2, ncol = 2, common.legend = TRUE, legend = c("bottom"),font.label = list(size = 12, face = "bold"))
+ggarrange(heatmap_winter_passive, heatmap_winter, heatmap_spring_passive, heatmap_spring, labels = c("AUTO"), nrow = 2, ncol = 2, common.legend = TRUE, legend = c("bottom"),font.label = list(size = 12, face = "bold"))
 
@@
 
-## Point distribution
-
-```{r point plots wrangle, message=FALSE, echo=FALSE, warning=FALSE}
-
-total_summer_point <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_path(data = total_summer %>% 
-              #filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
-  geom_point(data = total_summer, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
-  facet_wrap("type") +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme_bw() +
-  #ggtitle("Kelp Particles")
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-total_autumn_point <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_path(data = total_autumn %>% 
-              #filter(!type =='Passive'), aes(x = lon, y = lat, linetype = type))+
-  geom_point(data = total_autumn, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
-  facet_wrap("type") +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-total_winter_point <- ggplot()+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_path(data = total_winter %>% 
-              #filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
-  geom_point(data = total_winter, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
-  facet_wrap("type") +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme_bw() +
-  #ggtitle("Kelp Particles")
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-total_spring_point <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_path(data = total_spring %>% 
-             # filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
-  geom_point(data = total_spring, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
-  facet_wrap('type') +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-# Passive particles
-
-total_summer_point_pass <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_path(data = total_spring %>% 
-             # filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
-  geom_point(data = total_passive_summer, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
-  facet_wrap('type') +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme_bw() +
-  #ggtitle("Passive Particles")
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-total_autumn_point_pass <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_path(data = total_spring %>% 
-             # filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
-  geom_point(data = total_passive_autumn, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
-  facet_wrap('type') +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-total_winter_point_pass <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_path(data = total_spring %>% 
-             # filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
-  geom_point(data = total_passive_winter, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
-  facet_wrap('type') +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme_bw() +
-  ggtitle("Passive Particles")
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-total_spring_point_pass <- ggplot() +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 0.4,     colour = "black",     fill = "grey") +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  #geom_path(data = total_spring %>% 
-             # filter(!type == 'Passive'), aes(x = lon, y = lat, linetype = type))+
-  geom_point(data = total_passive_spring, aes(x = lon, y = lat, colour = age), show.legend = FALSE) +
-  facet_wrap('type') +
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  guides(colour=guide_legend(title="Simulation")) +
-  coord_fixed(ratio=1, clip = "on", xlim = c(11, 20), ylim = c(-36, -28)) + # fix the aspect ratio of plot
-  theme(plot.title = element_text(hjust = 0.5)) +
-  theme_bw() +
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-```
-
-```{r point plot, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE, fig.cap="Figure 6: Point plots shoing the age of all passive and kelp particles at the end of each simulation run."}
-
-ggarrange(total_summer_point, total_summer_point_pass, total_autumn_point, total_autumn_point_pass,total_winter_point,total_winter_point_pass, total_spring_point,total_spring_point_pass, labels = c("AUTO"), nrow = 2, ncol = 4, common.legend = TRUE, legend = c("bottom"),font.label = list(size = 12, face = "bold"))
-
-ggsave(filename = "figure_age.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 6, width = 6, scale = 1, dpi = 300)
-```
-
 ## Comparison of distances
 
-```{r distance plot processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+The results for the comparison of distance travelled by passive and kelp particles across all scenarios and seasons are presented in Figure 6.
+There were no significant differences (Figure 6; p > 0.05) for distance travelled across all kelp particles drag scenarios and seasons.
+Kelp particles travelled the least distance in Austral winter (Figure 6C) compared to other seasons and across passive particle scenarios.
+With regards to passive particles, clear differences and variation in the mean distance travelled was evident.
+For Austral Summer, the currents-wind scenario for passive particles travelled significantly less distance compared to the other scenarios (currents, currents-Stokes drift, currents-wind-Stokes drift, Figure 6B).
+There were no significant differences in distance travelled between all passive scenarios for Austral Autumn (p > 0.05; Figure 6D).
+For Austral winter, all passive particle scenarios with Stokes drift  included travelled significantly less distance compared to other passive particle scenarios (p < 0.05; Figure 6F).
+The same was true for Austal spring, which also showed that passive particles with Stokes drift flow travelled significantly less distance compared to other passive particle scenarios (p < 0.05; Figure 6H).
+
+```{r distance chp3 processing, message=FALSE, warning=FALSE, include=FALSE}
 # Kelp particles
@@
 
-```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, include=TRUE, fig.cap="Figure 6: Boxplots of total distance travelled by all particles from the release site over the course of the simulations for scenarios 1 - 52.
+```{r distance plot chp3, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 6: Boxplots of total distance travelled by all particles from the release site over the course of the simulations for scenarios 1 - 64.
 The *x*-axis represents the scenario number, and the *y*-axis is the distance travelled in kilometers.
@@
 The whiskers represent the range, solid black lines represent the median and black dots are outliers.
-The reader is referred to Tables 1 and 2 for details pertaining to each scenario.", fig.height=7, fig.width=7}
-
-ggarrange(dis_plot_summer, dis_plot_summer_pass, dis_plot_autumn, dis_plot_autumn_pass, dis_plot_winter, dis_plot_winter_pass,dis_plot_spring, dis_plot_spring_pass, label = "AUTO", ncol = 2, nrow = 4, labels = "AUTO")
-ggsave(filename = "figure_7.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 8, scale = 1, dpi = 300)
-```
-
-## Eddy/Line experiment
-
+The reader is referred to Tables 1 and 2 for details pertaining to each scenario."}
+
+ggarrange(dis_plot_summer, dis_plot_summer_pass, dis_plot_autumn, dis_plot_autumn_pass, dis_plot_winter, dis_plot_winter_pass,dis_plot_spring, dis_plot_spring_pass, label = "AUTO", ncol = 2, nrow = 4, legend = "none", common.legend = FALSE)
+
+ggsave(filename = "figure_6.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 8, scale = 1, dpi = 300)
+```
+
+## Line experiment
+
+The results from simulations which released single particles along the longitudinal extend of the study domain are presented in Figure 7 and 8.
+When comparing kelp particles with different drag scenarios, no differences or even variation in the trajectory of single particles was evident from the results.
+Across all seasons and drag scenarios, kelp particles located from the edge of the continental shelf towards the eastern part of the study domain showed entrainment patterns and/or flowed northward.
+Kelp particles released from the study site (Kommetjie) did not show any entrainment patterns and flowed north-westward.
 ```{r message=FALSE, echo=FALSE, warning=FALSE, include=FALSE}
-
 # summer
@@
   theme_bw() +
-  ggtitle("Kelp")
+  ggtitle("Kelp") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   theme_bw() +
-  ggtitle("Passive")
+  ggtitle("Passive") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2) +
-  geom_point(data = total_eddy_passive_currwindstks, aes(x = lon, y = lat, colour = trajectory), show.legend = FALSE) +
+  geom_point(data = total_eddy_passive_currwindstks, orientation = "y", aes(x = lon, y = lat, colour =       trajectory), show.legend = FALSE) +
   scale_colour_continuous(type = "viridis",limits = c(0, 9))  +
@@
 
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation for scenarios 53-60.
-For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=6, fig.width=8}
-
-ggarrange(eddy_hydro_plot, eddy_passive_plot_curr, eddy_hydrostksdrag_plot,eddy_passive_plot_currstks, common.legend = TRUE, legend = c("none"), ncol = 2, nrow = 2, labels = c("AUTO"), font.label = list(size = 12, face = "bold"), align = "hv")
+```{r Figure 7, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 7: Comparison of single kelp and passive particle trajectories for all seasons released along a line from the study site (Kommetjie), equidistant apart.
+Simulations 65-68 are represented in panel A, 81-84 are represented in panel B, 69-72 are represented in panel C, and 85-88 in panel D.", fig.height=8, fig.width=10}
+
+ggarrange(eddy_passive_plot_curr, eddy_hydro_plot, eddy_passive_plot_currstks, eddy_hydrostksdrag_plot, common.legend = TRUE, legend = c("none"), ncol = 2, nrow = 2, labels = c("AUTO"), font.label = list(size = 12, face = "bold"), align = "hv")
+
+ggsave(filename = "figure_7.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 6, width = 8, scale = 2, dpi = 300)
+```
+
+When comparing passive particles advected with no drag but by the same combination of flow fields as kelp particles, differences and variation was evident.
+Passive particle trajectories across all advection scenarios generally flowed north-westward and/or flowed on the boundary of mesoscale features (Figure 7B,D and Figure 8B,D).
+The advection scenarios which included Stokes drift showed trajectories closer to the coast compared to scenarios which did not include Stokes drift.
+The simulations which included Stokes drift for passive particles showed more particles travelling north westward compared to particles which did not.
+
+```{r Figure 8, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of single kelp and passive particle trajectories for all seasons released along a line from the study site (Kommetjie), equidistant apart.
+Simulations 73-76 are represented in panel A, 89-92 are represented in panel B, 77-80 are represented in panel C, and 93-96 in panel D.", fig.height=8, fig.width=10}
+
+ggarrange(eddy_passive_plot_currwind, eddy_hydrowinddrag_plot, eddy_passive_plot_currwindstks, eddy_hydrowindstksdrag_plot, common.legend = TRUE, legend = c("none"), ncol = 2, nrow = 2, labels = c("AUTO"), font.label = list(size = 12, face = "bold"), align = "hv")
 
@@
 ```
-
-```{r accumulation data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 8: Comparison of density between passive and kelp accumulation for scenarios 53-60.
-For each season, passive particles are plotted on the left column and kelp particles in the right.", fig.height=6, fig.width=8}
-
-ggarrange(eddy_hydro_plot, eddy_passive_plot_curr, eddy_hydrostksdrag_plot,eddy_passive_plot_currstks, common.legend = TRUE, legend = c("none"), ncol = 2, nrow = 2, labels = c("AUTO"), font.label = list(size = 12, face = "bold"), align = "hv")
-
-ggsave(filename = "figure_8.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 6, width = 8, scale = 2, dpi = 300)
-```
-The role of eddys that exist within the study domain during the study period play a clear role in the accumulation and dispersal of kelp particles.
-Passive particles tend to remain on the outer edge of the eddy and are then expelled on the other side and continue to follow the ocean current (Figure 9).
-The opposite is true for kelp particles which have the highest exposure level of hydrodynamic and wind drag.
-The kelp particles are entrained in the eddy (Figure 9).
-This is also reflected when comparing the age of particles at the end of the simulation.
-In Figure 10, younger particles are near the release site and older particles further away from the shore, while older kelp particles are retained within the eddy.
-
-## Comparison of coefficients
-
-The comparison of mean trajectories between the sphere and cylindrical shape coefficients revealed no difference (appendix).
-This was also seen when comparing distance traveled by particles and particle density plots (appendix).
-The simulation employing a cylindrical shape coefficient showed higher density in an eddy compared to the simulation which used a sphere for the shape coefficient (appendix).
 
@@
 In particular, the results show that consideration of drag and surface area are important when investigating macroalgal trajectory patterns.
-The treatment of drifting macroalgae as purely Lagrangian does not accurately determine patterns of passive dispersal, such as entrainment in eddys or vortices [@fossette2012; @miron2020].
-Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], past research has focused on large macroalgal-rafts which can vary greatly in size and surface area.
-None of the extant research has taken advantage of the advancement in numerical ocean models and Lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae.
-This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag.
-The mean virtual kelp trajectory deviated from the mean passive trajectory when including varying levels of hydrodynamic and wind drag scenarios.
-The results from this study show the inclusion of hydrodynamic and wind drag results in observable entrainment patterns and broadens our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case that of *E.
+The treatment of drifting kelp as purely Lagrangian does not accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @miron2020].
+Although past research has shown the inclusion of inertia and windage greatly increase observed kelp trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], past research has focused on large macroalgal-rafts which can vary greatly in size and surface area.
+None of the extant research has taken advantage of the advancement in numerical ocean models and Lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting kelp.
+This study compared particle trajectories with simulations that included no forms of drag with simulations that included various forms of drag.
+Kelp trajectory deviated from the passive trajectory when including varying levels of hydrodynamic and wind drag scenarios.
+The results from this study show the inclusion of hydrodynamic and wind drag results in observable entrainment patterns and broadens our understanding of how surface area plays a role in solitary drifting kelp trajectory, in this case that of *E.
 maxima*.
@@
 The results from this numerical experiment confirm that drifting macroalgae do follow the surface current in the study region when considering only hydrodynamic drag.
-In addition, not only are the trajectories similar but so are the end points at the end of simulation.
-However, when *ad hoc* wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely Lagrangian particle trajectories.
-The inclusion of wind drag causes the kelp particles to flow further away from the coast and with a greater distance travelled from the release location.
-The difference in trajectory and distance covered can be attributed to the inclusion of wind drag.
+In addition, not only are the trajectories similar but so are the endpoints at the end of simulation.
+Interestingly, when *ad hoc* wind drag (windage) is considered in combination with hydrodynamic drag the end locations do not differ either.
+It was expected that the inclusion of wind drag would cause kelp particles to flow further away from the coast and with a greater distance travelled from the release location.
 Wind drag adds to the effects of hydrodynamic drag when the wind direction and current direction are opposite to each other.
 However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object [@hackett2006; @putman2020].
-This causes the kelp particles to travel further as well as become exposed to different time varying flows.
+This causes objects to travel further as well as become exposed to different time varying flows.
+
 Past studies have shown this when including windage as a function of inertia [@putman2018;@brooks2019; @putman2020].
@@
 The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked *Sargassum sp* mats, which was partly due to the inclusion of momentum energy transfer from wind.
-This is also reflected in the analysis comparing distances travelled by all the particles in each simulation.
-The particles in the simulations that included wind drag traveled significantly greater distances compared to the passive or hydrodynamic only drag simulations.
-Similar patterns are seen in the simulations run for April (Austral autumn), July (Austral winter), and October (Austral spring); however unlike in January (Austral summer) where only kelp particles with combined drag (hydrodynamic and wind) flowed further, even kelp particles with only hydrodynamic drag flow further compared to passive particles in these times of the year.
-This may be a result of the kelp particles exposed to different time-varying flows, which results in the kelp particles following a different surface current vector compared to passive particles.
-This also results in different end-points and therefore these results further confirm that drag is an important factor to account for when investigating macroalgal drift patterns.
+The results from this study suggest that when considering solitary floating kelp the effect of wind in the form of momentum energy transfer is negligible, and does not influence the overall trajectory.
+The previously mentioned studies including wind drag focused on rafts which would have a significantly larger surface area exposed to the wind compared to a solitary individual.
+Therefore, in this study the surface area of kelp particles exposed to wind drag was not enough to influence the overall trajectory.
+A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region.
+The results suggested that the hatchling trajectories were mostly influenced by the ocean currents in the first year of hatching as juveniles do not swim fast enough to alter their overall trajectory.
+Another possible reason considered was the resolution of the ocean model used.
+The resolution of the underlying ocean model is an important aspect of Lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications.
+A study by @hart2018 assessed the inclusion of stochastic motion, wind, and currents into forecasting for search and rescue using the same ocean model used in this study.
+The authors found that the inclusion of Brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents, and stochastic motion greatly improved forecasting applications.
+In this study, an eddy-resolving model was used which greatly increases the sub-grid scale processes.
+Therefore, the authors argue that the lack of dissimilarity between trajectories of kelp particles is not as a result of the ocean model used.
+
 The different trajectory patterns between seasons reflects the underlying seasonal variation in regional oceanography.
-Transport in the study region is driven by a combination of variation of alongshore jets and wind induced Ekman transport of surface water [@rubio2009; @veitch2018; @ragoasha2019].
+Transport in the study region is driven by a combination of variation of alongshore jets and wind induced Ekman transport of surface water [@rubio2009;@Veitch2019; @ragoasha2019].
 @ragoasha2019 investigated the success of Lagrangian particles reaching St.
@@
 The authors showed that transport success was higher in spring due to the single coherent Benguela Jet, which is most intense over the 300m isobath [@ragoasha2019].
-During upwelling season (December - March) transport success was less efficient as particles tend to flow offshore.
-The move offshore was a result of the bifurification of the Benguela Jet, which splits because of shoaling and causes particles to be exposed to higher amounts of mesoscale activity.
+During the upwelling season (December - March) transport success was less efficient as particles tend to flow offshore.
+The move offshore was a result of the bifurcation of the Benguela Jet, which splits because of shoaling and causes particles to be exposed to higher amounts of mesoscale activity.
 In the winter months, the Benguela Jet was still split but the flow weakens which resulted in lower transport success [@ragoasha2019].
 Similar patterns were found in this study, suggesting that the difference in seasonal patterns reflects the regional oceanographic features, particularly the Benguela Jet.
-
-The addition of drag forces caused the kelp particles to cluster together along the mean trajectory compared to that of the passive particles which were more evenly dispersed.
-This was most likely due to the shared drag characteristics among the kelp particles which included the plants' mass and surface area into the calculation.
-Other studies investigating flotsam trajectory characteristics have found similar results [@breivik2011; @olascoaga2020; @miron2020].
-A study by @miron2020 conducted field experiments using a range of objects such as spheres, cubes, and cuboids to compare to investigate the effect of inertia on particles dynamics.
+The addition of drag forces caused the kelp particles to cluster together compared to that of the passive particles which were more evenly dispersed.
+This was most likely due to the shared drag characteristics among the kelp particles which included the plants’ mass and surface area into the calculation.
+Other studies investigating flotsam trajectory characteristics have found similar results [@breivik2011; @miron2020; @olascoaga2020].
+A study by @miron2020 conducted field experiments using a range of objects such as spheres, cubes, and cuboids to investigate the effect of inertia on particle dynamics.
 The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia.
-Although drag force and not inertia was used in this study, the same conclusion applies---drag is therefore an important characteristic to consider when investigating or predicting solitary drifting macroalgal trajectories.
-
-The ocean is comprised of time-varying flows and changes in an objects velocity that can cause the object to be exposed to different flow patterns over time.
-The lack of dissimilarity in trajectory between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area was negligible.
-A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region.
-The results suggested that the hatchling trajectories were mostly influenced by the ocean currents in the first year of hatching due as juveniles do not swim fast enough to alter their overall trajectory.
-Therefore, the differences in cross-sectional area types may not be significant enough to influence the relevant zonal and meridional velocity vectors.
-Another possible reason was the resolution of the ocean model used.
-The resolution of the underlying ocean model is an important aspect of Lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications.
-A study by @hart2018 assessed the inclusion of stochastic motion, wind, and currents into forecasting for search and rescue using the same model ocean model used in this study.
-The authors found that the inclusion of Brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents, and stochastic motion greatly improved forecasting applications.
-Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not because of the resolution of the model.
-
-The same is true when comparing different hydrodynamic and wind drag exposure scenarios.
-These scenarios reflected different buoyancy situations which resulted in different amounts of surface area exposed to hydrodynamic and/or wind drag.
-The similarities between the different combined drag scenarios suggest that higher magnitudes of wind exposure did not significantly alter trajectories.
-Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of modeling entrainment patterns.
+Although drag force and not inertia was used in this study, the same conclusion applies—drag is therefore an important characteristic to consider when investigating or predicting solitary drifting macroalgal trajectories.
+
+The similarities between the different combined drag scenarios suggest that wind exposure did not significantly alter trajectories.
+Instead, the results from this study suggest that the inclusion of drag forces (both hydrodynamic and wind) in simulating macroalgal trajectory may result in improving accuracy of modeling entrainment patterns.
 All the simulations that included any form of drag resulted in a high density of particles entrained in an eddy.
-The entrainment and expulsion from eddys and vortices are an important characteristic of floating macroalgae.
-Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddys has been investigated previously by @putman2018 and @putman2020, who showed that including windage greatly improved entrainment patterns when comparing virtual and real *Sargassum* rafts.
+The entrainment and expulsion from eddies and vortices are an important characteristic of floating kelp.
+Accuracy in predicting entrainment and expulsion of kelp-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who showed that including windage greatly improved entrainment patterns when comparing virtual and real Sargassum rafts.
 However, these studies investigated large rafts and not floating individuals.
@@
 Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered.
-The results from this study suggest that the inclusion of hydrodynamic and/or wind drag was an important component when simulating virtual kelp particles and greatly increased the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddys and vortices.
-
-The role of eddys and vortices in the entrainment of kelp is further supported by the accumulation simulations.
-These results clearly show that the inclusion of drag greatly increases the number of particles entrained in the eddy within the study domain.
-eddys and vortices may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in area below the eddy.
-A study by @woodborne1989 references a site offshore approximately 120 km from Cape Columbine where dredged kelp was found.
-The area coincides with the region where virtual kelp was entrained by an eddy.
-The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory.
-Instead, the inclusion of a different shape coefficient increased to the density of particles entrained in an eddy field.
-Therefore, more accurately parameterizing shape will therefore allow for further improvements in models of macroalgal dispersal patterns.
+The results from this study suggest that the inclusion of hydrodynamic and/or wind drag was an important component when simulating virtual kelp particles and greatly increased the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.
+
+These results clearly show that the inclusion of drag greatly increases the number of particles entrained in mesoscale features within the study domain.
+Eddies may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in an area below the eddy.
+A study by @woodborne1989 references a site offshore approximately 120 km from Cape Columbine where dredged kelp was found, which coincides with the region where virtual kelp was entrained by an eddy.
+Finally, the inclusion of Stokes drift within the simulations had no effect on overall trajectory or entrainment patterns of virtual kelp particles, which suggests that the influence of waves is negligible with regards to solitary floating *E.
+maxima*.
+Past research has shown that the length of a floating object in relation to the significant wave height determines the influence of waves on overall trajectory.
+Trajectories of objects with lengths shorter than the significant wave height will not be influenced by wave related processes, in this case Stokes drift [@allen1999; @breivik2011; @griffin2017].
 
@@
 
-Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape; however, prior to our study, no research existed to clarify the drifting characteristics of solitary macroalgae.
+Most of the past research has focused on large kelp-rafts which can vary greatly in size and shape; however, prior to our study, no research existed to clarify the drifting characteristics of solitary kelp.
 Understanding how to accurately model the distribution of solitary kelp will lead to better understanding of accumulation zones and sinks, and ultimately the ecological links pertaining to those processes.
@@
 maxima* particles tended to follow the prevailing surface currents and the inclusion of wind caused the trajectories to differ greatly from that of a purely Lagrangian particle.
-Furthermore, the inclusion of both hydrodynamic and wind drag caused clustering of particles along the trajectory as well caused particles to become entrained in eddys.
-Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E.
-maxima*.
-In addition, the use of a cylindrical shape coefficient had no effect on mean trajectory, but did cause higher density of particles to become entrained in an eddy.
-Overall, the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not.
-This study also provided an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately.
+Furthermore, the inclusion of both hydrodynamic and wind drag caused clustering of particles along the trajectory as well caused particles to become entrained in eddies.
+Finally, Stokes drift drag did not influence trajectory and dispersal patterns of virtual kelp particles, suggesting that the effect of waves can be ignored when modelling solitary floating kelp.
+Overall, the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary kelp, while differences in cross-sectional areas are not.
+This study also provided an approach which can be adapted to model any floating solitary kelp, provided the surface area can be estimated accurately.
 Our results further highlight the importance of regional oceanographic processes to better understand how kelp flotsam is dispersed in the ocean.
@@
 Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal.
-The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates.
-
-# References
-
-<div id="refs"></div>
+The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as estimates for kelp sinking rates which would aid in describing the spatial distribution on the seafloor.
+
+# Conflict of Interest Statement
+
+The remaining authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
+
+# Author Contributions
+
+RC conceptualized the scope of the research reported in this manuscript, collected samples, undertook the majority of numerical and statistical analyses, made the first round of interpretation, and did the majority of the writing.
+AS and CR provided the guidance in terms of writing, statistical analyses, and overall presentation of the manuscript.
+All authors contributed to the article and approved the submitted version.
+
+# Funding
+
+This research was funded by the South African National Research Foundation (http://www.nrf.ac.za).
+Funding number is CSRP170430229220.
+Aside from funding provided, the funders had no role in the study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+
+# Acknowledgments
+
+We thank A. J Smit and Christo Rautenbach for their guidance on this project.
 
@@
 
-```{r sim table, eval=TRUE, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
-
+```{r sim table, echo=FALSE, message=FALSE, warning=FALSE}
 set_flextable_defaults(fonts_ignore=TRUE)
 
-sim_table <- read.csv("tables/chapter_3/simulation_table.csv", sep = ";")
-
+sim_table <- read.csv("tables/chapter_3/simulation_table_A.csv", sep = ";")
 sim_table <- flextable(data = sim_table)
@@
 sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
-sim_table <- set_caption(sim_table, "Table 1: Details for each simulation run for the study.
-Type refers to either a passive particle with no drag behavior or a kelp particle that has been advected with any drag exposure.
-The drag exposure column refers to the parameterizations regarding the level of water and wind drag simulated.
-All simulations were run with a Runge-Kutte (4th) advection scheme, Brownian motion, 30 days, and included 1000 particles for a 2018 hindcast.")
+sim_table <- set_caption(sim_table, "Details for each simulation run for the study.
+Type refers to either a passive particle with no drag behavior or a kelp particle that has been advected with any type of drag exposure.
+The Number coloumn refers to the simulation number, the condition coloumn refers to the situation under which the particle is being advected, the advection column refers to the underlying velocity field or combination thereof, the simulation column refers to the type of particle (passive or kelp), and the season refers to the seasonal time period of the simulation.
+The run-time for summer simulations was 89, autumn 91, winter 91 and spring 91.
+The simulation output for all simulations was 3 hours and advected with a Runge-Kutta fourth order advection scheme.
+Particles were released every 3 hours for simulations which executed repeated releases.")
 sim_table
-
-```
-
-```{r test data wrangle, warning = FALSE, echo=FALSE, message=FALSE, eval=FALSE}
-
-kelp_hydro_A <- tidync("OceanParcels/chapter_3_output/kelp_hydro_testA.nc") %>% #1hr
+```
+
+```{r test data wrangle, warning = FALSE, echo=FALSE, message=FALSE, eval=TRUE}
+
+kelp_hydro_A <- tidync("OceanParcels/chapter_3_output/kelp_hydro_sense_1hr.nc") %>% #1hr
   hyper_tibble() %>%
@@
 
-kelp_hydro_B <- tidync("OceanParcels/chapter_3_output/kelp_hydro_testB.nc") %>% #3hr
+kelp_hydro_B <- tidync("OceanParcels/chapter_3_output/kelp_hydro_sense_3hr.nc") %>% #3hr
   hyper_tibble() %>%
@@
 
-kelp_hydro_C <- tidync("OceanParcels/chapter_3_output/kelp_hydro_testC.nc") %>% #6hr
+kelp_hydro_C <- tidync("OceanParcels/chapter_3_output/kelp_hydro_sense_6hr.nc") %>% #6hr
   hyper_tibble() %>%
@@
 
-kelp_hydro_D <- tidync("OceanParcels/chapter_3_output/kelp_hydro_testD.nc") %>% #12hr
+kelp_hydro_D <- tidync("OceanParcels/chapter_3_output/kelp_hydro_sense_12hr.nc") %>% #12hr
   hyper_tibble() %>%
@@
 
-```{r test data outputs, warning=FALSE, echo=FALSE, message=FALSE,eval=FALSE, fig.cap="Heatmaps comparing densities and patterns across time-release options.", fig.height=6, fig.width=5}
+```{r test data outputs, warning=FALSE, echo=FALSE, message=FALSE,eval=TRUE, fig.cap="Heatmaps comparing densities and patterns across time-release options.", fig.height=8, fig.width=8}
 
@@
 
-#ggsave(filename = "figure_9.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 10, width = 6, scale = 1, dpi = 300)
-```
+#ggsave(filename = "figure_sensetest.jpg", plot = last_plot(), path = "figures/chapter_3/", height = 8, #width = 8, scale = 1, dpi = 300)
+```
+
+# References

### 2022-08-31 thesis_chapter_4.Rmd
uploaded updated analysis files

@@
 ---
-title: "Individual-based numerical experiment to describe the distribution of floating kelp within the Southern Benguela Upwelling System."
+title: 'Individual-based numerical experiment to describe the distribution of floating kelp within the Southern Benguela Upwelling System.'
 author: 
-- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
+- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa]
+- AJ Smit^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa, South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
+- C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216]
+documentclass: article
+linestretch: 1.5
+fontfamily: accanthis #accanthis #txfonts
+pagestyle: plain
+geometry: "left=2.5cm,right=2.5cm,top=2cm,bottom=2cm"
+indent: false
 output:
@@
     reference_docx: markdown/wiley_template.docx
-  html_document: default
-  pdf_document: default
+  html_document:
+  pdf_document:
+    latex_engine: xelatex
+    toc: no
 bibliography: references.bib
@@
 
-The inter- to sub-tidal habitats of temperate shorelines are usually dominated by species of macroalgae, specifically brown seaweeds of the order Laminariales; commonly referred to as kelp [@Steneck2002;@Schiel2015-vr].
-Kelps are highly productive organisms [@Mann1973;@Miller2011-pm; @Dayton1998-xw;@Krumhansl2012;@Pedersen2012-vr] which provide a 3-dimensional habitat for various fauna and flora [@anderson1997;@Leliaert2000a;@Steneck2002; @Norderhaug2012-ks; @Krumhansl2016].
-Kelp are also capable of altering the local wave and light environment by wave dampening [@Mork1996-md; @Mia2013] and shading of the understorey habitat [@duggins1990; @watanabe1992; @clark2004], as well as providing important organic subsidy through erosion of fronds and fragmentation of plant structures [@Krumhansl2012; @Filbee-Dexter2018-wp; @DeBettignies2013; @Simonson2015; @ramirez2021].
+The rocky inter- to sub-tidal habitats of temperate shorelines are usually dominated by species of macroalgae, specifically brown seaweeds of the order Laminariales; commonly referred to as kelp [@Steneck2002;@Schiel2015-vr].
+Kelps are highly productive organisms [@Mann1973;@Miller2011-pm; @Dayton1998-xw;@Krumhansl2012;@Pedersen2012a] which provide a 3-dimensional habitat for various fauna and flora [@anderson1997;@Leliaert2000a;@Steneck2002; @Norderhaug2012-ks; @Krumhansl2016].
+Kelp are also capable of altering the local wave and light environment by wave dampening [@Mork1996-md; @Mia2013] and shading of the understorey habitat [@duggins1990; @watanabe1992; @clark2004], as well as providing important organic subsidy through erosion of fronds and fragmentation of plant structures [@Krumhansl2012; @filbee2018; @DeBettignies2013; @Simonson2015; @ramirez2021].
 Due to kelps’ ecological role in the nearshore, they are regarded as ecological engineers as well as important influencer of biodiversity along temperate shorelines [@FowlerWalker2005;@graham2004;@Harley2012;@Smale2017; @Wing2007-tv;@Wernberg2008].
@@
 During times of high wave energy kelps can become dislodged from the substrate [@seymour1989; @Thomsen2004] and are transported via ocean currents, eventually deposited on beaches or offshore areas [@hobday2000a].
-The kelp deposited on beaches provides important organic subsidies for coastal invertebrate communities, as well as a resource which can be harvested for economic purposes [@kirkman1997;@troell2006; @holden2018].
+The kelp deposited on beaches provides important organic subsidies for coastal invertebrate communities [@griffiths1983; @michaud2019; @gilson2021], as well as a resource which can be harvested for economic purposes [@kirkman1997;@troell2006; @holden2018].
 Along with the organic input are 'hitchhikers', which are organisms that remain attached during offshore transport once dislodged [@highsmith1985; @holmquist1994; @macaya2016; @thiel2005].
 The transport of other organisms attached to kelp is regarded as an important processes in invasive species intrusion [@lewis2005; @McCormick2008; @nikula2010; @batista2018] and marine connectivity [@norton1992, @ingolfsson1995; @hobday2000a; @fraser2011].
-The formation of kelp-rafts can also occur, providing habitat complexity in the pelagic environment for various fish species [@kingsford1992; @kingsford1995], Kelps which are transported offshore can become entangled to form a kelp-raft, but there is no clear mechanism for kelp-raft formation.
-The structural integrity of individual kelp or rafts will degrade over time through a combination of biotic and abiotic factors which ultimately results in sinking.
-Once settled on the seafloor, the kelp will degrade further over time resulting in organic input into offshore areas; which may be transported back towards the coast through upwelling processes [@Krumhansl2012; @Filbee-Dexter2018-wp; @ramirez2021].
+Kelps which are transported offshore can become entangled to form a kelp-raft, providing habitat complexity in the pelagic environment for various fish species [@kingsford1992; @kingsford1995].
+However, there is no clear mechanism for kelp-raft formation.
+The structural integrity of individual kelp or rafts will degrade over time through a combination of biotic and abiotic factors which ultimately results in sinking [@graiff2016; @macaya2016; @smith2002; @thiel2005].
+Once settled on the seafloor, the kelp will degrade further over time resulting in organic input into offshore areas; which may be transported back towards the coast through upwelling processes [@Krumhansl2012; @filbee2018; @ramirez2021].
 
@@
 Ocean currents are the main influencer of trajectory as well as drag and windage, which is the direct effect of wind on a floating object [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008].
-More recent work by @coppin2021 has highlighted the importance of both hydrodynamic and wind drag in entrainment patterns within oceanographic mesoscale processes, more specifically eddies.
+More recent work by @coppin2022 has highlighted the importance of both hydrodynamic and wind drag in entrainment patterns within oceanographic mesoscale processes, more specifically eddies.
 
@@
 For example, kelps are highly morphologically diverse which may influence buoyancy and overall drag characteristics.
-Morphological characteristics, sea surface temperatures, and light also influence the time which kelps remain afloat at the surface; known as raft-times.
+Morphological characteristics, sea surface temperatures, and light also influence the time which kelps remain afloat at the surface; known as raft-time.
 Macroalgae rafts also vary greatly in size, therefore studies investigating a specific raft size limits our understanding of that particular scenario as the findings cannot be inferred for other sizes of kelp rafts.
 Finally, it has been well established that morphological characteristics of macroalgae are influenced by local abiotic factors, which can vary greatly along coastlines.
-Although factors such as light, turbidity, and nutrients have been shown to influence macroalgae morphology; temperature and waves are regarded as the main influencers [@Hurd2000;@Denny2002a;@Thomsen2004;@FowlerWalker2005;@Wernberg2005c;@mabin2013;@bekkby2014; @coppin2021].
+Although factors such as light, turbidity, and nutrients have been shown to influence macroalgae morphology; temperature and waves are regarded as the main influencers [@Hurd2000;@Denny2002a;@Thomsen2004;@FowlerWalker2005;@Wernberg2005c;@mabin2013;@bekkby2014; @coppin2022].
 How the variable response in morphological characteristics influences sink rates, buoyancy, and drag is also yet to be established.
-There is also no current work where rafts may form on the surface and the role of eddies in this regard.
-It is most likely that eddies might play a role in this regard as they have shown to cause collection of other flotsam at the surface via entrainment @onink2019; @dasaro2018; @brach2018.
-For example, a study by @beron2019 showed that the inclusion of inertia in numerical experiments resulted in entrainment patterns of flotsam in eddies which more closely reflected observational data.
+There is also no current work where rafts may form on the surface.
+It is most likely that eddies might play a role in this regard as they have shown to cause collection of other flotsam at the surface via entrainment [@onink2019; @dasaro2018; @brach2018].
+For example, a study by @beron2019 showed that the inclusion of inertia in numerical experiments resulted in entrainment patterns of flotsam in eddies which more closely reflected in observational data.
 
@@
 In addition, since individual kelp has a higher probability of dislodgement compared to kelps in which holdfasts are clustered together, an individual-based approach will reflect real-world scenarios [@Thomsen2004].
-The novel approach of including hydrodynamic and wind drag based on site-specific morphological characteristics by @coppin2021 will be used and the current work will build on this approach by the inclusion of morphological specific buoyancy and sinking rates.
+The approach of including hydrodynamic and wind drag based on site-specific morphological characteristics by @coppin2022 will be used and the current work will build on this approach by the inclusion of morphological specific buoyancy and sinking rates.
 
 # Methods
+
+## Study site
+
+The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A).
+Kommetjie, located on the Atlantic side of the Cape Peninsula (Figure 2B), was chosen as a release site.
+The hydrodynamics in the study region are characterized by a complex interaction of wind, sea waves and ocean swell, and are also influenced by larger-scale regional oceanic processes.
+The larger processes in the region can be attributed to the Agulhas Current (AC) and the Benguela Current (BC) [@lutjeharms2000;@lutjeharms2007;@garzoli1996;@hutchings2009;@rubio2009].
+The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@lutjeharms2000;@lutjeharms2007].
+
+The region where the retroflection occurs is known as the Agulhas Retroflection and consists of high levels of mesoscale variability [@garzoli1996].
+The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988;@lutjeharms2006; @lutjeharms2007].
+Agulhas Rings are the main contributors to the Agulhas leakage which is the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015].
+The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009].
+The Benguela region ---known as the Benguela Upwelling System (BUS)--- is characterized by strong coastal upwelling which is primarily driven by equatorward winds [@shannon1996;@hardman2003;@veitch2018; @veitch2017].
+The BUS starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996].
+The region is also characterized by large amounts of mesoscale variability in the form of eddys and filaments [@blanke2002; @blanke2005; @rubio2009].
+Inshore, the BUS is dominated by a northwestward flow, known as the Benguela Jet, due to topographical steering, wind stress and interactions with passing Agulhas Rings and eddys [@stramma1989; @veitch2010; @veitch2017; @veitch2018; @hutchings2009].
+The stability of the jet varies seasonally with more stable flows in winter; in January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results in a weak narrow inshore branch that flows along the shelf edge and a strong flowing north-westwards offshore branch.
+The area is also characterized by high upwelling activity driven by strong westerly winds [@hutchings2009].
+These winds are most intense during Austral summer (December-March) which favours upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet which favours upwelling waters contributing to the across-shore density gradient [@ragoasha2019].
+In Austral winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela Jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].
+
+```{r map chp4, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Map of the study domain and the main features pertaining to transport in the region.
+In panel A, the map of the western portion of the South African coastline and the main currents is presented.
+The resulting eddies and filaments forming from the Agulhas retroflection are indicated, and the position of the study area is drawn in as a black rectangle.
+Details of the study area are given in panel B.
+Here the splitting of the Benguela Jet, eddy features, and the general flow within the study domain are presented.
+The release site of simulated kelp ‘particles’ is represented by a black triangle."}
+
+library(png)
+library(grid)
+library(gridExtra)
+
+#map_chp3A <- rasterGrob(as.raster(readPNG("map/chp3_mapA.png")))
+#map_chp3B <- rasterGrob(as.raster(readPNG("map/chp3_mapB.png")))
+#ggarrange(map_chp3A, map_chp3B, ncol = 2, nrow = 1, labels = "AUTO", align = c("h"))
+map_chp3_combined <- rasterGrob(as.raster(readPNG("map/chp3_map_final_v2.png")))
+ggarrange(map_chp3_combined)
+
+```
+
+## Simulations
+
+Simulations used daily outputs of velocity vectors (meridonal and zonal) and salinity (psu) for varying depths for the study region.
+To account for the role of morphological characteristics in influencing drift trajectory in terms of hydrodynamic and wind drag [@coppin2022] simulations were run separately for each site per season.
+Morphological characteristics of macroalgae has been shown to be based on the local wave and temperature environment, hence the need for site specific parameters with regards to morphological characteristics.
+Running simulations per season allowed incorporation of seasonal raft-times identified in pervious work [@graiff2016 and @graiff2013].
+The outputs from each simulation (site) was pooled based on season.
+
+Simulations were run using the OceanParcels software (available from https://oceanparcels.org/).
+All simulations were run using a Runge-Kutte fourth order advection scheme, while custom kernels simulated the effects of hydrodynamic and wind drag on overall trajectory, seasonal raft-times, sinking and settling on seafloor of virtual kelp particles; all important processes in the determination of surface and bottom sinks.
+The simulations were run for two years, which were divided into the relevant seasons (see table 2).
+For incorporating drag, the site specific buoyancy calculations were used to determine the magnitude of hydrodynamic (portion below surface) and wind (portion above the surface) while virtual kelp particles were on the surface and then converted to full hydrodynamic drag exposure (100%) once the virtual kelp started sinking; as wind would no longer act as a direct influencer of trajectory once below the ocean surface.
+
+Although OceanParcels is a robust and flexible tool, it does have some limitations when combining mixed field of wind and ocean vectors with sinking behaviour.
+Essentially, problems arise when including wind fields with ocean vector fields when simulations include sinking.
+This creates limitations when running simulations and requires the approach to be adapted.
+It was decided to incorporate an approach used in previous work by Putman et al.
+2018, Putman et al.
+2020, need more here, such that windage is estimated as 50% of the current speed.
+Therefore, the total potential wind was estimated as 50% of the current speed to attain the velocity vectors for the estimated wind field, which was then used to calculate the total effect of windage based on buoyancy characteristics (magnitude of surface area exposed to wind and hydrodynamic drag respectively.
+
+Kelp floating on the surface of the ocean will begin to sink once the structural integrity of air-filled structures of the plants fails and becomes flooded.
+It is also possible for sinking to begin due to an increase in epibiont biomass which overwhelms the buoyancy of kelp, however, this is beyond the scope of this study.
+The target species of this study, *Ecklonia maxima* (Osbeck) Papenfuss, contains a hollow stipe and a single air-filled pneumatocyst which acts to keep the plant upright in the water column, allowing the light-saturated surface waters to be reached.
+These structures also cause the plant to float if dislodged and therefore were taken into account when calculating the overall volume of the plant; which in turn was used to calculate other important parameters (buoyancy, drag, and sink rate).
 
@@
 In instances where the morphological data needed were not available, reasonable estimates were used based on previous studies and experience.
-To calculate the surface area and volume for kelps from each site, geometric shapes were used which most closely matched the shape of the relevant kelp morphological characteristics as per @coppin2021.
+
+The exception was the air-filled stipe, which helps to keep the plant upright within the water column, where ad-hoc fieldwork was performed to attain better estimates of the amount of air within the stipe.
+The hollowness of E.
+maxima has been investigated as a part of a previously by Rothman et al.
+(2017) who showed that the relative hollowness did not significantly differ among sites, with 86-90% of the total stipe length containing air.
+Therefore, based on these results, it is assumed that *E.
+maxima* individuals will exhibit the same degree of hollowness amongst populations along South African coastline.
+An estimate of stipe thickness of the hollow sections was used in conjunction with the 90% of the stipe length shown by @Rothman2017-pn to calculate a reasonable estimate for the volume of air in the stipe.
+This is an important aspect as it directly effects the buoyancy of an individual floating kelp which ultimately determines the magnitude of exposure to hydrodynamic and wind drag.
+
+To calculate the surface area and volume for kelps from each site, geometric shapes were used which most closely matched the shape of the relevant kelp morphological characteristics as per @coppin2022.
 The shapes assigned to each of the morphological features were used in the calculation of buoyancy, drag, and sinking rate.
@@
 To effectively model the trajectory of a floating object, buoyancy needs to be calculated.
-The buoyancy will determine the amount of exposure to, and ultimately magnitude, of hydrodynamic and wind drag which in turn influences the trajectory of the object to the variable current and wind fields.
-The portion of an object above and below the surface of the water is determined by the net effect of forces acting on the object, namely buoyancy force (BF) and the weight of the object.
+The buoyancy will determine the amount of exposure to, and ultimately magnitude, of hydrodynamic and wind drag which in turn influences the trajectory of the object to the variable current and wind fields [@dietrich1982; @lobelle2021; @kooi2017].
+The portion of an object above and below the surface of the water is determined by the net effect of forces acting on the object, namely buoyancy force (BF) and the weight of the object[@dietrich1982; @lobelle2021; @kooi2017].
 The BF is the force pushing the object out of the water and the weight is the force pulling the object down.
@@
 This in turn can be used to calculate the percentage of the object above the surface of the water (exposed to wind).
-The percentage of the object below and above the water surface was used in the calculation of hydrodynamic and wind drag when the virtual kelp is at the surface.
+The percentage of the object below and above the water surface was used in the calculation of hydrodynamic and wind drag when the virtual kelp is at the surface, while only hydrodynamic drag is considered for horizontal transport when sinking.
 Since morphological characteristics are influenced by the local wave and temperature environment, the buoyancy of virtual kelp particles may vary from site to site.
@@
 ## Sink Rate
-
-Kelp floating on the surface of the ocean will begin to sink once the structural integrity of air-filled structures of the plants fails and becomes flooded.
-It is also possible for sinking to begin due to an increase in epibiont biomass which overwhelms the buoyancy of kelp, however, this is beyond the scope of this study.
-The target species of this study, *E.
-maxima*, contains a hollow stipe and a single air-filled pneumatocyst which serve to keep the plant upright in the water column, allowing the light-saturated surface waters to be reached.
-These structures also cause the plant to float if dislodged and therefore was taken into account when calculating the overall volume of the plant; which in turn was used to calculate other important parameters (buoyancy, drag, and sink rate).
-
-## Vertical Velocity
 
@@
 
-Where $p_s$ is the density of the particle, $p__l$ the density of the medium, $g$ is the gravitational acceleration constant, and $V$ is the total volume of the particle.
+Where $p_s$ is the density of the particle, $p_l$ the density of the medium, $g$ is the gravitational acceleration constant, and $V$ is the total volume of the particle.
 In order to accurately calculate the density of the kelp based on volume and mass, the volume of the hollow section of the stipe was subtracted from the overall volume.
@@
 
-## Simulations
-
-Simulations used daily outputs of velocity vectors (meridonal and zonal) and salinity (psu) for varying depths for the study region.
-To account for the role of morphological characteristics in influencing drift trajectory in terms of hydrodynamic and wind drag [@coppin2021] simulations were run separately for each site per season.
-Morphological characteristics of macroalgae has been shown to be based on the local wave and temperature environment, hence the need for site specific parameters with regards to morphological characteristics.
-Running simulations per season allowed incorporation of seasonal raft-times identified in pervious work [@graiff2016 and @graiff2013].
-The outputs from each simulation (site) was pooled based on season.
-
-Simulations were run using the OceanParcels software (available from https://oceanparcels.org/).
-All simulations were run using a Runge-Kutte fourth order advection scheme, while custom kernels simulated the effects of hydrodynamic and wind drag on overall trajectory, seasonal raft-times, sinking and settling on seafloor of virtual kelp particles; all important processes in the determination of surface and bottom sinks.
-The simulations were run for 2019, which was divided into various seasons (see table 2).
-For incorporating drag, the site specific buoyancy calculations were used to determine the magnitude of hydrodynamic (portion below surface) and wind (portion above the surface) while virtual kelp particles were on the surface and then converted to full hydrodynamic drag exposure (100%) once the virtual kelp started sinking; as wind would no longer act as a direct influencer of trajectory once below the ocean surface.
-
 ## Raft-time
@@
 Raft-time, also referred to as raft longevity, is determined by epibiont load, macroalgal growth rate and the loss of kelp biomass over time through fragmentation.
-Epibiont load consists of epiphytic and bryozoan species which tend to grow more rapidly in higher light and temperature environments, such as the ocean surface [@thiel2005; @thiel2005; @graiff2016].
+Epibiont load consists of epiphytic and bryozoan species which tend to grow more rapidly in higher light and temperature environments, such as the ocean surface [@smith2002; @thiel2005; @graiff2016].
 The increase in epibiont load reduces the buoyancy of floating macroalgal rafts over time eventually causing them to sink to the seafloor, while also potentially increasing drag [@hobday2000; @tala2013; @craw2018].
 Fragmentation of kelp biomass over time may also reduce raft times further, while growth may counteract the effects of epibiont load and fragmentation provided the environmental conditions are favorable [@graiff2016; @macaya2016].
-For instance a study by @rothausler2011 investigated the abiotic and biotic factors influencing raft time and dispersal potential of the giant kelp *M.
-pyrifera* by a combination of tethering experiments and field surveys.
+For instance a study by @rothausler2011 investigated the abiotic and biotic factors influencing raft time and dispersal potential of the giant kelp *Macrocystis pyrifera* (Linnaeus) C.Agardh by a combination of tethering experiments and field surveys.
 The results showed the physiological performance of kelp declined with increasing epiphyte biomass and that higher temperatures increased growth rate of epibionts and fragmentation of kelp.
@@
 ```{r morphology surface area table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
+set_flextable_defaults(fonts_ignore=FALSE)
 site_surface_area <- set_caption(site_surface_area, "Table 1:Summary of the attributes and estimates used to calculate the overall site-specific surface area of kelp individuals used in the various simulations.")
@@
 site_surface_area
-
-```
-
-## Simulations
-
-Simulations used daily outputs of velocity vectors (meridonal and zonal) and salinity (psu) for varying depths for the study region.
-To account for the role of morphological characteristics in influencing drift trajectory in terms of hydrodynamic and wind drag [@coppin2021] simulations were run separately for each site per season.
-Morphological characteristics of macroalgae has been shown to be based on the local wave and temperature environment, hence the need for site specific parameters with regards to morphological characteristics.
-Running simulations per season allowed incorporation of seasonal raft-times identified in pervious work [@graiff2016 and @graiff2013].
-The outputs from each simulation (site) was pooled based on season.
-
-Simulations were run using the OceanParcels software (available from https://oceanparcels.org/).
-All simulations were run using a Runge-Kutte fourth order advection scheme, while custom kernels simulated the effects of hydrodynamic and wind drag on overall trajectory, seasonal raft-times, sinking and settling on seafloor of virtual kelp particles; all important processes in the determination of surface and bottom sinks.
-The simulations were run for two years, which were divided into the relevant seasons seasons (see table 2).
-
-```{r simulation table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="TBD."}
-
+```
+
+```{r simulation table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
+set_flextable_defaults(fonts_ignore=FALSE)
 chapter4_table2 <- read_csv2(file = "tables/chapter_4/simulation_table.csv")
@@
 The start of each year was the beginning of Austral summer.")
-chapter4_table2
 ```
@@
 
-To compare virtual kelp trajectories between sites the mean trajectories were calculated, and density distribution maps were produced.
+To compare virtual kelp trajectories between sites the mean trajectories were calculated, and density distribution maps were produced for each phase of the virtual kelp life cycle (surface, subsurface, beached and sunk).
 Density distribution maps are an effective method for assessing pathways and ocean connectivity [@vanSebille2018].
-The distance particles traveled was also measured and used to produce boxplots.
+To aid in seasonal comparison of trajectories, proportions of each kelp phase at the end of each simulation run was calculated and compared.
+The different phases were surface, subsurface, beached, sunk and out of bounds.
+
+In addition to the above analyses, the distance particles traveled was also measured and used to produce boxplots to aid in discerning patterns and distributions from the density distribution maps.
 Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particles [@vanSebille2018].
 If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity because of momentum energy transfer [@putman2020; @putman2018; @vanSebille2018].
-To aid in seasonal comparison of trajectories, proportions of each kelp phase at the end of each simulation run was calculated and compared.
-The different phases were surface, subsurface, beached, sunk and out of bounds.
+Distance boxplots are presented in the appendix.
 
@@
   #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
+  coord_cartesian(xlim = c(17.5, 19), ylim = c(-34.5, -32.5)) +
   labs(x = expression(Longitude*degree*S)) +
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  ggtitle("Year 1") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  ggtitle("Year 2") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  coord_cartesian(xlim = c(17.5, 19), ylim = c(-34.5, -32.5)) +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   #geom_path(data = total_autumn_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  coord_cartesian(xlim = c(17.5, 19), ylim = c(-34.5, -32.5)) +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  coord_cartesian(xlim = c(17.5, 19), ylim = c(-34.5, -32.5)) +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   #geom_path(data = total_winter_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  coord_cartesian(xlim = c(17.5, 19), ylim = c(-34.5, -32.5)) +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  coord_cartesian(xlim = c(17.5, 19), ylim = c(-34.5, -32.5)) +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
   #geom_path(data = total_spring_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(12, 20), ylim = c(-37, -29)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  coord_cartesian(xlim = c(17.5, 19), ylim = c(-34.5, -32.5)) +
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
   guides(fill=guide_legend(title="Number of particles")) +
@@
 The overall particle direction during summer for both years was north-westward.
-The only visible differences between years 1 and 2 was entrainment of particles in an eddy in year 1, as well as a higher concentration of particles closer to the coast (figure 2).
+The only visible differences between years 1 and 2 was entrainment of particles in an eddy in year 1, as well as a higher concentration of particles closer to the coast.
 The concentration of particles in the north-westward direction was lower during  year 1 of autumn compared to year 2, as well as no entrainment patterns.
@@
 
-```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=10, fig.width=6, fig.cap="Figure 2: Hexagonal heatmaps for all virtual surface kelp for both years and all seasons.
+```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=5, fig.width=6, fig.cap="Hexagonal heatmaps for all virtual surface kelp for both years and all seasons.
 The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon."}
@@
 
-ggarrange(summer_heatmap_surface, summer2_heatmap_surface, autumn_heatmap_surface, autumn2_heatmap_surface, winter_heatmap_surface, winter2_heatmap_surface, spring_heatmap_surface, spring2_heatmap_surface, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
-
-ggsave(filename = "figures/chapter_4/figure_2.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+ggarrange(summer_heatmap_surface, summer2_heatmap_surface, autumn_heatmap_surface, autumn2_heatmap_surface, winter_heatmap_surface, winter2_heatmap_surface, spring_heatmap_surface, spring2_heatmap_surface, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 4, nrow = 2, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_2.jpeg", plot = last_plot(), scale = 1, dpi = 300, width = 10, height = 6)
 
@@
 
-```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=10, fig.width=6,fig.cap="Figure 3: Hexagonal heatmaps for all virtual sub-surface kelp for both years and all seasons.
+```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=9, fig.width=5.5,fig.cap="Hexagonal heatmaps for all virtual sub-surface kelp for both years and all seasons.
 The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon."}
@@
 
-ggarrange(summer_heatmap_subsurface, summer2_heatmap_subsurface, autumn_heatmap_subsurface, autumn2_heatmap_subsurface, winter_heatmap_subsurface, winter2_heatmap_subsurface, spring_heatmap_subsurface, spring2_heatmap_subsurface, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
-
-ggsave(filename = "figures/chapter_4/figure_3.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+ggarrange(summer_heatmap_subsurface, summer2_heatmap_subsurface, autumn_heatmap_subsurface, autumn2_heatmap_subsurface, winter_heatmap_subsurface, winter2_heatmap_subsurface, spring_heatmap_subsurface, spring2_heatmap_subsurface, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 4, nrow = 2, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_3.jpeg", plot = last_plot(), scale = 1, dpi = 300, width = 10, height = 6)
 ```
@@
 
-```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=10, fig.width=6, fig.cap="Figure 4: Hexagonal heatmaps for all virtual sunken kelp for both years and all seasons.
+```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=9, fig.width=5.5, fig.cap="Hexagonal heatmaps for all virtual sunken kelp for both years and all seasons.
 The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon."}
@@
 
-ggarrange(summer_heatmap_sink, summer2_heatmap_sink, autumn_heatmap_sink, autumn2_heatmap_sink, winter_heatmap_sink, winter2_heatmap_sink, spring_heatmap_sink, spring2_heatmap_sink, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
-
-ggsave(filename = "figures/chapter_4/figure_4.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+ggarrange(summer_heatmap_sink, summer2_heatmap_sink, autumn_heatmap_sink, autumn2_heatmap_sink, winter_heatmap_sink, winter2_heatmap_sink, spring_heatmap_sink, spring2_heatmap_sink, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 4, nrow = 2, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_4.jpeg", plot = last_plot(), scale = 1, dpi = 300, width = 10, height = 6)
 ```
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  ggtitle("Year 1") +
   guides(fill=guide_legend(title="Site")) +
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
+  ggtitle("Year 2") +
   guides(fill=guide_legend(title="Site")) +
@@
 
-```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=10, fig.width=6, fig.cap="Figure 5: Hexagonal heatmaps for all virtual beached kelp for both years and all seasons.
+```{r message=FALSE, echo=FALSE, include=TRUE, warning=FALSE, fig.height=9, fig.width=5, fig.cap="Hexagonal heatmaps for all virtual beached kelp for both years and all seasons.
 The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon."}
@@
 
-ggarrange(summer_heatmap_beached, summer2_heatmap_beached, autumn_heatmap_beached, autumn2_heatmap_beached, winter_heatmap_beached, winter2_heatmap_beached, spring_heatmap_beached, spring2_heatmap_beached, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
-
-ggsave(filename = "figures/chapter_4/figure_5.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+ggarrange(summer_heatmap_beached, summer2_heatmap_beached, autumn_heatmap_beached, autumn2_heatmap_beached, winter_heatmap_beached, winter2_heatmap_beached, spring_heatmap_beached, spring2_heatmap_beached, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 4, nrow = 2, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_5.jpeg", plot = last_plot(), scale = 1, dpi = 300, width = 10, height = 6)
 ```
@@
   ggtitle("Year 1") +
-  guides(fill=guide_legend(title="Number of particles")) +
+  guides(linetype=guide_legend(title="Site")) +
   theme_bw() +
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Year 1") +
-  guides(fill=guide_legend(title="Number of particles")) +
+  ggtitle("Year 2") +
+  guides(linetype=guide_legend(title="Site")) +
   theme_bw() +
@@
   labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Year 1") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-```
-
-```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.height=10,fig.width=6,fig.cap="Figure 6: The mean point distrubtion for depth of all surface virtual kelp for both years and all seasons.
+  ggtitle("Year 2") +
+  guides(linetype=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.width = 6, fig.asp=1.5, fig.cap="The mean point distrubtion for depth of all surface virtual kelp for both years and all seasons.
 The surface kelp distributions are grouped per site and are represented by varying shapes."}
 
-ggarrange(summer_point, summer2_point, autumn_point, autumn2_point, winter_point, winter2_point, spring_point, spring2_point, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
-
-ggsave(filename = "figures/chapter_4/figure_6.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+ggarrange(summer_point, summer2_point, autumn_point, autumn2_point, winter_point, winter2_point, spring_point, spring2_point, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 4, nrow = 2, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_6.jpeg", plot = last_plot(), scale = 1.5, dpi = 300, width = 10, height = 6)
 ```
@@
 Some sites from False Bay (Baboon rock, Miller’s point, A-frame and Kalk Bay) were absent from sink locations.
+
+```{r mean sink locations, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}
+
+total_summer_sink_mean <- total_summer_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_summer2_sink_mean <- total_summer2_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_autumn_sink_mean <- total_autumn_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_autumn2_sink_mean <- total_autumn2_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_winter_sink_mean <- total_winter_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_winter2_sink_mean <- total_winter2_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_spring_sink_mean <- total_spring_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+
+total_spring2_sink_mean <- total_spring2_sink %>% # calculate mean trajectory
+  group_by(t, site) %>%
+  summarise(across(lon:lat, mean, na.rm= TRUE))
+# summer
+
+summer_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_summer_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(shape=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+summer2_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_summer2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
+  guides(shape=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# autumn
+
+autumn_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_autumn_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(shape=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+autumn2_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_autumn2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
+  guides(shape=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# winter
+
+winter_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_winter_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(shape=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+winter2_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_winter2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
+  guides(shape=guide_legend(title="Site")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+# spring
+
+spring_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_spring_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 1") +
+  guides(shape=guide_legend(title="Site")) +
+  #theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12))
+
+spring2_depth_sink <- ggplot() +
+  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
+  geom_point(data = total_spring2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
+  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
+  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
+               size = 0.4, colour = "black", fill = "gray") +
+  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
+  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
+  labs(x = expression(Longitude*degree*S)) +
+  labs(y = expression(Latitude*degree*E)) +
+  ggtitle("Year 2") +
+  guides(shape=guide_legend(title="Site")) +
+  #theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))+
+  theme(axis.text=element_text(size=12),
+        axis.title=element_text(size=12)) 
+```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.width=6, fig.asp=1.5, fig.cap="Point distrubtion of all sunken virtual kelp for both years and all seasons.The sunken kelp distributions are grouped per site and are represented by varying shapes."}
+
+ggarrange(summer_depth_sink, summer2_depth_sink, autumn_depth_sink, autumn2_depth_sink, winter_depth_sink, winter2_depth_sink, spring_depth_sink, spring2_depth_sink, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 4, nrow = 2, labels = c("Summer","","Autumn","","Winter","","Spring",""))
+
+ggsave(filename = "figures/chapter_4/figure_7.jpeg", plot = last_plot(), scale = 1.5, dpi = 300, width = 10, height = 6)
+
+```
 
@@
 
-```{r mean sink locations, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}
-
-total_summer_sink_mean <- total_summer_sink %>% # calculate mean trajectory
-  group_by(t, site) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-total_summer2_sink_mean <- total_summer2_sink %>% # calculate mean trajectory
-  group_by(t, site) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-total_autumn_sink_mean <- total_autumn_sink %>% # calculate mean trajectory
-  group_by(t, site) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-total_autumn2_sink_mean <- total_autumn2_sink %>% # calculate mean trajectory
-  group_by(t, site) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-total_winter_sink_mean <- total_winter_sink %>% # calculate mean trajectory
-  group_by(t, site) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-total_winter2_sink_mean <- total_winter2_sink %>% # calculate mean trajectory
-  group_by(t, site) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-total_spring_sink_mean <- total_spring_sink %>% # calculate mean trajectory
-  group_by(t, site) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-
-total_spring2_sink_mean <- total_spring2_sink %>% # calculate mean trajectory
-  group_by(t, site) %>%
-  summarise(across(lon:lat, mean, na.rm= TRUE))
-# summer
-
-summer_depth_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_summer_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
-  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Year 1") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-summer2_depth_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_summer2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
-  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("Year 2") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-# autumn
-
-autumn_depth_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_autumn_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
-  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-autumn2_depth_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_autumn2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
-  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-# winter
-
-winter_depth_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_winter_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
-  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-winter2_depth_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_winter2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
-  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-# spring
-
-spring_depth_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_spring_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
-  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  #theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12))
-
-spring2_depth_sink <- ggplot() +
-  geom_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), colour = "grey", size = 0.2)+
-  geom_point(data = total_spring2_sink_mean, aes(x = lon, y = lat, shape = site), size = 2.5) +
-  scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13))+
-  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
-               size = 0.4, colour = "black", fill = "gray") +
-  #geom_path(data = total_summer_mean,aes(x = lon, y = lat, color = 'red'), show.legend =     FALSE) +
-  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
-  labs(x = expression(Longitude*degree*S)) +
-  labs(y = expression(Latitude*degree*E)) +
-  ggtitle("") +
-  guides(fill=guide_legend(title="Number of particles")) +
-  #theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))+
-  theme(axis.text=element_text(size=12),
-        axis.title=element_text(size=12)) 
-```
-
-```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.height=10, fig.width=6,fig.cap="Figure 7: Point distrubtion of all sunken virtual kelp for both years and all seasons.The sunken kelp distributions are grouped per site and are represented by varying shapes."}
-
-ggarrange(summer_depth_sink, summer2_depth_sink, autumn_depth_sink, autumn2_depth_sink, winter_depth_sink, winter2_depth_sink, spring_depth_sink, spring2_depth_sink, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
-
-ggsave(filename = "figures/chapter_4/figure_7.jpeg", plot = last_plot(), scale = 1, dpi = 300)
-
-```
-
-There are clear differences between the amount of kelp that eventually sank, beached or traveled outside the study domain, between seasons with variation between years.
-During summer months, no particles traveled outside the study domain.
-The highest proportion of particles that traveled outside the domain occurred during spring months and to a lesser magnitude during autumn and winter months.
-The highest proportion of kelp particles that beached occurred during winter months, in particular year 2, where almost all the particles beached.
-Kelps released from Buffelsbaai were the only population to travel outside of False Bay during all times of the year, while the remaining sites (Baboon rock, Bachata rock, Miller’s point, and Kalk Bay) only traveled outside the bay during winter of year 2.
-Kelps released from Betty’s Bay (southern site extreme) usually become beached, while kelps from St.
-Helena Bay (northern extreme) had similar proportions of beached and sunken kelp across seasons and years.
-
 ```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
@@
 
-```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.height=10, fig.width=8, fig.cap="Figure 8: Porportion of particles for each phase at the end of the simulation run for each respective season and year.
-The size of the point is related to the amount of particles within that life cycle phase at the end of each simulation run."}
-
-ggarrange(dot_phase_summer, dot_phase_summer2, dot_phase_autumn,dot_phase_autumn2, dot_phase_winter, dot_phase_winter2, dot_phase_spring, dot_phase_spring2, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 4, labels = c("Summer","","Autumn","","Winter","","Spring",""))
-
-ggsave(filename = "figures/chapter_4/figure_8.jpeg", plot = last_plot(), scale = 2, dpi = 300)
-```
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Porportion of particles for each phase at the end of the simulation run for Austral Summer and Autumn for each respective year.
+The size of the point is related to the amount of particles within that life cycle phase at the end of each simulation run.", fig.height=8, fig.width=8}
+
+ggarrange(dot_phase_summer, dot_phase_summer2, dot_phase_autumn,dot_phase_autumn2, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 2, labels = c("A","","B","","C","","D",""))
+
+#ggsave(filename = "figures/chapter_4/figure_8.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+```
+
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Porportion of particles for each phase at the end of the simulation run for Austral Winter and Spring for each respective year.
+The size of the point is related to the amount of particles within that life cycle phase at the end of each simulation run.", fig.height=8, fig.width=8}
+
+ggarrange(dot_phase_winter, dot_phase_winter2, dot_phase_spring, dot_phase_spring2, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = TRUE, ncol = 2, nrow = 2, labels = c("C","","D",""))
+
+#ggsave(filename = "figures/chapter_4/figure_8.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+```
+
+There are clear differences between the amount of kelp that eventually sank, beached or traveled outside the study domain, between seasons with variation between years.
+During summer months, no particles traveled outside the study domain.
+The highest proportion of particles that traveled outside the domain occurred during spring months and to a lesser magnitude during autumn and winter months.
+The highest proportion of kelp particles that beached occurred during winter months, in particular year 2, where almost all the particles beached.
+Kelps released from Buffelsbaai were the only population to travel outside of False Bay during all times of the year, while the remaining sites (Baboon rock, Bachata rock, Miller’s point, and Kalk Bay) only traveled outside the bay during winter of year 2.
+Kelps released from Betty’s Bay (southern site extreme) usually become beached, while kelps from St.
+Helena Bay (northern extreme) had similar proportions of beached and sunken kelp across seasons and years.
 
@@
 The findings from this study deepens our understanding the role mesoscale features play in influencing macroalgal trajectory patterns and distribution.
-The importance of including drag components in relation to surface area as shown by @coppin2021 was confirmed, however, the study did not find the site-specific morphological characteristics influenced mean trajectory of virtual kelp particles.
+The importance of including drag components in relation to surface area as shown by @coppin2022 was confirmed, however, the study did not find the site-specific morphological characteristics influenced mean trajectory of virtual kelp particles.
 Instead, the inclusion of site-specific morphological characteristics were important in describing entrainment patterns of virtual kelp particles.
@@
 
-Previous work investigating drifting macroalgal trajectory patterns show that macroalgae tend to follow the ocean currents [@hobday2000; @fraser2011, @coppin2021] and drag components are important for characterising entrainment patterns within mesoscale oceanographic features [@coppin2021;@brooks2019; @putman2020; @beron2016; @beron2019].
+Previous work investigating drifting macroalgal trajectory patterns show that macroalgae tend to follow the ocean currents [@hobday2000; @fraser2011, @coppin2022] and drag components are important for characterising entrainment patterns within mesoscale oceanographic features [@coppin2022;@brooks2019; @putman2020; @beron2016; @beron2019].
 This was evident from the analysis of surface and subsurface kelp distributions, which show high concentrations of virtual kelp particles within mesoscale features during times of high activity (Austral summer and winter).
@@
 Virtual kelp particles remain entrained when sinking which was as a result of the incorporated drag components.
-@coppin2021 found that purely Lagrangian particles remained on the boundary of mesoscale features with very little evidence of entrainment, while virtual kelp exhibited high entrainment patterns and focused towards the centre.
+@coppin2022 found that purely Lagrangian particles remained on the boundary of mesoscale features with very little evidence of entrainment, while virtual kelp exhibited high entrainment patterns and focused towards the centre.
 This trend has been reported in previous work by @beron2019; @miron2020).
@@
 Although the study by @rubio2009 investigated fluid parcels of water and not biological flotsam, the mechanism of weaker dynamics explains the leaking of virtual kelp particles from mesoscale features within the study domain.
+The mechanism of leakage and transport towards the coast may also be considered to be a form of barrier to offshore transport [@bracco2000].
 
@@
 Since eddies are structured vertically through the water column, this aided in the vertical transport and distribution towards the seafloor.
-Work by @mcgillicuddy2003 using an eddy-resolving model to investigate the role of mesoscale features as sources and sinks of nutrients in the North Atlantic.
-The authors showed that the mesoscale features has a significant influence of nutrients within the euphotic zone.
+Work by @mcgillicuddy2003 used an eddy-resolving model to investigate the role of mesoscale features as sources and sinks of nutrients in the North Atlantic.
+The authors showed that the mesoscale features have a significant influence of nutrients within the euphotic zone.
 In fact, the downwelling of mesoscale features is of such high magnitudes that it can counter-balance Ekman driven upwelling in the subpolar gyre.
-Although virtual kelp particles and nutrients, similar patterns are seen in the results for the sub-surface and sunken kelp particles.
+Although biological flotsam is considered in this study and nutrients, similar patterns are seen in the results for the sub-surface and sunken kelp particles.
 High concentrations of kelp particles are located in areas on the seafloor which coincide with subsurface mesoscale entrainment patterns.
@@
 @woodborne1989 characterised kelp-rafted geological clasts along the West Coast cites previous work by @gentle1987 who described a possible kelp sink location.
-@ Gentle describes kelp-rafted clasts that were recovered from dredging an area approximately 120km northwest of  Cape Columbine at an approximate depth of 350m deep.
+@gentle1987 describes kelp-rafted clasts that were recovered from dredging an area approximately 120km northwest of Cape Columbine at an approximate depth of 350m deep.
 The dredged samples consisted of sand and pebbles held together by a loose mat of seaweed.
 Unfortunately, @gentle1987 did not describe the species of seaweed but simply as kelp, but based on the most common species of kelp in South Africa, the species could be *E.
-maxima*, *Ecklonia radiata* , *Laminaria pallida* and *Macrocystis Pyrifera*.
+maxima*, *Ecklonia radiata* (C.Agardh) J.Agardh, *Laminaria pallida* Greville and *M.
+Pyrifera*.
 Mesoscale features within the study domain also influence the proportions of virtual kelp particles which beach by acting as barriers to offshore transport.
@@
 The trajectory of the kelp particle is therefore determined by the trajectory of the eddy-filament which was either north-westward or shoreward causing higher proportions of beaching during the winter months of both years.
-Mesoscale and sub-mesoscale features have been shown to act as barriers to offshore transport of other flotsam.
-
+
+Distributions of sunken virtual kelp tend to be along the shelf edge or continental shelf, patterns which are similar to those observed in natural conditions.
+A study by @filbee2016 aimed to identify predictors and describe spatial patterns of drift algae.
+The results showed that bathymetry, distance to coast, and slope were the best predictors of the spatial distributions observed during surveys.
+Drift algae was often found in depressions or flats.
+The authors attributed the spatial patterns to flow-regimes, which are slower on flats than on slopes, and also noted high accumulations on slopes and ledges close to shore.
+Similar spatial distributions are observed in the current study.
+Although sunken virtual kelp did accumulate along steep bathymetry lines, most accumulated along the continental shelf.
+The continental shelf's flat covers a larger area compared to deeper offshore areas within the study domain.
+Currents in flat areas tend to travel slower which may explain the spatial patterns in this study, as slower currents ain in accumulation of drift kelp or in this case virtual kelp [@filbee2016].
 ## Conclusion
@@
 vector_summer_surface <- ggplot() +
-  geom_line(data = total_summer_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_summer_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_summer_surface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_summer_surface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
   guides(shape=guide_legend(title="Sites")) +
-  xlab("Date") +
+  xlab("Hours") +
   ylab("Kelp Velocity (m/s)") 
 vector_summer_subsurface <- ggplot() +
-  geom_line(data = total_summer_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_summer_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_summer_subsurface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_summer_subsurface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
   guides(shape=guide_legend(title="Sites")) +
-  xlab("Date") +
+  xlab("Hours") +
   ylab("Kelp Velocity (m/s)")
@@
 vector_autumn_surface <- ggplot() +
-  geom_line(data = total_autumn_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_autumn_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_autumn_surface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_autumn_surface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
   guides(shape=guide_legend(title="Sites")) +
-  xlab("Date") +
+  xlab("Hours") +
   ylab("Kelp Velocity (m/s)") 
 vector_autumn_subsurface <- ggplot() +
-  geom_line(data = total_autumn_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_autumn_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_autumn_subsurface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_autumn_subsurface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
 vector_winter_surface <- ggplot() +
-  geom_line(data = total_winter_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_winter_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_winter_surface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_winter_surface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
   guides(shape=guide_legend(title="Sites")) +
-  xlab("Date") +
+  xlab("Hours") +
   ylab("Kelp Velocity (m/s)") 
 vector_winter_subsurface <- ggplot() +
-  geom_line(data = total_winter_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_winter_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_winter_subsurface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_winter_subsurface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
 vector_spring_surface <- ggplot() +
-  geom_line(data = total_spring_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_spring_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_spring_surface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_spring_surface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
   guides(shape=guide_legend(title="Sites")) +
-  xlab("Date") +
+  xlab("Hours") +
   ylab("Kelp Velocity (m/s)") 
 vector_spring_subsurface <- ggplot() +
-  geom_line(data = total_spring_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_spring_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_spring_subsurface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_spring_subsurface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
 vector_summer2_surface <- ggplot() +
-  geom_line(data = total_summer2_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_summer2_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_summer2_surface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_summer2_surface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
   guides(shape=guide_legend(title="Sites")) +
-  xlab("Date") +
+  xlab("Hours") +
   ylab("Kelp Velocity (m/s)") 
 vector_summer2_subsurface <- ggplot() +
-  geom_line(data = total_summer2_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_summer2_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_summer2_subsurface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_summer2_subsurface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
 vector_autumn2_surface <- ggplot() +
-  geom_line(data = total_autumn2_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_autumn2_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_autumn2_surface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_autumn2_surface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
   guides(shape=guide_legend(title="Sites")) +
-  xlab("Date") +
+  xlab("Hours") +
   ylab("Kelp Velocity (m/s)") 
 vector_autumn2_subsurface <- ggplot() +
-  geom_line(data = total_autumn2_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_autumn2_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_autumn2_subsurface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_autumn2_subsurface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
 vector_winter2_surface <- ggplot() +
-  geom_line(data = total_winter2_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_winter2_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_winter2_surface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_winter2_surface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
   guides(shape=guide_legend(title="Sites")) +
-  xlab("Date") +
+  xlab("Hours") +
   ylab("Kelp Velocity (m/s)") 
 vector_winter2_subsurface <- ggplot() +
-  geom_line(data = total_winter2_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_winter2_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_winter2_subsurface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_winter2_subsurface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
 vector_spring2_surface <- ggplot() +
-  geom_line(data = total_spring2_surface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_spring2_surface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_spring2_surface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_spring2_surface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
   guides(shape=guide_legend(title="Sites")) +
-  xlab("Date") +
+  xlab("Hours") +
   ylab("Kelp Velocity (m/s)") 
 vector_spring2_subsurface <- ggplot() +
-  geom_line(data = total_spring2_subsurface_mean, aes(x = t, y = kelpvelocity, group = site))+
-  geom_point(data = total_spring2_subsurface_mean, aes(x = t, y = kelpvelocity, shape = site)) +
+  geom_line(data = total_spring2_subsurface_mean, aes(x = age, y = kelpvelocity, group = site))+
+  geom_point(data = total_spring2_subsurface_mean, aes(x = age, y = kelpvelocity, shape = site)) +
   scale_shape_manual(values=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)) +
@@
 
-```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Figure 9: Time-series of the mean velocity of surface kelp particles grouped by site for each season and year.", fig.height=10, fig.width=8}
+```{r message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Time-series of the mean velocity of surface kelp particles grouped by site for each season and year.", fig.height=10, fig.width=8}
 
@@
 
-ggsave(filename = "figures/chapter_4/figure_9.jpeg", plot = last_plot(), scale = 0.8, dpi = 300)
+ggsave(filename = "figures/chapter_4/figure_9.jpeg", plot = last_plot(), scale = 1, dpi = 300)
 ```
@@
 
-```{r chp4 figure 10, message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Figure 11: Boxplots of the total distance travelled by all particles for both years and all seasons.
+```{r chp4 figure 10, message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Boxplots of the total distance travelled by all particles for both years and all seasons.
 The lower and upper hinges correspond to the first and third quartiles.
-The whiskers represent the range, solid black lines represent the median and black dots are outliers.", fig.height=9, fig.width=8}
+The whiskers represent the range, solid black lines represent the median and black dots are outliers.", fig.height=10, fig.width=8}
 
@@
 
-```{r chp4 figure 11, message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Figure 11: Hexagonal heatmaps for all  passive surface particles for year 1 and all seasons.
+```{r chp4 figure 11, message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Hexagonal heatmaps for all  passive surface particles for year 1 and all seasons.
+The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon.", fig.height=11, fig.width=8}
+
+ggarrange(passive_heatmap_surface, passive_heatmap_subsurface, passive_heatmap_sink, passive_heatmap_beached, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = FALSE, ncol = 2, nrow = 2, labels = c("Surface","Subsurface","Sunk","Beached"))
+
+ggsave(filename = "figures/chapter_4/figure_11.jpeg", plot = last_plot(), scale = 1, dpi = 300)
+```
+
+```{r chp4 figure 12, message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Hexagonal heatmaps for all  phases passive surface particles for year 2 and all seasons.
 The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon.", fig.height=10, fig.width=8}
 
-ggarrange(passive_heatmap_surface, passive_heatmap_subsurface, passive_heatmap_sink, passive_heatmap_beached, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = FALSE, ncol = 2, nrow = 2, labels = c("Surface","Subsurface","Sunk","Beached"))
-
-ggsave(filename = "figures/chapter_4/figure_11.jpeg", plot = last_plot(), scale = 1, dpi = 300)
-```
-
-```{r chp4 figure 12, message=FALSE, echo=FALSE, warning=FALSE, include=TRUE, fig.cap="Figure 13: Hexagonal heatmaps for all  phases passive surface particles for year 2 and all seasons.
-The plane has been divided into regular hexagons and then are filled according to the number of particles within each hexagon.", fig.height=10, fig.width=8}
-
 ggarrange(passive_heatmap_surface2, passive_heatmap_subsurface2, passive_heatmap_sink2, passive_heatmap_beached2, font.label = list(size = 12, face = "bold"), legend = c("bottom"), common.legend = FALSE, ncol = 2, nrow = 2, labels = c("Surface","Subsurface","Sunk","Beached"))
@@
 
-```{r vector maps, eval=FALSE, message=FALSE, echo=FALSE, include=TRUE}
+```{r vector maps, eval=FALSE, message=FALSE, echo=FALSE, include=FALSE}
 

### 2022-08-31 thesis_chapter_3_datawrangle.R

@@ -1,5 +1,335 @@
-##-- New passive simulations
+# Single trajectory
+#---Single simulations
+# Kelp simulations single
+
+kelp_hydro_summer_single <- tidync("OceanParcels/chapter_3_output/single_simulations/kelp_hydro_summer_single.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, age, Phase, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, waterdragU, waterdragV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+kelp_hydro_summer_single$type <- "Hydro" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_hydro_summer_single$velocity <- with(kelp_hydro_summer_single, sqrt(KelpVelU^2 + KelpVelV^2))
+
+kelp_hydrostks_summer_single<- tidync("OceanParcels/chapter_3_output/single_simulations/kelp_hydrostksdrag_summer_single.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, age, Phase, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, waterdragU, waterdragV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
+kelp_hydrostks_summer_single$type <- "Hydro and Stokes" # add extra column 'type' to dataframe and assigned 'kelp' label
+kelp_hydrostks_summer_single$velocity <- with(kelp_hydrostks_summer_single, sqrt(KelpVelU^2 + KelpVelV^2))
+
+kelp_hydrowind_summer_single<- tidync("OceanParcels/chapter_3_output/single_simulations/kelp_hydrowinddrag_summer_single.nc") %>%
+  hyper_tibble() %>%
+  dplyr::select(time, trajectory, lon, lat, distance, age, Phase, KelpVelU, KelpVelV, VelU, VelV, WindU, WindV, waterdragU, waterdragV) %>%
+  dplyr::rename(t = time) %>%
+  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
+
[…]
